### 2017-03-14 Latex/article.tex
Initial move to GitHub

\documentclass[USenglish]{article}

%\usepackage[utf8]{inputenc}
\RequirePackage[no-math]{fontspec}

\usepackage[small]{dgruyter}
\usepackage{microtype}

\newcommand{\wrt}{w.\,r.\,t.\ }
\newcommand{\Wrt}{W.\,r.\,t.\ }

\newcommand{\eg}{e.\,g.,}
\newcommand{\Eg}{E.\,g.,}

\newcommand{\ie}{i.\,e.,}
\newcommand{\Ie}{I.\,e.,}

\newcommand{\CxMdn}{\ensuremath{\mathsf{Ms_{[DN]}}}}
\newcommand{\CxMn}{\ensuremath{\mathsf{Ms_{[N]}}}}
\newcommand{\CxMan}{\ensuremath{\mathsf{Ms_{[AN]}}}}

\newcommand{\Sub}[1]{\ensuremath{\mathrm{_{#1}}}}
\newcommand{\Sup}[1]{\ensuremath{\mathrm{^{#1}}}}
\newcommand{\Subsf}[1]{\ensuremath{\mathsf{_{#1}}}}
\newcommand{\Supsf}[1]{\ensuremath{\mathsf{^{#1}}}}

\newcommand*\Rot{\rotatebox{75}}

\usepackage{gb4e-}

\usepackage{bbding}
\newcommand{\CheckIt}{\CheckmarkBold}

\usepackage{natbib}
\bibliographystyle{./unified}

\begin{document}

  %\articletype{...}

  \author*[1]{Roland Schäfer}
  %\runningauthor{...}
  \affil[1]{Freie Universität Berlin}
  \title{Competing Constructions for German Measure Phrases}
  \runningtitle{Competing Constructions for German Measure Phrases}
  %\subtitle{...}
  \abstract{In German, nominal morpho-syntax\ldots}
  \keywords{alternations, hierarchical models, corpus methods and experimental methods, measure constructions, German}
  %\classification[PACS]{...}
  %\communicated{...}
  %\dedication{...}
  %\received{...}
  %\accepted{...}
  \journalname{Submitted to Cognitive Linguistics}
  \journalyear{2017}
  %\journalvolume{...}
  %\journalissue{...}
  %\startpage{...}
  %\aop
  %\DOI{...}

\maketitle

\section{Case assignment in German measure NPs}
\label{sec:germanmeasurenps}

This paper deals with a morpho-syntactic alternation that occurs only in a very specific syntactic measure noun phrase construction.
By \textit{alternation} I refer to a situation where two or more forms or constructions are available with no clear difference in acceptability, function, or meaning.%
\footnote{In German grammar, the term \textit{Zweifelsfälle} `cases of doubt' is often used \citep{Klein2009}.}
I use it as an example case to introduce a protocol for the estimation and validation of corpus-based models in cognitively oriented corpus linguistics.
Therefore, I briefly introduce and illustrate the relevant construction (sub-)types in this section.
I also review the small body of existing literature on this specific alternation phenomenon, which is mostly descriptive.

I use the term \textit{measure noun phrase} (MNP) to refer to noun phrases (NPs) where a kind-denoting (count or mass) noun depends on another noun specifying a quantity of the objects or the substance denoted by the kind noun.
I call the kind-denoting noun simply the \textit{kind noun} and the quantity-denoting noun the \textit{measure noun}.%
\footnote{It would be more technically appropriate---but also more cumbersome---to speak of the \textit{kind noun phrase} because the embedded noun can be modified.}
Measure nouns can be all sorts of nouns which genuinely denote a quantity (such as \textit{litre} or \textit{amount}) but also nouns denoting containers, collections, etc. (such as \textit{glass} or \textit{bucket}).
Descriptively, three different situations need to be distinguished \wrt case assignment in German measure noun phrases.
If the kind noun comes with a determiner, the construction resembles (and is usually called) a \textit{pseudo-partitive} (on partitives and pseudo-partitives, e.\,g., \citealp{Barker1998,Selkirk1977,Stickney2007,Vos1999}; for a recent application of the terminology to German, see \citealp{Gerstenberger2015}), where the kind noun phrase is always in the genitive.
The embedded kind noun cannot agree in case with the measure head noun in this case.
I use * to denote general unacceptability, and the acceptability judgements given in (\ref{ex:intro:pseudopartitive}) are completely undisputed.
I refer to the construction in (\ref{ex:intro:pseudopartitive1}) as the \textit{Pseudo-partitive Genitive Construction} (PGC).

\begin{exe}
  \ex\label{ex:intro:pseudopartitive}
  \begin{xlist}
    \ex[ ]{\label{ex:intro:pseudopartitive1}\gll Wir trinken [[eine Tasse]\Sub{Acc} [des leckeren Kaffees]\Sub{Gen}]\Sub{Acc}.\\
    we drink a cup the tasty coffee\\
    \glt We drink a cup of the tasty coffee.}
    \ex[*]{\label{ex:intro:pseudopartitive2} Wir trinken [[eine Tasse]\Sub{Acc} [den leckeren Kaffee]\Sub{Acc}]\Sub{Acc}.}
  \end{xlist}
\end{exe}

When the kind noun is bare---\ie, when it comes without a determiner and without pre-nominal modification---the acceptability of the case patterns is reversed, and the only acceptable construction is usually classified as a \textit{Narrow Apposition Construction} \citep{Loebel1986}, henceforth NAC, where the kind noun agrees in case with the measure noun, see (\ref{ex:intro:narrowapposition}).

\begin{exe}
  \ex\label{ex:intro:narrowapposition}
  \begin{xlist}
    \ex[*]{\label{ex:intro:narrowapposition1} Wir trinken [[eine Tasse]\Sub{Acc} [Kaffees]\Sub{Gen}]\Sub{Acc}.}
    \ex[ ]{\label{ex:intro:narrowapposition2}\gll Wir trinken [[eine Tasse]\Sub{Acc} [Kaffee]\Sub{Acc}]\Sub{Acc}.\\
    we drink a cup coffee\\
    \glt We drink a cup of coffee.}
  \end{xlist}
\end{exe}

The construction as in (\ref{ex:intro:narrowapposition2}) is also referred to as the \textit{Direct Partitive Construction} for other Germanic languages where the PGC with the synthetic genitive is not available.
This nomenclature makes sense in contrast to the \textit{Indirect Partitive Construction} with the prepositional linkers translating to `of'---\ie, analytic genitives---in such languages, see \cite{HankamerMikkelsen2008} for Danish.
For German, this terminology is not distinctive enough, which is why I use the terms NAC and PGC.

The actual alternation can be observed \textit{only} when the kind noun occurs with an attributive adjective but without a determiner, as in (\ref{ex:intro:alternation}), where both the NAC in (\ref{ex:intro:alternation1}) and the PGC in (\ref{ex:intro:alternation2}) are not only both acceptable but also functionally and semantically identical.%
\footnote{Some more normative grammars take stronger positions w.\,r.\,t. the grammaticality of these two options.
  See \cite{Eschenbach1993,Zimmer2015} for analyses of the stances taken in more or less descriptive or normative grammars of German.
  As the usage and experimental data presented below should render it clear, there might be preferences, but we cannot assume either construction to be \textit{ungrammatical} or, for those who reject the notion of grammaticality, \textit{(virtually) unanimously unacceptable}.}

\begin{exe}
  \ex\label{ex:intro:alternation}
  \begin{xlist}
    \ex[ ]{\label{ex:intro:alternation2}\gll Wir trinken [[eine Tasse]\Sub{Acc} [heißen Kaffee]\Sub{Acc}]\Sub{Acc}.\\
    we drink a cup hot coffee\\
    \glt We drink a cup of hot coffee.}
    \ex[ ]{\label{ex:intro:alternation1} Wir trinken [[eine Tasse]\Sub{Acc} [heißen Kaffees]\Sub{Gen}]\Sub{Acc}.}
  \end{xlist}
\end{exe}

The distribution of the case patterns of NAC (case identity between measure noun and kind noun) an PGC (genitive on kind noun) depending on the structure of the kind NP are summarised in Table~\ref{tab:constructions}.

\begin{table}
  \centering
  \begin{tabular}{lccc}
    \multicolumn{1}{r}{kind NP:} & bare noun & with adjective & with determiner \\
    \midrule
    narrow apposition         & \CheckIt & \CheckIt & \\
    pseudo-partitive genitive &          & \CheckIt & \CheckIt \\
  \end{tabular}
  \caption{Distribution of the NAC and PGC constructions in different NP structures}
  \label{tab:constructions}
\end{table}

In the remainder of this chapter, I briefly discuss some questions regarding the situation shown in Table~\ref{tab:constructions}, given existing literature.
First, a major interest in generatively oriented linguistics (and formal semantics) is to determine the syntactic structure of the constructions in question.
In the case at hand, the question

% left/right-headed (Vos: DPC and DCC)
% role of the adjective in the NP
% everything-as-measure-noun and grammaticalisation (Brems)
% factors influencing (Hentschel)
% some remarks on the history: Zimmer

% Cite
% q-as-a-functional-category-loebel
% eschenbach__massangaben-im-kontext_KOMMENTIERT
% hentschel_1993_zfgl.1993.21.3.320
% loebel__apposition-in-der-quantifizierung
% Bergenholtz_1985_PPN345203720_0107_log9
% brems_ijcl2003-measure-nouns
% [] FUNKTIONALE KATEGORIEN IN NOMINALPHRASEN
% gallmann-2004
% Rutkowski_Grammaticalized_Partitives_Pseudo-Partitives
% Vater_H_2006

% On Case-OCP (from Götz Keydana):
% Und da habe ich dann auch einen Verweis auf den locus classicus gefunden:
% STEFAN A. FRISCH, JANET B. PIERREHUMBERT and MICHAEL B. BROE: SIMILARITY AVOIDANCE AND THE OCP. Natural Language & Linguistic Theory 22: 179–228, 2004.
% Dazu gäbe es dann noch, ebenfalls von Pierrehumbert,
% Dissimilarity in the Arabic Verbal Roots. In Amy Schafer, ed.,Proceedings of NELS 23 Amherst, MA: Graduate Linguistics Student Association. 1993. pp 367-381 
% Darin stellt sie wohl erstmals explizit einen Zusammenahng zwischen OCP und Kognition her.

% Brems (Language Sciences 53 (2016) 160–176) also on grammaticalisation of SINGLE measure noun.

%\footnote{More work on the syntax of such constructions from various theoretical points of view can be found in \cite{Bhatt1990}, \cite{GallmannLindauer1994}, \cite{Demske2001}, \cite{Gerstenberger2015}, etc.}
%\footnote{A similar broad notion of measure noun is used, for example, in \cite{Brems2003}.}
%\footnote{In many cases, a construction with \textit{von} `of' is also possible and maybe even more natural.
%I do not consider this construction because it is structurally different from the cases where the alternation occurs, cf.\ below.
%Interestingly, constructions of the type \textit{NP of\slash von NP} are the ones mostly discussed in the literature on (pseudo-)partitives and }

\section{Theory and method}

\subsection{Cognitively oriented corpus linguistics}
\label{ssec:cocl}

Following the descriptive introduction to the measure noun phrase alternation in Section~\ref{sec:germanmeasurenps}, this section situates the work presented here within the broader theoretical context of what can be called \textit{cognitively oriented corpus linguistics}.

Unless there are strong, numeric, and cognitively motivated hypotheses, some experimental validation is required to make a corpus study \textit{cognitive} in any sense.
Assuming the \textit{corpus as input} hypothesis without experimental validation is completely unsound in as much as researcher could extract any arbitrary---often spurious---usage pattern from a (usually noisy) corpus and claim that it has a cognitive reality.

% Cite:
% [Cognitive Linguistics] Similarity in linguistic categorization- The importance of necessary properties
% 10.1.1.980.9935
% __IMP__Measurenouns__cog-2015-0101
% Entrenchment_Stefanowitsch and Flach for submission

% Maybe:
% [Cognitive Linguistics] The processing of verb-argument constructions is sensitive to form, function, frequency, contingency and prototypicality
% [Cognitive Linguistics] Alternation-based generalizations are stored in the mental grammar- Evidence from a sorting task experiment //// along with CApelle's allostructions

% Method:
% Noble

\subsection{Statistical issues}
\label{ssec:issuesstatistics}

% Cite:
% __IMP__Measurenouns__cog-2015-0054 (Levshina)
% [Cognitive Linguistics] Towards cognitively plausible data science in language research (NDL)

% Stats:
% BarrEa_RandomEffectsStructure_2014_JMemLang
% ContentServer (1)
% importatn_quotes
% Menard_2000-Coefficients_of_Determination_for_Multiple_Logistic_Regression_Analysis

In the following, I examine the case alternation in German measure NPs under a cognitively oriented corpus linguistic perspective.
Since, however, the statistical tool kit (hierarchical generalised linear models, general tools such as $p$ values and confidence intervals, resampling techniques, model selection and model evaluation strategies, alternative algorithms for estimation, etc.) is heterogeneous and often criticised (especially under noisy conditions and when researcher run the risk of doing data mining instead of proper theory testing) I present a careful protocol for specifying, estimating, validating, and interpreting such models in Section \ref{sec:corpusmodels}.
Before working out the models, I show how I made sure that the relevant data is contained in the sample in Section \ref{sec:gettingdata}.
Furthermore, to provide external validation, I present two experiments in Section \ref{sec:externalvalidation} in which I attempted to find correlations between the corpus-based findings and speakers' behaviour.

\section{Getting the right data}
\label{sec:gettingdata}

% TODO Quote:
% Brems (2016:171--175) for using GOOGLE data, including COUNTS!

Before I turn to the concrete sampling procedure, a few words are in order regarding the choice of the corpus.
I used the German \textit{Corpus from the Web} (COW) DECOW14A (\citealp{SchaeferBildhauer2012shorti,Schaefer2015b} and \citealp{BiemannEa2013,SchaeferBildhauer2013} for overviews of web corpora in general and the methodology of their construction), which contains 21 billion tokens, for two main reasons.%
\footnote{The corpora are made available for free at \url{https://www.webcorpora.org}.
At the time of this writing, the DECOW16A corpus had already been released.}
First, the external validity of the study is increased through a higher heterogeneity of the sample \citep[30]{MaxwellDelaney2004}, and the DECOW14 corpus has clearly a much more heterogeneous composition compared to the only other very large corpus of German, the DeReKo \citep{KupietzEa2010} of the Institute for the German Language (IDS), which contains almost exclusively newspaper texts.%
\footnote{It was shown as a by-product in \cite{W16-2601} that, for example, the spread of topics is much smaller in DeReKo compared to DECOW14.}
Second, it was already mentioned that normative grammars disagree on whether both NAC and PGC are grammatical in the case at hand.
Thus, newspaper text or any other text that conforms strongly to normative standards might not represent the alternation phenomenon fully.
Web corpora, on the other hand, contain at least some amount of non-standard language from forums and similar sources.
For these or similar reasons, COW corpora have been used in a number of peer-reviewed publications, for example \cite{VanGoethemHiligsmann2014,VanGoethemHuening2015,MuellerS2014,Schaefer2016c,SchaeferSayatz2014,SchaeferSayatz2016,StefanowitschFlach2016,Zimmer2015}. 
Therefore, DECOW14A can be considered the obvious choice for this study.

In Section~\ref{sec:germanmeasurenps} it was mentioned that there are vague intuitions about lemma-specific preference effects (related to the kind noun) for either the genitive or the appositive construction to be used in measure NPs.
There are no previous empirical studies or substantive theoretical hypotheses that would justify taking a corpus sample based on a predetermined list of lemmas, neither for the measure noun nor the kind noun slot.
Instead, it appears necessary to obtain sample in which all (or at least most of the highly frequent) actually occurring combinations of kind noun and measure noun are represented.
This section describes the three-stage bootstrap process that was used to obtain such a sample, which was then used to estimate the influences of various contextual factors on the alternation (cf.\ Section~\ref{sec:corpusmodels}).
The bootstrap process takes into account the problems of imperfect automatically generated annotations.

Automatically generated annotations (such as lemma, part-of-speech, or morphological information) are never perfect, especially if the material deviates from the standard as understood by the creators of the tools used for the annotation (cf. \citealp{BeisswengerEa2016,GiesbrechtEvert2009} and chapter 4 of \citealp{SchaeferBildhauer2013}).
For examining grammatical alternations, relying on existing annotations is therefore unwise because there is no systematic way of assessing the accuracy of the annotation tools on the target constructions.
Therefore, I did not rely on the annotations for part-of-speech, noun phrase chunks, or noun case available in DECOW14A wherever it was crucial to sample with a near-perfect recall.%
\footnote{The recall is the percentage of retrieved data relative to the data that should have been found.
If, for example, the corpus contains one hundred targets and a query retrieves only 80 of them, the recall is 80\% (or 0.8).}

Instead, I bootstrapped a sample which most likely contained all reasonably frequent combinations of measure nouns and mass nouns, and which was almost perfectly clean (\ie it did contain virtually no erroneous non-target results).
The bootstrapping consisted of three steps:\\

\begin{enumerate}
  \item\label{enum:data:step1} bootstrapping a list of the one hundred most frequent mass nouns,
  \item\label{enum:data:step2} bootstrapping a list of all measure nouns with which the mass nouns co-occur in the unambiguous measure construction with a determiner, and
  \item\label{enum:data:step3} sampling the target constructions by querying each combination of mass noun and measure noun found in step \ref{enum:data:step2}.
\end{enumerate}

In step \ref{enum:data:step1}, I sampled a list of all nouns in the DECOW14A01 corpus sorted by their token frequency and manually went through it from the most frequent noun downwards, selecting the first one hundred mass nouns that occurred in the list.%
\footnote{DECOW14A01 is the first slice (roughly a twentieth) of the complete DECOW14A corpus.
It contains just over one billion tokens.
The creators assigned the documents randomly to the slices, such that each slice can be considered representative of the whole corpus \cite{Schaefer2015b}.}
For this step, I relied on the noun part-of-speech label because for frequent nouns, high tagging accuracy can be expected, especially given the fact that German nouns are always spelled with an initial capital letter.
Abstract nouns which partially behave like mass nouns (like \textit{Spaß} `fun’ or \textit{Gefahr} `danger’) were excluded because they are usually not quantified in the same way as concrete mass nouns.%
\footnote{They behave like mass nouns mostly in that they occur without a determiner in the singular.}
The hundredth selected mass noun was \textit{Schmuck} `jewelry’, which is the 3,054th most frequent noun in the original frequency list.

This resulting list of mass nouns was used in step \ref{enum:data:step2} to bootstrap a list of measure nouns co-occurring with the mass nouns from DECOW14A01.
In order to generate this list, I utilized the fact that a direct sequence of two bare nouns very often instantiates the bare-noun NAC if the second noun is a mass noun.
In other words, I searched for all combinations [N\Sub{1} N\Sub{2}] with N\Sub{2} instantiated as a mass noun lemma.
Then, the resulting lists of noun--noun combinations were each sorted by frequency in descending order and sieved manually to remove erroneous hits.%
\footnote{In order to speed up the procedure without losing much material, I stopped when I had reached noun–noun combinations with an absolute frequency of 1 if twenty combinations for the respective mass noun (with a frequency of one or higher) had already been found.}
The result was a list of 2,365 individual combinations of measure nouns and mass nouns.

In step \ref{enum:data:step3}, each of these 2,365 noun–noun combinations was queried in the target construction individually in each of the first ten slices of DECOW14A (roughly 10 billion tokens).
The scripts generating the actual queries were crafted so as to make sure that adjectives and nouns were searched in the appropriate forms.
(\ref{ex:data:query01}) shows an example query for a neuter mass noun like \textit{Wasser} (`water’) which finds relevant sequences such as the ones shown in (\ref{ex:data:query02}), which all mean `litre of clear water’.
The versions with the genitive singular \textit{Liters} and the dative plural \textit{Litern} are not shown in (2) for space reasons but are also found by the query in (1).
In terms of context, single sentences containing the target were exported.

\begin{exe}
  \ex{\label{ex:data:query01} \texttt{\small[ lemma="Liter" ] [ pos="ADJA" \& word=".+(en|em|es)" ] [ lemma="Wasser" ]}}
  \ex\label{ex:data:query02}
  \begin{xlist}
    \ex{\label{ex:data:query02a} Liter [klaren Wassers]\Sub{Gen}}
    \ex{\label{ex:data:query02b} Liter [klares Wasser]\Sub{Nom\slash Acc}}
    \ex{\label{ex:data:query02c} Liter [klarem Wasser]\Sub{Dat}}
  \end{xlist}
\end{exe}

To reduce the sample size for the manual annotation process, the final concordance was sampled from the results of these 2,365 queries.
Since the mass nouns in the sample were distributed according to the usual power law, I used all hits for mass nouns occurring less than one hundred times, but randomly sampled one hundred hits for each mass noun that occurred one hundred or more times.
The final sample contained 7,547 sentences.
Given the careful bootstrapping and sampling procedure described in this section, we can be highly sure that it contains all relevant and reasonably frequent noun–noun combinations in the target construction.

%%%  TODO  %%%
% – excluded "zwei Glas Wein" etc. because not part of alternation; NOT disambiguated for "ein Glas" (Vos 1999)

\section{Corpus-based models}
\label{sec:corpusmodels}

\subsection{Tools for modelling alternations}
\label{sec:righttools}

In this section, I discuss the relevant statistical approaches to modelling alternations based on corpus data.
I mainly focus on well-established Generalised Linear Models (GLMs) and their \textit{multi-level} or \textit{hierarchical} extensions, which are also called \textit{mixed models} (GLMMs).

% TODO CONTENT HERE

In Section~\ref{ssec:corpusnonhierarchicalmodel}, I briefly discuss why a non-hierarchical model -- while it is an option in principle -- has serious disadvantages for the data set under discussion (and similar data sets).
In Section~\ref{ssec:corpushierarchicalmodel}, I present the hierarchical model which is actually interpreted as the primary result of this study.
Section~\ref{ssec:bayesian} takes a short detour and shows (by example) that results from Bayesian estimation are not intrinsically different from those obtained from Maximum Likelihood methods, although Bayesian approaches have recently been touted to solve various problems facing applied statistics in various fields, even including cognitively oriented corpus linguistics.
Finally, I summarise and interpret the results of the corpus study in Section~\ref{ssec:modelssummary}.

\subsection{Descriptive statstics and non-hierarchical models}
\label{ssec:corpusnonhierarchicalmodel}

% > summary(pl$Casedrop)
%   0   1 
%  67 794 
% > summary(fem$Casedrop)
%   0   1 
% 652 462 
% > summary(mn$Genitive)
%    0    1 
% 3277  672

% Why didn't we use "mit einem Liter guten Benzin"?
% Generally noisy sample, impossible to tell whether it's an omitted genitive -s on the noun!
% If ChZ did automatic queries, most likely incorrect results because of noise.

\subsection{Hierarchical models}
\label{ssec:corpushierarchicalmodel}

In this section, I apply standard Maximum Likelihood methods for estimating the parameters of a multilevel model.
As suggested in Section~\ref{ssec:corpusnonhierarchicalmodel}, hierarchical models are better suited than non-hierarchical models given that we can expect lemma-specific effects, and that the respective factors have a high number of levels.

\begin{figure}[h]
\centering
\includegraphics[width=0.9\textwidth]{figures/corpus/04_glmm_fixef.pdf}
\caption{Fixed effects in GLMMs for Masculine\slash Neuter and Feminine with 95\% bootstrapped confidence intervals (1,000 replicates)}
\label{fig:glmmfixef}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=0.6\textwidth]{figures/corpus/04_glmm_fixeff_mn_Measureclass}
\caption{Effect plot for the masculine\slash neuter GLMM: Measure lemma class effect}
\label{fig:glmm:fixef:measureclass}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_mn_Attraction}~\includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_fem_Attraction}
\caption{Effect plot for the two GLMMs: attraction effect (centered and standardised) from the NAC construction with a bare kind noun and the PGC with a determiner; scales are not aligned between left and right panel}
\label{fig:glmm:fixef:attraction}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_mn_Matchlength}~\includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_fem_Matchlength}
\caption{Effect plot for the two GLMMs: effect of the (centered and standardised) length of the whole construction; scales are not aligned between left and right panel}
\label{fig:glmm:fixef:matchlength}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_mn_Measurecase}~\includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_fem_Measurecase}
\caption{Effect plot for the two GLMMs: effects of the case of the head (kind) noun; scales are not aligned between left and right panel}
\label{fig:glmm:fixef:measureclass}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_mn_Measurenumber}~\includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_fem_Measurenumber}
\caption{Effect plot for the two GLMMs: effect of the grammatical number of the head (kind) noun; scales are not aligned between left and right panel}
\label{fig:glmm:fixef:measurenumber}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_mn_Minus1pos}~\includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_fem_Minus1pos}
\caption{Effect plot for the two GLMMs: effect of part-of-speech preceding the construction; C = cardinal, D = determiner, O = \textit{other}, A = adjective, P = preposition; scales are not aligned between left and right panel}
\label{fig:glmm:fixef:minus1pos}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=0.5\textwidth]{figures/corpus/04_glmm_raneff_mn}~\hspace{0.05\textwidth}~\includegraphics[width=0.5\textwidth]{figures/corpus/04_glmm_raneff_fem}
\caption{Sample random effects for the masculine\slash neuter and feminine GLMMs}
\label{fig:glmm:fixef:minus1pos}
\end{figure}

\subsection{A note on Bayesian estimation}
\label{ssec:bayesian}

\subsection{Summary and conclusions}
\label{ssec:modelssummary}

\input{bigtable}

\section{Experimental cross-check of the corpus-based models}
\label{sec:externalvalidation}

% TODO SURPRISAL
%
% Bybee, Joan und Thompson, Sandra. 1997. “Three Frequency Effects in Syntax”, Proceedings of the Twenty-Third Annual Meeting of the Berkeley Linguistics Society, 378-388.
% Bybee, Joan. 2006. “From Usage to Grammar: The Mind‘s Response to Repetition”. Language 82(4): 711-733.
% Levy, Roger .2008. “Expectation-based syntactic comprehension.” Cognition 106: 1126–1177.
% Smith, Nathaniel, and Levy, Roger. 2013. “The effect of word predictability on reading time is logarithmic.” Cognition 128: 302–319.

\subsection{Experiment 1: forced choice}

As was pointed out in Section \ref{ssec:cocl}, there can be substantial doubt that corpus linguistic methods in themselves can be called \textit{cognitively oriented} in any meaningful way.
At the same time, it is a highly plausible assumption in a usage-based framework that large and varied corpora ideally represent language users' relevant input of (at least) written language, and that there should be reflexes of the distributions found in such corpora in the linguistic behaviour of language users.
Therefore, this section and the next present the results of two experiments wherein I cross-checked the corpus-based findings.

First, a very simple computer-aided forced-choice experiment was conducted.%
\footnote{The experiment is \textit{simple} compared to the sophisticated origins of forced choice experiments and their interpretation in psychophysics \citep[166--179]{MacmillanCreelman2005}.}
From a linguistic point of view, the forced choice setting has the advantage over rating studies that is does not force participants into making an absolute acceptability judgement on some scale.
Instead, they express a preference for one of two constructions (or words, etc.) both of which are acceptable.
While this does not totally exclude normative thinking---to which native speakers of German in so-called \textit{cases of doubt} are typically prone---

There were 24 participants aged 19 to 30 who were recruited from introductory linguistics courses at Freie Universität Berlin.
Although the experiment was conducted in the last four weeks of their first semester, participants had no deeper explicit knowledge of linguistics or grammar.%
\footnote{Even though the ratio of target sentences to fillers was relatively high (see below) and some of the fillers were also related to case alternations, no participant used the word \textit{Kasus} `case' when asked whether they had an idea what the experiment was about in informal interviews after the experiment.}
Only participants with an L1 background in German were recruited.

As stimuli, I used attested (but simplified) sentences from the corpus study in an approach similar to the one taken in \cite{DivjakEa2016}.
The rationale was that it is virtually impossible to test the large and complex sets of regressors used in the corpus study in a traditional randomized experiments where usually just one relevant factor (ideally something like treatment vs.\ non-treatment) is of interest.
Therefore, attested sentences were used, and the predictions of the corpus models were tested against participants' preferences.

Concretely, I sampled 16 attested sentences from the corpus and simplified them (mostly removing embedded clauses, complex adverbials, etc.) and standardized their orthography.
It was made sure that the simplifications and normalisations did not affect any of the regressors used in the corpus study, of course.
Eight sentences contained masculine or neuter kind nouns, the other eight contained feminine kind nouns.
Furthermore, in each of the M\slash N and F groups, four sentences with a case drop (\ie genitive or oblique case on the kind noun) and four withe case identity were chosen.
More precisely, the four sentences were sampled as highly prototypical examples of case drop (high logits) and case identity (low logits), respectively, according to the corpus based models.%
\footnote{Sentences were sorted by their corresponding $\hat{y}$ calculated from the model.
For the M\slash N sentences, sentences were sampled from the top and bottom 15\% of the resulting vector.
For the F sentences, the top and bottom 25\% were used because the overall sample size was smaller.}
It was made sure that the highly predictive features according to Section \ref{ssec:modelssummary} occurred in unique combinations among the four sentences in each cell.
This is summarised in Table \ref{tab:experiment1:design}.

\begin{table}
  \centering
  \begin{tabular}[h]{rcc}
     & M\slash N & F \\
     \midrule
     \textbf{high prob.\ for case drop} & 4 sentences\Supsf{a} & 4 sentences\Supsf{a} \\
     \textbf{low prob.\ for case drop} & 4 sentences\Supsf{a} & 4 sentences\Supsf{a} \\
  \end{tabular}
  \caption{The four groups of attested sentences chosen as stimuli (\Supsf{a}Among the 4 sentences, combinations of important factor values were made unique.)}
  \label{tab:experiment1:design}
\end{table}

The pairs of stimuli were the original attested sentence containing the preferred construction and a modified version containing the dispreferred construction.
They were presented next to each other, and a 20 second time limit for each choice was set.%
\footnote{No participant ever exceeded the time limit.}
The position on the screen (left\slash right) and the order of sentences were randomized for each participant.
As fillers, 23 pairs of sentences exemplifying similar alternation phenomena from German morpho-syntax were used.
Thus, participants saw 39 pairs of sentences and 78 sentences in total.
They were instructed to select from each pair of sentences the one that seemed more natural to them in the sense that they would use it rather than the other one.
The experiment was conducted using \textit{PsychoPy} \citep{Peirce2007}.

\begin{figure}[h]
\centering
\includegraphics[width=0.6\textwidth]{figures/experiment/2afc_proportions}
\caption{Participants' responses in the forced-choice task plotted against the predictions from the corpus-based model}
\label{fig:afc:proportions}
\end{figure}

Descriptively, Figure \ref{fig:afc:proportions} shows the relationship between the logits calculated from the corpus-based model and the proportional choices made by participants.
Visually, it looks like there is a close correlation between the corpus-derived preferences and preferences in the forced choice experiment.

Then, a multilevel logistic regression was specified as in Equation \ref{eq:model:afc}.

\begin{equation}
  \begin{split}
  Pr(y_i=1) & = logit^{-1}(\beta^0+\beta^1\cdot x_i^1+\beta^2\cdot x_i^2+\beta^{1.2}\cdot x_i^1\cdot x_i^2+ \\
  & + \alpha^{\text{Item}}_{j_i} + \alpha^{\text{Participant}}_{k_i}) \\
  \text{where}\ \ & x^1 = \text{\it Modelpred} \\
  & x^2 = \text{\it KindgenderMN} \\
  %ResponseDrop~Modelpred*Kindgender+(1|Item)+(1|Participant)
  \end{split}
  \label{eq:model:afc}
\end{equation}

Coefficients were estimated using Maximum Likelihood, and the coefficient table is given in Table \ref{tab:afc:model}, where instead of approximate standard errors and $z$ and $p$ values the lower and the upper bound of a bootstrapped 95\% confidence interval are specified.

\begin{table}
  \centering
  \begin{tabular}{lrrr}
    Effect & Estimate & 95\% CI lower & 95\% CI upper \\
    \midrule
    (Intercept)            &  0.748  &  0.242  &  1.265 \\
    Modelpred              &  1.154  &  0.905  &  1.497 \\
    KindgenderMN           &  0.135  & -0.416  &  0.632 \\
    Modelpred:KindgenderMN & -0.567  & -0.843  & -0.298 \\
  \end{tabular}
  \caption{Coefficient table for the GLMM specified in equation \ref{eq:model:afc}, analysing the results of the forced choice experiment}
  \label{tab:afc:model}
\end{table}

The number of observations was $n=768$.
The random effects for participant ($\sigma=0.7841$, 24 groups) and item ($\sigma=0.4869$, 16 groups) show considerable variance.
Accordingly, there is a notable difference between the marginal and conditional Nakagawa \& Schielzeth's $R^2_{m}=0.383$ and $R^2_{c}=0.510$, while both values are quite high.
Dropping the term of interest ($\beta^1\cdot x_i^1+\beta^2\cdot x_i^2+\beta^{1.2}\cdot x_i^1\cdot x_i^2$) leads to a significantly worse fit of the nested model both in a traditional Likelihood Ratio Test ($LR=55.172$, $p<0.0001$) and its bootstrapped PB alternative ($p = 0.002$, 493 of 500 requested samples used).
The effect display for the term is given in Figure \ref{fig:afc:effects}.

\begin{figure}[h]
\centering
\includegraphics[width=0.6\textwidth]{figures/experiment/2afc_effects}
\caption{Effect plot for the multilevel logistic regression in the forced-choice experiment: predictability of participants' choices from the logits of the corpus-based GLMM (cf.\ Section~\ref{ssec:corpushierarchicalmodel})}
\label{fig:afc:effects}
\end{figure}

In summary, the forced choice experiment clearly succeeded in validating the results from the corpus study in as much as the preferences extracted from the corpus correspond to native speakers' choices.
We can take this as an indication that there is some relation between the process generating the corpus data and the behaviour of individual speakers.
In the next section, I report the results of a self-paced reading experiment which yielded (maybe expectedly) less clear results.

\subsection{Experiment 2: self-paced reading}

%\begin{figure}[h]
%\centering
%\includegraphics[width=0.6\textwidth]{figures/experiment/spr_effects_modelPredDichotomousReg}
%\caption{Effect plot for the LMM in the self-paced reading experiment: modeling participants' residualised log reading times on the dichotomised predictions of the corpus-based GLMM (cf.\ Section~\ref{ssec:corpushierarchicalmodel})}
%\label{fig:spr:dichotomous:effects}
%\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=0.6\textwidth]{figures/experiment/spr_effects_case+modelPred+gender}
\caption{Effect plot for the LMM in the self-paced reading experiment: modeling participants' residualised log reading times on the predictions (logits) of the corpus-based GLMM (cf.\ Section~\ref{ssec:corpushierarchicalmodel})}
\label{fig:spr:continuous:effects}
\end{figure}

\section{Conclusions}

\begin{acknowledgement}
  I want to thank (in alphabetical order) Felix Bildhauer, Götz Keydana, and Ulrike Sayatz for stimulating discussions on relevant aspects of this paper.
  I also thank Ulrike Sayatz for recruiting the participants for the experiments.
  I am grateful to Kim Maser for her work on the annotation of the concordances.
  Furthermore, Samuel Reichert substantially helped to develop the guiding ideas behind this study by writing an excellent BA theses on the subject under my supervision.
  The research presented here was made possible in part through funding from the \textit{Deutsche Forschungsgemeinschaft} (DFG, personal grant SCHA1916/1-1).
\end{acknowledgement}

\bibliography{rs,cow}
\end{document}

### 2017-03-14 Latex/book.tex
Initial move to GitHub

\documentclass[openany,ngerman]{book}

\usepackage[utf8]{inputenc}%(only for the pdftex engine)
%\RequirePackage[no-math]{fontspec}%(only for the luatex or the xetex engine)
\usepackage[small]{dgruyter}
\usepackage{microtype}

\author{Max Mustermann}
\title{Funktion als Teil der literarischen Darstellung}
\distributionseries{De\,Gruyter Studium}
\seriestitle{Übergreifende Fragestellung zu Methode und Darstellungsweise}
\seriessubtitle{Studien und Vorunter suchungen zur kommentierten Überlieferung und Wirkung der Abweichungen vom Text~der Ausgabe und Fassung}
\serieseditor{Max Mustermann und Moritz Maier}
\seriesvolume{5/31}
\subtitle{Unterscheidung als Hilfs- und Orientierungsbegriff zur Dekodierung}
\editor{Max Mustermann als Editor}
\collaborator{Mustermann Stiftung}
\edition{2. Auf\/lage}
%\publisherlogo{dg-mouton}
\classification[Mathematics Subject Classification 2010]{35-02, 65-02, 65C30, 65C05, 65N35, 65N75, 65N80}
\authorinfo{Auch beim Konzept der ästhetischen Bildung von Wissen und dessen möglichst rasche und erfolgsorientierte Anwendung verspielen Einsichten und Gewinne ohne den Bezug auf die um 1900 entwickelten Argumentationen.\\\vskip\baselineskip Auch umfasst die Untersuchung in der Hauptsache den Zeitraum zwischen dem Inkrafttreten und der Darstellung in seiner heute geltenden Fassung. Ihre Funktion als Teil der literarischen Darstellung und narrativen Technik enthält auch einen vollständigen textkritischen Apparat und ein Verzeichnis der Stellen.}
\copyrighttext{Copyright-Text}
\isbn{978-3-11-021808-4}
\eisbnpdf{978-3-11-021809-1}
\issn{0179-0986}
\cover{Cover-Firma}
\typesetter{le-tex publishing services GmbH, Leipzig}
\printbind{Druckerei XYZ}

%\includeonly{%
%preface,
%chapter01,
%...
%}

\begin{document}
\frontmatter
\maketitle
\dedication{...\\ ...}
%\include{preface}

\mainmatter

\backmatter
%\bibliographystyle{plain}
%\printbibliography[env=bibnumeric]
\end{document}

%here comes an example for a contribution header:
\contribution

  \contributionauthor{Max Müller}
  \affil{Institute of Marine Biology, National Taiwan Ocean University, 2 Pei-Ning Road Keelung
    20224, Taiwan (R.O.C), e-mail: klpang@ntou.edu.tw}
  \runningauthor{Müller}
  \contributiontitle{Was ist so anders am Neuroenhancement? Pharmakologische und mentale Selbstveränderung im ethischen Vergleich}
  \runningtitle{Was ist so anders am Neuroenhancement?}
  \contributionsubtitle{Pharmakologische und mentale Selbstveränderung im ethischen Vergleich}
  \abstract{Konzept der ästhetischen Bildung von Wissen und dessen möglichst rasche und
    erfolgsorientierte Anwendung verspielen Einsichten und Gewinne ohne den Bezug auf die um 1900
    entwickelten Argumentationen. Auch umfasst der literarischen Übergreifende Fragestellung zur
    besseren Darstellungsweise enthält auch einen vollständigen textkritischen Apparat und ein
    Verzeichnis der Stellen.}
  \keywords{Funktion, Überlieferung, Untersuchung, Hauptsache, Zeitraum, Darstellung, Inkrafttreten,
    Fassung}
  \classification{65C05, 62M20, 93E11, 62F15, 86A22}

\makecontributiontitle
  \DOI{10.1515/futur-2012-0001}

### 2017-03-14 R/00_wrapper.R

@@ -1,0 +1,104 @@
+#
+# This calls all scripts in the right order to replicate results from my paper
+# 'Competing Constructions for German Measure Phrases' (working title)
+#
+# Warning: This takes quite a long times to run, especially with 
+#  =>  model coef & LR BOOTSTRAPPING
+#  =>  BAYESIAN / MCMC estimation
+
+# Comment this line if you want to keep your old workspace.
+rm(list = ls())
+
+# Let this point to location of R script sources.
+setwd('~/Massangaben/')
+
+# OPTIONS
+
+set.seed(2398651)            # Do NOT change, or results will differ.
+save.persistent <- T         # write to files instead of screen/console
+out.dir         <- "output/" # put generated files here
+
+round.in.big.table <- 2
+
+# Comment the next line if you want to keep old output files.
+unlink(paste(out.dir, '*', sep=""))
+
+# ----------   Don't touch anything past this line   ---------- #
+# ---------- If you do, results might not replicate. ---------- #
+
+# HELPERS
+
+source('glmtools.R')
+
+
+# SCRIPTS
+
+cat("\014")
+
+if (save.persistent) sink(paste(out.dir, "00_wrapper.txt", ""), append=F)
+cat("\nRoland Schäfer, Freie Universität Berlin\n")
+cat("roland.schaefer@fu-berlin.de\n")
+cat("http://orcid.org/0000-0003-3233-787\n")
+cat("'Competing Constructions for German Measure Phrases'\n")
+cat("Full results of corpus analysis as produced by R.\n")
+cat("Run 'wrapper.R' for 100% reproducible results!")
+cat("\n\nThis run BEGAN at ")
+print(Sys.time())
+if (save.persistent) sink()
+
+# Steps 1 and 2 are required for the subsequent steps.
+
+t.start <- Sys.time()
+
+cat("\n\n LOADING ...\n")
+source('01_load.R')       # imports and preparation/transformation of the data
+
+t.loaded <- Sys.time()
+
+cat("\n\n ATTRACTION ...\n")
+source('02_attraction.R') # calculate attraction factors
+
+t.attracted <- Sys.time()
+
+cat("\n\n EXPLORE ...\n")
+source('03_explore.R')    # exploratory analysis (determine order of adding regressors)
+
[…]

### 2017-03-14 R/01_load.R

@@ -1,0 +1,225 @@
+# 
+#  LOAD DATA
+#
+
+require(gnumeric)
+require(plyr)
+
+linMap <- function(x, from, to)
+  (x - min(x)) / max(x - min(x)) * (to - from) + from
+
+m <- read.gnumeric.sheet(file = "data/nan_sample_KM.ods", head = TRUE, sheet.name = "Data")
+pl <- read.gnumeric.sheet(file = "data/countnouns_sample_KM.ods", head = TRUE, sheet.name = "Data")
+
+# Backup original sentences.
+
+sentences <- m[, c("Leftcontext", "Match", "Rightcontext", "Originalsort", "Url")]
+
+# Remove some columns. These are the ones we KEEP:
+
+m <- m[,c("Measurecase",
+          "Kindlemma",
+          "Kindgender",
+          "Kindcase",
+          "Kindlength", 
+          "Kindorigin",
+          "Kindfinal",
+          "Kindconsistency",
+          "Kindedible",
+          "Kindfreq",
+          "Measurelemma",
+          "Measuregender",
+          "Measurenumber",
+          "Measureabbreviated",
+          "Measurelength",
+          "Measureclass",
+          "Measurefreq",
+          "Minus1pos",
+          "Minus2pos",
+          "Matchlength",
+          "Genitives",
+          "Badness",
+          "Originalsort"
+)]
+
+pl <- pl[,c("Measurecase",
+          "Kindlemma",
+          "Kindgender",
+          "Kindcase",
+          "Kindlength",
+          # "Kindorigin",
+          "Kindfinal",
+          # "Kindconsistency",
+          # "Kindedible",
+          "Kindfreq",
+          "Measurelemma",
+          "Measuregender",
+          "Measurenumber",
+          "Measureabbreviated",
+          "Measurelength",
+          # "Measureclass",
+          "Measurefreq",
+          "Minus1pos",
+          "Minus2pos",
+          "Matchlength",
+          "Genitives",
+          "Badness",
[…]

### 2017-03-14 R/02_attraction.R

@@ -1,0 +1,169 @@
+# This creates the quotient database for lemmas in the NN and NDN construction.
+# It also creates some plots and a text file for illustration.
+
+require(boot)
+
+# How the plural lemma pair list was extracted from the filtered list in data frame pl:
+# write.table(unique(sort(apply(pl[,c("Measurelemma", "Kindlemma")], 1, function(x) paste(x[1], x[2])))), "plural_pairs.txt", sep="\t", quote=F, col.names=F, row.names=F)
+
+prop.func <- function(v, smooth = 1) {
+  log( (v[2]+smooth) / (v[1]+smooth), 10)
+}
+
+ndn.kind <- read.delim("data/ndn_kind_freq.tsv", header=FALSE, quote="")
+nn.kind <- read.delim("data/nn_kind_freq.tsv", header=FALSE, quote="")
+
+# Pl has been pre-processed differently, can load in one table.
+countnouns_attracfreq <- read.delim("data/countnouns_attracfreq.tsv", quote="")
+
+# A cleanup.
+nn.kind <- nn.kind[-which(nn.kind$Lem=="rege"),]
+colnames(ndn.kind) <- c("Freq","Lem")
+colnames(nn.kind) <- c("Freq","Lem")
+
+# Get all lemmas for master table.
+# Background: We want values for all nouns, even those which occur in
+# only one of the NN/NDN tables. If not in one, count will just be 0.
+cx.kind.lem <- c(as.character(nn.kind$Lem), as.character(ndn.kind$Lem))
+cx.kind.lem <- unique(cx.kind.lem[order(cx.kind.lem)])
+
+# Function to extract freqs for both constructions with "0 tolerance" (haha).
+get_freq <- function(data, lem) {
+  n1 <- data[which(data$Lem == lem),1]
+  if (length(n1) == 0) n1 <- 0
+  n1
+}
+
+# Generate full table.
+cx.kind <- cbind(cx.kind.lem, lapply(cx.kind.lem, function(lem){get_freq(nn.kind,lem)}), lapply(cx.kind.lem, function(lem){get_freq(ndn.kind,lem)}))
+
[…]

### 2017-03-14 R/03_explore.R

@@ -1,0 +1,179 @@
+#
+# Explore by putting each potential regressor in simple GLM.
+# ( Just to determine an order in which to try and add them to GLMM. )
+#
+
+
+# A function to do simple cross-validation.
+
+crossval.glm <- function(model, K = 10) {
+  data <- model$data
+  response <- all.vars(model$formula)[1]
+  n <- nrow(data)
+  z <- n%/%K
+  mod <- n%%K
+  
+  # Randomize order of data.
+  idx <- sample(1:n, n, replace=F)
+  data <- data[idx,]
+  
+  corrs <- c()
+  
+  # Iterate over the folds.
+  for (i in 1:K) {
+    
+    if (i==K)
+      idx.heldout <- seq((i-1)*z+1, i*z+mod, 1)
+    else
+      idx.heldout <- seq((i-1)*z+1, i*z, 1)
+    
+    data.train    <- data[-idx.heldout,]
+    data.heldout  <- data[idx.heldout,]
+    
+    tryCatch({
+      model.new <- glm(model$formula, family=model$family, data=data.train)
+      preds <- ifelse(predict(model.new, newdata=data.heldout) < 0.5, 0, 1)
+      confmat <- table(data.heldout[,response], preds)
+      corr <- sum(diag(confmat))
+      rate <- 1-corr/sum(confmat)
+      corrs <- c(corrs, rate)
+    },
+    error = function(err) {
+      cat("Hoppala! Convergence error...")
+      # Yes, we just ignore errors! It's all about non-converging GLMs anyway.
+    })
+  }
+  corr.armean <- 1/mean(1/corrs)
+  corr.armean
+}
+
+
+# A function that performs a simple glm for each potential regressor.
+# Should give a hint in which order to try and add regressors to complex model.
+
+order.regressors <- function(df, target, ignores) {
+  require(fmsb)
+  require(boot)
+  
+  mat <- matrix(, nrow=length(names(df)), ncol=3)
+  colnames(mat) <- c("R2", "p(LRT)", "Error(10CV)")
+  rownames(mat) <- names(df)
+  
+  model.0 <- glm(df[,target]~1, family=binomial(link=logit))
+  i = 1
+  for (n in names(df)) {
+    if (!(n %in% ignores)) {
[…]

### 2017-03-14 R/04_glm.R

@@ -1,0 +1,141 @@
+require(fmsb)
+
+
+###### MASCULINE ######
+
+mn.glm <- glm(Genitive~1
+              +Measurelemma
+              #+Kindlemma       # prob 0 or 1 
+              +Genitives
+              +Minus1pos
+              #+Kindlength       # unstable in bootstrap
+              #+Measureclass     # prob 0 or 1
+              +Matchlength
+              +Measureabbreviated
+              +Measurelength
+              +Attraction
+              +Measurenumber
+              +Kindfinal
+              +Kindedible
+              +Badness
+              +Kindconsistency
+              +Kindorigin
+              #+Measuregender   # convergence
+              +Kindfreq
+              #+Minus2pos       # prob 0 or 1 
+              +Measurecase
+              +Kindint
+              +Kindgender
+#              +Measurefreq
+              , data=mn,  family=binomial(link=logit))
+
+mn.glm.r2 <- NagelkerkeR2(mn.glm)$R2
+mn.glm.pred <- ifelse(predict(mn.glm) < 0, 0, 1)
+mn.glm.cmat <- table(mn.glm.pred, mn$Genitive)
+mn.glm.corr <- sum(diag(mn.glm.cmat))/sum(mn.glm.cmat)
+mn.glm.base <- length(which(mn$Genitive == 0))/length(mn$Genitive)
+mn.glm.pre <- (mn.glm.corr-mn.glm.base)/(1-mn.glm.base)
+
+if (save.persistent) sink(paste(out.dir, "04_glm.txt", sep=""))
+cat("\n\n\nMASCULINE\n\n")
+print(summary(mn.glm, correlation=F))
+cat("\n\n")
+print(mn.glm.r2)
+cat("\n\n")
+cat("correct", mn.glm.corr)
+cat("\n\n")
+cat("λ", mn.glm.pre)
+cat("\n\n")
+if (save.persistent) sink()
+
+
+
+###### FEMININE ######
+
+
+fem.glm <- glm(Casedrop~1
+              +Measurelemma
+              #+Kindlemma       # prob 0 or 1
+              +Measureclass
+              +Minus1pos
+              +Measureabbreviated
+              +Kindconsistency
+              +Matchlength
+              #+Kindfinal        # no hypothesis
+              +Genitives
+              +Kindedible
[…]

### 2017-03-14 R/05_glmm.R

@@ -1,0 +1,489 @@
+require(lme4)
+require(gridExtra)
+require(MuMIn)
+require(effects)
+require(gridExtra)
+
+
+# ######### Some settings, conveniently at top of file. #########
+
+the.nAGQ <- 0         # 0 for speed, 1 for precision and slow(!) computation.
+ci.boot.nsim <- 1000  # Even 100 can be extremely slow! SET TO MIN. 1000 FOR PRODUCTION!!!
+ci.boot.modelcomp.nsim <- 500 # This takes ages, even with 10. Set to 500 for PRODUCTION!!!
+
+# ###############################################################
+
+
+# Handling convergence warnings:
+# https://rstudio-pubs-static.s3.amazonaws.com/33653_57fc7b8e5d484c909b615d8633c01d51.html
+
+# A faster BOBYQA optimizer.
+library(nloptr)
+defaultControl <- list(algorithm="NLOPT_LN_BOBYQA",xtol_rel=1e-6,maxeval=1e5)
+nloptwrap2 <- function(fn,par,lower,upper,control=list(),...) {
+  for (n in names(defaultControl)) 
+    if (is.null(control[[n]])) control[[n]] <- defaultControl[[n]]
+  res <- nloptr(x0=par,eval_f=fn,lb=lower,ub=upper,opts=control,...)
+  with(res,list(par=solution,
+                fval=objective,
+                feval=iterations,
+                conv=if (status>0) 0 else status,
+                message=message))
+}
+
+
+# Function to get prop of correct predictions.
+corr.prop <- function(model, observed, cutoff) {
+  preds <- ifelse(predict(model) < cutoff, 0, 1)
+  cmat <- table(preds, observed)
+  sum(diag(cmat))/sum(cmat)
+}
+
+
+# Function to get LRT for (G)LMMMs with bootstrapped CIs. This takes AGES!
+lmer.modelcomparison <- function(model, regressors, formula.target, ci = 0.95, nsim = 4, print.updated = F, print.diag = F) {
+  require(pbkrtest)
+
+  result = NULL
+  
+  for (i in 1:length(regressors)) {
+    formula.parts <- c(formula.target, regressors[1:i-1], regressors[i+1:length(regressors)])
+    formula.parts <- formula.parts[which(!is.na(formula.parts))]
[…]

### 2017-03-14 R/06_glmm-mcmc.R

@@ -1,0 +1,154 @@
+require(rstanarm)
+require(rstan)
+
+rstan_options(auto_write = TRUE)
+options(mc.cores = parallel::detectCores())
+
+# MASCULINE GLMM
+
+chains <- 4
+seed   <- 6976
+iter   <- 1000
+
+# MN
+
+mn.glmm.mcmc <- stan_glmer(Genitive~1
+                           +(1|Measurelemma)
+                           +(1|Kindlemma)
+                           +Genitives
+                           +Minus1pos
+                           #+Kindlength          # unstable in bootstrap
+                           +Measureclass
+                           +Matchlength
+                           +Measureabbreviated
+                           #+Measurelength       # virtual non-convergence
+                           +Attraction
+                           +Measurenumber
+                           +Kindfinal
+                           +Kindedible
+                           +Badness
+                           #+Kindconsistency     # virtual non-convergence
+                           #+Kindorigin          # virtual non-convergence
+                           +Kindfreq
+                           +Measurecase
+                           +Kindint
+                           +Measurefreq
+                           ,
+                 data=mn, family=binomial(link=logit),
+                 chains=chains, seed=seed, iter=iter,
+                 prior = normal(0, 2.5), prior_intercept = normal(0, 10),
+                 prior_ops = prior_options(prior_scale_for_dispersion = 5, min_prior_scale = 1e-12, scaled = TRUE),
+                 prior_covariance = decov(regularization = 1, concentration = 1, shape = 1, scale = 1),
+                 prior_PD = F
+                 )
+
+mn.baysum <- summary(mn.glmm.mcmc)
+
+
+# FEM
+
+fem.glmm.mcmc <- stan_glmer(Casedrop~1
+                            +(1|Measurelemma)
+                            +(1|Kindlemma)
+                            #+Measureclass         # non-convergence
+                            +Minus1pos
+                            +Measureabbreviated
+                            +Kindconsistency
+                            +Matchlength
+                            #+Kindfinal             # no hypothesis
+                            +Genitives
+                            +Kindedible
+                            #+Kindorigin           # fixeff matrix rank deficient
[…]

### 2017-03-14 R/10_sel_exp_stimuli.R

@@ -1,0 +1,105 @@
+library(plyr)
+
+mn.selectors <- c("Minus1pos", "Measurenumber", "Measurecase")
+prop.mn <- 0.15
+
+fem.selectors <- c("Minus1pos", "Measurenumber", "Measurecase")
+prop.fem <- 0.25
+
+sel.no <- 7
+per.cmb <- 2
+
+exp_sample <- function(df, model, selectors, do.tail = F, prop = 0.25, find = 5, per.combo = 1, sel.feature = "Genitive") {
+  pred.order <- order(predict(model))
+
+  if (do.tail) {
+    pred.osort <- df[tail(pred.order, prop*nrow(df)), "Originalsort"]
+  } else {
+    pred.osort <- df[head(pred.order, prop*nrow(df)), "Originalsort"]
+  }
+
+  exp.sel <- df[which(df$Originalsort %in% pred.osort), ]          # The actual data (part of big data frame).
+  exp.conf <- ddply(exp.sel, selectors, nrow)                      # Just the feat/val configs of top pred cases.
+  exp.conf  <- exp.conf[order(exp.conf$V1, decreasing = T),]       # The top summary/counting of feat/val configs.
+  cat("Found feature configurations: ", nrow(exp.conf), " (showing 5)\n")
+  print(head(exp.conf, 5))
+
+  seen.knouns <- character()
+  seen.mnouns <- character()
+  
+  found <- 0
+  i <- 0
+  
+  # For each feature permutation, find unseen kind/measure lemmas.
+  while (found < find) {
+    i <- i + 1
+        
+    cat("\nFeatures selected: ", paste(as.character(unlist(exp.conf[i, 1:ncol(exp.conf)-1])), collapse="_"), "\n")
+    
+    if (do.tail) { sel.feature.val <- 1 } else { sel.feature.val <- 0 }
+    tmp.sel <- exp.sel[which(exp.sel[,sel.feature] == sel.feature.val & unname(apply(exp.sel[,selectors], 1, function(x) { paste(as.character(x), collapse="_") })) == paste(as.character(unlist(exp.conf[i, 1:ncol(exp.conf)-1])), collapse="_")),]
+    
+    tmp.top.kind <- ddply(tmp.sel, "Kindlemma", nrow)
[…]

### 2017-03-14 R/11_stimuli_predict.R

@@ -1,0 +1,83 @@
+require(lme4)
+
+stimuli <- read.csv("~/Massangaben/data/stimuli.csv", header=TRUE)
+
+stims.fax  <- c("Measurecase",
+          "Kindlemma",
+          "Kindfinal",
+          "Kindconsistency",
+          "Kindedible",
+          "Kindint",
+          "Measurelemma",
+          "Measuregender", 
+          "Measurenumber",
+          "Measureabbreviated",
+          "Measureclass",
+          "Minus1pos")
+
+
+# Separate MN and F.
+
+stimuli.mn <- stimuli[1:8,]
+stimuli.fem <- stimuli[9:16,]
+
+
+# Make factors factors, using levels from original model df.
+
+for (fac in stims.fax) {
+  stimuli.mn[, fac]  <- factor(stimuli.mn[, fac], levels = levels(mn[, fac]))
+  stimuli.fem[, fac]  <- factor(stimuli.fem[, fac], levels = levels(fem[, fac]))
+}
+
+
+# Pull numeric values from original table because of standardization.
+
+get.value <- function(df, index.column, index.value, value.column) {
+  df[which(df[, index.column] == index.value)[1], value.column]
+}
+
+stimuli.mn$Kindfreq <- unlist(lapply(as.character(stimuli.mn$Kindlemma), function(x) get.value(mn, "Kindlemma", x, "Kindfreq")))
+stimuli.mn$Measurelength <- unlist(lapply(as.character(stimuli.mn$Measurelemma), function(x) get.value(mn, "Measurelemma", x, "Measurelength")))
+stimuli.mn$Measurefreq <- unlist(lapply(as.character(stimuli.mn$Measurelemma), function(x) get.value(mn, "Measurelemma", x, "Measurefreq")))
+stimuli.mn$Attraction <- unlist(lapply(as.character(stimuli.mn$Kindlemma), function(x) get.value(mn, "Kindlemma", x, "Attraction")))
+
+stimuli.fem$Kindfreq <- unlist(lapply(as.character(stimuli.fem$Kindlemma), function(x) get.value(fem, "Kindlemma", x, "Kindfreq")))
+stimuli.fem$Measurelength <- unlist(lapply(as.character(stimuli.fem$Measurelemma), function(x) get.value(fem, "Measurelemma", x, "Measurelength")))
[…]

### 2017-03-14 R/glmtools.R

@@ -1,0 +1,157 @@
+join.factors <- function(...) {
+  unlist(list(...))
+}
+
+factorify <- function(f) {
+  levelsize <- function(l) { s = length(which(f==l)) }
+  f <- factor(f)
+  o = levels(f)[order(unlist(lapply(levels(f), levelsize)), decreasing = T)]
+  f <- factor(f, o)
+}
+
+# Sort factor to put most r(esponse)=0 in f(actor) first,
+factorify.to.zero <- function(f, r, dec=F) {
+  f <- factor(f)
+  r <- factor(r)
+  levs <- colnames(table(r, f))[order(table(r, f)[1,]/table(r, f)[2,],decreasing=dec)]
+  f <- factor(f, levels=levs)
+}
+
+
+siglev <- function(p)
+{
+  if (p <= 0.1)
+    if (p <= 0.05)
+      if (p <= 0.01)
+        if (p <= 0.001)
+          sig <- "***"
+  else
+    sig <- "**"
+  else
+    sig <- "*"
+  else
+    sig <- "."
+  else
+    sig <- ""
+}
+
+siglev <- function(p)
+{
+  if (p <= 0.1)
+    if (p <= 0.05)
+      if (p <= 0.01)
+        if (p <= 0.001)
+          sig <- "***"
+  else
+    sig <- "**"
+  else
+    sig <- "*"
+  else
+    sig <- "."
+  else
+    sig <- ""
+}
+
+# This functions makes a plot of the COEFFICIENTS of significant factors.
+plot.coef <- function(glm, alpha = 0.05, pr="Pr(>|z|)", cex.labels = 1, pos.labels = 0.4, ...)
+{
+  # Extract p values and coefs.
+  p <- coef(summary(glm))[,pr]
+  c <- coef(glm)
+  
+  # Separate intercept from other coefs.
+  p.int <- as.numeric(p[1])
+  c.int <- as.numeric(c[1])
+  p <- p[-1]
+  c <- c[-1]
+  
+  # Find the order in which to put coefficients.
+  c <- c[which(p <= alpha)]
+  p <- p[which(p <= alpha)]
+  
+  # Find the order according to coef.
+  order <- order(c)
+  c <- c[order]
+  p <- p[order]
+  names <- names(c)
+  
+  # Put intercept and coefs back together.
+  names <- c("(Intercept)", names)
+  vals <- c(round(c.int, 3), round(c, 3))
[…]

### 2017-03-15 Latex/article.tex
finished descriptive section

@@
 
-\section{Case assignment in German measure NPs}
-\label{sec:germanmeasurenps}
+\section{Cognitively oriented corpus linguistics: theories and case studies}
+\label{sec:cogocl}
 
@@
 I use it as an example case to introduce a protocol for the estimation and validation of corpus-based models in cognitively oriented corpus linguistics.
-Therefore, I briefly introduce and illustrate the relevant construction (sub-)types in this section.
-I also review the small body of existing literature on this specific alternation phenomenon, which is mostly descriptive.
+
+\section{Case assignment in German measure NPs}
+\label{sec:germanmeasurenps}
+
+\subsection{Two stable cases and a case alternation}
+
+The alternation phenomenon to be discussed in this paper appears to be a simple case alternation.
+However, without a solid knowledge of German nominal morpho-syntax, some of the more subtle distinctions to be made cudl confuse readers.
+Therefore, I introduce and illustrate the relevant constructions rather thoroughly in this section, including a review of the relevant literature.
+The discussion defines the narrow range of syntactic constellations in which the alternation occurs and thus motivates the focus on \textit{only} this narrow range rather than, for example, the whole range of nominal constructions expressing quantities.
 
@@
 I call the kind-denoting noun simply the \textit{kind noun} and the quantity-denoting noun the \textit{measure noun}.%
-\footnote{It would be more technically appropriate---but also more cumbersome---to speak of the \textit{kind noun phrase} because the embedded noun can be modified.}
+\footnote{It would be more technically appropriate -- but also more cumbersome -- to speak of the \textit{kind noun phrase} because the embedded noun can be modified, cf.\ below.}
 Measure nouns can be all sorts of nouns which genuinely denote a quantity (such as \textit{litre} or \textit{amount}) but also nouns denoting containers, collections, etc.
 (such as \textit{glass} or \textit{bucket}).
+In a similar vein, \cite[284]{Brems2003} also considers nouns as measure nouns \textit{which, strictly speaking, do not designate a `measure', but display a more nebulous (sic!) potential for quantification}.
+
 Descriptively, three different situations need to be distinguished \wrt case assignment in German measure noun phrases.
-If the kind noun comes with a determiner, the construction resembles (and is usually called) a \textit{pseudo-partitive} (on partitives and pseudo-partitives, e.\,g., \citealp{Barker1998,Selkirk1977,Stickney2007,Vos1999}; for a recent application of the terminology to German, see \citealp{Gerstenberger2015}), where the kind noun phrase is always in the genitive.
-The embedded kind noun cannot agree in case with the measure head noun in this case.
+If the kind noun comes with a determiner, the construction resembles (and is usually called) a \textit{pseudo-partitive} (on partitives and pseudo-partitives, e.\,g., \citealp{Barker1998,Selkirk1977,Stickney2007,Vos1999}; for a recent application of the terminology to German, see \citealp{Gerstenberger2015}).%
+\footnote{Wheras partitives are constructions denoting a proper part-of relation such as \textit{a sip of the wine}, pseudo partitives -- albeit syntactically similar and diachronically related to partitives in many languages -- merely denote quantities such as \textit{a sip of wine}.
+See the references just given.}
 I use * to denote general unacceptability, and the acceptability judgements given in (\ref{ex:intro:pseudopartitive}) are completely undisputed.
+Notice that the kind noun is obligatory in the genitive and that case agreement with the measure noun is not an option (but see immediately below).
 I refer to the construction in (\ref{ex:intro:pseudopartitive1}) as the \textit{Pseudo-partitive Genitive Construction} (PGC).
@@
   \begin{xlist}
-    \ex[ ]{\label{ex:intro:pseudopartitive1}\gll Wir trinken [[eine Tasse]\Sub{Acc} [des leckeren Kaffees]\Sub{Gen}]\Sub{Acc}.\\
-    we drink a cup the tasty coffee\\
-    \glt We drink a cup of the tasty coffee.}
-    \ex[*]{\label{ex:intro:pseudopartitive2} Wir trinken [[eine Tasse]\Sub{Acc} [den leckeren Kaffee]\Sub{Acc}]\Sub{Acc}.}
+    \ex[ ]{\label{ex:intro:pseudopartitive1}\gll Wir trinken [[eine Tasse]\Sub{Acc} [eines leckeren Kaffees]\Sub{Gen}]\Sub{Acc}.\\
+    we drink a cup a tasty coffee\\
+    \glt We drink a cup of a tasty coffee.}
+    \ex[*]{\label{ex:intro:pseudopartitive2} Wir trinken [[eine Tasse]\Sub{Acc} [einen leckeren Kaffee]\Sub{Acc}]\Sub{Acc}.}
   \end{xlist}
@@
 
-When the kind noun is bare---\ie, when it comes without a determiner and without pre-nominal modification---the acceptability of the case patterns is reversed, and the only acceptable construction is usually classified as a \textit{Narrow Apposition Construction} \citep{Loebel1986}, henceforth NAC, where the kind noun agrees in case with the measure noun, see (\ref{ex:intro:narrowapposition}).
+When the kind noun is bare -- \ie\ when it comes without a determiner and without modifying adjectives -- the acceptability of the case patterns is reversed, and the only acceptable construction is usually classified as a \textit{Narrow Apposition Construction} \citep{Loebel1986}, henceforth NAC.
+In the NAC, the kind noun agrees in case with the measure noun, see (\ref{ex:intro:narrowapposition}).
+Notice that the unavailability of the genitive on the kind noun follows from a quirky general constraint that genitive NPs in German require the presence of some strongly case-marked element (determiner or adjective) in addition to the head noun in order to be acceptable \citep{GallmannLindauer1994,Schachtl1989}.
 
@@
 
-The construction as in (\ref{ex:intro:narrowapposition2}) is also referred to as the \textit{Direct Partitive Construction} for other Germanic languages where the PGC with the synthetic genitive is not available.
-This nomenclature makes sense in contrast to the \textit{Indirect Partitive Construction} with the prepositional linkers translating to `of'---\ie, analytic genitives---in such languages, see \cite{HankamerMikkelsen2008} for Danish.
-For German, this terminology is not distinctive enough, which is why I use the terms NAC and PGC.
+The construction as in (\ref{ex:intro:narrowapposition2}) is also referred to as the \textit{Direct Partitive Construction} for other Germanic languages where the PGC with the synthetic genitive is not available.%
+\footnote{This nomenclature makes sense in contrast to the \textit{Indirect Partitive Construction} with prepositional linkers translating to `of' -- \ie\ analytic genitives -- in such languages, see \cite{HankamerMikkelsen2008} for Danish.
+For German, this terminology is not distinctive enough, which is why I use the terms NAC and PGC.}
+While only the accusative is shown, case agreement of the kind noun is obligatory also with nominative and dative measure nouns.
 
-The actual alternation can be observed \textit{only} when the kind noun occurs with an attributive adjective but without a determiner, as in (\ref{ex:intro:alternation}), where both the NAC in (\ref{ex:intro:alternation1}) and the PGC in (\ref{ex:intro:alternation2}) are not only both acceptable but also functionally and semantically identical.%
+The actual alternation can be observed \textit{only when the kind noun occurs with an attributive adjective but without a determiner}, as in (\ref{ex:intro:alternation}), where both the NAC in (\ref{ex:intro:alternation1}) and the PGC in (\ref{ex:intro:alternation2}) are not only both acceptable but also functionally and semantically identical.%
 \footnote{Some more normative grammars take stronger positions w.\,r.\,t. the grammaticality of these two options.
-  See \cite{Eschenbach1993,Zimmer2015} for analyses of the stances taken in more or less descriptive or normative grammars of German.
-  As the usage and experimental data presented below should render it clear, there might be preferences, but we cannot assume either construction to be \textit{ungrammatical} or, for those who reject the notion of grammaticality, \textit{(virtually) unanimously unacceptable}.}
+  See \cite{Hentschel1993,Zimmer2015} for analyses of the stances taken in more or less descriptive or normative grammars of German.
+  As the usage and experimental data presented below should render it clear, there might be preferences under certain circumstances, but we cannot assume either construction to be \textit{ungrammatical} or -- for readers who reject the notion of grammaticality: \textit{(virtually) unanimously unacceptable}.}
+The distribution of the case patterns of NAC (case identity between measure noun and kind noun) an PGC (genitive on kind noun) depending on the structure of the kind NP are summarised in Table~\ref{tab:constructions}.
 
@@
 \end{exe}
-
-The distribution of the case patterns of NAC (case identity between measure noun and kind noun) an PGC (genitive on kind noun) depending on the structure of the kind NP are summarised in Table~\ref{tab:constructions}.
 
@@
 
-In the remainder of this chapter, I briefly discuss some questions regarding the situation shown in Table~\ref{tab:constructions}, given existing literature.
+I now turn to some subtler points related to the measure noun case alternation, namely:
+
+\begin{enumerate}
+  \item{\label{it:intro:datsg} alternative forms of dative singular neuter adjectives}
+  \item{\label{it:intro:femsg} case syncretism in feminine NPs}
+  \item{\label{it:intro:plurl} the parallel plural/collective construction(s)}
+  \item{\label{it:intro:noifl} grammaticalised non-inflected measure nouns}
+  \item{\label{it:intro:other} alternative constructions for expressing quantities}
+\end{enumerate}
+
+As for \ref{it:intro:datsg}, a complication with neuter kind nouns in the dative is mentioned by \cite[20--22]{Zimmer2015}.
+He reports a high number of occurrences of adjectives being inflected according to the \textit{weak} inflectional pattern, which is normally only used if a strongly inflected determiner precedes the adjective: \textit{rot-en Wein} instead of \textit{rot-em Wein}.
+In my corpus sample (to be made available after publication of this paper), this tendency was not nearly as clear, and there was a high number of very noisy sentences among those potentially showing this pattern.
+Consequently, I do not discuss the problem further.
+
+With feminine kind nouns, \ref{it:intro:femsg} comes into play.
+Whereas the singular masculine and neuter nominal subsystem still represents a four case system with a distinguishable genitive, feminine singular NPs show nominative-accusative and dative-genitive syncretisms.
+Table~\ref{tab:syncretisms} also shows the partial syncretism in the plural, where no gender distinctions are made (for an overview, see \citealp[XYZ]{Eisenberg2013a}).
+This means that with dative measure nouns, the alternation is unobservable if the kind noun is feminine, and it is problematic to speak of a \textit{genitive} kind noun construction in this case.
+This will be taken into account in Section~\ref{sec:corpusmodels} when the corpus data are discussed.
+
+\begin{table}
+  \centering
+  \begin{tabular}{llll}
+     & Masculine\slash Neuter & Feminine & Plural \\
+     \midrule
+     nominative & rot-er Wein    & \multirow{2}{*}{frisch-e Sahne}   & \multirow{2}{*}{klein-e Äpfel} \\
+     accusative & rot-en Wein    &                                   &                                \\
+     dative     & rot-em Wein    & \multirow{2}{*}{frisch-er Sahne}  & klein-en Äpfel-n               \\
+     genitive   & rot-en Wein-es &                                   & klein-er Äpfel                 \\
+  \end{tabular}
+  \caption{Case syncretisms in NPs with an adjective and without a determiner (\textit{roter Wein} `red wine', \textit{frische Sahne} `fresh cream', \textit{kleine Äpfel} `small apples')}
+  \label{tab:syncretisms}
+\end{table}
+
+Turning to \ref{it:intro:plurl}, readers might habe notices that so far only MNPs denoting quantities of substances (mass kind nouns such as \textit{ein Glas roter Wein\slash roten Weines} `a glass of red wine') have been discussed.
+If the kind noun is a plural count noun with a collective denotation such as \textit{ein Sack kleine Äpfel\slash kleiner Äpfel} `a bag of small apples' (see Table~\ref{tab:syncretisms} for the inflection), a similar alternation between PGC and NAC can be observed.%
+\footnote{On the collective interpretation of such constructions, see \cite{Eschenbach1994}.}
+In line with experimental results reported in \cite[15--16]{Zimmer2015}, I found that the PGC is so dominant with plural kind nouns (over 92\%, cf.\ Section~\ref{sec:corpusmodels}) that the alternation is virtually unanalysable.
+While this will play a role in the interpretation of the corpus findings, such plural\slash collective MNPs will not be dealt with prominently in the remainder of the paper.
+
+Turning to \ref{it:intro:noifl}, some measure nouns have been grammaticalised in a way that they always appear non-inflected.
+They are typical measure nouns like \textit{Gramm} `gram', \textit{Pfund} `pound' or \textit{Prozent} `percent'.
+I treat them like any other measure noun because they enter into both the NAC as in (\ref{ex:uninflectedmeasures:a}) and the PGC as in (\ref{ex:uninflectedmeasures:b}).
+The statistical models presented in Section~\ref{sec:corpusmodels} will, however, show whether they (or other highly grammaticalised measure nouns) favor the NAC or the PGC.
+
+\begin{exe}
+  \ex\label{ex:uninflectedmeasures}
+  \begin{xlist}
+    \ex{\label{ex:uninflectedmeasures:a} zwei Gramm brauner Zucker}
+    \ex{\label{ex:uninflectedmeasures:b}\gll zwei Gramm braunen Zuckers\\
+                                         two gram brown sugar\\
+				          two grams of brown sugar}
+  \end{xlist}
+\end{exe}
+
+Finally, \ref{it:intro:other} suggests that there are alternative ways of expressing similar quantifications.
+First, the analytic (pseudo-)partitive with \textit{von} `of' is only available as an alternative to the PGC if the kind noun phrase contains a determiner, cf. (\ref{ex:analyticalpartitive}).
+
+\begin{exe}
+  \ex\label{ex:analyticalpartitive} 
+  \begin{xlist}
+    \ex[ ]{\label{ex:analyticalpartitive:a}\gll ein Glas von dem roten Wein\\
+                                                a glass of the red wine\\
+                                                \glt a glass of the red wine
+       }
+    \ex[*]{\label{ex:analyticalpartitive:b}\gll ein Glas von rotem Wein\\
+                                                a glass of red wine\\
+                                                \glt a glass of red wine
+       }
+  \end{xlist}
+\end{exe}
+
+However, if the measure noun denotes a container, a construction with \textit{voll} (\textit{von})\slash\textit{voller} `full (of)' is available, see (\ref{ex:voll}).
+This construction has very specific properties discussed in \cite{Zeldes2018}.
+While subtle interactions between the \textit{voll}\slash\textit{voller} constructions and NAC\slash PGC cannot be excluded, there are no clear and motivated hypotheses about such interactions, and I do not discuss the construction further.
+
+\begin{exe}
+  \ex\label{ex:voll}
+  \begin{xlist}
+    \ex{\label{ex:voll:a}\gll ein Glas voll (von) rotem Wein\\
+                              a glass full of red wine\\
+  			      \glt a glass full of red wine}
+    \ex{\label{ex:voll:b}\gll ein Glas voller rotem Wein\\
+                              a glass full-of red wine\\
+  			      \glt a glass full of red wine}
+  \end{xlist}
+\end{exe}
+
+This concludes the descriptive overview of the alternation.
+
+\subsection{Analyses of the constructions}
+
+As stated in Section~\ref{sec:cogocl}, I consider it essential not to disregard generalisations discovered in descriptive, generative or formal frameworks, even if the accompanying formulations do not fit into probabilistic settings.
+This section therefore reviews some existing analyses.
 First, a major interest in generatively oriented linguistics (and formal semantics) is to determine the syntactic structure of the constructions in question.
@@
 Unless there are strong, numeric, and cognitively motivated hypotheses, some experimental validation is required to make a corpus study \textit{cognitive} in any sense.
-Assuming the \textit{corpus as input} hypothesis without experimental validation is completely unsound in as much as researcher could extract any arbitrary---often spurious---usage pattern from a (usually noisy) corpus and claim that it has a cognitive reality.
+Assuming the \textit{corpus as input} hypothesis without experimental validation is completely unsound in as much as researcher could extract any arbitrary -- often spurious -- usage pattern from a (usually noisy) corpus and claim that it has a cognitive reality.
 
@@
 Before I turn to the concrete sampling procedure, a few words are in order regarding the choice of the corpus.
-I used the German \textit{Corpus from the Web} (COW) DECOW14A (\citealp{SchaeferBildhauer2012shorti,Schaefer2015b} and \citealp{BiemannEa2013,SchaeferBildhauer2013} for overviews of web corpora in general and the methodology of their construction), which contains 21 billion tokens, for two main reasons.%
+I used the German \textit{Corpus from the Web} (COW) DECOW14(A) (\citealp{SchaeferBildhauer2012full,Schaefer2015b} and \citealp{BiemannEa2013,SchaeferBildhauer2013} for overviews of web corpora in general and the methodology of their construction), which contains 21 billion tokens, for two main reasons.%
 \footnote{The corpora are made available for free at \url{https://www.webcorpora.org}.
-At the time of this writing, the DECOW16A corpus had already been released.}
+At the time of this writing, the DECOW16 corpus had already been released.}
 First, the external validity of the study is increased through a higher heterogeneity of the sample \citep[30]{MaxwellDelaney2004}, and the DECOW14 corpus has clearly a much more heterogeneous composition compared to the only other very large corpus of German, the DeReKo \citep{KupietzEa2010} of the Institute for the German Language (IDS), which contains almost exclusively newspaper texts.%
-\footnote{It was shown as a by-product in \cite{W16-2601} that, for example, the spread of topics is much smaller in DeReKo compared to DECOW14.}
+\footnote{It was shown in \cite{W16-2601} that, for example, the spread of topics is much smaller in DeReKo compared to DECOW14.}
 Second, it was already mentioned that normative grammars disagree on whether both NAC and PGC are grammatical in the case at hand.
-Thus, newspaper text or any other text that conforms strongly to normative standards might not represent the alternation phenomenon fully.
+Thus, newspaper text or any other text that conforms strongly to normative standards might not represent the alternation phenomenon fully (and without bias) because authors and proof-readers might favour one alternative or the other.
 Web corpora, on the other hand, contain at least some amount of non-standard language from forums and similar sources.
 For these or similar reasons, COW corpora have been used in a number of peer-reviewed publications, for example \cite{VanGoethemHiligsmann2014,VanGoethemHuening2015,MuellerS2014,Schaefer2016c,SchaeferSayatz2014,SchaeferSayatz2016,StefanowitschFlach2016,Zimmer2015}.
-Therefore, DECOW14A can be considered the obvious choice for this study.
+Therefore, DECOW14 can be considered the obvious choice for this study.
 
-In Section~\ref{sec:germanmeasurenps} it was mentioned that there are vague intuitions about lemma-specific preference effects (related to the kind noun) for either the genitive or the appositive construction to be used in measure NPs.
-There are no previous empirical studies or substantive theoretical hypotheses that would justify taking a corpus sample based on a predetermined list of lemmas, neither for the measure noun nor the kind noun slot.
-Instead, it appears necessary to obtain sample in which all (or at least most of the highly frequent) actually occurring combinations of kind noun and measure noun are represented.
+In Section~\ref{sec:germanmeasurenps} it was mentioned that there are vague intuitions about lemma-specific preference effects w.\,r.\,t.\ the kind noun.
+Also, degrees of grammaticalisation of the measure noun might influence the choice of one of the constructions.
+Therefore, it was necessary to obtain sample in which all (or at least most of the highly frequent) actually occurring combinations of kind noun and measure noun were represented.
 This section describes the three-stage bootstrap process that was used to obtain such a sample, which was then used to estimate the influences of various contextual factors on the alternation (cf.\ Section~\ref{sec:corpusmodels}).
-The bootstrap process takes into account the problems of imperfect automatically generated annotations.
+The bootstrap process takes into account the imperfection of automatically generated annotations in corpora.
 
-Automatically generated annotations (such as lemma, part-of-speech, or morphological information) are never perfect, especially if the material deviates from the standard as understood by the creators of the tools used for the annotation (cf.
+Automatic linguistic annotation (for lemma, part-of-speech, morphological information, etc.) is never perfect, especially if the material deviates from the standard as understood by the creators of the tools used for the annotation (cf.
 \citealp{BeisswengerEa2016,GiesbrechtEvert2009} and chapter 4 of \citealp{SchaeferBildhauer2013}).
 For examining grammatical alternations, relying on existing annotations is therefore unwise because there is no systematic way of assessing the accuracy of the annotation tools on the target constructions.
-Therefore, I did not rely on the annotations for part-of-speech, noun phrase chunks, or noun case available in DECOW14A wherever it was crucial to sample with a near-perfect recall.%
+Therefore, I did not rely on the annotations for part-of-speech, noun phrase chunks, or noun case available in DECOW14 wherever it was crucial to sample with a near-perfect recall.%
 \footnote{The recall is the percentage of retrieved data relative to the data that should have been found.
@@
 
-Instead, I bootstrapped a sample which most likely contained all reasonably frequent combinations of measure nouns and mass nouns, and which was almost perfectly clean (\ie it did contain virtually no erroneous non-target results).
+Instead, I bootstrapped a sample which most likely contained all reasonably frequent combinations of measure nouns and mass nouns, and which was almost perfectly clean (\ie\ it did contain virtually no erroneous non-target results).
 The bootstrapping consisted of three steps:\\
@@
 
-In step \ref{enum:data:step3}, each of these 2,365 noun–noun combinations was queried in the target construction individually in each of the first ten slices of DECOW14A (roughly 10 billion tokens).
+In step \ref{enum:data:step3}, each of these 2,365 noun–noun combinations was queried in the target construction individually in each of the first ten slices of DECOW14 (roughly 10 billion tokens).
 The scripts generating the actual queries were crafted so as to make sure that adjectives and nouns were searched in the appropriate forms.
@@
 Instead, they express a preference for one of two constructions (or words, etc.) both of which are acceptable.
-While this does not totally exclude normative thinking---to which native speakers of German in so-called \textit{cases of doubt} are typically prone---
+While this does not totally exclude normative thinking -- to which native speakers of German in so-called \textit{cases of doubt} are typically prone --
 
@@
 Eight sentences contained masculine or neuter kind nouns, the other eight contained feminine kind nouns.
-Furthermore, in each of the M\slash N and F groups, four sentences with a case drop (\ie genitive or oblique case on the kind noun) and four withe case identity were chosen.
+Furthermore, in each of the M\slash N and F groups, four sentences with a case drop (\ie\ genitive or oblique case on the kind noun) and four withe case identity were chosen.
 More precisely, the four sentences were sampled as highly prototypical examples of case drop (high logits) and case identity (low logits), respectively, according to the corpus based models.%
@@
 
+\pagebreak
+
 \bibliography{rs,cow}

### 2017-03-15 Latex/article.tex
some work on previous analyses

@@
 \usepackage{microtype}
+\usepackage{enumerate}
 
@@
 (such as \textit{glass} or \textit{bucket}).
-In a similar vein, \cite[284]{Brems2003} also considers nouns as measure nouns \textit{which, strictly speaking, do not designate a `measure', but display a more nebulous (sic!) potential for quantification}.
+In a similar vein, \cite[284]{Brems2003} also considers nouns as measure nouns \textit{which, strictly speaking, do not designate a `measure', but display a more nebulous} (\textit{sic!}) \textit{potential for quantification}.
 
@@
 If the kind noun comes with a determiner, the construction resembles (and is usually called) a \textit{pseudo-partitive} (on partitives and pseudo-partitives, e.\,g., \citealp{Barker1998,Selkirk1977,Stickney2007,Vos1999}; for a recent application of the terminology to German, see \citealp{Gerstenberger2015}).%
-\footnote{Wheras partitives are constructions denoting a proper part-of relation such as \textit{a sip of the wine}, pseudo partitives -- albeit syntactically similar and diachronically related to partitives in many languages -- merely denote quantities such as \textit{a sip of wine}.
-See the references just given.}
+\footnote{Whereas partitives are constructions denoting a proper part-of relation such as \textit{a sip of the wine}, pseudo partitives -- albeit syntactically similar and diachronically related to partitives in many languages -- merely denote quantities such as \textit{a sip of wine}.
+See the references just given.
+In the literature on German, some authors incorrectly call the pseudo-partitive a \textit{partitive} \citep{Zimmer2015} while some realise the difference \citep[225]{Eschenbach1994}.}
 I use * to denote general unacceptability, and the acceptability judgements given in (\ref{ex:intro:pseudopartitive}) are completely undisputed.
@@
 
-\begin{enumerate}
+\begin{enumerate}[i.]
   \item{\label{it:intro:datsg} alternative forms of dative singular neuter adjectives}
@@
 If the kind noun is a plural count noun with a collective denotation such as \textit{ein Sack kleine Äpfel\slash kleiner Äpfel} `a bag of small apples' (see Table~\ref{tab:syncretisms} for the inflection), a similar alternation between PGC and NAC can be observed.%
-\footnote{On the collective interpretation of such constructions, see \cite{Eschenbach1994}.}
+\footnote{On the collective interpretation of such constructions, see \cite{Eschenbach1994}.
+It is, however, very surprising that she does not notice the morpho-syntactic distribution of PGC and NAC, even when she gets very close to listing the relevant data on p.\ 217.}
 In line with experimental results reported in \cite[15--16]{Zimmer2015}, I found that the PGC is so dominant with plural kind nouns (over 92\%, cf.\ Section~\ref{sec:corpusmodels}) that the alternation is virtually unanalysable.
@@
 Turning to \ref{it:intro:noifl}, some measure nouns have been grammaticalised in a way that they always appear non-inflected.
-They are typical measure nouns like \textit{Gramm} `gram', \textit{Pfund} `pound' or \textit{Prozent} `percent'.
-I treat them like any other measure noun because they enter into both the NAC as in (\ref{ex:uninflectedmeasures:a}) and the PGC as in (\ref{ex:uninflectedmeasures:b}).
+They are typical measure nouns like \textit{Gramm} `gram', \textit{Pfund} `pound' or \textit{Prozent} `percent'.%
+\footnote{With neglectable frequency, this effect is extended to container nouns like \textit{Glas} `glass'.
+A non-inflected variant \textit{zwei Glas roter Wein} and an inflected variant \textit{zwei Gläser roter Wein} `two glasses of wine' co-exist.
+The non-inflected variant denotes the quantity of wine fitting into two glasses, the inflected variant denotes two glasses filled with wine.}
+I treat these cases like any other measure noun because they enter into both the NAC as in (\ref{ex:uninflectedmeasures:a}) and the PGC as in (\ref{ex:uninflectedmeasures:b}).
 The statistical models presented in Section~\ref{sec:corpusmodels} will, however, show whether they (or other highly grammaticalised measure nouns) favor the NAC or the PGC.
@@
 As stated in Section~\ref{sec:cogocl}, I consider it essential not to disregard generalisations discovered in descriptive, generative or formal frameworks, even if the accompanying formulations do not fit into probabilistic settings.
-This section therefore reviews some existing analyses.
-First, a major interest in generatively oriented linguistics (and formal semantics) is to determine the syntactic structure of the constructions in question.
+This section therefore reviews some existing analyses of the OGC-NAC alternation or related issues, including generative approaches.
+
+% Grammaticalisation (Brems, Rutkowski)
+% Syntactic structure: headedness (Galmann>Löbel, Eschenbach p. 213, Vos: DPC/DCC), difference to English (Brems)
+% Syntactic structure: adjectives-as-determiners and layers in NP, role assignment to nouns in PGC/NAC
+%     Gallmann>Bhatt1990 DP[gen] vs NP[agree]
+% Influencing factors: Hentschel, Zimmer
+% Squeeze in Anttila & Fong
+
+A major interest in generatively oriented linguistics (and formal semantics) is to determine the syntactic structure of the constructions in question.
 In the case at hand, the question
-
-% left/right-headed (Vos: DPC and DCC)
-% role of the adjective in the NP
-% everything-as-measure-noun and grammaticalisation (Brems)
-% factors influencing (Hentschel)
-% some remarks on the history: Zimmer
-
-% Cite
-% q-as-a-functional-category-loebel
-% eschenbach__massangaben-im-kontext_KOMMENTIERT
-% hentschel_1993_zfgl.1993.21.3.320
-% loebel__apposition-in-der-quantifizierung
-% Bergenholtz_1985_PPN345203720_0107_log9
-% brems_ijcl2003-measure-nouns
-% [] FUNKTIONALE KATEGORIEN IN NOMINALPHRASEN
-% gallmann-2004
-% Rutkowski_Grammaticalized_Partitives_Pseudo-Partitives
-% Vater_H_2006
-
-% On Case-OCP (from Götz Keydana):
-% Und da habe ich dann auch einen Verweis auf den locus classicus gefunden:
-% STEFAN A. FRISCH, JANET B. PIERREHUMBERT and MICHAEL B. BROE: SIMILARITY AVOIDANCE AND THE OCP. Natural Language & Linguistic Theory 22: 179–228, 2004.
-% Dazu gäbe es dann noch, ebenfalls von Pierrehumbert,
-% Dissimilarity in the Arabic Verbal Roots. In Amy Schafer, ed.,Proceedings of NELS 23 Amherst, MA: Graduate Linguistics Student Association. 1993. pp 367-381 
-% Darin stellt sie wohl erstmals explizit einen Zusammenahng zwischen OCP und Kognition her.
-
-% Brems (Language Sciences 53 (2016) 160–176) also on grammaticalisation of SINGLE measure noun.
-
-%\footnote{More work on the syntax of such constructions from various theoretical points of view can be found in \cite{Bhatt1990}, \cite{GallmannLindauer1994}, \cite{Demske2001}, \cite{Gerstenberger2015}, etc.}
-%\footnote{A similar broad notion of measure noun is used, for example, in \cite{Brems2003}.}
-%\footnote{In many cases, a construction with \textit{von} `of' is also possible and maybe even more natural.
-%I do not consider this construction because it is structurally different from the cases where the alternation occurs, cf.\ below.
-%Interestingly, constructions of the type \textit{NP of\slash von NP} are the ones mostly discussed in the literature on (pseudo-)partitives and }
 
@@
 
-\begin{enumerate}
+\begin{enumerate}[i.]
   \item\label{enum:data:step1} bootstrapping a list of the one hundred most frequent mass nouns,
@@
 
-\subsection{Descriptive statstics and non-hierarchical models}
+\subsection{Descriptive statistics and non-hierarchical models}
 \label{ssec:corpusnonhierarchicalmodel}
@@
 
-% Why didn't we use "mit einem Liter guten Benzin"?
-% Generally noisy sample, impossible to tell whether it's an omitted genitive -s on the noun!
-% If ChZ did automatic queries, most likely incorrect results because of noise.
+% On Case-OCP (from Götz Keydana):
+% Und da habe ich dann auch einen Verweis auf den locus classicus gefunden:
+% STEFAN A. FRISCH, JANET B. PIERREHUMBERT and MICHAEL B. BROE: SIMILARITY AVOIDANCE AND THE OCP. Natural Language & Linguistic Theory 22: 179–228, 2004.
+% Dazu gäbe es dann noch, ebenfalls von Pierrehumbert,
+% Dissimilarity in the Arabic Verbal Roots. In Amy Schafer, ed.,Proceedings of NELS 23 Amherst, MA: Graduate Linguistics Student Association. 1993. pp 367-381 
+% Darin stellt sie wohl erstmals explizit einen Zusammenahng zwischen OCP und Kognition her.
 
@@
 
-\pagebreak
-
 \bibliography{rs,cow}

### 2017-03-15 Latex/article.tex
some work on previous analyses

@@
 \usepackage{enumerate}
+\usepackage{bm}
 
@@
 \newcommand{\Wrt}{W.\,r.\,t.\ }
-
 \newcommand{\eg}{e.\,g.,}
 \newcommand{\Eg}{E.\,g.,}
-
 \newcommand{\ie}{i.\,e.,}
 \newcommand{\Ie}{I.\,e.,}
-
-\newcommand{\CxMdn}{\ensuremath{\mathsf{Ms_{[DN]}}}}
-\newcommand{\CxMn}{\ensuremath{\mathsf{Ms_{[N]}}}}
-\newcommand{\CxMan}{\ensuremath{\mathsf{Ms_{[AN]}}}}
-
 \newcommand{\Sub}[1]{\ensuremath{\mathrm{_{#1}}}}
@@
 As stated in Section~\ref{sec:cogocl}, I consider it essential not to disregard generalisations discovered in descriptive, generative or formal frameworks, even if the accompanying formulations do not fit into probabilistic settings.
-This section therefore reviews some existing analyses of the OGC-NAC alternation or related issues, including generative approaches.
+This section therefore reviews some existing analyses of the PGC-NAC alternation or related issues, including generative approaches.
 
@@
 As suggested in Section~\ref{ssec:corpusnonhierarchicalmodel}, hierarchical models are better suited than non-hierarchical models given that we can expect lemma-specific effects, and that the respective factors have a high number of levels.
+
+\begin{equation}
+  \begin{split}
+    Pr(y_i=1)       & = logit^{-1}(\pmb{\beta}\mathbf{X_i} + \alpha^{\text{measurelemma}}_{j_i} + \alpha^{\text{kindlemma}}_{k_i}) \\
+    \text{where}\ \ & \mathbf{X_i}\ \text{is a vector of values for Badness, Genitives, } \\
+                    & \text{Matchlength, Measureabbreviated, Measurecase,} \\
+                    & \text{Measeurenumber, Minus1pos for}\ i \\
+    \text{and}\ \   & \pmb{\beta}\ \text{a corresponding coefficient vector} \\
+    \text{and}\ \   & \alpha^{\text{measurelemma}}_{j_i}\sim\mathcal{N}(\mu_j,\sigma_1) \\
+    \text{and}\ \   & \alpha^{\text{kindlemma}}_{k_i}\sim\mathcal{N}(\mu_k,\sigma_2) \\
+    \text{with}\ \  & \mu_1 \\
+  \end{split}
+  \label{eq:model:mnglmm}
+\end{equation}
 

### 2017-03-16 Latex/article.tex
work on previous analyses

@@
 \footnote{In German grammar, the term \textit{Zweifelsfälle} `cases of doubt' is often used \citep{Klein2009}.}
-I use it as an example case to introduce a protocol for the estimation and validation of corpus-based models in cognitively oriented corpus linguistics.
+I use it as an example case to introduce a protocol for the estimation and validation of corpus-based models in cognitively oriented corpus linguistics which I briefly motivate and describe in this introduction.
+
+% Add to CogL special issue: Look at what non-cognitivists do
+% => their have interesting data and interesting generalisations (which are not intrinsically nonsense or incompatible with CogLi)
+% => where they lack is modelling probabilistic effects, and that should be one of the strengths of cognitively oriented (corpus) linguistics
+% => FUZZINESS and SIMILARITY in classification (prototypes and exemplars)
+% => advantage: testable with corpus and experiment (ideally both)
+
+% This is a case study and more.
+% => gathering a descriptive picture from non-cognitive literature
+% => motivating the phenomenon as relevant for CogLi
+% => careful protocol for:
+%  =>> motivating range of alternants under examination and selecting hypotheses (sound and/or slightly less sound, but MAKE CLEAR what they are!)
+%  =>> motivate whether larger utterance context plays a role or not (sociolinguistic & register(?) commitment)
+%  =>> choosing right data
+%  =>> choosing "the right" – NO, "some appropriate" statistics
+%  =>> getting minimal external validation with experiments
+
+% This article attempts to follow this protocol as well as possible given the available analyses/data sources/nature of the phenomenon.
 
@@
 \subsection{Two stable cases and a case alternation}
+\label{sec:descriptive}
 
 The alternation phenomenon to be discussed in this paper appears to be a simple case alternation.
-However, without a solid knowledge of German nominal morpho-syntax, some of the more subtle distinctions to be made cudl confuse readers.
-Therefore, I introduce and illustrate the relevant constructions rather thoroughly in this section, including a review of the relevant literature.
-The discussion defines the narrow range of syntactic constellations in which the alternation occurs and thus motivates the focus on \textit{only} this narrow range rather than, for example, the whole range of nominal constructions expressing quantities.
+However, without a solid knowledge of German nominal morpho-syntax, some of the more subtle distinctions to be made can be confusing.
+Therefore, I introduce and illustrate the relevant constructions in some detail in this section, including a review of the relevant literature.
+The discussion defines the narrow range of syntactic constellations in which the alternation occurs and thus motivates the focus on \textit{only} this narrow range rather than the whole range of nominal constructions expressing quantities.
 
-I use the term \textit{measure noun phrase} (MNP) to refer to noun phrases (NPs) where a kind-denoting (count or mass) noun depends on another noun specifying a quantity of the objects or the substance denoted by the kind noun.
+I use the term \textit{measure noun phrase} (MNP) to refer a noun phrase (NP) in which a kind-denoting (count or mass) noun depends on another noun specifying a quantity of the objects or the substance denoted by the kind noun.
 I call the kind-denoting noun simply the \textit{kind noun} and the quantity-denoting noun the \textit{measure noun}.%
-\footnote{It would be more technically appropriate -- but also more cumbersome -- to speak of the \textit{kind noun phrase} because the embedded noun can be modified, cf.\ below.}
+\footnote{It would be more technically appropriate -- but also more cumbersome -- to speak of the \textit{kind noun phrase} because the embedded noun can be modified and form a full noun phrase, cf.\ immediately below and Section~\ref{sec:analyses} on the phrasal status of the constituents.}
 Measure nouns can be all sorts of nouns which genuinely denote a quantity (such as \textit{litre} or \textit{amount}) but also nouns denoting containers, collections, etc.
 (such as \textit{glass} or \textit{bucket}).
-In a similar vein, \cite[284]{Brems2003} also considers nouns as measure nouns \textit{which, strictly speaking, do not designate a `measure', but display a more nebulous} (\textit{sic!}) \textit{potential for quantification}.
+In a similar vein, \cite[284]{Brems2003} considers nouns as measure nouns \textit{which, strictly speaking, do not designate a `measure', but display a more nebulous} (\textit{sic!}) \textit{potential for quantification} (see also \citealp[530]{Koptjevskaja2001}, and \citealp[338]{Rutkowski2007}).
+For illustration purposes, in the English \textit{a cup of fine coffee}, \textit{cup} is the measure noun, and \textit{coffee} is the kind noun.
 
-Descriptively, three different situations need to be distinguished \wrt case assignment in German measure noun phrases.
-If the kind noun comes with a determiner, the construction resembles (and is usually called) a \textit{pseudo-partitive} (on partitives and pseudo-partitives, e.\,g., \citealp{Barker1998,Selkirk1977,Stickney2007,Vos1999}; for a recent application of the terminology to German, see \citealp{Gerstenberger2015}).%
+Descriptively, three different syntactic configurations need to be distinguished \wrt case assignment inside German measure noun phrases.
+If the kind noun comes with a determiner, the construction resembles (and is usually called) a \textit{pseudo-partitive} (on partitives and pseudo-partitives see, e.\,g., \citealp{Barker1998,Selkirk1977,Stickney2007,Vos1999}; for a recent application of the terminology to German, see \citealp{Gerstenberger2015}).%
 \footnote{Whereas partitives are constructions denoting a proper part-of relation such as \textit{a sip of the wine}, pseudo partitives -- albeit syntactically similar and diachronically related to partitives in many languages -- merely denote quantities such as \textit{a sip of wine}.
-See the references just given.
 In the literature on German, some authors incorrectly call the pseudo-partitive a \textit{partitive} \citep{Zimmer2015} while some realise the difference \citep[225]{Eschenbach1994}.}
@@
 In the NAC, the kind noun agrees in case with the measure noun, see (\ref{ex:intro:narrowapposition}).
-Notice that the unavailability of the genitive on the kind noun follows from a quirky general constraint that genitive NPs in German require the presence of some strongly case-marked element (determiner or adjective) in addition to the head noun in order to be acceptable \citep{GallmannLindauer1994,Schachtl1989}.
+Notice that the unavailability of the genitive on the kind noun follows from a quirky constraint that genitive NPs in German require the presence of some strongly case-marked element (determiner or adjective) in addition to the head noun in order to be acceptable \citep{GallmannLindauer1994,Schachtl1989}.
 
@@
 For German, this terminology is not distinctive enough, which is why I use the terms NAC and PGC.}
-While only the accusative is shown, case agreement of the kind noun is obligatory also with nominative and dative measure nouns.
+While only the accusative is shown, the kind noun obligatory agrees in case also with nominative and dative measure nouns.
 
-The actual alternation can be observed \textit{only when the kind noun occurs with an attributive adjective but without a determiner}, as in (\ref{ex:intro:alternation}), where both the NAC in (\ref{ex:intro:alternation1}) and the PGC in (\ref{ex:intro:alternation2}) are not only both acceptable but also functionally and semantically identical.%
-\footnote{Some more normative grammars take stronger positions w.\,r.\,t. the grammaticality of these two options.
-  See \cite{Hentschel1993,Zimmer2015} for analyses of the stances taken in more or less descriptive or normative grammars of German.
-  As the usage and experimental data presented below should render it clear, there might be preferences under certain circumstances, but we cannot assume either construction to be \textit{ungrammatical} or -- for readers who reject the notion of grammaticality: \textit{(virtually) unanimously unacceptable}.}
-The distribution of the case patterns of NAC (case identity between measure noun and kind noun) an PGC (genitive on kind noun) depending on the structure of the kind NP are summarised in Table~\ref{tab:constructions}.
+The actual alternation can be observed \textit{only when the kind noun occurs with an attributive adjective but without a determiner}, as in (\ref{ex:intro:alternation}), where both the NAC in (\ref{ex:intro:alternation1}) and the PGC in (\ref{ex:intro:alternation2}) are equally acceptable.
+They are also functionally and semantically equivalent.%
+\footnote{Some descriptive and normative grammars take stronger positions w.\,r.\,t. the grammaticality of these two options.
+See \cite{Hentschel1993,Zimmer2015} for analyses of the sometimes absurd stances taken in grammars of German.
+As the usage and experimental data presented below should render it clear, there might be preferences under certain circumstances, but we cannot assume either construction to be unacceptable.}
 
@@
     \multicolumn{1}{r}{kind NP:} & bare noun & with adjective & with determiner \\
+    & [\ldots N\Subsf{meas} [N\Subsf{kind}]] & [\ldots N\Subsf{meas} [AP N\Subsf{kind}]] & [\ldots N\Subsf{meas} [D N\Subsf{kind}]] \\
     \midrule
@@
 
-I now turn to some subtler points related to the measure noun case alternation, namely:
+The distribution of the case patterns in the NAC (case identity between the measure noun and the kind noun) and in the PGC (genitive on the kind noun) depending on the structure of the kind NP are summarised in Table~\ref{tab:constructions}.
+I now turn to some more subtle issue related to the measure noun case alternation, namely:
 
 \begin{enumerate}[i.]
-  \item{\label{it:intro:datsg} alternative forms of dative singular neuter adjectives}
+  \item{\label{it:intro:datsg} alternative \textit{weak} forms of dative singular neuter adjectives}
   \item{\label{it:intro:femsg} case syncretism in feminine NPs}
-  \item{\label{it:intro:plurl} the parallel plural/collective construction(s)}
+  \item{\label{it:intro:plurl} similar plural/collective construction(s)}
   \item{\label{it:intro:noifl} grammaticalised non-inflected measure nouns}
@@
 
-As for \ref{it:intro:datsg}, a complication with neuter kind nouns in the dative is mentioned by \cite[20--22]{Zimmer2015}.
-He reports a high number of occurrences of adjectives being inflected according to the \textit{weak} inflectional pattern, which is normally only used if a strongly inflected determiner precedes the adjective: \textit{rot-en Wein} instead of \textit{rot-em Wein}.
-In my corpus sample (to be made available after publication of this paper), this tendency was not nearly as clear, and there was a high number of very noisy sentences among those potentially showing this pattern.
+As for issue \ref{it:intro:datsg}, a complication with neuter kind nouns in the dative is mentioned by \cite[20--22]{Zimmer2015}.
+He reports a high number of occurrences of adjectives being inflected according to the \textit{weak} inflectional pattern, which is normally only used if a strongly inflected determiner precedes the adjective: \textit{kalt-en Wasser} `cold water' instead of \textit{kalt-em Wasser}.%
+\footnote{The \textit{strong} case and number suffixes of the adjectives are (with one exception) the suffixes of articles and pronouns, and they are used when no strongly inflected pronoun or determiner precedes the adjective.}
+In my corpus sample (released with this article), this tendency was not nearly as clear, and there was a high number of very noisy sentences among those potentially showing this pattern.
+Also, in my syntactic\slash semantic interpretation of the alternation (see Section~\ref{sec:corpusmodels}), the alternative weak forms are not expected to influence the alternation.
 Consequently, I do not discuss the problem further.
 
-With feminine kind nouns, \ref{it:intro:femsg} comes into play.
-Whereas the singular masculine and neuter nominal subsystem still represents a four case system with a distinguishable genitive, feminine singular NPs show nominative-accusative and dative-genitive syncretisms.
-Table~\ref{tab:syncretisms} also shows the partial syncretism in the plural, where no gender distinctions are made (for an overview, see \citealp[XYZ]{Eisenberg2013a}).
+\label{page:femininesyncretism}
+With feminine kind nouns, issue \ref{it:intro:femsg} becomes relevant.
+Whereas the singular masculine and neuter nominal subsystem is a four case system (at least for NPs with adjectives and determiners), feminine singular NPs show nominative-accusative and dative-genitive syncretisms.
+Table~\ref{tab:syncretisms} shows this and the partial syncretism in the plural, where no gender distinctions are made (for an overview, see \citealp[XYZ]{Eisenberg2013a}).
 This means that with dative measure nouns, the alternation is unobservable if the kind noun is feminine, and it is problematic to speak of a \textit{genitive} kind noun construction in this case.
@@
 
-Turning to \ref{it:intro:plurl}, readers might habe notices that so far only MNPs denoting quantities of substances (mass kind nouns such as \textit{ein Glas roter Wein\slash roten Weines} `a glass of red wine') have been discussed.
+Turning to \ref{it:intro:plurl}, readers might have notices that so far only MNPs denoting quantities of substances (mass kind nouns such as \textit{ein Glas roter Wein\slash roten Weines} `a glass of red wine') have been discussed.
 If the kind noun is a plural count noun with a collective denotation such as \textit{ein Sack kleine Äpfel\slash kleiner Äpfel} `a bag of small apples' (see Table~\ref{tab:syncretisms} for the inflection), a similar alternation between PGC and NAC can be observed.%
-\footnote{On the collective interpretation of such constructions, see \cite{Eschenbach1994}.
-It is, however, very surprising that she does not notice the morpho-syntactic distribution of PGC and NAC, even when she gets very close to listing the relevant data on p.\ 217.}
-In line with experimental results reported in \cite[15--16]{Zimmer2015}, I found that the PGC is so dominant with plural kind nouns (over 92\%, cf.\ Section~\ref{sec:corpusmodels}) that the alternation is virtually unanalysable.
+\footnote{\label{fn:eschenbash}On the collective interpretation of such constructions, see \cite{Eschenbach1994}.
+It is, however, very surprising that Eschenbach does not notice and discuss the morpho-syntactic distribution of PGC and NAC, even when she gets very close to listing the relevant data \citep[217]{Eschenbach1994}.}
+In line with experimental results reported in \cite[15--16]{Zimmer2015}, I found that the PGC is so dominant with plural kind nouns (over 92\%, cf.\ Section~\ref{sec:corpusmodels}) that the alternation cannot be analysed in the same way as in the singular case.
 While this will play a role in the interpretation of the corpus findings, such plural\slash collective MNPs will not be dealt with prominently in the remainder of the paper.
+A statistical model for plural kind nouns will be reported in passing in Section~\ref{sec:corpusmodels}, however.
 
@@
 They are typical measure nouns like \textit{Gramm} `gram', \textit{Pfund} `pound' or \textit{Prozent} `percent'.%
-\footnote{With neglectable frequency, this effect is extended to container nouns like \textit{Glas} `glass'.
-A non-inflected variant \textit{zwei Glas roter Wein} and an inflected variant \textit{zwei Gläser roter Wein} `two glasses of wine' co-exist.
+\footnote{With neglectable frequency, this effect is extended to container nouns like \textit{Glas} `glass', but with a twist.
+A non-inflected variant \textit{zwei Glas roter Wein} `two glasses of wine' and an inflected variant \textit{zwei Gläser roter Wein} co-exist.
 The non-inflected variant denotes the quantity of wine fitting into two glasses, the inflected variant denotes two glasses filled with wine.}
 I treat these cases like any other measure noun because they enter into both the NAC as in (\ref{ex:uninflectedmeasures:a}) and the PGC as in (\ref{ex:uninflectedmeasures:b}).
-The statistical models presented in Section~\ref{sec:corpusmodels} will, however, show whether they (or other highly grammaticalised measure nouns) favor the NAC or the PGC.
+The statistical models presented in Section~\ref{sec:corpusmodels} will, however, show whether they (or other highly grammaticalised measure nouns) favor the NAC or the PGC through the estimation of lemma-specific random intercepts.
 
@@
 
-Finally, \ref{it:intro:other} suggests that there are alternative ways of expressing similar quantifications.
-First, the analytic (pseudo-)partitive with \textit{von} `of' is only available as an alternative to the PGC if the kind noun phrase contains a determiner, cf. (\ref{ex:analyticalpartitive}).
+Finally, \ref{it:intro:other} suggests that there are alternative ways of expressing quantification.
+First, the analytic (pseudo-)partitive with \textit{von} `of' is only available as an alternative to the PGC if the kind noun phrase contains a (definite or indefinite) determiner as in (\ref{ex:analyticalpartitive}).
 
@@
 
-However, if the measure noun denotes a container, a construction with \textit{voll} (\textit{von})\slash\textit{voller} `full (of)' is available, see (\ref{ex:voll}).
-This construction has very specific properties discussed in \cite{Zeldes2018}.
-While subtle interactions between the \textit{voll}\slash\textit{voller} constructions and NAC\slash PGC cannot be excluded, there are no clear and motivated hypotheses about such interactions, and I do not discuss the construction further.
+Second, however, if the measure noun denotes a container, a construction with \textit{voll} (\textit{von})\slash\textit{voller} `full (of)' is available, see (\ref{ex:voll}).
 
@@
 
+This construction has very specific properties discussed in \cite{Zeldes2018}, and obviously is .
+While subtle interactions between the \textit{voll}\slash\textit{voller} constructions and NAC\slash PGC cannot be excluded, there are no clear and motivated hypotheses about such interactions, and I do not discuss the construction further.
 This concludes the descriptive overview of the alternation.
 
-\subsection{Analyses of the constructions}
+\subsection{Previous analyses}
+\label{sec:analyses}
 
-As stated in Section~\ref{sec:cogocl}, I consider it essential not to disregard generalisations discovered in descriptive, generative or formal frameworks, even if the accompanying formulations do not fit into probabilistic settings.
-This section therefore reviews some existing analyses of the PGC-NAC alternation or related issues, including generative approaches.
+This section reviews some existing analyses of the PGC-NAC alternation or related issues.
+I do \textit{not} review any of the semantics literature on partitives vs.\ pseudo-partitives because the constructions of interest here are all semantically pseudo-partitives, and thus none of the variance in the choice of the PGC and NAC can be expected to be explained by a discussion along these lines.
+The same is mostly true for the work on the grammaticalisation paths leading to the formation of pseudo-partitives (\eg\ \citealp{Brems2003,DeclerckBrems2016,Koptjevskaja2001,Rutkowski2007}).
+It is often assumed that pseudo-partitives arise as a form of grammaticalised partitives (\eg\ \citealp[536--539]{Koptjevskaja2001} for Finnish and Estonian, \citealp[559]{Koptjevskaja2001} for European languages per se).%
+\footnote{Notice that the description of German in \cite[547--549]{Koptjevskaja2001} (mostly referring \citealp{Eschenbach1993}) is partly incorrect, maybe because her main source \cite{Eschenbach1993} is not as clear as it could be (see also Footnote \ref{fn:eschenbash} above).
+Her example (29b) from \cite[71]{Eschenbach1993}, repeated here verbatim as (\ref{ex:koptjewskaja}) is misclassified as a clear case of a dative on the measure noun and a genitive on the kind noun.
 
-% Grammaticalisation (Brems, Rutkowski)
+\begin{exe}
+  \ex{\label{ex:koptjewskaja}\gll trotz einer Flasche gut-en Wein-s\\
+  in.spite.of one-\textsc{fem.gen} bottle good-\textsc{masc.sg.gen} wine-\textsc{gen}\\
+  \glt in spite of one bottle of good wine}
+\end{exe}
+
+It could, however, equally well be a genitive on the measure noun because of the syncretisms discussed here on p.\ \ref{page:femininesyncretism} and an ambiguity of the preposition \textit{troz}.
+Then, she does not clearly distinguish acceptable from non-acceptable constructions (cf.\ Section \ref{sec:descriptive}) on her pp.\ 548 and 549, including a rather confusing \textit{???} rating of her (31b), which is clearly a non-acceptable construction of the type illustrated in (\ref{ex:analyticalpartitive:b}) above.
+}
+
 % Syntactic structure: headedness (Galmann>Löbel, Eschenbach p. 213, Vos: DPC/DCC), difference to English (Brems)
@@
 In the case at hand, the question
+
+% Final statement: This is about choosing case in a specific configuration, not about choosing one construction from a broader range of constructions expressing quantification/measure.
+Maybe mention variable/variant and suggest alternation vs.
+variation as terminology.
+Mention that there are NO specific hypotheses about socio-linguistic influences and maybe weaker style/register hypotheses MIXED with semantic hypotheses about classes of kind nouns (which leads nicely to next section).
 
@@
 
-\subsection{Statistical issues}
-\label{ssec:issuesstatistics}
+\subsection{The right statistics}
+\label{ssec:rightstatistics}
+
+In this section, I briefly motivate the choice of statistical models which I use later in Section~\ref{sec:corpusmodels}.
+In the spirit of the careful and methodologically sound approach to cognitively oriented corpus linguistics laid out in Section~\ref{sec:cogocl}, authors should always provide at least a minimal motivation as part of the standard protocol, especially if there are alternative methods.
+
+% NO to exploratory methods if we have hypotheses
+% Testing (hypotheses about signs of regressors)
+% Estimation => bootstrapping
+% Model selection
+%  => random vs fixed is not so simple, rather matter of getting model to converge
+%  => Maximality
+%  => NO to Gries 2015
+% Model evaluation
+% VS: Bayesian hype (Levshina)
+% VS: alternative math (R-W)
 
@@
 
-In the following, I examine the case alternation in German measure NPs under a cognitively oriented corpus linguistic perspective.
-Since, however, the statistical tool kit (hierarchical generalised linear models, general tools such as $p$ values and confidence intervals, resampling techniques, model selection and model evaluation strategies, alternative algorithms for estimation, etc.) is heterogeneous and often criticised (especially under noisy conditions and when researcher run the risk of doing data mining instead of proper theory testing) I present a careful protocol for specifying, estimating, validating, and interpreting such models in Section \ref{sec:corpusmodels}.
-Before working out the models, I show how I made sure that the relevant data is contained in the sample in Section \ref{sec:gettingdata}.
-Furthermore, to provide external validation, I present two experiments in Section \ref{sec:externalvalidation} in which I attempted to find correlations between the corpus-based findings and speakers' behaviour.
+%In the following, I examine the case alternation in German measure NPs under a cognitively oriented corpus linguistic perspective.
+%Since, however, the statistical tool kit (hierarchical generalised linear models, general tools such as $p$ values and confidence intervals, resampling techniques, model selection and model evaluation strategies, alternative algorithms for estimation, etc.) is heterogeneous and often criticised (especially under noisy conditions and when researcher run the risk of doing data mining instead of proper theory testing) I present a careful protocol for specifying, estimating, validating, and interpreting such models in Section \ref{sec:corpusmodels}.
+
+Before discussing the results in Section~\ref{sec:corpusmodels}, I show how I made sure that the relevant data is contained in the sample in Section \ref{sec:gettingdata}.
 
@@
 
-\subsection{Tools for modelling alternations}
-\label{sec:righttools}
-
-In this section, I discuss the relevant statistical approaches to modelling alternations based on corpus data.
-I mainly focus on well-established Generalised Linear Models (GLMs) and their \textit{multi-level} or \textit{hierarchical} extensions, which are also called \textit{mixed models} (GLMMs).
-
-% TODO CONTENT HERE
-
-In Section~\ref{ssec:corpusnonhierarchicalmodel}, I briefly discuss why a non-hierarchical model -- while it is an option in principle -- has serious disadvantages for the data set under discussion (and similar data sets).
-In Section~\ref{ssec:corpushierarchicalmodel}, I present the hierarchical model which is actually interpreted as the primary result of this study.
-Section~\ref{ssec:bayesian} takes a short detour and shows (by example) that results from Bayesian estimation are not intrinsically different from those obtained from Maximum Likelihood methods, although Bayesian approaches have recently been touted to solve various problems facing applied statistics in various fields, even including cognitively oriented corpus linguistics.
-Finally, I summarise and interpret the results of the corpus study in Section~\ref{ssec:modelssummary}.
+%\subsection{Tools for modelling alternations}
+%\label{sec:righttools}
+%
+%In this section, I discuss the relevant statistical approaches to modelling alternations based on corpus data.
+%I mainly focus on well-established Generalised Linear Models (GLMs) and their \textit{multi-level} or \textit{hierarchical} extensions, which are also called \textit{mixed models} (GLMMs).
+%
+%% TODO CONTENT HERE
+%
+%In Section~\ref{ssec:corpusnonhierarchicalmodel}, I briefly discuss why a non-hierarchical model -- while it is an option in principle -- has serious disadvantages for the data set under discussion (and similar data sets).
+%In Section~\ref{ssec:corpushierarchicalmodel}, I present the hierarchical model which is actually interpreted as the primary result of this study.
+%Section~\ref{ssec:bayesian} takes a short detour and shows (by example) that results from Bayesian estimation are not intrinsically different from those obtained from Maximum Likelihood methods, although Bayesian approaches have recently been touted to solve various problems facing applied statistics in various fields, even including cognitively oriented corpus linguistics.
+%Finally, I summarise and interpret the results of the corpus study in Section~\ref{ssec:modelssummary}.
 

### 2017-03-16 Latex/article.tex
work on previous analyses

@@
 I treat these cases like any other measure noun because they enter into both the NAC as in (\ref{ex:uninflectedmeasures:a}) and the PGC as in (\ref{ex:uninflectedmeasures:b}).
-The statistical models presented in Section~\ref{sec:corpusmodels} will, however, show whether they (or other highly grammaticalised measure nouns) favor the NAC or the PGC through the estimation of lemma-specific random intercepts.
+See also Section~\ref{sec:analyses}.
 
@@
 I do \textit{not} review any of the semantics literature on partitives vs.\ pseudo-partitives because the constructions of interest here are all semantically pseudo-partitives, and thus none of the variance in the choice of the PGC and NAC can be expected to be explained by a discussion along these lines.
-The same is mostly true for the work on the grammaticalisation paths leading to the formation of pseudo-partitives (\eg\ \citealp{Brems2003,DeclerckBrems2016,Koptjevskaja2001,Rutkowski2007}).
+The same is mostly true for the work on the grammaticalisation of pseudo-partitives and measure noun constructions (\eg\ \citealp{Brems2003,DeclerckBrems2016,Koptjevskaja2001,Rutkowski2007}).
 It is often assumed that pseudo-partitives arise as a form of grammaticalised partitives (\eg\ \citealp[536--539]{Koptjevskaja2001} for Finnish and Estonian, \citealp[559]{Koptjevskaja2001} for European languages per se).%
@@
 }
+Since the alternating constructions are both clearly pseudo-partitives and not in any way partitives, this diachronic development (now completed) is not expected to have any effect on the alternation.
+More pertinent are approaches like \cite{Brems2003,DeclerckBrems2016}.
+In English -- \cite{Brems2003} argues based on a view from \cite{Langacker1991} -- strongly lexicalised and thus semantically bleached measure nouns become quantifiers instead of nouns, leading to a reanalysis of the NP wherein the kind noun becomes the head.
+This is supported by convincing syntactic evidence, for example subject-verb agreement.
+In German MNPs, the measure noun remains the head, however (see immediately below), and I know of no strong hypotheses stating that the degree of grammaticalisation of the measure noun influences the alternation in any way.
+The statistical models presented in Section~\ref{sec:corpusmodels} could in principle show, however, whether they (or other highly grammaticalised measure nouns) favor the NAC or the PGC through the estimation of lemma-specific random intercepts.
+
+A major interest in generatively oriented linguistics (and formal semantics) is to determine the syntactic structure of the constructions in question.
+In the case at hand, the question
 
@@
 %     Gallmann>Bhatt1990 DP[gen] vs NP[agree]
+%     Mention problem of no case assigner for NAC, so sub-syntactic process or theta theory wrong
 % Influencing factors: Hentschel, Zimmer
 % Squeeze in Anttila & Fong
-
-A major interest in generatively oriented linguistics (and formal semantics) is to determine the syntactic structure of the constructions in question.
-In the case at hand, the question
 
@@
 \label{ssec:cocl}
+
+% Mention: Corpus-as-input is OK as a hypothesis but NOT as a research hypothesis
+%  => otherwise any spurious pattern found in corpora (meeting certain criteria)
+%  => is automatically "cognitive"
 
@@
 
-\subsection{The right statistics}
+\subsection{There are no right statistics!}
 \label{ssec:rightstatistics}

### 2017-03-18 Latex/article.tex
theory/analyses

--- 2017-03-16 Latex/article.tex
@@
 \usepackage{enumerate}
-\usepackage{bm}
+\usepackage{eqnarray}
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{Cognitively oriented corpus linguistics: theories and case studies}
@@
 This paper deals with a morpho-syntactic alternation that occurs only in a very specific syntactic measure noun phrase construction.
-By \textit{alternation} I refer to a situation where two or more forms or constructions are available with no clear difference in acceptability, function, or meaning.%
-\footnote{In German grammar, the term \textit{Zweifelsfälle} `cases of doubt' is often used \citep{Klein2009}.}
-I use it as an example case to introduce a protocol for the estimation and validation of corpus-based models in cognitively oriented corpus linguistics which I briefly motivate and describe in this introduction.
+By \textit{alternation} I refer to a situation where two or more forms or constructions are available with no clear difference in acceptability, function, or meaning.
+I use it as an example case for a protocol for the estimation and validation of corpus-based models in cognitively oriented corpus linguistics.
+I briefly motivate this approach in this introduction, for details see Section~\ref{sec:cocl}.
+
+In German grammar, the term \textit{Zweifelsfälle} `cases of doubt' is often used \citep{Klein2009}.
 
@@
 
+%%%%%% From old structure:
+%
+%\subsection{Cognitively oriented corpus linguistics}
+%\label{sec:cocl}
+%
+% Mention: Corpus-as-input is OK as a hypothesis but NOT as a research programme
+% Entrenchment_Stefanowitsch and Flach for submission + Schmid C2C
+%  => otherwise any spurious pattern found in corpora (meeting certain criteria)
+%  => is automatically "cognitive"
+%
+%Following the descriptive introduction to the measure noun phrase alternation in Section~\ref{sec:germanmeasurenps}, this section situates the work presented here within the broader theoretical context of what can be called \textit{cognitively oriented corpus linguistics}.
+%
+%Unless there are strong, numeric, and cognitively motivated hypotheses, some experimental validation is required to make a corpus study \textit{cognitive} in any sense.
+%Assuming the \textit{corpus as input} hypothesis without experimental validation is completely unsound in as much as researcher could extract any arbitrary -- often spurious -- usage pattern from a (usually noisy) corpus and claim that it has a cognitive reality.
+%
+% Cite:
+% [Cognitive Linguistics] Similarity in linguistic categorization- The importance of necessary properties
+% 10.1.1.980.9935
+% __IMP__Measurenouns__cog-2015-0101
+%
+% Maybe:
+% [Cognitive Linguistics] The processing of verb-argument constructions is sensitive to form, function, frequency, contingency and prototypicality
+% [Cognitive Linguistics] Alternation-based generalizations are stored in the mental grammar- Evidence from a sorting task experiment //// along with CApelle's allostructions
+%
+% Method:
+% Noble
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{Case assignment in German measure NPs}
@@
 
-The alternation phenomenon to be discussed in this paper appears to be a simple case alternation.
+The alternation phenomenon to be discussed in this paper appears to be a straightforward case alternation at first sight.
 However, without a solid knowledge of German nominal morpho-syntax, some of the more subtle distinctions to be made can be confusing.
-Therefore, I introduce and illustrate the relevant constructions in some detail in this section, including a review of the relevant literature.
-The discussion defines the narrow range of syntactic constellations in which the alternation occurs and thus motivates the focus on \textit{only} this narrow range rather than the whole range of nominal constructions expressing quantities.
+Therefore, I introduce and illustrate the relevant constructions in some detail in this section, followed by a review of some relevant literature in Section~\ref{sec:analyses}.
+I define the narrow range of syntactic constellations in which the alternation occurs and thus motivate the focus on \textit{only} this narrow range rather than the whole range of nominal constructions expressing quantities.
+This entails that I also have to discuss certain constructions and factors that might be seen as relevant, but which I exclude from my own study.
 
@@
 (such as \textit{glass} or \textit{bucket}).
-In a similar vein, \cite[284]{Brems2003} considers nouns as measure nouns \textit{which, strictly speaking, do not designate a `measure', but display a more nebulous} (\textit{sic!}) \textit{potential for quantification} (see also \citealp[530]{Koptjevskaja2001}, and \citealp[338]{Rutkowski2007}).
+In a similar vein, \citet[284]{Brems2003} considers nouns as measure nouns \textit{which, strictly speaking, do not designate a `measure', but display a more nebulous} (\textit{sic!}) \textit{potential for quantification} (see also \citealp[530]{Koptjevskaja2001}, and \citealp[338]{Rutkowski2007}).
 For illustration purposes, in the English \textit{a cup of fine coffee}, \textit{cup} is the measure noun, and \textit{coffee} is the kind noun.
@@
 If the kind noun comes with a determiner, the construction resembles (and is usually called) a \textit{pseudo-partitive} (on partitives and pseudo-partitives see, e.\,g., \citealp{Barker1998,Selkirk1977,Stickney2007,Vos1999}; for a recent application of the terminology to German, see \citealp{Gerstenberger2015}).%
-\footnote{Whereas partitives are constructions denoting a proper part-of relation such as \textit{a sip of the wine}, pseudo partitives -- albeit syntactically similar and diachronically related to partitives in many languages -- merely denote quantities such as \textit{a sip of wine}.
-In the literature on German, some authors incorrectly call the pseudo-partitive a \textit{partitive} \citep{Zimmer2015} while some realise the difference \citep[225]{Eschenbach1994}.}
-I use * to denote general unacceptability, and the acceptability judgements given in (\ref{ex:intro:pseudopartitive}) are completely undisputed.
+\footnote{Whereas partitives are constructions denoting a proper part-of relation strictly requiring a definite kind noun as in \textit{a sip of the wine}, pseudo partitives -- albeit syntactically similar and diachronically related to partitives in many languages -- merely denote quantities and contain indefinite kind nouns as in \textit{a sip of wine}.
+In the literature on German, some authors incorrectly call the pseudo-partitive a \textit{partitive} \cite{Hentschel1993} while some realise the difference and at least mention it \cite{Eschenbach1994,GallmannLindauer1994,Loebel1989,Zimmer2015}.}
+I use \textdagger\ to denote general unacceptability, and the acceptability judgements given in (\ref{ex:intro:pseudopartitive}) are completely undisputed.
 Notice that the kind noun is obligatory in the genitive and that case agreement with the measure noun is not an option (but see immediately below).
@@
     \glt We drink a cup of a tasty coffee.}
-    \ex[*]{\label{ex:intro:pseudopartitive2} Wir trinken [[eine Tasse]\Sub{Acc} [einen leckeren Kaffee]\Sub{Acc}]\Sub{Acc}.}
+    \ex[\textdagger]{\label{ex:intro:pseudopartitive2} Wir trinken [[eine Tasse]\Sub{Acc} [einen leckeren Kaffee]\Sub{Acc}]\Sub{Acc}.}
   \end{xlist}
@@
 
-When the kind noun is bare -- \ie\ when it comes without a determiner and without modifying adjectives -- the acceptability of the case patterns is reversed, and the only acceptable construction is usually classified as a \textit{Narrow Apposition Construction} \citep{Loebel1986}, henceforth NAC.
-In the NAC, the kind noun agrees in case with the measure noun, see (\ref{ex:intro:narrowapposition}).
-Notice that the unavailability of the genitive on the kind noun follows from a quirky constraint that genitive NPs in German require the presence of some strongly case-marked element (determiner or adjective) in addition to the head noun in order to be acceptable \citep{GallmannLindauer1994,Schachtl1989}.
+When the kind noun is bare -- \ie\ when it comes without a determiner and without modifying adjectives -- the kind noun has to agree in case with the measure noun, and the genitive seen in the PGC is not acceptable, see (\ref{ex:intro:narrowapposition}).
+This construction is usually classified as a \textit{Narrow Apposition Construction} \citep{Loebel1986}, henceforth NAC.%
+\footnote{Notice that the unavailability of the genitive on the kind noun can be seen as following from a quirky constraint that genitive NPs in German require the presence of some strongly case-marked element (determiner or adjective) in addition to the head noun in order to be acceptable \citep{GallmannLindauer1994,Schachtl1989}.}
 
@@
   \begin{xlist}
-    \ex[*]{\label{ex:intro:narrowapposition1} Wir trinken [[eine Tasse]\Sub{Acc} [Kaffees]\Sub{Gen}]\Sub{Acc}.}
+    \ex[\textdagger]{\label{ex:intro:narrowapposition1} Wir trinken [[eine Tasse]\Sub{Acc} [Kaffees]\Sub{Gen}]\Sub{Acc}.}
     \ex[ ]{\label{ex:intro:narrowapposition2}\gll Wir trinken [[eine Tasse]\Sub{Acc} [Kaffee]\Sub{Acc}]\Sub{Acc}.\\
@@
 For German, this terminology is not distinctive enough, which is why I use the terms NAC and PGC.}
-While only the accusative is shown, the kind noun obligatory agrees in case also with nominative and dative measure nouns.
+While only the accusative is shown in (\ref{ex:intro:narrowapposition}), the kind noun obligatory agrees in case also with nominative and dative measure nouns.
 
@@
 They are also functionally and semantically equivalent.%
-\footnote{Some descriptive and normative grammars take stronger positions w.\,r.\,t. the grammaticality of these two options.
+\footnote{Some descriptive and normative grammars take stronger positions w.\,r.\,t.\ the acceptability of these two options.
 See \cite{Hentschel1993,Zimmer2015} for analyses of the sometimes absurd stances taken in grammars of German.
@@
 \begin{enumerate}[i.]
+  \item{\label{it:intro:idiot} alleged alternatives to the NAC and the PGC}
   \item{\label{it:intro:datsg} alternative \textit{weak} forms of dative singular neuter adjectives}
@@
 
-As for issue \ref{it:intro:datsg}, a complication with neuter kind nouns in the dative is mentioned by \cite[20--22]{Zimmer2015}.
+\vspace{-1\baselineskip}
+
+First of all, (\ref{it:intro:idiot}) refers to claims found in some grammars that a generic nominative, accusative, and even dative on the kind noun are used instead of the genitive or case agreement (overviews in \citealp{Hentschel1993,Zimmer2015}).
+It was shown empirically in \cite{Hentschel1993} that such variants are de facto not acceptable.
+Also, in my corpus sample, they simply did not occur.
+Even if they are accepted by some speakers, their extremely low frequency makes it impossible to study them using either corpus methods (Section~\ref{sec:corpusstudies}) or experimental approaches (Section~\ref{sec:externalvalidation}), and I consequently ignore them.
+
+As for (\ref{it:intro:datsg}), a complication with neuter kind nouns in the dative is mentioned by \citet[20--22]{Zimmer2015}.
 He reports a high number of occurrences of adjectives being inflected according to the \textit{weak} inflectional pattern, which is normally only used if a strongly inflected determiner precedes the adjective: \textit{kalt-en Wasser} `cold water' instead of \textit{kalt-em Wasser}.%
-\footnote{The \textit{strong} case and number suffixes of the adjectives are (with one exception) the suffixes of articles and pronouns, and they are used when no strongly inflected pronoun or determiner precedes the adjective.}
+\footnote{\label{fn:strongweak}The \textit{strong} case and number suffixes of the adjectives are identical (with one exception) to the suffixes of determiners, and they are used when no strongly inflected determiner precedes the adjective.}
 In my corpus sample (released with this article), this tendency was not nearly as clear, and there was a high number of very noisy sentences among those potentially showing this pattern.
-Also, in my syntactic\slash semantic interpretation of the alternation (see Section~\ref{sec:corpusmodels}), the alternative weak forms are not expected to influence the alternation.
-Consequently, I do not discuss the problem further.
+Hence, I do not discuss these forms much, although I will offer a possible interpretation in Section~\ref{sec:analyses}.
 
 \label{page:femininesyncretism}
-With feminine kind nouns, issue \ref{it:intro:femsg} becomes relevant.
+With feminine kind nouns, (\ref{it:intro:femsg}) is relevant.
 Whereas the singular masculine and neuter nominal subsystem is a four case system (at least for NPs with adjectives and determiners), feminine singular NPs show nominative-accusative and dative-genitive syncretisms.
@@
 This means that with dative measure nouns, the alternation is unobservable if the kind noun is feminine, and it is problematic to speak of a \textit{genitive} kind noun construction in this case.
-This will be taken into account in Section~\ref{sec:corpusmodels} when the corpus data are discussed.
+This is taken into account in Section~\ref{sec:corpusstudies} in the model design and the discussion.
 
@@
 
-Turning to \ref{it:intro:plurl}, readers might have notices that so far only MNPs denoting quantities of substances (mass kind nouns such as \textit{ein Glas roter Wein\slash roten Weines} `a glass of red wine') have been discussed.
+Turning to (\ref{it:intro:plurl}), readers might have notices that so far only MNPs denoting quantities of substances (mass kind nouns such as \textit{ein Glas roter Wein\slash roten Weines} `a glass of red wine') have been discussed.
 If the kind noun is a plural count noun with a collective denotation such as \textit{ein Sack kleine Äpfel\slash kleiner Äpfel} `a bag of small apples' (see Table~\ref{tab:syncretisms} for the inflection), a similar alternation between PGC and NAC can be observed.%
 \footnote{\label{fn:eschenbash}On the collective interpretation of such constructions, see \cite{Eschenbach1994}.
-It is, however, very surprising that Eschenbach does not notice and discuss the morpho-syntactic distribution of PGC and NAC, even when she gets very close to listing the relevant data \citep[217]{Eschenbach1994}.}
-In line with experimental results reported in \cite[15--16]{Zimmer2015}, I found that the PGC is so dominant with plural kind nouns (over 92\%, cf.\ Section~\ref{sec:corpusmodels}) that the alternation cannot be analysed in the same way as in the singular case.
+It is, however, unfortunate that Eschenbach does not notice and discuss the morpho-syntactic distribution of PGC and NAC, even when she gets very close to listing the relevant data \citep[217]{Eschenbach1994}.}
+In line with experimental results reported in \citet[15--16]{Zimmer2015}, I found that the PGC is so dominant with plural kind nouns (over 92\%, cf.\ Section~\ref{sec:corpusstudies}) that the alternation cannot be analysed in the same way as in the singular case.
 While this will play a role in the interpretation of the corpus findings, such plural\slash collective MNPs will not be dealt with prominently in the remainder of the paper.
-A statistical model for plural kind nouns will be reported in passing in Section~\ref{sec:corpusmodels}, however.
+A statistical model for plural kind nouns will be reported in passing in Section~\ref{sec:corpusstudies}, however.
 
-Turning to \ref{it:intro:noifl}, some measure nouns have been grammaticalised in a way that they always appear non-inflected.
-They are typical measure nouns like \textit{Gramm} `gram', \textit{Pfund} `pound' or \textit{Prozent} `percent'.%
+Turning to issue (\ref{it:intro:noifl}) now, some measure nouns have been grammaticalised in a way that they always appear non-inflected.
+They are typical measure nouns like \textit{Gramm} `gram', \textit{Pfund} `pound' or \textit{Prozent} `percent', which do not have plural forms at all.%
 \footnote{With neglectable frequency, this effect is extended to container nouns like \textit{Glas} `glass', but with a twist.
@@
 
-Finally, \ref{it:intro:other} suggests that there are alternative ways of expressing quantification.
-First, the analytic (pseudo-)partitive with \textit{von} `of' is only available as an alternative to the PGC if the kind noun phrase contains a (definite or indefinite) determiner as in (\ref{ex:analyticalpartitive}).
+Finally, (\ref{it:intro:other}) suggests that there are alternative ways of expressing quantification.
+First of all, the analytic (pseudo-)partitive with \textit{von} `of' is only available as an alternative to the PGC if the kind noun phrase contains a (definite or indefinite) determiner as in (\ref{ex:analyticalpartitive}).
 
@@
        }
-    \ex[*]{\label{ex:analyticalpartitive:b}\gll ein Glas von rotem Wein\\
+    \ex[\textdagger]{\label{ex:analyticalpartitive:b}\gll ein Glas von rotem Wein\\
                                                 a glass of red wine\\
@@
 
-Second, however, if the measure noun denotes a container, a construction with \textit{voll} (\textit{von})\slash\textit{voller} `full (of)' is available, see (\ref{ex:voll}).
+However, if the measure noun denotes a container, constructions with \textit{voll} (\textit{von})\slash\textit{voller} `full (of)' and \textit{mit} `with' are available, see (\ref{ex:alternatives}).
 
 \begin{exe}
-  \ex\label{ex:voll}
+  \ex\label{ex:alternatives}
   \begin{xlist}
-    \ex{\label{ex:voll:a}\gll ein Glas voll (von) rotem Wein\\
+  \ex{\label{ex:voll}\gll   ein Glas voll von rotem Wein\\
                             a glass full of red wine\\
 			    \glt a glass full of red wine}
-    \ex{\label{ex:voll:b}\gll ein Glas voller rotem Wein\\
+  \ex{\label{ex:voller}\gll ein Glas {voll\slash voller} {rotem\slash roter} Wein\\
                             a glass full-of red wine\\
 			    \glt a glass full of red wine}
+  \ex{\label{ex:mit}\gll    ein Glas mit rotem Wein\\
+                            a glass with red wine\\
+			    \glt a glass with red wine in it}
   \end{xlist}
@@
 
-This construction has very specific properties discussed in \cite{Zeldes2018}, and obviously is .
-While subtle interactions between the \textit{voll}\slash\textit{voller} constructions and NAC\slash PGC cannot be excluded, there are no clear and motivated hypotheses about such interactions, and I do not discuss the construction further.
+These construction have very specific properties (the construction with \textit{voller} are discussed in \citealp{Zeldes2018}, the construction with \textit{mit} is discussed in \citealp{Bhatt1990}), they are not semantically equivalent to NAC and PGC, and -- most importantly -- they are only available for a small subset of measure nouns.
+Subtle interactions between these constructions and NAC\slash PGC cannot be excluded, especially because the former might act as a (partial) alternative and thus reduce the count of exemplars of the NAC, the PGC, or both for container measure nouns.
+There are no clear and motivated hypotheses about such interactions, however.
+Nevertheless, the corpus-based models in Section~\ref{sec:corpusstudies} are specified in a way that they could detect preferences for either NAC or PGC for those nouns for which the alternative constructions illustrated in (\ref{ex:alternatives}) is available.
+
 This concludes the descriptive overview of the alternation.
+I have shown that there is an alternation between two very specific measure noun constructions under quite specific syntactic conditions, and that the two constructions can be distinguished clearly by looking at the case of the kind noun.
+It should be kept in mind that this paper is about this specific alternation and not about a possibly broader range of nominal constructions expressing quantification or measure.
+I turn to some more theory-oriented discussion in the next section.
 
-\subsection{Previous analyses}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{A theoretical assessment}
 \label{sec:analyses}
 
-This section reviews some existing analyses of the PGC-NAC alternation or related issues.
-I do \textit{not} review any of the semantics literature on partitives vs.\ pseudo-partitives because the constructions of interest here are all semantically pseudo-partitives, and thus none of the variance in the choice of the PGC and NAC can be expected to be explained by a discussion along these lines.
-The same is mostly true for the work on the grammaticalisation paths leading to the formation of pseudo-partitives (\eg\ \citealp{Brems2003,DeclerckBrems2016,Koptjevskaja2001,Rutkowski2007}).
-It is often assumed that pseudo-partitives arise as a form of grammaticalised partitives (\eg\ \citealp[536--539]{Koptjevskaja2001} for Finnish and Estonian, \citealp[559]{Koptjevskaja2001} for European languages per se).%
-\footnote{Notice that the description of German in \cite[547--549]{Koptjevskaja2001} (mostly referring \citealp{Eschenbach1993}) is partly incorrect, maybe because her main source \cite{Eschenbach1993} is not as clear as it could be (see also Footnote \ref{fn:eschenbash} above).
-Her example (29b) from \cite[71]{Eschenbach1993}, repeated here verbatim as (\ref{ex:koptjewskaja}) is misclassified as a clear case of a dative on the measure noun and a genitive on the kind noun.
+This section reviews some existing analyses of the PGC-NAC alternation or related issues, and I develop my own analysis and the appropriate hypotheses for the studies in Sections~\ref{sec:corpusstudies} and \ref{sec:externalvalidation}.
+While the constructions and the alternation are often covered in descriptive grammars such as \citet[XYZ]{Eisenberg2013a}, the theoretical literature has little specific discussion to offer.
+
+A caveat first:
+I do \textit{not} review the semantics literature on partitives and pseudo-partitives because the constructions of interest in the present paper are all pseudo-partitives, and thus none of the variance in the choice of the PGC and NAC can be expected to be explained by a discussion along these lines.
+The same is mostly true for the work on the grammaticalisation of pseudo-partitives and measure noun constructions, which has two main streams.
+On one hand, it is often assumed that pseudo-partitives arise as a form of grammaticalised partitives (\eg\ \citealp[536--539]{Koptjevskaja2001} for Finnish and Estonian, \citealp[559]{Koptjevskaja2001} for European languages per se).%
+\footnote{Notice that the description of German in \citet[547--549]{Koptjevskaja2001} is vague and partly incorrect, maybe because her main source \cite{Eschenbach1993} is not as clear as it could be (see also Footnote \ref{fn:eschenbash} above).
+Her example (29b) from \citet[71]{Eschenbach1993}, repeated here verbatim as (\ref{ex:koptjewskaja}) is misclassified as a clear case of a dative on the measure noun and a genitive on the kind noun.
 
@@
 
-It could, however, equally well be a genitive on the measure noun because of the syncretisms discussed here on p.\ \ref{page:femininesyncretism} and an ambiguity of the preposition \textit{troz}.
+It could, however, equally well be a genitive on the measure noun because of the syncretisms discussed here on p.\ \pageref{page:femininesyncretism} and an ambiguous case requirement of the preposition \textit{troz}.
 Then, she does not clearly distinguish acceptable from non-acceptable constructions (cf.\ Section \ref{sec:descriptive}) on her pp.\ 548 and 549, including a rather confusing \textit{???} rating of her (31b), which is clearly a non-acceptable construction of the type illustrated in (\ref{ex:analyticalpartitive:b}) above.
 }
+Since the alternating constructions are both clearly pseudo-partitives and not in any way partitives, this diachronic development (now completed) is not expected to have any effect on the alternation.
+On the other hand, there are more pertinent approaches like \cite{Brems2003,DeclerckBrems2016}.
+In English -- \cite{Brems2003} argues based on a view from \cite{Langacker1991} -- strongly lexicalised and thus semantically bleached measure nouns turn from nouns into quantifiers, leading to a reanalysis of the NP, wherein the kind noun becomes the head.
+This is supported by convincing syntactic evidence, for example subject-verb agreement.
+In German MNPs, the measure noun always remains the syntactic head, however (see immediately below for references), and I know of no strong hypotheses stating that the degree of grammaticalisation of the measure noun influences the alternation in any way.
+The statistical models presented in Section~\ref{sec:corpusstudies} could in principle show, however, whether highly grammaticalised measure nouns favor the NAC or the PGC through the estimation of lemma-specific random intercepts.
 
-% Syntactic structure: headedness (Galmann>Löbel, Eschenbach p. 213, Vos: DPC/DCC), difference to English (Brems)
+A major interest in traditional and generative linguistics is to find a plausible analysis of the hierarchical syntactic structure of the constructions in question.
+In the case at hand, the first question is which noun constitutes the head of the whole construction.
+This question was answered quite clearly by \cite{Loebel1986} already (see also \citealp[213]{Eschenbach1994} or \citealp[16]{GallmannLindauer1994}).
+Subject-verb agreement is always realised on the measure noun, and we can therefore assume that the measure noun is the head.
+Turning to the MNP-internal syntax, it cannot be understood properly without taking into account the strong\slash weak inflection patterns in German described in Section~\ref{sec:descriptive} (especially Footnote~\ref{fn:strongweak}).
+In NPs with a strongly inflected determiner, attributive adjectives inflect according to the so-called \textit{weak} pattern showing massive syncretism.
+If there is no determiner (as is the case in the alternating constructions), attributive adjectives inflect like determiners.
+Thus, the adjectives in the target constructions have properties of adjectives (they are lexical adjectives) and determiners (inflection and leftmost position in the NP) in a sense.
+In this light, the pattern shown in Table~\ref{tab:constructions} makes perfect sense, especially in a probabilistic framework:
+If speakers use classify the adjective as a determiner, they have to use the PGC because with a determiner, the PGC is the only option.
+If speakers classify the adjective as an adjective, the kind NP has no determiner, and the NAC is the only option.
+While the generative analysis presented in \cite{Bhatt1990} cannot deal with such probabilistic effects, she comes quite close to this interpretation by analysing the kind NP in the GPC as a DP and in the NAC as an NP.
+
 % Syntactic structure: adjectives-as-determiners and layers in NP, role assignment to nouns in PGC/NAC
-%     Gallmann>Bhatt1990 DP[gen] vs NP[agree]
+%     Gallmann => Bhatt1990 DP[gen] vs NP[agree]
+%     Mention problem of no case assigner for NAC, so sub-syntactic process or theta theory wrong
+
+Finally, I collect some hypotheses (or maybe mostly observations) from the existing literature as to which factors influence the choice of NAC and PGC.
+
 % Influencing factors: Hentschel, Zimmer
@@
 
-A major interest in generatively oriented linguistics (and formal semantics) is to determine the syntactic structure of the constructions in question.
-In the case at hand, the question
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-% Final statement: This is about choosing case in a specific configuration, not about choosing one construction from a broader range of constructions expressing quantification/measure.
-Maybe mention variable/variant and suggest alternation vs.
-variation as terminology.
-Mention that there are NO specific hypotheses about socio-linguistic influences and maybe weaker style/register hypotheses MIXED with semantic hypotheses about classes of kind nouns (which leads nicely to next section).
+\section{Corpus study}
+\label{sec:corpusstudies}
 
-\section{Theory and method}
-
-\subsection{Cognitively oriented corpus linguistics}
-\label{ssec:cocl}
-
-Following the descriptive introduction to the measure noun phrase alternation in Section~\ref{sec:germanmeasurenps}, this section situates the work presented here within the broader theoretical context of what can be called \textit{cognitively oriented corpus linguistics}.
-
-Unless there are strong, numeric, and cognitively motivated hypotheses, some experimental validation is required to make a corpus study \textit{cognitive} in any sense.
-Assuming the \textit{corpus as input} hypothesis without experimental validation is completely unsound in as much as researcher could extract any arbitrary -- often spurious -- usage pattern from a (usually noisy) corpus and claim that it has a cognitive reality.
-
-% Cite:
-% [Cognitive Linguistics] Similarity in linguistic categorization- The importance of necessary properties
-% 10.1.1.980.9935
-% __IMP__Measurenouns__cog-2015-0101
-% Entrenchment_Stefanowitsch and Flach for submission
-
-% Maybe:
-% [Cognitive Linguistics] The processing of verb-argument constructions is sensitive to form, function, frequency, contingency and prototypicality
-% [Cognitive Linguistics] Alternation-based generalizations are stored in the mental grammar- Evidence from a sorting task experiment //// along with CApelle's allostructions
-
-% Method:
-% Noble
-
-\subsection{The right statistics}
-\label{ssec:rightstatistics}
-
-In this section, I briefly motivate the choice of statistical models which I use later in Section~\ref{sec:corpusmodels}.
-In the spirit of the careful and methodologically sound approach to cognitively oriented corpus linguistics laid out in Section~\ref{sec:cogocl}, authors should always provide at least a minimal motivation as part of the standard protocol, especially if there are alternative methods.
-
-% NO to exploratory methods if we have hypotheses
-% Testing (hypotheses about signs of regressors)
-% Estimation => bootstrapping
-% Model selection
-%  => random vs fixed is not so simple, rather matter of getting model to converge
-%  => Maximality
-%  => NO to Gries 2015
-% Model evaluation
-% VS: Bayesian hype (Levshina)
-% VS: alternative math (R-W)
-
-% Cite:
-% __IMP__Measurenouns__cog-2015-0054 (Levshina)
-% [Cognitive Linguistics] Towards cognitively plausible data science in language research (NDL)
-
-% Stats:
-% BarrEa_RandomEffectsStructure_2014_JMemLang
-% ContentServer (1)
-% importatn_quotes
-% Menard_2000-Coefficients_of_Determination_for_Multiple_Logistic_Regression_Analysis
-
-%In the following, I examine the case alternation in German measure NPs under a cognitively oriented corpus linguistic perspective.
-%Since, however, the statistical tool kit (hierarchical generalised linear models, general tools such as $p$ values and confidence intervals, resampling techniques, model selection and model evaluation strategies, alternative algorithms for estimation, etc.) is heterogeneous and often criticised (especially under noisy conditions and when researcher run the risk of doing data mining instead of proper theory testing) I present a careful protocol for specifying, estimating, validating, and interpreting such models in Section \ref{sec:corpusmodels}.
-
-Before discussing the results in Section~\ref{sec:corpusmodels}, I show how I made sure that the relevant data is contained in the sample in Section \ref{sec:gettingdata}.
-
-\section{Getting the right data}
+\subsection{Getting the right data}
 \label{sec:gettingdata}
@@
 Therefore, it was necessary to obtain sample in which all (or at least most of the highly frequent) actually occurring combinations of kind noun and measure noun were represented.
-This section describes the three-stage bootstrap process that was used to obtain such a sample, which was then used to estimate the influences of various contextual factors on the alternation (cf.\ Section~\ref{sec:corpusmodels}).
+This section describes the three-stage bootstrap process that was used to obtain such a sample, which was then used to estimate the influences of various contextual factors on the alternation (cf.\ Section~\ref{sec:corpusstudies}).
 The bootstrap process takes into account the imperfection of automatically generated annotations in corpora.
@@
 
-%%%  TODO  %%%
-% – excluded "zwei Glas Wein" etc. because not part of alternation; NOT disambiguated for "ein Glas" (Vos 1999)
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\section{Corpus-based models}
-\label{sec:corpusmodels}
+\subsection{A prelude to the statistical analysis: There are no right statistics!}
+\label{sec:rightstatistics}
+
+In this section, I briefly motivate the choice of statistical models which I use later in Section~\ref{sec:corpusstudies}.
+In the spirit of the careful and methodologically sound approach to cognitively oriented corpus linguistics laid out in Section~\ref{sec:cogocl}, authors should always provide at least a minimal motivation as part of the standard protocol, especially if there are alternative methods.
+
+% NO to exploratory methods if we have hypotheses
+% Testing (hypotheses about signs of regressors)
+% Estimation => bootstrapping
+% Model selection
+%  => random vs fixed is not so simple, rather matter of getting model to converge
+%  => Maximality
+%  => NO to Gries 2015
+% Model evaluation
+% VS: Bayesian hype (Levshina)
+% VS: alternative math (R-W)
+
+% Cite:
+% __IMP__Measurenouns__cog-2015-0054 (Levshina)
+% [Cognitive Linguistics] Towards cognitively plausible data science in language research (NDL)
+
+% Stats:
+% BarrEa_RandomEffectsStructure_2014_JMemLang
+% ContentServer (1)
+% importatn_quotes
+% Menard_2000-Coefficients_of_Determination_for_Multiple_Logistic_Regression_Analysis
+
+%In the following, I examine the case alternation in German measure NPs under a cognitively oriented corpus linguistic perspective.
+%Since, however, the statistical tool kit (hierarchical generalised linear models, general tools such as $p$ values and confidence intervals, resampling techniques, model selection and model evaluation strategies, alternative algorithms for estimation, etc.) is heterogeneous and often criticised (especially under noisy conditions and when researcher run the risk of doing data mining instead of proper theory testing) I present a careful protocol for specifying, estimating, validating, and interpreting such models in Section \ref{sec:corpusstudies}.
+
+Before discussing the results in Section~\ref{sec:corpusstudies}, I show how I made sure that the relevant data is contained in the sample in Section \ref{sec:gettingdata}.
 
@@
 %
-%In Section~\ref{ssec:corpusnonhierarchicalmodel}, I briefly discuss why a non-hierarchical model -- while it is an option in principle -- has serious disadvantages for the data set under discussion (and similar data sets).
-%In Section~\ref{ssec:corpushierarchicalmodel}, I present the hierarchical model which is actually interpreted as the primary result of this study.
-%Section~\ref{ssec:bayesian} takes a short detour and shows (by example) that results from Bayesian estimation are not intrinsically different from those obtained from Maximum Likelihood methods, although Bayesian approaches have recently been touted to solve various problems facing applied statistics in various fields, even including cognitively oriented corpus linguistics.
-%Finally, I summarise and interpret the results of the corpus study in Section~\ref{ssec:modelssummary}.
+%In Section~\ref{sec:corpusnonhierarchicalmodel}, I briefly discuss why a non-hierarchical model -- while it is an option in principle -- has serious disadvantages for the data set under discussion (and similar data sets).
+%In Section~\ref{sec:corpushierarchicalmodel}, I present the hierarchical model which is actually interpreted as the primary result of this study.
+%Section~\ref{sec:bayesian} takes a short detour and shows (by example) that results from Bayesian estimation are not intrinsically different from those obtained from Maximum Likelihood methods, although Bayesian approaches have recently been touted to solve various problems facing applied statistics in various fields, even including cognitively oriented corpus linguistics.
+%Finally, I summarise and interpret the results of the corpus study in Section~\ref{sec:modelssummary}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 \subsection{Descriptive statistics and non-hierarchical models}
-\label{ssec:corpusnonhierarchicalmodel}
+\label{sec:corpusnonhierarchicalmodel}
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \subsection{Hierarchical models}
-\label{ssec:corpushierarchicalmodel}
+\label{sec:corpushierarchicalmodel}
 
 In this section, I apply standard Maximum Likelihood methods for estimating the parameters of a multilevel model.
-As suggested in Section~\ref{ssec:corpusnonhierarchicalmodel}, hierarchical models are better suited than non-hierarchical models given that we can expect lemma-specific effects, and that the respective factors have a high number of levels.
+As suggested in Section~\ref{sec:corpusnonhierarchicalmodel}, hierarchical models are better suited than non-hierarchical models given that we can expect lemma-specific effects, and that the respective factors have a high number of levels.
 
-\begin{equation}
-  \begin{split}
-    Pr(y_i=1)       & = logit^{-1}(\pmb{\beta}\mathbf{X_i} + \alpha^{\text{measurelemma}}_{j_i} + \alpha^{\text{kindlemma}}_{k_i}) \\
-    \text{where}\ \ & \mathbf{X_i}\ \text{is a vector of values for Badness, Genitives, } \\
-                    & \text{Matchlength, Measureabbreviated, Measurecase,} \\
-                    & \text{Measeurenumber, Minus1pos for}\ i \\
-    \text{and}\ \   & \pmb{\beta}\ \text{a corresponding coefficient vector} \\
-    \text{and}\ \   & \alpha^{\text{measurelemma}}_{j_i}\sim\mathcal{N}(\mu_j,\sigma_1) \\
-    \text{and}\ \   & \alpha^{\text{kindlemma}}_{k_i}\sim\mathcal{N}(\mu_k,\sigma_2) \\
-    \text{with}\ \  & \mu_1 \\
-  \end{split}
+Equations~\ref{eq:model:mnglmm} through \ref{eq:model:mnglmm:alphakind} specify the basic model fitted to the masculine\slash neuter and the feminine dataset.
+I use the same notation as used in \cite{GelmanHill2006} with the exception that I use vector notation for regressor variables and coefficients if the number of regressors is large in order to make the model specification more compact.%
+\footnote{
+$\beta_0$ is the general intercept, $B$ a vector of first-level coefficients, and $X_i$ a vector of regressor variables for sentence $i$.
+$\alpha^{\text{measure}}_{j[i]}$ is the random intercept for measure lemma $j$, chosen appropriately for the measure lemma occurring in sentence $i$ (similarly, $\alpha^{\text{kind}}_{k[i]}$ for kind lemma $k$).
+$\gamma_0^{\text{measure}}$ is the intercept of the second-level model for measure lemmas, $\Gamma^{\text{measure}}$ a vector of second-level coefficients for measure lemmas, $U_j$ a vector of regressor variables for measure lemma $j$, and $\sigma_{\text{measure}}^2$ the variance of the measure lemma intercepts.
+The second-level model for kind lemmas is specified in the same manner.
+}
+
+\begin{eqnarray}
     \label{eq:model:mnglmm}
-\end{equation}
+    Pr(y_i=1) &=& logit^{-1}(\beta_0 + B\cdot X_i + \alpha^{\text{measure}}_{j[i]} + \alpha^{\text{kind}}_{k[i]}) \\
+    \alpha^{\text{measure}}_j &\sim& \mathcal{N}(\gamma_0^{\text{measure}} + \Gamma^{\text{measure}}\cdot U_j,\sigma^2_{\text{measure}}) \\
+    \alpha^{\text{kind}}_k &\sim& \mathcal{N}(\gamma_0^{\text{kind}} + \Gamma^{\text{kind}}\cdot U_k,\sigma^2_{\text{kind}}) 
+  \label{eq:model:mnglmm:alphakind}
+\end{eqnarray}
 
@@
 
-\subsection{A note on Bayesian estimation}
-\label{ssec:bayesian}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Bayesian estimation}
+\label{sec:bayesian}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 \subsection{Summary and conclusions}
-\label{ssec:modelssummary}
+\label{sec:modelssummary}
 
@@
 
-\section{Experimental cross-check of the corpus-based models}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Experimental validation of the corpus-based models}
 \label{sec:externalvalidation}
+
+\subsection{Experiment 1: forced choice}
 
@@
 
-\subsection{Experiment 1: forced choice}
-
-As was pointed out in Section \ref{ssec:cocl}, there can be substantial doubt that corpus linguistic methods in themselves can be called \textit{cognitively oriented} in any meaningful way.
+As was pointed out in Section \ref{sec:cocl}, there can be substantial doubt that corpus linguistic methods in themselves can be called \textit{cognitively oriented} in any meaningful way.
 At the same time, it is a highly plausible assumption in a usage-based framework that large and varied corpora ideally represent language users' relevant input of (at least) written language, and that there should be reflexes of the distributions found in such corpora in the linguistic behaviour of language users.
@@
 For the F sentences, the top and bottom 25\% were used because the overall sample size was smaller.}
-It was made sure that the highly predictive features according to Section \ref{ssec:modelssummary} occurred in unique combinations among the four sentences in each cell.
+It was made sure that the highly predictive features according to Section \ref{sec:modelssummary} occurred in unique combinations among the four sentences in each cell.
 This is summarised in Table \ref{tab:experiment1:design}.
@@
 \includegraphics[width=0.6\textwidth]{figures/experiment/2afc_effects}
-\caption{Effect plot for the multilevel logistic regression in the forced-choice experiment: predictability of participants' choices from the logits of the corpus-based GLMM (cf.\ Section~\ref{ssec:corpushierarchicalmodel})}
+\caption{Effect plot for the multilevel logistic regression in the forced-choice experiment: predictability of participants' choices from the logits of the corpus-based GLMM (cf.\ Section~\ref{sec:corpushierarchicalmodel})}
 \label{fig:afc:effects}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \subsection{Experiment 2: self-paced reading}
@@
 %\includegraphics[width=0.6\textwidth]{figures/experiment/spr_effects_modelPredDichotomousReg}
-%\caption{Effect plot for the LMM in the self-paced reading experiment: modeling participants' residualised log reading times on the dichotomised predictions of the corpus-based GLMM (cf.\ Section~\ref{ssec:corpushierarchicalmodel})}
+%\caption{Effect plot for the LMM in the self-paced reading experiment: modeling participants' residualised log reading times on the dichotomised predictions of the corpus-based GLMM (cf.\ Section~\ref{sec:corpushierarchicalmodel})}
 %\label{fig:spr:dichotomous:effects}
@@
 \includegraphics[width=0.6\textwidth]{figures/experiment/spr_effects_case+modelPred+gender}
-\caption{Effect plot for the LMM in the self-paced reading experiment: modeling participants' residualised log reading times on the predictions (logits) of the corpus-based GLMM (cf.\ Section~\ref{ssec:corpushierarchicalmodel})}
+\caption{Effect plot for the LMM in the self-paced reading experiment: modeling participants' residualised log reading times on the predictions (logits) of the corpus-based GLMM (cf.\ Section~\ref{sec:corpushierarchicalmodel})}
 \label{fig:spr:continuous:effects}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{Conclusions}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \bibliography{rs,cow}

### 2017-03-18 Latex/article.tex
adding new data files for slightly revised models

@@
 
-Turning to (\ref{it:intro:plurl}), readers might have notices that so far only MNPs denoting quantities of substances (mass kind nouns such as \textit{ein Glas roter Wein\slash roten Weines} `a glass of red wine') have been discussed.
+Turning to (\ref{it:intro:plurl}), readers might have noticed that so far only MNPs denoting quantities of substances (mass kind nouns such as \textit{ein Glas roter Wein\slash roten Weines} `a glass of red wine') have been discussed.
 If the kind noun is a plural count noun with a collective denotation such as \textit{ein Sack kleine Äpfel\slash kleiner Äpfel} `a bag of small apples' (see Table~\ref{tab:syncretisms} for the inflection), a similar alternation between PGC and NAC can be observed.%

### 2017-03-18 R/00_wrapper.R
modified R files (Corpus), final

@@ -101,4 +101,6 @@ if (save.persistent) sink()
 # Save the state after all computations for debugging/ensuring reproducibility.
 save(list = ls(), file=paste(out.dir, "workspace_", end.time, ".RData", sep=""))
 
+# Add-on: Run prediction for experiment stimuli.
+source('11_stimuli_predict.R')
 

### 2017-03-18 R/01_load.R

@@ -198,6 +198,7 @@ for (fak in fax) {
 
 # Hm, different level ordering looks odd in plot, so make Minus1pos same order.
 mn$Minus1pos <- factor(mn$Minus1pos, levels = c("C", "D", "O", "A", "P"))
+mn$Measureclass <- factor(mn$Measureclass, levels = c("Square", "Length", "Currency", "Capacity", "Weight", "Rest", "Vessel", "Amount", "Layer", "Portion" ))
 fem$Minus1pos <- factor(fem$Minus1pos, levels = c("C", "D", "O", "A", "P"))
 
 

### 2017-03-18 R/02_attraction.R

@@ -13,13 +13,15 @@ prop.func <- function(v, smooth = 1) {
 ndn.kind <- read.delim("data/ndn_kind_freq.tsv", header=FALSE, quote="")
 nn.kind <- read.delim("data/nn_kind_freq.tsv", header=FALSE, quote="")
 
-# Pl has been pre-processed differently, can load in one table.
+# Pl & Measure have been pre-processed differently, can load in one table.
 countnouns_attracfreq <- read.delim("data/countnouns_attracfreq.tsv", quote="")
+measurenouns_attracfreq <- read.delim("data/measurenouns_attracfreq.tsv", quote="")
 
 # A cleanup.
 nn.kind <- nn.kind[-which(nn.kind$Lem=="rege"),]
 colnames(ndn.kind) <- c("Freq","Lem")
 colnames(nn.kind) <- c("Freq","Lem")
+
 
 # Get all lemmas for master table.
 # Background: We want values for all nouns, even those which occur in
@@ -27,7 +29,7 @@ colnames(nn.kind) <- c("Freq","Lem")
 cx.kind.lem <- c(as.character(nn.kind$Lem), as.character(ndn.kind$Lem))
 cx.kind.lem <- unique(cx.kind.lem[order(cx.kind.lem)])
 
-# Function to extract freqs for both constructions with "0 tolerance" (haha).
+# Function to extract freqs for both constructions with "add 1".
 get_freq <- function(data, lem) {
   n1 <- data[which(data$Lem == lem),1]
   if (length(n1) == 0) n1 <- 0
@@ -47,29 +49,39 @@ cx.kind <- as.data.frame(cx.kind)
 cx.kind$Pv <- unlist(apply(cx.kind[,2:3], 1, function(cs){attraction(cs)} ))
 cx.kind$Lpv <- log(cx.kind$Pv, 10)
 
-# PLURAL is simpler because the table was important done & ready.
+# PLURAL & MEASURE are simpler because the tables were imported done & ready.
+cx.measure <- measurenouns_attracfreq
+cx.measure$Pv <- unlist(apply(cx.measure[,2:3], 1, function(cs){attraction(cs)} ))
[…]

### 2017-03-18 R/03_explore.R

@@ -133,32 +133,7 @@ cat("\n\nRank correlation between regressors ordered by monofactorial GLM.\n\n")
 mnf.corr <- cor(order(rownames(mn.explore)[-1]), order(rownames(fem.explore)[-1]), method="sp")
 cat('\n\nRank correlation (Spearman) MN/F : ', mnf.corr)
 
-fem.explore.subs <- fem.explore[rownames(pl.explore),]
-fem.explore.subs <- fem.explore.subs[order(fem.explore.subs[,1], decreasing = T),]
-plf.corr <- cor(order(rownames(pl.explore)[-1]), order(rownames(fem.explore.subs)[-1]), method="sp")
-cat('\n\nRank correlation (Spearman) PL/F : ', plf.corr)
-
-mn.explore.subs <- mn.explore[rownames(pl.explore),]
-mn.explore.subs <- mn.explore.subs[order(mn.explore.subs[,1], decreasing = T),]
-plmn.corr <- cor(order(rownames(pl.explore)), order(rownames(mn.explore.subs)), method="sp")
-cat('\n\nRank correlation (Spearman) PL/F : ', plf.corr)
-
 if (save.persistent) sink()
-
-if (save.persistent) pdf(paste(out.dir, "regressor_monofact_corr.pdf", sep=""))
-
-plot(fem.explore[-1,1]~mn.explore[-1,1], pch=15, col="darkblue", xlim=c(0,0.45), ylim=c(0,0.45), cex=1.5,
-     xlab="R²", ylab="R²", main="Correlation: monofactorial pre-analysis of regressors")
-#abline(lm(fem.explore[-1,1]~mn.explore[-1,1]), col="darkblue", lty=1, lwd=2)
-points(pl.explore[-1,1]~mn.explore.subs[-1,1], pch=16, col="orange", cex=1.5)
-#abline(lm(pl.explore[-1,1]~mn.explore.subs[-1,1]), col="orange", lty=3, lwd=2)
-points(fem.explore.subs[-1,1]~pl.explore[-1,1], pch=17, col="darkgreen", type="p", cex=1.5)
-#abline(lm(fem.explore.subs[-1,1]~pl.explore[-1,1]), col="darkgreen", lty=2, lwd=2)
-#abline(0,1, lwd=2, lty=2)
-legend("bottomright", legend = c("Fem~Masc/Neut", "Pl~Masc/Neut", "Pl~Fem"), col=c("darkblue","orange","darkgreen"), pch=15:17)
[…]

### 2017-03-18 R/04_glm.R

@@ -12,8 +12,9 @@ mn.glm <- glm(Genitive~1
               #+Measureclass     # prob 0 or 1
               +Matchlength
               +Measureabbreviated
+              +Measureattraction
               +Measurelength
-              +Attraction
+              +Kindattraction
               +Measurenumber
               +Kindfinal
               +Kindedible
@@ -61,6 +62,7 @@ fem.glm <- glm(Casedrop~1
               +Measureabbreviated
               +Kindconsistency
               +Matchlength
+              #+Measureattraction# prob 0 or 1
               #+Kindfinal        # no hypothesis
               +Genitives
               +Kindedible
@@ -74,7 +76,7 @@ fem.glm <- glm(Casedrop~1
               +Measurefreq
               +Measurecase
               +Kindfreq
-              #+Attraction     # prob 0 or 1
+              #+Kindattraction # prob 0 or 1
               #+Kindlength     # unstable in bootstrap
               , data=fem,  family=binomial(link=logit))
 fem.glm.r2 <- NagelkerkeR2(fem.glm)$R2

### 2017-03-18 R/05_glmm.R

@@ -92,8 +92,9 @@ mn.glmm <- glmer(Genitive~1
                 +Measureclass
                 +Matchlength
                 +Measureabbreviated
+                +Measureattraction
                 #+Measurelength       # virtual non-convergence
-                +Attraction
+                +Kindattraction
                 +Measurenumber
                 +Kindfinal
                 +Kindedible
@@ -111,9 +112,9 @@ mn.glmm <- glmer(Genitive~1
 # Regressors to be used in getting LRT w/ bootstrapped CIs.
 
 mn.bootcomp.regs <- c("(1|Measurelemma)", "(1|Kindlemma)", "Genitives", "Minus1pos",
-                   "Measureclass", "Matchlength", "Measureabbreviated", "Attraction",
-                   "Measurenumber", "Kindfinal", "Kindedible", "Badness", "Kindfreq",
-                   "Measurecase", "Kindint", "Measurefreq")
+                   "Measureclass", "Matchlength", "Measureabbreviated", "Measureattraction",
+                   "Kindattraction", "Measurenumber", "Kindfinal", "Kindedible",
+                   "Badness", "Kindfreq", "Measurecase", "Kindint", "Measurefreq")
 
 # Get PB test.
 mn.modelcomp <- lmer.modelcomparison(model = mn.glmm, regressors = mn.bootcomp.regs,
@@ -171,6 +172,7 @@ fem.glmm <- glmer(Casedrop~1
                  +Measureabbreviated
                  +Kindconsistency
                  +Matchlength
+                 +Measureattraction
                  #+Kindfinal             # no hypothesis
                  +Genitives
                  +Kindedible
@@ -183,7 +185,7 @@ fem.glmm <- glmer(Casedrop~1
                  +Measurefreq
                  +Measurecase
                  +Kindfreq
-                 +Attraction
+                 +Kindattraction
                  #+Kindlength         # unstable in bootstrap
[…]

### 2017-03-18 R/06_glmm-mcmc.R

@@ -8,7 +8,7 @@ options(mc.cores = parallel::detectCores())
 
 chains <- 4
 seed   <- 6976
-iter   <- 1000
+iter   <- 1000 # Set to 1000 for production.
 
 # MN
 
@@ -21,8 +21,9 @@ mn.glmm.mcmc <- stan_glmer(Genitive~1
                            +Measureclass
                            +Matchlength
                            +Measureabbreviated
+                           +Measureattraction
                            #+Measurelength       # virtual non-convergence
-                           +Attraction
+                           +Kindattraction
                            +Measurenumber
                            +Kindfinal
                            +Kindedible
@@ -55,6 +56,7 @@ fem.glmm.mcmc <- stan_glmer(Casedrop~1
                             +Measureabbreviated
                             +Kindconsistency
                             +Matchlength
+                            +Measureattraction
                             #+Kindfinal             # no hypothesis
                             +Genitives
                             +Kindedible
@@ -67,7 +69,7 @@ fem.glmm.mcmc <- stan_glmer(Casedrop~1
                             +Measurefreq
                             +Measurecase
                             +Kindfreq
-                            +Attraction
+                            +Kindattraction
                             #+Kindlength         # unstable in bootstrap
                   ,
                   data=fem, family=binomial(link=logit),

### 2017-03-18 Latex/article.tex
swapping more graphics, final

@@
 \centering
-\includegraphics[width=0.9\textwidth]{figures/corpus/04_glmm_fixef.pdf}
-\caption{Fixed effects in GLMMs for Masculine\slash Neuter and Feminine with 95\% bootstrapped confidence intervals (1,000 replicates)}
+\includegraphics[width=0.9\textwidth]{figures/corpus/04_glmm_fixef_firstlevel}
+\caption{Estimated first-level effects in GLMMs for Masculine\slash Neuter and Feminine}
+\label{fig:glmmfixef}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.9\textwidth]{figures/corpus/04_glmm_fixef_secondlevel_kind}
+\caption{Estimated second-level effects for kind lemmas in GLMMs for Masculine\slash Neuter and Feminine}
+\label{fig:glmmfixef}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.9\textwidth]{figures/corpus/04_glmm_fixef_secondlevel_measure}
+\caption{Estimated second-level effects for measure lemmas in GLMMs for Masculine\slash Neuter and Feminine}
 \label{fig:glmmfixef}

### 2017-03-18 Latex/article.tex
theory part almost done

@@
 
-However, if the measure noun denotes a container, constructions with \textit{voll} (\textit{von})\slash\textit{voller} `full (of)' and \textit{mit} `with' are available, see (\ref{ex:alternatives}).
+However, if the measure noun denotes a container or vessel, constructions with \textit{voll} (\textit{von})\slash\textit{voller} `full (of)' and \textit{mit} `with' are available, see (\ref{ex:alternatives}).
 
@@
 This section reviews some existing analyses of the PGC-NAC alternation or related issues, and I develop my own analysis and the appropriate hypotheses for the studies in Sections~\ref{sec:corpusstudies} and \ref{sec:externalvalidation}.
-While the constructions and the alternation are often covered in descriptive grammars such as \citet[XYZ]{Eisenberg2013a}, the theoretical literature has little specific discussion to offer.
+While the constructions and the alternation are often covered in descriptive grammars such as \citet[XYZ]{Eisenberg2013a}, the available linguistic literature contains little specific discussion of the situation in German.
+In The remainder of this section, I discuss as factors that might influence the alternation:
+degrees of grammaticalisation of the measure noun, the different syntactic structures of the variants and their syntactic prototypes, and potential effects of semantic prototypicality.
 
-A caveat first:
-I do \textit{not} review the semantics literature on partitives and pseudo-partitives because the constructions of interest in the present paper are all pseudo-partitives, and thus none of the variance in the choice of the PGC and NAC can be expected to be explained by a discussion along these lines.
-The same is mostly true for the work on the grammaticalisation of pseudo-partitives and measure noun constructions, which has two main streams.
-On one hand, it is often assumed that pseudo-partitives arise as a form of grammaticalised partitives (\eg\ \citealp[536--539]{Koptjevskaja2001} for Finnish and Estonian, \citealp[559]{Koptjevskaja2001} for European languages per se).%
+Turning to grammaticalisation, it is often assumed that pseudo-partitives arise as a form of grammaticalised partitives (\eg\ \citealp[536--539]{Koptjevskaja2001} for Finnish and Estonian, \citealp[559]{Koptjevskaja2001} for European languages per se).%
 \footnote{Notice that the description of German in \citet[547--549]{Koptjevskaja2001} is vague and partly incorrect, maybe because her main source \cite{Eschenbach1993} is not as clear as it could be (see also Footnote \ref{fn:eschenbash} above).
-Her example (29b) from \citet[71]{Eschenbach1993}, repeated here verbatim as (\ref{ex:koptjewskaja}) is misclassified as a clear case of a dative on the measure noun and a genitive on the kind noun.
+Her example (29b) from \citet[71]{Eschenbach1993}, repeated here verbatim as (\ref{ex:koptjewskaja}) is misclassified as a clear case of a dative on the measure noun and a genitive on the kind noun (hence, PGC).
 
@@
 
-It could, however, equally well be a genitive on the measure noun because of the syncretisms discussed here on p.\ \pageref{page:femininesyncretism} and an ambiguous case requirement of the preposition \textit{troz}.
+It could, however, equally well be a genitive on the measure noun because of the syncretisms discussed here on p.\ \pageref{page:femininesyncretism} and an ambiguous case requirement of the preposition \textit{troz}, hence an instance of NAC.
 Then, she does not clearly distinguish acceptable from non-acceptable constructions (cf.\ Section \ref{sec:descriptive}) on her pp.\ 548 and 549, including a rather confusing \textit{???} rating of her (31b), which is clearly a non-acceptable construction of the type illustrated in (\ref{ex:analyticalpartitive:b}) above.
 }
-Since the alternating constructions are both clearly pseudo-partitives and not in any way partitives, this diachronic development (now completed) is not expected to have any effect on the alternation.
-On the other hand, there are more pertinent approaches like \cite{Brems2003,DeclerckBrems2016}.
-In English -- \cite{Brems2003} argues based on a view from \cite{Langacker1991} -- strongly lexicalised and thus semantically bleached measure nouns turn from nouns into quantifiers, leading to a reanalysis of the NP, wherein the kind noun becomes the head.
-This is supported by convincing syntactic evidence, for example subject-verb agreement.
-In German MNPs, the measure noun always remains the syntactic head, however (see immediately below for references), and I know of no strong hypotheses stating that the degree of grammaticalisation of the measure noun influences the alternation in any way.
-The statistical models presented in Section~\ref{sec:corpusstudies} could in principle show, however, whether highly grammaticalised measure nouns favor the NAC or the PGC through the estimation of lemma-specific random intercepts.
+The alternating constructions in German are both clearly pseudo-partitives, but the grammaticalisation paths assumed by \cite[esp.\ 526--530]{Koptjevskaja2001}, which I interpret freely here, are clearly pertinent.
+The development usually begins with constructions involving two referential nouns (not necessarily forming a single NP) and a \textit{separative} meaning as in \textit{(cut) two slices from the cake} \citep[535]{Koptjevskaja2001}.
+The \textit{part-of} meaning of true partitives as in \textit{a slice of the case} represents the first stage of a development wherein the measure noun already tends to loose some semantic content.
+The pseudo-partitive stage finally instantiates a \textit{quantity-of} relation, potentially even leading to fully grammaticalised quantifiers as in English \textit{a lot}.
+Since in German, the PGC is clearly the older variant \citep{Zimmer2015} with the overt \textit{quantity-of} marker in the form of the genitive suffix and at least the potential to form a true partitive (if the kind noun is definite), which the NAC lacks, we can expect more strongly grammaticalised measure nouns to prefer the NAC.
+Since clearly non-referential nouns like \textit{Gramm} `gram' and \textit{Meter} `metre' should favor the NAC according to this hypothesis, the effect can be detected by including information about measure noun classes in the corpus-based models in Section~\ref{sec:corpusstudies}.
 
-A major interest in traditional and generative linguistics is to find a plausible analysis of the hierarchical syntactic structure of the constructions in question.
+On the other hand, there are case studies like \cite{Brems2003,DeclerckBrems2016} where syntactic evidence for grammaticalisation plays a prominent role.
+In English -- \cite{Brems2003} argues based on a view from \cite{Langacker1991} -- strongly lexicalised and thus semantically bleached measure nouns turn from nouns into quantifiers, leading to a syntactic reanalysis of the NP, wherein the kind noun becomes the head.
+The argument is supported by convincing syntactic evidence, for example subject-verb agreement.
+In German MNPs, the measure noun always remains the syntactic head, however (see immediately below for references), and there is no similar clear criterion to detect grammaticalisation.
+The following analysis of the morpho-syntax of German MNPs will however, provide hypothetical cues along similar lines.
+
+A major interest in descriptive and generative linguistics is to find a plausible analysis of the hierarchical syntactic structure of the constructions in question.
 In the case at hand, the first question is which noun constitutes the head of the whole construction.
-This question was answered quite clearly by \cite{Loebel1986} already (see also \citealp[213]{Eschenbach1994} or \citealp[16]{GallmannLindauer1994}).
-Subject-verb agreement is always realised on the measure noun, and we can therefore assume that the measure noun is the head.
-Turning to the MNP-internal syntax, it cannot be understood properly without taking into account the strong\slash weak inflection patterns in German described in Section~\ref{sec:descriptive} (especially Footnote~\ref{fn:strongweak}).
-In NPs with a strongly inflected determiner, attributive adjectives inflect according to the so-called \textit{weak} pattern showing massive syncretism.
-If there is no determiner (as is the case in the alternating constructions), attributive adjectives inflect like determiners.
-Thus, the adjectives in the target constructions have properties of adjectives (they are lexical adjectives) and determiners (inflection and leftmost position in the NP) in a sense.
-In this light, the pattern shown in Table~\ref{tab:constructions} makes perfect sense, especially in a probabilistic framework:
-If speakers use classify the adjective as a determiner, they have to use the PGC because with a determiner, the PGC is the only option.
-If speakers classify the adjective as an adjective, the kind NP has no determiner, and the NAC is the only option.
-While the generative analysis presented in \cite{Bhatt1990} cannot deal with such probabilistic effects, she comes quite close to this interpretation by analysing the kind NP in the GPC as a DP and in the NAC as an NP.
+This question was answered quite clearly by \cite{Loebel1986} already (see also \citealp[213]{Eschenbach1994}, and \citealp[16]{GallmannLindauer1994}).
+As hinted above, subject-verb agreement is always realised on the measure noun, and we can therefore assume that the measure noun is the head.
 
-% Syntactic structure: adjectives-as-determiners and layers in NP, role assignment to nouns in PGC/NAC
-%     Gallmann => Bhatt1990 DP[gen] vs NP[agree]
-%     Mention problem of no case assigner for NAC, so sub-syntactic process or theta theory wrong
+I therefore turn to the MNP-internal syntax, which will show itself to be highly revealing.
+It cannot be understood properly without taking into account the strong\slash weak inflection patterns of adjectives described in Section~\ref{sec:descriptive} (especially Footnote~\ref{fn:strongweak}).
+In NPs with a strongly inflected determiner, attributive adjectives inflect according to the massively syncretistic \textit{weak} pattern.%
+\footnote{The structural cases (nominative and accusative) in the singular are marked with \textit{-e}, the oblique cases (dative and genitive) as well as the entire plural are marked with \textit{-en}.
+The only exception to this rule is the masculine accusative singular, which is also marked with \textit{-en}.}
+If there is no determiner (as is the case in the alternating constructions), attributive adjectives inflect like determiners, which is called the \textit{strong} pattern (see Table~\ref{tab:syncretisms}).
+Thus, the adjectives in the target constructions have properties of adjectives as well as determiners in a sense:
+they are lexical adjectives and function as attributive modifiers, but they are inflected as determiners and are the leftmost element in the NP.
+This unusual double nature of adjectives in NPs without a determiner leads to a plausible interpretation of the pattern shown in Table~\ref{tab:constructions}.
+Whenever speakers classify the adjective in the kind noun phrase as a determiner, they have to use the PGC because, if there is a determiner, the PGC is the only option.
+When they classify the adjective as an adjective, the kind NP has no determiner, and the NAC is the only option.%
+\footnote{While the generative analysis presented in \cite{Bhatt1990} cannot properly deal with probabilistic effects, Bhatt comes close to this interpretation by analysing the kind NP in the GPC as a DP and in the NAC as an NP.
+Within the confines of her framework, this is the appropriate was of encoding the relevant generalisation.}
+In a framework that assumes similarity-based probabilistic classification, such as Prototype Theory, we can now regard the PGC with a determiner in the kind noun phrase as instantiating the PGC prototype and the NAC with just a bare kind noun as the NAC prototype.
+The choice of the alternating variants is then expected to be influenced by the degree of similarity to either of these prototypes.
 
-Finally, I collect some hypotheses (or maybe mostly observations) from the existing literature as to which factors influence the choice of NAC and PGC.
+The situation is made difficult by the fact that the semantic differences between the variants are at best subtle, which means the prototypes are not associated with nice and clearly detectable differences in meaning.
+However, both the kind noun lemmas and the measure noun lemmas probably appear with different relative frequencies in the NAC with a bare kind noun and the PGC with a determiner inside the kind noun phrase (in part possibly due to their meaning and usage).
+Their relative frequencies in these non-alternating configurations should be indicative of their preference to appear in one of the two variants.
+In other words, lemmas themselves might have preferences for one prototype or the other.
 
-% Influencing factors: Hentschel, Zimmer
-% Squeeze in Anttila & Fong
+%Related to the , dative measure nouns favor the PGC more than nominative and accusative (also \citealp{Hentschel1993,Zimmer2015}).
+%As an example, \textit{mit einem Stück frischen Brots} `with a piece of fresh bread' (PGC) would be preferred over \textit{mit einem Stück frischem Brot} (NAC).
+%I propose
+
+Furthermore, the grammaticalisation path as described above leads from NPs denoting highly individuated objects standing in a \textit{separated-from} relation to a construction with a \textit{part-of} meaning (still available in the prototypical PGC) to a construction with a more diffuse \textit{quantity-of} relation (characteristic of the prototypical NAC).
+The best evidence for the relevance of this factor in an ongoing grammaticalisation process is that plural kind nouns almost always trigger PGC and virtually never NAC (Section~\ref{sec:descriptive}).
+More indirectly, plurality of the measure noun could favor the PGC over the NAC, and exact quantification in the form of cardinals even more so.%
+\footnote{The effect of the plural was empirically shown but not explained in \cite{Hentschel1993}.}
+As a highly illustrative example (using a measure noun which has virtually developed two variants, one of which grammaticalsed to a much higher degree), \textit{ein Haufen französische Watte} `a heap\slash lot of French cotton' (NAC) has essentially two uses.
+It can be used to denote an unspecific amount of cotton in the sense of \textit{a lot of French cotton} in situations when the exact quantity of cotton is not of interest.
+Under the appropriate circumstances, it could also be used to denote a highly specific quantity of cotton, for example when a cotton merchant always sells her French cotton in \textit{heaps} weighing two stone each.
+The former is highly restricted for \textit{zwei Haufen französischer Watte} `two heaps/\textdagger lots of French cotton' (GPC).
+Such an NP is less likely to be used in situations where the exact quantity of cotton is irrelevant, even if there are actually two heaps of cotton.
+In other words, a precisely quantified measure noun leads to a higher individuation (the creation of portions) of the substance denoted by the kind noun, which is more appropriate for the PGC prototype.%
+\footnote{In the most extreme, some measure nouns cannot be pluralised, such as \textit{eine Menge} `a lot of'.
+Also note that the presentation of the examples in NAC and PGC is suggestive.
+Native speakers will confirm that both NPs sound highly acceptable in both GPC and NAC.
+It was made clear from the beginning and throughout that the preferences are indeed subtle.}
+
+Finally, it is claimed that the PGC is more typical of higher registers or even exclusive to written language (see \citealp[320--323]{Hentschel1993}, where she also refers to descriptive grammars).
+This is not surprising in as much as the genitive -- an intrinsic part of the PGC -- is generally underrepresented in colloquial and vernacular variants of German, and thus overrepresented to the same degree as colloquial, vernacular, and spoken variants are overrepresented in most corpora of written German.
+There is no reason to assume that such preferences for cannot be part of the construction prototypes.
+In the corpus study, the alternation tendencies controlled by this complex factor related to register, medium, and lectal variation will be tested using certain very simple proxies.
+
+To summarise the discussion and core hypotheses, it is expected that:
+
+\begin{enumerate}[i.]
+  \item classes of more clearly grammaticalised measure nouns favor the NAC,
+  \item the relative frequencies with which measure noun lemmas and kind noun lemmas appear in the prototypical (unequivocal) PGC and NAC are positively correlated with the relative frequency the PGC and the NAC are chosen in the alternating case,
+  \item MNPs with masculine dative kind nouns disfavour the GPC comparatively less
+  \item pluralic measure nouns and measure nouns quantified by cardinals favor the PGC variant
+  \item .
+\end{enumerate}
+
+In the next section, I report the corpus study that was designed to test these hypotheses on usage data.
 
@@
 
-% On Case-OCP (from Götz Keydana):
-% Und da habe ich dann auch einen Verweis auf den locus classicus gefunden:
-% STEFAN A. FRISCH, JANET B. PIERREHUMBERT and MICHAEL B. BROE: SIMILARITY AVOIDANCE AND THE OCP. Natural Language & Linguistic Theory 22: 179–228, 2004.
-% Dazu gäbe es dann noch, ebenfalls von Pierrehumbert,
-% Dissimilarity in the Arabic Verbal Roots. In Amy Schafer, ed.,Proceedings of NELS 23 Amherst, MA: Graduate Linguistics Student Association. 1993. pp 367-381 
-% Darin stellt sie wohl erstmals explizit einen Zusammenahng zwischen OCP und Kognition her.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.5\textwidth]{figures/corpus/04_glmm_raneff_mn}~\hspace{0.05\textwidth}~\includegraphics[width=0.5\textwidth]{figures/corpus/04_glmm_raneff_fem}
-\caption{Sample random effects for the masculine\slash neuter and feminine GLMMs}
-\label{fig:glmm:fixef:minus1pos}
-\end{figure}
+%\begin{figure}[h]
+%\centering
+%\includegraphics[width=0.5\textwidth]{figures/corpus/04_glmm_raneff_mn}~\hspace{0.05\textwidth}~\includegraphics[width=0.5\textwidth]{figures/corpus/04_glmm_raneff_fem}
+%\caption{Sample random effects for the masculine\slash neuter and feminine GLMMs}
+%\label{fig:glmm:fixef:minus1pos}
+%\end{figure}
 
@@
 \end{document}
+
+%\footnote{\cite{AnttilaFong2000} propose an \textit{Obligatory Contour Principle for case} (Case-OCP) in a superficially similar but fundamentally unrelated case alternation in Finnish partitives.
+%}
+%
+% On Case-OCP (from Götz Keydana):
+% Und da habe ich dann auch einen Verweis auf den locus classicus gefunden:
+% STEFAN A. FRISCH, JANET B. PIERREHUMBERT and MICHAEL B. BROE: SIMILARITY AVOIDANCE AND THE OCP. Natural Language & Linguistic Theory 22: 179–228, 2004.
+% Dazu gäbe es dann noch, ebenfalls von Pierrehumbert,
+% Dissimilarity in the Arabic Verbal Roots. In Amy Schafer, ed.,Proceedings of NELS 23 Amherst, MA: Graduate Linguistics Student Association. 1993. pp 367-381 
+% Darin stellt sie wohl erstmals explizit einen Zusammenahng zwischen OCP und Kognition her.

### 2017-03-19 R/04_glm.R
final-final-final version of R stuff

@@ -16,12 +16,10 @@ mn.glm <- glm(Genitive~1
               +Measurelength
               +Kindattraction
               +Measurenumber
-              +Kindfinal
               +Kindedible
               +Badness
               +Kindconsistency
               +Kindorigin
-              #+Measuregender   # convergence
               +Kindfreq
               #+Minus2pos       # prob 0 or 1 
               +Measurecase
@@ -63,12 +61,10 @@ fem.glm <- glm(Casedrop~1
               +Kindconsistency
               +Matchlength
               #+Measureattraction# prob 0 or 1
-              #+Kindfinal        # no hypothesis
               +Genitives
               +Kindedible
               +Kindorigin
               +Measurelength
-              +Measuregender
               +Measurenumber
               +Kindint
               +Badness
@@ -108,7 +104,6 @@ pl.glm <- glm(Casedrop~1
               #+Minus1pos     # prob 0 or 1
               #+Attraction    # prob 0 or 1
               +Genitives
-              #+Kindfinal     # no hypothesis
               #+Kindgender    # prob 0 or 1
               #+Matchlength   # prob 0 or 1
               #+Measurefreq   # prob 0 or 1
@@ -118,7 +113,6 @@ pl.glm <- glm(Casedrop~1
               +Badness
               #+Minus2pos     # prob 0 or 1
               +Kindlength
-              +Measuregender
               #+Measurenumber # 1 level
               #+Measureabbreviated # 1 level
               , data=pl,  family=binomial(link=logit)

### 2017-03-19 R/06_glmm-mcmc.R

@@ -25,7 +25,6 @@ mn.glmm.mcmc <- stan_glmer(Genitive~1
                            #+Measurelength       # virtual non-convergence
                            +Kindattraction
                            +Measurenumber
-                           +Kindfinal
                            +Kindedible
                            +Badness
                            #+Kindconsistency     # virtual non-convergence
@@ -57,12 +56,10 @@ fem.glmm.mcmc <- stan_glmer(Casedrop~1
                             +Kindconsistency
                             +Matchlength
                             +Measureattraction
-                            #+Kindfinal             # no hypothesis
                             +Genitives
                             +Kindedible
                             #+Kindorigin           # fixeff matrix rank deficient
                             #+Measurelength        # unstable in bootstrap
-                            +Measuregender
                             +Measurenumber
                             +Kindint
                             +Badness
@@ -91,7 +88,6 @@ pl.glmm.mcmc <- stan_glmer(Casedrop~1
                            #+Minus1pos           # model unidentifiable
                            +Attraction
                            #+Genitives           # model unidentifiable
-                           #+Kindfinal           # model unidentifiable
                            #+Kindgender          # Hessian singular
                            #+Matchlength         # model unidentifiable
                            +Measurefreq
@@ -101,7 +97,6 @@ pl.glmm.mcmc <- stan_glmer(Casedrop~1
                            #+Badness             # model unidentifiable
                            #+Minus2pos           # model unidentifiable
                            #+Kindlength          # non-convergence
-                           +Measuregender
                            #+Measurenumber       # makes no sense
                            #+Measureabbreviated  # only one level
                  ,

### 2017-03-19 Latex/article.tex
revisions of theory, more on corpus study

@@
 
-\section{Cognitively oriented corpus linguistics: theories and case studies}
+\section{Cognitively oriented corpus linguistics: theories, case studies, and methods}
 \label{sec:cogocl}
+
+\subsection{Corpora and cognition}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\section{Case assignment in German measure NPs}
-\label{sec:germanmeasurenps}
-
-\subsection{Two stable cases and a case alternation}
-\label{sec:descriptive}
-
-The alternation phenomenon to be discussed in this paper appears to be a straightforward case alternation at first sight.
-However, without a solid knowledge of German nominal morpho-syntax, some of the more subtle distinctions to be made can be confusing.
-Therefore, I introduce and illustrate the relevant constructions in some detail in this section, followed by a review of some relevant literature in Section~\ref{sec:analyses}.
-I define the narrow range of syntactic constellations in which the alternation occurs and thus motivate the focus on \textit{only} this narrow range rather than the whole range of nominal constructions expressing quantities.
-This entails that I also have to discuss certain constructions and factors that might be seen as relevant, but which I exclude from my own study.
-
-I use the term \textit{measure noun phrase} (MNP) to refer a noun phrase (NP) in which a kind-denoting (count or mass) noun depends on another noun specifying a quantity of the objects or the substance denoted by the kind noun.
-I call the kind-denoting noun simply the \textit{kind noun} and the quantity-denoting noun the \textit{measure noun}.%
-\footnote{It would be more technically appropriate -- but also more cumbersome -- to speak of the \textit{kind noun phrase} because the embedded noun can be modified and form a full noun phrase, cf.\ immediately below and Section~\ref{sec:analyses} on the phrasal status of the constituents.}
-Measure nouns can be all sorts of nouns which genuinely denote a quantity (such as \textit{litre} or \textit{amount}) but also nouns denoting containers, collections, etc.
-(such as \textit{glass} or \textit{bucket}).
-In a similar vein, \citet[284]{Brems2003} considers nouns as measure nouns \textit{which, strictly speaking, do not designate a `measure', but display a more nebulous} (\textit{sic!}) \textit{potential for quantification} (see also \citealp[530]{Koptjevskaja2001}, and \citealp[338]{Rutkowski2007}).
-For illustration purposes, in the English \textit{a cup of fine coffee}, \textit{cup} is the measure noun, and \textit{coffee} is the kind noun.
-
-Descriptively, three different syntactic configurations need to be distinguished \wrt case assignment inside German measure noun phrases.
-If the kind noun comes with a determiner, the construction resembles (and is usually called) a \textit{pseudo-partitive} (on partitives and pseudo-partitives see, e.\,g., \citealp{Barker1998,Selkirk1977,Stickney2007,Vos1999}; for a recent application of the terminology to German, see \citealp{Gerstenberger2015}).%
-\footnote{Whereas partitives are constructions denoting a proper part-of relation strictly requiring a definite kind noun as in \textit{a sip of the wine}, pseudo partitives -- albeit syntactically similar and diachronically related to partitives in many languages -- merely denote quantities and contain indefinite kind nouns as in \textit{a sip of wine}.
-In the literature on German, some authors incorrectly call the pseudo-partitive a \textit{partitive} \cite{Hentschel1993} while some realise the difference and at least mention it \cite{Eschenbach1994,GallmannLindauer1994,Loebel1989,Zimmer2015}.}
-I use \textdagger\ to denote general unacceptability, and the acceptability judgements given in (\ref{ex:intro:pseudopartitive}) are completely undisputed.
-Notice that the kind noun is obligatory in the genitive and that case agreement with the measure noun is not an option (but see immediately below).
-I refer to the construction in (\ref{ex:intro:pseudopartitive1}) as the \textit{Pseudo-partitive Genitive Construction} (PGC).
-
-\begin{exe}
-  \ex\label{ex:intro:pseudopartitive}
-  \begin{xlist}
-    \ex[ ]{\label{ex:intro:pseudopartitive1}\gll Wir trinken [[eine Tasse]\Sub{Acc} [eines leckeren Kaffees]\Sub{Gen}]\Sub{Acc}.\\
-    we drink a cup a tasty coffee\\
-    \glt We drink a cup of a tasty coffee.}
-    \ex[\textdagger]{\label{ex:intro:pseudopartitive2} Wir trinken [[eine Tasse]\Sub{Acc} [einen leckeren Kaffee]\Sub{Acc}]\Sub{Acc}.}
-  \end{xlist}
-\end{exe}
-
-When the kind noun is bare -- \ie\ when it comes without a determiner and without modifying adjectives -- the kind noun has to agree in case with the measure noun, and the genitive seen in the PGC is not acceptable, see (\ref{ex:intro:narrowapposition}).
-This construction is usually classified as a \textit{Narrow Apposition Construction} \citep{Loebel1986}, henceforth NAC.%
-\footnote{Notice that the unavailability of the genitive on the kind noun can be seen as following from a quirky constraint that genitive NPs in German require the presence of some strongly case-marked element (determiner or adjective) in addition to the head noun in order to be acceptable \citep{GallmannLindauer1994,Schachtl1989}.}
-
-\begin{exe}
-  \ex\label{ex:intro:narrowapposition}
-  \begin{xlist}
-    \ex[\textdagger]{\label{ex:intro:narrowapposition1} Wir trinken [[eine Tasse]\Sub{Acc} [Kaffees]\Sub{Gen}]\Sub{Acc}.}
-    \ex[ ]{\label{ex:intro:narrowapposition2}\gll Wir trinken [[eine Tasse]\Sub{Acc} [Kaffee]\Sub{Acc}]\Sub{Acc}.\\
-    we drink a cup coffee\\
-    \glt We drink a cup of coffee.}
-  \end{xlist}
-\end{exe}
-
-The construction as in (\ref{ex:intro:narrowapposition2}) is also referred to as the \textit{Direct Partitive Construction} for other Germanic languages where the PGC with the synthetic genitive is not available.%
-\footnote{This nomenclature makes sense in contrast to the \textit{Indirect Partitive Construction} with prepositional linkers translating to `of' -- \ie\ analytic genitives -- in such languages, see \cite{HankamerMikkelsen2008} for Danish.
-For German, this terminology is not distinctive enough, which is why I use the terms NAC and PGC.}
-While only the accusative is shown in (\ref{ex:intro:narrowapposition}), the kind noun obligatory agrees in case also with nominative and dative measure nouns.
-
-The actual alternation can be observed \textit{only when the kind noun occurs with an attributive adjective but without a determiner}, as in (\ref{ex:intro:alternation}), where both the NAC in (\ref{ex:intro:alternation1}) and the PGC in (\ref{ex:intro:alternation2}) are equally acceptable.
-They are also functionally and semantically equivalent.%
-\footnote{Some descriptive and normative grammars take stronger positions w.\,r.\,t.\ the acceptability of these two options.
-See \cite{Hentschel1993,Zimmer2015} for analyses of the sometimes absurd stances taken in grammars of German.
-As the usage and experimental data presented below should render it clear, there might be preferences under certain circumstances, but we cannot assume either construction to be unacceptable.}
-
-\begin{exe}
-  \ex\label{ex:intro:alternation}
-  \begin{xlist}
-    \ex[ ]{\label{ex:intro:alternation2}\gll Wir trinken [[eine Tasse]\Sub{Acc} [heißen Kaffee]\Sub{Acc}]\Sub{Acc}.\\
-    we drink a cup hot coffee\\
-    \glt We drink a cup of hot coffee.}
-    \ex[ ]{\label{ex:intro:alternation1} Wir trinken [[eine Tasse]\Sub{Acc} [heißen Kaffees]\Sub{Gen}]\Sub{Acc}.}
-  \end{xlist}
-\end{exe}
-
-\begin{table}
-  \centering
-  \begin{tabular}{lccc}
-    \multicolumn{1}{r}{kind NP:} & bare noun & with adjective & with determiner \\
-    & [\ldots N\Subsf{meas} [N\Subsf{kind}]] & [\ldots N\Subsf{meas} [AP N\Subsf{kind}]] & [\ldots N\Subsf{meas} [D N\Subsf{kind}]] \\
-    \midrule
-    narrow apposition         & \CheckIt & \CheckIt &          \\
-    pseudo-partitive genitive &          & \CheckIt & \CheckIt \\
-  \end{tabular}
-  \caption{Distribution of the NAC and PGC constructions in different NP structures}
-  \label{tab:constructions}
-\end{table}
-
-The distribution of the case patterns in the NAC (case identity between the measure noun and the kind noun) and in the PGC (genitive on the kind noun) depending on the structure of the kind NP are summarised in Table~\ref{tab:constructions}.
-I now turn to some more subtle issue related to the measure noun case alternation, namely:
-
-\begin{enumerate}[i.]
-  \item{\label{it:intro:idiot} alleged alternatives to the NAC and the PGC}
-  \item{\label{it:intro:datsg} alternative \textit{weak} forms of dative singular neuter adjectives}
-  \item{\label{it:intro:femsg} case syncretism in feminine NPs}
-  \item{\label{it:intro:plurl} similar plural/collective construction(s)}
-  \item{\label{it:intro:noifl} grammaticalised non-inflected measure nouns}
-  \item{\label{it:intro:other} alternative constructions for expressing quantities}
-\end{enumerate}
-
-\vspace{-1\baselineskip}
-
-First of all, (\ref{it:intro:idiot}) refers to claims found in some grammars that a generic nominative, accusative, and even dative on the kind noun are used instead of the genitive or case agreement (overviews in \citealp{Hentschel1993,Zimmer2015}).
-It was shown empirically in \cite{Hentschel1993} that such variants are de facto not acceptable.
-Also, in my corpus sample, they simply did not occur.
-Even if they are accepted by some speakers, their extremely low frequency makes it impossible to study them using either corpus methods (Section~\ref{sec:corpusstudies}) or experimental approaches (Section~\ref{sec:externalvalidation}), and I consequently ignore them.
-
-As for (\ref{it:intro:datsg}), a complication with neuter kind nouns in the dative is mentioned by \citet[20--22]{Zimmer2015}.
-He reports a high number of occurrences of adjectives being inflected according to the \textit{weak} inflectional pattern, which is normally only used if a strongly inflected determiner precedes the adjective: \textit{kalt-en Wasser} `cold water' instead of \textit{kalt-em Wasser}.%
-\footnote{\label{fn:strongweak}The \textit{strong} case and number suffixes of the adjectives are identical (with one exception) to the suffixes of determiners, and they are used when no strongly inflected determiner precedes the adjective.}
-In my corpus sample (released with this article), this tendency was not nearly as clear, and there was a high number of very noisy sentences among those potentially showing this pattern.
-Hence, I do not discuss these forms much, although I will offer a possible interpretation in Section~\ref{sec:analyses}.
-
-\label{page:femininesyncretism}
-With feminine kind nouns, (\ref{it:intro:femsg}) is relevant.
-Whereas the singular masculine and neuter nominal subsystem is a four case system (at least for NPs with adjectives and determiners), feminine singular NPs show nominative-accusative and dative-genitive syncretisms.
-Table~\ref{tab:syncretisms} shows this and the partial syncretism in the plural, where no gender distinctions are made (for an overview, see \citealp[XYZ]{Eisenberg2013a}).
-This means that with dative measure nouns, the alternation is unobservable if the kind noun is feminine, and it is problematic to speak of a \textit{genitive} kind noun construction in this case.
-This is taken into account in Section~\ref{sec:corpusstudies} in the model design and the discussion.
-
-\begin{table}
-  \centering
-  \begin{tabular}{llll}
-     & Masculine\slash Neuter & Feminine & Plural \\
-     \midrule
-     nominative & rot-er Wein    & \multirow{2}{*}{frisch-e Sahne}   & \multirow{2}{*}{klein-e Äpfel} \\
-     accusative & rot-en Wein    &                                   &                                \\
-     dative     & rot-em Wein    & \multirow{2}{*}{frisch-er Sahne}  & klein-en Äpfel-n               \\
-     genitive   & rot-en Wein-es &                                   & klein-er Äpfel                 \\
-  \end{tabular}
-  \caption{Case syncretisms in NPs with an adjective and without a determiner (\textit{roter Wein} `red wine', \textit{frische Sahne} `fresh cream', \textit{kleine Äpfel} `small apples')}
-  \label{tab:syncretisms}
-\end{table}
-
-Turning to (\ref{it:intro:plurl}), readers might have noticed that so far only MNPs denoting quantities of substances (mass kind nouns such as \textit{ein Glas roter Wein\slash roten Weines} `a glass of red wine') have been discussed.
-If the kind noun is a plural count noun with a collective denotation such as \textit{ein Sack kleine Äpfel\slash kleiner Äpfel} `a bag of small apples' (see Table~\ref{tab:syncretisms} for the inflection), a similar alternation between PGC and NAC can be observed.%
-\footnote{\label{fn:eschenbash}On the collective interpretation of such constructions, see \cite{Eschenbach1994}.
-It is, however, unfortunate that Eschenbach does not notice and discuss the morpho-syntactic distribution of PGC and NAC, even when she gets very close to listing the relevant data \citep[217]{Eschenbach1994}.}
-In line with experimental results reported in \citet[15--16]{Zimmer2015}, I found that the PGC is so dominant with plural kind nouns (over 92\%, cf.\ Section~\ref{sec:corpusstudies}) that the alternation cannot be analysed in the same way as in the singular case.
-While this will play a role in the interpretation of the corpus findings, such plural\slash collective MNPs will not be dealt with prominently in the remainder of the paper.
-A statistical model for plural kind nouns will be reported in passing in Section~\ref{sec:corpusstudies}, however.
-
-Turning to issue (\ref{it:intro:noifl}) now, some measure nouns have been grammaticalised in a way that they always appear non-inflected.
-They are typical measure nouns like \textit{Gramm} `gram', \textit{Pfund} `pound' or \textit{Prozent} `percent', which do not have plural forms at all.%
-\footnote{With neglectable frequency, this effect is extended to container nouns like \textit{Glas} `glass', but with a twist.
-A non-inflected variant \textit{zwei Glas roter Wein} `two glasses of wine' and an inflected variant \textit{zwei Gläser roter Wein} co-exist.
-The non-inflected variant denotes the quantity of wine fitting into two glasses, the inflected variant denotes two glasses filled with wine.}
-I treat these cases like any other measure noun because they enter into both the NAC as in (\ref{ex:uninflectedmeasures:a}) and the PGC as in (\ref{ex:uninflectedmeasures:b}).
-See also Section~\ref{sec:analyses}.
-
-\begin{exe}
-  \ex\label{ex:uninflectedmeasures}
-  \begin{xlist}
-    \ex{\label{ex:uninflectedmeasures:a} zwei Gramm brauner Zucker}
-    \ex{\label{ex:uninflectedmeasures:b}\gll zwei Gramm braunen Zuckers\\
-                                         two gram brown sugar\\
-				         two grams of brown sugar}
-  \end{xlist}
-\end{exe}
-
-Finally, (\ref{it:intro:other}) suggests that there are alternative ways of expressing quantification.
-First of all, the analytic (pseudo-)partitive with \textit{von} `of' is only available as an alternative to the PGC if the kind noun phrase contains a (definite or indefinite) determiner as in (\ref{ex:analyticalpartitive}).
-
-\begin{exe}
-  \ex\label{ex:analyticalpartitive} 
-  \begin{xlist}
-    \ex[ ]{\label{ex:analyticalpartitive:a}\gll ein Glas von dem roten Wein\\
-                                                a glass of the red wine\\
-                                                \glt a glass of the red wine
-       }
-    \ex[\textdagger]{\label{ex:analyticalpartitive:b}\gll ein Glas von rotem Wein\\
-                                                a glass of red wine\\
-                                                \glt a glass of red wine
-       }
-  \end{xlist}
-\end{exe}
-
-However, if the measure noun denotes a container or vessel, constructions with \textit{voll} (\textit{von})\slash\textit{voller} `full (of)' and \textit{mit} `with' are available, see (\ref{ex:alternatives}).
-
-\begin{exe}
-  \ex\label{ex:alternatives}
-  \begin{xlist}
-  \ex{\label{ex:voll}\gll   ein Glas voll von rotem Wein\\
-                            a glass full of red wine\\
-			    \glt a glass full of red wine}
-  \ex{\label{ex:voller}\gll ein Glas {voll\slash voller} {rotem\slash roter} Wein\\
-                            a glass full-of red wine\\
-			    \glt a glass full of red wine}
-  \ex{\label{ex:mit}\gll    ein Glas mit rotem Wein\\
-                            a glass with red wine\\
-			    \glt a glass with red wine in it}
-  \end{xlist}
-\end{exe}
-
-These construction have very specific properties (the construction with \textit{voller} are discussed in \citealp{Zeldes2018}, the construction with \textit{mit} is discussed in \citealp{Bhatt1990}), they are not semantically equivalent to NAC and PGC, and -- most importantly -- they are only available for a small subset of measure nouns.
-Subtle interactions between these constructions and NAC\slash PGC cannot be excluded, especially because the former might act as a (partial) alternative and thus reduce the count of exemplars of the NAC, the PGC, or both for container measure nouns.
-There are no clear and motivated hypotheses about such interactions, however.
-Nevertheless, the corpus-based models in Section~\ref{sec:corpusstudies} are specified in a way that they could detect preferences for either NAC or PGC for those nouns for which the alternative constructions illustrated in (\ref{ex:alternatives}) is available.
-
-This concludes the descriptive overview of the alternation.
-I have shown that there is an alternation between two very specific measure noun constructions under quite specific syntactic conditions, and that the two constructions can be distinguished clearly by looking at the case of the kind noun.
-It should be kept in mind that this paper is about this specific alternation and not about a possibly broader range of nominal constructions expressing quantification or measure.
-I turn to some more theory-oriented discussion in the next section.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{A theoretical assessment}
-\label{sec:analyses}
-
-This section reviews some existing analyses of the PGC-NAC alternation or related issues, and I develop my own analysis and the appropriate hypotheses for the studies in Sections~\ref{sec:corpusstudies} and \ref{sec:externalvalidation}.
-While the constructions and the alternation are often covered in descriptive grammars such as \citet[XYZ]{Eisenberg2013a}, the available linguistic literature contains little specific discussion of the situation in German.
-In The remainder of this section, I discuss as factors that might influence the alternation:
-degrees of grammaticalisation of the measure noun, the different syntactic structures of the variants and their syntactic prototypes, and potential effects of semantic prototypicality.
-
-Turning to grammaticalisation, it is often assumed that pseudo-partitives arise as a form of grammaticalised partitives (\eg\ \citealp[536--539]{Koptjevskaja2001} for Finnish and Estonian, \citealp[559]{Koptjevskaja2001} for European languages per se).%
-\footnote{Notice that the description of German in \citet[547--549]{Koptjevskaja2001} is vague and partly incorrect, maybe because her main source \cite{Eschenbach1993} is not as clear as it could be (see also Footnote \ref{fn:eschenbash} above).
-Her example (29b) from \citet[71]{Eschenbach1993}, repeated here verbatim as (\ref{ex:koptjewskaja}) is misclassified as a clear case of a dative on the measure noun and a genitive on the kind noun (hence, PGC).
-
-\begin{exe}
-  \ex{\label{ex:koptjewskaja}\gll trotz einer Flasche gut-en Wein-s\\
-  in.spite.of one-\textsc{fem.gen} bottle good-\textsc{masc.sg.gen} wine-\textsc{gen}\\
-  \glt in spite of one bottle of good wine}
-\end{exe}
-
-It could, however, equally well be a genitive on the measure noun because of the syncretisms discussed here on p.\ \pageref{page:femininesyncretism} and an ambiguous case requirement of the preposition \textit{troz}, hence an instance of NAC.
-Then, she does not clearly distinguish acceptable from non-acceptable constructions (cf.\ Section \ref{sec:descriptive}) on her pp.\ 548 and 549, including a rather confusing \textit{???} rating of her (31b), which is clearly a non-acceptable construction of the type illustrated in (\ref{ex:analyticalpartitive:b}) above.
-}
-The alternating constructions in German are both clearly pseudo-partitives, but the grammaticalisation paths assumed by \cite[esp.\ 526--530]{Koptjevskaja2001}, which I interpret freely here, are clearly pertinent.
-The development usually begins with constructions involving two referential nouns (not necessarily forming a single NP) and a \textit{separative} meaning as in \textit{(cut) two slices from the cake} \citep[535]{Koptjevskaja2001}.
-The \textit{part-of} meaning of true partitives as in \textit{a slice of the case} represents the first stage of a development wherein the measure noun already tends to loose some semantic content.
-The pseudo-partitive stage finally instantiates a \textit{quantity-of} relation, potentially even leading to fully grammaticalised quantifiers as in English \textit{a lot}.
-Since in German, the PGC is clearly the older variant \citep{Zimmer2015} with the overt \textit{quantity-of} marker in the form of the genitive suffix and at least the potential to form a true partitive (if the kind noun is definite), which the NAC lacks, we can expect more strongly grammaticalised measure nouns to prefer the NAC.
-Since clearly non-referential nouns like \textit{Gramm} `gram' and \textit{Meter} `metre' should favor the NAC according to this hypothesis, the effect can be detected by including information about measure noun classes in the corpus-based models in Section~\ref{sec:corpusstudies}.
-
-On the other hand, there are case studies like \cite{Brems2003,DeclerckBrems2016} where syntactic evidence for grammaticalisation plays a prominent role.
-In English -- \cite{Brems2003} argues based on a view from \cite{Langacker1991} -- strongly lexicalised and thus semantically bleached measure nouns turn from nouns into quantifiers, leading to a syntactic reanalysis of the NP, wherein the kind noun becomes the head.
-The argument is supported by convincing syntactic evidence, for example subject-verb agreement.
-In German MNPs, the measure noun always remains the syntactic head, however (see immediately below for references), and there is no similar clear criterion to detect grammaticalisation.
-The following analysis of the morpho-syntax of German MNPs will however, provide hypothetical cues along similar lines.
-
-A major interest in descriptive and generative linguistics is to find a plausible analysis of the hierarchical syntactic structure of the constructions in question.
-In the case at hand, the first question is which noun constitutes the head of the whole construction.
-This question was answered quite clearly by \cite{Loebel1986} already (see also \citealp[213]{Eschenbach1994}, and \citealp[16]{GallmannLindauer1994}).
-As hinted above, subject-verb agreement is always realised on the measure noun, and we can therefore assume that the measure noun is the head.
-
-I therefore turn to the MNP-internal syntax, which will show itself to be highly revealing.
-It cannot be understood properly without taking into account the strong\slash weak inflection patterns of adjectives described in Section~\ref{sec:descriptive} (especially Footnote~\ref{fn:strongweak}).
-In NPs with a strongly inflected determiner, attributive adjectives inflect according to the massively syncretistic \textit{weak} pattern.%
-\footnote{The structural cases (nominative and accusative) in the singular are marked with \textit{-e}, the oblique cases (dative and genitive) as well as the entire plural are marked with \textit{-en}.
-The only exception to this rule is the masculine accusative singular, which is also marked with \textit{-en}.}
-If there is no determiner (as is the case in the alternating constructions), attributive adjectives inflect like determiners, which is called the \textit{strong} pattern (see Table~\ref{tab:syncretisms}).
-Thus, the adjectives in the target constructions have properties of adjectives as well as determiners in a sense:
-they are lexical adjectives and function as attributive modifiers, but they are inflected as determiners and are the leftmost element in the NP.
-This unusual double nature of adjectives in NPs without a determiner leads to a plausible interpretation of the pattern shown in Table~\ref{tab:constructions}.
-Whenever speakers classify the adjective in the kind noun phrase as a determiner, they have to use the PGC because, if there is a determiner, the PGC is the only option.
-When they classify the adjective as an adjective, the kind NP has no determiner, and the NAC is the only option.%
-\footnote{While the generative analysis presented in \cite{Bhatt1990} cannot properly deal with probabilistic effects, Bhatt comes close to this interpretation by analysing the kind NP in the GPC as a DP and in the NAC as an NP.
-Within the confines of her framework, this is the appropriate was of encoding the relevant generalisation.}
-In a framework that assumes similarity-based probabilistic classification, such as Prototype Theory, we can now regard the PGC with a determiner in the kind noun phrase as instantiating the PGC prototype and the NAC with just a bare kind noun as the NAC prototype.
-The choice of the alternating variants is then expected to be influenced by the degree of similarity to either of these prototypes.
-
-The situation is made difficult by the fact that the semantic differences between the variants are at best subtle, which means the prototypes are not associated with nice and clearly detectable differences in meaning.
-However, both the kind noun lemmas and the measure noun lemmas probably appear with different relative frequencies in the NAC with a bare kind noun and the PGC with a determiner inside the kind noun phrase (in part possibly due to their meaning and usage).
-Their relative frequencies in these non-alternating configurations should be indicative of their preference to appear in one of the two variants.
-In other words, lemmas themselves might have preferences for one prototype or the other.
-
-%Related to the , dative measure nouns favor the PGC more than nominative and accusative (also \citealp{Hentschel1993,Zimmer2015}).
-%As an example, \textit{mit einem Stück frischen Brots} `with a piece of fresh bread' (PGC) would be preferred over \textit{mit einem Stück frischem Brot} (NAC).
-%I propose
-
-Furthermore, the grammaticalisation path as described above leads from NPs denoting highly individuated objects standing in a \textit{separated-from} relation to a construction with a \textit{part-of} meaning (still available in the prototypical PGC) to a construction with a more diffuse \textit{quantity-of} relation (characteristic of the prototypical NAC).
-The best evidence for the relevance of this factor in an ongoing grammaticalisation process is that plural kind nouns almost always trigger PGC and virtually never NAC (Section~\ref{sec:descriptive}).
-More indirectly, plurality of the measure noun could favor the PGC over the NAC, and exact quantification in the form of cardinals even more so.%
-\footnote{The effect of the plural was empirically shown but not explained in \cite{Hentschel1993}.}
-As a highly illustrative example (using a measure noun which has virtually developed two variants, one of which grammaticalsed to a much higher degree), \textit{ein Haufen französische Watte} `a heap\slash lot of French cotton' (NAC) has essentially two uses.
-It can be used to denote an unspecific amount of cotton in the sense of \textit{a lot of French cotton} in situations when the exact quantity of cotton is not of interest.
-Under the appropriate circumstances, it could also be used to denote a highly specific quantity of cotton, for example when a cotton merchant always sells her French cotton in \textit{heaps} weighing two stone each.
-The former is highly restricted for \textit{zwei Haufen französischer Watte} `two heaps/\textdagger lots of French cotton' (GPC).
-Such an NP is less likely to be used in situations where the exact quantity of cotton is irrelevant, even if there are actually two heaps of cotton.
-In other words, a precisely quantified measure noun leads to a higher individuation (the creation of portions) of the substance denoted by the kind noun, which is more appropriate for the PGC prototype.%
-\footnote{In the most extreme, some measure nouns cannot be pluralised, such as \textit{eine Menge} `a lot of'.
-Also note that the presentation of the examples in NAC and PGC is suggestive.
-Native speakers will confirm that both NPs sound highly acceptable in both GPC and NAC.
-It was made clear from the beginning and throughout that the preferences are indeed subtle.}
-
-Finally, it is claimed that the PGC is more typical of higher registers or even exclusive to written language (see \citealp[320--323]{Hentschel1993}, where she also refers to descriptive grammars).
-This is not surprising in as much as the genitive -- an intrinsic part of the PGC -- is generally underrepresented in colloquial and vernacular variants of German, and thus overrepresented to the same degree as colloquial, vernacular, and spoken variants are overrepresented in most corpora of written German.
-There is no reason to assume that such preferences for cannot be part of the construction prototypes.
-In the corpus study, the alternation tendencies controlled by this complex factor related to register, medium, and lectal variation will be tested using certain very simple proxies.
-
-To summarise the discussion and core hypotheses, it is expected that:
-
-\begin{enumerate}[i.]
-  \item classes of more clearly grammaticalised measure nouns favor the NAC,
-  \item the relative frequencies with which measure noun lemmas and kind noun lemmas appear in the prototypical (unequivocal) PGC and NAC are positively correlated with the relative frequency the PGC and the NAC are chosen in the alternating case,
-  \item MNPs with masculine dative kind nouns disfavour the GPC comparatively less
-  \item pluralic measure nouns and measure nouns quantified by cardinals favor the PGC variant
-  \item .
-\end{enumerate}
-
-In the next section, I report the corpus study that was designed to test these hypotheses on usage data.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Corpus study}
-\label{sec:corpusstudies}
-
-\subsection{Getting the right data}
-\label{sec:gettingdata}
-
-% TODO Quote:
-% Brems (2016:171--175) for using GOOGLE data, including COUNTS!
-
-Before I turn to the concrete sampling procedure, a few words are in order regarding the choice of the corpus.
-I used the German \textit{Corpus from the Web} (COW) DECOW14(A) (\citealp{SchaeferBildhauer2012full,Schaefer2015b} and \citealp{BiemannEa2013,SchaeferBildhauer2013} for overviews of web corpora in general and the methodology of their construction), which contains 21 billion tokens, for two main reasons.%
-\footnote{The corpora are made available for free at \url{https://www.webcorpora.org}.
-At the time of this writing, the DECOW16 corpus had already been released.}
-First, the external validity of the study is increased through a higher heterogeneity of the sample \citep[30]{MaxwellDelaney2004}, and the DECOW14 corpus has clearly a much more heterogeneous composition compared to the only other very large corpus of German, the DeReKo \citep{KupietzEa2010} of the Institute for the German Language (IDS), which contains almost exclusively newspaper texts.%
-\footnote{It was shown in \cite{W16-2601} that, for example, the spread of topics is much smaller in DeReKo compared to DECOW14.}
-Second, it was already mentioned that normative grammars disagree on whether both NAC and PGC are grammatical in the case at hand.
-Thus, newspaper text or any other text that conforms strongly to normative standards might not represent the alternation phenomenon fully (and without bias) because authors and proof-readers might favour one alternative or the other.
-Web corpora, on the other hand, contain at least some amount of non-standard language from forums and similar sources.
-For these or similar reasons, COW corpora have been used in a number of peer-reviewed publications, for example \cite{VanGoethemHiligsmann2014,VanGoethemHuening2015,MuellerS2014,Schaefer2016c,SchaeferSayatz2014,SchaeferSayatz2016,StefanowitschFlach2016,Zimmer2015}.
-Therefore, DECOW14 can be considered the obvious choice for this study.
-
-In Section~\ref{sec:germanmeasurenps} it was mentioned that there are vague intuitions about lemma-specific preference effects w.\,r.\,t.\ the kind noun.
-Also, degrees of grammaticalisation of the measure noun might influence the choice of one of the constructions.
-Therefore, it was necessary to obtain sample in which all (or at least most of the highly frequent) actually occurring combinations of kind noun and measure noun were represented.
-This section describes the three-stage bootstrap process that was used to obtain such a sample, which was then used to estimate the influences of various contextual factors on the alternation (cf.\ Section~\ref{sec:corpusstudies}).
-The bootstrap process takes into account the imperfection of automatically generated annotations in corpora.
-
-Automatic linguistic annotation (for lemma, part-of-speech, morphological information, etc.) is never perfect, especially if the material deviates from the standard as understood by the creators of the tools used for the annotation (cf.
-\citealp{BeisswengerEa2016,GiesbrechtEvert2009} and chapter 4 of \citealp{SchaeferBildhauer2013}).
-For examining grammatical alternations, relying on existing annotations is therefore unwise because there is no systematic way of assessing the accuracy of the annotation tools on the target constructions.
-Therefore, I did not rely on the annotations for part-of-speech, noun phrase chunks, or noun case available in DECOW14 wherever it was crucial to sample with a near-perfect recall.%
-\footnote{The recall is the percentage of retrieved data relative to the data that should have been found.
-If, for example, the corpus contains one hundred targets and a query retrieves only 80 of them, the recall is 80\% (or 0.8).}
-
-Instead, I bootstrapped a sample which most likely contained all reasonably frequent combinations of measure nouns and mass nouns, and which was almost perfectly clean (\ie\ it did contain virtually no erroneous non-target results).
-The bootstrapping consisted of three steps:\\
-
-\begin{enumerate}[i.]
-  \item\label{enum:data:step1} bootstrapping a list of the one hundred most frequent mass nouns,
-  \item\label{enum:data:step2} bootstrapping a list of all measure nouns with which the mass nouns co-occur in the unambiguous measure construction with a determiner, and
-  \item\label{enum:data:step3} sampling the target constructions by querying each combination of mass noun and measure noun found in step \ref{enum:data:step2}.
-\end{enumerate}
-
-In step \ref{enum:data:step1}, I sampled a list of all nouns in the DECOW14A01 corpus sorted by their token frequency and manually went through it from the most frequent noun downwards, selecting the first one hundred mass nouns that occurred in the list.%
-\footnote{DECOW14A01 is the first slice (roughly a twentieth) of the complete DECOW14A corpus.
-It contains just over one billion tokens.
-The creators assigned the documents randomly to the slices, such that each slice can be considered representative of the whole corpus \cite{Schaefer2015b}.}
-For this step, I relied on the noun part-of-speech label because for frequent nouns, high tagging accuracy can be expected, especially given the fact that German nouns are always spelled with an initial capital letter.
-Abstract nouns which partially behave like mass nouns (like \textit{Spaß} `fun’ or \textit{Gefahr} `danger’) were excluded because they are usually not quantified in the same way as concrete mass nouns.%
-\footnote{They behave like mass nouns mostly in that they occur without a determiner in the singular.}
-The hundredth selected mass noun was \textit{Schmuck} `jewelry’, which is the 3,054th most frequent noun in the original frequency list.
-
-This resulting list of mass nouns was used in step \ref{enum:data:step2} to bootstrap a list of measure nouns co-occurring with the mass nouns from DECOW14A01.
-In order to generate this list, I utilized the fact that a direct sequence of two bare nouns very often instantiates the bare-noun NAC if the second noun is a mass noun.
-In other words, I searched for all combinations [N\Sub{1} N\Sub{2}] with N\Sub{2} instantiated as a mass noun lemma.
-Then, the resulting lists of noun--noun combinations were each sorted by frequency in descending order and sieved manually to remove erroneous hits.%
-\footnote{In order to speed up the procedure without losing much material, I stopped when I had reached noun–noun combinations with an absolute frequency of 1 if twenty combinations for the respective mass noun (with a frequency of one or higher) had already been found.}
-The result was a list of 2,365 individual combinations of measure nouns and mass nouns.
-
-In step \ref{enum:data:step3}, each of these 2,365 noun–noun combinations was queried in the target construction individually in each of the first ten slices of DECOW14 (roughly 10 billion tokens).
-The scripts generating the actual queries were crafted so as to make sure that adjectives and nouns were searched in the appropriate forms.
-(\ref{ex:data:query01}) shows an example query for a neuter mass noun like \textit{Wasser} (`water’) which finds relevant sequences such as the ones shown in (\ref{ex:data:query02}), which all mean `litre of clear water’.
-The versions with the genitive singular \textit{Liters} and the dative plural \textit{Litern} are not shown in (2) for space reasons but are also found by the query in (1).
-In terms of context, single sentences containing the target were exported.
-
-\begin{exe}
-  \ex{\label{ex:data:query01} \texttt{\small[ lemma="Liter" ] [ pos="ADJA" \& word=".+(en|em|es)" ] [ lemma="Wasser" ]}}
-  \ex\label{ex:data:query02}
-  \begin{xlist}
-    \ex{\label{ex:data:query02a} Liter [klaren Wassers]\Sub{Gen}}
-    \ex{\label{ex:data:query02b} Liter [klares Wasser]\Sub{Nom\slash Acc}}
-    \ex{\label{ex:data:query02c} Liter [klarem Wasser]\Sub{Dat}}
-  \end{xlist}
-\end{exe}
-
-To reduce the sample size for the manual annotation process, the final concordance was sampled from the results of these 2,365 queries.
-Since the mass nouns in the sample were distributed according to the usual power law, I used all hits for mass nouns occurring less than one hundred times, but randomly sampled one hundred hits for each mass noun that occurred one hundred or more times.
-The final sample contained 7,547 sentences.
-Given the careful bootstrapping and sampling procedure described in this section, we can be highly sure that it contains all relevant and reasonably frequent noun–noun combinations in the target construction.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{A prelude to the statistical analysis: There are no right statistics!}
+\subsection{A word on statistical analysis}
 \label{sec:rightstatistics}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\subsection{Descriptive statistics and non-hierarchical models}
-\label{sec:corpusnonhierarchicalmodel}
+\section{Case assignment in German measure NPs}
+\label{sec:germanmeasurenps}
+
+\subsection{Two stable cases and a case alternation}
+\label{sec:descriptive}
+
+The alternation phenomenon to be discussed in this paper appears to be a straightforward case alternation at first sight.
+However, without a solid knowledge of German nominal morpho-syntax, some of the more subtle distinctions that have to be made in its analysis can be confusing.
+Therefore, I introduce and illustrate the relevant constructions in some detail in this section.
+The narrow range of syntactic constellations in which the alternation occurs is described, and the focus on \textit{only} this narrow (range rather than, for example, the whole range of nominal constructions expressing quantities) is motivated.
+This entails that I also have to discuss certain constructions and factors that might be seen as relevant, but which I exclude from my own study.
+
+I use the term \textit{measure noun phrase} (MNP) to refer a noun phrase (NP) in which a kind-denoting (count or mass) noun depends on another noun specifying a quantity of the objects or the substance denoted by the kind noun.
+I call the kind-denoting noun simply the \textit{kind noun} and the quantity-denoting noun the \textit{measure noun}.
+Measure nouns can be all sorts of nouns which genuinely denote a quantity (such as \textit{litre} or \textit{amount}) but also nouns denoting containers, collections, etc.
+(such as \textit{glass} or \textit{bucket}).
+In a similar vein, \citet[284]{Brems2003} considers nouns as measure nouns \textit{which, strictly speaking, do not designate a `measure', but display a more nebulous} (\textit{sic!}) \textit{potential for quantification} (see also \citealp[530]{Koptjevskaja2001}, and \citealp[338]{Rutkowski2007}).
+For illustration purposes, in the English \textit{a cup of fine coffee}, \textit{cup} is the measure noun, and \textit{coffee} is the kind noun.
+
+In the case at hand, three different syntactic configurations need to be distinguished \wrt case assignment inside German measure noun phrases.
+If the kind noun forms an NP with a determiner, the construction resembles (and is usually called) a \textit{pseudo-partitive} (on partitives and pseudo-partitives see, e.\,g., \citealp{Barker1998,Selkirk1977,Stickney2007,Vos1999}; for a recent application of the terminology to German, see \citealp{Gerstenberger2015}).%
+\footnote{Whereas partitives are constructions denoting a proper part-of relation strictly requiring a definite kind noun as in \textit{a sip of the wine}, pseudo partitives -- albeit syntactically similar and diachronically related to partitives in many languages -- merely denote quantities and contain indefinite kind nouns as in \textit{a sip of wine}.
+In the literature on German, some authors incorrectly call the pseudo-partitive a \textit{partitive} \cite{Hentschel1993} while some realise the difference and at least mention it \citep{Eschenbach1994,GallmannLindauer1994,Loebel1989,Zimmer2015}.}
+I use \textdagger\ to denote general unacceptability, and the acceptability judgements given in (\ref{ex:intro:pseudopartitive}) are completely undisputed.
+Notice that the kind noun is obligatorily in the genitive and that case agreement with the measure noun is not an option (but see immediately below).
+I refer to the construction in (\ref{ex:intro:pseudopartitive1}) as the \textit{Pseudo-partitive Genitive Construction} (PGC).
+
+\begin{exe}
+  \ex\label{ex:intro:pseudopartitive}
+  \begin{xlist}
+    \ex[ ]{\label{ex:intro:pseudopartitive1}\gll Wir trinken [[eine Tasse]\Sub{Acc} [eines leckeren Kaffees]\Sub{Gen}]\Sub{Acc}.\\
+    we drink a cup a tasty coffee\\
+    \glt We drink a cup of a tasty coffee.}
+    \ex[\textdagger]{\label{ex:intro:pseudopartitive2} Wir trinken [[eine Tasse]\Sub{Acc} [einen leckeren Kaffee]\Sub{Acc}]\Sub{Acc}.}
+  \end{xlist}
+\end{exe}
+
+When the kind noun is bare -- \ie\ when it comes without a determiner and without modifying adjectives -- the kind noun has to agree in case with the measure noun, and the genitive seen in the PGC is not acceptable, see (\ref{ex:intro:narrowapposition}).
+This construction is usually classified as a \textit{Narrow Apposition Construction} \citep{Loebel1986}, henceforth NAC.%
+\footnote{Notice that the unavailability of the genitive on the kind noun can be seen as following from a rather quirky constraint that genitive NPs in German require the presence of some strongly case-marked element (determiner or adjective) in addition to the head noun in order to be acceptable \citep{GallmannLindauer1994,Schachtl1989}.}
+
+\begin{exe}
+  \ex\label{ex:intro:narrowapposition}
+  \begin{xlist}
+    \ex[\textdagger]{\label{ex:intro:narrowapposition1} Wir trinken [[eine Tasse]\Sub{Acc} [Kaffees]\Sub{Gen}]\Sub{Acc}.}
+    \ex[ ]{\label{ex:intro:narrowapposition2}\gll Wir trinken [[eine Tasse]\Sub{Acc} [Kaffee]\Sub{Acc}]\Sub{Acc}.\\
+    we drink a cup coffee\\
+    \glt We drink a cup of coffee.}
+  \end{xlist}
+\end{exe}
+
+The construction as in (\ref{ex:intro:narrowapposition2}) is also referred to as the \textit{Direct Partitive Construction} for other Germanic languages in which the PGC with the synthetic genitive is not available.%
+\footnote{This nomenclature makes sense in contrast to the \textit{Indirect Partitive Construction} with prepositional linkers translating to `of' -- \ie\ analytic genitives -- in such languages, see \cite{HankamerMikkelsen2008} for Danish.
+For German, this terminology is not distinctive enough, which is why I use the terms NAC and PGC.}
+While only an accusative is shown in (\ref{ex:intro:narrowapposition}), the kind noun obligatory agrees in case also with nominative and dative measure nouns.
+
+The actual alternation can be observed \textit{only when the kind noun occurs with an attributive adjective but without a determiner}, as in (\ref{ex:intro:alternation}), where both the NAC in (\ref{ex:intro:alternation1}) and the PGC in (\ref{ex:intro:alternation2}) are equally acceptable.
+They are by-and-large functionally and semantically equivalent.
+However, Section~\ref{sec:analyses} is devoted to developing hypotheses about subtle differences between them.%
+\footnote{Some descriptive and normative grammars take stronger positions w.\,r.\,t.\ the acceptability of the two options.
+See \cite{Hentschel1993,Zimmer2015} for analyses of the sometimes absurd stances taken in grammars of German.
+As the usage and experimental data presented below should render it clear, there might be preferences under certain circumstances, but we cannot assume either construction to be unacceptable.}
+
+\begin{exe}
+  \ex\label{ex:intro:alternation}
+  \begin{xlist}
+    \ex[ ]{\label{ex:intro:alternation1}\gll Wir trinken [[eine Tasse]\Sub{Acc} [heißen Kaffee]\Sub{Acc}]\Sub{Acc}.\\
+    we drink a cup hot coffee\\
+    \glt We drink a cup of hot coffee.}
+    \ex[ ]{\label{ex:intro:alternation2} Wir trinken [[eine Tasse]\Sub{Acc} [heißen Kaffees]\Sub{Gen}]\Sub{Acc}.}
+  \end{xlist}
+\end{exe}
+
+\begin{table}
+  \centering
+  \begin{tabular}{lccc}
+    \multicolumn{1}{r}{kind NP is:} & bare noun NP & NP with adjective & NP with determiner \\
+    & [\ldots N\Subsf{meas} [N\Subsf{kind}]] & [\ldots N\Subsf{meas} [AP N\Subsf{kind}]] & [\ldots N\Subsf{meas} [D N\Subsf{kind}]] \\
+    \midrule
+    narrow apposition         & \CheckIt & \CheckIt &          \\
+    pseudo-partitive genitive &          & \CheckIt & \CheckIt \\
+  \end{tabular}
+  \caption{Distribution of the NAC and PGC constructions in different NP structures}
+  \label{tab:constructions}
+\end{table}
+
+The distribution of the case patterns in the NAC (case identity between the measure noun and the kind noun) and in the PGC (genitive on the kind noun) depending on the structure of the kind NP are summarised in Table~\ref{tab:constructions}.
+I now turn to some more subtle issue related to the measure noun case alternation, namely:
+
+\begin{enumerate}[i.]
+  \item{\label{it:intro:idiot} alleged alternatives to the NAC and the PGC}
+  \item{\label{it:intro:datsg} alternative \textit{weak} forms of dative singular neuter adjectives}
+  \item{\label{it:intro:femsg} case syncretism in feminine NPs}
+  \item{\label{it:intro:plurl} similar constructions with plural/collective kind nouns}
+  \item{\label{it:intro:noifl} grammaticalised non-inflected measure nouns}
+  \item{\label{it:intro:other} alternative constructions for expressing quantities}
+\end{enumerate}
+
+\vspace{-1\baselineskip}
+
+First of all, (\ref{it:intro:idiot}) refers to claims found in some grammars that a generic nominative, accusative, and even dative on the kind noun are used instead of the genitive or case agreement (overviews in \citealp{Hentschel1993,Zimmer2015}).
+It was shown empirically in \cite{Hentschel1993} that such variants are de facto not acceptable.
+Also, in my corpus sample, they simply did not occur.
+Even if they are accepted by some speakers, their extremely low frequency makes it virtually impossible to study them using either corpus methods (Section~\ref{sec:corpusstudies}) or experimental approaches (Section~\ref{sec:externalvalidation}), and I consequently ignore them.
+
+As for (\ref{it:intro:datsg}), a complication with neuter kind nouns in the dative is mentioned by \citet[20--22]{Zimmer2015}.
+He reports a high number of occurrences of adjectives being inflected according to the \textit{weak} inflectional pattern, which is normally only used if a strongly inflected determiner precedes the adjective: \textit{kalt-en Wasser} `cold water' instead of \textit{kalt-em Wasser}.%
+\footnote{See Section~\ref{sec:analyses} for more discussion of adjectival inflection.}
+In my corpus sample (released with this article), this tendency was not nearly as clear, and there was a high number of very noisy sentences among those potentially showing this pattern.
+Hence, I do not discuss these forms much, although I will offer a possible interpretation in Section~\ref{sec:analyses}.
+
+\label{page:femininesyncretism}
+With feminine kind nouns, (\ref{it:intro:femsg}) is relevant.
+Whereas the singular masculine and neuter nominal subsystem is a four case system (at least for NPs with adjectives and determiners), feminine singular NPs show nominative-accusative and dative-genitive syncretisms.
+Table~\ref{tab:syncretisms} shows this as well as the partial syncretism in the plural, where no gender distinctions are made (for an overview, see \citealp[XYZ]{Eisenberg2013a}).
+This means that with dative measure nouns, the alternation is unobservable if the kind noun is feminine, and it is problematic to speak of a \textit{genitive} kind noun construction in this case.
+This is taken into account in the model design and the discussion presented in Section~\ref{sec:corpusstudies}.
+
+\begin{table}
+  \centering
+  \begin{tabular}{llll}
+     & Masculine\slash Neuter & Feminine & Plural \\
+     \midrule
+     nominative & rot-er Wein    & \multirow{2}{*}{frisch-e Sahne}   & \multirow{2}{*}{klein-e Äpfel} \\
+     accusative & rot-en Wein    &                                   &                                \\
+     dative     & rot-em Wein    & \multirow{2}{*}{frisch-er Sahne}  & klein-en Äpfel-n               \\
+     genitive   & rot-en Wein-es &                                   & klein-er Äpfel                 \\
+  \end{tabular}
+  \caption{Case syncretisms in NPs with an adjective and without a determiner (\textit{roter Wein} `red wine', \textit{frische Sahne} `fresh cream', \textit{kleine Äpfel} `small apples')}
+  \label{tab:syncretisms}
+\end{table}
+
+Turning to (\ref{it:intro:plurl}), readers might have noticed that so far only MNPs denoting quantities of substances (mass kind nouns such as \textit{ein Glas roter Wein\slash roten Weines} `a glass of red wine') have been discussed.
+If the kind noun is a plural count noun with a collective denotation such as \textit{ein Sack kleine Äpfel\slash kleiner Äpfel} `a bag of small apples' (see Table~\ref{tab:syncretisms} for the inflection pattern), a similar alternation between PGC and NAC can be observed.%
+\footnote{\label{fn:eschenbash}On the collective interpretation of such constructions, see \cite{Eschenbach1994}.
+It is, however, unfortunate that Eschenbach does not notice and discuss the morpho-syntactic distribution of PGC and NAC, even when she gets very close to listing the relevant data \citep[217]{Eschenbach1994}.}
+In line with experimental results reported in \citet[15--16]{Zimmer2015}, I found that the PGC is so dominant with plural kind nouns (over 92\%, cf.\ Section~\ref{sec:corpusstudies}) that the alternation cannot be analysed in the same way as in the singular case.
+While this will play a role in the interpretation of the corpus findings, MNPs with plural\slash collective kind nouns will not be dealt with prominently in the remainder of the paper.
+A statistical model for plural kind nouns will be reported in passing in Section~\ref{sec:corpusstudies}, however.
+
+Turning to issue (\ref{it:intro:noifl}), some measure nouns have been grammaticalised in a way that they always appear non-inflected.
+They are typical measure nouns like \textit{Gramm} `gram', \textit{Pfund} `pound' or \textit{Prozent} `percent', which do not have plural forms at all.%
+\footnote{With neglectable frequency, this effect is extended to container nouns like \textit{Glas} `glass', but with a twist.
+A non-inflected variant \textit{zwei Glas roter Wein} `two glasses of wine' and an inflected variant \textit{zwei Gläser roter Wein} co-exist.
+The non-inflected variant denotes the quantity of wine fitting into two glasses, the inflected variant denotes two glasses filled with wine.}
+I treat these cases like any other measure noun because they enter into both the NAC as in (\ref{ex:uninflectedmeasures:a}) and the PGC as in (\ref{ex:uninflectedmeasures:b}).
+As will be explained in Section~\ref{sec:analyses}, they are expected to exert a specific influence on the choice of NAC and PGC.
+
+\begin{exe}
+  \ex\label{ex:uninflectedmeasures}
+  \begin{xlist}
+    \ex{\label{ex:uninflectedmeasures:a} zwei Gramm brauner Zucker}
+    \ex{\label{ex:uninflectedmeasures:b}\gll zwei Gramm braunen Zuckers\\
+                                         two gram brown sugar\\
+				         two grams of brown sugar}
+  \end{xlist}
+\end{exe}
+
+Finally, (\ref{it:intro:other}) suggests that there are alternative ways of expressing similar quantificational meanings.
+First of all, the analytic (pseudo-)partitive with \textit{von} `of' is only available as an alternative to the PGC if the kind noun phrase contains a (definite or indefinite) determiner as in (\ref{ex:analyticalpartitive}).
+
+\begin{exe}
+  \ex\label{ex:analyticalpartitive} 
+  \begin{xlist}
+    \ex[ ]{\label{ex:analyticalpartitive:a}\gll ein Glas von dem roten Wein\\
+                                                a glass of the red wine\\
+                                                \glt a glass of the red wine
+       }
+    \ex[\textdagger]{\label{ex:analyticalpartitive:b}\gll ein Glas von rotem Wein\\
+                                                a glass of red wine\\
+                                                \glt a glass of red wine
+       }
+  \end{xlist}
+\end{exe}
+
+However, if the measure noun denotes a container or vessel, constructions with \textit{voll} (\textit{von})\slash\textit{voller} `full (of)' and \textit{mit} `with' are available, see (\ref{ex:alternatives}).
+
+\begin{exe}
+  \ex\label{ex:alternatives}
+  \begin{xlist}
+  \ex{\label{ex:voll}\gll   ein Glas voll von rotem Wein\\
+                            a glass full of red wine\\
+			    \glt a glass full of red wine}
+  \ex{\label{ex:voller}\gll ein Glas {voll\slash voller} {rotem\slash roter} Wein\\
+                            a glass full-of red wine\\
+			    \glt a glass full of red wine}
+  \ex{\label{ex:mit}\gll    ein Glas mit rotem Wein\\
+                            a glass with red wine\\
+			    \glt a glass with red wine in it\slash a glass filled with red wine}
+  \end{xlist}
+\end{exe}
+
+These construction have very idiosyncratic properties (the construction with \textit{voller} is discussed in \citealp{Zeldes2018}, the construction with \textit{mit} is discussed in \citealp{Bhatt1990}), they are not semantically equivalent to NAC and PGC, and -- most importantly -- they are only available for a small subset of measure nouns.
+Subtle interactions between these constructions and NAC\slash PGC cannot be excluded, especially because the former might act as a (partial) alternative and thus reduce the count of exemplars of the NAC, the PGC, or both for container measure nouns.
+There are no clear and motivated hypotheses about such interactions, however.
+Nevertheless, the corpus-based models in Section~\ref{sec:corpusstudies} are specified in a way that they could detect preferences for either NAC or PGC for the class of container or vessel measure nouns, for which the alternative constructions illustrated in (\ref{ex:alternatives}) is available.
+
+This concludes the descriptive overview of the alternation.
+I have shown that there is an alternation between two measure noun constructions in a narrow syntactic configuration (kind NP with an adjective but without a determiner), and that the two constructions differ in the case of the kind noun (case agreement with the measure noun or genitive).
+I turn to some more theory-oriented discussion in the next section.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{A theoretical assessment}
+\label{sec:analyses}
+
+This section reviews some existing analyses of the PGC-NAC alternation or related issues, and I develop my own analysis and the appropriate hypotheses for the empirical studies presented in Sections~\ref{sec:corpusstudies} and \ref{sec:externalvalidation}.
+While the constructions and the alternation are standardly covered in descriptive grammars such as \citet[XYZ]{Eisenberg2013a}, the research literature contains little specific discussion of the situation in German.
+In The remainder of this section, I discuss as factors that might influence the alternation:
+degrees of grammaticalisation of the measure noun, the different syntactic structures of the variants and their syntactic prototypes, and potential effects of semantic prototypicality.
+
+Turning to grammaticalisation first, it is often assumed that pseudo-partitives arise as a form of grammaticalised partitives (\eg\ \citealp[536--539]{Koptjevskaja2001} for Finnish and Estonian, \citealp[559]{Koptjevskaja2001} for European languages per se).%
+\footnote{Notice that the description of German in \citet[547--549]{Koptjevskaja2001} is vague and partly incorrect, maybe because her main source \cite{Eschenbach1993} is not as clear as it could be (see also Footnote \ref{fn:eschenbash} above).
+Her example (29b) from \citet[71]{Eschenbach1993}, repeated here verbatim as (\ref{ex:koptjewskaja}) is misclassified as a clear case of a dative on the measure noun and a genitive on the kind noun (hence, PGC).
+
+\begin{exe}
+  \ex{\label{ex:koptjewskaja}\gll trotz einer Flasche gut-en Wein-s\\
+  in.spite.of one-\textsc{fem.gen} bottle good-\textsc{masc.sg.gen} wine-\textsc{gen}\\
+  \glt in spite of one bottle of good wine}
+\end{exe}
+
+It could, however, equally well be a genitive on the measure noun because of the syncretisms discussed here on p.\ \pageref{page:femininesyncretism} and an ambiguous case requirement of the preposition \textit{troz}, hence an instance of NAC.
+Then, she does not clearly distinguish acceptable from non-acceptable constructions (cf.\ Section \ref{sec:descriptive}) on her pp.\ 548 and 549, including a rather confusing \textit{???} rating of her (31b), which is clearly a non-acceptable construction of the type illustrated in (\ref{ex:analyticalpartitive:b}) above.
+}
+The alternating constructions in German are both clearly pseudo-partitives, but the grammaticalisation paths uncovered by \cite[esp.\ 526--530]{Koptjevskaja2001}, which I interpret freely here, are clearly relevant for an analysis of the alternation.
+The grammaticalisation path can start out with constructions involving two referential nouns (not necessarily forming a single NP) and a \textit{separative} meaning as in \textit{(cut) two slices from the cake} \citep[535]{Koptjevskaja2001}.
+The \textit{part-of} meaning of true partitives as in \textit{a slice of the case} represents the first stage of a development wherein the measure noun already tends to loose some semantic content.
+The pseudo-partitive stage finally instantiates a \textit{quantity-of} relation, potentially even leading to fully grammaticalised quantifiers such as \textit{a lot}.
+In German, the PGC is clearly the older variant \citep{Zimmer2015} with an overt \textit{quantity-of} marker in the form of the genitive suffix, and at least the potential to form a true partitive (if the kind noun is definite), which the NAC lacks.
+Hence, we can expect more strongly grammaticalised measure nouns to prefer the NAC.
+Since clearly non-referential nouns like \textit{Gramm} `gram' and \textit{Meter} `metre' should favor the NAC according to this hypothesis, the effect will be detected by including information about measure noun classes in the corpus-based models in Section~\ref{sec:corpusstudies}.
+
+On the other hand, there are case studies like \cite{Brems2003,DeclerckBrems2016} where syntactic evidence for grammaticalisation plays a prominent role.
+In English -- \cite{Brems2003} argues based on a view from \cite{Langacker1991} -- strongly lexicalised and thus semantically bleached measure nouns turn from nouns into quantifiers, leading to a syntactic reanalysis of the NP, wherein the kind noun becomes the head.
+The argument is supported by convincing syntactic evidence, for example subject-verb agreement.
+In German MNPs, the measure noun always remains the syntactic head, however (see immediately below for references), and there is no similar clear criterion to detect grammaticalisation.
+The analysis of the morpho-syntax of German MNPs, to which I turn now, will provide hypothetical cues along similar lines, however.
+
+A major interest in descriptive and generative linguistics is to find a plausible analysis of the hierarchical syntactic structure of the constructions in question.
+In the case at hand, the first question is which noun constitutes the head of the whole construction.
+This question was answered quite clearly by \cite{Loebel1986} already (see also \citealp[213]{Eschenbach1994}, and \citealp[16]{GallmannLindauer1994}).
+As hinted above, subject-verb agreement is always realised on the measure noun, and we can therefore assume that the measure noun is the head.
+
+I therefore turn to the MNP-internal syntax, which will show itself to be highly revealing.
+It cannot be understood properly without taking into account the strong\slash weak inflection patterns of adjectives briefly described in Section~\ref{sec:descriptive} already.
+In NPs with a strongly inflected determiner, attributive adjectives inflect according to the massively syncretistic \textit{weak} pattern.%
+\footnote{The weak pattern has only two suffixes.
+The structural cases (nominative and accusative) in the singular are marked with \textit{-e}, the oblique cases (dative and genitive) as well as the entire plural are marked with \textit{-en}.
+The only exception to this rule is the masculine accusative singular, which is also marked with \textit{-en}.}
+If there is no determiner (as is the case in the alternating constructions), attributive adjectives inflect like determiners, which is called the \textit{strong} pattern (see Table~\ref{tab:syncretisms}).
+Thus, the adjectives in the target constructions have properties of adjectives as well as determiners.
+They are lexical adjectives and function as attributive modifiers, but they are inflected as determiners and are the leftmost element in the NP, which is typical of determiners.
+This unusual double nature of adjectives in determinerless NPs leads to a plausible interpretation of the pattern shown in Table~\ref{tab:constructions}.
+Whenever speakers classify the adjective in the kind noun phrase as a determiner, they \textit{have} to use the PGC because, if there is a determiner, the PGC is the only option.
+When they classify the adjective as an adjective, the kind NP has no determiner, and they \textit{have} to use the NAC.%
+\footnote{While the generative analysis presented in \cite{Bhatt1990} cannot properly deal with probabilistic effects, Bhatt comes close to this interpretation by analysing the kind NP in the GPC as a DP and in the NAC as an NP.
+Within the confines of her framework, this is the appropriate was of encoding the relevant generalisation.}
+Now, in a framework that assumes similarity-based probabilistic classification, such as Prototype Theory, we can regard the PGC with a determiner in the kind noun phrase as instantiating the PGC prototype and the NAC with just a bare kind noun as the NAC prototype.
+The choice of the alternating variants is then expected to be influenced by the degree of similarity to either of these prototypes.
+
+Any attempt to substantiate or test this hypothesis is rendered difficult by the fact that the semantic differences between the variants are at best subtle, which means the prototypes are not associated with clearly detectable differences in meaning.
+However, measure noun lemmas all have individual and group-wise lexical-semantic properties which make them more appropriate for one prototype or the other.
+As a result, kind noun lemmas and measure noun lemmas probably appear with different relative frequencies in the prototypical NAC (with a bare kind noun) and the prototypical PGC (with a determiner inside the kind noun phrase).
+Their relative frequencies in these non-alternating configurations should be indicative of their preference to appear in one of the two variants.
+In other words, lemmas themselves might have specific degrees of affinity to prototype or the other, and this affinity, measured by looking at the non-alternating constructions, should be predictive of their frequencies in the alternating cases.
+
+%Related to the , dative measure nouns favor the PGC more than nominative and accusative (also \citealp{Hentschel1993,Zimmer2015}).
+%As an example, \textit{mit einem Stück frischen Brots} `with a piece of fresh bread' (PGC) would be preferred over \textit{mit einem Stück frischem Brot} (NAC).
+%I propose
+%
+
+Furthermore, the grammaticalisation path as described above leads from NPs denoting highly individuated objects standing in a \textit{separated-from} relation to a construction with a \textit{part-of} meaning (still available in the prototypical PGC) to a construction with a more diffuse \textit{quantity-of} relation (characteristic of the prototypical NAC).
+The best evidence for the relevance of this factor in an ongoing grammaticalisation process is that plural kind nouns almost always (but not exclusively) trigger PGC and virtually never NAC (Section~\ref{sec:descriptive}).
+More indirectly, plurality of the measure noun could also favor the PGC over the NAC, and exact quantification in the form of cardinals even more so.%
+\footnote{That the plural favors PGC was empirically shown but not theoretically motivated in \cite{Hentschel1993}.}
+As a highly illustrative example (using a measure noun which has virtually developed two variants, one of which grammaticalsed to a much higher degree), \textit{ein Haufen französische Watte} `a heap\slash lot of French cotton' (NAC) has essentially two uses.
+It can be used to denote an unspecific amount of cotton in the sense of \textit{a lot of French cotton} in situations when the exact quantity of cotton is not of interest.
+Under the appropriate circumstances, it could also be used to denote a highly specific quantity of cotton, for example when a cotton merchant always sells her French cotton in \textit{heaps} weighing two stone each.
+The former use, however, is highly restricted for \textit{zwei Haufen französischer Watte} `two heaps of French cotton' (GPC) -- not `two lots of French cotton'.
+Such an NP is less likely to be used in situations where the exact quantity of cotton is irrelevant, even if there are actually two heaps of cotton.
+In other words, a precisely quantified measure noun leads to a higher individuation (the creation of discrete portions) of the substance denoted by the kind noun, which is more appropriate for the PGC prototype.%
+\footnote{In the extreme, some measure nouns cannot be pluralised, such as \textit{eine Menge} `a lot of'.
+Also note that the presentation of the examples in NAC and PGC is suggestive.
+Native speakers will confirm that both NPs sound highly acceptable in both GPC and NAC.
+However, it was made clear from the beginning and throughout that the preferences are indeed subtle.}
+
+Finally, it has been iterated very often that the PGC is more typical of higher registers or even exclusive to written language (see \citealp[320--323]{Hentschel1993}, where she also refers to descriptive grammars).
+This is not surprising in as much as the genitive -- an intrinsic part of the PGC -- is generally underrepresented in colloquial and vernacular variants of German, and thus overrepresented to the same degree as colloquial, vernacular, and spoken variants are overrepresented in most corpora of written German.
+Under the integral view outlined in Section~\ref{sec:cogocl}, such preferences are assumed to be part of the construction prototypes.
+In the corpus study, the alternation tendencies controlled by this complex factor related to register, medium, and lectal variation will be tested using certain very simple proxies.
+
+To summarise the discussion and core hypotheses, I assume that the alternation is controlled by the similarity of the chosen lemmas (including the degree of grammaticalisation of measure nouns), certain morpho-syntactic choices, as well as the larger utterance context to two prototypes instantiated more straightforwardly by the two non-alternating cases of NAC and PGC.
+Concretely, I predict that:
+
+\begin{enumerate}[i.]
+  \item Classes of more clearly grammaticalised measure nouns favor the NAC.
+  \item The relative frequencies with which measure noun lemmas and kind noun lemmas appear in the prototypical (unequivocal) PGC and NAC are positively correlated with the relative frequency the PGC and the NAC are chosen in the alternating case.
+  \item MNPs with masculine dative kind nouns disfavour the GPC less.
+  \item Plural measure nouns and measure nouns quantified by cardinals favor the PGC variant.
+  \item The NAC is associated more strongly with higher registers.
+\end{enumerate}
+
+\vspace{-1\baselineskip}
+
+In the next section, I report the corpus study that was designed to test these hypotheses on usage data.
+The resulting models will then be validated using experimental methods in Section~\ref{sec:externalvalidation}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Corpus study}
+\label{sec:corpusstudies}
+
+\subsection{Getting the right data}
+\label{sec:gettingdata}
+
+% TODO Quote:
+% Brems (2016:171--175) for using GOOGLE data, including COUNTS!
+
+For the present study, I used the German \textit{Corpus from the Web} (COW) in its 2014 version DECOW14A (\citealp{SchaeferBildhauer2012full,Schaefer2015b} and \citealp{BiemannEa2013,SchaeferBildhauer2013} for overviews of web corpora in general and the methodology of their construction), which contains 21 billion tokens.
+I will simply refer to it as DECOW in the remainder of the paper except where clarification is necessary.
+I chose this corpus for two main reasons.%
+\footnote{The corpora are made available for free at \url{https://www.webcorpora.org}.
+At the time of this writing, a newer 2016 version DECOW16 had already been released.}
+First, the external validity of any study is increased through a higher heterogeneity of the sample \citep[30]{MaxwellDelaney2004}, and the DECOW corpus has clearly a much more heterogeneous composition compared to the only other very large corpus of German, the DeReKo \citep{KupietzEa2010} of the Institute for the German Language (IDS), which contains almost exclusively newspaper texts.%
+\footnote{It was shown in \cite{W16-2601} that, for example, the spread of topics is much smaller in DeReKo compared to DECOW.}
+Second, it was already mentioned that normative grammars often adopt clear positions on the grammaticality of the NAC or the PGC.
+Thus, newspaper text or any other text that conforms strongly to normative grammars might not represent the alternation phenomenon fully (and without bias) because authors and proof-readers might favour one alternative or the other.
+Web corpora, on the other hand, contain at least some amount of non-standard language from forums and similar sources.
+For these or similar reasons, COW corpora have been used in a number of peer-reviewed publications, for example \cite{VanGoethemHiligsmann2014,VanGoethemHuening2015,MuellerS2014,Schaefer2016c,SchaeferSayatz2014,SchaeferSayatz2016,Zimmer2015}.
+Therefore, DECOW can be considered the obvious choice for this study.
+
+I now turn to the sampling procedure applied to obtain concordances for manual annotation and statistical analysis.
+Among the factors potentially influencing the alternation (see Section~\ref{sec:germanmeasurenps}) were lemma-specific preference effects.
+Therefore, it was highly desirable to obtain a sample in which all (or at least most of the highly frequent) actually occurring combinations of kind noun and measure noun were represented.
+I applied a three-stage bootstrap process in order to obtain such a sample.
+It consisted of three steps:
+
+\begin{enumerate}[i.]
+  \item\label{enum:data:step1} bootstrapping a list of the one hundred most frequent mass nouns,
+  \item\label{enum:data:step2} bootstrapping a list of all measure nouns with which the mass nouns co-occur in the bare-noun NAC 
+  \item\label{enum:data:step3} sampling the target constructions by querying each combination of mass noun and measure noun found in step (\ref{enum:data:step2}).
+\end{enumerate}
+
+\vspace{-1\baselineskip}
+
+In step (\ref{enum:data:step1}), I exported a list of all nouns in the DECOW14A01 sub-corpus sorted by their token frequency and manually went through it from the most frequent noun downwards, selecting the first one hundred mass nouns that occurred in the list.%
+\footnote{DECOW14A01 is the first slice (roughly a twentieth) of the complete DECOW14A corpus.
+It contains just over one billion tokens.
+The creators assigned the documents randomly to the slices, such that each slice can be considered representative of the whole corpus \cite{Schaefer2015b}.}
+For this step, I relied on the noun part-of-speech label because for frequent nouns, high tagging accuracy can be expected, especially given the fact that German nouns are always spelled with an initial capital letter.
+Abstract nouns which partially behave like mass nouns (like \textit{Spaß} `fun’ or \textit{Gefahr} `danger’) were excluded because they are usually not quantified in the same way as concrete mass nouns.%
+\footnote{They behave like mass nouns mostly in that they occur without a determiner in the singular.}
+The hundredth selected mass noun was \textit{Schmuck} `jewelry’, which is the 3,054th most frequent noun in the original frequency list.
+
+This resulting list of mass nouns was used in step (\ref{enum:data:step2}) to bootstrap a list of measure nouns co-occurring with the mass nouns from DECOW14A01.
+In order to generate this list, I utilized the fact that a direct sequence of two nouns almost always instantiates the bare-noun NAC if the second noun is a mass noun.
+Hence, I searched for all sequences N\Sub{1}N\Sub{2} where N\Sub{2} was one of the mass noun lemmas extracted in step (\ref{enum:data:step1}).
+Then, the resulting 100 lists of noun-noun combinations were each sorted by frequency in descending order and sieved manually to remove erroneous hits.
+From each of the 100 lists, I also removed noun-noun combinations that had a frequency below 2, except if the individual list would have otherwise been shorter than 20 noun-noun combinations
+The result was a list of the most frequent 2,365 individual combinations of a measure noun and a mass noun.
+
+In step (\ref{enum:data:step3}), each of these 2,365 noun–noun combinations was queried in the target constructions (PGC and NAC with just an adjective) individually in each of the first ten slices of DECOW (roughly 10 billion tokens).
+The scripts generating the actual queries were crafted so as to make sure that adjectives and nouns were searched in the appropriate forms.
+The CQL query in (\ref{ex:data:query01}) shows an example query for a neuter mass noun like \textit{Wasser} `water' which finds relevant sequences such as the ones shown in (\ref{ex:data:query02}).%
+\footnote{The COW creators used to make corpora available in the IMS Open Corpus Query Workbench \citep{EvertHardie2011}, which has a query syntax informally referred to as CQL.
+Documentation can be obtained at \url{http://cwb.sourceforge.net}.}
+The versions with the genitive singular \textit{Liters} and the dative plural \textit{Litern} are not shown in (2) for space reasons but are also found by the query in (1).
+In terms of context, single sentences containing the target were exported.
+
+\begin{exe}
+  \ex{\label{ex:data:query01} \texttt{\small[ lemma="Liter" ] [ pos="ADJA" \& word=".+(en|em|es)" ] [ lemma="Wasser" ]}}
+  \ex\label{ex:data:query02}
+  \begin{xlist}
+    \ex{\label{ex:data:query02a} Liter [klaren Wassers]\Sub{Gen}}
+    \ex{\label{ex:data:query02b} Liter [klares Wasser]\Sub{Nom\slash Acc}}
+    \ex{\label{ex:data:query02c} Liter [klarem Wasser]\Sub{Dat}}
+  \end{xlist}
+\end{exe}
+
+To reduce the sample size for the manual annotation process, the final concordance was sampled from the results of these 2,365 queries.
+Since the mass nouns in the sample were distributed according to the usual power law, I used all hits for mass nouns occurring less than one hundred times, but randomly sampled one hundred hits for each mass noun that occurred one hundred or more times.
+The final sample contained 7,547 sentences.
+Given the careful bootstrapping and sampling procedure described in this section, we can be highly sure that it contains all relevant and reasonably frequent noun–noun combinations in the target constructions.%
+\footnote{In a similar fashion, the 100 most frequent measure nouns occurring with plural kind nouns were bootstrapped and queried.
+Bootstrapping measure nouns first was more efficient in the plural for reasons to complicated to discuss here, especially because the data was not properly analysed anyway.
+As stated in Section~\ref{sec:germanmeasurenps}, the NAC is virtually never used with plural kind nouns.
+Consequently, it does not make sense to model the influencing factors in the same way as for singular kind nouns.}
+
+Finally, two auxiliary samples were also drawn.
+As mentioned in Section~\ref{sec:analyses}, the distribution of the measure noun and kind noun lemmas in the bare-noun NAC and the PGC with a determiner will be modelled as factors influencing the alternation.
+Therefore, all noun-noun pairs from the bootstrap process were also queried in the two non-alternating constructions, resulting in 17,252 hits for the PGC with a determiner and 315,635 hits for the bar-noun NAC.
+Notice that the PGC with a determiner accounts for a mere 5.2\% of all non-alternating MNPs.
+
+In the next section, I describe the annotation process for the main sample.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Annotation}
+\label{sec:annotation}
+
+A mixture of automatic pre-annotation based on the annotations available in DECOW and manual annotations was used.
+Importantly, the case of the measure noun was annotated completely by hand because case syncretisms make automatic case annotations unreliable. 
+The full set of variables is given in Table~\ref{tab:variables}, and I briefly discuss them now.
+Notice that \textit{Genitive} is the boolean (dichotomous) response variable (`dependent variable') which is 1 when the PGC is chosen in a given sentence, and 0 when the NAC is chosen.
+
+\begin{table}
+  \centering
+  \begin{tabular}{llll}
+    Unit of reference & Variable                  & Type    & Levels (for factors only)   \\
+    \midrule
+    Document       & Badness                      & numeric &                             \\
+                   & Genitives                    & numeric &                             \\
+    Sentence       & \textbf{Genitive (response)} & boolean &                             \\
+                   & Matchlength                  & numeric &                             \\
+                   & Measureabbreviated           & boolean &                             \\
+                   & Measurecase                  & factor  & Nom, Acc, Dat               \\
+                   & Measurenumber                & factor  & Pl, Sg                      \\
+                   & Minus1pos                    & factor  & A, C, D, O, P               \\
+    Kind lemma     & Kindattraction               & numeric &                             \\
+                   & Kindconsistency              & factor  & Gaseous, Hard, Immaterial,  \\
+                   &                              &         & Liquid, Mixed, Powder, Soft \\
+                   & Kindedible                   & boolean &                             \\
+                   & Kindfreq                     & numeric &                             \\
+                   & Kindint                      & boolean &                             \\
+    Measure lemma  & Measureattraction            & numeric &                             \\
+                   & Measureclass                 & factor  & Amount, Capacity, Currency, \\
+                   &                              &         & Layer, Length, Portion,     \\
+                   &                              &         & Rest, Vessel, Weight        \\
+                   & Measurefreq                  & numeric &                             \\
+  \end{tabular}
+  \caption{Annotated variables in the main sample}
+  \label{tab:variables}
+\end{table}
+
+To capture the influence of register or style mentioned in Section~\ref{sec:analyses}, several proxy variables were used.
+At the document level, the DECOW corpus has an annotation for \textit{Badness}.
+As described in \cite{SchaeferEa2013}, \textit{Badness} measures how well the distribution of highly frequent short words in the document matches a pre-generated language model for German.
+Documents with higher Badness usually contain more incoherent language, shorter sentences, etc.
+If the PGC actually favors higher registers and styles, a high \textit{Badness} should be correlated with fewer occurrences of the PGC.
+
+Documents in DECOW14 have also been annotated with a variable called \textit{Genitives}.
+The higher the values of this variable, the lower the proportion of genitives among all case-bearing forms is.
+While more genitives are also indicative of higher registers and styles, this variable might be considered problematic in the present study.
+Since the PGC contains a genitive itself, the regressor variable \textit{Genitive} and the document-level variable \textit{Genitives} are not independent.
+However, since the PGCs make up for only a minute fraction of all genitives, I still use \textit{Genitives} as a regressor with the appropriate caveats.
+
+Since word length is also a good indicator of higher registers and styles, the length of the match was added as another proxy.%
+\footnote{The new version of the corpus contains over 100 document-level variables in the spirit of \cite{Biber1988}.
+Unfortunately, these annotations were not available when my samples were taken, and somewhat more crude proxies like the length of the query match had to be used.}
+Furthermore, abbreviated measure nouns like \textit{kg} or \textit{m} might be indicative of less elaborate technical styles, and hence the variable \textit{Measureabbreviated} was included, which is 1 if the measure noun is abbreviated.
+
+The variables \textit{Measurecase} and \textit{Measurenumber} simply encode the grammatical case and number of the measure noun.
+The variable \textit{Minus1pos} encodes the part-of-speech of the word preceding the measure noun.
+The level names stand for \textit{adjective}, \textit{cardinal}, \textit{determiner}, \textit{other}, and \textit{preposition} (in the same order as in Table~\ref{tab:variables}).
+The purpose of this variable is mainly to see whether cardinals really favor the PGC as hypothesised in Section~\ref{sec:analyses}.
+
+The variables \textit{Kindattraction} and \textit{Measureattraction} encode the ratio with which a given kind noun lemma or measure noun lemma occurs in the bare-noun NAC and the PGC with a determiner.
+They were calculated from the auxiliary samples described at the end of Section~\ref{sec:gettingdata}.
+The higher the value, the more often the noun occurs in the PGC with a determiner (proportionally).
+
+In Section~\ref{sec:analyses}, it was hypothesised that classes of lemmas might have different preferences for the two variants.
+For the kind nouns, it must be admitted that semantic classes were defined ad hoc based on intuitions of colleagues and native speakers, especially because classes available in lexical-semantic databases like GermaNet \citep{HampFeldweg1997} are too coarse for specialised purposes like the present study.
+As a result, \textit{Kindconsistency} with the self-explanatory factor levels in Table~\ref{tab:variables}) was annotated.
+Also, \textit{Kindedible} encodes whether the substance denoted by the kind noun is edible.
+Finally, \textit{Kindint} encodes whether the substance is an artifact made with intent (like \textit{Suppe} `soup') or not (like \textit{Meerwasser} `sea water').
+For the measure nouns, relevant classification seemed much clearer, and it was designed inspired by the list in \cite[530]{Koptjevskaja2001}.
+With typical examples, the classes are:
+\textit{Amount} (\textit{Menge} `amount'), 
+\textit{Capacity} (\textit{Liter} `litre'), 
+\textit{Currency} (\textit{Euro}), 
+\textit{Layer} (\textit{Schicht} `layer')
+\textit{Length} (\textit{Meter} `metre'), 
+\textit{Portion} (natural portions like \textit{Happen} `bite' or \textit{Krümel} `crumb'), 
+\textit{Vessel} (\textit{Eimer} `bucket'), 
+\textit{Weight} (\textit{Gramm} `gram').
+The few lemmas that did not fit into either of these classes were labeled \textit{Rest}.
+
+Last, to control for lemma frequency effects, the logarithm-transformed frequency per 1,000,000 words of each lemma was added from the frequency lists distributed by the DECOW designers (variables \textit{Kindfreq}, \textit{Measurefreq}).
+All numeric variables were standardised and centered (z-transformed) to facilitate their interpretation in the
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Hierarchical models}
+\label{sec:corpushierarchicalmodel}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Hierarchical models}
-\label{sec:corpushierarchicalmodel}
-
 In this section, I apply standard Maximum Likelihood methods for estimating the parameters of a multilevel model.
@@
 \footnote{
+The response $y_i=1$ if the PGC is chosen in sentence $i$, else $0$.
 $\beta_0$ is the general intercept, $B$ a vector of first-level coefficients, and $X_i$ a vector of regressor variables for sentence $i$.
@@
 \centering
-\includegraphics[width=0.9\textwidth]{figures/corpus/04_glmm_fixef_firstlevel}
+\includegraphics[width=0.8\textwidth]{figures/corpus/04_glmm_fixef_firstlevel}
 \caption{Estimated first-level effects in GLMMs for Masculine\slash Neuter and Feminine}
@@
 \centering
-\includegraphics[width=0.9\textwidth]{figures/corpus/04_glmm_fixef_secondlevel_kind}
+\includegraphics[width=0.8\textwidth]{figures/corpus/04_glmm_fixef_secondlevel_kind}
 \caption{Estimated second-level effects for kind lemmas in GLMMs for Masculine\slash Neuter and Feminine}
@@
 \centering
-\includegraphics[width=0.9\textwidth]{figures/corpus/04_glmm_fixef_secondlevel_measure}
+\includegraphics[width=0.8\textwidth]{figures/corpus/04_glmm_fixef_secondlevel_measure}
 \caption{Estimated second-level effects for measure lemmas in GLMMs for Masculine\slash Neuter and Feminine}
@@
 
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.6\textwidth]{figures/corpus/04_glmm_fixeff_mn_Measureclass}
-\caption{Effect plot for the masculine\slash neuter GLMM: Measure lemma class effect}
-\label{fig:glmm:fixef:measureclass}
-\end{figure}
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_mn_Attraction}~\includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_fem_Attraction}
-\caption{Effect plot for the two GLMMs: attraction effect (centered and standardised) from the NAC construction with a bare kind noun and the PGC with a determiner; scales are not aligned between left and right panel}
-\label{fig:glmm:fixef:attraction}
-\end{figure}
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_mn_Matchlength}~\includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_fem_Matchlength}
-\caption{Effect plot for the two GLMMs: effect of the (centered and standardised) length of the whole construction; scales are not aligned between left and right panel}
-\label{fig:glmm:fixef:matchlength}
-\end{figure}
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_mn_Measurecase}~\includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_fem_Measurecase}
-\caption{Effect plot for the two GLMMs: effects of the case of the head (kind) noun; scales are not aligned between left and right panel}
-\label{fig:glmm:fixef:measureclass}
-\end{figure}
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_mn_Measurenumber}~\includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_fem_Measurenumber}
-\caption{Effect plot for the two GLMMs: effect of the grammatical number of the head (kind) noun; scales are not aligned between left and right panel}
-\label{fig:glmm:fixef:measurenumber}
-\end{figure}
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_mn_Minus1pos}~\includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_fem_Minus1pos}
-\caption{Effect plot for the two GLMMs: effect of part-of-speech preceding the construction; C = cardinal, D = determiner, O = \textit{other}, A = adjective, P = preposition; scales are not aligned between left and right panel}
-\label{fig:glmm:fixef:minus1pos}
-\end{figure}
-
+%\begin{figure}[h]
+%\centering
+%\includegraphics[width=0.6\textwidth]{figures/corpus/04_glmm_fixeff_mn_Measureclass}
+%\caption{Effect plot for the masculine\slash neuter GLMM: Measure lemma class effect}
+%\label{fig:glmm:fixef:measureclass}
+%\end{figure}
+%
+%\begin{figure}[h]
+%\centering
+%\includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_mn_Attraction}~\includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_fem_Attraction}
+%\caption{Effect plot for the two GLMMs: attraction effect (centered and standardised) from the NAC construction with a bare kind noun and the PGC with a determiner; scales are not aligned between left and right panel}
+%\label{fig:glmm:fixef:attraction}
+%\end{figure}
+%
+%\begin{figure}[h]
+%\centering
+%\includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_mn_Matchlength}~\includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_fem_Matchlength}
+%\caption{Effect plot for the two GLMMs: effect of the (centered and standardised) length of the whole construction; scales are not aligned between left and right panel}
+%\label{fig:glmm:fixef:matchlength}
+%\end{figure}
+%
+%\begin{figure}[h]
+%\centering
+%\includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_mn_Measurecase}~\includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_fem_Measurecase}
+%\caption{Effect plot for the two GLMMs: effects of the case of the head (kind) noun; scales are not aligned between left and right panel}
+%\label{fig:glmm:fixef:measureclass}
+%\end{figure}
+%
+%\begin{figure}[h]
+%\centering
+%\includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_mn_Measurenumber}~\includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_fem_Measurenumber}
+%\caption{Effect plot for the two GLMMs: effect of the grammatical number of the head (kind) noun; scales are not aligned between left and right panel}
+%\label{fig:glmm:fixef:measurenumber}
+%\end{figure}
+%
+%\begin{figure}[h]
+%\centering
+%\includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_mn_Minus1pos}~\includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_fem_Minus1pos}
+%\caption{Effect plot for the two GLMMs: effect of part-of-speech preceding the construction; C = cardinal, D = determiner, O = \textit{other}, A = adjective, P = preposition; scales are not aligned between left and right panel}
+%\label{fig:glmm:fixef:minus1pos}
+%\end{figure}
+%
 %\begin{figure}[h]
@@
 %\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Bayesian estimation}
-\label{sec:bayesian}
 

### 2017-03-20 Latex/article.tex
new big table

@@
 
-Turning to grammaticalisation first, it is often assumed that pseudo-partitives arise as a form of grammaticalised partitives (\eg\ \citealp[536--539]{Koptjevskaja2001} for Finnish and Estonian, \citealp[559]{Koptjevskaja2001} for European languages per se).%
-\footnote{Notice that the description of German in \citet[547--549]{Koptjevskaja2001} is vague and partly incorrect, maybe because her main source \cite{Eschenbach1993} is not as clear as it could be (see also Footnote \ref{fn:eschenbash} above).
-Her example (29b) from \citet[71]{Eschenbach1993}, repeated here verbatim as (\ref{ex:koptjewskaja}) is misclassified as a clear case of a dative on the measure noun and a genitive on the kind noun (hence, PGC).
-
-\begin{exe}
-  \ex{\label{ex:koptjewskaja}\gll trotz einer Flasche gut-en Wein-s\\
-  in.spite.of one-\textsc{fem.gen} bottle good-\textsc{masc.sg.gen} wine-\textsc{gen}\\
-  \glt in spite of one bottle of good wine}
-\end{exe}
-
-It could, however, equally well be a genitive on the measure noun because of the syncretisms discussed here on p.\ \pageref{page:femininesyncretism} and an ambiguous case requirement of the preposition \textit{troz}, hence an instance of NAC.
-Then, she does not clearly distinguish acceptable from non-acceptable constructions (cf.\ Section \ref{sec:descriptive}) on her pp.\ 548 and 549, including a rather confusing \textit{???} rating of her (31b), which is clearly a non-acceptable construction of the type illustrated in (\ref{ex:analyticalpartitive:b}) above.
-}
+Turning to grammaticalisation first, it is often assumed that pseudo-partitives arise as a form of grammaticalised partitives (\eg\ \citealp[536--539]{Koptjevskaja2001} for Finnish and Estonian, \citealp[559]{Koptjevskaja2001} for European languages per se).
+%
+%\footnote{Notice that the description of German in \citet[547--549]{Koptjevskaja2001} is vague and partly incorrect, maybe because her main source \cite{Eschenbach1993} is not as clear as it could be (see also Footnote \ref{fn:eschenbash} above).
+%Her example (29b) from \citet[71]{Eschenbach1993}, repeated here verbatim as (\ref{ex:koptjewskaja}) is misclassified as a clear case of a dative on the measure noun and a genitive on the kind noun (hence, PGC).
+%
+%\begin{exe}
+%  \ex{\label{ex:koptjewskaja}\gll trotz einer Flasche gut-en Wein-s\\
+%  in.spite.of one-\textsc{fem.gen} bottle good-\textsc{masc.sg.gen} wine-\textsc{gen}\\
+%  \glt in spite of one bottle of good wine}
+%\end{exe}
+%
+%It could, however, equally well be a genitive on the measure noun because of the syncretisms discussed here on p.\ \pageref{page:femininesyncretism} and an ambiguous case requirement of the preposition \textit{troz}, hence an instance of NAC.
+%Then, she does not clearly distinguish acceptable from non-acceptable constructions (cf.\ Section \ref{sec:descriptive}) on her pp.\ 548 and 549, including a rather confusing \textit{???} rating of her (31b), which is clearly a non-acceptable construction of the type illustrated in (\ref{ex:analyticalpartitive:b}) above.
+%}
+%
 The alternating constructions in German are both clearly pseudo-partitives, but the grammaticalisation paths uncovered by \cite[esp.\ 526--530]{Koptjevskaja2001}, which I interpret freely here, are clearly relevant for an analysis of the alternation.

### 2017-03-22 Latex/article.tex
progress in statistics section

@@
 \newcommand{\Supsf}[1]{\ensuremath{\mathsf{^{#1}}}}
+
+\newcommand{\NACb}{NAC\Sub{bare}}
+\newcommand{\NACa}{NAC\Sub{adj}}
+\newcommand{\PGCd}{PGC\Sub{det}}
+\newcommand{\PGCa}{PGC\Sub{adj}}
 
@@
   %\accepted{...}
-  \journalname{Submitted to Cognitive Linguistics}
+  \journalname{(submitted)}
   \journalyear{2017}
@@
 Therefore, I introduce and illustrate the relevant constructions in some detail in this section.
-The narrow range of syntactic constellations in which the alternation occurs is described, and the focus on \textit{only} this narrow (range rather than, for example, the whole range of nominal constructions expressing quantities) is motivated.
-This entails that I also have to discuss certain constructions and factors that might be seen as relevant, but which I exclude from my own study.
+I describe the narrow range of syntactic constellations in which the alternation occurs, and I motivate the focus on \textit{only} this narrow (range rather than, for example, the whole range of nominal constructions expressing quantities).
 
@@
 If the kind noun forms an NP with a determiner, the construction resembles (and is usually called) a \textit{pseudo-partitive} (on partitives and pseudo-partitives see, e.\,g., \citealp{Barker1998,Selkirk1977,Stickney2007,Vos1999}; for a recent application of the terminology to German, see \citealp{Gerstenberger2015}).%
-\footnote{Whereas partitives are constructions denoting a proper part-of relation strictly requiring a definite kind noun as in \textit{a sip of the wine}, pseudo partitives -- albeit syntactically similar and diachronically related to partitives in many languages -- merely denote quantities and contain indefinite kind nouns as in \textit{a sip of wine}.
-In the literature on German, some authors incorrectly call the pseudo-partitive a \textit{partitive} \cite{Hentschel1993} while some realise the difference and at least mention it \citep{Eschenbach1994,GallmannLindauer1994,Loebel1989,Zimmer2015}.}
-I use \textdagger\ to denote general unacceptability, and the acceptability judgements given in (\ref{ex:intro:pseudopartitive}) are completely undisputed.
-Notice that the kind noun is obligatorily in the genitive and that case agreement with the measure noun is not an option (but see immediately below).
-I refer to the construction in (\ref{ex:intro:pseudopartitive1}) as the \textit{Pseudo-partitive Genitive Construction} (PGC).
+\footnote{If the kind noun is definite, the construction instantiates a true partitive.
+Whereas partitives are constructions denoting a proper part-of relation strictly requiring a definite kind noun as in \textit{a sip of the wine}, pseudo partitives -- albeit syntactically similar and diachronically related to partitives in many languages -- merely denote quantities and contain indefinite kind nouns as in \textit{a sip of wine}.
+%In the literature on German, some authors incorrectly call the pseudo-partitive a \textit{partitive} \cite{Hentschel1993} while some realise the difference and at least mention it \citep{Eschenbach1994,GallmannLindauer1994,Loebel1989,Zimmer2015}.
+}
+%I use * to denote general unacceptability, and the acceptability judgements given in (\ref{ex:intro:pseudopartitive}) are completely undisputed.
+Here, the kind noun is in the genitive, and I refer to the construction in (\ref{ex:intro:pseudopartitive1}) as the \textit{Pseudo-partitive Genitive Construction} (PGC).
 
@@
     \glt We drink a cup of a tasty coffee.}
-    \ex[\textdagger]{\label{ex:intro:pseudopartitive2} Wir trinken [[eine Tasse]\Sub{Acc} [einen leckeren Kaffee]\Sub{Acc}]\Sub{Acc}.}
+    \ex[*]{\label{ex:intro:pseudopartitive2} Wir trinken [[eine Tasse]\Sub{Acc} [einen leckeren Kaffee]\Sub{Acc}]\Sub{Acc}.}
   \end{xlist}
@@
 
-When the kind noun is bare -- \ie\ when it comes without a determiner and without modifying adjectives -- the kind noun has to agree in case with the measure noun, and the genitive seen in the PGC is not acceptable, see (\ref{ex:intro:narrowapposition}).
+When the kind noun is bare -- \ie, without a determiner and without modifying adjectives -- the kind noun has to agree in case with the measure noun, and the genitive seen in the PGC is not acceptable, see (\ref{ex:intro:narrowapposition}).
 This construction is usually classified as a \textit{Narrow Apposition Construction} \citep{Loebel1986}, henceforth NAC.%
 \footnote{Notice that the unavailability of the genitive on the kind noun can be seen as following from a rather quirky constraint that genitive NPs in German require the presence of some strongly case-marked element (determiner or adjective) in addition to the head noun in order to be acceptable \citep{GallmannLindauer1994,Schachtl1989}.}
+%The construction as in (\ref{ex:intro:narrowapposition2}) is also referred to as the \textit{Direct Partitive Construction} for other Germanic languages in which the PGC with the synthetic genitive is not available.%
+%\footnote{This nomenclature makes sense in contrast to the \textit{Indirect Partitive Construction} with prepositional linkers translating to `of' -- \ie\ analytic genitives -- in such languages, see \cite{HankamerMikkelsen2008} for Danish.
+%For German, this terminology is not distinctive enough, which is why I use the terms NAC and PGC.}
+While only an accusative is shown in (\ref{ex:intro:narrowapposition}), the kind noun obligatory agrees in case also with nominative and dative measure nouns.
 
@@
   \begin{xlist}
-    \ex[\textdagger]{\label{ex:intro:narrowapposition1} Wir trinken [[eine Tasse]\Sub{Acc} [Kaffees]\Sub{Gen}]\Sub{Acc}.}
+    \ex[*]{\label{ex:intro:narrowapposition1} Wir trinken [[eine Tasse]\Sub{Acc} [Kaffees]\Sub{Gen}]\Sub{Acc}.}
     \ex[ ]{\label{ex:intro:narrowapposition2}\gll Wir trinken [[eine Tasse]\Sub{Acc} [Kaffee]\Sub{Acc}]\Sub{Acc}.\\
@@
 
-The construction as in (\ref{ex:intro:narrowapposition2}) is also referred to as the \textit{Direct Partitive Construction} for other Germanic languages in which the PGC with the synthetic genitive is not available.%
-\footnote{This nomenclature makes sense in contrast to the \textit{Indirect Partitive Construction} with prepositional linkers translating to `of' -- \ie\ analytic genitives -- in such languages, see \cite{HankamerMikkelsen2008} for Danish.
-For German, this terminology is not distinctive enough, which is why I use the terms NAC and PGC.}
-While only an accusative is shown in (\ref{ex:intro:narrowapposition}), the kind noun obligatory agrees in case also with nominative and dative measure nouns.
-
 The actual alternation can be observed \textit{only when the kind noun occurs with an attributive adjective but without a determiner}, as in (\ref{ex:intro:alternation}), where both the NAC in (\ref{ex:intro:alternation1}) and the PGC in (\ref{ex:intro:alternation2}) are equally acceptable.
@@
 See \cite{Hentschel1993,Zimmer2015} for analyses of the sometimes absurd stances taken in grammars of German.
-As the usage and experimental data presented below should render it clear, there might be preferences under certain circumstances, but we cannot assume either construction to be unacceptable.}
+%As the usage and experimental data presented below should render it clear, there might be preferences under certain circumstances, but we cannot assume either construction to be unacceptable.
+}
 
@@
 The distribution of the case patterns in the NAC (case identity between the measure noun and the kind noun) and in the PGC (genitive on the kind noun) depending on the structure of the kind NP are summarised in Table~\ref{tab:constructions}.
-I now turn to some more subtle issue related to the measure noun case alternation, namely:
+I call the narrow apposition construction with a bare kind noun the \NACb, the partitive genitive with a determiner in the kind noun phrase the \PGCd.
+I call the alternating variants with an adjective but no determiner in the kind noun phrase \NACa, and \PGCa, respectively.
+
+I now turn to some more subtle issues related to the measure noun case alternation, namely:
 
@@
 \footnote{See Section~\ref{sec:analyses} for more discussion of adjectival inflection.}
-In my corpus sample (released with this article), this tendency was not nearly as clear, and there was a high number of very noisy sentences among those potentially showing this pattern.
+In my corpus sample, this tendency was not nearly as clear, and there was a high number of very noisy sentences among those potentially showing this pattern.
 Hence, I do not discuss these forms much, although I will offer a possible interpretation in Section~\ref{sec:analyses}.
@@
 Whereas the singular masculine and neuter nominal subsystem is a four case system (at least for NPs with adjectives and determiners), feminine singular NPs show nominative-accusative and dative-genitive syncretisms.
-Table~\ref{tab:syncretisms} shows this as well as the partial syncretism in the plural, where no gender distinctions are made (for an overview, see \citealp[XYZ]{Eisenberg2013a}).
+Table~\ref{tab:syncretisms} shows this as well as the partial syncretism in the plural, where no gender distinctions are made.
 This means that with dative measure nouns, the alternation is unobservable if the kind noun is feminine, and it is problematic to speak of a \textit{genitive} kind noun construction in this case.
-This is taken into account in the model design and the discussion presented in Section~\ref{sec:corpusstudies}.
+This is taken into account in the statistical modelling and the discussion presented in Section~\ref{sec:corpusstudies}.
 
@@
 Turning to (\ref{it:intro:plurl}), readers might have noticed that so far only MNPs denoting quantities of substances (mass kind nouns such as \textit{ein Glas roter Wein\slash roten Weines} `a glass of red wine') have been discussed.
-If the kind noun is a plural count noun with a collective denotation such as \textit{ein Sack kleine Äpfel\slash kleiner Äpfel} `a bag of small apples' (see Table~\ref{tab:syncretisms} for the inflection pattern), a similar alternation between PGC and NAC can be observed.%
-\footnote{\label{fn:eschenbash}On the collective interpretation of such constructions, see \cite{Eschenbach1994}.
-It is, however, unfortunate that Eschenbach does not notice and discuss the morpho-syntactic distribution of PGC and NAC, even when she gets very close to listing the relevant data \citep[217]{Eschenbach1994}.}
-In line with experimental results reported in \citet[15--16]{Zimmer2015}, I found that the PGC is so dominant with plural kind nouns (over 92\%, cf.\ Section~\ref{sec:corpusstudies}) that the alternation cannot be analysed in the same way as in the singular case.
-While this will play a role in the interpretation of the corpus findings, MNPs with plural\slash collective kind nouns will not be dealt with prominently in the remainder of the paper.
-A statistical model for plural kind nouns will be reported in passing in Section~\ref{sec:corpusstudies}, however.
+If the kind noun is a plural count noun as in \textit{ein Sack kleine Äpfel\slash kleiner Äpfel} `a bag of small apples' (see Table~\ref{tab:syncretisms} for the inflection pattern), a similar alternation between PGC and NAC can be observed.
+%\footnote{\label{fn:eschenbash}On the collective interpretation of such constructions, see \cite{Eschenbach1994}.
+%It is, however, unfortunate that Eschenbach does not notice and discuss the morpho-syntactic distribution of PGC and NAC, even when she gets very close to listing the relevant data \citep[217]{Eschenbach1994}.}
+In line with experimental results reported in \citet[15--16]{Zimmer2015}, I found that the PGC is so dominant with plural kind nouns (67 of 861 cases or over 92\%, cf.\ Section~\ref{sec:corpusstudies}) that the alternation cannot be analysed in the same way as in the singular case.
+While this will play a role in the interpretation of the corpus findings, MNPs with plural ve kind nouns will not be dealt with prominently in the remainder of the paper.
+%A statistical model for plural kind nouns will be reported in passing in Section~\ref{sec:corpusstudies}, however.
 
-Turning to issue (\ref{it:intro:noifl}), some measure nouns have been grammaticalised in a way that they always appear non-inflected.
+As for (\ref{it:intro:noifl}), some measure nouns have been grammaticalised in a way that they always appear non-inflected.
 They are typical measure nouns like \textit{Gramm} `gram', \textit{Pfund} `pound' or \textit{Prozent} `percent', which do not have plural forms at all.%
@@
 The non-inflected variant denotes the quantity of wine fitting into two glasses, the inflected variant denotes two glasses filled with wine.}
-I treat these cases like any other measure noun because they enter into both the NAC as in (\ref{ex:uninflectedmeasures:a}) and the PGC as in (\ref{ex:uninflectedmeasures:b}).
-As will be explained in Section~\ref{sec:analyses}, they are expected to exert a specific influence on the choice of NAC and PGC.
+I treat these cases like any other measure noun because they enter into both the \NACa\ as in (\ref{ex:uninflectedmeasures:a}) and the \PGCa\ as in (\ref{ex:uninflectedmeasures:b}).
+In Section~\ref{sec:analyses}, degrees of grammaticalisation as a factor influencing the alternation will be discussed, however.
 
@@
 Finally, (\ref{it:intro:other}) suggests that there are alternative ways of expressing similar quantificational meanings.
-First of all, the analytic (pseudo-)partitive with \textit{von} `of' is only available as an alternative to the PGC if the kind noun phrase contains a (definite or indefinite) determiner as in (\ref{ex:analyticalpartitive}).
-
-\begin{exe}
-  \ex\label{ex:analyticalpartitive} 
-  \begin{xlist}
-    \ex[ ]{\label{ex:analyticalpartitive:a}\gll ein Glas von dem roten Wein\\
-                                                a glass of the red wine\\
-                                                \glt a glass of the red wine
-       }
-    \ex[\textdagger]{\label{ex:analyticalpartitive:b}\gll ein Glas von rotem Wein\\
-                                                a glass of red wine\\
-                                                \glt a glass of red wine
-       }
-  \end{xlist}
-\end{exe}
-
-However, if the measure noun denotes a container or vessel, constructions with \textit{voll} (\textit{von})\slash\textit{voller} `full (of)' and \textit{mit} `with' are available, see (\ref{ex:alternatives}).
+%First of all, the analytic (pseudo-)partitive with \textit{von} `of' is only available as an alternative to the PGC if the kind noun phrase contains a (definite or indefinite) determiner as in (\ref{ex:analyticalpartitive}).
+%
+%\begin{exe}
+%  \ex\label{ex:analyticalpartitive} 
+%  \begin{xlist}
+%    \ex[ ]{\label{ex:analyticalpartitive:a}\gll ein Glas von dem roten Wein\\
+%                                                a glass of the red wine\\
+%                                                \glt a glass of the red wine
+%       }
+%    \ex[*]{\label{ex:analyticalpartitive:b}\gll ein Glas von rotem Wein\\
+%                                                a glass of red wine\\
+%                                                \glt a glass of red wine
+%       }
+%  \end{xlist}
+%\end{exe}
+%
+If the measure noun denotes a container or vessel, constructions with \textit{voll} (\textit{von})\slash\textit{voller} `full (of)' and \textit{mit} `with' are available, see (\ref{ex:alternatives}).
 
@@
 
-These construction have very idiosyncratic properties (the construction with \textit{voller} is discussed in \citealp{Zeldes2018}, the construction with \textit{mit} is discussed in \citealp{Bhatt1990}), they are not semantically equivalent to NAC and PGC, and -- most importantly -- they are only available for a small subset of measure nouns.
+These construction have very idiosyncratic properties (the construction with \textit{voller} is discussed in \citealp{Zeldes2018}, the construction with \textit{mit} is discussed in \citealp{Bhatt1990}), they are not semantically equivalent to the NAC and the PGC, and -- most importantly -- they are only available for a small subset of measure nouns.
 Subtle interactions between these constructions and NAC\slash PGC cannot be excluded, especially because the former might act as a (partial) alternative and thus reduce the count of exemplars of the NAC, the PGC, or both for container measure nouns.
-There are no clear and motivated hypotheses about such interactions, however.
-Nevertheless, the corpus-based models in Section~\ref{sec:corpusstudies} are specified in a way that they could detect preferences for either NAC or PGC for the class of container or vessel measure nouns, for which the alternative constructions illustrated in (\ref{ex:alternatives}) is available.
+There are no clear hypotheses about such interactions, however.
+Nevertheless, the corpus-based models in Section~\ref{sec:corpusstudies} are specified in a way that they could detect preferences for either NAC or PGC for the class of container or vessel measure nouns, for which the alternative constructions illustrated in (\ref{ex:alternatives}) are available.
 
-This concludes the descriptive overview of the alternation.
+This concludes the descriptive overview of the phenomenon.
 I have shown that there is an alternation between two measure noun constructions in a narrow syntactic configuration (kind NP with an adjective but without a determiner), and that the two constructions differ in the case of the kind noun (case agreement with the measure noun or genitive).
@@
 
-\subsection{A theoretical assessment}
+\subsection{Theoretical assessment}
 \label{sec:analyses}
 
-This section reviews some existing analyses of the PGC-NAC alternation or related issues, and I develop my own analysis and the appropriate hypotheses for the empirical studies presented in Sections~\ref{sec:corpusstudies} and \ref{sec:externalvalidation}.
-While the constructions and the alternation are standardly covered in descriptive grammars such as \citet[XYZ]{Eisenberg2013a}, the research literature contains little specific discussion of the situation in German.
-In The remainder of this section, I discuss as factors that might influence the alternation:
-degrees of grammaticalisation of the measure noun, the different syntactic structures of the variants and their syntactic prototypes, and potential effects of semantic prototypicality.
+This section briefly reviews existing analyses of the \PGCa-\NACa\ alternation and related issues.
+I also develop my own analysis and the appropriate hypotheses for the empirical studies presented in Sections~\ref{sec:corpusstudies} and \ref{sec:externalvalidation}.
+%While the constructions and the alternation are standardly covered in descriptive grammars such as \citet[XYZ]{Eisenberg2013a}, the research literature contains little specific discussion of the situation in German.
+I discuss four main factors that might influence the alternation:
+the different syntactic structures of the variants \NACa\ and \PGCa\ and their syntactic prototypes \NACb\ and \PGCd,
+degrees of grammaticalisation of the measure noun associated with the prototypes,
+a semantic prototypicality effect of the degrees of grammaticalisation (plurality),
+register.
 
-Turning to grammaticalisation first, it is often assumed that pseudo-partitives arise as a form of grammaticalised partitives (\eg\ \citealp[536--539]{Koptjevskaja2001} for Finnish and Estonian, \citealp[559]{Koptjevskaja2001} for European languages per se).
+First, I discuss the analysis of the morpho-syntax of German MNPs.
+A major interest in descriptive and generative linguistics is to find a plausible analysis of the hierarchical syntactic structure of the constructions in question.
+The first question is which noun constitutes the head of the whole construction.
+This question was answered quite clearly by \cite{Loebel1986} already (see also \citealp[213]{Eschenbach1994}, and \citealp[16]{GallmannLindauer1994}).
+Subject-verb agreement is always realised on the measure noun, and we can therefore assume that the measure noun is the head.
+However, the MNP-internal structure has more interesting cues to offer if the strong\slash weak inflection patterns of adjectives already mentioned in Section~\ref{sec:descriptive} are taken into account.
+In NPs with a strongly inflected determiner, attributive adjectives inflect according to the massively syncretistic \textit{weak} pattern.
+%\footnote{The weak pattern has only two suffixes.
+%The structural cases (nominative and accusative) in the singular are marked with \textit{-e}, the oblique cases (dative and genitive) as well as the entire plural are marked with \textit{-en}.
+%The only exception to this rule is the masculine accusative singular, which is also marked with \textit{-en}.}
+If there is no determiner (as is the case in the alternating constructions), attributive adjectives inflect like determiners, which is called the \textit{strong} pattern (see Table~\ref{tab:syncretisms}).
+Thus, the adjectives in the \NACa\ and the \PGCa\ have properties of adjectives as well as determiners.
+They are lexical adjectives and function as attributive modifiers.
+On the other hand, they are inflected like determiners, and they are the leftmost element in the NP, which is typical of determiners.
+This unusual double nature of adjectives in NPs without determiners leads to a plausible probabilistic interpretation of the pattern shown in Table~\ref{tab:constructions}.
+Whenever speakers classify the adjective in the kind noun phrase more as a determiner, they \textit{have} to use the \PGCa\ because, if there is a determiner, the PGC (just like in the case of the \PGCd) is the only option.
+When they classify the adjective more as an adjective, the kind NP has no determiner, and they \textit{have} to use the \NACa\ (like in the case of the \NACb).%
+\footnote{While the generative analysis presented in \cite{Bhatt1990} cannot properly deal with probabilistic effects, Bhatt comes close to this interpretation by analysing the kind NP in the GPC as a DP and in the NAC as an NP.}
+In a framework that assumes similarity-based probabilistic classification, such as Prototype Theory, we can regard the \PGCd\ as instantiating the PGC prototype and the \NACb\ as the NAC prototype.
+The choice between \PGCa\ and \NACa\ is then expected to be influenced by the degree of similarity to either of these prototypes.
+
+Any attempt to substantiate or test this hypothesis is rendered difficult by the fact that the semantic differences between the variants are at best subtle, which means that the prototypes are not associated with clearly detectable differences in meaning.
+However, if there is a semantic difference, kind and measure noun lemmas (and maybe semantically of functionally defined classes thereof) could be attracted more strongly to one of the prototypes.
+As a result, kind noun lemmas and measure noun lemmas probably appear with different relative frequencies in the prototypical \NACb\ and the prototypical \PGCd.
+Their relative frequencies in these non-alternating configurations should then be indicative of their preference to appear in one of the two variants.
+This predicted effect will be quantified in Section~\ref{sec:corpusstudies}.
+
+%Related to the , dative measure nouns favor the PGC more than nominative and accusative (also \citealp{Hentschel1993,Zimmer2015}).
+%As an example, \textit{mit einem Stück frischen Brots} `with a piece of fresh bread' (PGC) would be preferred over \textit{mit einem Stück frischem Brot} (NAC).
+%I propose
+%
+
+Turning to grammaticalisation now, it is often assumed that pseudo-partitives arise as a form of grammaticalised partitives (\eg\ \citealp[536--539]{Koptjevskaja2001} for Finnish and Estonian, \citealp[559]{Koptjevskaja2001} for European languages per se).
 %
@@
 %
-The alternating constructions in German are both clearly pseudo-partitives, but the grammaticalisation paths uncovered by \cite[esp.\ 526--530]{Koptjevskaja2001}, which I interpret freely here, are clearly relevant for an analysis of the alternation.
-The grammaticalisation path can start out with constructions involving two referential nouns (not necessarily forming a single NP) and a \textit{separative} meaning as in \textit{(cut) two slices from the cake} \citep[535]{Koptjevskaja2001}.
-The \textit{part-of} meaning of true partitives as in \textit{a slice of the case} represents the first stage of a development wherein the measure noun already tends to loose some semantic content.
+The alternating constructions in German are both clearly pseudo-partitives, but the grammaticalisation paths uncovered by \cite[esp.\ 526--530]{Koptjevskaja2001} -- which I interpret freely here -- are clearly relevant for an analysis of the alternation.
+The grammaticalisation path can start out (in some languages) with constructions involving two referential nouns (not necessarily forming a single NP) and a \textit{separative} meaning as in \textit{(cut) two slices from the cake} \citep[535]{Koptjevskaja2001}.
+The \textit{part-of} meaning of true partitives as in \textit{a slice of the cake} represents the first stage of a development wherein the measure noun already tends to loose some semantic content.
 The pseudo-partitive stage finally instantiates a \textit{quantity-of} relation, potentially even leading to fully grammaticalised quantifiers such as \textit{a lot}.
-In German, the PGC is clearly the older variant \citep{Zimmer2015} with an overt \textit{quantity-of} marker in the form of the genitive suffix, and at least the potential to form a true partitive (if the kind noun is definite), which the NAC lacks.
+In German, the PGC is clearly the older variant \citep{Zimmer2015} with an overt \textit{quantity-of} marker in the form of the genitive suffix.
+Also, it still has the potential to form a true partitive (if the kind noun is definite).
+Conversely, the NAC completely lacks this ability to form true partitives.
 Hence, we can expect more strongly grammaticalised measure nouns to prefer the NAC.
-Since clearly non-referential nouns like \textit{Gramm} `gram' and \textit{Meter} `metre' should favor the NAC according to this hypothesis, the effect will be detected by including information about measure noun classes in the corpus-based models in Section~\ref{sec:corpusstudies}.
+For example, strongly grammaticalised non-referential nouns like \textit{Gramm} `gram' and \textit{Meter} `metre' should occur proportionally more often in the \NACa\ than in the \PGCa.
 
-On the other hand, there are case studies like \cite{Brems2003,DeclerckBrems2016} where syntactic evidence for grammaticalisation plays a prominent role.
+The question arises whether there might other cues to the degree of grammaticalisation of individual measure nouns or classes of them.
+In case studies like \cite{Brems2003,DeclerckBrems2016} syntactic evidence for grammaticalisation plays a prominent role.
 In English -- \cite{Brems2003} argues based on a view from \cite{Langacker1991} -- strongly lexicalised and thus semantically bleached measure nouns turn from nouns into quantifiers, leading to a syntactic reanalysis of the NP, wherein the kind noun becomes the head.
 The argument is supported by convincing syntactic evidence, for example subject-verb agreement.
-In German MNPs, the measure noun always remains the syntactic head, however (see immediately below for references), and there is no similar clear criterion to detect grammaticalisation.
-The analysis of the morpho-syntax of German MNPs, to which I turn now, will provide hypothetical cues along similar lines, however.
+As argued above, in German MNPs the measure noun always remains the syntactic head, however, and there is no similar clear criterion to detect grammaticalisation.
+Therefore, in the analysis of the corpus finding in Section~\ref{sec:corpusstudies}, effects of different degrees of grammaticalisation will be taken into account by interpreting lemma-specific and lemma class-specific effects.
 
-A major interest in descriptive and generative linguistics is to find a plausible analysis of the hierarchical syntactic structure of the constructions in question.
-In the case at hand, the first question is which noun constitutes the head of the whole construction.
-This question was answered quite clearly by \cite{Loebel1986} already (see also \citealp[213]{Eschenbach1994}, and \citealp[16]{GallmannLindauer1994}).
-As hinted above, subject-verb agreement is always realised on the measure noun, and we can therefore assume that the measure noun is the head.
-
-I therefore turn to the MNP-internal syntax, which will show itself to be highly revealing.
-It cannot be understood properly without taking into account the strong\slash weak inflection patterns of adjectives briefly described in Section~\ref{sec:descriptive} already.
-In NPs with a strongly inflected determiner, attributive adjectives inflect according to the massively syncretistic \textit{weak} pattern.%
-\footnote{The weak pattern has only two suffixes.
-The structural cases (nominative and accusative) in the singular are marked with \textit{-e}, the oblique cases (dative and genitive) as well as the entire plural are marked with \textit{-en}.
-The only exception to this rule is the masculine accusative singular, which is also marked with \textit{-en}.}
-If there is no determiner (as is the case in the alternating constructions), attributive adjectives inflect like determiners, which is called the \textit{strong} pattern (see Table~\ref{tab:syncretisms}).
-Thus, the adjectives in the target constructions have properties of adjectives as well as determiners.
-They are lexical adjectives and function as attributive modifiers, but they are inflected as determiners and are the leftmost element in the NP, which is typical of determiners.
-This unusual double nature of adjectives in determinerless NPs leads to a plausible interpretation of the pattern shown in Table~\ref{tab:constructions}.
-Whenever speakers classify the adjective in the kind noun phrase as a determiner, they \textit{have} to use the PGC because, if there is a determiner, the PGC is the only option.
-When they classify the adjective as an adjective, the kind NP has no determiner, and they \textit{have} to use the NAC.%
-\footnote{While the generative analysis presented in \cite{Bhatt1990} cannot properly deal with probabilistic effects, Bhatt comes close to this interpretation by analysing the kind NP in the GPC as a DP and in the NAC as an NP.
-Within the confines of her framework, this is the appropriate was of encoding the relevant generalisation.}
-Now, in a framework that assumes similarity-based probabilistic classification, such as Prototype Theory, we can regard the PGC with a determiner in the kind noun phrase as instantiating the PGC prototype and the NAC with just a bare kind noun as the NAC prototype.
-The choice of the alternating variants is then expected to be influenced by the degree of similarity to either of these prototypes.
-
-Any attempt to substantiate or test this hypothesis is rendered difficult by the fact that the semantic differences between the variants are at best subtle, which means the prototypes are not associated with clearly detectable differences in meaning.
-However, measure noun lemmas all have individual and group-wise lexical-semantic properties which make them more appropriate for one prototype or the other.
-As a result, kind noun lemmas and measure noun lemmas probably appear with different relative frequencies in the prototypical NAC (with a bare kind noun) and the prototypical PGC (with a determiner inside the kind noun phrase).
-Their relative frequencies in these non-alternating configurations should be indicative of their preference to appear in one of the two variants.
-In other words, lemmas themselves might have specific degrees of affinity to prototype or the other, and this affinity, measured by looking at the non-alternating constructions, should be predictive of their frequencies in the alternating cases.
-
-%Related to the , dative measure nouns favor the PGC more than nominative and accusative (also \citealp{Hentschel1993,Zimmer2015}).
-%As an example, \textit{mit einem Stück frischen Brots} `with a piece of fresh bread' (PGC) would be preferred over \textit{mit einem Stück frischem Brot} (NAC).
-%I propose
-%
-
-Furthermore, the grammaticalisation path as described above leads from NPs denoting highly individuated objects standing in a \textit{separated-from} relation to a construction with a \textit{part-of} meaning (still available in the prototypical PGC) to a construction with a more diffuse \textit{quantity-of} relation (characteristic of the prototypical NAC).
-The best evidence for the relevance of this factor in an ongoing grammaticalisation process is that plural kind nouns almost always (but not exclusively) trigger PGC and virtually never NAC (Section~\ref{sec:descriptive}).
-More indirectly, plurality of the measure noun could also favor the PGC over the NAC, and exact quantification in the form of cardinals even more so.%
-\footnote{That the plural favors PGC was empirically shown but not theoretically motivated in \cite{Hentschel1993}.}
-As a highly illustrative example (using a measure noun which has virtually developed two variants, one of which grammaticalsed to a much higher degree), \textit{ein Haufen französische Watte} `a heap\slash lot of French cotton' (NAC) has essentially two uses.
+Furthermore, the grammaticalisation path as described above leads from NPs denoting individuated objects standing in a \textit{part-of} relation to a construction with a more diffuse \textit{quantity-of} relation.
+The best evidence for the relevance of this factor in an ongoing grammaticalisation process is that plural kind nouns almost always (but not exclusively) occur in the \PGCa\ and virtually never in the \NACa\ (see Section~\ref{sec:descriptive}).
+More indirectly, however, plurality of the measure noun could also favor the \PGCa\ over the \NACa, and exact quantification in the form of cardinals even more so.%
+\footnote{This preference was observed empirically by \cite{Hentschel1993}.}
+I provide a highly illustrative example using a measure noun which has virtually developed two variants, one of which grammaticalised to a much higher degree, namely \textit{Haufen} `heap\slash lot'.
+The phrase \textit{ein Haufen französische Watte} `a heap\slash lot of French cotton' (\NACa) has essentially two uses.
 It can be used to denote an unspecific amount of cotton in the sense of \textit{a lot of French cotton} in situations when the exact quantity of cotton is not of interest.
 Under the appropriate circumstances, it could also be used to denote a highly specific quantity of cotton, for example when a cotton merchant always sells her French cotton in \textit{heaps} weighing two stone each.
-The former use, however, is highly restricted for \textit{zwei Haufen französischer Watte} `two heaps of French cotton' (GPC) -- not `two lots of French cotton'.
+The former use, however, is highly restricted for \textit{zwei Haufen französischer Watte} `two heaps of French cotton' (\PGCa) -- not `two lots of French cotton'.
 Such an NP is less likely to be used in situations where the exact quantity of cotton is irrelevant, even if there are actually two heaps of cotton.
-In other words, a precisely quantified measure noun leads to a higher individuation (the creation of discrete portions) of the substance denoted by the kind noun, which is more appropriate for the PGC prototype.%
-\footnote{In the extreme, some measure nouns cannot be pluralised, such as \textit{eine Menge} `a lot of'.
-Also note that the presentation of the examples in NAC and PGC is suggestive.
-Native speakers will confirm that both NPs sound highly acceptable in both GPC and NAC.
-However, it was made clear from the beginning and throughout that the preferences are indeed subtle.}
+In other words, a precisely quantified measure noun leads to a higher individuation (the creation of discrete portions) of the substance denoted by the kind noun, which is more appropriate for the PGC prototype.
+The statistical models in Section~\ref{sec:corpusstudies} will be designed to detect such a subtle effect, if it exists.
+%\footnote{In the extreme, some measure nouns cannot be pluralised, such as \textit{eine Menge} `a lot of'.
+%Also note that the presentation of the examples in NAC and PGC is suggestive.
+%Native speakers will confirm that both NPs sound highly acceptable in both GPC and NAC.
+%However, it was made clear from the beginning and throughout that the preferences are indeed subtle.}
 
 Finally, it has been iterated very often that the PGC is more typical of higher registers or even exclusive to written language (see \citealp[320--323]{Hentschel1993}, where she also refers to descriptive grammars).
-This is not surprising in as much as the genitive -- an intrinsic part of the PGC -- is generally underrepresented in colloquial and vernacular variants of German, and thus overrepresented to the same degree as colloquial, vernacular, and spoken variants are overrepresented in most corpora of written German.
-Under the integral view outlined in Section~\ref{sec:cogocl}, such preferences are assumed to be part of the construction prototypes.
+This is not surprising in as much as the genitive -- an intrinsic part of the PGC -- is generally underrepresented in colloquial and vernacular variants of German.
+Under an integral view, such preferences can be to be part of the construction prototypes.
 In the corpus study, the alternation tendencies controlled by this complex factor related to register, medium, and lectal variation will be tested using certain very simple proxies.
@@
 \begin{enumerate}[i.]
-  \item Classes of more clearly grammaticalised measure nouns favor the NAC.
   \item The relative frequencies with which measure noun lemmas and kind noun lemmas appear in the prototypical (unequivocal) PGC and NAC are positively correlated with the relative frequency the PGC and the NAC are chosen in the alternating case.
   \item MNPs with masculine dative kind nouns disfavour the GPC less.
+  \item Classes of more clearly grammaticalised measure nouns favor the NAC.
   \item Plural measure nouns and measure nouns quantified by cardinals favor the PGC variant.
@@
 
-For the present study, I used the German \textit{Corpus from the Web} (COW) in its 2014 version DECOW14A (\citealp{SchaeferBildhauer2012full,Schaefer2015b} and \citealp{BiemannEa2013,SchaeferBildhauer2013} for overviews of web corpora in general and the methodology of their construction), which contains 21 billion tokens.
-I will simply refer to it as DECOW in the remainder of the paper except where clarification is necessary.
-I chose this corpus for two main reasons.%
+For the present study, I used the German \textit{Corpus from the Web} (COW) in its 2014 version DECOW14A (\citealp{SchaeferBildhauer2012full,Schaefer2015b} and \citealp{BiemannEa2013,SchaeferBildhauer2013} for overviews of web corpora in general and the methodology of their construction), which contains almost 21 billion tokens.%
 \footnote{The corpora are made available for free at \url{https://www.webcorpora.org}.
+I chose this corpus for two main reasons.
 At the time of this writing, a newer 2016 version DECOW16 had already been released.}
@@
 \footnote{DECOW14A01 is the first slice (roughly a twentieth) of the complete DECOW14A corpus.
-It contains just over one billion tokens.
-The creators assigned the documents randomly to the slices, such that each slice can be considered representative of the whole corpus \cite{Schaefer2015b}.}
-For this step, I relied on the noun part-of-speech label because for frequent nouns, high tagging accuracy can be expected, especially given the fact that German nouns are always spelled with an initial capital letter.
-Abstract nouns which partially behave like mass nouns (like \textit{Spaß} `fun’ or \textit{Gefahr} `danger’) were excluded because they are usually not quantified in the same way as concrete mass nouns.%
-\footnote{They behave like mass nouns mostly in that they occur without a determiner in the singular.}
+It contains just over one billion tokens.}
+Abstract nouns which partially behave like mass nouns (like \textit{Spaß} `fun’ or \textit{Gefahr} `danger’) were excluded because they are usually not quantified in the same way as concrete mass nouns.
+%\footnote{They behave like mass nouns mostly in that they occur without a determiner in the singular.}
 The hundredth selected mass noun was \textit{Schmuck} `jewelry’, which is the 3,054th most frequent noun in the original frequency list.
@@
 
-In step (\ref{enum:data:step3}), each of these 2,365 noun–noun combinations was queried in the target constructions (PGC and NAC with just an adjective) individually in each of the first ten slices of DECOW (roughly 10 billion tokens).
-The scripts generating the actual queries were crafted so as to make sure that adjectives and nouns were searched in the appropriate forms.
-The CQL query in (\ref{ex:data:query01}) shows an example query for a neuter mass noun like \textit{Wasser} `water' which finds relevant sequences such as the ones shown in (\ref{ex:data:query02}).%
-\footnote{The COW creators used to make corpora available in the IMS Open Corpus Query Workbench \citep{EvertHardie2011}, which has a query syntax informally referred to as CQL.
-Documentation can be obtained at \url{http://cwb.sourceforge.net}.}
-The versions with the genitive singular \textit{Liters} and the dative plural \textit{Litern} are not shown in (2) for space reasons but are also found by the query in (1).
-In terms of context, single sentences containing the target were exported.
-
-\begin{exe}
-  \ex{\label{ex:data:query01} \texttt{\small[ lemma="Liter" ] [ pos="ADJA" \& word=".+(en|em|es)" ] [ lemma="Wasser" ]}}
-  \ex\label{ex:data:query02}
-  \begin{xlist}
-    \ex{\label{ex:data:query02a} Liter [klaren Wassers]\Sub{Gen}}
-    \ex{\label{ex:data:query02b} Liter [klares Wasser]\Sub{Nom\slash Acc}}
-    \ex{\label{ex:data:query02c} Liter [klarem Wasser]\Sub{Dat}}
-  \end{xlist}
-\end{exe}
-
-To reduce the sample size for the manual annotation process, the final concordance was sampled from the results of these 2,365 queries.
+In step (\ref{enum:data:step3}), each of these 2,365 noun–noun combinations was queried in the target constructions (\PGCa\ and \NACa) individually in each of the first ten slices of DECOW (roughly 10 billion tokens).
+%The scripts generating the actual queries were crafted so as to make sure that adjectives and nouns were searched in the appropriate forms.
+%The CQL query in (\ref{ex:data:query01}) shows an example query for a neuter mass noun like \textit{Wasser} `water' which finds relevant sequences such as the ones shown in (\ref{ex:data:query02}).%
+%\footnote{The COW creators used to make corpora available in the IMS Open Corpus Query Workbench \citep{EvertHardie2011}, which has a query syntax informally referred to as CQL.
+%Documentation can be obtained at \url{http://cwb.sourceforge.net}.}
+%The versions with the genitive singular \textit{Liters} and the dative plural \textit{Litern} are not shown in (2) for space reasons but are also found by the query in (1).
+%In terms of context, single sentences containing the target were exported.
+%
+%\begin{exe}
+%  \ex{\label{ex:data:query01} \texttt{\small[ lemma="Liter" ] [ pos="ADJA" \& word=".+(en|em|es)" ] [ lemma="Wasser" ]}}
+%  \ex\label{ex:data:query02}
+%  \begin{xlist}
+%    \ex{\label{ex:data:query02a} Liter [klaren Wassers]\Sub{Gen}}
+%    \ex{\label{ex:data:query02b} Liter [klares Wasser]\Sub{Nom\slash Acc}}
+%    \ex{\label{ex:data:query02c} Liter [klarem Wasser]\Sub{Dat}}
+%  \end{xlist}
+%\end{exe}
+In order to reduce the sample size for the manual annotation process, the final concordance was sampled from the results of these 2,365 queries.
 Since the mass nouns in the sample were distributed according to the usual power law, I used all hits for mass nouns occurring less than one hundred times, but randomly sampled one hundred hits for each mass noun that occurred one hundred or more times.
-The final sample contained 7,547 sentences.
+The final sample contained 6,843 sentences.
 Given the careful bootstrapping and sampling procedure described in this section, we can be highly sure that it contains all relevant and reasonably frequent noun–noun combinations in the target constructions.%
-\footnote{In a similar fashion, the 100 most frequent measure nouns occurring with plural kind nouns were bootstrapped and queried.
-Bootstrapping measure nouns first was more efficient in the plural for reasons to complicated to discuss here, especially because the data was not properly analysed anyway.
-As stated in Section~\ref{sec:germanmeasurenps}, the NAC is virtually never used with plural kind nouns.
-Consequently, it does not make sense to model the influencing factors in the same way as for singular kind nouns.}
+\footnote{In a similar fashion, the 100 most frequent measure nouns occurring with plural kind nouns were bootstrapped and queried, resulting in a sample of 1,876 sentences.
+%Bootstrapping measure nouns first was more efficient in the plural for reasons to complicated to discuss here, especially because the data was not properly analysed anyway.
+As stated in Section~\ref{sec:germanmeasurenps}, the NAC is virtually never used with plural kind nouns, and this sample was not used for full analysis.
+%Consequently, it does not make sense to model the influencing factors in the same way as for singular kind nouns.
+}
 
 Finally, two auxiliary samples were also drawn.
-As mentioned in Section~\ref{sec:analyses}, the distribution of the measure noun and kind noun lemmas in the bare-noun NAC and the PGC with a determiner will be modelled as factors influencing the alternation.
-Therefore, all noun-noun pairs from the bootstrap process were also queried in the two non-alternating constructions, resulting in 17,252 hits for the PGC with a determiner and 315,635 hits for the bar-noun NAC.
-Notice that the PGC with a determiner accounts for a mere 5.2\% of all non-alternating MNPs.
-
-In the next section, I describe the annotation process for the main sample.
+As mentioned in Section~\ref{sec:analyses}, the distribution of the measure noun and kind noun lemmas in the \NACb\ and the \PGCd\ with a determiner will be modelled as factors influencing the alternation.
+Therefore, all noun-noun pairs from the bootstrap process were also queried in the two non-alternating constructions, resulting in 17,252 hits for the \PGCd\ and 315,635 hits for the \NACb.
 
@@
 
-A mixture of automatic pre-annotation based on the annotations available in DECOW and manual annotations was used.
-Importantly, the case of the measure noun was annotated completely by hand because case syncretisms make automatic case annotations unreliable. 
-The full set of variables is given in Table~\ref{tab:variables}, and I briefly discuss them now.
+%A mixture of automatic pre-annotation based on the annotations available in DECOW and manual annotations was used.
+%Importantly, the case of the measure noun was annotated completely by hand because case syncretisms make automatic case annotations unreliable. 
+The full set of manually annotated variables for the main sample is given in Table~\ref{tab:variables}, and I briefly discuss them now.
 Notice that \textit{Genitive} is the boolean (dichotomous) response variable (`dependent variable') which is 1 when the PGC is chosen in a given sentence, and 0 when the NAC is chosen.
@@
 If the PGC actually favors higher registers and styles, a high \textit{Badness} should be correlated with fewer occurrences of the PGC.
-
 Documents in DECOW14 have also been annotated with a variable called \textit{Genitives}.
 The higher the values of this variable, the lower the proportion of genitives among all case-bearing forms is.
-While more genitives are also indicative of higher registers and styles, this variable might be considered problematic in the present study.
-Since the PGC contains a genitive itself, the regressor variable \textit{Genitive} and the document-level variable \textit{Genitives} are not independent.
+While more genitives are also indicative of higher registers, the use of this variable as a regressor in the present study might be considered problematic.
+Since the PGC contains a genitive itself, the regressor variable \textit{Genitive} and the document-level variable \textit{Genitives} are not fully independent.
 However, since the PGCs make up for only a minute fraction of all genitives, I still use \textit{Genitives} as a regressor with the appropriate caveats.
-
-Since word length is also a good indicator of higher registers and styles, the length of the match was added as another proxy.%
+Since word length is also a good indicator of higher registers and styles, the length of the match in characters (for example 17 for \textit{Kilo tolle Schoki} `a kilo of nice choc' and 49 for \textit{ein Kilogramm vorzüglicher Edelbitter-Schokolade} ` a kilogram of excellent fine dark chocolate') was added as another proxy to register.%
 \footnote{The new version of the corpus contains over 100 document-level variables in the spirit of \cite{Biber1988}.
-Unfortunately, these annotations were not available when my samples were taken, and somewhat more crude proxies like the length of the query match had to be used.}
+Unfortunately, these annotations were not available when my samples were taken, and even cruder proxies to register than Biberian features (like the length of the query match) had to be used.
+All in all, all variables used as indicators of register here have to be interpreted with extreme caution.
+}
 Furthermore, abbreviated measure nouns like \textit{kg} or \textit{m} might be indicative of less elaborate technical styles, and hence the variable \textit{Measureabbreviated} was included, which is 1 if the measure noun is abbreviated.
@@
 The level names stand for \textit{adjective}, \textit{cardinal}, \textit{determiner}, \textit{other}, and \textit{preposition} (in the same order as in Table~\ref{tab:variables}).
-The purpose of this variable is mainly to see whether cardinals really favor the PGC as hypothesised in Section~\ref{sec:analyses}.
+The purpose of this variable is mainly to see whether cardinals really favor the \PGCa\ as hypothesised in Section~\ref{sec:analyses}.
 
-The variables \textit{Kindattraction} and \textit{Measureattraction} encode the ratio with which a given kind noun lemma or measure noun lemma occurs in the bare-noun NAC and the PGC with a determiner.
-They were calculated from the auxiliary samples described at the end of Section~\ref{sec:gettingdata}.
+The variables \textit{Kindattraction} and \textit{Measureattraction} encode the ratio with which a given kind noun lemma or measure noun lemma occurs in the \NACb\ and the \PGCd.
+They were calculated from the auxiliary samples described at the end of Section~\ref{sec:gettingdata} as a quotient with one added to numerator and denominator to avoid zero divisions.
 The higher the value, the more often the noun occurs in the PGC with a determiner (proportionally).
@@
 In Section~\ref{sec:analyses}, it was hypothesised that classes of lemmas might have different preferences for the two variants.
+To capture this, class information was annotated for measure and kind lemmas.
 For the kind nouns, it must be admitted that semantic classes were defined ad hoc based on intuitions of colleagues and native speakers, especially because classes available in lexical-semantic databases like GermaNet \citep{HampFeldweg1997} are too coarse for specialised purposes like the present study.
-As a result, \textit{Kindconsistency} with the self-explanatory factor levels in Table~\ref{tab:variables}) was annotated.
+As a result, \textit{Kindconsistency} with the self-explanatory factor levels in Table~\ref{tab:variables} was annotated.
 Also, \textit{Kindedible} encodes whether the substance denoted by the kind noun is edible.
-Finally, \textit{Kindint} encodes whether the substance is an artifact made with intent (like \textit{Suppe} `soup') or not (like \textit{Meerwasser} `sea water').
+Finally, \textit{Kindint} encodes whether that substance is an artifact made with intent (like \textit{Suppe} `soup') or not (like \textit{Meerwasser} `sea water').
 For the measure nouns, relevant classification seemed much clearer, and it was designed inspired by the list in \cite[530]{Koptjevskaja2001}.
@@
 Last, to control for lemma frequency effects, the logarithm-transformed frequency per 1,000,000 words of each lemma was added from the frequency lists distributed by the DECOW designers (variables \textit{Kindfreq}, \textit{Measurefreq}).
-All numeric variables were standardised and centered (z-transformed) to facilitate their interpretation in the
+All numeric variables were standardised and centered (z-transformed) to facilitate their interpretation in the regression models reported 
+in the next section.
 
@@
 
-\subsection{Hierarchical models}
+\subsection{Hierarchical models of the measure noun alternation}
 \label{sec:corpushierarchicalmodel}
 
-% > summary(pl$Casedrop)
-%   0   1 
-%  67 794 
-% > summary(fem$Casedrop)
-%   0   1 
-% 652 462 
-% > summary(mn$Genitive)
-%    0    1 
-% 3277  672
+In this section, I report the results of fitting a multilevel model to the data using R \citep{R}, \textit{lme4} \citep{lme4} for Maximum Likelihood Estimation, and \textit{rstanarm} \citep{rstanarm} for Markov-Chain Monte Carlo estimation.
+The purpose is to model the influence of the regressors specified in Table~\ref{tab:variables} on the probability that either the \NACa\ or the \PGCa\ is chosen.
+As explained in Section~\ref{sec:descriptive}, due to extensive case syncrestism in feminine NPs, feminine kind nouns in MNPs like \textit{(mit) einem Liter kalter Milch} `with a litre of cold milk' can either be datives (case agreement, \NACa) or genitives (\PGCa) whenever the measure noun is in the dative.
+All dative MNPs with feminine kind nouns had to be removed, and the regressor \textit{Measurecase} consequently only had two levels when the kind noun was feminine.%
+\footnote{Together with other cleanups (erroneous hits, extremely noisy material), the removal of such uninformative sentences reduced the total size of the sample to 5,063 sentences.}
+Therefore, two separate models for masculine\slash neuter and feminine kind nouns were fitted.
+For the masculine\slash neuter model, the final sample size was $n_{\text{MN}}=3,949$ with 672 cases of \PGCa\ and 3,277 cases of \NACa.
+For the feminine model, it was $n_{\text{F}}=1,114$ with 462 cases of \PGCa\ and 652 cases of $\NACa$.
+All regressors from Table~\ref{tab:variables} were included to begin with, and the measure lemma and the kind noun lemma were specified as varying-intercept random effects.
+The regressors listed in Table~\ref{tab:variables} with the measure lemma as their unit of reference have no within-measure lemma variance, and the \textit{glmer} function automatically estimates them as \textit{group level predictors} (or \textit{second-level effects}), cf.\ \cite[265--269,302--304]{GelmanHill2006}.
+The same goes for those listed with the kind lemma as their unit of reference.
 
-In this section, I apply standard Maximum Likelihood methods for estimating the parameters of a multilevel model.
-As suggested in Section~\ref{sec:corpusnonhierarchicalmodel}, hierarchical models are better suited than non-hierarchical models given that we can expect lemma-specific effects, and that the respective factors have a high number of levels.
+\input{bigtable}
 
-Equations~\ref{eq:model:mnglmm} through \ref{eq:model:mnglmm:alphakind} specify the basic model fitted to the masculine\slash neuter and the feminine dataset.
-I use the same notation as used in \cite{GelmanHill2006} with the exception that I use vector notation for regressor variables and coefficients if the number of regressors is large in order to make the model specification more compact.%
-\footnote{
-The response $y_i=1$ if the PGC is chosen in sentence $i$, else $0$.
-$\beta_0$ is the general intercept, $B$ a vector of first-level coefficients, and $X_i$ a vector of regressor variables for sentence $i$.
-$\alpha^{\text{measure}}_{j[i]}$ is the random intercept for measure lemma $j$, chosen appropriately for the measure lemma occurring in sentence $i$ (similarly, $\alpha^{\text{kind}}_{k[i]}$ for kind lemma $k$).
-$\gamma_0^{\text{measure}}$ is the intercept of the second-level model for measure lemmas, $\Gamma^{\text{measure}}$ a vector of second-level coefficients for measure lemmas, $U_j$ a vector of regressor variables for measure lemma $j$, and $\sigma_{\text{measure}}^2$ the variance of the measure lemma intercepts.
-The second-level model for kind lemmas is specified in the same manner.
-}
+%Second-level models were specified for kind nouns lemmas and measure noun lemmas.
+%Equations~\ref{eq:model:mnglmm} through \ref{eq:model:mnglmm:alphakind} show the specified structure of the model.
+%I use the same notation as used in \cite{GelmanHill2006} with the exception that I use vector notation for regressor variables and coefficients if the number of regressors is large in order to make the model specification more compact.%
+%\footnote{
+%The response $y_i=1$ if the PGC is chosen in sentence $i$, else $0$.
+%$\beta_0$ is the general intercept, $B$ a vector of first-level coefficients, and $X_i$ a vector of regressor variables for sentence $i$ (see Table~\ref{tab:bigtable}).
+%$\alpha^{\text{measure}}_{j[i]}$ is the random intercept for measure lemma $j$, chosen appropriately for the measure lemma occurring in sentence $i$ (similarly, $\alpha^{\text{kind}}_{k[i]}$ for kind lemma $k$).
+%$\gamma_0^{\text{measure}}$ is the intercept of the second-level model for measure lemmas, $\Gamma^{\text{measure}}$ a vector of second-level coefficients for measure lemmas, $U_j$ a vector of regressor variables for measure lemma $j$, and $\sigma_{\text{measure}}^2$ the variance of the measure lemma intercepts.
+%The second-level model for kind lemmas is specified in the same manner.
+%$\alpha\sim\mathcal{N}(\mu,\sigma^2)$ should be read \textit{$\alpha$ is normally distributed with mean $\mu$ and variance $\sigma^2$}.
+%}
+%
+%\begin{eqnarray}
+%    \label{eq:model:mnglmm}
+%    Pr(y_i=1) &=& logit^{-1}(\beta_0 + B\cdot X_i + \alpha^{\text{measure}}_{j[i]} + \alpha^{\text{kind}}_{k[i]}) \\
+%    \alpha^{\text{measure}}_j &\sim& \mathcal{N}(\gamma_0^{\text{measure}} + \Gamma^{\text{measure}}\cdot U_j,\sigma^2_{\text{measure}}) \\
+%    \alpha^{\text{kind}}_k &\sim& \mathcal{N}(\gamma_0^{\text{kind}} + \Gamma^{\text{kind}}\cdot U_k,\sigma^2_{\text{kind}}) 
+%  \label{eq:model:mnglmm:alphakind}
+%\end{eqnarray}
 
-\begin{eqnarray}
-    \label{eq:model:mnglmm}
-    Pr(y_i=1) &=& logit^{-1}(\beta_0 + B\cdot X_i + \alpha^{\text{measure}}_{j[i]} + \alpha^{\text{kind}}_{k[i]}) \\
-    \alpha^{\text{measure}}_j &\sim& \mathcal{N}(\gamma_0^{\text{measure}} + \Gamma^{\text{measure}}\cdot U_j,\sigma^2_{\text{measure}}) \\
-    \alpha^{\text{kind}}_k &\sim& \mathcal{N}(\gamma_0^{\text{kind}} + \Gamma^{\text{kind}}\cdot U_k,\sigma^2_{\text{kind}}) 
-  \label{eq:model:mnglmm:alphakind}
-\end{eqnarray}
+Regressors were removed if no convergence could be achieved (even after tuning the optimizer) or perfect separation occurred.
+Beyond that, no step-down model selection was performed (see Section~\ref{sec:rightstatistics}).
+Generalized variance inflation factors for the remaining regressors were calculated to check for multicollinearity \citep{FoxMonette1992}.
+Except for the levels of lemma noun class factors \textit{Kindconsistency} and \textit{Measureclass}, virtually no regressor has a GVIF over 3.
+Exceptions were GVIF(\textit{Kindattraction})=3.17 and GVIF(\textit{Kindfreq})=3.92, both still being far from dramatic.
+The high GVIFs for the noun lemma classes (up to 46.2) correlate with a low degree of theoretical motivation (at least of the concrete classes as annotated).
+However, re-estimating the models with these variables removed did \textit{not} alter the sign or magnitude of any of the other regressors, and I left them in the model.
 
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.8\textwidth]{figures/corpus/04_glmm_fixef_firstlevel}
-\caption{Estimated first-level effects in GLMMs for Masculine\slash Neuter and Feminine}
-\label{fig:glmmfixef}
-\end{figure}
+%\begin{table}
+%  \centering
+%  \begin{tabular}{lrr}
+%    Regressor (level)         &  Masculine\slash Neuter &   Feminine \\
+%    \midrule
+%    Badness                   &   1.02                  &   1.04     \\
+%    Genitives                 &   1.11                  &   1.13     \\
+%    Kindattraction            &   2.17                  &   3.17     \\
+%    KindconsistencyGaseous    &                         &   6.27     \\
+%    KindconsistencyHard       &                         &   5.86     \\
+%    KindconsistencyImmaterial &                         &   4.32     \\
+%    KindconsistencyLiquid     &                         &   8.45     \\
+%    KindconsistencyMixed      &                         &   7.28     \\
+%    KindconsistencyPowder     &                         &   3.87     \\
+%    KindconsistencySoft       &                         &  15.09     \\
+%    Kindedible0               &  (1.32)                 &   2.25     \\
+%    Kindfreq                  &   2.13                  &   3.92     \\
+%    KindintNonint             &  (1.27)                 &   2.45     \\
+%    Matchlength               &   1.17                  &   1.62     \\
+%    Measureabbreviated0       &   1.28                  &   1.29     \\
+%    Measureattraction         &   1.99                  &   1.06     \\
+%    MeasurecaseAcc            &   1.35                  &   1.07     \\
+%    MeasurecaseDat            &   1.37                  &            \\
+%    MeasureclassAmount        &  46.20                  &            \\
+%    MeasureclassCapacity      &  16.32                  &            \\
+%    MeasureclassCurrency      &   3.50                  &            \\
+%    MeasureclassLayer         &   7.92                  &            \\
+%    MeasureclassLength        &   2.14                  &            \\
+%    MeasureclassPortion       &  36.91                  &            \\
+%    MeasureclassRest          &   4.54                  &            \\
+%    MeasureclassVessel        &  18.81                  &            \\
+%    MeasureclassWeight        &  25.40                  &            \\
+%    Measurefreq               &   2.27                  &  (1.21)    \\
+%    MeasurenumberSg           &   1.90                  &   1.65     \\
+%    Minus1posA                &   3.77                  &   4.19     \\
+%    Minus1posD                &   4.05                  &   5.25     \\
+%    Minus1posO                &   1.30                  &   1.47     \\
+%    Minus1posP                &   1.21                  &   1.10     \\
+%  \end{tabular}
+%  \caption{Generalized variance inflation factors for the regressors of the hierarchical models}
+%  \label{tab:vif}
+%\end{table}
 
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.8\textwidth]{figures/corpus/04_glmm_fixef_secondlevel_kind}
-\caption{Estimated second-level effects for kind lemmas in GLMMs for Masculine\slash Neuter and Feminine}
-\label{fig:glmmfixef}
-\end{figure}
+Standard model diagnostics show that the models perform very well.
+For the masculine\slash neuter model, Nakagawa \& Schielzeth's determination coefficient is $R_m^2=0.495$ and $R^2_c=0.544$ (see \citealp{Gries2015} for a linguist-friendly introduction to these $R^2$ measures, or else \citealp{NakagawaSchielzeth2013}).
+The rate of correct predictions is $0.872$, which means an increase over the baseline of $\lambda=0.247$.
+For the feminine model, $R_m^2=0.468$ and $R^2_c=0.574$ are achieved, as well as a correct prediction rate of $0.815$, which means an improvement over the baseline of $\lambda=0.554$.
 
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.8\textwidth]{figures/corpus/04_glmm_fixef_secondlevel_measure}
-\caption{Estimated second-level effects for measure lemmas in GLMMs for Masculine\slash Neuter and Feminine}
-\label{fig:glmmfixef}
-\end{figure}
+In Table~\ref{tab:bigtable}, there are groups of columns for the results obtained from Maximum Likelihood estimation and Markov-Chain Monte Carlo estimation, both for the masculine\slash neuter and the feminine models.
+In the remainder of this section, I describe how these results were obtained.
+The coefficient estimates are specified for each regressor (level) in the columns labelled \textit{Coef}.%
+\footnote{Empty cells in the table correspond to regressors and their levels which had to be removed because of convergence problems.}
+For a robust estimation of the precision of the estimation, I ran a parametric bootstrap (using the \textit{confint.merMod} function from \textit{lme4}) with 1,000 replications and the bootstrap percentile method.
+The resulting 95\% bootstrap intervals are reported in Table~\ref{tab:bigtable} in the columns labelled \textit{Low}, \textit{High} (for the upper and lower 2.5th percentile) and \textit{Width} (for the width of the interval).
+The column \textit{Not 0} shows an asterisk for those intervals that do not include 0 as a more reasonable replacement of `significance levels'.
 
+As mentioned above, not step-down model selection was performed.
+However, for each regressor, a $p$ value was obtained by dropping it from the full model, re-eastimating the nested model and comparing it to the full model.
+Instead of inexact Wald approximations or Likelihood Ratio Tests, I used a drop-in bootstrap replacement for the Likelihood Ratio Test with the function \textit{PBmodcomp} from the \textit{pbkrtest} package \citep{HalekohHojsgaard2014}.
+These bootstrapped $p$ values are given in the columns labelled \textit{p\Sub{BP}} in Table~\ref{tab:bigtable}.
+
+Finally, MCMC TODO.
+
+%\begin{figure}[h]
+%\centering
+%\includegraphics[width=0.8\textwidth]{figures/corpus/04_glmm_fixef_firstlevel}
+%\caption{Estimated first-level effects in GLMMs for Masculine\slash Neuter and Feminine}
+%\label{fig:glmmfixef}
+%\end{figure}
+%
+%\begin{figure}[h]
+%\centering
+%\includegraphics[width=0.8\textwidth]{figures/corpus/04_glmm_fixef_secondlevel_kind}
+%\caption{Estimated second-level effects for kind lemmas in GLMMs for Masculine\slash Neuter and Feminine}
+%\label{fig:glmmfixef}
+%\end{figure}
+%
+%\begin{figure}[h]
+%\centering
+%\includegraphics[width=0.8\textwidth]{figures/corpus/04_glmm_fixef_secondlevel_measure}
+%\caption{Estimated second-level effects for measure lemmas in GLMMs for Masculine\slash Neuter and Feminine}
+%\label{fig:glmmfixef}
+%\end{figure}
+%
+%
 %\begin{figure}[h]
@@
 \label{sec:modelssummary}
-
-\input{bigtable}
 

### 2017-03-22 R/05_glmm.R

@@ -81,7 +81,7 @@ lmer.modelcomparison <- function(model, regressors, formula.target, ci = 0.95, n
 # for the remaning factors substantially.
 # Code:
 # source("mer-utils.R")
-# vif.mer(mn.glmm)
+# vif.mer(fem.glmm)
 
 mn.glmm <- glmer(Genitive~1
                 +(1|Measurelemma)
@@ -524,3 +524,10 @@ print(p)
 if (save.persistent) dev.off()
 
 
+if (save.persistent) sink(paste(out.dir, "05_glmm.txt", sep=""), append = T)
+source("mer-utils.R")
+cat("\n\n MULTICOLLINEARITY\nM/N\n")
+print(vif.mer(mn.glmm))
+cat("\n\n F\n")
+print(vif.mer(fem.glmm))
+if (save.persistent) sink()

### 2017-03-24 Latex/article.tex
Work on interpretation

@@
 %Section~\ref{sec:bayesian} takes a short detour and shows (by example) that results from Bayesian estimation are not intrinsically different from those obtained from Maximum Likelihood methods, although Bayesian approaches have recently been touted to solve various problems facing applied statistics in various fields, even including cognitively oriented corpus linguistics.
-%Finally, I summarise and interpret the results of the corpus study in Section~\ref{sec:modelssummary}.
+%Finally, I summarise and interpret the results of the corpus study in Section~\ref{sec:interpretation}.
 
@@
 
-%Related to the , dative measure nouns favor the PGC more than nominative and accusative (also \citealp{Hentschel1993,Zimmer2015}).
-%As an example, \textit{mit einem Stück frischen Brots} `with a piece of fresh bread' (PGC) would be preferred over \textit{mit einem Stück frischem Brot} (NAC).
-%I propose
-%
-
 Turning to grammaticalisation now, it is often assumed that pseudo-partitives arise as a form of grammaticalised partitives (\eg\ \citealp[536--539]{Koptjevskaja2001} for Finnish and Estonian, \citealp[559]{Koptjevskaja2001} for European languages per se).
@@
 The best evidence for the relevance of this factor in an ongoing grammaticalisation process is that plural kind nouns almost always (but not exclusively) occur in the \PGCa\ and virtually never in the \NACa\ (see Section~\ref{sec:descriptive}).
-More indirectly, however, plurality of the measure noun could also favor the \PGCa\ over the \NACa, and exact quantification in the form of cardinals even more so.%
+More indirectly, however, plurality of the measure noun could also favour the \PGCa\ over the \NACa, and exact quantification in the form of cardinals even more so.%
 \footnote{This preference was observed empirically by \cite{Hentschel1993}.}
@@
 \begin{enumerate}[i.]
-  \item The relative frequencies with which measure noun lemmas and kind noun lemmas appear in the prototypical (unequivocal) PGC and NAC are positively correlated with the relative frequency the PGC and the NAC are chosen in the alternating case.
-  \item MNPs with masculine dative kind nouns disfavour the GPC less.
-  \item Classes of more clearly grammaticalised measure nouns favor the NAC.
-  \item Plural measure nouns and measure nouns quantified by cardinals favor the PGC variant.
+  \item The relative frequencies with which measure noun lemmas and kind noun lemmas appear in the prototypical (non-alternating) \PGCd\ and \NACb\ are positively correlated with the relative frequency the alternating \PGCa\ and the \NACa\ are chosen.
+  \item (Classes of) more strongly grammaticalised measure nouns favour the NAC.
+  \item Plural measure nouns and measure nouns quantified by cardinals favour the PGC variant.
   \item The NAC is associated more strongly with higher registers.
@@
 
-\subsection{Annotation}
+\subsection{Variables and annotation}
 \label{sec:annotation}
@@
 Documents with higher Badness usually contain more incoherent language, shorter sentences, etc.
-If the PGC actually favors higher registers and styles, a high \textit{Badness} should be correlated with fewer occurrences of the PGC.
+If the PGC actually favours higher registers and styles, a high \textit{Badness} should be correlated with fewer occurrences of the PGC.
 Documents in DECOW14 have also been annotated with a variable called \textit{Genitives}.
@@
 The level names stand for \textit{adjective}, \textit{cardinal}, \textit{determiner}, \textit{other}, and \textit{preposition} (in the same order as in Table~\ref{tab:variables}).
-The purpose of this variable is mainly to see whether cardinals really favor the \PGCa\ as hypothesised in Section~\ref{sec:analyses}.
+The purpose of this variable is mainly to see whether cardinals really favour the \PGCa\ as hypothesised in Section~\ref{sec:analyses}.
 
@@
 
-Last, to control for lemma frequency effects, the logarithm-transformed frequency per 1,000,000 words of each lemma was added from the frequency lists distributed by the DECOW designers (variables \textit{Kindfreq}, \textit{Measurefreq}).
+Finally, two variables were added as more or less nuisance variables in the context of the present study.
+First, it was reported in the literature that MNPs in the dative and with a masculine or neuter kind noun favour the \PGCa\ more than the corresponding nominative and accusative MNPs \citep{Hentschel1993,Zimmer2015}.
+As an example, \textit{mit einem Stück frischen Brots} `with a piece of fresh bread' (\PGCa) would be preferred over \textit{mit einem Stück frischem Brot} (\NACa).
+As with all the examples, native speakers of German will most likely notice that differences are subtle.
+To control for this variable, the case of the measure noun was manually annotated (variable \textit{Measurecase}).
+Second, to control for lemma frequency effects, the logarithm-transformed frequency per 1,000,000 words of each lemma was added from the frequency lists distributed by the DECOW designers (variables \textit{Kindfreq}, \textit{Measurefreq}).
 All numeric variables were standardised and centered (z-transformed) to facilitate their interpretation in the regression models reported 
@@
 As explained in Section~\ref{sec:descriptive}, due to extensive case syncrestism in feminine NPs, feminine kind nouns in MNPs like \textit{(mit) einem Liter kalter Milch} `with a litre of cold milk' can either be datives (case agreement, \NACa) or genitives (\PGCa) whenever the measure noun is in the dative.
-All dative MNPs with feminine kind nouns had to be removed, and the regressor \textit{Measurecase} consequently only had two levels when the kind noun was feminine.%
+All dative MNPs with feminine kind nouns had to be removed, and the regressor \textit{Measurecase} consequently only had two levels with feminine kind nouns.%
 \footnote{Together with other cleanups (erroneous hits, extremely noisy material), the removal of such uninformative sentences reduced the total size of the sample to 5,063 sentences.}
@@
 The same goes for those listed with the kind lemma as their unit of reference.
+The results of the estimation are shown in Table~\ref{tab:bigtable}.
 
@@
 Exceptions were GVIF(\textit{Kindattraction})=3.17 and GVIF(\textit{Kindfreq})=3.92, both still being far from dramatic.
-The high GVIFs for the noun lemma classes (up to 46.2) correlate with a low degree of theoretical motivation (at least of the concrete classes as annotated).
+The high GVIFs for the noun lemma classes (up to 46.2) correlate with a low degree of theoretical motivation (at least for the concrete classes as annotated).
 However, re-estimating the models with these variables removed did \textit{not} alter the sign or magnitude of any of the other regressors, and I left them in the model.
@@
 
-Standard model diagnostics show that the models perform very well.
-For the masculine\slash neuter model, Nakagawa \& Schielzeth's determination coefficient is $R_m^2=0.495$ and $R^2_c=0.544$ (see \citealp{Gries2015} for a linguist-friendly introduction to these $R^2$ measures, or else \citealp{NakagawaSchielzeth2013}).
-The rate of correct predictions is $0.872$, which means an increase over the baseline of $\lambda=0.247$.
-For the feminine model, $R_m^2=0.468$ and $R^2_c=0.574$ are achieved, as well as a correct prediction rate of $0.815$, which means an improvement over the baseline of $\lambda=0.554$.
-
-In Table~\ref{tab:bigtable}, there are groups of columns for the results obtained from Maximum Likelihood estimation and Markov-Chain Monte Carlo estimation, both for the masculine\slash neuter and the feminine models.
+Standard diagnostics show that the models perform very well.
+For the masculine\slash neuter model, Nakagawa \& Schielzeth's pseudo-coefficient of determination is $R_m^2=0.495$ and $R^2_c=0.544$ (see \citealp{Gries2015} for a linguist-friendly introduction to these $R^2$ measures, or else \citealp{NakagawaSchielzeth2013}).
+The rate of correct predictions is $0.872$, which means a proportional reduction of error of $\lambda=0.247$.
+For the feminine model, $R_m^2=0.468$ and $R^2_c=0.574$ are achieved, as well as a correct prediction rate of $0.815$, which means a proportional reduction of error of $\lambda=0.554$.
+The measure and kind lemma intercepts are distributed with the following standard deviations:
+Masculine\slash neuter $\sigma_{\text{Measurelemma}}>0.001$ (131 lemmas), $\sigma_{\text{Kindlemma}}>0.591$ (68 lemmas);
+Feminine $\sigma_{\text{Measurelemma}}=0.905$ (73 lemmas), $\sigma_{\text{Kindlemma}}=0.000$ (19 lemmas).
+While in the masculine\slash neuter model individual measure lemmas do not play much of role, kind lemmas (and -- as will be shown -- the second level model for kind lemmas) are irrelevant in the feminine model, most likely because of the low number of levels.%
+\footnote{This low number is expected given that most mass nouns are masculine and neuter in German.}
+This will be discussed in Section~\ref{sec:interpretation}.
+
+In Table~\ref{tab:bigtable}, there are groups of columns for the results obtained from Maximum Likelihood estimation and Markov-Chain Monte Carlo estimation, both for the masculine\slash neuter and the feminine models and including confidence intervals.
 In the remainder of this section, I describe how these results were obtained.
@@
 \footnote{Empty cells in the table correspond to regressors and their levels which had to be removed because of convergence problems.}
-For a robust estimation of the precision of the estimation, I ran a parametric bootstrap (using the \textit{confint.merMod} function from \textit{lme4}) with 1,000 replications and the bootstrap percentile method.
-The resulting 95\% bootstrap intervals are reported in Table~\ref{tab:bigtable} in the columns labelled \textit{Low}, \textit{High} (for the upper and lower 2.5th percentile) and \textit{Width} (for the width of the interval).
-The column \textit{Not 0} shows an asterisk for those intervals that do not include 0 as a more reasonable replacement of `significance levels'.
-
-As mentioned above, not step-down model selection was performed.
-However, for each regressor, a $p$ value was obtained by dropping it from the full model, re-eastimating the nested model and comparing it to the full model.
-Instead of inexact Wald approximations or Likelihood Ratio Tests, I used a drop-in bootstrap replacement for the Likelihood Ratio Test with the function \textit{PBmodcomp} from the \textit{pbkrtest} package \citep{HalekohHojsgaard2014}.
-These bootstrapped $p$ values are given in the columns labelled \textit{p\Sub{BP}} in Table~\ref{tab:bigtable}.
-
-Finally, MCMC TODO.
+For a robust quantification of the precision of the estimation, I ran a parametric bootstrap (using the \textit{confint.merMod} function from \textit{lme4}) with 1,000 replications, and using the percentile method for the calculation of the intervals.
+The resulting 95\% bootstrap confidence intervals are reported in Table~\ref{tab:bigtable} in the columns labelled \textit{Low}, \textit{High} (for the upper and lower 2.5th percentile) and \textit{Width} (for the width of the interval).
+The column \textit{Not 0} shows an asterisk for those intervals that do not include 0 as a more modest replacement of significance levels.
+
+As mentioned above, no step-down model selection was performed.
+However, for each regressor, a $p$ value was obtained by dropping the regressor from the full model, re-estimating the nested model and comparing it to the full model.
+Instead of inexact Wald approximations or Likelihood Ratio Tests, I used a drop-in bootstrap replacement for the Likelihood Ratio Test from the function \textit{PBmodcomp} from the \textit{pbkrtest} package \citep{HalekohHojsgaard2014}.
+These bootstrapped $p$ values are given in the columns labelled p\Sub{BP} in Table~\ref{tab:bigtable}.
+
+Finally, I show that Bayesian methods in the form of Markov-Chain Monte Carlo estimation do not necessarily (and in this case predictably) lead to different results.
+Both the masculine\slash neuter model and the feminine model were re-estimated using the \textit{stan\_glmer} function from the \textit{rstanarm} package, which provides an \textit{lme4}-compatible syntax for estimating common model types with the \textit{stan} software \citep{CarpenterEa2017}.
+The algorithm was run using 4 chains and 1,000 iterations, and instead of guessing prior distributions or using implausible uniform prior distributions \cite[251--252]{Levshina2016}, I used plausible default priors.
+Most notably, priors for coefficients were specified as $\mathcal{N}(0,10)$ because coefficients higher than $10$ or lower than $-10$ are extremely rare in well-specified models on appropriate data.%
+\footnote{Consider that with a coefficient of 10, each increase by $1$ in the regressor variable means an increase in odds of $exp(10)=22026.47$.
+To reliably estimate such coefficients, extremely large samples would be required.}
+The algorithm converged, and for all coefficients, the $\hat{R}$ diagnostic was exactly $1$.
+The resulting coefficients and intervals as well as an * for \textit{interval does not contain $0$} are also given in Table~\ref{tab:bigtable} in the groups of columns labelled \textit{MCMC}.
+In the interpretation of the results in the next section, it will become apparent that both methods of estimation lead to the exact same theoretically relevant results.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 %
-%
-%\begin{figure}[h]
-%\centering
-%\includegraphics[width=0.6\textwidth]{figures/corpus/04_glmm_fixeff_mn_Measureclass}
-%\caption{Effect plot for the masculine\slash neuter GLMM: Measure lemma class effect}
-%\label{fig:glmm:fixef:measureclass}
-%\end{figure}
-%
-%\begin{figure}[h]
-%\centering
-%\includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_mn_Attraction}~\includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_fem_Attraction}
-%\caption{Effect plot for the two GLMMs: attraction effect (centered and standardised) from the NAC construction with a bare kind noun and the PGC with a determiner; scales are not aligned between left and right panel}
-%\label{fig:glmm:fixef:attraction}
-%\end{figure}
-%
-%\begin{figure}[h]
-%\centering
-%\includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_mn_Matchlength}~\includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_fem_Matchlength}
-%\caption{Effect plot for the two GLMMs: effect of the (centered and standardised) length of the whole construction; scales are not aligned between left and right panel}
-%\label{fig:glmm:fixef:matchlength}
-%\end{figure}
-%
-%\begin{figure}[h]
-%\centering
-%\includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_mn_Measurecase}~\includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_fem_Measurecase}
-%\caption{Effect plot for the two GLMMs: effects of the case of the head (kind) noun; scales are not aligned between left and right panel}
-%\label{fig:glmm:fixef:measureclass}
-%\end{figure}
-%
-%\begin{figure}[h]
-%\centering
-%\includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_mn_Measurenumber}~\includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_fem_Measurenumber}
-%\caption{Effect plot for the two GLMMs: effect of the grammatical number of the head (kind) noun; scales are not aligned between left and right panel}
-%\label{fig:glmm:fixef:measurenumber}
-%\end{figure}
-%
-%\begin{figure}[h]
-%\centering
-%\includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_mn_Minus1pos}~\includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_fem_Minus1pos}
-%\caption{Effect plot for the two GLMMs: effect of part-of-speech preceding the construction; C = cardinal, D = determiner, O = \textit{other}, A = adjective, P = preposition; scales are not aligned between left and right panel}
-%\label{fig:glmm:fixef:minus1pos}
-%\end{figure}
-%
-%\begin{figure}[h]
-%\centering
-%\includegraphics[width=0.5\textwidth]{figures/corpus/04_glmm_raneff_mn}~\hspace{0.05\textwidth}~\includegraphics[width=0.5\textwidth]{figures/corpus/04_glmm_raneff_fem}
-%\caption{Sample random effects for the masculine\slash neuter and feminine GLMMs}
-%\label{fig:glmm:fixef:minus1pos}
-%\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Summary and conclusions}
-\label{sec:modelssummary}
+
+\subsection{Interpretation}
+\label{sec:interpretation}
+
+The results reported in Section~\ref{sec:corpushierarchicalmodel} generally confirm the hypotheses from Section~\ref{sec:analyses}.
+Before I turn to some details, it should be noticed that the results from standard MLE (with bootstrap) and MCMC (with default priors) obviously differ numerically but not enough to result in different interpretations.
+The signs and magnitudes of the coefficients are identical with the two methods, and the confidence intervals have roughly the same width and symmetry properties.
+In only one single case (masculine\slash neuter, \textit{Measureattraction}) the MLE bootstrap confidence interval does not include $0$ while the MCMC interval does (barely, with $-0.01$).
+While I only interpret the MLE results, virtually the same interpretation could be made for the MCMC results.
+Furthermore, the masculine\slash neuter model and the feminine model agree in many relevant aspects.
+Disagreement between the models will be discussed below.
+
+% ATTRACTION
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_mn_Measureattraction}~\includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_fem_Measureattraction}
+  \caption{Effect plot for \textit{Measureattraction}}
+  \label{fig:eff:measureattraction}
+\end{figure}
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_mn_Kindattraction}~\includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_fem_Kindattraction}
+  \caption{Effect plot for \textit{Kindattraction}}
+  \label{fig:eff:kindattraction}
+\end{figure}
+
+First, the prototypicality effect related to the non-alternating \PGCd\ and \NACb\ can be shown, see the effect plots in Figure~\ref{fig:eff:measureattraction} and Figure~\ref{fig:eff:kindattraction}.%
+\footnote{Effect plots were created using the \textit{effects} package \citep{Fox2003}.
+It shows the change in probability for the outcome (y axis) dependent on values of a regressor (x axis),assuming typical values of all other regressors.}
+The effect is mostly as expected:
+if a lemma appears relatively more often in the \PGCd\ (compared to its frequency in the \NACb), the more often the \PGCa\ is chosen over the \NACa\ with this specific lemma.
+Only the feminine model \textit{Kindattraction} regressor is clearly useless (`non-significant') in the PBmodcomp test, and the effect plot shows a corresponding picture.
+This is related to the fact that in the feminine model the kind lemma effect and its second-level model turned out to be uninformative relative to the alternation (remember that $\sigma_{\text{Kindlemma}}=0.000$).
+
+% MN
+% $\sigma_{\text{Measurelemma}}<0.001$
+% $\sigma_{\text{Kindlemma}}>0.591$
+% F
+% $\sigma_{\text{Measurelemma}}=0.905$
+% $\sigma_{\text{Kindlemma}}=0.000$
+
+% GRAMMATICALISATION
+% => lemmas and lemma classes
+
+% PLURAL & CARDINALS
+
+% REGISTER
+% => Badness
+% => Genitives
+% => Matchlength
+% => Measureabbreviated
+
+% NUISANCE
+% => dative effect
+% => frequency
 
@@
 For the F sentences, the top and bottom 25\% were used because the overall sample size was smaller.}
-It was made sure that the highly predictive features according to Section \ref{sec:modelssummary} occurred in unique combinations among the four sentences in each cell.
+It was made sure that the highly predictive features according to Section \ref{sec:interpretation} occurred in unique combinations among the four sentences in each cell.
 This is summarised in Table \ref{tab:experiment1:design}.

### 2017-03-25 Latex/article.tex
changes to statistics section

@@
 As with all the examples, native speakers of German will most likely notice that differences are subtle.
-To control for this variable, the case of the measure noun was manually annotated (variable \textit{Measurecase}).
+To control for this effect, the case of the measure noun was manually annotated (variable \textit{Measurecase}).
 Second, to control for lemma frequency effects, the logarithm-transformed frequency per 1,000,000 words of each lemma was added from the frequency lists distributed by the DECOW designers (variables \textit{Kindfreq}, \textit{Measurefreq}).
@@
 The results of the estimation are shown in Table~\ref{tab:bigtable}.
+For the response variable, \PGCa\ was coded as $1$, such that coefficients leaning to the positive side can be interpreted as favouring the \NACa.
 
@@
 
-Standard diagnostics show that the models perform very well.
+Standard diagnostics show that the model quality is quite high.
 For the masculine\slash neuter model, Nakagawa \& Schielzeth's pseudo-coefficient of determination is $R_m^2=0.495$ and $R^2_c=0.544$ (see \citealp{Gries2015} for a linguist-friendly introduction to these $R^2$ measures, or else \citealp{NakagawaSchielzeth2013}).
@@
 The measure and kind lemma intercepts are distributed with the following standard deviations:
-Masculine\slash neuter $\sigma_{\text{Measurelemma}}>0.001$ (131 lemmas), $\sigma_{\text{Kindlemma}}>0.591$ (68 lemmas);
+Masculine\slash neuter $\sigma_{\text{Measurelemma}}>0.001$ (131 lemmas), $\sigma_{\text{Kindlemma}}=0.591$ (68 lemmas);
 Feminine $\sigma_{\text{Measurelemma}}=0.905$ (73 lemmas), $\sigma_{\text{Kindlemma}}=0.000$ (19 lemmas).
-While in the masculine\slash neuter model individual measure lemmas do not play much of role, kind lemmas (and -- as will be shown -- the second level model for kind lemmas) are irrelevant in the feminine model, most likely because of the low number of levels.%
+In the masculine\slash neuter model individual measure lemmas do not play much of role, mostly because the second-level regressors successfully model the between-lemma variance.
+In the feminine model, kind lemmas (and -- as will be shown -- the second level model for kind lemmas) are irrelevant, most likely because of the low number of levels.%
 \footnote{This low number is expected given that most mass nouns are masculine and neuter in German.}
@@
 The algorithm converged, and for all coefficients, the $\hat{R}$ diagnostic was exactly $1$.
-The resulting coefficients and intervals as well as an * for \textit{interval does not contain $0$} are also given in Table~\ref{tab:bigtable} in the groups of columns labelled \textit{MCMC}.
+The resulting coefficients and intervals as well as an * for \textit{interval does not contain $0$} (column \textit{Not 0}) are also given in Table~\ref{tab:bigtable} in the groups of columns labelled \textit{MCMC}.
 In the interpretation of the results in the next section, it will become apparent that both methods of estimation lead to the exact same theoretically relevant results.
@@
 The results reported in Section~\ref{sec:corpushierarchicalmodel} generally confirm the hypotheses from Section~\ref{sec:analyses}.
-Before I turn to some details, it should be noticed that the results from standard MLE (with bootstrap) and MCMC (with default priors) obviously differ numerically but not enough to result in different interpretations.
+Before I turn to some details, it should be noticed that the results from standard MLE (with bootstrap) and MCMC (with default priors) obviously differ in numerical details -- but not enough to result in different interpretations.
 The signs and magnitudes of the coefficients are identical with the two methods, and the confidence intervals have roughly the same width and symmetry properties.
@@
   \centering
+  \includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_mn_Kindattraction}~\includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_fem_Kindattraction}
+  \caption{Effect plot for the regressor \textit{Kindattraction} for the masculine\slash neuter and the feminine model (y axes have different scales)}
+  \label{fig:eff:kindattraction}
+\end{figure}
+
+\begin{figure}[h!]
+  \centering
   \includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_mn_Measureattraction}~\includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_fem_Measureattraction}
-  \caption{Effect plot for \textit{Measureattraction}}
+  \caption{Effect plot for the regressor \textit{Measureattraction} for the masculine\slash neuter and the feminine model (y axes have different scales)}
   \label{fig:eff:measureattraction}
@@
 
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_mn_Kindattraction}~\includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_fem_Kindattraction}
-  \caption{Effect plot for \textit{Kindattraction}}
-  \label{fig:eff:kindattraction}
-\end{figure}
-
-First, the prototypicality effect related to the non-alternating \PGCd\ and \NACb\ can be shown, see the effect plots in Figure~\ref{fig:eff:measureattraction} and Figure~\ref{fig:eff:kindattraction}.%
+First, the prototypicality effect related to the non-alternating \PGCd\ and \NACb\ can be shown, see the effect plots in Figure~\ref{fig:eff:kindattraction} and Figure~\ref{fig:eff:measureattraction}.%
 \footnote{Effect plots were created using the \textit{effects} package \citep{Fox2003}.
-It shows the change in probability for the outcome (y axis) dependent on values of a regressor (x axis),assuming typical values of all other regressors.}
+They show the changes in probability for the outcome (y axis) dependent on values of a regressor (x axis),assuming typical values of all other regressors.}
 The effect is mostly as expected:
@@
 This is related to the fact that in the feminine model the kind lemma effect and its second-level model turned out to be uninformative relative to the alternation (remember that $\sigma_{\text{Kindlemma}}=0.000$).
-
-% MN
-% $\sigma_{\text{Measurelemma}}<0.001$
-% $\sigma_{\text{Kindlemma}}>0.591$
-% F
-% $\sigma_{\text{Measurelemma}}=0.905$
-% $\sigma_{\text{Kindlemma}}=0.000$
+The effect plots show a picture in line with a \textit{divide by 4} interpretation of the coefficients as shown in Table~\ref{tab:bigtable}.%
+\footnote{Coefficients are on the logit scale, which does not linearly correspond to the probability scale.
+If coefficients are reasonably small, however, dividing the coefficient by $4$ gives a good approximation of the change on the probability scale for each increment of the response variable \citep[82]{GelmanHill2006}.}
+Given that all regressors have been z-transformed, we can make statements such as the following:
+If a masculine kind noun is one standard deviation more frequently encountered in the \PGCd\ (compared to its frequency in the \NACb) than the kind nouns on average, the probability that the \PGCa\ is chosen instead of the \NACa\ with this kind noun in it rises by $\beta^{MN}_{\text{Kindattraction}}/4=0.49/4=0.1225$, etc.
+For measure nouns in MNPs with masculine kind nouns, it is $\beta^{MN}_{\text{Measureattraction}}/4=0.21/4=0.0525$, and for measure nouns in MNPs with feminine kind nouns, it is $\beta^{F}_{\text{Measureattraction}}/4=0.43/4=0,1075$.
+As readers can observe, these numbers correspond approximately to the slopes of the effect plots.
 
@@
 
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=0.6\textwidth]{figures/corpus/04_glmm_fixeff_mn_Measureclass}
+  \caption{Effect plot for the regressor \textit{Measureclass} for the masculine\slash neuter model}
+  \label{fig:eff:measureattraction}
+\end{figure}
+
 % PLURAL & CARDINALS
+%
+%\begin{figure}[h!]
+%  \centering
+%  \includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_mn_Minus1pos}~\includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_fem_Minus1pos}
+%  \caption{Effect plot for the regressor \textit{Minus1pos} for the masculine\slash neuter and the feminine model (y axes have different scales)}
+%  \label{fig:eff:measureattraction}
+%\end{figure}
+%
+%\begin{figure}[h!]
+%  \centering
+%  \includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_mn_Measurenumber}~\includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_fem_Measurenumber}
+%  \caption{Effect plot for the regressor \textit{Measurenumber} for the masculine\slash neuter and the feminine model (y axes have different scales)}
+%  \label{fig:eff:measureattraction}
+%\end{figure}
 
@@
 
-\section{Experimental validation of the corpus-based models}
+\section{Experimental validation}
 \label{sec:externalvalidation}

### 2017-03-27 Latex/article.tex
new statistics, part 1

@@
 
-\subsection{Getting the right data}
+\subsection{Corpus choice and sampling}
 \label{sec:gettingdata}
@@
 \footnote{The corpora are made available for free at \url{https://www.webcorpora.org}.
+At the time of this writing, a newer 2016 version DECOW16 had already been released.}
 I chose this corpus for two main reasons.
-At the time of this writing, a newer 2016 version DECOW16 had already been released.}
 First, the external validity of any study is increased through a higher heterogeneity of the sample \citep[30]{MaxwellDelaney2004}, and the DECOW corpus has clearly a much more heterogeneous composition compared to the only other very large corpus of German, the DeReKo \citep{KupietzEa2010} of the Institute for the German Language (IDS), which contains almost exclusively newspaper texts.%
@@
 Therefore, two separate models for masculine\slash neuter and feminine kind nouns were fitted.
-For the masculine\slash neuter model, the final sample size was $n_{\text{MN}}=3,949$ with 672 cases of \PGCa\ and 3,277 cases of \NACa.
-For the feminine model, it was $n_{\text{F}}=1,114$ with 462 cases of \PGCa\ and 652 cases of $\NACa$.
+For the masculine\slash neuter model (henceforth \textit{MN model}), the final sample size was $n_{\text{MN}}=3,949$ with 672 cases of \PGCa\ and 3,277 cases of \NACa.
+For the feminine model (henceforth \textit{F model}), it was $n_{\text{F}}=1,114$ with 462 cases of \PGCa\ and 652 cases of $\NACa$.
 All regressors from Table~\ref{tab:variables} were included to begin with, and the measure lemma and the kind noun lemma were specified as varying-intercept random effects.
@@
 Standard diagnostics show that the model quality is quite high.
-For the masculine\slash neuter model, Nakagawa \& Schielzeth's pseudo-coefficient of determination is $R_m^2=0.495$ and $R^2_c=0.544$ (see \citealp{Gries2015} for a linguist-friendly introduction to these $R^2$ measures, or else \citealp{NakagawaSchielzeth2013}).
+For the MN model, Nakagawa \& Schielzeth's pseudo-coefficient of determination is $R_m^2=0.495$ and $R^2_c=0.544$ (see \citealp{Gries2015} for a linguist-friendly introduction to these $R^2$ measures, or else \citealp{NakagawaSchielzeth2013}).
 The rate of correct predictions is $0.872$, which means a proportional reduction of error of $\lambda=0.247$.
-For the feminine model, $R_m^2=0.468$ and $R^2_c=0.574$ are achieved, as well as a correct prediction rate of $0.815$, which means a proportional reduction of error of $\lambda=0.554$.
+For the F model, $R_m^2=0.468$ and $R^2_c=0.574$ are achieved, as well as a correct prediction rate of $0.815$, which means a proportional reduction of error of $\lambda=0.554$.
 The measure and kind lemma intercepts are distributed with the following standard deviations:
-Masculine\slash neuter $\sigma_{\text{Measurelemma}}>0.001$ (131 lemmas), $\sigma_{\text{Kindlemma}}=0.591$ (68 lemmas);
-Feminine $\sigma_{\text{Measurelemma}}=0.905$ (73 lemmas), $\sigma_{\text{Kindlemma}}=0.000$ (19 lemmas).
-In the masculine\slash neuter model individual measure lemmas do not play much of role, mostly because the second-level regressors successfully model the between-lemma variance.
-In the feminine model, kind lemmas (and -- as will be shown -- the second level model for kind lemmas) are irrelevant, most likely because of the low number of levels.%
+MN model $\sigma_{\text{Measurelemma}}<0.001$ (131 lemmas), $\sigma_{\text{Kindlemma}}=0.591$ (68 lemmas);
+F model $\sigma_{\text{Measurelemma}}=0.905$ (73 lemmas), $\sigma_{\text{Kindlemma}}=0.000$ (19 lemmas).
+In the MN model, variance between individual measure lemmas do not play much of role, mostly because the second-level regressors successfully model the between-lemma variance.
+In the F model, kind lemmas (and -- as will be shown -- the second level model for kind lemmas) could not be estimated with great success, most likely because of the low number of levels.%
 \footnote{This low number is expected given that most mass nouns are masculine and neuter in German.}
@@
 
-In Table~\ref{tab:bigtable}, there are groups of columns for the results obtained from Maximum Likelihood estimation and Markov-Chain Monte Carlo estimation, both for the masculine\slash neuter and the feminine models and including confidence intervals.
+In Table~\ref{tab:bigtable}, there are groups of columns for the results obtained from Maximum Likelihood estimation and Markov-Chain Monte Carlo estimation, both for the MN model and F models and including confidence intervals.
 In the remainder of this section, I describe how these results were obtained.
@@
 Instead of inexact Wald approximations or Likelihood Ratio Tests, I used a drop-in bootstrap replacement for the Likelihood Ratio Test from the function \textit{PBmodcomp} from the \textit{pbkrtest} package \citep{HalekohHojsgaard2014}.
-These bootstrapped $p$ values are given in the columns labelled p\Sub{BP} in Table~\ref{tab:bigtable}.
+These bootstrapped $p$ values are given in the columns labelled p\Sub{PB} in Table~\ref{tab:bigtable}.
 
 Finally, I show that Bayesian methods in the form of Markov-Chain Monte Carlo estimation do not necessarily (and in this case predictably) lead to different results.
-Both the masculine\slash neuter model and the feminine model were re-estimated using the \textit{stan\_glmer} function from the \textit{rstanarm} package, which provides an \textit{lme4}-compatible syntax for estimating common model types with the \textit{stan} software \citep{CarpenterEa2017}.
+Both the MN model and the F model were re-estimated using the \textit{stan\_glmer} function from the \textit{rstanarm} package, which provides an \textit{lme4}-compatible syntax for estimating common model types with the \textit{stan} software \citep{CarpenterEa2017}.
 The algorithm was run using 4 chains and 1,000 iterations, and instead of guessing prior distributions or using implausible uniform prior distributions \cite[251--252]{Levshina2016}, I used plausible default priors.
@@
 The signs and magnitudes of the coefficients are identical with the two methods, and the confidence intervals have roughly the same width and symmetry properties.
-In only one single case (masculine\slash neuter, \textit{Measureattraction}) the MLE bootstrap confidence interval does not include $0$ while the MCMC interval does (barely, with $-0.01$).
+In only one single case (MN model, \textit{Measureattraction}) the MLE bootstrap confidence interval does not include $0$ while the MCMC interval does (barely, with $-0.01$).
 While I only interpret the MLE results, virtually the same interpretation could be made for the MCMC results.
-Furthermore, the masculine\slash neuter model and the feminine model agree in many relevant aspects.
+Furthermore, the MN model and the F model agree in many relevant aspects.
 Disagreement between the models will be discussed below.
@@
   \includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_mn_Kindattraction}~\includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_fem_Kindattraction}
-  \caption{Effect plot for the regressor \textit{Kindattraction} for the masculine\slash neuter and the feminine model (y axes have different scales)}
+  \caption{Effect plot for the regressor \textit{Kindattraction} for the MN model and the F model (y axes have different scales)}
   \label{fig:eff:kindattraction}
@@
   \includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_mn_Measureattraction}~\includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_fem_Measureattraction}
-  \caption{Effect plot for the regressor \textit{Measureattraction} for the masculine\slash neuter and the feminine model (y axes have different scales)}
+  \caption{Effect plot for the regressor \textit{Measureattraction} for the MN model and the F model (y axes have different scales)}
   \label{fig:eff:measureattraction}
@@
 if a lemma appears relatively more often in the \PGCd\ (compared to its frequency in the \NACb), the more often the \PGCa\ is chosen over the \NACa\ with this specific lemma.
-Only the feminine model \textit{Kindattraction} regressor is clearly useless (`non-significant') in the PBmodcomp test, and the effect plot shows a corresponding picture.
-This is related to the fact that in the feminine model the kind lemma effect and its second-level model turned out to be uninformative relative to the alternation (remember that $\sigma_{\text{Kindlemma}}=0.000$).
+Only the F model \textit{Kindattraction} regressor is clearly useless (`non-significant') in the PBmodcomp test, and the effect plot shows a corresponding picture.
+This is related to the fact that in the F model the kind lemma effect and the corresponding second-level model could not be estimated successfully (remember that $\sigma_{\text{Kindlemma}}=0.000$).
 The effect plots show a picture in line with a \textit{divide by 4} interpretation of the coefficients as shown in Table~\ref{tab:bigtable}.%
@@
   \includegraphics[width=0.6\textwidth]{figures/corpus/04_glmm_fixeff_mn_Measureclass}
-  \caption{Effect plot for the regressor \textit{Measureclass} for the masculine\slash neuter model}
+  \caption{Effect plot for the regressor \textit{Measureclass} for the MN model}
   \label{fig:eff:measureattraction}
@@
 
+In Section~\ref{sec:analyses}, it was also hypothesised that there may be effects specific to classes of lemmas and classes of lemmas, and that they might be related to degrees of grammaticalisation of the measure noun and\slash or semantic prototypicality effects.
+The picture is clear for the MN model.
+Lemma-specific effects were not detected ($\sigma_{\text{Measurelemma}}<0.001$), but instead the lemma class second-level predictor was successfully estimated ($p=0.004$ in the PB test).
+The large confidence intervals around the estimates (between $15.6$ and $28.37$) indicate that maybe the sample size needs to be higher, or a slightly modified classification might be even more adequate.
+Looking at the effect plot in Figure~\ref{fig:eff:measureattraction}, however, it is evident that the more abstract physical measure nouns with a high degree of grammaticalisation favour the \NACa\ (\textit{Square}, \textit{Length}, \textit{Currency}, \textit{Capacity}, \textit{Weight}) whereas the more concrete and less grammaticalised classes of measure nouns (which usually have an alternative use outside of MNPs) favour the \PGCa\ (\textit{Layer}, \textit{Amount}, \textit{Vessel}, \textit{Portion}).
+Ideally, there should be more differentiation in the second group, but given the large confidence intervals, significantly larger samples would be needed for more precise estimation.
+As mentioned before, the second level model for measure nouns in the F model could not estimated with meaningful results.
+
+On the other hand, the heuristic semantic classification for kind nouns in \textit{Kindconsistency} is much less useful.
+The MN model did not even converge with this regressor, and for the F model an insufficient $p=0.07$ was achieved in the PB test.
+Also, \textit{Kindint} does not contribute anything useful in either model (MN model $p=0.432$; F model $p=0.326$).
+The only regressor that has a considerable influence is \textit{Kindedible}, \ie, whether the substance denoted by the kind noun is edible.
+Since these regressors were added heuristically, there is no interpretation in terms of substantive theory, however, and I am hesitant to count it as very meaningful.
+In sum, no consistent picture emerges for kind noun-specific semantic prototypicality effects, but for measure nouns there is a clear indication that -- as predicted -- classes of more strongly grammaticalised nouns occur proportionally more frequently in the \PGCa.
+
 % PLURAL & CARDINALS
-%
-%\begin{figure}[h!]
-%  \centering
-%  \includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_mn_Minus1pos}~\includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_fem_Minus1pos}
-%  \caption{Effect plot for the regressor \textit{Minus1pos} for the masculine\slash neuter and the feminine model (y axes have different scales)}
-%  \label{fig:eff:measureattraction}
-%\end{figure}
-%
-%\begin{figure}[h!]
-%  \centering
-%  \includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_mn_Measurenumber}~\includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_fem_Measurenumber}
-%  \caption{Effect plot for the regressor \textit{Measurenumber} for the masculine\slash neuter and the feminine model (y axes have different scales)}
-%  \label{fig:eff:measureattraction}
-%\end{figure}
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_mn_Minus1pos}~\includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_fem_Minus1pos}
+  \caption{Effect plot for the regressor \textit{Minus1pos} for the MN model and the F model (y axes have different scales)}
+  \label{fig:eff:measureattraction}
+\end{figure}
+
+I now turn to the predicted effect of plural measure nouns and cardinals as quantifiers.
+Plurality has a clear effect in the expected direction (plural favours the \PGCa) in the MN model ($p=0.002$ for \textit{Measurenumber}).
+In the F model, the effect has the same direction, but the estimate is quite imprecise ($p=0.160$), again likely due to the low sample size.
+Figure~\ref{fig:eff:measureattraction} shows, however, that the left context of the measure noun influences the choice of the variant (in both models, the regressor has $p=0.002$).
+Cardinals have a strong tendency to co-occur with the \NACa\ in both models.
 
@@
 
+The register-related proxy variables point into the expected direction.
+Increased \textit{Badness} favours the \NACa\ (only in the MN model $\beta=-0.2$, $p=0.002$), and so does a lesser density of genitives (MN model $\beta=-0.66$, $p=0.004$; F model $\beta=-0.52$, $p=0.002$).
+A higher length of the NP (\textit{Matchlength}) favours the \PGCa\ (MN model $\beta=0.54$, $p=0.002$; F model $\beta=0.63$, $p=0.002$) as predicted, but abbreviated measure nouns (\textit{Measureabbreviated}) do not seem to have much of an effect (MN model $p=0.730$; F model $p=0.267$).
+
 % NUISANCE
@@
 % => frequency
+
+Finally, the nuisance variables \textit{Measurefreq} and \textit{Kindfreq} have no clear effect, except for maybe a $p=0.041$ for \textit{Measurefreq} in the MN model.
+The coefficient is $-0.16$ meaning that with each increase of the lemma log-frequency (per one million tokens) of the measure noun of one standard deviation, the probability that the \PGCa\ is chosen is lowered by approximately $-0.16/4=-0.04$.
+Since token frequency is no direct proxy for grammaticalisation, I hesitate to interpret this effect.
+The effect for \textit{Measurecase} is as predicted by previous studies (see Section~\ref{sec:analyses}).
+The dative favours the \PGCa\ in the MN model with $p=0.002$ and a coefficient for the dative (vs.\ the nominative, which is on the intercept) of $0.63$.
+In the F model, \textit{Dative} was not a level of \textit{Measurecase} due to the syncretisms mentioned in Section~\ref{sec:descriptive}.
+Although \textit{Measurecase} is a nuisance variable in the context of this study, convergence with previous work strengthens its validity.
 

### 2017-03-27 R/00_wrapper.R

@@ -60,16 +60,6 @@ source('02_attraction.R') # calculate attraction factors
 
 t.attracted <- Sys.time()
 
-cat("\n\n EXPLORE ...\n")
-source('03_explore.R')    # exploratory analysis (determine order of adding regressors)
-
-t.explored <- Sys.time()
-
-cat("\n\n GLM ...\n")
-source('04_glm.R')        # GLMs
-
-t.glm <- Sys.time()
-
 cat("\n\n GLMM ...\n")
 source('05_glmm.R')       # full GLMMs
 
@@ -87,9 +77,7 @@ cat("Total:", t.mcmc - t.start)
 
 cat("... Loading:", t.loaded - t.start, "\n")
 cat("... Attraction:", t.attracted - t.loaded, "\n")
-cat("... Exploring:", t.explored - t.attracted, "\n")
-cat("... GLMs:", t.glm - t.explored, "\n")
-cat("... GLMMs:", t.glmm - t.glm, "\n")
+cat("... GLMMs:", t.glmm - t.attracted, "\n")
 cat("... MCMC:", t.mcmc - t.glmm, "\n\n")
 
 cat("\n\nThis run ENDED at ")

### 2017-03-27 R/01_load.R

@@ -9,7 +9,6 @@ linMap <- function(x, from, to)
   (x - min(x)) / max(x - min(x)) * (to - from) + from
 
 m <- read.gnumeric.sheet(file = "data/nan_sample_KM.ods", head = TRUE, sheet.name = "Data")
-pl <- read.gnumeric.sheet(file = "data/countnouns_sample_KM.ods", head = TRUE, sheet.name = "Data")
 
 # Backup original sentences.
 
@@ -22,42 +21,15 @@ m <- m[,c("Measurecase",
           "Kindgender",
           "Kindcase",
           "Kindlength", 
-          "Kindorigin",
           "Kindconsistency",
           "Kindedible",
           "Kindfreq",
           "Measurelemma",
           "Measurenumber",
           "Measureabbreviated",
-          "Measurelength",
           "Measureclass",
           "Measurefreq",
           "Minus1pos",
-          "Minus2pos",
-          "Matchlength",
-          "Genitives",
-          "Badness",
-          "Originalsort"
-)]
-
-pl <- pl[,c("Measurecase",
-          "Kindlemma",
-          "Kindgender",
-          "Kindcase",
-          "Kindlength",
-          # "Kindorigin",
-          # "Kindconsistency",
-          # "Kindedible",
-          "Kindfreq",
-          "Measurelemma",
-          "Measurenumber",
-          "Measureabbreviated",
-          "Measurelength",
-          # "Measureclass",
-          "Measurefreq",
-          "Minus1pos",
-          "Minus2pos",
-          "Matchlength",
           "Genitives",
           "Badness",
           "Originalsort"
@@ -69,10 +41,6 @@ pl <- pl[,c("Measurecase",
 m <- m[which(as.character(m$Measurecase) %in% c("Nom","Acc","Dat")),]
 m <- m[which(as.character(m$Kindcase) %in% c("Acc", "Dat", "Nom", "Gen", "Obl", "Str", "Nomacc")),]
 m <- m[-which(m$Kindlemma %in% c("Sonnenschein","Silber","Obst","Strom")),]
[…]

### 2017-03-27 R/02_attraction.R

@@ -13,15 +13,13 @@ prop.func <- function(v, smooth = 1) {
 ndn.kind <- read.delim("data/ndn_kind_freq.tsv", header=FALSE, quote="")
 nn.kind <- read.delim("data/nn_kind_freq.tsv", header=FALSE, quote="")
 
-# Pl & Measure have been pre-processed differently, can load in one table.
-countnouns_attracfreq <- read.delim("data/countnouns_attracfreq.tsv", quote="")
+# Measure nousn have been pre-processed differently, can load in one table.
 measurenouns_attracfreq <- read.delim("data/measurenouns_attracfreq.tsv", quote="")
 
 # A cleanup.
-nn.kind <- nn.kind[-which(nn.kind$Lem=="rege"),]
 colnames(ndn.kind) <- c("Freq","Lem")
 colnames(nn.kind) <- c("Freq","Lem")
-
+nn.kind <- nn.kind[-which(nn.kind$Lem=="rege"),]
 
 # Get all lemmas for master table.
 # Background: We want values for all nouns, even those which occur in
@@ -29,7 +27,7 @@ colnames(nn.kind) <- c("Freq","Lem")
 cx.kind.lem <- c(as.character(nn.kind$Lem), as.character(ndn.kind$Lem))
 cx.kind.lem <- unique(cx.kind.lem[order(cx.kind.lem)])
 
-# Function to extract freqs for both constructions with "add 1".
+# Function to extract freqs for both constructions.
 get_freq <- function(data, lem) {
   n1 <- data[which(data$Lem == lem),1]
   if (length(n1) == 0) n1 <- 0
@@ -39,190 +37,35 @@ get_freq <- function(data, lem) {
 # Generate full table.
 cx.kind <- cbind(cx.kind.lem, lapply(cx.kind.lem, function(lem){get_freq(nn.kind,lem)}), lapply(cx.kind.lem, function(lem){get_freq(ndn.kind,lem)}))
 
-# A rather simple "attraction" function, i.e., quotients.
-attraction <- function(cs, smooth=1) {
-  (unlist(cs)[2]+smooth) / (unlist(cs)[1]+smooth)
+# A rather simple "attraction" function, i.e., proportion.
[…]

### 2017-03-27 R/03_explore.R

@@ -1,154 +1,0 @@
-#
-# Explore by putting each potential regressor in simple GLM.
-# ( Just to determine an order in which to try and add them to GLMM. )
-#
-
-
-# A function to do simple cross-validation.
-
-crossval.glm <- function(model, K = 10) {
-  data <- model$data
-  response <- all.vars(model$formula)[1]
-  n <- nrow(data)
-  z <- n%/%K
-  mod <- n%%K
-  
-  # Randomize order of data.
-  idx <- sample(1:n, n, replace=F)
-  data <- data[idx,]
-  
-  corrs <- c()
-  
-  # Iterate over the folds.
-  for (i in 1:K) {
-    
-    if (i==K)
-      idx.heldout <- seq((i-1)*z+1, i*z+mod, 1)
-    else
-      idx.heldout <- seq((i-1)*z+1, i*z, 1)
-    
-    data.train    <- data[-idx.heldout,]
-    data.heldout  <- data[idx.heldout,]
-    
-    tryCatch({
-      model.new <- glm(model$formula, family=model$family, data=data.train)
-      preds <- ifelse(predict(model.new, newdata=data.heldout) < 0.5, 0, 1)
-      confmat <- table(data.heldout[,response], preds)
-      corr <- sum(diag(confmat))
-      rate <- 1-corr/sum(confmat)
-      corrs <- c(corrs, rate)
-    },
-    error = function(err) {
-      cat("Hoppala! Convergence error...")
-      # Yes, we just ignore errors! It's all about non-converging GLMs anyway.
-    })
-  }
-  corr.armean <- 1/mean(1/corrs)
-  corr.armean
-}
-
-
-# A function that performs a simple glm for each potential regressor.
-# Should give a hint in which order to try and add regressors to complex model.
-
-order.regressors <- function(df, target, ignores) {
-  require(fmsb)
-  require(boot)
-  
-  mat <- matrix(, nrow=length(names(df)), ncol=3)
-  colnames(mat) <- c("R2", "p(LRT)", "Error(10CV)")
-  rownames(mat) <- names(df)
-  
-  model.0 <- glm(df[,target]~1, family=binomial(link=logit))
-  i = 1
-  for (n in names(df)) {
-    if (!(n %in% ignores)) {
-      #cat(n, "\n") # To Debug models that dont work.
[…]

### 2017-03-27 R/04_glm.R

@@ -1,137 +1,0 @@
-require(fmsb)
-
-
-###### MASCULINE ######
-
-mn.glm <- glm(Genitive~1
-              +Measurelemma
-              #+Kindlemma       # prob 0 or 1 
-              +Genitives
-              +Minus1pos
-              #+Kindlength       # unstable in bootstrap
-              #+Measureclass     # prob 0 or 1
-              +Matchlength
-              +Measureabbreviated
-              +Measureattraction
-              +Measurelength
-              +Kindattraction
-              +Measurenumber
-              +Kindedible
-              +Badness
-              +Kindconsistency
-              +Kindorigin
-              +Kindfreq
-              #+Minus2pos       # prob 0 or 1 
-              +Measurecase
-              +Kindint
-              +Kindgender
-#              +Measurefreq
-              , data=mn,  family=binomial(link=logit))
-
-mn.glm.r2 <- NagelkerkeR2(mn.glm)$R2
-mn.glm.pred <- ifelse(predict(mn.glm) < 0, 0, 1)
-mn.glm.cmat <- table(mn.glm.pred, mn$Genitive)
-mn.glm.corr <- sum(diag(mn.glm.cmat))/sum(mn.glm.cmat)
-mn.glm.base <- length(which(mn$Genitive == 0))/length(mn$Genitive)
-mn.glm.pre <- (mn.glm.corr-mn.glm.base)/(1-mn.glm.base)
-
-if (save.persistent) sink(paste(out.dir, "04_glm.txt", sep=""))
-cat("\n\n\nMASCULINE\n\n")
-print(summary(mn.glm, correlation=F))
-cat("\n\n")
-print(mn.glm.r2)
-cat("\n\n")
-cat("correct", mn.glm.corr)
-cat("\n\n")
-cat("λ", mn.glm.pre)
-cat("\n\n")
-if (save.persistent) sink()
-
-
-
-###### FEMININE ######
-
-
-fem.glm <- glm(Casedrop~1
-              +Measurelemma
-              #+Kindlemma       # prob 0 or 1
-              +Measureclass
-              +Minus1pos
-              +Measureabbreviated
-              +Kindconsistency
-              +Matchlength
-              #+Measureattraction# prob 0 or 1
-              +Genitives
-              +Kindedible
-              +Kindorigin
[…]

### 2017-03-27 R/05_glmm.R

@@ -88,21 +88,16 @@ mn.glmm <- glmer(Genitive~1
                 +(1|Kindlemma)
                 +Genitives
                 +Minus1pos
-                #+Kindlength          # unstable in bootstrap
                 +Measureclass
-                +Matchlength
                 +Measureabbreviated
                 +Measureattraction
-                #+Measurelength       # virtual non-convergence
                 +Kindattraction
                 +Measurenumber
                 +Kindedible
                 +Badness
-                #+Kindconsistency     # virtual non-convergence
-                #+Kindorigin          # virtual non-convergence
+                +Kindconsistency
                 +Kindfreq
                 +Measurecase
-                +Kindint
                 +Measurefreq
               , data=mn, family=binomial(link=logit), na.action = na.fail, nAGQ=the.nAGQ,
               control=glmerControl(optimizer="nloptwrap2", optCtrl=list(maxfun=2e5)))
@@ -111,9 +106,9 @@ mn.glmm <- glmer(Genitive~1
 # Regressors to be used in getting LRT w/ bootstrapped CIs.
 
 mn.bootcomp.regs <- c("(1|Measurelemma)", "(1|Kindlemma)", "Genitives", "Minus1pos",
-                   "Measureclass", "Matchlength", "Measureabbreviated", "Measureattraction",
+                   "Measureclass", "Measureabbreviated", "Measureattraction",
                    "Kindattraction", "Measurenumber", "Kindedible",
-                   "Badness", "Kindfreq", "Measurecase", "Kindint", "Measurefreq")
+                   "Badness", "Kindconsistency", "Kindfreq", "Measurecase", "Measurefreq")
 
 # Get PB test.
 mn.modelcomp <- lmer.modelcomparison(model = mn.glmm, regressors = mn.bootcomp.regs,
@@ -166,31 +161,26 @@ if (save.persistent) sink()
 fem.glmm <- glmer(Casedrop~1
                  +(1|Measurelemma)
                  +(1|Kindlemma)
[…]

### 2017-03-27 R/06_glmm-mcmc.R

@@ -17,21 +17,16 @@ mn.glmm.mcmc <- stan_glmer(Genitive~1
                            +(1|Kindlemma)
                            +Genitives
                            +Minus1pos
-                           #+Kindlength          # unstable in bootstrap
                            +Measureclass
-                           +Matchlength
                            +Measureabbreviated
                            +Measureattraction
-                           #+Measurelength       # virtual non-convergence
                            +Kindattraction
                            +Measurenumber
                            +Kindedible
                            +Badness
-                           #+Kindconsistency     # virtual non-convergence
-                           #+Kindorigin          # virtual non-convergence
+                           +Kindconsistency
                            +Kindfreq
                            +Measurecase
-                           +Kindint
                            +Measurefreq
                            ,
                  data=mn, family=binomial(link=logit),
@@ -50,24 +45,19 @@ mn.baysum <- summary(mn.glmm.mcmc)
 fem.glmm.mcmc <- stan_glmer(Casedrop~1
                             +(1|Measurelemma)
                             +(1|Kindlemma)
-                            #+Measureclass         # non-convergence
+                            +Measureclass
                             +Minus1pos
                             +Measureabbreviated
                             +Kindconsistency
-                            +Matchlength
                             +Measureattraction
                             +Genitives
                             +Kindedible
-                            #+Kindorigin           # fixeff matrix rank deficient
-                            #+Measurelength        # unstable in bootstrap
                             +Measurenumber
-                            +Kindint
                             +Badness
                             +Measurefreq
                             +Measurecase
                             +Kindfreq
                             +Kindattraction
-                            #+Kindlength         # unstable in bootstrap
                   ,
                   data=fem, family=binomial(link=logit),
                   chains=chains, seed=seed, iter=iter,
@@ -80,37 +70,6 @@ fem.glmm.mcmc <- stan_glmer(Casedrop~1
 fem.baysum <- summary(fem.glmm.mcmc)
 
 
-# PL
-
-pl.glmm.mcmc <- stan_glmer(Casedrop~1
-                           +(1|Kindlemma)
-                           #+(1|Measurelemma)    # model unidentifiable
-                           #+Minus1pos           # model unidentifiable
-                           +Attraction
-                           #+Genitives           # model unidentifiable
[…]

### 2017-03-27 R/11_stimuli_predict.R

@@ -6,7 +6,6 @@ stims.fax  <- c("Measurecase",
           "Kindlemma",
           "Kindconsistency",
           "Kindedible",
-          "Kindint",
           "Measurelemma",
           "Measurenumber",
           "Measureabbreviated",
@@ -35,24 +34,15 @@ get.value <- function(df, index.column, index.value, value.column) {
 }
 
 stimuli.mn$Kindfreq <- unlist(lapply(as.character(stimuli.mn$Kindlemma), function(x) get.value(mn, "Kindlemma", x, "Kindfreq")))
-stimuli.mn$Measurelength <- unlist(lapply(as.character(stimuli.mn$Measurelemma), function(x) get.value(mn, "Measurelemma", x, "Measurelength")))
 stimuli.mn$Measurefreq <- unlist(lapply(as.character(stimuli.mn$Measurelemma), function(x) get.value(mn, "Measurelemma", x, "Measurefreq")))
-stimuli.mn$Attraction <- unlist(lapply(as.character(stimuli.mn$Kindlemma), function(x) get.value(mn, "Kindlemma", x, "Attraction")))
+stimuli.mn$Kindattraction <- unlist(lapply(as.character(stimuli.mn$Kindlemma), function(x) get.value(mn, "Kindlemma", x, "Kindattraction")))
+stimuli.mn$Measureattraction <- unlist(lapply(as.character(stimuli.mn$Measurelemma), function(x) get.value(mn, "Measurelemma", x, "Measureattraction")))
 
 stimuli.fem$Kindfreq <- unlist(lapply(as.character(stimuli.fem$Kindlemma), function(x) get.value(fem, "Kindlemma", x, "Kindfreq")))
-stimuli.fem$Measurelength <- unlist(lapply(as.character(stimuli.fem$Measurelemma), function(x) get.value(fem, "Measurelemma", x, "Measurelength")))
 stimuli.fem$Measurefreq <- unlist(lapply(as.character(stimuli.fem$Measurelemma), function(x) get.value(fem, "Measurelemma", x, "Measurefreq")))
-stimuli.fem$Attraction <- unlist(lapply(as.character(stimuli.fem$Kindlemma), function(x) get.value(fem, "Kindlemma", x, "Attraction")))
[…]

### 2017-03-27 R/00_wrapper.R
stats reset

@@ -1,90 +1,0 @@
-#
-# This calls all scripts in the right order to replicate results from my paper
-# 'Competing Constructions for German Measure Phrases' (working title)
-#
-# Warning: This takes quite a long times to run, especially with 
-#  =>  model coef & LR BOOTSTRAPPING
-#  =>  BAYESIAN / MCMC estimation
-
-# Comment this line if you want to keep your old workspace.
-rm(list = ls())
-
-# Set this to the location of R script sources.
-setwd('~/Massangaben/')
-
-# OPTIONS
-
-set.seed(2398651)            # Do NOT change, or results will differ.
-save.persistent <- T         # write to files instead of screen/console
-out.dir         <- "output/" # put generated files here
-
-round.in.big.table <- 2
-
-# Comment the next line if you want to keep old output files.
-unlink(paste(out.dir, '*', sep=""))
-
-# ----------   Don't touch anything past this line   ---------- #
-# ---------- If you do, results might not replicate. ---------- #
-
-# HELPERS
-
-source('glmtools.R')
-
-
-# SCRIPTS
-
-cat("\014")
-
-if (save.persistent) sink(paste(out.dir, "00_wrapper.txt", ""), append=F)
-cat("\nRoland Schäfer, Freie Universität Berlin\n")
-cat("roland.schaefer@fu-berlin.de\n")
-cat("http://orcid.org/0000-0003-3233-787\n")
-cat("'Competing Constructions for German Measure Phrases'\n")
-cat("Full results of corpus analysis as produced by R.\n")
-cat("Run 'wrapper.R' for 100% reproducible results!")
-cat("\n\nThis run BEGAN at ")
-print(Sys.time())
-if (save.persistent) sink()
-
-# Steps 1 and 2 are required for the subsequent steps.
-
-t.start <- Sys.time()
-
-cat("\n\n LOADING ...\n")
-source('01_load.R')       # imports and preparation/transformation of the data
-
-t.loaded <- Sys.time()
-
-cat("\n\n ATTRACTION ...\n")
-source('02_attraction.R') # calculate attraction factors
-
-t.attracted <- Sys.time()
-
-cat("\n\n GLMM ...\n")
-source('05_glmm.R')       # full GLMMs
-
-t.glmm <- Sys.time()
-
-cat("\n\n GLMM with MCMC ...\n")
[…]

### 2017-03-27 R/01_load.R

@@ -1,139 +1,0 @@
-# 
-#  LOAD DATA
-#
-
-require(gnumeric)
-require(plyr)
-
-linMap <- function(x, from, to)
-  (x - min(x)) / max(x - min(x)) * (to - from) + from
-
-m <- read.gnumeric.sheet(file = "data/nan_sample_KM.ods", head = TRUE, sheet.name = "Data")
-
-# Backup original sentences.
-
-sentences <- m[, c("Leftcontext", "Match", "Rightcontext", "Originalsort", "Url")]
-
-# Remove some columns. These are the ones we KEEP:
-
-m <- m[,c("Measurecase",
-          "Kindlemma",
-          "Kindgender",
-          "Kindcase",
-          "Kindlength", 
-          "Kindconsistency",
-          "Kindedible",
-          "Kindfreq",
-          "Measurelemma",
-          "Measurenumber",
-          "Measureabbreviated",
-          "Measureclass",
-          "Measurefreq",
-          "Minus1pos",
-          "Genitives",
-          "Badness",
-          "Originalsort"
-)]
-
-
-# Filter data points that we cannot use.
-
-m <- m[which(as.character(m$Measurecase) %in% c("Nom","Acc","Dat")),]
-m <- m[which(as.character(m$Kindcase) %in% c("Acc", "Dat", "Nom", "Gen", "Obl", "Str", "Nomacc")),]
-m <- m[-which(m$Kindlemma %in% c("Sonnenschein","Silber","Obst","Strom")),]
-
-
-# Simplify POS tags.
-
-m$Minus1pos <- as.character(m$Minus1pos)
-m$Minus1pos <- ifelse(m$Minus1pos %in% c("Art", "Pdat", "Pds", "Piat", "Pis", "Pposat"), "D", m$Minus1pos)
-m$Minus1pos <- ifelse(m$Minus1pos %in% c("Appr", "Apprart"), "P", m$Minus1pos)
-m$Minus1pos <- ifelse(m$Minus1pos == "Adja", "A", m$Minus1pos)
-m$Minus1pos <- ifelse(m$Minus1pos == "Card", "C", m$Minus1pos)
-m$Minus1pos <- ifelse(m$Minus1pos %in% c("D", "P", "A", "C"), m$Minus1pos, "O")
-m$Minus1pos <- as.factor(m$Minus1pos)
-
-
-# Simplify "Measureclass".
-
[…]

### 2017-03-27 R/02_attraction.R

@@ -1,71 +1,0 @@
-# This creates the quotient database for lemmas in the NN and NDN construction.
-# It also creates some plots and a text file for illustration.
-
-require(boot)
-
-# How the plural lemma pair list was extracted from the filtered list in data frame pl:
-# write.table(unique(sort(apply(pl[,c("Measurelemma", "Kindlemma")], 1, function(x) paste(x[1], x[2])))), "plural_pairs.txt", sep="\t", quote=F, col.names=F, row.names=F)
-
-prop.func <- function(v, smooth = 1) {
-  log( (v[2]+smooth) / (v[1]+smooth), 10)
-}
-
-ndn.kind <- read.delim("data/ndn_kind_freq.tsv", header=FALSE, quote="")
-nn.kind <- read.delim("data/nn_kind_freq.tsv", header=FALSE, quote="")
-
-# Measure nousn have been pre-processed differently, can load in one table.
-measurenouns_attracfreq <- read.delim("data/measurenouns_attracfreq.tsv", quote="")
-
-# A cleanup.
-colnames(ndn.kind) <- c("Freq","Lem")
-colnames(nn.kind) <- c("Freq","Lem")
-nn.kind <- nn.kind[-which(nn.kind$Lem=="rege"),]
-
-# Get all lemmas for master table.
-# Background: We want values for all nouns, even those which occur in
-# only one of the NN/NDN tables. If not in one, count will just be 0.
-cx.kind.lem <- c(as.character(nn.kind$Lem), as.character(ndn.kind$Lem))
-cx.kind.lem <- unique(cx.kind.lem[order(cx.kind.lem)])
-
-# Function to extract freqs for both constructions.
-get_freq <- function(data, lem) {
-  n1 <- data[which(data$Lem == lem),1]
-  if (length(n1) == 0) n1 <- 0
-  n1
-}
-
-# Generate full table.
-cx.kind <- cbind(cx.kind.lem, lapply(cx.kind.lem, function(lem){get_freq(nn.kind,lem)}), lapply(cx.kind.lem, function(lem){get_freq(ndn.kind,lem)}))
-
-# A rather simple "attraction" function, i.e., proportion.
-attraction <- function(cs) {
-  unlist(cs)[2] / (unlist(cs)[1] + unlist(cs)[2])
-}
-
-
-
-# Generate attraction quotients for full table.
-cx.kind <- as.data.frame(cx.kind)
[…]

### 2017-03-27 R/05_glmm.R

@@ -1,470 +1,0 @@
-require(lme4)
-require(gridExtra)
-require(MuMIn)
-require(effects)
-require(gridExtra)
-
-
-# ######### Some settings, conveniently at top of file. #########
-
-the.nAGQ <- 0         # 0 for speed, 1 for precision and slow(!) computation.
-ci.boot.nsim <- 1000  # Even 100 can be extremely slow! SET TO MIN. 1000 FOR PRODUCTION!!!
-ci.boot.modelcomp.nsim <- 500 # This takes ages, even with 10. Set to 500 for PRODUCTION!!!
-
-# ###############################################################
-
-
-# Handling convergence warnings:
-# https://rstudio-pubs-static.s3.amazonaws.com/33653_57fc7b8e5d484c909b615d8633c01d51.html
-
-# A faster BOBYQA optimizer.
-library(nloptr)
-defaultControl <- list(algorithm="NLOPT_LN_BOBYQA",xtol_rel=1e-6,maxeval=1e5)
-nloptwrap2 <- function(fn,par,lower,upper,control=list(),...) {
-  for (n in names(defaultControl)) 
-    if (is.null(control[[n]])) control[[n]] <- defaultControl[[n]]
-  res <- nloptr(x0=par,eval_f=fn,lb=lower,ub=upper,opts=control,...)
-  with(res,list(par=solution,
-                fval=objective,
-                feval=iterations,
-                conv=if (status>0) 0 else status,
-                message=message))
-}
-
-
-# Function to get prop of correct predictions.
-corr.prop <- function(model, observed, cutoff) {
-  preds <- ifelse(predict(model) < cutoff, 0, 1)
-  cmat <- table(preds, observed)
-  sum(diag(cmat))/sum(cmat)
-}
-
-
-# Function to get LRT for (G)LMMMs with bootstrapped CIs. This takes AGES!
-lmer.modelcomparison <- function(model, regressors, formula.target, ci = 0.95, nsim = 4, print.updated = F, print.diag = F) {
-  require(pbkrtest)
-
-  result = NULL
-  
-  for (i in 1:length(regressors)) {
-    formula.parts <- c(formula.target, regressors[1:i-1], regressors[i+1:length(regressors)])
[…]

### 2017-03-27 R/06_glmm-mcmc.R

@@ -1,109 +1,0 @@
-require(rstanarm)
-require(rstan)
-
-rstan_options(auto_write = TRUE)
-options(mc.cores = parallel::detectCores())
-
-# MASCULINE GLMM
-
-chains <- 4
-seed   <- 6976
-iter   <- 1000 # Set to 1000 for production.
-
-# MN
-
-mn.glmm.mcmc <- stan_glmer(Genitive~1
-                           +(1|Measurelemma)
-                           +(1|Kindlemma)
-                           +Genitives
-                           +Minus1pos
-                           +Measureclass
-                           +Measureabbreviated
-                           +Measureattraction
-                           +Kindattraction
-                           +Measurenumber
-                           +Kindedible
-                           +Badness
-                           +Kindconsistency
-                           +Kindfreq
-                           +Measurecase
-                           +Measurefreq
-                           ,
-                 data=mn, family=binomial(link=logit),
-                 chains=chains, seed=seed, iter=iter,
-                 prior = normal(0, 2.5), prior_intercept = normal(0, 10),
-                 prior_ops = prior_options(prior_scale_for_dispersion = 5, min_prior_scale = 1e-12, scaled = TRUE),
-                 prior_covariance = decov(regularization = 1, concentration = 1, shape = 1, scale = 1),
-                 prior_PD = F
-                 )
-
-mn.baysum <- summary(mn.glmm.mcmc)
-
-
-# FEM
-
-fem.glmm.mcmc <- stan_glmer(Casedrop~1
-                            +(1|Measurelemma)
-                            +(1|Kindlemma)
-                            +Measureclass
-                            +Minus1pos
-                            +Measureabbreviated
-                            +Kindconsistency
-                            +Measureattraction
-                            +Genitives
-                            +Kindedible
-                            +Measurenumber
-                            +Badness
-                            +Measurefreq
-                            +Measurecase
-                            +Kindfreq
-                            +Kindattraction
[…]

### 2017-03-27 R/10_sel_exp_stimuli.R

@@ -1,105 +1,0 @@
-library(plyr)
-
-mn.selectors <- c("Minus1pos", "Measurenumber", "Measurecase")
-prop.mn <- 0.15
-
-fem.selectors <- c("Minus1pos", "Measurenumber", "Measurecase")
-prop.fem <- 0.25
-
-sel.no <- 7
-per.cmb <- 2
-
-exp_sample <- function(df, model, selectors, do.tail = F, prop = 0.25, find = 5, per.combo = 1, sel.feature = "Genitive") {
-  pred.order <- order(predict(model))
-
-  if (do.tail) {
-    pred.osort <- df[tail(pred.order, prop*nrow(df)), "Originalsort"]
-  } else {
-    pred.osort <- df[head(pred.order, prop*nrow(df)), "Originalsort"]
-  }
-
-  exp.sel <- df[which(df$Originalsort %in% pred.osort), ]          # The actual data (part of big data frame).
-  exp.conf <- ddply(exp.sel, selectors, nrow)                      # Just the feat/val configs of top pred cases.
-  exp.conf  <- exp.conf[order(exp.conf$V1, decreasing = T),]       # The top summary/counting of feat/val configs.
-  cat("Found feature configurations: ", nrow(exp.conf), " (showing 5)\n")
-  print(head(exp.conf, 5))
-
-  seen.knouns <- character()
-  seen.mnouns <- character()
-  
-  found <- 0
-  i <- 0
-  
-  # For each feature permutation, find unseen kind/measure lemmas.
-  while (found < find) {
-    i <- i + 1
-        
-    cat("\nFeatures selected: ", paste(as.character(unlist(exp.conf[i, 1:ncol(exp.conf)-1])), collapse="_"), "\n")
-    
-    if (do.tail) { sel.feature.val <- 1 } else { sel.feature.val <- 0 }
-    tmp.sel <- exp.sel[which(exp.sel[,sel.feature] == sel.feature.val & unname(apply(exp.sel[,selectors], 1, function(x) { paste(as.character(x), collapse="_") })) == paste(as.character(unlist(exp.conf[i, 1:ncol(exp.conf)-1])), collapse="_")),]
-    
-    tmp.top.kind <- ddply(tmp.sel, "Kindlemma", nrow)
[…]

### 2017-03-27 R/11_stimuli_predict.R

@@ -1,71 +1,0 @@
-require(lme4)
-
-stimuli <- read.csv("~/Massangaben/data/stimuli.csv", header=TRUE)
-
-stims.fax  <- c("Measurecase",
-          "Kindlemma",
-          "Kindconsistency",
-          "Kindedible",
-          "Measurelemma",
-          "Measurenumber",
-          "Measureabbreviated",
-          "Measureclass",
-          "Minus1pos")
-
-
-# Separate MN and F.
-
-stimuli.mn <- stimuli[1:8,]
-stimuli.fem <- stimuli[9:16,]
-
-
-# Make factors factors, using levels from original model df.
-
-for (fac in stims.fax) {
-  stimuli.mn[, fac]  <- factor(stimuli.mn[, fac], levels = levels(mn[, fac]))
-  stimuli.fem[, fac]  <- factor(stimuli.fem[, fac], levels = levels(fem[, fac]))
-}
-
-
-# Pull numeric values from original table because of standardization.
-
-get.value <- function(df, index.column, index.value, value.column) {
-  df[which(df[, index.column] == index.value)[1], value.column]
-}
-
-stimuli.mn$Kindfreq <- unlist(lapply(as.character(stimuli.mn$Kindlemma), function(x) get.value(mn, "Kindlemma", x, "Kindfreq")))
-stimuli.mn$Measurefreq <- unlist(lapply(as.character(stimuli.mn$Measurelemma), function(x) get.value(mn, "Measurelemma", x, "Measurefreq")))
-stimuli.mn$Kindattraction <- unlist(lapply(as.character(stimuli.mn$Kindlemma), function(x) get.value(mn, "Kindlemma", x, "Kindattraction")))
-stimuli.mn$Measureattraction <- unlist(lapply(as.character(stimuli.mn$Measurelemma), function(x) get.value(mn, "Measurelemma", x, "Measureattraction")))
-
-stimuli.fem$Kindfreq <- unlist(lapply(as.character(stimuli.fem$Kindlemma), function(x) get.value(fem, "Kindlemma", x, "Kindfreq")))
-stimuli.fem$Measurefreq <- unlist(lapply(as.character(stimuli.fem$Measurelemma), function(x) get.value(fem, "Measurelemma", x, "Measurefreq")))
[…]

### 2017-03-27 R/glmtools.R

@@ -1,157 +1,0 @@
-join.factors <- function(...) {
-  unlist(list(...))
-}
-
-factorify <- function(f) {
-  levelsize <- function(l) { s = length(which(f==l)) }
-  f <- factor(f)
-  o = levels(f)[order(unlist(lapply(levels(f), levelsize)), decreasing = T)]
-  f <- factor(f, o)
-}
-
-# Sort factor to put most r(esponse)=0 in f(actor) first,
-factorify.to.zero <- function(f, r, dec=F) {
-  f <- factor(f)
-  r <- factor(r)
-  levs <- colnames(table(r, f))[order(table(r, f)[1,]/table(r, f)[2,],decreasing=dec)]
-  f <- factor(f, levels=levs)
-}
-
-
-siglev <- function(p)
-{
-  if (p <= 0.1)
-    if (p <= 0.05)
-      if (p <= 0.01)
-        if (p <= 0.001)
-          sig <- "***"
-  else
-    sig <- "**"
-  else
-    sig <- "*"
-  else
-    sig <- "."
-  else
-    sig <- ""
-}
-
-siglev <- function(p)
-{
-  if (p <= 0.1)
-    if (p <= 0.05)
-      if (p <= 0.01)
-        if (p <= 0.001)
-          sig <- "***"
-  else
-    sig <- "**"
-  else
-    sig <- "*"
-  else
-    sig <- "."
-  else
-    sig <- ""
-}
-
-# This functions makes a plot of the COEFFICIENTS of significant factors.
-plot.coef <- function(glm, alpha = 0.05, pr="Pr(>|z|)", cex.labels = 1, pos.labels = 0.4, ...)
-{
-  # Extract p values and coefs.
-  p <- coef(summary(glm))[,pr]
-  c <- coef(glm)
-  
-  # Separate intercept from other coefs.
-  p.int <- as.numeric(p[1])
-  c.int <- as.numeric(c[1])
-  p <- p[-1]
-  c <- c[-1]
-  
-  # Find the order in which to put coefficients.
-  c <- c[which(p <= alpha)]
-  p <- p[which(p <= alpha)]
-  
-  # Find the order according to coef.
-  order <- order(c)
-  c <- c[order]
-  p <- p[order]
-  names <- names(c)
-  
-  # Put intercept and coefs back together.
-  names <- c("(Intercept)", names)
-  vals <- c(round(c.int, 3), round(c, 3))
[…]

### 2017-03-27 R/0_wrapper.R
new stats

@@ -1,0 +1,49 @@
+#
+# This calls all scripts in the right order to replicate results from my paper
+# 'Competing Constructions for German Measure Phrases' (working title)
+#
+# Warning: This takes quite a long times to run, especially with 
+#  =>  model coef & LR BOOTSTRAPPING
+#  =>  BAYESIAN / MCMC estimation
+
+# Comment this line if you want to keep your old workspace.
+rm(list = ls())
+
+# Set this to the location of R script sources.
+setwd('/Users/user/Workingcopies/measurenps/GitHub/R')
+
+# OPTIONS
+
+set.seed(2398651)            # Do NOT change, or results will differ.
+save.persistent <- F         # write to files instead of screen/console
+out.dir         <- "output/" # put generated files here
+
+round.in.big.table <- 2
+
+# Comment the next line if you want to keep old output files.
+unlink(paste(out.dir, '*', sep=""))
+
+# ----------   Don't touch anything past this line   ---------- #
+# ---------- If you do, results might not replicate. ---------- #
+
+# HELPERS
+
+source('glmtools.R')
+
+
+# SCRIPTS
+
+cat("\014")
+
+cat("\n\n LOADING ...\n")
+source('1_load.R')       # imports and preparation/transformation of the data
+
+cat("\n\n ATTRACTION ...\n")
+source('2_attraction.R') # calculate attraction factors
+
+cat("\n\n GLMM ...\n")
+source('3_glmm.R')       # full GLMMs
+
+#cat("\n\n GLMM with MCMC ...\n")
+#source('4_glmm-mcmc.R')  # same GLMMs estimated with Stan (Bayesian/MCMC)
+

### 2017-03-27 R/1_load.R

@@ -1,0 +1,109 @@
+require(gnumeric)
+require(plyr)
+
+
+# Read data.
+m <- read.csv(file = "data/nan_sample.csv", head = TRUE, sep = '\t', quote = "")
+
+
+# Backup original sentences.
+sentences <- m[, c("Leftcontext", "Match", "Rightcontext", "Originalsort", "Url")]
+
+
+# Remove some columns. These are the ones we KEEP:
+m <- m[,c("Measurecase",
+          "Kindlemma",
+          "Kindgender",
+          "Kindcase",
+          "Kindfreq",
+          "Measurelemma",
+          "Measurenumber",
+          "Measureclass",
+          "Measurefreq",
+          "Minus1pos",
+          "Genitives",
+          "Badness",
+          "Originalsort"
+)]
+
+
+
+
+# Rename some columns for better readability in paper.
+colnames(m)[which(colnames(m) == "Minus1pos")] <- "Leftcontext"
+
+
+# Filter data points that we cannot use.
+m <- m[which(as.character(m$Measurecase) %in% c("Nom","Acc","Dat")),]
+m <- m[which(as.character(m$Kindcase) %in% c("Acc", "Dat", "Nom", "Gen", "Obl", "Str", "Nomacc")),]
+m <- m[-which(m$Kindlemma %in% c("Sonnenschein","Silber","Obst","Strom")),]
+
+
+# Simplify POS tags.
+m$Leftcontext <- as.character(m$Leftcontext)
+m$Leftcontext <- ifelse(m$Leftcontext %in% c("Art", "Pdat", "Pds", "Piat", "Pis", "Pposat"), "D", m$Leftcontext)
+m$Leftcontext <- ifelse(m$Leftcontext %in% c("Appr", "Apprart"), "P", m$Leftcontext)
+m$Leftcontext <- ifelse(m$Leftcontext == "Adja", "A", m$Leftcontext)
+m$Leftcontext <- ifelse(m$Leftcontext == "Card", "C", m$Leftcontext)
+m$Leftcontext <- ifelse(m$Leftcontext %in% c("D", "P", "A", "C"), m$Leftcontext, "O")
+m$Leftcontext <- as.factor(m$Leftcontext)
+
+
+# Simplify "Measureclass".
[…]

### 2017-03-27 R/2_attraction.R

@@ -1,0 +1,69 @@
+# This creates the quotient database for lemmas in the NN and NDN construction.
+
+
+prop.func <- function(v, smooth = 1) {
+  log( (v[2]+smooth) / (v[1]+smooth), 10)
+}
+
+ndn.kind <- read.delim("data/ndn_kind_freq.tsv", header=FALSE, quote="")
+nn.kind <- read.delim("data/nn_kind_freq.tsv", header=FALSE, quote="")
+
+
+# Measure nousn have been pre-processed differently, can load in one table.
+measurenouns_attracfreq <- read.delim("data/measurenouns_attracfreq.tsv", quote="")
+
+
+# A cleanup.
+colnames(ndn.kind) <- c("Freq","Lem")
+colnames(nn.kind) <- c("Freq","Lem")
+nn.kind <- nn.kind[-which(nn.kind$Lem=="rege"),]
+
+
+# Get all lemmas for master table.
+# Background: We want values for all nouns, even those which occur in
+# only one of the NN/NDN tables. If not in one, count will just be 0.
+cx.kind.lem <- c(as.character(nn.kind$Lem), as.character(ndn.kind$Lem))
+cx.kind.lem <- unique(cx.kind.lem[order(cx.kind.lem)])
+
+
+# Function to extract freqs for both constructions.
+get_freq <- function(data, lem) {
+  n1 <- data[which(data$Lem == lem),1]
+  if (length(n1) == 0) n1 <- 0
+  n1
+}
+
+
+# Generate full table.
+cx.kind <- cbind(cx.kind.lem, lapply(cx.kind.lem, function(lem){get_freq(nn.kind,lem)}), lapply(cx.kind.lem, function(lem){get_freq(ndn.kind,lem)}))
+
+
+# A rather simple "attraction" function, i.e., proportion.
+attraction <- function(cs) {
+  unlist(cs)[2] / (unlist(cs)[1] + unlist(cs)[2])
+}
+
+
+# Generate attraction quotients for full table.
+cx.kind <- as.data.frame(cx.kind)
+cx.kind$Pv <- unlist(apply(cx.kind[,2:3], 1, function(cs){attraction(cs)} ))
+
+
+# Measure nouns are simpler because the tables were imported done & ready.
+cx.measure <- measurenouns_attracfreq
+cx.measure$Pv <- unlist(apply(cx.measure[,2:3], 1, function(cs){attraction(cs)} ))
+
+
[…]

### 2017-03-27 R/3_glmm.R

@@ -1,0 +1,99 @@
+require(lme4)
+require(gridExtra)
+require(MuMIn)
+require(effects)
+require(gridExtra)
+require(car)
+
+
+the.nAGQ <- 0                 # 0 for speed, 1 for precision and EXTREMELY slow computation.
+ci.boot.nsim <- 2            # Even 100 can be extremely slow! SET TO MIN. 1000 FOR PRODUCTION!!!
+ci.boot.modelcomp.nsim <- 2  # This takes ages, even with 10. Set to 500 for PRODUCTION!!!
+
+
+
+# Estimate model.
+measure.glmm <- glmer(Construction~1
+                 
+                 +(1|Measurelemma)
+                 +(1|Kindlemma)
+                 
+                 +Badness                 
+                 +Genitives
+                 +Leftcontext
+                 +Measurecase
+                 +Measurenumber
+                 
+                 +Kindattraction
+                 +Kindfreq
+                 +Kindgender
+
+                 +Measureattraction
+                 +Measureclass
+                 +Measurefreq
+
+                 , data=m, family=binomial(link=logit), na.action = na.fail, nAGQ=the.nAGQ,
+                 control=glmerControl(optimizer="nloptwrap2", optCtrl=list(maxfun=2e5)))
+
+
+# Standard diagnostics.
+measure.glmm.wald <- Anova(measure.glmm, type="II")
+measure.glmm.r2 <- r.squaredGLMM(measure.glmm)
+
+
+# Predict and analyze with optimal cutoff.
+all.cut           <- seq(-2,2,0.01)[which.max(unlist(lapply(seq(-2,2,0.01), function(x) {corr.prop(measure.glmm, m$Construction, x)})))]
+measure.glmm.pred <- ifelse(predict(measure.glmm) < all.cut, 0, 1)
+measure.glmm.cmat <- table(measure.glmm.pred, m$Construction, dnn=c("Pred","Obs"))
+measure.glmm.corr <- sum(diag(measure.glmm.cmat))/sum(measure.glmm.cmat)
+measure.glmm.base <- length(which(m$Construction == "NACa"))/length(m$Construction)
+measure.glmm.pre  <- (measure.glmm.corr-measure.glmm.base)/(1-measure.glmm.base)
+
+
+# Extract random effects.
+measure.glmm.lemrand.order <- order(ranef(measure.glmm)$Kindlemma)
[…]

### 2017-03-27 R/4_glmm-mcmc.R

@@ -1,0 +1,109 @@
+require(rstanarm)
+require(rstan)
+
+rstan_options(auto_write = TRUE)
+options(mc.cores = parallel::detectCores())
+
+# MASCULINE GLMM
+
+chains <- 4
+seed   <- 6976
+iter   <- 1000 # Set to 1000 for production.
+
+# MN
+
+mn.glmm.mcmc <- stan_glmer(Genitive~1
+                           +(1|Measurelemma)
+                           +(1|Kindlemma)
+                           +Genitives
+                           +Minus1pos
+                           +Measureclass
+                           +Measureabbreviated
+                           +Measureattraction
+                           +Kindattraction
+                           +Measurenumber
+                           +Kindedible
+                           +Badness
+                           +Kindconsistency
+                           +Kindfreq
+                           +Measurecase
+                           +Measurefreq
+                           ,
+                 data=mn, family=binomial(link=logit),
+                 chains=chains, seed=seed, iter=iter,
+                 prior = normal(0, 2.5), prior_intercept = normal(0, 10),
+                 prior_ops = prior_options(prior_scale_for_dispersion = 5, min_prior_scale = 1e-12, scaled = TRUE),
+                 prior_covariance = decov(regularization = 1, concentration = 1, shape = 1, scale = 1),
+                 prior_PD = F
+                 )
+
+mn.baysum <- summary(mn.glmm.mcmc)
+
+
+# FEM
+
+fem.glmm.mcmc <- stan_glmer(Casedrop~1
+                            +(1|Measurelemma)
+                            +(1|Kindlemma)
+                            +Measureclass
+                            +Minus1pos
+                            +Measureabbreviated
+                            +Kindconsistency
+                            +Measureattraction
+                            +Genitives
+                            +Kindedible
+                            +Measurenumber
+                            +Badness
+                            +Measurefreq
+                            +Measurecase
+                            +Kindfreq
+                            +Kindattraction
[…]

### 2017-03-27 R/5_stimuli_predict.R

@@ -1,0 +1,71 @@
+require(lme4)
+
+stimuli <- read.csv("~/Massangaben/data/stimuli.csv", header=TRUE)
+
+stims.fax  <- c("Measurecase",
+          "Kindlemma",
+          "Kindconsistency",
+          "Kindedible",
+          "Measurelemma",
+          "Measurenumber",
+          "Measureabbreviated",
+          "Measureclass",
+          "Minus1pos")
+
+
+# Separate MN and F.
+
+stimuli.mn <- stimuli[1:8,]
+stimuli.fem <- stimuli[9:16,]
+
+
+# Make factors factors, using levels from original model df.
+
+for (fac in stims.fax) {
+  stimuli.mn[, fac]  <- factor(stimuli.mn[, fac], levels = levels(mn[, fac]))
+  stimuli.fem[, fac]  <- factor(stimuli.fem[, fac], levels = levels(fem[, fac]))
+}
+
+
+# Pull numeric values from original table because of standardization.
+
+get.value <- function(df, index.column, index.value, value.column) {
+  df[which(df[, index.column] == index.value)[1], value.column]
+}
+
+stimuli.mn$Kindfreq <- unlist(lapply(as.character(stimuli.mn$Kindlemma), function(x) get.value(mn, "Kindlemma", x, "Kindfreq")))
+stimuli.mn$Measurefreq <- unlist(lapply(as.character(stimuli.mn$Measurelemma), function(x) get.value(mn, "Measurelemma", x, "Measurefreq")))
+stimuli.mn$Kindattraction <- unlist(lapply(as.character(stimuli.mn$Kindlemma), function(x) get.value(mn, "Kindlemma", x, "Kindattraction")))
+stimuli.mn$Measureattraction <- unlist(lapply(as.character(stimuli.mn$Measurelemma), function(x) get.value(mn, "Measurelemma", x, "Measureattraction")))
+
+stimuli.fem$Kindfreq <- unlist(lapply(as.character(stimuli.fem$Kindlemma), function(x) get.value(fem, "Kindlemma", x, "Kindfreq")))
+stimuli.fem$Measurefreq <- unlist(lapply(as.character(stimuli.fem$Measurelemma), function(x) get.value(fem, "Measurelemma", x, "Measurefreq")))
[…]

### 2017-03-27 R/glmtools.R

@@ -1,0 +1,218 @@
+# Handling convergence warnings:
+# https://rstudio-pubs-static.s3.amazonaws.com/33653_57fc7b8e5d484c909b615d8633c01d51.html
+
+# A faster BOBYQA optimizer.
+library(nloptr)
+defaultControl <- list(algorithm="NLOPT_LN_BOBYQA",xtol_rel=1e-6,maxeval=1e5)
+nloptwrap2 <- function(fn,par,lower,upper,control=list(),...) {
+  for (n in names(defaultControl)) 
+    if (is.null(control[[n]])) control[[n]] <- defaultControl[[n]]
+  res <- nloptr(x0=par,eval_f=fn,lb=lower,ub=upper,opts=control,...)
+  with(res,list(par=solution,
+                fval=objective,
+                feval=iterations,
+                conv=if (status>0) 0 else status,
+                message=message))
+}
+
+
+# Function to get prop of correct predictions.
+corr.prop <- function(model, observed, cutoff) {
+  preds <- ifelse(predict(model) < cutoff, 0, 1)
+  cmat <- table(preds, observed)
+  sum(diag(cmat))/sum(cmat)
+}
+
+
+# Function to get LRT for (G)LMMMs with bootstrapped CIs. This takes AGES!
+lmer.modelcomparison <- function(model, regressors, formula.target, ci = 0.95, nsim = 4, print.updated = F, print.diag = F) {
+  require(pbkrtest)
+  
+  result = NULL
+  
+  for (i in 1:length(regressors)) {
+    formula.parts <- c(formula.target, regressors[1:i-1], regressors[i+1:length(regressors)])
+    formula.parts <- formula.parts[which(!is.na(formula.parts))]
+    formula <- as.formula(paste(formula.parts, collapse="+"))
+    
+    model.0 <- update(model, formula = formula)
+    pbmc <- PBmodcomp(model, model.0, nsim = nsim)
+    
+    if (print.diag) {
+      cat("\n\n\n")
+      print(regressors[i])
+      print(pbmc)
+    }
+    
+    res.tmp <- data.frame(Regressor = regressors[i], PB.p = pbmc$test$p.value[2],
+                          LRT.p = pbmc$test$p.value[1], LR = unname(pbmc$LRTstat[1]),
[…]

### 2017-03-27 R/0_wrapper.R
final stats; my gosh, so simple!

@@ -15,7 +15,7 @@ setwd('/Users/user/Workingcopies/measurenps/GitHub/R')
 # OPTIONS
 
 set.seed(2398651)            # Do NOT change, or results will differ.
-save.persistent <- F         # write to files instead of screen/console
+save.persistent <- T         # write to files instead of screen/console
 out.dir         <- "output/" # put generated files here
 
 round.in.big.table <- 2
@@ -44,6 +44,6 @@ source('2_attraction.R') # calculate attraction factors
 cat("\n\n GLMM ...\n")
 source('3_glmm.R')       # full GLMMs
 
-#cat("\n\n GLMM with MCMC ...\n")
-#source('4_glmm-mcmc.R')  # same GLMMs estimated with Stan (Bayesian/MCMC)
+cat("\n\n GLMM with MCMC ...\n")
+source('4_glmm-mcmc.R')  # same GLMMs estimated with Stan (Bayesian/MCMC)
 

### 2017-03-27 R/1_load.R

@@ -3,15 +3,15 @@ require(plyr)
 
 
 # Read data.
-m <- read.csv(file = "data/nan_sample.csv", head = TRUE, sep = '\t', quote = "")
+measure <- read.csv(file = "data/nan_sample.csv", head = TRUE, sep = '\t', quote = "")
 
 
 # Backup original sentences.
-sentences <- m[, c("Leftcontext", "Match", "Rightcontext", "Originalsort", "Url")]
+sentences <- measure[, c("Leftcontext", "Match", "Rightcontext", "Originalsort", "Url")]
 
 
 # Remove some columns. These are the ones we KEEP:
-m <- m[,c("Measurecase",
+measure <- measure[,c("Measurecase",
           "Kindlemma",
           "Kindgender",
           "Kindcase",
@@ -30,44 +30,44 @@ m <- m[,c("Measurecase",
 
 
 # Rename some columns for better readability in paper.
-colnames(m)[which(colnames(m) == "Minus1pos")] <- "Leftcontext"
+colnames(measure)[which(colnames(measure) == "Minus1pos")] <- "Leftcontext"
 
 
 # Filter data points that we cannot use.
-m <- m[which(as.character(m$Measurecase) %in% c("Nom","Acc","Dat")),]
-m <- m[which(as.character(m$Kindcase) %in% c("Acc", "Dat", "Nom", "Gen", "Obl", "Str", "Nomacc")),]
-m <- m[-which(m$Kindlemma %in% c("Sonnenschein","Silber","Obst","Strom")),]
+measure <- measure[which(as.character(measure$Measurecase) %in% c("Nom","Acc","Dat")),]
+measure <- measure[which(as.character(measure$Kindcase) %in% c("Acc", "Dat", "Nom", "Gen", "Obl", "Str", "Nomacc")),]
+measure <- measure[-which(measure$Kindlemma %in% c("Sonnenschein","Silber","Obst","Strom")),]
 
 
 # Simplify POS tags.
-m$Leftcontext <- as.character(m$Leftcontext)
-m$Leftcontext <- ifelse(m$Leftcontext %in% c("Art", "Pdat", "Pds", "Piat", "Pis", "Pposat"), "D", m$Leftcontext)
-m$Leftcontext <- ifelse(m$Leftcontext %in% c("Appr", "Apprart"), "P", m$Leftcontext)
-m$Leftcontext <- ifelse(m$Leftcontext == "Adja", "A", m$Leftcontext)
[…]

### 2017-03-27 R/2_attraction.R

@@ -40,7 +40,8 @@ cx.kind <- cbind(cx.kind.lem, lapply(cx.kind.lem, function(lem){get_freq(nn.kind
 
 # A rather simple "attraction" function, i.e., proportion.
 attraction <- function(cs) {
-  unlist(cs)[2] / (unlist(cs)[1] + unlist(cs)[2])
+#  unlist(cs)[2] / (unlist(cs)[1] + unlist(cs)[2])
+  log((unlist(cs)[2] + 1) / (unlist(cs)[1] + 1))
 }
 
 
@@ -60,10 +61,10 @@ colnames(cx.measure) <- c("Lemma", "Nn", "Ndn", "Pv")
 
 
 # Pull attraction strength for each observation.
-m$Kindattraction <- unlist(lapply(m$Kindlemma, function(x) { cx.kind[which(cx.kind$Lemma == x), "Pv"] } ))
-m$Measureattraction <- unlist(lapply(m$Measurelemma, function(x) { cx.measure[which(as.character(cx.measure$Lemma) == as.character(x)), "Pv"] } ))
+measure$Kindattraction <- unlist(lapply(measure$Kindlemma, function(x) { cx.kind[which(cx.kind$Lemma == x), "Pv"] } ))
+measure$Measureattraction <- unlist(lapply(measure$Measurelemma, function(x) { cx.measure[which(as.character(cx.measure$Lemma) == as.character(x)), "Pv"] } ))
 
 
 # Center
-m$Kindattraction    <- z.transform(m$Kindattraction)
-m$Measureattraction <- z.transform(m$Measureattraction)
+measure$Kindattraction    <- z.transform(measure$Kindattraction)
+measure$Measureattraction <- z.transform(measure$Measureattraction)

### 2017-03-27 R/3_glmm.R

@@ -6,13 +6,12 @@ require(gridExtra)
 require(car)
 
 
-the.nAGQ <- 0                 # 0 for speed, 1 for precision and EXTREMELY slow computation.
-ci.boot.nsim <- 2            # Even 100 can be extremely slow! SET TO MIN. 1000 FOR PRODUCTION!!!
-ci.boot.modelcomp.nsim <- 2  # This takes ages, even with 10. Set to 500 for PRODUCTION!!!
+the.nAGQ               <- 0      # 0 for speed, 1 for precision and EXTREMELY slow computation.
+ci.boot.nsim           <- 100   # Even 100 can be extremely slow! SET TO MIN. 1000 FOR PRODUCTION!!!
+ci.boot.modelcomp.nsim <- 100   # This takes ages, even with 10. Set to 1000 for PRODUCTION!!!
 
 
 
-# Estimate model.
 measure.glmm <- glmer(Construction~1
                  
                  +(1|Measurelemma)
@@ -32,7 +31,7 @@ measure.glmm <- glmer(Construction~1
                  +Measureclass
                  +Measurefreq
 
-                 , data=m, family=binomial(link=logit), na.action = na.fail, nAGQ=the.nAGQ,
+                 , data=measure, family=binomial(link=logit), na.action = na.fail, nAGQ=the.nAGQ,
                  control=glmerControl(optimizer="nloptwrap2", optCtrl=list(maxfun=2e5)))
 
 
@@ -42,11 +41,11 @@ measure.glmm.r2 <- r.squaredGLMM(measure.glmm)
 
 
 # Predict and analyze with optimal cutoff.
-all.cut           <- seq(-2,2,0.01)[which.max(unlist(lapply(seq(-2,2,0.01), function(x) {corr.prop(measure.glmm, m$Construction, x)})))]
+all.cut           <- seq(-2,2,0.01)[which.max(unlist(lapply(seq(-2,2,0.01), function(x) {corr.prop(measure.glmm, measure$Construction, x)})))]
 measure.glmm.pred <- ifelse(predict(measure.glmm) < all.cut, 0, 1)
-measure.glmm.cmat <- table(measure.glmm.pred, m$Construction, dnn=c("Pred","Obs"))
+measure.glmm.cmat <- table(measure.glmm.pred, measure$Construction, dnn=c("Pred","Obs"))
[…]

### 2017-03-27 R/4_glmm-mcmc.R

@@ -4,32 +4,32 @@ require(rstan)
 rstan_options(auto_write = TRUE)
 options(mc.cores = parallel::detectCores())
 
-# MASCULINE GLMM
 
 chains <- 4
 seed   <- 6976
-iter   <- 1000 # Set to 1000 for production.
+iter   <- 100 # Set to 1000 for production.
 
-# MN
 
-mn.glmm.mcmc <- stan_glmer(Genitive~1
+measure.glmm.mcmc <- stan_glmer(Construction~1
+                           
                            +(1|Measurelemma)
                            +(1|Kindlemma)
+                           
+                           +Badness                 
                            +Genitives
-                           +Minus1pos
-                           +Measureclass
-                           +Measureabbreviated
-                           +Measureattraction
-                           +Kindattraction
-                           +Measurenumber
-                           +Kindedible
-                           +Badness
-                           +Kindconsistency
-                           +Kindfreq
+                           +Leftcontext
                            +Measurecase
+                           +Measurenumber
+                           
+                           +Kindattraction
+                           +Kindfreq
+                           +Kindgender
+                           
+                           +Measureattraction
+                           +Measureclass
                            +Measurefreq
                            ,
-                 data=mn, family=binomial(link=logit),
+                 data=measure, family=binomial(link=logit),
                  chains=chains, seed=seed, iter=iter,
                  prior = normal(0, 2.5), prior_intercept = normal(0, 10),
                  prior_ops = prior_options(prior_scale_for_dispersion = 5, min_prior_scale = 1e-12, scaled = TRUE),
@@ -37,72 +37,32 @@ mn.glmm.mcmc <- stan_glmer(Genitive~1
                  prior_PD = F
                  )
 
-mn.baysum <- summary(mn.glmm.mcmc)
-
-
-# FEM
-
-fem.glmm.mcmc <- stan_glmer(Casedrop~1
-                            +(1|Measurelemma)
-                            +(1|Kindlemma)
[…]

### 2017-03-27 R/0_wrapper.R
final statistics, wow!

@@ -1,37 +1,38 @@
 #
 # This calls all scripts in the right order to replicate results from my paper
-# 'Competing Constructions for German Measure Phrases' (working title)
-#
-# Warning: This takes quite a long times to run, especially with 
-#  =>  model coef & LR BOOTSTRAPPING
-#  =>  BAYESIAN / MCMC estimation
+# 'Competing Constructions for German Measure Phrases'
 
-# Comment this line if you want to keep your old workspace.
+
 rm(list = ls())
+
 
 # Set this to the location of R script sources.
 setwd('/Users/user/Workingcopies/measurenps/GitHub/R')
 
-# OPTIONS
 
-set.seed(2398651)            # Do NOT change, or results will differ.
-save.persistent <- T         # write to files instead of screen/console
-out.dir         <- "output/" # put generated files here
+# Options.
+set.seed(2398651)                   # Do NOT change, or results will differ.
+save.persistent        <- T         # write to files instead of screen/console.
+out.dir                <- "output/" # Put generated files here.
+precision              <- 3         # Precision in result table.
+the.nAGQ               <- 0         # 0 for speed, 1 for precision and eytremely slow computation.
+ci.boot.nsim           <- 1000      # Even 100 can be extremely slow. Set to 1000 for production.
+ci.boot.modelcomp.nsim <- 1000      # This takes ages, even with 10. Set to 1000 for for production.
+chains                 <- 4         # MCMC.
+seed                   <- 6976      # MCMC.
+iter                   <- 1000      # MCMC. Set to 1000 for production.
 
-round.in.big.table <- 2
 
-# Comment the next line if you want to keep old output files.
+# Delete old output.
 unlink(paste(out.dir, '*', sep=""))
+
 
 # ----------   Don't touch anything past this line   ---------- #
 # ---------- If you do, results might not replicate. ---------- #
 
-# HELPERS
 
+# Helpers.
 source('glmtools.R')
-
-
-# SCRIPTS
 
 cat("\014")
 
@@ -47,3 +48,5 @@ source('3_glmm.R')       # full GLMMs
 cat("\n\n GLMM with MCMC ...\n")
 source('4_glmm-mcmc.R')  # same GLMMs estimated with Stan (Bayesian/MCMC)
 
+cat("\n\n Prediction for experimental stimuli ...\n")
+source('5_stimuli_predict.R')

### 2017-03-27 R/1_load.R

@@ -1,4 +1,3 @@
-require(gnumeric)
 require(plyr)
 
 

### 2017-03-27 R/2_attraction.R

@@ -1,9 +1,3 @@
-# This creates the quotient database for lemmas in the NN and NDN construction.
-
-
-prop.func <- function(v, smooth = 1) {
-  log( (v[2]+smooth) / (v[1]+smooth), 10)
-}
 
 ndn.kind <- read.delim("data/ndn_kind_freq.tsv", header=FALSE, quote="")
 nn.kind <- read.delim("data/nn_kind_freq.tsv", header=FALSE, quote="")
@@ -38,9 +32,8 @@ get_freq <- function(data, lem) {
 cx.kind <- cbind(cx.kind.lem, lapply(cx.kind.lem, function(lem){get_freq(nn.kind,lem)}), lapply(cx.kind.lem, function(lem){get_freq(ndn.kind,lem)}))
 
 
-# A rather simple "attraction" function, i.e., proportion.
+# A rather simple "attraction" function, i.e., log ratio.
 attraction <- function(cs) {
-#  unlist(cs)[2] / (unlist(cs)[1] + unlist(cs)[2])
   log((unlist(cs)[2] + 1) / (unlist(cs)[1] + 1))
 }
 

### 2017-03-27 R/3_glmm.R

@@ -6,12 +6,6 @@ require(gridExtra)
 require(car)
 
 
-the.nAGQ               <- 0      # 0 for speed, 1 for precision and EXTREMELY slow computation.
-ci.boot.nsim           <- 100   # Even 100 can be extremely slow! SET TO MIN. 1000 FOR PRODUCTION!!!
-ci.boot.modelcomp.nsim <- 100   # This takes ages, even with 10. Set to 1000 for PRODUCTION!!!
-
-
-
 measure.glmm <- glmer(Construction~1
                  
                  +(1|Measurelemma)
@@ -75,12 +69,13 @@ cat("\n\n")
 print(measure.glmm.wald)
 cat("\n\nModel comparison with LR and PB test\n")
 print(modelcomp)
+print(round(modelcomp$PB.p, precision))
 cat("\n\n")
-print(measure.glmm.r2)
+print(round(measure.glmm.r2, precision))
 cat("\n\n")
-cat("correct", measure.glmm.corr)
+cat("correct", round(measure.glmm.corr, precision))
 cat("\n\n")
-cat("λ", measure.glmm.pre)
+cat("λ", round(measure.glmm.pre, precision))
 cat("\n\n")
 if (save.persistent) sink()
 
@@ -149,15 +144,4 @@ do.call(ranef.plot, c(list(measure.glmm, measure, "Measurelemma", n.select, main
 do.call(ranef.plot, c(list(measure.glmm, measure, "Kindlemma", n.select, main = main.dotchart.kind), opts.dotchart))
 par(mfrow=c(1,1))
 if (save.persistent) dev.off()
-
-
-
-
-# Make table.
-measure.glmm.table <- data.frame(
-  Coefficient = round(fixef(measure.glmm)[-1], round.in.big.table),
-  Low         = rev(round(measure.ci.95[,1], round.in.big.table)),
-  High        = rev(round(measure.ci.95[,2], round.in.big.table)),
-  Not0        = rev(ifelse( (measure.ci.95[,1] < 0 & measure.ci.95[,2] < 0) | (measure.ci.95[,1] > 0 & measure.ci.95[,2] > 0) , "*", ""))
-)
 

### 2017-03-27 R/4_glmm-mcmc.R

@@ -1,13 +1,9 @@
 require(rstanarm)
 require(rstan)
 
+
 rstan_options(auto_write = TRUE)
 options(mc.cores = parallel::detectCores())
-
-
-chains <- 4
-seed   <- 6976
-iter   <- 100 # Set to 1000 for production.
 
 
 measure.glmm.mcmc <- stan_glmer(Construction~1
@@ -50,20 +46,23 @@ if (save.persistent) sink()
 
 # Make table.
 measure.glmm.table <- data.frame(
-  ML_Coef  = round(fixef(measure.glmm)[-1], round.in.big.table),
-  ML_Low   = rev(round(measure.ci.95[,1], round.in.big.table)),
-  ML_High  = rev(round(measure.ci.95[,2], round.in.big.table)),
+  ML_Coef  = round(fixef(measure.glmm)[-1], precision),
+  MC_Coef  = round(measure.baysum[,1][names(fixef(measure.glmm)[-1])], precision),
+  
+  ML_Low   = rev(round(measure.ci.95[,1], precision)),
+  MC_Low   = round(measure.baysum[,4][names(fixef(measure.glmm)[-1])], precision),
+
+  ML_High  = rev(round(measure.ci.95[,2], precision)),
+  MC_High  = round(measure.baysum[,8][names(fixef(measure.glmm)[-1])], precision),
+
   ML_Not0  = rev(ifelse( (measure.ci.95[,1] < 0 & measure.ci.95[,2] < 0) | (measure.ci.95[,1] > 0 & measure.ci.95[,2] > 0) , "*", "")),
-  MC_Coef  = round(measure.baysum[,1][names(fixef(measure.glmm)[-1])], round.in.big.table),
-  MC_Low   = round(measure.baysum[,4][names(fixef(measure.glmm)[-1])], round.in.big.table),
-  MC_High  = round(measure.baysum[,8][names(fixef(measure.glmm)[-1])], round.in.big.table),
   MC_Not0  = ifelse( (measure.baysum[,4][names(fixef(measure.glmm)[-1])] < 0 & measure.baysum[,8][names(fixef(measure.glmm)[-1])] < 0) |
                        (measure.baysum[,4][names(fixef(measure.glmm)[-1])] > 0 & measure.baysum[,8][names(fixef(measure.glmm)[-1])] > 0) , "*", "")
 )
 
 
 # Print table.
-if (save.persistent) sink(paste(out.dir, "06_glmm-mcmc.txt", sep=""), append = T)
[…]

### 2017-03-27 R/5_stimuli_predict.R

@@ -1,71 +1,44 @@
 require(lme4)
 
-stimuli <- read.csv("~/Massangaben/data/stimuli.csv", header=TRUE)
 
-stims.fax  <- c("Measurecase",
-          "Kindlemma",
-          "Kindconsistency",
-          "Kindedible",
-          "Measurelemma",
-          "Measurenumber",
-          "Measureabbreviated",
-          "Measureclass",
-          "Minus1pos")
+stimuli <- read.csv("data/stimuli.csv", header=TRUE)
 
 
-# Separate MN and F.
-
-stimuli.mn <- stimuli[1:8,]
-stimuli.fem <- stimuli[9:16,]
+stims.fax  <- c(
+  "Measurecase",
+  "Kindlemma",
+  "Measurelemma",
+  "Measurenumber",
+  "Measureclass",
+  "Leftcontext",
+  "Kindgender"
+  )
 
 
 # Make factors factors, using levels from original model df.
-
 for (fac in stims.fax) {
-  stimuli.mn[, fac]  <- factor(stimuli.mn[, fac], levels = levels(mn[, fac]))
-  stimuli.fem[, fac]  <- factor(stimuli.fem[, fac], levels = levels(fem[, fac]))
+  stimuli[, fac]  <- factor(stimuli[, fac], levels = levels(measure[, fac]))
 }
 
 
 # Pull numeric values from original table because of standardization.
-
 get.value <- function(df, index.column, index.value, value.column) {
   df[which(df[, index.column] == index.value)[1], value.column]
 }
 
-stimuli.mn$Kindfreq <- unlist(lapply(as.character(stimuli.mn$Kindlemma), function(x) get.value(mn, "Kindlemma", x, "Kindfreq")))
-stimuli.mn$Measurefreq <- unlist(lapply(as.character(stimuli.mn$Measurelemma), function(x) get.value(mn, "Measurelemma", x, "Measurefreq")))
-stimuli.mn$Kindattraction <- unlist(lapply(as.character(stimuli.mn$Kindlemma), function(x) get.value(mn, "Kindlemma", x, "Kindattraction")))
-stimuli.mn$Measureattraction <- unlist(lapply(as.character(stimuli.mn$Measurelemma), function(x) get.value(mn, "Measurelemma", x, "Measureattraction")))
-
[…]

### 2017-03-28 R/0_wrapper.R
hopefully final statistics run

@@ -2,6 +2,7 @@
 # This calls all scripts in the right order to replicate results from my paper
 # 'Competing Constructions for German Measure Phrases'
 
+# Scripts for the analysis of the experiments need to be run manually.
 
 rm(list = ls())
 
@@ -13,7 +14,7 @@ setwd('/Users/user/Workingcopies/measurenps/GitHub/R')
 # Options.
 set.seed(2398651)                          # Do NOT change, or results will differ.
 save.persistent        <- T                # write to files instead of screen/console.
-out.dir                <- "output/corpus_" # Put generated files here.
+out.dir                <- "output/"        # Put generated files here.
 precision              <- 3                # Precision in result table.
 the.nAGQ               <- 0                # 0 for speed, 1 for precision and eytremely slow computation.
 ci.boot.nsim           <- 1000             # Even 100 can be extremely slow. Set to 1000 for production.
@@ -25,6 +26,7 @@ iter                   <- 1000             # MCMC. Set to 1000 for production.
 
 # Delete old output.
 unlink(paste(out.dir, '*', sep=""))
+out.dir <- paste(out.dir, "corpus_", sep="")
 
 
 # ----------   Don't touch anything past this line   ---------- #

### 2017-03-28 R/6_exp-fc.R

@@ -10,7 +10,7 @@ set.seed(239865)
 setwd('/Users/user/Workingcopies/measurenps/GitHub/R')
 
 ci.method        <- "boot" # boot
-modcomp.nsim     <- 500    # 500
+modcomp.nsim     <- 1000   # 500
 boot.nsim        <- 1000   # 1000
 out.dir          <- "output/fc_"
 save.persistent  <- T

### 2017-03-28 R/7_exp-spr.R

@@ -11,8 +11,8 @@ set.seed(239865)
 setwd('/Users/user/Workingcopies/measurenps/GitHub/R')
 
 
-ci.method       <- "boot"
-modcomp.nsim    <- 500    # 500
+ci.method       <- "boot" # boot
+modcomp.nsim    <- 1000   # 500
 boot.nsim       <- 1000   # 1000
 iqr.factor      <- 3      # For removing outliers in final data set.
 out.dir         <- "output/spr_"

### 2017-03-28 Latex/article.tex
massive improvements in stats

@@
 \newcommand{\Supsf}[1]{\ensuremath{\mathsf{^{#1}}}}
+\newcommand{\pPB}{p\Subsf{PB}}
 
@@
 I discuss four main factors that might influence the alternation:
-the different syntactic structures of the variants \NACa\ and \PGCa\ and their syntactic prototypes \NACb\ and \PGCd,
-degrees of grammaticalisation of the measure noun associated with the prototypes,
-a semantic prototypicality effect of the degrees of grammaticalisation (plurality),
-register.
-
-First, I discuss the analysis of the morpho-syntax of German MNPs.
+first, the different syntactic structures of the variants \NACa\ and \PGCa\ in relation to their syntactic prototypes \NACb\ and \PGCd;
+second, degrees of grammaticalisation of the measure noun associated with the two prototypes;
+third, a semantic prototypicality effect of the degrees of grammaticalisation (plurality);
+fourth, preferences in different registers.
+
+The first aspect is related to analyses of the morpho-syntax of German MNPs.
 A major interest in descriptive and generative linguistics is to find a plausible analysis of the hierarchical syntactic structure of the constructions in question.
 The first question is which noun constitutes the head of the whole construction.
-This question was answered quite clearly by \cite{Loebel1986} already (see also \citealp[213]{Eschenbach1994}, and \citealp[16]{GallmannLindauer1994}).
+This question was answered for the case at hand quite clearly by \cite{Loebel1986} already (see also \citealp[213]{Eschenbach1994}, and \citealp[16]{GallmannLindauer1994}).
 Subject-verb agreement is always realised on the measure noun, and we can therefore assume that the measure noun is the head.
-However, the MNP-internal structure has more interesting cues to offer if the strong\slash weak inflection patterns of adjectives already mentioned in Section~\ref{sec:descriptive} are taken into account.
+However, the MNP-internal structure has more interesting cues to offer if the strong\slash weak inflection patterns of adjectives (already mentioned in Section~\ref{sec:descriptive}) are taken into account.
 In NPs with a strongly inflected determiner, attributive adjectives inflect according to the massively syncretistic \textit{weak} pattern.
@@
 %The only exception to this rule is the masculine accusative singular, which is also marked with \textit{-en}.}
-If there is no determiner (as is the case in the alternating constructions), attributive adjectives inflect like determiners, which is called the \textit{strong} pattern (see Table~\ref{tab:syncretisms}).
+If there is no determiner (as is the case in the alternating constructions), attributive adjectives inflect like determiners, however.
+This is called the \textit{strong} inflectional pattern (see Table~\ref{tab:syncretisms}).
 Thus, the adjectives in the \NACa\ and the \PGCa\ have properties of adjectives as well as determiners.
@@
 This unusual double nature of adjectives in NPs without determiners leads to a plausible probabilistic interpretation of the pattern shown in Table~\ref{tab:constructions}.
-Whenever speakers classify the adjective in the kind noun phrase more as a determiner, they \textit{have} to use the \PGCa\ because, if there is a determiner, the PGC (just like in the case of the \PGCd) is the only option.
-When they classify the adjective more as an adjective, the kind NP has no determiner, and they \textit{have} to use the \NACa\ (like in the case of the \NACb).%
+Whenever speakers classify the adjective in the kind noun phrase more as a determiner, they have to use the \PGCa\ because, if there is a determiner, the PGC (just like in the case of the \PGCd) is the only option.
+When they classify the adjective more as an adjective, the kind NP has no determiner, and they have to use the \NACa\ (like in the case of the \NACb).%
 \footnote{While the generative analysis presented in \cite{Bhatt1990} cannot properly deal with probabilistic effects, Bhatt comes close to this interpretation by analysing the kind NP in the GPC as a DP and in the NAC as an NP.}
@@
 Any attempt to substantiate or test this hypothesis is rendered difficult by the fact that the semantic differences between the variants are at best subtle, which means that the prototypes are not associated with clearly detectable differences in meaning.
-However, if there is a semantic difference, kind and measure noun lemmas (and maybe semantically of functionally defined classes thereof) could be attracted more strongly to one of the prototypes.
-As a result, kind noun lemmas and measure noun lemmas probably appear with different relative frequencies in the prototypical \NACb\ and the prototypical \PGCd.
-Their relative frequencies in these non-alternating configurations should then be indicative of their preference to appear in one of the two variants.
+However, if there is a semantic difference, kind and measure noun lemmas should be attracted more strongly to one of the prototypes.
+As a result, kind noun lemmas and measure noun lemmas are expected to appear with different relative frequencies in the prototypical \NACb\ and the prototypical \PGCd.
+Their relative frequencies in these non-alternating configurations should then be indicative of their preference to appear in one of the two alternating variants \NACa\ and \PGCa.
 This predicted effect will be quantified in Section~\ref{sec:corpusstudies}.
@@
 The alternating constructions in German are both clearly pseudo-partitives, but the grammaticalisation paths uncovered by \cite[esp.\ 526--530]{Koptjevskaja2001} -- which I interpret freely here -- are clearly relevant for an analysis of the alternation.
-The grammaticalisation path can start out (in some languages) with constructions involving two referential nouns (not necessarily forming a single NP) and a \textit{separative} meaning as in \textit{(cut) two slices from the cake} \citep[535]{Koptjevskaja2001}.
+The grammaticalisation path can start out (in some languages) with constructions involving two referential nouns (not necessarily forming a single and contiguous NP) and a \textit{separative} meaning as in \textit{(cut) two slices from the cake} \citep[535]{Koptjevskaja2001}.
 The \textit{part-of} meaning of true partitives as in \textit{a slice of the cake} represents the first stage of a development wherein the measure noun already tends to loose some semantic content.
@@
 Also, it still has the potential to form a true partitive (if the kind noun is definite).
-Conversely, the NAC completely lacks this ability to form true partitives.
+Conversely, the NAC completely lacks this ability to form true partitives, and it forms a more restrictive environment with less combinatory and semantic flexibility.
 Hence, we can expect more strongly grammaticalised measure nouns to prefer the NAC.
-For example, strongly grammaticalised non-referential nouns like \textit{Gramm} `gram' and \textit{Meter} `metre' should occur proportionally more often in the \NACa\ than in the \PGCa.
-
-The question arises whether there might other cues to the degree of grammaticalisation of individual measure nouns or classes of them.
-In case studies like \cite{Brems2003,DeclerckBrems2016} syntactic evidence for grammaticalisation plays a prominent role.
+For example, strongly grammaticalised non-referential nouns like \textit{Gramm} `gram' and \textit{Meter} `metre' should occur proportionally more often in the \NACa\ than in the \PGCa.%
+\footnote{In case studies like \cite{Brems2003,DeclerckBrems2016} syntactic evidence for grammaticalisation plays a prominent role.
 In English -- \cite{Brems2003} argues based on a view from \cite{Langacker1991} -- strongly lexicalised and thus semantically bleached measure nouns turn from nouns into quantifiers, leading to a syntactic reanalysis of the NP, wherein the kind noun becomes the head.
 The argument is supported by convincing syntactic evidence, for example subject-verb agreement.
-As argued above, in German MNPs the measure noun always remains the syntactic head, however, and there is no similar clear criterion to detect grammaticalisation.
-Therefore, in the analysis of the corpus finding in Section~\ref{sec:corpusstudies}, effects of different degrees of grammaticalisation will be taken into account by interpreting lemma-specific and lemma class-specific effects.
+As argued above, in German MNPs the measure noun always remains the syntactic head, and there is no similarly handy criterion to diagnose a high degree of grammaticalisation.}
 
@@
 More indirectly, however, plurality of the measure noun could also favour the \PGCa\ over the \NACa, and exact quantification in the form of cardinals even more so.%
-\footnote{This preference was observed empirically by \cite{Hentschel1993}.}
-I provide a highly illustrative example using a measure noun which has virtually developed two variants, one of which grammaticalised to a much higher degree, namely \textit{Haufen} `heap\slash lot'.
-The phrase \textit{ein Haufen französische Watte} `a heap\slash lot of French cotton' (\NACa) has essentially two uses.
+\footnote{Without offering a theoretical analysis, such a preference was observed empirically by \cite{Hentschel1993}.}
+There are highly illustrative examples with measure nouns that have virtually developed two variants, one of which grammaticalised to a much higher degree, \eg\ \textit{Haufen} `heap\slash lot'.
+The phrase \textit{ein Haufen französische Watte} `a heap\slash lot of French cotton' (\NACa) has two uses.
 It can be used to denote an unspecific amount of cotton in the sense of \textit{a lot of French cotton} in situations when the exact quantity of cotton is not of interest.
-Under the appropriate circumstances, it could also be used to denote a highly specific quantity of cotton, for example when a cotton merchant always sells her French cotton in \textit{heaps} weighing two stone each.
+Under the appropriate circumstances, it could also be used to denote a heap-shaped, potentially highly specific quantity of cotton, for example when a cotton merchant always sells her French cotton in \textit{heaps} weighing two stone each.
 The former use, however, is highly restricted for \textit{zwei Haufen französischer Watte} `two heaps of French cotton' (\PGCa) -- not `two lots of French cotton'.
@@
 Under an integral view, such preferences can be to be part of the construction prototypes.
-In the corpus study, the alternation tendencies controlled by this complex factor related to register, medium, and lectal variation will be tested using certain very simple proxies.
+In the corpus study in Section~\ref{sec:corpusstudies}, register effects will therefore be modelled, even if only using some very simple proxies.
 
@@
 Then, the resulting 100 lists of noun-noun combinations were each sorted by frequency in descending order and sieved manually to remove erroneous hits.
-From each of the 100 lists, I also removed noun-noun combinations that had a frequency below 2, except if the individual list would have otherwise been shorter than 20 noun-noun combinations
+From each of the 100 lists, I also removed noun-noun combinations that had a frequency below 2, except if the individual list would have otherwise been shorter than 20 noun-noun combinations.
 The result was a list of the most frequent 2,365 individual combinations of a measure noun and a mass noun.
@@
 Since the mass nouns in the sample were distributed according to the usual power law, I used all hits for mass nouns occurring less than one hundred times, but randomly sampled one hundred hits for each mass noun that occurred one hundred or more times.
-The final sample contained 6,843 sentences.
+The final sample contained 6,843 sentences, which was reduced to 5,063 in the manual annotation process due to removal of noisy material, erroneous hits and uninformative cases where the measure noun was in the genitive, in which case the \NACa\ cannot be distinguished from the \PGCa.
 Given the careful bootstrapping and sampling procedure described in this section, we can be highly sure that it contains all relevant and reasonably frequent noun–noun combinations in the target constructions.%
@@
 
-%A mixture of automatic pre-annotation based on the annotations available in DECOW and manual annotations was used.
-%Importantly, the case of the measure noun was annotated completely by hand because case syncretisms make automatic case annotations unreliable. 
-The full set of manually annotated variables for the main sample is given in Table~\ref{tab:variables}, and I briefly discuss them now.
-Notice that \textit{Genitive} is the boolean (dichotomous) response variable (`dependent variable') which is 1 when the PGC is chosen in a given sentence, and 0 when the NAC is chosen.
+The full set of manually annotated variables for the main sample is given in Table~\ref{tab:variables}, and I briefly discuss them now.%
+\footnote{All numeric variables were also z-transformed (\ie, centered to the mean and rescaled such that they have a standard deviation of $1$) to facilitate their interpretation in the regression models reported in the next section.}
+Notice that \textit{Construction} is the response variable (or `dependent variable') with the values \textit{PGCa} and \textit{NACa}.
 
@@
                    & Genitives                        & numeric &                              \\
-    Sentence       & \textbf{Genitive (response)} & boolean &                             \\
-                   & Matchlength                  & numeric &                             \\
-                   & Measureabbreviated           & boolean &                             \\
+    Sentence       & \textbf{Construction (response)} & factor  & NACa, PGCa                   \\
                    & Measurecase                      & factor  & Nom, Acc, Dat                \\
                    & Measurenumber                    & factor  & Pl, Sg                       \\
-                   & Minus1pos                    & factor  & A, C, D, O, P               \\
+                   & Leftcontext                      & factor  & C, D, O, A, P                \\
     Kind lemma     & Kindattraction                   & numeric &                              \\
-                   & Kindconsistency              & factor  & Gaseous, Hard, Immaterial,  \\
-                   &                              &         & Liquid, Mixed, Powder, Soft \\
-                   & Kindedible                   & boolean &                             \\
                    & Kindfreq                         & numeric &                              \\
-                   & Kindint                      & boolean &                             \\
     Measure lemma  & Measureattraction                & numeric &                              \\
-                   & Measureclass                 & factor  & Amount, Capacity, Currency, \\
-                   &                              &         & Layer, Length, Portion,     \\
-                   &                              &         & Rest, Vessel, Weight        \\
+                   & Measureclass                     & factor  & Physical, Container, Amount, \\
+                   &                                  &         & Portion, Rest                \\
                    & Measurefreq                      & numeric &                              \\
   \end{tabular}
-  \caption{Annotated variables in the main sample}
+  \caption{Annotated variables for the main sample}
   \label{tab:variables}
@@
 
-To capture the influence of register or style mentioned in Section~\ref{sec:analyses}, several proxy variables were used.
+The variables \textit{Kindattraction} and \textit{Measureattraction} encode the ratio with which a given kind noun lemma or measure noun lemma occurs in the \NACb\ and the \PGCd.
+They were calculated from the auxiliary samples described at the end of Section~\ref{sec:gettingdata} as a log-transformed quotient.
+The higher the value, the more often the noun occurs in the \PGCa (proportionally).
+\textit{Kindfreq} and \textit{Measurefreq} are the logarithm-transformed frequencies per 1,000,000 words of each lemma, extracted from the frequency lists distributed by the DECOW corpus creators on their web page.
+
+In Section~\ref{sec:analyses}, it was hypothesised that classes of measure lemmas might have different preferences for the two variants.
+To capture this, class information was annotated for measure lemmas.
+The classification was designed inspired by the list in \cite[530]{Koptjevskaja2001}, but due to the frequencies of the respective nouns, very coarse classes were necessary.
+With typical examples and the number of data points in the final sample, the classes are:
+\textit{Physical} (abstract precisely measurable units such as \textit{Liter} `litre', \textit{Meter} `metre', \textit{Gramm} `gram'; size 1,968),
+\textit{Container} (\textit{Eimer} `bucket'; size 740), 
+\textit{Amount} (\textit{Menge} `amount'; size 1,364), 
+\textit{Portion} (natural portions like \textit{Happen} `bite' or \textit{Krümel} `crumb'; size 713).
+The few lemmas that did not fit into either of these classes were labeled \textit{Rest} (size 278).
+
+The variable \textit{Leftcontext} encodes the coarse-grained part-of-speech of the word preceding the measure noun.
+The level names stand for (again with class sizes in the final sample) \textit{cardinal} (size 1,939), \textit{determiner} (size 1,971), \textit{other} (size 221), \textit{adjective} (size 845), and \textit{preposition} (size 87).
+The purpose of this variable is mainly to test whether cardinals really favour the \PGCa\ as hypothesised in Section~\ref{sec:analyses}.
+Along the same lines, \textit{Measurenumber} encodes the grammatical case and number of the measure noun.
+
+To capture the influence of register or style mentioned in Section~\ref{sec:analyses}, two proxy variables were used.
 At the document level, the DECOW corpus has an annotation for \textit{Badness}.
@@
 However, since the PGCs make up for only a minute fraction of all genitives, I still use \textit{Genitives} as a regressor with the appropriate caveats.
-Since word length is also a good indicator of higher registers and styles, the length of the match in characters (for example 17 for \textit{Kilo tolle Schoki} `a kilo of nice choc' and 49 for \textit{ein Kilogramm vorzüglicher Edelbitter-Schokolade} ` a kilogram of excellent fine dark chocolate') was added as another proxy to register.%
-\footnote{The new version of the corpus contains over 100 document-level variables in the spirit of \cite{Biber1988}.
-Unfortunately, these annotations were not available when my samples were taken, and even cruder proxies to register than Biberian features (like the length of the query match) had to be used.
-All in all, all variables used as indicators of register here have to be interpreted with extreme caution.
-}
-Furthermore, abbreviated measure nouns like \textit{kg} or \textit{m} might be indicative of less elaborate technical styles, and hence the variable \textit{Measureabbreviated} was included, which is 1 if the measure noun is abbreviated.
-
-The variables \textit{Measurecase} and \textit{Measurenumber} simply encode the grammatical case and number of the measure noun.
-The variable \textit{Minus1pos} encodes the part-of-speech of the word preceding the measure noun.
-The level names stand for \textit{adjective}, \textit{cardinal}, \textit{determiner}, \textit{other}, and \textit{preposition} (in the same order as in Table~\ref{tab:variables}).
-The purpose of this variable is mainly to see whether cardinals really favour the \PGCa\ as hypothesised in Section~\ref{sec:analyses}.
-
-The variables \textit{Kindattraction} and \textit{Measureattraction} encode the ratio with which a given kind noun lemma or measure noun lemma occurs in the \NACb\ and the \PGCd.
-They were calculated from the auxiliary samples described at the end of Section~\ref{sec:gettingdata} as a quotient with one added to numerator and denominator to avoid zero divisions.
-The higher the value, the more often the noun occurs in the PGC with a determiner (proportionally).
-
-In Section~\ref{sec:analyses}, it was hypothesised that classes of lemmas might have different preferences for the two variants.
-To capture this, class information was annotated for measure and kind lemmas.
-For the kind nouns, it must be admitted that semantic classes were defined ad hoc based on intuitions of colleagues and native speakers, especially because classes available in lexical-semantic databases like GermaNet \citep{HampFeldweg1997} are too coarse for specialised purposes like the present study.
-As a result, \textit{Kindconsistency} with the self-explanatory factor levels in Table~\ref{tab:variables} was annotated.
-Also, \textit{Kindedible} encodes whether the substance denoted by the kind noun is edible.
-Finally, \textit{Kindint} encodes whether that substance is an artifact made with intent (like \textit{Suppe} `soup') or not (like \textit{Meerwasser} `sea water').
-For the measure nouns, relevant classification seemed much clearer, and it was designed inspired by the list in \cite[530]{Koptjevskaja2001}.
-With typical examples, the classes are:
-\textit{Amount} (\textit{Menge} `amount'), 
-\textit{Capacity} (\textit{Liter} `litre'), 
-\textit{Currency} (\textit{Euro}), 
-\textit{Layer} (\textit{Schicht} `layer')
-\textit{Length} (\textit{Meter} `metre'), 
-\textit{Portion} (natural portions like \textit{Happen} `bite' or \textit{Krümel} `crumb'), 
-\textit{Vessel} (\textit{Eimer} `bucket'), 
-\textit{Weight} (\textit{Gramm} `gram').
-The few lemmas that did not fit into either of these classes were labeled \textit{Rest}.
-
-Finally, two variables were added as more or less nuisance variables in the context of the present study.
-First, it was reported in the literature that MNPs in the dative and with a masculine or neuter kind noun favour the \PGCa\ more than the corresponding nominative and accusative MNPs \citep{Hentschel1993,Zimmer2015}.
+
+Finally, one variable as added as nuisance variable in the context of the present study.
+It was reported in the literature that MNPs in the dative and with a masculine or neuter kind noun favour the \PGCa\ more than the corresponding nominative and accusative MNPs \citep{Hentschel1993,Zimmer2015}.
 As an example, \textit{mit einem Stück frischen Brots} `with a piece of fresh bread' (\PGCa) would be preferred over \textit{mit einem Stück frischem Brot} (\NACa).
@@
 To control for this effect, the case of the measure noun was manually annotated (variable \textit{Measurecase}).
-Second, to control for lemma frequency effects, the logarithm-transformed frequency per 1,000,000 words of each lemma was added from the frequency lists distributed by the DECOW designers (variables \textit{Kindfreq}, \textit{Measurefreq}).
-All numeric variables were standardised and centered (z-transformed) to facilitate their interpretation in the regression models reported 
-in the next section.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Hierarchical models of the measure noun alternation}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{A hierarchical model of the measure noun alternation}
 \label{sec:corpushierarchicalmodel}
@@
 The purpose is to model the influence of the regressors specified in Table~\ref{tab:variables} on the probability that either the \NACa\ or the \PGCa\ is chosen.
-As explained in Section~\ref{sec:descriptive}, due to extensive case syncrestism in feminine NPs, feminine kind nouns in MNPs like \textit{(mit) einem Liter kalter Milch} `with a litre of cold milk' can either be datives (case agreement, \NACa) or genitives (\PGCa) whenever the measure noun is in the dative.
-All dative MNPs with feminine kind nouns had to be removed, and the regressor \textit{Measurecase} consequently only had two levels with feminine kind nouns.%
-\footnote{Together with other cleanups (erroneous hits, extremely noisy material), the removal of such uninformative sentences reduced the total size of the sample to 5,063 sentences.}
-Therefore, two separate models for masculine\slash neuter and feminine kind nouns were fitted.
-For the masculine\slash neuter model (henceforth \textit{MN model}), the final sample size was $n_{\text{MN}}=3,949$ with 672 cases of \PGCa\ and 3,277 cases of \NACa.
-For the feminine model (henceforth \textit{F model}), it was $n_{\text{F}}=1,114$ with 462 cases of \PGCa\ and 652 cases of $\NACa$.
 All regressors from Table~\ref{tab:variables} were included to begin with, and the measure lemma and the kind noun lemma were specified as varying-intercept random effects.
+The sample size was $n=5,063$ with $1,1344$ cases of \PGCa\ and $3,929$ cases of \NACa.
 The regressors listed in Table~\ref{tab:variables} with the measure lemma as their unit of reference have no within-measure lemma variance, and the \textit{glmer} function automatically estimates them as \textit{group level predictors} (or \textit{second-level effects}), cf.\ \cite[265--269,302--304]{GelmanHill2006}.
@@
 The results of the estimation are shown in Table~\ref{tab:bigtable}.
-For the response variable, \PGCa\ was coded as $1$, such that coefficients leaning to the positive side can be interpreted as favouring the \NACa.
-
-\input{bigtable}
-
-%Second-level models were specified for kind nouns lemmas and measure noun lemmas.
-%Equations~\ref{eq:model:mnglmm} through \ref{eq:model:mnglmm:alphakind} show the specified structure of the model.
-%I use the same notation as used in \cite{GelmanHill2006} with the exception that I use vector notation for regressor variables and coefficients if the number of regressors is large in order to make the model specification more compact.%
-%\footnote{
-%The response $y_i=1$ if the PGC is chosen in sentence $i$, else $0$.
-%$\beta_0$ is the general intercept, $B$ a vector of first-level coefficients, and $X_i$ a vector of regressor variables for sentence $i$ (see Table~\ref{tab:bigtable}).
-%$\alpha^{\text{measure}}_{j[i]}$ is the random intercept for measure lemma $j$, chosen appropriately for the measure lemma occurring in sentence $i$ (similarly, $\alpha^{\text{kind}}_{k[i]}$ for kind lemma $k$).
-%$\gamma_0^{\text{measure}}$ is the intercept of the second-level model for measure lemmas, $\Gamma^{\text{measure}}$ a vector of second-level coefficients for measure lemmas, $U_j$ a vector of regressor variables for measure lemma $j$, and $\sigma_{\text{measure}}^2$ the variance of the measure lemma intercepts.
-%The second-level model for kind lemmas is specified in the same manner.
-%$\alpha\sim\mathcal{N}(\mu,\sigma^2)$ should be read \textit{$\alpha$ is normally distributed with mean $\mu$ and variance $\sigma^2$}.
-%}
-%
-%\begin{eqnarray}
-%    \label{eq:model:mnglmm}
-%    Pr(y_i=1) &=& logit^{-1}(\beta_0 + B\cdot X_i + \alpha^{\text{measure}}_{j[i]} + \alpha^{\text{kind}}_{k[i]}) \\
-%    \alpha^{\text{measure}}_j &\sim& \mathcal{N}(\gamma_0^{\text{measure}} + \Gamma^{\text{measure}}\cdot U_j,\sigma^2_{\text{measure}}) \\
-%    \alpha^{\text{kind}}_k &\sim& \mathcal{N}(\gamma_0^{\text{kind}} + \Gamma^{\text{kind}}\cdot U_k,\sigma^2_{\text{kind}}) 
-%  \label{eq:model:mnglmm:alphakind}
-%\end{eqnarray}
-
-Regressors were removed if no convergence could be achieved (even after tuning the optimizer) or perfect separation occurred.
-Beyond that, no step-down model selection was performed (see Section~\ref{sec:rightstatistics}).
-Generalized variance inflation factors for the remaining regressors were calculated to check for multicollinearity \citep{FoxMonette1992}.
-Except for the levels of lemma noun class factors \textit{Kindconsistency} and \textit{Measureclass}, virtually no regressor has a GVIF over 3.
-Exceptions were GVIF(\textit{Kindattraction})=3.17 and GVIF(\textit{Kindfreq})=3.92, both still being far from dramatic.
-The high GVIFs for the noun lemma classes (up to 46.2) correlate with a low degree of theoretical motivation (at least for the concrete classes as annotated).
-However, re-estimating the models with these variables removed did \textit{not} alter the sign or magnitude of any of the other regressors, and I left them in the model.
-
-%\begin{table}
-%  \centering
-%  \begin{tabular}{lrr}
-%    Regressor (level)         &  Masculine\slash Neuter &   Feminine \\
-%    \midrule
-%    Badness                   &   1.02                  &   1.04     \\
-%    Genitives                 &   1.11                  &   1.13     \\
-%    Kindattraction            &   2.17                  &   3.17     \\
-%    KindconsistencyGaseous    &                         &   6.27     \\
-%    KindconsistencyHard       &                         &   5.86     \\
-%    KindconsistencyImmaterial &                         &   4.32     \\
-%    KindconsistencyLiquid     &                         &   8.45     \\
-%    KindconsistencyMixed      &                         &   7.28     \\
-%    KindconsistencyPowder     &                         &   3.87     \\
-%    KindconsistencySoft       &                         &  15.09     \\
-%    Kindedible0               &  (1.32)                 &   2.25     \\
-%    Kindfreq                  &   2.13                  &   3.92     \\
-%    KindintNonint             &  (1.27)                 &   2.45     \\
-%    Matchlength               &   1.17                  &   1.62     \\
-%    Measureabbreviated0       &   1.28                  &   1.29     \\
-%    Measureattraction         &   1.99                  &   1.06     \\
-%    MeasurecaseAcc            &   1.35                  &   1.07     \\
-%    MeasurecaseDat            &   1.37                  &            \\
-%    MeasureclassAmount        &  46.20                  &            \\
-%    MeasureclassCapacity      &  16.32                  &            \\
-%    MeasureclassCurrency      &   3.50                  &            \\
-%    MeasureclassLayer         &   7.92                  &            \\
-%    MeasureclassLength        &   2.14                  &            \\
-%    MeasureclassPortion       &  36.91                  &            \\
-%    MeasureclassRest          &   4.54                  &            \\
-%    MeasureclassVessel        &  18.81                  &            \\
-%    MeasureclassWeight        &  25.40                  &            \\
-%    Measurefreq               &   2.27                  &  (1.21)    \\
-%    MeasurenumberSg           &   1.90                  &   1.65     \\
-%    Minus1posA                &   3.77                  &   4.19     \\
-%    Minus1posD                &   4.05                  &   5.25     \\
-%    Minus1posO                &   1.30                  &   1.47     \\
-%    Minus1posP                &   1.21                  &   1.10     \\
-%  \end{tabular}
-%  \caption{Generalized variance inflation factors for the regressors of the hierarchical models}
-%  \label{tab:vif}
-%\end{table}
-
-Standard diagnostics show that the model quality is quite high.
-For the MN model, Nakagawa \& Schielzeth's pseudo-coefficient of determination is $R_m^2=0.495$ and $R^2_c=0.544$ (see \citealp{Gries2015} for a linguist-friendly introduction to these $R^2$ measures, or else \citealp{NakagawaSchielzeth2013}).
-The rate of correct predictions is $0.872$, which means a proportional reduction of error of $\lambda=0.247$.
-For the F model, $R_m^2=0.468$ and $R^2_c=0.574$ are achieved, as well as a correct prediction rate of $0.815$, which means a proportional reduction of error of $\lambda=0.554$.
-The measure and kind lemma intercepts are distributed with the following standard deviations:
-MN model $\sigma_{\text{Measurelemma}}<0.001$ (131 lemmas), $\sigma_{\text{Kindlemma}}=0.591$ (68 lemmas);
-F model $\sigma_{\text{Measurelemma}}=0.905$ (73 lemmas), $\sigma_{\text{Kindlemma}}=0.000$ (19 lemmas).
-In the MN model, variance between individual measure lemmas do not play much of role, mostly because the second-level regressors successfully model the between-lemma variance.
-In the F model, kind lemmas (and -- as will be shown -- the second level model for kind lemmas) could not be estimated with great success, most likely because of the low number of levels.%
-\footnote{This low number is expected given that most mass nouns are masculine and neuter in German.}
-This will be discussed in Section~\ref{sec:interpretation}.
-
-In Table~\ref{tab:bigtable}, there are groups of columns for the results obtained from Maximum Likelihood estimation and Markov-Chain Monte Carlo estimation, both for the MN model and F models and including confidence intervals.
-In the remainder of this section, I describe how these results were obtained.
-The coefficient estimates are specified for each regressor (level) in the columns labelled \textit{Coef}.%
-\footnote{Empty cells in the table correspond to regressors and their levels which had to be removed because of convergence problems.}
+Given the coding of the response variable, coefficients leaning to the positive side can be interpreted as favouring the \PGCa.
+
+\begin{sidewaystable}
+  \centering
+    \begin{tabular}{lllrrrrrrrcc}
+    Level     & Regressor         & Level     & \pPB  & \multicolumn{2}{l}{Coefficient} & \multicolumn{2}{l}{CI Low} & \multicolumn{2}{l}{CI High} & \multicolumn{2}{l}{CI contains 0} \\
+    &                   &           &       & \multicolumn{1}{l}{MLE} & \multicolumn{1}{l}{MCMC} & \multicolumn{1}{l}{MLE} & \multicolumn{1}{l}{MCMC} & \multicolumn{1}{l}{MLE} & \multicolumn{1}{l}{MCMC} & \multicolumn{1}{l}{MLE} & \multicolumn{1}{l}{MCMC} \\\midrule
+
+    First     & Badness           & (numeric) & 0.001 & -0.165 & -0.169 & -0.264 & -0.264 & -0.074 & -0.075 &      * &      * \\
+              & Genitives         & (numeric) & 0.001 & -0.630 & -0.645 & -0.710 & -0.739 & -0.523 & -0.552 &      * &      * \\
+              & Leftcontext       & D         & 0.001 &  1.178 &  1.173 &  0.751 &  0.784 &  1.548 &  1.546 &      * &      * \\
+              &                   & O         &       &  1.561 &  1.588 &  1.093 &  1.179 &  1.919 &  2.007 &      * &      * \\
+              &                   & A         &       &  2.300 &  2.340 &  1.872 &  1.959 &  2.644 &  2.726 &      * &      * \\
+              &                   & P         &       &  2.353 &  2.371 &  1.693 &  1.766 &  2.900 &  2.978 &      * &      * \\
+              & Measurecase       & Acc       & 0.001 &  0.065 &  0.070 & -0.130 & -0.119 &  0.268 &  0.269 &        &        \\
+              &                   & Dat       &       &  0.707 &  0.735 &  0.413 &  0.480 &  0.959 &  1.002 &      * &      * \\
+	      & Measurenumber     & Pl        & 0.001 &  0.814 &  0.825 &  0.534 &  0.574 &  1.080 &  1.079 &      * &      * \\[0.5\baselineskip]
+
+    Second    & Kindattraction    & (numeric) & 0.033 &  0.204 &  0.216 &  0.030 &  0.014 &  0.366 &  0.412 &      * &      * \\
+    (Kind)    & Kindfreq          & (numeric) & 0.066 &  0.161 &  0.173 & -0.008 &  0.009 &  0.333 &  0.350 &        &      * \\
+              & Kindgender        & Neut      & 0.001 & -0.007 & -0.022 & -0.407 & -0.424 &  0.408 &  0.379 &        &        \\
+	      &                   & Fem       &       &  1.357 &  1.379 &  0.831 &  0.872 &  1.808 &  1.897 &      * &      * \\[0.5\baselineskip]
+
+    Second    & Measureattraction & (numeric) & 0.015 &  0.247 &  0.275 &  0.058 &  0.078 &  0.413 &  0.476 &      * &      * \\
+    (Measure) & Measureclass      & Rest      & 0.001 &  0.564 &  0.488 & -0.140 & -0.309 &  1.197 &  1.236 &        &        \\
+              &                   & Container &       &  0.609 &  0.618 &  0.050 &  0.018 &  1.145 &  1.215 &      * &      * \\
+              &                   & Amount    &       &  0.927 &  0.959 &  0.184 &  0.189 &  1.566 &  1.718 &      * &      * \\
+              &                   & Portion   &       &  1.472 &  1.503 &  0.916 &  0.956 &  1.938 &  2.064 &      * &      * \\
+              & Measurefreq       & (numeric) & 0.003 & -0.257 & -0.264 & -0.413 & -0.435 & -0.098 & -0.099 &      * &      * \\
+  \end{tabular}
+  \smallskip
+  \caption{Coefficient table comparing Maximum Likelihood Estimation (MLE, with 95\% bootstrap confidence interval) and `Bayesian' Markov-Chain Monte Carlo estimation (MCMC); for details see text}
+  \label{tab:bigtable}
+\end{sidewaystable}
+
+Standard diagnostics for the Maximum Likelihood Estimation (MLE) show that the model quality is quite high.
+Generalized variance inflation factors for the regressors were calculated to check for multicollinearity \citep{FoxMonette1992,ZuurEa2010}, and none of the corrected $\text{GVIF}^{1/2\text{df}}$ was higher than $2$.
+Nakagawa \& Schielzeth's pseudo-coefficient of determination is $R_m^2=0.423$ and $R^2_c=0.515$ (see \citealp{Gries2015} for a linguist-friendly introduction to these $R^2$ measures, or else \citealp{NakagawaSchielzeth2013}).
+The rate of correct predictions is $0.855$, which means a proportional reduction of error of $\lambda=0.354$.
+The measure lemma intercepts have a standard deviation of $\sigma_{\text{Measurelemma}}=0.623$, the kind lemma intercept $\sigma_{\text{Kindlemma}}=0.488$
+
+The coefficient estimates are specified in Table~\ref{tab:bigtable} for each regressor (or regressor level) in the columns labelled \textit{Coefficient}.
 For a robust quantification of the precision of the estimation, I ran a parametric bootstrap (using the \textit{confint.merMod} function from \textit{lme4}) with 1,000 replications, and using the percentile method for the calculation of the intervals.
-The resulting 95\% bootstrap confidence intervals are reported in Table~\ref{tab:bigtable} in the columns labelled \textit{Low}, \textit{High} (for the upper and lower 2.5th percentile) and \textit{Width} (for the width of the interval).
-The column \textit{Not 0} shows an asterisk for those intervals that do not include 0 as a more modest replacement of significance levels.
-
-As mentioned above, no step-down model selection was performed.
-However, for each regressor, a $p$ value was obtained by dropping the regressor from the full model, re-estimating the nested model and comparing it to the full model.
+The resulting 95\% bootstrap confidence intervals are reported in Table~\ref{tab:bigtable} in the columns labelled \textit{CI Low} and \textit{CI High} (for the upper and lower 2.5th percentiles).
+The column \textit{CI contains 0} shows an asterisk for those intervals that do not include 0 as a more modest replacement of significance levels.
+Furthermore, for each regressor, a $p$ value was obtained by dropping the regressor from the full model, re-estimating the nested model and comparing it to the full model.
 Instead of inexact Wald approximations or Likelihood Ratio Tests, I used a drop-in bootstrap replacement for the Likelihood Ratio Test from the function \textit{PBmodcomp} from the \textit{pbkrtest} package \citep{HalekohHojsgaard2014}.
 These bootstrapped $p$ values are given in the columns labelled p\Sub{PB} in Table~\ref{tab:bigtable}.
+All regressors have low \pPB\ values of magnitude $10^{-2}$ and $10^{-3}$, and only \textit{Kindattraction} ($p_{\text{PB}}=0.033$) and \textit{Kindfreq} ($p_{\text{PB}}=0.066$) can be seen as slightly too high to be convincing (`non-significant').
 
@@
 The algorithm converged, and for all coefficients, the $\hat{R}$ diagnostic was exactly $1$.
-The resulting coefficients and intervals as well as an * for \textit{interval does not contain $0$} (column \textit{Not 0}) are also given in Table~\ref{tab:bigtable} in the groups of columns labelled \textit{MCMC}.
-In the interpretation of the results in the next section, it will become apparent that both methods of estimation lead to the exact same theoretically relevant results.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%\begin{figure}[h]
-%\centering
-%\includegraphics[width=0.8\textwidth]{figures/corpus/04_glmm_fixef_firstlevel}
-%\caption{Estimated first-level effects in GLMMs for Masculine\slash Neuter and Feminine}
-%\label{fig:glmmfixef}
-%\end{figure}
-%
-%\begin{figure}[h]
-%\centering
-%\includegraphics[width=0.8\textwidth]{figures/corpus/04_glmm_fixef_secondlevel_kind}
-%\caption{Estimated second-level effects for kind lemmas in GLMMs for Masculine\slash Neuter and Feminine}
-%\label{fig:glmmfixef}
-%\end{figure}
-%
-%\begin{figure}[h]
-%\centering
-%\includegraphics[width=0.8\textwidth]{figures/corpus/04_glmm_fixef_secondlevel_measure}
-%\caption{Estimated second-level effects for measure lemmas in GLMMs for Masculine\slash Neuter and Feminine}
-%\label{fig:glmmfixef}
-%\end{figure}
-%
+The resulting coefficients and intervals as well as an * are also given in Table~\ref{tab:bigtable} in the columns labelled \textit{MCMC}.
+Since both methods lead to exactly the same results (minus negligible numerical differences) as expected (see Section~\ref{sec:rightstatistics}), I only interpret the MLE model in the next section.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
   \centering
-  \includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_mn_Kindattraction}~\includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_fem_Kindattraction}
-  \caption{Effect plot for the regressor \textit{Kindattraction} for the MN model and the F model (y axes have different scales)}
-  \label{fig:eff:kindattraction}
-\end{figure}
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_mn_Measureattraction}~\includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_fem_Measureattraction}
-  \caption{Effect plot for the regressor \textit{Measureattraction} for the MN model and the F model (y axes have different scales)}
+  \includegraphics[width=0.75\textwidth]{../R/output/corpus_Measureattraction}
+  \caption{Effect plot for the regressor \textit{Measureattraction}}
   \label{fig:eff:measureattraction}
@@
 
-First, the prototypicality effect related to the non-alternating \PGCd\ and \NACb\ can be shown, see the effect plots in Figure~\ref{fig:eff:kindattraction} and Figure~\ref{fig:eff:measureattraction}.%
+First, the prototypicality effect related to the non-alternating \PGCd\ and \NACb\ can be shown, see the effect plots in Figure~\ref{fig:eff:measureattraction}.%
 \footnote{Effect plots were created using the \textit{effects} package \citep{Fox2003}.
@@
   \centering
-  \includegraphics[width=0.6\textwidth]{figures/corpus/04_glmm_fixeff_mn_Measureclass}
-  \caption{Effect plot for the regressor \textit{Measureclass} for the MN model}
+  \includegraphics[width=0.75\textwidth]{../R/output/corpus_Measureclass}
+  \caption{Effect plot for the regressor \textit{Measureclass}}
   \label{fig:eff:measureattraction}
@@
   \centering
-  \includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_mn_Minus1pos}~\includegraphics[width=0.4\textwidth]{figures/corpus/04_glmm_fixeff_fem_Minus1pos}
-  \caption{Effect plot for the regressor \textit{Minus1pos} for the MN model and the F model (y axes have different scales)}
-  \label{fig:eff:measureattraction}
+  \includegraphics[width=0.75\textwidth]{../R/output/corpus_Leftcontext}
+  \caption{Effect plot for the regressor \textit{Leftcontext}}
+  \label{fig:eff:leftcontext}
 \end{figure}
@@
 In the F model, the effect has the same direction, but the estimate is quite imprecise ($p=0.160$), again likely due to the low sample size.
-Figure~\ref{fig:eff:measureattraction} shows, however, that the left context of the measure noun influences the choice of the variant (in both models, the regressor has $p=0.002$).
+Figure~\ref{fig:eff:leftcontext} shows, however, that the left context of the measure noun influences the choice of the variant (in both models, the regressor has $p=0.002$).
 Cardinals have a strong tendency to co-occur with the \NACa\ in both models.
@@
 % => Genitives
-% => Matchlength
-% => Measureabbreviated
 
@@
 
-First, a very simple computer-aided forced-choice experiment was conducted.%
-\footnote{The experiment is \textit{simple} compared to the sophisticated origins of forced choice experiments and their interpretation in psychophysics \citep[166--179]{MacmillanCreelman2005}.}
-From a linguistic point of view, the forced choice setting has the advantage over rating studies that is does not force participants into making an absolute acceptability judgement on some scale.
-Instead, they express a preference for one of two constructions (or words, etc.) both of which are acceptable.
-While this does not totally exclude normative thinking -- to which native speakers of German in so-called \textit{cases of doubt} are typically prone --
-
-There were 24 participants aged 19 to 30 who were recruited from introductory linguistics courses at Freie Universität Berlin.
-Although the experiment was conducted in the last four weeks of their first semester, participants had no deeper explicit knowledge of linguistics or grammar.%
-\footnote{Even though the ratio of target sentences to fillers was relatively high (see below) and some of the fillers were also related to case alternations, no participant used the word \textit{Kasus} `case' when asked whether they had an idea what the experiment was about in informal interviews after the experiment.}
-Only participants with an L1 background in German were recruited.
+First, a forced-choice experiment was conducted.
+There were 24 participants (L1 speakers of German without reading or writing disability) aged 19 to 30 who were recruited from introductory linguistics courses at Freie Universität Berlin.
+Although the experiment was conducted in the last four weeks of their first semester, participants had no deeper explicit knowledge of linguistics, grammar, or experimental methods.
+None had never participated in a forced-choice experiment before.
+Participation was voluntary, but participants received credit in partial fulfillment of course requirements.
 
@@
 Concretely, I sampled 16 attested sentences from the corpus and simplified them (mostly removing embedded clauses, complex adverbials, etc.) and standardized their orthography.
-It was made sure that the simplifications and normalisations did not affect any of the regressors used in the corpus study, of course.
+It was made sure that the simplifications and normalisations did not affect any of the regressors used in the corpus study.
 Eight sentences contained masculine or neuter kind nouns, the other eight contained feminine kind nouns.
@@
 It was made sure that the highly predictive features according to Section \ref{sec:interpretation} occurred in unique combinations among the four sentences in each cell.
-This is summarised in Table \ref{tab:experiment1:design}.
+The design is summarised in Table \ref{tab:experiment1:design}.
 
@@
 \centering
-\includegraphics[width=0.6\textwidth]{figures/experiment/2afc_proportions}
+\includegraphics[width=0.75\textwidth]{../R/output/fc_proportions}
 \caption{Participants' responses in the forced-choice task plotted against the predictions from the corpus-based model}
@@
 
-Descriptively, Figure \ref{fig:afc:proportions} shows the relationship between the logits calculated from the corpus-based model and the proportional choices made by participants.
+Figure \ref{fig:afc:proportions} shows the relationship between the logits calculated from the corpus-based model and the proportional choices made by participants.
 Visually, it looks like there is a close correlation between the corpus-derived preferences and preferences in the forced choice experiment.
 
-Then, a multilevel logistic regression was specified as in Equation \ref{eq:model:afc}.
-
-\begin{equation}
-  \begin{split}
-  Pr(y_i=1) & = logit^{-1}(\beta^0+\beta^1\cdot x_i^1+\beta^2\cdot x_i^2+\beta^{1.2}\cdot x_i^1\cdot x_i^2+ \\
-  & + \alpha^{\text{Item}}_{j_i} + \alpha^{\text{Participant}}_{k_i}) \\
-  \text{where}\ \ & x^1 = \text{\it Modelpred} \\
-  & x^2 = \text{\it KindgenderMN} \\
-  %ResponseDrop~Modelpred*Kindgender+(1|Item)+(1|Participant)
-  \end{split}
-  \label{eq:model:afc}
-\end{equation}
-
+Then, a multilevel logistic regression was specified.
 Coefficients were estimated using Maximum Likelihood, and the coefficient table is given in Table \ref{tab:afc:model}, where instead of approximate standard errors and $z$ and $p$ values the lower and the upper bound of a bootstrapped 95\% confidence interval are specified.
@@
 \centering
-\includegraphics[width=0.6\textwidth]{figures/experiment/2afc_effects}
+\includegraphics[width=0.75\textwidth]{../R/output/fc_effects}
 \caption{Effect plot for the multilevel logistic regression in the forced-choice experiment: predictability of participants' choices from the logits of the corpus-based GLMM (cf.\ Section~\ref{sec:corpushierarchicalmodel})}
@@
 \centering
-\includegraphics[width=0.6\textwidth]{figures/experiment/spr_effects_case+modelPred+gender}
+\includegraphics[width=0.75\textwidth]{../R/output/spr_effects}
 \caption{Effect plot for the LMM in the self-paced reading experiment: modeling participants' residualised log reading times on the predictions (logits) of the corpus-based GLMM (cf.\ Section~\ref{sec:corpushierarchicalmodel})}

### 2017-03-28 R/0_wrapper.R

@@ -54,4 +54,4 @@ cat("\n\n Prediction for experimental stimuli ...\n")
 source('5_stimuli_predict.R')
 
 # Save workspace.
-save(list = ls(), file=paste(out.dir, "workspace_", Sys.time, ".RData", sep=""))
+save(list = ls(), file=paste(out.dir, "workspace_", Sys.time(), ".RData", sep=""))

### 2017-03-28 R/3_glmm.R

@@ -117,7 +117,7 @@ for (eff in effs) {
   if (save.persistent) pdf(paste(out.dir, eff, ".pdf", sep=""))
   p <- plot(effect(eff, measure.glmm),
             rug=F, 
-            main = paste("EFfect plot: ", eff),
+            main = paste("Effect plot: ", eff),
             ylab = "Probability of PGCa",
             colors = c("black", "darkblue"), cex = fixeff.pl.cex)
   print(p)
@@ -140,3 +140,10 @@ do.call(ranef.plot, c(list(measure.glmm, measure, "Kindlemma", n.select, main = 
 par(mfrow=c(1,1))
 if (save.persistent) dev.off()
 
+
+
+source('highstat.r')
+if (save.persistent) sink(paste(out.dir, "glmm.txt", sep=""), append = T)
+cat("\n\n Variance inflation diagnostics\n")
+print(myvif(measure.glmm))
+if (save.persistent) sink()

### 2017-03-28 R/6_exp-fc.R

@@ -2,22 +2,27 @@ require(lme4)
 require(effects)
 require(MuMIn)
 require(pbkrtest)
+require(boot)
 
-# rm(list = ls())
+rm(list = ls())
 
 set.seed(239865)
 
 setwd('/Users/user/Workingcopies/measurenps/GitHub/R')
 
 ci.method        <- "boot" # boot
-modcomp.nsim     <- 1000   # 500
+modcomp.nsim     <- 1000   # 1000
 boot.nsim        <- 1000   # 1000
+the.nAGQ         <- 0 
 out.dir          <- "output/fc_"
 save.persistent  <- T
 data.dir         <- "data/fc/"
 data.files.names <- list.files(data.dir)
 data.files.names <- data.files.names[grep('csv$', data.files.names)]
 data.files       <- paste(data.dir, data.files.names, sep="")
+
+# Helpers.
+source('glmtools.R')
 
 load("output/corpus_stims_pred.RData")
 
@@ -41,7 +46,7 @@ reconstruct.rating <- function(c) {
 
 # These are the predictions from the actual model.
 responses.df <- data.frame(
-    Modelpred     = rep(stims.pred, ncol(results))
+    Modelpred     = rep(inv.logit(stims.pred), ncol(results))
   , Kindgender    = factor(rep(c(rep("MN", 8), rep("F", 8)), ncol(results)))
   , ResponseDrop  = factor(as.vector(apply(results, 2, reconstruct.rating)))
   , Item          = as.factor(rep(c(rep(c(1:8), 2), rep(c(9:16), 2)), ncol(results)))
@@ -61,16 +66,15 @@ if (save.persistent) sink(paste(out.dir, "results.txt", sep=""))
 # Actual model.
 cat("\n\nGLMM, predicting reactions from corpus model\n\n")
 model.fc <- glmer(Chosenconstruction~Modelprediction+(1|Item)+(1|Participant),
-                  family = binomial(link = "logit"), data = responses.df)
+                  family = binomial(link = "logit"), data = responses.df, nAGQ=the.nAGQ,
+                  control=glmerControl(optimizer="nloptwrap2", optCtrl=list(maxfun=2e5)))
 print(summary(model.fc))
 
 # Model comparison.
-cat("\n\nANOVA on nested models w & w/o Modelpred\n\n")
-model.fc.0 <- glmer(Chosenconstruction~Kindgender+(1|Item)+(1|Participant),
-                    family = binomial(link = "logit"), data = responses.df)
[…]

### 2017-03-28 R/7_exp-spr.R

@@ -3,6 +3,7 @@ require(effects)
 require(MuMIn)
 require(pbkrtest)
 require(plyr)
+require(boot)
 
 rm(list = ls())
 
@@ -25,11 +26,12 @@ data.files.names <- list.files(data.dir)
 data.files.names <- data.files.names[grep('csv$', data.files.names)]
 data.files       <- paste(data.dir, data.files.names, sep="")
 
+
 load("output/corpus_stims_pred.RData")
 
 assigs <- read.delim(assigs.fn, header = T, sep = ",", quote = "", stringsAsFactors = F)
 stims  <- read.delim(stims.fn, header = T, sep = "\t", quote = "", stringsAsFactors = F)
-stims$Modelprediction <- rep(c(stims.pred, rep(0, 33)), 2)
+stims$Modelprediction <- rep(c(inv.logit(stims.pred), rep(0, 33)), 2)
 
 parts.rt.var <- NULL
 
@@ -152,7 +154,7 @@ if (save.persistent) sink()
 # Make formula and estimate model.
 # ################################
 
-# Model with coprpus-based logits as input.
+# Model with coprpus-based probs as input.
 formule <- paste(target, "~Construction*Modelprediction+Position+(1|Participant)", sep = "")
 formule.0 <- paste(target, "~Position+(1|Participant)", sep = "")
 rt.model <- lmer(formule, data = rt.all, REML = F)
@@ -165,7 +167,7 @@ rt.model.0 <- lmer(formule.0, data = rt.all, REML = F)
 
 if (save.persistent) sink(paste(out.dir, "results.txt", sep=""), append = T)
 
-cat("\n\n##### Model w/corpus-based logits #####\n\n")
+cat("\n\n##### Model w/corpus-based probabilities #####\n\n")
 
 print(summary(rt.model))
 
@@ -187,7 +189,7 @@ if (save.persistent) sink()
 # #############
 
 p <- plot(effect("Construction:Modelprediction", rt.model, KR = T), rug=F, colors = c("black", "darkblue"),
-          ylab="Residual log. reading time", xlab = "Logits from corpus-based model",
+          ylab="Residual log. reading time", xlab = "Probability for PGCa from corpus-based model",
           main="Predicting reading times from\ncorpus-based model")
 if (save.persistent) pdf(paste(out.dir, "effects.pdf", sep=""))
 print(p)
[…]

### 2017-03-29 Latex/article.tex
corpus finished; exp 1 progress

@@
 %
-If the measure noun denotes a container or vessel, constructions with \textit{voll} (\textit{von})\slash\textit{voller} `full (of)' and \textit{mit} `with' are available, see (\ref{ex:alternatives}).
+If the measure noun denotes a container, constructions with \textit{voll} (\textit{von})\slash\textit{voller} `full (of)' and \textit{mit} `with' are available, see (\ref{ex:alternatives}).
 
@@
 There are no clear hypotheses about such interactions, however.
-Nevertheless, the corpus-based models in Section~\ref{sec:corpusstudies} are specified in a way that they could detect preferences for either NAC or PGC for the class of container or vessel measure nouns, for which the alternative constructions illustrated in (\ref{ex:alternatives}) are available.
+Nevertheless, the corpus-based models in Section~\ref{sec:corpusstudies} are specified in a way that they could detect preferences for either NAC or PGC for the class of container measure nouns, for which the alternative constructions illustrated in (\ref{ex:alternatives}) are available.
 
@@
 I also develop my own analysis and the appropriate hypotheses for the empirical studies presented in Sections~\ref{sec:corpusstudies} and \ref{sec:externalvalidation}.
-%While the constructions and the alternation are standardly covered in descriptive grammars such as \citet[XYZ]{Eisenberg2013a}, the research literature contains little specific discussion of the situation in German.
 I discuss four main factors that might influence the alternation:
@@
 Abstract nouns which partially behave like mass nouns (like \textit{Spaß} `fun’ or \textit{Gefahr} `danger’) were excluded because they are usually not quantified in the same way as concrete mass nouns.
-%\footnote{They behave like mass nouns mostly in that they occur without a determiner in the singular.}
 The hundredth selected mass noun was \textit{Schmuck} `jewelry’, which is the 3,054th most frequent noun in the original frequency list.
@@
 In step (\ref{enum:data:step3}), each of these 2,365 noun–noun combinations was queried in the target constructions (\PGCa\ and \NACa) individually in each of the first ten slices of DECOW (roughly 10 billion tokens).
-%The scripts generating the actual queries were crafted so as to make sure that adjectives and nouns were searched in the appropriate forms.
-%The CQL query in (\ref{ex:data:query01}) shows an example query for a neuter mass noun like \textit{Wasser} `water' which finds relevant sequences such as the ones shown in (\ref{ex:data:query02}).%
-%\footnote{The COW creators used to make corpora available in the IMS Open Corpus Query Workbench \citep{EvertHardie2011}, which has a query syntax informally referred to as CQL.
-%Documentation can be obtained at \url{http://cwb.sourceforge.net}.}
-%The versions with the genitive singular \textit{Liters} and the dative plural \textit{Litern} are not shown in (2) for space reasons but are also found by the query in (1).
-%In terms of context, single sentences containing the target were exported.
-%
-%\begin{exe}
-%  \ex{\label{ex:data:query01} \texttt{\small[ lemma="Liter" ] [ pos="ADJA" \& word=".+(en|em|es)" ] [ lemma="Wasser" ]}}
-%  \ex\label{ex:data:query02}
-%  \begin{xlist}
-%    \ex{\label{ex:data:query02a} Liter [klaren Wassers]\Sub{Gen}}
-%    \ex{\label{ex:data:query02b} Liter [klares Wasser]\Sub{Nom\slash Acc}}
-%    \ex{\label{ex:data:query02c} Liter [klarem Wasser]\Sub{Dat}}
-%  \end{xlist}
-%\end{exe}
 In order to reduce the sample size for the manual annotation process, the final concordance was sampled from the results of these 2,365 queries.
@@
 \footnote{In a similar fashion, the 100 most frequent measure nouns occurring with plural kind nouns were bootstrapped and queried, resulting in a sample of 1,876 sentences.
-%Bootstrapping measure nouns first was more efficient in the plural for reasons to complicated to discuss here, especially because the data was not properly analysed anyway.
 As stated in Section~\ref{sec:germanmeasurenps}, the NAC is virtually never used with plural kind nouns, and this sample was not used for full analysis.
-%Consequently, it does not make sense to model the influencing factors in the same way as for singular kind nouns.
 }
@@
 To capture this, class information was annotated for measure lemmas.
-The classification was designed inspired by the list in \cite[530]{Koptjevskaja2001}, but due to the frequencies of the respective nouns, very coarse classes were necessary.
+The classification was designed inspired by the list in \cite[530]{Koptjevskaja2001}, but due to the low frequencies of many of the potential classes, a very coarse classification was finally used.
 With typical examples and the number of data points in the final sample, the classes are:
@@
 
+\begin{figure}[hb!]
+  \centering
+  \includegraphics[width=0.85\textwidth]{../R/output/corpus_fixeffs}
+  \caption{Coefficients (MLE) with 95\% parametric bootstrap confidence intervals; intercept (\textit{Leftcontext=C}, \textit{Measurecase=Nom}, \textit{Measurenumber=Sg}, \textit{Kindgender=Masc}, \textit{Measureclass=Physical}; 0 for all numeric z-transformed regressors) is -4.328}
+  \label{fig:fixeffs}
+\end{figure}
+
 In this section, I report the results of fitting a multilevel model to the data using R \citep{R}, \textit{lme4} \citep{lme4} for Maximum Likelihood Estimation, and \textit{rstanarm} \citep{rstanarm} for Markov-Chain Monte Carlo estimation.
@@
 The sample size was $n=5,063$ with $1,1344$ cases of \PGCa\ and $3,929$ cases of \NACa.
-The regressors listed in Table~\ref{tab:variables} with the measure lemma as their unit of reference have no within-measure lemma variance, and the \textit{glmer} function automatically estimates them as \textit{group level predictors} (or \textit{second-level effects}), cf.\ \cite[265--269,302--304]{GelmanHill2006}.
+The results of the estimation are shown in Table~\ref{tab:bigtable} and in Figure~\ref{fig:fixeffs}.
+The regressors with the measure lemma as their unit of reference have no within-measure lemma variance, and the \textit{glmer} function automatically estimates them as \textit{group level predictors} (or \textit{second-level effects}), cf.\ \cite[265--269,302--304]{GelmanHill2006}.
 The same goes for those listed with the kind lemma as their unit of reference.
-The results of the estimation are shown in Table~\ref{tab:bigtable}.
 Given the coding of the response variable, coefficients leaning to the positive side can be interpreted as favouring the \PGCa.
@@
   \centering
-    \begin{tabular}{lllrrrrrrrcc}
-    Level     & Regressor         & Level     & \pPB  & \multicolumn{2}{l}{Coefficient} & \multicolumn{2}{l}{CI Low} & \multicolumn{2}{l}{CI High} & \multicolumn{2}{l}{CI contains 0} \\
+    \begin{tabular}{llrlrrrrrrcc}
+      Level & Regressor & \multicolumn{1}{l}{\pPB} & Level  & \multicolumn{2}{l}{Coefficient} & \multicolumn{2}{l}{CI Low} & \multicolumn{2}{l}{CI High} & \multicolumn{2}{l}{CI contains 0} \\
               &                   &       &          & \multicolumn{1}{l}{MLE} & \multicolumn{1}{l}{MCMC} & \multicolumn{1}{l}{MLE} & \multicolumn{1}{l}{MCMC} & \multicolumn{1}{l}{MLE} & \multicolumn{1}{l}{MCMC} & \multicolumn{1}{l}{MLE} & \multicolumn{1}{l}{MCMC} \\\midrule
 
-    First     & Badness           & (numeric) & 0.001 & -0.165 & -0.169 & -0.264 & -0.264 & -0.074 & -0.075 &      * &      * \\
-              & Genitives         & (numeric) & 0.001 & -0.630 & -0.645 & -0.710 & -0.739 & -0.523 & -0.552 &      * &      * \\
-              & Leftcontext       & D         & 0.001 &  1.178 &  1.173 &  0.751 &  0.784 &  1.548 &  1.546 &      * &      * \\
-              &                   & O         &       &  1.561 &  1.588 &  1.093 &  1.179 &  1.919 &  2.007 &      * &      * \\
-              &                   & A         &       &  2.300 &  2.340 &  1.872 &  1.959 &  2.644 &  2.726 &      * &      * \\
-              &                   & P         &       &  2.353 &  2.371 &  1.693 &  1.766 &  2.900 &  2.978 &      * &      * \\
-              & Measurecase       & Acc       & 0.001 &  0.065 &  0.070 & -0.130 & -0.119 &  0.268 &  0.269 &        &        \\
-              &                   & Dat       &       &  0.707 &  0.735 &  0.413 &  0.480 &  0.959 &  1.002 &      * &      * \\
-	      & Measurenumber     & Pl        & 0.001 &  0.814 &  0.825 &  0.534 &  0.574 &  1.080 &  1.079 &      * &      * \\[0.5\baselineskip]
+    First     & Badness           & 0.001 & (numeric) & -0.165 & -0.169 & -0.264 & -0.264 & -0.074 & -0.075 &      * &      * \\
+              & Genitives         & 0.001 & (numeric) & -0.630 & -0.645 & -0.710 & -0.739 & -0.523 & -0.552 &      * &      * \\
+              & Leftcontext       & 0.001 & D         &  1.178 &  1.173 &  0.751 &  0.784 &  1.548 &  1.546 &      * &      * \\
+              &                   &       & O         &  1.561 &  1.588 &  1.093 &  1.179 &  1.919 &  2.007 &      * &      * \\
+              &                   &       & A         &  2.300 &  2.340 &  1.872 &  1.959 &  2.644 &  2.726 &      * &      * \\
+              &                   &       & P         &  2.353 &  2.371 &  1.693 &  1.766 &  2.900 &  2.978 &      * &      * \\
+              & Measurecase       & 0.001 & Acc       &  0.065 &  0.070 & -0.130 & -0.119 &  0.268 &  0.269 &        &        \\
+              &                   &       & Dat       &  0.707 &  0.735 &  0.413 &  0.480 &  0.959 &  1.002 &      * &      * \\
+	      & Measurenumber     & 0.001 & Pl        &  0.814 &  0.825 &  0.534 &  0.574 &  1.080 &  1.079 &      * &      * \\[0.5\baselineskip]
 
-    Second    & Kindattraction    & (numeric) & 0.033 &  0.204 &  0.216 &  0.030 &  0.014 &  0.366 &  0.412 &      * &      * \\
-    (Kind)    & Kindfreq          & (numeric) & 0.066 &  0.161 &  0.173 & -0.008 &  0.009 &  0.333 &  0.350 &        &      * \\
-              & Kindgender        & Neut      & 0.001 & -0.007 & -0.022 & -0.407 & -0.424 &  0.408 &  0.379 &        &        \\
-	      &                   & Fem       &       &  1.357 &  1.379 &  0.831 &  0.872 &  1.808 &  1.897 &      * &      * \\[0.5\baselineskip]
+    Second    & Kindattraction    & 0.033 & (numeric) &  0.204 &  0.216 &  0.030 &  0.014 &  0.366 &  0.412 &      * &      * \\
+    (Kind)    & Kindfreq          & 0.066 & (numeric) &  0.161 &  0.173 & -0.008 &  0.009 &  0.333 &  0.350 &        &      * \\
+              & Kindgender        & 0.001 & Neut      & -0.007 & -0.022 & -0.407 & -0.424 &  0.408 &  0.379 &        &        \\
+	      &                   &       & Fem       &  1.357 &  1.379 &  0.831 &  0.872 &  1.808 &  1.897 &      * &      * \\[0.5\baselineskip]
 
-    Second    & Measureattraction & (numeric) & 0.015 &  0.247 &  0.275 &  0.058 &  0.078 &  0.413 &  0.476 &      * &      * \\
-    (Measure) & Measureclass      & Rest      & 0.001 &  0.564 &  0.488 & -0.140 & -0.309 &  1.197 &  1.236 &        &        \\
-              &                   & Container &       &  0.609 &  0.618 &  0.050 &  0.018 &  1.145 &  1.215 &      * &      * \\
-              &                   & Amount    &       &  0.927 &  0.959 &  0.184 &  0.189 &  1.566 &  1.718 &      * &      * \\
-              &                   & Portion   &       &  1.472 &  1.503 &  0.916 &  0.956 &  1.938 &  2.064 &      * &      * \\
-              & Measurefreq       & (numeric) & 0.003 & -0.257 & -0.264 & -0.413 & -0.435 & -0.098 & -0.099 &      * &      * \\
+    Second    & Measureattraction & 0.015 & (numeric) &  0.247 &  0.275 &  0.058 &  0.078 &  0.413 &  0.476 &      * &      * \\
+    (Measure) & Measureclass      & 0.001 & Rest      &  0.564 &  0.488 & -0.140 & -0.309 &  1.197 &  1.236 &        &        \\
+              &                   &       & Container &  0.609 &  0.618 &  0.050 &  0.018 &  1.145 &  1.215 &      * &      * \\
+              &                   &       & Amount    &  0.927 &  0.959 &  0.184 &  0.189 &  1.566 &  1.718 &      * &      * \\
+              &                   &       & Portion   &  1.472 &  1.503 &  0.916 &  0.956 &  1.938 &  2.064 &      * &      * \\
+              & Measurefreq       & 0.003 & (numeric) & -0.257 & -0.264 & -0.413 & -0.435 & -0.098 & -0.099 &      * &      * \\
   \end{tabular}
@@
 The rate of correct predictions is $0.855$, which means a proportional reduction of error of $\lambda=0.354$.
-The measure lemma intercepts have a standard deviation of $\sigma_{\text{Measurelemma}}=0.623$, the kind lemma intercept $\sigma_{\text{Kindlemma}}=0.488$
+The measure lemma intercepts have a standard deviation of $\sigma_{\text{Measurelemma}}=0.623$, the kind lemma intercept $\sigma_{\text{Kindlemma}}=0.488$.
 
@@
 For a robust quantification of the precision of the estimation, I ran a parametric bootstrap (using the \textit{confint.merMod} function from \textit{lme4}) with 1,000 replications, and using the percentile method for the calculation of the intervals.
-The resulting 95\% bootstrap confidence intervals are reported in Table~\ref{tab:bigtable} in the columns labelled \textit{CI Low} and \textit{CI High} (for the upper and lower 2.5th percentiles).
-The column \textit{CI contains 0} shows an asterisk for those intervals that do not include 0 as a more modest replacement of significance levels.
+The resulting 95\% bootstrap confidence intervals are reported in Table~\ref{tab:bigtable} in the columns labelled \textit{CI Low} and \textit{CI High} (= upper and lower 2.5th percentiles).
+The column \textit{CI contains 0} shows an asterisk for those intervals that do \textit{not} include 0.
 Furthermore, for each regressor, a $p$ value was obtained by dropping the regressor from the full model, re-estimating the nested model and comparing it to the full model.
@@
 These bootstrapped $p$ values are given in the columns labelled p\Sub{PB} in Table~\ref{tab:bigtable}.
-All regressors have low \pPB\ values of magnitude $10^{-2}$ and $10^{-3}$, and only \textit{Kindattraction} ($p_{\text{PB}}=0.033$) and \textit{Kindfreq} ($p_{\text{PB}}=0.066$) can be seen as slightly too high to be convincing (`non-significant').
+All regressors have low p\Sub{PB} values of magnitude $10^{-2}$ or $10^{-3}$.
+Only \textit{Kindfreq} ($p_{\text{PB}}=0.066$) can be seen as slightly too high to be convincing (`non-significant').
 
@@
 Both the MN model and the F model were re-estimated using the \textit{stan\_glmer} function from the \textit{rstanarm} package, which provides an \textit{lme4}-compatible syntax for estimating common model types with the \textit{stan} software \citep{CarpenterEa2017}.
-The algorithm was run using 4 chains and 1,000 iterations, and instead of guessing prior distributions or using implausible uniform prior distributions \cite[251--252]{Levshina2016}, I used plausible default priors.
+The algorithm was run with $4$ chains and $1,000$ iterations, and instead of guessing prior distributions or using implausible uniform prior distributions \cite[251--252]{Levshina2016}, I used plausible default priors.
 Most notably, priors for coefficients were specified as $\mathcal{N}(0,10)$ because coefficients higher than $10$ or lower than $-10$ are extremely rare in well-specified models on appropriate data.%
-\footnote{Consider that with a coefficient of 10, each increase by $1$ in the regressor variable means an increase in odds of $exp(10)=22026.47$.
+\footnote{Consider that with a coefficient of $10$, each increase by $1$ in the regressor variable means an increase in odds of $exp(10)=22,026.47$.
 To reliably estimate such coefficients, extremely large samples would be required.}
@@
 The resulting coefficients and intervals as well as an * are also given in Table~\ref{tab:bigtable} in the columns labelled \textit{MCMC}.
-Since both methods lead to exactly the same results (minus negligible numerical differences) as expected (see Section~\ref{sec:rightstatistics}), I only interpret the MLE model in the next section.
+Both methods lead to exactly the same results (minus negligible numerical differences) as expected (see Section~\ref{sec:rightstatistics}).
+The signs and magnitudes of the coefficients are identical, and confidence intervals have the same width and symmetry properties.
+If one were interested in significance at $\alpha=0.05$, then the decision based on the confidence intervals differs for \textit{Kindfreq} ($p_{\text{PB}}=0.066$, MLE lower bound is $-0.008$ and MCMC lower bound is $0.009$).
+With MCMC, the regressor is `not significant' at the standard $\alpha$ level, but it is with MLE.
+This might only go to show how absurd strict significance testing is in complex modelling situations:
+both methods essentially estimate the same sign and magnitude for the coefficient (MLE $0.161$, MCMC $0.173$) and do not differ substantially with regard to the precision of the estimate in the form of the width of the confidence interval (both intervals have a width of exactly $0.341$).
+Hence, I only interpret the MLE model in the next section.
 
@@
 
-The results reported in Section~\ref{sec:corpushierarchicalmodel} generally confirm the hypotheses from Section~\ref{sec:analyses}.
-Before I turn to some details, it should be noticed that the results from standard MLE (with bootstrap) and MCMC (with default priors) obviously differ in numerical details -- but not enough to result in different interpretations.
-The signs and magnitudes of the coefficients are identical with the two methods, and the confidence intervals have roughly the same width and symmetry properties.
-In only one single case (MN model, \textit{Measureattraction}) the MLE bootstrap confidence interval does not include $0$ while the MCMC interval does (barely, with $-0.01$).
-While I only interpret the MLE results, virtually the same interpretation could be made for the MCMC results.
-Furthermore, the MN model and the F model agree in many relevant aspects.
-Disagreement between the models will be discussed below.
-
 % ATTRACTION
@@
   \centering
-  \includegraphics[width=0.75\textwidth]{../R/output/corpus_Measureattraction}
-  \caption{Effect plot for the regressor \textit{Measureattraction}}
-  \label{fig:eff:measureattraction}
+  \includegraphics[width=0.5\textwidth]{../R/output/corpus_Measureattraction}~\includegraphics[width=0.5\textwidth]{../R/output/corpus_Kindattraction}
+  \caption{Effect plots for the regressors \textit{Measureattraction} and \textit{Kindattraction}; y axes are not aligned}
+  \label{fig:eff:attraction}
 \end{figure}
 
-First, the prototypicality effect related to the non-alternating \PGCd\ and \NACb\ can be shown, see the effect plots in Figure~\ref{fig:eff:measureattraction}.%
+The results reported in Section~\ref{sec:corpushierarchicalmodel} generally confirm the hypotheses from Section~\ref{sec:analyses}.
+First, the prototypicality effect related to the non-alternating \PGCd\ and \NACb\ can be shown, see the effect plots in Figure~\ref{fig:eff:attraction}.%
 \footnote{Effect plots were created using the \textit{effects} package \citep{Fox2003}.
-They show the changes in probability for the outcome (y axis) dependent on values of a regressor (x axis),assuming typical values of all other regressors.}
+They show the changes in probability for the outcome (y axis) dependent on values of a regressor (x axis), assuming typical values of all other regressors.}
 The effect is mostly as expected:
 if a lemma appears relatively more often in the \PGCd\ (compared to its frequency in the \NACb), the more often the \PGCa\ is chosen over the \NACa\ with this specific lemma.
-Only the F model \textit{Kindattraction} regressor is clearly useless (`non-significant') in the PBmodcomp test, and the effect plot shows a corresponding picture.
-This is related to the fact that in the F model the kind lemma effect and the corresponding second-level model could not be estimated successfully (remember that $\sigma_{\text{Kindlemma}}=0.000$).
-The effect plots show a picture in line with a \textit{divide by 4} interpretation of the coefficients as shown in Table~\ref{tab:bigtable}.%
-\footnote{Coefficients are on the logit scale, which does not linearly correspond to the probability scale.
-If coefficients are reasonably small, however, dividing the coefficient by $4$ gives a good approximation of the change on the probability scale for each increment of the response variable \citep[82]{GelmanHill2006}.}
-Given that all regressors have been z-transformed, we can make statements such as the following:
-If a masculine kind noun is one standard deviation more frequently encountered in the \PGCd\ (compared to its frequency in the \NACb) than the kind nouns on average, the probability that the \PGCa\ is chosen instead of the \NACa\ with this kind noun in it rises by $\beta^{MN}_{\text{Kindattraction}}/4=0.49/4=0.1225$, etc.
-For measure nouns in MNPs with masculine kind nouns, it is $\beta^{MN}_{\text{Measureattraction}}/4=0.21/4=0.0525$, and for measure nouns in MNPs with feminine kind nouns, it is $\beta^{F}_{\text{Measureattraction}}/4=0.43/4=0,1075$.
-As readers can observe, these numbers correspond approximately to the slopes of the effect plots.
+The effect for measure nouns is stronger, and it was estimated with higher precision.
 
@@
 
-In Section~\ref{sec:analyses}, it was also hypothesised that there may be effects specific to classes of lemmas and classes of lemmas, and that they might be related to degrees of grammaticalisation of the measure noun and\slash or semantic prototypicality effects.
-The picture is clear for the MN model.
-Lemma-specific effects were not detected ($\sigma_{\text{Measurelemma}}<0.001$), but instead the lemma class second-level predictor was successfully estimated ($p=0.004$ in the PB test).
-The large confidence intervals around the estimates (between $15.6$ and $28.37$) indicate that maybe the sample size needs to be higher, or a slightly modified classification might be even more adequate.
-Looking at the effect plot in Figure~\ref{fig:eff:measureattraction}, however, it is evident that the more abstract physical measure nouns with a high degree of grammaticalisation favour the \NACa\ (\textit{Square}, \textit{Length}, \textit{Currency}, \textit{Capacity}, \textit{Weight}) whereas the more concrete and less grammaticalised classes of measure nouns (which usually have an alternative use outside of MNPs) favour the \PGCa\ (\textit{Layer}, \textit{Amount}, \textit{Vessel}, \textit{Portion}).
-Ideally, there should be more differentiation in the second group, but given the large confidence intervals, significantly larger samples would be needed for more precise estimation.
-As mentioned before, the second level model for measure nouns in the F model could not estimated with meaningful results.
-
-On the other hand, the heuristic semantic classification for kind nouns in \textit{Kindconsistency} is much less useful.
-The MN model did not even converge with this regressor, and for the F model an insufficient $p=0.07$ was achieved in the PB test.
-Also, \textit{Kindint} does not contribute anything useful in either model (MN model $p=0.432$; F model $p=0.326$).
-The only regressor that has a considerable influence is \textit{Kindedible}, \ie, whether the substance denoted by the kind noun is edible.
-Since these regressors were added heuristically, there is no interpretation in terms of substantive theory, however, and I am hesitant to count it as very meaningful.
-In sum, no consistent picture emerges for kind noun-specific semantic prototypicality effects, but for measure nouns there is a clear indication that -- as predicted -- classes of more strongly grammaticalised nouns occur proportionally more frequently in the \PGCa.
+In Section~\ref{sec:analyses}, it was also hypothesised that there may be effects specific to classes of measure lemmas, and that they might be related to degrees of grammaticalisation of the measure noun.
+Transcending the effects of individual lemmas (captured in the random intercepts), the \textit{Measureclass} second-level predictor was successfully estimated ($p_{\text{PB}}=0.001$).
+Looking at the effect plot in Figure~\ref{fig:eff:measureattraction}, it is evident that abstract non-referential physical measure nouns (such as \textit{Gramm} `gram' or \textit{Liter} `litre') with a high degree of grammaticalisation favour the \NACa.
+At the other end of the scale, nouns denoting natural portions like \textit{Haufen} `heap', \textit{Bündel} `bundle', \textit{Schluck} `gulp' favour the \PGCa.
+These are referential nouns, confirming the hypothesis that it is prototypical of the PGC to host two referential nouns and the NAC only one.
 
@@
 I now turn to the predicted effect of plural measure nouns and cardinals as quantifiers.
-Plurality has a clear effect in the expected direction (plural favours the \PGCa) in the MN model ($p=0.002$ for \textit{Measurenumber}).
-In the F model, the effect has the same direction, but the estimate is quite imprecise ($p=0.160$), again likely due to the low sample size.
-Figure~\ref{fig:eff:leftcontext} shows, however, that the left context of the measure noun influences the choice of the variant (in both models, the regressor has $p=0.002$).
+Plurality has a clear effect in the expected direction (plural favours the \PGCa) ($\beta_{\text{MeasurenumberPl}}=0.814$, $p=0.001$ for \textit{Measurenumber}).
+Figure~\ref{fig:eff:leftcontext} shows that the left context of the measure noun influences the choice of the variant ($p=0.001$).
 Cardinals have a strong tendency to co-occur with the \NACa\ in both models.
@@
 The register-related proxy variables point into the expected direction.
-Increased \textit{Badness} favours the \NACa\ (only in the MN model $\beta=-0.2$, $p=0.002$), and so does a lesser density of genitives (MN model $\beta=-0.66$, $p=0.004$; F model $\beta=-0.52$, $p=0.002$).
-A higher length of the NP (\textit{Matchlength}) favours the \PGCa\ (MN model $\beta=0.54$, $p=0.002$; F model $\beta=0.63$, $p=0.002$) as predicted, but abbreviated measure nouns (\textit{Measureabbreviated}) do not seem to have much of an effect (MN model $p=0.730$; F model $p=0.267$).
+Increased \textit{Badness} of the document favours the \NACa\ ($\beta_{\text{Badness}}=-0.165$, $p=0.001$), and so does a lesser density of genitives ($\beta=-0.630$, $p=0.001$).
+While these are poor proxies to register (and partially circular in the case of \textit{Genitives}), this result can at least encourage future work into register effects
 
@@
 
-Finally, the nuisance variables \textit{Measurefreq} and \textit{Kindfreq} have no clear effect, except for maybe a $p=0.041$ for \textit{Measurefreq} in the MN model.
-The coefficient is $-0.16$ meaning that with each increase of the lemma log-frequency (per one million tokens) of the measure noun of one standard deviation, the probability that the \PGCa\ is chosen is lowered by approximately $-0.16/4=-0.04$.
-Since token frequency is no direct proxy for grammaticalisation, I hesitate to interpret this effect.
-The effect for \textit{Measurecase} is as predicted by previous studies (see Section~\ref{sec:analyses}).
-The dative favours the \PGCa\ in the MN model with $p=0.002$ and a coefficient for the dative (vs.\ the nominative, which is on the intercept) of $0.63$.
-In the F model, \textit{Dative} was not a level of \textit{Measurecase} due to the syncretisms mentioned in Section~\ref{sec:descriptive}.
+The effect of \textit{Measurecase} ($p=0.001$) is as predicted by previous studies (see Section~\ref{sec:analyses}).
+A measure noun in the dative favours the \PGCa\ with $\beta_{\text{MeasurecaseDat}}=0.707$ (compared to the nominative, which is on the intercept).
 Although \textit{Measurecase} is a nuisance variable in the context of this study, convergence with previous work strengthens its validity.
@@
 
-As was pointed out in Section \ref{sec:cocl}, there can be substantial doubt that corpus linguistic methods in themselves can be called \textit{cognitively oriented} in any meaningful way.
-At the same time, it is a highly plausible assumption in a usage-based framework that large and varied corpora ideally represent language users' relevant input of (at least) written language, and that there should be reflexes of the distributions found in such corpora in the linguistic behaviour of language users.
+As was pointed out in Section \ref{sec:cogocl}, there is a strong interest in validating corpus-based findings through experiments in order to substantiate \textit{cognitively oriented usage-based corpus linguistics} as a research programme.
 Therefore, this section and the next present the results of two experiments wherein I cross-checked the corpus-based findings.
+Both experiments use attested sentences from the corpus sample (see Section~\ref{sec:gettingdata}) and the probabilities that the corpus-based statistical model (see Section~\ref{sec:corpushierarchicalmodel} assigns to the competing constructions in the sentences.
+First, a forced-choice experiment was conducted.
+Participants had to chose between two sentences differing only <++>
 
-First, a forced-choice experiment was conducted.
 There were 24 participants (L1 speakers of German without reading or writing disability) aged 19 to 30 who were recruited from introductory linguistics courses at Freie Universität Berlin.
@@
 
-As stimuli, I used attested (but simplified) sentences from the corpus study in an approach similar to the one taken in \cite{DivjakEa2016}.
-The rationale was that it is virtually impossible to test the large and complex sets of regressors used in the corpus study in a traditional randomized experiments where usually just one relevant factor (ideally something like treatment vs.\ non-treatment) is of interest.
-Therefore, attested sentences were used, and the predictions of the corpus models were tested against participants' preferences.
-
-Concretely, I sampled 16 attested sentences from the corpus and simplified them (mostly removing embedded clauses, complex adverbials, etc.) and standardized their orthography.
+As stimuli, attested (but simplified) sentences from the corpus study were used in an approach similar to the one taken in \cite{DivjakEa2016}.
+I sampled 16 attested sentences from the corpus and simplified them (mostly removing embedded clauses, complex adverbials, etc.) and standardized their orthography.
 It was made sure that the simplifications and normalisations did not affect any of the regressors used in the corpus study.

### 2017-03-29 R/3_glmm.R

@@ -96,7 +96,7 @@ dotchart(measure.fixeffs, pch=20,
          xlim = c(x.lower, x.upper),
          lcolor = "gray",
          cex = 1.2,
-         main=paste("Fixed effects with bootstrapped\n 95% CIs (", ci.boot.nsim, " simulations)", sep=""))
+         main=paste("Coefficient estimates\n with bootstrapped 95% CI", sep=""))
 lines(c(0,0), c(0,length(measure.ci.95)), col="gray")
 
 for (i in 1:nrow(measure.ci.95)) {

### 2017-03-29 Latex/article.tex
FC experiment section finished

@@
   \affil[1]{Freie Universität Berlin}
-  \title{Competing Constructions for German Measure Phrases}
-  \runningtitle{Competing Constructions for German Measure Phrases}
+  \title{A Case Alternation in German Measure Noun Phrases: from Usage Data to Experimental Validation}
+  \runningtitle{A Case Alternation in German Measure Noun Phrases}
   %\subtitle{...}
-  \abstract{In German, nominal morpho-syntax\ldots}
-  \keywords{alternations, hierarchical models, corpus methods and experimental methods, measure constructions, German}
+  \abstract{}
+  \keywords{corpus methods and experimental methods, alternations, hierarchical models, measure constructions, German}
   %\classification[PACS]{...}
@@
 By \textit{alternation} I refer to a situation where two or more forms or constructions are available with no clear difference in acceptability, function, or meaning.
-I use it as an example case for a protocol for the estimation and validation of corpus-based models in cognitively oriented corpus linguistics.
-I briefly motivate this approach in this introduction, for details see Section~\ref{sec:cocl}.
 
@@
     \begin{tabular}{llrlrrrrrrcc}
-      Level & Regressor & \multicolumn{1}{l}{\pPB} & Level  & \multicolumn{2}{l}{Coefficient} & \multicolumn{2}{l}{CI Low} & \multicolumn{2}{l}{CI High} & \multicolumn{2}{l}{CI contains 0} \\
+      Level & Regressor & \multicolumn{1}{l}{\pPB} & Level  & \multicolumn{2}{l}{Coefficient} & \multicolumn{2}{l}{CI low} & \multicolumn{2}{l}{CI high} & \multicolumn{2}{l}{CI contains 0} \\
               &                   &       &          & \multicolumn{1}{l}{MLE} & \multicolumn{1}{l}{MCMC} & \multicolumn{1}{l}{MLE} & \multicolumn{1}{l}{MCMC} & \multicolumn{1}{l}{MLE} & \multicolumn{1}{l}{MCMC} & \multicolumn{1}{l}{MLE} & \multicolumn{1}{l}{MCMC} \\\midrule
@@
 For a robust quantification of the precision of the estimation, I ran a parametric bootstrap (using the \textit{confint.merMod} function from \textit{lme4}) with 1,000 replications, and using the percentile method for the calculation of the intervals.
-The resulting 95\% bootstrap confidence intervals are reported in Table~\ref{tab:bigtable} in the columns labelled \textit{CI Low} and \textit{CI High} (= upper and lower 2.5th percentiles).
+The resulting 95\% bootstrap confidence intervals are reported in Table~\ref{tab:bigtable} in the columns labelled \textit{CI low} and \textit{CI high} (= upper and lower 2.5th percentiles).
 The column \textit{CI contains 0} shows an asterisk for those intervals that do \textit{not} include 0.
@@
 both methods essentially estimate the same sign and magnitude for the coefficient (MLE $0.161$, MCMC $0.173$) and do not differ substantially with regard to the precision of the estimate in the form of the width of the confidence interval (both intervals have a width of exactly $0.341$).
-Hence, I only interpret the MLE model in the next section.
+Since both estimators converge, I only interpret the MLE model in the next section.
 
@@
 Therefore, this section and the next present the results of two experiments wherein I cross-checked the corpus-based findings.
-Both experiments use attested sentences from the corpus sample (see Section~\ref{sec:gettingdata}) and the probabilities that the corpus-based statistical model (see Section~\ref{sec:corpushierarchicalmodel} assigns to the competing constructions in the sentences.
+Both experiments use attested sentences from the corpus sample (see Section~\ref{sec:gettingdata}) and the probabilities that the corpus-based statistical model (see Section~\ref{sec:corpushierarchicalmodel}) assigns to the competing constructions in the sentences.
 First, a forced-choice experiment was conducted.
-Participants had to chose between two sentences differing only <++>
+Participants had to chose between two sentences differing only in that one contained the \NACa\ and the other contained the \PGCa.
 
-There were 24 participants (L1 speakers of German without reading or writing disability) aged 19 to 30 who were recruited from introductory linguistics courses at Freie Universität Berlin.
+There were 24 participants (L1 speakers of German without reading or writing disability) aged 19 to 30 who were recruited from introductory linguistics courses at [name of university anonymised].
 Although the experiment was conducted in the last four weeks of their first semester, participants had no deeper explicit knowledge of linguistics, grammar, or experimental methods.
-None had never participated in a forced-choice experiment before.
+None of them had ever participated in a forced-choice experiment before.
 Participation was voluntary, but participants received credit in partial fulfillment of course requirements.
@@
 Eight sentences contained masculine or neuter kind nouns, the other eight contained feminine kind nouns.
-Furthermore, in each of the M\slash N and F groups, four sentences with a case drop (\ie\ genitive or oblique case on the kind noun) and four withe case identity were chosen.
-More precisely, the four sentences were sampled as highly prototypical examples of case drop (high logits) and case identity (low logits), respectively, according to the corpus based models.%
-\footnote{Sentences were sorted by their corresponding $\hat{y}$ calculated from the model.
-For the M\slash N sentences, sentences were sampled from the top and bottom 15\% of the resulting vector.
-For the F sentences, the top and bottom 25\% were used because the overall sample size was smaller.}
-It was made sure that the highly predictive features according to Section \ref{sec:interpretation} occurred in unique combinations among the four sentences in each cell.
+Furthermore, in each of the masculine\slash neuter and feminine groups, four sentences originally containing the \NACa\ and four sentences originally containing the \PGCa\ were chosen.
+More precisely, the sentences were sampled as \textit{highly prototypical examples} of \PGCa\ (high probability assigned by GLMM) and \NACa\ (low probability assigned by GLMM), respectively.%
+\footnote{Remember from Section~\ref{sec:corpusstudies} that the model predicts the probability that the \PGCa\ is chosen over the \NACa.}
+High and low probability were defined as the top and bottom 20\% of all probabilities assigned by the GLMM.
+Lemmas and feature combinations were made unique within each group whenever possible.
 The design is summarised in Table \ref{tab:experiment1:design}.
@@
   \begin{tabular}[h]{rcc}
-     & M\slash N & F \\
+     & Masculine\slash Neuter & Feminine \\
      \midrule
-     \textbf{high prob.\ for case drop} & 4 sentences\Supsf{a} & 4 sentences\Supsf{a} \\
-     \textbf{low prob.\ for case drop} & 4 sentences\Supsf{a} & 4 sentences\Supsf{a} \\
+     \textbf{high prob.\ for PGC\Subsf{adj}} & 4 sentences\Supsf{a} & 4 sentences\Supsf{a} \\
+     \textbf{low prob.\ for PGC\Subsf{adj}} & 4 sentences\Supsf{a} & 4 sentences\Supsf{a} \\
   \end{tabular}
-  \caption{The four groups of attested sentences chosen as stimuli (\Supsf{a}Among the 4 sentences, combinations of important factor values were made unique.)}
+  \caption{The four groups of attested sentences chosen as stimuli (\Supsf{a}Among the 4 sentences, combinations of important factor values were made unique whenever possible.)}
   \label{tab:experiment1:design}
 \end{table}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.75\textwidth]{../R/output/fc_proportions}
+\caption{Participants' responses in the forced-choice task plotted against the predictions from the corpus-based model}
+\label{fig:afc:proportions}
+\end{figure}
 
@@
 The experiment was conducted using \textit{PsychoPy} \citep{Peirce2007}.
+Figure \ref{fig:afc:proportions} shows the relationship between the probabilities calculated from the corpus-based model and the proportional choices made by participants.
+The close correlation between the probabilities derived from usage data and the choices in the experiment are evident, and the relation appears to be linear.
 
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.75\textwidth]{../R/output/fc_proportions}
-\caption{Participants' responses in the forced-choice task plotted against the predictions from the corpus-based model}
-\label{fig:afc:proportions}
-\end{figure}
-
-Figure \ref{fig:afc:proportions} shows the relationship between the logits calculated from the corpus-based model and the proportional choices made by participants.
-Visually, it looks like there is a close correlation between the corpus-derived preferences and preferences in the forced choice experiment.
-
-Then, a multilevel logistic regression was specified.
-Coefficients were estimated using Maximum Likelihood, and the coefficient table is given in Table \ref{tab:afc:model}, where instead of approximate standard errors and $z$ and $p$ values the lower and the upper bound of a bootstrapped 95\% confidence interval are specified.
-
-\begin{table}
-  \centering
-  \begin{tabular}{lrrr}
-    Effect & Estimate & 95\% CI lower & 95\% CI upper \\
-    \midrule
-    (Intercept)            &  0.748  &  0.242  &  1.265 \\
-    Modelpred              &  1.154  &  0.905  &  1.497 \\
-    KindgenderMN           &  0.135  & -0.416  &  0.632 \\
-    Modelpred:KindgenderMN & -0.567  & -0.843  & -0.298 \\
-  \end{tabular}
-  \caption{Coefficient table for the GLMM specified in equation \ref{eq:model:afc}, analysing the results of the forced choice experiment}
-  \label{tab:afc:model}
-\end{table}
-
-The number of observations was $n=768$.
-The random effects for participant ($\sigma=0.7841$, 24 groups) and item ($\sigma=0.4869$, 16 groups) show considerable variance.
-Accordingly, there is a notable difference between the marginal and conditional Nakagawa \& Schielzeth's $R^2_{m}=0.383$ and $R^2_{c}=0.510$, while both values are quite high.
-Dropping the term of interest ($\beta^1\cdot x_i^1+\beta^2\cdot x_i^2+\beta^{1.2}\cdot x_i^1\cdot x_i^2$) leads to a significantly worse fit of the nested model both in a traditional Likelihood Ratio Test ($LR=55.172$, $p<0.0001$) and its bootstrapped PB alternative ($p = 0.002$, 493 of 500 requested samples used).
-The effect display for the term is given in Figure \ref{fig:afc:effects}.
-
-\begin{figure}[h]
+\begin{figure}[hb!]
 \centering
 \includegraphics[width=0.75\textwidth]{../R/output/fc_effects}
-\caption{Effect plot for the multilevel logistic regression in the forced-choice experiment: predictability of participants' choices from the logits of the corpus-based GLMM (cf.\ Section~\ref{sec:corpushierarchicalmodel})}
+\caption{Effect plot for the multilevel logistic regression in the forced-choice experiment: predictability of participants' choices using the probabilities derived from the corpus-based GLMM}
 \label{fig:afc:effects}
@@
 
-In summary, the forced choice experiment clearly succeeded in validating the results from the corpus study in as much as the preferences extracted from the corpus correspond to native speakers' choices.
-We can take this as an indication that there is some relation between the process generating the corpus data and the behaviour of individual speakers.
-In the next section, I report the results of a self-paced reading experiment which yielded (maybe expectedly) less clear results.
+Then, a multilevel logistic regression was specified with the probability predicted by the corpus-based GLMM (\textit{Modelprediction}) as the only fixed effect.
+A random effect for the sentence ($16$ levels) was added to catch idiosyncrasies of single sentences (or badly selected sentences).
+Also, a random effect for the participant ($24$ levels) was added.
+Coefficients were estimated with Maximum Likelihood Estimation (\textit{lmer} function from \textit{lme4}).
+The number of observations was $n=768$.
+
+The regressor \textit{Modelprediction} achieves $p_{\text{PB}}=0.001$ ($1,000$ replications) and is estimated at $3.953$ relative to an intercept of $-1.343$.
+The confidence interval from a parametric bootstrap ($1,000$ replications) for the regressor is acceptable with a lower bound of $2.836$ and an upper bound of $5.023$.
+The random effects for participant ($\sigma=0.768$) and item ($\sigma=0.700$) show considerable variance.
+Accordingly, there is a notable difference between the marginal and conditional Nakagawa \& Schielzeth's $R^2_{m}=0.216$ and $R^2_{c}=0.409$.
+The effect display for the term is given in Figure \ref{fig:afc:effects}.
+It confirms the effect intuitively noticeable from Figure~\ref{fig:afc:proportions}:
+the higher the probability of the \PGCa\ predicted from usage-data, the more often participants chose the \PGCa\ variant in the forced-choice task.
+In summary, the forced choice experiment clearly succeeded in validating the results from the corpus study in as much as the preferences extracted from usage data correspond to native speakers' choices.
 

### 2017-03-29 Latex/article.tex
cleaning stats

@@
 
+% TODO SURPRISAL
+%
+% Bybee, Joan und Thompson, Sandra. 1997. “Three Frequency Effects in Syntax”, Proceedings of the Twenty-Third Annual Meeting of the Berkeley Linguistics Society, 378-388.
+% Bybee, Joan. 2006. “From Usage to Grammar: The Mind‘s Response to Repetition”. Language 82(4): 711-733.
+% Levy, Roger .2008. “Expectation-based syntactic comprehension.” Cognition 106: 1126–1177.
+% Smith, Nathaniel, and Levy, Roger. 2013. “The effect of word predictability on reading time is logarithmic.” Cognition 128: 302–319.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 }
-%I use * to denote general unacceptability, and the acceptability judgements given in (\ref{ex:intro:pseudopartitive}) are completely undisputed.
 Here, the kind noun is in the genitive, and I refer to the construction in (\ref{ex:intro:pseudopartitive1}) as the \textit{Pseudo-partitive Genitive Construction} (PGC).
@@
 \footnote{Notice that the unavailability of the genitive on the kind noun can be seen as following from a rather quirky constraint that genitive NPs in German require the presence of some strongly case-marked element (determiner or adjective) in addition to the head noun in order to be acceptable \citep{GallmannLindauer1994,Schachtl1989}.}
-%The construction as in (\ref{ex:intro:narrowapposition2}) is also referred to as the \textit{Direct Partitive Construction} for other Germanic languages in which the PGC with the synthetic genitive is not available.%
-%\footnote{This nomenclature makes sense in contrast to the \textit{Indirect Partitive Construction} with prepositional linkers translating to `of' -- \ie\ analytic genitives -- in such languages, see \cite{HankamerMikkelsen2008} for Danish.
-%For German, this terminology is not distinctive enough, which is why I use the terms NAC and PGC.}
+The construction as in (\ref{ex:intro:narrowapposition2}) is also referred to as the \textit{Direct Partitive Construction} for other Germanic languages in which the PGC with the synthetic genitive is not available.%
+\footnote{This nomenclature makes sense in contrast to the \textit{Indirect Partitive Construction} with prepositional linkers translating to `of' -- \ie\ analytic genitives -- in such languages, see \cite{HankamerMikkelsen2008} for Danish.
+For German, this terminology is not distinctive enough, which is why I use the terms NAC and PGC.}
 While only an accusative is shown in (\ref{ex:intro:narrowapposition}), the kind noun obligatory agrees in case also with nominative and dative measure nouns.
@@
   \centering
-  \includegraphics[width=0.85\textwidth]{../R/output/corpus_fixeffs}
-  \caption{Coefficients (MLE) with 95\% parametric bootstrap confidence intervals; intercept (\textit{Leftcontext=C}, \textit{Measurecase=Nom}, \textit{Measurenumber=Sg}, \textit{Kindgender=Masc}, \textit{Measureclass=Physical}; 0 for all numeric z-transformed regressors) is -4.328}
+  \includegraphics[width=0.85\textwidth]{../R/output/corpus_fixeffs_mle+mcmc}
+  \caption{Coefficients (MLE and MCMC) with 95\% confidence intervals (for details see text); intercept (\textit{Leftcontext=C}, \textit{Measurecase=Nom}, \textit{Measurenumber=Sg}, \textit{Kindgender=Masc}, \textit{Measureclass=Physical}; 0 for all numeric z-transformed regressors) is -4.328 (MLE) and -4.441 (MCMC)}
   \label{fig:fixeffs}
@@
 \footnote{Effect plots were created using the \textit{effects} package \citep{Fox2003}.
-They show the changes in probability for the outcome (y axis) dependent on values of a regressor (x axis), assuming typical values of all other regressors.}
+They show the changes in probability for the outcome (y axis) dependent on values of a regressor (x axis), at typical values of all other regressors.}
 The effect is mostly as expected:
@@
 The effect for measure nouns is stronger, and it was estimated with higher precision.
+
+An interesting picture emerges for the lemma frequencies.
+A higher than average lemma frequency of measure nouns favours the \NACa\ as expected ($\beta_{\text{Measurefreq}}=-0.257$, $p_{\text{PB}}=0.003$).
+With kind nouns, higher frequency seems to favour the \PGCa ($\beta_{\text{Kindfreq}}=0.161$, $p_{\text{PB}}=0.066$).
 
@@
 \subsection{Experiment 1: forced choice}
-
-% TODO SURPRISAL
-%
-% Bybee, Joan und Thompson, Sandra. 1997. “Three Frequency Effects in Syntax”, Proceedings of the Twenty-Third Annual Meeting of the Berkeley Linguistics Society, 378-388.
-% Bybee, Joan. 2006. “From Usage to Grammar: The Mind‘s Response to Repetition”. Language 82(4): 711-733.
-% Levy, Roger .2008. “Expectation-based syntactic comprehension.” Cognition 106: 1126–1177.
-% Smith, Nathaniel, and Levy, Roger. 2013. “The effect of word predictability on reading time is logarithmic.” Cognition 128: 302–319.
 
@@
 
-%\begin{figure}[h]
-%\centering
-%\includegraphics[width=0.6\textwidth]{figures/experiment/spr_effects_modelPredDichotomousReg}
-%\caption{Effect plot for the LMM in the self-paced reading experiment: modeling participants' residualised log reading times on the dichotomised predictions of the corpus-based GLMM (cf.\ Section~\ref{sec:corpushierarchicalmodel})}
-%\label{fig:spr:dichotomous:effects}
-%\end{figure}
-
 \begin{figure}[h]

### 2017-03-29 R/0_wrapper.R

@@ -54,4 +54,4 @@ cat("\n\n Prediction for experimental stimuli ...\n")
 source('5_stimuli_predict.R')
 
 # Save workspace.
-save(list = ls(), file=paste(out.dir, "workspace_", Sys.time(), ".RData", sep=""))
+save(list = ls(), file=paste(out.dir, "workspace.RData", sep=""))

### 2017-03-29 R/4_glmm-mcmc.R

@@ -66,3 +66,27 @@ if (save.persistent) sink(paste(out.dir, "mcmc.txt", sep=""), append = T)
 cat("\n\nTable comparing coefficient estimates with ML and MCMC\n\n")
 print(measure.glmm.table)
 if (save.persistent) sink()
+
+
+# Print new fixeff plot with MCMC.
+if (save.persistent) pdf(paste(out.dir, "fixeffs_mle+mcmc.pdf", sep=""))
+dotchart(rep(-100, length(measure.fixeffs)), pch=20,
+         xlim = c(x.lower, x.upper),
+         labels = names(measure.fixeffs),
+         lcolor = "gray",
+         cex = 1.2,
+         main=paste("MLE and MCMC oefficient estimates\n with 95% confidence intervals", sep=""))
+lines(c(0,0), c(0,length(measure.ci.95)), col="gray")
+
+for (i in 1:nrow(measure.ci.95)) {
+  points(measure.fixeffs[i], i+0.25, pch=25, cex=0.75, bg="black", col = "black")
+  lines(measure.ci.95[i,c(1,2)], c(i+0.25, i+0.25), col="black", lwd=2)
+  
+  # MCMC.
+  points(measure.glmm.table[names(measure.fixeffs[i]), "MC_Coef"], i-0.25, pch=24, cex=0.75, bg="gray", col = "gray")
+  lines(measure.glmm.table[names(measure.fixeffs[i]), c("MC_Low", "MC_High")], c(i-0.25,i-0.25), col="gray", lwd=2)
+}
+
+legend("topright", legend = c("MLE", "MCMC"), cex = 1.2, col = c("black", "gray"), pt.bg = c("black", "gray"),
+       lty = 1, lwd = 2, pch = c(25, 24), pt.cex = 0.75, bg = "white")
+if (save.persistent) dev.off()

### 2017-03-29 R/6_exp-fc.R

@@ -130,5 +130,5 @@ if (save.persistent) dev.off()
 
 
 # Save workspace.
-save(list = ls(), file=paste(out.dir, "workspace_", Sys.time(), ".RData", sep=""))
+save(list = ls(), file=paste(out.dir, "workspace.RData", sep=""))
 

### 2017-03-29 R/7_exp-spr.R

@@ -292,4 +292,4 @@ if (save.persistent) dev.off()
 
 
 # Save workspace.
-save(list = ls(), file=paste(out.dir, "workspace_", Sys.time(), ".RData", sep=""))
+save(list = ls(), file=paste(out.dir, "workspace.RData", sep=""))

### 2017-03-29 Latex/article.tex
Feierabend für heute.

@@
 \subsection{Experiment 1: forced choice}
+\label{sec:exp:fc}
 
@@
 Then, a multilevel logistic regression was specified with the probability predicted by the corpus-based GLMM (\textit{Modelprediction}) as the only fixed effect.
-A random effect for the sentence ($16$ levels) was added to catch idiosyncrasies of single sentences (or badly selected sentences).
-Also, a random effect for the participant ($24$ levels) was added.
+A random effect for the sentence was added to catch idiosyncrasies of single sentences (or badly selected sentences).
+Also, a random effect for the participant was added.
 Coefficients were estimated with Maximum Likelihood Estimation (\textit{lmer} function from \textit{lme4}).
-The number of observations was $n=768$.
+The number of observations was $n=384$.
 
-The regressor \textit{Modelprediction} achieves $p_{\text{PB}}=0.001$ ($1,000$ replications) and is estimated at $3.953$ relative to an intercept of $-1.343$.
-The confidence interval from a parametric bootstrap ($1,000$ replications) for the regressor is acceptable with a lower bound of $2.836$ and an upper bound of $5.023$.
-The random effects for participant ($\sigma=0.768$) and item ($\sigma=0.700$) show considerable variance.
-Accordingly, there is a notable difference between the marginal and conditional Nakagawa \& Schielzeth's $R^2_{m}=0.216$ and $R^2_{c}=0.409$.
-The effect display for the term is given in Figure \ref{fig:afc:effects}.
+A certain amount of the variance can be accounted for by idiosyncrasies of single sentences ($\sigma_{\text{Sentence}}=0.599$, $16$ levels).
+Also, among participants, there are clearly different preferences ($\sigma_{\text{Participant}}=0.407$, $24$ levels).
+On the extreme sides, one participant chose the \PGCa\ in $13$ of $16$ cases, and two participants only chose it in $5$ of $16$ cases.
+However, the regressor \textit{Modelprediction} still achieves $p_{\text{PB}}=0.001$ ($1,000$ replications) and is estimated at $4.555$ relative to an intercept of $-1.511$.
+The confidence interval from a parametric bootstrap ($1,000$ replications) for the regressor is acceptable with a lower bound of $2.746$ and an upper bound of $6.330$.
+Nakagawa \& Schielzeth's pseudo-coefficients of determination are $R^2_{m}=0.271$ and $R^2_{c}=0.442$, which means that roughly 27\% of the variance in the data can be explained by considering the predictions from the corpus-based GLMM.
+The effect display for the single fixed regressor \textit{Modelprediction} is given in Figure \ref{fig:afc:effects}.
 It confirms the effect intuitively noticeable from Figure~\ref{fig:afc:proportions}:
@@
 \subsection{Experiment 2: self-paced reading}
+\label{sec:exp:spr}
+
+Finally, I report the results of a self-paced reading experiment also designed to test whether the statistical model derived from usage data (see Section~\ref{sec:corpusnonhierarchicalmodel}) can be used to predict speakers' behaviour.
+As opposed to the forced choice experiment reported in the last section, self-paced reading is completely implicit (see Section~\ref{sec:cogocl} for more on the motivation).
+
+I used exactly the same stimuli as in the forced choice experiments.
+Each participant read the sentence with the variant preferred by the corpus model and the modified sentence with the dispreferred variant.
+To avoid and\slash or randomise repetition effects, the stimuli for each participant were separated into two blocks of $16$ targets and $33$ fillers.
+It was made sure that from each sentence pair, one sentence was assigned to the first block, and the other sentence to the other block.
+This assignment was randomised for each participant, and the order within the block was also randomised while keeping the sentences from each pair as far apart as possible.
+In total, each participant read $98$ sentences.
+There were $38$ (different) participants recruited in exactly the same manner as for the experiment reported in Section~\ref{sec:exp:fc}.
 

### 2017-03-29 R/6_exp-fc.R

@@ -49,7 +49,7 @@ responses.df <- data.frame(
     Modelpred     = rep(inv.logit(stims.pred), ncol(results))
   , Kindgender    = factor(rep(c(rep("MN", 8), rep("F", 8)), ncol(results)))
   , ResponseDrop  = factor(as.vector(apply(results, 2, reconstruct.rating)))
-  , Item          = as.factor(rep(c(rep(c(1:8), 2), rep(c(9:16), 2)), ncol(results)))
+  , Item          = as.factor(rep(1:16, ncol(results)))
   , Participant   = as.factor(unlist(lapply(1:ncol(results), function(x) rep(x, 16))))
 )
 

### 2017-03-31 Latex/article.tex
experiment sections done

--- 2017-03-29 Latex/article.tex
@@
   \affil[1]{Freie Universität Berlin}
-  \title{A Case Alternation in German Measure Noun Phrases: from Usage Data to Experimental Validation}
-  \runningtitle{A Case Alternation in German Measure Noun Phrases}
+  \title{Competing Constructions for German Measure Noun Phrases: from Usage Data to Experimental Validation}
+  \runningtitle{Competing Constructions for German Measure Noun Phrases}
   %\subtitle{...}
@@
 %
-%In Section~\ref{sec:corpusnonhierarchicalmodel}, I briefly discuss why a non-hierarchical model -- while it is an option in principle -- has serious disadvantages for the data set under discussion (and similar data sets).
 %In Section~\ref{sec:corpushierarchicalmodel}, I present the hierarchical model which is actually interpreted as the primary result of this study.
@@
 %Finally, I summarise and interpret the results of the corpus study in Section~\ref{sec:interpretation}.
+
+% VanderPlas: Bayesian
 
@@
 Therefore, this section and the next present the results of two experiments wherein I cross-checked the corpus-based findings.
-Both experiments use attested sentences from the corpus sample (see Section~\ref{sec:gettingdata}) and the probabilities that the corpus-based statistical model (see Section~\ref{sec:corpushierarchicalmodel}) assigns to the competing constructions in the sentences.
+Both experiments use sentences containing attested MNPs from the corpus sample embedded into radiacally simplified sentences, and the probabilities that the corpus-based statistical model (see Section~\ref{sec:corpushierarchicalmodel}) assigns to the competing constructions in these sentences.
 First, a forced-choice experiment was conducted.
@@
 
-There were 24 participants (L1 speakers of German without reading or writing disability) aged 19 to 30 who were recruited from introductory linguistics courses at [name of university anonymised].
+There were 24 participants (native speakers of German without reading or writing disabilities) aged 19 to 30 who were recruited from introductory linguistics courses at [name of university anonymised].
 Although the experiment was conducted in the last four weeks of their first semester, participants had no deeper explicit knowledge of linguistics, grammar, or experimental methods.
@@
 
-As stimuli, attested (but simplified) sentences from the corpus study were used in an approach similar to the one taken in \cite{DivjakEa2016}.
-I sampled 16 attested sentences from the corpus and simplified them (mostly removing embedded clauses, complex adverbials, etc.) and standardized their orthography.
-It was made sure that the simplifications and normalisations did not affect any of the regressors used in the corpus study.
+As stimuli, attested MNPs from the corpus study were used, but the sentences were radically simplified to avoid .
+I sampled 16 MNPs from the corpus, and it was made sure that the simplifications and normalisations did not affect any of the regressors used in the corpus study.
+For example, the case, number, etc. of the MNP remained the same, as well as the choice of lexical material within the MNP.
 Eight sentences contained masculine or neuter kind nouns, the other eight contained feminine kind nouns.
@@
   \end{tabular}
-  \caption{The four groups of attested sentences chosen as stimuli (\Supsf{a}Among the 4 sentences, combinations of important factor values were made unique whenever possible.)}
+  \caption{The four groups of sentences chosen as stimuli (\Supsf{a}Among the 4 sentences, combinations of important factor values were made unique whenever possible.)}
   \label{tab:experiment1:design}
@@
 
-The pairs of stimuli were the original attested sentence containing the preferred construction and a modified version containing the dispreferred construction.
+The pairs of stimuli were the sentence containing the preferred construction (according to the corpus GLMM) and a modified version containing the dispreferred construction.
 They were presented next to each other, and a 20 second time limit for each choice was set.%
@@
 
-Then, a multilevel logistic regression was specified with the probability predicted by the corpus-based GLMM (\textit{Modelprediction}) as the only fixed effect.
-A random effect for the sentence ($16$ levels) was added to catch idiosyncrasies of single sentences (or badly selected sentences).
-Also, a random effect for the participant ($24$ levels) was added.
+Then, a multilevel logistic regression was specified with the probability of the \PGCa\ predicted for each sentence by the corpus-based GLMM (\textit{Modelprediction}) as the only fixed effect.%
+\footnote{The document-level variables \textit{Badness} and \textit{Genitives} were set to $0$ (which is the mean for z-transformed variables).}
+A random intercept and slope were added for the individual sentence (item) in order to catch idiosyncrasies of single sentences.
+Also, a random intercept and slope for participants was added.%
+\footnote{The random slopes were added to comply with \citet[257]{BarrEa2013} who predict \textit{catastrophically high Type I error rates} for designs with within-subject manipulations if random effects structures are not kept maximal.
+Notice that the model reported here was estimated with conceptually identical results with regard to the predictor of interest (\textit{Modelprediction}) if only random intercepts were used.}
 Coefficients were estimated with Maximum Likelihood Estimation (\textit{lmer} function from \textit{lme4}).
-The number of observations was $n=768$.
+The number of observations was $n=384$.
 
-The regressor \textit{Modelprediction} achieves $p_{\text{PB}}=0.001$ ($1,000$ replications) and is estimated at $3.953$ relative to an intercept of $-1.343$.
-The confidence interval from a parametric bootstrap ($1,000$ replications) for the regressor is acceptable with a lower bound of $2.836$ and an upper bound of $5.023$.
-The random effects for participant ($\sigma=0.768$) and item ($\sigma=0.700$) show considerable variance.
-Accordingly, there is a notable difference between the marginal and conditional Nakagawa \& Schielzeth's $R^2_{m}=0.216$ and $R^2_{c}=0.409$.
-The effect display for the term is given in Figure \ref{fig:afc:effects}.
+A certain amount of the variance can be accounted for by idiosyncrasies of single sentences ($\sigma_{\text{Sentence}}=0.861$, $\sigma^{\text{Modelprediction}}_{\text{Sentence}}=0.250$, $16$ levels).%
+\footnote{I use $\sigma^f_r$ to denote the standard deviation of the of the random intercepts for the fixed effect $f$ varying by random effect $r$.}
+Also, among participants, there are clearly different preferences ($\sigma_{\text{Participant}}=0.530$, $\sigma^{\text{Modelprediction}}_{\text{Participant}}=0.076$, $24$ levels).
+On the extreme sides, one participant chose the \PGCa\ in $13$ of $16$ cases, and two participants only chose it in $5$ of $16$ cases.
+The regressor \textit{Modelprediction} achieves $p_{\text{PB}}=0.002$ ($1,000$ replications) and is estimated at $4.500$ relative to an intercept of $-1.520$.
+The confidence interval from a parametric bootstrap ($1,000$ replications) for the regressor is acceptable with a lower bound of $2.738$ and an upper bound of $6.390$.
+Nakagawa \& Schielzeth's pseudo-coefficients of determination are $R^2_{m}=0.267$ and $R^2_{c}=0.435$, which means that roughly 27\% of the variance in the data can be explained by considering the predictions from the corpus-based GLMM.
+The effect display for the single fixed regressor \textit{Modelprediction} is given in Figure \ref{fig:afc:effects}.
 It confirms the effect intuitively noticeable from Figure~\ref{fig:afc:proportions}:
@@
 \subsection{Experiment 2: self-paced reading}
+\label{sec:exp:spr}
+
+Finally, I report the results of a self-paced reading experiment also designed to test whether the statistical model derived from usage data (see Section~\ref{sec:corpushierarchicalmodel}) can be used to predict speakers' behaviour.
+It is expected that reading the \textit{un}expected or less prototypical variant (the one assigned a low probability by the corpus-derived model) in a given context and with given lexical material might incur a processing overhead \citep{Kaiser2013}.
+In a very similar fashion, \cite{DivjakEa2016} apply the self-paced reading paradigm in the validation of corpus-based models.
+
+I used exactly the same stimuli as in the forced choice experiments.
+Each participant read both the $16$ sentences with the variant predicted by the corpus model and the $16$ modified sentence with the variant that the corpus model would not predict.%
+\footnote{Notice that lemmas and their frequencies as well as lemma classes are part of the corpus-based GLMM, and there was no additional controlling of lemma frequencies, etc.}
+To avoid repetition effects, the stimuli for each participant were separated into two blocks of $16$ targets and $33$ fillers per block.
+It was made sure that from each target sentence pair, one sentence was assigned to the first block, and the other sentence to the other block.
+The assignment of members of the individual sentence pairs to one of the blocks was randomised for each participant individually, and so was the order within the block.
+The sentences from each pair of variants were kept as far apart as possible.
+The fillers also came in pairs such that the second block contained only sentences to which participants had been exposed in the first block in slightly modified form.
+In total, each participant read $98$ sentences.
+After each sentence, participants had to answer simple (non-metalinguistic) yes-no questions about the previous sentence as distractors.
+The distractor questions were different ones in the first and the second block.
+There were $38$ (different) participants recruited in exactly the same manner as for the experiment reported in Section~\ref{sec:exp:fc}.
+The experiment was conducted using \textit{PsychoPy}.
+
+The reading times were residualised per speaker based on the reading times of all words (not just the targets) by that speaker.
+The adjective and kind noun (\ie\ the constituents bearing the critical case markers) were used as the target region, such as the bracketed words in the example \textit{zwei Gläser} [\textit{sprudelndes Wasser}] `two glasses of sparkling water'.
+Outliers farther than 2 inter-quartile ranges from the mean logarithmised residual time were reamoved ($64$ data points), resulting in a total number of $n=1,152$ observations.
+An LMM was specified with the logarithmised residual reading times as the response variable and the probabilities derived from the corpus GLMM (\textit{Modelprediction}) as the main regressor.
+Since the corpus GLMM predicts the probability of the \PGCa, it is expected that reading times are positively correlated (longer reading times) with the probability if the stimulus actually contains the \NACa, and negatively correlated (shorter reading times) if the stimulus actually contains the \PGCa.
+To account for this, an interaction between \textit{Modelprediction} and \textit{Construction} (levels \textit{PGCa} and \textit{NACa}) was added to the model.
+Furthermore, the position of the sentence in the individual experiment run was included as a fixed effect to control for increasing reading speed during experiment runs.
+Random intercepts were specified for \textit{Participant} and \textit{Item} (the $16$ sentence pairs are one \textit{Item}).%
+\footnote{I tried random slopes, but it was impossible to get the algorithm to converge.}
+
+\begin{table}
+  \centering
+  \begin{tabular}{lrrrc}
+    Regressor & \multicolumn{1}{r}{Coefficient} & \multicolumn{1}{r}{CI low} & \multicolumn{1}{r}{CI high} & \multicolumn{1}{r}{0 in CI} \\ \midrule
+    ConstructionPGCa                 &  0.062 &  0.023 &  0.103 &  *  \\ 
+    Modelprediction                  &  0.025 & -0.065 &  0.122 &     \\ 
+    Position                         & -0.005 & -0.005 &  0.004 &     \\ 
+    ConstructionPGCa:Modelprediction & -0.132 & -0.229 & -0.041 &  *  \\ 
+  \end{tabular}
+  \caption{Fixed effect coefficient table for the LMM used to analyse the self-paced reading experiment}
+  \label{tab:exp:spr}
+\end{table}
+
+Table~\ref{tab:exp:spr} shows the coefficient estimates with a 95\% parametric bootstrap confidence interval ($1,000$ replications).
+The standard deviation of the participant intercepts is $\sigma_{\text{Participant}}=0.079$ and of the item intercepts $\sigma_{\text{Item}}=0.039$.
+Comparing the full model to a model without the main regressor \textit{Modelprediction} (and consequently also the interaction with \textit{Construction}) in a PB test gives $p_{\text{PB}}=0.018$.
+Nakagawa and Schielzeth's pseudo-determination coefficients are $R^2_m=0.238$ and $R^2_c=0.347$.
 
@@
 \includegraphics[width=0.75\textwidth]{../R/output/spr_effects}
-\caption{Effect plot for the LMM in the self-paced reading experiment: modeling participants' residualised log reading times on the predictions (logits) of the corpus-based GLMM (cf.\ Section~\ref{sec:corpushierarchicalmodel})}
-\label{fig:spr:continuous:effects}
+\caption{Effect plot for the LMM in the self-paced reading experiment: modeling participants' residualised log reading times on the probabilities given by the corpus-based GLMM}
+\label{fig:spr:effects}
 \end{figure}
+
+While the overall model quality is acceptable, the relevant effect is estimated with some uncertainty, which is also visible in the effect plot in Figure~\ref{fig:spr:effects}.
+While the estimate for the sentences with \NACa\ is clearly too imprecise, there is a clear tendency in the sentences with \PGCa, which is also confirmed by the `significant' results from the bootstrapped confidence intervals (see Table~\ref{tab:exp:spr}) and from the PB test reported above.
+The \PGCa\ brings about an increased reading time ($\beta_{\text{ConstructionPGCa}}=0.062$), which is plausible because it is the much rarer construction (see Section~\ref{sec:corpusstudies}).
+However, if it occurs in a prototypical situation, reading times drop ($\beta_{\text{ConstructionPGCa:Modelprediction}}=-0.132$), which is also reflected in the slope of curve in the right panel of Figure~\ref{fig:spr:effects}.
+This fits into the general picture in as much as the construction with the lower frequency might be developing towards a more sharply defined prototype.
+In this context, it should be remembered from Section~\ref{sec:gettingdata} that even the \PGCd\ is much rarer that the \NACb\ ($17,252$ vs.\ $315,635$ occurrences in the auxiliary corpus samples).
+Conversely, the \NACa\ (like the NAC in general) might be the highly frequent default which does not incur reading time penalty, even if it is not the optimal choice in the given context and with the given lexical material.
+This possibility will be considered, among other things, in the next section.
 
@@
 \section{Conclusions}
+\label{sec:conclusion}
 
@@
 \begin{acknowledgement}
-  I want to thank (in alphabetical order) Felix Bildhauer, Götz Keydana, and Ulrike Sayatz for stimulating discussions on relevant aspects of this paper.
+  I want to thank (in alphabetical order) Felix Bildhauer, Samuel Reichert, and Ulrike Sayatz for discussions on relevant aspects of this paper.
   I also thank Ulrike Sayatz for recruiting the participants for the experiments.
   I am grateful to Kim Maser for her work on the annotation of the concordances.
-  Furthermore, Samuel Reichert substantially helped to develop the guiding ideas behind this study by writing an excellent BA theses on the subject under my supervision.
   The research presented here was made possible in part through funding from the \textit{Deutsche Forschungsgemeinschaft} (DFG, personal grant SCHA1916/1-1).

### 2017-03-31 R/0_wrapper.R

@@ -13,7 +13,7 @@ setwd('/Users/user/Workingcopies/measurenps/GitHub/R')
 
 # Options.
 set.seed(2398651)                          # Do NOT change, or results will differ.
-save.persistent        <- T                # write to files instead of screen/console.
+save.persistent        <- F                # Write to files instead of screen/console.
 out.dir                <- "output/"        # Put generated files here.
 precision              <- 3                # Precision in result table.
 the.nAGQ               <- 0                # 0 for speed, 1 for precision and eytremely slow computation.
@@ -54,4 +54,4 @@ cat("\n\n Prediction for experimental stimuli ...\n")
 source('5_stimuli_predict.R')
 
 # Save workspace.
-save(list = ls(), file=paste(out.dir, "workspace.RData", sep=""))
+if (save.persistent) save(list = ls(), file=paste(out.dir, "workspace.RData", sep=""))

### 2017-03-31 R/6_exp-fc.R

@@ -11,8 +11,8 @@ set.seed(239865)
 setwd('/Users/user/Workingcopies/measurenps/GitHub/R')
 
 ci.method        <- "boot" # boot
-modcomp.nsim     <- 1000   # 1000
-boot.nsim        <- 1000   # 1000
+modcomp.nsim     <- 1000 # 1000
+boot.nsim        <- 1000 # 1000
 the.nAGQ         <- 0 
 out.dir          <- "output/fc_"
 save.persistent  <- T
@@ -65,7 +65,7 @@ if (save.persistent) sink(paste(out.dir, "results.txt", sep=""))
 
 # Actual model.
 cat("\n\nGLMM, predicting reactions from corpus model\n\n")
-model.fc <- glmer(Chosenconstruction~Modelprediction+(1|Item)+(1|Participant),
+model.fc <- glmer(Chosenconstruction~Modelprediction+(1+Modelprediction|Item)+(1+Modelprediction|Participant),
                   family = binomial(link = "logit"), data = responses.df, nAGQ=the.nAGQ,
                   control=glmerControl(optimizer="nloptwrap2", optCtrl=list(maxfun=2e5)))
 print(summary(model.fc))
@@ -89,7 +89,7 @@ print(ci.95.fc)
 
 if (save.persistent) sink()
 
-# Effect plot for main interaction.
+# Effect plot.
 p <- plot(effect("Modelprediction", model.fc, KR = T), rug=F, colors = c("black", "darkblue"),
           main="Forced choice experiment",
           ylab="Probability that PGCa is chosen",
@@ -130,5 +130,5 @@ if (save.persistent) dev.off()
 
 
 # Save workspace.
-save(list = ls(), file=paste(out.dir, "workspace.RData", sep=""))
+if (save.persistent) save(list = ls(), file=paste(out.dir, "workspace.RData", sep=""))
 

### 2017-03-31 R/7_exp-spr.R

@@ -11,11 +11,11 @@ set.seed(239865)
 
 setwd('/Users/user/Workingcopies/measurenps/GitHub/R')
 
-
-ci.method       <- "boot" # boot
-modcomp.nsim    <- 1000   # 500
-boot.nsim       <- 1000   # 1000
-iqr.factor      <- 3      # For removing outliers in final data set.
+target          <- "tf.rt.resid" # Response variable.
+ci.method       <- "boot"        # boot
+modcomp.nsim    <- 1000          # 1000
+boot.nsim       <- 1000          # 1000
+iqr.factor      <- 2             # For removing outliers in final data set.
 out.dir         <- "output/spr_"
 save.persistent <- T
 data.dir        <- "data/spr/"
@@ -56,7 +56,7 @@ for (dafi in data.files) {
   
   # Get the stims for this part in part order and make right format (split words).
   part.stims.order   <- unname(unlist(assigs[which(assigs$Participant == part), 2:99]))
-  part.stims         <- stims[part.stims.order, c("Sentence", "Target", "Kindgender", "Construction", "Modelpred", "Modelprediction", "Expectedcorr", "Adjtarget", "Wordcount", "Number", "Expno")]
+  part.stims         <- stims[part.stims.order, c("Sentence", "Target", "Kindgender", "Construction", "Modelpred", "Modelprediction", "Expectedcorr", "Adjtarget", "Wordcount", "Number", "Expno", "Item")]
   part.stims$woffset <- head(c(1, cumsum(part.stims$Wordcount)+1), -1)
   
   part.stims$Position = 1:nrow(part.stims)
@@ -81,20 +81,25 @@ for (dafi in data.files) {
   part.stims$t2        <- apply(part.stims, 1, function(r) { ifelse(!r['Adjtarget'] == 'x', unlist(unname(strsplit(r["Sentence"], " ", fixed = T)))[as.numeric(r["Adjtarget"])+1 ], NA) } )
   part.stims$t1.length <- apply(part.stims, 1, function(r) { ifelse(!is.na(r["t1"]), nchar(r["t1"]), NA) } )
   part.stims$t2.length <- apply(part.stims, 1, function(r) { ifelse(!is.na(r["t2"]), nchar(r["t2"]), NA) } )
[…]

### 2017-04-01 Latex/article.tex
milestone: done except intro/conclusion

@@
 \usepackage{eqnarray}
+\usepackage{colortbl}
 
@@
 \newcommand{\pPB}{p\Subsf{PB}}
+\newcommand{\mpPB}{\ensuremath{p_{\text{PB}}}}
 
@@
 
+\definecolor{lg}{rgb}{.8,.8,.8}
+\newcommand{\Dim}{\cellcolor{lg}}
+
 \usepackage{gb4e-}
@@
 \newcommand{\CheckIt}{\CheckmarkBold}
+\newcommand{\CrossIt}{\XSolidBold}
 
@@
   \abstract{}
-  \keywords{corpus methods and experimental methods, alternations, hierarchical models, measure constructions, German}
+  \keywords{corpus methods and experimental methods, self-paced reading, forced choice, alternations, hierarchical models, measure constructions, German}
   %\classification[PACS]{...}
@@
 
-\section{Cognitively oriented corpus linguistics: theories, case studies, and methods}
+\section{Cognitively oriented corpus linguistics}
 \label{sec:cogocl}
@@
 
-The alternation phenomenon to be discussed in this paper appears to be a straightforward case alternation at first sight.
-However, without a solid knowledge of German nominal morpho-syntax, some of the more subtle distinctions that have to be made in its analysis can be confusing.
-Therefore, I introduce and illustrate the relevant constructions in some detail in this section.
-I describe the narrow range of syntactic constellations in which the alternation occurs, and I motivate the focus on \textit{only} this narrow (range rather than, for example, the whole range of nominal constructions expressing quantities).
+In this section, I introduce and illustrate the relevant alternating constructions.
+I describe the narrowly defined syntactic configuration in which the alternation occurs, and I motivate the focus on \textit{only} this narrow (range rather than, for example, the whole range of nominal constructions expressing quantities).
 
-I use the term \textit{measure noun phrase} (MNP) to refer a noun phrase (NP) in which a kind-denoting (count or mass) noun depends on another noun specifying a quantity of the objects or the substance denoted by the kind noun.
+I use the term \textit{measure noun phrase} (MNP) to refer a noun phrase (NP) in which a kind-denoting (count or mass) noun depends on another noun that specifies a quantity of the objects or the substance denoted by the kind noun.
 I call the kind-denoting noun simply the \textit{kind noun} and the quantity-denoting noun the \textit{measure noun}.
@@
 \footnote{If the kind noun is definite, the construction instantiates a true partitive.
-Whereas partitives are constructions denoting a proper part-of relation strictly requiring a definite kind noun as in \textit{a sip of the wine}, pseudo partitives -- albeit syntactically similar and diachronically related to partitives in many languages -- merely denote quantities and contain indefinite kind nouns as in \textit{a sip of wine}.
+Whereas partitives are constructions denoting a proper part-of relation as in \textit{a sip of the wine}, pseudo partitives -- albeit syntactically similar and diachronically related to partitives in many languages -- merely denote quantities and contain indefinite kind nouns as in \textit{a sip of wine}.
 %In the literature on German, some authors incorrectly call the pseudo-partitive a \textit{partitive} \cite{Hentschel1993} while some realise the difference and at least mention it \citep{Eschenbach1994,GallmannLindauer1994,Loebel1989,Zimmer2015}.
@@
 
-When the kind noun is bare -- \ie, without a determiner and without modifying adjectives -- the kind noun has to agree in case with the measure noun, and the genitive seen in the PGC is not acceptable, see (\ref{ex:intro:narrowapposition}).
-This construction is usually classified as a \textit{Narrow Apposition Construction} \citep{Loebel1986}, henceforth NAC.%
-\footnote{Notice that the unavailability of the genitive on the kind noun can be seen as following from a rather quirky constraint that genitive NPs in German require the presence of some strongly case-marked element (determiner or adjective) in addition to the head noun in order to be acceptable \citep{GallmannLindauer1994,Schachtl1989}.}
-The construction as in (\ref{ex:intro:narrowapposition2}) is also referred to as the \textit{Direct Partitive Construction} for other Germanic languages in which the PGC with the synthetic genitive is not available.%
-\footnote{This nomenclature makes sense in contrast to the \textit{Indirect Partitive Construction} with prepositional linkers translating to `of' -- \ie\ analytic genitives -- in such languages, see \cite{HankamerMikkelsen2008} for Danish.
-For German, this terminology is not distinctive enough, which is why I use the terms NAC and PGC.}
-While only an accusative is shown in (\ref{ex:intro:narrowapposition}), the kind noun obligatory agrees in case also with nominative and dative measure nouns.
+If the kind noun is bare -- \ie\ if it does neither come with a determiner nor a modifying adjective -- it has to agree in case with the measure noun, and the genitive seen in the PGC is not acceptable, see (\ref{ex:intro:narrowapposition}).
 
@@
 
+This construction is usually classified as a \textit{Narrow Apposition Construction} \citep{Loebel1986}, henceforth NAC.%
+\footnote{The construction as in (\ref{ex:intro:narrowapposition2}) is also referred to as the \textit{Direct Partitive Construction} for other Germanic languages in which the PGC with the synthetic genitive is not available.
+This nomenclature makes sense in contrast to the \textit{Indirect Partitive Construction} with prepositional linkers translating to `of' -- \ie\ analytic genitives -- in such languages, see \cite{HankamerMikkelsen2008} for Danish.
+For German, this terminology is not distinctive enough, which is why I use the terms NAC and PGC.}
+Notice that the unavailability of the genitive on the kind noun can be seen as following from a rather quirky constraint that genitive NPs in German require the presence of some strongly case-marked element (determiner or adjective) in addition to the head noun in order to be acceptable \citep{GallmannLindauer1994,Schachtl1989}.
+While only an accusative is shown in (\ref{ex:intro:narrowapposition}), the kind noun obligatory agrees in case also with nominative and dative measure nouns.
+
 The actual alternation can be observed \textit{only when the kind noun occurs with an attributive adjective but without a determiner}, as in (\ref{ex:intro:alternation}), where both the NAC in (\ref{ex:intro:alternation1}) and the PGC in (\ref{ex:intro:alternation2}) are equally acceptable.
@@
 See \cite{Hentschel1993,Zimmer2015} for analyses of the sometimes absurd stances taken in grammars of German.
-%As the usage and experimental data presented below should render it clear, there might be preferences under certain circumstances, but we cannot assume either construction to be unacceptable.
+As the usage and experimental data presented below (especially Section~\ref{sec:exp:fc}) should render it clear, there might be preferences under certain circumstances, but we cannot assume either construction to be unacceptable.
 }
@@
     \midrule
-    narrow apposition         & \CheckIt & \CheckIt &          \\
-    pseudo-partitive genitive &          & \CheckIt & \CheckIt \\
+    narrow apposition         & \NACb & \NACa & ---   \\
+    pseudo-partitive genitive & ---   & \PGCa & \PGCd \\
   \end{tabular}
@@
 
-The distribution of the case patterns in the NAC (case identity between the measure noun and the kind noun) and in the PGC (genitive on the kind noun) depending on the structure of the kind NP are summarised in Table~\ref{tab:constructions}.
+The distribution of the case patterns in the NAC (case identity between the measure noun and the kind noun) and in the PGC (genitive on the kind noun, regardless of the case of the measure noun) depending on the structure of the kind NP are summarised in Table~\ref{tab:constructions}.
 I call the narrow apposition construction with a bare kind noun the \NACb, the partitive genitive with a determiner in the kind noun phrase the \PGCd.
@@
   \item{\label{it:intro:datsg} alternative \textit{weak} forms of dative singular neuter adjectives}
-  \item{\label{it:intro:femsg} case syncretism in feminine NPs}
+%  \item{\label{it:intro:femsg} case syncretism in feminine NPs}
   \item{\label{it:intro:plurl} similar constructions with plural/collective kind nouns}
@@
 As for (\ref{it:intro:datsg}), a complication with neuter kind nouns in the dative is mentioned by \citet[20--22]{Zimmer2015}.
-He reports a high number of occurrences of adjectives being inflected according to the \textit{weak} inflectional pattern, which is normally only used if a strongly inflected determiner precedes the adjective: \textit{kalt-en Wasser} `cold water' instead of \textit{kalt-em Wasser}.%
+In the \NACa, the adjective normally inflects with the so-called \textit{strong} case and number suffixes, which are used when no determiner with strong inflection is present in the NP.
+\cite{Zimmer2015} reports a high number of occurrences of adjectives being inflected with the \textit{weak} inflectional marker, which are normally only used if a strongly inflected determiner precedes the adjective: \textit{kalt-en Wasser} `cold water' instead of \textit{kalt-em Wasser}.%
 \footnote{See Section~\ref{sec:analyses} for more discussion of adjectival inflection.}
 In my corpus sample, this tendency was not nearly as clear, and there was a high number of very noisy sentences among those potentially showing this pattern.
-Hence, I do not discuss these forms much, although I will offer a possible interpretation in Section~\ref{sec:analyses}.
+Hence, I do not discuss these forms.%
+\footnote{All corpus data and scripts used for this paper will be released freely, so further examination of the few cases maybe showing this kind of inflection is possible.}
 
-\label{page:femininesyncretism}
-With feminine kind nouns, (\ref{it:intro:femsg}) is relevant.
-Whereas the singular masculine and neuter nominal subsystem is a four case system (at least for NPs with adjectives and determiners), feminine singular NPs show nominative-accusative and dative-genitive syncretisms.
-Table~\ref{tab:syncretisms} shows this as well as the partial syncretism in the plural, where no gender distinctions are made.
-This means that with dative measure nouns, the alternation is unobservable if the kind noun is feminine, and it is problematic to speak of a \textit{genitive} kind noun construction in this case.
-This is taken into account in the statistical modelling and the discussion presented in Section~\ref{sec:corpusstudies}.
-
-\begin{table}
-  \centering
-  \begin{tabular}{llll}
-     & Masculine\slash Neuter & Feminine & Plural \\
-     \midrule
-     nominative & rot-er Wein    & \multirow{2}{*}{frisch-e Sahne}   & \multirow{2}{*}{klein-e Äpfel} \\
-     accusative & rot-en Wein    &                                   &                                \\
-     dative     & rot-em Wein    & \multirow{2}{*}{frisch-er Sahne}  & klein-en Äpfel-n               \\
-     genitive   & rot-en Wein-es &                                   & klein-er Äpfel                 \\
-  \end{tabular}
-  \caption{Case syncretisms in NPs with an adjective and without a determiner (\textit{roter Wein} `red wine', \textit{frische Sahne} `fresh cream', \textit{kleine Äpfel} `small apples')}
-  \label{tab:syncretisms}
-\end{table}
+% \label{page:femininesyncretism}
+% With feminine kind nouns, (\ref{it:intro:femsg}) is relevant.
+% Whereas the singular masculine and neuter nominal subsystem is a four case system (at least for NPs with adjectives and determiners), feminine singular NPs show nominative-accusative and dative-genitive syncretisms.
+% Table~\ref{tab:syncretisms} shows this as well as the partial syncretism in the plural, where no gender distinctions are made.
+% This means that with dative measure nouns, the alternation is unobservable if the kind noun is feminine, and it is problematic to speak of a \textit{genitive} kind noun construction in this case.
+% This is taken into account in the statistical modelling and the discussion presented in Section~\ref{sec:corpusstudies}.
+% 
+% \begin{table}
+%   \centering
+%   \begin{tabular}{llll}
+%      & Masculine\slash Neuter & Feminine & Plural \\
+%      \midrule
+%      nominative & rot-er Wein    & \multirow{2}{*}{frisch-e Sahne}   & \multirow{2}{*}{klein-e Äpfel} \\
+%      accusative & rot-en Wein    &                                   &                                \\
+%      dative     & rot-em Wein    & \multirow{2}{*}{frisch-er Sahne}  & klein-en Äpfel-n               \\
+%      genitive   & rot-en Wein-es &                                   & klein-er Äpfel                 \\
+%   \end{tabular}
+%   \caption{Case syncretisms in NPs with an adjective and without a determiner (\textit{roter Wein} `red wine', \textit{frische Sahne} `fresh cream', \textit{kleine Äpfel} `small apples')}
+%   \label{tab:syncretisms}
+% \end{table}
 
 Turning to (\ref{it:intro:plurl}), readers might have noticed that so far only MNPs denoting quantities of substances (mass kind nouns such as \textit{ein Glas roter Wein\slash roten Weines} `a glass of red wine') have been discussed.
-If the kind noun is a plural count noun as in \textit{ein Sack kleine Äpfel\slash kleiner Äpfel} `a bag of small apples' (see Table~\ref{tab:syncretisms} for the inflection pattern), a similar alternation between PGC and NAC can be observed.
-%\footnote{\label{fn:eschenbash}On the collective interpretation of such constructions, see \cite{Eschenbach1994}.
-%It is, however, unfortunate that Eschenbach does not notice and discuss the morpho-syntactic distribution of PGC and NAC, even when she gets very close to listing the relevant data \citep[217]{Eschenbach1994}.}
-In line with experimental results reported in \citet[15--16]{Zimmer2015}, I found that the PGC is so dominant with plural kind nouns (67 of 861 cases or over 92\%, cf.\ Section~\ref{sec:corpusstudies}) that the alternation cannot be analysed in the same way as in the singular case.
-While this will play a role in the interpretation of the corpus findings, MNPs with plural ve kind nouns will not be dealt with prominently in the remainder of the paper.
-%A statistical model for plural kind nouns will be reported in passing in Section~\ref{sec:corpusstudies}, however.
+If the kind noun is a plural count noun as in \textit{ein Sack kleine Äpfel} (\NACa) or \textit{ein Sack kleiner Äpfel} (\PGCa) `a bag of small apples', a similar alternation between PGC and NAC can be observed.
+In line with experimental results reported in \citet[15--16]{Zimmer2015}, I found that the PGC is so dominant with plural kind nouns (67 of 861 cases or over 92\%, cf.\ Section~\ref{sec:corpusstudies}) that the alternation cannot be analysed in the same way as in the singular case.%
+\footnote{More concretely, I these cases were included in the regression analysis of the corpus data along with a factor encoding the number of the kind noun, this factor would most assuredly override any other regressor for the data points with a plural kind noun.}
+While this will play a role in the interpretation of the corpus findings, MNPs with plural kind nouns will not be included in the corpus study and the experiments reported here.
 
 As for (\ref{it:intro:noifl}), some measure nouns have been grammaticalised in a way that they always appear non-inflected.
-They are typical measure nouns like \textit{Gramm} `gram', \textit{Pfund} `pound' or \textit{Prozent} `percent', which do not have plural forms at all.%
-\footnote{With neglectable frequency, this effect is extended to container nouns like \textit{Glas} `glass', but with a twist.
-A non-inflected variant \textit{zwei Glas roter Wein} `two glasses of wine' and an inflected variant \textit{zwei Gläser roter Wein} co-exist.
-The non-inflected variant denotes the quantity of wine fitting into two glasses, the inflected variant denotes two glasses filled with wine.}
-I treat these cases like any other measure noun because they enter into both the \NACa\ as in (\ref{ex:uninflectedmeasures:a}) and the \PGCa\ as in (\ref{ex:uninflectedmeasures:b}).
+They are typical measure nouns like \textit{Gramm} `gram', \textit{Pfund} `pound' or \textit{Prozent} `percent', which do not have plural forms at all.
+%\footnote{With neglectable frequency, this effect is extended to container nouns like \textit{Glas} `glass', but with a twist.
+%A non-inflected variant \textit{zwei Glas roter Wein} `two glasses of wine' and an inflected variant \textit{zwei Gläser roter Wein} co-exist.
+%The non-inflected variant denotes the quantity of wine fitting into two glasses, the inflected variant denotes two glasses filled with wine.}
+I treat these cases like other measure nouns because they enter into both the \NACa\ as in (\ref{ex:uninflectedmeasures:a}) and the \PGCa\ as in (\ref{ex:uninflectedmeasures:b}).
 In Section~\ref{sec:analyses}, degrees of grammaticalisation as a factor influencing the alternation will be discussed, however.
@@
 Finally, (\ref{it:intro:other}) suggests that there are alternative ways of expressing similar quantificational meanings.
-%First of all, the analytic (pseudo-)partitive with \textit{von} `of' is only available as an alternative to the PGC if the kind noun phrase contains a (definite or indefinite) determiner as in (\ref{ex:analyticalpartitive}).
-%
-%\begin{exe}
-%  \ex\label{ex:analyticalpartitive} 
-%  \begin{xlist}
-%    \ex[ ]{\label{ex:analyticalpartitive:a}\gll ein Glas von dem roten Wein\\
-%                                                a glass of the red wine\\
-%                                                \glt a glass of the red wine
-%       }
-%    \ex[*]{\label{ex:analyticalpartitive:b}\gll ein Glas von rotem Wein\\
-%                                                a glass of red wine\\
-%                                                \glt a glass of red wine
-%       }
-%  \end{xlist}
-%\end{exe}
-%
-If the measure noun denotes a container, constructions with \textit{voll} (\textit{von})\slash\textit{voller} `full (of)' and \textit{mit} `with' are available, see (\ref{ex:alternatives}).
+In the variationist tradition, which is strongly influenced by Labovian sociolinguistics, the \textit{principle of accountability} would dictate that proper studies should examine a variationist \textit{variable}, \ie\ all different \textit{ways of saying the same thing} (\citealp{Labov1966}, \citealp{Labov1969}, for an overview see \citealp{Tagliamonte2012}).
+In the case at hand, the variable might be something like \textit{measuring substances and collections}.
+I argue that it is fully justified to focus narrowly on \NACa\ and \PGCa\ with their well-defined morph-syntactic properties, mostly because the alternative constructions are not used in the same range of contexts.
+Two major alternatives might be considered.
+First of all, the analytic (pseudo-)partitive with \textit{von} `of' is only available as an alternative to the PGC if the kind noun phrase contains a (definite or indefinite) determiner as in (\ref{ex:analyticalpartitive}).
+
+\begin{exe}
+  \ex\label{ex:analyticalpartitive} 
+  \begin{xlist}
+    \ex[ ]{\label{ex:analyticalpartitive:a}\gll ein Glas von dem roten Wein\\
+                                                a glass of the red wine\\
+                                                \glt a glass of the red wine
+       }
+    \ex[*]{\label{ex:analyticalpartitive:b}\gll ein Glas von rotem Wein\\
+                                                a glass of red wine\\
+                                                \glt a glass of red wine
+       }
+  \end{xlist}
+\end{exe}
+
+This means that the \textit{von} (pseudo-)partitive does not compete with the \NACa\ and the \PGCa.
+In fact, there is not a single context in which they can be interchanged.
+
+Second constructions with \textit{voll} (\textit{von})\slash\textit{voller} `full (of)' and \textit{mit} `with' are available, see (\ref{ex:alternatives}).
 
@@
 
-These construction have very idiosyncratic properties (the construction with \textit{voller} is discussed in \citealp{Zeldes2018}, the construction with \textit{mit} is discussed in \citealp{Bhatt1990}), they are not semantically equivalent to the NAC and the PGC, and -- most importantly -- they are only available for a small subset of measure nouns.
-Subtle interactions between these constructions and NAC\slash PGC cannot be excluded, especially because the former might act as a (partial) alternative and thus reduce the count of exemplars of the NAC, the PGC, or both for container measure nouns.
-There are no clear hypotheses about such interactions, however.
-Nevertheless, the corpus-based models in Section~\ref{sec:corpusstudies} are specified in a way that they could detect preferences for either NAC or PGC for the class of container measure nouns, for which the alternative constructions illustrated in (\ref{ex:alternatives}) are available.
+These construction have very idiosyncratic properties (the construction with \textit{voller} is discussed in \citealp{Zeldes2018}, the construction with \textit{mit} is discussed in \citealp{Bhatt1990}), they are not semantically equivalent to the NAC and the PGC, and they are only available for a small subset of measure nouns.
+All of these constructions are only used with measure nouns denoting containers, and the whole NP always refers to the container, and never to the quantity contained in it.
+They are consequently incompatible with measure nouns denoting natural portions such as \textit{Schluck} `gulp' or \textit{Haufen} `heap' and strongly grammaticalised nouns such as \textit{Gramm} `gram'.
+This disqualifies them as alternatives to the \NACa\ or \PGCa\ in most contexts, and they are thus decidedly \textit{not} more ways of saying the same thing. 
+It is therefore reasonable to focus on the two well-defined variants.
 
@@
 
-This section briefly reviews existing analyses of the \PGCa-\NACa\ alternation and related issues.
+This section briefly reviews existing analyses of the \PGCa vs.\ \NACa\ alternation and related issues.
 I also develop my own analysis and the appropriate hypotheses for the empirical studies presented in Sections~\ref{sec:corpusstudies} and \ref{sec:externalvalidation}.
@@
 The first question is which noun constitutes the head of the whole construction.
-This question was answered for the case at hand quite clearly by \cite{Loebel1986} already (see also \citealp[213]{Eschenbach1994}, and \citealp[16]{GallmannLindauer1994}).
+This was answered quite clearly by \cite{Loebel1986} already (see also \citealp[213]{Eschenbach1994}, and \citealp[16]{GallmannLindauer1994}).
 Subject-verb agreement is always realised on the measure noun, and we can therefore assume that the measure noun is the head.
-However, the MNP-internal structure has more interesting cues to offer if the strong\slash weak inflection patterns of adjectives (already mentioned in Section~\ref{sec:descriptive}) are taken into account.
+However, the MNP-internal structure has more interesting cues to offer if the strong\slash weak inflection patterns of adjectives are taken into account.
 In NPs with a strongly inflected determiner, attributive adjectives inflect according to the massively syncretistic \textit{weak} pattern.
@@
 %The only exception to this rule is the masculine accusative singular, which is also marked with \textit{-en}.}
-If there is no determiner (as is the case in the alternating constructions), attributive adjectives inflect like determiners, however.
-This is called the \textit{strong} inflectional pattern (see Table~\ref{tab:syncretisms}).
+If there is no determiner (as is the case in the alternating constructions), attributive adjectives inflect like determiners themselves, however.
+This is called the \textit{strong} inflectional pattern.
 Thus, the adjectives in the \NACa\ and the \PGCa\ have properties of adjectives as well as determiners.
@@
 The choice between \PGCa\ and \NACa\ is then expected to be influenced by the degree of similarity to either of these prototypes.
+I will argue below that the prototypes differ mainly in that they attract different (classes of) measure nouns more or less strongly, mostly because of degrees of grammaticalisation.
+However, if lemmas are attracted with different strengths by the two prototypes, then we can expect a baseline attraction effect:
+nouns which occur more often in the \PGCd\ should favour the \PGCa, and nouns which occur more often in the \NACb\ should favour the \NACa.
+Put differently, a noun's relative frequencies in the prototypical non-alternating constructions should be indicative of same noun's preference to appear in the two alternating variants.
+This baseline attraction effect will be quantified in the corpus study Section~\ref{sec:corpusstudies}.
 
-Any attempt to substantiate or test this hypothesis is rendered difficult by the fact that the semantic differences between the variants are at best subtle, which means that the prototypes are not associated with clearly detectable differences in meaning.
-However, if there is a semantic difference, kind and measure noun lemmas should be attracted more strongly to one of the prototypes.
-As a result, kind noun lemmas and measure noun lemmas are expected to appear with different relative frequencies in the prototypical \NACb\ and the prototypical \PGCd.
-Their relative frequencies in these non-alternating configurations should then be indicative of their preference to appear in one of the two alternating variants \NACa\ and \PGCa.
-This predicted effect will be quantified in Section~\ref{sec:corpusstudies}.
-
-Turning to grammaticalisation now, it is often assumed that pseudo-partitives arise as a form of grammaticalised partitives (\eg\ \citealp[536--539]{Koptjevskaja2001} for Finnish and Estonian, \citealp[559]{Koptjevskaja2001} for European languages per se).
+Turning to grammaticalisation, it is often assumed that pseudo-partitives arise as a form of grammaticalised partitives (\eg\ \citealp[536--539]{Koptjevskaja2001} for Finnish and Estonian, \citealp[559]{Koptjevskaja2001} for European languages per se).
 %
@@
 %
-The alternating constructions in German are both clearly pseudo-partitives, but the grammaticalisation paths uncovered by \cite[esp.\ 526--530]{Koptjevskaja2001} -- which I interpret freely here -- are clearly relevant for an analysis of the alternation.
+The alternating constructions in German are both clearly pseudo-partitives, but the grammaticalisation paths uncovered by \cite[esp.\ 526--530]{Koptjevskaja2001} are relevant for an analysis of the alternation.
 The grammaticalisation path can start out (in some languages) with constructions involving two referential nouns (not necessarily forming a single and contiguous NP) and a \textit{separative} meaning as in \textit{(cut) two slices from the cake} \citep[535]{Koptjevskaja2001}.
@@
 The pseudo-partitive stage finally instantiates a \textit{quantity-of} relation, potentially even leading to fully grammaticalised quantifiers such as \textit{a lot}.
-In German, the PGC is clearly the older variant \citep{Zimmer2015} with an overt \textit{quantity-of} marker in the form of the genitive suffix.
+In German, the PGC is clearly the older variant \citep{Zimmer2015}, still having an overt \textit{quantity-of} marker in the form of the genitive suffix.
 Also, it still has the potential to form a true partitive (if the kind noun is definite).
@@
 Hence, we can expect more strongly grammaticalised measure nouns to prefer the NAC.
-For example, strongly grammaticalised non-referential nouns like \textit{Gramm} `gram' and \textit{Meter} `metre' should occur proportionally more often in the \NACa\ than in the \PGCa.%
+For example, highly grammaticalised non-referential nouns like \textit{Gramm} `gram' and \textit{Meter} `metre' should occur proportionally more often in the \NACa\ than in the \PGCa.%
 \footnote{In case studies like \cite{Brems2003,DeclerckBrems2016} syntactic evidence for grammaticalisation plays a prominent role.
@@
 In other words, a precisely quantified measure noun leads to a higher individuation (the creation of discrete portions) of the substance denoted by the kind noun, which is more appropriate for the PGC prototype.
-The statistical models in Section~\ref{sec:corpusstudies} will be designed to detect such a subtle effect, if it exists.
+The statistical models in Section~\ref{sec:corpusstudies} will be designed to detect such an effect.
 %\footnote{In the extreme, some measure nouns cannot be pluralised, such as \textit{eine Menge} `a lot of'.
@@
 \footnote{It was shown in \cite{W16-2601} that, for example, the spread of topics is much smaller in DeReKo compared to DECOW.}
-Second, it was already mentioned that normative grammars often adopt clear positions on the grammaticality of the NAC or the PGC.
+Second, it was already mentioned that normative grammars often adopt clear positions regarding the grammaticality of either the \NACa\ or the \PGCa.
 Thus, newspaper text or any other text that conforms strongly to normative grammars might not represent the alternation phenomenon fully (and without bias) because authors and proof-readers might favour one alternative or the other.
@@
 Among the factors potentially influencing the alternation (see Section~\ref{sec:germanmeasurenps}) were lemma-specific preference effects.
-Therefore, it was highly desirable to obtain a sample in which all (or at least most of the highly frequent) actually occurring combinations of kind noun and measure noun were represented.
+Therefore, it was highly desirable to obtain a sample in which most of the highly frequent actually occurring combinations of kind nouns and measure nouns were represented.
 I applied a three-stage bootstrap process in order to obtain such a sample.
@@
   \item\label{enum:data:step1} bootstrapping a list of the one hundred most frequent mass nouns,
-  \item\label{enum:data:step2} bootstrapping a list of all measure nouns with which the mass nouns co-occur in the bare-noun NAC 
+  \item\label{enum:data:step2} bootstrapping a list of all measure nouns with which the mass nouns co-occur in the \NACb 
   \item\label{enum:data:step3} sampling the target constructions by querying each combination of mass noun and measure noun found in step (\ref{enum:data:step2}).
@@
 
-This resulting list of mass nouns was used in step (\ref{enum:data:step2}) to bootstrap a list of measure nouns co-occurring with the mass nouns from DECOW14A01.
-In order to generate this list, I utilized the fact that a direct sequence of two nouns almost always instantiates the bare-noun NAC if the second noun is a mass noun.
+This resulting list of mass nouns was used in step (\ref{enum:data:step2}) to bootstrap a list of measure nouns co-occurring with the mass nouns. 
+In order to generate this list, I utilised the fact that a direct sequence of two nouns almost always instantiates the bare-noun NAC if the second noun is a mass noun.
 Hence, I searched for all sequences N\Sub{1}N\Sub{2} where N\Sub{2} was one of the mass noun lemmas extracted in step (\ref{enum:data:step1}).
@@
 Given the careful bootstrapping and sampling procedure described in this section, we can be highly sure that it contains all relevant and reasonably frequent noun–noun combinations in the target constructions.%
-\footnote{In a similar fashion, the 100 most frequent measure nouns occurring with plural kind nouns were bootstrapped and queried, resulting in a sample of 1,876 sentences.
-As stated in Section~\ref{sec:germanmeasurenps}, the NAC is virtually never used with plural kind nouns, and this sample was not used for full analysis.
+\footnote{In a similar fashion, the 100 most frequent measure nouns occurring with plural kind nouns were bootstrapped and queried, resulting in a sample of 871 sentences.
+As stated in Section~\ref{sec:germanmeasurenps}, the \NACa\ is virtually never used with plural kind nouns, and this sample was not used except for quantifying the frequency of occurrence of the constructions (67 times \NACa\ and 794 times \PGCa).
+The sample is distributed in the data package accompanying this paper, however.
 }
@@
 
-The full set of manually annotated variables for the main sample is given in Table~\ref{tab:variables}, and I briefly discuss them now.%
+The full set of manually annotated variables for the main sample is given in Table~\ref{tab:variables}, and I briefly discuss it now.%
 \footnote{All numeric variables were also z-transformed (\ie, centered to the mean and rescaled such that they have a standard deviation of $1$) to facilitate their interpretation in the regression models reported in the next section.}
@@
 To capture this, class information was annotated for measure lemmas.
-The classification was designed inspired by the list in \cite[530]{Koptjevskaja2001}, but due to the low frequencies of many of the potential classes, a very coarse classification was finally used.
-With typical examples and the number of data points in the final sample, the classes are:
-\textit{Physical} (abstract precisely measurable units such as \textit{Liter} `litre', \textit{Meter} `metre', \textit{Gramm} `gram'; size 1,968),
-\textit{Container} (\textit{Eimer} `bucket'; size 740), 
-\textit{Amount} (\textit{Menge} `amount'; size 1,364), 
-\textit{Portion} (natural portions like \textit{Happen} `bite' or \textit{Krümel} `crumb'; size 713).
-The few lemmas that did not fit into either of these classes were labeled \textit{Rest} (size 278).
+The classification was inspired by the list in \cite[530]{Koptjevskaja2001}, but due to the low frequencies of many of the potential classes, a very coarse classification was finally used.
+With typical examples and their frequencies in the final sample, the classes are:
+\textit{Physical} (abstract precisely measurable units such as \textit{Liter} `litre', \textit{Meter} `metre', \textit{Gramm} `gram'; $f=1,968$),
+\textit{Container} (\textit{Eimer} `bucket'; $f=740$), 
+\textit{Amount} (\textit{Menge} `amount'; $f=1,364$), 
+\textit{Portion} (natural portions like \textit{Happen} `bite' or \textit{Krümel} `crumb'; $f=713$).
+The few lemmas that did not fit into either of these classes were labeled \textit{Rest} ($f=278$).
 
 The variable \textit{Leftcontext} encodes the coarse-grained part-of-speech of the word preceding the measure noun.
-The level names stand for (again with class sizes in the final sample) \textit{cardinal} (size 1,939), \textit{determiner} (size 1,971), \textit{other} (size 221), \textit{adjective} (size 845), and \textit{preposition} (size 87).
+The level names stand for (again with their frequency in the final sample) \textit{cardinal} ($f=1,939$), \textit{determiner} ($f=1,971$), \textit{other} ($f=221$), \textit{adjective} ($f=845$), and \textit{preposition} ($f=87$).
 The purpose of this variable is mainly to test whether cardinals really favour the \PGCa\ as hypothesised in Section~\ref{sec:analyses}.
@@
   \includegraphics[width=0.85\textwidth]{../R/output/corpus_fixeffs_mle+mcmc}
-  \caption{Coefficients (MLE and MCMC) with 95\% confidence intervals (for details see text); intercept (\textit{Leftcontext=C}, \textit{Measurecase=Nom}, \textit{Measurenumber=Sg}, \textit{Kindgender=Masc}, \textit{Measureclass=Physical}; 0 for all numeric z-transformed regressors) is -4.328 (MLE) and -4.441 (MCMC)}
+  \caption{Coefficients (MLE and MCMC) with 95\% confidence intervals (for details see text); the intercept (\textit{Leftcontext=C}, \textit{Measurecase=Nom}, \textit{Measurenumber=Sg}, \textit{Kindgender=Masc}, \textit{Measureclass=Physical}; 0 for all numeric z-transformed regressors) is -4.328 (MLE) and -4.441 (MCMC)}
   \label{fig:fixeffs}
@@
 The purpose is to model the influence of the regressors specified in Table~\ref{tab:variables} on the probability that either the \NACa\ or the \PGCa\ is chosen.
-All regressors from Table~\ref{tab:variables} were included to begin with, and the measure lemma and the kind noun lemma were specified as varying-intercept random effects.
+All regressors from Table~\ref{tab:variables} were included, and the measure lemma and the kind noun lemma were specified as varying-intercept random effects.
 The sample size was $n=5,063$ with $1,1344$ cases of \PGCa\ and $3,929$ cases of \NACa.
@@
   \centering
-    \begin{tabular}{llrlrrrrrrcc}
-      Level & Regressor & \multicolumn{1}{l}{\pPB} & Level  & \multicolumn{2}{l}{Coefficient} & \multicolumn{2}{l}{CI low} & \multicolumn{2}{l}{CI high} & \multicolumn{2}{l}{CI contains 0} \\
-              &                   &       &          & \multicolumn{1}{l}{MLE} & \multicolumn{1}{l}{MCMC} & \multicolumn{1}{l}{MLE} & \multicolumn{1}{l}{MCMC} & \multicolumn{1}{l}{MLE} & \multicolumn{1}{l}{MCMC} & \multicolumn{1}{l}{MLE} & \multicolumn{1}{l}{MCMC} \\\midrule
+  \resizebox{\textheight}{!}{
+  \begin{tabular}{llrlp{0.5em}rrp{0.5em}rrp{0.5em}rrp{0.5em}cc}
+      Level & Regressor & \multicolumn{1}{l}{\pPB} & Level  && \multicolumn{2}{l}{Coefficient} && \multicolumn{2}{l}{CI low} && \multicolumn{2}{l}{CI high} && \multicolumn{2}{l}{CI contains 0} \\
+              &                   &       &           && \multicolumn{1}{l}{MLE} & \multicolumn{1}{l}{MCMC} && \multicolumn{1}{l}{MLE} & \multicolumn{1}{l}{MCMC} && \multicolumn{1}{l}{MLE} & \multicolumn{1}{l}{MCMC} && \multicolumn{1}{l}{MLE} & \multicolumn{1}{l}{MCMC} \\\midrule
 
-    First     & Badness           & 0.001 & (numeric) & -0.165 & -0.169 & -0.264 & -0.264 & -0.074 & -0.075 &      * &      * \\
-              & Genitives         & 0.001 & (numeric) & -0.630 & -0.645 & -0.710 & -0.739 & -0.523 & -0.552 &      * &      * \\
-              & Leftcontext       & 0.001 & D         &  1.178 &  1.173 &  0.751 &  0.784 &  1.548 &  1.546 &      * &      * \\
-              &                   &       & O         &  1.561 &  1.588 &  1.093 &  1.179 &  1.919 &  2.007 &      * &      * \\
-              &                   &       & A         &  2.300 &  2.340 &  1.872 &  1.959 &  2.644 &  2.726 &      * &      * \\
-              &                   &       & P         &  2.353 &  2.371 &  1.693 &  1.766 &  2.900 &  2.978 &      * &      * \\
-              & Measurecase       & 0.001 & Acc       &  0.065 &  0.070 & -0.130 & -0.119 &  0.268 &  0.269 &        &        \\
-              &                   &       & Dat       &  0.707 &  0.735 &  0.413 &  0.480 &  0.959 &  1.002 &      * &      * \\
-	      & Measurenumber     & 0.001 & Pl        &  0.814 &  0.825 &  0.534 &  0.574 &  1.080 &  1.079 &      * &      * \\[0.5\baselineskip]
+    First     & Badness           & 0.001 & (numeric) && -0.165 & -0.169 && -0.264 & -0.264 && -0.074 & -0.075 &&      * &      * \\
+              & Genitives         & 0.001 & (numeric) && -0.630 & -0.645 && -0.710 & -0.739 && -0.523 & -0.552 &&      * &      * \\
+              & Leftcontext       & 0.001 & D         &&  1.178 &  1.173 &&  0.751 &  0.784 &&  1.548 &  1.546 &&      * &      * \\
+              &                   &       & O         &&  1.561 &  1.588 &&  1.093 &  1.179 &&  1.919 &  2.007 &&      * &      * \\
+              &                   &       & A         &&  2.300 &  2.340 &&  1.872 &  1.959 &&  2.644 &  2.726 &&      * &      * \\
+              &                   &       & P         &&  2.353 &  2.371 &&  1.693 &  1.766 &&  2.900 &  2.978 &&      * &      * \\
+              & Measurecase       & 0.001 & Acc       &&  0.065 &  0.070 && -0.130 & -0.119 &&  0.268 &  0.269 &&        &        \\
+              &                   &       & Dat       &&  0.707 &  0.735 &&  0.413 &  0.480 &&  0.959 &  1.002 &&      * &      * \\
+	      & Measurenumber     & 0.001 & Pl        &&  0.814 &  0.825 &&  0.534 &  0.574 &&  1.080 &  1.079 &&      * &      * \\[0.5\baselineskip]
 
-    Second    & Kindattraction    & 0.033 & (numeric) &  0.204 &  0.216 &  0.030 &  0.014 &  0.366 &  0.412 &      * &      * \\
-    (Kind)    & Kindfreq          & 0.066 & (numeric) &  0.161 &  0.173 & -0.008 &  0.009 &  0.333 &  0.350 &        &      * \\
-              & Kindgender        & 0.001 & Neut      & -0.007 & -0.022 & -0.407 & -0.424 &  0.408 &  0.379 &        &        \\
-	      &                   &       & Fem       &  1.357 &  1.379 &  0.831 &  0.872 &  1.808 &  1.897 &      * &      * \\[0.5\baselineskip]
+    Second    & Kindattraction    & 0.033 & (numeric) &&  0.204 &  0.216 &&  0.030 &  0.014 &&  0.366 &  0.412 &&      * &      * \\
+    (Kind)    & Kindfreq          & 0.066 & (numeric) &&  0.161 &  0.173 && -0.008 &  0.009 &&  0.333 &  0.350 &&        &      * \\
+              & Kindgender        & 0.001 & Neut      && -0.007 & -0.022 && -0.407 & -0.424 &&  0.408 &  0.379 &&        &        \\
+	      &                   &       & Fem       &&  1.357 &  1.379 &&  0.831 &  0.872 &&  1.808 &  1.897 &&      * &      * \\[0.5\baselineskip]
 
-    Second    & Measureattraction & 0.015 & (numeric) &  0.247 &  0.275 &  0.058 &  0.078 &  0.413 &  0.476 &      * &      * \\
-    (Measure) & Measureclass      & 0.001 & Rest      &  0.564 &  0.488 & -0.140 & -0.309 &  1.197 &  1.236 &        &        \\
-              &                   &       & Container &  0.609 &  0.618 &  0.050 &  0.018 &  1.145 &  1.215 &      * &      * \\
-              &                   &       & Amount    &  0.927 &  0.959 &  0.184 &  0.189 &  1.566 &  1.718 &      * &      * \\
-              &                   &       & Portion   &  1.472 &  1.503 &  0.916 &  0.956 &  1.938 &  2.064 &      * &      * \\
-              & Measurefreq       & 0.003 & (numeric) & -0.257 & -0.264 & -0.413 & -0.435 & -0.098 & -0.099 &      * &      * \\
+    Second    & Measureattraction & 0.015 & (numeric) &&  0.247 &  0.275 &&  0.058 &  0.078 &&  0.413 &  0.476 &&      * &      * \\
+    (Measure) & Measureclass      & 0.001 & Rest      &&  0.564 &  0.488 && -0.140 & -0.309 &&  1.197 &  1.236 &&        &        \\
+              &                   &       & Container &&  0.609 &  0.618 &&  0.050 &  0.018 &&  1.145 &  1.215 &&      * &      * \\
+              &                   &       & Amount    &&  0.927 &  0.959 &&  0.184 &  0.189 &&  1.566 &  1.718 &&      * &      * \\
+              &                   &       & Portion   &&  1.472 &  1.503 &&  0.916 &  0.956 &&  1.938 &  2.064 &&      * &      * \\
+              & Measurefreq       & 0.003 & (numeric) && -0.257 & -0.264 && -0.413 & -0.435 && -0.098 & -0.099 &&      * &      * \\
   \end{tabular}
-  \smallskip
+  }
   \caption{Coefficient table comparing Maximum Likelihood Estimation (MLE, with 95\% bootstrap confidence interval) and `Bayesian' Markov-Chain Monte Carlo estimation (MCMC); for details see text}
@@
 Standard diagnostics for the Maximum Likelihood Estimation (MLE) show that the model quality is quite high.
-Generalized variance inflation factors for the regressors were calculated to check for multicollinearity \citep{FoxMonette1992,ZuurEa2010}, and none of the corrected $\text{GVIF}^{1/2\text{df}}$ was higher than $2$.
+Generalised variance inflation factors for the regressors were calculated to check for multicollinearity \citep{FoxMonette1992,ZuurEa2010}, and none of the corrected $\text{GVIF}^{1/2\text{df}}$ was higher than $2$.
 Nakagawa \& Schielzeth's pseudo-coefficient of determination is $R_m^2=0.423$ and $R^2_c=0.515$ (see \citealp{Gries2015} for a linguist-friendly introduction to these $R^2$ measures, or else \citealp{NakagawaSchielzeth2013}).
@@
 All regressors have low p\Sub{PB} values of magnitude $10^{-2}$ or $10^{-3}$.
-Only \textit{Kindfreq} ($p_{\text{PB}}=0.066$) can be seen as slightly too high to be convincing (`non-significant').
+Only \textit{Kindfreq} ($\mpPB=0.066$) can be seen as slightly too high to be convincing (`non-significant').
 
@@
 The signs and magnitudes of the coefficients are identical, and confidence intervals have the same width and symmetry properties.
-If one were interested in significance at $\alpha=0.05$, then the decision based on the confidence intervals differs for \textit{Kindfreq} ($p_{\text{PB}}=0.066$, MLE lower bound is $-0.008$ and MCMC lower bound is $0.009$).
+Figure~\ref{fig:fixeffs} illustrates this by also showing both estimates.
+If we were interested in significance at $\alpha=0.05$, then the decision based on the confidence intervals differs for \textit{Kindfreq} ($\mpPB=0.066$, MLE lower bound is $-0.008$ and MCMC lower bound is $0.009$).
 With MCMC, the regressor is `not significant' at the standard $\alpha$ level, but it is with MLE.
@@
 An interesting picture emerges for the lemma frequencies.
-A higher than average lemma frequency of measure nouns favours the \NACa\ as expected ($\beta_{\text{Measurefreq}}=-0.257$, $p_{\text{PB}}=0.003$).
-With kind nouns, higher frequency seems to favour the \PGCa ($\beta_{\text{Kindfreq}}=0.161$, $p_{\text{PB}}=0.066$).
+A higher than average lemma frequency of measure nouns favours the \NACa\ ($\beta_{\text{Measurefreq}}=-0.257$, $\mpPB=0.003$) as expected if we assume at least a tendency for highly grammaticalised items to be more frequent.
+With kind nouns, higher frequency seems to favour the \PGCa ($\beta_{\text{Kindfreq}}=0.161$, $\mpPB=0.066$).
+However, there is no clear theoretical interpretation (see Section~\ref{sec:analyses}), and the estimate is imprecise (`not significant at $\alpha=0.05$', see above).
+The effect can therefore be ignored or treated as a nuisance variable.
 
@@
 In Section~\ref{sec:analyses}, it was also hypothesised that there may be effects specific to classes of measure lemmas, and that they might be related to degrees of grammaticalisation of the measure noun.
-Transcending the effects of individual lemmas (captured in the random intercepts), the \textit{Measureclass} second-level predictor was successfully estimated ($p_{\text{PB}}=0.001$).
+Transcending the effects of individual lemmas (captured in the random intercepts), the \textit{Measureclass} second-level predictor was successfully estimated ($\mpPB=0.001$).
 Looking at the effect plot in Figure~\ref{fig:eff:measureattraction}, it is evident that abstract non-referential physical measure nouns (such as \textit{Gramm} `gram' or \textit{Liter} `litre') with a high degree of grammaticalisation favour the \NACa.
 At the other end of the scale, nouns denoting natural portions like \textit{Haufen} `heap', \textit{Bündel} `bundle', \textit{Schluck} `gulp' favour the \PGCa.
-These are referential nouns, confirming the hypothesis that it is prototypical of the PGC to host two referential nouns and the NAC only one.
+These are referential nouns, confirming the hypothesis that it is prototypical of the PGC to contain two referential nouns, while the NAC only contains one (the kind noun).
 
@@
 I now turn to the predicted effect of plural measure nouns and cardinals as quantifiers.
-Plurality has a clear effect in the expected direction (plural favours the \PGCa) ($\beta_{\text{MeasurenumberPl}}=0.814$, $p=0.001$ for \textit{Measurenumber}).
-Figure~\ref{fig:eff:leftcontext} shows that the left context of the measure noun influences the choice of the variant ($p=0.001$).
-Cardinals have a strong tendency to co-occur with the \NACa\ in both models.
+Plurality has a clear effect in the expected direction (plural favours the \PGCa) ($\beta_{\text{MeasurenumberPl}}=0.814$, $\mpPB=0.001$ for \textit{Measurenumber}).
+Figure~\ref{fig:eff:leftcontext} shows that the left context of the measure noun influences the choice of the variant ($\mpPB=0.001$).
+Cardinals have a strong tendency to co-occur with the \NACa.
+Both effects were predicted in Section~\ref{sec:analyses}.
 
@@
 The register-related proxy variables point into the expected direction.
-Increased \textit{Badness} of the document favours the \NACa\ ($\beta_{\text{Badness}}=-0.165$, $p=0.001$), and so does a lesser density of genitives ($\beta=-0.630$, $p=0.001$).
-While these are poor proxies to register (and partially circular in the case of \textit{Genitives}), this result can at least encourage future work into register effects
+Increased \textit{Badness} of the document favours the \NACa\ ($\beta_{\text{Badness}}=-0.165$, $\mpPB=0.001$), and so does a lesser density of genitives ($\beta=-0.630$, $\mpPB=0.001$).
+While these are poor proxies to register (and partially circular in the case of \textit{Genitives}), this result can at least encourage future work into register effects.
 
@@
 
-The effect of \textit{Measurecase} ($p=0.001$) is as predicted by previous studies (see Section~\ref{sec:analyses}).
+The influence of \textit{Measurecase} ($\mpPB=0.001$) is as predicted in previous analyses (see Section~\ref{sec:analyses}).
 A measure noun in the dative favours the \PGCa\ with $\beta_{\text{MeasurecaseDat}}=0.707$ (compared to the nominative, which is on the intercept).
@@
 Therefore, this section and the next present the results of two experiments wherein I cross-checked the corpus-based findings.
-Both experiments use sentences containing attested MNPs from the corpus sample embedded into radiacally simplified sentences, and the probabilities that the corpus-based statistical model (see Section~\ref{sec:corpushierarchicalmodel}) assigns to the competing constructions in these sentences.
+Both experiments use sentences containing attested MNPs from the corpus sample (embedded into simplified sentences) as stimuli.
+Also, the probabilities that the corpus-based model assigns to the two variants in these sentences is used as the main regressor in both studies.
 First, a forced-choice experiment was conducted.
@@
 
-As stimuli, attested MNPs from the corpus study were used, but the sentences were radically simplified to avoid .
+As stimuli, attested MNPs from the corpus study were used, but the sentences were radically simplified to avoid influences from contextual nuisance variables as much as possible.
+The approach is also justified because according to the theoretical assessment in Section~\ref{sec:analyses}, the choice of variants depends mostly on a very local constructional context.
 I sampled 16 MNPs from the corpus, and it was made sure that the simplifications and normalisations did not affect any of the regressors used in the corpus study.
-For example, the case, number, etc. of the MNP remained the same, as well as the choice of lexical material within the MNP.
+In the simplified sentences, the case, number, etc. of the MNP remained the same as in the attested sentence, as well as the choice of lexical material within the MNP.
 Eight sentences contained masculine or neuter kind nouns, the other eight contained feminine kind nouns.
@@
 \footnote{No participant ever exceeded the time limit.}
-The position on the screen (left\slash right) and the order of sentences were randomized for each participant.
+The position on the screen (left\slash right) and the order of sentences were randomised for each participant.
 As fillers, 23 pairs of sentences exemplifying similar alternation phenomena from German morpho-syntax were used.
@@
 The experiment was conducted using \textit{PsychoPy} \citep{Peirce2007}.
-Figure \ref{fig:afc:proportions} shows the relationship between the probabilities calculated from the corpus-based model and the proportional choices made by participants.
-The close correlation between the probabilities derived from usage data and the choices in the experiment are evident, and the relation appears to be linear.
+Figure \ref{fig:afc:proportions} shows the relationship between the probabilities calculated from the corpus-based model and the choices made by participants.
+The close correlation between the probabilities derived from usage data and the choices in the experiment are evident.
 
@@
 
-Then, a multilevel logistic regression was specified with the probability of the \PGCa\ predicted for each sentence by the corpus-based GLMM (\textit{Modelprediction}) as the only fixed effect.%
-\footnote{The document-level variables \textit{Badness} and \textit{Genitives} were set to $0$ (which is the mean for z-transformed variables).}
+Then, a multilevel logistic regression was specified with the probability of the \PGCa\ predicted for each sentence by the corpus-based GLMM as the only fixed effect \textit{Modelprediction}.%
+\footnote{The document-level variables \textit{Badness} and \textit{Genitives} were set to $0$, which is the mean for z-transformed variables.}
 A random intercept and slope were added for the individual sentence (item) in order to catch idiosyncrasies of single sentences.
 Also, a random intercept and slope for participants was added.%
-\footnote{The random slopes were added to comply with \citet[257]{BarrEa2013} who predict \textit{catastrophically high Type I error rates} for designs with within-subject manipulations if random effects structures are not kept maximal.
+\footnote{The random slopes were added to comply with \citet[257]{BarrEa2013} who predict \textit{catastrophically high Type I error rates} for experimental designs with within-subject manipulations if random effects structures are not kept maximal.
 Notice that the model reported here was estimated with conceptually identical results with regard to the predictor of interest (\textit{Modelprediction}) if only random intercepts were used.}
@@
 On the extreme sides, one participant chose the \PGCa\ in $13$ of $16$ cases, and two participants only chose it in $5$ of $16$ cases.
-The regressor \textit{Modelprediction} achieves $p_{\text{PB}}=0.002$ ($1,000$ replications) and is estimated at $4.500$ relative to an intercept of $-1.520$.
-The confidence interval from a parametric bootstrap ($1,000$ replications) for the regressor is acceptable with a lower bound of $2.738$ and an upper bound of $6.390$.
-Nakagawa \& Schielzeth's pseudo-coefficients of determination are $R^2_{m}=0.267$ and $R^2_{c}=0.435$, which means that roughly 27\% of the variance in the data can be explained by considering the predictions from the corpus-based GLMM.
+The regressor \textit{Modelprediction} achieves $\mpPB=0.002$ ($1,000$ replications) and is estimated at $4.500$ relative to an intercept of $-1.520$.
+The confidence interval from a parametric bootstrap ($1,000$ replications, percentile method) for the regressor is acceptable with a lower bound of $2.738$ and an upper bound of $6.390$.
+Nakagawa \& Schielzeth's pseudo-coefficients of determination are $R^2_{m}=0.267$ and $R^2_{c}=0.435$, which means that roughly 27\% of the variance in the data can be explained by considering only the predictions from the corpus-based GLMM.
 The effect display for the single fixed regressor \textit{Modelprediction} is given in Figure \ref{fig:afc:effects}.
-It confirms the effect intuitively noticeable from Figure~\ref{fig:afc:proportions}:
+It confirms the effect descriptively shown in Figure~\ref{fig:afc:proportions}:
 the higher the probability of the \PGCa\ predicted from usage-data, the more often participants chose the \PGCa\ variant in the forced-choice task.
@@
 Finally, I report the results of a self-paced reading experiment also designed to test whether the statistical model derived from usage data (see Section~\ref{sec:corpushierarchicalmodel}) can be used to predict speakers' behaviour.
-It is expected that reading the \textit{un}expected or less prototypical variant (the one assigned a low probability by the corpus-derived model) in a given context and with given lexical material might incur a processing overhead \citep{Kaiser2013}.
+It is expected that reading the less prototypical variant (the one assigned a low probability by the corpus-derived model) in a given context and with given lexical material incurs a processing overhead for the reader (\citealp{Kaiser2013}; see also Section~\ref{sec:cogocl}).
 In a very similar fashion, \cite{DivjakEa2016} apply the self-paced reading paradigm in the validation of corpus-based models.
@@
 I used exactly the same stimuli as in the forced choice experiments.
-Each participant read both the $16$ sentences with the variant predicted by the corpus model and the $16$ modified sentence with the variant that the corpus model would not predict.%
+Each participant read both the $16$ sentences with the variant predicted by the corpus model and the $16$ modified sentences with the variant that the corpus model does not predict.%
 \footnote{Notice that lemmas and their frequencies as well as lemma classes are part of the corpus-based GLMM, and there was no additional controlling of lemma frequencies, etc.}
@@
 The sentences from each pair of variants were kept as far apart as possible.
-The fillers also came in pairs such that the second block contained only sentences to which participants had been exposed in the first block in slightly modified form.
+The fillers also came in pairs such that the second block exclusively contained sentences to which participants had been exposed in the first block in slightly modified form.
 In total, each participant read $98$ sentences.
@@
 Outliers farther than 2 inter-quartile ranges from the mean logarithmised residual time were reamoved ($64$ data points), resulting in a total number of $n=1,152$ observations.
-An LMM was specified with the logarithmised residual reading times as the response variable and the probabilities derived from the corpus GLMM (\textit{Modelprediction}) as the main regressor.
-Since the corpus GLMM predicts the probability of the \PGCa, it is expected that reading times are positively correlated (longer reading times) with the probability if the stimulus actually contains the \NACa, and negatively correlated (shorter reading times) if the stimulus actually contains the \PGCa.
+An LMM was specified with the logarithmised residual reading times as the response variable, and the probabilities derived from the corpus GLMM (\textit{Modelprediction}) as the main regressor.
+Since the corpus GLMM predicts the probability of the \PGCa (vs.\ the \NACa), it is expected that reading times are positively correlated (longer reading times) with the probability if the stimulus actually contains the \NACa, and negatively correlated (shorter reading times) if the stimulus actually contains the \PGCa.
 To account for this, an interaction between \textit{Modelprediction} and \textit{Construction} (levels \textit{PGCa} and \textit{NACa}) was added to the model.
 Furthermore, the position of the sentence in the individual experiment run was included as a fixed effect to control for increasing reading speed during experiment runs.
-Random intercepts were specified for \textit{Participant} and \textit{Item} (the $16$ sentence pairs are one \textit{Item}).%
-\footnote{I tried random slopes, but it was impossible to get the algorithm to converge.}
+Random intercepts were specified for \textit{Participant} and \textit{Item} (the $16$ sentence pairs are one \textit{Item} each).%
+\footnote{I tried random slopes in order to keep the random effect structure maximal \citep{BarrEa2013}, but it was impossible to get the algorithm to converge.}
 
@@
   \end{tabular}
-  \caption{Fixed effect coefficient table for the LMM used to analyse the self-paced reading experiment}
+  \caption{Fixed effect coefficient table for the LMM used to analyse the self-paced reading experiment; the intercept is 0.819}
   \label{tab:exp:spr}
@@
 
-Table~\ref{tab:exp:spr} shows the coefficient estimates with a 95\% parametric bootstrap confidence interval ($1,000$ replications).
+Table~\ref{tab:exp:spr} shows the coefficient estimates with a 95\% parametric bootstrap confidence interval ($1,000$ replications, percentile method).
 The standard deviation of the participant intercepts is $\sigma_{\text{Participant}}=0.079$ and of the item intercepts $\sigma_{\text{Item}}=0.039$.
-Comparing the full model to a model without the main regressor \textit{Modelprediction} (and consequently also the interaction with \textit{Construction}) in a PB test gives $p_{\text{PB}}=0.018$.
+Comparing the full model to a model without the main regressor \textit{Modelprediction} (and consequently also without the interaction with \textit{Construction}) in a PB test gives $\mpPB=0.018$.
 Nakagawa and Schielzeth's pseudo-determination coefficients are $R^2_m=0.238$ and $R^2_c=0.347$.
@@
 While the overall model quality is acceptable, the relevant effect is estimated with some uncertainty, which is also visible in the effect plot in Figure~\ref{fig:spr:effects}.
-While the estimate for the sentences with \NACa\ is clearly too imprecise, there is a clear tendency in the sentences with \PGCa, which is also confirmed by the `significant' results from the bootstrapped confidence intervals (see Table~\ref{tab:exp:spr}) and from the PB test reported above.
+While the estimate for the sentences with \NACa\ is obviously very imprecise, there is a clearer tendency in the sentences with \PGCa, which is also confirmed by the `significant' results from the bootstrapped confidence intervals (see Table~\ref{tab:exp:spr}) and from the PB test reported above.
 The \PGCa\ brings about an increased reading time ($\beta_{\text{ConstructionPGCa}}=0.062$), which is plausible because it is the much rarer construction (see Section~\ref{sec:corpusstudies}).
-However, if it occurs in a prototypical situation, reading times drop ($\beta_{\text{ConstructionPGCa:Modelprediction}}=-0.132$), which is also reflected in the slope of curve in the right panel of Figure~\ref{fig:spr:effects}.
-This fits into the general picture in as much as the construction with the lower frequency might be developing towards a more sharply defined prototype.
-In this context, it should be remembered from Section~\ref{sec:gettingdata} that even the \PGCd\ is much rarer that the \NACb\ ($17,252$ vs.\ $315,635$ occurrences in the auxiliary corpus samples).
+However, if it occurs in a prototypical situation, reading times drop ($\beta_{\text{ConstructionPGCa:Modelprediction}}=-0.132$).
+This is noticeable in the downward slope of curve in the right panel of Figure~\ref{fig:spr:effects}.
+This fits into the general picture in as much as the construction with the lower frequency might be developing towards a more sharply defined prototype.%
+\footnote{In this context, it should be remembered from Section~\ref{sec:gettingdata} that even the \PGCd\ is much rarer that the \NACb\ ($17,252$ vs.\ $315,635$ occurrences in the auxiliary corpus samples).}
 Conversely, the \NACa\ (like the NAC in general) might be the highly frequent default which does not incur reading time penalty, even if it is not the optimal choice in the given context and with the given lexical material.
-This possibility will be considered, among other things, in the next section.
+This interpretation will be considered, among other things, in the next and final section.
 
@@
   I want to thank (in alphabetical order) Felix Bildhauer, Samuel Reichert, and Ulrike Sayatz for discussions on relevant aspects of this paper.
-  I also thank Ulrike Sayatz for recruiting the participants for the experiments.
+  I also thank Ulrike Sayatz for helping to recruit the participants for the experiments.
   I am grateful to Kim Maser for her work on the annotation of the concordances.

### 2017-04-02 R/0_wrapper.R
cleaning out old figures dir

@@ -13,7 +13,7 @@ setwd('/Users/user/Workingcopies/measurenps/GitHub/R')
 
 # Options.
 set.seed(2398651)                          # Do NOT change, or results will differ.
-save.persistent        <- F                # Write to files instead of screen/console.
+save.persistent        <- T                # Write to files instead of screen/console.
 out.dir                <- "output/"        # Put generated files here.
 precision              <- 3                # Precision in result table.
 the.nAGQ               <- 0                # 0 for speed, 1 for precision and eytremely slow computation.

### 2017-04-02 R/3_glmm.R

@@ -4,6 +4,8 @@ require(MuMIn)
 require(effects)
 require(gridExtra)
 require(car)
+require(pbkrtest)
+
 
 
 measure.glmm <- glmer(Construction~1
@@ -16,7 +18,7 @@ measure.glmm <- glmer(Construction~1
                  +Leftcontext
                  +Measurecase
                  +Measurenumber
-                 
+
                  +Kindattraction
                  +Kindfreq
                  +Kindgender
@@ -58,6 +60,7 @@ bootcomp.regs <- c("(1|Measurelemma)", "(1|Kindlemma)",
 modelcomp <- lmer.modelcomparison(model = measure.glmm, regressors = bootcomp.regs,
                                      formula.target = "Construction~1", nsim = ci.boot.modelcomp.nsim,
                                      print.updated = T)
+
 
 
 
@@ -141,9 +144,43 @@ par(mfrow=c(1,1))
 if (save.persistent) dev.off()
 
 
-
+# Multicollinearity.
 source('highstat.r')
 if (save.persistent) sink(paste(out.dir, "glmm.txt", sep=""), append = T)
 cat("\n\n Variance inflation diagnostics\n")
 print(myvif(measure.glmm))
+if (save.persistent) sink()
+
+
+# Nachschlag: PB test for interaction.
+measure.glmm.plus <- glmer(Construction~1
+                           
+                           +(1|Measurelemma)
+                           +(1|Kindlemma)
+                           
+                           +Badness                 
+                           +Genitives
+                           +Leftcontext
+                           +Measurecase
+                           +Measurenumber
+                           
+                           +Measurenumber:Leftcontext
+                           
+                           +Kindattraction
+                           +Kindfreq
+                           +Kindgender
+                           
+                           +Measureattraction
+                           +Measureclass
+                           +Measurefreq
+                           
+                           , data=measure, family=binomial(link=logit), na.action = na.fail, nAGQ=the.nAGQ,
+                           control=glmerControl(optimizer="nloptwrap2", optCtrl=list(maxfun=2e5)))
+
+
+modcomp.interact <- PBmodcomp(measure.glmm.plus, measure.glmm, nsim = ci.boot.modelcomp.nsim)
+if (save.persistent) sink(paste(out.dir, "glmm.txt", sep=""), append = T)
+cat("\n\nModel comparison with LR and PB test for INTERACTION\n")
+print(modcomp.interact)
+cat("\n\n")
 if (save.persistent) sink()

### 2017-04-02 R/4_glmm-mcmc.R

@@ -16,7 +16,7 @@ measure.glmm.mcmc <- stan_glmer(Construction~1
                            +Leftcontext
                            +Measurecase
                            +Measurenumber
-                           
+
                            +Kindattraction
                            +Kindfreq
                            +Kindgender

### 2017-04-02 Latex/article.tex
fixed a serious flaw in theory

@@
 In this section, I briefly motivate the choice of statistical models which I use later in Section~\ref{sec:corpusstudies}.
-In the spirit of the careful and methodologically sound approach to cognitively oriented corpus linguistics laid out in Section~\ref{sec:cogocl}, authors should always provide at least a minimal motivation as part of the standard protocol, especially if there are alternative methods.
 
@@
 \footnote{If the kind noun is definite, the construction instantiates a true partitive.
-Whereas partitives are constructions denoting a proper part-of relation as in \textit{a sip of the wine}, pseudo partitives -- albeit syntactically similar and diachronically related to partitives in many languages -- merely denote quantities and contain indefinite kind nouns as in \textit{a sip of wine}.
-%In the literature on German, some authors incorrectly call the pseudo-partitive a \textit{partitive} \cite{Hentschel1993} while some realise the difference and at least mention it \citep{Eschenbach1994,GallmannLindauer1994,Loebel1989,Zimmer2015}.
+Whereas partitives are constructions denoting a proper part-of relation as in \textit{a sip of the wine}, pseudo partitives -- albeit syntactically similar and diachronically related to partitives in many languages -- merely denote quantities and contain indefinite kind nouns as in \textit{a sip of wine}.%
+In the literature on German, some authors incorrectly call the pseudo-partitive a \textit{partitive} \cite{Hentschel1993} while some realise the difference and at least mention it \citep{Eschenbach1994,GallmannLindauer1994,Loebel1989,Zimmer2015}.
 }
@@
 I call the alternating variants with an adjective but no determiner in the kind noun phrase \NACa, and \PGCa, respectively.
+This paper is about the middle column of Table~\ref{tab:constructions}, \ie, the syntactic configuration in which two different case patterns are acceptable.
 
@@
 \begin{enumerate}[i.]
-  \item{\label{it:intro:idiot} alleged alternatives to the NAC and the PGC}
+  \item{\label{it:intro:idiot} alleged alternatives to the \NACa\ and the \PGCa}
   \item{\label{it:intro:datsg} alternative \textit{weak} forms of dative singular neuter adjectives}
@@
 
-First of all, (\ref{it:intro:idiot}) refers to claims found in some grammars that a generic nominative, accusative, and even dative on the kind noun are used instead of the genitive or case agreement (overviews in \citealp{Hentschel1993,Zimmer2015}).
+First of all, (\ref{it:intro:idiot}) refers to claims found in some grammars that a generic nominative, accusative, and even dative on the kind noun are used instead of the genitive (\PGCa) or case agreement (\NACa).
+Overviews can be found in \citealp{Hentschel1993,Zimmer2015}.
 It was shown empirically in \cite{Hentschel1993} that such variants are de facto not acceptable.
@@
 In the variationist tradition, which is strongly influenced by Labovian sociolinguistics, the \textit{principle of accountability} would dictate that proper studies should examine a variationist \textit{variable}, \ie\ all different \textit{ways of saying the same thing} (\citealp{Labov1966}, \citealp{Labov1969}, for an overview see \citealp{Tagliamonte2012}).
-In the case at hand, the variable might be something like \textit{measuring substances and collections}.
+In the case at hand, the variable might be something like \textit{measuring quantities of substances and collections}.
 I argue that it is fully justified to focus narrowly on \NACa\ and \PGCa\ with their well-defined morph-syntactic properties, mostly because the alternative constructions are not used in the same range of contexts.
@@
 I also develop my own analysis and the appropriate hypotheses for the empirical studies presented in Sections~\ref{sec:corpusstudies} and \ref{sec:externalvalidation}.
-I discuss four main factors that might influence the alternation:
+I discuss four main aspects:
 first, the different syntactic structures of the variants \NACa\ and \PGCa\ in relation to their syntactic prototypes \NACb\ and \PGCd;
 second, degrees of grammaticalisation of the measure noun associated with the two prototypes;
-third, a semantic prototypicality effect of the degrees of grammaticalisation (plurality);
+third, a semantic prototypicality effect of the degrees of grammaticalisation (preference to occur with cardinals);
 fourth, preferences in different registers.
 
-The first aspect is related to analyses of the morpho-syntax of German MNPs.
-A major interest in descriptive and generative linguistics is to find a plausible analysis of the hierarchical syntactic structure of the constructions in question.
-The first question is which noun constitutes the head of the whole construction.
+% MORPHO-SYNTAX AND PROTOTYPES
+
+My whole analysis is based on the idea that the syntactic structure [N\Sub{1}~[A~N\Sub{2}]\Sub{NP\Sub{2}}]\Sub{NP\Sub{1}} as instantiated in the \NACa\ and the \PGCa\ is ambiguous.
+To show this, the first question is which noun constitutes the head of the whole construction.
 This was answered quite clearly by \cite{Loebel1986} already (see also \citealp[213]{Eschenbach1994}, and \citealp[16]{GallmannLindauer1994}).
@@
 \footnote{While the generative analysis presented in \cite{Bhatt1990} cannot properly deal with probabilistic effects, Bhatt comes close to this interpretation by analysing the kind NP in the GPC as a DP and in the NAC as an NP.}
-In a framework that assumes similarity-based probabilistic classification, such as Prototype Theory, we can regard the \PGCd\ as instantiating the PGC prototype and the \NACb\ as the NAC prototype.
-The choice between \PGCa\ and \NACa\ is then expected to be influenced by the degree of similarity to either of these prototypes.
+
+Put differently, the \NACa\ is a \NACb in disguise, and the \PGCa\ is a \PGCd in disguise.
+In a probabilistic framework, the crucial question is what controls speakers' decisions to use a a \NACb in disguise or a \PGCd in disguise.
+If the \PGCd\ instantiates the PGC prototype and the \NACb\ the NAC prototype, the choice between \PGCa\ and \NACa\ is expected to be influenced by the degree of similarity to either of these prototypes.
 I will argue below that the prototypes differ mainly in that they attract different (classes of) measure nouns more or less strongly, mostly because of degrees of grammaticalisation.
-However, if lemmas are attracted with different strengths by the two prototypes, then we can expect a baseline attraction effect:
-nouns which occur more often in the \PGCd\ should favour the \PGCa, and nouns which occur more often in the \NACb\ should favour the \NACa.
+However, if lemmas are attracted with different strengths by the two prototypes, then we can expect a baseline attraction effect.
+Nouns which occur more often in the \PGCd\ should favour the \PGCa, and nouns which occur more often in the \NACb\ should favour the \NACa.%
+\footnote{Such associations between constructions and words or lemmas are the main object of study in approaches such as Collexeme Analysis \citep{GriesStefanowitsch2004}.}
 Put differently, a noun's relative frequencies in the prototypical non-alternating constructions should be indicative of same noun's preference to appear in the two alternating variants.
 This baseline attraction effect will be quantified in the corpus study Section~\ref{sec:corpusstudies}.
+
+% GRAMMATICALISATION
 
@@
 
+% CARDINAL QUANTIFIERS
+
 Furthermore, the grammaticalisation path as described above leads from NPs denoting individuated objects standing in a \textit{part-of} relation to a construction with a more diffuse \textit{quantity-of} relation.
-The best evidence for the relevance of this factor in an ongoing grammaticalisation process is that plural kind nouns almost always (but not exclusively) occur in the \PGCa\ and virtually never in the \NACa\ (see Section~\ref{sec:descriptive}).
-More indirectly, however, plurality of the measure noun could also favour the \PGCa\ over the \NACa, and exact quantification in the form of cardinals even more so.%
-\footnote{Without offering a theoretical analysis, such a preference was observed empirically by \cite{Hentschel1993}.}
-There are highly illustrative examples with measure nouns that have virtually developed two variants, one of which grammaticalised to a much higher degree, \eg\ \textit{Haufen} `heap\slash lot'.
-The phrase \textit{ein Haufen französische Watte} `a heap\slash lot of French cotton' (\NACa) has two uses.
-It can be used to denote an unspecific amount of cotton in the sense of \textit{a lot of French cotton} in situations when the exact quantity of cotton is not of interest.
-Under the appropriate circumstances, it could also be used to denote a heap-shaped, potentially highly specific quantity of cotton, for example when a cotton merchant always sells her French cotton in \textit{heaps} weighing two stone each.
-The former use, however, is highly restricted for \textit{zwei Haufen französischer Watte} `two heaps of French cotton' (\PGCa) -- not `two lots of French cotton'.
-Such an NP is less likely to be used in situations where the exact quantity of cotton is irrelevant, even if there are actually two heaps of cotton.
-In other words, a precisely quantified measure noun leads to a higher individuation (the creation of discrete portions) of the substance denoted by the kind noun, which is more appropriate for the PGC prototype.
-The statistical models in Section~\ref{sec:corpusstudies} will be designed to detect such an effect.
-%\footnote{In the extreme, some measure nouns cannot be pluralised, such as \textit{eine Menge} `a lot of'.
-%Also note that the presentation of the examples in NAC and PGC is suggestive.
-%Native speakers will confirm that both NPs sound highly acceptable in both GPC and NAC.
-%However, it was made clear from the beginning and throughout that the preferences are indeed subtle.}
-
-Finally, it has been iterated very often that the PGC is more typical of higher registers or even exclusive to written language (see \citealp[320--323]{Hentschel1993}, where she also refers to descriptive grammars).
-This is not surprising in as much as the genitive -- an intrinsic part of the PGC -- is generally underrepresented in colloquial and vernacular variants of German.
-Under an integral view, such preferences can be to be part of the construction prototypes.
-In the corpus study in Section~\ref{sec:corpusstudies}, register effects will therefore be modelled, even if only using some very simple proxies.
+Both types of relations can be numerically quantified -- in as much as a precise number of \textit{parts} or a numerically exact \textit{quantity} can be specified --, but it is much more prototypical of quantities to be specified with numerical precision.
+Since the \NACa\ is more closely associated with the \textit{quantity-of} relation, cardinals as attributes of the measure noun are expected to have a higher proportional frequency in the \NACa.
+For illustration, (\ref{ex:determiners}) shows the expected variants under this hypothesis.
+In (\ref{ex:determiners:nac}), the measure noun is modified by a cardinal \textit{drei} `three', and hence the \NACa\ (kind NP shows case agreement) is preferred.
+In (\ref{ex:determiners:pgc}), the measure noun is modified by a non-cardinal determiner \textit{einige} `some', and the \PGCa\ (kind NP is in the genitive) is preferred.
+
+\begin{exe}
+  \ex\label{ex:determiners} 
+  \begin{xlist}
+    \ex{\label{ex:determiners:nac}\gll [[Drei Löffel]\Sub{Nom} [heißer Rum]\Sub{Nom}]\Sub{Nom} sind genug.\\
+                                       three spoons hot rum are enough.\\
+				     \glt Three spoonful of hot rum are enough.}
+    \ex{\label{ex:determiners:pgc}\gll [[Einige Löffel]\Sub{Nom} [heißen Rums]\Sub{Gen}]\Sub{Nom} sind genug\\
+                                       some   spoons hot    rum  are   enough.\\
+				     \glt A few spoonful of hot rum are enough.}
+  \end{xlist}
+\end{exe}
+
+The statistical models presented in Section~\ref{sec:corpusstudies} will be specified in a way that they could detect such a preference.
+
+Finally, it has often been iterated that the \PGCa\ is more typical of higher registers or even exclusive to written language (see \citealp[320--323]{Hentschel1993}, where she also refers to descriptive and normative grammars).
+This is not surprising in as much as the genitive -- an intrinsic part of the PGC -- is generally underrepresented in colloquial vernacular variants of German as a result of a diachronic process wherein many (but not all) uses of the genitive are replaced by other cases or periphrastic constructions \citep{FleischerSchallert2011}.
+Under an integral view of prototypes, which incorporate effects related to larger contexts and registers, such preferences can be to be part of the construction prototypes.
+In the corpus study in Section~\ref{sec:corpusstudies}, register effects will therefore be modelled -- even if only using some very simple proxies.
 
@@
   \item (Classes of) more strongly grammaticalised measure nouns favour the NAC.
-  \item Plural measure nouns and measure nouns quantified by cardinals favour the PGC variant.
+  \item Measure nouns quantified by cardinals favour the PGC variant.
   \item The NAC is associated more strongly with higher registers.
@@
 For these or similar reasons, COW corpora have been used in a number of peer-reviewed publications, for example \cite{VanGoethemHiligsmann2014,VanGoethemHuening2015,MuellerS2014,Schaefer2016c,SchaeferSayatz2014,SchaeferSayatz2016,Zimmer2015}.
-Therefore, DECOW can be considered the obvious choice for this study.
+Therefore, DECOW can be considered the obvious choice for this study.%
+\footnote{The use of web data for linguistic research does require explicit and careful justification.
+Due to the noisy nature and unknown composition of the web, only carefully designed and established web corpora like the COW corpora or the SketchEngine corpora \citep{KilgarriffEa2014} should be used.
+Clearly, using search engine results is \textit{bad science} for many reasons, most prominently total irreproducibility of results, as \cite{Kilgarriff2006} pointed out more than ten years ago.
+Careless use of search engine results is still found, however, see \citet[171--175]{DeclerckBrems2016}.}
 
@@
 Each participant read both the $16$ sentences with the variant predicted by the corpus model and the $16$ modified sentences with the variant that the corpus model does not predict.%
-\footnote{Notice that lemmas and their frequencies as well as lemma classes are part of the corpus-based GLMM, and there was no additional controlling of lemma frequencies, etc.}
-To avoid repetition effects, the stimuli for each participant were separated into two blocks of $16$ targets and $33$ fillers per block.
+\footnote{Notice that lemmas and their frequencies as well as lemma classes are included as regressors in the corpus-based GLMM, and there was consequently no additional controlling of lemma frequencies, etc.}
+To minimise repetition effects, the stimuli for each participant were separated into two blocks of $16$ targets and $33$ fillers per block.
+In the experiment, participants first read all sentences from the first block, then all sentences from the second block.
 It was made sure that from each target sentence pair, one sentence was assigned to the first block, and the other sentence to the other block.
-The assignment of members of the individual sentence pairs to one of the blocks was randomised for each participant individually, and so was the order within the block.
+The assignment of members of the individual sentence pairs to the blocks was randomised for each participant individually, and so was the order within each block.
 The sentences from each pair of variants were kept as far apart as possible.
@@
 The reading times were residualised per speaker based on the reading times of all words (not just the targets) by that speaker.
-The adjective and kind noun (\ie\ the constituents bearing the critical case markers) were used as the target region, such as the bracketed words in the example \textit{zwei Gläser} [\textit{sprudelndes Wasser}] `two glasses of sparkling water'.
-Outliers farther than 2 inter-quartile ranges from the mean logarithmised residual time were reamoved ($64$ data points), resulting in a total number of $n=1,152$ observations.
+The adjective and the kind noun (\ie\ the constituents bearing the critical case markers) were used as the target region, such as the bracketed words in the example \textit{zwei Gläser} [\textit{sprudelndes Wasser}] `two glasses of sparkling water'.
+Outliers farther than 2 inter-quartile ranges from the mean logarithmised residual time were removed ($64$ data points), resulting in a total number of $n=1,152$ observations.
 An LMM was specified with the logarithmised residual reading times as the response variable, and the probabilities derived from the corpus GLMM (\textit{Modelprediction}) as the main regressor.
@@
 To account for this, an interaction between \textit{Modelprediction} and \textit{Construction} (levels \textit{PGCa} and \textit{NACa}) was added to the model.
-Furthermore, the position of the sentence in the individual experiment run was included as a fixed effect to control for increasing reading speed during experiment runs.
+Furthermore, the position ($1$--$98$) of the sentence in the individual experiment (\textit{Position} was included as a fixed effect to control for increasing reading speed during experiment runs.
 Random intercepts were specified for \textit{Participant} and \textit{Item} (the $16$ sentence pairs are one \textit{Item} each).%
-\footnote{I tried random slopes in order to keep the random effect structure maximal \citep{BarrEa2013}, but it was impossible to get the algorithm to converge.}
+\footnote{I tried random slopes in order to keep the random effect structure maximal \citep{BarrEa2013}, but it was impossible to get the algorithm to converge due to the added complexity of the interaction.}
 
@@
 
-While the overall model quality is acceptable, the relevant effect is estimated with some uncertainty, which is also visible in the effect plot in Figure~\ref{fig:spr:effects}.
-While the estimate for the sentences with \NACa\ is obviously very imprecise, there is a clearer tendency in the sentences with \PGCa, which is also confirmed by the `significant' results from the bootstrapped confidence intervals (see Table~\ref{tab:exp:spr}) and from the PB test reported above.
+The overall model quality is acceptable, and the effect plot for the main effect is shown in Figure~\ref{fig:spr:effects}.
+The estimate for the sentences with \NACa\ is obviously imprecise, although pointing into the right direction (longer reading times when the corpus model predicts the \PGCa).
+There is a clearer effect in the sentences with \PGCa, which is also confirmed by the `significant' results from the bootstrapped confidence intervals (see Table~\ref{tab:exp:spr}) and from the PB test reported above.
 The \PGCa\ brings about an increased reading time ($\beta_{\text{ConstructionPGCa}}=0.062$), which is plausible because it is the much rarer construction (see Section~\ref{sec:corpusstudies}).
-However, if it occurs in a prototypical situation, reading times drop ($\beta_{\text{ConstructionPGCa:Modelprediction}}=-0.132$).
-This is noticeable in the downward slope of curve in the right panel of Figure~\ref{fig:spr:effects}.
+However, if it occurs in a prototypical context and with prototypical lexical material, reading times drop ($\beta_{\text{ConstructionPGCa:Modelprediction}}=-0.132$).
+This can be seen in the downward slope of curve in the right panel of Figure~\ref{fig:spr:effects}.
 This fits into the general picture in as much as the construction with the lower frequency might be developing towards a more sharply defined prototype.%

### 2017-04-02 R/2_attraction.R

@@ -29,7 +29,11 @@ get_freq <- function(data, lem) {
 
 
 # Generate full table.
-cx.kind <- cbind(cx.kind.lem, lapply(cx.kind.lem, function(lem){get_freq(nn.kind,lem)}), lapply(cx.kind.lem, function(lem){get_freq(ndn.kind,lem)}))
+cx.kind <- data.frame(
+  Lemma = as.character(cx.kind.lem),
+  Nn    = unlist(lapply(cx.kind.lem, function(lem){get_freq(nn.kind,lem)})),
+  Ndn   = unlist(lapply(cx.kind.lem, function(lem){get_freq(ndn.kind,lem)}))
+)
 
 
 # A rather simple "attraction" function, i.e., log ratio.
@@ -37,27 +41,43 @@ attraction <- function(cs) {
   log((unlist(cs)[2] + 1) / (unlist(cs)[1] + 1))
 }
 
+collottraction <- function(cs, alls) {
+  .cs <- as.numeric(cs)
+  .alls <- as.numeric(alls)
+  .mat<- matrix(c(.cs[2], .cs[1], .alls[2]-.cs[2], .alls[1]-.cs[1]), nrow = 2)
+  .p <- fisher.test(.mat)$p.value
+  log(.p, 10)
+}
+
+
 
 # Generate attraction quotients for full table.
+kind.all.nn  <- sum(unlist((cx.kind$Nn)))
+kind.all.ndn <- sum(unlist((cx.kind$Ndn)))
 cx.kind <- as.data.frame(cx.kind)
 cx.kind$Pv <- unlist(apply(cx.kind[,2:3], 1, function(cs){attraction(cs)} ))
+cx.kind$Collo <- unlist(apply(cx.kind[,2:3], 1, function(cs){collottraction(cs, c(kind.all.nn, kind.all.ndn))} ))
 
 
 # Measure nouns are simpler because the tables were imported done & ready.
 cx.measure <- measurenouns_attracfreq
+measure.all.nn  <- sum(unlist((cx.measure$Nn)))
+measure.all.ndn <- sum(unlist((cx.measure$Ndn)))
 cx.measure$Pv <- unlist(apply(cx.measure[,2:3], 1, function(cs){attraction(cs)} ))
+cx.measure$Collo <- unlist(apply(cx.measure[,2:3], 1, function(cs){collottraction(cs, c(measure.all.nn, measure.all.ndn))} ))
 
-
[…]

### 2017-04-02 R/3_glmm.R

@@ -20,10 +20,12 @@ measure.glmm <- glmer(Construction~1
                  +Measurenumber
 
                  +Kindattraction
+                 #+Kindcollo
                  +Kindfreq
                  +Kindgender
 
                  +Measureattraction
+                 #+Measurecollo
                  +Measureclass
                  +Measurefreq
 
@@ -66,7 +68,7 @@ modelcomp <- lmer.modelcomparison(model = measure.glmm, regressors = bootcomp.re
 
 # Output.
 if (save.persistent) sink(paste(out.dir, "glmm.txt", sep=""))
-cat("\n\n\nMASCULINE\n\n")
+cat("\n\n\nGLMM summary\n\n")
 print(summary(measure.glmm), correlation=F)
 cat("\n\n")
 print(measure.glmm.wald)

### 2017-04-03 Latex/article.tex
new stats, yet againmake all

@@
 
-\subsection{Theoretical assessment}
+\subsection{Factors controlling the alternation}
 \label{sec:analyses}
@@
 Thus, the adjectives in the \NACa\ and the \PGCa\ have properties of adjectives as well as determiners.
-They are lexical adjectives and function as attributive modifiers.
+On the one hand, they are lexical adjectives and function as attributive modifiers.
 On the other hand, they are inflected like determiners, and they are the leftmost element in the NP, which is typical of determiners.
@@
 
-Put differently, the \NACa\ is a \NACb in disguise, and the \PGCa\ is a \PGCd in disguise.
-In a probabilistic framework, the crucial question is what controls speakers' decisions to use a a \NACb in disguise or a \PGCd in disguise.
-If the \PGCd\ instantiates the PGC prototype and the \NACb\ the NAC prototype, the choice between \PGCa\ and \NACa\ is expected to be influenced by the degree of similarity to either of these prototypes.
-I will argue below that the prototypes differ mainly in that they attract different (classes of) measure nouns more or less strongly, mostly because of degrees of grammaticalisation.
-However, if lemmas are attracted with different strengths by the two prototypes, then we can expect a baseline attraction effect.
-Nouns which occur more often in the \PGCd\ should favour the \PGCa, and nouns which occur more often in the \NACb\ should favour the \NACa.%
-\footnote{Such associations between constructions and words or lemmas are the main object of study in approaches such as Collexeme Analysis \citep{GriesStefanowitsch2004}.}
-Put differently, a noun's relative frequencies in the prototypical non-alternating constructions should be indicative of same noun's preference to appear in the two alternating variants.
-This baseline attraction effect will be quantified in the corpus study Section~\ref{sec:corpusstudies}.
+This morpho-syntactic ambiguity means that the \NACa\ is in fact a \NACb in disguise, and the \PGCa\ is a \PGCd in disguise.
+This should ideally be reflected in the following basic selection effect:
+It is known from frameworks like Collexeme Analysis \citep{GriesStefanowitsch2004} that lemmas are attracted with different strengths by competing constructions.
+If the \NACa\ is highly similar to the \NACb\ and the \PGCa\ is highly similar to the \PGCd, the probability with which individual noun lemmas appear in the alternating constructions should be predictable at least partly from the relative frequency with which the same lemmas are used in the non-alternating constructions.
+Nouns which occur proportionally more often in the \PGCd\ should favour the \PGCa, and nouns which occur proportionally more often in the \NACb\ should favour the \NACa.
+This basic attraction effect will be quantified in the corpus study Section~\ref{sec:corpusstudies}.
 
@@
 
-Turning to grammaticalisation, it is often assumed that pseudo-partitives arise as a form of grammaticalised partitives (\eg\ \citealp[536--539]{Koptjevskaja2001} for Finnish and Estonian, \citealp[559]{Koptjevskaja2001} for European languages per se).
+In a probabilistic framework like Prototype Theory, the crucial question is, however, what controls speakers' decisions to use either variant beyond simple lemma preference effects. 
+I argue that the NAC and the PGC are prototypes associated with different degrees of the grammaticalisation of the measure noun, morpho-syntactic properties, and register effects.
+The degree of similarity of a given instance to either of the two prototypes makes speakers chose the \NACa\ or the \PGCa.
+I first turn to grammaticalisation.
+It is often assumed that pseudo-partitives arise as a form of grammaticalised partitives (\eg\ \citealp[536--539]{Koptjevskaja2001} for Finnish and Estonian, \citealp[559]{Koptjevskaja2001} for European languages per se).
 %
@@
 %
-The alternating constructions in German are both clearly pseudo-partitives, but the grammaticalisation paths uncovered by \cite[esp.\ 526--530]{Koptjevskaja2001} are relevant for an analysis of the alternation.
+The alternating constructions in German are both clearly pseudo-partitives, but the grammaticalisation paths uncovered by \cite[esp.\ 526--530]{Koptjevskaja2001} are still relevant in the case at hand.
 The grammaticalisation path can start out (in some languages) with constructions involving two referential nouns (not necessarily forming a single and contiguous NP) and a \textit{separative} meaning as in \textit{(cut) two slices from the cake} \citep[535]{Koptjevskaja2001}.
@@
 The pseudo-partitive stage finally instantiates a \textit{quantity-of} relation, potentially even leading to fully grammaticalised quantifiers such as \textit{a lot}.
-In German, the PGC is clearly the older variant \citep{Zimmer2015}, still having an overt \textit{quantity-of} marker in the form of the genitive suffix.
-Also, it still has the potential to form a true partitive (if the kind noun is definite).
+In German, the PGC is clearly the older variant \citep{Zimmer2015}.
+It still has the potential to form a true partitive (if the kind noun is definite).
 Conversely, the NAC completely lacks this ability to form true partitives, and it forms a more restrictive environment with less combinatory and semantic flexibility.
-Hence, we can expect more strongly grammaticalised measure nouns to prefer the NAC.
-For example, highly grammaticalised non-referential nouns like \textit{Gramm} `gram' and \textit{Meter} `metre' should occur proportionally more often in the \NACa\ than in the \PGCa.%
-\footnote{In case studies like \cite{Brems2003,DeclerckBrems2016} syntactic evidence for grammaticalisation plays a prominent role.
-In English -- \cite{Brems2003} argues based on a view from \cite{Langacker1991} -- strongly lexicalised and thus semantically bleached measure nouns turn from nouns into quantifiers, leading to a syntactic reanalysis of the NP, wherein the kind noun becomes the head.
-The argument is supported by convincing syntactic evidence, for example subject-verb agreement.
-As argued above, in German MNPs the measure noun always remains the syntactic head, and there is no similarly handy criterion to diagnose a high degree of grammaticalisation.}
+Hence, we can expect the \NACa\ to be a prototypical hosting construction for more strongly grammaticalised measure nouns.
+For example, highly grammaticalised non-referential nouns like \textit{Gramm} `gram' and \textit{Meter} `metre' should occur proportionally more often in the \NACa\ than in the \PGCa.
+% \footnote{In case studies like \cite{Brems2003,DeclerckBrems2016} syntactic evidence for grammaticalisation plays a prominent role.
+% In English -- \cite{Brems2003} argues based on a view from \cite{Langacker1991} -- strongly lexicalised and thus semantically bleached measure nouns turn from nouns into quantifiers, leading to a syntactic reanalysis of the NP, wherein the kind noun becomes the head.
+% The argument is supported by convincing syntactic evidence, for example subject-verb agreement.
+% As argued above, in German MNPs the measure noun always remains the syntactic head, and there is no similarly handy criterion to diagnose a high degree of grammaticalisation.}
 
@@
 For illustration, (\ref{ex:determiners}) shows the expected variants under this hypothesis.
-In (\ref{ex:determiners:nac}), the measure noun is modified by a cardinal \textit{drei} `three', and hence the \NACa\ (kind NP shows case agreement) is preferred.
-In (\ref{ex:determiners:pgc}), the measure noun is modified by a non-cardinal determiner \textit{einige} `some', and the \PGCa\ (kind NP is in the genitive) is preferred.
+In (\ref{ex:determiners:nac}), the measure noun is modified by a cardinal \textit{drei} `three', and hence the \NACa\ (the kind NP shows case agreement) is preferred.
+In (\ref{ex:determiners:pgc}), the measure noun is modified by a non-cardinal determiner \textit{einige} `some', and the \PGCa\ (the kind NP is in the genitive) is preferred.
 
@@
 
-Finally, it has often been iterated that the \PGCa\ is more typical of higher registers or even exclusive to written language (see \citealp[320--323]{Hentschel1993}, where she also refers to descriptive and normative grammars).
+Finally, it has often been iterated that the \PGCa\ is more typical of higher registers or even exclusive to written language (see \citealp[320--323]{Hentschel1993}).
 This is not surprising in as much as the genitive -- an intrinsic part of the PGC -- is generally underrepresented in colloquial vernacular variants of German as a result of a diachronic process wherein many (but not all) uses of the genitive are replaced by other cases or periphrastic constructions \citep{FleischerSchallert2011}.
@@
 \begin{enumerate}[i.]
-  \item The relative frequencies with which measure noun lemmas and kind noun lemmas appear in the prototypical (non-alternating) \PGCd\ and \NACb\ are positively correlated with the relative frequency the alternating \PGCa\ and the \NACa\ are chosen.
-  \item (Classes of) more strongly grammaticalised measure nouns favour the NAC.
-  \item Measure nouns quantified by cardinals favour the PGC variant.
-  \item The NAC is associated more strongly with higher registers.
+  \item The relative frequencies with which measure noun lemmas and kind noun lemmas appear in the prototypical (non-alternating) \PGCd\ and \NACb\ are predictive of the probability with which the alternating \PGCa\ and the \NACa\ are chosen.
+  \item (Classes of) more strongly grammaticalised measure nouns favour the \NACa.
+  \item Measure nouns modified by cardinals favour the \NACa.
+  \item The \NACa\ is associated with higher registers.
 \end{enumerate}
@@
                    & Genitives                        & numeric &                           \\
-    Sentence       & \textbf{Construction (response)} & factor  & NACa, PGCa                   \\
+    Sentence       & Cardinal                         & factor  & Yes, No                   \\
+                   & \textbf{Construction (response)} & factor  & NACa, PGCa                \\
                    & Measurecase                      & factor  & Nom, Acc, Dat             \\
-                   & Measurenumber                    & factor  & Pl, Sg                       \\
-                   & Leftcontext                      & factor  & C, D, O, A, P                \\
     Kind lemma     & Kindattraction                   & numeric &                           \\
@@
     Measure lemma  & Measureattraction                & numeric &                           \\
-                   & Measureclass                     & factor  & Physical, Container, Amount, \\
-                   &                                  &         & Portion, Rest                \\
+                   & Measureclass                     & factor  & Physical, Container,      \\
+                   &                                  &         & Amount, Portion, Rest     \\
                    & Measurefreq                      & numeric &                           \\
@@
 They were calculated from the auxiliary samples described at the end of Section~\ref{sec:gettingdata} as a log-transformed quotient.
-The higher the value, the more often the noun occurs in the \PGCa (proportionally).
+The higher the value, the more often the noun occurs in the \PGCa (proportionally).%
+\footnote{One could argue that some more advanced measure of attraction strength should be used, as is done in Collostructional (or Collexeme) Analysis \citep{GriesStefanowitsch2004}.
+Two main points speak against such an approach.
+First, the attraction values will be used as regressors in a GLMM, and the values resulting from collostructional approaches, \ie\ logarithmised Fisher p values, have a very unfavourable distribution for such a use.
+They cluster around $-\infty$ and $0$, especially when (as in this case) cells contain $0$ count values.
+This cannot be remedied on principled grounds, since, for example, a z transformation is inadequate for such a strongly bimodal distribution (even if we clamp $-\infty$ to a very low numeric value).
+Second, their main use in collostructional analyses is to \textit{sort} a list of collexemes and interpret the order.
+Their concrete numerical value thus does not have a solid cognitive interpretation.
+For the present study, it suffices to estimate the relative frequency with which speakers have encountered a specific lemma in the \NACa\ and the \PGCa.
+}
 \textit{Kindfreq} and \textit{Measurefreq} are the logarithm-transformed frequencies per 1,000,000 words of each lemma, extracted from the frequency lists distributed by the DECOW corpus creators on their web page.
@@
 
-The variable \textit{Leftcontext} encodes the coarse-grained part-of-speech of the word preceding the measure noun.
-The level names stand for (again with their frequency in the final sample) \textit{cardinal} ($f=1,939$), \textit{determiner} ($f=1,971$), \textit{other} ($f=221$), \textit{adjective} ($f=845$), and \textit{preposition} ($f=87$).
-The purpose of this variable is mainly to test whether cardinals really favour the \PGCa\ as hypothesised in Section~\ref{sec:analyses}.
-Along the same lines, \textit{Measurenumber} encodes the grammatical case and number of the measure noun.
+The variable \textit{Cardinal} encodes whether the measure noun is modified by a cardinal ($f=1,939$) or not ($f=3,124$).
+The purpose of this variable is to test whether cardinals really favour the \NACa\ as hypothesised in Section~\ref{sec:analyses}.
 
@@
   \includegraphics[width=0.85\textwidth]{../R/output/corpus_fixeffs_mle+mcmc}
-  \caption{Coefficients (MLE and MCMC) with 95\% confidence intervals (for details see text); the intercept (\textit{Leftcontext=C}, \textit{Measurecase=Nom}, \textit{Measurenumber=Sg}, \textit{Kindgender=Masc}, \textit{Measureclass=Physical}; 0 for all numeric z-transformed regressors) is -4.328 (MLE) and -4.441 (MCMC)}
+  \caption{Coefficients (MLE and MCMC) with 95\% confidence intervals (for details see text); the intercept (\textit{Cardinal=Yes}, \textit{Measurecase=Nom}, \textit{Kindgender=Masc}, \textit{Measureclass=Physical}; 0 for all numeric z-transformed regressors) is -4.328 (MLE) and -4.441 (MCMC)}
   \label{fig:fixeffs}
@@
               &                   &       &           && \multicolumn{1}{l}{MLE} & \multicolumn{1}{l}{MCMC} && \multicolumn{1}{l}{MLE} & \multicolumn{1}{l}{MCMC} && \multicolumn{1}{l}{MLE} & \multicolumn{1}{l}{MCMC} && \multicolumn{1}{l}{MLE} & \multicolumn{1}{l}{MCMC} \\\midrule
-
-    First     & Badness           & 0.001 & (numeric) && -0.165 & -0.169 && -0.264 & -0.264 && -0.074 & -0.075 &&      * &      * \\
-              & Genitives         & 0.001 & (numeric) && -0.630 & -0.645 && -0.710 & -0.739 && -0.523 & -0.552 &&      * &      * \\
-              & Leftcontext       & 0.001 & D         &&  1.178 &  1.173 &&  0.751 &  0.784 &&  1.548 &  1.546 &&      * &      * \\
-              &                   &       & O         &&  1.561 &  1.588 &&  1.093 &  1.179 &&  1.919 &  2.007 &&      * &      * \\
-              &                   &       & A         &&  2.300 &  2.340 &&  1.872 &  1.959 &&  2.644 &  2.726 &&      * &      * \\
-              &                   &       & P         &&  2.353 &  2.371 &&  1.693 &  1.766 &&  2.900 &  2.978 &&      * &      * \\
-              & Measurecase       & 0.001 & Acc       &&  0.065 &  0.070 && -0.130 & -0.119 &&  0.268 &  0.269 &&        &        \\
-              &                   &       & Dat       &&  0.707 &  0.735 &&  0.413 &  0.480 &&  0.959 &  1.002 &&      * &      * \\
-	      & Measurenumber     & 0.001 & Pl        &&  0.814 &  0.825 &&  0.534 &  0.574 &&  1.080 &  1.079 &&      * &      * \\[0.5\baselineskip]
-
-    Second    & Kindattraction    & 0.033 & (numeric) &&  0.204 &  0.216 &&  0.030 &  0.014 &&  0.366 &  0.412 &&      * &      * \\
-    (Kind)    & Kindfreq          & 0.066 & (numeric) &&  0.161 &  0.173 && -0.008 &  0.009 &&  0.333 &  0.350 &&        &      * \\
-              & Kindgender        & 0.001 & Neut      && -0.007 & -0.022 && -0.407 & -0.424 &&  0.408 &  0.379 &&        &        \\
-	      &                   &       & Fem       &&  1.357 &  1.379 &&  0.831 &  0.872 &&  1.808 &  1.897 &&      * &      * \\[0.5\baselineskip]
-
-    Second    & Measureattraction & 0.015 & (numeric) &&  0.247 &  0.275 &&  0.058 &  0.078 &&  0.413 &  0.476 &&      * &      * \\
-    (Measure) & Measureclass      & 0.001 & Rest      &&  0.564 &  0.488 && -0.140 & -0.309 &&  1.197 &  1.236 &&        &        \\
-              &                   &       & Container &&  0.609 &  0.618 &&  0.050 &  0.018 &&  1.145 &  1.215 &&      * &      * \\
-              &                   &       & Amount    &&  0.927 &  0.959 &&  0.184 &  0.189 &&  1.566 &  1.718 &&      * &      * \\
-              &                   &       & Portion   &&  1.472 &  1.503 &&  0.916 &  0.956 &&  1.938 &  2.064 &&      * &      * \\
-              & Measurefreq       & 0.003 & (numeric) && -0.257 & -0.264 && -0.413 & -0.435 && -0.098 & -0.099 &&      * &      * \\
+       First     & Badness           &  0.002 &           && -0.152 & -0.155 && -0.247 & -0.247 && -0.061 & -0.065 && * & * \\
+                 & Cardinal          &  0.001 & No        &&  1.189 &  1.222 &&  0.862 &  0.927 &&  1.466 &  1.496 && * & * \\
+                 & Genitives         &  0.001 &           && -0.693 & -0.711 && -0.768 & -0.801 && -0.592 & -0.616 && * & * \\
+                 & Measurecase       &  0.001 & Acc       &&  0.030 &  0.031 && -0.150 & -0.159 &&  0.212 &  0.222 &&   &   \\
+                 &                   &        & Dat       &&  0.705 &  0.729 &&  0.455 &  0.465 &&  0.944 &  0.995 && * & * \\[0.5\baselineskip]
+
+       Second    & Kindattraction    &  0.020 &           &&  0.225 &  0.244 &&  0.049 &  0.056 &&  0.393 &  0.422 && * & * \\
+       (Kind)    & Kindfreq          &  0.095 &           &&  0.146 &  0.164 && -0.023 & -0.016 &&  0.301 &  0.341 &&   &   \\
+                 & Kindgender        &  0.001 & Neut      &&  0.021 &  0.013 && -0.367 & -0.409 &&  0.392 &  0.435 &&   &   \\
+                 &                   &        & Fem       &&  1.269 &  1.289 &&  0.800 &  0.788 &&  1.709 &  1.783 && * & * \\[0.5\baselineskip]
+
+       Second    & Measureattraction &  0.001 &           &&  0.282 &  0.299 &&  0.106 &  0.102 &&  0.447 &  0.515 && * & * \\
+       (Measure) & Measureclass      &  0.001 & Container &&  0.252 &  0.257 && -0.265 & -0.303 &&  0.788 &  0.813 &&   &   \\
+                 &                   &        & Rest      &&  0.421 &  0.379 && -0.209 & -0.378 &&  1.063 &  1.091 &&   &   \\
+                 &                   &        & Amount    &&  0.831 &  0.889 &&  0.215 &  0.220 &&  1.432 &  1.569 && * & * \\
+                 &                   &        & Portion   &&  1.217 &  1.253 &&  0.675 &  0.689 &&  1.684 &  1.840 && * & * \\
+                 & Measurefreq       &  0.005 &           && -0.231 & -0.232 && -0.363 & -0.395 && -0.079 & -0.073 && * & * \\
+
   \end{tabular}
   }
-  \caption{Coefficient table comparing Maximum Likelihood Estimation (MLE, with 95\% bootstrap confidence interval) and `Bayesian' Markov-Chain Monte Carlo estimation (MCMC); for details see text}
+  \caption{Coefficient table comparing Maximum Likelihood Estimation (MLE, with 95\% bootstrap confidence interval) and `Bayesian' Markov-Chain Monte Carlo estimation (MCMC); the intercept (\textit{Cardinal=Yes}, \textit{Measurecase=Nom}, \textit{Kindgender=Masc}, \textit{Measureclass=Physical}; 0 for all numeric z-transformed regressors) is -3.548 (MLE) and -3.700 (MCMC)}
   \label{tab:bigtable}
@@
 Standard diagnostics for the Maximum Likelihood Estimation (MLE) show that the model quality is quite high.
-Generalised variance inflation factors for the regressors were calculated to check for multicollinearity \citep{FoxMonette1992,ZuurEa2010}, and none of the corrected $\text{GVIF}^{1/2\text{df}}$ was higher than $2$.
-Nakagawa \& Schielzeth's pseudo-coefficient of determination is $R_m^2=0.423$ and $R^2_c=0.515$ (see \citealp{Gries2015} for a linguist-friendly introduction to these $R^2$ measures, or else \citealp{NakagawaSchielzeth2013}).
-The rate of correct predictions is $0.855$, which means a proportional reduction of error of $\lambda=0.354$.
-The measure lemma intercepts have a standard deviation of $\sigma_{\text{Measurelemma}}=0.623$, the kind lemma intercept $\sigma_{\text{Kindlemma}}=0.488$.
+Generalised variance inflation factors for the regressors were calculated to check for multicollinearity \citep{FoxMonette1992,ZuurEa2010}, and none of the corrected $\text{GVIF}^{1/2\text{df}}$ was higher than $1.6$.
+Nakagawa \& Schielzeth's pseudo-coefficient of determination is $R_m^2=0.409$ and $R^2_c=0.495$ (see \citealp{Gries2015} for a linguist-friendly introduction to these $R^2$ measures, or else \citealp{NakagawaSchielzeth2013}).
+The rate of correct predictions is $0.843$, which means a proportional reduction of error of $\lambda=0.297$.
+The measure lemma intercepts have a standard deviation of $\sigma_{\text{Measurelemma}}=0.448$, the kind lemma intercept $\sigma_{\text{Kindlemma}}=0.604$.
 
 The coefficient estimates are specified in Table~\ref{tab:bigtable} for each regressor (or regressor level) in the columns labelled \textit{Coefficient}.
-For a robust quantification of the precision of the estimation, I ran a parametric bootstrap (using the \textit{confint.merMod} function from \textit{lme4}) with 1,000 replications, and using the percentile method for the calculation of the intervals.
+For a robust quantification of the precision of the estimation, I ran a parametric bootstrap (using the \textit{confint.merMod} function from \textit{lme4}) with $1,000$ replications, and using the percentile method for the calculation of the intervals.
 The resulting 95\% bootstrap confidence intervals are reported in Table~\ref{tab:bigtable} in the columns labelled \textit{CI low} and \textit{CI high} (= upper and lower 2.5th percentiles).
@@
 Furthermore, for each regressor, a $p$ value was obtained by dropping the regressor from the full model, re-estimating the nested model and comparing it to the full model.
-Instead of inexact Wald approximations or Likelihood Ratio Tests, I used a drop-in bootstrap replacement for the Likelihood Ratio Test from the function \textit{PBmodcomp} from the \textit{pbkrtest} package \citep{HalekohHojsgaard2014}.
+Instead of inexact Wald approximations or Likelihood Ratio Tests, I used a drop-in bootstrap replacement for the Likelihood Ratio Test from the function \textit{PBmodcomp} from the \textit{pbkrtest} package \citep{HalekohHojsgaard2014}, hence I call the value $p_{\text{PB}}$.
 These bootstrapped $p$ values are given in the columns labelled p\Sub{PB} in Table~\ref{tab:bigtable}.
-All regressors have low p\Sub{PB} values of magnitude $10^{-2}$ or $10^{-3}$.
-Only \textit{Kindfreq} ($\mpPB=0.066$) can be seen as slightly too high to be convincing (`non-significant').
+Only \textit{Kindfreq} ($\mpPB=0.095$) can be seen as slightly too high to be convincing (`non-significant').
 
 Finally, I show that Bayesian methods in the form of Markov-Chain Monte Carlo estimation do not necessarily (and in this case predictably) lead to different results.
-Both the MN model and the F model were re-estimated using the \textit{stan\_glmer} function from the \textit{rstanarm} package, which provides an \textit{lme4}-compatible syntax for estimating common model types with the \textit{stan} software \citep{CarpenterEa2017}.
+The model was re-estimated using the \textit{stan\_glmer} function from the \textit{rstanarm} package, which provides an \textit{lme4}-compatible syntax for estimating common model types with the \textit{stan} software \citep{CarpenterEa2017}.
 The algorithm was run with $4$ chains and $1,000$ iterations, and instead of guessing prior distributions or using implausible uniform prior distributions \cite[251--252]{Levshina2016}, I used plausible default priors.
@@
 Figure~\ref{fig:fixeffs} illustrates this by also showing both estimates.
-If we were interested in significance at $\alpha=0.05$, then the decision based on the confidence intervals differs for \textit{Kindfreq} ($\mpPB=0.066$, MLE lower bound is $-0.008$ and MCMC lower bound is $0.009$).
-With MCMC, the regressor is `not significant' at the standard $\alpha$ level, but it is with MLE.
-This might only go to show how absurd strict significance testing is in complex modelling situations:
-both methods essentially estimate the same sign and magnitude for the coefficient (MLE $0.161$, MCMC $0.173$) and do not differ substantially with regard to the precision of the estimate in the form of the width of the confidence interval (both intervals have a width of exactly $0.341$).
 Since both estimators converge, I only interpret the MLE model in the next section.
@@
 
-% PLURAL & CARDINALS
+% CARDINALS
 
@@
   \centering
-  \includegraphics[width=0.75\textwidth]{../R/output/corpus_Leftcontext}
-  \caption{Effect plot for the regressor \textit{Leftcontext}}
+  \includegraphics[width=0.75\textwidth]{../R/output/corpus_Cardinal}
+  \caption{Effect plot for the regressor \textit{Cardinal}}
   \label{fig:eff:leftcontext}
@@
 
-I now turn to the predicted effect of plural measure nouns and cardinals as quantifiers.
-Plurality has a clear effect in the expected direction (plural favours the \PGCa) ($\beta_{\text{MeasurenumberPl}}=0.814$, $\mpPB=0.001$ for \textit{Measurenumber}).
-Figure~\ref{fig:eff:leftcontext} shows that the left context of the measure noun influences the choice of the variant ($\mpPB=0.001$).
-Cardinals have a strong tendency to co-occur with the \NACa.
-Both effects were predicted in Section~\ref{sec:analyses}.
+I now turn to the predicted effect of cardinals as modifiers of the measure noun.
+Figure~\ref{fig:eff:leftcontext} shows that cardinals indeed influence the choice of the variant ($\mpPB=0.001$), and that cardinals have a strong tendency to co-occur with the \NACa.
+This effect was predicted in Section~\ref{sec:analyses}.
 
@@
 
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.75\textwidth]{../R/output/fc_proportions}
-\caption{Participants' responses in the forced-choice task plotted against the predictions from the corpus-based model}
-\label{fig:afc:proportions}
-\end{figure}
-
 The pairs of stimuli were the sentence containing the preferred construction (according to the corpus GLMM) and a modified version containing the dispreferred construction.
@@
 The experiment was conducted using \textit{PsychoPy} \citep{Peirce2007}.
-Figure \ref{fig:afc:proportions} shows the relationship between the probabilities calculated from the corpus-based model and the choices made by participants.
-The close correlation between the probabilities derived from usage data and the choices in the experiment are evident.
 
@@
 
-A certain amount of the variance can be accounted for by idiosyncrasies of single sentences ($\sigma_{\text{Sentence}}=0.861$, $\sigma^{\text{Modelprediction}}_{\text{Sentence}}=0.250$, $16$ levels).%
+A certain amount of the variance can be accounted for by idiosyncrasies of single sentences ($\sigma_{\text{Sentence}}=1.785$, $\sigma^{\text{Modelprediction}}_{\text{Sentence}}=5.996$, $16$ levels).%
 \footnote{I use $\sigma^f_r$ to denote the standard deviation of the of the random intercepts for the fixed effect $f$ varying by random effect $r$.}
-Also, among participants, there are clearly different preferences ($\sigma_{\text{Participant}}=0.530$, $\sigma^{\text{Modelprediction}}_{\text{Participant}}=0.076$, $24$ levels).
+Also, among participants, there are clearly different preferences ($\sigma_{\text{Participant}}=0.781$, $\sigma^{\text{Modelprediction}}_{\text{Participant}}=0.484$, $24$ levels).
 On the extreme sides, one participant chose the \PGCa\ in $13$ of $16$ cases, and two participants only chose it in $5$ of $16$ cases.
-The regressor \textit{Modelprediction} achieves $\mpPB=0.002$ ($1,000$ replications) and is estimated at $4.500$ relative to an intercept of $-1.520$.
-The confidence interval from a parametric bootstrap ($1,000$ replications, percentile method) for the regressor is acceptable with a lower bound of $2.738$ and an upper bound of $6.390$.
-Nakagawa \& Schielzeth's pseudo-coefficients of determination are $R^2_{m}=0.267$ and $R^2_{c}=0.435$, which means that roughly 27\% of the variance in the data can be explained by considering only the predictions from the corpus-based GLMM.
+The regressor \textit{Modelprediction} achieves $\mpPB=0.007$ ($1,000$ replications) and is estimated at $5.408$ relative to an intercept of $-1.304$.
+The confidence interval from a parametric bootstrap ($1,000$ replications, percentile method) for the regressor is acceptable but slightly large with a lower bound of $1.626$ and an upper bound of $8.397$.
+Nakagawa \& Schielzeth's pseudo-coefficients of determination are $R^2_{m}=0.227$ and $R^2_{c}=0.561$, which means that over 22\% of the variance in the data can be explained by considering only the predictions from the corpus-based GLMM.
 The effect display for the single fixed regressor \textit{Modelprediction} is given in Figure \ref{fig:afc:effects}.
-It confirms the effect descriptively shown in Figure~\ref{fig:afc:proportions}:
-the higher the probability of the \PGCa\ predicted from usage-data, the more often participants chose the \PGCa\ variant in the forced-choice task.
+The result is very clear.
+The higher the probability of the \PGCa\ predicted from usage-data, the more often participants chose the \PGCa\ variant in the forced-choice task.
 In summary, the forced choice experiment clearly succeeded in validating the results from the corpus study in as much as the preferences extracted from usage data correspond to native speakers' choices.
@@
     Regressor & \multicolumn{1}{r}{Coefficient} & \multicolumn{1}{r}{CI low} & \multicolumn{1}{r}{CI high} & \multicolumn{1}{r}{0 in CI} \\ \midrule
-    ConstructionPGCa                 &  0.062 &  0.023 &  0.103 &  *  \\ 
-    Modelprediction                  &  0.025 & -0.065 &  0.122 &     \\ 
+    ConstructionPGCa                 &  0.054 &  0.012 &  0.095 &  *  \\ 
+    Modelprediction                  & -0.006 & -0.113 &  0.110 &     \\ 
     Position                         & -0.005 & -0.005 &  0.004 &     \\ 
-    ConstructionPGCa:Modelprediction & -0.132 & -0.229 & -0.041 &  *  \\ 
+    ConstructionPGCa:Modelprediction & -0.125 & -0.234 & -0.023 &  *  \\ 
   \end{tabular}
-  \caption{Fixed effect coefficient table for the LMM used to analyse the self-paced reading experiment; the intercept is 0.819}
+  \caption{Fixed effect coefficient table for the LMM used to analyse the self-paced reading experiment; the intercept is 0.829}
   \label{tab:exp:spr}
@@
 Table~\ref{tab:exp:spr} shows the coefficient estimates with a 95\% parametric bootstrap confidence interval ($1,000$ replications, percentile method).
-The standard deviation of the participant intercepts is $\sigma_{\text{Participant}}=0.079$ and of the item intercepts $\sigma_{\text{Item}}=0.039$.
-Comparing the full model to a model without the main regressor \textit{Modelprediction} (and consequently also without the interaction with \textit{Construction}) in a PB test gives $\mpPB=0.018$.
-Nakagawa and Schielzeth's pseudo-determination coefficients are $R^2_m=0.238$ and $R^2_c=0.347$.
+The standard deviation of the participant intercepts is $\sigma_{\text{Participant}}=0.079$ and of the item intercepts $\sigma_{\text{Item}}=0.037$.
+Comparing the full model to a model without the main regressor \textit{Modelprediction} (and consequently also without the interaction with \textit{Construction}) in a PB test gives $\mpPB=0.036$.
+Nakagawa and Schielzeth's pseudo-determination coefficients are $R^2_m=0.237$ and $R^2_c=0.346$.
 
@@
 There is a clearer effect in the sentences with \PGCa, which is also confirmed by the `significant' results from the bootstrapped confidence intervals (see Table~\ref{tab:exp:spr}) and from the PB test reported above.
-The \PGCa\ brings about an increased reading time ($\beta_{\text{ConstructionPGCa}}=0.062$), which is plausible because it is the much rarer construction (see Section~\ref{sec:corpusstudies}).
-However, if it occurs in a prototypical context and with prototypical lexical material, reading times drop ($\beta_{\text{ConstructionPGCa:Modelprediction}}=-0.132$).
+The \PGCa\ brings about an increased reading time ($\beta_{\text{ConstructionPGCa}}=0.054$), which is plausible because it is the much rarer construction (see Section~\ref{sec:corpusstudies}).
+However, if it occurs in a prototypical context and with prototypical lexical material, reading times drop ($\beta_{\text{ConstructionPGCa:Modelprediction}}=-0.125$).
 This can be seen in the downward slope of curve in the right panel of Figure~\ref{fig:spr:effects}.
@@
 
-\begin{acknowledgement}
-  I want to thank (in alphabetical order) Felix Bildhauer, Samuel Reichert, and Ulrike Sayatz for discussions on relevant aspects of this paper.
-  I also thank Ulrike Sayatz for helping to recruit the participants for the experiments.
-  I am grateful to Kim Maser for her work on the annotation of the concordances.
-  The research presented here was made possible in part through funding from the \textit{Deutsche Forschungsgemeinschaft} (DFG, personal grant SCHA1916/1-1).
-\end{acknowledgement}
+% \begin{acknowledgement}
+%   I want to thank (in alphabetical order) Felix Bildhauer, Samuel Reichert, and Ulrike Sayatz for discussions on relevant aspects of this paper.
+%   I also thank Ulrike Sayatz for helping to recruit the participants for the experiments.
+%   I am grateful to Kim Maser for her work on the annotation of the concordances.
+%   The research presented here was made possible in part through funding from the \textit{Deutsche Forschungsgemeinschaft} (DFG, personal grant SCHA1916/1-1).
+% \end{acknowledgement}
 

### 2017-04-03 R/0_wrapper.R

@@ -25,7 +25,7 @@ iter                   <- 1000             # MCMC. Set to 1000 for production.
 
 
 # Delete old output.
-unlink(paste(out.dir, '*', sep=""))
+if (save.persistent) unlink(paste(out.dir, '*', sep=""))
 out.dir <- paste(out.dir, "corpus_", sep="")
 
 

### 2017-04-03 R/1_load.R

@@ -16,9 +16,9 @@ measure <- measure[,c("Measurecase",
           "Kindcase",
           "Kindfreq",
           "Measurelemma",
-          "Measurenumber",
           "Measureclass",
           "Measurefreq",
+          "Measurenumber",
           "Minus1pos",
           "Genitives",
           "Badness",
@@ -47,6 +47,8 @@ measure$Leftcontext <- ifelse(measure$Leftcontext == "Card", "C", measure$Leftco
 measure$Leftcontext <- ifelse(measure$Leftcontext %in% c("D", "P", "A", "C"), measure$Leftcontext, "O")
 measure$Leftcontext <- as.factor(measure$Leftcontext)
 
+# Even simpler.
+measure$Cardinal <- factor(ifelse(measure$Leftcontext == "C", "Yes", "No"), levels = c("Yes", "No"))
 
 # Simplify "Measureclass".
 measure$Measureclass <- factor(ifelse(as.character(measure$Measureclass) %in% c("Natural", "Trace", "Transport", "Object", "Unknown", "Unit", "Currency", "Layer"), "Rest", as.character(measure$Measureclass)))
@@ -77,8 +79,8 @@ fax  <- c("Measurecase",
           "Kindgender",
           "Kindcase",
           "Measurelemma",
-          "Measurenumber",
           "Measureclass",
+          "Measurenumber",
           "Leftcontext")
 
 for (fak in fax) {
@@ -88,7 +90,7 @@ for (fak in fax) {
 
 # Hm, different level ordering looks odd in plot, so make Leftcontext same order.
 measure$Leftcontext  <- factor(measure$Leftcontext, levels = c("C", "D", "O", "A", "P"))
-measure$Measureclass <- factor(measure$Measureclass, levels = c("Physical", "Rest", "Container", "Amount", "Portion" ))
+measure$Measureclass <- factor(measure$Measureclass, levels = c("Physical", "Container", "Rest", "Amount", "Portion" ))
 measure$Kindgender   <- factor(measure$Kindgender, levels = c("Masc", "Neut", "Fem"))
 
 

### 2017-04-03 R/3_glmm.R

@@ -14,8 +14,8 @@ measure.glmm <- glmer(Construction~1
                  +(1|Kindlemma)
                  
                  +Badness                 
+                 +Cardinal
                  +Genitives
-                 +Leftcontext
                  +Measurecase
                  +Measurenumber
 
@@ -56,7 +56,7 @@ measure.glmm.lemrand.names <- rownames(coef(measure.glmm)$Kindlemma)[measure.glm
 
 # PB test.
 bootcomp.regs <- c("(1|Measurelemma)", "(1|Kindlemma)", 
-                   "Badness", "Genitives", "Leftcontext", "Measurecase", "Measurenumber", 
+                   "Badness", "Genitives", "Cardinal", "Measurecase", "Measurenumber",
                    "Kindattraction", "Kindfreq", "Kindgender", 
                    "Measureattraction", "Measureclass", "Measurefreq")
 modelcomp <- lmer.modelcomparison(model = measure.glmm, regressors = bootcomp.regs,
@@ -114,7 +114,7 @@ if (save.persistent) dev.off()
 
 # Effect plots.
 fixeff.pl.cex <- 1.5
-effs <- c("Badness", "Genitives", "Leftcontext", "Measurecase", "Measurenumber", 
+effs <- c("Badness", "Genitives", "Cardinal", "Measurecase", "Measurenumber",
           "Kindattraction", "Kindfreq", "Kindgender", 
           "Measureattraction", "Measureclass", "Measurefreq")
 
@@ -151,38 +151,4 @@ source('highstat.r')
 if (save.persistent) sink(paste(out.dir, "glmm.txt", sep=""), append = T)
 cat("\n\n Variance inflation diagnostics\n")
 print(myvif(measure.glmm))
-if (save.persistent) sink()
-
-
-# Nachschlag: PB test for interaction.
-measure.glmm.plus <- glmer(Construction~1
-                           
-                           +(1|Measurelemma)
-                           +(1|Kindlemma)
-                           
-                           +Badness                 
-                           +Genitives
-                           +Leftcontext
-                           +Measurecase
-                           +Measurenumber
-                           
-                           +Measurenumber:Leftcontext
-                           
-                           +Kindattraction
-                           +Kindfreq
-                           +Kindgender
-                           
-                           +Measureattraction
-                           +Measureclass
-                           +Measurefreq
-                           
[…]

### 2017-04-03 R/4_glmm-mcmc.R

@@ -7,23 +7,25 @@ options(mc.cores = parallel::detectCores())
 
 
 measure.glmm.mcmc <- stan_glmer(Construction~1
-                           
-                           +(1|Measurelemma)
-                           +(1|Kindlemma)
-                           
-                           +Badness                 
-                           +Genitives
-                           +Leftcontext
-                           +Measurecase
-                           +Measurenumber
-
-                           +Kindattraction
-                           +Kindfreq
-                           +Kindgender
-                           
-                           +Measureattraction
-                           +Measureclass
-                           +Measurefreq
+                                
+                                +(1|Measurelemma)
+                                +(1|Kindlemma)
+                                
+                                +Badness                 
+                                +Cardinal
+                                +Genitives
+                                +Measurecase
+                                +Measurenumber
+                                
+                                +Kindattraction
+                                #+Kindcollo
+                                +Kindfreq
+                                +Kindgender
+                                
+                                +Measureattraction
+                                #+Measurecollo
+                                +Measureclass
+                                +Measurefreq
                            ,
                  data=measure, family=binomial(link=logit),
                  chains=chains, seed=seed, iter=iter,

### 2017-04-03 R/5_stimuli_predict.R

@@ -8,9 +8,9 @@ stims.fax  <- c(
   "Measurecase",
   "Kindlemma",
   "Measurelemma",
-  "Measurenumber",
   "Measureclass",
-  "Leftcontext",
+  "Measurenumber",
+  "Cardinal",
   "Kindgender"
   )
 

### 2017-04-03 Latex/article.tex
evening sync

@@
 
-This paper deals with a morpho-syntactic alternation that occurs only in a very specific syntactic measure noun phrase construction.
+This paper deals with a morpho-syntactic alternation that occurs only in a very specific syntactic measure noun phrase construction in German.
 By \textit{alternation} I refer to a situation where two or more forms or constructions are available with no clear difference in acceptability, function, or meaning.
 
-In German grammar, the term \textit{Zweifelsfälle} `cases of doubt' is often used \citep{Klein2009}.
+% In German grammar, the term \textit{Zweifelsfälle} `cases of doubt' is often used \citep{Klein2009}.
 
@@
 %
-%Following the descriptive introduction to the measure noun phrase alternation in Section~\ref{sec:germanmeasurenps}, this section situates the work presented here within the broader theoretical context of what can be called \textit{cognitively oriented corpus linguistics}.
-%
-%Unless there are strong, numeric, and cognitively motivated hypotheses, some experimental validation is required to make a corpus study \textit{cognitive} in any sense.
-%Assuming the \textit{corpus as input} hypothesis without experimental validation is completely unsound in as much as researcher could extract any arbitrary -- often spurious -- usage pattern from a (usually noisy) corpus and claim that it has a cognitive reality.
-%
 % Cite:
@@
 
-In this section, I briefly motivate the choice of statistical models which I use later in Section~\ref{sec:corpusstudies}.
+In this section, I briefly motivate the choice of statistical models which I use later in Section~\ref{sec:corpusstudies}, namely Generalised Linear Mixed Models (GLMMs) using Maximum Likelihood Estimation.%
+\footnote{It would be technically more precise to speak of Hierarchical Logistic Regression rather than a GLMM in the case of this study.}
+Many readers might think that the use of this class of statistical methods in the modeling of grammatical alternations does not require much motivation.
+After all, GLMMs have been established as the major tool in the analysis of (\citealp{BresnanEa2007,BresnanHay2010,BresnanFord2010,DivjakArppe2013,Gries2015,NessetJanda2010}, to name only a few publications).
+However, over the past years alternative methods have been proposed.
+Within the Generalised Linear Modeling framework, multimodel averaging and multimodel inference (\citealp{BurnhamAnderson2002}, linguistic applications include \citealp{BarthKapatsinski2014,KupermanBresnan2012}) have been suggested to deal with model uncertainty and the sometimes problematic necessity to settle on one specific model structure.
+Also, alternative estimators have been introduced into the linguistics literature, most notably Bayesian estimation (a comprehensive introduction in \citealp{GelmanEa2014}, see \citealp{Levshina2016,Divjak2016} for praises by linguists).
+On the other hand, completely different modeling approaches have been proposed, for example Trees \& Forests \citep{StroblEa2009} and Naive Discriminative Learning \citep{Baayen2011,MilinEa2016}.
+The last two methods are summarised and compared to regression methods in \cite{BaayenEa2013}, and \cite{TheijssenEa2013} provides a similar comparison with even more methods.
 
-% NO to exploratory methods if we have hypotheses
-% Testing (hypotheses about signs of regressors)
-% Estimation => bootstrapping
-% Model selection
-%  => random vs fixed is not so simple, rather matter of getting model to converge
-%  => Maximality
-%  => NO to Gries 2015
-% Model evaluation
-% VS: Bayesian hype (Levshina)
-% VS: alternative math (R-W)
-
-% Cite:
-% __IMP__Measurenouns__cog-2015-0054 (Levshina)
-% [Cognitive Linguistics] Towards cognitively plausible data science in language research (NDL)
-
-% Stats:
-% BarrEa_RandomEffectsStructure_2014_JMemLang
-% ContentServer (1)
-% importatn_quotes
-% Menard_2000-Coefficients_of_Determination_for_Multiple_Logistic_Regression_Analysis
-
-%In the following, I examine the case alternation in German measure NPs under a cognitively oriented corpus linguistic perspective.
-%Since, however, the statistical tool kit (hierarchical generalised linear models, general tools such as $p$ values and confidence intervals, resampling techniques, model selection and model evaluation strategies, alternative algorithms for estimation, etc.) is heterogeneous and often criticised (especially under noisy conditions and when researcher run the risk of doing data mining instead of proper theory testing) I present a careful protocol for specifying, estimating, validating, and interpreting such models in Section \ref{sec:corpusstudies}.
-
-Before discussing the results in Section~\ref{sec:corpusstudies}, I show how I made sure that the relevant data is contained in the sample in Section \ref{sec:gettingdata}.
-
-%\subsection{Tools for modelling alternations}
-%\label{sec:righttools}
-%
-%In this section, I discuss the relevant statistical approaches to modelling alternations based on corpus data.
-%I mainly focus on well-established Generalised Linear Models (GLMs) and their \textit{multi-level} or \textit{hierarchical} extensions, which are also called \textit{mixed models} (GLMMs).
-%
-%% TODO CONTENT HERE
-%
-%In Section~\ref{sec:corpushierarchicalmodel}, I present the hierarchical model which is actually interpreted as the primary result of this study.
-%Section~\ref{sec:bayesian} takes a short detour and shows (by example) that results from Bayesian estimation are not intrinsically different from those obtained from Maximum Likelihood methods, although Bayesian approaches have recently been touted to solve various problems facing applied statistics in various fields, even including cognitively oriented corpus linguistics.
-%Finally, I summarise and interpret the results of the corpus study in Section~\ref{sec:interpretation}.
-
-% VanderPlas: Bayesian
+Often, linguists provide convincing arguments why adopting a statistical tool might beneficial.
+For example, \citep{Baayen2011,MilinEa2016} argue that Naive Discriminative Learning (NDL) is more adequate for modeling in strands of linguistics where cognitive reality is desired.
+For example, the fact that regression models cannot properly deal with redundancy in the form of multicollinearity makes them implausible given some cognitive research, while NDL can deal with redundant information \citep[508]{MilinEa2016}.
+Furthermore, the fact that likelihood maximisation is a form of optimsation is criticised \citep[508]{MilinEa2016}, although in NDL and Trees \& Forests are praised for \textit{arriv}[\textit{ing}] \textit{at their optimal solutions on their own} \cite[256]{BaayenEa2013}.
+More fundamentally, regression modeling has been criticised because it usually relies on high-level linguistic abstractions, and NDL (or other methods) can achieve at least the same adequacy by considering low-level units such as words or even character ngrams <++>
 
@@
 
-In a probabilistic framework like Prototype Theory, the crucial question is, however, what controls speakers' decisions to use either variant beyond simple lemma preference effects. 
-I argue that the NAC and the PGC are prototypes associated with different degrees of the grammaticalisation of the measure noun, morpho-syntactic properties, and register effects.
+In a probabilistic framework like Prototype Theory, the crucial question (beyond simple lemma preference effects) is, however, what controls speakers' decisions to use either variant. 
+In the remainder of this section, I argue that the NAC and the PGC are prototypes associated with different degrees of the grammaticalisation of the measure noun, a related morpho-syntactic property, and register effects.
 The degree of similarity of a given instance to either of the two prototypes makes speakers chose the \NACa\ or the \PGCa.
@@
 For illustration, (\ref{ex:determiners}) shows the expected variants under this hypothesis.
-In (\ref{ex:determiners:nac}), the measure noun is modified by a cardinal \textit{drei} `three', and hence the \NACa\ (the kind NP shows case agreement) is preferred.
-In (\ref{ex:determiners:pgc}), the measure noun is modified by a non-cardinal determiner \textit{einige} `some', and the \PGCa\ (the kind NP is in the genitive) is preferred.
+In (\ref{ex:determiners:nac}), the measure noun is modified by a cardinal \textit{drei} `three', and hence the \NACa\ is preferred.
+In (\ref{ex:determiners:pgc}), the measure noun is modified by a non-cardinal determiner \textit{einige} `some', and the \PGCa\ is preferred.
 
@@
 Under an integral view of prototypes, which incorporate effects related to larger contexts and registers, such preferences can be to be part of the construction prototypes.
-In the corpus study in Section~\ref{sec:corpusstudies}, register effects will therefore be modelled -- even if only using some very simple proxies.
+In the corpus study in Section~\ref{sec:corpusstudies}, register effects will therefore be modelled -- even if only using two very simple proxy variables.
 
@@
 At the time of this writing, a newer 2016 version DECOW16 had already been released.}
-I chose this corpus for two main reasons.
+I chose this corpus for two main reasons.%
+\footnote{The use of web data for linguistic research does require explicit and careful justification.
+Due to the noisy nature and unknown composition of the web, only carefully designed and established web corpora like the COW corpora or the SketchEngine corpora \citep{KilgarriffEa2014} should be used.
+Clearly, using search engine results is \textit{bad science} for many reasons, most prominently total irreproducibility of results, as \cite{Kilgarriff2006} pointed out more than ten years ago.
+Careless use of search engine results is still found, however, see \citet[171--175]{DeclerckBrems2016}.}
 First, the external validity of any study is increased through a higher heterogeneity of the sample \citep[30]{MaxwellDelaney2004}, and the DECOW corpus has clearly a much more heterogeneous composition compared to the only other very large corpus of German, the DeReKo \citep{KupietzEa2010} of the Institute for the German Language (IDS), which contains almost exclusively newspaper texts.%
@@
 For these or similar reasons, COW corpora have been used in a number of peer-reviewed publications, for example \cite{VanGoethemHiligsmann2014,VanGoethemHuening2015,MuellerS2014,Schaefer2016c,SchaeferSayatz2014,SchaeferSayatz2016,Zimmer2015}.
-Therefore, DECOW can be considered the obvious choice for this study.%
-\footnote{The use of web data for linguistic research does require explicit and careful justification.
-Due to the noisy nature and unknown composition of the web, only carefully designed and established web corpora like the COW corpora or the SketchEngine corpora \citep{KilgarriffEa2014} should be used.
-Clearly, using search engine results is \textit{bad science} for many reasons, most prominently total irreproducibility of results, as \cite{Kilgarriff2006} pointed out more than ten years ago.
-Careless use of search engine results is still found, however, see \citet[171--175]{DeclerckBrems2016}.}
+Therefore, DECOW can be considered the obvious choice for this study.
 
@@
 Furthermore, for each regressor, a $p$ value was obtained by dropping the regressor from the full model, re-estimating the nested model and comparing it to the full model.
-Instead of inexact Wald approximations or Likelihood Ratio Tests, I used a drop-in bootstrap replacement for the Likelihood Ratio Test from the function \textit{PBmodcomp} from the \textit{pbkrtest} package \citep{HalekohHojsgaard2014}, hence I call the value $p_{\text{PB}}$.
+Instead of inexact Wald approximations or Likelihood Ratio Tests, I used a drop-in bootstrap replacement for the Likelihood Ratio Test from the function \textit{PBmodcomp} from the \textit{pbkrtest} package \citep{HalekohHojsgaard2014}, hence I call the corresponding value $p_{\text{PB}}$.
 These bootstrapped $p$ values are given in the columns labelled p\Sub{PB} in Table~\ref{tab:bigtable}.

### 2017-04-04 Latex/article.tex
finished stats intro

@@
 
-% In German grammar, the term \textit{Zweifelsfälle} `cases of doubt' is often used \citep{Klein2009}.
+\vspace{\baselineskip}
 
-% Add to CogL special issue: Look at what non-cognitivists do
-% => their have interesting data and interesting generalisations (which are not intrinsically nonsense or incompatible with CogLi)
-% => where they lack is modelling probabilistic effects, and that should be one of the strengths of cognitively oriented (corpus) linguistics
-% => FUZZINESS and SIMILARITY in classification (prototypes and exemplars)
-% => advantage: testable with corpus and experiment (ideally both)
-
-% This is a case study and more.
-% => gathering a descriptive picture from non-cognitive literature
-% => motivating the phenomenon as relevant for CogLi
-% => careful protocol for:
-%  =>> motivating range of alternants under examination and selecting hypotheses (sound and/or slightly less sound, but MAKE CLEAR what they are!)
-%  =>> motivate whether larger utterance context plays a role or not (sociolinguistic & register(?) commitment)
-%  =>> choosing right data
-%  =>> choosing "the right" – NO, "some appropriate" statistics
-%  =>> getting minimal external validation with experiments
-
-% This article attempts to follow this protocol as well as possible given the available analyses/data sources/nature of the phenomenon.
-
-%%%%%% From old structure:
-%
-%\subsection{Cognitively oriented corpus linguistics}
-%\label{sec:cocl}
-%
-% Mention: Corpus-as-input is OK as a hypothesis but NOT as a research programme
-% Entrenchment_Stefanowitsch and Flach for submission + Schmid C2C
-%  => otherwise any spurious pattern found in corpora (meeting certain criteria)
-%  => is automatically "cognitive"
-%
-% Cite:
-% [Cognitive Linguistics] Similarity in linguistic categorization- The importance of necessary properties
-% 10.1.1.980.9935
-% __IMP__Measurenouns__cog-2015-0101
-%
-% Maybe:
-% [Cognitive Linguistics] The processing of verb-argument constructions is sensitive to form, function, frequency, contingency and prototypicality
-% [Cognitive Linguistics] Alternation-based generalizations are stored in the mental grammar- Evidence from a sorting task experiment //// along with CApelle's allostructions
-%
-% Method:
-% Noble
-
-% TODO SURPRISAL
-%
-% Bybee, Joan und Thompson, Sandra. 1997. “Three Frequency Effects in Syntax”, Proceedings of the Twenty-Third Annual Meeting of the Berkeley Linguistics Society, 378-388.
-% Bybee, Joan. 2006. “From Usage to Grammar: The Mind‘s Response to Repetition”. Language 82(4): 711-733.
-% Levy, Roger .2008. “Expectation-based syntactic comprehension.” Cognition 106: 1126–1177.
-% Smith, Nathaniel, and Levy, Roger. 2013. “The effect of word predictability on reading time is logarithmic.” Cognition 128: 302–319.
+\begin{itemize}
+  \item usage-based/cognitive corpus linguistics and the study of alternations
+  \item validationist turn; 2016 special issue CoGL, but Bresnan did it all the time!
+  \item also \cite{Gries2015a}
+  \item the value of case studies
+  \item `Zweifelsfälle' \citep{Klein2009} understood as ideal cases cor cognitive modeling
+  \item why forced choice and why reading times?
+\end{itemize}
 
@@
 
-In this section, I briefly motivate the choice of statistical models which I use later in Section~\ref{sec:corpusstudies}, namely Generalised Linear Mixed Models (GLMMs) using Maximum Likelihood Estimation.%
-\footnote{It would be technically more precise to speak of Hierarchical Logistic Regression rather than a GLMM in the case of this study.}
-Many readers might think that the use of this class of statistical methods in the modeling of grammatical alternations does not require much motivation.
-After all, GLMMs have been established as the major tool in the analysis of (\citealp{BresnanEa2007,BresnanHay2010,BresnanFord2010,DivjakArppe2013,Gries2015,NessetJanda2010}, to name only a few publications).
-However, over the past years alternative methods have been proposed.
-Within the Generalised Linear Modeling framework, multimodel averaging and multimodel inference (\citealp{BurnhamAnderson2002}, linguistic applications include \citealp{BarthKapatsinski2014,KupermanBresnan2012}) have been suggested to deal with model uncertainty and the sometimes problematic necessity to settle on one specific model structure.
-Also, alternative estimators have been introduced into the linguistics literature, most notably Bayesian estimation (a comprehensive introduction in \citealp{GelmanEa2014}, see \citealp{Levshina2016,Divjak2016} for praises by linguists).
-On the other hand, completely different modeling approaches have been proposed, for example Trees \& Forests \citep{StroblEa2009} and Naive Discriminative Learning \citep{Baayen2011,MilinEa2016}.
-The last two methods are summarised and compared to regression methods in \cite{BaayenEa2013}, and \cite{TheijssenEa2013} provides a similar comparison with even more methods.
+In this section, I motivate the choice of statistical models which I use later in Section~\ref{sec:corpusstudies}.
+Readers might think that the methods used here for modeling a grammatical alternation -- namely Hierarchical Regression and Generalised Linear Mixed Models (GLMMs) -- does not require much motivation.
+After all, GLMMs have been established as the major tool in the analysis of alternation phenomena (\citealp{BresnanEa2007,BresnanHay2010,BresnanFord2010,DivjakArppe2013,Gries2015,NessetJanda2010}, to name just a few publications).
+Over the past few years, however, modified or alternative methods have been proposed, and I want to make a few remarks on Bayesian estimation (see \citealp{GelmanEa2014}) as suggested in \cite{Levshina2016,Divjak2016a}, for example.
+Conceptually, I see three points of discussion that should be kept apart.
+First, Bayesian methods are sometimes touted as superior tools for scientific inference compared to frequentist methods.
+Second, it has been proposed that the Bayesian interpretation of probability is more cognitively adequate for the modeling of linguistic data.
+Third, and very specific to this paper, given established methods in the modeling of alternation and variation, it has to be decided whether so-called Bayesian methods lead to substantially different results.
 
-Often, linguists provide convincing arguments why adopting a statistical tool might beneficial.
-For example, \citep{Baayen2011,MilinEa2016} argue that Naive Discriminative Learning (NDL) is more adequate for modeling in strands of linguistics where cognitive reality is desired.
-For example, the fact that regression models cannot properly deal with redundancy in the form of multicollinearity makes them implausible given some cognitive research, while NDL can deal with redundant information \citep[508]{MilinEa2016}.
-Furthermore, the fact that likelihood maximisation is a form of optimsation is criticised \citep[508]{MilinEa2016}, although in NDL and Trees \& Forests are praised for \textit{arriv}[\textit{ing}] \textit{at their optimal solutions on their own} \cite[256]{BaayenEa2013}.
-More fundamentally, regression modeling has been criticised because it usually relies on high-level linguistic abstractions, and NDL (or other methods) can achieve at least the same adequacy by considering low-level units such as words or even character ngrams <++>
+As for the first question, this is not quite the place to discuss it fully.
+The basic distinction is a philosophical one and related to direct and inverse probability (\eg\ \citealp{Senn2011}).
+Frequentists assume that models and parameters are fixed, for example a model specifiying that a coin is fair.
+Given we know the distribution of the values measured, we can calculate for observed data (for example a measurement of $3$ heads in an experiment with $10$ tosses) how often such a result or a more extreme result would occur if the model were true and we repeated the experiment arbitrarily often.
+This is essentially the frequentist notion of direct probability, \ie\ long-run frequencies under replication.
+What makes this attractive is that a model -- specified with reference to a substantive theory of the given phenomenon -- can fail on given data, and if it does not, the outcome of the experiment contributes to the overall well-testedness of the theory.
+Bayesian approaches, on the other hand, condition on the data and quantify (inductively) the probability of models given specific observations.
+The models are thus not fixed, and the probability is usually equated with researchers posterior beliefs about models.
+There is actually a debate among Bayesians whether and how a notion of theory testing is compatible (or even already contained) in the Bayesian approach.
+In \citet[10]{GelmanShalizi2013}, the authors acknowledge this and state about the standard inductive interpretation of Bayesianism that \textit{most of this received view of Bayesian inference is wrong}, and they develop a Bayesian notion of p values (see also \citealp{Mayo2013}, for a frequentist reply; also \citealp{Senn2011} on different strands of Bayesianism and their stance on inductive vs.\ deductive reasoning).
+Clearly, in such quarrels between and among camps of philosophers of science and statisticians it is virtually impossible for mere practitioners to take sides based on a sufficiently informed judgement.
+
+These quarrels relate to the second question, however.
+\citet[301--302]{Divjak2016a} speaks favourably of Bayesian methods because the Bayesian concept of probability is more adequate for cognitive modeling compared to the frequentist one.
+This criticism is part of a larger body of literature asking for cognitively plausible modeling techniques, for example Naive Discriminative Learning (NDL; \citealp{Baayen2011,BaayenEa2013,MilinEa2016,TheijssenEa2013}).
+Frequentist methods were, however, never designed as cognitive models (see above).
+The actual question that emerges here is what our statistical models (estimated on corpus data) should represent.
+Are they inductive models of cognitive representations that learners would infer from being exposed to the corpus data?%
+\footnote{In which case we are doing \textit{data science in language research} in the words of \citealp{MilinEa2016}.}
+Or are they tests of theories that are pre-specified and merely tested for predictive accuracy on output data.
+In the former case, we should definitely resort to methods like NDL, which most likely would entail tossing most previous work done in (cognitive) linguistics into the bin and abandoning all high-level generalisations that most existing studies have been based upon.
+In fact, arguments to this extreme effect and against using high-level generalisation have been made, for example in \cite{BaayenEa2016}, \citet[299--300]{Divjak2016a}, \cite{RamscarPort2016}, \cite{TheijssenEa2013}.
+In the latter case, the cognitive commitment does not necessarily extend to the statistical methods used.
+These methods then do not need to be any more cognitively plausible than an ANOVA used to analyse the results from an experiment.
+I view my own work in the tradition of testing theories (embracing high-level generalisations) on data (see Sections~\ref{sec:germanmeasurenps}--\ref{sec:externalvalidation}), but I am fundamentally agnostic with respect to this question.
+The best strategy for cognitive linguistics as a field might be to cultivate many methods while making sure that each method is applied carefully and competently.
+
+The third point is the most practically relevant in the context of this paper.
+In her otherwise excellent study, \citet[251--252]{Levshina2016} argues for Bayesian estimation in mixed regression settings.
+First, she claims that \textit{while frequentist statistics only allows one to test whether the null hypothesis can be rejected, Bayesian statistics enables one both to test the null hypothesis and to estimate the probability of specific parameter values given the data}.
+This does not do justice to frequentist methods in that rejection of the null hypothesis is characteristic of Fisher's approach, but in the Neyman-Pearson approach, results ideally \textit{favour} the main hypothesis (vs.\ the alternative), cf.\ \cite{Lehmann1993,Lehmann2011,Perezgonzalez2015}.
+She then explains that a \textit{distinctive feature of Bayesian statistics is the use of so-called priors} and that \textit{posterior probabilities depend on both the prior beliefs and the data, whereas the results of a frequentist model depend only on the data} \cite[252]{Levshina2016}.
+Remarkably, however, she does \textit{not} use informative priors, and in her footnote 8 \cite[252]{Levshina2016} admits that priors were probed using trial and error, such that the proclaimed advantage of Bayesian modeling was apparently not taken advantage of.%
+\footnote{In the words of \cite{Senn2011}: \textit{You may believe you are a Bayesian but you are probably wrong}.
+\cite[347--348]{GelmanHill2006} \textit{view any noninformative prior distribution as inherently provisional}.}
+Now, Maximum Likelihood Estimation (MLE), the traditional method which could have been used here, is not exactly \textit{frequentist} in the sense of Neyman-Pearson test theory.
+MLE also conditions on the data in as much as it finds the most likely set of parameters given the data.
+What is more, Bayesian estimators in fact incorporate the Likelihood function and merely multiply it by the prior \citep[PAGENUMBER??]{GelmanEa2014}.
+If the prior is flat, results converge (see also \citealp[347]{GelmanHill2006}).
+The same is true if the sample size is large compared to the number of parameters, at least for finite-dimensional models \citep[1119--1120]{Freedman1999}, a well-established result known as the Bernstein-von Mises theorem.
+With a modest model structure including $17$ fixed effects and $2,646$ data points in \cite{Levshina2016}, it is highly likely that the same results would have obtained with Maximum Likelihood methods.
+If not, especially in case of convergence problems, this would have been an ideal occasion to demonstrate the superiority of Bayesian estimation.
+There are situations where Bayesian estimators can be more robust, namely with heavily censored data, empty cells, complex hierarchical models, perfect separation (see \citealp{Freedman1999}, \cite[345--348]{GelmanHill2006}).
+I want to reiterate that these points do not in any way invalidate the study presented by \cite{Levshina2016}.
+However, being \textit{Bayesian} is not among its selling points.
+In Section~\ref{sec:corpushierarchicalmodel}, I compare my own hierarchical models estimated with Bayesian and MLE estimators to demonstrate convergence.
+
+This concludes the general introduction.
+In the remainder of the paper, I introduce the alternation phenomenon (Section~\ref{sec:germanmeasurenps}) and suggest a theory-driven set of factors influencing the alternation (Section~\ref{sec:analyses}).
+Then, the corpus study is presented including statistical analysis (Section~\ref{sec:corpusstudies}).
+The two experiments validating the corpus study are then reported (Section~\ref{sec:externalvalidation} before I sum up the paper in Section~\ref{sec:conclusion}.
 
@@
 % This means that with dative measure nouns, the alternation is unobservable if the kind noun is feminine, and it is problematic to speak of a \textit{genitive} kind noun construction in this case.
-% This is taken into account in the statistical modelling and the discussion presented in Section~\ref{sec:corpusstudies}.
+% This is taken into account in the statistical modeling and the discussion presented in Section~\ref{sec:corpusstudies}.
 % 
@@
 Under an integral view of prototypes, which incorporate effects related to larger contexts and registers, such preferences can be to be part of the construction prototypes.
-In the corpus study in Section~\ref{sec:corpusstudies}, register effects will therefore be modelled -- even if only using two very simple proxy variables.
+In the corpus study in Section~\ref{sec:corpusstudies}, register effects will therefore be modeled -- even if only using two very simple proxy variables.
 
@@
 Finally, two auxiliary samples were also drawn.
-As mentioned in Section~\ref{sec:analyses}, the distribution of the measure noun and kind noun lemmas in the \NACb\ and the \PGCd\ with a determiner will be modelled as factors influencing the alternation.
+As mentioned in Section~\ref{sec:analyses}, the distribution of the measure noun and kind noun lemmas in the \NACb\ and the \PGCd\ with a determiner will be modeled as factors influencing the alternation.
 Therefore, all noun-noun pairs from the bootstrap process were also queried in the two non-alternating constructions, resulting in 17,252 hits for the \PGCd\ and 315,635 hits for the \NACb.
@@
 They cluster around $-\infty$ and $0$, especially when (as in this case) cells contain $0$ count values.
-This cannot be remedied on principled grounds, since, for example, a z transformation is inadequate for such a strongly bimodal distribution (even if we clamp $-\infty$ to a very low numeric value).
+This cannot be remedied on principled grounds, since, for example, a z-transformation is inadequate for such a strongly bimodal distribution (even if we clamp $-\infty$ to a very low numeric value).
 Second, their main use in collostructional analyses is to \textit{sort} a list of collexemes and interpret the order.
@@
 The model was re-estimated using the \textit{stan\_glmer} function from the \textit{rstanarm} package, which provides an \textit{lme4}-compatible syntax for estimating common model types with the \textit{stan} software \citep{CarpenterEa2017}.
-The algorithm was run with $4$ chains and $1,000$ iterations, and instead of guessing prior distributions or using implausible uniform prior distributions \cite[251--252]{Levshina2016}, I used plausible default priors.
+The algorithm was run with $4$ chains and $1,000$ iterations, and I used plausible default priors.
 Most notably, priors for coefficients were specified as $\mathcal{N}(0,10)$ because coefficients higher than $10$ or lower than $-10$ are extremely rare in well-specified models on appropriate data.%
@@
 
+I have shown!
+Oh, yes, I have shown\ldots
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2017-04-04 Latex/article.tex
minor fixes

@@
 I call the alternating variants with an adjective but no determiner in the kind noun phrase \NACa, and \PGCa, respectively.
-This paper is about the middle column of Table~\ref{tab:constructions}, \ie, the syntactic configuration in which two different case patterns are acceptable.
+This paper is about the middle column of Table~\ref{tab:constructions}, \ie\ the syntactic configuration in which two different case patterns are acceptable.
 
@@
 In the case at hand, the variable might be something like \textit{measuring quantities of substances and collections}.
-I argue that it is fully justified to focus narrowly on \NACa\ and \PGCa\ with their well-defined morph-syntactic properties, mostly because the alternative constructions are not used in the same range of contexts.
+I argue that it is fully justified to focus narrowly on \NACa\ and \PGCa\ with their well-defined morpho-syntactic properties, mostly because the alternative constructions are not used in the same range of contexts.
 Two major alternatives might be considered.
@@
 It is known from frameworks like Collexeme Analysis \citep{GriesStefanowitsch2004} that lemmas are attracted with different strengths by competing constructions.
-If the \NACa\ is highly similar to the \NACb\ and the \PGCa\ is highly similar to the \PGCd, the probability with which individual noun lemmas appear in the alternating constructions should be predictable at least partly from the relative frequency with which the same lemmas are used in the non-alternating constructions.
+If the \NACa\ is highly similar to the \NACb\ and the \PGCa\ is highly similar to the \PGCd, the probability with which individual noun lemmas appear in the alternating constructions should be predictable at least partly from the relative frequency with which the same lemmas are used in the non-alternating constructions (see \citealp[246--249]{Levshina2016} for a similar idea in a different context).
 Nouns which occur proportionally more often in the \PGCd\ should favour the \PGCa, and nouns which occur proportionally more often in the \NACb\ should favour the \NACa.
@@
 The full set of manually annotated variables for the main sample is given in Table~\ref{tab:variables}, and I briefly discuss it now.%
-\footnote{All numeric variables were also z-transformed (\ie, centered to the mean and rescaled such that they have a standard deviation of $1$) to facilitate their interpretation in the regression models reported in the next section.}
+\footnote{All numeric variables were also z-transformed (\ie\ centered to the mean and rescaled such that they have a standard deviation of $1$) to facilitate their interpretation in the regression models reported in the next section.}
 Notice that \textit{Construction} is the response variable (or `dependent variable') with the values \textit{PGCa} and \textit{NACa}.

### 2017-04-04 Latex/article.tex
minor fixes

@@
 After all, GLMMs have been established as the major tool in the analysis of alternation phenomena (\citealp{BresnanEa2007,BresnanHay2010,BresnanFord2010,DivjakArppe2013,Gries2015,NessetJanda2010}, to name just a few publications).
-Over the past few years, however, modified or alternative methods have been proposed, and I want to make a few remarks on Bayesian estimation (see \citealp{GelmanEa2014}) as suggested in \cite{Levshina2016,Divjak2016a}, for example.
+Over the past few years, however, modified or alternative methods have been proposed, and I want to make a few remarks on Bayesian estimation (see \citealp{GelmanEa2014}) as suggested in \cite{Levshina2016,Divjak2016a}.
 Conceptually, I see three points of discussion that should be kept apart.
@@
 Bayesian approaches, on the other hand, condition on the data and quantify (inductively) the probability of models given specific observations.
-The models are thus not fixed, and the probability is usually equated with researchers posterior beliefs about models.
-There is actually a debate among Bayesians whether and how a notion of theory testing is compatible (or even already contained) in the Bayesian approach.
+The models are thus not fixed, and the probability is usually equated with researchers' posterior beliefs about models.
+There is actually a debate among Bayesians whether and how a notion of hypothesis testing is compatible (or even already contained) in the Bayesian approach.
 In \citet[10]{GelmanShalizi2013}, the authors acknowledge this and state about the standard inductive interpretation of Bayesianism that \textit{most of this received view of Bayesian inference is wrong}, and they develop a Bayesian notion of p values (see also \citealp{Mayo2013}, for a frequentist reply; also \citealp{Senn2011} on different strands of Bayesianism and their stance on inductive vs.\ deductive reasoning).
-Clearly, in such quarrels between and among camps of philosophers of science and statisticians it is virtually impossible for mere practitioners to take sides based on a sufficiently informed judgement.
+Clearly, in such quarrels between and among camps of philosophers of science and statisticians it is virtually impossible for mere practitioners to take sides.
 
@@
 \citet[301--302]{Divjak2016a} speaks favourably of Bayesian methods because the Bayesian concept of probability is more adequate for cognitive modeling compared to the frequentist one.
-This criticism is part of a larger body of literature asking for cognitively plausible modeling techniques, for example Naive Discriminative Learning (NDL; \citealp{Baayen2011,BaayenEa2013,MilinEa2016,TheijssenEa2013}).
+Her argument is part of a larger body of literature asking for cognitively plausible modeling techniques, for example Naive Discriminative Learning (NDL; \citealp{Baayen2011,BaayenEa2013,MilinEa2016,TheijssenEa2013}).
 Frequentist methods were, however, never designed as cognitive models (see above).
-The actual question that emerges here is what our statistical models (estimated on corpus data) should represent.
+The actual question that emerges here is how we interpret our statistical models (estimated on corpus data).
 Are they inductive models of cognitive representations that learners would infer from being exposed to the corpus data?%
 \footnote{In which case we are doing \textit{data science in language research} in the words of \citealp{MilinEa2016}.}
-Or are they tests of theories that are pre-specified and merely tested for predictive accuracy on output data.
+Or are they tests of theories that are pre-specified and merely tested for predictive accuracy on linguistic output data contained in corpora.
 In the former case, we should definitely resort to methods like NDL, which most likely would entail tossing most previous work done in (cognitive) linguistics into the bin and abandoning all high-level generalisations that most existing studies have been based upon.
@@
 
-The third point is the most practically relevant in the context of this paper.
+The third point, then, is the most practically relevant in the context of this paper.
 In her otherwise excellent study, \citet[251--252]{Levshina2016} argues for Bayesian estimation in mixed regression settings.
 First, she claims that \textit{while frequentist statistics only allows one to test whether the null hypothesis can be rejected, Bayesian statistics enables one both to test the null hypothesis and to estimate the probability of specific parameter values given the data}.
-This does not do justice to frequentist methods in that rejection of the null hypothesis is characteristic of Fisher's approach, but in the Neyman-Pearson approach, results ideally \textit{favour} the main hypothesis (vs.\ the alternative), cf.\ \cite{Lehmann1993,Lehmann2011,Perezgonzalez2015}.
+This does not do justice to frequentist methods in that mere rejection of the null hypothesis is characteristic only of Fisher's approach, but in the Neyman-Pearson approach, results ideally \textit{favour} the main hypothesis (vs.\ the alternative), cf.\ \cite{Lehmann1993,Lehmann2011,Perezgonzalez2015}.
 She then explains that a \textit{distinctive feature of Bayesian statistics is the use of so-called priors} and that \textit{posterior probabilities depend on both the prior beliefs and the data, whereas the results of a frequentist model depend only on the data} \cite[252]{Levshina2016}.
@@
 \cite[347--348]{GelmanHill2006} \textit{view any noninformative prior distribution as inherently provisional}.}
-Now, Maximum Likelihood Estimation (MLE), the traditional method which could have been used here, is not exactly \textit{frequentist} in the sense of Neyman-Pearson test theory.
+Now, Maximum Likelihood Estimation (MLE), the traditional method which could have been used here, is not exactly \textit{frequentist} in the sense of Neyman-Pearson testing theory.
 MLE also conditions on the data in as much as it finds the most likely set of parameters given the data.
@@
 If the prior is flat, results converge (see also \citealp[347]{GelmanHill2006}).
-The same is true if the sample size is large compared to the number of parameters, at least for finite-dimensional models \citep[1119--1120]{Freedman1999}, a well-established result known as the Bernstein-von Mises theorem.
-With a modest model structure including $17$ fixed effects and $2,646$ data points in \cite{Levshina2016}, it is highly likely that the same results would have obtained with Maximum Likelihood methods.
+The same is true if the sample size is large compared to the number of parameters, at least for finite-dimensional models \citep[1119--1120]{Freedman1999}, a well-established result known as the Bernstein-von Mises theorem.%
+\footnote{Relating back to cognitive plausibility, this means that Bayesian methods cannot be more cognitively plausible in these situations than Maximum Likelihood Estimation.}
+With a modest model structure including $17$ fixed effects and $2,646$ data points in \cite{Levshina2016}, it is highly likely that the same results would have obtained with Maximum Likelihood methods.%
+\footnote{In fact, she admits that changing the priors did not lead to substantially different results in her footnote 8.
+  This is a clear sign that the prior is \textit{swamped by the data} (QUOTE??).}
 If not, especially in case of convergence problems, this would have been an ideal occasion to demonstrate the superiority of Bayesian estimation.
-There are situations where Bayesian estimators can be more robust, namely with heavily censored data, empty cells, complex hierarchical models, perfect separation (see \citealp{Freedman1999}, \cite[345--348]{GelmanHill2006}).
-I want to reiterate that these points do not in any way invalidate the study presented by \cite{Levshina2016}.
+After all, there are situations where Bayesian estimators can be more robust, namely with heavily censored data, complex hierarchical models, perfect separation, etc.
+(see \citealp{Freedman1999}, \citealp[345--348]{GelmanHill2006}).
+I want to reiterate that these points do not in any way invalidate the study presented in \cite{Levshina2016}.
 However, being \textit{Bayesian} is not among its selling points.
-In Section~\ref{sec:corpushierarchicalmodel}, I compare my own hierarchical models estimated with Bayesian and MLE estimators to demonstrate convergence.
+That said, I want to voice the concern that most likely many practitioners struggle with getting an adequate grasp of advanced statistical methods, and that it might therefore be wise to use the more conservative and better understood method if the alternative method is not absolutely required for substantive reasons.
+In Section~\ref{sec:corpushierarchicalmodel}, I compare my own hierarchical models estimated with Bayesian and MLE estimators to demonstrate their expectable convergence.
 
 This concludes the general introduction.
-In the remainder of the paper, I introduce the alternation phenomenon (Section~\ref{sec:germanmeasurenps}) and suggest a theory-driven set of factors influencing the alternation (Section~\ref{sec:analyses}).
+In the remainder of the paper, I introduce the alternation phenomenon (Section~\ref{sec:descriptive}) and suggest a theory-driven set of factors influencing the alternation (Section~\ref{sec:analyses}).
 Then, the corpus study is presented including statistical analysis (Section~\ref{sec:corpusstudies}).
-The two experiments validating the corpus study are then reported (Section~\ref{sec:externalvalidation} before I sum up the paper in Section~\ref{sec:conclusion}.
+The two experiments validating the corpus study are then reported (Section~\ref{sec:externalvalidation}) before I sum up the paper in Section~\ref{sec:conclusion}.
 

### 2017-04-05 Latex/article.tex
another claenup pass before finishing writing

@@
 In this section, I motivate the choice of statistical models which I use later in Section~\ref{sec:corpusstudies}.
-Readers might think that the methods used here for modeling a grammatical alternation -- namely Hierarchical Regression and Generalised Linear Mixed Models (GLMMs) -- does not require much motivation.
+Readers might think that the method used here for modeling grammatical alternations -- namely Hierarchical Logistic Regression\slash Generalised Linear Mixed Models (GLMMs) -- does not require much motivation.
 After all, GLMMs have been established as the major tool in the analysis of alternation phenomena (\citealp{BresnanEa2007,BresnanHay2010,BresnanFord2010,DivjakArppe2013,Gries2015,NessetJanda2010}, to name just a few publications).
-Over the past few years, however, modified or alternative methods have been proposed, and I want to make a few remarks on Bayesian estimation (see \citealp{GelmanEa2014}) as suggested in \cite{Levshina2016,Divjak2016a}.
+Over the past few years, however, modified or alternative methods have been proposed.
+From among these methods, I just make a few remarks on Bayesian estimation (see \citealp{GelmanEa2014}), as it was proposed in \cite{Levshina2016,Divjak2016a}.
 Conceptually, I see three points of discussion that should be kept apart.
@@
 As for the first question, this is not quite the place to discuss it fully.
-The basic distinction is a philosophical one and related to direct and inverse probability (\eg\ \citealp{Senn2011}).
+The basic distinction is a philosophical one and related to the concepts of \textit{direct} and \textit{inverse probability} (\eg\ \citealp{Senn2011}).
 Frequentists assume that models and parameters are fixed, for example a model specifiying that a coin is fair.
-Given we know the distribution of the values measured, we can calculate for observed data (for example a measurement of $3$ heads in an experiment with $10$ tosses) how often such a result or a more extreme result would occur if the model were true and we repeated the experiment arbitrarily often.
+We can then calculate for observed data (for example a measurement of $3$ heads in an experiment with $10$ tosses) how often such a result or a more extreme result would occur if the model were true and we repeated the experiment arbitrarily often.
 This is essentially the frequentist notion of direct probability, \ie\ long-run frequencies under replication.
-What makes this attractive is that a model -- specified with reference to a substantive theory of the given phenomenon -- can fail on given data, and if it does not, the outcome of the experiment contributes to the overall well-testedness of the theory.
-Bayesian approaches, on the other hand, condition on the data and quantify (inductively) the probability of models given specific observations.
-The models are thus not fixed, and the probability is usually equated with researchers' posterior beliefs about models.
+What makes frequentism attractive is that it lends itself to a theory of testing whether a theoretical prediction (a fixed model) was successful or not.
+Such tests, if passed, contribute to the overall well-testedness of the theory.
+Bayesian approaches, on the other hand, condition on the data and quantify inductively (in the standard interpretation) the probability of models given specific observations.
+The models are thus not fixed, and the probability is usually equated with researchers' posterior beliefs about models or model parameters.
 There is actually a debate among Bayesians whether and how a notion of hypothesis testing is compatible (or even already contained) in the Bayesian approach.
-In \citet[10]{GelmanShalizi2013}, the authors acknowledge this and state about the standard inductive interpretation of Bayesianism that \textit{most of this received view of Bayesian inference is wrong}, and they develop a Bayesian notion of p values (see also \citealp{Mayo2013}, for a frequentist reply; also \citealp{Senn2011} on different strands of Bayesianism and their stance on inductive vs.\ deductive reasoning).
-Clearly, in such quarrels between and among camps of philosophers of science and statisticians it is virtually impossible for mere practitioners to take sides.
+In \citet[10]{GelmanShalizi2013}, the authors -- prominent Bayesians -- acknowledge that theory testing is a desideratum, and they state about the standard inductive interpretation of Bayesianism that \textit{most of this received view of Bayesian inference is wrong}, and they develop a Bayesian notion of p values (see also \citealp{Mayo2013}, for a frequentist reply; also \citealp{Senn2011} on different strands of Bayesianism and their stance on inductive vs.\ deductive reasoning).
+Clearly, in such quarrels between and among camps of philosophers of science and statisticians it is difficult for mere practitioners to take sides.
 
@@
 Her argument is part of a larger body of literature asking for cognitively plausible modeling techniques, for example Naive Discriminative Learning (NDL; \citealp{Baayen2011,BaayenEa2013,MilinEa2016,TheijssenEa2013}).
-Frequentist methods were, however, never designed as cognitive models (see above).
+Frequentist methods were, however, never designed as cognitive models, but as systems of inference for researchers (see above).
 The actual question that emerges here is how we interpret our statistical models (estimated on corpus data).
@@
 Or are they tests of theories that are pre-specified and merely tested for predictive accuracy on linguistic output data contained in corpora.
-In the former case, we should definitely resort to methods like NDL, which most likely would entail tossing most previous work done in (cognitive) linguistics into the bin and abandoning all high-level generalisations that most existing studies have been based upon.
+In the former case, we should definitely resort to methods like NDL.
+However, this would most likely require us to toss most previous work done in (cognitive) linguistics into the bin and abandoning all high-level generalisations that most existing studies have been based upon.
 In fact, arguments to this extreme effect and against using high-level generalisation have been made, for example in \cite{BaayenEa2016}, \citet[299--300]{Divjak2016a}, \cite{RamscarPort2016}, \cite{TheijssenEa2013}.
-In the latter case, the cognitive commitment does not necessarily extend to the statistical methods used.
+In the latter and less extreme case, the cognitive commitment does, however, not necessarily extend to the statistical methods used.
 These methods then do not need to be any more cognitively plausible than an ANOVA used to analyse the results from an experiment.
-I view my own work in the tradition of testing theories (embracing high-level generalisations) on data (see Sections~\ref{sec:germanmeasurenps}--\ref{sec:externalvalidation}), but I am fundamentally agnostic with respect to this question.
+I view my own work (see Sections~\ref{sec:germanmeasurenps}--\ref{sec:externalvalidation}) in the tradition of testing theories that embrace high-level generalisations in a probabilistic framework.
+I remain fully agnostic with respect to the question which approach is `right'.
 The best strategy for cognitive linguistics as a field might be to cultivate many methods while making sure that each method is applied carefully and competently.
@@
 First, she claims that \textit{while frequentist statistics only allows one to test whether the null hypothesis can be rejected, Bayesian statistics enables one both to test the null hypothesis and to estimate the probability of specific parameter values given the data}.
-This does not do justice to frequentist methods in that mere rejection of the null hypothesis is characteristic only of Fisher's approach, but in the Neyman-Pearson approach, results ideally \textit{favour} the main hypothesis (vs.\ the alternative), cf.\ \cite{Lehmann1993,Lehmann2011,Perezgonzalez2015}.
+This does not do justice to frequentist methods in that mere rejection of the null hypothesis is characteristic only of Fisher's approach, but in the Neyman-Pearson approach, results ideally \textit{favour} the main hypothesis vis-à-vis the alternative hypothesis (cf.\ \citealp{Lehmann1993,Lehmann2011,Perezgonzalez2015}).
 She then explains that a \textit{distinctive feature of Bayesian statistics is the use of so-called priors} and that \textit{posterior probabilities depend on both the prior beliefs and the data, whereas the results of a frequentist model depend only on the data} \cite[252]{Levshina2016}.
@@
 \footnote{In the words of \cite{Senn2011}: \textit{You may believe you are a Bayesian but you are probably wrong}.
-\cite[347--348]{GelmanHill2006} \textit{view any noninformative prior distribution as inherently provisional}.}
-Now, Maximum Likelihood Estimation (MLE), the traditional method which could have been used here, is not exactly \textit{frequentist} in the sense of Neyman-Pearson testing theory.
+\citet[347--348]{GelmanHill2006} \textit{view any noninformative prior distribution as inherently provisional} and give recommendations how to proceed once posteriors have been obtained from noninformative priors.}
+Now, Maximum Likelihood Estimation (MLE) -- the traditional method which could have been used instead -- is not exactly \textit{frequentist} in the sense of Neyman-Pearson testing theory.
 MLE also conditions on the data in as much as it finds the most likely set of parameters given the data.
@@
 If the prior is flat, results converge (see also \citealp[347]{GelmanHill2006}).
-The same is true if the sample size is large compared to the number of parameters, at least for finite-dimensional models \citep[1119--1120]{Freedman1999}, a well-established result known as the Bernstein-von Mises theorem.%
-\footnote{Relating back to cognitive plausibility, this means that Bayesian methods cannot be more cognitively plausible in these situations than Maximum Likelihood Estimation.}
+The same is true if the sample size is large compared to the number of parameters, at least for finite-dimensional models \citep[1119--1120]{Freedman1999}, a well-established result known as the Bernstein-von Mises theorem.
 With a modest model structure including $17$ fixed effects and $2,646$ data points in \cite{Levshina2016}, it is highly likely that the same results would have obtained with Maximum Likelihood methods.%
 \footnote{In fact, she admits that changing the priors did not lead to substantially different results in her footnote 8.
-  This is a clear sign that the prior is \textit{swamped by the data} (QUOTE??).}
+  This is a clear sign that the prior is \textit{swamped by the data} (QUOTE??).
+  Relating back to cognitive plausibility, this means that in the case at hand Bayesian methods cannot be more cognitively plausible than Maximum Likelihood Estimation.}
 If not, especially in case of convergence problems, this would have been an ideal occasion to demonstrate the superiority of Bayesian estimation.
@@
 However, being \textit{Bayesian} is not among its selling points.
-That said, I want to voice the concern that most likely many practitioners struggle with getting an adequate grasp of advanced statistical methods, and that it might therefore be wise to use the more conservative and better understood method if the alternative method is not absolutely required for substantive reasons.
+That said, I want to voice the concern that most likely many practitioners are already struggling with getting an adequate grasp of advanced statistical methods, and that it might therefore be wise to use the more conservative and better understood method if the alternative method is not absolutely required for substantive reasons.
 In Section~\ref{sec:corpushierarchicalmodel}, I compare my own hierarchical models estimated with Bayesian and MLE estimators to demonstrate their expectable convergence.
@@
 In this section, I introduce and illustrate the relevant alternating constructions.
-I describe the narrowly defined syntactic configuration in which the alternation occurs, and I motivate the focus on \textit{only} this narrow (range rather than, for example, the whole range of nominal constructions expressing quantities).
+I describe the narrowly defined syntactic configuration in which the alternation occurs, and I motivate the focus on \textit{only} this narrow range (rather than, for example, the whole range of nominal constructions expressing quantities).
 
@@
 I call the kind-denoting noun simply the \textit{kind noun} and the quantity-denoting noun the \textit{measure noun}.
+For illustration purposes, in the English \textit{a cup of fine coffee}, \textit{cup} is the measure noun, and \textit{coffee} is the kind noun.
 Measure nouns can be all sorts of nouns which genuinely denote a quantity (such as \textit{litre} or \textit{amount}) but also nouns denoting containers, collections, etc.
@@
 In a similar vein, \citet[284]{Brems2003} considers nouns as measure nouns \textit{which, strictly speaking, do not designate a `measure', but display a more nebulous} (\textit{sic!}) \textit{potential for quantification} (see also \citealp[530]{Koptjevskaja2001}, and \citealp[338]{Rutkowski2007}).
-For illustration purposes, in the English \textit{a cup of fine coffee}, \textit{cup} is the measure noun, and \textit{coffee} is the kind noun.
 
@@
 \begin{exe}
-  \ex\label{ex:intro:pseudopartitive}
-  \begin{xlist}
+%  \ex\label{ex:intro:pseudopartitive}
+%  \begin{xlist}
     \ex[ ]{\label{ex:intro:pseudopartitive1}\gll Wir trinken [[eine Tasse]\Sub{Acc} [eines leckeren Kaffees]\Sub{Gen}]\Sub{Acc}.\\
@@
     \glt We drink a cup of a tasty coffee.}
-    \ex[*]{\label{ex:intro:pseudopartitive2} Wir trinken [[eine Tasse]\Sub{Acc} [einen leckeren Kaffee]\Sub{Acc}]\Sub{Acc}.}
-  \end{xlist}
+%    \ex[*]{\label{ex:intro:pseudopartitive2} Wir trinken [[eine Tasse]\Sub{Acc} [einen leckeren Kaffee]\Sub{Acc}]\Sub{Acc}.}
+%  \end{xlist}
 \end{exe}
 
-If the kind noun is bare -- \ie\ if it does neither come with a determiner nor a modifying adjective -- it has to agree in case with the measure noun, and the genitive seen in the PGC is not acceptable, see (\ref{ex:intro:narrowapposition}).
+If the kind noun is bare -- \ie\ if it does neither come with a determiner nor a modifying adjective -- it has to agree in case with the measure noun, and the genitive seen in the PGC is not acceptable, see (\ref{ex:intro:narrowapposition2}).
 
 \begin{exe}
-  \ex\label{ex:intro:narrowapposition}
-  \begin{xlist}
-    \ex[*]{\label{ex:intro:narrowapposition1} Wir trinken [[eine Tasse]\Sub{Acc} [Kaffees]\Sub{Gen}]\Sub{Acc}.}
+%  \ex\label{ex:intro:narrowapposition}
+%  \begin{xlist}
+%    \ex[*]{\label{ex:intro:narrowapposition1} Wir trinken [[eine Tasse]\Sub{Acc} [Kaffees]\Sub{Gen}]\Sub{Acc}.}
     \ex[ ]{\label{ex:intro:narrowapposition2}\gll Wir trinken [[eine Tasse]\Sub{Acc} [Kaffee]\Sub{Acc}]\Sub{Acc}.\\
@@
     \glt We drink a cup of coffee.}
-  \end{xlist}
+%  \end{xlist}
 \end{exe}
@@
 Notice that the unavailability of the genitive on the kind noun can be seen as following from a rather quirky constraint that genitive NPs in German require the presence of some strongly case-marked element (determiner or adjective) in addition to the head noun in order to be acceptable \citep{GallmannLindauer1994,Schachtl1989}.
-While only an accusative is shown in (\ref{ex:intro:narrowapposition}), the kind noun obligatory agrees in case also with nominative and dative measure nouns.
+While only an accusative is shown in (\ref{ex:intro:narrowapposition2}), the kind noun obligatory agrees in case also with nominative and dative measure nouns.
 
@@
 Also, in my corpus sample, they simply did not occur.
-Even if they are accepted by some speakers, their extremely low frequency makes it virtually impossible to study them using either corpus methods (Section~\ref{sec:corpusstudies}) or experimental approaches (Section~\ref{sec:externalvalidation}), and I consequently ignore them.
+Even if they are accepted by some speakers, their extremely low frequency makes it virtually impossible to study them using either corpus methods (Section~\ref{sec:corpusstudies}), and I consequently ignore them.
 
@@
 In line with experimental results reported in \citet[15--16]{Zimmer2015}, I found that the PGC is so dominant with plural kind nouns (67 of 861 cases or over 92\%, cf.\ Section~\ref{sec:corpusstudies}) that the alternation cannot be analysed in the same way as in the singular case.%
-\footnote{More concretely, I these cases were included in the regression analysis of the corpus data along with a factor encoding the number of the kind noun, this factor would most assuredly override any other regressor for the data points with a plural kind noun.}
+\footnote{More concretely, if these cases were included in the regression analysis of the corpus data along with a factor encoding the number of the kind noun, this factor would most assuredly override any other regressor for the data points with a plural kind noun.}
 While this will play a role in the interpretation of the corpus findings, MNPs with plural kind nouns will not be included in the corpus study and the experiments reported here.
@@
 
-Second constructions with \textit{voll} (\textit{von})\slash\textit{voller} `full (of)' and \textit{mit} `with' are available, see (\ref{ex:alternatives}).
+Second, constructions with \textit{voll} (\textit{von})\slash\textit{voller} `full (of)' and \textit{mit} `with' are available, see (\ref{ex:alternatives}).
 
@@
 
-My whole analysis is based on the idea that the syntactic structure [N\Sub{1}~[A~N\Sub{2}]\Sub{NP\Sub{2}}]\Sub{NP\Sub{1}} as instantiated in the \NACa\ and the \PGCa\ is ambiguous.
+My analysis is based on the idea that the relevant syntactic structure [N\Sub{1}~[A~N\Sub{2}]\Sub{NP\Sub{2}}]\Sub{NP\Sub{1}} as instantiated in the \NACa\ and the \PGCa\ is ambiguous.
 To show this, the first question is which noun constitutes the head of the whole construction.
@@
 This unusual double nature of adjectives in NPs without determiners leads to a plausible probabilistic interpretation of the pattern shown in Table~\ref{tab:constructions}.
-Whenever speakers classify the adjective in the kind noun phrase more as a determiner, they have to use the \PGCa\ because, if there is a determiner, the PGC (just like in the case of the \PGCd) is the only option.
-When they classify the adjective more as an adjective, the kind NP has no determiner, and they have to use the \NACa\ (like in the case of the \NACb).%
+Whenever speakers classify the adjective in the kind noun phrase more as a determiner, they have to use the \PGCa\ because, if there is a determiner, the PGC is the only option.
+When they classify the adjective more as an adjective, the kind NP has no determiner, and they have to use the \NACa.%
 \footnote{While the generative analysis presented in \cite{Bhatt1990} cannot properly deal with probabilistic effects, Bhatt comes close to this interpretation by analysing the kind NP in the GPC as a DP and in the NAC as an NP.}
 
-This morpho-syntactic ambiguity means that the \NACa\ is in fact a \NACb in disguise, and the \PGCa\ is a \PGCd in disguise.
+This morpho-syntactic ambiguity means that the \NACa\ is in fact a \NACb\ in disguise, and the \PGCa\ is a \PGCd\ in disguise.
 This should ideally be reflected in the following basic selection effect:
@@
 It still has the potential to form a true partitive (if the kind noun is definite).
-Conversely, the NAC completely lacks this ability to form true partitives, and it forms a more restrictive environment with less combinatory and semantic flexibility.
+Conversely, the NAC completely lacks this ability to form true partitives, and it thus forms the more prototypical environment for expressing measurement (as opposed to partitivity).
 Hence, we can expect the \NACa\ to be a prototypical hosting construction for more strongly grammaticalised measure nouns.
@@
 
-Furthermore, the grammaticalisation path as described above leads from NPs denoting individuated objects standing in a \textit{part-of} relation to a construction with a more diffuse \textit{quantity-of} relation.
+As described above, the grammaticalisation path leads from NPs denoting individuated objects standing in a \textit{part-of} relation to a construction with a more diffuse \textit{quantity-of} relation.
 Both types of relations can be numerically quantified -- in as much as a precise number of \textit{parts} or a numerically exact \textit{quantity} can be specified --, but it is much more prototypical of quantities to be specified with numerical precision.
@@
 
-The statistical models presented in Section~\ref{sec:corpusstudies} will be specified in a way that they could detect such a preference.
+The statistical model presented in Section~\ref{sec:corpusstudies} will be specified in a way that it can detect such a preference.
 
-Finally, it has often been iterated that the \PGCa\ is more typical of higher registers or even exclusive to written language (see \citealp[320--323]{Hentschel1993}).
-This is not surprising in as much as the genitive -- an intrinsic part of the PGC -- is generally underrepresented in colloquial vernacular variants of German as a result of a diachronic process wherein many (but not all) uses of the genitive are replaced by other cases or periphrastic constructions \citep{FleischerSchallert2011}.
+Finally, it was found that the \PGCa\ is more typical of higher registers or even exclusive to written language (see \citealp[320--323]{Hentschel1993}).
+This is not surprising in as much as the genitive -- an intrinsic part of the PGC -- is generally underrepresented in colloquial vernacular variants of German as a result of a diachronic process wherein many (but by no means all) uses of the genitive are replaced by other cases or periphrastic constructions \citep{FleischerSchallert2011}.
 Under an integral view of prototypes, which incorporate effects related to larger contexts and registers, such preferences can be to be part of the construction prototypes.
@@
 To summarise the discussion and core hypotheses, I assume that the alternation is controlled by the similarity of the chosen lemmas (including the degree of grammaticalisation of measure nouns), certain morpho-syntactic choices, as well as the larger utterance context to two prototypes instantiated more straightforwardly by the two non-alternating cases of NAC and PGC.
-Concretely, I predict that:
+Concretely:
 
@@
 For the present study, I used the German \textit{Corpus from the Web} (COW) in its 2014 version DECOW14A (\citealp{SchaeferBildhauer2012full,Schaefer2015b} and \citealp{BiemannEa2013,SchaeferBildhauer2013} for overviews of web corpora in general and the methodology of their construction), which contains almost 21 billion tokens.%
-\footnote{The corpora are made available for free at \url{https://www.webcorpora.org}.
+\footnote{The COW corpora (Dutch, English, French, German, Spanish, Swedish) are made available for free at \url{https://www.webcorpora.org}.
 At the time of this writing, a newer 2016 version DECOW16 had already been released.}
@@
 Clearly, using search engine results is \textit{bad science} for many reasons, most prominently total irreproducibility of results, as \cite{Kilgarriff2006} pointed out more than ten years ago.
-Careless use of search engine results is still found, however, see \citet[171--175]{DeclerckBrems2016}.}
+Careless use of search engine results is still found, however, see for example \citet[171--175]{DeclerckBrems2016}.}
 First, the external validity of any study is increased through a higher heterogeneity of the sample \citep[30]{MaxwellDelaney2004}, and the DECOW corpus has clearly a much more heterogeneous composition compared to the only other very large corpus of German, the DeReKo \citep{KupietzEa2010} of the Institute for the German Language (IDS), which contains almost exclusively newspaper texts.%
@@
 \footnote{All numeric variables were also z-transformed (\ie\ centered to the mean and rescaled such that they have a standard deviation of $1$) to facilitate their interpretation in the regression models reported in the next section.}
-Notice that \textit{Construction} is the response variable (or `dependent variable') with the values \textit{PGCa} and \textit{NACa}.
+Notice first that \textit{Construction} is the response variable (or `dependent variable') with the values \textit{PGCa} and \textit{NACa}.
 
@@
 The higher the value, the more often the noun occurs in the \PGCa (proportionally).%
-\footnote{One could argue that some more advanced measure of attraction strength should be used, as is done in Collostructional (or Collexeme) Analysis \citep{GriesStefanowitsch2004}.
+\footnote{
+  It could be argued that some more advanced measure of attraction strength should be used, as is done in Collostructional (or Collexeme) Analysis \citep{GriesStefanowitsch2004}.
   Two main points speak against such an approach.
@@
   Their concrete numerical value thus does not have a solid cognitive interpretation.
-For the present study, it suffices to estimate the relative frequency with which speakers have encountered a specific lemma in the \NACa\ and the \PGCa.
 }
-\textit{Kindfreq} and \textit{Measurefreq} are the logarithm-transformed frequencies per 1,000,000 words of each lemma, extracted from the frequency lists distributed by the DECOW corpus creators on their web page.
+Additionally, \textit{Kindfreq} and \textit{Measurefreq} are the logarithm-transformed frequencies per 1,000,000 words of each lemma, extracted from the frequency lists distributed by the DECOW corpus creators on their web page.
+They were added to control for basic frequency effects.
 
@@
 Documents with higher Badness usually contain more incoherent language, shorter sentences, etc.
-If the PGC actually favours higher registers and styles, a high \textit{Badness} should be correlated with fewer occurrences of the PGC.
+If the \PGCa\ actually favours higher registers and styles, a high \textit{Badness} should be correlated with fewer occurrences.
 Documents in DECOW14 have also been annotated with a variable called \textit{Genitives}.
@@
 While more genitives are also indicative of higher registers, the use of this variable as a regressor in the present study might be considered problematic.
-Since the PGC contains a genitive itself, the regressor variable \textit{Genitive} and the document-level variable \textit{Genitives} are not fully independent.
-However, since the PGCs make up for only a minute fraction of all genitives, I still use \textit{Genitives} as a regressor with the appropriate caveats.
+Since the \PGCa\ contains a genitive itself, the regressor variable \textit{Genitive} and the document-level variable \textit{Genitives} are not fully independent.
+However, since instances of the \PGCa\ make up for only a minute fraction of all genitives, I still use \textit{Genitives} as a regressor with the appropriate caveats.
 
@@
 
-In this section, I report the results of fitting a multilevel model to the data using R \citep{R}, \textit{lme4} \citep{lme4} for Maximum Likelihood Estimation, and \textit{rstanarm} \citep{rstanarm} for Markov-Chain Monte Carlo estimation.
-The purpose is to model the influence of the regressors specified in Table~\ref{tab:variables} on the probability that either the \NACa\ or the \PGCa\ is chosen.
+In this section, I report the results of fitting a multilevel model to the data using R \citep{R}, \textit{lme4} \citep{lme4} for Maximum Likelihood Estimation, and \textit{rstanarm} \citep{rstanarm} for Markov-Chain Monte Carlo estimation (see Section~\ref{sec:rightstatistics}).
+The purpose is to model the influence of the regressors specified in Table~\ref{tab:variables} on the probability that either the \PGCa\ is chosen over the \NACa.
 All regressors from Table~\ref{tab:variables} were included, and the measure lemma and the kind noun lemma were specified as varying-intercept random effects.
@@
 The coefficient estimates are specified in Table~\ref{tab:bigtable} for each regressor (or regressor level) in the columns labelled \textit{Coefficient}.
-For a robust quantification of the precision of the estimation, I ran a parametric bootstrap (using the \textit{confint.merMod} function from \textit{lme4}) with $1,000$ replications, and using the percentile method for the calculation of the intervals.
+For a robust quantification of the precision of the estimation, I ran a parametric bootstrap (using the \mbox{\textit{confint.merMod}} function from \textit{lme4}) with $1,000$ replications, and using the percentile method for the calculation of the intervals.
 The resulting 95\% bootstrap confidence intervals are reported in Table~\ref{tab:bigtable} in the columns labelled \textit{CI low} and \textit{CI high} (= upper and lower 2.5th percentiles).
@@
 These bootstrapped $p$ values are given in the columns labelled p\Sub{PB} in Table~\ref{tab:bigtable}.
-Only \textit{Kindfreq} ($\mpPB=0.095$) can be seen as slightly too high to be convincing (`non-significant').
+Only \textit{Kindfreq} ($\mpPB=0.095$) can be seen as slightly too high to be convincing (non-significant).
 
@@
 The resulting coefficients and intervals as well as an * are also given in Table~\ref{tab:bigtable} in the columns labelled \textit{MCMC}.
-Both methods lead to exactly the same results (minus negligible numerical differences) as expected (see Section~\ref{sec:rightstatistics}).
+Both methods lead to exactly the same results (minus negligible numerical differences) as expected given the modest complexity of the model structure and the large sample size (see Section~\ref{sec:rightstatistics}).
 The signs and magnitudes of the coefficients are identical, and confidence intervals have the same width and symmetry properties.
@@
 An interesting picture emerges for the lemma frequencies.
-A higher than average lemma frequency of measure nouns favours the \NACa\ ($\beta_{\text{Measurefreq}}=-0.257$, $\mpPB=0.003$) as expected if we assume at least a tendency for highly grammaticalised items to be more frequent.
-With kind nouns, higher frequency seems to favour the \PGCa ($\beta_{\text{Kindfreq}}=0.161$, $\mpPB=0.066$).
-However, there is no clear theoretical interpretation (see Section~\ref{sec:analyses}), and the estimate is imprecise (`not significant at $\alpha=0.05$', see above).
+A higher than average lemma frequency of measure nouns favours the \NACa\ ($\beta_{\text{Measurefreq}}=-0.231$, $\mpPB=0.005$) as expected if we assume at least a tendency for highly grammaticalised items to be more frequent.
+With kind nouns, higher frequency seems to favour the \PGCa ($\beta_{\text{Kindfreq}}=0.146$, $\mpPB=0.095$).
+However, there is no clear theoretical interpretation (see Section~\ref{sec:analyses}), and the estimate is imprecise (not significant at $\alpha=0.05$, see above).
 The effect can therefore be ignored or treated as a nuisance variable.
@@
 The register-related proxy variables point into the expected direction.
-Increased \textit{Badness} of the document favours the \NACa\ ($\beta_{\text{Badness}}=-0.165$, $\mpPB=0.001$), and so does a lesser density of genitives ($\beta=-0.630$, $\mpPB=0.001$).
+Increased \textit{Badness} of the document favours the \NACa\ ($\beta_{\text{Badness}}=-0.152$, $\mpPB=0.002$), and so does a lesser density of genitives ($\beta=-0.693$, $\mpPB=0.001$).
 While these are poor proxies to register (and partially circular in the case of \textit{Genitives}), this result can at least encourage future work into register effects.
@@
 The influence of \textit{Measurecase} ($\mpPB=0.001$) is as predicted in previous analyses (see Section~\ref{sec:analyses}).
-A measure noun in the dative favours the \PGCa\ with $\beta_{\text{MeasurecaseDat}}=0.707$ (compared to the nominative, which is on the intercept).
+A measure noun in the dative favours the \PGCa\ with $\beta_{\text{MeasurecaseDat}}=0.705$ (compared to the nominative, which is on the intercept).
 Although \textit{Measurecase} is a nuisance variable in the context of this study, convergence with previous work strengthens its validity.
@@
 
-There were 24 participants (native speakers of German without reading or writing disabilities) aged 19 to 30 who were recruited from introductory linguistics courses at [name of university anonymised].
+There were 24 participants (native speakers of German without reading or writing disabilities) aged 19 to 30 living permanently in [name of city anonymised], who were recruited from introductory linguistics courses at [name of university anonymised].
 Although the experiment was conducted in the last four weeks of their first semester, participants had no deeper explicit knowledge of linguistics, grammar, or experimental methods.
@@
 There were $38$ (different) participants recruited in exactly the same manner as for the experiment reported in Section~\ref{sec:exp:fc}.
+None of them had ever participated in any kind of reading experiment.
 The experiment was conducted using \textit{PsychoPy}.
@@
 An LMM was specified with the logarithmised residual reading times as the response variable, and the probabilities derived from the corpus GLMM (\textit{Modelprediction}) as the main regressor.
-Since the corpus GLMM predicts the probability of the \PGCa (vs.\ the \NACa), it is expected that reading times are positively correlated (longer reading times) with the probability if the stimulus actually contains the \NACa, and negatively correlated (shorter reading times) if the stimulus actually contains the \PGCa.
+Since the corpus GLMM predicts the probability of the \PGCa\ (vis-à-vis the \NACa), it is expected that reading times are positively correlated (longer reading times) with the probability if the stimulus actually contains the \NACa, and negatively correlated (shorter reading times) if the stimulus actually contains the \PGCa.
 To account for this, an interaction between \textit{Modelprediction} and \textit{Construction} (levels \textit{PGCa} and \textit{NACa}) was added to the model.
@@
 
-The overall model quality is acceptable, and the effect plot for the main effect is shown in Figure~\ref{fig:spr:effects}.
-The estimate for the sentences with \NACa\ is obviously imprecise, although pointing into the right direction (longer reading times when the corpus model predicts the \PGCa).
-There is a clearer effect in the sentences with \PGCa, which is also confirmed by the `significant' results from the bootstrapped confidence intervals (see Table~\ref{tab:exp:spr}) and from the PB test reported above.
+The overall model quality is thus good, and the effect plot for the main effect is shown in Figure~\ref{fig:spr:effects}.
+The estimate for the sentences with \NACa\ is obviously imprecise, and no differences in reading times are observed.
+There is a clearer effect in the sentences with \PGCa, which is also confirmed by the significant results from the bootstrapped confidence intervals (see Table~\ref{tab:exp:spr}) and from the PB test reported above.
 The \PGCa\ brings about an increased reading time ($\beta_{\text{ConstructionPGCa}}=0.054$), which is plausible because it is the much rarer construction (see Section~\ref{sec:corpusstudies}).
 However, if it occurs in a prototypical context and with prototypical lexical material, reading times drop ($\beta_{\text{ConstructionPGCa:Modelprediction}}=-0.125$).
-This can be seen in the downward slope of curve in the right panel of Figure~\ref{fig:spr:effects}.
+This can be seen in the downward slope in the right panel of Figure~\ref{fig:spr:effects}.
 This fits into the general picture in as much as the construction with the lower frequency might be developing towards a more sharply defined prototype.%

### 2017-04-05 Latex/article.tex
DONEmake quick (with draft... minus abstract and conclusion)

@@
   \affil[1]{Freie Universität Berlin}
-  \title{Competing Constructions for German Measure Noun Phrases: from Usage Data to Experimental Validation}
+  \title{Competing Constructions for German Measure Noun Phrases: from Usage Data to Experiment}
   \runningtitle{Competing Constructions for German Measure Noun Phrases}
@@
 
-\subsection{Corpora and cognition}
-
-This paper deals with a morpho-syntactic alternation that occurs only in a very specific syntactic measure noun phrase construction in German.
+\subsection{Corpora, cognition, and alternations}
+
+This paper deals with a morpho-syntactic alternation between two constructions that occurs only in a very specific type of measure noun phrase in German.
 By \textit{alternation} I refer to a situation where two or more forms or constructions are available with no clear difference in acceptability, function, or meaning.
-
-\vspace{\baselineskip}
-
-\begin{itemize}
-  \item usage-based/cognitive corpus linguistics and the study of alternations
-  \item validationist turn; 2016 special issue CoGL, but Bresnan did it all the time!
-  \item also \cite{Gries2015a}
-  \item the value of case studies
-  \item `Zweifelsfälle' \citep{Klein2009} understood as ideal cases cor cognitive modeling
-  \item why forced choice and why reading times?
-\end{itemize}
+The study of alternations has a now longstanding tradition in cognitively oriented corpus linguistics (\citealp{BresnanEa2007,BresnanHay2010,BresnanFord2010,DivjakArppe2013,Gries2015,NessetJanda2010}, to name just a few publications).
+This strand of research is based on the assumption that language is a probabilistic phenomenon \cite{Bresnan2007}, and choices of variants are neither chosen deterministically nor fully at random.
+Instead, multifactorial models are constructed, which incorporate influencing factors from diverse levels, including contextual ones.
+The estimation of the model quantifies the influence the factors have on the probability that either variant is chosen.
+
+For such probabilistic generalisations estimated on corpus data, there has always been an interest in correlating the results with results from experimental work (for example, \citealp{ArppeJaervikivi2007,BresnanEa2007,BresnanFord2010,DivjakGries2008,DivjakEa2016,FordBresnan2013}).
+This often called a \textit{validation} of the corpus-derived findings, but \citet[303]{Divjak2016} criticises this choice of words \textit{because it creates the impression that behavioral experimental data is inherently more valuable than textual data}, citing \cite{TummersEa2005}, who state that a corpus is \textit{a sample of spontaneous language use that is} (\textit{generally}) \textit{realized by native speakers}.
+This position, which I adopt by and large, does not, however, imply that we can in some way \textit{deduce mental representations from patterns of use}, \ie\ from corpus data \citep[486--487]{Dabrowska2016}.%
+\footnote{I suggest that the word \textit{induce} would be more appropriate than \textit{deduce}.}
+In fact, it would be surprising if we could!
+Nobody assumes that we can induce mental representations from experiments, which -- as opposed to corpus studies -- even provide a direct access to the cognitive agent and offer much better options to control and experimental conditions and nuisance variables.
+Instead, a theory of cognitive representation is pre-specified, and predictions are derived from this theory \textit{before} the experiment or the corpus study is conducted.
+Technical problems of statistical inference aside (but see Section~\ref{sec:rightstatistics}), a successful experiment corroborates the theory, and a failed experiment weakens the theory.
+
+That said, the present study is conducted within the general paradigm of cognitive corpus linguistics and includes a comparison of the corpus findings with results from two experiments.
+As a theoretical framework, I assume a model of similarity-based classification in the sense of Prototype Theory.
+The approach within cognitive corpus linguistics (\eg\ \citealp{DivjakArppe2013,Gries2003}) assumes that the variables annotated in a corpus study on an alternation phenomenon define prototypes for the alternating variants.
+The similarity of a given feature vector (in a concrete sentence where one of the variants is used) determines to a large extent which variant is chosen by speakers.
+This entails that a variant of Prototype Theory with \textit{features} is assumed \citep{Rosch1978}, instead of the monolithic prototypes assumed in earlier versions.
+While Prototype Theory is well suited for modeling constructional choices, it is just one similarity-based framework, the most prominent other framework being Exemplar Theory (\citealp{MedinSchaffer1978,Hintzman1986}; see \citealp{StormsEa2000} for a comparison of the theories in experimental settings).
+Prototype Theory and Exemplar Theory differ mainly in whether they assume higher-level abstractions as part of the cognitive representation (Prototype Theory) or not (Exemplar Theory).
+I am sceptical whether corpus analysis alone could ever decide which theory is more adequate, given that substantial doubt has been voiced whether even experimental methods are ultimately able to do so \citep{Barsalou1990}.%
+\footnote{In \cite{DivjakArppe2013}, a fruitful attempt was made to show how both models form a converging picture based on corpus data.}
+Prototype Theory (with feature abstractions) is simply preferred here because it fits the established alternation modeling paradigm, which relies on features being weighted (in a sense) in statistical models.
+In Section~\ref{sec:analyses}, a prototype-theoretical parlance is therefore adopted.
+
+Before moving on to the description of the phenomenon and the corpus study in Sections~\ref{sec:descriptive} and \ref{sec:analyses}, I comment on the statistical methods used in the corpus study in Section~\ref{sec:rightstatistics} because the discussion is closely related to basic questions of the cognitive approach in corpus linguistics.
 
@@
 In this section, I motivate the choice of statistical models which I use later in Section~\ref{sec:corpusstudies}.
-Readers might think that the method used here for modeling grammatical alternations -- namely Hierarchical Logistic Regression\slash Generalised Linear Mixed Models (GLMMs) -- does not require much motivation.
-After all, GLMMs have been established as the major tool in the analysis of alternation phenomena (\citealp{BresnanEa2007,BresnanHay2010,BresnanFord2010,DivjakArppe2013,Gries2015,NessetJanda2010}, to name just a few publications).
+Readers might think that the method used here for modeling grammatical alternations -- namely (Hierarchical) Logistic Regression\slash Generalised Linear (Mixed) Models or GLM(M)s -- does not require much motivation.
+After all, GLMMs have been established as the major tool in the analysis of alternation phenomena.
+All studies mentioned at the outset of Section~\ref{sec:cogocl} use GLM(M)s \citep{BresnanEa2007,BresnanHay2010,BresnanFord2010,DivjakArppe2013,Gries2015,NessetJanda2010}.
 Over the past few years, however, modified or alternative methods have been proposed.
@@
 The basic distinction is a philosophical one and related to the concepts of \textit{direct} and \textit{inverse probability} (\eg\ \citealp{Senn2011}).
-Frequentists assume that models and parameters are fixed, for example a model specifiying that a coin is fair.
+Frequentists assume that models and parameters are fixed, for example a model specifying that a coin is fair.
 We can then calculate for observed data (for example a measurement of $3$ heads in an experiment with $10$ tosses) how often such a result or a more extreme result would occur if the model were true and we repeated the experiment arbitrarily often.
 This is essentially the frequentist notion of direct probability, \ie\ long-run frequencies under replication.
-What makes frequentism attractive is that it lends itself to a theory of testing whether a theoretical prediction (a fixed model) was successful or not.
-Such tests, if passed, contribute to the overall well-testedness of the theory.
-Bayesian approaches, on the other hand, condition on the data and quantify inductively (in the standard interpretation) the probability of models given specific observations.
-The models are thus not fixed, and the probability is usually equated with researchers' posterior beliefs about models or model parameters.
-There is actually a debate among Bayesians whether and how a notion of hypothesis testing is compatible (or even already contained) in the Bayesian approach.
-In \citet[10]{GelmanShalizi2013}, the authors -- prominent Bayesians -- acknowledge that theory testing is a desideratum, and they state about the standard inductive interpretation of Bayesianism that \textit{most of this received view of Bayesian inference is wrong}, and they develop a Bayesian notion of p values (see also \citealp{Mayo2013}, for a frequentist reply; also \citealp{Senn2011} on different strands of Bayesianism and their stance on inductive vs.\ deductive reasoning).
+Standard tests in the Fisher and Neyman-Pearson traditions as well as Neyman confidence intervals are based on this notion.
+Bayesian approaches (in the now commonly found interpretation), on the other hand, condition on the particular data and quantify inductively the probability of model parameters given the data.
+The parameters are thus not fixed, and the probability is usually equated with researchers' posterior beliefs about model parameters.
+There is actually a debate among Bayesians about the proper interpretation of Bayesian methods, and whether a notion of hypothesis testing is compatible (or even already contained) in the Bayesian approach.
+In \citet[10]{GelmanShalizi2013}, the authors -- prominent Bayesians themselves -- acknowledge that a theory of testing is a desideratum, and they state about the standard inductive interpretation of Bayesianism that \textit{most of this received view of Bayesian inference is wrong}, and they develop a Bayesian notion of p values (see also \citealp{Mayo2013}, for a frequentist reply; also \citealp{Senn2011} on different strands of Bayesianism and their stance on inductive vs.\ deductive reasoning).
 Clearly, in such quarrels between and among camps of philosophers of science and statisticians it is difficult for mere practitioners to take sides.
@@
 Her argument is part of a larger body of literature asking for cognitively plausible modeling techniques, for example Naive Discriminative Learning (NDL; \citealp{Baayen2011,BaayenEa2013,MilinEa2016,TheijssenEa2013}).
-Frequentist methods were, however, never designed as cognitive models, but as systems of inference for researchers (see above).
-The actual question that emerges here is how we interpret our statistical models (estimated on corpus data).
+Yet, neither frequentist nor Bayesian methods were conceived as cognitive models, but as systems of inference for researchers (see above, and see also \citealp[302]{Divjak2016a}).
+The fundamental question that lurks behind such arguments is how we interpret our statistical models (estimated on corpus data).
 Are they inductive models of cognitive representations that learners would infer from being exposed to the corpus data?%
@@
 Or are they tests of theories that are pre-specified and merely tested for predictive accuracy on linguistic output data contained in corpora.
-In the former case, we should definitely resort to methods like NDL.
+In the former case, we adopt a strong \textit{corpus as input} hypothesis and should definitely resort to methods like NDL.
 However, this would most likely require us to toss most previous work done in (cognitive) linguistics into the bin and abandoning all high-level generalisations that most existing studies have been based upon.
@@
 These methods then do not need to be any more cognitively plausible than an ANOVA used to analyse the results from an experiment.
-I view my own work (see Sections~\ref{sec:germanmeasurenps}--\ref{sec:externalvalidation}) in the tradition of testing theories that embrace high-level generalisations in a probabilistic framework.
+I view my own work (see Sections~\ref{sec:germanmeasurenps}--\ref{sec:experimental}) in the tradition of testing theories that embrace high-level generalisations, and I agree to provisionally use Likelihood methods (see below).
 I remain fully agnostic with respect to the question which approach is `right'.
@@
 Now, Maximum Likelihood Estimation (MLE) -- the traditional method which could have been used instead -- is not exactly \textit{frequentist} in the sense of Neyman-Pearson testing theory.
-MLE also conditions on the data in as much as it finds the most likely set of parameters given the data.
-What is more, Bayesian estimators in fact incorporate the Likelihood function and merely multiply it by the prior \citep[PAGENUMBER??]{GelmanEa2014}.
+MLE, like inductive Bayesianism, conditions on the data in as much as it searches for the most likely set of parameters given the data.
+What is more, Bayesian estimators are in fact based the Likelihood function and merely multiply it by the prior \citep[6--8]{GelmanEa2014}.
 If the prior is flat, results converge (see also \citealp[347]{GelmanHill2006}).
 The same is true if the sample size is large compared to the number of parameters, at least for finite-dimensional models \citep[1119--1120]{Freedman1999}, a well-established result known as the Bernstein-von Mises theorem.
-With a modest model structure including $17$ fixed effects and $2,646$ data points in \cite{Levshina2016}, it is highly likely that the same results would have obtained with Maximum Likelihood methods.%
-\footnote{In fact, she admits that changing the priors did not lead to substantially different results in her footnote 8.
-  This is a clear sign that the prior is \textit{swamped by the data} (QUOTE??).
-  Relating back to cognitive plausibility, this means that in the case at hand Bayesian methods cannot be more cognitively plausible than Maximum Likelihood Estimation.}
-If not, especially in case of convergence problems, this would have been an ideal occasion to demonstrate the superiority of Bayesian estimation.
+With a modest model structure including $17$ fixed effects and $2,646$ data points in \cite{Levshina2016}, it is highly likely that the same results would have obtained with Maximum Likelihood methods.
+In fact, she admits that changing the priors did not lead to substantially different results in her footnote 8.
+This is a clear sign that the prior is \textit{swamped} by the data \citep[1119]{Freedman1999}.
+Going back to the argument that Bayesian methods are allegedly more cognitively plausible than `frequentist' methods, the claim might still be true.
+However, most similar studies (back to \citealp{BresnanEa2007}) use an estimator (MLE) that is not frequentist, and most likely as cognitively plausible as Bayesian estimators.
+If for Levshina's study, results did, in fact, \textit{not} converge, it would have been an ideal occasion to demonstrate the superiority of the algorithms used in Bayesian estimation.
 After all, there are situations where Bayesian estimators can be more robust, namely with heavily censored data, complex hierarchical models, perfect separation, etc.
 (see \citealp{Freedman1999}, \citealp[345--348]{GelmanHill2006}).
-I want to reiterate that these points do not in any way invalidate the study presented in \cite{Levshina2016}.
-However, being \textit{Bayesian} is not among its selling points.
-That said, I want to voice the concern that most likely many practitioners are already struggling with getting an adequate grasp of advanced statistical methods, and that it might therefore be wise to use the more conservative and better understood method if the alternative method is not absolutely required for substantive reasons.
+I want to reiterate that these points do not in any way invalidate the results presented in \cite{Levshina2016}.
+However, being \textit{Bayesian} is most likely not among its selling points.
+That said, I want to voice the concern that probably many practitioners are already struggling with getting an adequate grasp of advanced statistical methods, and that it might therefore be wise to use the more conservative and better understood method if the alternative method is not absolutely required for substantive reasons.
 In Section~\ref{sec:corpushierarchicalmodel}, I compare my own hierarchical models estimated with Bayesian and MLE estimators to demonstrate their expectable convergence.
@@
 In the remainder of the paper, I introduce the alternation phenomenon (Section~\ref{sec:descriptive}) and suggest a theory-driven set of factors influencing the alternation (Section~\ref{sec:analyses}).
-Then, the corpus study is presented including statistical analysis (Section~\ref{sec:corpusstudies}).
-The two experiments validating the corpus study are then reported (Section~\ref{sec:externalvalidation}) before I sum up the paper in Section~\ref{sec:conclusion}.
+Then, the corpus study is presented including an appropriate statistical analysis (Section~\ref{sec:corpusstudies}).
+The two experiments which correlate results from the corpus study with experimental results are then reported (Section~\ref{sec:experimental}) before I sum up the paper in Section~\ref{sec:conclusion}.
 
@@
 This section briefly reviews existing analyses of the \PGCa vs.\ \NACa\ alternation and related issues.
-I also develop my own analysis and the appropriate hypotheses for the empirical studies presented in Sections~\ref{sec:corpusstudies} and \ref{sec:externalvalidation}.
+I also develop my own analysis and the appropriate hypotheses for the empirical studies presented in Sections~\ref{sec:corpusstudies} and \ref{sec:experimental}.
 I discuss four main aspects:
@@
 In the next section, I report the corpus study that was designed to test these hypotheses on usage data.
-The resulting models will then be validated using experimental methods in Section~\ref{sec:externalvalidation}.
+The resulting models will then be compared to experimental methods in Section~\ref{sec:experimental}.
 
@@
 \footnote{
-  It could be argued that some more advanced measure of attraction strength should be used, as is done in Collostructional (or Collexeme) Analysis \citep{GriesStefanowitsch2004}.
-  Two main points speak against such an approach.
-  First, the attraction values will be used as regressors in a GLMM, and the values resulting from collostructional approaches, \ie\ logarithmised Fisher p values, have a very unfavourable distribution for such a use.
-  They cluster around $-\infty$ and $0$, especially when (as in this case) cells contain $0$ count values.
-  This cannot be remedied on principled grounds, since, for example, a z-transformation is inadequate for such a strongly bimodal distribution (even if we clamp $-\infty$ to a very low numeric value).
-  Second, their main use in collostructional analyses is to \textit{sort} a list of collexemes and interpret the order.
-  Their concrete numerical value thus does not have a solid cognitive interpretation.
+  It could be argued that some more advanced measure of attraction strength should be used, as is done in Collostructional (or Collexeme) Analysis \citep{GriesStefanowitsch2004}, see also \cite{Gries2015a}.
+  Three main points speak against such an approach in the present case (but see \citealp[246--249]{Levshina2016} for a successful more refined approach).
+  First, the attraction values will be used as regressors in a GLMM, and the values resulting from collostructional approaches, \ie\ logarithmised Fisher p values, have a very unfavourable distribution in the case at hand.
+  They cluster around $0$, and they include values of $-\infty$.
+  Second, their main use in collostructional analyses is to \textit{sort} a list of collexemes and interpret their order.
+  Their concrete numerical value might not have a solid cognitive interpretation in the given context.
+  Third, I tried using collexeme strength as a regressor, and the results were unsatisfactory.
 }
@@
 
-\section{Experimental validation}
-\label{sec:externalvalidation}
+\section{Correlating corpus findings with experiments}
+\label{sec:experimental}
+
+\subsection{Preliminary remarks}
+
+In the two experiments reported in this section, I use probabilities for the alternating constructions calculated for attested material, and I correlate these probabilities with the participants' reactions.
+Thus, a direct link can be established between output material found in corpora and the behaviour of linguistic agents.
+The first experiment tests preferences for constructions explicitly.
+\cite{FordBresnan2013} use the \textit{split-100} task in which participants have to distribute $100$ points between the alternatives, assigning more points to more natural sounding alternative.
+In essence, participants distribute a probability mass between two variants, which is intended to produce more subtle results compared to what is essentially a two-alternative forced-choice task \cite{Rosenbach2013}.
+The split-100 paradigm has been criticised in \cite{ArppeJaervikivi2007}, reiterated in \cite{DivjakEa2016}.
+In \cite{VerhoevenTemme2017}, it was shown that results from forced-choice and split-100 experiments mostly converge, which is expected in the long run if a probability is turned into a binary decision.
+I present a forced-choice experiment in Section~\ref{sec:exp:fc} and not a split-100 experiment, mainly because in a dry-run of the experiment, participants complained about the unnaturalness of the task of assigning a probability mass to two variants.
+The analysis compares the probabilities assigned to stimuli by the corpus-derived model with the frequency with which participants chose the variants for the same stimuli.
+
+The second experiment tests preferences more implicitly.
+It is expected that reading less prototypical variants (the one assigned a low probability by the corpus-derived model) in a given context and with given lexical material incurs a processing overhead for the reader (\citealp{Kaiser2013}).
+In Section~\ref{sec:exp:spr}, a self-paced reading experiment is therefore presented.
+In a very similar fashion, \cite{DivjakEa2016} apply the self-paced reading paradigm in the validation of corpus-based models.
+The analysis compares the corpus-derived probabilities with potential lags in reading time for sentences with the preferred and the non-preferred constructions.
 
@@
 
-As was pointed out in Section \ref{sec:cogocl}, there is a strong interest in validating corpus-based findings through experiments in order to substantiate \textit{cognitively oriented usage-based corpus linguistics} as a research programme.
-Therefore, this section and the next present the results of two experiments wherein I cross-checked the corpus-based findings.
+As was pointed out in Section \ref{sec:cogocl}, there is a strong interest in combining corpus-based findings with experiments.
+Therefore, this section and the next present the results of two experiments wherein I compare results from the corpus study with experimental results.
 Both experiments use sentences containing attested MNPs from the corpus sample (embedded into simplified sentences) as stimuli.
@@
 The higher the probability of the \PGCa\ predicted from usage-data, the more often participants chose the \PGCa\ variant in the forced-choice task.
-In summary, the forced choice experiment clearly succeeded in validating the results from the corpus study in as much as the preferences extracted from usage data correspond to native speakers' choices.
+In summary, the forced choice experiment clearly succeeded in corroborating the results from the corpus study in as much as the preferences extracted from usage data correspond to native speakers' choices.
 
@@
 Finally, I report the results of a self-paced reading experiment also designed to test whether the statistical model derived from usage data (see Section~\ref{sec:corpushierarchicalmodel}) can be used to predict speakers' behaviour.
-It is expected that reading the less prototypical variant (the one assigned a low probability by the corpus-derived model) in a given context and with given lexical material incurs a processing overhead for the reader (\citealp{Kaiser2013}; see also Section~\ref{sec:cogocl}).
-In a very similar fashion, \cite{DivjakEa2016} apply the self-paced reading paradigm in the validation of corpus-based models.
-
 I used exactly the same stimuli as in the forced choice experiments.

### 2017-04-06 Latex/article.tex
Even more DONEmake quick

@@
   \abstract{}
-  \keywords{corpus methods and experimental methods, self-paced reading, forced choice, alternations, hierarchical models, measure constructions, German}
+  \keywords{corpus methods and experimental methods, self-paced reading, forced-choice, alternations, hierarchical models, measure constructions, German}
   %\classification[PACS]{...}
@@
 For such probabilistic generalisations estimated on corpus data, there has always been an interest in correlating the results with results from experimental work (for example, \citealp{ArppeJaervikivi2007,BresnanEa2007,BresnanFord2010,DivjakGries2008,DivjakEa2016,FordBresnan2013}).
-This often called a \textit{validation} of the corpus-derived findings, but \citet[303]{Divjak2016} criticises this choice of words \textit{because it creates the impression that behavioral experimental data is inherently more valuable than textual data}, citing \cite{TummersEa2005}, who state that a corpus is \textit{a sample of spontaneous language use that is} (\textit{generally}) \textit{realized by native speakers}.
+This often called a \textit{validation} of the corpus-derived findings, but \citet[303]{Divjak2016a} criticises this choice of words \textit{because it creates the impression that behavioral experimental data is inherently more valuable than textual data}, citing \cite{TummersEa2005}, who state that a corpus is \textit{a sample of spontaneous language use that is} (\textit{generally}) \textit{realized by native speakers}.
 This position, which I adopt by and large, does not, however, imply that we can in some way \textit{deduce mental representations from patterns of use}, \ie\ from corpus data \citep[486--487]{Dabrowska2016}.%
@@
 Technical problems of statistical inference aside (but see Section~\ref{sec:rightstatistics}), a successful experiment corroborates the theory, and a failed experiment weakens the theory.
+The simplest assumption would be that the same approach is adequate for corpus studies.
 
 That said, the present study is conducted within the general paradigm of cognitive corpus linguistics and includes a comparison of the corpus findings with results from two experiments.
-As a theoretical framework, I assume a model of similarity-based classification in the sense of Prototype Theory.
+As a theoretical framework, I assume a model of similarity-based classification in the form of Prototype Theory.
 The approach within cognitive corpus linguistics (\eg\ \citealp{DivjakArppe2013,Gries2003}) assumes that the variables annotated in a corpus study on an alternation phenomenon define prototypes for the alternating variants.
@@
 Prototype Theory and Exemplar Theory differ mainly in whether they assume higher-level abstractions as part of the cognitive representation (Prototype Theory) or not (Exemplar Theory).
-I am sceptical whether corpus analysis alone could ever decide which theory is more adequate, given that substantial doubt has been voiced whether even experimental methods are ultimately able to do so \citep{Barsalou1990}.%
+I am sceptical that corpus analysis alone could ever decide which theory is more adequate, given that substantial doubt has been voiced whether even experimental methods are ultimately able to do so \citep{Barsalou1990}.%
 \footnote{In \cite{DivjakArppe2013}, a fruitful attempt was made to show how both models form a converging picture based on corpus data.}
@@
 
-\subsection{A word on statistical analysis}
+\subsection{On statistical analysis}
 \label{sec:rightstatistics}
@@
 After all, GLMMs have been established as the major tool in the analysis of alternation phenomena.
-All studies mentioned at the outset of Section~\ref{sec:cogocl} use GLM(M)s \citep{BresnanEa2007,BresnanHay2010,BresnanFord2010,DivjakArppe2013,Gries2015,NessetJanda2010}.
+All studies mentioned at the outset of Section~\ref{sec:cogocl} use GLM(M)s \citep{BresnanEa2007,BresnanHay2010,BresnanFord2010,DivjakArppe2013,Gries2015,NessetJanda2010} use some form of regression\slash GLM(M).
 Over the past few years, however, modified or alternative methods have been proposed.
-From among these methods, I just make a few remarks on Bayesian estimation (see \citealp{GelmanEa2014}), as it was proposed in \cite{Levshina2016,Divjak2016a}.
+From among these methods, I just make a few remarks on Bayesian estimation (see \citealp{GelmanEa2014}), as it was proposed in \cite{Levshina2016,Divjak2016a}, for example.
 Conceptually, I see three points of discussion that should be kept apart.
@@
 Standard tests in the Fisher and Neyman-Pearson traditions as well as Neyman confidence intervals are based on this notion.
-Bayesian approaches (in the now commonly found interpretation), on the other hand, condition on the particular data and quantify inductively the probability of model parameters given the data.
+Bayesian approaches (in the now common interpretation), on the other hand, condition on the particular data and quantify inductively the probability of model parameters given the data.
 The parameters are thus not fixed, and the probability is usually equated with researchers' posterior beliefs about model parameters.
 There is actually a debate among Bayesians about the proper interpretation of Bayesian methods, and whether a notion of hypothesis testing is compatible (or even already contained) in the Bayesian approach.
-In \citet[10]{GelmanShalizi2013}, the authors -- prominent Bayesians themselves -- acknowledge that a theory of testing is a desideratum, and they state about the standard inductive interpretation of Bayesianism that \textit{most of this received view of Bayesian inference is wrong}, and they develop a Bayesian notion of p values (see also \citealp{Mayo2013}, for a frequentist reply; also \citealp{Senn2011} on different strands of Bayesianism and their stance on inductive vs.\ deductive reasoning).
+In \citet[10]{GelmanShalizi2013}, the authors -- prominent Bayesians themselves -- acknowledge that a theory of testing is a desideratum, and they state about the standard inductive interpretation of Bayesianism that \textit{most of this received view of Bayesian inference is wrong}, and they develop a Bayesian notion of $p$ values (see also \citealp{Mayo2013}, for a frequentist reply; also \citealp{Senn2011} on different strands of Bayesianism and their stance on inductive vs.\ deductive reasoning).
 Clearly, in such quarrels between and among camps of philosophers of science and statisticians it is difficult for mere practitioners to take sides.
@@
 Her argument is part of a larger body of literature asking for cognitively plausible modeling techniques, for example Naive Discriminative Learning (NDL; \citealp{Baayen2011,BaayenEa2013,MilinEa2016,TheijssenEa2013}).
-Yet, neither frequentist nor Bayesian methods were conceived as cognitive models, but as systems of inference for researchers (see above, and see also \citealp[302]{Divjak2016a}).
+On p.\ 303 of \cite{Divjak2016a}, the author explicitly mentions NDL.
+Yet, neither frequentist nor Bayesian methods were conceived as cognitive models, but as systems of inference for scientists (see above, and see also \citealp[302]{Divjak2016a}).
 The fundamental question that lurks behind such arguments is how we interpret our statistical models (estimated on corpus data).
-Are they inductive models of cognitive representations that learners would infer from being exposed to the corpus data?%
+Are they inductive models of cognitive representations that also human learners would infer from being exposed to the corpus data?%
 \footnote{In which case we are doing \textit{data science in language research} in the words of \citealp{MilinEa2016}.}
@@
 This does not do justice to frequentist methods in that mere rejection of the null hypothesis is characteristic only of Fisher's approach, but in the Neyman-Pearson approach, results ideally \textit{favour} the main hypothesis vis-à-vis the alternative hypothesis (cf.\ \citealp{Lehmann1993,Lehmann2011,Perezgonzalez2015}).
+Also, especially Neyman-style frequentism has well-known extensions to estimation (see \citealp{GreenlandEa2016}, esp.\ p.\ 340).
 She then explains that a \textit{distinctive feature of Bayesian statistics is the use of so-called priors} and that \textit{posterior probabilities depend on both the prior beliefs and the data, whereas the results of a frequentist model depend only on the data} \cite[252]{Levshina2016}.
-Remarkably, however, she does \textit{not} use informative priors, and in her footnote 8 \cite[252]{Levshina2016} admits that priors were probed using trial and error, such that the proclaimed advantage of Bayesian modeling was apparently not taken advantage of.%
+Remarkably, given this statement, she does \textit{not} use informative priors, and in her footnote 8 \cite[252]{Levshina2016} admits that priors were probed using trial and error.
+So, the proclaimed major advantage of Bayesian modeling was apparently not taken advantage of.%
 \footnote{In the words of \cite{Senn2011}: \textit{You may believe you are a Bayesian but you are probably wrong}.
@@
 Now, Maximum Likelihood Estimation (MLE) -- the traditional method which could have been used instead -- is not exactly \textit{frequentist} in the sense of Neyman-Pearson testing theory.
-MLE, like inductive Bayesianism, conditions on the data in as much as it searches for the most likely set of parameters given the data.
+MLE, like inductive Bayesianism, conditions on the data in as much as it searches for the most likely set of parameters given the particular data.
 What is more, Bayesian estimators are in fact based the Likelihood function and merely multiply it by the prior \citep[6--8]{GelmanEa2014}.
@@
 Going back to the argument that Bayesian methods are allegedly more cognitively plausible than `frequentist' methods, the claim might still be true.
-However, most similar studies (back to \citealp{BresnanEa2007}) use an estimator (MLE) that is not frequentist, and most likely as cognitively plausible as Bayesian estimators.
-If for Levshina's study, results did, in fact, \textit{not} converge, it would have been an ideal occasion to demonstrate the superiority of the algorithms used in Bayesian estimation.
+However, most similar studies (back to \citealp{BresnanEa2007}) use an estimator (MLE) that is not frequentist, and as cognitively plausible as Bayesian estimators because it most likely produces the same results in these studies.
+If for Levshina's study, results from Bayesian and MLE methods did, in fact, \textit{not} converge, it would have been an ideal occasion to demonstrate the superiority of the algorithms used in Bayesian estimation.
 After all, there are situations where Bayesian estimators can be more robust, namely with heavily censored data, complex hierarchical models, perfect separation, etc.
@@
 second, degrees of grammaticalisation of the measure noun associated with the two prototypes;
-third, a semantic prototypicality effect of the degrees of grammaticalisation (preference to occur with cardinals);
+third, a semantic prototypicality effect of the degrees of grammaticalisation (a preference to occur with cardinals);
 fourth, preferences in different registers.
@@
   Three main points speak against such an approach in the present case (but see \citealp[246--249]{Levshina2016} for a successful more refined approach).
-  First, the attraction values will be used as regressors in a GLMM, and the values resulting from collostructional approaches, \ie\ logarithmised Fisher p values, have a very unfavourable distribution in the case at hand.
+  First, the attraction values will be used as regressors in a GLMM, and the values resulting from collostructional approaches, \ie\ logarithmised Fisher $p$ values, have a very unfavourable distribution in the case at hand.
   They cluster around $0$, and they include values of $-\infty$.
@@
 Furthermore, for each regressor, a $p$ value was obtained by dropping the regressor from the full model, re-estimating the nested model and comparing it to the full model.
-Instead of inexact Wald approximations or Likelihood Ratio Tests, I used a drop-in bootstrap replacement for the Likelihood Ratio Test from the function \textit{PBmodcomp} from the \textit{pbkrtest} package \citep{HalekohHojsgaard2014}, hence I call the corresponding value $p_{\text{PB}}$.
-These bootstrapped $p$ values are given in the columns labelled p\Sub{PB} in Table~\ref{tab:bigtable}.
+Instead of inexact Wald approximations or Likelihood Ratio Tests, I used a drop-in bootstrap replacement for the Likelihood Ratio Test from the function \textit{PBmodcomp} from the \textit{pbkrtest} package \citep{HalekohHojsgaard2014}.
+I call the corresponding value $p_{\text{PB}}$, and it is given in the appropriately labeled columns in Table~\ref{tab:bigtable}.
 Only \textit{Kindfreq} ($\mpPB=0.095$) can be seen as slightly too high to be convincing (non-significant).
@@
 
-\subsection{Preliminary remarks}
+\subsection{Experiment 1: forced-choice}
+\label{sec:exp:fc}
 
 In the two experiments reported in this section, I use probabilities for the alternating constructions calculated for attested material, and I correlate these probabilities with the participants' reactions.
-Thus, a direct link can be established between output material found in corpora and the behaviour of linguistic agents.
+Thus, a direct link can be established between output material found in corpora and the behaviour of linguistic agents (see also Section~\ref{sec:cogocl}).
+Both experiments use sentences containing attested MNPs from the corpus sample (embedded into simplified sentences) as stimuli.
+Also, the probabilities that the corpus-based model assigns to the two variants in these sentences is used as the main regressor in both studies.
+
 The first experiment tests preferences for constructions explicitly.
 \cite{FordBresnan2013} use the \textit{split-100} task in which participants have to distribute $100$ points between the alternatives, assigning more points to more natural sounding alternative.
-In essence, participants distribute a probability mass between two variants, which is intended to produce more subtle results compared to what is essentially a two-alternative forced-choice task \cite{Rosenbach2013}.
-The split-100 paradigm has been criticised in \cite{ArppeJaervikivi2007}, reiterated in \cite{DivjakEa2016}.
+In essence, participants distribute a probability mass between two variants, which is intended to produce more subtle results compared to a two-alternative forced-choice task \cite{Rosenbach2013}, where participants have to choose one of two variants.
+The split-100 paradigm has been criticised in \cite{ArppeJaervikivi2007}.
+The criticism was reiterated in \cite{DivjakEa2016}, where they use a forced-choice task.
 In \cite{VerhoevenTemme2017}, it was shown that results from forced-choice and split-100 experiments mostly converge, which is expected in the long run if a probability is turned into a binary decision.
-I present a forced-choice experiment in Section~\ref{sec:exp:fc} and not a split-100 experiment, mainly because in a dry-run of the experiment, participants complained about the unnaturalness of the task of assigning a probability mass to two variants.
+I present a forced-choice experiment in Section~\ref{sec:exp:fc} and not a split-100 experiment, mainly because in a dry-run of the experiment, participants complained about the unnaturalness of the task of distributing a probability mass across two variants.
+Participants had to chose between two sentences differing only in that one contained the \NACa\ and the other contained the \PGCa.
 The analysis compares the probabilities assigned to stimuli by the corpus-derived model with the frequency with which participants chose the variants for the same stimuli.
-
-The second experiment tests preferences more implicitly.
-It is expected that reading less prototypical variants (the one assigned a low probability by the corpus-derived model) in a given context and with given lexical material incurs a processing overhead for the reader (\citealp{Kaiser2013}).
-In Section~\ref{sec:exp:spr}, a self-paced reading experiment is therefore presented.
-In a very similar fashion, \cite{DivjakEa2016} apply the self-paced reading paradigm in the validation of corpus-based models.
-The analysis compares the corpus-derived probabilities with potential lags in reading time for sentences with the preferred and the non-preferred constructions.
-
-\subsection{Experiment 1: forced choice}
-\label{sec:exp:fc}
-
-As was pointed out in Section \ref{sec:cogocl}, there is a strong interest in combining corpus-based findings with experiments.
-Therefore, this section and the next present the results of two experiments wherein I compare results from the corpus study with experimental results.
-Both experiments use sentences containing attested MNPs from the corpus sample (embedded into simplified sentences) as stimuli.
-Also, the probabilities that the corpus-based model assigns to the two variants in these sentences is used as the main regressor in both studies.
-First, a forced-choice experiment was conducted.
-Participants had to chose between two sentences differing only in that one contained the \NACa\ and the other contained the \PGCa.
 
@@
 The higher the probability of the \PGCa\ predicted from usage-data, the more often participants chose the \PGCa\ variant in the forced-choice task.
-In summary, the forced choice experiment clearly succeeded in corroborating the results from the corpus study in as much as the preferences extracted from usage data correspond to native speakers' choices.
+In summary, the forced-choice experiment clearly succeeded in corroborating the results from the corpus study in as much as the preferences extracted from usage data correspond to native speakers' choices.
 
@@
 
-Finally, I report the results of a self-paced reading experiment also designed to test whether the statistical model derived from usage data (see Section~\ref{sec:corpushierarchicalmodel}) can be used to predict speakers' behaviour.
-I used exactly the same stimuli as in the forced choice experiments.
+The second experiment tests preferences more implicitly.
+It is expected that reading less prototypical variants (the one assigned a low probability by the corpus-derived model) in a given context and with given lexical material incurs a processing overhead for the reader (\citealp{Kaiser2013}).
+In Section~\ref{sec:exp:spr}, a self-paced reading experiment is therefore presented.
+In a very similar fashion, \cite{DivjakEa2016} apply the self-paced reading paradigm in the validation of corpus-based models.
+The analysis compares the corpus-derived probabilities with potential lags in reading time for sentences with the preferred and the non-preferred constructions.
+
+Conretely, the exact same stimuli as in the forced-choice experiments were used.
 Each participant read both the $16$ sentences with the variant predicted by the corpus model and the $16$ modified sentences with the variant that the corpus model does not predict.%
@@
 There were $38$ (different) participants recruited in exactly the same manner as for the experiment reported in Section~\ref{sec:exp:fc}.
-None of them had ever participated in any kind of reading experiment.
+None of them had ever participated in any kind of reading experiment, and none of them took part in the first experiment (and vice versa).
 The experiment was conducted using \textit{PsychoPy}.
@@
 Conversely, the \NACa\ (like the NAC in general) might be the highly frequent default which does not incur reading time penalty, even if it is not the optimal choice in the given context and with the given lexical material.
-This interpretation will be considered, among other things, in the next and final section.
+
+This concludes the report of the two experiments.
+In the final section, I take stock and summarise the contribution of the present study to the research on alternations in cognitive linguistics.
 
@@
 
-I have shown!
-Oh, yes, I have shown\ldots
+This paper stands in a now ten year-old tradition of research on grammatical alternations using corpus and experimental data.
+In my view, the main tenets of this line of research are:
+(i)~Language, viewed from a cognitively realistic angle, is a probabilistic phenomenon and cannot be modeled appropriately within Aristotelian frameworks that assume discrete and clearly delimited categories.
+(ii)~Corpora are collections of usage events (language production) and can therefore be used to evaluate both the claim made in (i) and specific theoretical claims (in case studies) about factors influencing speakers' decisions to use specific forms or constructions.
+(iii)~Given (ii), we expect results from corpus analyses and from appropriate experiments to yield similar results, not necessarily as a form of validation of the corpus-based findings, but as converging evidence.
+I consider these three points to be of utmost importance because they clearly set this brand of cognitive linguistics apart from \textit{both} Aristotelian frameworks of the generative flavour \textit{and} introspective, non-empirical, and anti-quantitative versions of cognitive linguistics (see \citealp{Dabrowska2016} for an impressive philippic against such approaches).
+
+The present paper adds to the evidence that all of the aforementioned three points are correct.
+A grammatical alternation in German measure NPs was examined using corpus data and based on factors partly derived from existing accounts and formulated in terms of construction prototypes.
+The preferences extracted from the DECOW web corpus were confirmed in a forced-choice experiment, in which participants explicitly chose variants in line with the corpus-derived model.
+In a more implicit self-paced reading experiment, it was shown that the much rarer variant brings about a reading time penalty except in cases for which the corpus model predicts very high probability for this variant.
+
+Future work could extend these results and provide a general picture of the constructions expressing measurements (see Section~\ref{sec:descriptive}).
+This would be a much more complicated task given that the choices then would no longer be binary, and that the meaning of the alternative ways of expressing measurements are semantically more varied.
+Finally, I want to point out that German is mildly under-researched in the specific framework used here.
+This is quite surprising given the fact that German morpho-syntax is famous for its alternations, which are usually called \textit{Zweifelsfälle} (`cases of doubt') in the traditional literature \citep{Klein2009}.
+Instead of drowning them in normative, descriptive, or didactic discussions, they could serve as ideal test cases in cognitive linguistics.
 

### 2017-04-06 Latex/article.tex
Draft version for colleagues.

@@
 The study of alternations has a now longstanding tradition in cognitively oriented corpus linguistics (\citealp{BresnanEa2007,BresnanHay2010,BresnanFord2010,DivjakArppe2013,Gries2015,NessetJanda2010}, to name just a few publications).
-This strand of research is based on the assumption that language is a probabilistic phenomenon \cite{Bresnan2007}, and choices of variants are neither chosen deterministically nor fully at random.
+This strand of research is based on the assumption that language is a probabilistic phenomenon \citep{Bresnan2007}, and choices of variants are neither chosen deterministically nor fully at random.
 Instead, multifactorial models are constructed, which incorporate influencing factors from diverse levels, including contextual ones.
@@
 This often called a \textit{validation} of the corpus-derived findings, but \citet[303]{Divjak2016a} criticises this choice of words \textit{because it creates the impression that behavioral experimental data is inherently more valuable than textual data}, citing \cite{TummersEa2005}, who state that a corpus is \textit{a sample of spontaneous language use that is} (\textit{generally}) \textit{realized by native speakers}.
-This position, which I adopt by and large, does not, however, imply that we can in some way \textit{deduce mental representations from patterns of use}, \ie\ from corpus data \citep[486--487]{Dabrowska2016}.%
+This position, which I adopt, does not, however, imply that we can in some way \textit{deduce mental representations from patterns of use}, \ie\ from corpus data \citep[486--487]{Dabrowska2016}.%
 \footnote{I suggest that the word \textit{induce} would be more appropriate than \textit{deduce}.}
@@
 Instead, a theory of cognitive representation is pre-specified, and predictions are derived from this theory \textit{before} the experiment or the corpus study is conducted.
-Technical problems of statistical inference aside (but see Section~\ref{sec:rightstatistics}), a successful experiment corroborates the theory, and a failed experiment weakens the theory.
+Technical problems of statistical inference aside (see also Section~\ref{sec:rightstatistics}), a successful experiment corroborates the theory, and a failed experiment weakens the theory.
 The simplest assumption would be that the same approach is adequate for corpus studies.
@@
 As a theoretical framework, I assume a model of similarity-based classification in the form of Prototype Theory.
-The approach within cognitive corpus linguistics (\eg\ \citealp{DivjakArppe2013,Gries2003}) assumes that the variables annotated in a corpus study on an alternation phenomenon define prototypes for the alternating variants.
+Protoype Theory has been used in alternation research in cognitively oriented corpus linguistics, see \cite{DivjakArppe2013,Gries2003}.
+It is assumed that the variables annotated in a corpus study on an alternation phenomenon define prototypes for the alternating variants.
 The similarity of a given feature vector (in a concrete sentence where one of the variants is used) determines to a large extent which variant is chosen by speakers.
@@
 This is essentially the frequentist notion of direct probability, \ie\ long-run frequencies under replication.
-Standard tests in the Fisher and Neyman-Pearson traditions as well as Neyman confidence intervals are based on this notion.
+Standard tests in the Fisher and Neyman-Pearson traditions as well as Neyman confidence intervals are based on this concept of probability.
 Bayesian approaches (in the now common interpretation), on the other hand, condition on the particular data and quantify inductively the probability of model parameters given the data.
@@
 Her argument is part of a larger body of literature asking for cognitively plausible modeling techniques, for example Naive Discriminative Learning (NDL; \citealp{Baayen2011,BaayenEa2013,MilinEa2016,TheijssenEa2013}).
-On p.\ 303 of \cite{Divjak2016a}, the author explicitly mentions NDL.
+On p.\ 303 of \cite{Divjak2016a}, the author goes on to explicitly mention NDL.
 Yet, neither frequentist nor Bayesian methods were conceived as cognitive models, but as systems of inference for scientists (see above, and see also \citealp[302]{Divjak2016a}).
@@
 In this section, I introduce and illustrate the relevant alternating constructions.
-I describe the narrowly defined syntactic configuration in which the alternation occurs, and I motivate the focus on \textit{only} this narrow range (rather than, for example, the whole range of nominal constructions expressing quantities).
+I describe the narrowly defined syntactic configuration in which the alternation occurs, and I motivate the focus on \textit{only} this narrow range rather than, for example, the whole range of nominal constructions expressing quantities.
 
@@
 For German, this terminology is not distinctive enough, which is why I use the terms NAC and PGC.}
-Notice that the unavailability of the genitive on the kind noun can be seen as following from a rather quirky constraint that genitive NPs in German require the presence of some strongly case-marked element (determiner or adjective) in addition to the head noun in order to be acceptable \citep{GallmannLindauer1994,Schachtl1989}.
+Notice that the unavailability of the genitive on the kind noun can be seen as following from a rather quirky constraint that genitive NPs in German require the presence of some strongly case-marked element (determiner or adjective) in addition to the head noun in order to be acceptable (\citealp{GallmannLindauer1994,Schachtl1989}; see also \citealp[160]{Eisenberg2013b}).
 While only an accusative is shown in (\ref{ex:intro:narrowapposition2}), the kind noun obligatory agrees in case also with nominative and dative measure nouns.
@@
   \item{\label{it:intro:idiot} alleged alternatives to the \NACa\ and the \PGCa}
-  \item{\label{it:intro:datsg} alternative \textit{weak} forms of dative singular neuter adjectives}
+%  \item{\label{it:intro:datsg} alternative \textit{weak} forms of dative singular neuter adjectives}
 %  \item{\label{it:intro:femsg} case syncretism in feminine NPs}
@@
 First of all, (\ref{it:intro:idiot}) refers to claims found in some grammars that a generic nominative, accusative, and even dative on the kind noun are used instead of the genitive (\PGCa) or case agreement (\NACa).
-Overviews can be found in \citealp{Hentschel1993,Zimmer2015}.
+Overviews can be found in \citealp{Hentschel1993,Zimmer2015}, and (\ref{ex:intro:gennom}) shows a putative generic nominative on the kind noun inside an accusative MNP.
+
+\begin{exe}
+    \ex[*]{\label{ex:intro:gennom} Wir trinken [[eine Tasse]\Sub{Acc} [heißer Kaffee]\Sub{Nom}]\Sub{Acc}.}
+\end{exe}
+
 It was shown empirically in \cite{Hentschel1993} that such variants are de facto not acceptable.
@@
 
-As for (\ref{it:intro:datsg}), a complication with neuter kind nouns in the dative is mentioned by \citet[20--22]{Zimmer2015}.
-In the \NACa, the adjective normally inflects with the so-called \textit{strong} case and number suffixes, which are used when no determiner with strong inflection is present in the NP.
-\cite{Zimmer2015} reports a high number of occurrences of adjectives being inflected with the \textit{weak} inflectional marker, which are normally only used if a strongly inflected determiner precedes the adjective: \textit{kalt-en Wasser} `cold water' instead of \textit{kalt-em Wasser}.%
-\footnote{See Section~\ref{sec:analyses} for more discussion of adjectival inflection.}
-In my corpus sample, this tendency was not nearly as clear, and there was a high number of very noisy sentences among those potentially showing this pattern.
-Hence, I do not discuss these forms.%
-\footnote{All corpus data and scripts used for this paper will be released freely, so further examination of the few cases maybe showing this kind of inflection is possible.}
-
+% As for (\ref{it:intro:datsg}), a complication with neuter kind nouns in the dative is mentioned by \citet[20--22]{Zimmer2015}.
+% In the \NACa, the adjective normally inflects with the so-called \textit{strong} case and number suffixes, which are used when no determiner with strong inflection is present in the NP.
+% \cite{Zimmer2015} reports a high number of occurrences of adjectives being inflected with the \textit{weak} inflectional marker, which are normally only used if a strongly inflected determiner precedes the adjective: \textit{kalt-en Wasser} `cold water' instead of \textit{kalt-em Wasser}.%
+% \footnote{See Section~\ref{sec:analyses} for more discussion of adjectival inflection.}
+% In my corpus sample, this tendency was not nearly as clear, and there was a high number of very noisy sentences among those potentially showing this pattern.
+% Hence, I do not discuss these forms.%
+% \footnote{All corpus data and scripts used for this paper will be released freely, so further examination of the few cases maybe showing this kind of inflection is possible.}
+% 
 % \label{page:femininesyncretism}
@@
                                        three spoons hot rum are enough.\\
-				     \glt Three spoonful of hot rum are enough.}
+				     \glt Three spoonful of hot rum is enough.}
     \ex{\label{ex:determiners:pgc}\gll [[Einige Löffel]\Sub{Nom} [heißen Rums]\Sub{Gen}]\Sub{Nom} sind genug\\
                                        some   spoons hot    rum  are   enough.\\
-				     \glt A few spoonful of hot rum are enough.}
+				     \glt A few spoonful of hot rum is enough.}
   \end{xlist}
@@
   It could be argued that some more advanced measure of attraction strength should be used, as is done in Collostructional (or Collexeme) Analysis \citep{GriesStefanowitsch2004}, see also \cite{Gries2015a}.
-  Three main points speak against such an approach in the present case (but see \citealp[246--249]{Levshina2016} for a successful more refined approach).
+  Three main points speak against such an approach in the present case (but see \citealp[246--249]{Levshina2016} for a different approach).
   First, the attraction values will be used as regressors in a GLMM, and the values resulting from collostructional approaches, \ie\ logarithmised Fisher $p$ values, have a very unfavourable distribution in the case at hand.
@@
 
-In Section~\ref{sec:analyses}, it was also hypothesised that there may be effects specific to classes of measure lemmas, and that they might be related to degrees of grammaticalisation of the measure noun.
-Transcending the effects of individual lemmas (captured in the random intercepts), the \textit{Measureclass} second-level predictor was successfully estimated ($\mpPB=0.001$).
+In Section~\ref{sec:analyses}, it was also hypothesised that classes of measure nouns with a higher degree of grammaticalisation should favour the \PGCa.
+The \textit{Measureclass} second-level predictor was successfully estimated ($\mpPB=0.001$).
 Looking at the effect plot in Figure~\ref{fig:eff:measureattraction}, it is evident that abstract non-referential physical measure nouns (such as \textit{Gramm} `gram' or \textit{Liter} `litre') with a high degree of grammaticalisation favour the \NACa.
@@
 In my view, the main tenets of this line of research are:
-(i)~Language, viewed from a cognitively realistic angle, is a probabilistic phenomenon and cannot be modeled appropriately within Aristotelian frameworks that assume discrete and clearly delimited categories.
+(i)~Language, viewed from a cognitively realistic angle, is a probabilistic phenomenon and cannot be modeled appropriately within Aristotelian frameworks that assume discrete categories.
 (ii)~Corpora are collections of usage events (language production) and can therefore be used to evaluate both the claim made in (i) and specific theoretical claims (in case studies) about factors influencing speakers' decisions to use specific forms or constructions.
@@
 The present paper adds to the evidence that all of the aforementioned three points are correct.
-A grammatical alternation in German measure NPs was examined using corpus data and based on factors partly derived from existing accounts and formulated in terms of construction prototypes.
+A grammatical alternation in German measure NPs was examined using corpus data based on factors partly derived from existing accounts, formulated in terms of construction prototypes.
 The preferences extracted from the DECOW web corpus were confirmed in a forced-choice experiment, in which participants explicitly chose variants in line with the corpus-derived model.
@@
 This is quite surprising given the fact that German morpho-syntax is famous for its alternations, which are usually called \textit{Zweifelsfälle} (`cases of doubt') in the traditional literature \citep{Klein2009}.
-Instead of drowning them in normative, descriptive, or didactic discussions, they could serve as ideal test cases in cognitive linguistics.
+Instead of being drowned in normative, descriptive, or didactic discussions, they could serve as ideal test cases in cognitive linguistics.
 
@@
 \end{document}
-
-%\footnote{\cite{AnttilaFong2000} propose an \textit{Obligatory Contour Principle for case} (Case-OCP) in a superficially similar but fundamentally unrelated case alternation in Finnish partitives.
-%}
-%
-% On Case-OCP (from Götz Keydana):
-% Und da habe ich dann auch einen Verweis auf den locus classicus gefunden:
-% STEFAN A. FRISCH, JANET B. PIERREHUMBERT and MICHAEL B. BROE: SIMILARITY AVOIDANCE AND THE OCP. Natural Language & Linguistic Theory 22: 179–228, 2004.
-% Dazu gäbe es dann noch, ebenfalls von Pierrehumbert,
-% Dissimilarity in the Arabic Verbal Roots. In Amy Schafer, ed.,Proceedings of NELS 23 Amherst, MA: Graduate Linguistics Student Association. 1993. pp 367-381 
-% Darin stellt sie wohl erstmals explizit einen Zusammenahng zwischen OCP und Kognition her.

### 2017-04-06 Latex/article.tex
minor fixes, new plot

@@
 If the prior is flat, results converge (see also \citealp[347]{GelmanHill2006}).
-The same is true if the sample size is large compared to the number of parameters, at least for finite-dimensional models \citep[1119--1120]{Freedman1999}, a well-established result known as the Bernstein-von Mises theorem.
+The same is true if the sample size is large compared to the number of parameters, at least for finite-dimensional parameter models \citep[1119--1120]{Freedman1999}, a well-established result known as the Bernstein-von Mises theorem.
 With a modest model structure including $17$ fixed effects and $2,646$ data points in \cite{Levshina2016}, it is highly likely that the same results would have obtained with Maximum Likelihood methods.
 In fact, she admits that changing the priors did not lead to substantially different results in her footnote 8.
-This is a clear sign that the prior is \textit{swamped} by the data \citep[1119]{Freedman1999}.
+This is a clear sign that the prior is \textit{swamped by the data} \citep[1119]{Freedman1999}.
 Going back to the argument that Bayesian methods are allegedly more cognitively plausible than `frequentist' methods, the claim might still be true.
-However, most similar studies (back to \citealp{BresnanEa2007}) use an estimator (MLE) that is not frequentist, and as cognitively plausible as Bayesian estimators because it most likely produces the same results in these studies.
+However, most similar studies (back to \citealp{BresnanEa2007}) use an estimator (MLE) that is not frequentist in a narrow sense, and which is as cognitively plausible as Bayesian estimators because it most likely produces the same results in these studies.
 If for Levshina's study, results from Bayesian and MLE methods did, in fact, \textit{not} converge, it would have been an ideal occasion to demonstrate the superiority of the algorithms used in Bayesian estimation.
@@
 
-Finally, I show that Bayesian methods in the form of Markov-Chain Monte Carlo estimation do not necessarily (and in this case predictably) lead to different results.
+Finally, I show that Bayesian methods in the form of Markov-Chain Monte Carlo sampling (MCMC) do not necessarily (and in this case predictably) lead to different results.
 The model was re-estimated using the \textit{stan\_glmer} function from the \textit{rstanarm} package, which provides an \textit{lme4}-compatible syntax for estimating common model types with the \textit{stan} software \citep{CarpenterEa2017}.

### 2017-04-06 R/4_glmm-mcmc.R

@@ -89,6 +89,6 @@ for (i in 1:nrow(measure.ci.95)) {
   lines(measure.glmm.table[names(measure.fixeffs[i]), c("MC_Low", "MC_High")], c(i-0.25,i-0.25), col="gray", lwd=2)
 }
 
-legend("topright", legend = c("MLE", "MCMC"), cex = 1.2, col = c("black", "gray"), pt.bg = c("black", "gray"),
+legend(1.11, 12.5, legend = c("MLE", "MCMC"), cex = 1.2, col = c("black", "gray"), pt.bg = c("black", "gray"),
        lty = 1, lwd = 2, pch = c(25, 24), pt.cex = 0.75, bg = "white")
 if (save.persistent) dev.off()

### 2017-04-09 Latex/article.tex
Minora.

@@
 After all, GLMMs have been established as the major tool in the analysis of alternation phenomena.
-All studies mentioned at the outset of Section~\ref{sec:cogocl} use GLM(M)s \citep{BresnanEa2007,BresnanHay2010,BresnanFord2010,DivjakArppe2013,Gries2015,NessetJanda2010} use some form of regression\slash GLM(M).
+All studies mentioned at the outset of Section~\ref{sec:cogocl} \citep{BresnanEa2007,BresnanHay2010,BresnanFord2010,DivjakArppe2013,Gries2015,NessetJanda2010} use some form of regression\slash GLM(M).
 Over the past few years, however, modified or alternative methods have been proposed.
@@
 There is actually a debate among Bayesians about the proper interpretation of Bayesian methods, and whether a notion of hypothesis testing is compatible (or even already contained) in the Bayesian approach.
-In \citet[10]{GelmanShalizi2013}, the authors -- prominent Bayesians themselves -- acknowledge that a theory of testing is a desideratum, and they state about the standard inductive interpretation of Bayesianism that \textit{most of this received view of Bayesian inference is wrong}, and they develop a Bayesian notion of $p$ values (see also \citealp{Mayo2013}, for a frequentist reply; also \citealp{Senn2011} on different strands of Bayesianism and their stance on inductive vs.\ deductive reasoning).
+In \citet[10]{GelmanShalizi2013}, the authors -- prominent Bayesians themselves -- acknowledge that a theory of testing is a desideratum, and they state about the standard inductive interpretation of Bayesianism that \textit{most of this received view of Bayesian inference is wrong}, and they develop a Bayesian notion of $p$ values (see also \citealp{Mayo2013}, for a frequentist reply; also \citealp{Senn2011} on different strands of Bayesianism and their stance on inductive vs.\ deductive reasoning, and \citealp{Mayo2011}, for a critical reply).
 Clearly, in such quarrels between and among camps of philosophers of science and statisticians it is difficult for mere practitioners to take sides.
@@
 In the former case, we adopt a strong \textit{corpus as input} hypothesis and should definitely resort to methods like NDL.
-However, this would most likely require us to toss most previous work done in (cognitive) linguistics into the bin and abandoning all high-level generalisations that most existing studies have been based upon.
+However, this would most likely require us to toss most previous work done in (cognitive) linguistics into the bin and to abandon all high-level generalisations that most existing studies have been based upon.
 In fact, arguments to this extreme effect and against using high-level generalisation have been made, for example in \cite{BaayenEa2016}, \citet[299--300]{Divjak2016a}, \cite{RamscarPort2016}, \cite{TheijssenEa2013}.
@@
 These methods then do not need to be any more cognitively plausible than an ANOVA used to analyse the results from an experiment.
-I view my own work (see Sections~\ref{sec:germanmeasurenps}--\ref{sec:experimental}) in the tradition of testing theories that embrace high-level generalisations, and I agree to provisionally use Likelihood methods (see below).
+I view my own work (see Sections~\ref{sec:germanmeasurenps}--\ref{sec:experimental}) in the tradition of testing theories (which embrace high-level generalisations), and I agree to provisionally use Likelihood methods (see below).
 I remain fully agnostic with respect to the question which approach is `right'.

### 2017-04-09 Latex/article.tex
Minora.

@@
 
-As for the first question, this is not quite the place to discuss it fully.
+As for the first point, this is not quite the place to discuss it fully.
 The basic distinction is a philosophical one and related to the concepts of \textit{direct} and \textit{inverse probability} (\eg\ \citealp{Senn2011}).
@@
 There is actually a debate among Bayesians about the proper interpretation of Bayesian methods, and whether a notion of hypothesis testing is compatible (or even already contained) in the Bayesian approach.
-In \citet[10]{GelmanShalizi2013}, the authors -- prominent Bayesians themselves -- acknowledge that a theory of testing is a desideratum, and they state about the standard inductive interpretation of Bayesianism that \textit{most of this received view of Bayesian inference is wrong}, and they develop a Bayesian notion of $p$ values (see also \citealp{Mayo2013}, for a frequentist reply; also \citealp{Senn2011} on different strands of Bayesianism and their stance on inductive vs.\ deductive reasoning, and \citealp{Mayo2011}, for a critical reply).
+In \citet[10]{GelmanShalizi2013}, the authors -- prominent Bayesians themselves -- acknowledge that a theory of testing is a desideratum, and they state about the standard inductive interpretation of Bayesianism that \textit{most of this received view of Bayesian inference is wrong}, and they develop a Bayesian notion of $p$ values (see also \citealp{Mayo2013}, for a frequentist reply; also \citealp{Senn2011} on different strands of Bayesianism and their stance on inductive vs.\ deductive reasoning, and \citealp{Mayo2011}, for a critical reply to \citealp{Senn2011}).
 Clearly, in such quarrels between and among camps of philosophers of science and statisticians it is difficult for mere practitioners to take sides.
 
-These quarrels relate to the second question, however.
+These quarrels relate to the second point, however.
 \citet[301--302]{Divjak2016a} speaks favourably of Bayesian methods because the Bayesian concept of probability is more adequate for cognitive modeling compared to the frequentist one.
 Her argument is part of a larger body of literature asking for cognitively plausible modeling techniques, for example Naive Discriminative Learning (NDL; \citealp{Baayen2011,BaayenEa2013,MilinEa2016,TheijssenEa2013}).
-On p.\ 303 of \cite{Divjak2016a}, the author goes on to explicitly mention NDL.
+On p.\ 303 of \cite{Divjak2016a}, the author goes on to explicitly mention NDL as well.
 Yet, neither frequentist nor Bayesian methods were conceived as cognitive models, but as systems of inference for scientists (see above, and see also \citealp[302]{Divjak2016a}).
@@
 Then, the corpus study is presented including an appropriate statistical analysis (Section~\ref{sec:corpusstudies}).
-The two experiments which correlate results from the corpus study with experimental results are then reported (Section~\ref{sec:experimental}) before I sum up the paper in Section~\ref{sec:conclusion}.
+Two experiments that correlate results from the corpus study with experimental results are then reported (Section~\ref{sec:experimental}) before I sum up the paper in Section~\ref{sec:conclusion}.
 
@@
 \footnote{If the kind noun is definite, the construction instantiates a true partitive.
-Whereas partitives are constructions denoting a proper part-of relation as in \textit{a sip of the wine}, pseudo partitives -- albeit syntactically similar and diachronically related to partitives in many languages -- merely denote quantities and contain indefinite kind nouns as in \textit{a sip of wine}.%
+Whereas partitives are constructions denoting a proper part-of relation as in \textit{a sip of the wine}, pseudo partitives -- albeit syntactically similar and diachronically related to partitives in many languages -- merely denote quantities and contain indefinite kind nouns as in \textit{a sip of wine}.
 In the literature on German, some authors incorrectly call the pseudo-partitive a \textit{partitive} \cite{Hentschel1993} while some realise the difference and at least mention it \citep{Eschenbach1994,GallmannLindauer1994,Loebel1989,Zimmer2015}.
@@
 
-If the kind noun is bare -- \ie\ if it does neither come with a determiner nor a modifying adjective -- it has to agree in case with the measure noun, and the genitive seen in the PGC is not acceptable, see (\ref{ex:intro:narrowapposition2}).
+If the kind noun is bare -- \ie\ if it comes neither with a determiner nor a modifying adjective -- it has to agree in case with the measure noun, and the genitive as seen in the PGC is not acceptable, see (\ref{ex:intro:narrowapposition2}).
 
@@
 If the kind noun is a plural count noun as in \textit{ein Sack kleine Äpfel} (\NACa) or \textit{ein Sack kleiner Äpfel} (\PGCa) `a bag of small apples', a similar alternation between PGC and NAC can be observed.
-In line with experimental results reported in \citet[15--16]{Zimmer2015}, I found that the PGC is so dominant with plural kind nouns (67 of 861 cases or over 92\%, cf.\ Section~\ref{sec:corpusstudies}) that the alternation cannot be analysed in the same way as in the singular case.%
+In line with experimental results reported in \citet[15--16]{Zimmer2015}, I found that the PGC is so dominant with plural kind nouns (67 of 861 cases or over 92\%, cf.\ Section~\ref{sec:corpusstudies}) that the alternation cannot be analysed in the same way as in the singular.%
 \footnote{More concretely, if these cases were included in the regression analysis of the corpus data along with a factor encoding the number of the kind noun, this factor would most assuredly override any other regressor for the data points with a plural kind noun.}
-While this will play a role in the interpretation of the corpus findings, MNPs with plural kind nouns will not be included in the corpus study and the experiments reported here.
+While this will play a role in the interpretation of the corpus findings, MNPs with plural kind nouns will not be included in the corpus study and the experiments reported in Sections~\ref{sec:corpusstudies} and \ref{sec:experimental}.
 
@@
 These construction have very idiosyncratic properties (the construction with \textit{voller} is discussed in \citealp{Zeldes2018}, the construction with \textit{mit} is discussed in \citealp{Bhatt1990}), they are not semantically equivalent to the NAC and the PGC, and they are only available for a small subset of measure nouns.
-All of these constructions are only used with measure nouns denoting containers, and the whole NP always refers to the container, and never to the quantity contained in it.
+All of these constructions are used only with measure nouns denoting containers, and the whole NP always refers to the container, and never to the quantity contained in it.
 They are consequently incompatible with measure nouns denoting natural portions such as \textit{Schluck} `gulp' or \textit{Haufen} `heap' and strongly grammaticalised nouns such as \textit{Gramm} `gram'.
@@
     \ex{\label{ex:determiners:nac}\gll [[Drei Löffel]\Sub{Nom} [heißer Rum]\Sub{Nom}]\Sub{Nom} sind genug.\\
-                                       three spoons hot rum are enough.\\
+                                       three spoons hot rum are enough\\
 				     \glt Three spoonful of hot rum is enough.}
-    \ex{\label{ex:determiners:pgc}\gll [[Einige Löffel]\Sub{Nom} [heißen Rums]\Sub{Gen}]\Sub{Nom} sind genug\\
-                                       some   spoons hot    rum  are   enough.\\
+    \ex{\label{ex:determiners:pgc}\gll [[Einige Löffel]\Sub{Nom} [heißen Rums]\Sub{Gen}]\Sub{Nom} sind genug.\\
+                                       some   spoons hot    rum  are   enough\\
 				     \glt A few spoonful of hot rum is enough.}

### 2017-04-10 Latex/article.tex
Some Bildhauer comments.

@@
 Instead, multifactorial models are constructed, which incorporate influencing factors from diverse levels, including contextual ones.
-The estimation of the model quantifies the influence the factors have on the probability that either variant is chosen.
+The estimation of the model coefficients quantifies the influence that the factors have on the probability that either variant is chosen.
 
@@
 This often called a \textit{validation} of the corpus-derived findings, but \citet[303]{Divjak2016a} criticises this choice of words \textit{because it creates the impression that behavioral experimental data is inherently more valuable than textual data}, citing \cite{TummersEa2005}, who state that a corpus is \textit{a sample of spontaneous language use that is} (\textit{generally}) \textit{realized by native speakers}.
-This position, which I adopt, does not, however, imply that we can in some way \textit{deduce mental representations from patterns of use}, \ie\ from corpus data \citep[486--487]{Dabrowska2016}.%
+This position, which I adopt, does not, however, imply that we can in some way \textit{deduce mental representations from patterns of use}, \ie\ from corpus data, as  \citet[486--487]{Dabrowska2016} convincingly argues.%
 \footnote{I suggest that the word \textit{induce} would be more appropriate than \textit{deduce}.}
-In fact, it would be surprising if we could!
-Nobody assumes that we can induce mental representations from experiments, which -- as opposed to corpus studies -- even provide a direct access to the cognitive agent and offer much better options to control and experimental conditions and nuisance variables.
+It would be highly surprising if this were possible.
+Nobody assumes that we can induce mental representations from experiments, which -- as opposed to corpus studies -- even allow for direct access to the cognitive agent and offer much better options to control experimental conditions and nuisance variables.
 Instead, a theory of cognitive representation is pre-specified, and predictions are derived from this theory \textit{before} the experiment or the corpus study is conducted.
@@
 The similarity of a given feature vector (in a concrete sentence where one of the variants is used) determines to a large extent which variant is chosen by speakers.
-This entails that a variant of Prototype Theory with \textit{features} is assumed \citep{Rosch1978}, instead of the monolithic prototypes assumed in earlier versions.
-While Prototype Theory is well suited for modeling constructional choices, it is just one similarity-based framework, the most prominent other framework being Exemplar Theory (\citealp{MedinSchaffer1978,Hintzman1986}; see \citealp{StormsEa2000} for a comparison of the theories in experimental settings).
+This entails that a variant of Prototype Theory with \textit{features} is assumed \citep{Rosch1978}, instead of the monolithic prototypes of earlier versions.
+While Prototype Theory is well suited for modeling constructional choices, it is just one of several similarity-based theories of classification, the most prominent other framework being Exemplar Theory (\citealp{MedinSchaffer1978,Hintzman1986}; see \citealp{StormsEa2000} for a comparison of the theories in experimental settings).
 Prototype Theory and Exemplar Theory differ mainly in whether they assume higher-level abstractions as part of the cognitive representation (Prototype Theory) or not (Exemplar Theory).
@@
 \footnote{In \cite{DivjakArppe2013}, a fruitful attempt was made to show how both models form a converging picture based on corpus data.}
-Prototype Theory (with feature abstractions) is simply preferred here because it fits the established alternation modeling paradigm, which relies on features being weighted (in a sense) in statistical models.
+Prototype Theory (with feature abstractions) is simply preferred here because it fits the established alternation modeling paradigm, which relies on features being weighted in statistical models.
 In Section~\ref{sec:analyses}, a prototype-theoretical parlance is therefore adopted.
@@
 After all, GLMMs have been established as the major tool in the analysis of alternation phenomena.
-All studies mentioned at the outset of Section~\ref{sec:cogocl} \citep{BresnanEa2007,BresnanHay2010,BresnanFord2010,DivjakArppe2013,Gries2015,NessetJanda2010} use some form of regression\slash GLM(M).
+All studies mentioned at the outset of Section~\ref{sec:cogocl} use some form of regression\slash GLM(M).
 Over the past few years, however, modified or alternative methods have been proposed.
@@
 There is actually a debate among Bayesians about the proper interpretation of Bayesian methods, and whether a notion of hypothesis testing is compatible (or even already contained) in the Bayesian approach.
-In \citet[10]{GelmanShalizi2013}, the authors -- prominent Bayesians themselves -- acknowledge that a theory of testing is a desideratum, and they state about the standard inductive interpretation of Bayesianism that \textit{most of this received view of Bayesian inference is wrong}, and they develop a Bayesian notion of $p$ values (see also \citealp{Mayo2013}, for a frequentist reply; also \citealp{Senn2011} on different strands of Bayesianism and their stance on inductive vs.\ deductive reasoning, and \citealp{Mayo2011}, for a critical reply to \citealp{Senn2011}).
-Clearly, in such quarrels between and among camps of philosophers of science and statisticians it is difficult for mere practitioners to take sides.
+In \citet[10]{GelmanShalizi2013}, the authors -- prominent Bayesians themselves -- acknowledge that a theory of statistical testing is a desideratum, and they state about the standard inductive interpretation of Bayesianism that \textit{most of this received view of Bayesian inference is wrong}, and they develop a Bayesian notion of $p$ values (see also \citealp{Mayo2013}, for a frequentist reply; also \citealp{Senn2011} on different strands of Bayesianism and their stance on inductive vs.\ deductive reasoning, and \citealp{Mayo2011}, for a critical reply to \citealp{Senn2011}).
+Clearly, in such quarrels between and among camps of philosophers of science and statisticians, it is difficult for mere practitioners to take sides.
 
@@
 Are they inductive models of cognitive representations that also human learners would infer from being exposed to the corpus data?%
-\footnote{In which case we are doing \textit{data science in language research} in the words of \citealp{MilinEa2016}.}
-Or are they tests of theories that are pre-specified and merely tested for predictive accuracy on linguistic output data contained in corpora.
+\footnote{In which case we are doing \textit{data science in language research} in the words of \citealp{MilinEa2016}.
+I see this as standing in contradiction to the view advocated in \cite{Dabrowska2016} as cited above.}
+Or are they tests of theories that are pre-specified and merely tested for predictive accuracy on linguistic output data contained in corpora?
 In the former case, we adopt a strong \textit{corpus as input} hypothesis and should definitely resort to methods like NDL.
@@
 First, she claims that \textit{while frequentist statistics only allows one to test whether the null hypothesis can be rejected, Bayesian statistics enables one both to test the null hypothesis and to estimate the probability of specific parameter values given the data}.
-This does not do justice to frequentist methods in that mere rejection of the null hypothesis is characteristic only of Fisher's approach, but in the Neyman-Pearson approach, results ideally \textit{favour} the main hypothesis vis-à-vis the alternative hypothesis (cf.\ \citealp{Lehmann1993,Lehmann2011,Perezgonzalez2015}).
-Also, especially Neyman-style frequentism has well-known extensions to estimation (see \citealp{GreenlandEa2016}, esp.\ p.\ 340).
+This does not do justice to frequentist methods in that mere rejection of the null hypothesis is characteristic only of Fisher's approach.
+In the Neyman-Pearson approach, results ideally \textit{favour} the main hypothesis vis-à-vis the alternative hypothesis (cf.\ \citealp{Lehmann1993,Lehmann2011,Perezgonzalez2015}).
+Also, especially Neyman-style frequentism has well-known extensions to estimation, for example in the form of confidence intervals (see \citealp{GreenlandEa2016}, esp.\ p.\ 340).
 She then explains that a \textit{distinctive feature of Bayesian statistics is the use of so-called priors} and that \textit{posterior probabilities depend on both the prior beliefs and the data, whereas the results of a frequentist model depend only on the data} \cite[252]{Levshina2016}.
@@
 Now, Maximum Likelihood Estimation (MLE) -- the traditional method which could have been used instead -- is not exactly \textit{frequentist} in the sense of Neyman-Pearson testing theory.
-MLE, like inductive Bayesianism, conditions on the data in as much as it searches for the most likely set of parameters given the particular data.
-What is more, Bayesian estimators are in fact based the Likelihood function and merely multiply it by the prior \citep[6--8]{GelmanEa2014}.
+MLE, like inductive Bayesianism, conditions on the particular data in as much as it searches for the most likely set of parameters given the data.
+What is more, Bayesian estimators are in fact based the Likelihood and merely multiply it by the prior \citep[6--8]{GelmanEa2014}.
 If the prior is flat, results converge (see also \citealp[347]{GelmanHill2006}).
@@
 Going back to the argument that Bayesian methods are allegedly more cognitively plausible than `frequentist' methods, the claim might still be true.
-However, most similar studies (back to \citealp{BresnanEa2007}) use an estimator (MLE) that is not frequentist in a narrow sense, and which is as cognitively plausible as Bayesian estimators because it most likely produces the same results in these studies.
+However, most similar studies (dating back to \citealp{BresnanEa2007}) use an estimator (MLE) that is not frequentist in a narrow sense, and which is as cognitively plausible as Bayesian estimators because it most likely produces the same results in these studies.
 If for Levshina's study, results from Bayesian and MLE methods did, in fact, \textit{not} converge, it would have been an ideal occasion to demonstrate the superiority of the algorithms used in Bayesian estimation.
@@
 (such as \textit{glass} or \textit{bucket}).
-In a similar vein, \citet[284]{Brems2003} considers nouns as measure nouns \textit{which, strictly speaking, do not designate a `measure', but display a more nebulous} (\textit{sic!}) \textit{potential for quantification} (see also \citealp[530]{Koptjevskaja2001}, and \citealp[338]{Rutkowski2007}).
+In a similar vein, \citet[284]{Brems2003} considers such nouns denoting containers etc.\ as measure nouns \textit{which, strictly speaking, do not designate a `measure', but display a more nebulous} (sic!) \textit{potential for quantification} (see also \citealp[530]{Koptjevskaja2001}, and \citealp[338]{Rutkowski2007}).
 
@@
 Whereas partitives are constructions denoting a proper part-of relation as in \textit{a sip of the wine}, pseudo partitives -- albeit syntactically similar and diachronically related to partitives in many languages -- merely denote quantities and contain indefinite kind nouns as in \textit{a sip of wine}.
-In the literature on German, some authors incorrectly call the pseudo-partitive a \textit{partitive} \cite{Hentschel1993} while some realise the difference and at least mention it \citep{Eschenbach1994,GallmannLindauer1994,Loebel1989,Zimmer2015}.
+In the literature on German, some authors incorrectly call the pseudo-partitive a \textit{partitive} \citep{Hentschel1993} while some realise the difference and at least mention it \citep{Eschenbach1994,GallmannLindauer1994,Loebel1989,Zimmer2015}.
 }
@@
 Notice that the unavailability of the genitive on the kind noun can be seen as following from a rather quirky constraint that genitive NPs in German require the presence of some strongly case-marked element (determiner or adjective) in addition to the head noun in order to be acceptable (\citealp{GallmannLindauer1994,Schachtl1989}; see also \citealp[160]{Eisenberg2013b}).
-While only an accusative is shown in (\ref{ex:intro:narrowapposition2}), the kind noun obligatory agrees in case also with nominative and dative measure nouns.
+%While only an accusative is shown in (\ref{ex:intro:narrowapposition2}), the kind noun obligatory agrees in case also with nominative and dative measure nouns.
 
@@
 Also, in my corpus sample, they simply did not occur.
-Even if they are accepted by some speakers, their extremely low frequency makes it virtually impossible to study them using either corpus methods (Section~\ref{sec:corpusstudies}), and I consequently ignore them.
+Even if they are accepted by some speakers, their extremely low frequency makes it virtually impossible to study them using corpus linguistic methods (Section~\ref{sec:corpusstudies}), and I consequently ignore them.
 
@@
 If the kind noun is a plural count noun as in \textit{ein Sack kleine Äpfel} (\NACa) or \textit{ein Sack kleiner Äpfel} (\PGCa) `a bag of small apples', a similar alternation between PGC and NAC can be observed.
-In line with experimental results reported in \citet[15--16]{Zimmer2015}, I found that the PGC is so dominant with plural kind nouns (67 of 861 cases or over 92\%, cf.\ Section~\ref{sec:corpusstudies}) that the alternation cannot be analysed in the same way as in the singular.%
+In line with experimental results reported in \citet[15--16]{Zimmer2015}, I found that the PGC is so dominant with plural kind nouns ($794$ of $861$ cases, or over 92\%, cf.\ Section~\ref{sec:corpusstudies}) that the alternation cannot be analysed in the same way as in the singular.%
 \footnote{More concretely, if these cases were included in the regression analysis of the corpus data along with a factor encoding the number of the kind noun, this factor would most assuredly override any other regressor for the data points with a plural kind noun.}
@@
 As for (\ref{it:intro:noifl}), some measure nouns have been grammaticalised in a way that they always appear non-inflected.
-They are typical measure nouns like \textit{Gramm} `gram', \textit{Pfund} `pound' or \textit{Prozent} `percent', which do not have plural forms at all.
+They are typical measure nouns like \textit{Gramm} `gram', \textit{Pfund} `pound' or \textit{Prozent} `percent', which have no normal plural forms at all.%
+\footnote{Plurals \textit{Pfunde} and \textit{Prozente} have special meanings and have very restricted uses, mostly in idiomatic expressions and prefabs.
+They cannot be used in MNPs.}
 %\footnote{With neglectable frequency, this effect is extended to container nouns like \textit{Glas} `glass', but with a twist.
@@
 This concludes the descriptive overview of the phenomenon.
-I have shown that there is an alternation between two measure noun constructions in a narrow syntactic configuration (kind NP with an adjective but without a determiner), and that the two constructions differ in the case of the kind noun (case agreement with the measure noun or genitive).
+I have demonstrated that there is an alternation between two measure noun constructions in a narrow syntactic configuration (kind NP with an adjective but without a determiner), and that the two constructions differ in the case of the kind noun (case agreement with the measure noun or genitive).
 I turn to some more theory-oriented discussion in the next section.
@@
 When they classify the adjective more as an adjective, the kind NP has no determiner, and they have to use the \NACa.%
-\footnote{While the generative analysis presented in \cite{Bhatt1990} cannot properly deal with probabilistic effects, Bhatt comes close to this interpretation by analysing the kind NP in the GPC as a DP and in the NAC as an NP.}
+\footnote{While the generative analysis presented in \cite{Bhatt1990} cannot properly deal with probabilistic effects, Bhatt comes close to this interpretation by analysing the kind NP in the PGC as a DP and in the NAC as an NP.}
 
@@
 The degree of similarity of a given instance to either of the two prototypes makes speakers chose the \NACa\ or the \PGCa.
+
 I first turn to grammaticalisation.
-It is often assumed that pseudo-partitives arise as a form of grammaticalised partitives (\eg\ \citealp[536--539]{Koptjevskaja2001} for Finnish and Estonian, \citealp[559]{Koptjevskaja2001} for European languages per se).
+It is often assumed that pseudo-partitives arise as a form of grammaticalised partitives (\eg\ \citealp[536--539]{Koptjevskaja2001} for Finnish and Estonian, \citealp[559]{Koptjevskaja2001} for European languages in general).
 %
@@
 %
-The alternating constructions in German are both clearly pseudo-partitives, but the grammaticalisation paths uncovered by \cite[esp.\ 526--530]{Koptjevskaja2001} are still relevant in the case at hand.
+The alternating constructions in German are both clearly pseudo-partitives, but the grammaticalisation paths uncovered by \citet[esp.\ 526--530]{Koptjevskaja2001} are still relevant in the case at hand.
 The grammaticalisation path can start out (in some languages) with constructions involving two referential nouns (not necessarily forming a single and contiguous NP) and a \textit{separative} meaning as in \textit{(cut) two slices from the cake} \citep[535]{Koptjevskaja2001}.
@@
 As described above, the grammaticalisation path leads from NPs denoting individuated objects standing in a \textit{part-of} relation to a construction with a more diffuse \textit{quantity-of} relation.
-Both types of relations can be numerically quantified -- in as much as a precise number of \textit{parts} or a numerically exact \textit{quantity} can be specified --, but it is much more prototypical of quantities to be specified with numerical precision.
+Both types of relations can be numerically quantified -- in as much as a precise number of \textit{parts} or a numerically exact \textit{quantity} can be specified.
+However, it is much more prototypical of quantities to be specified with numerical precision.
 Since the \NACa\ is more closely associated with the \textit{quantity-of} relation, cardinals as attributes of the measure noun are expected to have a higher proportional frequency in the \NACa.
 For illustration, (\ref{ex:determiners}) shows the expected variants under this hypothesis.
-In (\ref{ex:determiners:nac}), the measure noun is modified by a cardinal \textit{drei} `three', and hence the \NACa\ is preferred.
-In (\ref{ex:determiners:pgc}), the measure noun is modified by a non-cardinal determiner \textit{einige} `some', and the \PGCa\ is preferred.
 
@@
   \begin{xlist}
-    \ex{\label{ex:determiners:nac}\gll [[Drei Löffel]\Sub{Nom} [heißer Rum]\Sub{Nom}]\Sub{Nom} sind genug.\\
-                                       three spoons hot rum are enough\\
-				     \glt Three spoonful of hot rum is enough.}
-    \ex{\label{ex:determiners:pgc}\gll [[Einige Löffel]\Sub{Nom} [heißen Rums]\Sub{Gen}]\Sub{Nom} sind genug.\\
-                                       some   spoons hot    rum  are   enough\\
-				     \glt A few spoonful of hot rum is enough.}
+    \ex{\label{ex:determiners:nac}\gll [[Drei Centiliter]\Sub{Nom} [heißer Rum]\Sub{Nom}]\Sub{Nom} sind genug.\\
+                                       three centilitres hot rum are enough\\
+				     \glt Three centilitres of hot rum is enough.}
+    \ex{\label{ex:determiners:pgc}\gll [[Einige Centiliter]\Sub{Nom} [heißen Rums]\Sub{Gen}]\Sub{Nom} sind genug.\\
+                                       some   centilitres hot    rum  are   enough\\
+				     \glt A few centilitres of hot rum is enough.}
   \end{xlist}
@@
 
+In (\ref{ex:determiners:nac}), the measure noun is modified by a cardinal \textit{drei} `three', and hence the \NACa\ is preferred.
+In (\ref{ex:determiners:pgc}), the measure noun is modified by a non-cardinal determiner \textit{einige} `some', and the \PGCa\ is preferred.
+Especially with abstract physical measure nouns (like \textit{centilitre}), exact numerical quantification is invited.
 The statistical model presented in Section~\ref{sec:corpusstudies} will be specified in a way that it can detect such a preference.

### 2017-04-14 Latex/article.tex
More Bildhauer comments, PhilStat moved to Sec 3.

@@
 
-Before moving on to the description of the phenomenon and the corpus study in Sections~\ref{sec:descriptive} and \ref{sec:analyses}, I comment on the statistical methods used in the corpus study in Section~\ref{sec:rightstatistics} because the discussion is closely related to basic questions of the cognitive approach in corpus linguistics.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{On statistical analysis}
-\label{sec:rightstatistics}
-
-In this section, I motivate the choice of statistical models which I use later in Section~\ref{sec:corpusstudies}.
-Readers might think that the method used here for modeling grammatical alternations -- namely (Hierarchical) Logistic Regression\slash Generalised Linear (Mixed) Models or GLM(M)s -- does not require much motivation.
-After all, GLMMs have been established as the major tool in the analysis of alternation phenomena.
-All studies mentioned at the outset of Section~\ref{sec:cogocl} use some form of regression\slash GLM(M).
-Over the past few years, however, modified or alternative methods have been proposed.
-From among these methods, I just make a few remarks on Bayesian estimation (see \citealp{GelmanEa2014}), as it was proposed in \cite{Levshina2016,Divjak2016a}, for example.
-Conceptually, I see three points of discussion that should be kept apart.
-First, Bayesian methods are sometimes touted as superior tools for scientific inference compared to frequentist methods.
-Second, it has been proposed that the Bayesian interpretation of probability is more cognitively adequate for the modeling of linguistic data.
-Third, and very specific to this paper, given established methods in the modeling of alternation and variation, it has to be decided whether so-called Bayesian methods lead to substantially different results.
-
-As for the first point, this is not quite the place to discuss it fully.
-The basic distinction is a philosophical one and related to the concepts of \textit{direct} and \textit{inverse probability} (\eg\ \citealp{Senn2011}).
-Frequentists assume that models and parameters are fixed, for example a model specifying that a coin is fair.
-We can then calculate for observed data (for example a measurement of $3$ heads in an experiment with $10$ tosses) how often such a result or a more extreme result would occur if the model were true and we repeated the experiment arbitrarily often.
-This is essentially the frequentist notion of direct probability, \ie\ long-run frequencies under replication.
-Standard tests in the Fisher and Neyman-Pearson traditions as well as Neyman confidence intervals are based on this concept of probability.
-Bayesian approaches (in the now common interpretation), on the other hand, condition on the particular data and quantify inductively the probability of model parameters given the data.
-The parameters are thus not fixed, and the probability is usually equated with researchers' posterior beliefs about model parameters.
-There is actually a debate among Bayesians about the proper interpretation of Bayesian methods, and whether a notion of hypothesis testing is compatible (or even already contained) in the Bayesian approach.
-In \citet[10]{GelmanShalizi2013}, the authors -- prominent Bayesians themselves -- acknowledge that a theory of statistical testing is a desideratum, and they state about the standard inductive interpretation of Bayesianism that \textit{most of this received view of Bayesian inference is wrong}, and they develop a Bayesian notion of $p$ values (see also \citealp{Mayo2013}, for a frequentist reply; also \citealp{Senn2011} on different strands of Bayesianism and their stance on inductive vs.\ deductive reasoning, and \citealp{Mayo2011}, for a critical reply to \citealp{Senn2011}).
-Clearly, in such quarrels between and among camps of philosophers of science and statisticians, it is difficult for mere practitioners to take sides.
-
-These quarrels relate to the second point, however.
-\citet[301--302]{Divjak2016a} speaks favourably of Bayesian methods because the Bayesian concept of probability is more adequate for cognitive modeling compared to the frequentist one.
-Her argument is part of a larger body of literature asking for cognitively plausible modeling techniques, for example Naive Discriminative Learning (NDL; \citealp{Baayen2011,BaayenEa2013,MilinEa2016,TheijssenEa2013}).
-On p.\ 303 of \cite{Divjak2016a}, the author goes on to explicitly mention NDL as well.
-Yet, neither frequentist nor Bayesian methods were conceived as cognitive models, but as systems of inference for scientists (see above, and see also \citealp[302]{Divjak2016a}).
-The fundamental question that lurks behind such arguments is how we interpret our statistical models (estimated on corpus data).
-Are they inductive models of cognitive representations that also human learners would infer from being exposed to the corpus data?%
-\footnote{In which case we are doing \textit{data science in language research} in the words of \citealp{MilinEa2016}.
-I see this as standing in contradiction to the view advocated in \cite{Dabrowska2016} as cited above.}
-Or are they tests of theories that are pre-specified and merely tested for predictive accuracy on linguistic output data contained in corpora?
-In the former case, we adopt a strong \textit{corpus as input} hypothesis and should definitely resort to methods like NDL.
-However, this would most likely require us to toss most previous work done in (cognitive) linguistics into the bin and to abandon all high-level generalisations that most existing studies have been based upon.
-In fact, arguments to this extreme effect and against using high-level generalisation have been made, for example in \cite{BaayenEa2016}, \citet[299--300]{Divjak2016a}, \cite{RamscarPort2016}, \cite{TheijssenEa2013}.
-In the latter and less extreme case, the cognitive commitment does, however, not necessarily extend to the statistical methods used.
-These methods then do not need to be any more cognitively plausible than an ANOVA used to analyse the results from an experiment.
-I view my own work (see Sections~\ref{sec:germanmeasurenps}--\ref{sec:experimental}) in the tradition of testing theories (which embrace high-level generalisations), and I agree to provisionally use Likelihood methods (see below).
-I remain fully agnostic with respect to the question which approach is `right'.
-The best strategy for cognitive linguistics as a field might be to cultivate many methods while making sure that each method is applied carefully and competently.
-
-The third point, then, is the most practically relevant in the context of this paper.
-In her otherwise excellent study, \citet[251--252]{Levshina2016} argues for Bayesian estimation in mixed regression settings.
-First, she claims that \textit{while frequentist statistics only allows one to test whether the null hypothesis can be rejected, Bayesian statistics enables one both to test the null hypothesis and to estimate the probability of specific parameter values given the data}.
-This does not do justice to frequentist methods in that mere rejection of the null hypothesis is characteristic only of Fisher's approach.
-In the Neyman-Pearson approach, results ideally \textit{favour} the main hypothesis vis-à-vis the alternative hypothesis (cf.\ \citealp{Lehmann1993,Lehmann2011,Perezgonzalez2015}).
-Also, especially Neyman-style frequentism has well-known extensions to estimation, for example in the form of confidence intervals (see \citealp{GreenlandEa2016}, esp.\ p.\ 340).
-She then explains that a \textit{distinctive feature of Bayesian statistics is the use of so-called priors} and that \textit{posterior probabilities depend on both the prior beliefs and the data, whereas the results of a frequentist model depend only on the data} \cite[252]{Levshina2016}.
-Remarkably, given this statement, she does \textit{not} use informative priors, and in her footnote 8 \cite[252]{Levshina2016} admits that priors were probed using trial and error.
-So, the proclaimed major advantage of Bayesian modeling was apparently not taken advantage of.%
-\footnote{In the words of \cite{Senn2011}: \textit{You may believe you are a Bayesian but you are probably wrong}.
-\citet[347--348]{GelmanHill2006} \textit{view any noninformative prior distribution as inherently provisional} and give recommendations how to proceed once posteriors have been obtained from noninformative priors.}
-Now, Maximum Likelihood Estimation (MLE) -- the traditional method which could have been used instead -- is not exactly \textit{frequentist} in the sense of Neyman-Pearson testing theory.
-MLE, like inductive Bayesianism, conditions on the particular data in as much as it searches for the most likely set of parameters given the data.
-What is more, Bayesian estimators are in fact based the Likelihood and merely multiply it by the prior \citep[6--8]{GelmanEa2014}.
-If the prior is flat, results converge (see also \citealp[347]{GelmanHill2006}).
-The same is true if the sample size is large compared to the number of parameters, at least for finite-dimensional parameter models \citep[1119--1120]{Freedman1999}, a well-established result known as the Bernstein-von Mises theorem.
-With a modest model structure including $17$ fixed effects and $2,646$ data points in \cite{Levshina2016}, it is highly likely that the same results would have obtained with Maximum Likelihood methods.
-In fact, she admits that changing the priors did not lead to substantially different results in her footnote 8.
-This is a clear sign that the prior is \textit{swamped by the data} \citep[1119]{Freedman1999}.
-Going back to the argument that Bayesian methods are allegedly more cognitively plausible than `frequentist' methods, the claim might still be true.
-However, most similar studies (dating back to \citealp{BresnanEa2007}) use an estimator (MLE) that is not frequentist in a narrow sense, and which is as cognitively plausible as Bayesian estimators because it most likely produces the same results in these studies.
-If for Levshina's study, results from Bayesian and MLE methods did, in fact, \textit{not} converge, it would have been an ideal occasion to demonstrate the superiority of the algorithms used in Bayesian estimation.
-After all, there are situations where Bayesian estimators can be more robust, namely with heavily censored data, complex hierarchical models, perfect separation, etc.
-(see \citealp{Freedman1999}, \citealp[345--348]{GelmanHill2006}).
-I want to reiterate that these points do not in any way invalidate the results presented in \cite{Levshina2016}.
-However, being \textit{Bayesian} is most likely not among its selling points.
-That said, I want to voice the concern that probably many practitioners are already struggling with getting an adequate grasp of advanced statistical methods, and that it might therefore be wise to use the more conservative and better understood method if the alternative method is not absolutely required for substantive reasons.
-In Section~\ref{sec:corpushierarchicalmodel}, I compare my own hierarchical models estimated with Bayesian and MLE estimators to demonstrate their expectable convergence.
-
-This concludes the general introduction.
 In the remainder of the paper, I introduce the alternation phenomenon (Section~\ref{sec:descriptive}) and suggest a theory-driven set of factors influencing the alternation (Section~\ref{sec:analyses}).
@@
 This is not surprising in as much as the genitive -- an intrinsic part of the PGC -- is generally underrepresented in colloquial vernacular variants of German as a result of a diachronic process wherein many (but by no means all) uses of the genitive are replaced by other cases or periphrastic constructions \citep{FleischerSchallert2011}.
-Under an integral view of prototypes, which incorporate effects related to larger contexts and registers, such preferences can be to be part of the construction prototypes.
+Under an integral view of prototypes, which incorporate effects related to larger contexts and registers, such preferences can be part of what defines the construction prototypes.
 In the corpus study in Section~\ref{sec:corpusstudies}, register effects will therefore be modeled -- even if only using two very simple proxy variables.
@@
 In order to reduce the sample size for the manual annotation process, the final concordance was sampled from the results of these 2,365 queries.
-Since the mass nouns in the sample were distributed according to the usual power law, I used all hits for mass nouns occurring less than one hundred times, but randomly sampled one hundred hits for each mass noun that occurred one hundred or more times.
+Since the mass nouns in the sample were distributed according to the usual power law, I used all hits for nouns with a frequency up to $100$, and a sample of $100$ of all those with higher frequency.
 The final sample contained 6,843 sentences, which was reduced to 5,063 in the manual annotation process due to removal of noisy material, erroneous hits and uninformative cases where the measure noun was in the genitive, in which case the \NACa\ cannot be distinguished from the \PGCa.
@@
 
-The variables \textit{Kindattraction} and \textit{Measureattraction} encode the ratio with which a given kind noun lemma or measure noun lemma occurs in the \NACb\ and the \PGCd.
+The variables \textit{Kindattraction} and \textit{Measureattraction} encode the ratio with which a given kind noun lemma or measure noun lemma occurs in the \PGCd\ and the  \NACb.
 They were calculated from the auxiliary samples described at the end of Section~\ref{sec:gettingdata} as a log-transformed quotient.
-The higher the value, the more often the noun occurs in the \PGCa (proportionally).%
+The higher the value, the more often the noun occurs in the \PGCd (proportionally).%
 \footnote{
@@
 To capture this, class information was annotated for measure lemmas.
-The classification was inspired by the list in \cite[530]{Koptjevskaja2001}, but due to the low frequencies of many of the potential classes, a very coarse classification was finally used.
+The classification was inspired by the list in \citet[530]{Koptjevskaja2001}, but due to the low frequencies of many of the potential classes, a very coarse classification was finally used.
 With typical examples and their frequencies in the final sample, the classes are:
@@
 
-Finally, one variable as added as nuisance variable in the context of the present study.
+Finally, one variable was added as nuisance variable in the context of the present study.
 It was reported in the literature that MNPs in the dative and with a masculine or neuter kind noun favour the \PGCa\ more than the corresponding nominative and accusative MNPs \citep{Hentschel1993,Zimmer2015}.
@@
 To control for this effect, the case of the measure noun was manually annotated (variable \textit{Measurecase}).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{On statistical analysis}
+\label{sec:rightstatistics}
+
+In this section, I motivate the choice of statistical models which I use in Section~\ref{sec:corpushierarchicalmodel}.
+Readers might think that the method used here for modeling grammatical alternations -- namely (Hierarchical) Logistic Regression\slash Generalised Linear (Mixed) Models or GLM(M)s -- does not require much motivation.
+After all, GLMMs have been established as the major tool in the analysis of alternation phenomena.
+All studies mentioned at the outset of Section~\ref{sec:cogocl} use some form of regression\slash GLM(M).
+Over the past few years, however, modified or alternative methods have been proposed.
+From among these methods, I just make a few remarks on Bayesian estimation (see \citealp{GelmanEa2014}), as it was proposed in \cite{Levshina2016,Divjak2016a}, for example.
+Conceptually, I see three points of discussion that should be kept apart.
+First, Bayesian methods are sometimes touted as superior tools for scientific inference compared to frequentist methods.
+Second, it has been proposed that the Bayesian interpretation of probability is more cognitively adequate for the modeling of linguistic data.
+Third, and very specific to this paper, given established methods in the modeling of alternation and variation, it has to be decided whether so-called Bayesian methods lead to substantially different results.
+
+As for the first point, this is not quite the place to discuss it fully.
+The basic distinction is a philosophical one and related to the concepts of \textit{direct} and \textit{inverse probability} (\eg\ \citealp{Senn2011}).
+Frequentists assume that models and parameters are fixed, for example a model specifying that a coin is fair.
+We can then calculate for observed data (for example a measurement of $3$ heads in an experiment with $10$ tosses) how often such a result or a more extreme result would occur if the model were true and we repeated the experiment arbitrarily often.
+This is essentially the frequentist notion of direct probability, \ie\ long-run frequencies under replication.
+Standard tests in the Fisher and Neyman-Pearson traditions as well as Neyman confidence intervals are based on this concept of probability.
+Bayesian approaches (in the now common interpretation), on the other hand, condition on the particular data and quantify inductively the probability of model parameters given the data.
+The parameters are thus not fixed, and the probability is usually equated with researchers' posterior beliefs about model parameters.
+There is actually a debate among Bayesians about the proper interpretation of Bayesian methods, and whether a notion of hypothesis testing is compatible (or even already contained) in the Bayesian approach.
+In \citet[10]{GelmanShalizi2013}, the authors -- prominent Bayesians themselves -- acknowledge that a theory of statistical testing is a desideratum, and they state about the standard inductive interpretation of Bayesianism that \textit{most of this received view of Bayesian inference is wrong}, and they develop a Bayesian notion of $p$ values (see also \citealp{Mayo2013}, for a frequentist reply; also \citealp{Senn2011} on different strands of Bayesianism and their stance on inductive vs.\ deductive reasoning, and \citealp{Mayo2011}, for a critical reply to \citealp{Senn2011}).
+Clearly, in such quarrels between and among camps of philosophers of science and statisticians, it is difficult for mere practitioners to take sides.
+
+These quarrels relate to the second point, however.
+\citet[301--302]{Divjak2016a} speaks favourably of Bayesian methods because the Bayesian concept of probability is more adequate for cognitive modeling compared to the frequentist one.
+Her argument is part of a larger body of literature asking for cognitively plausible modeling techniques, for example Naive Discriminative Learning (NDL; \citealp{Baayen2011,BaayenEa2013,MilinEa2016,TheijssenEa2013}).
+On p.\ 303 of \cite{Divjak2016a}, the author goes on to explicitly mention NDL as well.
+Yet, neither frequentist nor Bayesian methods were conceived as cognitive models, but as systems of inference for scientists (see above, and see also \citealp[302]{Divjak2016a}).
+The fundamental question that lurks behind such arguments is how we interpret our statistical models (estimated on corpus data).
+Are they inductive models of cognitive representations that also human learners would infer from being exposed to the corpus data?%
+\footnote{In which case we are doing \textit{data science in language research} in the words of \citealp{MilinEa2016}.
+I see this as standing in contradiction to the view advocated in \cite{Dabrowska2016} as cited above.}
+Or are they tests of theories that are pre-specified and merely tested for predictive accuracy on linguistic output data contained in corpora?
+In the former case, we adopt a strong \textit{corpus as input} hypothesis and should definitely resort to methods like NDL.
+However, this would most likely require us to toss most previous work done in (cognitive) linguistics into the bin and to abandon all high-level generalisations that most existing studies have been based upon.
+In fact, arguments to this extreme effect and against using high-level generalisation have been made, for example in \cite{BaayenEa2016}, \citet[299--300]{Divjak2016a}, \cite{RamscarPort2016}, \cite{TheijssenEa2013}.
+In the latter and less extreme case, the cognitive commitment does, however, not necessarily extend to the statistical methods used.
+These methods then do not need to be any more cognitively plausible than an ANOVA used to analyse the results from an experiment.
+I view my own work (see Sections~\ref{sec:germanmeasurenps}--\ref{sec:experimental}) in the tradition of testing theories (which embrace high-level generalisations), and I agree to provisionally use Likelihood methods (see below).
+I remain fully agnostic with respect to the question which approach is `right'.
+The best strategy for cognitive linguistics as a field might be to cultivate many methods while making sure that each method is applied carefully and competently.
+
+The third point, then, is the most practically relevant in the context of this paper.
+In her otherwise excellent study, \citet[251--252]{Levshina2016} argues for Bayesian estimation in mixed regression settings.
+First, she claims that \textit{while frequentist statistics only allows one to test whether the null hypothesis can be rejected, Bayesian statistics enables one both to test the null hypothesis and to estimate the probability of specific parameter values given the data}.
+This does not do justice to frequentist methods in that mere rejection of the null hypothesis is characteristic only of Fisher's approach.
+In the Neyman-Pearson approach, results ideally \textit{favour} the main hypothesis vis-à-vis the alternative hypothesis (cf.\ \citealp{Lehmann1993,Lehmann2011,Perezgonzalez2015}).
+Also, especially Neyman-style frequentism has well-known extensions to estimation, for example in the form of confidence intervals (see \citealp{GreenlandEa2016}, esp.\ p.\ 340).
+She then explains that a \textit{distinctive feature of Bayesian statistics is the use of so-called priors} and that \textit{posterior probabilities depend on both the prior beliefs and the data, whereas the results of a frequentist model depend only on the data} \cite[252]{Levshina2016}.
+Remarkably, given this statement, she does \textit{not} use informative priors, and in her footnote 8 \cite[252]{Levshina2016} admits that priors were probed using trial and error.
+So, the proclaimed major advantage of Bayesian modeling was apparently not taken advantage of.%
+\footnote{In the words of \cite{Senn2011}: \textit{You may believe you are a Bayesian but you are probably wrong}.
+\citet[347--348]{GelmanHill2006} \textit{view any noninformative prior distribution as inherently provisional} and give recommendations how to proceed once posteriors have been obtained from noninformative priors.}
+Now, Maximum Likelihood Estimation (MLE) -- the traditional method which could have been used instead -- is not exactly \textit{frequentist} in the sense of Neyman-Pearson testing theory.
+MLE, like inductive Bayesianism, conditions on the particular data in as much as it searches for the most likely set of parameters given the data.
+What is more, Bayesian estimators are in fact based the Likelihood and merely multiply it by the prior \citep[6--8]{GelmanEa2014}.
+If the prior is flat, results converge (see also \citealp[347]{GelmanHill2006}).
+The same is true if the sample size is large compared to the number of parameters, at least for finite-dimensional parameter models \citep[1119--1120]{Freedman1999}, a well-established result known as the Bernstein-von Mises theorem.
+With a modest model structure including $17$ fixed effects and $2,646$ data points in \cite{Levshina2016}, it is highly likely that the same results would have obtained with Maximum Likelihood methods.
+In fact, she admits that changing the priors did not lead to substantially different results in her footnote 8.
+This is a clear sign that the prior is \textit{swamped by the data} \citep[1119]{Freedman1999}.
+Going back to the argument that Bayesian methods are allegedly more cognitively plausible than `frequentist' methods, the claim might still be true.
+However, most similar studies (dating back to \citealp{BresnanEa2007}) use an estimator (MLE) that is not frequentist in a narrow sense, and which is as cognitively plausible as Bayesian estimators because it most likely produces the same results in these studies.
+If for Levshina's study, results from Bayesian and MLE methods did, in fact, \textit{not} converge, it would have been an ideal occasion to demonstrate the superiority of the algorithms used in Bayesian estimation.
+After all, there are situations where Bayesian estimators can be more robust, namely with heavily censored data, complex hierarchical models, perfect separation, etc.
+(see \citealp{Freedman1999}, \citealp[345--348]{GelmanHill2006}).
+I want to reiterate that these points do not in any way invalidate the results presented in \cite{Levshina2016}.
+However, being \textit{Bayesian} is most likely not among its selling points.
+That said, I want to voice the concern that probably many practitioners are already struggling with getting an adequate grasp of advanced statistical methods, and that it might therefore be wise to use the more conservative and better understood method if the alternative method is not absolutely required for substantive reasons.
+In Section~\ref{sec:corpushierarchicalmodel}, I compare my own hierarchical models estimated with Bayesian and MLE estimators to demonstrate their expectable convergence.
 
@@
 In this section, I report the results of fitting a multilevel model to the data using R \citep{R}, \textit{lme4} \citep{lme4} for Maximum Likelihood Estimation, and \textit{rstanarm} \citep{rstanarm} for Markov-Chain Monte Carlo estimation (see Section~\ref{sec:rightstatistics}).
-The purpose is to model the influence of the regressors specified in Table~\ref{tab:variables} on the probability that either the \PGCa\ is chosen over the \NACa.
+The purpose is to model the influence of the regressors specified in Table~\ref{tab:variables} on the probability that the \PGCa\ is chosen over the \NACa.
 All regressors from Table~\ref{tab:variables} were included, and the measure lemma and the kind noun lemma were specified as varying-intercept random effects.
@@
 The results of the estimation are shown in Table~\ref{tab:bigtable} and in Figure~\ref{fig:fixeffs}.
-The regressors with the measure lemma as their unit of reference have no within-measure lemma variance, and the \textit{glmer} function automatically estimates them as \textit{group level predictors} (or \textit{second-level effects}), cf.\ \cite[265--269,302--304]{GelmanHill2006}.
+The regressors with the measure lemma as their unit of reference have no within-measure lemma variance, and the \textit{glmer} function automatically estimates them as \textit{group level predictors} (or \textit{second-level effects}), cf.\ \citet[265--269,302--304]{GelmanHill2006}.
 The same goes for those listed with the kind lemma as their unit of reference.
@@
 Generalised variance inflation factors for the regressors were calculated to check for multicollinearity \citep{FoxMonette1992,ZuurEa2010}, and none of the corrected $\text{GVIF}^{1/2\text{df}}$ was higher than $1.6$.
-Nakagawa \& Schielzeth's pseudo-coefficient of determination is $R_m^2=0.409$ and $R^2_c=0.495$ (see \citealp{Gries2015} for a linguist-friendly introduction to these $R^2$ measures, or else \citealp{NakagawaSchielzeth2013}).
+Nakagawa \& Schielzeth's pseudo-coefficient of determination is $R_m^2=0.409$ and $R^2_c=0.495$ (see \citealp{Gries2015} for a basic introduction to these $R^2$ measures, or else \citealp{NakagawaSchielzeth2013}).
 The rate of correct predictions is $0.843$, which means a proportional reduction of error of $\lambda=0.297$.
-The measure lemma intercepts have a standard deviation of $\sigma_{\text{Measurelemma}}=0.448$, the kind lemma intercept $\sigma_{\text{Kindlemma}}=0.604$.
+The lemma intercepts have standard deviations of $\sigma_{\text{Measurelemma}}=0.448$ and $\sigma_{\text{Kindlemma}}=0.604$.
 
@@
 \footnote{Effect plots were created using the \textit{effects} package \citep{Fox2003}.
-They show the changes in probability for the outcome (y axis) dependent on values of a regressor (x axis), at typical values of all other regressors.}
+They show the changes in probability for the outcome (y axis) dependent on values of a regressor (x axis), at typical values of all other regressors.
+The vertical bars (categorical variables) and grey areas (continuous variables) are asymptotic $95\%$ confidence intervals calculated from \textit{glmer}.
+They are not bootstrapped.}
 The effect is mostly as expected:
@@
 \cite{FordBresnan2013} use the \textit{split-100} task in which participants have to distribute $100$ points between the alternatives, assigning more points to more natural sounding alternative.
-In essence, participants distribute a probability mass between two variants, which is intended to produce more subtle results compared to a two-alternative forced-choice task \cite{Rosenbach2013}, where participants have to choose one of two variants.
+In essence, participants distribute a probability mass between two variants, which is intended to produce more subtle results compared to a two-alternative forced-choice task \citep{Rosenbach2013}, where participants have to choose one of two variants.
 The split-100 paradigm has been criticised in \cite{ArppeJaervikivi2007}.
@@
      \midrule
-     \textbf{high prob.\ for PGC\Subsf{adj}} & 4 sentences\Supsf{a} & 4 sentences\Supsf{a} \\
-     \textbf{low prob.\ for PGC\Subsf{adj}} & 4 sentences\Supsf{a} & 4 sentences\Supsf{a} \\
+     \textbf{high prob.\ for PGC\Subsf{adj}} & 4 sentences & 4 sentences \\
+     \textbf{low prob.\ for PGC\Subsf{adj}} & 4 sentences & 4 sentences \\
   \end{tabular}
-  \caption{The four groups of sentences chosen as stimuli (\Supsf{a}Among the 4 sentences, combinations of important factor values were made unique whenever possible.)}
+  \caption{The four groups of sentences chosen as stimuli; in each group of four sentences, combinations of important factor values were made unique whenever possible}
   \label{tab:experiment1:design}
@@
 A certain amount of the variance can be accounted for by idiosyncrasies of single sentences ($\sigma_{\text{Sentence}}=1.785$, $\sigma^{\text{Modelprediction}}_{\text{Sentence}}=5.996$, $16$ levels).%
-\footnote{I use $\sigma^f_r$ to denote the standard deviation of the of the random intercepts for the fixed effect $f$ varying by random effect $r$.}
+\footnote{I use $\sigma^f_r$ to denote the standard deviation of the random intercepts for the fixed effect $f$ varying by random effect $r$.}
 Also, among participants, there are clearly different preferences ($\sigma_{\text{Participant}}=0.781$, $\sigma^{\text{Modelprediction}}_{\text{Participant}}=0.484$, $24$ levels).
@@
 There were $38$ (different) participants recruited in exactly the same manner as for the experiment reported in Section~\ref{sec:exp:fc}.
-None of them had ever participated in any kind of reading experiment, and none of them took part in the first experiment (and vice versa).
+None of them had ever participated in any kind of reading experiment, and none of them took part in the first experiment.
 The experiment was conducted using \textit{PsychoPy}.
@@
 The reading times were residualised per speaker based on the reading times of all words (not just the targets) by that speaker.
-The adjective and the kind noun (\ie\ the constituents bearing the critical case markers) were used as the target region, such as the bracketed words in the example \textit{zwei Gläser} [\textit{sprudelndes Wasser}] `two glasses of sparkling water'.
+The adjective and the kind noun (\ie\ the constituents bearing the critical case markers) were used as the target region, such as the bracketed words in the example \textit{zwei Gläser} [\textit{spru\-deln\-des Wasser}] `two glasses of sparkling water'.
 Outliers farther than 2 inter-quartile ranges from the mean logarithmised residual time were removed ($64$ data points), resulting in a total number of $n=1,152$ observations.

### 2017-04-14 Latex/article.tex
Minor fixes.

@@
 \label{sec:cogocl}
-
-\subsection{Corpora, cognition, and alternations}
 
@@
 If the prior is flat, results converge (see also \citealp[347]{GelmanHill2006}).
-The same is true if the sample size is large compared to the number of parameters, at least for finite-dimensional parameter models \citep[1119--1120]{Freedman1999}, a well-established result known as the Bernstein-von Mises theorem.
+The same is true if the sample size is large compared to the number of parameters, at least for finite-dimensional parameter models \citep[1119--1120]{Freedman1999}, a well-established result known as the \textit{Bernstein-von Mises theorem}.
 With a modest model structure including $17$ fixed effects and $2,646$ data points in \cite{Levshina2016}, it is highly likely that the same results would have obtained with Maximum Likelihood methods.
@@
 
-In this section, I report the results of fitting a multilevel model to the data using R \citep{R}, \textit{lme4} \citep{lme4} for Maximum Likelihood Estimation, and \textit{rstanarm} \citep{rstanarm} for Markov-Chain Monte Carlo estimation (see Section~\ref{sec:rightstatistics}).
+In this section, I report the results of fitting a multilevel model to the data using R \citep{R}, \textit{lme4} \citep{lme4} for Maximum Likelihood Estimation, and \textit{rstanarm} \citep{rstanarm} for `Bayesian' Markov-Chain Monte Carlo estimation (see Section~\ref{sec:rightstatistics}).
 The purpose is to model the influence of the regressors specified in Table~\ref{tab:variables} on the probability that the \PGCa\ is chosen over the \NACa.
@@
 
-Standard diagnostics for the Maximum Likelihood Estimation (MLE) show that the model quality is quite high.
+Standard diagnostics for MLE show that the model quality is quite good.
 Generalised variance inflation factors for the regressors were calculated to check for multicollinearity \citep{FoxMonette1992,ZuurEa2010}, and none of the corrected $\text{GVIF}^{1/2\text{df}}$ was higher than $1.6$.
@@
 At the other end of the scale, nouns denoting natural portions like \textit{Haufen} `heap', \textit{Bündel} `bundle', \textit{Schluck} `gulp' favour the \PGCa.
-These are referential nouns, confirming the hypothesis that it is prototypical of the PGC to contain two referential nouns, while the NAC only contains one (the kind noun).
+These are referential nouns, confirming the hypothesis that it is prototypical of the PGC to contain two referential nouns, while the NAC prototypically only contains one (the kind noun).
 
@@
 In \cite{VerhoevenTemme2017}, it was shown that results from forced-choice and split-100 experiments mostly converge, which is expected in the long run if a probability is turned into a binary decision.
-I present a forced-choice experiment in Section~\ref{sec:exp:fc} and not a split-100 experiment, mainly because in a dry-run of the experiment, participants complained about the unnaturalness of the task of distributing a probability mass across two variants.
+I present a forced-choice experiment in this Section and not a split-100 experiment, mainly because in a dry-run of the experiment, participants complained about the unnaturalness of the task of distributing a probability mass across two variants.
 Participants had to chose between two sentences differing only in that one contained the \NACa\ and the other contained the \PGCa.
@@
 
-As stimuli, attested MNPs from the corpus study were used, but the sentences were radically simplified to avoid influences from contextual nuisance variables as much as possible.
+As stimuli, attested MNPs from the corpus study were used, but the sentences were radically simplified to avoid influences from contextual nuisance factors as much as possible.
 The approach is also justified because according to the theoretical assessment in Section~\ref{sec:analyses}, the choice of variants depends mostly on a very local constructional context.
@@
 The confidence interval from a parametric bootstrap ($1,000$ replications, percentile method) for the regressor is acceptable but slightly large with a lower bound of $1.626$ and an upper bound of $8.397$.
-Nakagawa \& Schielzeth's pseudo-coefficients of determination are $R^2_{m}=0.227$ and $R^2_{c}=0.561$, which means that over 22\% of the variance in the data can be explained by considering only the predictions from the corpus-based GLMM.
+The pseudo-coefficients of determination are $R^2_{m}=0.227$ and $R^2_{c}=0.561$, which means that over 22\% of the variance in the data can be explained by considering only the predictions from the corpus-based GLMM.
 The effect display for the single fixed regressor \textit{Modelprediction} is given in Figure \ref{fig:afc:effects}.
@@
 It is expected that reading less prototypical variants (the one assigned a low probability by the corpus-derived model) in a given context and with given lexical material incurs a processing overhead for the reader (\citealp{Kaiser2013}).
-In Section~\ref{sec:exp:spr}, a self-paced reading experiment is therefore presented.
+In this Section, a self-paced reading experiment is therefore presented.
 In a very similar fashion, \cite{DivjakEa2016} apply the self-paced reading paradigm in the validation of corpus-based models.
@@
 The adjective and the kind noun (\ie\ the constituents bearing the critical case markers) were used as the target region, such as the bracketed words in the example \textit{zwei Gläser} [\textit{spru\-deln\-des Wasser}] `two glasses of sparkling water'.
-Outliers farther than 2 inter-quartile ranges from the mean logarithmised residual time were removed ($64$ data points), resulting in a total number of $n=1,152$ observations.
+Outliers farther than 2 inter-quartile ranges from the mean logarithmised residualised reading time were removed ($64$ data points), resulting in a total number of $n=1,152$ observations.
 An LMM was specified with the logarithmised residual reading times as the response variable, and the probabilities derived from the corpus GLMM (\textit{Modelprediction}) as the main regressor.
@@
 Comparing the full model to a model without the main regressor \textit{Modelprediction} (and consequently also without the interaction with \textit{Construction}) in a PB test gives $\mpPB=0.036$.
-Nakagawa and Schielzeth's pseudo-determination coefficients are $R^2_m=0.237$ and $R^2_c=0.346$.
+The pseudo-coefficients of determination are $R^2_m=0.237$ and $R^2_c=0.346$.
 
@@
 
-The overall model quality is thus good, and the effect plot for the main effect is shown in Figure~\ref{fig:spr:effects}.
+The overall model quality is high, and the effect plot for the main effect is shown in Figure~\ref{fig:spr:effects}.
 The estimate for the sentences with \NACa\ is obviously imprecise, and no differences in reading times are observed.

### 2017-04-15 Latex/article.tex
Off to the Lektorat.

@@
   %\subtitle{...}
-  \abstract{}
+  \abstract{In this paper, a case alternation occurring in German measure noun phrases is examined from the perspective of cognitively oriented corpus linguistics.
+    In pseudo-partitive constructions such as \textit{eine Tasse guter Kaffe} `a cup of good coffee', the embedded kind noun either agrees in its case with the measure head noun, or it is a genitive as in \textit{eine Tasse guten Kaffees}.
+    This alternation occurs only under highly specific syntactic conditions.
+    The analysis is formulated in terms of Prototype Theory. 
+    It is assumed the variants are modelled after non-alternating neighbouring constructions, which represent the prototypes more clearly.
+    Mainly, it is argued the frequencies with which individual lemmas appear in the prototypical constructions partially predicts which variant is chosen in the alternating case, and that the genitive construction is prototypical of more strongly grammaticalised measure nouns.
+    A large-scale corpus study using the DECOW web corpus is reported to support the theoretical analysis.
+    In the statistical analysis of the corpus study, it is discussed whether Bayesian estimators are intrinsically better than Likelihood Maximisation methods, and all results are estimated with Maximum Likelihood and Bayesian methods for comparison.
+    Finally, a forced choice experiment and a self-paced reading experiment are reported.
+    Both experiments successfully correlate the usage-based corpus analysis with the behaviour of cognitive agents.
+    The present study therefore contributes to a now well-established strand of research into alternations using both corpus and experimental methods.}
   \keywords{corpus methods and experimental methods, self-paced reading, forced-choice, alternations, hierarchical models, measure constructions, German}
@@
 For such probabilistic generalisations estimated on corpus data, there has always been an interest in correlating the results with results from experimental work (for example, \citealp{ArppeJaervikivi2007,BresnanEa2007,BresnanFord2010,DivjakGries2008,DivjakEa2016,FordBresnan2013}).
-This often called a \textit{validation} of the corpus-derived findings, but \citet[303]{Divjak2016a} criticises this choice of words \textit{because it creates the impression that behavioral experimental data is inherently more valuable than textual data}, citing \cite{TummersEa2005}, who state that a corpus is \textit{a sample of spontaneous language use that is} (\textit{generally}) \textit{realized by native speakers}.
+This often called a \textit{validation} of the corpus-derived findings, but \citet[303]{Divjak2016a} rightly criticises this choice of words \textit{because it creates the impression that behavioral experimental data is inherently more valuable than textual data}, citing \cite{TummersEa2005}, who state that a corpus is \textit{a sample of spontaneous language use that is} (\textit{generally}) \textit{realized by native speakers}.
 This position, which I adopt, does not, however, imply that we can in some way \textit{deduce mental representations from patterns of use}, \ie\ from corpus data, as  \citet[486--487]{Dabrowska2016} convincingly argues.%
@@
 Under an integral view of prototypes, which incorporate effects related to larger contexts and registers, such preferences can be part of what defines the construction prototypes.
-In the corpus study in Section~\ref{sec:corpusstudies}, register effects will therefore be modeled -- even if only using two very simple proxy variables.
+In the corpus study in Section~\ref{sec:corpusstudies}, register effects will therefore be modelled -- even if only using two very simple proxy variables.
 
@@
 Finally, two auxiliary samples were also drawn.
-As mentioned in Section~\ref{sec:analyses}, the distribution of the measure noun and kind noun lemmas in the \NACb\ and the \PGCd\ with a determiner will be modeled as factors influencing the alternation.
+As mentioned in Section~\ref{sec:analyses}, the distribution of the measure noun and kind noun lemmas in the \NACb\ and the \PGCd\ with a determiner will be modelled as factors influencing the alternation.
 Therefore, all noun-noun pairs from the bootstrap process were also queried in the two non-alternating constructions, resulting in 17,252 hits for the \PGCd\ and 315,635 hits for the \NACb.
@@
 \textit{Portion} (natural portions like \textit{Happen} `bite' or \textit{Krümel} `crumb'; $f=713$).
-The few lemmas that did not fit into either of these classes were labeled \textit{Rest} ($f=278$).
+The few lemmas that did not fit into either of these classes were labelled \textit{Rest} ($f=278$).
 
@@
 First, Bayesian methods are sometimes touted as superior tools for scientific inference compared to frequentist methods.
-Second, it has been proposed that the Bayesian interpretation of probability is more cognitively adequate for the modeling of linguistic data.
+Second, it has been proposed that the Bayesian interpretation of probability is more cognitively adequate for the modeling of linguistic data \cite[301--302]{Divjak2016a}.
 Third, and very specific to this paper, given established methods in the modeling of alternation and variation, it has to be decided whether so-called Bayesian methods lead to substantially different results.
@@
 
+%The first and second point cannot be discussed in detail here.%
+%\footnote{However, among the Bayesian hype, there is also a number of critical papers by statisticians (even prominent Bayesians) and philosophers of science in which the extreme credibility that has been assigned to Bayesian approaches (and especially the standard inductivist interpretation of Bayesianism) is questioned (for example, \citealp{GelmanShalizi2013,Mayo2011,Senn2011}).}
 The third point, then, is the most practically relevant in the context of this paper.
@@
   \begin{tabular}{llrlp{0.5em}rrp{0.5em}rrp{0.5em}rrp{0.5em}cc}
-      Level & Regressor & \multicolumn{1}{l}{\pPB} & Level  && \multicolumn{2}{l}{Coefficient} && \multicolumn{2}{l}{CI low} && \multicolumn{2}{l}{CI high} && \multicolumn{2}{l}{CI contains 0} \\
+      Level & Regressor & \multicolumn{1}{l}{\pPB} & Level  && \multicolumn{2}{l}{Coefficient} && \multicolumn{2}{l}{CI low} && \multicolumn{2}{l}{CI high} && \multicolumn{2}{l}{CI excludes 0} \\
               &                   &       &           && \multicolumn{1}{l}{MLE} & \multicolumn{1}{l}{MCMC} && \multicolumn{1}{l}{MLE} & \multicolumn{1}{l}{MCMC} && \multicolumn{1}{l}{MLE} & \multicolumn{1}{l}{MCMC} && \multicolumn{1}{l}{MLE} & \multicolumn{1}{l}{MCMC} \\\midrule
@@
 The resulting 95\% bootstrap confidence intervals are reported in Table~\ref{tab:bigtable} in the columns labelled \textit{CI low} and \textit{CI high} (= upper and lower 2.5th percentiles).
-The column \textit{CI contains 0} shows an asterisk for those intervals that do \textit{not} include 0.
+The column \textit{CI contains 0} shows an asterisk for those intervals that do \textit{not} include $0$.
 Furthermore, for each regressor, a $p$ value was obtained by dropping the regressor from the full model, re-estimating the nested model and comparing it to the full model.
 Instead of inexact Wald approximations or Likelihood Ratio Tests, I used a drop-in bootstrap replacement for the Likelihood Ratio Test from the function \textit{PBmodcomp} from the \textit{pbkrtest} package \citep{HalekohHojsgaard2014}.
-I call the corresponding value $p_{\text{PB}}$, and it is given in the appropriately labeled columns in Table~\ref{tab:bigtable}.
+I call the corresponding value $p_{\text{PB}}$, and it is given in the appropriately labelled columns in Table~\ref{tab:bigtable}.
 Only \textit{Kindfreq} ($\mpPB=0.095$) can be seen as slightly too high to be convincing (non-significant).
@@
 Outliers farther than 2 inter-quartile ranges from the mean logarithmised residualised reading time were removed ($64$ data points), resulting in a total number of $n=1,152$ observations.
-An LMM was specified with the logarithmised residual reading times as the response variable, and the probabilities derived from the corpus GLMM (\textit{Modelprediction}) as the main regressor.
-Since the corpus GLMM predicts the probability of the \PGCa\ (vis-à-vis the \NACa), it is expected that reading times are positively correlated (longer reading times) with the probability if the stimulus actually contains the \NACa, and negatively correlated (shorter reading times) if the stimulus actually contains the \PGCa.
+An LMM was specified with the logarithmised residual reading times as the response variable.
+
+The probabilities derived from the corpus GLMM (\textit{Modelprediction}) were added as the main regressor of interest.
+It should be remembered that the corpus GLMM predicts the probability of the \PGCa, such that the higher the GLMM prediction, the more prototypical the sentence is for containing the \PGCa.
+Therefore, it is expected that reading times are higher when model \textit{Modelprediction} is higher and the sentence contains the \NACa.
+However, when the sentence contains the \PGCa, reading times should be lower when \textit{Modelprediction} is higher.
 To account for this, an interaction between \textit{Modelprediction} and \textit{Construction} (levels \textit{PGCa} and \textit{NACa}) was added to the model.
-Furthermore, the position ($1$--$98$) of the sentence in the individual experiment (\textit{Position} was included as a fixed effect to control for increasing reading speed during experiment runs.
+
+Furthermore, the position ($1$--$98$) of the sentence in the individual experiment (\textit{Position}) was included as a fixed effect to control for the usual increase in reading speed during an experiment run.
 Random intercepts were specified for \textit{Participant} and \textit{Item} (the $16$ sentence pairs are one \textit{Item} each).%
@@
   \begin{tabular}{lrrrc}
-    Regressor & \multicolumn{1}{r}{Coefficient} & \multicolumn{1}{r}{CI low} & \multicolumn{1}{r}{CI high} & \multicolumn{1}{r}{0 in CI} \\ \midrule
+    Regressor & \multicolumn{1}{r}{Coefficient} & \multicolumn{1}{r}{CI low} & \multicolumn{1}{r}{CI high} & \multicolumn{1}{r}{0 not in CI} \\ \midrule
     ConstructionPGCa                 &  0.054 &  0.012 &  0.095 &  *  \\ 
@@
 \footnote{In this context, it should be remembered from Section~\ref{sec:gettingdata} that even the \PGCd\ is much rarer that the \NACb\ ($17,252$ vs.\ $315,635$ occurrences in the auxiliary corpus samples).}
-Conversely, the \NACa\ (like the NAC in general) might be the highly frequent default which does not incur reading time penalty, even if it is not the optimal choice in the given context and with the given lexical material.
+Conversely, the \NACa\ (like the NAC in general) might be the highly frequent default which does not incur a reading time penalty, even if it is not the optimal choice in the given context and with the given lexical material.
 
@@
 In my view, the main tenets of this line of research are:
-(i)~Language, viewed from a cognitively realistic angle, is a probabilistic phenomenon and cannot be modeled appropriately within Aristotelian frameworks that assume discrete categories.
+(i)~Language, viewed from a cognitively realistic angle, is a probabilistic phenomenon and cannot be modelled appropriately within Aristotelian frameworks that assume discrete categories.
 (ii)~Corpora are collections of usage events (language production) and can therefore be used to evaluate both the claim made in (i) and specific theoretical claims (in case studies) about factors influencing speakers' decisions to use specific forms or constructions.
@@
 Finally, I want to point out that German is mildly under-researched in the specific framework used here.
-This is quite surprising given the fact that German morpho-syntax is famous for its alternations, which are usually called \textit{Zweifelsfälle} (`cases of doubt') in the traditional literature \citep{Klein2009}.
+This is quite surprising given the fact that German morpho-syntax is famous for its alternations, which are usually called \textit{Zweifelsfälle} (`cases of doubt') in the traditional literature \citep{Duden09,Klein2009}.
 Instead of being drowned in normative, descriptive, or didactic discussions, they could serve as ideal test cases in cognitive linguistics.

### 2017-04-16 Latex/article.tex
First round of Lektorat.

@@
   \abstract{In this paper, a case alternation occurring in German measure noun phrases is examined from the perspective of cognitively oriented corpus linguistics.
-    In pseudo-partitive constructions such as \textit{eine Tasse guter Kaffe} `a cup of good coffee', the embedded kind noun either agrees in its case with the measure head noun, or it is a genitive as in \textit{eine Tasse guten Kaffees}.
+    In pseudo-partitive constructions such as \textit{eine Tasse guter Kaffe} `a cup of good coffee', the embedded kind noun either agrees in its case with the measure head noun or it is a genitive as in \textit{eine Tasse guten Kaffees}.
     This alternation occurs only under highly specific syntactic conditions.
     The analysis is formulated in terms of Prototype Theory. 
-    It is assumed the variants are modelled after non-alternating neighbouring constructions, which represent the prototypes more clearly.
-    Mainly, it is argued the frequencies with which individual lemmas appear in the prototypical constructions partially predicts which variant is chosen in the alternating case, and that the genitive construction is prototypical of more strongly grammaticalised measure nouns.
-    A large-scale corpus study using the DECOW web corpus is reported to support the theoretical analysis.
+    I assume that the variants are modelled after non-alternating neighbouring constructions, which represent the prototypes more clearly.
+    Primarily, it is argued that the frequencies with which individual lemmas appear in the prototypical constructions partially predicts which variant is chosen in the alternating case, and that the genitive construction is prototypical of more strongly grammaticalised measure nouns.
+    I present a large-scale corpus study using the DECOW web corpus which supports the theoretical analysis.
     In the statistical analysis of the corpus study, it is discussed whether Bayesian estimators are intrinsically better than Likelihood Maximisation methods, and all results are estimated with Maximum Likelihood and Bayesian methods for comparison.
     Finally, a forced choice experiment and a self-paced reading experiment are reported.
-    Both experiments successfully correlate the usage-based corpus analysis with the behaviour of cognitive agents.
-    The present study therefore contributes to a now well-established strand of research into alternations using both corpus and experimental methods.}
+    In both experiments, the usage-based corpus analysis is successfully correlated with the behaviour of cognitive agents.
+    The present study therefore contributes to a now well-established area of research into alternations using both corpus and experimental methods.}
   \keywords{corpus methods and experimental methods, self-paced reading, forced-choice, alternations, hierarchical models, measure constructions, German}
@@
 By \textit{alternation} I refer to a situation where two or more forms or constructions are available with no clear difference in acceptability, function, or meaning.
-The study of alternations has a now longstanding tradition in cognitively oriented corpus linguistics (\citealp{BresnanEa2007,BresnanHay2010,BresnanFord2010,DivjakArppe2013,Gries2015,NessetJanda2010}, to name just a few publications).
-This strand of research is based on the assumption that language is a probabilistic phenomenon \citep{Bresnan2007}, and choices of variants are neither chosen deterministically nor fully at random.
-Instead, multifactorial models are constructed, which incorporate influencing factors from diverse levels, including contextual ones.
+The study of alternations has a long history in cognitively oriented corpus linguistics (\citealp{BresnanEa2007,BresnanHay2010,BresnanFord2010,DivjakArppe2013,Gries2015,NessetJanda2010}, to name just a few publications).
+This area of research is based on the assumption that language is a probabilistic phenomenon \citep{Bresnan2007} where choices of variants are chosen neither deterministically nor fully at random.
+Instead, multifactorial models are constructed which incorporate influencing factors from diverse levels, including contextual factors.
 The estimation of the model coefficients quantifies the influence that the factors have on the probability that either variant is chosen.
 
-For such probabilistic generalisations estimated on corpus data, there has always been an interest in correlating the results with results from experimental work (for example, \citealp{ArppeJaervikivi2007,BresnanEa2007,BresnanFord2010,DivjakGries2008,DivjakEa2016,FordBresnan2013}).
-This often called a \textit{validation} of the corpus-derived findings, but \citet[303]{Divjak2016a} rightly criticises this choice of words \textit{because it creates the impression that behavioral experimental data is inherently more valuable than textual data}, citing \cite{TummersEa2005}, who state that a corpus is \textit{a sample of spontaneous language use that is} (\textit{generally}) \textit{realized by native speakers}.
+For such probabilistic generalisations estimated on corpus data, there has always been an interest in correlating the findings with results from experimental work (for example, \citealp{ArppeJaervikivi2007,BresnanEa2007,BresnanFord2010,DivjakGries2008,DivjakEa2016,FordBresnan2013}).
+This often called a \textit{validation} of the corpus-derived findings, but \citet[303]{Divjak2016a} rightly criticises this choice of words ``because it creates the impression that behavioral experimental data is inherently more valuable than textual data'', citing \cite{TummersEa2005}, who state that a corpus is ``a sample of spontaneous language use that is'' (\textit{generally}) \textit{realized by native speakers}.
 This position, which I adopt, does not, however, imply that we can in some way \textit{deduce mental representations from patterns of use}, \ie\ from corpus data, as  \citet[486--487]{Dabrowska2016} convincingly argues.%
-\footnote{I suggest that the word \textit{induce} would be more appropriate than \textit{deduce}.}
+\footnote{In its colloquial meaning, the verb \textit{deduce} seems appropriate here.
+However, in terms of the scientific method, we are most likely talking about \textit{induction} rather than \textit{deduction}.}
 It would be highly surprising if this were possible.
-Nobody assumes that we can induce mental representations from experiments, which -- as opposed to corpus studies -- even allow for direct access to the cognitive agent and offer much better options to control experimental conditions and nuisance variables.
-Instead, a theory of cognitive representation is pre-specified, and predictions are derived from this theory \textit{before} the experiment or the corpus study is conducted.
+Nobody assumes that we can inductively infer mental representations from experiments, which -- as opposed to corpus studies -- even allow for direct access to the cognitive agent and offer much better possibilities to control experimental conditions and nuisance variables.
+Under the standard approach, we pre-specify a theory of cognitive representation.
+Then, predictions are derived from this theory \textit{before} the experiment or the corpus study is conducted to \textit{test} the theory.
 Technical problems of statistical inference aside (see also Section~\ref{sec:rightstatistics}), a successful experiment corroborates the theory, and a failed experiment weakens the theory.
@@
 
-That said, the present study is conducted within the general paradigm of cognitive corpus linguistics and includes a comparison of the corpus findings with results from two experiments.
+The present study is conducted within the general paradigm of cognitive corpus linguistics and includes a comparison of the corpus findings with results from two experiments.
 As a theoretical framework, I assume a model of similarity-based classification in the form of Prototype Theory.
@@
 I am sceptical that corpus analysis alone could ever decide which theory is more adequate, given that substantial doubt has been voiced whether even experimental methods are ultimately able to do so \citep{Barsalou1990}.%
-\footnote{In \cite{DivjakArppe2013}, a fruitful attempt was made to show how both models form a converging picture based on corpus data.}
-Prototype Theory (with feature abstractions) is simply preferred here because it fits the established alternation modeling paradigm, which relies on features being weighted in statistical models.
+\footnote{In \cite{DivjakArppe2013}, it was shown using corpus data how both models form a converging picture.}
+Prototype Theory (with feature abstractions) is preferred here simply because it fits the established alternation modeling paradigm, which relies on features being weighted in statistical models.
 In Section~\ref{sec:analyses}, a prototype-theoretical parlance is therefore adopted.
@@
 In the remainder of the paper, I introduce the alternation phenomenon (Section~\ref{sec:descriptive}) and suggest a theory-driven set of factors influencing the alternation (Section~\ref{sec:analyses}).
-Then, the corpus study is presented including an appropriate statistical analysis (Section~\ref{sec:corpusstudies}).
+Then, the corpus study is presented, including an appropriate statistical analysis (Section~\ref{sec:corpusstudies}).
 Two experiments that correlate results from the corpus study with experimental results are then reported (Section~\ref{sec:experimental}) before I sum up the paper in Section~\ref{sec:conclusion}.
@@
 In this section, I introduce and illustrate the relevant alternating constructions.
-I describe the narrowly defined syntactic configuration in which the alternation occurs, and I motivate the focus on \textit{only} this narrow range rather than, for example, the whole range of nominal constructions expressing quantities.
-
-I use the term \textit{measure noun phrase} (MNP) to refer a noun phrase (NP) in which a kind-denoting (count or mass) noun depends on another noun that specifies a quantity of the objects or the substance denoted by the kind noun.
+I describe the narrowly defined syntactic configuration in which the alternation occurs and I motivate the focus on \textit{only} this narrow range rather than, for example, the whole range of nominal constructions expressing quantities.
+
+I use the term \textit{measure noun phrase} (MNP) to refer a noun phrase (NP) in which a kind-denoting (count or mass) noun depends on another noun that specifies a quantity of the objects or the substance denoted by the kind-denoting noun.
 I call the kind-denoting noun simply the \textit{kind noun} and the quantity-denoting noun the \textit{measure noun}.
-For illustration purposes, in the English \textit{a cup of fine coffee}, \textit{cup} is the measure noun, and \textit{coffee} is the kind noun.
-Measure nouns can be all sorts of nouns which genuinely denote a quantity (such as \textit{litre} or \textit{amount}) but also nouns denoting containers, collections, etc.
+For illustration purposes, in the English \textit{a glass of good wine}, \textit{glass} is the measure noun, and \textit{wine} is the kind noun.
+Measure nouns can be all sorts of nouns which denote a quantity (such as \textit{litre} or \textit{amount}) but also those denoting containers, collections, etc.
 (such as \textit{glass} or \textit{bucket}).
-In a similar vein, \citet[284]{Brems2003} considers such nouns denoting containers etc.\ as measure nouns \textit{which, strictly speaking, do not designate a `measure', but display a more nebulous} (sic!) \textit{potential for quantification} (see also \citealp[530]{Koptjevskaja2001}, and \citealp[338]{Rutkowski2007}).
-
-In the case at hand, three different syntactic configurations need to be distinguished \wrt case assignment inside German measure noun phrases.
+In a similar vein, \citet[284]{Brems2003} considers such nouns denoting containers etc.\ as measure nouns ``which, strictly speaking, do not designate a `measure', but display a more nebulous potential for quantification'' (see also \citealp[530]{Koptjevskaja2001}, and \citealp[338]{Rutkowski2007}).
+
+In the case at hand, three different syntactic configurations need to be distinguished \wrt case assignment inside German MNPs.
 If the kind noun forms an NP with a determiner, the construction resembles (and is usually called) a \textit{pseudo-partitive} (on partitives and pseudo-partitives see, e.\,g., \citealp{Barker1998,Selkirk1977,Stickney2007,Vos1999}; for a recent application of the terminology to German, see \citealp{Gerstenberger2015}).%
@@
 %  \begin{xlist}
-    \ex[ ]{\label{ex:intro:pseudopartitive1}\gll Wir trinken [[eine Tasse]\Sub{Acc} [eines leckeren Kaffees]\Sub{Gen}]\Sub{Acc}.\\
-    we drink a cup a tasty coffee\\
-    \glt We drink a cup of a tasty coffee.}
+    \ex[ ]{\label{ex:intro:pseudopartitive1}\gll Wir trinken [[eine Glas]\Sub{Acc} [dieses Weins]\Sub{Gen}]\Sub{Acc}.\\
+    we drink a glass this wine\\
+    \glt We drink a glass of this wine.}
 %    \ex[*]{\label{ex:intro:pseudopartitive2} Wir trinken [[eine Tasse]\Sub{Acc} [einen leckeren Kaffee]\Sub{Acc}]\Sub{Acc}.}
@@
 
-If the kind noun is bare -- \ie\ if it comes neither with a determiner nor a modifying adjective -- it has to agree in case with the measure noun, and the genitive as seen in the PGC is not acceptable, see (\ref{ex:intro:narrowapposition2}).
+If the kind noun is bare -- \ie\ if it comes neither with a determiner nor a modifying adjective -- it has to agree in case with the measure noun as in (\ref{ex:intro:narrowapposition1}), and the genitive as seen in the PGC is not acceptable, see (\ref{ex:intro:narrowapposition2}).
 
 \begin{exe}
-%  \ex\label{ex:intro:narrowapposition}
-%  \begin{xlist}
-%    \ex[*]{\label{ex:intro:narrowapposition1} Wir trinken [[eine Tasse]\Sub{Acc} [Kaffees]\Sub{Gen}]\Sub{Acc}.}
-    \ex[ ]{\label{ex:intro:narrowapposition2}\gll Wir trinken [[eine Tasse]\Sub{Acc} [Kaffee]\Sub{Acc}]\Sub{Acc}.\\
-    we drink a cup coffee\\
-    \glt We drink a cup of coffee.}
-%  \end{xlist}
+  \ex\label{ex:intro:narrowapposition}
+  \begin{xlist}
+    \ex[ ]{\label{ex:intro:narrowapposition1}\gll Wir trinken [[ein Glas]\Sub{Acc} [Wein]\Sub{Acc}]\Sub{Acc}.\\
+    we drink a glass wine\\
+    \glt We drink a glass of wine.}
+    \ex[*]{\label{ex:intro:narrowapposition2} Wir trinken [[ein Glas]\Sub{Acc} [Weins]\Sub{Gen}]\Sub{Acc}.}
+  \end{xlist}
 \end{exe}
@@
 This construction is usually classified as a \textit{Narrow Apposition Construction} \citep{Loebel1986}, henceforth NAC.%
-\footnote{The construction as in (\ref{ex:intro:narrowapposition2}) is also referred to as the \textit{Direct Partitive Construction} for other Germanic languages in which the PGC with the synthetic genitive is not available.
-This nomenclature makes sense in contrast to the \textit{Indirect Partitive Construction} with prepositional linkers translating to `of' -- \ie\ analytic genitives -- in such languages, see \cite{HankamerMikkelsen2008} for Danish.
+\footnote{The construction as in (\ref{ex:intro:narrowapposition1}) is also referred to as the \textit{Direct Partitive Construction} for other Germanic languages in which the PGC with the synthetic genitive is not available.
+This nomenclature makes sense in contrast to the \textit{Indirect Partitive Construction} with prepositional linkers translating to \textit{of} -- \ie\ analytic genitives -- in such languages, see \cite{HankamerMikkelsen2008} for Danish.
 For German, this terminology is not distinctive enough, which is why I use the terms NAC and PGC.}
 Notice that the unavailability of the genitive on the kind noun can be seen as following from a rather quirky constraint that genitive NPs in German require the presence of some strongly case-marked element (determiner or adjective) in addition to the head noun in order to be acceptable (\citealp{GallmannLindauer1994,Schachtl1989}; see also \citealp[160]{Eisenberg2013b}).
-%While only an accusative is shown in (\ref{ex:intro:narrowapposition2}), the kind noun obligatory agrees in case also with nominative and dative measure nouns.
-
-The actual alternation can be observed \textit{only when the kind noun occurs with an attributive adjective but without a determiner}, as in (\ref{ex:intro:alternation}), where both the NAC in (\ref{ex:intro:alternation1}) and the PGC in (\ref{ex:intro:alternation2}) are equally acceptable.
+%While only an accusative is shown in (\ref{ex:intro:narrowapposition1}), the kind noun obligatory agrees in case also with nominative and dative measure nouns.
+
+The alternation can be observed \textit{only when the kind noun occurs with an attributive adjective but without a determiner}, as in (\ref{ex:intro:alternation}), where both the PGC in (\ref{ex:intro:alternation1}) and the NAC in (\ref{ex:intro:alternation2}) are equally acceptable.
 They are by-and-large functionally and semantically equivalent.
@@
   \begin{xlist}
-    \ex[ ]{\label{ex:intro:alternation1}\gll Wir trinken [[eine Tasse]\Sub{Acc} [heißen Kaffee]\Sub{Acc}]\Sub{Acc}.\\
-    we drink a cup hot coffee\\
-    \glt We drink a cup of hot coffee.}
-    \ex[ ]{\label{ex:intro:alternation2} Wir trinken [[eine Tasse]\Sub{Acc} [heißen Kaffees]\Sub{Gen}]\Sub{Acc}.}
+    \ex[ ]{\label{ex:intro:alternation1} \gll Wir trinken [[ein Glas]\Sub{Acc} [guten Weins]\Sub{Gen}]\Sub{Acc}.\\
+    we drink a glass good wine \\
+    \glt We drink a glass of good wine.}
+    \ex[ ]{\label{ex:intro:alternation2} Wir trinken [[ein Glas]\Sub{Acc} [guten Wein]\Sub{Acc}]\Sub{Acc}.}
   \end{xlist}
@@
   \centering
-  \begin{tabular}{lccc}
+  \begin{tabular}{llll}
     \multicolumn{1}{r}{kind NP is:} & bare noun NP & NP with adjective & NP with determiner \\
@@
     \midrule
-    narrow apposition         & \NACb & \NACa & ---   \\
-    pseudo-partitive genitive & ---   & \PGCa & \PGCd \\
+    \multirow{2}{*}{narrow apposition}
+                & \NACb                                                 & \NACa                                                   & \multirow{2}{*}{---}       \\
+		& (\ref{ex:intro:narrowapposition1}) \textit{Glas Wein} & (\ref{ex:intro:alternation2}) \textit{Glas guten Wein}  &                            \\
+    \midrule
+
+    \multirow{2}{*}{pseudo-partitive genitive} 
+                & \multirow{2}{*}{---}                                  & \PGCa                                                   & \PGCd                      \\
+                &                                                       & (\ref{ex:intro:alternation1}) \textit{Glas guten Weins} & (\ref{ex:intro:pseudopartitive1}) \textit{Glas dieses Weins} \\
   \end{tabular}
-  \caption{Distribution of the NAC and PGC constructions in different NP structures}
+  \caption{Distribution of the NAC and PGC constructions in different NP structures with examples and references to full example sentences}
   \label{tab:constructions}
@@
 \begin{enumerate}[i.]
-  \item{\label{it:intro:idiot} alleged alternatives to the \NACa\ and the \PGCa}
-%  \item{\label{it:intro:datsg} alternative \textit{weak} forms of dative singular neuter adjectives}
-%  \item{\label{it:intro:femsg} case syncretism in feminine NPs}
-  \item{\label{it:intro:plurl} similar constructions with plural/collective kind nouns}
-  \item{\label{it:intro:noifl} grammaticalised non-inflected measure nouns}
-  \item{\label{it:intro:other} alternative constructions for expressing quantities}
+  \item{\label{it:intro:idiot} alleged alternatives to the \NACa\ and the \PGCa,}
+%  \item{\label{it:intro:datsg} alternative \textit{weak} forms of dative singular neuter adjectives,}
+%  \item{\label{it:intro:femsg} case syncretism in feminine NPs,}
+  \item{\label{it:intro:plurl} similar constructions with plural/collective kind nouns,}
+  \item{\label{it:intro:noifl} grammaticalised non-inflected measure nouns,}
+  \item{\label{it:intro:other} alternative constructions for expressing quantities.}
 \end{enumerate}
@@
 First of all, (\ref{it:intro:idiot}) refers to claims found in some grammars that a generic nominative, accusative, and even dative on the kind noun are used instead of the genitive (\PGCa) or case agreement (\NACa).
-Overviews can be found in \citealp{Hentschel1993,Zimmer2015}, and (\ref{ex:intro:gennom}) shows a putative generic nominative on the kind noun inside an accusative MNP.
+Overviews can be found in \citealp{Hentschel1993} and \citealp{Zimmer2015}.
+Sentence (\ref{ex:intro:gennom}) shows a putative generic nominative on the kind noun inside an accusative MNP.
 
@@
 I treat these cases like other measure nouns because they enter into both the \NACa\ as in (\ref{ex:uninflectedmeasures:a}) and the \PGCa\ as in (\ref{ex:uninflectedmeasures:b}).
-In Section~\ref{sec:analyses}, degrees of grammaticalisation as a factor influencing the alternation will be discussed, however.
+In Section~\ref{sec:analyses}, however, degrees of grammaticalisation as a factor influencing the alternation will be discussed.
 
@@
 
-These construction have very idiosyncratic properties (the construction with \textit{voller} is discussed in \citealp{Zeldes2018}, the construction with \textit{mit} is discussed in \citealp{Bhatt1990}), they are not semantically equivalent to the NAC and the PGC, and they are only available for a small subset of measure nouns.
+These construction have very idiosyncratic properties (the construction with \textit{voller} is discussed in \citealp{Zeldes2018}.
+The construction with \textit{mit} is discussed in \citealp{Bhatt1990}), they are not semantically equivalent to the NAC and the PGC, and they are only available for a small subset of measure nouns.
 All of these constructions are used only with measure nouns denoting containers, and the whole NP always refers to the container, and never to the quantity contained in it.
@@
 
-My analysis is based on the idea that the relevant syntactic structure [N\Sub{1}~[A~N\Sub{2}]\Sub{NP\Sub{2}}]\Sub{NP\Sub{1}} as instantiated in the \NACa\ and the \PGCa\ is ambiguous.
-To show this, the first question is which noun constitutes the head of the whole construction.
-This was answered quite clearly by \cite{Loebel1986} already (see also \citealp[213]{Eschenbach1994}, and \citealp[16]{GallmannLindauer1994}).
-Subject-verb agreement is always realised on the measure noun, and we can therefore assume that the measure noun is the head.
-However, the MNP-internal structure has more interesting cues to offer if the strong\slash weak inflection patterns of adjectives are taken into account.
+My analysis is based on the idea that in the relevant syntactic structure [N\Sub{1}~[A~N\Sub{2}]\Sub{NP\Sub{2}}]\Sub{NP\Sub{1}} (as instantiated in the \NACa\ and the \PGCa), the adjective is morpho-syntactically ambiguous between a determiner and a modifying adjective.
+To show this, %the first question is which noun constitutes the head of the whole construction.
+%This was answered quite clearly by \cite{Loebel1986} already (see also \citealp[213]{Eschenbach1994}, and \citealp[16]{GallmannLindauer1994}).
+%Subject-verb agreement is always realised on the measure noun, and we can therefore assume that the measure noun is the head.
+%Furthermore, the internal structure of MNPs can only be understood if the strong\slash weak inflection patterns of adjectives are taken into account.
+the strong\slash weak inflection patterns of adjectives needs to be taken into account.
 In NPs with a strongly inflected determiner, attributive adjectives inflect according to the massively syncretistic \textit{weak} pattern.
@@
 %The only exception to this rule is the masculine accusative singular, which is also marked with \textit{-en}.}
-If there is no determiner (as is the case in the alternating constructions), attributive adjectives inflect like determiners themselves, however.
+If there is no determiner (as is the case in the alternating constructions), on the other hand, attributive adjectives inflect like determiners themselves.
 This is called the \textit{strong} inflectional pattern.
@@
 In a probabilistic framework like Prototype Theory, the crucial question (beyond simple lemma preference effects) is, however, what controls speakers' decisions to use either variant. 
-In the remainder of this section, I argue that the NAC and the PGC are prototypes associated with different degrees of the grammaticalisation of the measure noun, a related morpho-syntactic property, and register effects.
+In the remainder of this section, I argue that the NAC and the PGC are prototypes associated with different degrees of the grammaticalisation of the measure noun and related morpho-syntactic properties, as well as register effects.
 The degree of similarity of a given instance to either of the two prototypes makes speakers chose the \NACa\ or the \PGCa.
 
-I first turn to grammaticalisation.
+I first focus on grammaticalisation.
 It is often assumed that pseudo-partitives arise as a form of grammaticalised partitives (\eg\ \citealp[536--539]{Koptjevskaja2001} for Finnish and Estonian, \citealp[559]{Koptjevskaja2001} for European languages in general).
@@
 %
-The alternating constructions in German are both clearly pseudo-partitives, but the grammaticalisation paths uncovered by \citet[esp.\ 526--530]{Koptjevskaja2001} are still relevant in the case at hand.
+The grammaticalisation paths uncovered by \citet[esp.\ 526--530]{Koptjevskaja2001} are relevant in the case at hand.
 The grammaticalisation path can start out (in some languages) with constructions involving two referential nouns (not necessarily forming a single and contiguous NP) and a \textit{separative} meaning as in \textit{(cut) two slices from the cake} \citep[535]{Koptjevskaja2001}.
-The \textit{part-of} meaning of true partitives as in \textit{a slice of the cake} represents the first stage of a development wherein the measure noun already tends to loose some semantic content.
+The \textit{part-of} meaning of true partitives as in \textit{a slice of the cake} represents the first stage of a development wherein the measure noun can already lose some semantic content, when for example words like \textit{bite} are no longer necessarily interpreted as a piece bitten out of something.
 The pseudo-partitive stage finally instantiates a \textit{quantity-of} relation, potentially even leading to fully grammaticalised quantifiers such as \textit{a lot}.
@@
 As described above, the grammaticalisation path leads from NPs denoting individuated objects standing in a \textit{part-of} relation to a construction with a more diffuse \textit{quantity-of} relation.
-Both types of relations can be numerically quantified -- in as much as a precise number of \textit{parts} or a numerically exact \textit{quantity} can be specified.
+Both types of relations can be numerically quantified -- inasmuch as a precise number of \textit{parts} or a numerically exact \textit{quantity} can be specified.
 However, it is much more prototypical of quantities to be specified with numerical precision.
@@
 Finally, it was found that the \PGCa\ is more typical of higher registers or even exclusive to written language (see \citealp[320--323]{Hentschel1993}).
-This is not surprising in as much as the genitive -- an intrinsic part of the PGC -- is generally underrepresented in colloquial vernacular variants of German as a result of a diachronic process wherein many (but by no means all) uses of the genitive are replaced by other cases or periphrastic constructions \citep{FleischerSchallert2011}.
+This is not surprising inasmuch as the genitive -- an intrinsic part of the PGC -- is generally underrepresented in colloquial vernacular variants of German as a result of a diachronic process wherein many (but by no means all) uses of the genitive are replaced by other cases or periphrastic constructions \citep{FleischerSchallert2011}.
 Under an integral view of prototypes, which incorporate effects related to larger contexts and registers, such preferences can be part of what defines the construction prototypes.
@@
 Due to the noisy nature and unknown composition of the web, only carefully designed and established web corpora like the COW corpora or the SketchEngine corpora \citep{KilgarriffEa2014} should be used.
-Clearly, using search engine results is \textit{bad science} for many reasons, most prominently total irreproducibility of results, as \cite{Kilgarriff2006} pointed out more than ten years ago.
+Clearly, using search engine results is `bad science' for many reasons, most prominently total irreproducibility of results, as \cite{Kilgarriff2006} pointed out more than ten years ago.
 Careless use of search engine results is still found, however, see for example \citet[171--175]{DeclerckBrems2016}.}
@@
 Among the factors potentially influencing the alternation (see Section~\ref{sec:germanmeasurenps}) were lemma-specific preference effects.
-Therefore, it was highly desirable to obtain a sample in which most of the highly frequent actually occurring combinations of kind nouns and measure nouns were represented.
+Therefore, it was highly desirable to obtain a sample in which most of the highly frequent actually-occurring combinations of kind nouns and measure nouns were represented.
 I applied a three-stage bootstrap process in order to obtain such a sample.
@@
 Abstract nouns which partially behave like mass nouns (like \textit{Spaß} `fun’ or \textit{Gefahr} `danger’) were excluded because they are usually not quantified in the same way as concrete mass nouns.
-The hundredth selected mass noun was \textit{Schmuck} `jewelry’, which is the 3,054th most frequent noun in the original frequency list.
+The hundredth selected mass noun was \textit{Schmuck} `jewellery’, which is the $3,054$th most frequent noun in the original frequency list.
 
@@
   Three main points speak against such an approach in the present case (but see \citealp[246--249]{Levshina2016} for a different approach).
-  First, the attraction values will be used as regressors in a GLMM, and the values resulting from collostructional approaches, \ie\ logarithmised Fisher $p$ values, have a very unfavourable distribution in the case at hand.
+  First, the attraction values will be used as regressors in a hierarchical logistic regression, and the values resulting from collostructional approaches, \ie\ logarithmised Fisher $p$ values, have a very unfavourable distribution in the case at hand.
   They cluster around $0$, and they include values of $-\infty$.
@@
 
-In this section, I motivate the choice of statistical models which I use in Section~\ref{sec:corpushierarchicalmodel}.
-Readers might think that the method used here for modeling grammatical alternations -- namely (Hierarchical) Logistic Regression\slash Generalised Linear (Mixed) Models or GLM(M)s -- does not require much motivation.
+In this section, I justify the choice of statistical models which I use in Section~\ref{sec:corpushierarchicalmodel}.
+Readers might think that the method used here for modeling grammatical alternations -- namely (Hierarchical) Logistic Regression\slash Generalised Linear (Mixed) Models or GL(M)Ms -- does not require much justification.
 After all, GLMMs have been established as the major tool in the analysis of alternation phenomena.
-All studies mentioned at the outset of Section~\ref{sec:cogocl} use some form of regression\slash GLM(M).
+All studies mentioned at the outset of Section~\ref{sec:cogocl} use some form of regression\slash GL(M)M.
 Over the past few years, however, modified or alternative methods have been proposed.
-From among these methods, I just make a few remarks on Bayesian estimation (see \citealp{GelmanEa2014}), as it was proposed in \cite{Levshina2016,Divjak2016a}, for example.
+From among these methods, I just make a few remarks on Bayesian estimation (see \citealp{GelmanEa2014}), as it was proposed in \cite{Levshina2016} and \cite{Divjak2016a}, for example.
 Conceptually, I see three points of discussion that should be kept apart.
@@
 There is actually a debate among Bayesians about the proper interpretation of Bayesian methods, and whether a notion of hypothesis testing is compatible (or even already contained) in the Bayesian approach.
-In \citet[10]{GelmanShalizi2013}, the authors -- prominent Bayesians themselves -- acknowledge that a theory of statistical testing is a desideratum, and they state about the standard inductive interpretation of Bayesianism that \textit{most of this received view of Bayesian inference is wrong}, and they develop a Bayesian notion of $p$ values (see also \citealp{Mayo2013}, for a frequentist reply; also \citealp{Senn2011} on different strands of Bayesianism and their stance on inductive vs.\ deductive reasoning, and \citealp{Mayo2011}, for a critical reply to \citealp{Senn2011}).
+In \citet[10]{GelmanShalizi2013}, the authors -- prominent Bayesians themselves -- acknowledge that a theory of statistical testing is a desideratum, and they state about the standard inductive interpretation of Bayesianism that ``most of this received view of Bayesian inference is wrong'', and they develop a Bayesian notion of $p$ values (see also \citealp{Mayo2013}, for a frequentist reply; also \citealp{Senn2011} on different strands of Bayesianism and their stance on inductive vs.\ deductive reasoning, and \citealp{Mayo2011}, for a critical reply to \citealp{Senn2011}).
 Clearly, in such quarrels between and among camps of philosophers of science and statisticians, it is difficult for mere practitioners to take sides.
@@
 Are they inductive models of cognitive representations that also human learners would infer from being exposed to the corpus data?%
-\footnote{In which case we are doing \textit{data science in language research} in the words of \citealp{MilinEa2016}.
+\footnote{In which case we are doing ``data science in language research'' in the words of \citealp{MilinEa2016}.
 I see this as standing in contradiction to the view advocated in \cite{Dabrowska2016} as cited above.}
@@
 I view my own work (see Sections~\ref{sec:germanmeasurenps}--\ref{sec:experimental}) in the tradition of testing theories (which embrace high-level generalisations), and I agree to provisionally use Likelihood methods (see below).
-I remain fully agnostic with respect to the question which approach is `right'.
+I remain fully agnostic with respect to the question, which approach is \textit{right}.
 The best strategy for cognitive linguistics as a field might be to cultivate many methods while making sure that each method is applied carefully and competently.
@@
 In her otherwise excellent study, \citet[251--252]{Levshina2016} argues for Bayesian estimation in mixed regression settings.
-First, she claims that \textit{while frequentist statistics only allows one to test whether the null hypothesis can be rejected, Bayesian statistics enables one both to test the null hypothesis and to estimate the probability of specific parameter values given the data}.
+First, she claims that ``while frequentist statistics only allows one to test whether the null hypothesis can be rejected, Bayesian statistics enables one both to test the null hypothesis and to estimate the probability of specific parameter values given the data.''
 This does not do justice to frequentist methods in that mere rejection of the null hypothesis is characteristic only of Fisher's approach.
@@
 Also, especially Neyman-style frequentism has well-known extensions to estimation, for example in the form of confidence intervals (see \citealp{GreenlandEa2016}, esp.\ p.\ 340).
-She then explains that a \textit{distinctive feature of Bayesian statistics is the use of so-called priors} and that \textit{posterior probabilities depend on both the prior beliefs and the data, whereas the results of a frequentist model depend only on the data} \cite[252]{Levshina2016}.
+She then explains that a ``distinctive feature of Bayesian statistics is the use of so-called priors'' and that ``posterior probabilities depend on both the prior beliefs and the data, whereas the results of a frequentist model depend only on the data'' \cite[252]{Levshina2016}.
 Remarkably, given this statement, she does \textit{not} use informative priors, and in her footnote 8 \cite[252]{Levshina2016} admits that priors were probed using trial and error.
 So, the proclaimed major advantage of Bayesian modeling was apparently not taken advantage of.%
-\footnote{In the words of \cite{Senn2011}: \textit{You may believe you are a Bayesian but you are probably wrong}.
-\citet[347--348]{GelmanHill2006} \textit{view any noninformative prior distribution as inherently provisional} and give recommendations how to proceed once posteriors have been obtained from noninformative priors.}
+\footnote{In the words of \cite{Senn2011}: ``You may believe you are a Bayesian but you are probably wrong.''
+\citet[347--348]{GelmanHill2006} ``view any noninformative prior distribution as inherently provisional'' and give recommendations how to proceed once posteriors have been obtained from noninformative priors.}
 Now, Maximum Likelihood Estimation (MLE) -- the traditional method which could have been used instead -- is not exactly \textit{frequentist} in the sense of Neyman-Pearson testing theory.
-MLE, like inductive Bayesianism, conditions on the particular data in as much as it searches for the most likely set of parameters given the data.
+MLE, like inductive Bayesianism, conditions on the particular data inasmuch as it searches for the most likely set of parameters given the data.
 What is more, Bayesian estimators are in fact based the Likelihood and merely multiply it by the prior \citep[6--8]{GelmanEa2014}.
@@
 The same is true if the sample size is large compared to the number of parameters, at least for finite-dimensional parameter models \citep[1119--1120]{Freedman1999}, a well-established result known as the \textit{Bernstein-von Mises theorem}.
-With a modest model structure including $17$ fixed effects and $2,646$ data points in \cite{Levshina2016}, it is highly likely that the same results would have obtained with Maximum Likelihood methods.
+With a modest model structure including $17$ fixed effects and $2,646$ data points in \cite{Levshina2016}, it is highly likely that the same results would have been obtained with Maximum Likelihood methods.
 In fact, she admits that changing the priors did not lead to substantially different results in her footnote 8.
-This is a clear sign that the prior is \textit{swamped by the data} \citep[1119]{Freedman1999}.
+This is a clear sign that the prior is ``swamped by the data'' \citep[1119]{Freedman1999}.
 Going back to the argument that Bayesian methods are allegedly more cognitively plausible than `frequentist' methods, the claim might still be true.
@@
 However, being \textit{Bayesian} is most likely not among its selling points.
-That said, I want to voice the concern that probably many practitioners are already struggling with getting an adequate grasp of advanced statistical methods, and that it might therefore be wise to use the more conservative and better understood method if the alternative method is not absolutely required for substantive reasons.
+That said, I want to voice the concern that probably, many practitioners are already struggling with getting an adequate grasp of advanced statistical methods and that it might therefore be wise to use the more conservative and better understood method if the alternative method is not absolutely required for substantive reasons.
 In Section~\ref{sec:corpushierarchicalmodel}, I compare my own hierarchical models estimated with Bayesian and MLE estimators to demonstrate their expectable convergence.
@@
 
-Finally, I show that Bayesian methods in the form of Markov-Chain Monte Carlo sampling (MCMC) do not necessarily (and in this case predictably) lead to different results.
+Finally, I show that Bayesian methods in the form of Markov-Chain Monte Carlo sampling (MCMC) do not necessarily lead to different results.
+Actually, given the low complexity of the model and the large sample size, it would be surprising if they did (see Section~\ref{sec:rightstatistics}).
 The model was re-estimated using the \textit{stan\_glmer} function from the \textit{rstanarm} package, which provides an \textit{lme4}-compatible syntax for estimating common model types with the \textit{stan} software \citep{CarpenterEa2017}.
@@
 The results reported in Section~\ref{sec:corpushierarchicalmodel} generally confirm the hypotheses from Section~\ref{sec:analyses}.
-First, the prototypicality effect related to the non-alternating \PGCd\ and \NACb\ can be shown, see the effect plots in Figure~\ref{fig:eff:attraction}.%
+First, the prototypicality effect related to the non-alternating \PGCd\ and \NACb\ can be shown (see the effect plots in Figure~\ref{fig:eff:attraction}).%
 \footnote{Effect plots were created using the \textit{effects} package \citep{Fox2003}.
@@
 With kind nouns, higher frequency seems to favour the \PGCa ($\beta_{\text{Kindfreq}}=0.146$, $\mpPB=0.095$).
-However, there is no clear theoretical interpretation (see Section~\ref{sec:analyses}), and the estimate is imprecise (not significant at $\alpha=0.05$, see above).
+However, there is no clear theoretical interpretation (see Section~\ref{sec:analyses}) and the estimate is imprecise (not significant at $\alpha=0.05$, see above).
 The effect can therefore be ignored or treated as a nuisance variable.
@@
 
-Conretely, the exact same stimuli as in the forced-choice experiments were used.
-Each participant read both the $16$ sentences with the variant predicted by the corpus model and the $16$ modified sentences with the variant that the corpus model does not predict.%
+Concretely, the exact same stimuli as in the forced-choice experiments were used.
+Each participant read both the $16$ sentences with the variant predicted by the corpus model and the $16$ modified sentences with the variant that the corpus model did not predict.%
 \footnote{Notice that lemmas and their frequencies as well as lemma classes are included as regressors in the corpus-based GLMM, and there was consequently no additional controlling of lemma frequencies, etc.}
@@
 In the experiment, participants first read all sentences from the first block, then all sentences from the second block.
-It was made sure that from each target sentence pair, one sentence was assigned to the first block, and the other sentence to the other block.
-The assignment of members of the individual sentence pairs to the blocks was randomised for each participant individually, and so was the order within each block.
+From each target sentence pair, one sentence was assigned to the first block and the other sentence to the other block.
+The assignment of members of the individual sentence pairs to the blocks was randomised for each participant individually and so was the order within each block.
 The sentences from each pair of variants were kept as far apart as possible.
@@
 After each sentence, participants had to answer simple (non-metalinguistic) yes-no questions about the previous sentence as distractors.
-The distractor questions were different ones in the first and the second block.
+The distractor questions were different between the first and the second blocks.
 There were $38$ (different) participants recruited in exactly the same manner as for the experiment reported in Section~\ref{sec:exp:fc}.
@@
 The probabilities derived from the corpus GLMM (\textit{Modelprediction}) were added as the main regressor of interest.
-It should be remembered that the corpus GLMM predicts the probability of the \PGCa, such that the higher the GLMM prediction, the more prototypical the sentence is for containing the \PGCa.
+It should be remembered that the corpus GLMM predicts the probability of the \PGCa.
+As a consequence, the higher the GLMM prediction is, the more prototypical the sentence is for containing the \PGCa.
 Therefore, it is expected that reading times are higher when model \textit{Modelprediction} is higher and the sentence contains the \NACa.
@@
 This can be seen in the downward slope in the right panel of Figure~\ref{fig:spr:effects}.
-This fits into the general picture in as much as the construction with the lower frequency might be developing towards a more sharply defined prototype.%
+This fits into the general picture inasmuch as the construction with the lower frequency might be developing towards a more sharply defined prototype.%
 \footnote{In this context, it should be remembered from Section~\ref{sec:gettingdata} that even the \PGCd\ is much rarer that the \NACb\ ($17,252$ vs.\ $315,635$ occurrences in the auxiliary corpus samples).}

### 2017-04-16 Latex/article.tex
Another release candidate.

@@
   \abstract{In this paper, a case alternation occurring in German measure noun phrases is examined from the perspective of cognitively oriented corpus linguistics.
-    In pseudo-partitive constructions such as \textit{eine Tasse guter Kaffe} `a cup of good coffee', the embedded kind noun either agrees in its case with the measure head noun or it is a genitive as in \textit{eine Tasse guten Kaffees}.
+    In pseudo-partitive constructions such as \textit{eine Tasse guter Kaffe} `a cup of good coffee', the embedded kind-denoting noun either agrees in its case with the measure head noun or it is a genitive as in \textit{eine Tasse guten Kaffees}.
     This alternation occurs only under highly specific syntactic conditions.
     The analysis is formulated in terms of Prototype Theory. 
-    I assume that the variants are modelled after non-alternating neighbouring constructions, which represent the prototypes more clearly.
-    Primarily, it is argued that the frequencies with which individual lemmas appear in the prototypical constructions partially predicts which variant is chosen in the alternating case, and that the genitive construction is prototypical of more strongly grammaticalised measure nouns.
+    I assume that the variants are modelled after non-alternating neighbouring constructions, which represent the prototypes more purely.
+    Primarily, it is argued that the frequencies with which individual lemmas occur in the prototypical constructions partially predict which variant is chosen in the alternating case.
+    Another major prediction is that the genitive construction is prototypical of more strongly grammaticalised measure nouns.
     I present a large-scale corpus study using the DECOW web corpus which supports the theoretical analysis.
-    In the statistical analysis of the corpus study, it is discussed whether Bayesian estimators are intrinsically better than Likelihood Maximisation methods, and all results are estimated with Maximum Likelihood and Bayesian methods for comparison.
+    In the statistical analysis of the corpus study, it is discussed whether Bayesian estimators lead to substantially different results compared to Likelihood Maximisation methods in studies like this one.
+    The model is estimated with Maximum Likelihood and Bayesian methods for comparison.
     Finally, a forced choice experiment and a self-paced reading experiment are reported.
-    In both experiments, the usage-based corpus analysis is successfully correlated with the behaviour of cognitive agents.
+    The usage-based findings from the corpus study show a clear correlation with the behaviour of cognitive agents in both experiments.
     The present study therefore contributes to a now well-established area of research into alternations using both corpus and experimental methods.}
@@
 If the kind noun is a plural count noun as in \textit{ein Sack kleine Äpfel} (\NACa) or \textit{ein Sack kleiner Äpfel} (\PGCa) `a bag of small apples', a similar alternation between PGC and NAC can be observed.
-In line with experimental results reported in \citet[15--16]{Zimmer2015}, I found that the PGC is so dominant with plural kind nouns ($794$ of $861$ cases, or over 92\%, cf.\ Section~\ref{sec:corpusstudies}) that the alternation cannot be analysed in the same way as in the singular.%
+In line with experimental results reported in \citet[15--16]{Zimmer2015}, I found that the PGC is so dominant with plural kind nouns (794 of 861 cases, or over 92\%, cf.\ Section~\ref{sec:corpusstudies}) that the alternation cannot be analysed in the same way as in the singular.%
 \footnote{More concretely, if these cases were included in the regression analysis of the corpus data along with a factor encoding the number of the kind noun, this factor would most assuredly override any other regressor for the data points with a plural kind noun.}
@@
 It is known from frameworks like Collexeme Analysis \citep{GriesStefanowitsch2004} that lemmas are attracted with different strengths by competing constructions.
-If the \NACa\ is highly similar to the \NACb\ and the \PGCa\ is highly similar to the \PGCd, the probability with which individual noun lemmas appear in the alternating constructions should be predictable at least partly from the relative frequency with which the same lemmas are used in the non-alternating constructions (see \citealp[246--249]{Levshina2016} for a similar idea in a different context).
+If the \NACa\ is highly similar to the \NACb\ and the \PGCa\ is highly similar to the \PGCd, the probability with which individual noun lemmas appear in the alternating constructions should be predictable at least partly from the relative frequency with which the same lemmas are used in the non-alternating constructions.
 Nouns which occur proportionally more often in the \PGCd\ should favour the \PGCa, and nouns which occur proportionally more often in the \NACb\ should favour the \NACa.
@@
 Abstract nouns which partially behave like mass nouns (like \textit{Spaß} `fun’ or \textit{Gefahr} `danger’) were excluded because they are usually not quantified in the same way as concrete mass nouns.
-The hundredth selected mass noun was \textit{Schmuck} `jewellery’, which is the $3,054$th most frequent noun in the original frequency list.
+The hundredth selected mass noun was \textit{Schmuck} `jewellery’, which is the 3,054th most frequent noun in the original frequency list.
 
@@
 In order to reduce the sample size for the manual annotation process, the final concordance was sampled from the results of these 2,365 queries.
-Since the mass nouns in the sample were distributed according to the usual power law, I used all hits for nouns with a frequency up to $100$, and a sample of $100$ of all those with higher frequency.
+Since the mass nouns in the sample were distributed according to the usual power law, I used all hits for nouns with a frequency up to 100, and a sample of 100 of all those with higher frequency.
 The final sample contained 6,843 sentences, which was reduced to 5,063 in the manual annotation process due to removal of noisy material, erroneous hits and uninformative cases where the measure noun was in the genitive, in which case the \NACa\ cannot be distinguished from the \PGCa.
@@
 The full set of manually annotated variables for the main sample is given in Table~\ref{tab:variables}, and I briefly discuss it now.%
-\footnote{All numeric variables were also z-transformed (\ie\ centered to the mean and rescaled such that they have a standard deviation of $1$) to facilitate their interpretation in the regression models reported in the next section.}
+\footnote{All numeric variables were also z-transformed (\ie\ centered to the mean and rescaled such that they have a standard deviation of 1) to facilitate their interpretation in the regression models reported in the next section.}
 Notice first that \textit{Construction} is the response variable (or `dependent variable') with the values \textit{PGCa} and \textit{NACa}.
@@
   Three main points speak against such an approach in the present case (but see \citealp[246--249]{Levshina2016} for a different approach).
-  First, the attraction values will be used as regressors in a hierarchical logistic regression, and the values resulting from collostructional approaches, \ie\ logarithmised Fisher $p$ values, have a very unfavourable distribution in the case at hand.
-  They cluster around $0$, and they include values of $-\infty$.
+  First, the attraction values will be used as regressors in a hierarchical logistic regression, and the values resulting from collostructional approaches, \ie\ logarithmised Fisher p values, have a very unfavourable distribution in the case at hand.
+  They cluster around 0, and they include values of $-\infty$.
   Second, their main use in collostructional analyses is to \textit{sort} a list of collexemes and interpret their order.
@@
 With typical examples and their frequencies in the final sample, the classes are:
-\textit{Physical} (abstract precisely measurable units such as \textit{Liter} `litre', \textit{Meter} `metre', \textit{Gramm} `gram'; $f=1,968$),
-\textit{Container} (\textit{Eimer} `bucket'; $f=740$), 
-\textit{Amount} (\textit{Menge} `amount'; $f=1,364$), 
-\textit{Portion} (natural portions like \textit{Happen} `bite' or \textit{Krümel} `crumb'; $f=713$).
-The few lemmas that did not fit into either of these classes were labelled \textit{Rest} ($f=278$).
-
-The variable \textit{Cardinal} encodes whether the measure noun is modified by a cardinal ($f=1,939$) or not ($f=3,124$).
+\textit{Physical} (abstract precisely measurable units such as \textit{Liter} `litre', \textit{Meter} `metre', \textit{Gramm} `gram'; \textit{f=}1,968),
+\textit{Container} (\textit{Eimer} `bucket'; \textit{f=}740),
+\textit{Amount} (\textit{Menge} `amount'; \textit{f=}1,364), 
+\textit{Portion} (natural portions like \textit{Happen} `bite' or \textit{Krümel} `crumb'; \textit{f=}713).
+The few lemmas that did not fit into either of these classes were labelled \textit{Rest} (\textit{f=}278).
+
+The variable \textit{Cardinal} encodes whether the measure noun is modified by a cardinal (\textit{f=}1,939) or not (\textit{f=}3,124).
 The purpose of this variable is to test whether cardinals really favour the \NACa\ as hypothesised in Section~\ref{sec:analyses}.
@@
 Frequentists assume that models and parameters are fixed, for example a model specifying that a coin is fair.
-We can then calculate for observed data (for example a measurement of $3$ heads in an experiment with $10$ tosses) how often such a result or a more extreme result would occur if the model were true and we repeated the experiment arbitrarily often.
+We can then calculate for observed data (for example a measurement of 3 heads in an experiment with 10 tosses) how often such a result or a more extreme result would occur if the model were true and we repeated the experiment arbitrarily often.
 This is essentially the frequentist notion of direct probability, \ie\ long-run frequencies under replication.
@@
 There is actually a debate among Bayesians about the proper interpretation of Bayesian methods, and whether a notion of hypothesis testing is compatible (or even already contained) in the Bayesian approach.
-In \citet[10]{GelmanShalizi2013}, the authors -- prominent Bayesians themselves -- acknowledge that a theory of statistical testing is a desideratum, and they state about the standard inductive interpretation of Bayesianism that ``most of this received view of Bayesian inference is wrong'', and they develop a Bayesian notion of $p$ values (see also \citealp{Mayo2013}, for a frequentist reply; also \citealp{Senn2011} on different strands of Bayesianism and their stance on inductive vs.\ deductive reasoning, and \citealp{Mayo2011}, for a critical reply to \citealp{Senn2011}).
+In \citet[10]{GelmanShalizi2013}, the authors -- prominent Bayesians themselves -- acknowledge that a theory of statistical testing is a desideratum, and they state about the standard inductive interpretation of Bayesianism that ``most of this received view of Bayesian inference is wrong'', and they develop a Bayesian notion of p values (see also \citealp{Mayo2013}, for a frequentist reply; also \citealp{Senn2011} on different strands of Bayesianism and their stance on inductive vs.\ deductive reasoning, and \citealp{Mayo2011}, for a critical reply to \citealp{Senn2011}).
 Clearly, in such quarrels between and among camps of philosophers of science and statisticians, it is difficult for mere practitioners to take sides.
@@
 The same is true if the sample size is large compared to the number of parameters, at least for finite-dimensional parameter models \citep[1119--1120]{Freedman1999}, a well-established result known as the \textit{Bernstein-von Mises theorem}.
-With a modest model structure including $17$ fixed effects and $2,646$ data points in \cite{Levshina2016}, it is highly likely that the same results would have been obtained with Maximum Likelihood methods.
+With a modest model structure including 17 fixed effects and 2,646 data points in \cite{Levshina2016}, it is highly likely that the same results would have been obtained with Maximum Likelihood methods.
 In fact, she admits that changing the priors did not lead to substantially different results in her footnote 8.
@@
 All regressors from Table~\ref{tab:variables} were included, and the measure lemma and the kind noun lemma were specified as varying-intercept random effects.
-The sample size was $n=5,063$ with $1,1344$ cases of \PGCa\ and $3,929$ cases of \NACa.
+The sample size was \textit{n=}5,063 with 1,1344 cases of \PGCa\ and 3,929 cases of \NACa.
 The results of the estimation are shown in Table~\ref{tab:bigtable} and in Figure~\ref{fig:fixeffs}.
@@
 Standard diagnostics for MLE show that the model quality is quite good.
-Generalised variance inflation factors for the regressors were calculated to check for multicollinearity \citep{FoxMonette1992,ZuurEa2010}, and none of the corrected $\text{GVIF}^{1/2\text{df}}$ was higher than $1.6$.
+Generalised variance inflation factors for the regressors were calculated to check for multicollinearity \citep{FoxMonette1992,ZuurEa2010}, and none of the corrected $\text{GVIF}^{1/2\text{df}}$ was higher than 1.6.
 Nakagawa \& Schielzeth's pseudo-coefficient of determination is $R_m^2=0.409$ and $R^2_c=0.495$ (see \citealp{Gries2015} for a basic introduction to these $R^2$ measures, or else \citealp{NakagawaSchielzeth2013}).
-The rate of correct predictions is $0.843$, which means a proportional reduction of error of $\lambda=0.297$.
+The rate of correct predictions is 0.843, which means a proportional reduction of error of $\lambda=0.297$.
 The lemma intercepts have standard deviations of $\sigma_{\text{Measurelemma}}=0.448$ and $\sigma_{\text{Kindlemma}}=0.604$.
@@
 The coefficient estimates are specified in Table~\ref{tab:bigtable} for each regressor (or regressor level) in the columns labelled \textit{Coefficient}.
-For a robust quantification of the precision of the estimation, I ran a parametric bootstrap (using the \mbox{\textit{confint.merMod}} function from \textit{lme4}) with $1,000$ replications, and using the percentile method for the calculation of the intervals.
+For a robust quantification of the precision of the estimation, I ran a parametric bootstrap (using the \mbox{\textit{confint.merMod}} function from \textit{lme4}) with 1,000 replications, and using the percentile method for the calculation of the intervals.
 The resulting 95\% bootstrap confidence intervals are reported in Table~\ref{tab:bigtable} in the columns labelled \textit{CI low} and \textit{CI high} (= upper and lower 2.5th percentiles).
-The column \textit{CI contains 0} shows an asterisk for those intervals that do \textit{not} include $0$.
-Furthermore, for each regressor, a $p$ value was obtained by dropping the regressor from the full model, re-estimating the nested model and comparing it to the full model.
+The column \textit{CI contains 0} shows an asterisk for those intervals that do \textit{not} include 0.
+Furthermore, for each regressor, a p value was obtained by dropping the regressor from the full model, re-estimating the nested model and comparing it to the full model.
 Instead of inexact Wald approximations or Likelihood Ratio Tests, I used a drop-in bootstrap replacement for the Likelihood Ratio Test from the function \textit{PBmodcomp} from the \textit{pbkrtest} package \citep{HalekohHojsgaard2014}.
@@
 The model was re-estimated using the \textit{stan\_glmer} function from the \textit{rstanarm} package, which provides an \textit{lme4}-compatible syntax for estimating common model types with the \textit{stan} software \citep{CarpenterEa2017}.
-The algorithm was run with $4$ chains and $1,000$ iterations, and I used plausible default priors.
-Most notably, priors for coefficients were specified as $\mathcal{N}(0,10)$ because coefficients higher than $10$ or lower than $-10$ are extremely rare in well-specified models on appropriate data.%
-\footnote{Consider that with a coefficient of $10$, each increase by $1$ in the regressor variable means an increase in odds of $exp(10)=22,026.47$.
+The algorithm was run with 4 chains and 1,000 iterations, and I used plausible default priors.
+Most notably, priors for coefficients were specified as $\mathcal{N}(0,10)$ because coefficients higher than 10 or lower than $-10$ are extremely rare in well-specified models on appropriate data.%
+\footnote{Consider that with a coefficient of 10, each increase by 1 in the regressor variable means an increase in odds of $exp(10)=22,026.47$.
 To reliably estimate such coefficients, extremely large samples would be required.}
-The algorithm converged, and for all coefficients, the $\hat{R}$ diagnostic was exactly $1$.
+The algorithm converged, and for all coefficients, the $\hat{R}$ diagnostic was exactly 1.
 The resulting coefficients and intervals as well as an * are also given in Table~\ref{tab:bigtable} in the columns labelled \textit{MCMC}.
@@
 They show the changes in probability for the outcome (y axis) dependent on values of a regressor (x axis), at typical values of all other regressors.
-The vertical bars (categorical variables) and grey areas (continuous variables) are asymptotic $95\%$ confidence intervals calculated from \textit{glmer}.
+The vertical bars (categorical variables) and grey areas (continuous variables) are asymptotic 95\% confidence intervals calculated from \textit{glmer}.
 They are not bootstrapped.}
@@
 The first experiment tests preferences for constructions explicitly.
-\cite{FordBresnan2013} use the \textit{split-100} task in which participants have to distribute $100$ points between the alternatives, assigning more points to more natural sounding alternative.
+\cite{FordBresnan2013} use the \textit{split-100} task in which participants have to distribute 100 points between the alternatives, assigning more points to more natural sounding alternative.
 In essence, participants distribute a probability mass between two variants, which is intended to produce more subtle results compared to a two-alternative forced-choice task \citep{Rosenbach2013}, where participants have to choose one of two variants.
@@
   \centering
-  \begin{tabular}[h]{rcc}
+  \begin{tabular}[h]{lll}
      & Masculine\slash Neuter & Feminine \\
@@
 Then, a multilevel logistic regression was specified with the probability of the \PGCa\ predicted for each sentence by the corpus-based GLMM as the only fixed effect \textit{Modelprediction}.%
-\footnote{The document-level variables \textit{Badness} and \textit{Genitives} were set to $0$, which is the mean for z-transformed variables.}
+\footnote{The document-level variables \textit{Badness} and \textit{Genitives} were set to 0, which is the mean for z-transformed variables.}
 A random intercept and slope were added for the individual sentence (item) in order to catch idiosyncrasies of single sentences.
@@
 Coefficients were estimated with Maximum Likelihood Estimation (\textit{lmer} function from \textit{lme4}).
-The number of observations was $n=384$.
-
-A certain amount of the variance can be accounted for by idiosyncrasies of single sentences ($\sigma_{\text{Sentence}}=1.785$, $\sigma^{\text{Modelprediction}}_{\text{Sentence}}=5.996$, $16$ levels).%
+The number of observations was \textit{n=}384.
+
+A certain amount of the variance can be accounted for by idiosyncrasies of single sentences ($\sigma_{\text{Sentence}}=1.785$, $\sigma^{\text{Modelprediction}}_{\text{Sentence}}=5.996$, 16 levels).%
 \footnote{I use $\sigma^f_r$ to denote the standard deviation of the random intercepts for the fixed effect $f$ varying by random effect $r$.}
-Also, among participants, there are clearly different preferences ($\sigma_{\text{Participant}}=0.781$, $\sigma^{\text{Modelprediction}}_{\text{Participant}}=0.484$, $24$ levels).
-On the extreme sides, one participant chose the \PGCa\ in $13$ of $16$ cases, and two participants only chose it in $5$ of $16$ cases.
-The regressor \textit{Modelprediction} achieves $\mpPB=0.007$ ($1,000$ replications) and is estimated at $5.408$ relative to an intercept of $-1.304$.
-The confidence interval from a parametric bootstrap ($1,000$ replications, percentile method) for the regressor is acceptable but slightly large with a lower bound of $1.626$ and an upper bound of $8.397$.
+Also, among participants, there are clearly different preferences ($\sigma_{\text{Participant}}=0.781$, $\sigma^{\text{Modelprediction}}_{\text{Participant}}=0.484$, 24 levels).
+On the extreme sides, one participant chose the \PGCa\ in 13 of 16 cases, and two participants only chose it in 5 of 16 cases.
+The regressor \textit{Modelprediction} achieves $\mpPB=0.007$ (1,000 replications) and is estimated at 5.408 relative to an intercept of -1.304.
+The confidence interval from a parametric bootstrap (1,000 replications, percentile method) for the regressor is acceptable but slightly large with a lower bound of 1.626 and an upper bound of 8.397.
 The pseudo-coefficients of determination are $R^2_{m}=0.227$ and $R^2_{c}=0.561$, which means that over 22\% of the variance in the data can be explained by considering only the predictions from the corpus-based GLMM.
@@
 Concretely, the exact same stimuli as in the forced-choice experiments were used.
-Each participant read both the $16$ sentences with the variant predicted by the corpus model and the $16$ modified sentences with the variant that the corpus model did not predict.%
+Each participant read both the 16 sentences with the variant predicted by the corpus model and the 16 modified sentences with the variant that the corpus model did not predict.%
 \footnote{Notice that lemmas and their frequencies as well as lemma classes are included as regressors in the corpus-based GLMM, and there was consequently no additional controlling of lemma frequencies, etc.}
-To minimise repetition effects, the stimuli for each participant were separated into two blocks of $16$ targets and $33$ fillers per block.
+To minimise repetition effects, the stimuli for each participant were separated into two blocks of 16 targets and 33 fillers per block.
 In the experiment, participants first read all sentences from the first block, then all sentences from the second block.
@@
 The fillers also came in pairs such that the second block exclusively contained sentences to which participants had been exposed in the first block in slightly modified form.
-In total, each participant read $98$ sentences.
+In total, each participant read 98 sentences.
 After each sentence, participants had to answer simple (non-metalinguistic) yes-no questions about the previous sentence as distractors.
 The distractor questions were different between the first and the second blocks.
-There were $38$ (different) participants recruited in exactly the same manner as for the experiment reported in Section~\ref{sec:exp:fc}.
+There were 38 participants recruited in exactly the same manner as for the experiment reported in Section~\ref{sec:exp:fc}.
 None of them had ever participated in any kind of reading experiment, and none of them took part in the first experiment.
@@
 The adjective and the kind noun (\ie\ the constituents bearing the critical case markers) were used as the target region, such as the bracketed words in the example \textit{zwei Gläser} [\textit{spru\-deln\-des Wasser}] `two glasses of sparkling water'.
-Outliers farther than 2 inter-quartile ranges from the mean logarithmised residualised reading time were removed ($64$ data points), resulting in a total number of $n=1,152$ observations.
+Outliers farther than 2 inter-quartile ranges from the mean logarithmised residualised reading time were removed (64 data points), resulting in a total number of \textit{n=}1,152 observations.
 An LMM was specified with the logarithmised residual reading times as the response variable.
@@
 As a consequence, the higher the GLMM prediction is, the more prototypical the sentence is for containing the \PGCa.
-Therefore, it is expected that reading times are higher when model \textit{Modelprediction} is higher and the sentence contains the \NACa.
+Therefore, it is expected that reading times are higher when the value of \textit{Modelprediction} is higher but the sentence contains the \NACa.
 However, when the sentence contains the \PGCa, reading times should be lower when \textit{Modelprediction} is higher.
@@
 
-Furthermore, the position ($1$--$98$) of the sentence in the individual experiment (\textit{Position}) was included as a fixed effect to control for the usual increase in reading speed during an experiment run.
-Random intercepts were specified for \textit{Participant} and \textit{Item} (the $16$ sentence pairs are one \textit{Item} each).%
+Furthermore, the position (1--98) of the sentence in the individual experiment (\textit{Position}) was included as a fixed effect to control for the usual increase in reading speed during an experiment run.
+Random intercepts were specified for \textit{Participant} and \textit{Item} (the 16 sentence pairs are one \textit{Item} each).%
 \footnote{I tried random slopes in order to keep the random effect structure maximal \citep{BarrEa2013}, but it was impossible to get the algorithm to converge due to the added complexity of the interaction.}
@@
 
-Table~\ref{tab:exp:spr} shows the coefficient estimates with a 95\% parametric bootstrap confidence interval ($1,000$ replications, percentile method).
+Table~\ref{tab:exp:spr} shows the coefficient estimates with a 95\% parametric bootstrap confidence interval (1,000 replications, percentile method).
 The standard deviation of the participant intercepts is $\sigma_{\text{Participant}}=0.079$ and of the item intercepts $\sigma_{\text{Item}}=0.037$.
@@
 This fits into the general picture inasmuch as the construction with the lower frequency might be developing towards a more sharply defined prototype.%
-\footnote{In this context, it should be remembered from Section~\ref{sec:gettingdata} that even the \PGCd\ is much rarer that the \NACb\ ($17,252$ vs.\ $315,635$ occurrences in the auxiliary corpus samples).}
+\footnote{In this context, it should be remembered from Section~\ref{sec:gettingdata} that even the \PGCd\ is much rarer that the \NACb\ (17,252 vs.\ 315,635 occurrences in the auxiliary corpus samples).}
 Conversely, the \NACa\ (like the NAC in general) might be the highly frequent default which does not incur a reading time penalty, even if it is not the optimal choice in the given context and with the given lexical material.

### 2017-04-16 Latex/article.tex
Another release candidate.

@@
 \footnote{The random slopes were added to comply with \citet[257]{BarrEa2013} who predict \textit{catastrophically high Type I error rates} for experimental designs with within-subject manipulations if random effects structures are not kept maximal.
-Notice that the model reported here was estimated with conceptually identical results with regard to the predictor of interest (\textit{Modelprediction}) if only random intercepts were used.}
+Notice that the model reported here was estimated with conceptually identical results \wrt the predictor of interest (\textit{Modelprediction}) if only random intercepts were used.}
 Coefficients were estimated with Maximum Likelihood Estimation (\textit{lmer} function from \textit{lme4}).

### 2017-04-17 Latex/article.tex
Draft watermark.
Susanne's comments.

--- 2017-04-16 Latex/article.tex
@@
 
-\newcommand{\wrt}{w.\,r.\,t.\ }
-\newcommand{\Wrt}{W.\,r.\,t.\ }
+% \usepackage[firstpage]{draftwatermark}
+% \SetWatermarkFontSize{40pt}
+% \SetWatermarkHorCenter{300pt}
+% \SetWatermarkColor[rgb]{1,0.2,0.2}
+% \SetWatermarkText{\textsf{Submitted April 2017.}}
+
 \newcommand{\eg}{e.\,g.,}
@@
     The analysis is formulated in terms of Prototype Theory. 
-    I assume that the variants are modelled after non-alternating neighbouring constructions, which represent the prototypes more purely.
-    Primarily, it is argued that the frequencies with which individual lemmas occur in the prototypical constructions partially predict which variant is chosen in the alternating case.
+    I assume that the alternants are modelled after non-alternating neighbouring constructions, which represent the prototypes more purely.
+    Primarily, it is argued that the frequencies with which individual lemmas occur in the prototypical constructions partially predict which alternant is chosen in the alternating case.
     Another major prediction is that the genitive construction is prototypical of more strongly grammaticalised measure nouns.
@@
   %\accepted{...}
-  \journalname{(submitted)}
-  \journalyear{2017}
+  \journalname{Submitted April 2017. Do not quote.}
+  \journalyear{}
   %\journalvolume{...}
@@
 The study of alternations has a long history in cognitively oriented corpus linguistics (\citealp{BresnanEa2007,BresnanHay2010,BresnanFord2010,DivjakArppe2013,Gries2015,NessetJanda2010}, to name just a few publications).
-This area of research is based on the assumption that language is a probabilistic phenomenon \citep{Bresnan2007} where choices of variants are chosen neither deterministically nor fully at random.
+This area of research is based on the assumption that language is a probabilistic phenomenon \citep{Bresnan2007} where alternants are chosen neither deterministically nor fully at random.
 Instead, multifactorial models are constructed which incorporate influencing factors from diverse levels, including contextual factors.
-The estimation of the model coefficients quantifies the influence that the factors have on the probability that either variant is chosen.
+The estimation of the model coefficients quantifies the influence that the factors have on the probability that either alternant is chosen.
 
@@
 However, in terms of the scientific method, we are most likely talking about \textit{induction} rather than \textit{deduction}.}
-It would be highly surprising if this were possible.
+It would be highly surprising if this were possible, and it is the same for experimental methods.
 Nobody assumes that we can inductively infer mental representations from experiments, which -- as opposed to corpus studies -- even allow for direct access to the cognitive agent and offer much better possibilities to control experimental conditions and nuisance variables.
-Under the standard approach, we pre-specify a theory of cognitive representation.
+Rather, under the standard approach, a theory of cognitive representation is pre-specified.
 Then, predictions are derived from this theory \textit{before} the experiment or the corpus study is conducted to \textit{test} the theory.
@@
 As a theoretical framework, I assume a model of similarity-based classification in the form of Prototype Theory.
-Protoype Theory has been used in alternation research in cognitively oriented corpus linguistics, see \cite{DivjakArppe2013,Gries2003}.
-It is assumed that the variables annotated in a corpus study on an alternation phenomenon define prototypes for the alternating variants.
-The similarity of a given feature vector (in a concrete sentence where one of the variants is used) determines to a large extent which variant is chosen by speakers.
-This entails that a variant of Prototype Theory with \textit{features} is assumed \citep{Rosch1978}, instead of the monolithic prototypes of earlier versions.
-While Prototype Theory is well suited for modeling constructional choices, it is just one of several similarity-based theories of classification, the most prominent other framework being Exemplar Theory (\citealp{MedinSchaffer1978,Hintzman1986}; see \citealp{StormsEa2000} for a comparison of the theories in experimental settings).
-Prototype Theory and Exemplar Theory differ mainly in whether they assume higher-level abstractions as part of the cognitive representation (Prototype Theory) or not (Exemplar Theory).
+Protoype Theory has been used in alternation research in cognitively oriented corpus linguistics (see \citealp{DivjakArppe2013}, \citealp{Gries2003}).
+It is assumed that the variables annotated in a corpus study on an alternation phenomenon define prototypes for the alternants.
+The similarity of a given feature vector (in a concrete sentence where one of the alternants is used) influences to a large extent which alternant is chosen by speakers.
+This entails that a variant of Prototype Theory with features is assumed \citep{Rosch1978}, instead of the monolithic prototypes of earlier versions.
+Prototype Theory is well suited for modeling constructional choices, but it is just one of several similarity-based theories of classification, the most prominent other framework being Exemplar Theory (\citealp{MedinSchaffer1978,Hintzman1986}; see \citealp{StormsEa2000} for a comparison of the theories in experimental settings).
+Prototype Theory and Exemplar Theory model essentially the same types off effects and differ mainly in whether they assume higher-level abstractions as part of the cognitive representation (Prototype Theory) or try to do without them (Exemplar Theory).
 I am sceptical that corpus analysis alone could ever decide which theory is more adequate, given that substantial doubt has been voiced whether even experimental methods are ultimately able to do so \citep{Barsalou1990}.%
@@
 
-In the case at hand, three different syntactic configurations need to be distinguished \wrt case assignment inside German MNPs.
+In the case at hand, three different syntactic configurations need to be distinguished with regard to case assignment inside German MNPs.
 If the kind noun forms an NP with a determiner, the construction resembles (and is usually called) a \textit{pseudo-partitive} (on partitives and pseudo-partitives see, e.\,g., \citealp{Barker1998,Selkirk1977,Stickney2007,Vos1999}; for a recent application of the terminology to German, see \citealp{Gerstenberger2015}).%
@@
 %  \begin{xlist}
-    \ex[ ]{\label{ex:intro:pseudopartitive1}\gll Wir trinken [[eine Glas]\Sub{Acc} [dieses Weins]\Sub{Gen}]\Sub{Acc}.\\
+    \ex[ ]{\label{ex:intro:pseudopartitive1}\gll Wir trinken [[ein Glas]\Sub{Acc} [dieses Weins]\Sub{Gen}]\Sub{Acc}.\\
     we drink a glass this wine\\
@@
 I call the narrow apposition construction with a bare kind noun the \NACb, the partitive genitive with a determiner in the kind noun phrase the \PGCd.
-I call the alternating variants with an adjective but no determiner in the kind noun phrase \NACa, and \PGCa, respectively.
-This paper is about the middle column of Table~\ref{tab:constructions}, \ie\ the syntactic configuration in which two different case patterns are acceptable.
+I call the alternants with an adjective but no determiner in the kind noun phrase \NACa, and \PGCa, respectively.
+In principle, this paper is about the middle column of Table~\ref{tab:constructions}, \ie\ the syntactic configuration in which two different case patterns are acceptable.
+However, in Section~\ref{sec:analyses}, the outer columns (\NACb\ and \PGCd) will still play a major role when the factors controlling the alternation are named.
 
@@
 First of all, (\ref{it:intro:idiot}) refers to claims found in some grammars that a generic nominative, accusative, and even dative on the kind noun are used instead of the genitive (\PGCa) or case agreement (\NACa).
-Overviews can be found in \citealp{Hentschel1993} and \citealp{Zimmer2015}.
+Overviews can be found in \cite{Hentschel1993} and \cite{Zimmer2015}.
 Sentence (\ref{ex:intro:gennom}) shows a putative generic nominative on the kind noun inside an accusative MNP.
@@
 
-It was shown empirically in \cite{Hentschel1993} that such variants are de facto not acceptable.
+It was shown empirically in \cite{Hentschel1993} that such sentences are de facto not acceptable.
 Also, in my corpus sample, they simply did not occur.
@@
 %\footnote{With neglectable frequency, this effect is extended to container nouns like \textit{Glas} `glass', but with a twist.
-%A non-inflected variant \textit{zwei Glas roter Wein} `two glasses of wine' and an inflected variant \textit{zwei Gläser roter Wein} co-exist.
-%The non-inflected variant denotes the quantity of wine fitting into two glasses, the inflected variant denotes two glasses filled with wine.}
+%A non-inflected version \textit{zwei Glas roter Wein} `two glasses of wine' and an inflected version \textit{zwei Gläser roter Wein} co-exist.
+%The non-inflected version denotes the quantity of wine fitting into two glasses, the inflected version denotes two glasses filled with wine.}
 I treat these cases like other measure nouns because they enter into both the \NACa\ as in (\ref{ex:uninflectedmeasures:a}) and the \PGCa\ as in (\ref{ex:uninflectedmeasures:b}).
@@
 
-These construction have very idiosyncratic properties (the construction with \textit{voller} is discussed in \citealp{Zeldes2018}.
-The construction with \textit{mit} is discussed in \citealp{Bhatt1990}), they are not semantically equivalent to the NAC and the PGC, and they are only available for a small subset of measure nouns.
+These construction have very idiosyncratic properties.
+The construction with \textit{voller} is discussed in \cite{Zeldes2018} and the construction with \textit{mit} is discussed in \cite{Bhatt1990}.
+They are not semantically equivalent to the NAC and the PGC, and they are only available for a small subset of measure nouns.
 All of these constructions are used only with measure nouns denoting containers, and the whole NP always refers to the container, and never to the quantity contained in it.
@@
 This disqualifies them as alternatives to the \NACa\ or \PGCa\ in most contexts, and they are thus decidedly \textit{not} more ways of saying the same thing. 
-It is therefore reasonable to focus on the two well-defined variants.
+It is therefore reasonable to focus on the two well-defined alternants.
 
@@
 I discuss four main aspects:
-first, the different syntactic structures of the variants \NACa\ and \PGCa\ in relation to their syntactic prototypes \NACb\ and \PGCd;
+first, the different syntactic structures of the alternants \NACa\ and \PGCa\ in relation to their syntactic prototypes \NACb\ and \PGCd;
 second, degrees of grammaticalisation of the measure noun associated with the two prototypes;
@@
 
-My analysis is based on the idea that in the relevant syntactic structure [N\Sub{1}~[A~N\Sub{2}]\Sub{NP\Sub{2}}]\Sub{NP\Sub{1}} (as instantiated in the \NACa\ and the \PGCa), the adjective is morpho-syntactically ambiguous between a determiner and a modifying adjective.
+First, my analysis is based on the idea that in the relevant syntactic structure [N\Sub{1}~[A~N\Sub{2}]\Sub{NP\Sub{2}}]\Sub{NP\Sub{1}} (as instantiated in the \NACa\ and the \PGCa), the adjective is morpho-syntactically ambiguous between a determiner and a modifying adjective.
 To show this, %the first question is which noun constitutes the head of the whole construction.
@@
 This is called the \textit{strong} inflectional pattern.
+For example, in the dative (governed here by the preposition \textit{mit}), the strong suffix \textit{-em} is on the determiner in (\ref{ex:starkschwach:a}) and the adjective bears the weak suffix \textit{-en}.
+In (\ref{ex:starkschwach:b}), the strong suffix \textit{-em} appears on the adjective because there is no determiner.
+
+\begin{exe}
+  \ex\label{ex:starkschwach} 
+  \begin{xlist}
+    \ex{\label{ex:starkschwach:a}\gll mit ein\textbf{-em} stark\textbf{-en} Kaffe\\
+                                      with a strong coffee\\}
+    \ex{\label{ex:starkschwach:b}\gll mit stark\textbf{-em} Kaffe\\
+                                      with strong coffee\\}
+  \end{xlist}
+\end{exe}
+
 Thus, the adjectives in the \NACa\ and the \PGCa\ have properties of adjectives as well as determiners.
@@
 This should ideally be reflected in the following basic selection effect:
-It is known from frameworks like Collexeme Analysis \citep{GriesStefanowitsch2004} that lemmas are attracted with different strengths by competing constructions.
+It is known from frameworks like Collostructional Analysis \citep{GriesStefanowitsch2004} that lemmas are attracted with different strengths by constructions.
 If the \NACa\ is highly similar to the \NACb\ and the \PGCa\ is highly similar to the \PGCd, the probability with which individual noun lemmas appear in the alternating constructions should be predictable at least partly from the relative frequency with which the same lemmas are used in the non-alternating constructions.
 Nouns which occur proportionally more often in the \PGCd\ should favour the \PGCa, and nouns which occur proportionally more often in the \NACb\ should favour the \NACa.
-This basic attraction effect will be quantified in the corpus study Section~\ref{sec:corpusstudies}.
+This basic attraction effect will be used in the corpus study Section~\ref{sec:corpusstudies}.
 
@@
 
-In a probabilistic framework like Prototype Theory, the crucial question (beyond simple lemma preference effects) is, however, what controls speakers' decisions to use either variant. 
+In a probabilistic framework like Prototype Theory, the crucial question (beyond simple lemma preference effects) is, however, what controls speakers' decisions to use either alternant. 
 In the remainder of this section, I argue that the NAC and the PGC are prototypes associated with different degrees of the grammaticalisation of the measure noun and related morpho-syntactic properties, as well as register effects.
@@
 
-I first focus on grammaticalisation.
+Second, I now focus on grammaticalisation.
 It is often assumed that pseudo-partitives arise as a form of grammaticalised partitives (\eg\ \citealp[536--539]{Koptjevskaja2001} for Finnish and Estonian, \citealp[559]{Koptjevskaja2001} for European languages in general).
@@
 %
-The grammaticalisation paths uncovered by \citet[esp.\ 526--530]{Koptjevskaja2001} are relevant in the case at hand.
+The grammaticalisation paths uncovered by \citet[esp.\ 526--530]{Koptjevskaja2001} are relevant for the case at hand.
 The grammaticalisation path can start out (in some languages) with constructions involving two referential nouns (not necessarily forming a single and contiguous NP) and a \textit{separative} meaning as in \textit{(cut) two slices from the cake} \citep[535]{Koptjevskaja2001}.
@@
 The pseudo-partitive stage finally instantiates a \textit{quantity-of} relation, potentially even leading to fully grammaticalised quantifiers such as \textit{a lot}.
-In German, the PGC is clearly the older variant \citep{Zimmer2015}.
+In German, the PGC is clearly the older construction \citep{Zimmer2015}.
 It still has the potential to form a true partitive (if the kind noun is definite).
-Conversely, the NAC completely lacks this ability to form true partitives, and it thus forms the more prototypical environment for expressing measurement (as opposed to partitivity).
+Conversely, the NAC lacks this ability to form true partitives, and it thus forms the more prototypical environment for expressing measurement (as opposed to partitivity).
 Hence, we can expect the \NACa\ to be a prototypical hosting construction for more strongly grammaticalised measure nouns.
@@
 
-As described above, the grammaticalisation path leads from NPs denoting individuated objects standing in a \textit{part-of} relation to a construction with a more diffuse \textit{quantity-of} relation.
+Third, as described above, the grammaticalisation path leads from NPs denoting individuated objects standing in a \textit{part-of} relation to a construction with a more diffuse \textit{quantity-of} relation.
 Both types of relations can be numerically quantified -- inasmuch as a precise number of \textit{parts} or a numerically exact \textit{quantity} can be specified.
 However, it is much more prototypical of quantities to be specified with numerical precision.
+This is most obvious with the highly grammaticalised physical measure nouns like \textit{centilitre} which are very prototypically used with exact numeral instead of unspecific quantifiers although both options are available (\textit{three centilitres} vs.\ \textit{several centilitres}).
 Since the \NACa\ is more closely associated with the \textit{quantity-of} relation, cardinals as attributes of the measure noun are expected to have a higher proportional frequency in the \NACa.
-For illustration, (\ref{ex:determiners}) shows the expected variants under this hypothesis.
+For illustration, (\ref{ex:determiners}) shows the expected alternants under this hypothesis.
 
@@
 
-Finally, it was found that the \PGCa\ is more typical of higher registers or even exclusive to written language (see \citealp[320--323]{Hentschel1993}).
+Fourth and finally, it was found that the \PGCa\ is more typical of higher registers or even exclusive to written language (see \citealp[320--323]{Hentschel1993}).
 This is not surprising inasmuch as the genitive -- an intrinsic part of the PGC -- is generally underrepresented in colloquial vernacular variants of German as a result of a diachronic process wherein many (but by no means all) uses of the genitive are replaced by other cases or periphrastic constructions \citep{FleischerSchallert2011}.
@@
   \item Measure nouns modified by cardinals favour the \NACa.
-  \item The \NACa\ is associated with higher registers.
+  \item The \PGCa\ is associated with higher registers.
 \end{enumerate}
@@
 Second, it was already mentioned that normative grammars often adopt clear positions regarding the grammaticality of either the \NACa\ or the \PGCa.
-Thus, newspaper text or any other text that conforms strongly to normative grammars might not represent the alternation phenomenon fully (and without bias) because authors and proof-readers might favour one alternative or the other.
+Thus, newspaper text or any other text that conforms strongly to normative grammars might not represent the alternation phenomenon fully (and without bias) because authors and proof-readers might favour one alternative or the other explicitly because the have to follow normative guidelines.
 Web corpora, on the other hand, contain at least some amount of non-standard language from forums and similar sources.
@@
 Therefore, it was highly desirable to obtain a sample in which most of the highly frequent actually-occurring combinations of kind nouns and measure nouns were represented.
-I applied a three-stage bootstrap process in order to obtain such a sample.
+I applied a three-stage process in order to obtain such a sample.
 It consisted of three steps:
@@
 \begin{enumerate}[i.]
-  \item\label{enum:data:step1} bootstrapping a list of the one hundred most frequent mass nouns,
-  \item\label{enum:data:step2} bootstrapping a list of all measure nouns with which the mass nouns co-occur in the \NACb 
+  \item\label{enum:data:step1} generating a list of the one hundred most frequent mass nouns,
+  \item\label{enum:data:step2} deriving a list of all measure nouns with which the mass nouns co-occur in the \NACb 
   \item\label{enum:data:step3} sampling the target constructions by querying each combination of mass noun and measure noun found in step (\ref{enum:data:step2}).
@@
 
-This resulting list of mass nouns was used in step (\ref{enum:data:step2}) to bootstrap a list of measure nouns co-occurring with the mass nouns. 
+This resulting list of mass nouns was used in step (\ref{enum:data:step2}) to derive a list of measure nouns co-occurring with the mass nouns. 
 In order to generate this list, I utilised the fact that a direct sequence of two nouns almost always instantiates the bare-noun NAC if the second noun is a mass noun.
@@
 The final sample contained 6,843 sentences, which was reduced to 5,063 in the manual annotation process due to removal of noisy material, erroneous hits and uninformative cases where the measure noun was in the genitive, in which case the \NACa\ cannot be distinguished from the \PGCa.
-Given the careful bootstrapping and sampling procedure described in this section, we can be highly sure that it contains all relevant and reasonably frequent noun–noun combinations in the target constructions.%
-\footnote{In a similar fashion, the 100 most frequent measure nouns occurring with plural kind nouns were bootstrapped and queried, resulting in a sample of 871 sentences.
+Given the careful sampling procedure described in this section, we can be highly sure that it contains all relevant and reasonably frequent noun–noun combinations in the target constructions.%
+\footnote{In a similar fashion, the 100 most frequent measure nouns occurring with plural kind nouns were listed and queried, resulting in a sample of 871 sentences.
 As stated in Section~\ref{sec:germanmeasurenps}, the \NACa\ is virtually never used with plural kind nouns, and this sample was not used except for quantifying the frequency of occurrence of the constructions (67 times \NACa\ and 794 times \PGCa).
@@
 As mentioned in Section~\ref{sec:analyses}, the distribution of the measure noun and kind noun lemmas in the \NACb\ and the \PGCd\ with a determiner will be modelled as factors influencing the alternation.
-Therefore, all noun-noun pairs from the bootstrap process were also queried in the two non-alternating constructions, resulting in 17,252 hits for the \PGCd\ and 315,635 hits for the \NACb.
+Therefore, all noun-noun pairs from the process described above were also queried in the two non-alternating constructions, resulting in 17,252 hits for the \PGCd\ and 315,635 hits for the \NACb.
 
@@
 \footnote{
-  It could be argued that some more advanced measure of attraction strength should be used, as is done in Collostructional (or Collexeme) Analysis \citep{GriesStefanowitsch2004}, see also \cite{Gries2015a}.
-  Three main points speak against such an approach in the present case (but see \citealp[246--249]{Levshina2016} for a different approach).
-  First, the attraction values will be used as regressors in a hierarchical logistic regression, and the values resulting from collostructional approaches, \ie\ logarithmised Fisher p values, have a very unfavourable distribution in the case at hand.
+  It could be argued that some more advanced measure of attraction strength should be used, as is done in Collostructional Analysis \citep{StefanowitschGries2003,GriesStefanowitsch2004}, see also \cite{Gries2015a}.
+  Three main points speak against such an approach in the present case (but see \citealp[246--249]{Levshina2016} for a more comprehensive approach albeit for a different purpose).
+  First, the goal here is to quantify how often lemmas occur in the \PGCd\ and the \NACb, and these construction do not compete at all but are rather mutually exclusive.
+  Collostructional approaches are not made such scenarios.
+  Second, the attraction values will be used as regressors in a hierarchical logistic regression, and the values resulting from collostructional analysis, \ie\ logarithmised Fisher p values, have a very unfavourable distribution in the case at hand.
   They cluster around 0, and they include values of $-\infty$.
-  Second, their main use in collostructional analyses is to \textit{sort} a list of collexemes and interpret their order.
-  Their concrete numerical value might not have a solid cognitive interpretation in the given context.
-  Third, I tried using collexeme strength as a regressor, and the results were unsatisfactory.
+  Third, I tried using collexeme strength as a regressor, and the results were unsatisfactory compared to the simple quotient used here.
 }
@@
 
-In Section~\ref{sec:analyses}, it was hypothesised that classes of measure lemmas might have different preferences for the two variants.
+In Section~\ref{sec:analyses}, it was hypothesised that classes of measure lemmas might have different preferences for the two alternants.
 To capture this, class information was annotated for measure lemmas.
@@
 The higher the values of this variable, the lower the proportion of genitives among all case-bearing forms is.
-While more genitives are also indicative of higher registers, the use of this variable as a regressor in the present study might be considered problematic.
+A high number of genitives is indicative of higher registers.
+However, the use of this variable as a regressor in the present study might be considered problematic.
 Since the \PGCa\ contains a genitive itself, the regressor variable \textit{Genitive} and the document-level variable \textit{Genitives} are not fully independent.
@@
 
-As for the first point, this is not quite the place to discuss it fully.
-The basic distinction is a philosophical one and related to the concepts of \textit{direct} and \textit{inverse probability} (\eg\ \citealp{Senn2011}).
-Frequentists assume that models and parameters are fixed, for example a model specifying that a coin is fair.
-We can then calculate for observed data (for example a measurement of 3 heads in an experiment with 10 tosses) how often such a result or a more extreme result would occur if the model were true and we repeated the experiment arbitrarily often.
-This is essentially the frequentist notion of direct probability, \ie\ long-run frequencies under replication.
-Standard tests in the Fisher and Neyman-Pearson traditions as well as Neyman confidence intervals are based on this concept of probability.
-Bayesian approaches (in the now common interpretation), on the other hand, condition on the particular data and quantify inductively the probability of model parameters given the data.
-The parameters are thus not fixed, and the probability is usually equated with researchers' posterior beliefs about model parameters.
-There is actually a debate among Bayesians about the proper interpretation of Bayesian methods, and whether a notion of hypothesis testing is compatible (or even already contained) in the Bayesian approach.
-In \citet[10]{GelmanShalizi2013}, the authors -- prominent Bayesians themselves -- acknowledge that a theory of statistical testing is a desideratum, and they state about the standard inductive interpretation of Bayesianism that ``most of this received view of Bayesian inference is wrong'', and they develop a Bayesian notion of p values (see also \citealp{Mayo2013}, for a frequentist reply; also \citealp{Senn2011} on different strands of Bayesianism and their stance on inductive vs.\ deductive reasoning, and \citealp{Mayo2011}, for a critical reply to \citealp{Senn2011}).
-Clearly, in such quarrels between and among camps of philosophers of science and statisticians, it is difficult for mere practitioners to take sides.
-
-These quarrels relate to the second point, however.
-\citet[301--302]{Divjak2016a} speaks favourably of Bayesian methods because the Bayesian concept of probability is more adequate for cognitive modeling compared to the frequentist one.
-Her argument is part of a larger body of literature asking for cognitively plausible modeling techniques, for example Naive Discriminative Learning (NDL; \citealp{Baayen2011,BaayenEa2013,MilinEa2016,TheijssenEa2013}).
-On p.\ 303 of \cite{Divjak2016a}, the author goes on to explicitly mention NDL as well.
-Yet, neither frequentist nor Bayesian methods were conceived as cognitive models, but as systems of inference for scientists (see above, and see also \citealp[302]{Divjak2016a}).
-The fundamental question that lurks behind such arguments is how we interpret our statistical models (estimated on corpus data).
-Are they inductive models of cognitive representations that also human learners would infer from being exposed to the corpus data?%
-\footnote{In which case we are doing ``data science in language research'' in the words of \citealp{MilinEa2016}.
-I see this as standing in contradiction to the view advocated in \cite{Dabrowska2016} as cited above.}
-Or are they tests of theories that are pre-specified and merely tested for predictive accuracy on linguistic output data contained in corpora?
-In the former case, we adopt a strong \textit{corpus as input} hypothesis and should definitely resort to methods like NDL.
-However, this would most likely require us to toss most previous work done in (cognitive) linguistics into the bin and to abandon all high-level generalisations that most existing studies have been based upon.
-In fact, arguments to this extreme effect and against using high-level generalisation have been made, for example in \cite{BaayenEa2016}, \citet[299--300]{Divjak2016a}, \cite{RamscarPort2016}, \cite{TheijssenEa2013}.
-In the latter and less extreme case, the cognitive commitment does, however, not necessarily extend to the statistical methods used.
-These methods then do not need to be any more cognitively plausible than an ANOVA used to analyse the results from an experiment.
-I view my own work (see Sections~\ref{sec:germanmeasurenps}--\ref{sec:experimental}) in the tradition of testing theories (which embrace high-level generalisations), and I agree to provisionally use Likelihood methods (see below).
-I remain fully agnostic with respect to the question, which approach is \textit{right}.
-The best strategy for cognitive linguistics as a field might be to cultivate many methods while making sure that each method is applied carefully and competently.
-
-%The first and second point cannot be discussed in detail here.%
-%\footnote{However, among the Bayesian hype, there is also a number of critical papers by statisticians (even prominent Bayesians) and philosophers of science in which the extreme credibility that has been assigned to Bayesian approaches (and especially the standard inductivist interpretation of Bayesianism) is questioned (for example, \citealp{GelmanShalizi2013,Mayo2011,Senn2011}).}
+% As for the first point, this is not quite the place to discuss it fully.
+% The basic distinction is a philosophical one and related to the concepts of \textit{direct} and \textit{inverse probability} (\eg\ \citealp{Senn2011}).
+% Frequentists assume that models and parameters are fixed, for example a model specifying that a coin is fair.
+% We can then calculate for observed data (for example a measurement of 3 heads in an experiment with 10 tosses) how often such a result or a more extreme result would occur if the model were true and we repeated the experiment arbitrarily often.
+% This is essentially the frequentist notion of direct probability, \ie\ long-run frequencies under replication.
+% Standard tests in the Fisher and Neyman-Pearson traditions as well as Neyman confidence intervals are based on this concept of probability.
+% Bayesian approaches (in the now common interpretation), on the other hand, condition on the particular data and quantify inductively the probability of model parameters given the data.
+% The parameters are thus not fixed, and the probability is usually equated with researchers' posterior beliefs about model parameters.
+% There is actually a debate among Bayesians about the proper interpretation of Bayesian methods, and whether a notion of hypothesis testing is compatible (or even already contained) in the Bayesian approach.
+% In \citet[10]{GelmanShalizi2013}, the authors -- prominent Bayesians themselves -- acknowledge that a theory of statistical testing is a desideratum, and they state about the standard inductive interpretation of Bayesianism that ``most of this received view of Bayesian inference is wrong'', and they develop a Bayesian notion of p values (see also \citealp{Mayo2013}, for a frequentist reply; also \citealp{Senn2011} on different strands of Bayesianism and their stance on inductive vs.\ deductive reasoning, and \citealp{Mayo2011}, for a critical reply to \citealp{Senn2011}).
+% Clearly, in such quarrels between and among camps of philosophers of science and statisticians, it is difficult for mere practitioners to take sides.
+% 
+% These quarrels relate to the second point, however.
+% \citet[301--302]{Divjak2016a} speaks favourably of Bayesian methods because the Bayesian concept of probability is more adequate for cognitive modeling compared to the frequentist one.
+% Her argument is part of a larger body of literature asking for cognitively plausible modeling techniques, for example Naive Discriminative Learning (NDL; \citealp{Baayen2011,BaayenEa2013,MilinEa2016,TheijssenEa2013}).
+% On p.\ 303 of \cite{Divjak2016a}, the author goes on to explicitly mention NDL as well.
+% Yet, neither frequentist nor Bayesian methods were conceived as cognitive models, but as systems of inference for scientists (see above, and see also \citealp[302]{Divjak2016a}).
+% The fundamental question that lurks behind such arguments is how we interpret our statistical models (estimated on corpus data).
+% Are they inductive models of cognitive representations that also human learners would infer from being exposed to the corpus data?%
+% \footnote{In which case we are doing ``data science in language research'' in the words of \citealp{MilinEa2016}.
+% I see this as standing in contradiction to the view advocated in \cite{Dabrowska2016} as cited above.}
+% Or are they tests of theories that are pre-specified and merely tested for predictive accuracy on linguistic output data contained in corpora?
+% In the former case, we adopt a strong \textit{corpus as input} hypothesis and should definitely resort to methods like NDL.
+% However, this would most likely require us to toss most previous work done in (cognitive) linguistics into the bin and to abandon all high-level generalisations that most existing studies have been based upon.
+% In fact, arguments to this extreme effect and against using high-level generalisation have been made, for example in \cite{BaayenEa2016}, \citet[299--300]{Divjak2016a}, \cite{RamscarPort2016}, \cite{TheijssenEa2013}.
+% In the latter and less extreme case, the cognitive commitment does, however, not necessarily extend to the statistical methods used.
+% These methods then do not need to be any more cognitively plausible than an ANOVA used to analyse the results from an experiment.
+% I view my own work (see Sections~\ref{sec:germanmeasurenps}--\ref{sec:experimental}) in the tradition of testing theories (which embrace high-level generalisations), and I agree to provisionally use Likelihood methods (see below).
+% I remain fully agnostic with respect to the question, which approach is \textit{right}.
+% The best strategy for cognitive linguistics as a field might be to cultivate many methods while making sure that each method is applied carefully and competently.
+
+The first and second point cannot be discussed in detail here.%
+\footnote{However, among the Bayesian hype, there is also a number of critical papers by statisticians (even prominent Bayesians) and philosophers of science in which the extreme credibility that has been assigned to Bayesian approaches (and especially the standard inductivist interpretation of Bayesianism) is questioned (for example, \citealp{GelmanShalizi2013,Mayo2011,Senn2011}).}
 The third point, then, is the most practically relevant in the context of this paper.
@@
 This is a clear sign that the prior is ``swamped by the data'' \citep[1119]{Freedman1999}.
-Going back to the argument that Bayesian methods are allegedly more cognitively plausible than `frequentist' methods, the claim might still be true.
-However, most similar studies (dating back to \citealp{BresnanEa2007}) use an estimator (MLE) that is not frequentist in a narrow sense, and which is as cognitively plausible as Bayesian estimators because it most likely produces the same results in these studies.
-If for Levshina's study, results from Bayesian and MLE methods did, in fact, \textit{not} converge, it would have been an ideal occasion to demonstrate the superiority of the algorithms used in Bayesian estimation.
+% Going back to the argument that Bayesian methods are allegedly more cognitively plausible than `frequentist' methods, the claim might still be true.
+H% owever, most similar studies (dating back to \citealp{BresnanEa2007}) use an estimator (MLE) that is not frequentist in a narrow sense, and which is as cognitively plausible as Bayesian estimators because it most likely produces the same results in these studies.
+
+If in Levshina's study, results from Bayesian and MLE methods did, in fact, \textit{not} converge, it would have been an ideal occasion to demonstrate the superiority of the algorithms used in Bayesian estimation.
 After all, there are situations where Bayesian estimators can be more robust, namely with heavily censored data, complex hierarchical models, perfect separation, etc.
@@
 I now turn to the predicted effect of cardinals as modifiers of the measure noun.
-Figure~\ref{fig:eff:leftcontext} shows that cardinals indeed influence the choice of the variant ($\mpPB=0.001$), and that cardinals have a strong tendency to co-occur with the \NACa.
+Figure~\ref{fig:eff:leftcontext} shows that cardinals indeed influence the choice of the alternant ($\mpPB=0.001$), and that cardinals have a strong tendency to co-occur with the \NACa.
 This effect was predicted in Section~\ref{sec:analyses}.
@@
 Both experiments use sentences containing attested MNPs from the corpus sample (embedded into simplified sentences) as stimuli.
-Also, the probabilities that the corpus-based model assigns to the two variants in these sentences is used as the main regressor in both studies.
+Also, the probabilities that the corpus-based model assigns to the two alternants in these sentences is used as the main regressor in both studies.
 
@@
 \cite{FordBresnan2013} use the \textit{split-100} task in which participants have to distribute 100 points between the alternatives, assigning more points to more natural sounding alternative.
-In essence, participants distribute a probability mass between two variants, which is intended to produce more subtle results compared to a two-alternative forced-choice task \citep{Rosenbach2013}, where participants have to choose one of two variants.
+In essence, participants distribute a probability mass between two alternants, which is intended to produce more subtle results compared to a two-alternative forced-choice task \citep{Rosenbach2013}, where participants have to choose one of two alternants.
 The split-100 paradigm has been criticised in \cite{ArppeJaervikivi2007}.
@@
 In \cite{VerhoevenTemme2017}, it was shown that results from forced-choice and split-100 experiments mostly converge, which is expected in the long run if a probability is turned into a binary decision.
-I present a forced-choice experiment in this Section and not a split-100 experiment, mainly because in a dry-run of the experiment, participants complained about the unnaturalness of the task of distributing a probability mass across two variants.
+I present a forced-choice experiment in this Section and not a split-100 experiment, mainly because in a dry-run of the experiment, participants complained about the unnaturalness of the task of distributing a probability mass across two alternants.
 Participants had to chose between two sentences differing only in that one contained the \NACa\ and the other contained the \PGCa.
-The analysis compares the probabilities assigned to stimuli by the corpus-derived model with the frequency with which participants chose the variants for the same stimuli.
+The analysis compares the probabilities assigned to stimuli by the corpus-derived model with the frequency with which participants chose the alternants for the same stimuli.
 
@@
 As stimuli, attested MNPs from the corpus study were used, but the sentences were radically simplified to avoid influences from contextual nuisance factors as much as possible.
-The approach is also justified because according to the theoretical assessment in Section~\ref{sec:analyses}, the choice of variants depends mostly on a very local constructional context.
+The approach is also justified because according to the theoretical assessment in Section~\ref{sec:analyses}, the choice of alternants depends mostly on a very local constructional context.
 I sampled 16 MNPs from the corpus, and it was made sure that the simplifications and normalisations did not affect any of the regressors used in the corpus study.
@@
 The result is very clear.
-The higher the probability of the \PGCa\ predicted from usage-data, the more often participants chose the \PGCa\ variant in the forced-choice task.
+The higher the probability of the \PGCa\ predicted from usage-data, the more often participants chose the \PGCa\ alternant in the forced-choice task.
 In summary, the forced-choice experiment clearly succeeded in corroborating the results from the corpus study in as much as the preferences extracted from usage data correspond to native speakers' choices.
@@
 The second experiment tests preferences more implicitly.
-It is expected that reading less prototypical variants (the one assigned a low probability by the corpus-derived model) in a given context and with given lexical material incurs a processing overhead for the reader (\citealp{Kaiser2013}).
+It is expected that reading less prototypical alternants (the one assigned a low probability by the corpus-derived model) in a given context and with given lexical material incurs a processing overhead for the reader (\citealp{Kaiser2013}).
 In this Section, a self-paced reading experiment is therefore presented.
@@
 Concretely, the exact same stimuli as in the forced-choice experiments were used.
-Each participant read both the 16 sentences with the variant predicted by the corpus model and the 16 modified sentences with the variant that the corpus model did not predict.%
+Each participant read both the 16 sentences with the alternant predicted by the corpus model and the 16 modified sentences with the alternant that the corpus model did not predict.%
 \footnote{Notice that lemmas and their frequencies as well as lemma classes are included as regressors in the corpus-based GLMM, and there was consequently no additional controlling of lemma frequencies, etc.}
@@
 The assignment of members of the individual sentence pairs to the blocks was randomised for each participant individually and so was the order within each block.
-The sentences from each pair of variants were kept as far apart as possible.
+The sentences from each pair of alternants were kept as far apart as possible.
 The fillers also came in pairs such that the second block exclusively contained sentences to which participants had been exposed in the first block in slightly modified form.
@@
 A grammatical alternation in German measure NPs was examined using corpus data based on factors partly derived from existing accounts, formulated in terms of construction prototypes.
-The preferences extracted from the DECOW web corpus were confirmed in a forced-choice experiment, in which participants explicitly chose variants in line with the corpus-derived model.
-In a more implicit self-paced reading experiment, it was shown that the much rarer variant brings about a reading time penalty except in cases for which the corpus model predicts very high probability for this variant.
+The preferences extracted from the DECOW web corpus were confirmed in a forced-choice experiment, in which participants explicitly chose alternants in line with the corpus-derived model.
+In a more implicit self-paced reading experiment, it was shown that the much rarer alternant brings about a reading time penalty except in cases for which the corpus model predicts very high probability for this alternant.
 
@@
 
-% \begin{acknowledgement}
-%   I want to thank (in alphabetical order) Felix Bildhauer, Samuel Reichert, and Ulrike Sayatz for discussions on relevant aspects of this paper.
-%   I also thank Ulrike Sayatz for helping to recruit the participants for the experiments.
-%   I am grateful to Kim Maser for her work on the annotation of the concordances.
-%   The research presented here was made possible in part through funding from the \textit{Deutsche Forschungsgemeinschaft} (DFG, personal grant SCHA1916/1-1).
-% \end{acknowledgement}
+\begin{acknowledgement}
+  I thank (in alphabetical order) Felix Bildhauer, Susanne Flach, Elizabeth Pankratz, Samuel Reichert, and Ulrike Sayatz for valuable discussions and comments.
+  Elizabeth thankfully also fixed my English.
+  Furthermore, I thank Ulrike for helping me to recruit the participants for the experiments.
+  Finally, I am grateful to Kim Maser for her work on the annotation of the concordances.
+  The research presented here was made possible in part through funding from the \textit{Deutsche Forschungsgemeinschaft} (DFG, personal grant SCHA1916/1-1).
+\end{acknowledgement}
 

### 2017-04-18 Latex/article.tex
Virtually done.

@@
 
-% \usepackage[firstpage]{draftwatermark}
-% \SetWatermarkFontSize{40pt}
-% \SetWatermarkHorCenter{300pt}
-% \SetWatermarkColor[rgb]{1,0.2,0.2}
-% \SetWatermarkText{\textsf{Submitted April 2017.}}
+\usepackage[firstpage]{draftwatermark}
+\SetWatermarkFontSize{40pt}
+\SetWatermarkHorCenter{300pt}
+\SetWatermarkColor[rgb]{1,0.2,0.2}
+\SetWatermarkText{\textsf{Submitted April 2017.}}
 
@@
   \author*[1]{Roland Schäfer}
-  %\runningauthor{...}
+  \runningauthor{Roland Schäfer. Submitted April 2017. Do not quote.}
   \affil[1]{Freie Universität Berlin}
@@
   \abstract{In this paper, a case alternation occurring in German measure noun phrases is examined from the perspective of cognitively oriented corpus linguistics.
-    In pseudo-partitive constructions such as \textit{eine Tasse guter Kaffe} `a cup of good coffee', the embedded kind-denoting noun either agrees in its case with the measure head noun or it is a genitive as in \textit{eine Tasse guten Kaffees}.
-    This alternation occurs only under highly specific syntactic conditions.
+    In pseudo-partitive constructions such as \textit{eine Tasse guter Kaffee} `a cup of good coffee', the embedded kind-denoting noun either agrees in its case with the measure head noun or it stands in the genitive as in \textit{eine Tasse guten Kaffees} with no difference in meaning.
+    This alternation occurs only under specific syntactic conditions.
     The analysis is formulated in terms of Prototype Theory. 
-    I assume that the alternants are modelled after non-alternating neighbouring constructions, which represent the prototypes more purely.
+    I assume that the alternants are modelled after non-alternating neighbouring constructions, which represent the prototypes more directly.
     Primarily, it is argued that the frequencies with which individual lemmas occur in the prototypical constructions partially predict which alternant is chosen in the alternating case.
-    Another major prediction is that the genitive construction is prototypical of more strongly grammaticalised measure nouns.
-    I present a large-scale corpus study using the DECOW web corpus which supports the theoretical analysis.
-    In the statistical analysis of the corpus study, it is discussed whether Bayesian estimators lead to substantially different results compared to Likelihood Maximisation methods in studies like this one.
-    The model is estimated with Maximum Likelihood and Bayesian methods for comparison.
+    Another major assumption is that the genitive construction is prototypical of more strongly grammaticalised measure nouns.
+    I present a large-scale corpus study using the DECOW web corpus which supports the theory.
+    In the statistical analysis of the corpus study, it is discussed whether Bayesian estimators lead to substantially different results than Likelihood Maximisation methods in this and similar studies.
+    The model parameters are estimated with both methods for comparison.
     Finally, a forced choice experiment and a self-paced reading experiment are reported.
-    The usage-based findings from the corpus study show a clear correlation with the behaviour of cognitive agents in both experiments.
+    The usage-based findings from the corpus study show a correlation with the behaviour of cognitive agents in both experiments.
     The present study therefore contributes to a now well-established area of research into alternations using both corpus and experimental methods.}
@@
 For such probabilistic generalisations estimated on corpus data, there has always been an interest in correlating the findings with results from experimental work (for example, \citealp{ArppeJaervikivi2007,BresnanEa2007,BresnanFord2010,DivjakGries2008,DivjakEa2016,FordBresnan2013}).
-This often called a \textit{validation} of the corpus-derived findings, but \citet[303]{Divjak2016a} rightly criticises this choice of words ``because it creates the impression that behavioral experimental data is inherently more valuable than textual data'', citing \cite{TummersEa2005}, who state that a corpus is ``a sample of spontaneous language use that is'' (\textit{generally}) \textit{realized by native speakers}.
-This position, which I adopt, does not, however, imply that we can in some way \textit{deduce mental representations from patterns of use}, \ie\ from corpus data, as  \citet[486--487]{Dabrowska2016} convincingly argues.%
+This often called a \textit{validation} of the corpus-derived findings, but \citet[303]{Divjak2016a} rightly criticises this choice of words ``because it creates the impression that behavioral experimental data is inherently more valuable than textual data,'' citing \cite{TummersEa2005}, who state that a corpus is ``a sample of spontaneous language use that is (generally) realized by native speakers.''
+However, as \citet[486--487]{Dabrowska2016} convincingly argues, this does not imply that we can in some way ``deduce mental representations from patterns of use,'' \ie\ from corpus data.%
 \footnote{In its colloquial meaning, the verb \textit{deduce} seems appropriate here.
 However, in terms of the scientific method, we are most likely talking about \textit{induction} rather than \textit{deduction}.}
-It would be highly surprising if this were possible, and it is the same for experimental methods.
+It would be highly surprising if this were possible and the same holds for experimental methods.
 Nobody assumes that we can inductively infer mental representations from experiments, which -- as opposed to corpus studies -- even allow for direct access to the cognitive agent and offer much better possibilities to control experimental conditions and nuisance variables.
@@
 This entails that a variant of Prototype Theory with features is assumed \citep{Rosch1978}, instead of the monolithic prototypes of earlier versions.
-Prototype Theory is well suited for modeling constructional choices, but it is just one of several similarity-based theories of classification, the most prominent other framework being Exemplar Theory (\citealp{MedinSchaffer1978,Hintzman1986}; see \citealp{StormsEa2000} for a comparison of the theories in experimental settings).
+Prototype Theory is well suited for modeling constructional choices but it is just one of several similarity-based theories of classification, the most prominent other framework being Exemplar Theory (\citealp{MedinSchaffer1978,Hintzman1986}; see \citealp{StormsEa2000} for a comparison of the theories in experimental settings).
 Prototype Theory and Exemplar Theory model essentially the same types off effects and differ mainly in whether they assume higher-level abstractions as part of the cognitive representation (Prototype Theory) or try to do without them (Exemplar Theory).
-I am sceptical that corpus analysis alone could ever decide which theory is more adequate, given that substantial doubt has been voiced whether even experimental methods are ultimately able to do so \citep{Barsalou1990}.%
+I am sceptical that corpus analysis alone could ever decide which theory is more suitable, given that substantial doubt has been voiced whether even experimental methods are ultimately able to do so \citep{Barsalou1990}.%
 \footnote{In \cite{DivjakArppe2013}, it was shown using corpus data how both models form a converging picture.}
@@
 In the remainder of the paper, I introduce the alternation phenomenon (Section~\ref{sec:descriptive}) and suggest a theory-driven set of factors influencing the alternation (Section~\ref{sec:analyses}).
-Then, the corpus study is presented, including an appropriate statistical analysis (Section~\ref{sec:corpusstudies}).
+Then the corpus study is presented, including an appropriate statistical analysis (Section~\ref{sec:corpusstudies}).
 Two experiments that correlate results from the corpus study with experimental results are then reported (Section~\ref{sec:experimental}) before I sum up the paper in Section~\ref{sec:conclusion}.
@@
 I use the term \textit{measure noun phrase} (MNP) to refer a noun phrase (NP) in which a kind-denoting (count or mass) noun depends on another noun that specifies a quantity of the objects or the substance denoted by the kind-denoting noun.
-I call the kind-denoting noun simply the \textit{kind noun} and the quantity-denoting noun the \textit{measure noun}.
-For illustration purposes, in the English \textit{a glass of good wine}, \textit{glass} is the measure noun, and \textit{wine} is the kind noun.
+I call the kind-denoting noun the \textit{kind noun} and the quantity-denoting noun the \textit{measure noun}.
+For illustration purposes, in the English \textit{a glass of good wine}, \textit{glass} is the measure noun and \textit{wine} is the kind noun.
 Measure nouns can be all sorts of nouns which denote a quantity (such as \textit{litre} or \textit{amount}) but also those denoting containers, collections, etc.
 (such as \textit{glass} or \textit{bucket}).
-In a similar vein, \citet[284]{Brems2003} considers such nouns denoting containers etc.\ as measure nouns ``which, strictly speaking, do not designate a `measure', but display a more nebulous potential for quantification'' (see also \citealp[530]{Koptjevskaja2001}, and \citealp[338]{Rutkowski2007}).
+Like \citet[284]{Brems2003}, I consider nouns as measure nouns ``which, strictly speaking, do not designate a `measure', but display a more nebulous potential for quantification'' (also \citealp[530]{Koptjevskaja2001}, and \citealp[338]{Rutkowski2007}).
 
@@
 
+\enlargethispage{1\baselineskip}
+
 \begin{exe}
@@
 However, Section~\ref{sec:analyses} is devoted to developing hypotheses about subtle differences between them.%
-\footnote{Some descriptive and normative grammars take stronger positions w.\,r.\,t.\ the acceptability of the two options.
+\footnote{Some descriptive and normative grammars take stronger positions with regard to the acceptability of the two options.
 See \cite{Hentschel1993,Zimmer2015} for analyses of the sometimes absurd stances taken in grammars of German.
-As the usage and experimental data presented below (especially Section~\ref{sec:exp:fc}) should render it clear, there might be preferences under certain circumstances, but we cannot assume either construction to be unacceptable.
+As the usage and experimental data presented below (especially Section~\ref{sec:exp:fc}) should render it clear, there might be preferences, but we cannot assume either construction to be unacceptable.
 }
@@
 
-The distribution of the case patterns in the NAC (case identity between the measure noun and the kind noun) and in the PGC (genitive on the kind noun, regardless of the case of the measure noun) depending on the structure of the kind NP are summarised in Table~\ref{tab:constructions}.
-I call the narrow apposition construction with a bare kind noun the \NACb, the partitive genitive with a determiner in the kind noun phrase the \PGCd.
-I call the alternants with an adjective but no determiner in the kind noun phrase \NACa, and \PGCa, respectively.
+The distribution of the case patterns in the NAC (case identity between the measure noun and the kind noun) and in the PGC (genitive on the kind noun, regardless of the case of the measure noun) depending on the structure of the kind NP is summarised in Table~\ref{tab:constructions}.
+I call the narrow apposition construction with a bare kind noun the \NACb\ and the partitive genitive with a determiner in the kind NP the \PGCd.
+For the alternants with an adjective but no determiner in the kind noun phrase I use the terms \NACa and \PGCa.
 In principle, this paper is about the middle column of Table~\ref{tab:constructions}, \ie\ the syntactic configuration in which two different case patterns are acceptable.
@@
   \item{\label{it:intro:plurl} similar constructions with plural/collective kind nouns,}
-  \item{\label{it:intro:noifl} grammaticalised non-inflected measure nouns,}
+  \item{\label{it:intro:noifl} grammaticalised non-inflected measure nouns, and}
   \item{\label{it:intro:other} alternative constructions for expressing quantities.}
@@
 Also, in my corpus sample, they simply did not occur.
-Even if they are accepted by some speakers, their extremely low frequency makes it virtually impossible to study them using corpus linguistic methods (Section~\ref{sec:corpusstudies}), and I consequently ignore them.
+Even if they are accepted by some speakers, their extremely low frequency makes it virtually impossible to study them using corpus linguistic methods (Section~\ref{sec:corpusstudies}) and I consequently ignore them.
 
@@
 If the kind noun is a plural count noun as in \textit{ein Sack kleine Äpfel} (\NACa) or \textit{ein Sack kleiner Äpfel} (\PGCa) `a bag of small apples', a similar alternation between PGC and NAC can be observed.
-In line with experimental results reported in \citet[15--16]{Zimmer2015}, I found that the PGC is so dominant with plural kind nouns (794 of 861 cases, or over 92\%, cf.\ Section~\ref{sec:corpusstudies}) that the alternation cannot be analysed in the same way as in the singular.%
-\footnote{More concretely, if these cases were included in the regression analysis of the corpus data along with a factor encoding the number of the kind noun, this factor would most assuredly override any other regressor for the data points with a plural kind noun.}
+In line with experimental results reported in \citet[15--16]{Zimmer2015}, I found that the PGC is so dominant with plural kind nouns (794 of 861 cases, or over 92\%, cf.\ Section~\ref{sec:corpusstudies}) that the alternation cannot be analysed in the same way as in the singular.
+%\footnote{More concretely, if these cases were included in the regression analysis of the corpus data along with a factor encoding the number of the kind noun, this factor would most assuredly override any other regressor for the data points with a plural kind noun.}
 While this will play a role in the interpretation of the corpus findings, MNPs with plural kind nouns will not be included in the corpus study and the experiments reported in Sections~\ref{sec:corpusstudies} and \ref{sec:experimental}.
 
-As for (\ref{it:intro:noifl}), some measure nouns have been grammaticalised in a way that they always appear non-inflected.
+As for (\ref{it:intro:noifl}), some measure nouns have been grammaticalised in a way that they always appear in their non-inflected form.
 They are typical measure nouns like \textit{Gramm} `gram', \textit{Pfund} `pound' or \textit{Prozent} `percent', which have no normal plural forms at all.%
@@
 The construction with \textit{voller} is discussed in \cite{Zeldes2018} and the construction with \textit{mit} is discussed in \cite{Bhatt1990}.
-They are not semantically equivalent to the NAC and the PGC, and they are only available for a small subset of measure nouns.
-All of these constructions are used only with measure nouns denoting containers, and the whole NP always refers to the container, and never to the quantity contained in it.
+They are not semantically equivalent to the NAC and the PGC and they are only available for a small subset of measure nouns.
+All of these constructions are used only with measure nouns denoting containers (where the whole NP always refers to the container) and never to the quantity contained in it.
 They are consequently incompatible with measure nouns denoting natural portions such as \textit{Schluck} `gulp' or \textit{Haufen} `heap' and strongly grammaticalised nouns such as \textit{Gramm} `gram'.
-This disqualifies them as alternatives to the \NACa\ or \PGCa\ in most contexts, and they are thus decidedly \textit{not} more ways of saying the same thing. 
+This disqualifies them as alternatives to the \NACa\ or \PGCa\ in most contexts and they are thus decidedly \textit{not} more ways of saying the same thing. 
 It is therefore reasonable to focus on the two well-defined alternants.
@@
 I have demonstrated that there is an alternation between two measure noun constructions in a narrow syntactic configuration (kind NP with an adjective but without a determiner), and that the two constructions differ in the case of the kind noun (case agreement with the measure noun or genitive).
-I turn to some more theory-oriented discussion in the next section.
+I turn to more theory-oriented discussion in the next section.
 
@@
 the strong\slash weak inflection patterns of adjectives needs to be taken into account.
-In NPs with a strongly inflected determiner, attributive adjectives inflect according to the massively syncretistic \textit{weak} pattern.
+In NPs with a strongly inflected determiner, attributive adjectives inflect according to the massively syncretic \textit{weak} pattern.
 %\footnote{The weak pattern has only two suffixes.
@@
 %The only exception to this rule is the masculine accusative singular, which is also marked with \textit{-en}.}
-If there is no determiner (as is the case in the alternating constructions), on the other hand, attributive adjectives inflect like determiners themselves.
+If there is no determiner (as is the case in the alternating constructions), attributive adjectives inflect like determiners themselves.
 This is called the \textit{strong} inflectional pattern.
@@
   \begin{xlist}
-    \ex{\label{ex:starkschwach:a}\gll mit ein\textbf{-em} stark\textbf{-en} Kaffe\\
+    \ex{\label{ex:starkschwach:a}\gll mit ein\textbf{-em} stark\textbf{-en} Kaffee\\
                                       with a strong coffee\\}
-    \ex{\label{ex:starkschwach:b}\gll mit stark\textbf{-em} Kaffe\\
+    \ex{\label{ex:starkschwach:b}\gll mit stark\textbf{-em} Kaffee\\
                                       with strong coffee\\}
@@
 It is known from frameworks like Collostructional Analysis \citep{GriesStefanowitsch2004} that lemmas are attracted with different strengths by constructions.
-If the \NACa\ is highly similar to the \NACb\ and the \PGCa\ is highly similar to the \PGCd, the probability with which individual noun lemmas appear in the alternating constructions should be predictable at least partly from the relative frequency with which the same lemmas are used in the non-alternating constructions.
+If the \NACa\ is highly similar to the \NACb\ and the \PGCa\ is highly similar to the \PGCd, the probability with which individual noun lemmas appear in the alternating constructions should be predictable at least in part from the relative frequency with which the same lemmas are used in the non-alternating constructions.
 Nouns which occur proportionally more often in the \PGCd\ should favour the \PGCa, and nouns which occur proportionally more often in the \NACb\ should favour the \NACa.
@@
 
-Second, I now focus on grammaticalisation.
+Second, I focus on grammaticalisation.
 It is often assumed that pseudo-partitives arise as a form of grammaticalised partitives (\eg\ \citealp[536--539]{Koptjevskaja2001} for Finnish and Estonian, \citealp[559]{Koptjevskaja2001} for European languages in general).
@@
 The grammaticalisation path can start out (in some languages) with constructions involving two referential nouns (not necessarily forming a single and contiguous NP) and a \textit{separative} meaning as in \textit{(cut) two slices from the cake} \citep[535]{Koptjevskaja2001}.
-The \textit{part-of} meaning of true partitives as in \textit{a slice of the cake} represents the first stage of a development wherein the measure noun can already lose some semantic content, when for example words like \textit{bite} are no longer necessarily interpreted as a piece bitten out of something.
+The \textit{part-of} meaning of true partitives as in \textit{a slice of the cake} represents the first stage of a development wherein the measure noun can already lose some semantic content, when for example words like \textit{bite} are no longer necessarily interpreted as a piece literally bitten out of something.
 The pseudo-partitive stage finally instantiates a \textit{quantity-of} relation, potentially even leading to fully grammaticalised quantifiers such as \textit{a lot}.
@@
 However, it is much more prototypical of quantities to be specified with numerical precision.
-This is most obvious with the highly grammaticalised physical measure nouns like \textit{centilitre} which are very prototypically used with exact numeral instead of unspecific quantifiers although both options are available (\textit{three centilitres} vs.\ \textit{several centilitres}).
+This is most obvious with the highly grammaticalised physical measure nouns like \textit{centilitre} which are very prototypically used with exact numeral instead of unspecific quantifiers, although both options are available (\textit{three centilitres} vs.\ \textit{several centilitres}).
 Since the \NACa\ is more closely associated with the \textit{quantity-of} relation, cardinals as attributes of the measure noun are expected to have a higher proportional frequency in the \NACa.
@@
 Fourth and finally, it was found that the \PGCa\ is more typical of higher registers or even exclusive to written language (see \citealp[320--323]{Hentschel1993}).
-This is not surprising inasmuch as the genitive -- an intrinsic part of the PGC -- is generally underrepresented in colloquial vernacular variants of German as a result of a diachronic process wherein many (but by no means all) uses of the genitive are replaced by other cases or periphrastic constructions \citep{FleischerSchallert2011}.
+This is not surprising, given that the genitive -- an intrinsic part of the PGC -- is generally underrepresented in colloquial vernacular variants of German as a result of a diachronic process wherein many (but by no means all) uses of the genitive are replaced by other cases or periphrastic constructions \citep{FleischerSchallert2011}.
 Under an integral view of prototypes, which incorporate effects related to larger contexts and registers, such preferences can be part of what defines the construction prototypes.
@@
 \footnote{The COW corpora (Dutch, English, French, German, Spanish, Swedish) are made available for free at \url{https://www.webcorpora.org}.
-At the time of this writing, a newer 2016 version DECOW16 had already been released.}
+At the time of this writing, a newer 2016 version DECOW16 has already been released.}
 I chose this corpus for two main reasons.%
@@
 Second, it was already mentioned that normative grammars often adopt clear positions regarding the grammaticality of either the \NACa\ or the \PGCa.
-Thus, newspaper text or any other text that conforms strongly to normative grammars might not represent the alternation phenomenon fully (and without bias) because authors and proof-readers might favour one alternative or the other explicitly because the have to follow normative guidelines.
+Thus, newspaper text or any other text that conforms strongly to normative grammars might not represent the alternation phenomenon fully (and without bias) because authors and proofreaders who must adhere to normative guidelines might favour one alternative or the other explicitly.
 Web corpora, on the other hand, contain at least some amount of non-standard language from forums and similar sources.
 For these or similar reasons, COW corpora have been used in a number of peer-reviewed publications, for example \cite{VanGoethemHiligsmann2014,VanGoethemHuening2015,MuellerS2014,Schaefer2016c,SchaeferSayatz2014,SchaeferSayatz2016,Zimmer2015}.
-Therefore, DECOW can be considered the obvious choice for this study.
+Therefore, DECOW is the obvious choice for this study.
 
@@
 Therefore, it was highly desirable to obtain a sample in which most of the highly frequent actually-occurring combinations of kind nouns and measure nouns were represented.
-I applied a three-stage process in order to obtain such a sample.
-It consisted of three steps:
+I applied a three-stage process in order to obtain such a sample, which consisted of the following steps:
 
@@
   \item\label{enum:data:step1} generating a list of the one hundred most frequent mass nouns,
-  \item\label{enum:data:step2} deriving a list of all measure nouns with which the mass nouns co-occur in the \NACb 
+  \item\label{enum:data:step2} deriving a list of all measure nouns with which the mass nouns co-occur in the \NACb, and 
   \item\label{enum:data:step3} sampling the target constructions by querying each combination of mass noun and measure noun found in step (\ref{enum:data:step2}).
@@
 In order to reduce the sample size for the manual annotation process, the final concordance was sampled from the results of these 2,365 queries.
-Since the mass nouns in the sample were distributed according to the usual power law, I used all hits for nouns with a frequency up to 100, and a sample of 100 of all those with higher frequency.
+Since the mass nouns in the sample were distributed according to the usual power law, I used all hits for nouns with a frequency up to 100 and a sample of 100 of all those with higher frequency.
 The final sample contained 6,843 sentences, which was reduced to 5,063 in the manual annotation process due to removal of noisy material, erroneous hits and uninformative cases where the measure noun was in the genitive, in which case the \NACa\ cannot be distinguished from the \PGCa.
@@
   It could be argued that some more advanced measure of attraction strength should be used, as is done in Collostructional Analysis \citep{StefanowitschGries2003,GriesStefanowitsch2004}, see also \cite{Gries2015a}.
-  Three main points speak against such an approach in the present case (but see \citealp[246--249]{Levshina2016} for a more comprehensive approach albeit for a different purpose).
-  First, the goal here is to quantify how often lemmas occur in the \PGCd\ and the \NACb, and these construction do not compete at all but are rather mutually exclusive.
-  Collostructional approaches are not made such scenarios.
-  Second, the attraction values will be used as regressors in a hierarchical logistic regression, and the values resulting from collostructional analysis, \ie\ logarithmised Fisher p values, have a very unfavourable distribution in the case at hand.
-  They cluster around 0, and they include values of $-\infty$.
-  Third, I tried using collexeme strength as a regressor, and the results were unsatisfactory compared to the simple quotient used here.
+  Three main points speak against such an approach in the present case.
+  First, the goal here is to quantify how often lemmas occur in the \PGCd\ and the \NACb, and these constructions do not compete at all but are rather mutually exclusive.
+  Collostructional approaches are not made for such scenarios.
+  Second, the attraction values will be used as regressors in a hierarchical logistic regression and the values resulting from collostructional analysis, \ie\ logarithmised Fisher p-values, have a very unfavourable distribution in the case at hand.
+  They cluster around 0 and they include values of $-\infty$.
+  Third, I tried using collexeme strength as a regressor (with smoothing to remedy the mathematical problems) and the results were unsatisfactory compared to the simple quotient used here.
 }
@@
 To capture this, class information was annotated for measure lemmas.
-The classification was inspired by the list in \citet[530]{Koptjevskaja2001}, but due to the low frequencies of many of the potential classes, a very coarse classification was finally used.
+The classification was inspired by the list in \citet[530]{Koptjevskaja2001} but due to the low frequencies of many of the potential classes, a very coarse classification was finally used.
 With typical examples and their frequencies in the final sample, the classes are:
@@
 After all, GLMMs have been established as the major tool in the analysis of alternation phenomena.
-All studies mentioned at the outset of Section~\ref{sec:cogocl} use some form of regression\slash GL(M)M.
+All studies mentioned at the outset of Section~\ref{sec:cogocl} use some form of regression\slash GLMM.
 Over the past few years, however, modified or alternative methods have been proposed.
-From among these methods, I just make a few remarks on Bayesian estimation (see \citealp{GelmanEa2014}), as it was proposed in \cite{Levshina2016} and \cite{Divjak2016a}, for example.
+From among these methods, I just make a few remarks on Bayesian estimation (see \citealp{GelmanEa2014}) as it was proposed in \cite{Levshina2016} and \cite{Divjak2016a}, for example.
 Conceptually, I see three points of discussion that should be kept apart.
@@
 Second, it has been proposed that the Bayesian interpretation of probability is more cognitively adequate for the modeling of linguistic data \cite[301--302]{Divjak2016a}.
-Third, and very specific to this paper, given established methods in the modeling of alternation and variation, it has to be decided whether so-called Bayesian methods lead to substantially different results.
+Third and very specific to this paper, given established methods in the modeling of alternation and variation, it has to be decided whether so-called Bayesian methods lead to substantially different results.
 
@@
 % There is actually a debate among Bayesians about the proper interpretation of Bayesian methods, and whether a notion of hypothesis testing is compatible (or even already contained) in the Bayesian approach.
-% In \citet[10]{GelmanShalizi2013}, the authors -- prominent Bayesians themselves -- acknowledge that a theory of statistical testing is a desideratum, and they state about the standard inductive interpretation of Bayesianism that ``most of this received view of Bayesian inference is wrong'', and they develop a Bayesian notion of p values (see also \citealp{Mayo2013}, for a frequentist reply; also \citealp{Senn2011} on different strands of Bayesianism and their stance on inductive vs.\ deductive reasoning, and \citealp{Mayo2011}, for a critical reply to \citealp{Senn2011}).
+% In \citet[10]{GelmanShalizi2013}, the authors -- prominent Bayesians themselves -- acknowledge that a theory of statistical testing is a desideratum, and they state about the standard inductive interpretation of Bayesianism that ``most of this received view of Bayesian inference is wrong'', and they develop a Bayesian notion of p-values (see also \citealp{Mayo2013}, for a frequentist reply; also \citealp{Senn2011} on different strands of Bayesianism and their stance on inductive vs.\ deductive reasoning, and \citealp{Mayo2011}, for a critical reply to \citealp{Senn2011}).
 % Clearly, in such quarrels between and among camps of philosophers of science and statisticians, it is difficult for mere practitioners to take sides.
@@
 The first and second point cannot be discussed in detail here.%
-\footnote{However, among the Bayesian hype, there is also a number of critical papers by statisticians (even prominent Bayesians) and philosophers of science in which the extreme credibility that has been assigned to Bayesian approaches (and especially the standard inductivist interpretation of Bayesianism) is questioned (for example, \citealp{GelmanShalizi2013,Mayo2011,Senn2011}).}
-The third point, then, is the most practically relevant in the context of this paper.
+\footnote{However, there is a number of critical papers by statisticians (even prominent Bayesians) and philosophers of science in which the extreme credibility that has recently been assigned to (subjective and automatic) Bayesian approaches (and especially the standard inductivist interpretation of Bayesianism) is questioned (for example, \citealp{GelmanShalizi2013,Mayo2011,Senn2011}).}
 In her otherwise excellent study, \citet[251--252]{Levshina2016} argues for Bayesian estimation in mixed regression settings.
@@
 She then explains that a ``distinctive feature of Bayesian statistics is the use of so-called priors'' and that ``posterior probabilities depend on both the prior beliefs and the data, whereas the results of a frequentist model depend only on the data'' \cite[252]{Levshina2016}.
-Remarkably, given this statement, she does \textit{not} use informative priors, and in her footnote 8 \cite[252]{Levshina2016} admits that priors were probed using trial and error.
+Remarkably, given this statement, she does \textit{not} use informative priors and in her footnote 8 \cite[252]{Levshina2016} admits that priors were probed using trial and error.
 So, the proclaimed major advantage of Bayesian modeling was apparently not taken advantage of.%
@@
 MLE, like inductive Bayesianism, conditions on the particular data inasmuch as it searches for the most likely set of parameters given the data.
-What is more, Bayesian estimators are in fact based the Likelihood and merely multiply it by the prior \citep[6--8]{GelmanEa2014}.
+What is more, Bayesian estimators are in fact based on the Likelihood and merely multiply it by the prior \citep[6--8]{GelmanEa2014}.
 If the prior is flat, results converge (see also \citealp[347]{GelmanHill2006}).
@@
 % Going back to the argument that Bayesian methods are allegedly more cognitively plausible than `frequentist' methods, the claim might still be true.
-H% owever, most similar studies (dating back to \citealp{BresnanEa2007}) use an estimator (MLE) that is not frequentist in a narrow sense, and which is as cognitively plausible as Bayesian estimators because it most likely produces the same results in these studies.
-
-If in Levshina's study, results from Bayesian and MLE methods did, in fact, \textit{not} converge, it would have been an ideal occasion to demonstrate the superiority of the algorithms used in Bayesian estimation.
+% However, most similar studies (dating back to \citealp{BresnanEa2007}) use an estimator (MLE) that is not frequentist in a narrow sense, and which is as cognitively plausible as Bayesian estimators because it most likely produces the same results in these studies.
+
+If there had been evidence in Levshina's study that Bayesian and MLE methods would \textit{not} converge, it would have been an occasion to demonstrate the selective superiority of the algorithms used in Bayesian estimation.
 After all, there are situations where Bayesian estimators can be more robust, namely with heavily censored data, complex hierarchical models, perfect separation, etc.
 (see \citealp{Freedman1999}, \citealp[345--348]{GelmanHill2006}).
-I want to reiterate that these points do not in any way invalidate the results presented in \cite{Levshina2016}.
+I want to state clearly that these points do not in any way invalidate the results presented in \cite{Levshina2016}.
 However, being \textit{Bayesian} is most likely not among its selling points.
-That said, I want to voice the concern that probably, many practitioners are already struggling with getting an adequate grasp of advanced statistical methods and that it might therefore be wise to use the more conservative and better understood method if the alternative method is not absolutely required for substantive reasons.
-In Section~\ref{sec:corpushierarchicalmodel}, I compare my own hierarchical models estimated with Bayesian and MLE estimators to demonstrate their expectable convergence.
+Additionally, I want to voice the concern that probably, many practitioners are already struggling with getting an adequate grasp of advanced statistical methods and that it might therefore be wise to use the more conservative and better understood method if the alternative method is not absolutely required for substantive reasons.
+In Section~\ref{sec:corpushierarchicalmodel}, I estimate the parameters of my hierarchical model with MLE and Markov-Chain Monte Carlo (MCMC) methods (the currently most prominent estimator used in Bayesian settings) to demonstrate their expectable convergence.
 
@@
 
-In this section, I report the results of fitting a multilevel model to the data using R \citep{R}, \textit{lme4} \citep{lme4} for Maximum Likelihood Estimation, and \textit{rstanarm} \citep{rstanarm} for `Bayesian' Markov-Chain Monte Carlo estimation (see Section~\ref{sec:rightstatistics}).
+In this section, I report the results of fitting a multilevel model to the data using R \citep{R}, \textit{lme4} \citep{lme4} for Maximum Likelihood Estimation (MLE) and \textit{rstanarm} \citep{rstanarm} for ``Bayesian'' Markov-Chain Monte Carlo (MCMC) estimation (see Section~\ref{sec:rightstatistics}).
 The purpose is to model the influence of the regressors specified in Table~\ref{tab:variables} on the probability that the \PGCa\ is chosen over the \NACa.
@@
 The coefficient estimates are specified in Table~\ref{tab:bigtable} for each regressor (or regressor level) in the columns labelled \textit{Coefficient}.
-For a robust quantification of the precision of the estimation, I ran a parametric bootstrap (using the \mbox{\textit{confint.merMod}} function from \textit{lme4}) with 1,000 replications, and using the percentile method for the calculation of the intervals.
+For a robust quantification of the precision of the estimation, I ran a parametric bootstrap (using the \mbox{\textit{confint.merMod}} function from \textit{lme4}) with 1,000 replications and using the percentile method for the calculation of the intervals.
 The resulting 95\% bootstrap confidence intervals are reported in Table~\ref{tab:bigtable} in the columns labelled \textit{CI low} and \textit{CI high} (= upper and lower 2.5th percentiles).
 The column \textit{CI contains 0} shows an asterisk for those intervals that do \textit{not} include 0.
-Furthermore, for each regressor, a p value was obtained by dropping the regressor from the full model, re-estimating the nested model and comparing it to the full model.
-Instead of inexact Wald approximations or Likelihood Ratio Tests, I used a drop-in bootstrap replacement for the Likelihood Ratio Test from the function \textit{PBmodcomp} from the \textit{pbkrtest} package \citep{HalekohHojsgaard2014}.
-I call the corresponding value $p_{\text{PB}}$, and it is given in the appropriately labelled columns in Table~\ref{tab:bigtable}.
+Furthermore, for each regressor, a p-value was obtained by dropping the regressor from the full model, re-estimating the nested model, and comparing it to the full model.
+Instead of inexact Wald approximations and Likelihood Ratio Tests, I used a drop-in bootstrap replacement for the Likelihood Ratio Test from the function \textit{PBmodcomp} from the \textit{pbkrtest} package \citep{HalekohHojsgaard2014}.
+I call the corresponding value $p_{\text{PB}}$\ and it is given in the appropriately labelled columns in Table~\ref{tab:bigtable}.
 Only \textit{Kindfreq} ($\mpPB=0.095$) can be seen as slightly too high to be convincing (non-significant).
 
-Finally, I show that Bayesian methods in the form of Markov-Chain Monte Carlo sampling (MCMC) do not necessarily lead to different results.
+Finally, I show that MCMC does not necessarily lead to different results.
 Actually, given the low complexity of the model and the large sample size, it would be surprising if they did (see Section~\ref{sec:rightstatistics}).
 The model was re-estimated using the \textit{stan\_glmer} function from the \textit{rstanarm} package, which provides an \textit{lme4}-compatible syntax for estimating common model types with the \textit{stan} software \citep{CarpenterEa2017}.
-The algorithm was run with 4 chains and 1,000 iterations, and I used plausible default priors.
+The algorithm was run with 4 chains and 1,000 iterations and I used plausible default priors.
 Most notably, priors for coefficients were specified as $\mathcal{N}(0,10)$ because coefficients higher than 10 or lower than $-10$ are extremely rare in well-specified models on appropriate data.%
@@
 To reliably estimate such coefficients, extremely large samples would be required.}
-The algorithm converged, and for all coefficients, the $\hat{R}$ diagnostic was exactly 1.
-The resulting coefficients and intervals as well as an * are also given in Table~\ref{tab:bigtable} in the columns labelled \textit{MCMC}.
+The algorithm converged and for all coefficients, the $\hat{R}$ diagnostic was exactly 1.
+The resulting coefficients and intervals as well as an * for ``confidence interval does not contain 0'' are also given in Table~\ref{tab:bigtable} in the columns labelled \textit{MCMC}.
 Both methods lead to exactly the same results (minus negligible numerical differences) as expected given the modest complexity of the model structure and the large sample size (see Section~\ref{sec:rightstatistics}).
-The signs and magnitudes of the coefficients are identical, and confidence intervals have the same width and symmetry properties.
+The signs and magnitudes of the coefficients are identical and confidence intervals have the same width and symmetry properties.
 Figure~\ref{fig:fixeffs} illustrates this by also showing both estimates.
@@
   \includegraphics[width=0.5\textwidth]{../R/output/corpus_Measureattraction}~\includegraphics[width=0.5\textwidth]{../R/output/corpus_Kindattraction}
-  \caption{Effect plots for the regressors \textit{Measureattraction} and \textit{Kindattraction}; y axes are not aligned}
+  \caption{Effect plots for the regressors \textit{Measureattraction} and \textit{Kindattraction}; y-axes are not aligned}
   \label{fig:eff:attraction}
@@
 \footnote{Effect plots were created using the \textit{effects} package \citep{Fox2003}.
-They show the changes in probability for the outcome (y axis) dependent on values of a regressor (x axis), at typical values of all other regressors.
+They show the changes in probability for the outcome (y-axis) dependent on values of a regressor (x-axis) at typical values of all other regressors.
 The vertical bars (categorical variables) and grey areas (continuous variables) are asymptotic 95\% confidence intervals calculated from \textit{glmer}.
 They are not bootstrapped.}
-The effect is mostly as expected:
-if a lemma appears relatively more often in the \PGCd\ (compared to its frequency in the \NACb), the more often the \PGCa\ is chosen over the \NACa\ with this specific lemma.
+The effect is as expected:
+if a lemma appears relatively more often in the \PGCd\ (compared to its frequency in the \NACb), the \PGCa\ tends to be chosen over the \NACa\ with this specific lemma.
 The effect for measure nouns is stronger, and it was estimated with higher precision.
@@
 An interesting picture emerges for the lemma frequencies.
-A higher than average lemma frequency of measure nouns favours the \NACa\ ($\beta_{\text{Measurefreq}}=-0.231$, $\mpPB=0.005$) as expected if we assume at least a tendency for highly grammaticalised items to be more frequent.
+A higher-than-average lemma frequency of measure nouns favours the \NACa\ ($\beta_{\text{Measurefreq}}=-0.231$, $\mpPB=0.005$), which is as expected if we assume at least a tendency for highly grammaticalised items to be more frequent.
 With kind nouns, higher frequency seems to favour the \PGCa ($\beta_{\text{Kindfreq}}=0.146$, $\mpPB=0.095$).
@@
 
-The register-related proxy variables point into the expected direction.
-Increased \textit{Badness} of the document favours the \NACa\ ($\beta_{\text{Badness}}=-0.152$, $\mpPB=0.002$), and so does a lesser density of genitives ($\beta=-0.693$, $\mpPB=0.001$).
+The register-related proxy variables point to the expected direction.
+Increased \textit{Badness} of the document favours the \NACa\ ($\beta_{\text{Badness}}=-0.152$, $\mpPB=0.002$) and so does a lower density of genitives ($\beta=-0.693$, $\mpPB=0.001$).
 While these are poor proxies to register (and partially circular in the case of \textit{Genitives}), this result can at least encourage future work into register effects.
@@
 
-In the two experiments reported in this section, I use probabilities for the alternating constructions calculated for attested material, and I correlate these probabilities with the participants' reactions.
+In the two experiments reported in this section, I use probabilities for the alternating constructions calculated for attested material and I correlate these probabilities with the participants' reactions.
 Thus, a direct link can be established between output material found in corpora and the behaviour of linguistic agents (see also Section~\ref{sec:cogocl}).
 Both experiments use sentences containing attested MNPs from the corpus sample (embedded into simplified sentences) as stimuli.
-Also, the probabilities that the corpus-based model assigns to the two alternants in these sentences is used as the main regressor in both studies.
+Also, the probabilities that the corpus-based model assigns to the two alternants in these sentences are used as the main regressor in both studies.
 
@@
 In \cite{VerhoevenTemme2017}, it was shown that results from forced-choice and split-100 experiments mostly converge, which is expected in the long run if a probability is turned into a binary decision.
-I present a forced-choice experiment in this Section and not a split-100 experiment, mainly because in a dry-run of the experiment, participants complained about the unnaturalness of the task of distributing a probability mass across two alternants.
-Participants had to chose between two sentences differing only in that one contained the \NACa\ and the other contained the \PGCa.
+I present a forced-choice experiment in this Section and not a split-100 experiment, mainly because in a dry run of the experiment, participants complained about the unnaturalness of distributing a probability mass across two alternants.
+Participants had to choose between two sentences that differed only in that one contained the \NACa\ and the other contained the \PGCa.
 The analysis compares the probabilities assigned to stimuli by the corpus-derived model with the frequency with which participants chose the alternants for the same stimuli.
 
-There were 24 participants (native speakers of German without reading or writing disabilities) aged 19 to 30 living permanently in [name of city anonymised], who were recruited from introductory linguistics courses at [name of university anonymised].
+There were 24 participants (native speakers of German without reading or writing disabilities) aged 19 to 30 living permanently in [name of city redacted], who were recruited from introductory linguistics courses at [name of university redacted].
 Although the experiment was conducted in the last four weeks of their first semester, participants had no deeper explicit knowledge of linguistics, grammar, or experimental methods.
 None of them had ever participated in a forced-choice experiment before.
-Participation was voluntary, but participants received credit in partial fulfillment of course requirements.
+Participation was voluntary but participants received credit in partial fulfillment of course requirements.
 
@@
 The approach is also justified because according to the theoretical assessment in Section~\ref{sec:analyses}, the choice of alternants depends mostly on a very local constructional context.
-I sampled 16 MNPs from the corpus, and it was made sure that the simplifications and normalisations did not affect any of the regressors used in the corpus study.
-In the simplified sentences, the case, number, etc. of the MNP remained the same as in the attested sentence, as well as the choice of lexical material within the MNP.
-Eight sentences contained masculine or neuter kind nouns, the other eight contained feminine kind nouns.
+I sampled 16 MNPs from the concordance and made sure that the simplifications and normalisations did not affect any of the regressors used in the corpus study.
+In the simplified sentences, the case, number, etc. of the MNP remained the same as in the attested sentence, as did the choice of lexical material within the MNP.
+Eight sentences contained masculine or neuter kind nouns and the other eight contained feminine kind nouns.
 Furthermore, in each of the masculine\slash neuter and feminine groups, four sentences originally containing the \NACa\ and four sentences originally containing the \PGCa\ were chosen.
@@
 The pairs of stimuli were the sentence containing the preferred construction (according to the corpus GLMM) and a modified version containing the dispreferred construction.
-They were presented next to each other, and a 20 second time limit for each choice was set.%
+They were presented next to each other and a 20 second time limit for each choice was set.%
 \footnote{No participant ever exceeded the time limit.}
 The position on the screen (left\slash right) and the order of sentences were randomised for each participant.
-As fillers, 23 pairs of sentences exemplifying similar alternation phenomena from German morpho-syntax were used.
+As fillers, 23 pairs of sentences exemplifying similar but unrelated alternation phenomena from German morpho-syntax were used.
 Thus, participants saw 39 pairs of sentences and 78 sentences in total.
@@
 Also, among participants, there are clearly different preferences ($\sigma_{\text{Participant}}=0.781$, $\sigma^{\text{Modelprediction}}_{\text{Participant}}=0.484$, 24 levels).
-On the extreme sides, one participant chose the \PGCa\ in 13 of 16 cases, and two participants only chose it in 5 of 16 cases.
+On the extreme ends, one participant chose the \PGCa\ in 13 of 16 cases and two participants only chose it in 5 of 16 cases.
 The regressor \textit{Modelprediction} achieves $\mpPB=0.007$ (1,000 replications) and is estimated at 5.408 relative to an intercept of -1.304.
@@
 The result is very clear.
-The higher the probability of the \PGCa\ predicted from usage-data, the more often participants chose the \PGCa\ alternant in the forced-choice task.
+The higher the probability of the \PGCa\ predicted from usage data, the more often participants chose the \PGCa\ alternant in the forced-choice task.
 In summary, the forced-choice experiment clearly succeeded in corroborating the results from the corpus study in as much as the preferences extracted from usage data correspond to native speakers' choices.
@@
 It is expected that reading less prototypical alternants (the one assigned a low probability by the corpus-derived model) in a given context and with given lexical material incurs a processing overhead for the reader (\citealp{Kaiser2013}).
-In this Section, a self-paced reading experiment is therefore presented.
+In this section, a self-paced reading experiment is therefore presented.
 In a very similar fashion, \cite{DivjakEa2016} apply the self-paced reading paradigm in the validation of corpus-based models.
@@
 Each participant read both the 16 sentences with the alternant predicted by the corpus model and the 16 modified sentences with the alternant that the corpus model did not predict.%
-\footnote{Notice that lemmas and their frequencies as well as lemma classes are included as regressors in the corpus-based GLMM, and there was consequently no additional controlling of lemma frequencies, etc.}
+\footnote{Notice that lemmas and their frequencies as well as lemma classes are included as regressors in the corpus-based GLMM and there was consequently no additional controlling of lemma frequencies, etc.}
 To minimise repetition effects, the stimuli for each participant were separated into two blocks of 16 targets and 33 fillers per block.
@@
 From each target sentence pair, one sentence was assigned to the first block and the other sentence to the other block.
-The assignment of members of the individual sentence pairs to the blocks was randomised for each participant individually and so was the order within each block.
+The assignment of members of the individual sentence pairs to the blocks was randomised for each participant individually, as was the order within each block.
 The sentences from each pair of alternants were kept as far apart as possible.
@@
 The reading times were residualised per speaker based on the reading times of all words (not just the targets) by that speaker.
-The adjective and the kind noun (\ie\ the constituents bearing the critical case markers) were used as the target region, such as the bracketed words in the example \textit{zwei Gläser} [\textit{spru\-deln\-des Wasser}] `two glasses of sparkling water'.
+The adjective and the kind noun (\ie\ the constituents bearing the critical case markers) were used as the target region, for instance the bracketed words in the example \textit{zwei Gläser} [\textit{spru\-deln\-des Wasser}] `two glasses of sparkling water'.
 Outliers farther than 2 inter-quartile ranges from the mean logarithmised residualised reading time were removed (64 data points), resulting in a total number of \textit{n=}1,152 observations.
@@
 As a consequence, the higher the GLMM prediction is, the more prototypical the sentence is for containing the \PGCa.
-Therefore, it is expected that reading times are higher when the value of \textit{Modelprediction} is higher but the sentence contains the \NACa.
+It is therefore expected that reading times are higher when the value of \textit{Modelprediction} is higher but the sentence contains the \NACa.
 However, when the sentence contains the \PGCa, reading times should be lower when \textit{Modelprediction} is higher.
@@
 Random intercepts were specified for \textit{Participant} and \textit{Item} (the 16 sentence pairs are one \textit{Item} each).%
-\footnote{I tried random slopes in order to keep the random effect structure maximal \citep{BarrEa2013}, but it was impossible to get the algorithm to converge due to the added complexity of the interaction.}
+\footnote{I tried random slopes in order to keep the random effect structure maximal \citep{BarrEa2013} but it was impossible to get the algorithm to converge due to the added complexity of the interaction.}
 
@@
 
-The overall model quality is high, and the effect plot for the main effect is shown in Figure~\ref{fig:spr:effects}.
-The estimate for the sentences with \NACa\ is obviously imprecise, and no differences in reading times are observed.
+The overall model quality is high and the effect plot for the main effect is shown in Figure~\ref{fig:spr:effects}.
+The estimate for the sentences with \NACa\ is obviously imprecise and no differences in reading times are observed.
 There is a clearer effect in the sentences with \PGCa, which is also confirmed by the significant results from the bootstrapped confidence intervals (see Table~\ref{tab:exp:spr}) and from the PB test reported above.
@@
 Future work could extend these results and provide a general picture of the constructions expressing measurements (see Section~\ref{sec:descriptive}).
-This would be a much more complicated task given that the choices then would no longer be binary, and that the meaning of the alternative ways of expressing measurements are semantically more varied.
+This would be a much more complicated task given that the choices then would no longer be binary and that the meaning of the alternative ways of expressing measurements are semantically more varied.
 Finally, I want to point out that German is mildly under-researched in the specific framework used here.

### 2017-04-18 Latex/article.tex
Submitted version.

@@
   \runningauthor{Roland Schäfer. Submitted April 2017. Do not quote.}
+  %\affil[1]{}
   \affil[1]{Freie Universität Berlin}
@@
   %\subtitle{...}
-  \abstract{In this paper, a case alternation occurring in German measure noun phrases is examined from the perspective of cognitively oriented corpus linguistics.
-    In pseudo-partitive constructions such as \textit{eine Tasse guter Kaffee} `a cup of good coffee', the embedded kind-denoting noun either agrees in its case with the measure head noun or it stands in the genitive as in \textit{eine Tasse guten Kaffees} with no difference in meaning.
-    This alternation occurs only under specific syntactic conditions.
-    The analysis is formulated in terms of Prototype Theory. 
-    I assume that the alternants are modelled after non-alternating neighbouring constructions, which represent the prototypes more directly.
-    Primarily, it is argued that the frequencies with which individual lemmas occur in the prototypical constructions partially predict which alternant is chosen in the alternating case.
-    Another major assumption is that the genitive construction is prototypical of more strongly grammaticalised measure nouns.
-    I present a large-scale corpus study using the DECOW web corpus which supports the theory.
-    In the statistical analysis of the corpus study, it is discussed whether Bayesian estimators lead to substantially different results than Likelihood Maximisation methods in this and similar studies.
-    The model parameters are estimated with both methods for comparison.
-    Finally, a forced choice experiment and a self-paced reading experiment are reported.
-    The usage-based findings from the corpus study show a correlation with the behaviour of cognitive agents in both experiments.
-    The present study therefore contributes to a now well-established area of research into alternations using both corpus and experimental methods.}
-  \keywords{corpus methods and experimental methods, self-paced reading, forced-choice, alternations, hierarchical models, measure constructions, German}
+  \abstract{In this paper, a case alternation in German measure noun phrases is examined from the perspective of cognitively oriented corpus linguistics.
+    In certain pseudo-partitive constructions, the embedded kind-denoting noun either agrees in its case with the measure head noun (\textit{eine Tasse guter Kaffee} `a cup of good coffee') or it stands in the genitive (\textit{eine Tasse guten Kaffees}).
+    My analysis is formulated in terms of Prototype Theory.
+    I assume that the choice of alternants is influenced by non-alternating neighbouring constructions representing the prototypes more directly.
+    I argue that the frequencies with which individual lemmas occur in these prototypical non-alternating constructions partially predict which alternant is chosen and that the genitive alternant is prototypical of strongly grammaticalised measure nouns.
+    I present a large-scale corpus study using the DECOW corpus to support this theory.
+    In the statistical analysis, I compare Maximum Likelihood estimators to Bayesian estimators (recently proposed for similar studies), showing that results \textit{predictably} do not differ.
+    Finally, two experiments (forced choice and self-paced reading) are reported.
+    The usage-based findings show a clear correlation with the behaviour of cognitive agents in both experiments.
+    The present study therefore contributes to the well-established field of research into alternations using corpus and experimental methods.}
+  \keywords{corpus methods and experimental methods, alternations, hierarchical models, pseudo-partitives, German}
   %\classification[PACS]{...}
@@
 If the kind noun is bare -- \ie\ if it comes neither with a determiner nor a modifying adjective -- it has to agree in case with the measure noun as in (\ref{ex:intro:narrowapposition1}), and the genitive as seen in the PGC is not acceptable, see (\ref{ex:intro:narrowapposition2}).
-
-\enlargethispage{1\baselineskip}
 
@@
 It was reported in the literature that MNPs in the dative and with a masculine or neuter kind noun favour the \PGCa\ more than the corresponding nominative and accusative MNPs \citep{Hentschel1993,Zimmer2015}.
-As an example, \textit{mit einem Stück frischen Brots} `with a piece of fresh bread' (\PGCa) would be preferred over \textit{mit einem Stück frischem Brot} (\NACa).
+As an example, \textit{mit einem Stück frischen Brots} `with a piece of fresh bread' (\PGCa) would be preferred more strongly against \textit{mit einem Stück frischem Brot} (\NACa).
 As with all the examples, native speakers of German will most likely notice that differences are subtle.
@@
 % There is actually a debate among Bayesians about the proper interpretation of Bayesian methods, and whether a notion of hypothesis testing is compatible (or even already contained) in the Bayesian approach.
-% In \citet[10]{GelmanShalizi2013}, the authors -- prominent Bayesians themselves -- acknowledge that a theory of statistical testing is a desideratum, and they state about the standard inductive interpretation of Bayesianism that ``most of this received view of Bayesian inference is wrong'', and they develop a Bayesian notion of p-values (see also \citealp{Mayo2013}, for a frequentist reply; also \citealp{Senn2011} on different strands of Bayesianism and their stance on inductive vs.\ deductive reasoning, and \citealp{Mayo2011}, for a critical reply to \citealp{Senn2011}).
+% In \citet[10]{GelmanShalizi2013}, the authors -- prominent Bayesians themselves -- acknowledge that a theory of statistical testing is a desideratum, and they state about the standard inductive interpretation of Bayesianism that ``most of this received view of Bayesian inference is wrong,'' and they develop a Bayesian notion of p-values (see also \citealp{Mayo2013}, for a frequentist reply; also \citealp{Senn2011} on different strands of Bayesianism and their stance on inductive vs.\ deductive reasoning, and \citealp{Mayo2011}, for a critical reply to \citealp{Senn2011}).
 % Clearly, in such quarrels between and among camps of philosophers of science and statisticians, it is difficult for mere practitioners to take sides.
@@
 \footnote{However, there is a number of critical papers by statisticians (even prominent Bayesians) and philosophers of science in which the extreme credibility that has recently been assigned to (subjective and automatic) Bayesian approaches (and especially the standard inductivist interpretation of Bayesianism) is questioned (for example, \citealp{GelmanShalizi2013,Mayo2011,Senn2011}).}
-In her otherwise excellent study, \citet[251--252]{Levshina2016} argues for Bayesian estimation in mixed regression settings.
+With regard to the third point, \citet[251--252]{Levshina2016} argues for Bayesian estimation in mixed regression settings.
 First, she claims that ``while frequentist statistics only allows one to test whether the null hypothesis can be rejected, Bayesian statistics enables one both to test the null hypothesis and to estimate the probability of specific parameter values given the data.''
@@
 
-If there had been evidence in Levshina's study that Bayesian and MLE methods would \textit{not} converge, it would have been an occasion to demonstrate the selective superiority of the algorithms used in Bayesian estimation.
+If there had been evidence in Levshina's study that Bayesian and MLE methods did \textit{not} converge, it would have been an occasion to demonstrate the selective superiority of the algorithms used in Bayesian estimation.
 After all, there are situations where Bayesian estimators can be more robust, namely with heavily censored data, complex hierarchical models, perfect separation, etc.
@@
 I want to state clearly that these points do not in any way invalidate the results presented in \cite{Levshina2016}.
-However, being \textit{Bayesian} is most likely not among its selling points.
-Additionally, I want to voice the concern that probably, many practitioners are already struggling with getting an adequate grasp of advanced statistical methods and that it might therefore be wise to use the more conservative and better understood method if the alternative method is not absolutely required for substantive reasons.
+However, being ``Bayesian'' is most likely not among its selling points.
+Additionally, I want to voice the concern that many practitioners are probably already struggling with getting an adequate grasp of advanced statistical methods and that it might therefore be wise to use the more conservative and better understood method if the alternative method is not absolutely required for substantive reasons.
 In Section~\ref{sec:corpushierarchicalmodel}, I estimate the parameters of my hierarchical model with MLE and Markov-Chain Monte Carlo (MCMC) methods (the currently most prominent estimator used in Bayesian settings) to demonstrate their expectable convergence.
@@
 
-In this section, I report the results of fitting a multilevel model to the data using R \citep{R}, \textit{lme4} \citep{lme4} for Maximum Likelihood Estimation (MLE) and \textit{rstanarm} \citep{rstanarm} for ``Bayesian'' Markov-Chain Monte Carlo (MCMC) estimation (see Section~\ref{sec:rightstatistics}).
+In this section, I report the results of fitting a multilevel model to the data using R \citep{R}, \textit{lme4} \citep{lme4} for Maximum Likelihood Estimation (MLE) and \textit{rstanarm} \citep{rstanarm} for Bayesian Markov-Chain Monte Carlo (MCMC) estimation (see Section~\ref{sec:rightstatistics}).
 The purpose is to model the influence of the regressors specified in Table~\ref{tab:variables} on the probability that the \PGCa\ is chosen over the \NACa.
@@
 
-\begin{acknowledgement}
-  I thank (in alphabetical order) Felix Bildhauer, Susanne Flach, Elizabeth Pankratz, Samuel Reichert, and Ulrike Sayatz for valuable discussions and comments.
-  Elizabeth thankfully also fixed my English.
-  Furthermore, I thank Ulrike for helping me to recruit the participants for the experiments.
-  Finally, I am grateful to Kim Maser for her work on the annotation of the concordances.
-  The research presented here was made possible in part through funding from the \textit{Deutsche Forschungsgemeinschaft} (DFG, personal grant SCHA1916/1-1).
-\end{acknowledgement}
+% \begin{acknowledgement}
+%   I thank (in alphabetical order) Felix Bildhauer, Susanne Flach, Elizabeth Pankratz, Samuel Reichert, and Ulrike Sayatz for valuable discussions and comments.
+%   Elizabeth thankfully also fixed my English.
+%   Furthermore, I thank Ulrike for helping me to recruit the participants for the experiments.
+%   Finally, I am grateful to Kim Maser for her work on the annotation of the concordances.
+%   The research presented here was made possible in part through funding from the \textit{Deutsche Forschungsgemeinschaft} (DFG, personal grant SCHA1916/1-1).
+% \end{acknowledgement}
 

### 2017-04-18 Latex/article.tex
Submitted version.

@@
 
-% \begin{acknowledgement}
-%   I thank (in alphabetical order) Felix Bildhauer, Susanne Flach, Elizabeth Pankratz, Samuel Reichert, and Ulrike Sayatz for valuable discussions and comments.
-%   Elizabeth thankfully also fixed my English.
-%   Furthermore, I thank Ulrike for helping me to recruit the participants for the experiments.
-%   Finally, I am grateful to Kim Maser for her work on the annotation of the concordances.
-%   The research presented here was made possible in part through funding from the \textit{Deutsche Forschungsgemeinschaft} (DFG, personal grant SCHA1916/1-1).
-% \end{acknowledgement}
+\begin{acknowledgement}
+  I thank (in alphabetical order) Felix Bildhauer, Susanne Flach, Elizabeth Pankratz, Samuel Reichert, and Ulrike Sayatz for valuable discussions and comments.
+  Elizabeth thankfully also fixed my English.
+  Furthermore, I thank Ulrike for helping me to recruit the participants for the experiments.
+  Finally, I am grateful to Kim Maser for her work on the annotation of the concordances.
+  The research presented here was made possible in part through funding from the \textit{Deutsche Forschungsgemeinschaft} (DFG, personal grant SCHA1916/1-1).
+\end{acknowledgement}
 

### 2017-04-19 Latex/article.tex
Submitted version plus fixed typos etc.

@@
 I call the narrow apposition construction with a bare kind noun the \NACb\ and the partitive genitive with a determiner in the kind NP the \PGCd.
-For the alternants with an adjective but no determiner in the kind noun phrase I use the terms \NACa and \PGCa.
+For the alternants with an adjective but no determiner in the kind noun phrase I use the terms \NACa\ and \PGCa.
 In principle, this paper is about the middle column of Table~\ref{tab:constructions}, \ie\ the syntactic configuration in which two different case patterns are acceptable.
@@
 They are typical measure nouns like \textit{Gramm} `gram', \textit{Pfund} `pound' or \textit{Prozent} `percent', which have no normal plural forms at all.%
-\footnote{Plurals \textit{Pfunde} and \textit{Prozente} have special meanings and have very restricted uses, mostly in idiomatic expressions and prefabs.
+\footnote{Plurals like \textit{Pfunde} and \textit{Prozente} have special meanings and have very restricted uses, mostly in idiomatic expressions and prefabs.
 They cannot be used in MNPs.}
@@
 They are not semantically equivalent to the NAC and the PGC and they are only available for a small subset of measure nouns.
-All of these constructions are used only with measure nouns denoting containers (where the whole NP always refers to the container) and never to the quantity contained in it.
+All of these constructions are used only with measure nouns denoting containers, where the whole NP always refers to the container and never to the quantity contained in it.
 They are consequently incompatible with measure nouns denoting natural portions such as \textit{Schluck} `gulp' or \textit{Haufen} `heap' and strongly grammaticalised nouns such as \textit{Gramm} `gram'.
@@
 
-Second, I focus on grammaticalisation.
+Second, therefore, I focus on grammaticalisation.
 It is often assumed that pseudo-partitives arise as a form of grammaticalised partitives (\eg\ \citealp[536--539]{Koptjevskaja2001} for Finnish and Estonian, \citealp[559]{Koptjevskaja2001} for European languages in general).
@@
 Due to the noisy nature and unknown composition of the web, only carefully designed and established web corpora like the COW corpora or the SketchEngine corpora \citep{KilgarriffEa2014} should be used.
-Clearly, using search engine results is `bad science' for many reasons, most prominently total irreproducibility of results, as \cite{Kilgarriff2006} pointed out more than ten years ago.
+Clearly, using search engine results is ``bad science'' for many reasons, most prominently total irreproducibility of results, as \cite{Kilgarriff2006} pointed out more than ten years ago.
 Careless use of search engine results is still found, however, see for example \citet[171--175]{DeclerckBrems2016}.}
@@
 Since the \PGCa\ contains a genitive itself, the regressor variable \textit{Genitive} and the document-level variable \textit{Genitives} are not fully independent.
-However, since instances of the \PGCa\ make up for only a minute fraction of all genitives, I still use \textit{Genitives} as a regressor with the appropriate caveats.
+Since instances of the \PGCa\ make up for only a minute fraction of all genitives, I still use \textit{Genitives} as a regressor with the appropriate caveats.
 
@@
 The resulting 95\% bootstrap confidence intervals are reported in Table~\ref{tab:bigtable} in the columns labelled \textit{CI low} and \textit{CI high} (= upper and lower 2.5th percentiles).
-The column \textit{CI contains 0} shows an asterisk for those intervals that do \textit{not} include 0.
+The column \textit{CI excludes 0} shows an asterisk for those intervals that do not include 0.
 Furthermore, for each regressor, a p-value was obtained by dropping the regressor from the full model, re-estimating the nested model, and comparing it to the full model.
@@
 The resulting coefficients and intervals as well as an * for ``confidence interval does not contain 0'' are also given in Table~\ref{tab:bigtable} in the columns labelled \textit{MCMC}.
-Both methods lead to exactly the same results (minus negligible numerical differences) as expected given the modest complexity of the model structure and the large sample size (see Section~\ref{sec:rightstatistics}).
+Both methods lead to exactly the same results (minus negligible numerical differences) as expected given the modest complexity of the model structure and the large sample size.
 The signs and magnitudes of the coefficients are identical and confidence intervals have the same width and symmetry properties.
@@
   \begin{tabular}{lrrrc}
-    Regressor & \multicolumn{1}{r}{Coefficient} & \multicolumn{1}{r}{CI low} & \multicolumn{1}{r}{CI high} & \multicolumn{1}{r}{0 not in CI} \\ \midrule
+    Regressor & \multicolumn{1}{r}{Coefficient} & \multicolumn{1}{r}{CI low} & \multicolumn{1}{r}{CI high} & \multicolumn{1}{r}{CI excludes 0} \\ \midrule
     ConstructionPGCa                 &  0.054 &  0.012 &  0.095 &  *  \\ 

### 2017-05-01 Latex/article.tex
Minor fixes after submission.

@@
 Then, predictions are derived from this theory \textit{before} the experiment or the corpus study is conducted to \textit{test} the theory.
-Technical problems of statistical inference aside (see also Section~\ref{sec:rightstatistics}), a successful experiment corroborates the theory, and a failed experiment weakens the theory.
+Technical problems of statistical inference aside, a successful experiment corroborates the theory, and a failed experiment weakens the theory.
 The simplest assumption would be that the same approach is adequate for corpus studies.
@@
 Prototype Theory is well suited for modeling constructional choices but it is just one of several similarity-based theories of classification, the most prominent other framework being Exemplar Theory (\citealp{MedinSchaffer1978,Hintzman1986}; see \citealp{StormsEa2000} for a comparison of the theories in experimental settings).
-Prototype Theory and Exemplar Theory model essentially the same types off effects and differ mainly in whether they assume higher-level abstractions as part of the cognitive representation (Prototype Theory) or try to do without them (Exemplar Theory).
+Prototype Theory and Exemplar Theory model essentially the same types of effects and differ mainly in whether they assume higher-level abstractions as part of the cognitive representation (Prototype Theory) or try to do without them (Exemplar Theory).
 I am sceptical that corpus analysis alone could ever decide which theory is more suitable, given that substantial doubt has been voiced whether even experimental methods are ultimately able to do so \citep{Barsalou1990}.%
@@
 \footnote{Some descriptive and normative grammars take stronger positions with regard to the acceptability of the two options.
-See \cite{Hentschel1993,Zimmer2015} for analyses of the sometimes absurd stances taken in grammars of German.
+See \cite{Hentschel1993} and \cite{Zimmer2015} for analyses of the sometimes absurd stances taken in grammars of German.
 As the usage and experimental data presented below (especially Section~\ref{sec:exp:fc}) should render it clear, there might be preferences, but we cannot assume either construction to be unacceptable.
@@
 They are typical measure nouns like \textit{Gramm} `gram', \textit{Pfund} `pound' or \textit{Prozent} `percent', which have no normal plural forms at all.%
-\footnote{Plurals like \textit{Pfunde} and \textit{Prozente} have special meanings and have very restricted uses, mostly in idiomatic expressions and prefabs.
-They cannot be used in MNPs.}
+\footnote{Plurals like \textit{Pfunde} and \textit{Prozente} have special meanings and have very restricted uses, mostly in idiomatic expressions and prefabs such as \textit{Pfunde verlieren} `lose pounds.'
+They cannot be used in normal MNPs.}
 %\footnote{With neglectable frequency, this effect is extended to container nouns like \textit{Glas} `glass', but with a twist.
@@
 I treat these cases like other measure nouns because they enter into both the \NACa\ as in (\ref{ex:uninflectedmeasures:a}) and the \PGCa\ as in (\ref{ex:uninflectedmeasures:b}).
-In Section~\ref{sec:analyses}, however, degrees of grammaticalisation as a factor influencing the alternation will be discussed.
+In Section~\ref{sec:analyses}, however, degrees of grammaticalisation as a factor influencing the alternation will be discussed prominently.
 
@@
 
-This section briefly reviews existing analyses of the \PGCa vs.\ \NACa\ alternation and related issues.
+This section briefly reviews existing analyses of the \PGCa\ vs.\ \NACa\ alternation and related issues.
 I also develop my own analysis and the appropriate hypotheses for the empirical studies presented in Sections~\ref{sec:corpusstudies} and \ref{sec:experimental}.
@@
 However, it is much more prototypical of quantities to be specified with numerical precision.
-This is most obvious with the highly grammaticalised physical measure nouns like \textit{centilitre} which are very prototypically used with exact numeral instead of unspecific quantifiers, although both options are available (\textit{three centilitres} vs.\ \textit{several centilitres}).
+This is most obviously so with the highly grammaticalised physical measure nouns like \textit{centilitre} which are very prototypically used with exact numeral instead of unspecific quantifiers, although both options are available (\textit{three centilitres} vs.\ \textit{several centilitres}).
 Since the \NACa\ is more closely associated with the \textit{quantity-of} relation, cardinals as attributes of the measure noun are expected to have a higher proportional frequency in the \NACa.
@@
 Especially with abstract physical measure nouns (like \textit{centilitre}), exact numerical quantification is invited.
+By hpothesis, however, this goes beyond a selection effect tied to measure lemmas and cardinal quantifiers are expected to co-occur relatively more often with the \NACa.
 The statistical model presented in Section~\ref{sec:corpusstudies} will be specified in a way that it can detect such a preference.
@@
 
-For the present study, I used the German \textit{Corpus from the Web} (COW) in its 2014 version DECOW14A (\citealp{SchaeferBildhauer2012full,Schaefer2015b} and \citealp{BiemannEa2013,SchaeferBildhauer2013} for overviews of web corpora in general and the methodology of their construction), which contains almost 21 billion tokens.%
+For the present study, I used the German \textit{Corpus from the Web} (COW) in its 2014 version DECOW14A (\citealp{SchaeferBildhauer2012full}, and \citealp{Schaefer2015b}, as well as \citealp{BiemannEa2013}, and \citealp{SchaeferBildhauer2013}, for overviews of web corpora in general and the methodology of their construction), which contains almost 21 billion tokens.%
 \footnote{The COW corpora (Dutch, English, French, German, Spanish, Swedish) are made available for free at \url{https://www.webcorpora.org}.
@@
 Web corpora, on the other hand, contain at least some amount of non-standard language from forums and similar sources.
-For these or similar reasons, COW corpora have been used in a number of peer-reviewed publications, for example \cite{VanGoethemHiligsmann2014,VanGoethemHuening2015,MuellerS2014,Schaefer2016c,SchaeferSayatz2014,SchaeferSayatz2016,Zimmer2015}.
+For these or similar reasons, COW corpora have been used in a number of peer-reviewed publications, for example \cite{VanGoethemHiligsmann2014}, \cite{VanGoethemHuening2015}, \cite{MuellerS2014}, \cite{Schaefer2016c}, \cite{SchaeferSayatz2014}, \cite{SchaeferSayatz2016}, and \cite{Zimmer2015}.
 Therefore, DECOW is the obvious choice for this study.
@@
 As stated in Section~\ref{sec:germanmeasurenps}, the \NACa\ is virtually never used with plural kind nouns, and this sample was not used except for quantifying the frequency of occurrence of the constructions (67 times \NACa\ and 794 times \PGCa).
-The sample is distributed in the data package accompanying this paper, however.
+However, the sample is distributed in the data package accompanying this paper.
 }
@@
 First, she claims that ``while frequentist statistics only allows one to test whether the null hypothesis can be rejected, Bayesian statistics enables one both to test the null hypothesis and to estimate the probability of specific parameter values given the data.''
-This does not do justice to frequentist methods in that mere rejection of the null hypothesis is characteristic only of Fisher's approach.
+This does not do justice to frequentist methods in that a strong focus on the rejection of the null hypothesis is characteristic only of Fisher's approach.
 In the Neyman-Pearson approach, results ideally \textit{favour} the main hypothesis vis-à-vis the alternative hypothesis (cf.\ \citealp{Lehmann1993,Lehmann2011,Perezgonzalez2015}).
@@
 Most notably, priors for coefficients were specified as $\mathcal{N}(0,10)$ because coefficients higher than 10 or lower than $-10$ are extremely rare in well-specified models on appropriate data.%
-\footnote{Consider that with a coefficient of 10, each increase by 1 in the regressor variable means an increase in odds of $exp(10)=22,026.47$.
+\footnote{Consider that with a coefficient of 10, each increase by 1 in the regressor variable increases the odds by $exp(10)=22,026.47$.
 To reliably estimate such coefficients, extremely large samples would be required.}
@@
 \cite{FordBresnan2013} use the \textit{split-100} task in which participants have to distribute 100 points between the alternatives, assigning more points to more natural sounding alternative.
-In essence, participants distribute a probability mass between two alternants, which is intended to produce more subtle results compared to a two-alternative forced-choice task \citep{Rosenbach2013}, where participants have to choose one of two alternants.
+In essence, participants distribute a probability mass between two alternants, which is intended to produce more subtle results compared to a two-alternative forced-choice task such as in \cite{Rosenbach2013}, where participants have to choose one of two alternants.
 The split-100 paradigm has been criticised in \cite{ArppeJaervikivi2007}.
@@
 
-The pairs of stimuli were the sentence containing the preferred construction (according to the corpus GLMM) and a modified version containing the dispreferred construction.
+The final pairs of stimuli were the sentence containing the attested and preferred alternant (according to the corpus GLMM) on the one hand and a modified version containing the dispreferred alternant on the other hand.
 They were presented next to each other and a 20 second time limit for each choice was set.%
@@
 (iii)~Given (ii), we expect results from corpus analyses and from appropriate experiments to yield similar results, not necessarily as a form of validation of the corpus-based findings, but as converging evidence.
-I consider these three points to be of utmost importance because they clearly set this brand of cognitive linguistics apart from \textit{both} Aristotelian frameworks of the generative flavour \textit{and} introspective, non-empirical, and anti-quantitative versions of cognitive linguistics (see \citealp{Dabrowska2016} for an impressive philippic against such approaches).
+I consider these three points to be of utmost importance because they clearly set this brand of cognitive linguistics apart from \textit{both} Aristotelian frameworks of the generative flavour \textit{and} introspective, non-empirical, and anti-quantitative versions of cognitive linguistics (see \citealp{Dabrowska2016} for a pithy philippic against such approaches).
 
@@
 \begin{acknowledgement}
-  I thank (in alphabetical order) Felix Bildhauer, Susanne Flach, Elizabeth Pankratz, Samuel Reichert, and Ulrike Sayatz for valuable discussions and comments.
+  I thank (in alphabetical order) Felix Bildhauer, Susanne Flach, Elizabeth Pankratz, Samuel Reichert, Ulrike Sayatz, and Christian Zimmer for valuable discussions and comments.
   Elizabeth thankfully also fixed my English.

### 2017-05-10 Latex/article.tex
Minor fixes, references.

@@
 \usepackage[firstpage]{draftwatermark}
-\SetWatermarkFontSize{40pt}
+\SetWatermarkFontSize{30pt}
 \SetWatermarkHorCenter{300pt}
-\SetWatermarkColor[rgb]{1,0.2,0.2}
-\SetWatermarkText{\textsf{Submitted April 2017.}}
+\SetWatermarkColor[rgb]{0.9,0.5,0.5}
+\SetWatermarkText{\textsf{Submitted April 2017. Do not quote.}}
 
@@
 The criticism was reiterated in \cite{DivjakEa2016}, where they use a forced-choice task.
-In \cite{VerhoevenTemme2017}, it was shown that results from forced-choice and split-100 experiments mostly converge, which is expected in the long run if a probability is turned into a binary decision.
+In \cite{VerhoevenTemme2017}, it was shown that results from forced-choice and split-100 experiments mostly converge (with some numeric intricacies related to the non-linear distribution of preferences for alternants).
 I present a forced-choice experiment in this Section and not a split-100 experiment, mainly because in a dry run of the experiment, participants complained about the unnaturalness of distributing a probability mass across two alternants.
@@
 
-There were 24 participants (native speakers of German without reading or writing disabilities) aged 19 to 30 living permanently in [name of city redacted], who were recruited from introductory linguistics courses at [name of university redacted].
+There were 24 participants (native speakers of German without reading or writing disabilities) aged 19 to 30 living permanently in Berlin, who were recruited from introductory linguistics courses at Freie Universität Berlin.
 Although the experiment was conducted in the last four weeks of their first semester, participants had no deeper explicit knowledge of linguistics, grammar, or experimental methods.

### 2017-12-19 Latex/article.tex
Beginnig revisison (R3).

@@
 
-\usepackage[firstpage]{draftwatermark}
-\SetWatermarkFontSize{30pt}
-\SetWatermarkHorCenter{300pt}
-\SetWatermarkColor[rgb]{0.9,0.5,0.5}
-\SetWatermarkText{\textsf{Submitted April 2017. Do not quote.}}
+% \usepackage[firstpage]{draftwatermark}
+% \SetWatermarkFontSize{30pt}
+% \SetWatermarkHorCenter{300pt}
+% \SetWatermarkColor[rgb]{0.9,0.5,0.5}
+% \SetWatermarkText{\textsf{Submitted April 2017. Do not quote.}}
 
@@
 For such probabilistic generalisations estimated on corpus data, there has always been an interest in correlating the findings with results from experimental work (for example, \citealp{ArppeJaervikivi2007,BresnanEa2007,BresnanFord2010,DivjakGries2008,DivjakEa2016,FordBresnan2013}).
-This often called a \textit{validation} of the corpus-derived findings, but \citet[303]{Divjak2016a} rightly criticises this choice of words ``because it creates the impression that behavioral experimental data is inherently more valuable than textual data,'' citing \cite{TummersEa2005}, who state that a corpus is ``a sample of spontaneous language use that is (generally) realized by native speakers.''
+This is often called a \textit{validation} of the corpus-derived findings, but \citet[303]{Divjak2016a} rightly criticises this choice of words ``because it creates the impression that behavioral experimental data is inherently more valuable than textual data,'' citing \cite{TummersEa2005}, who state that a corpus is ``a sample of spontaneous language use that is (generally) realized by native speakers.''
 However, as \citet[486--487]{Dabrowska2016} convincingly argues, this does not imply that we can in some way ``deduce mental representations from patterns of use,'' \ie\ from corpus data.%
@@
 The present study is conducted within the general paradigm of cognitive corpus linguistics and includes a comparison of the corpus findings with results from two experiments.
-As a theoretical framework, I assume a model of similarity-based classification in the form of Prototype Theory.
+I assume a model of similarity-based classification in the form of Prototype Theory for modeling alternation.
 Protoype Theory has been used in alternation research in cognitively oriented corpus linguistics (see \citealp{DivjakArppe2013}, \citealp{Gries2003}).
@@
 Then the corpus study is presented, including an appropriate statistical analysis (Section~\ref{sec:corpusstudies}).
-Two experiments that correlate results from the corpus study with experimental results are then reported (Section~\ref{sec:experimental}) before I sum up the paper in Section~\ref{sec:conclusion}.
+Two experiments which corroborate the corpus-based findings are then reported in Section~\ref{sec:experimental} before I sum up the paper in Section~\ref{sec:conclusion}.
 
@@
 
-I use the term \textit{measure noun phrase} (MNP) to refer a noun phrase (NP) in which a kind-denoting (count or mass) noun depends on another noun that specifies a quantity of the objects or the substance denoted by the kind-denoting noun.
+I use the term \textit{measure noun phrase} (MNP) to refer to a noun phrase (NP) in which a kind-denoting (count or mass) noun depends on another noun that specifies a quantity of the objects or the substance denoted by the kind-denoting noun.
 I call the kind-denoting noun the \textit{kind noun} and the quantity-denoting noun the \textit{measure noun}.
@@
   \end{tabular}
-  \caption{Distribution of the NAC and PGC constructions in different NP structures with examples and references to full example sentences}
+  \caption{NAC and PGC constructions in different NP structures with examples and references to full example sentences}
   \label{tab:constructions}
@@
 In principle, this paper is about the middle column of Table~\ref{tab:constructions}, \ie\ the syntactic configuration in which two different case patterns are acceptable.
-However, in Section~\ref{sec:analyses}, the outer columns (\NACb\ and \PGCd) will still play a major role when the factors controlling the alternation are named.
+However, in Section~\ref{sec:analyses}, the outer columns (\NACb\ and \PGCd) will still play a major role when the factors controlling the alternation are discussed.
 
@@
 They are typical measure nouns like \textit{Gramm} `gram', \textit{Pfund} `pound' or \textit{Prozent} `percent', which have no normal plural forms at all.%
-\footnote{Plurals like \textit{Pfunde} and \textit{Prozente} have special meanings and have very restricted uses, mostly in idiomatic expressions and prefabs such as \textit{Pfunde verlieren} `lose pounds.'
+\footnote{Plurals like \textit{Pfunde} and \textit{Prozente} have special meanings and have very restricted uses, mostly in idiomatic expressions and prefabs such as \textit{Pfunde verlieren} `lose pounds' and \textit{Prozente machen} `make a profit'.
 They cannot be used in normal MNPs.}
@@
 They are consequently incompatible with measure nouns denoting natural portions such as \textit{Schluck} `gulp' or \textit{Haufen} `heap' and strongly grammaticalised nouns such as \textit{Gramm} `gram'.
-This disqualifies them as alternatives to the \NACa\ or \PGCa\ in most contexts and they are thus decidedly \textit{not} more ways of saying the same thing. 
+This disqualifies them as alternatives to the \NACa\ or \PGCa\ in most contexts and they are thus clearly not just more ways of saying the same thing. 
 It is therefore reasonable to focus on the two well-defined alternants.
@@
 I have demonstrated that there is an alternation between two measure noun constructions in a narrow syntactic configuration (kind NP with an adjective but without a determiner), and that the two constructions differ in the case of the kind noun (case agreement with the measure noun or genitive).
-I turn to more theory-oriented discussion in the next section.
+I turn to more theory-oriented discussion of the alternation in the next section.
 
@@
 third, a semantic prototypicality effect of the degrees of grammaticalisation (a preference to occur with cardinals);
-fourth, preferences in different registers.
+fourth, preferences in different styles.
 
@@
 For example, in the dative (governed here by the preposition \textit{mit}), the strong suffix \textit{-em} is on the determiner in (\ref{ex:starkschwach:a}) and the adjective bears the weak suffix \textit{-en}.
-In (\ref{ex:starkschwach:b}), the strong suffix \textit{-em} appears on the adjective because there is no determiner.
+In (\ref{ex:starkschwach:b}), the strong suffix \textit{-em} appears on the adjective because there is no determiner.%
+\footnote{In the masculine and neuter singular genitive, the strong and weak forms are indistinguishable.
+Since the alternation itself does not occur in the genitive (see Section~\ref{sec:descriptive}), this does not create any complications for the present study.}
 
@@
 Nouns which occur proportionally more often in the \PGCd\ should favour the \PGCa, and nouns which occur proportionally more often in the \NACb\ should favour the \NACa.
-This basic attraction effect will be used in the corpus study Section~\ref{sec:corpusstudies}.
+This basic attraction effect will be quantified in the corpus study Section~\ref{sec:corpusstudies}.
 
@@
 
-In a probabilistic framework like Prototype Theory, the crucial question (beyond simple lemma preference effects) is, however, what controls speakers' decisions to use either alternant. 
-In the remainder of this section, I argue that the NAC and the PGC are prototypes associated with different degrees of the grammaticalisation of the measure noun and related morpho-syntactic properties, as well as register effects.
+In a probabilistic framework like Prototype Theory, the crucial question (beyond simple lemma preference effects) is, however, which factors control speakers' decisions to use either alternant. 
+In the remainder of this section, I argue that the NAC and the PGC are prototypes associated with different degrees of the grammaticalisation of the measure noun and related morpho-syntactic properties, as well as stylistic effects.
 The degree of similarity of a given instance to either of the two prototypes makes speakers chose the \NACa\ or the \PGCa.
@@
 In (\ref{ex:determiners:pgc}), the measure noun is modified by a non-cardinal determiner \textit{einige} `some', and the \PGCa\ is preferred.
-Especially with abstract physical measure nouns (like \textit{centilitre}), exact numerical quantification is invited.
-By hpothesis, however, this goes beyond a selection effect tied to measure lemmas and cardinal quantifiers are expected to co-occur relatively more often with the \NACa.
-The statistical model presented in Section~\ref{sec:corpusstudies} will be specified in a way that it can detect such a preference.
-
-Fourth and finally, it was found that the \PGCa\ is more typical of higher registers or even exclusive to written language (see \citealp[320--323]{Hentschel1993}).
+Especially with exact physical measure nouns (like \textit{centilitre}), exact numerical quantification is invited.
+By hypothesis, however, this goes beyond a selection effect tied to measure lemmas and cardinal quantifiers are expected to co-occur relatively more often with the \NACa.
+This hypothesis will be reflected in the specification of the statistical model presented in Section~\ref{sec:corpusstudies}.
+
+Fourth and finally, it was found that the \PGCa\ is more typical of higher stylistic levels (distinctly edited, closer to the non-regional standard, more formal) and\slash or even exclusive to written language (see \citealp[320--323]{Hentschel1993}).
 This is not surprising, given that the genitive -- an intrinsic part of the PGC -- is generally underrepresented in colloquial vernacular variants of German as a result of a diachronic process wherein many (but by no means all) uses of the genitive are replaced by other cases or periphrastic constructions \citep{FleischerSchallert2011}.
-Under an integral view of prototypes, which incorporate effects related to larger contexts and registers, such preferences can be part of what defines the construction prototypes.
-In the corpus study in Section~\ref{sec:corpusstudies}, register effects will therefore be modelled -- even if only using two very simple proxy variables.
-
-To summarise the discussion and core hypotheses, I assume that the alternation is controlled by the similarity of the chosen lemmas (including the degree of grammaticalisation of measure nouns), certain morpho-syntactic choices, as well as the larger utterance context to two prototypes instantiated more straightforwardly by the two non-alternating cases of NAC and PGC.
-Concretely:
+Under an integral view of prototypes, which incorporate effects related to larger contexts, registers, and styles, such preferences can be part of what defines the construction prototypes.
+In the corpus study in Section~\ref{sec:corpusstudies}, stylistic effects will therefore be modelled -- even if only using two very simple proxy variables.
+
+To summarise, I assume that the alternation is controlled by the following factors:
 
@@
   \item Measure nouns modified by cardinals favour the \NACa.
-  \item The \PGCa\ is associated with higher registers.
+  \item The \PGCa\ is associated with higher stylistic levels.
 \end{enumerate}
@@
 It contains just over one billion tokens.}
+Mass nouns were defined as concrete nouns which denote a substance in the broad sense, combine with uninflected mass quantifiers such as \textit{viel} `much' and \textit{wenig} `little' (\textit{viel Bier} `much beer'), and form only sortal and unit plurals (such as the plural \textit{Biere} `types of beer' or `glasses of beer').
 Abstract nouns which partially behave like mass nouns (like \textit{Spaß} `fun’ or \textit{Gefahr} `danger’) were excluded because they are usually not quantified in the same way as concrete mass nouns.
@@
 In order to reduce the sample size for the manual annotation process, the final concordance was sampled from the results of these 2,365 queries.
-Since the mass nouns in the sample were distributed according to the usual power law, I used all hits for nouns with a frequency up to 100 and a sample of 100 of all those with higher frequency.
+Since the mass nouns in the sample were distributed according to the usual power law (often referred to as a \textit{Zipfian} distribution), I used all hits for nouns with a frequency up to 100 and a sample of 100 of all those with higher frequency.
 The final sample contained 6,843 sentences, which was reduced to 5,063 in the manual annotation process due to removal of noisy material, erroneous hits and uninformative cases where the measure noun was in the genitive, in which case the \NACa\ cannot be distinguished from the \PGCa.
@@
 \textit{Portion} (natural portions like \textit{Happen} `bite' or \textit{Krümel} `crumb'; \textit{f=}713).
-The few lemmas that did not fit into either of these classes were labelled \textit{Rest} (\textit{f=}278).
+The lemmas that did not fit into either of these classes were labelled \textit{Rest} (\textit{f=}278).
 
@@
 
-To capture the influence of register or style mentioned in Section~\ref{sec:analyses}, two proxy variables were used.
+To capture the influence of style mentioned in Section~\ref{sec:analyses}, two proxy variables were used.
 At the document level, the DECOW corpus has an annotation for \textit{Badness}.
@@
 Documents with higher Badness usually contain more incoherent language, shorter sentences, etc.
-If the \PGCa\ actually favours higher registers and styles, a high \textit{Badness} should be correlated with fewer occurrences.
+If the \PGCa\ actually favours higher stylistic levels, a high \textit{Badness} should be correlated with fewer occurrences.
 Documents in DECOW14 have also been annotated with a variable called \textit{Genitives}.
 The higher the values of this variable, the lower the proportion of genitives among all case-bearing forms is.
-A high number of genitives is indicative of higher registers.
+A high number of genitives is indicative of higher levels of style.
 However, the use of this variable as a regressor in the present study might be considered problematic.
@@
 Additionally, I want to voice the concern that many practitioners are probably already struggling with getting an adequate grasp of advanced statistical methods and that it might therefore be wise to use the more conservative and better understood method if the alternative method is not absolutely required for substantive reasons.
-In Section~\ref{sec:corpushierarchicalmodel}, I estimate the parameters of my hierarchical model with MLE and Markov-Chain Monte Carlo (MCMC) methods (the currently most prominent estimator used in Bayesian settings) to demonstrate their expectable convergence.
+In the next section, I estimate the parameters of my hierarchical model with MLE and Markov-Chain Monte Carlo (MCMC) methods (the currently most prominent estimator used in Bayesian settings) to demonstrate their expectable convergence.
 
@@
 All regressors from Table~\ref{tab:variables} were included, and the measure lemma and the kind noun lemma were specified as varying-intercept random effects.
-The sample size was \textit{n=}5,063 with 1,1344 cases of \PGCa\ and 3,929 cases of \NACa.
+The sample size was \textit{n=}5,063 with 1,134 cases of \PGCa\ and 3,929 cases of \NACa.
 The results of the estimation are shown in Table~\ref{tab:bigtable} and in Figure~\ref{fig:fixeffs}.
@@
 
-The register-related proxy variables point to the expected direction.
+The style-related proxy variables point to the expected direction.
 Increased \textit{Badness} of the document favours the \NACa\ ($\beta_{\text{Badness}}=-0.152$, $\mpPB=0.002$) and so does a lower density of genitives ($\beta=-0.693$, $\mpPB=0.001$).
-While these are poor proxies to register (and partially circular in the case of \textit{Genitives}), this result can at least encourage future work into register effects.
+While these are merely proxies to style (and partially circular in the case of \textit{Genitives}), this result can at least encourage future work into stylistic effects.
 
@@
 
-\bibliography{rs,cow}
+\bibliography{rs}
 \end{document}

### 2017-12-19 Latex/article.tex
More comments from reviewers implemented.

@@
 I assume a model of similarity-based classification in the form of Prototype Theory for modeling alternation.
-Protoype Theory has been used in alternation research in cognitively oriented corpus linguistics (see \citealp{DivjakArppe2013}, \citealp{Gries2003}).
+Protoype Theory has been used in alternation research in cognitively oriented corpus linguistics (see \citealp{DivjakArppe2013}; \citealp{Gries2003}).
 It is assumed that the variables annotated in a corpus study on an alternation phenomenon define prototypes for the alternants.
@@
 It still has the potential to form a true partitive (if the kind noun is definite).
-Conversely, the NAC lacks this ability to form true partitives, and it thus forms the more prototypical environment for expressing measurement (as opposed to partitivity).
-Hence, we can expect the \NACa\ to be a prototypical hosting construction for more strongly grammaticalised measure nouns.
+Conversely, the NAC lacks this ability to form true partitives, and it thus forms the more typical environment for expressing measurement (as opposed to partitivity).
+Hence, we can expect the \NACa\ to be a typical hosting construction for more strongly grammaticalised measure nouns.
 For example, highly grammaticalised non-referential nouns like \textit{Gramm} `gram' and \textit{Meter} `metre' should occur proportionally more often in the \NACa\ than in the \PGCa.
@@
 Both types of relations can be numerically quantified -- inasmuch as a precise number of \textit{parts} or a numerically exact \textit{quantity} can be specified.
-However, it is much more prototypical of quantities to be specified with numerical precision.
-This is most obviously so with the highly grammaticalised physical measure nouns like \textit{centilitre} which are very prototypically used with exact numeral instead of unspecific quantifiers, although both options are available (\textit{three centilitres} vs.\ \textit{several centilitres}).
+However, it is much more typical of quantities to be specified with numerical precision.
+This is most obviously so with the highly grammaticalised physical measure nouns like \textit{centilitre} which are very typically used with exact numeral instead of unspecific quantifiers, although both options are available (\textit{three centilitres} vs.\ \textit{several centilitres}).
 Since the \NACa\ is more closely associated with the \textit{quantity-of} relation, cardinals as attributes of the measure noun are expected to have a higher proportional frequency in the \NACa.
@@
 Fourth and finally, it was found that the \PGCa\ is more typical of higher stylistic levels (distinctly edited, closer to the non-regional standard, more formal) and\slash or even exclusive to written language (see \citealp[320--323]{Hentschel1993}).
-This is not surprising, given that the genitive -- an intrinsic part of the PGC -- is generally underrepresented in colloquial vernacular variants of German as a result of a diachronic process wherein many (but by no means all) uses of the genitive are replaced by other cases or periphrastic constructions \citep{FleischerSchallert2011}.
+This is not surprising, given that the genitive -- an intrinsic part of the PGC -- is generally underrepresented in colloquial vernacular variants of German as a result of a diachronic process wherein many (but by no means all) uses of the genitive are being replaced by other cases or periphrastic constructions \citep{FleischerSchallert2011}.
 Under an integral view of prototypes, which incorporate effects related to larger contexts, registers, and styles, such preferences can be part of what defines the construction prototypes.
@@
 Due to the noisy nature and unknown composition of the web, only carefully designed and established web corpora like the COW corpora or the SketchEngine corpora \citep{KilgarriffEa2014} should be used.
-Clearly, using search engine results is ``bad science'' for many reasons, most prominently total irreproducibility of results, as \cite{Kilgarriff2006} pointed out more than ten years ago.
+Clearly, using search engine results is ``bad science'' for many reasons, most prominently total non-replicability of results, as \cite{Kilgarriff2006} pointed out more than ten years ago.
 Careless use of search engine results is still found, however, see for example \citet[171--175]{DeclerckBrems2016}.}
 First, the external validity of any study is increased through a higher heterogeneity of the sample \citep[30]{MaxwellDelaney2004}, and the DECOW corpus has clearly a much more heterogeneous composition compared to the only other very large corpus of German, the DeReKo \citep{KupietzEa2010} of the Institute for the German Language (IDS), which contains almost exclusively newspaper texts.%
-\footnote{It was shown in \cite{W16-2601} that, for example, the spread of topics is much smaller in DeReKo compared to DECOW.}
+\footnote{It was shown in \cite{W16-2601} that, for example, the range of topics covered is much smaller in DeReKo compared to DECOW.}
 Second, it was already mentioned that normative grammars often adopt clear positions regarding the grammaticality of either the \NACa\ or the \PGCa.
@@
 For these or similar reasons, COW corpora have been used in a number of peer-reviewed publications, for example \cite{VanGoethemHiligsmann2014}, \cite{VanGoethemHuening2015}, \cite{MuellerS2014}, \cite{Schaefer2016c}, \cite{SchaeferSayatz2014}, \cite{SchaeferSayatz2016}, and \cite{Zimmer2015}.
-Therefore, DECOW is the obvious choice for this study.
+Therefore, DECOW is a valid choice for this study.
 
@@
 
-This resulting list of mass nouns was used in step (\ref{enum:data:step2}) to derive a list of measure nouns co-occurring with the mass nouns. 
+This list of mass nouns was used in step (\ref{enum:data:step2}) to derive a list of measure nouns co-occurring with the mass nouns. 
 In order to generate this list, I utilised the fact that a direct sequence of two nouns almost always instantiates the bare-noun NAC if the second noun is a mass noun.
@@
 As stated in Section~\ref{sec:germanmeasurenps}, the \NACa\ is virtually never used with plural kind nouns, and this sample was not used except for quantifying the frequency of occurrence of the constructions (67 times \NACa\ and 794 times \PGCa).
-However, the sample is distributed in the data package accompanying this paper.
+However, the sample is distributed with the data package accompanying this paper.
 }
@@
 
-In this section, I justify the choice of statistical models which I use in Section~\ref{sec:corpushierarchicalmodel}.
+In this section, I justify my choice of statistical models used in Section~\ref{sec:corpushierarchicalmodel}.
 Readers might think that the method used here for modeling grammatical alternations -- namely (Hierarchical) Logistic Regression\slash Generalised Linear (Mixed) Models or GL(M)Ms -- does not require much justification.
@@
 Over the past few years, however, modified or alternative methods have been proposed.
-From among these methods, I just make a few remarks on Bayesian estimation (see \citealp{GelmanEa2014}) as it was proposed in \cite{Levshina2016} and \cite{Divjak2016a}, for example.
+While it is impossible to discuss all of these methods here, I just make a few remarks on Bayesian estimation (see \citealp{GelmanEa2014}) as it was proposed in \cite{Levshina2016} and \cite{Divjak2016a}, for example.
 Conceptually, I see three points of discussion that should be kept apart.
@@
   \includegraphics[width=0.85\textwidth]{../R/output/corpus_fixeffs_mle+mcmc}
-  \caption{Coefficients (MLE and MCMC) with 95\% confidence intervals (for details see text); the intercept (\textit{Cardinal=Yes}, \textit{Measurecase=Nom}, \textit{Kindgender=Masc}, \textit{Measureclass=Physical}; 0 for all numeric z-transformed regressors) is -4.328 (MLE) and -4.441 (MCMC)}
+  \caption{Coefficients (MLE and MCMC) with 95\% confidence intervals (for details see text); the intercept is -4.328 (MLE) and -4.441 (MCMC)}
   \label{fig:fixeffs}
@@
 The results of the estimation are shown in Table~\ref{tab:bigtable} and in Figure~\ref{fig:fixeffs}.
+The intercept comprises \textit{Cardinal=Yes}, \textit{Measurecase=Nom}, \textit{Kindgender=Masc}, \textit{Measureclass=Physical}, and 0 for all numeric z-transformed regressors.
+It is -4.328 for the ML estimate and -4.441 for the MCMC estimate.
+
 The regressors with the measure lemma as their unit of reference have no within-measure lemma variance, and the \textit{glmer} function automatically estimates them as \textit{group level predictors} (or \textit{second-level effects}), cf.\ \citet[265--269,302--304]{GelmanHill2006}.
@@
   \begin{tabular}{llrlp{0.5em}rrp{0.5em}rrp{0.5em}rrp{0.5em}cc}
-      Level & Regressor & \multicolumn{1}{l}{\pPB} & Level  && \multicolumn{2}{l}{Coefficient} && \multicolumn{2}{l}{CI low} && \multicolumn{2}{l}{CI high} && \multicolumn{2}{l}{CI excludes 0} \\
-              &                   &       &           && \multicolumn{1}{l}{MLE} & \multicolumn{1}{l}{MCMC} && \multicolumn{1}{l}{MLE} & \multicolumn{1}{l}{MCMC} && \multicolumn{1}{l}{MLE} & \multicolumn{1}{l}{MCMC} && \multicolumn{1}{l}{MLE} & \multicolumn{1}{l}{MCMC} \\\midrule
+      Model & Regressor & \multicolumn{1}{l}{\pPB} & Factor && \multicolumn{2}{l}{Coefficient} && \multicolumn{2}{l}{CI low} && \multicolumn{2}{l}{CI high} && \multicolumn{2}{l}{CI excludes 0} \\
+      level &                   &       & level     && \multicolumn{1}{l}{MLE} & \multicolumn{1}{l}{MCMC} && \multicolumn{1}{l}{MLE} & \multicolumn{1}{l}{MCMC} && \multicolumn{1}{l}{MLE} & \multicolumn{1}{l}{MCMC} && \multicolumn{1}{l}{MLE} & \multicolumn{1}{l}{MCMC} \\\midrule
        First     & Badness           &  0.002 &           && -0.152 & -0.155 && -0.247 & -0.247 && -0.061 & -0.065 && * & * \\
@@
   }
-  \caption{Coefficient table comparing Maximum Likelihood Estimation (MLE, with 95\% bootstrap confidence interval) and `Bayesian' Markov-Chain Monte Carlo estimation (MCMC); the intercept (\textit{Cardinal=Yes}, \textit{Measurecase=Nom}, \textit{Kindgender=Masc}, \textit{Measureclass=Physical}; 0 for all numeric z-transformed regressors) is -3.548 (MLE) and -3.700 (MCMC)}
+  \caption{Coefficient table comparing Maximum Likelihood Estimation (MLE, with 95\% bootstrap confidence interval) and `Bayesian' Markov-Chain Monte Carlo estimation (MCMC); the intercept is -4.328 (MLE) and -4.441 (MCMC)}
   \label{tab:bigtable}
@@
 Instead of inexact Wald approximations and Likelihood Ratio Tests, I used a drop-in bootstrap replacement for the Likelihood Ratio Test from the function \textit{PBmodcomp} from the \textit{pbkrtest} package \citep{HalekohHojsgaard2014}.
-I call the corresponding value $p_{\text{PB}}$\ and it is given in the appropriately labelled columns in Table~\ref{tab:bigtable}.
+I call the corresponding value $p_{\text{PB}}$\ and it is given in the respective columns in Table~\ref{tab:bigtable}.
 Only \textit{Kindfreq} ($\mpPB=0.095$) can be seen as slightly too high to be convincing (non-significant).
@@
   \centering
-  \includegraphics[width=0.75\textwidth]{../R/output/corpus_Measureclass}
+  \includegraphics[width=0.5\textwidth]{../R/output/corpus_Measureclass}
   \caption{Effect plot for the regressor \textit{Measureclass}}
@@
 
-In Section~\ref{sec:analyses}, it was also hypothesised that classes of measure nouns with a higher degree of grammaticalisation should favour the \PGCa.
+In Section~\ref{sec:analyses}, it was also hypothesised that classes of measure nouns with a higher degree of grammaticalisation should favour the \NACa.
 The \textit{Measureclass} second-level predictor was successfully estimated ($\mpPB=0.001$).
@@
   \centering
-  \includegraphics[width=0.75\textwidth]{../R/output/corpus_Cardinal}
+  \includegraphics[width=0.5\textwidth]{../R/output/corpus_Cardinal}
   \caption{Effect plot for the regressor \textit{Cardinal}}
@@
 
-In the two experiments reported in this section, I use probabilities for the alternating constructions calculated for attested material and I correlate these probabilities with the participants' reactions.
+In the two experiments reported in this section, I use probabilities for the alternating constructions calculated for attested material, and I correlate these probabilities with the participants' reactions.
 Thus, a direct link can be established between output material found in corpora and the behaviour of linguistic agents (see also Section~\ref{sec:cogocl}).
@@
 Participants had to choose between two sentences that differed only in that one contained the \NACa\ and the other contained the \PGCa.
-The analysis compares the probabilities assigned to stimuli by the corpus-derived model with the frequency with which participants chose the alternants for the same stimuli.
+The analysis compares the probabilities assigned to the stimuli by the corpus-derived model with the frequency with which participants chose the alternants for the same stimuli.
 
@@
 \footnote{Remember from Section~\ref{sec:corpusstudies} that the model predicts the probability that the \PGCa\ is chosen over the \NACa.}
-High and low probability were defined as the top and bottom 20\% of all probabilities assigned by the GLMM.
+High and low probabilities were defined as the top and bottom 20\% of all probabilities assigned by the GLMM.
 Lemmas and feature combinations were made unique within each group whenever possible.
@@
 \centering
-\includegraphics[width=0.75\textwidth]{../R/output/fc_effects}
+\includegraphics[width=0.5\textwidth]{../R/output/fc_effects}
 \caption{Effect plot for the multilevel logistic regression in the forced-choice experiment: predictability of participants' choices using the probabilities derived from the corpus-based GLMM}
@@
 \centering
-\includegraphics[width=0.75\textwidth]{../R/output/spr_effects}
+\includegraphics[width=0.5\textwidth]{../R/output/spr_effects}
 \caption{Effect plot for the LMM in the self-paced reading experiment: modeling participants' residualised log reading times on the probabilities given by the corpus-based GLMM}
@@
 (iii)~Given (ii), we expect results from corpus analyses and from appropriate experiments to yield similar results, not necessarily as a form of validation of the corpus-based findings, but as converging evidence.
-I consider these three points to be of utmost importance because they clearly set this brand of cognitive linguistics apart from \textit{both} Aristotelian frameworks of the generative flavour \textit{and} introspective, non-empirical, and anti-quantitative versions of cognitive linguistics (see \citealp{Dabrowska2016} for a pithy philippic against such approaches).
+I consider these three points to be of utmost importance because they clearly set this approach to linguistic research apart from \textit{both} Aristotelian frameworks \textit{and} introspective, non-empirical, and anti-quantitative versions of cognitive linguistics (see \citealp{Dabrowska2016} for a pithy philippic against such approaches).
 
@@
 A grammatical alternation in German measure NPs was examined using corpus data based on factors partly derived from existing accounts, formulated in terms of construction prototypes.
-The preferences extracted from the DECOW web corpus were confirmed in a forced-choice experiment, in which participants explicitly chose alternants in line with the corpus-derived model.
+The preferences extracted from the DECOW web corpus were confirmed in a forced-choice experiment, in which participants explicitly chose alternants in line with the probabilities derived from the corpus-based model.
 In a more implicit self-paced reading experiment, it was shown that the much rarer alternant brings about a reading time penalty except in cases for which the corpus model predicts very high probability for this alternant.

### 2017-12-21 Latex/article.tex
More reviewers' comments.

@@
-\documentclass[USenglish]{article}
+\documentclass[UKenglish]{article}
 
@@
 \usepackage{colortbl}
+\usepackage{soul}
 
@@
   \affil[1]{Freie Universität Berlin}
-  \title{Competing Constructions for German Measure Noun Phrases: from Usage Data to Experiment}
+  \title{Competing Constructions for German Measure Noun Phrases}
   \runningtitle{Competing Constructions for German Measure Noun Phrases}
@@
 However, in terms of the scientific method, we are most likely talking about \textit{induction} rather than \textit{deduction}.}
-It would be highly surprising if this were possible and the same holds for experimental methods.
+It would be highly surprising if this were possible, and the same holds for experimental methods.
 Nobody assumes that we can inductively infer mental representations from experiments, which -- as opposed to corpus studies -- even allow for direct access to the cognitive agent and offer much better possibilities to control experimental conditions and nuisance variables.
@@
 In this section, I introduce and illustrate the relevant alternating constructions.
-I describe the narrowly defined syntactic configuration in which the alternation occurs and I motivate the focus on \textit{only} this narrow range rather than, for example, the whole range of nominal constructions expressing quantities.
+I describe the narrowly defined syntactic configuration in which the alternation occurs, and I motivate the focus on \textit{only} this narrow range rather than, for example, the whole range of nominal constructions expressing quantities.
 
@@
 
-In the case at hand, three different syntactic configurations need to be distinguished with regard to case assignment inside German MNPs.
-If the kind noun forms an NP with a determiner, the construction resembles (and is usually called) a \textit{pseudo-partitive} (on partitives and pseudo-partitives see, e.\,g., \citealp{Barker1998,Selkirk1977,Stickney2007,Vos1999}; for a recent application of the terminology to German, see \citealp{Gerstenberger2015}).%
+\subsubsection{Core structures related to the alternation}
+
+Three different syntactic configurations within MNPs need to be distinguished, and the case alternation occurs only in one of them.
+It occurs only when the kind noun is modified by an attributive adjective and there is no determiner, as in (\ref{ex:intro:alternation}).
+Superficially, the sentences are functionally and semantically equivalent either with the kind noun in the genitive (\ref{ex:intro:alternation1}) or in the same case as the measure noun, an accusative in the case of (\ref{ex:intro:alternation2}).%
+\footnote{Some descriptive and normative grammars take stronger positions with regard to the acceptability of the two options.
+See \cite{Hentschel1993} and \cite{Zimmer2015} for analyses of the sometimes absurd stances taken in grammars of German.
+As will be shown (especially in Section~\ref{sec:exp:fc}), there might be preferences, but we cannot assume either construction to be unacceptable.
+}
+
+\begin{exe}
+  \ex\label{ex:intro:alternation}
+  \begin{xlist}
+    \ex[ ]{\label{ex:intro:alternation1} \gll Wir trinken [[ein Glas]\Sub{Acc} [guten Weins]\Sub{Gen}]\Sub{Acc}.\\
+    we drink a glass good wine \\
+    \glt We drink a glass of good wine.}
+    \ex[ ]{\label{ex:intro:alternation2} Wir trinken [[ein Glas]\Sub{Acc} [guten Wein]\Sub{Acc}]\Sub{Acc}.}
+  \end{xlist}
+\end{exe}
+
+This specific configuration has to be seen in the context of two other configurations, to which I turn now.
+First, if the kind noun forms an NP with a determiner, the construction resembles (and is usually called) a \textit{pseudo-partitive} (on partitives and pseudo-partitives see, e.\,g., \citealp{Barker1998,Selkirk1977,Stickney2007,Vos1999}; for a recent application of the terminology to German, see \citealp{Gerstenberger2015}).%
 \footnote{If the kind noun is definite, the construction instantiates a true partitive.
@@
 
-If the kind noun is bare -- \ie\ if it comes neither with a determiner nor a modifying adjective -- it has to agree in case with the measure noun as in (\ref{ex:intro:narrowapposition1}), and the genitive as seen in the PGC is not acceptable, see (\ref{ex:intro:narrowapposition2}).
+Second, if the kind noun is bare -- \ie\ if it comes neither with a determiner nor a modifying adjective -- it is uninflected as in (\ref{ex:intro:narrowapposition1}), and the genitive as seen in the PGC is not acceptable, see (\ref{ex:intro:narrowapposition2}).
 
@@
   \begin{xlist}
-    \ex[ ]{\label{ex:intro:narrowapposition1}\gll Wir trinken [[ein Glas]\Sub{Acc} [Wein]\Sub{Acc}]\Sub{Acc}.\\
+    \ex[ ]{\label{ex:intro:narrowapposition1}\gll Wir trinken [[ein Glas]\Sub{Acc} [Wein]\Sub{?}]\Sub{Acc}.\\
     we drink a glass wine\\
@@
 For German, this terminology is not distinctive enough, which is why I use the terms NAC and PGC.}
-Notice that the unavailability of the genitive on the kind noun can be seen as following from a rather quirky constraint that genitive NPs in German require the presence of some strongly case-marked element (determiner or adjective) in addition to the head noun in order to be acceptable (\citealp{GallmannLindauer1994,Schachtl1989}; see also \citealp[160]{Eisenberg2013b}).
-%While only an accusative is shown in (\ref{ex:intro:narrowapposition1}), the kind noun obligatory agrees in case also with nominative and dative measure nouns.
-
-The alternation can be observed \textit{only when the kind noun occurs with an attributive adjective but without a determiner}, as in (\ref{ex:intro:alternation}), where both the PGC in (\ref{ex:intro:alternation1}) and the NAC in (\ref{ex:intro:alternation2}) are equally acceptable.
-They are by-and-large functionally and semantically equivalent.
-However, Section~\ref{sec:analyses} is devoted to developing hypotheses about subtle differences between them.%
-\footnote{Some descriptive and normative grammars take stronger positions with regard to the acceptability of the two options.
-See \cite{Hentschel1993} and \cite{Zimmer2015} for analyses of the sometimes absurd stances taken in grammars of German.
-As the usage and experimental data presented below (especially Section~\ref{sec:exp:fc}) should render it clear, there might be preferences, but we cannot assume either construction to be unacceptable.
-}
+Notice that the unavailability of the genitive on the kind noun follows independently from a constraint that genitive NPs in German require the presence of some strongly case-marked element (determiner or adjective) in addition to the head noun in order to be acceptable (\citealp{GallmannLindauer1994,Schachtl1989}; see also \citealp[160]{Eisenberg2013b}).
+
+It is difficult to determine whether the bare kind noun in the narrow apposition construction as in (\ref{ex:intro:narrowapposition1}) bears no case at all, a generic case, or agrees in case with the measure noun.%
+\footnote{I would like to thank one of the reviewers for pointing this out.}
+When there is an adjective as in (\ref{ex:intro:alternation2}), the embedded kind NP clearly agrees in case, but due to the overall absence of markers of case in the singular, bare nouns mostly show no indication of their case.
+The only nouns which do have case markers in the singular are the so-called weak nouns \citep{Koepcke1995,Schaefer2016c}, which have something like a non-nominative \textit{-en} marker in the singular.
+Unfortunately, there are no genuine mass nouns among the weak nouns.
+However, a few of them can be coerced into a mass noun, such as \textit{Hase} `rabbit' (meaning `rabbit meat').
+It then appears as if the uninflected form is preferred, but the inflected form is not excluded.
+In (\ref{ex:macabre1}), the clearly acceptable form \textit{Hase} can only be a nominative singular or caseless.
+In (\ref{ex:macabre2}), the form \textit{Hasen} could be an accusative, dative, or genitive.
 
 \begin{exe}
-  \ex\label{ex:intro:alternation}
+  \ex\label{ex:macabre}
   \begin{xlist}
-    \ex[ ]{\label{ex:intro:alternation1} \gll Wir trinken [[ein Glas]\Sub{Acc} [guten Weins]\Sub{Gen}]\Sub{Acc}.\\
-    we drink a glass good wine \\
-    \glt We drink a glass of good wine.}
-    \ex[ ]{\label{ex:intro:alternation2} Wir trinken [[ein Glas]\Sub{Acc} [guten Wein]\Sub{Acc}]\Sub{Acc}.}
+    \ex[ ]{\label{ex:macabre1}\gll Niemand will [ein Stück [Hase]\Sub{Nom\slash caseless}]\Sub{Acc} essen.\\
+    nobody wants a piece rabbit eat\\
+  \trans Nobody wants to eat a piece of rabbit.}
+    \ex[?]{\label{ex:macabre2}\gll Niemand will [ein Stück [Hasen]\Sub{Acc\slash Dat\slash Gen}]\Sub{Acc} essen.\\
+    nobody wants a piece rabbit eat\\}
   \end{xlist}
 \end{exe}
+
+Even a full unacceptability of (\ref{ex:macabre2}) would not be conclusive, however, as a possible aversion of speakers towards the case-marked form might be due to the fact that it is at least potentially also a genitive, in which case the constraint against bare genitive nouns would apply.
+With plural kind nouns, the obligatory marking of the dative plural with \textit{-en} might provide some clues.
+However, plural kind nouns do not behave like singular ones in measure phrases anyway, as will be argued in Section~\ref{sec:neighbouringcases}.
+Also, judgements vary between (\ref{ex:datpl1}) and (\ref{ex:datpl2}), and both are found in corpora.%
+\footnote{For example, the variant in (\ref{ex:datpl1}) with the lemmas \textit{Sack} and \textit{Apfel} occurs four times, and the one in (\ref{ex:datpl2}) twice in the very large DECOW corpus (see Section~\ref{sec:gettingdata}).
+Similarly, for [\textit{Kiste} [\textit{Äpfel}]\Sub{Nom\slash Acc\slash Gen}]\Sub{Dat} `box of apples' (no case identity), I find three examples, and [\textit{Kiste} [\textit{Äpfeln}]\Sub{Dat}]\Sub{Dat} (clearly marked case identity), I find six examples.
+}
+
+\begin{exe}
+  \ex\label{ex:datpl} 
+  \begin{xlist}
+  \ex{\label{ex:datpl1}\gll mit [zwei Säcken [Äpfel]\Sub{Nom\slash Acc\slash Gen}]\Sub{Dat}\\
+     with a sack apples\\
+    \trans with a sack of apples}
+    \ex{\label{ex:datpl2}\gll mit [zwei Säcken [Äpfeln]\Sub{Dat}]\Sub{Dat}\\
+    with a sack apples\\}
+  \end{xlist}
+\end{exe}
+
+Descriptive grammars seem to favour an analysis in terms of caselessness (for example, \citealp[1981]{ZifonunEa1997c}).
+The hazy picture of the case of bare kind NPs is most likely due to the fact that case is so rarely marked on German nouns, where case is marked mostly on determiners and to some degree adjectives.
+The uncertainty in the few cases where case can be marked (weak nouns in the singular and dative plurals) would thus be a direct consequence of the fact that the construction is not very specific with respect to the case of the kind noun.
 
@@
 
-The distribution of the case patterns in the NAC (case identity between the measure noun and the kind noun) and in the PGC (genitive on the kind noun, regardless of the case of the measure noun) depending on the structure of the kind NP is summarised in Table~\ref{tab:constructions}.
+To summarise, the case patterns in the NAC and in the PGC (depending on the structure of the kind NP) is given in Table~\ref{tab:constructions}.
 I call the narrow apposition construction with a bare kind noun the \NACb\ and the partitive genitive with a determiner in the kind NP the \PGCd.
@@
 
-I now turn to some more subtle issues related to the measure noun case alternation, namely:
-
-\begin{enumerate}[i.]
-  \item{\label{it:intro:idiot} alleged alternatives to the \NACa\ and the \PGCa,}
-%  \item{\label{it:intro:datsg} alternative \textit{weak} forms of dative singular neuter adjectives,}
-%  \item{\label{it:intro:femsg} case syncretism in feminine NPs,}
-  \item{\label{it:intro:plurl} similar constructions with plural/collective kind nouns,}
-  \item{\label{it:intro:noifl} grammaticalised non-inflected measure nouns, and}
-  \item{\label{it:intro:other} alternative constructions for expressing quantities.}
-\end{enumerate}
-
-\vspace{-1\baselineskip}
-
-First of all, (\ref{it:intro:idiot}) refers to claims found in some grammars that a generic nominative, accusative, and even dative on the kind noun are used instead of the genitive (\PGCa) or case agreement (\NACa).
+\subsubsection{Neighbouring cases}
+\label{sec:neighbouringcases}
+
+% namely:
+% 
+% \begin{enumerate}[i.]
+%   \item{\label{it:intro:idiot} alleged alternatives to the \NACa\ and the \PGCa,}
+% %  \item{\label{it:intro:datsg} alternative \textit{weak} forms of dative singular neuter adjectives,}
+% %  \item{\label{it:intro:femsg} case syncretism in feminine NPs,}
+%   \item{\label{it:intro:plurl} similar constructions with plural/collective kind nouns,}
+%   \item{\label{it:intro:noifl} grammaticalised non-inflected measure nouns, and}
+%   \item{\label{it:intro:other} alternative constructions for expressing quantities.}
+% \end{enumerate}
+%
+% \vspace{-1\baselineskip}
+
+I now turn to some more subtle issues related to the measure noun case alternation in order to delimit the scope of the study.
+First, there is a claim found in some grammars that a generic nominative, accusative, and even dative on the kind noun are used instead of the genitive (\PGCa) or case agreement (\NACa).
 Overviews can be found in \cite{Hentschel1993} and \cite{Zimmer2015}.
@@
 Also, in my corpus sample, they simply did not occur.
-Even if they are accepted by some speakers, their extremely low frequency makes it virtually impossible to study them using corpus linguistic methods (Section~\ref{sec:corpusstudies}) and I consequently ignore them.
+Even if they are accepted by some speakers, their extremely low frequency makes it virtually impossible to study them using corpus linguistic methods (Section~\ref{sec:corpusstudies}), and I consequently do not discuss them further.
 
@@
 
-Turning to (\ref{it:intro:plurl}), readers might have noticed that so far only MNPs denoting quantities of substances (mass kind nouns such as \textit{ein Glas roter Wein\slash roten Weines} `a glass of red wine') have been discussed.
-If the kind noun is a plural count noun as in \textit{ein Sack kleine Äpfel} (\NACa) or \textit{ein Sack kleiner Äpfel} (\PGCa) `a bag of small apples', a similar alternation between PGC and NAC can be observed.
+Second, we find that, if the kind noun is a plural count noun as in \textit{ein Sack kleine Äpfel} (\NACa) or \textit{ein Sack kleiner Äpfel} (\PGCa) `a bag of small apples', a similar alternation between PGC and NAC can be observed.
 In line with experimental results reported in \citet[15--16]{Zimmer2015}, I found that the PGC is so dominant with plural kind nouns (794 of 861 cases, or over 92\%, cf.\ Section~\ref{sec:corpusstudies}) that the alternation cannot be analysed in the same way as in the singular.
@@
 
-As for (\ref{it:intro:noifl}), some measure nouns have been grammaticalised in a way that they always appear in their non-inflected form.
+Third, some measure nouns have been grammaticalised in a way that they always appear in their non-inflected form.
 They are typical measure nouns like \textit{Gramm} `gram', \textit{Pfund} `pound' or \textit{Prozent} `percent', which have no normal plural forms at all.%
-\footnote{Plurals like \textit{Pfunde} and \textit{Prozente} have special meanings and have very restricted uses, mostly in idiomatic expressions and prefabs such as \textit{Pfunde verlieren} `lose pounds' and \textit{Prozente machen} `make a profit'.
+\footnote{Plurals like \textit{Pfunde} and \textit{Prozente} have special meanings and have very restricted uses, mostly in idiomatic expressions such as \textit{Pfunde verlieren} `lose pounds' and \textit{Prozente machen} `make a profit'.
 They cannot be used in normal MNPs.}
@@
 
-Finally, (\ref{it:intro:other}) suggests that there are alternative ways of expressing similar quantificational meanings.
+Finally, there are alternative ways of expressing similar quantificational meanings.
 In the variationist tradition, which is strongly influenced by Labovian sociolinguistics, the \textit{principle of accountability} would dictate that proper studies should examine a variationist \textit{variable}, \ie\ all different \textit{ways of saying the same thing} (\citealp{Labov1966}, \citealp{Labov1969}, for an overview see \citealp{Tagliamonte2012}).
-In the case at hand, the variable might be something like \textit{measuring quantities of substances and collections}.
+In the case at hand, the variable might be something like \textit{measurement of quantities of substances and collections}.
 I argue that it is fully justified to focus narrowly on \NACa\ and \PGCa\ with their well-defined morpho-syntactic properties, mostly because the alternative constructions are not used in the same range of contexts.
 Two major alternatives might be considered.
-First of all, the analytic (pseudo-)partitive with \textit{von} `of' is only available as an alternative to the PGC if the kind noun phrase contains a (definite or indefinite) determiner as in (\ref{ex:analyticalpartitive}).
+There is an analytic (pseudo-)partitive with \textit{von} `of'.
+It is only available as an alternative to the PGC if the kind noun phrase contains a (definite or indefinite) determiner as in (\ref{ex:analyticalpartitive}).
 
@@
 
-Second, constructions with \textit{voll} (\textit{von})\slash\textit{voller} `full (of)' and \textit{mit} `with' are available, see (\ref{ex:alternatives}).
+Additionally, we find constructions with \textit{voll} (\textit{von})\slash\textit{voller} `full (of)' and \textit{mit} `with' are available, see (\ref{ex:alternatives}).
 
@@
 These construction have very idiosyncratic properties.
-The construction with \textit{voller} is discussed in \cite{Zeldes2018} and the construction with \textit{mit} is discussed in \cite{Bhatt1990}.
-They are not semantically equivalent to the NAC and the PGC and they are only available for a small subset of measure nouns.
-All of these constructions are used only with measure nouns denoting containers, where the whole NP always refers to the container and never to the quantity contained in it.
-They are consequently incompatible with measure nouns denoting natural portions such as \textit{Schluck} `gulp' or \textit{Haufen} `heap' and strongly grammaticalised nouns such as \textit{Gramm} `gram'.
-This disqualifies them as alternatives to the \NACa\ or \PGCa\ in most contexts and they are thus clearly not just more ways of saying the same thing. 
+The construction with \textit{voller} is discussed in \cite{Zeldes2018} and the construction with \textit{mit} in \cite{Bhatt1990}.
+They are not semantically equivalent to the NAC and the PGC, and they are only available for a small subset of measure nouns.
+All of these constructions are used only with measure nouns denoting containers, where the whole NP typically refers to the container and never to the quantity contained in it.
+They are incompatible with measure nouns denoting natural portions such as \textit{Schluck} `gulp' or \textit{Haufen} `heap' and strongly grammaticalised nouns such as \textit{Gramm} `gram'.
+This disqualifies them as alternatives to the \NACa\ or \PGCa\ in most contexts, and they are thus clearly not just more ways of saying the same thing. 
 It is therefore reasonable to focus on the two well-defined alternants.
@@
 I also develop my own analysis and the appropriate hypotheses for the empirical studies presented in Sections~\ref{sec:corpusstudies} and \ref{sec:experimental}.
-I discuss four main aspects:
-first, the different syntactic structures of the alternants \NACa\ and \PGCa\ in relation to their syntactic prototypes \NACb\ and \PGCd;
-second, degrees of grammaticalisation of the measure noun associated with the two prototypes;
-third, a semantic prototypicality effect of the degrees of grammaticalisation (a preference to occur with cardinals);
-fourth, preferences in different styles.
-
-% MORPHO-SYNTAX AND PROTOTYPES
-
-First, my analysis is based on the idea that in the relevant syntactic structure [N\Sub{1}~[A~N\Sub{2}]\Sub{NP\Sub{2}}]\Sub{NP\Sub{1}} (as instantiated in the \NACa\ and the \PGCa), the adjective is morpho-syntactically ambiguous between a determiner and a modifying adjective.
-To show this, %the first question is which noun constitutes the head of the whole construction.
-%This was answered quite clearly by \cite{Loebel1986} already (see also \citealp[213]{Eschenbach1994}, and \citealp[16]{GallmannLindauer1994}).
-%Subject-verb agreement is always realised on the measure noun, and we can therefore assume that the measure noun is the head.
-%Furthermore, the internal structure of MNPs can only be understood if the strong\slash weak inflection patterns of adjectives are taken into account.
-the strong\slash weak inflection patterns of adjectives needs to be taken into account.
-In NPs with a strongly inflected determiner, attributive adjectives inflect according to the massively syncretic \textit{weak} pattern.
-%\footnote{The weak pattern has only two suffixes.
-%The structural cases (nominative and accusative) in the singular are marked with \textit{-e}, the oblique cases (dative and genitive) as well as the entire plural are marked with \textit{-en}.
-%The only exception to this rule is the masculine accusative singular, which is also marked with \textit{-en}.}
-If there is no determiner (as is the case in the alternating constructions), attributive adjectives inflect like determiners themselves.
-This is called the \textit{strong} inflectional pattern.
-For example, in the dative (governed here by the preposition \textit{mit}), the strong suffix \textit{-em} is on the determiner in (\ref{ex:starkschwach:a}) and the adjective bears the weak suffix \textit{-en}.
-In (\ref{ex:starkschwach:b}), the strong suffix \textit{-em} appears on the adjective because there is no determiner.%
-\footnote{In the masculine and neuter singular genitive, the strong and weak forms are indistinguishable.
-Since the alternation itself does not occur in the genitive (see Section~\ref{sec:descriptive}), this does not create any complications for the present study.}
-
-\begin{exe}
-  \ex\label{ex:starkschwach} 
-  \begin{xlist}
-    \ex{\label{ex:starkschwach:a}\gll mit ein\textbf{-em} stark\textbf{-en} Kaffee\\
-                                      with a strong coffee\\}
-    \ex{\label{ex:starkschwach:b}\gll mit stark\textbf{-em} Kaffee\\
-                                      with strong coffee\\}
-  \end{xlist}
-\end{exe}
-
-Thus, the adjectives in the \NACa\ and the \PGCa\ have properties of adjectives as well as determiners.
-On the one hand, they are lexical adjectives and function as attributive modifiers.
-On the other hand, they are inflected like determiners, and they are the leftmost element in the NP, which is typical of determiners.
-This unusual double nature of adjectives in NPs without determiners leads to a plausible probabilistic interpretation of the pattern shown in Table~\ref{tab:constructions}.
-Whenever speakers classify the adjective in the kind noun phrase more as a determiner, they have to use the \PGCa\ because, if there is a determiner, the PGC is the only option.
-When they classify the adjective more as an adjective, the kind NP has no determiner, and they have to use the \NACa.%
-\footnote{While the generative analysis presented in \cite{Bhatt1990} cannot properly deal with probabilistic effects, Bhatt comes close to this interpretation by analysing the kind NP in the PGC as a DP and in the NAC as an NP.}
-
-This morpho-syntactic ambiguity means that the \NACa\ is in fact a \NACb\ in disguise, and the \PGCa\ is a \PGCd\ in disguise.
-This should ideally be reflected in the following basic selection effect:
-It is known from frameworks like Collostructional Analysis \citep{GriesStefanowitsch2004} that lemmas are attracted with different strengths by constructions.
-If the \NACa\ is highly similar to the \NACb\ and the \PGCa\ is highly similar to the \PGCd, the probability with which individual noun lemmas appear in the alternating constructions should be predictable at least in part from the relative frequency with which the same lemmas are used in the non-alternating constructions.
-Nouns which occur proportionally more often in the \PGCd\ should favour the \PGCa, and nouns which occur proportionally more often in the \NACb\ should favour the \NACa.
-This basic attraction effect will be quantified in the corpus study Section~\ref{sec:corpusstudies}.
-
-% GRAMMATICALISATION
-
-In a probabilistic framework like Prototype Theory, the crucial question (beyond simple lemma preference effects) is, however, which factors control speakers' decisions to use either alternant. 
-In the remainder of this section, I argue that the NAC and the PGC are prototypes associated with different degrees of the grammaticalisation of the measure noun and related morpho-syntactic properties, as well as stylistic effects.
-The degree of similarity of a given instance to either of the two prototypes makes speakers chose the \NACa\ or the \PGCa.
-
-Second, therefore, I focus on grammaticalisation.
-It is often assumed that pseudo-partitives arise as a form of grammaticalised partitives (\eg\ \citealp[536--539]{Koptjevskaja2001} for Finnish and Estonian, \citealp[559]{Koptjevskaja2001} for European languages in general).
-%
-%\footnote{Notice that the description of German in \citet[547--549]{Koptjevskaja2001} is vague and partly incorrect, maybe because her main source \cite{Eschenbach1993} is not as clear as it could be (see also Footnote \ref{fn:eschenbash} above).
-%Her example (29b) from \citet[71]{Eschenbach1993}, repeated here verbatim as (\ref{ex:koptjewskaja}) is misclassified as a clear case of a dative on the measure noun and a genitive on the kind noun (hence, PGC).
-%
-%\begin{exe}
-%  \ex{\label{ex:koptjewskaja}\gll trotz einer Flasche gut-en Wein-s\\
-%  in.spite.of one-\textsc{fem.gen} bottle good-\textsc{masc.sg.gen} wine-\textsc{gen}\\
-%  \glt in spite of one bottle of good wine}
-%\end{exe}
-%
-%It could, however, equally well be a genitive on the measure noun because of the syncretisms discussed here on p.\ \pageref{page:femininesyncretism} and an ambiguous case requirement of the preposition \textit{troz}, hence an instance of NAC.
-%Then, she does not clearly distinguish acceptable from non-acceptable constructions (cf.\ Section \ref{sec:descriptive}) on her pp.\ 548 and 549, including a rather confusing \textit{???} rating of her (31b), which is clearly a non-acceptable construction of the type illustrated in (\ref{ex:analyticalpartitive:b}) above.
-%}
-%
-The grammaticalisation paths uncovered by \citet[esp.\ 526--530]{Koptjevskaja2001} are relevant for the case at hand.
-The grammaticalisation path can start out (in some languages) with constructions involving two referential nouns (not necessarily forming a single and contiguous NP) and a \textit{separative} meaning as in \textit{(cut) two slices from the cake} \citep[535]{Koptjevskaja2001}.
-The \textit{part-of} meaning of true partitives as in \textit{a slice of the cake} represents the first stage of a development wherein the measure noun can already lose some semantic content, when for example words like \textit{bite} are no longer necessarily interpreted as a piece literally bitten out of something.
-The pseudo-partitive stage finally instantiates a \textit{quantity-of} relation, potentially even leading to fully grammaticalised quantifiers such as \textit{a lot}.
-In German, the PGC is clearly the older construction \citep{Zimmer2015}.
-It still has the potential to form a true partitive (if the kind noun is definite).
-Conversely, the NAC lacks this ability to form true partitives, and it thus forms the more typical environment for expressing measurement (as opposed to partitivity).
-Hence, we can expect the \NACa\ to be a typical hosting construction for more strongly grammaticalised measure nouns.
-For example, highly grammaticalised non-referential nouns like \textit{Gramm} `gram' and \textit{Meter} `metre' should occur proportionally more often in the \NACa\ than in the \PGCa.
-% \footnote{In case studies like \cite{Brems2003,DeclerckBrems2016} syntactic evidence for grammaticalisation plays a prominent role.
-% In English -- \cite{Brems2003} argues based on a view from \cite{Langacker1991} -- strongly lexicalised and thus semantically bleached measure nouns turn from nouns into quantifiers, leading to a syntactic reanalysis of the NP, wherein the kind noun becomes the head.
-% The argument is supported by convincing syntactic evidence, for example subject-verb agreement.
-% As argued above, in German MNPs the measure noun always remains the syntactic head, and there is no similarly handy criterion to diagnose a high degree of grammaticalisation.}
-
-% CARDINAL QUANTIFIERS
-
-Third, as described above, the grammaticalisation path leads from NPs denoting individuated objects standing in a \textit{part-of} relation to a construction with a more diffuse \textit{quantity-of} relation.
-Both types of relations can be numerically quantified -- inasmuch as a precise number of \textit{parts} or a numerically exact \textit{quantity} can be specified.
-However, it is much more typical of quantities to be specified with numerical precision.
-This is most obviously so with the highly grammaticalised physical measure nouns like \textit{centilitre} which are very typically used with exact numeral instead of unspecific quantifiers, although both options are available (\textit{three centilitres} vs.\ \textit{several centilitres}).
-Since the \NACa\ is more closely associated with the \textit{quantity-of} relation, cardinals as attributes of the measure noun are expected to have a higher proportional frequency in the \NACa.
-For illustration, (\ref{ex:determiners}) shows the expected alternants under this hypothesis.
-
-\begin{exe}
-  \ex\label{ex:determiners} 
-  \begin{xlist}
-    \ex{\label{ex:determiners:nac}\gll [[Drei Centiliter]\Sub{Nom} [heißer Rum]\Sub{Nom}]\Sub{Nom} sind genug.\\
-                                       three centilitres hot rum are enough\\
-				     \glt Three centilitres of hot rum is enough.}
-    \ex{\label{ex:determiners:pgc}\gll [[Einige Centiliter]\Sub{Nom} [heißen Rums]\Sub{Gen}]\Sub{Nom} sind genug.\\
-                                       some   centilitres hot    rum  are   enough\\
-				     \glt A few centilitres of hot rum is enough.}
-  \end{xlist}
-\end{exe}
-
-In (\ref{ex:determiners:nac}), the measure noun is modified by a cardinal \textit{drei} `three', and hence the \NACa\ is preferred.
-In (\ref{ex:determiners:pgc}), the measure noun is modified by a non-cardinal determiner \textit{einige} `some', and the \PGCa\ is preferred.
-Especially with exact physical measure nouns (like \textit{centilitre}), exact numerical quantification is invited.
-By hypothesis, however, this goes beyond a selection effect tied to measure lemmas and cardinal quantifiers are expected to co-occur relatively more often with the \NACa.
-This hypothesis will be reflected in the specification of the statistical model presented in Section~\ref{sec:corpusstudies}.
-
-Fourth and finally, it was found that the \PGCa\ is more typical of higher stylistic levels (distinctly edited, closer to the non-regional standard, more formal) and\slash or even exclusive to written language (see \citealp[320--323]{Hentschel1993}).
-This is not surprising, given that the genitive -- an intrinsic part of the PGC -- is generally underrepresented in colloquial vernacular variants of German as a result of a diachronic process wherein many (but by no means all) uses of the genitive are being replaced by other cases or periphrastic constructions \citep{FleischerSchallert2011}.
-Under an integral view of prototypes, which incorporate effects related to larger contexts, registers, and styles, such preferences can be part of what defines the construction prototypes.
-In the corpus study in Section~\ref{sec:corpusstudies}, stylistic effects will therefore be modelled -- even if only using two very simple proxy variables.
-
-To summarise, I assume that the alternation is controlled by the following factors:
-
-\begin{enumerate}[i.]
-  \item The relative frequencies with which measure noun lemmas and kind noun lemmas appear in the prototypical (non-alternating) \PGCd\ and \NACb\ are predictive of the probability with which the alternating \PGCa\ and the \NACa\ are chosen.
-  \item (Classes of) more strongly grammaticalised measure nouns favour the \NACa.
-  \item Measure nouns modified by cardinals favour the \NACa.
-  \item The \PGCa\ is associated with higher stylistic levels.
-\end{enumerate}
-
-\vspace{-1\baselineskip}
-
-In the next section, I report the corpus study that was designed to test these hypotheses on usage data.
-The resulting models will then be compared to experimental methods in Section~\ref{sec:experimental}.
+
+.%
+\footnote{A reviewer mentioned that she or he had the impression that dialectal variation is also a factor influencing the alternation.
+This is definitely true as some dialects (such as Alemannic) tend to have no genitive at all.
+While this is an interesting aspect for further research, the main obstacle is that there are no corpora of German which are both large enough and annotated with reliable meta data about regional variants.
+In the annotation of the corpus study, documents obviously written in a regional variant were excluded.
+}
 
@@
   They cluster around 0 and they include values of $-\infty$.
-  Third, I tried using collexeme strength as a regressor (with smoothing to remedy the mathematical problems) and the results were unsatisfactory compared to the simple quotient used here.
+  Third, I tried using collexeme strength as a regressor (with smoothing to remedy the mathematical problems), and the results were unsatisfactory compared to the simple quotient used here.
 }
@@
 Instead of inexact Wald approximations and Likelihood Ratio Tests, I used a drop-in bootstrap replacement for the Likelihood Ratio Test from the function \textit{PBmodcomp} from the \textit{pbkrtest} package \citep{HalekohHojsgaard2014}.
-I call the corresponding value $p_{\text{PB}}$\ and it is given in the respective columns in Table~\ref{tab:bigtable}.
+I call the corresponding value $p_{\text{PB}}$, and it is given in the respective columns in Table~\ref{tab:bigtable}.
 Only \textit{Kindfreq} ($\mpPB=0.095$) can be seen as slightly too high to be convincing (non-significant).
@@
 The model was re-estimated using the \textit{stan\_glmer} function from the \textit{rstanarm} package, which provides an \textit{lme4}-compatible syntax for estimating common model types with the \textit{stan} software \citep{CarpenterEa2017}.
-The algorithm was run with 4 chains and 1,000 iterations and I used plausible default priors.
+The algorithm was run with 4 chains and 1,000 iterations, and I used plausible default priors.
 Most notably, priors for coefficients were specified as $\mathcal{N}(0,10)$ because coefficients higher than 10 or lower than $-10$ are extremely rare in well-specified models on appropriate data.%
@@
 To reliably estimate such coefficients, extremely large samples would be required.}
-The algorithm converged and for all coefficients, the $\hat{R}$ diagnostic was exactly 1.
+The algorithm converged, and for all coefficients, the $\hat{R}$ diagnostic was exactly 1.
 The resulting coefficients and intervals as well as an * for ``confidence interval does not contain 0'' are also given in Table~\ref{tab:bigtable} in the columns labelled \textit{MCMC}.
 Both methods lead to exactly the same results (minus negligible numerical differences) as expected given the modest complexity of the model structure and the large sample size.
-The signs and magnitudes of the coefficients are identical and confidence intervals have the same width and symmetry properties.
+The signs and magnitudes of the coefficients are identical, and the confidence intervals have the same width and symmetry properties.
 Figure~\ref{fig:fixeffs} illustrates this by also showing both estimates.
@@
 They show the changes in probability for the outcome (y-axis) dependent on values of a regressor (x-axis) at typical values of all other regressors.
-The vertical bars (categorical variables) and grey areas (continuous variables) are asymptotic 95\% confidence intervals calculated from \textit{glmer}.
-They are not bootstrapped.}
+The vertical bars (categorical variables), and the grey areas (continuous variables) are asymptotic 95\% confidence intervals calculated from \textit{glmer}.
+They are not bootstrapped.
+Readers should be aware that the axes are specifically scaled so as to result in a linear plot, and that the range of the axes varies between plots.}
 The effect is as expected:
@@
 With kind nouns, higher frequency seems to favour the \PGCa ($\beta_{\text{Kindfreq}}=0.146$, $\mpPB=0.095$).
-However, there is no clear theoretical interpretation (see Section~\ref{sec:analyses}) and the estimate is imprecise (not significant at $\alpha=0.05$, see above).
+However, there is no clear theoretical interpretation (see Section~\ref{sec:analyses}), and the estimate is imprecise (not significant at $\alpha=0.05$, see above).
 The effect can therefore be ignored or treated as a nuisance variable.
@@
 The style-related proxy variables point to the expected direction.
-Increased \textit{Badness} of the document favours the \NACa\ ($\beta_{\text{Badness}}=-0.152$, $\mpPB=0.002$) and so does a lower density of genitives ($\beta=-0.693$, $\mpPB=0.001$).
+Increased \textit{Badness} of the document favours the \NACa\ ($\beta_{\text{Badness}}=-0.152$, $\mpPB=0.002$), and so does a lower density of genitives ($\beta=-0.693$, $\mpPB=0.001$).
 While these are merely proxies to style (and partially circular in the case of \textit{Genitives}), this result can at least encourage future work into stylistic effects.
@@
 In \cite{VerhoevenTemme2017}, it was shown that results from forced-choice and split-100 experiments mostly converge (with some numeric intricacies related to the non-linear distribution of preferences for alternants).
-I present a forced-choice experiment in this Section and not a split-100 experiment, mainly because in a dry run of the experiment, participants complained about the unnaturalness of distributing a probability mass across two alternants.
-Participants had to choose between two sentences that differed only in that one contained the \NACa\ and the other contained the \PGCa.
+I present a forced-choice experiment in this Section and not a split-100 experiment, mainly because in a dry run of the experiment, participants complained about the unnaturalness of distributing a probability mass across two alternants and tended to produce ratings of 0 and 100 (and to a lesser degree 50).
+Participants had to choose between two sentences that differed only in that one contained the \NACa\, and the other one contained the \PGCa.
 The analysis compares the probabilities assigned to the stimuli by the corpus-derived model with the frequency with which participants chose the alternants for the same stimuli.
@@
 In the simplified sentences, the case, number, etc. of the MNP remained the same as in the attested sentence, as did the choice of lexical material within the MNP.
-Eight sentences contained masculine or neuter kind nouns and the other eight contained feminine kind nouns.
+Eight sentences contained masculine or neuter kind nouns, and the other eight contained feminine kind nouns.
 Furthermore, in each of the masculine\slash neuter and feminine groups, four sentences originally containing the \NACa\ and four sentences originally containing the \PGCa\ were chosen.
@@
 The final pairs of stimuli were the sentence containing the attested and preferred alternant (according to the corpus GLMM) on the one hand and a modified version containing the dispreferred alternant on the other hand.
-They were presented next to each other and a 20 second time limit for each choice was set.%
+They were presented next to each other, and a 20 second time limit for each choice was set.%
 \footnote{No participant ever exceeded the time limit.}
@@
 Also, among participants, there are clearly different preferences ($\sigma_{\text{Participant}}=0.781$, $\sigma^{\text{Modelprediction}}_{\text{Participant}}=0.484$, 24 levels).
-On the extreme ends, one participant chose the \PGCa\ in 13 of 16 cases and two participants only chose it in 5 of 16 cases.
+On the extreme ends, one participant chose the \PGCa\ in 13 of 16 cases, and two participants only chose it in 5 of 16 cases.
 The regressor \textit{Modelprediction} achieves $\mpPB=0.007$ (1,000 replications) and is estimated at 5.408 relative to an intercept of -1.304.
@@
 Each participant read both the 16 sentences with the alternant predicted by the corpus model and the 16 modified sentences with the alternant that the corpus model did not predict.%
-\footnote{Notice that lemmas and their frequencies as well as lemma classes are included as regressors in the corpus-based GLMM and there was consequently no additional controlling of lemma frequencies, etc.}
+\footnote{Notice that lemmas and their frequencies as well as lemma classes are included as regressors in the corpus-based GLMM, and there was consequently no additional controlling of lemma frequencies, etc.}
 To minimise repetition effects, the stimuli for each participant were separated into two blocks of 16 targets and 33 fillers per block.
@@
 The pseudo-coefficients of determination are $R^2_m=0.237$ and $R^2_c=0.346$.
+There is thus some inter-subject variation and a noteworthy correlation between the reactions of the subjects and the corpus-derived probabilities.
+\hl{The fact that the correlation is not stronger can be explained in parts \ldots}
 
@@
 
-The overall model quality is high and the effect plot for the main effect is shown in Figure~\ref{fig:spr:effects}.
+The effect plot for the effect of interest is shown in Figure~\ref{fig:spr:effects}.
 The estimate for the sentences with \NACa\ is obviously imprecise and no differences in reading times are observed.
@@
   I thank (in alphabetical order) Felix Bildhauer, Susanne Flach, Elizabeth Pankratz, Samuel Reichert, Ulrike Sayatz, and Christian Zimmer for valuable discussions and comments.
-  Elizabeth thankfully also fixed my English.
-  Furthermore, I thank Ulrike for helping me to recruit the participants for the experiments.
-  Finally, I am grateful to Kim Maser for her work on the annotation of the concordances.
+  Also, I would like to thank two reviewers for Cognitive Linguistics with deep linguistics knowledge and knowledge of German grammar for comments which helped to improve the paper a lot.
+  I thank Dagmar Divjak who, in her role as editor for Cognitive Linguistics, also read the paper thoroughly and provided very useful comments.
+
+  Furthermore, I thank Ulrike Sayatz for helping me to recruit the participants for the experiments.
+  Elizabeth Pankratz thankfully also fixed my English.
+  Finally, I am grateful to my student assistants Kim Maser for her work on the annotation of the concordances and Luise Rißmann for supervising most of the experiments.
   The research presented here was made possible in part through funding from the \textit{Deutsche Forschungsgemeinschaft} (DFG, personal grant SCHA1916/1-1).

### 2017-12-21 R/3_glmm.R

@@ -5,7 +5,7 @@ require(effects)
 require(gridExtra)
 require(car)
 require(pbkrtest)
-
+require(lattice)
 
 
 measure.glmm <- glmer(Construction~1
@@ -101,7 +101,8 @@ dotchart(measure.fixeffs, pch=20,
          xlim = c(x.lower, x.upper),
          lcolor = "gray",
          cex = 1.2,
-         main=paste("Coefficient estimates\n with bootstrapped 95% CI", sep=""))
+         # main=paste("Coefficient estimates\n with bootstrapped 95% CI", sep=""))
+         main=NULL)
 lines(c(0,0), c(0,length(measure.ci.95)), col="gray")
 
 for (i in 1:nrow(measure.ci.95)) {
@@ -113,18 +114,23 @@ if (save.persistent) dev.off()
 
 
 # Effect plots.
-fixeff.pl.cex <- 1.5
 effs <- c("Badness", "Genitives", "Cardinal", "Measurecase", "Measurenumber",
           "Kindattraction", "Kindfreq", "Kindgender", 
           "Measureattraction", "Measureclass", "Measurefreq")
 
 for (eff in effs) {
-  if (save.persistent) pdf(paste(out.dir, eff, ".pdf", sep=""))
+  if (save.persistent) {
+    trellis.device(device = "pdf", file = paste0(out.dir, eff, ".pdf"))
+    trellis.par.set(list(axis.text = list(cex = 1.75)))
+    trellis.par.set(list(par.ylab.text = list(cex = 1.75)))
+    trellis.par.set(list(par.xlab.text = list(cex = 1.75)))
+  }
   p <- plot(effect(eff, measure.glmm),
-            rug=F, 
-            main = paste("Effect plot: ", eff),
+            rug=F,
+            main = NULL,
             ylab = "Probability of PGCa",
-            colors = c("black", "darkblue"), cex = fixeff.pl.cex)
+            colors = c("black", "darkblue")
+            )
   print(p)
   if (save.persistent) dev.off()
 }

### 2017-12-21 R/4_glmm-mcmc.R

@@ -77,7 +77,9 @@ dotchart(rep(-100, length(measure.fixeffs)), pch=20,
          labels = names(measure.fixeffs),
          lcolor = "gray",
          cex = 1.2,
-         main=paste("MLE and MCMC oefficient estimates\n with 95% confidence intervals", sep=""))
+#         main=paste("MLE and MCMC oefficient estimates\n with 95% confidence intervals", sep="")
+         main=NULL
+         )
 lines(c(0,0), c(0,length(measure.ci.95)), col="gray")
 
 for (i in 1:nrow(measure.ci.95)) {

### 2017-12-21 R/6_exp-fc.R

@@ -3,6 +3,7 @@ require(effects)
 require(MuMIn)
 require(pbkrtest)
 require(boot)
+require(lattice)
 
 rm(list = ls())
 
@@ -55,10 +56,9 @@ responses.df <- data.frame(
 
 
 # Rename some variables to make paper consistent and readable.
-
 colnames(responses.df)[which(colnames(responses.df) == "Modelpred")] <- "Modelprediction"
 colnames(responses.df)[which(colnames(responses.df) == "ResponseDrop")] <- "Chosenconstruction"
-responses.df$Chosenconstruction <- as.factor(ifelse(responses.df$Chosenconstruction == T, "PGCa", "NACa"))
+responses.df$Chosenconstruction <- as.factor(ifelse(responses.df$Chosenconstruction == T, "PGCadj", "NACadj"))
 
 # REPORT
 if (save.persistent) sink(paste(out.dir, "results.txt", sep=""))
@@ -91,11 +91,17 @@ if (save.persistent) sink()
 
 # Effect plot.
 p <- plot(effect("Modelprediction", model.fc, KR = T), rug=F, colors = c("black", "darkblue"),
-          main="Forced choice experiment",
-          ylab="Probability that PGCa is chosen",
-          xlab="Probability for PGCa from corpus-based model"
+          main=NULL,
+          ylab="Probability that PGCadj is chosen",
+          xlab="Probability for PGCadj from corpus-based model"
 )
-if (save.persistent) pdf(paste(out.dir, "effects.pdf", sep=""))
+
+if (save.persistent) {
+  trellis.device(device = "pdf", file = paste0(out.dir, "effects.pdf"))
+  trellis.par.set(list(axis.text = list(cex = 1.75)))
+  trellis.par.set(list(par.ylab.text = list(cex = 1.75)))
+  trellis.par.set(list(par.xlab.text = list(cex = 1.75)))
+}
 print(p)
 if (save.persistent) dev.off()
 
@@ -124,7 +130,7 @@ if (save.persistent) dev.off()
 if (save.persistent) pdf(paste(out.dir, "proportions.pdf", sep=""))
 plot(responses.df$Chosenconstruction~responses.df$Modelprediction,
      main = "Forced choice: distribution of responses\nby (binned) predictions from corpus-based model",
-     xlab = "Probability for PGCa from corpus-based model", ylab="Proportion of responses",
[…]

### 2017-12-21 R/7_exp-spr.R

@@ -4,6 +4,7 @@ require(MuMIn)
 require(pbkrtest)
 require(plyr)
 require(boot)
+require(lattice)
 
 rm(list = ls())
 
@@ -138,6 +139,9 @@ rt.all$Modelprediction <- as.numeric(rt.all$Modelprediction)
 rt.all$rt.tt.resid     <- as.numeric(rt.all$t1.rt.resid + rt.all$t1.rt.resid)
 
 
+# Fix factor levels.
+revalue(rt.all$Construction, c("NACa" = "NACadj", "PGCa" = "PGCadj"))
+
 
 # ################
 # Remove outliers.
@@ -198,9 +202,18 @@ if (save.persistent) sink()
 # #############
 
 p <- plot(effect("Construction:Modelprediction", rt.model, KR = T), rug=F, colors = c("black", "darkblue"),
-          ylab="Residual log. reading time", xlab = "Probability for PGCa from corpus-based model",
-          main="Predicting reading times from\ncorpus-based model")
-if (save.persistent) pdf(paste(out.dir, "effects.pdf", sep=""))
+          ylab="Residual log. reading time", xlab = "Probability for PGCadj from corpus-based model",
+          main=NULL)
+
+if (save.persistent) {
+  trellis.device(device = "pdf", file = paste0(out.dir, "effects.pdf"))
+  trellis.par.set(list(axis.text = list(cex = 1.75)))
+  trellis.par.set(list(par.ylab.text = list(cex = 1.75)))
+  trellis.par.set(list(par.xlab.text = list(cex = 1.75)))
+  trellis.par.set(list(par.zlab.text = list(cex = 1.75)))
+  trellis.par.set(list(add.text = list(cex = 1.6)))
+  trellis.par.set(list(par.sub.text = list(cex = 1.75)))
+}
 print(p)
 if (save.persistent) dev.off()
 

### 2017-12-22 Latex/article.tex
Sections in separate files.

@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Cognitively oriented corpus linguistics}
-\label{sec:cogocl}
-
-This paper deals with a morpho-syntactic alternation between two constructions that occurs only in a very specific type of measure noun phrase in German.
-By \textit{alternation} I refer to a situation where two or more forms or constructions are available with no clear difference in acceptability, function, or meaning.
-The study of alternations has a long history in cognitively oriented corpus linguistics (\citealp{BresnanEa2007,BresnanHay2010,BresnanFord2010,DivjakArppe2013,Gries2015,NessetJanda2010}, to name just a few publications).
-This area of research is based on the assumption that language is a probabilistic phenomenon \citep{Bresnan2007} where alternants are chosen neither deterministically nor fully at random.
-Instead, multifactorial models are constructed which incorporate influencing factors from diverse levels, including contextual factors.
-The estimation of the model coefficients quantifies the influence that the factors have on the probability that either alternant is chosen.
-
-For such probabilistic generalisations estimated on corpus data, there has always been an interest in correlating the findings with results from experimental work (for example, \citealp{ArppeJaervikivi2007,BresnanEa2007,BresnanFord2010,DivjakGries2008,DivjakEa2016,FordBresnan2013}).
-This is often called a \textit{validation} of the corpus-derived findings, but \citet[303]{Divjak2016a} rightly criticises this choice of words ``because it creates the impression that behavioral experimental data is inherently more valuable than textual data,'' citing \cite{TummersEa2005}, who state that a corpus is ``a sample of spontaneous language use that is (generally) realized by native speakers.''
-However, as \citet[486--487]{Dabrowska2016} convincingly argues, this does not imply that we can in some way ``deduce mental representations from patterns of use,'' \ie\ from corpus data.%
-\footnote{In its colloquial meaning, the verb \textit{deduce} seems appropriate here.
-However, in terms of the scientific method, we are most likely talking about \textit{induction} rather than \textit{deduction}.}
-It would be highly surprising if this were possible, and the same holds for experimental methods.
-Nobody assumes that we can inductively infer mental representations from experiments, which -- as opposed to corpus studies -- even allow for direct access to the cognitive agent and offer much better possibilities to control experimental conditions and nuisance variables.
-Rather, under the standard approach, a theory of cognitive representation is pre-specified.
-Then, predictions are derived from this theory \textit{before} the experiment or the corpus study is conducted to \textit{test} the theory.
-Technical problems of statistical inference aside, a successful experiment corroborates the theory, and a failed experiment weakens the theory.
-The simplest assumption would be that the same approach is adequate for corpus studies.
-
-The present study is conducted within the general paradigm of cognitive corpus linguistics and includes a comparison of the corpus findings with results from two experiments.
-I assume a model of similarity-based classification in the form of Prototype Theory for modeling alternation.
-Protoype Theory has been used in alternation research in cognitively oriented corpus linguistics (see \citealp{DivjakArppe2013}; \citealp{Gries2003}).
-It is assumed that the variables annotated in a corpus study on an alternation phenomenon define prototypes for the alternants.
-The similarity of a given feature vector (in a concrete sentence where one of the alternants is used) influences to a large extent which alternant is chosen by speakers.
-This entails that a variant of Prototype Theory with features is assumed \citep{Rosch1978}, instead of the monolithic prototypes of earlier versions.
-Prototype Theory is well suited for modeling constructional choices but it is just one of several similarity-based theories of classification, the most prominent other framework being Exemplar Theory (\citealp{MedinSchaffer1978,Hintzman1986}; see \citealp{StormsEa2000} for a comparison of the theories in experimental settings).
-Prototype Theory and Exemplar Theory model essentially the same types of effects and differ mainly in whether they assume higher-level abstractions as part of the cognitive representation (Prototype Theory) or try to do without them (Exemplar Theory).
-I am sceptical that corpus analysis alone could ever decide which theory is more suitable, given that substantial doubt has been voiced whether even experimental methods are ultimately able to do so \citep{Barsalou1990}.%
-\footnote{In \cite{DivjakArppe2013}, it was shown using corpus data how both models form a converging picture.}
-Prototype Theory (with feature abstractions) is preferred here simply because it fits the established alternation modeling paradigm, which relies on features being weighted in statistical models.
-In Section~\ref{sec:analyses}, a prototype-theoretical parlance is therefore adopted.
-
-In the remainder of the paper, I introduce the alternation phenomenon (Section~\ref{sec:descriptive}) and suggest a theory-driven set of factors influencing the alternation (Section~\ref{sec:analyses}).
-Then the corpus study is presented, including an appropriate statistical analysis (Section~\ref{sec:corpusstudies}).
-Two experiments which corroborate the corpus-based findings are then reported in Section~\ref{sec:experimental} before I sum up the paper in Section~\ref{sec:conclusion}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Case assignment in German measure NPs}
-\label{sec:germanmeasurenps}
-
-\subsection{Two stable cases and a case alternation}
-\label{sec:descriptive}
-
-In this section, I introduce and illustrate the relevant alternating constructions.
-I describe the narrowly defined syntactic configuration in which the alternation occurs, and I motivate the focus on \textit{only} this narrow range rather than, for example, the whole range of nominal constructions expressing quantities.
-
-I use the term \textit{measure noun phrase} (MNP) to refer to a noun phrase (NP) in which a kind-denoting (count or mass) noun depends on another noun that specifies a quantity of the objects or the substance denoted by the kind-denoting noun.
-I call the kind-denoting noun the \textit{kind noun} and the quantity-denoting noun the \textit{measure noun}.
-For illustration purposes, in the English \textit{a glass of good wine}, \textit{glass} is the measure noun and \textit{wine} is the kind noun.
-Measure nouns can be all sorts of nouns which denote a quantity (such as \textit{litre} or \textit{amount}) but also those denoting containers, collections, etc.
-(such as \textit{glass} or \textit{bucket}).
-Like \citet[284]{Brems2003}, I consider nouns as measure nouns ``which, strictly speaking, do not designate a `measure', but display a more nebulous potential for quantification'' (also \citealp[530]{Koptjevskaja2001}, and \citealp[338]{Rutkowski2007}).
-
-\subsubsection{Core structures related to the alternation}
-
-Three different syntactic configurations within MNPs need to be distinguished, and the case alternation occurs only in one of them.
-It occurs only when the kind noun is modified by an attributive adjective and there is no determiner, as in (\ref{ex:intro:alternation}).
-Superficially, the sentences are functionally and semantically equivalent either with the kind noun in the genitive (\ref{ex:intro:alternation1}) or in the same case as the measure noun, an accusative in the case of (\ref{ex:intro:alternation2}).%
-\footnote{Some descriptive and normative grammars take stronger positions with regard to the acceptability of the two options.
-See \cite{Hentschel1993} and \cite{Zimmer2015} for analyses of the sometimes absurd stances taken in grammars of German.
-As will be shown (especially in Section~\ref{sec:exp:fc}), there might be preferences, but we cannot assume either construction to be unacceptable.
-}
-
-\begin{exe}
-  \ex\label{ex:intro:alternation}
-  \begin{xlist}
-    \ex[ ]{\label{ex:intro:alternation1} \gll Wir trinken [[ein Glas]\Sub{Acc} [guten Weins]\Sub{Gen}]\Sub{Acc}.\\
-    we drink a glass good wine \\
-    \glt We drink a glass of good wine.}
-    \ex[ ]{\label{ex:intro:alternation2} Wir trinken [[ein Glas]\Sub{Acc} [guten Wein]\Sub{Acc}]\Sub{Acc}.}
-  \end{xlist}
-\end{exe}
-
-This specific configuration has to be seen in the context of two other configurations, to which I turn now.
-First, if the kind noun forms an NP with a determiner, the construction resembles (and is usually called) a \textit{pseudo-partitive} (on partitives and pseudo-partitives see, e.\,g., \citealp{Barker1998,Selkirk1977,Stickney2007,Vos1999}; for a recent application of the terminology to German, see \citealp{Gerstenberger2015}).%
-\footnote{If the kind noun is definite, the construction instantiates a true partitive.
-Whereas partitives are constructions denoting a proper part-of relation as in \textit{a sip of the wine}, pseudo partitives -- albeit syntactically similar and diachronically related to partitives in many languages -- merely denote quantities and contain indefinite kind nouns as in \textit{a sip of wine}.
-In the literature on German, some authors incorrectly call the pseudo-partitive a \textit{partitive} \citep{Hentschel1993} while some realise the difference and at least mention it \citep{Eschenbach1994,GallmannLindauer1994,Loebel1989,Zimmer2015}.
-}
-Here, the kind noun is in the genitive, and I refer to the construction in (\ref{ex:intro:pseudopartitive1}) as the \textit{Pseudo-partitive Genitive Construction} (PGC).
-
-\begin{exe}
-%  \ex\label{ex:intro:pseudopartitive}
-%  \begin{xlist}
-    \ex[ ]{\label{ex:intro:pseudopartitive1}\gll Wir trinken [[ein Glas]\Sub{Acc} [dieses Weins]\Sub{Gen}]\Sub{Acc}.\\
-    we drink a glass this wine\\
-    \glt We drink a glass of this wine.}
-%    \ex[*]{\label{ex:intro:pseudopartitive2} Wir trinken [[eine Tasse]\Sub{Acc} [einen leckeren Kaffee]\Sub{Acc}]\Sub{Acc}.}
-%  \end{xlist}
-\end{exe}
-
-Second, if the kind noun is bare -- \ie\ if it comes neither with a determiner nor a modifying adjective -- it is uninflected as in (\ref{ex:intro:narrowapposition1}), and the genitive as seen in the PGC is not acceptable, see (\ref{ex:intro:narrowapposition2}).
-
-\begin{exe}
-  \ex\label{ex:intro:narrowapposition}
-  \begin{xlist}
-    \ex[ ]{\label{ex:intro:narrowapposition1}\gll Wir trinken [[ein Glas]\Sub{Acc} [Wein]\Sub{?}]\Sub{Acc}.\\
-    we drink a glass wine\\
-    \glt We drink a glass of wine.}
-    \ex[*]{\label{ex:intro:narrowapposition2} Wir trinken [[ein Glas]\Sub{Acc} [Weins]\Sub{Gen}]\Sub{Acc}.}
-  \end{xlist}
-\end{exe}
-
-This construction is usually classified as a \textit{Narrow Apposition Construction} \citep{Loebel1986}, henceforth NAC.%
-\footnote{The construction as in (\ref{ex:intro:narrowapposition1}) is also referred to as the \textit{Direct Partitive Construction} for other Germanic languages in which the PGC with the synthetic genitive is not available.
-This nomenclature makes sense in contrast to the \textit{Indirect Partitive Construction} with prepositional linkers translating to \textit{of} -- \ie\ analytic genitives -- in such languages, see \cite{HankamerMikkelsen2008} for Danish.
-For German, this terminology is not distinctive enough, which is why I use the terms NAC and PGC.}
-Notice that the unavailability of the genitive on the kind noun follows independently from a constraint that genitive NPs in German require the presence of some strongly case-marked element (determiner or adjective) in addition to the head noun in order to be acceptable (\citealp{GallmannLindauer1994,Schachtl1989}; see also \citealp[160]{Eisenberg2013b}).
-
-It is difficult to determine whether the bare kind noun in the narrow apposition construction as in (\ref{ex:intro:narrowapposition1}) bears no case at all, a generic case, or agrees in case with the measure noun.%
-\footnote{I would like to thank one of the reviewers for pointing this out.}
-When there is an adjective as in (\ref{ex:intro:alternation2}), the embedded kind NP clearly agrees in case, but due to the overall absence of markers of case in the singular, bare nouns mostly show no indication of their case.
-The only nouns which do have case markers in the singular are the so-called weak nouns \citep{Koepcke1995,Schaefer2016c}, which have something like a non-nominative \textit{-en} marker in the singular.
-Unfortunately, there are no genuine mass nouns among the weak nouns.
-However, a few of them can be coerced into a mass noun, such as \textit{Hase} `rabbit' (meaning `rabbit meat').
-It then appears as if the uninflected form is preferred, but the inflected form is not excluded.
-In (\ref{ex:macabre1}), the clearly acceptable form \textit{Hase} can only be a nominative singular or caseless.
-In (\ref{ex:macabre2}), the form \textit{Hasen} could be an accusative, dative, or genitive.
-
-\begin{exe}
-  \ex\label{ex:macabre}
-  \begin{xlist}
-    \ex[ ]{\label{ex:macabre1}\gll Niemand will [ein Stück [Hase]\Sub{Nom\slash caseless}]\Sub{Acc} essen.\\
-    nobody wants a piece rabbit eat\\
-  \trans Nobody wants to eat a piece of rabbit.}
-    \ex[?]{\label{ex:macabre2}\gll Niemand will [ein Stück [Hasen]\Sub{Acc\slash Dat\slash Gen}]\Sub{Acc} essen.\\
-    nobody wants a piece rabbit eat\\}
-  \end{xlist}
-\end{exe}
-
-Even a full unacceptability of (\ref{ex:macabre2}) would not be conclusive, however, as a possible aversion of speakers towards the case-marked form might be due to the fact that it is at least potentially also a genitive, in which case the constraint against bare genitive nouns would apply.
-With plural kind nouns, the obligatory marking of the dative plural with \textit{-en} might provide some clues.
-However, plural kind nouns do not behave like singular ones in measure phrases anyway, as will be argued in Section~\ref{sec:neighbouringcases}.
-Also, judgements vary between (\ref{ex:datpl1}) and (\ref{ex:datpl2}), and both are found in corpora.%
-\footnote{For example, the variant in (\ref{ex:datpl1}) with the lemmas \textit{Sack} and \textit{Apfel} occurs four times, and the one in (\ref{ex:datpl2}) twice in the very large DECOW corpus (see Section~\ref{sec:gettingdata}).
-Similarly, for [\textit{Kiste} [\textit{Äpfel}]\Sub{Nom\slash Acc\slash Gen}]\Sub{Dat} `box of apples' (no case identity), I find three examples, and [\textit{Kiste} [\textit{Äpfeln}]\Sub{Dat}]\Sub{Dat} (clearly marked case identity), I find six examples.
-}
-
-\begin{exe}
-  \ex\label{ex:datpl} 
-  \begin{xlist}
-  \ex{\label{ex:datpl1}\gll mit [zwei Säcken [Äpfel]\Sub{Nom\slash Acc\slash Gen}]\Sub{Dat}\\
-     with a sack apples\\
-    \trans with a sack of apples}
-    \ex{\label{ex:datpl2}\gll mit [zwei Säcken [Äpfeln]\Sub{Dat}]\Sub{Dat}\\
-    with a sack apples\\}
-  \end{xlist}
-\end{exe}
-
-Descriptive grammars seem to favour an analysis in terms of caselessness (for example, \citealp[1981]{ZifonunEa1997c}).
-The hazy picture of the case of bare kind NPs is most likely due to the fact that case is so rarely marked on German nouns, where case is marked mostly on determiners and to some degree adjectives.
-The uncertainty in the few cases where case can be marked (weak nouns in the singular and dative plurals) would thus be a direct consequence of the fact that the construction is not very specific with respect to the case of the kind noun.
-
-\begin{table}
-  \centering
-  \begin{tabular}{llll}
-    \multicolumn{1}{r}{kind NP is:} & bare noun NP & NP with adjective & NP with determiner \\
-    & [\ldots N\Subsf{meas} [N\Subsf{kind}]] & [\ldots N\Subsf{meas} [AP N\Subsf{kind}]] & [\ldots N\Subsf{meas} [D N\Subsf{kind}]] \\
-    \midrule
-    \multirow{2}{*}{narrow apposition}
-                & \NACb                                                 & \NACa                                                   & \multirow{2}{*}{---}       \\
-		& (\ref{ex:intro:narrowapposition1}) \textit{Glas Wein} & (\ref{ex:intro:alternation2}) \textit{Glas guten Wein}  &                            \\
-    \midrule
-
-    \multirow{2}{*}{pseudo-partitive genitive} 
-                & \multirow{2}{*}{---}                                  & \PGCa                                                   & \PGCd                      \\
-                &                                                       & (\ref{ex:intro:alternation1}) \textit{Glas guten Weins} & (\ref{ex:intro:pseudopartitive1}) \textit{Glas dieses Weins} \\
-  \end{tabular}
-  \caption{NAC and PGC constructions in different NP structures with examples and references to full example sentences}
-  \label{tab:constructions}
-\end{table}
-
-To summarise, the case patterns in the NAC and in the PGC (depending on the structure of the kind NP) is given in Table~\ref{tab:constructions}.
-I call the narrow apposition construction with a bare kind noun the \NACb\ and the partitive genitive with a determiner in the kind NP the \PGCd.
-For the alternants with an adjective but no determiner in the kind noun phrase I use the terms \NACa\ and \PGCa.
-In principle, this paper is about the middle column of Table~\ref{tab:constructions}, \ie\ the syntactic configuration in which two different case patterns are acceptable.
-However, in Section~\ref{sec:analyses}, the outer columns (\NACb\ and \PGCd) will still play a major role when the factors controlling the alternation are discussed.
-
-\subsubsection{Neighbouring cases}
-\label{sec:neighbouringcases}
-
-% namely:
-% 
-% \begin{enumerate}[i.]
-%   \item{\label{it:intro:idiot} alleged alternatives to the \NACa\ and the \PGCa,}
-% %  \item{\label{it:intro:datsg} alternative \textit{weak} forms of dative singular neuter adjectives,}
-% %  \item{\label{it:intro:femsg} case syncretism in feminine NPs,}
-%   \item{\label{it:intro:plurl} similar constructions with plural/collective kind nouns,}
-%   \item{\label{it:intro:noifl} grammaticalised non-inflected measure nouns, and}
-%   \item{\label{it:intro:other} alternative constructions for expressing quantities.}
-% \end{enumerate}
-%
-% \vspace{-1\baselineskip}
-
-I now turn to some more subtle issues related to the measure noun case alternation in order to delimit the scope of the study.
-First, there is a claim found in some grammars that a generic nominative, accusative, and even dative on the kind noun are used instead of the genitive (\PGCa) or case agreement (\NACa).
-Overviews can be found in \cite{Hentschel1993} and \cite{Zimmer2015}.
-Sentence (\ref{ex:intro:gennom}) shows a putative generic nominative on the kind noun inside an accusative MNP.
-
-\begin{exe}
-    \ex[*]{\label{ex:intro:gennom} Wir trinken [[eine Tasse]\Sub{Acc} [heißer Kaffee]\Sub{Nom}]\Sub{Acc}.}
-\end{exe}
-
-It was shown empirically in \cite{Hentschel1993} that such sentences are de facto not acceptable.
-Also, in my corpus sample, they simply did not occur.
-Even if they are accepted by some speakers, their extremely low frequency makes it virtually impossible to study them using corpus linguistic methods (Section~\ref{sec:corpusstudies}), and I consequently do not discuss them further.
-
-% As for (\ref{it:intro:datsg}), a complication with neuter kind nouns in the dative is mentioned by \citet[20--22]{Zimmer2015}.
-% In the \NACa, the adjective normally inflects with the so-called \textit{strong} case and number suffixes, which are used when no determiner with strong inflection is present in the NP.
-% \cite{Zimmer2015} reports a high number of occurrences of adjectives being inflected with the \textit{weak} inflectional marker, which are normally only used if a strongly inflected determiner precedes the adjective: \textit{kalt-en Wasser} `cold water' instead of \textit{kalt-em Wasser}.%
-% \footnote{See Section~\ref{sec:analyses} for more discussion of adjectival inflection.}
-% In my corpus sample, this tendency was not nearly as clear, and there was a high number of very noisy sentences among those potentially showing this pattern.
-% Hence, I do not discuss these forms.%
-% \footnote{All corpus data and scripts used for this paper will be released freely, so further examination of the few cases maybe showing this kind of inflection is possible.}
-% 
-% \label{page:femininesyncretism}
-% With feminine kind nouns, (\ref{it:intro:femsg}) is relevant.
-% Whereas the singular masculine and neuter nominal subsystem is a four case system (at least for NPs with adjectives and determiners), feminine singular NPs show nominative-accusative and dative-genitive syncretisms.
-% Table~\ref{tab:syncretisms} shows this as well as the partial syncretism in the plural, where no gender distinctions are made.
-% This means that with dative measure nouns, the alternation is unobservable if the kind noun is feminine, and it is problematic to speak of a \textit{genitive} kind noun construction in this case.
-% This is taken into account in the statistical modeling and the discussion presented in Section~\ref{sec:corpusstudies}.
-% 
-% \begin{table}
-%   \centering
-%   \begin{tabular}{llll}
-%      & Masculine\slash Neuter & Feminine & Plural \\
-%      \midrule
-%      nominative & rot-er Wein    & \multirow{2}{*}{frisch-e Sahne}   & \multirow{2}{*}{klein-e Äpfel} \\
-%      accusative & rot-en Wein    &                                   &                                \\
-%      dative     & rot-em Wein    & \multirow{2}{*}{frisch-er Sahne}  & klein-en Äpfel-n               \\
-%      genitive   & rot-en Wein-es &                                   & klein-er Äpfel                 \\
-%   \end{tabular}
-%   \caption{Case syncretisms in NPs with an adjective and without a determiner (\textit{roter Wein} `red wine', \textit{frische Sahne} `fresh cream', \textit{kleine Äpfel} `small apples')}
-%   \label{tab:syncretisms}
-% \end{table}
-
-Second, we find that, if the kind noun is a plural count noun as in \textit{ein Sack kleine Äpfel} (\NACa) or \textit{ein Sack kleiner Äpfel} (\PGCa) `a bag of small apples', a similar alternation between PGC and NAC can be observed.
-In line with experimental results reported in \citet[15--16]{Zimmer2015}, I found that the PGC is so dominant with plural kind nouns (794 of 861 cases, or over 92\%, cf.\ Section~\ref{sec:corpusstudies}) that the alternation cannot be analysed in the same way as in the singular.
-%\footnote{More concretely, if these cases were included in the regression analysis of the corpus data along with a factor encoding the number of the kind noun, this factor would most assuredly override any other regressor for the data points with a plural kind noun.}
-While this will play a role in the interpretation of the corpus findings, MNPs with plural kind nouns will not be included in the corpus study and the experiments reported in Sections~\ref{sec:corpusstudies} and \ref{sec:experimental}.
-
-Third, some measure nouns have been grammaticalised in a way that they always appear in their non-inflected form.
-They are typical measure nouns like \textit{Gramm} `gram', \textit{Pfund} `pound' or \textit{Prozent} `percent', which have no normal plural forms at all.%
-\footnote{Plurals like \textit{Pfunde} and \textit{Prozente} have special meanings and have very restricted uses, mostly in idiomatic expressions such as \textit{Pfunde verlieren} `lose pounds' and \textit{Prozente machen} `make a profit'.
-They cannot be used in normal MNPs.}
-%\footnote{With neglectable frequency, this effect is extended to container nouns like \textit{Glas} `glass', but with a twist.
-%A non-inflected version \textit{zwei Glas roter Wein} `two glasses of wine' and an inflected version \textit{zwei Gläser roter Wein} co-exist.
-%The non-inflected version denotes the quantity of wine fitting into two glasses, the inflected version denotes two glasses filled with wine.}
-I treat these cases like other measure nouns because they enter into both the \NACa\ as in (\ref{ex:uninflectedmeasures:a}) and the \PGCa\ as in (\ref{ex:uninflectedmeasures:b}).
-In Section~\ref{sec:analyses}, however, degrees of grammaticalisation as a factor influencing the alternation will be discussed prominently.
-
-\begin{exe}
-  \ex\label{ex:uninflectedmeasures}
-  \begin{xlist}
-    \ex{\label{ex:uninflectedmeasures:a} zwei Gramm brauner Zucker}
-    \ex{\label{ex:uninflectedmeasures:b}\gll zwei Gramm braunen Zuckers\\
-                                         two gram brown sugar\\
-				         two grams of brown sugar}
-  \end{xlist}
-\end{exe}
-
-Finally, there are alternative ways of expressing similar quantificational meanings.
-In the variationist tradition, which is strongly influenced by Labovian sociolinguistics, the \textit{principle of accountability} would dictate that proper studies should examine a variationist \textit{variable}, \ie\ all different \textit{ways of saying the same thing} (\citealp{Labov1966}, \citealp{Labov1969}, for an overview see \citealp{Tagliamonte2012}).
-In the case at hand, the variable might be something like \textit{measurement of quantities of substances and collections}.
-I argue that it is fully justified to focus narrowly on \NACa\ and \PGCa\ with their well-defined morpho-syntactic properties, mostly because the alternative constructions are not used in the same range of contexts.
-Two major alternatives might be considered.
-There is an analytic (pseudo-)partitive with \textit{von} `of'.
-It is only available as an alternative to the PGC if the kind noun phrase contains a (definite or indefinite) determiner as in (\ref{ex:analyticalpartitive}).
-
-\begin{exe}
-  \ex\label{ex:analyticalpartitive} 
-  \begin{xlist}
-    \ex[ ]{\label{ex:analyticalpartitive:a}\gll ein Glas von dem roten Wein\\
-                                                a glass of the red wine\\
-                                                \glt a glass of the red wine
-       }
-    \ex[*]{\label{ex:analyticalpartitive:b}\gll ein Glas von rotem Wein\\
-                                                a glass of red wine\\
-                                                \glt a glass of red wine
-       }
-  \end{xlist}
-\end{exe}
-
-This means that the \textit{von} (pseudo-)partitive does not compete with the \NACa\ and the \PGCa.
-In fact, there is not a single context in which they can be interchanged.
-
-Additionally, we find constructions with \textit{voll} (\textit{von})\slash\textit{voller} `full (of)' and \textit{mit} `with' are available, see (\ref{ex:alternatives}).
-
-\begin{exe}
-  \ex\label{ex:alternatives}
-  \begin{xlist}
-  \ex{\label{ex:voll}\gll   ein Glas voll von rotem Wein\\
-                            a glass full of red wine\\
-			    \glt a glass full of red wine}
-  \ex{\label{ex:voller}\gll ein Glas {voll\slash voller} {rotem\slash roter} Wein\\
-                            a glass full-of red wine\\
-			    \glt a glass full of red wine}
-  \ex{\label{ex:mit}\gll    ein Glas mit rotem Wein\\
-                            a glass with red wine\\
-			    \glt a glass with red wine in it\slash a glass filled with red wine}
-  \end{xlist}
-\end{exe}
-
-These construction have very idiosyncratic properties.
-The construction with \textit{voller} is discussed in \cite{Zeldes2018} and the construction with \textit{mit} in \cite{Bhatt1990}.
-They are not semantically equivalent to the NAC and the PGC, and they are only available for a small subset of measure nouns.
-All of these constructions are used only with measure nouns denoting containers, where the whole NP typically refers to the container and never to the quantity contained in it.
-They are incompatible with measure nouns denoting natural portions such as \textit{Schluck} `gulp' or \textit{Haufen} `heap' and strongly grammaticalised nouns such as \textit{Gramm} `gram'.
-This disqualifies them as alternatives to the \NACa\ or \PGCa\ in most contexts, and they are thus clearly not just more ways of saying the same thing. 
-It is therefore reasonable to focus on the two well-defined alternants.
-
-This concludes the descriptive overview of the phenomenon.
-I have demonstrated that there is an alternation between two measure noun constructions in a narrow syntactic configuration (kind NP with an adjective but without a determiner), and that the two constructions differ in the case of the kind noun (case agreement with the measure noun or genitive).
-I turn to more theory-oriented discussion of the alternation in the next section.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Factors controlling the alternation}
-\label{sec:analyses}
-
-This section briefly reviews existing analyses of the \PGCa\ vs.\ \NACa\ alternation and related issues.
-I also develop my own analysis and the appropriate hypotheses for the empirical studies presented in Sections~\ref{sec:corpusstudies} and \ref{sec:experimental}.
-
-.%
-\footnote{A reviewer mentioned that she or he had the impression that dialectal variation is also a factor influencing the alternation.
-This is definitely true as some dialects (such as Alemannic) tend to have no genitive at all.
-While this is an interesting aspect for further research, the main obstacle is that there are no corpora of German which are both large enough and annotated with reliable meta data about regional variants.
-In the annotation of the corpus study, documents obviously written in a regional variant were excluded.
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Corpus study}
-\label{sec:corpusstudies}
-
-\subsection{Corpus choice and sampling}
-\label{sec:gettingdata}
-
-% TODO Quote:
-% Brems (2016:171--175) for using GOOGLE data, including COUNTS!
-
-For the present study, I used the German \textit{Corpus from the Web} (COW) in its 2014 version DECOW14A (\citealp{SchaeferBildhauer2012full}, and \citealp{Schaefer2015b}, as well as \citealp{BiemannEa2013}, and \citealp{SchaeferBildhauer2013}, for overviews of web corpora in general and the methodology of their construction), which contains almost 21 billion tokens.%
-\footnote{The COW corpora (Dutch, English, French, German, Spanish, Swedish) are made available for free at \url{https://www.webcorpora.org}.
-At the time of this writing, a newer 2016 version DECOW16 has already been released.}
-I chose this corpus for two main reasons.%
-\footnote{The use of web data for linguistic research does require explicit and careful justification.
-Due to the noisy nature and unknown composition of the web, only carefully designed and established web corpora like the COW corpora or the SketchEngine corpora \citep{KilgarriffEa2014} should be used.
-Clearly, using search engine results is ``bad science'' for many reasons, most prominently total non-replicability of results, as \cite{Kilgarriff2006} pointed out more than ten years ago.
-Careless use of search engine results is still found, however, see for example \citet[171--175]{DeclerckBrems2016}.}
-First, the external validity of any study is increased through a higher heterogeneity of the sample \citep[30]{MaxwellDelaney2004}, and the DECOW corpus has clearly a much more heterogeneous composition compared to the only other very large corpus of German, the DeReKo \citep{KupietzEa2010} of the Institute for the German Language (IDS), which contains almost exclusively newspaper texts.%
-\footnote{It was shown in \cite{W16-2601} that, for example, the range of topics covered is much smaller in DeReKo compared to DECOW.}
-Second, it was already mentioned that normative grammars often adopt clear positions regarding the grammaticality of either the \NACa\ or the \PGCa.
-Thus, newspaper text or any other text that conforms strongly to normative grammars might not represent the alternation phenomenon fully (and without bias) because authors and proofreaders who must adhere to normative guidelines might favour one alternative or the other explicitly.
-Web corpora, on the other hand, contain at least some amount of non-standard language from forums and similar sources.
-For these or similar reasons, COW corpora have been used in a number of peer-reviewed publications, for example \cite{VanGoethemHiligsmann2014}, \cite{VanGoethemHuening2015}, \cite{MuellerS2014}, \cite{Schaefer2016c}, \cite{SchaeferSayatz2014}, \cite{SchaeferSayatz2016}, and \cite{Zimmer2015}.
-Therefore, DECOW is a valid choice for this study.
-
-I now turn to the sampling procedure applied to obtain concordances for manual annotation and statistical analysis.
-Among the factors potentially influencing the alternation (see Section~\ref{sec:germanmeasurenps}) were lemma-specific preference effects.
-Therefore, it was highly desirable to obtain a sample in which most of the highly frequent actually-occurring combinations of kind nouns and measure nouns were represented.
-I applied a three-stage process in order to obtain such a sample, which consisted of the following steps:
-
-\begin{enumerate}[i.]
-  \item\label{enum:data:step1} generating a list of the one hundred most frequent mass nouns,
-  \item\label{enum:data:step2} deriving a list of all measure nouns with which the mass nouns co-occur in the \NACb, and 
-  \item\label{enum:data:step3} sampling the target constructions by querying each combination of mass noun and measure noun found in step (\ref{enum:data:step2}).
-\end{enumerate}
-
-\vspace{-1\baselineskip}
-
-In step (\ref{enum:data:step1}), I exported a list of all nouns in the DECOW14A01 sub-corpus sorted by their token frequency and manually went through it from the most frequent noun downwards, selecting the first one hundred mass nouns that occurred in the list.%
-\footnote{DECOW14A01 is the first slice (roughly a twentieth) of the complete DECOW14A corpus.
-It contains just over one billion tokens.}
-Mass nouns were defined as concrete nouns which denote a substance in the broad sense, combine with uninflected mass quantifiers such as \textit{viel} `much' and \textit{wenig} `little' (\textit{viel Bier} `much beer'), and form only sortal and unit plurals (such as the plural \textit{Biere} `types of beer' or `glasses of beer').
-Abstract nouns which partially behave like mass nouns (like \textit{Spaß} `fun’ or \textit{Gefahr} `danger’) were excluded because they are usually not quantified in the same way as concrete mass nouns.
-The hundredth selected mass noun was \textit{Schmuck} `jewellery’, which is the 3,054th most frequent noun in the original frequency list.
-
-This list of mass nouns was used in step (\ref{enum:data:step2}) to derive a list of measure nouns co-occurring with the mass nouns. 
-In order to generate this list, I utilised the fact that a direct sequence of two nouns almost always instantiates the bare-noun NAC if the second noun is a mass noun.
-Hence, I searched for all sequences N\Sub{1}N\Sub{2} where N\Sub{2} was one of the mass noun lemmas extracted in step (\ref{enum:data:step1}).
-Then, the resulting 100 lists of noun-noun combinations were each sorted by frequency in descending order and sieved manually to remove erroneous hits.
-From each of the 100 lists, I also removed noun-noun combinations that had a frequency below 2, except if the individual list would have otherwise been shorter than 20 noun-noun combinations.
-The result was a list of the most frequent 2,365 individual combinations of a measure noun and a mass noun.
-
-In step (\ref{enum:data:step3}), each of these 2,365 noun–noun combinations was queried in the target constructions (\PGCa\ and \NACa) individually in each of the first ten slices of DECOW (roughly 10 billion tokens).
-In order to reduce the sample size for the manual annotation process, the final concordance was sampled from the results of these 2,365 queries.
-Since the mass nouns in the sample were distributed according to the usual power law (often referred to as a \textit{Zipfian} distribution), I used all hits for nouns with a frequency up to 100 and a sample of 100 of all those with higher frequency.
-The final sample contained 6,843 sentences, which was reduced to 5,063 in the manual annotation process due to removal of noisy material, erroneous hits and uninformative cases where the measure noun was in the genitive, in which case the \NACa\ cannot be distinguished from the \PGCa.
-Given the careful sampling procedure described in this section, we can be highly sure that it contains all relevant and reasonably frequent noun–noun combinations in the target constructions.%
-\footnote{In a similar fashion, the 100 most frequent measure nouns occurring with plural kind nouns were listed and queried, resulting in a sample of 871 sentences.
-As stated in Section~\ref{sec:germanmeasurenps}, the \NACa\ is virtually never used with plural kind nouns, and this sample was not used except for quantifying the frequency of occurrence of the constructions (67 times \NACa\ and 794 times \PGCa).
-However, the sample is distributed with the data package accompanying this paper.
-}
-
-Finally, two auxiliary samples were also drawn.
-As mentioned in Section~\ref{sec:analyses}, the distribution of the measure noun and kind noun lemmas in the \NACb\ and the \PGCd\ with a determiner will be modelled as factors influencing the alternation.
-Therefore, all noun-noun pairs from the process described above were also queried in the two non-alternating constructions, resulting in 17,252 hits for the \PGCd\ and 315,635 hits for the \NACb.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Variables and annotation}
-\label{sec:annotation}
-
-The full set of manually annotated variables for the main sample is given in Table~\ref{tab:variables}, and I briefly discuss it now.%
-\footnote{All numeric variables were also z-transformed (\ie\ centered to the mean and rescaled such that they have a standard deviation of 1) to facilitate their interpretation in the regression models reported in the next section.}
-Notice first that \textit{Construction} is the response variable (or `dependent variable') with the values \textit{PGCa} and \textit{NACa}.
-
-\begin{table}
-  \centering
-  \begin{tabular}{llll}
-    Unit of reference & Variable                      & Type    & Levels (for factors only) \\
-    \midrule
-    Document       & Badness                          & numeric &                           \\
-                   & Genitives                        & numeric &                           \\
-    Sentence       & Cardinal                         & factor  & Yes, No                   \\
-                   & \textbf{Construction (response)} & factor  & NACa, PGCa                \\
-                   & Measurecase                      & factor  & Nom, Acc, Dat             \\
-    Kind lemma     & Kindattraction                   & numeric &                           \\
-                   & Kindfreq                         & numeric &                           \\
-    Measure lemma  & Measureattraction                & numeric &                           \\
-                   & Measureclass                     & factor  & Physical, Container,      \\
-                   &                                  &         & Amount, Portion, Rest     \\
-                   & Measurefreq                      & numeric &                           \\
-  \end{tabular}
-  \caption{Annotated variables for the main sample}
-  \label{tab:variables}
-\end{table}
-
-The variables \textit{Kindattraction} and \textit{Measureattraction} encode the ratio with which a given kind noun lemma or measure noun lemma occurs in the \PGCd\ and the  \NACb.
-They were calculated from the auxiliary samples described at the end of Section~\ref{sec:gettingdata} as a log-transformed quotient.
-The higher the value, the more often the noun occurs in the \PGCd (proportionally).%
-\footnote{
-  It could be argued that some more advanced measure of attraction strength should be used, as is done in Collostructional Analysis \citep{StefanowitschGries2003,GriesStefanowitsch2004}, see also \cite{Gries2015a}.
-  Three main points speak against such an approach in the present case.
-  First, the goal here is to quantify how often lemmas occur in the \PGCd\ and the \NACb, and these constructions do not compete at all but are rather mutually exclusive.
-  Collostructional approaches are not made for such scenarios.
-  Second, the attraction values will be used as regressors in a hierarchical logistic regression and the values resulting from collostructional analysis, \ie\ logarithmised Fisher p-values, have a very unfavourable distribution in the case at hand.
-  They cluster around 0 and they include values of $-\infty$.
-  Third, I tried using collexeme strength as a regressor (with smoothing to remedy the mathematical problems), and the results were unsatisfactory compared to the simple quotient used here.
-}
-Additionally, \textit{Kindfreq} and \textit{Measurefreq} are the logarithm-transformed frequencies per 1,000,000 words of each lemma, extracted from the frequency lists distributed by the DECOW corpus creators on their web page.
-They were added to control for basic frequency effects.
-
-In Section~\ref{sec:analyses}, it was hypothesised that classes of measure lemmas might have different preferences for the two alternants.
-To capture this, class information was annotated for measure lemmas.
-The classification was inspired by the list in \citet[530]{Koptjevskaja2001} but due to the low frequencies of many of the potential classes, a very coarse classification was finally used.
-With typical examples and their frequencies in the final sample, the classes are:
-\textit{Physical} (abstract precisely measurable units such as \textit{Liter} `litre', \textit{Meter} `metre', \textit{Gramm} `gram'; \textit{f=}1,968),
-\textit{Container} (\textit{Eimer} `bucket'; \textit{f=}740),
-\textit{Amount} (\textit{Menge} `amount'; \textit{f=}1,364), 
-\textit{Portion} (natural portions like \textit{Happen} `bite' or \textit{Krümel} `crumb'; \textit{f=}713).
-The lemmas that did not fit into either of these classes were labelled \textit{Rest} (\textit{f=}278).
-
-The variable \textit{Cardinal} encodes whether the measure noun is modified by a cardinal (\textit{f=}1,939) or not (\textit{f=}3,124).
-The purpose of this variable is to test whether cardinals really favour the \NACa\ as hypothesised in Section~\ref{sec:analyses}.
-
-To capture the influence of style mentioned in Section~\ref{sec:analyses}, two proxy variables were used.
-At the document level, the DECOW corpus has an annotation for \textit{Badness}.
-As described in \cite{SchaeferEa2013}, \textit{Badness} measures how well the distribution of highly frequent short words in the document matches a pre-generated language model for German.
-Documents with higher Badness usually contain more incoherent language, shorter sentences, etc.
-If the \PGCa\ actually favours higher stylistic levels, a high \textit{Badness} should be correlated with fewer occurrences.
-Documents in DECOW14 have also been annotated with a variable called \textit{Genitives}.
-The higher the values of this variable, the lower the proportion of genitives among all case-bearing forms is.
-A high number of genitives is indicative of higher levels of style.
-However, the use of this variable as a regressor in the present study might be considered problematic.
-Since the \PGCa\ contains a genitive itself, the regressor variable \textit{Genitive} and the document-level variable \textit{Genitives} are not fully independent.
-Since instances of the \PGCa\ make up for only a minute fraction of all genitives, I still use \textit{Genitives} as a regressor with the appropriate caveats.
-
-Finally, one variable was added as nuisance variable in the context of the present study.
-It was reported in the literature that MNPs in the dative and with a masculine or neuter kind noun favour the \PGCa\ more than the corresponding nominative and accusative MNPs \citep{Hentschel1993,Zimmer2015}.
-As an example, \textit{mit einem Stück frischen Brots} `with a piece of fresh bread' (\PGCa) would be preferred more strongly against \textit{mit einem Stück frischem Brot} (\NACa).
-As with all the examples, native speakers of German will most likely notice that differences are subtle.
-To control for this effect, the case of the measure noun was manually annotated (variable \textit{Measurecase}).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{On statistical analysis}
-\label{sec:rightstatistics}
-
-In this section, I justify my choice of statistical models used in Section~\ref{sec:corpushierarchicalmodel}.
-Readers might think that the method used here for modeling grammatical alternations -- namely (Hierarchical) Logistic Regression\slash Generalised Linear (Mixed) Models or GL(M)Ms -- does not require much justification.
-After all, GLMMs have been established as the major tool in the analysis of alternation phenomena.
-All studies mentioned at the outset of Section~\ref{sec:cogocl} use some form of regression\slash GLMM.
-Over the past few years, however, modified or alternative methods have been proposed.
-While it is impossible to discuss all of these methods here, I just make a few remarks on Bayesian estimation (see \citealp{GelmanEa2014}) as it was proposed in \cite{Levshina2016} and \cite{Divjak2016a}, for example.
-Conceptually, I see three points of discussion that should be kept apart.
-First, Bayesian methods are sometimes touted as superior tools for scientific inference compared to frequentist methods.
-Second, it has been proposed that the Bayesian interpretation of probability is more cognitively adequate for the modeling of linguistic data \cite[301--302]{Divjak2016a}.
-Third and very specific to this paper, given established methods in the modeling of alternation and variation, it has to be decided whether so-called Bayesian methods lead to substantially different results.
-
-% As for the first point, this is not quite the place to discuss it fully.
-% The basic distinction is a philosophical one and related to the concepts of \textit{direct} and \textit{inverse probability} (\eg\ \citealp{Senn2011}).
-% Frequentists assume that models and parameters are fixed, for example a model specifying that a coin is fair.
-% We can then calculate for observed data (for example a measurement of 3 heads in an experiment with 10 tosses) how often such a result or a more extreme result would occur if the model were true and we repeated the experiment arbitrarily often.
-% This is essentially the frequentist notion of direct probability, \ie\ long-run frequencies under replication.
-% Standard tests in the Fisher and Neyman-Pearson traditions as well as Neyman confidence intervals are based on this concept of probability.
-% Bayesian approaches (in the now common interpretation), on the other hand, condition on the particular data and quantify inductively the probability of model parameters given the data.
-% The parameters are thus not fixed, and the probability is usually equated with researchers' posterior beliefs about model parameters.
-% There is actually a debate among Bayesians about the proper interpretation of Bayesian methods, and whether a notion of hypothesis testing is compatible (or even already contained) in the Bayesian approach.
-% In \citet[10]{GelmanShalizi2013}, the authors -- prominent Bayesians themselves -- acknowledge that a theory of statistical testing is a desideratum, and they state about the standard inductive interpretation of Bayesianism that ``most of this received view of Bayesian inference is wrong,'' and they develop a Bayesian notion of p-values (see also \citealp{Mayo2013}, for a frequentist reply; also \citealp{Senn2011} on different strands of Bayesianism and their stance on inductive vs.\ deductive reasoning, and \citealp{Mayo2011}, for a critical reply to \citealp{Senn2011}).
-% Clearly, in such quarrels between and among camps of philosophers of science and statisticians, it is difficult for mere practitioners to take sides.
-% 
-% These quarrels relate to the second point, however.
-% \citet[301--302]{Divjak2016a} speaks favourably of Bayesian methods because the Bayesian concept of probability is more adequate for cognitive modeling compared to the frequentist one.
-% Her argument is part of a larger body of literature asking for cognitively plausible modeling techniques, for example Naive Discriminative Learning (NDL; \citealp{Baayen2011,BaayenEa2013,MilinEa2016,TheijssenEa2013}).
-% On p.\ 303 of \cite{Divjak2016a}, the author goes on to explicitly mention NDL as well.
-% Yet, neither frequentist nor Bayesian methods were conceived as cognitive models, but as systems of inference for scientists (see above, and see also \citealp[302]{Divjak2016a}).
-% The fundamental question that lurks behind such arguments is how we interpret our statistical models (estimated on corpus data).
-% Are they inductive models of cognitive representations that also human learners would infer from being exposed to the corpus data?%
-% \footnote{In which case we are doing ``data science in language research'' in the words of \citealp{MilinEa2016}.
-% I see this as standing in contradiction to the view advocated in \cite{Dabrowska2016} as cited above.}
-% Or are they tests of theories that are pre-specified and merely tested for predictive accuracy on linguistic output data contained in corpora?
-% In the former case, we adopt a strong \textit{corpus as input} hypothesis and should definitely resort to methods like NDL.
-% However, this would most likely require us to toss most previous work done in (cognitive) linguistics into the bin and to abandon all high-level generalisations that most existing studies have been based upon.
-% In fact, arguments to this extreme effect and against using high-level generalisation have been made, for example in \cite{BaayenEa2016}, \citet[299--300]{Divjak2016a}, \cite{RamscarPort2016}, \cite{TheijssenEa2013}.
-% In the latter and less extreme case, the cognitive commitment does, however, not necessarily extend to the statistical methods used.
-% These methods then do not need to be any more cognitively plausible than an ANOVA used to analyse the results from an experiment.
-% I view my own work (see Sections~\ref{sec:germanmeasurenps}--\ref{sec:experimental}) in the tradition of testing theories (which embrace high-level generalisations), and I agree to provisionally use Likelihood methods (see below).
-% I remain fully agnostic with respect to the question, which approach is \textit{right}.
-% The best strategy for cognitive linguistics as a field might be to cultivate many methods while making sure that each method is applied carefully and competently.
-
-The first and second point cannot be discussed in detail here.%
-\footnote{However, there is a number of critical papers by statisticians (even prominent Bayesians) and philosophers of science in which the extreme credibility that has recently been assigned to (subjective and automatic) Bayesian approaches (and especially the standard inductivist interpretation of Bayesianism) is questioned (for example, \citealp{GelmanShalizi2013,Mayo2011,Senn2011}).}
-With regard to the third point, \citet[251--252]{Levshina2016} argues for Bayesian estimation in mixed regression settings.
-First, she claims that ``while frequentist statistics only allows one to test whether the null hypothesis can be rejected, Bayesian statistics enables one both to test the null hypothesis and to estimate the probability of specific parameter values given the data.''
-This does not do justice to frequentist methods in that a strong focus on the rejection of the null hypothesis is characteristic only of Fisher's approach.
-In the Neyman-Pearson approach, results ideally \textit{favour} the main hypothesis vis-à-vis the alternative hypothesis (cf.\ \citealp{Lehmann1993,Lehmann2011,Perezgonzalez2015}).
-Also, especially Neyman-style frequentism has well-known extensions to estimation, for example in the form of confidence intervals (see \citealp{GreenlandEa2016}, esp.\ p.\ 340).
-She then explains that a ``distinctive feature of Bayesian statistics is the use of so-called priors'' and that ``posterior probabilities depend on both the prior beliefs and the data, whereas the results of a frequentist model depend only on the data'' \cite[252]{Levshina2016}.
-Remarkably, given this statement, she does \textit{not} use informative priors and in her footnote 8 \cite[252]{Levshina2016} admits that priors were probed using trial and error.
-So, the proclaimed major advantage of Bayesian modeling was apparently not taken advantage of.%
-\footnote{In the words of \cite{Senn2011}: ``You may believe you are a Bayesian but you are probably wrong.''
-\citet[347--348]{GelmanHill2006} ``view any noninformative prior distribution as inherently provisional'' and give recommendations how to proceed once posteriors have been obtained from noninformative priors.}
-Now, Maximum Likelihood Estimation (MLE) -- the traditional method which could have been used instead -- is not exactly \textit{frequentist} in the sense of Neyman-Pearson testing theory.
-MLE, like inductive Bayesianism, conditions on the particular data inasmuch as it searches for the most likely set of parameters given the data.
-What is more, Bayesian estimators are in fact based on the Likelihood and merely multiply it by the prior \citep[6--8]{GelmanEa2014}.
-If the prior is flat, results converge (see also \citealp[347]{GelmanHill2006}).
-The same is true if the sample size is large compared to the number of parameters, at least for finite-dimensional parameter models \citep[1119--1120]{Freedman1999}, a well-established result known as the \textit{Bernstein-von Mises theorem}.
-With a modest model structure including 17 fixed effects and 2,646 data points in \cite{Levshina2016}, it is highly likely that the same results would have been obtained with Maximum Likelihood methods.
-In fact, she admits that changing the priors did not lead to substantially different results in her footnote 8.
-This is a clear sign that the prior is ``swamped by the data'' \citep[1119]{Freedman1999}.
-% Going back to the argument that Bayesian methods are allegedly more cognitively plausible than `frequentist' methods, the claim might still be true.
-% However, most similar studies (dating back to \citealp{BresnanEa2007}) use an estimator (MLE) that is not frequentist in a narrow sense, and which is as cognitively plausible as Bayesian estimators because it most likely produces the same results in these studies.
-
-If there had been evidence in Levshina's study that Bayesian and MLE methods did \textit{not} converge, it would have been an occasion to demonstrate the selective superiority of the algorithms used in Bayesian estimation.
-After all, there are situations where Bayesian estimators can be more robust, namely with heavily censored data, complex hierarchical models, perfect separation, etc.
-(see \citealp{Freedman1999}, \citealp[345--348]{GelmanHill2006}).
-I want to state clearly that these points do not in any way invalidate the results presented in \cite{Levshina2016}.
-However, being ``Bayesian'' is most likely not among its selling points.
-Additionally, I want to voice the concern that many practitioners are probably already struggling with getting an adequate grasp of advanced statistical methods and that it might therefore be wise to use the more conservative and better understood method if the alternative method is not absolutely required for substantive reasons.
-In the next section, I estimate the parameters of my hierarchical model with MLE and Markov-Chain Monte Carlo (MCMC) methods (the currently most prominent estimator used in Bayesian settings) to demonstrate their expectable convergence.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{A hierarchical model of the measure noun alternation}
-\label{sec:corpushierarchicalmodel}
-
-\begin{figure}[hb!]
-  \centering
-  \includegraphics[width=0.85\textwidth]{../R/output/corpus_fixeffs_mle+mcmc}
-  \caption{Coefficients (MLE and MCMC) with 95\% confidence intervals (for details see text); the intercept is -4.328 (MLE) and -4.441 (MCMC)}
-  \label{fig:fixeffs}
-\end{figure}
-
-In this section, I report the results of fitting a multilevel model to the data using R \citep{R}, \textit{lme4} \citep{lme4} for Maximum Likelihood Estimation (MLE) and \textit{rstanarm} \citep{rstanarm} for Bayesian Markov-Chain Monte Carlo (MCMC) estimation (see Section~\ref{sec:rightstatistics}).
-The purpose is to model the influence of the regressors specified in Table~\ref{tab:variables} on the probability that the \PGCa\ is chosen over the \NACa.
-All regressors from Table~\ref{tab:variables} were included, and the measure lemma and the kind noun lemma were specified as varying-intercept random effects.
-The sample size was \textit{n=}5,063 with 1,134 cases of \PGCa\ and 3,929 cases of \NACa.
-The results of the estimation are shown in Table~\ref{tab:bigtable} and in Figure~\ref{fig:fixeffs}.
-The intercept comprises \textit{Cardinal=Yes}, \textit{Measurecase=Nom}, \textit{Kindgender=Masc}, \textit{Measureclass=Physical}, and 0 for all numeric z-transformed regressors.
-It is -4.328 for the ML estimate and -4.441 for the MCMC estimate.
-
-The regressors with the measure lemma as their unit of reference have no within-measure lemma variance, and the \textit{glmer} function automatically estimates them as \textit{group level predictors} (or \textit{second-level effects}), cf.\ \citet[265--269,302--304]{GelmanHill2006}.
-The same goes for those listed with the kind lemma as their unit of reference.
-Given the coding of the response variable, coefficients leaning to the positive side can be interpreted as favouring the \PGCa.
-
-\begin{sidewaystable}
-  \centering
-  \resizebox{\textheight}{!}{
-  \begin{tabular}{llrlp{0.5em}rrp{0.5em}rrp{0.5em}rrp{0.5em}cc}
-      Model & Regressor & \multicolumn{1}{l}{\pPB} & Factor && \multicolumn{2}{l}{Coefficient} && \multicolumn{2}{l}{CI low} && \multicolumn{2}{l}{CI high} && \multicolumn{2}{l}{CI excludes 0} \\
-      level &                   &       & level     && \multicolumn{1}{l}{MLE} & \multicolumn{1}{l}{MCMC} && \multicolumn{1}{l}{MLE} & \multicolumn{1}{l}{MCMC} && \multicolumn{1}{l}{MLE} & \multicolumn{1}{l}{MCMC} && \multicolumn{1}{l}{MLE} & \multicolumn{1}{l}{MCMC} \\\midrule
-       First     & Badness           &  0.002 &           && -0.152 & -0.155 && -0.247 & -0.247 && -0.061 & -0.065 && * & * \\
-                 & Cardinal          &  0.001 & No        &&  1.189 &  1.222 &&  0.862 &  0.927 &&  1.466 &  1.496 && * & * \\
-                 & Genitives         &  0.001 &           && -0.693 & -0.711 && -0.768 & -0.801 && -0.592 & -0.616 && * & * \\
-                 & Measurecase       &  0.001 & Acc       &&  0.030 &  0.031 && -0.150 & -0.159 &&  0.212 &  0.222 &&   &   \\
-                 &                   &        & Dat       &&  0.705 &  0.729 &&  0.455 &  0.465 &&  0.944 &  0.995 && * & * \\[0.5\baselineskip]
-
-       Second    & Kindattraction    &  0.020 &           &&  0.225 &  0.244 &&  0.049 &  0.056 &&  0.393 &  0.422 && * & * \\
-       (Kind)    & Kindfreq          &  0.095 &           &&  0.146 &  0.164 && -0.023 & -0.016 &&  0.301 &  0.341 &&   &   \\
-                 & Kindgender        &  0.001 & Neut      &&  0.021 &  0.013 && -0.367 & -0.409 &&  0.392 &  0.435 &&   &   \\
-                 &                   &        & Fem       &&  1.269 &  1.289 &&  0.800 &  0.788 &&  1.709 &  1.783 && * & * \\[0.5\baselineskip]
-
-       Second    & Measureattraction &  0.001 &           &&  0.282 &  0.299 &&  0.106 &  0.102 &&  0.447 &  0.515 && * & * \\
-       (Measure) & Measureclass      &  0.001 & Container &&  0.252 &  0.257 && -0.265 & -0.303 &&  0.788 &  0.813 &&   &   \\
-                 &                   &        & Rest      &&  0.421 &  0.379 && -0.209 & -0.378 &&  1.063 &  1.091 &&   &   \\
-                 &                   &        & Amount    &&  0.831 &  0.889 &&  0.215 &  0.220 &&  1.432 &  1.569 && * & * \\
-                 &                   &        & Portion   &&  1.217 &  1.253 &&  0.675 &  0.689 &&  1.684 &  1.840 && * & * \\
-                 & Measurefreq       &  0.005 &           && -0.231 & -0.232 && -0.363 & -0.395 && -0.079 & -0.073 && * & * \\
-
-  \end{tabular}
-  }
-  \caption{Coefficient table comparing Maximum Likelihood Estimation (MLE, with 95\% bootstrap confidence interval) and `Bayesian' Markov-Chain Monte Carlo estimation (MCMC); the intercept is -4.328 (MLE) and -4.441 (MCMC)}
-  \label{tab:bigtable}
-\end{sidewaystable}
-
-Standard diagnostics for MLE show that the model quality is quite good.
-Generalised variance inflation factors for the regressors were calculated to check for multicollinearity \citep{FoxMonette1992,ZuurEa2010}, and none of the corrected $\text{GVIF}^{1/2\text{df}}$ was higher than 1.6.
-Nakagawa \& Schielzeth's pseudo-coefficient of determination is $R_m^2=0.409$ and $R^2_c=0.495$ (see \citealp{Gries2015} for a basic introduction to these $R^2$ measures, or else \citealp{NakagawaSchielzeth2013}).
-The rate of correct predictions is 0.843, which means a proportional reduction of error of $\lambda=0.297$.
-The lemma intercepts have standard deviations of $\sigma_{\text{Measurelemma}}=0.448$ and $\sigma_{\text{Kindlemma}}=0.604$.
-
-The coefficient estimates are specified in Table~\ref{tab:bigtable} for each regressor (or regressor level) in the columns labelled \textit{Coefficient}.
-For a robust quantification of the precision of the estimation, I ran a parametric bootstrap (using the \mbox{\textit{confint.merMod}} function from \textit{lme4}) with 1,000 replications and using the percentile method for the calculation of the intervals.
-The resulting 95\% bootstrap confidence intervals are reported in Table~\ref{tab:bigtable} in the columns labelled \textit{CI low} and \textit{CI high} (= upper and lower 2.5th percentiles).
-The column \textit{CI excludes 0} shows an asterisk for those intervals that do not include 0.
-Furthermore, for each regressor, a p-value was obtained by dropping the regressor from the full model, re-estimating the nested model, and comparing it to the full model.
-Instead of inexact Wald approximations and Likelihood Ratio Tests, I used a drop-in bootstrap replacement for the Likelihood Ratio Test from the function \textit{PBmodcomp} from the \textit{pbkrtest} package \citep{HalekohHojsgaard2014}.
-I call the corresponding value $p_{\text{PB}}$, and it is given in the respective columns in Table~\ref{tab:bigtable}.
-Only \textit{Kindfreq} ($\mpPB=0.095$) can be seen as slightly too high to be convincing (non-significant).
-
-Finally, I show that MCMC does not necessarily lead to different results.
-Actually, given the low complexity of the model and the large sample size, it would be surprising if they did (see Section~\ref{sec:rightstatistics}).
-The model was re-estimated using the \textit{stan\_glmer} function from the \textit{rstanarm} package, which provides an \textit{lme4}-compatible syntax for estimating common model types with the \textit{stan} software \citep{CarpenterEa2017}.
-The algorithm was run with 4 chains and 1,000 iterations, and I used plausible default priors.
-Most notably, priors for coefficients were specified as $\mathcal{N}(0,10)$ because coefficients higher than 10 or lower than $-10$ are extremely rare in well-specified models on appropriate data.%
-\footnote{Consider that with a coefficient of 10, each increase by 1 in the regressor variable increases the odds by $exp(10)=22,026.47$.
-To reliably estimate such coefficients, extremely large samples would be required.}
-The algorithm converged, and for all coefficients, the $\hat{R}$ diagnostic was exactly 1.
-The resulting coefficients and intervals as well as an * for ``confidence interval does not contain 0'' are also given in Table~\ref{tab:bigtable} in the columns labelled \textit{MCMC}.
-Both methods lead to exactly the same results (minus negligible numerical differences) as expected given the modest complexity of the model structure and the large sample size.
-The signs and magnitudes of the coefficients are identical, and the confidence intervals have the same width and symmetry properties.
-Figure~\ref{fig:fixeffs} illustrates this by also showing both estimates.
-Since both estimators converge, I only interpret the MLE model in the next section.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Interpretation}
-\label{sec:interpretation}
-
-% ATTRACTION
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width=0.5\textwidth]{../R/output/corpus_Measureattraction}~\includegraphics[width=0.5\textwidth]{../R/output/corpus_Kindattraction}
-  \caption{Effect plots for the regressors \textit{Measureattraction} and \textit{Kindattraction}; y-axes are not aligned}
-  \label{fig:eff:attraction}
-\end{figure}
-
-The results reported in Section~\ref{sec:corpushierarchicalmodel} generally confirm the hypotheses from Section~\ref{sec:analyses}.
-First, the prototypicality effect related to the non-alternating \PGCd\ and \NACb\ can be shown (see the effect plots in Figure~\ref{fig:eff:attraction}).%
-\footnote{Effect plots were created using the \textit{effects} package \citep{Fox2003}.
-They show the changes in probability for the outcome (y-axis) dependent on values of a regressor (x-axis) at typical values of all other regressors.
-The vertical bars (categorical variables), and the grey areas (continuous variables) are asymptotic 95\% confidence intervals calculated from \textit{glmer}.
-They are not bootstrapped.
-Readers should be aware that the axes are specifically scaled so as to result in a linear plot, and that the range of the axes varies between plots.}
-The effect is as expected:
-if a lemma appears relatively more often in the \PGCd\ (compared to its frequency in the \NACb), the \PGCa\ tends to be chosen over the \NACa\ with this specific lemma.
-The effect for measure nouns is stronger, and it was estimated with higher precision.
-
-An interesting picture emerges for the lemma frequencies.
-A higher-than-average lemma frequency of measure nouns favours the \NACa\ ($\beta_{\text{Measurefreq}}=-0.231$, $\mpPB=0.005$), which is as expected if we assume at least a tendency for highly grammaticalised items to be more frequent.
-With kind nouns, higher frequency seems to favour the \PGCa ($\beta_{\text{Kindfreq}}=0.146$, $\mpPB=0.095$).
-However, there is no clear theoretical interpretation (see Section~\ref{sec:analyses}), and the estimate is imprecise (not significant at $\alpha=0.05$, see above).
-The effect can therefore be ignored or treated as a nuisance variable.
-
-% GRAMMATICALISATION
-% => lemmas and lemma classes
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width=0.5\textwidth]{../R/output/corpus_Measureclass}
-  \caption{Effect plot for the regressor \textit{Measureclass}}
-  \label{fig:eff:measureattraction}
-\end{figure}
-
-In Section~\ref{sec:analyses}, it was also hypothesised that classes of measure nouns with a higher degree of grammaticalisation should favour the \NACa.
-The \textit{Measureclass} second-level predictor was successfully estimated ($\mpPB=0.001$).
-Looking at the effect plot in Figure~\ref{fig:eff:measureattraction}, it is evident that abstract non-referential physical measure nouns (such as \textit{Gramm} `gram' or \textit{Liter} `litre') with a high degree of grammaticalisation favour the \NACa.
-At the other end of the scale, nouns denoting natural portions like \textit{Haufen} `heap', \textit{Bündel} `bundle', \textit{Schluck} `gulp' favour the \PGCa.
-These are referential nouns, confirming the hypothesis that it is prototypical of the PGC to contain two referential nouns, while the NAC prototypically only contains one (the kind noun).
-
-% CARDINALS
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width=0.5\textwidth]{../R/output/corpus_Cardinal}
-  \caption{Effect plot for the regressor \textit{Cardinal}}
-  \label{fig:eff:leftcontext}
-\end{figure}
-
-I now turn to the predicted effect of cardinals as modifiers of the measure noun.
-Figure~\ref{fig:eff:leftcontext} shows that cardinals indeed influence the choice of the alternant ($\mpPB=0.001$), and that cardinals have a strong tendency to co-occur with the \NACa.
-This effect was predicted in Section~\ref{sec:analyses}.
-
-% REGISTER
-% => Badness
-% => Genitives
-
-The style-related proxy variables point to the expected direction.
-Increased \textit{Badness} of the document favours the \NACa\ ($\beta_{\text{Badness}}=-0.152$, $\mpPB=0.002$), and so does a lower density of genitives ($\beta=-0.693$, $\mpPB=0.001$).
-While these are merely proxies to style (and partially circular in the case of \textit{Genitives}), this result can at least encourage future work into stylistic effects.
-
-% NUISANCE
-% => dative effect
-% => frequency
-
-The influence of \textit{Measurecase} ($\mpPB=0.001$) is as predicted in previous analyses (see Section~\ref{sec:analyses}).
-A measure noun in the dative favours the \PGCa\ with $\beta_{\text{MeasurecaseDat}}=0.705$ (compared to the nominative, which is on the intercept).
-Although \textit{Measurecase} is a nuisance variable in the context of this study, convergence with previous work strengthens its validity.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Correlating corpus findings with experiments}
-\label{sec:experimental}
-
-\subsection{Experiment 1: forced-choice}
-\label{sec:exp:fc}
-
-In the two experiments reported in this section, I use probabilities for the alternating constructions calculated for attested material, and I correlate these probabilities with the participants' reactions.
-Thus, a direct link can be established between output material found in corpora and the behaviour of linguistic agents (see also Section~\ref{sec:cogocl}).
-Both experiments use sentences containing attested MNPs from the corpus sample (embedded into simplified sentences) as stimuli.
-Also, the probabilities that the corpus-based model assigns to the two alternants in these sentences are used as the main regressor in both studies.
-
-The first experiment tests preferences for constructions explicitly.
-\cite{FordBresnan2013} use the \textit{split-100} task in which participants have to distribute 100 points between the alternatives, assigning more points to more natural sounding alternative.
-In essence, participants distribute a probability mass between two alternants, which is intended to produce more subtle results compared to a two-alternative forced-choice task such as in \cite{Rosenbach2013}, where participants have to choose one of two alternants.
-The split-100 paradigm has been criticised in \cite{ArppeJaervikivi2007}.
-The criticism was reiterated in \cite{DivjakEa2016}, where they use a forced-choice task.
-In \cite{VerhoevenTemme2017}, it was shown that results from forced-choice and split-100 experiments mostly converge (with some numeric intricacies related to the non-linear distribution of preferences for alternants).
-I present a forced-choice experiment in this Section and not a split-100 experiment, mainly because in a dry run of the experiment, participants complained about the unnaturalness of distributing a probability mass across two alternants and tended to produce ratings of 0 and 100 (and to a lesser degree 50).
-Participants had to choose between two sentences that differed only in that one contained the \NACa\, and the other one contained the \PGCa.
-The analysis compares the probabilities assigned to the stimuli by the corpus-derived model with the frequency with which participants chose the alternants for the same stimuli.
-
-There were 24 participants (native speakers of German without reading or writing disabilities) aged 19 to 30 living permanently in Berlin, who were recruited from introductory linguistics courses at Freie Universität Berlin.
-Although the experiment was conducted in the last four weeks of their first semester, participants had no deeper explicit knowledge of linguistics, grammar, or experimental methods.
-None of them had ever participated in a forced-choice experiment before.
-Participation was voluntary but participants received credit in partial fulfillment of course requirements.
-
-As stimuli, attested MNPs from the corpus study were used, but the sentences were radically simplified to avoid influences from contextual nuisance factors as much as possible.
-The approach is also justified because according to the theoretical assessment in Section~\ref{sec:analyses}, the choice of alternants depends mostly on a very local constructional context.
-I sampled 16 MNPs from the concordance and made sure that the simplifications and normalisations did not affect any of the regressors used in the corpus study.
-In the simplified sentences, the case, number, etc. of the MNP remained the same as in the attested sentence, as did the choice of lexical material within the MNP.
-Eight sentences contained masculine or neuter kind nouns, and the other eight contained feminine kind nouns.
-Furthermore, in each of the masculine\slash neuter and feminine groups, four sentences originally containing the \NACa\ and four sentences originally containing the \PGCa\ were chosen.
-More precisely, the sentences were sampled as \textit{highly prototypical examples} of \PGCa\ (high probability assigned by GLMM) and \NACa\ (low probability assigned by GLMM), respectively.%
-\footnote{Remember from Section~\ref{sec:corpusstudies} that the model predicts the probability that the \PGCa\ is chosen over the \NACa.}
-High and low probabilities were defined as the top and bottom 20\% of all probabilities assigned by the GLMM.
-Lemmas and feature combinations were made unique within each group whenever possible.
-The design is summarised in Table \ref{tab:experiment1:design}.
-
-\begin{table}
-  \centering
-  \begin{tabular}[h]{lll}
-     & Masculine\slash Neuter & Feminine \\
-     \midrule
-     \textbf{high prob.\ for PGC\Subsf{adj}} & 4 sentences & 4 sentences \\
-     \textbf{low prob.\ for PGC\Subsf{adj}} & 4 sentences & 4 sentences \\
-  \end{tabular}
-  \caption{The four groups of sentences chosen as stimuli; in each group of four sentences, combinations of important factor values were made unique whenever possible}
-  \label{tab:experiment1:design}
-\end{table}
-
-The final pairs of stimuli were the sentence containing the attested and preferred alternant (according to the corpus GLMM) on the one hand and a modified version containing the dispreferred alternant on the other hand.
-They were presented next to each other, and a 20 second time limit for each choice was set.%
-\footnote{No participant ever exceeded the time limit.}
-The position on the screen (left\slash right) and the order of sentences were randomised for each participant.
-As fillers, 23 pairs of sentences exemplifying similar but unrelated alternation phenomena from German morpho-syntax were used.
-Thus, participants saw 39 pairs of sentences and 78 sentences in total.
-They were instructed to select from each pair of sentences the one that seemed more natural to them in the sense that they would use it rather than the other one.
-The experiment was conducted using \textit{PsychoPy} \citep{Peirce2007}.
-
-\begin{figure}[hb!]
-\centering
-\includegraphics[width=0.5\textwidth]{../R/output/fc_effects}
-\caption{Effect plot for the multilevel logistic regression in the forced-choice experiment: predictability of participants' choices using the probabilities derived from the corpus-based GLMM}
-\label{fig:afc:effects}
-\end{figure}
-
-Then, a multilevel logistic regression was specified with the probability of the \PGCa\ predicted for each sentence by the corpus-based GLMM as the only fixed effect \textit{Modelprediction}.%
-\footnote{The document-level variables \textit{Badness} and \textit{Genitives} were set to 0, which is the mean for z-transformed variables.}
-A random intercept and slope were added for the individual sentence (item) in order to catch idiosyncrasies of single sentences.
-Also, a random intercept and slope for participants was added.%
-\footnote{The random slopes were added to comply with \citet[257]{BarrEa2013} who predict \textit{catastrophically high Type I error rates} for experimental designs with within-subject manipulations if random effects structures are not kept maximal.
-Notice that the model reported here was estimated with conceptually identical results with regard to the predictor of interest (\textit{Modelprediction}) if only random intercepts were used.}
-Coefficients were estimated with Maximum Likelihood Estimation (\textit{lmer} function from \textit{lme4}).
-The number of observations was \textit{n=}384.
-
-A certain amount of the variance can be accounted for by idiosyncrasies of single sentences ($\sigma_{\text{Sentence}}=1.785$, $\sigma^{\text{Modelprediction}}_{\text{Sentence}}=5.996$, 16 levels).%
-\footnote{I use $\sigma^f_r$ to denote the standard deviation of the random intercepts for the fixed effect $f$ varying by random effect $r$.}
-Also, among participants, there are clearly different preferences ($\sigma_{\text{Participant}}=0.781$, $\sigma^{\text{Modelprediction}}_{\text{Participant}}=0.484$, 24 levels).
-On the extreme ends, one participant chose the \PGCa\ in 13 of 16 cases, and two participants only chose it in 5 of 16 cases.
-The regressor \textit{Modelprediction} achieves $\mpPB=0.007$ (1,000 replications) and is estimated at 5.408 relative to an intercept of -1.304.
-The confidence interval from a parametric bootstrap (1,000 replications, percentile method) for the regressor is acceptable but slightly large with a lower bound of 1.626 and an upper bound of 8.397.
-The pseudo-coefficients of determination are $R^2_{m}=0.227$ and $R^2_{c}=0.561$, which means that over 22\% of the variance in the data can be explained by considering only the predictions from the corpus-based GLMM.
-The effect display for the single fixed regressor \textit{Modelprediction} is given in Figure \ref{fig:afc:effects}.
-The result is very clear.
-The higher the probability of the \PGCa\ predicted from usage data, the more often participants chose the \PGCa\ alternant in the forced-choice task.
-In summary, the forced-choice experiment clearly succeeded in corroborating the results from the corpus study in as much as the preferences extracted from usage data correspond to native speakers' choices.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Experiment 2: self-paced reading}
-\label{sec:exp:spr}
-
-The second experiment tests preferences more implicitly.
-It is expected that reading less prototypical alternants (the one assigned a low probability by the corpus-derived model) in a given context and with given lexical material incurs a processing overhead for the reader (\citealp{Kaiser2013}).
-In this section, a self-paced reading experiment is therefore presented.
-In a very similar fashion, \cite{DivjakEa2016} apply the self-paced reading paradigm in the validation of corpus-based models.
-The analysis compares the corpus-derived probabilities with potential lags in reading time for sentences with the preferred and the non-preferred constructions.
-
-Concretely, the exact same stimuli as in the forced-choice experiments were used.
-Each participant read both the 16 sentences with the alternant predicted by the corpus model and the 16 modified sentences with the alternant that the corpus model did not predict.%
-\footnote{Notice that lemmas and their frequencies as well as lemma classes are included as regressors in the corpus-based GLMM, and there was consequently no additional controlling of lemma frequencies, etc.}
-To minimise repetition effects, the stimuli for each participant were separated into two blocks of 16 targets and 33 fillers per block.
-In the experiment, participants first read all sentences from the first block, then all sentences from the second block.
-From each target sentence pair, one sentence was assigned to the first block and the other sentence to the other block.
-The assignment of members of the individual sentence pairs to the blocks was randomised for each participant individually, as was the order within each block.
-The sentences from each pair of alternants were kept as far apart as possible.
-The fillers also came in pairs such that the second block exclusively contained sentences to which participants had been exposed in the first block in slightly modified form.
-In total, each participant read 98 sentences.
-After each sentence, participants had to answer simple (non-metalinguistic) yes-no questions about the previous sentence as distractors.
-The distractor questions were different between the first and the second blocks.
-There were 38 participants recruited in exactly the same manner as for the experiment reported in Section~\ref{sec:exp:fc}.
-None of them had ever participated in any kind of reading experiment, and none of them took part in the first experiment.
-The experiment was conducted using \textit{PsychoPy}.
-
-The reading times were residualised per speaker based on the reading times of all words (not just the targets) by that speaker.
-The adjective and the kind noun (\ie\ the constituents bearing the critical case markers) were used as the target region, for instance the bracketed words in the example \textit{zwei Gläser} [\textit{spru\-deln\-des Wasser}] `two glasses of sparkling water'.
-Outliers farther than 2 inter-quartile ranges from the mean logarithmised residualised reading time were removed (64 data points), resulting in a total number of \textit{n=}1,152 observations.
-An LMM was specified with the logarithmised residual reading times as the response variable.
-
-The probabilities derived from the corpus GLMM (\textit{Modelprediction}) were added as the main regressor of interest.
-It should be remembered that the corpus GLMM predicts the probability of the \PGCa.
-As a consequence, the higher the GLMM prediction is, the more prototypical the sentence is for containing the \PGCa.
-It is therefore expected that reading times are higher when the value of \textit{Modelprediction} is higher but the sentence contains the \NACa.
-However, when the sentence contains the \PGCa, reading times should be lower when \textit{Modelprediction} is higher.
-To account for this, an interaction between \textit{Modelprediction} and \textit{Construction} (levels \textit{PGCa} and \textit{NACa}) was added to the model.
-
-Furthermore, the position (1--98) of the sentence in the individual experiment (\textit{Position}) was included as a fixed effect to control for the usual increase in reading speed during an experiment run.
-Random intercepts were specified for \textit{Participant} and \textit{Item} (the 16 sentence pairs are one \textit{Item} each).%
-\footnote{I tried random slopes in order to keep the random effect structure maximal \citep{BarrEa2013} but it was impossible to get the algorithm to converge due to the added complexity of the interaction.}
-
-\begin{table}
-  \centering
-  \begin{tabular}{lrrrc}
-    Regressor & \multicolumn{1}{r}{Coefficient} & \multicolumn{1}{r}{CI low} & \multicolumn{1}{r}{CI high} & \multicolumn{1}{r}{CI excludes 0} \\ \midrule
-    ConstructionPGCa                 &  0.054 &  0.012 &  0.095 &  *  \\ 
-    Modelprediction                  & -0.006 & -0.113 &  0.110 &     \\ 
-    Position                         & -0.005 & -0.005 &  0.004 &     \\ 
-    ConstructionPGCa:Modelprediction & -0.125 & -0.234 & -0.023 &  *  \\ 
-  \end{tabular}
-  \caption{Fixed effect coefficient table for the LMM used to analyse the self-paced reading experiment; the intercept is 0.829}
-  \label{tab:exp:spr}
-\end{table}
-
-Table~\ref{tab:exp:spr} shows the coefficient estimates with a 95\% parametric bootstrap confidence interval (1,000 replications, percentile method).
-The standard deviation of the participant intercepts is $\sigma_{\text{Participant}}=0.079$ and of the item intercepts $\sigma_{\text{Item}}=0.037$.
-Comparing the full model to a model without the main regressor \textit{Modelprediction} (and consequently also without the interaction with \textit{Construction}) in a PB test gives $\mpPB=0.036$.
-The pseudo-coefficients of determination are $R^2_m=0.237$ and $R^2_c=0.346$.
-There is thus some inter-subject variation and a noteworthy correlation between the reactions of the subjects and the corpus-derived probabilities.
-\hl{The fact that the correlation is not stronger can be explained in parts \ldots}
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.5\textwidth]{../R/output/spr_effects}
-\caption{Effect plot for the LMM in the self-paced reading experiment: modeling participants' residualised log reading times on the probabilities given by the corpus-based GLMM}
-\label{fig:spr:effects}
-\end{figure}
-
-The effect plot for the effect of interest is shown in Figure~\ref{fig:spr:effects}.
-The estimate for the sentences with \NACa\ is obviously imprecise and no differences in reading times are observed.
-There is a clearer effect in the sentences with \PGCa, which is also confirmed by the significant results from the bootstrapped confidence intervals (see Table~\ref{tab:exp:spr}) and from the PB test reported above.
-The \PGCa\ brings about an increased reading time ($\beta_{\text{ConstructionPGCa}}=0.054$), which is plausible because it is the much rarer construction (see Section~\ref{sec:corpusstudies}).
-However, if it occurs in a prototypical context and with prototypical lexical material, reading times drop ($\beta_{\text{ConstructionPGCa:Modelprediction}}=-0.125$).
-This can be seen in the downward slope in the right panel of Figure~\ref{fig:spr:effects}.
-This fits into the general picture inasmuch as the construction with the lower frequency might be developing towards a more sharply defined prototype.%
-\footnote{In this context, it should be remembered from Section~\ref{sec:gettingdata} that even the \PGCd\ is much rarer that the \NACb\ (17,252 vs.\ 315,635 occurrences in the auxiliary corpus samples).}
-Conversely, the \NACa\ (like the NAC in general) might be the highly frequent default which does not incur a reading time penalty, even if it is not the optimal choice in the given context and with the given lexical material.
-
-This concludes the report of the two experiments.
-In the final section, I take stock and summarise the contribution of the present study to the research on alternations in cognitive linguistics.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Conclusions}
-\label{sec:conclusion}
-
-This paper stands in a now ten year-old tradition of research on grammatical alternations using corpus and experimental data.
-In my view, the main tenets of this line of research are:
-(i)~Language, viewed from a cognitively realistic angle, is a probabilistic phenomenon and cannot be modelled appropriately within Aristotelian frameworks that assume discrete categories.
-(ii)~Corpora are collections of usage events (language production) and can therefore be used to evaluate both the claim made in (i) and specific theoretical claims (in case studies) about factors influencing speakers' decisions to use specific forms or constructions.
-(iii)~Given (ii), we expect results from corpus analyses and from appropriate experiments to yield similar results, not necessarily as a form of validation of the corpus-based findings, but as converging evidence.
-I consider these three points to be of utmost importance because they clearly set this approach to linguistic research apart from \textit{both} Aristotelian frameworks \textit{and} introspective, non-empirical, and anti-quantitative versions of cognitive linguistics (see \citealp{Dabrowska2016} for a pithy philippic against such approaches).
-
-The present paper adds to the evidence that all of the aforementioned three points are correct.
-A grammatical alternation in German measure NPs was examined using corpus data based on factors partly derived from existing accounts, formulated in terms of construction prototypes.
-The preferences extracted from the DECOW web corpus were confirmed in a forced-choice experiment, in which participants explicitly chose alternants in line with the probabilities derived from the corpus-based model.
-In a more implicit self-paced reading experiment, it was shown that the much rarer alternant brings about a reading time penalty except in cases for which the corpus model predicts very high probability for this alternant.
-
-Future work could extend these results and provide a general picture of the constructions expressing measurements (see Section~\ref{sec:descriptive}).
-This would be a much more complicated task given that the choices then would no longer be binary and that the meaning of the alternative ways of expressing measurements are semantically more varied.
-Finally, I want to point out that German is mildly under-researched in the specific framework used here.
-This is quite surprising given the fact that German morpho-syntax is famous for its alternations, which are usually called \textit{Zweifelsfälle} (`cases of doubt') in the traditional literature \citep{Duden09,Klein2009}.
-Instead of being drowned in normative, descriptive, or didactic discussions, they could serve as ideal test cases in cognitive linguistics.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\include{sections/1intro}
+\include{sections/2descriptive}
+\include{sections/3corpus}
+\include{sections/4experiments}
+\include{sections/5conclusions}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\bibliography{rs}
 
-\bibliography{rs}
 \end{document}

### 2017-12-22 Latex/sections/1intro.tex
Sections in separate files.

\section{Cognitively oriented corpus linguistics}
\label{sec:cogocl}

This paper deals with a morpho-syntactic alternation between two constructions that occurs only in a very specific type of measure noun phrase in German.
By \textit{alternation} I refer to a situation where two or more forms or constructions are available with no clear difference in acceptability, function, or meaning.
The study of alternations has a long history in cognitively oriented corpus linguistics (\citealp{BresnanEa2007,BresnanHay2010,BresnanFord2010,DivjakArppe2013,Gries2015,NessetJanda2010}, to name just a few publications).
This area of research is based on the assumption that language is a probabilistic phenomenon \citep{Bresnan2007} where alternants are chosen neither deterministically nor fully at random.
Instead, multifactorial models are constructed which incorporate influencing factors from diverse levels, including contextual factors.
The estimation of the model coefficients quantifies the influence that the factors have on the probability that either alternant is chosen.

For such probabilistic generalisations estimated on corpus data, there has always been an interest in correlating the findings with results from experimental work (for example, \citealp{ArppeJaervikivi2007,BresnanEa2007,BresnanFord2010,DivjakGries2008,DivjakEa2016,FordBresnan2013}).
This is often called a \textit{validation} of the corpus-derived findings, but \citet[303]{Divjak2016a} rightly criticises this choice of words ``because it creates the impression that behavioral experimental data is inherently more valuable than textual data,'' citing \cite{TummersEa2005}, who state that a corpus is ``a sample of spontaneous language use that is (generally) realized by native speakers.''
However, as \citet[486--487]{Dabrowska2016} convincingly argues, this does not imply that we can in some way ``deduce mental representations from patterns of use,'' \ie\ from corpus data.%
\footnote{In its colloquial meaning, the verb \textit{deduce} seems appropriate here.
However, in terms of the scientific method, we are most likely talking about \textit{induction} rather than \textit{deduction}.}
It would be highly surprising if this were possible, and the same holds for experimental methods.
Nobody assumes that we can inductively infer mental representations from experiments, which -- as opposed to corpus studies -- even allow for direct access to the cognitive agent and offer much better possibilities to control experimental conditions and nuisance variables.
Rather, under the standard approach, a theory of cognitive representation is pre-specified.
Then, predictions are derived from this theory \textit{before} the experiment or the corpus study is conducted to \textit{test} the theory.
Technical problems of statistical inference aside, a successful experiment corroborates the theory, and a failed experiment weakens the theory.
The simplest assumption would be that the same approach is adequate for corpus studies.

The present study is conducted within the general paradigm of cognitive corpus linguistics and includes a comparison of the corpus findings with results from two experiments.
I assume a model of similarity-based classification in the form of Prototype Theory for modeling alternation.
Protoype Theory has been used in alternation research in cognitively oriented corpus linguistics (see \citealp{DivjakArppe2013}; \citealp{Gries2003}).
It is assumed that the variables annotated in a corpus study on an alternation phenomenon define prototypes for the alternants.
The similarity of a given feature vector (in a concrete sentence where one of the alternants is used) influences to a large extent which alternant is chosen by speakers.
This entails that a variant of Prototype Theory with features is assumed \citep{Rosch1978}, instead of the monolithic prototypes of earlier versions.
Prototype Theory is well suited for modeling constructional choices but it is just one of several similarity-based theories of classification, the most prominent other framework being Exemplar Theory (\citealp{MedinSchaffer1978,Hintzman1986}; see \citealp{StormsEa2000} for a comparison of the theories in experimental settings).
Prototype Theory and Exemplar Theory model essentially the same types of effects and differ mainly in whether they assume higher-level abstractions as part of the cognitive representation (Prototype Theory) or try to do without them (Exemplar Theory).
I am sceptical that corpus analysis alone could ever decide which theory is more suitable, given that substantial doubt has been voiced whether even experimental methods are ultimately able to do so \citep{Barsalou1990}.%
\footnote{In \cite{DivjakArppe2013}, it was shown using corpus data how both models form a converging picture.}
Prototype Theory (with feature abstractions) is preferred here simply because it fits the established alternation modeling paradigm, which relies on features being weighted in statistical models.
In Section~\ref{sec:analyses}, a prototype-theoretical parlance is therefore adopted.

In the remainder of the paper, I introduce the alternation phenomenon (Section~\ref{sec:descriptive}) and suggest a theory-driven set of factors influencing the alternation (Section~\ref{sec:analyses}).
Then the corpus study is presented, including an appropriate statistical analysis (Section~\ref{sec:corpusstudies}).
Two experiments which corroborate the corpus-based findings are then reported in Section~\ref{sec:experimental} before I sum up the paper in Section~\ref{sec:conclusion}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2017-12-22 Latex/sections/2descriptive.tex
Sections in separate files.

\section{Case assignment in German measure NPs}
\label{sec:germanmeasurenps}

\subsection{Two stable cases and a case alternation}
\label{sec:descriptive}

In this section, I introduce and illustrate the relevant alternating constructions.
I describe the narrowly defined syntactic configuration in which the alternation occurs, and I motivate the focus on \textit{only} this narrow range rather than, for example, the whole range of nominal constructions expressing quantities.

I use the term \textit{measure noun phrase} (MNP) to refer to a noun phrase (NP) in which a kind-denoting (count or mass) noun depends on another noun that specifies a quantity of the objects or the substance denoted by the kind-denoting noun.
I call the kind-denoting noun the \textit{kind noun} and the quantity-denoting noun the \textit{measure noun}.
For illustration purposes, in the English \textit{a glass of good wine}, \textit{glass} is the measure noun and \textit{wine} is the kind noun.
Measure nouns can be all sorts of nouns which denote a quantity (such as \textit{litre} or \textit{amount}) but also those denoting containers, collections, etc. (such as \textit{glass} or \textit{bucket}).
Like \citet[284]{Brems2003}, I consider nouns as measure nouns ``which, strictly speaking, do not designate a `measure', but display a more nebulous potential for quantification'' (also \citealp[530]{Koptjevskaja2001}, and \citealp[338]{Rutkowski2007}).

\subsubsection{Core structures related to the alternation}

Three different syntactic configurations within MNPs need to be distinguished, and the case alternation occurs only in one of them.
It occurs only when the kind noun is modified by an attributive adjective and there is no determiner, as in (\ref{ex:intro:alternation}).
Superficially, the sentences are functionally and semantically equivalent either with the kind noun in the genitive (\ref{ex:intro:alternation1}) or in the same case as the measure noun, an accusative in the case of (\ref{ex:intro:alternation2}).%
\footnote{Some descriptive and normative grammars take stronger positions with regard to the acceptability of the two options.
See \cite{Hentschel1993} and \cite{Zimmer2015} for analyses of the sometimes absurd stances taken in grammars of German.
As will be shown (especially in Section~\ref{sec:exp:fc}), there might be preferences, but we cannot assume either construction to be unacceptable.
}

\begin{exe}
  \ex\label{ex:intro:alternation}
  \begin{xlist}
    \ex[ ]{\label{ex:intro:alternation1} \gll Wir trinken [[ein Glas]\Sub{Acc} [guten Weins]\Sub{Gen}]\Sub{Acc}.\\
    we drink a glass good wine \\
    \glt We drink a glass of good wine.}
    \ex[ ]{\label{ex:intro:alternation2} Wir trinken [[ein Glas]\Sub{Acc} [guten Wein]\Sub{Acc}]\Sub{Acc}.}
  \end{xlist}
\end{exe}

This specific configuration has to be seen in the context of two other configurations, to which I turn now.
First, if the kind noun forms an NP with a determiner, the construction resembles (and is usually called) a \textit{pseudo-partitive} (on partitives and pseudo-partitives see, e.\,g., \citealp{Barker1998,Selkirk1977,Stickney2007,Vos1999}; for a recent application of the terminology to German, see \citealp{Gerstenberger2015}).%
\footnote{If the kind noun is definite, the construction instantiates a true partitive.
Whereas partitives are constructions denoting a proper part-of relation as in \textit{a sip of the wine}, pseudo partitives -- albeit syntactically similar and diachronically related to partitives in many languages -- merely denote quantities and contain indefinite kind nouns as in \textit{a sip of wine}.
In the literature on German, some authors incorrectly call the pseudo-partitive a \textit{partitive} \citep{Hentschel1993} while some realise the difference and at least mention it \citep{Eschenbach1994,GallmannLindauer1994,Loebel1989,Zimmer2015}.
}
Here, the kind noun is in the genitive, and I refer to the construction in (\ref{ex:intro:pseudopartitive1}) as the \textit{Pseudo-partitive Genitive Construction} (PGC).

\begin{exe}
%  \ex\label{ex:intro:pseudopartitive}
%  \begin{xlist}
    \ex[ ]{\label{ex:intro:pseudopartitive1}\gll Wir trinken [[ein Glas]\Sub{Acc} [dieses Weins]\Sub{Gen}]\Sub{Acc}.\\
    we drink a glass this wine\\
    \glt We drink a glass of this wine.}
%    \ex[*]{\label{ex:intro:pseudopartitive2} Wir trinken [[eine Tasse]\Sub{Acc} [einen leckeren Kaffee]\Sub{Acc}]\Sub{Acc}.}
%  \end{xlist}
\end{exe}

Second, if the kind noun is bare -- \ie\ if it comes neither with a determiner nor a modifying adjective -- it is uninflected as in (\ref{ex:intro:narrowapposition1}), and the genitive as seen in the PGC is not acceptable, see (\ref{ex:intro:narrowapposition2}).

\begin{exe}
  \ex\label{ex:intro:narrowapposition}
  \begin{xlist}
    \ex[ ]{\label{ex:intro:narrowapposition1}\gll Wir trinken [[ein Glas]\Sub{Acc} [Wein]\Sub{?}]\Sub{Acc}.\\
    we drink a glass wine\\
    \glt We drink a glass of wine.}
    \ex[*]{\label{ex:intro:narrowapposition2} Wir trinken [[ein Glas]\Sub{Acc} [Weins]\Sub{Gen}]\Sub{Acc}.}
  \end{xlist}
\end{exe}

This construction is usually classified as a \textit{Narrow Apposition Construction} \citep{Loebel1986}, henceforth NAC.%
\footnote{The construction as in (\ref{ex:intro:narrowapposition1}) is also referred to as the \textit{Direct Partitive Construction} for other Germanic languages in which the PGC with the synthetic genitive is not available.
This nomenclature makes sense in contrast to the \textit{Indirect Partitive Construction} with prepositional linkers translating to \textit{of} -- \ie\ analytic genitives -- in such languages, see \cite{HankamerMikkelsen2008} for Danish.
For German, this terminology is not distinctive enough, which is why I use the terms NAC and PGC.}
Notice that the unavailability of the genitive on the kind noun follows independently from a constraint that genitive NPs in German require the presence of some strongly case-marked element (determiner or adjective) in addition to the head noun in order to be acceptable (\citealp{GallmannLindauer1994,Schachtl1989}; see also \citealp[160]{Eisenberg2013b}).

It is difficult to determine whether the bare kind noun in the narrow apposition construction as in (\ref{ex:intro:narrowapposition1}) bears no case at all, a generic case, or agrees in case with the measure noun.%
\footnote{I would like to thank one of the reviewers for pointing this out.}
When there is an adjective as in (\ref{ex:intro:alternation2}), the embedded kind NP clearly agrees in case, but due to the overall absence of markers of case in the singular, bare nouns mostly show no indication of their case.
The only nouns which do have case markers in the singular are the so-called weak nouns \citep{Koepcke1995,Schaefer2016c}, which have something like a non-nominative \textit{-en} marker in the singular.
Unfortunately, there are no genuine mass nouns among the weak nouns.
However, a few of them can be coerced into a mass noun, such as \textit{Hase} `rabbit' (meaning `rabbit meat').
It then appears as if the uninflected form is preferred, but the inflected form is not excluded.
In (\ref{ex:macabre1}), the clearly acceptable form \textit{Hase} can only be a nominative singular or caseless.
In (\ref{ex:macabre2}), the form \textit{Hasen} could be an accusative, dative, or genitive.

\begin{exe}
  \ex\label{ex:macabre}
  \begin{xlist}
    \ex[ ]{\label{ex:macabre1}\gll Niemand will [ein Stück [Hase]\Sub{Nom\slash caseless}]\Sub{Acc} essen.\\
    nobody wants a piece rabbit eat\\
  \trans Nobody wants to eat a piece of rabbit.}
    \ex[?]{\label{ex:macabre2}\gll Niemand will [ein Stück [Hasen]\Sub{Acc\slash Dat\slash Gen}]\Sub{Acc} essen.\\
    nobody wants a piece rabbit eat\\}
  \end{xlist}
\end{exe}

Even a full unacceptability of (\ref{ex:macabre2}) would not be conclusive, however, as a possible aversion of speakers towards the case-marked form might be due to the fact that it is at least potentially also a genitive, in which case the constraint against bare genitive nouns would apply.
With plural kind nouns, the obligatory marking of the dative plural with \textit{-en} might provide some clues.
However, plural kind nouns do not behave like singular ones in measure phrases anyway, as will be argued in Section~\ref{sec:neighbouringcases}.
Also, judgements vary between (\ref{ex:datpl1}) and (\ref{ex:datpl2}), and both are found in corpora.%
\footnote{For example, the variant in (\ref{ex:datpl1}) with the lemmas \textit{Sack} and \textit{Apfel} occurs four times, and the one in (\ref{ex:datpl2}) twice in the very large DECOW corpus (see Section~\ref{sec:gettingdata}).
Similarly, for [\textit{Kiste} [\textit{Äpfel}]\Sub{Nom\slash Acc\slash Gen}]\Sub{Dat} `box of apples' (no case identity), I find three examples, and [\textit{Kiste} [\textit{Äpfeln}]\Sub{Dat}]\Sub{Dat} (clearly marked case identity), I find six examples.
}

\begin{exe}
  \ex\label{ex:datpl} 
  \begin{xlist}
  \ex{\label{ex:datpl1}\gll mit [zwei Säcken [Äpfel]\Sub{Nom\slash Acc\slash Gen}]\Sub{Dat}\\
     with a sack apples\\
    \trans with a sack of apples}
    \ex{\label{ex:datpl2}\gll mit [zwei Säcken [Äpfeln]\Sub{Dat}]\Sub{Dat}\\
    with a sack apples\\}
  \end{xlist}
\end{exe}

Descriptive grammars seem to favour an analysis in terms of caselessness (for example, \citealp[1981]{ZifonunEa1997c}).
The hazy picture of the case of bare kind NPs is most likely due to the fact that case is so rarely marked on German nouns, where case is marked mostly on determiners and to some degree adjectives.
The uncertainty in the few cases where case can be marked (weak nouns in the singular and dative plurals) would thus be a direct consequence of the fact that the construction is not very specific with respect to the case of the kind noun.

\begin{table}
  \centering
  \begin{tabular}{llll}
    \multicolumn{1}{r}{kind NP is:} & bare noun NP & NP with adjective & NP with determiner \\
    & [\ldots N\Subsf{meas} [N\Subsf{kind}]] & [\ldots N\Subsf{meas} [AP N\Subsf{kind}]] & [\ldots N\Subsf{meas} [D N\Subsf{kind}]] \\
    \midrule
    \multirow{2}{*}{narrow apposition}
                & \NACb                                                 & \NACa                                                   & \multirow{2}{*}{---}       \\
		& (\ref{ex:intro:narrowapposition1}) \textit{Glas Wein} & (\ref{ex:intro:alternation2}) \textit{Glas guten Wein}  &                            \\
    \midrule

    \multirow{2}{*}{pseudo-partitive genitive} 
                & \multirow{2}{*}{---}                                  & \PGCa                                                   & \PGCd                      \\
                &                                                       & (\ref{ex:intro:alternation1}) \textit{Glas guten Weins} & (\ref{ex:intro:pseudopartitive1}) \textit{Glas dieses Weins} \\
  \end{tabular}
  \caption{NAC and PGC constructions in different NP structures with examples and references to full example sentences}
  \label{tab:constructions}
\end{table}

To summarise, the case patterns in the NAC and in the PGC (depending on the structure of the kind NP) is given in Table~\ref{tab:constructions}.
I call the narrow apposition construction with a bare kind noun the \NACb\ and the partitive genitive with a determiner in the kind NP the \PGCd.
For the alternants with an adjective but no determiner in the kind noun phrase I use the terms \NACa\ and \PGCa.
In principle, this paper is about the middle column of Table~\ref{tab:constructions}, \ie\ the syntactic configuration in which two different case patterns are acceptable.
However, in Section~\ref{sec:analyses}, the outer columns (\NACb\ and \PGCd) will still play a major role when the factors controlling the alternation are discussed.

\subsubsection{Neighbouring cases}
\label{sec:neighbouringcases}

% namely:
% 
% \begin{enumerate}[i.]
%   \item{\label{it:intro:idiot} alleged alternatives to the \NACa\ and the \PGCa,}
% %  \item{\label{it:intro:datsg} alternative \textit{weak} forms of dative singular neuter adjectives,}
% %  \item{\label{it:intro:femsg} case syncretism in feminine NPs,}
%   \item{\label{it:intro:plurl} similar constructions with plural/collective kind nouns,}
%   \item{\label{it:intro:noifl} grammaticalised non-inflected measure nouns, and}
%   \item{\label{it:intro:other} alternative constructions for expressing quantities.}
% \end{enumerate}
%
% \vspace{-1\baselineskip}

I now turn to some more subtle issues related to the measure noun case alternation in order to delimit the scope of the study.
First, there is a claim found in some grammars that a generic nominative, accusative, and even dative on the kind noun are used instead of the genitive (\PGCa) or case agreement (\NACa).
Overviews can be found in \cite{Hentschel1993} and \cite{Zimmer2015}.
Sentence (\ref{ex:intro:gennom}) shows a putative generic nominative on the kind noun inside an accusative MNP.

\begin{exe}
    \ex[*]{\label{ex:intro:gennom} Wir trinken [[eine Tasse]\Sub{Acc} [heißer Kaffee]\Sub{Nom}]\Sub{Acc}.}
\end{exe}

It was shown empirically in \cite{Hentschel1993} that such sentences are de facto not acceptable.
Also, in my corpus sample, they simply did not occur.
Even if they are accepted by some speakers, their extremely low frequency makes it virtually impossible to study them using corpus linguistic methods (Section~\ref{sec:corpusstudies}), and I consequently do not discuss them further.

% As for (\ref{it:intro:datsg}), a complication with neuter kind nouns in the dative is mentioned by \citet[20--22]{Zimmer2015}.
% In the \NACa, the adjective normally inflects with the so-called \textit{strong} case and number suffixes, which are used when no determiner with strong inflection is present in the NP.
% \cite{Zimmer2015} reports a high number of occurrences of adjectives being inflected with the \textit{weak} inflectional marker, which are normally only used if a strongly inflected determiner precedes the adjective: \textit{kalt-en Wasser} `cold water' instead of \textit{kalt-em Wasser}.%
% \footnote{See Section~\ref{sec:analyses} for more discussion of adjectival inflection.}
% In my corpus sample, this tendency was not nearly as clear, and there was a high number of very noisy sentences among those potentially showing this pattern.
% Hence, I do not discuss these forms.%
% \footnote{All corpus data and scripts used for this paper will be released freely, so further examination of the few cases maybe showing this kind of inflection is possible.}
% 
% \label{page:femininesyncretism}
% With feminine kind nouns, (\ref{it:intro:femsg}) is relevant.
% Whereas the singular masculine and neuter nominal subsystem is a four case system (at least for NPs with adjectives and determiners), feminine singular NPs show nominative-accusative and dative-genitive syncretisms.
% Table~\ref{tab:syncretisms} shows this as well as the partial syncretism in the plural, where no gender distinctions are made.
% This means that with dative measure nouns, the alternation is unobservable if the kind noun is feminine, and it is problematic to speak of a \textit{genitive} kind noun construction in this case.
% This is taken into account in the statistical modeling and the discussion presented in Section~\ref{sec:corpusstudies}.
% 
% \begin{table}
%   \centering
%   \begin{tabular}{llll}
%      & Masculine\slash Neuter & Feminine & Plural \\
%      \midrule
%      nominative & rot-er Wein    & \multirow{2}{*}{frisch-e Sahne}   & \multirow{2}{*}{klein-e Äpfel} \\
%      accusative & rot-en Wein    &                                   &                                \\
%      dative     & rot-em Wein    & \multirow{2}{*}{frisch-er Sahne}  & klein-en Äpfel-n               \\
%      genitive   & rot-en Wein-es &                                   & klein-er Äpfel                 \\
%   \end{tabular}
%   \caption{Case syncretisms in NPs with an adjective and without a determiner (\textit{roter Wein} `red wine', \textit{frische Sahne} `fresh cream', \textit{kleine Äpfel} `small apples')}
%   \label{tab:syncretisms}
% \end{table}

Second, we find that, if the kind noun is a plural count noun as in \textit{ein Sack kleine Äpfel} (\NACa) or \textit{ein Sack kleiner Äpfel} (\PGCa) `a bag of small apples', a similar alternation between PGC and NAC can be observed.
In line with experimental results reported in \citet[15--16]{Zimmer2015}, I found that the PGC is so dominant with plural kind nouns (794 of 861 cases, or over 92\%, cf.\ Section~\ref{sec:corpusstudies}) that the alternation cannot be analysed in the same way as in the singular.
%\footnote{More concretely, if these cases were included in the regression analysis of the corpus data along with a factor encoding the number of the kind noun, this factor would most assuredly override any other regressor for the data points with a plural kind noun.}
While this will play a role in the interpretation of the corpus findings, MNPs with plural kind nouns will not be included in the corpus study and the experiments reported in Sections~\ref{sec:corpusstudies} and \ref{sec:experimental}.

Third, some measure nouns have been grammaticalised in a way that they always appear in their non-inflected form.
They are typical measure nouns like \textit{Gramm} `gram', \textit{Pfund} `pound' or \textit{Prozent} `percent', which have no normal plural forms at all.%
\footnote{Plurals like \textit{Pfunde} and \textit{Prozente} have special meanings and have very restricted uses, mostly in idiomatic expressions such as \textit{Pfunde verlieren} `lose pounds' and \textit{Prozente machen} `make a profit'.
They cannot be used in normal MNPs.}
%\footnote{With neglectable frequency, this effect is extended to container nouns like \textit{Glas} `glass', but with a twist.
%A non-inflected version \textit{zwei Glas roter Wein} `two glasses of wine' and an inflected version \textit{zwei Gläser roter Wein} co-exist.
%The non-inflected version denotes the quantity of wine fitting into two glasses, the inflected version denotes two glasses filled with wine.}
I treat these cases like other measure nouns because they enter into both the \NACa\ as in (\ref{ex:uninflectedmeasures:a}) and the \PGCa\ as in (\ref{ex:uninflectedmeasures:b}).
In Section~\ref{sec:analyses}, however, degrees of grammaticalisation as a factor influencing the alternation will be discussed prominently.

\begin{exe}
  \ex\label{ex:uninflectedmeasures}
  \begin{xlist}
    \ex{\label{ex:uninflectedmeasures:a} zwei Gramm brauner Zucker}
    \ex{\label{ex:uninflectedmeasures:b}\gll zwei Gramm braunen Zuckers\\
                                         two gram brown sugar\\
				         two grams of brown sugar}
  \end{xlist}
\end{exe}

Finally, there are alternative ways of expressing similar quantificational meanings.
In the variationist tradition, which is strongly influenced by Labovian sociolinguistics, the \textit{principle of accountability} would dictate that proper studies should examine a variationist \textit{variable}, \ie\ all different \textit{ways of saying the same thing} (\citealp{Labov1966}, \citealp{Labov1969}, for an overview see \citealp{Tagliamonte2012}).
In the case at hand, the variable might be something like \textit{measurement of quantities of substances and collections}.
I argue that it is fully justified to focus narrowly on \NACa\ and \PGCa\ with their well-defined morpho-syntactic properties, mostly because the alternative constructions are not used in the same range of contexts.
Two major alternatives might be considered.
There is an analytic (pseudo-)partitive with \textit{von} `of'.
It is only available as an alternative to the PGC if the kind noun phrase contains a (definite or indefinite) determiner as in (\ref{ex:analyticalpartitive}).

\begin{exe}
  \ex\label{ex:analyticalpartitive} 
  \begin{xlist}
    \ex[ ]{\label{ex:analyticalpartitive:a}\gll ein Glas von dem roten Wein\\
                                                a glass of the red wine\\
                                                \glt a glass of the red wine
       }
    \ex[*]{\label{ex:analyticalpartitive:b}\gll ein Glas von rotem Wein\\
                                                a glass of red wine\\
                                                \glt a glass of red wine
       }
  \end{xlist}
\end{exe}

This means that the \textit{von} (pseudo-)partitive does not compete with the \NACa\ and the \PGCa.
In fact, there is not a single context in which they can be interchanged.

Additionally, we find constructions with \textit{voll} (\textit{von})\slash\textit{voller} `full (of)' and \textit{mit} `with' are available, see (\ref{ex:alternatives}).

\begin{exe}
  \ex\label{ex:alternatives}
  \begin{xlist}
  \ex{\label{ex:voll}\gll   ein Glas voll von rotem Wein\\
                            a glass full of red wine\\
			    \glt a glass full of red wine}
  \ex{\label{ex:voller}\gll ein Glas {voll\slash voller} {rotem\slash roter} Wein\\
                            a glass full-of red wine\\
			    \glt a glass full of red wine}
  \ex{\label{ex:mit}\gll    ein Glas mit rotem Wein\\
                            a glass with red wine\\
			    \glt a glass with red wine in it\slash a glass filled with red wine}
  \end{xlist}
\end{exe}

These construction have very idiosyncratic properties.
The construction with \textit{voller} is discussed in \cite{Zeldes2018} and the construction with \textit{mit} in \cite{Bhatt1990}.
They are not semantically equivalent to the NAC and the PGC, and they are only available for a small subset of measure nouns.
All of these constructions are used only with measure nouns denoting containers, where the whole NP typically refers to the container and never to the quantity contained in it.
They are incompatible with measure nouns denoting natural portions such as \textit{Schluck} `gulp' or \textit{Haufen} `heap' and strongly grammaticalised nouns such as \textit{Gramm} `gram'.
This disqualifies them as alternatives to the \NACa\ or \PGCa\ in most contexts, and they are thus clearly not just more ways of saying the same thing. 
It is therefore reasonable to focus on the two well-defined alternants.

This concludes the descriptive overview of the phenomenon.
I have demonstrated that there is an alternation between two measure noun constructions in a narrow syntactic configuration (kind NP with an adjective but without a determiner), and that the two constructions differ in the case of the kind noun (case agreement with the measure noun or genitive).
I turn to more theory-oriented discussion of the alternation in the next section.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Factors controlling the alternation}
\label{sec:analyses}

This section briefly reviews existing analyses of the \PGCa\ vs.\ \NACa\ alternation and related issues.
I also develop my own analysis and the appropriate hypotheses for the empirical studies presented in Sections~\ref{sec:corpusstudies} and \ref{sec:experimental}.

.%
\footnote{A reviewer mentioned that she or he had the impression that dialectal variation is also a factor influencing the alternation.
This is definitely true as some dialects (such as Alemannic) tend to have no genitive at all.
While this is an interesting aspect for further research, the main obstacle is that there are no corpora of German which are both large enough and annotated with reliable meta data about regional variants.
In the annotation of the corpus study, documents obviously written in a regional variant were excluded.
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2017-12-22 Latex/sections/3corpus.tex
Sections in separate files.

\section{Corpus study}
\label{sec:corpusstudies}

\subsection{Corpus choice and sampling}
\label{sec:gettingdata}

% TODO Quote:
% Brems (2016:171--175) for using GOOGLE data, including COUNTS!

For the present study, I used the German \textit{Corpus from the Web} (COW) in its 2014 version DECOW14A (\citealp{SchaeferBildhauer2012full}, and \citealp{Schaefer2015b}, as well as \citealp{BiemannEa2013}, and \citealp{SchaeferBildhauer2013}, for overviews of web corpora in general and the methodology of their construction), which contains almost 21 billion tokens.%
\footnote{The COW corpora (Dutch, English, French, German, Spanish, Swedish) are made available for free at \url{https://www.webcorpora.org}.
At the time of this writing, a newer 2016 version DECOW16 has already been released.}
I chose this corpus for two main reasons.%
\footnote{The use of web data for linguistic research does require explicit and careful justification.
Due to the noisy nature and unknown composition of the web, only carefully designed and established web corpora like the COW corpora or the SketchEngine corpora \citep{KilgarriffEa2014} should be used.
Clearly, using search engine results is ``bad science'' for many reasons, most prominently total non-replicability of results, as \cite{Kilgarriff2006} pointed out more than ten years ago.
Careless use of search engine results is still found, however, see for example \citet[171--175]{DeclerckBrems2016}.}
First, the external validity of any study is increased through a higher heterogeneity of the sample \citep[30]{MaxwellDelaney2004}, and the DECOW corpus has clearly a much more heterogeneous composition compared to the only other very large corpus of German, the DeReKo \citep{KupietzEa2010} of the Institute for the German Language (IDS), which contains almost exclusively newspaper texts.%
\footnote{It was shown in \cite{W16-2601} that, for example, the range of topics covered is much smaller in DeReKo compared to DECOW.}
Second, it was already mentioned that normative grammars often adopt clear positions regarding the grammaticality of either the \NACa\ or the \PGCa.
Thus, newspaper text or any other text that conforms strongly to normative grammars might not represent the alternation phenomenon fully (and without bias) because authors and proofreaders who must adhere to normative guidelines might favour one alternative or the other explicitly.
Web corpora, on the other hand, contain at least some amount of non-standard language from forums and similar sources.
For these or similar reasons, COW corpora have been used in a number of peer-reviewed publications, for example \cite{VanGoethemHiligsmann2014}, \cite{VanGoethemHuening2015}, \cite{MuellerS2014}, \cite{Schaefer2016c}, \cite{SchaeferSayatz2014}, \cite{SchaeferSayatz2016}, and \cite{Zimmer2015}. 
Therefore, DECOW is a valid choice for this study.

I now turn to the sampling procedure applied to obtain concordances for manual annotation and statistical analysis.
Among the factors potentially influencing the alternation (see Section~\ref{sec:germanmeasurenps}) were lemma-specific preference effects.
Therefore, it was highly desirable to obtain a sample in which most of the highly frequent actually-occurring combinations of kind nouns and measure nouns were represented.
I applied a three-stage process in order to obtain such a sample, which consisted of the following steps:

\begin{enumerate}[i.]
  \item\label{enum:data:step1} generating a list of the one hundred most frequent mass nouns,
  \item\label{enum:data:step2} deriving a list of all measure nouns with which the mass nouns co-occur in the \NACb, and 
  \item\label{enum:data:step3} sampling the target constructions by querying each combination of mass noun and measure noun found in step (\ref{enum:data:step2}).
\end{enumerate}

\vspace{-1\baselineskip}

In step (\ref{enum:data:step1}), I exported a list of all nouns in the DECOW14A01 sub-corpus sorted by their token frequency and manually went through it from the most frequent noun downwards, selecting the first one hundred mass nouns that occurred in the list.%
\footnote{DECOW14A01 is the first slice (roughly a twentieth) of the complete DECOW14A corpus.
It contains just over one billion tokens.}
Mass nouns were defined as concrete nouns which denote a substance in the broad sense, combine with uninflected mass quantifiers such as \textit{viel} `much' and \textit{wenig} `little' (\textit{viel Bier} `much beer'), and form only sortal and unit plurals (such as the plural \textit{Biere} `types of beer' or `glasses of beer').
Abstract nouns which partially behave like mass nouns (like \textit{Spaß} `fun’ or \textit{Gefahr} `danger’) were excluded because they are usually not quantified in the same way as concrete mass nouns.
The hundredth selected mass noun was \textit{Schmuck} `jewellery’, which is the 3,054th most frequent noun in the original frequency list.

This list of mass nouns was used in step (\ref{enum:data:step2}) to derive a list of measure nouns co-occurring with the mass nouns. 
In order to generate this list, I utilised the fact that a direct sequence of two nouns almost always instantiates the bare-noun NAC if the second noun is a mass noun.
Hence, I searched for all sequences N\Sub{1}N\Sub{2} where N\Sub{2} was one of the mass noun lemmas extracted in step (\ref{enum:data:step1}).
Then, the resulting 100 lists of noun-noun combinations were each sorted by frequency in descending order and sieved manually to remove erroneous hits.
From each of the 100 lists, I also removed noun-noun combinations that had a frequency below 2, except if the individual list would have otherwise been shorter than 20 noun-noun combinations.
The result was a list of the most frequent 2,365 individual combinations of a measure noun and a mass noun.

In step (\ref{enum:data:step3}), each of these 2,365 noun–noun combinations was queried in the target constructions (\PGCa\ and \NACa) individually in each of the first ten slices of DECOW (roughly 10 billion tokens).
In order to reduce the sample size for the manual annotation process, the final concordance was sampled from the results of these 2,365 queries.
Since the mass nouns in the sample were distributed according to the usual power law (often referred to as a \textit{Zipfian} distribution), I used all hits for nouns with a frequency up to 100 and a sample of 100 of all those with higher frequency.
The final sample contained 6,843 sentences, which was reduced to 5,063 in the manual annotation process due to removal of noisy material, erroneous hits and uninformative cases where the measure noun was in the genitive, in which case the \NACa\ cannot be distinguished from the \PGCa.
Given the careful sampling procedure described in this section, we can be highly sure that it contains all relevant and reasonably frequent noun–noun combinations in the target constructions.%
\footnote{In a similar fashion, the 100 most frequent measure nouns occurring with plural kind nouns were listed and queried, resulting in a sample of 871 sentences.
As stated in Section~\ref{sec:germanmeasurenps}, the \NACa\ is virtually never used with plural kind nouns, and this sample was not used except for quantifying the frequency of occurrence of the constructions (67 times \NACa\ and 794 times \PGCa).
However, the sample is distributed with the data package accompanying this paper.
}

Finally, two auxiliary samples were also drawn.
As mentioned in Section~\ref{sec:analyses}, the distribution of the measure noun and kind noun lemmas in the \NACb\ and the \PGCd\ with a determiner will be modelled as factors influencing the alternation.
Therefore, all noun-noun pairs from the process described above were also queried in the two non-alternating constructions, resulting in 17,252 hits for the \PGCd\ and 315,635 hits for the \NACb.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Variables and annotation}
\label{sec:annotation}

The full set of manually annotated variables for the main sample is given in Table~\ref{tab:variables}, and I briefly discuss it now.%
\footnote{All numeric variables were also z-transformed (\ie\ centered to the mean and rescaled such that they have a standard deviation of 1) to facilitate their interpretation in the regression models reported in the next section.}
Notice first that \textit{Construction} is the response variable (or `dependent variable') with the values \textit{PGCa} and \textit{NACa}.

\begin{table}
  \centering
  \begin{tabular}{llll}
    Unit of reference & Variable                      & Type    & Levels (for factors only) \\
    \midrule
    Document       & Badness                          & numeric &                           \\
                   & Genitives                        & numeric &                           \\
    Sentence       & Cardinal                         & factor  & Yes, No                   \\
                   & \textbf{Construction (response)} & factor  & NACa, PGCa                \\
                   & Measurecase                      & factor  & Nom, Acc, Dat             \\
    Kind lemma     & Kindattraction                   & numeric &                           \\
                   & Kindfreq                         & numeric &                           \\
    Measure lemma  & Measureattraction                & numeric &                           \\
                   & Measureclass                     & factor  & Physical, Container,      \\
                   &                                  &         & Amount, Portion, Rest     \\
                   & Measurefreq                      & numeric &                           \\
  \end{tabular}
  \caption{Annotated variables for the main sample}
  \label{tab:variables}
\end{table}

The variables \textit{Kindattraction} and \textit{Measureattraction} encode the ratio with which a given kind noun lemma or measure noun lemma occurs in the \PGCd\ and the  \NACb.
They were calculated from the auxiliary samples described at the end of Section~\ref{sec:gettingdata} as a log-transformed quotient.
The higher the value, the more often the noun occurs in the \PGCd (proportionally).%
\footnote{
  It could be argued that some more advanced measure of attraction strength should be used, as is done in Collostructional Analysis \citep{StefanowitschGries2003,GriesStefanowitsch2004}, see also \cite{Gries2015a}.
  Three main points speak against such an approach in the present case.
  First, the goal here is to quantify how often lemmas occur in the \PGCd\ and the \NACb, and these constructions do not compete at all but are rather mutually exclusive.
  Collostructional approaches are not made for such scenarios.
  Second, the attraction values will be used as regressors in a hierarchical logistic regression and the values resulting from collostructional analysis, \ie\ logarithmised Fisher p-values, have a very unfavourable distribution in the case at hand.
  They cluster around 0 and they include values of $-\infty$.
  Third, I tried using collexeme strength as a regressor (with smoothing to remedy the mathematical problems), and the results were unsatisfactory compared to the simple quotient used here.
}
Additionally, \textit{Kindfreq} and \textit{Measurefreq} are the logarithm-transformed frequencies per 1,000,000 words of each lemma, extracted from the frequency lists distributed by the DECOW corpus creators on their web page.
They were added to control for basic frequency effects.

In Section~\ref{sec:analyses}, it was hypothesised that classes of measure lemmas might have different preferences for the two alternants.
To capture this, class information was annotated for measure lemmas.
The classification was inspired by the list in \citet[530]{Koptjevskaja2001} but due to the low frequencies of many of the potential classes, a very coarse classification was finally used.
With typical examples and their frequencies in the final sample, the classes are:
\textit{Physical} (abstract precisely measurable units such as \textit{Liter} `litre', \textit{Meter} `metre', \textit{Gramm} `gram'; \textit{f=}1,968),
\textit{Container} (\textit{Eimer} `bucket'; \textit{f=}740),
\textit{Amount} (\textit{Menge} `amount'; \textit{f=}1,364), 
\textit{Portion} (natural portions like \textit{Happen} `bite' or \textit{Krümel} `crumb'; \textit{f=}713).
The lemmas that did not fit into either of these classes were labelled \textit{Rest} (\textit{f=}278).

The variable \textit{Cardinal} encodes whether the measure noun is modified by a cardinal (\textit{f=}1,939) or not (\textit{f=}3,124).
The purpose of this variable is to test whether cardinals really favour the \NACa\ as hypothesised in Section~\ref{sec:analyses}.

To capture the influence of style mentioned in Section~\ref{sec:analyses}, two proxy variables were used.
At the document level, the DECOW corpus has an annotation for \textit{Badness}.
As described in \cite{SchaeferEa2013}, \textit{Badness} measures how well the distribution of highly frequent short words in the document matches a pre-generated language model for German.
Documents with higher Badness usually contain more incoherent language, shorter sentences, etc.
If the \PGCa\ actually favours higher stylistic levels, a high \textit{Badness} should be correlated with fewer occurrences.
Documents in DECOW14 have also been annotated with a variable called \textit{Genitives}.
The higher the values of this variable, the lower the proportion of genitives among all case-bearing forms is.
A high number of genitives is indicative of higher levels of style.
However, the use of this variable as a regressor in the present study might be considered problematic.
Since the \PGCa\ contains a genitive itself, the regressor variable \textit{Genitive} and the document-level variable \textit{Genitives} are not fully independent.
Since instances of the \PGCa\ make up for only a minute fraction of all genitives, I still use \textit{Genitives} as a regressor with the appropriate caveats.

Finally, one variable was added as nuisance variable in the context of the present study.
It was reported in the literature that MNPs in the dative and with a masculine or neuter kind noun favour the \PGCa\ more than the corresponding nominative and accusative MNPs \citep{Hentschel1993,Zimmer2015}.
As an example, \textit{mit einem Stück frischen Brots} `with a piece of fresh bread' (\PGCa) would be preferred more strongly against \textit{mit einem Stück frischem Brot} (\NACa).
As with all the examples, native speakers of German will most likely notice that differences are subtle.
To control for this effect, the case of the measure noun was manually annotated (variable \textit{Measurecase}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{On statistical analysis}
\label{sec:rightstatistics}

In this section, I justify my choice of statistical models used in Section~\ref{sec:corpushierarchicalmodel}.
Readers might think that the method used here for modeling grammatical alternations -- namely (Hierarchical) Logistic Regression\slash Generalised Linear (Mixed) Models or GL(M)Ms -- does not require much justification.
After all, GLMMs have been established as the major tool in the analysis of alternation phenomena.
All studies mentioned at the outset of Section~\ref{sec:cogocl} use some form of regression\slash GLMM.
Over the past few years, however, modified or alternative methods have been proposed.
While it is impossible to discuss all of these methods here, I just make a few remarks on Bayesian estimation (see \citealp{GelmanEa2014}) as it was proposed in \cite{Levshina2016} and \cite{Divjak2016a}, for example.
Conceptually, I see three points of discussion that should be kept apart.
First, Bayesian methods are sometimes touted as superior tools for scientific inference compared to frequentist methods.
Second, it has been proposed that the Bayesian interpretation of probability is more cognitively adequate for the modeling of linguistic data \cite[301--302]{Divjak2016a}.
Third and very specific to this paper, given established methods in the modeling of alternation and variation, it has to be decided whether so-called Bayesian methods lead to substantially different results.

% As for the first point, this is not quite the place to discuss it fully.
% The basic distinction is a philosophical one and related to the concepts of \textit{direct} and \textit{inverse probability} (\eg\ \citealp{Senn2011}).
% Frequentists assume that models and parameters are fixed, for example a model specifying that a coin is fair.
% We can then calculate for observed data (for example a measurement of 3 heads in an experiment with 10 tosses) how often such a result or a more extreme result would occur if the model were true and we repeated the experiment arbitrarily often.
% This is essentially the frequentist notion of direct probability, \ie\ long-run frequencies under replication.
% Standard tests in the Fisher and Neyman-Pearson traditions as well as Neyman confidence intervals are based on this concept of probability.
% Bayesian approaches (in the now common interpretation), on the other hand, condition on the particular data and quantify inductively the probability of model parameters given the data.
% The parameters are thus not fixed, and the probability is usually equated with researchers' posterior beliefs about model parameters.
% There is actually a debate among Bayesians about the proper interpretation of Bayesian methods, and whether a notion of hypothesis testing is compatible (or even already contained) in the Bayesian approach.
% In \citet[10]{GelmanShalizi2013}, the authors -- prominent Bayesians themselves -- acknowledge that a theory of statistical testing is a desideratum, and they state about the standard inductive interpretation of Bayesianism that ``most of this received view of Bayesian inference is wrong,'' and they develop a Bayesian notion of p-values (see also \citealp{Mayo2013}, for a frequentist reply; also \citealp{Senn2011} on different strands of Bayesianism and their stance on inductive vs.\ deductive reasoning, and \citealp{Mayo2011}, for a critical reply to \citealp{Senn2011}).
% Clearly, in such quarrels between and among camps of philosophers of science and statisticians, it is difficult for mere practitioners to take sides.
% 
% These quarrels relate to the second point, however.
% \citet[301--302]{Divjak2016a} speaks favourably of Bayesian methods because the Bayesian concept of probability is more adequate for cognitive modeling compared to the frequentist one.
% Her argument is part of a larger body of literature asking for cognitively plausible modeling techniques, for example Naive Discriminative Learning (NDL; \citealp{Baayen2011,BaayenEa2013,MilinEa2016,TheijssenEa2013}). 
% On p.\ 303 of \cite{Divjak2016a}, the author goes on to explicitly mention NDL as well.
% Yet, neither frequentist nor Bayesian methods were conceived as cognitive models, but as systems of inference for scientists (see above, and see also \citealp[302]{Divjak2016a}).
% The fundamental question that lurks behind such arguments is how we interpret our statistical models (estimated on corpus data).
% Are they inductive models of cognitive representations that also human learners would infer from being exposed to the corpus data?%
% \footnote{In which case we are doing ``data science in language research'' in the words of \citealp{MilinEa2016}.
% I see this as standing in contradiction to the view advocated in \cite{Dabrowska2016} as cited above.}
% Or are they tests of theories that are pre-specified and merely tested for predictive accuracy on linguistic output data contained in corpora?
% In the former case, we adopt a strong \textit{corpus as input} hypothesis and should definitely resort to methods like NDL.
% However, this would most likely require us to toss most previous work done in (cognitive) linguistics into the bin and to abandon all high-level generalisations that most existing studies have been based upon.
% In fact, arguments to this extreme effect and against using high-level generalisation have been made, for example in \cite{BaayenEa2016}, \citet[299--300]{Divjak2016a}, \cite{RamscarPort2016}, \cite{TheijssenEa2013}.
% In the latter and less extreme case, the cognitive commitment does, however, not necessarily extend to the statistical methods used.
% These methods then do not need to be any more cognitively plausible than an ANOVA used to analyse the results from an experiment.
% I view my own work (see Sections~\ref{sec:germanmeasurenps}--\ref{sec:experimental}) in the tradition of testing theories (which embrace high-level generalisations), and I agree to provisionally use Likelihood methods (see below).
% I remain fully agnostic with respect to the question, which approach is \textit{right}.
% The best strategy for cognitive linguistics as a field might be to cultivate many methods while making sure that each method is applied carefully and competently.

The first and second point cannot be discussed in detail here.%
\footnote{However, there is a number of critical papers by statisticians (even prominent Bayesians) and philosophers of science in which the extreme credibility that has recently been assigned to (subjective and automatic) Bayesian approaches (and especially the standard inductivist interpretation of Bayesianism) is questioned (for example, \citealp{GelmanShalizi2013,Mayo2011,Senn2011}).}
With regard to the third point, \citet[251--252]{Levshina2016} argues for Bayesian estimation in mixed regression settings.
First, she claims that ``while frequentist statistics only allows one to test whether the null hypothesis can be rejected, Bayesian statistics enables one both to test the null hypothesis and to estimate the probability of specific parameter values given the data.''
This does not do justice to frequentist methods in that a strong focus on the rejection of the null hypothesis is characteristic only of Fisher's approach.
In the Neyman-Pearson approach, results ideally \textit{favour} the main hypothesis vis-à-vis the alternative hypothesis (cf.\ \citealp{Lehmann1993,Lehmann2011,Perezgonzalez2015}).
Also, especially Neyman-style frequentism has well-known extensions to estimation, for example in the form of confidence intervals (see \citealp{GreenlandEa2016}, esp.\ p.\ 340).
She then explains that a ``distinctive feature of Bayesian statistics is the use of so-called priors'' and that ``posterior probabilities depend on both the prior beliefs and the data, whereas the results of a frequentist model depend only on the data'' \cite[252]{Levshina2016}.
Remarkably, given this statement, she does \textit{not} use informative priors and in her footnote 8 \cite[252]{Levshina2016} admits that priors were probed using trial and error.
So, the proclaimed major advantage of Bayesian modeling was apparently not taken advantage of.%
\footnote{In the words of \cite{Senn2011}: ``You may believe you are a Bayesian but you are probably wrong.''
\citet[347--348]{GelmanHill2006} ``view any noninformative prior distribution as inherently provisional'' and give recommendations how to proceed once posteriors have been obtained from noninformative priors.}
Now, Maximum Likelihood Estimation (MLE) -- the traditional method which could have been used instead -- is not exactly \textit{frequentist} in the sense of Neyman-Pearson testing theory.
MLE, like inductive Bayesianism, conditions on the particular data inasmuch as it searches for the most likely set of parameters given the data.
What is more, Bayesian estimators are in fact based on the Likelihood and merely multiply it by the prior \citep[6--8]{GelmanEa2014}.
If the prior is flat, results converge (see also \citealp[347]{GelmanHill2006}).
The same is true if the sample size is large compared to the number of parameters, at least for finite-dimensional parameter models \citep[1119--1120]{Freedman1999}, a well-established result known as the \textit{Bernstein-von Mises theorem}.
With a modest model structure including 17 fixed effects and 2,646 data points in \cite{Levshina2016}, it is highly likely that the same results would have been obtained with Maximum Likelihood methods.
In fact, she admits that changing the priors did not lead to substantially different results in her footnote 8.
This is a clear sign that the prior is ``swamped by the data'' \citep[1119]{Freedman1999}.
% Going back to the argument that Bayesian methods are allegedly more cognitively plausible than `frequentist' methods, the claim might still be true.
% However, most similar studies (dating back to \citealp{BresnanEa2007}) use an estimator (MLE) that is not frequentist in a narrow sense, and which is as cognitively plausible as Bayesian estimators because it most likely produces the same results in these studies.

If there had been evidence in Levshina's study that Bayesian and MLE methods did \textit{not} converge, it would have been an occasion to demonstrate the selective superiority of the algorithms used in Bayesian estimation.
After all, there are situations where Bayesian estimators can be more robust, namely with heavily censored data, complex hierarchical models, perfect separation, etc. (see \citealp{Freedman1999}, \citealp[345--348]{GelmanHill2006}).
I want to state clearly that these points do not in any way invalidate the results presented in \cite{Levshina2016}.
However, being ``Bayesian'' is most likely not among its selling points.
Additionally, I want to voice the concern that many practitioners are probably already struggling with getting an adequate grasp of advanced statistical methods and that it might therefore be wise to use the more conservative and better understood method if the alternative method is not absolutely required for substantive reasons.
In the next section, I estimate the parameters of my hierarchical model with MLE and Markov-Chain Monte Carlo (MCMC) methods (the currently most prominent estimator used in Bayesian settings) to demonstrate their expectable convergence.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{A hierarchical model of the measure noun alternation}
\label{sec:corpushierarchicalmodel}

\begin{figure}[hb!]
  \centering
  \includegraphics[width=0.85\textwidth]{../R/output/corpus_fixeffs_mle+mcmc}
  \caption{Coefficients (MLE and MCMC) with 95\% confidence intervals (for details see text); the intercept is -4.328 (MLE) and -4.441 (MCMC)}
  \label{fig:fixeffs}
\end{figure}

In this section, I report the results of fitting a multilevel model to the data using R \citep{R}, \textit{lme4} \citep{lme4} for Maximum Likelihood Estimation (MLE) and \textit{rstanarm} \citep{rstanarm} for Bayesian Markov-Chain Monte Carlo (MCMC) estimation (see Section~\ref{sec:rightstatistics}).
The purpose is to model the influence of the regressors specified in Table~\ref{tab:variables} on the probability that the \PGCa\ is chosen over the \NACa.
All regressors from Table~\ref{tab:variables} were included, and the measure lemma and the kind noun lemma were specified as varying-intercept random effects.
The sample size was \textit{n=}5,063 with 1,134 cases of \PGCa\ and 3,929 cases of \NACa.
The results of the estimation are shown in Table~\ref{tab:bigtable} and in Figure~\ref{fig:fixeffs}.
The intercept comprises \textit{Cardinal=Yes}, \textit{Measurecase=Nom}, \textit{Kindgender=Masc}, \textit{Measureclass=Physical}, and 0 for all numeric z-transformed regressors.
It is -4.328 for the ML estimate and -4.441 for the MCMC estimate.

The regressors with the measure lemma as their unit of reference have no within-measure lemma variance, and the \textit{glmer} function automatically estimates them as \textit{group level predictors} (or \textit{second-level effects}), cf.\ \citet[265--269,302--304]{GelmanHill2006}.
The same goes for those listed with the kind lemma as their unit of reference.
Given the coding of the response variable, coefficients leaning to the positive side can be interpreted as favouring the \PGCa.

\begin{sidewaystable}
  \centering
  \resizebox{\textheight}{!}{
  \begin{tabular}{llrlp{0.5em}rrp{0.5em}rrp{0.5em}rrp{0.5em}cc}
      Model & Regressor & \multicolumn{1}{l}{\pPB} & Factor && \multicolumn{2}{l}{Coefficient} && \multicolumn{2}{l}{CI low} && \multicolumn{2}{l}{CI high} && \multicolumn{2}{l}{CI excludes 0} \\
      level &                   &       & level     && \multicolumn{1}{l}{MLE} & \multicolumn{1}{l}{MCMC} && \multicolumn{1}{l}{MLE} & \multicolumn{1}{l}{MCMC} && \multicolumn{1}{l}{MLE} & \multicolumn{1}{l}{MCMC} && \multicolumn{1}{l}{MLE} & \multicolumn{1}{l}{MCMC} \\\midrule
       First     & Badness           &  0.002 &           && -0.152 & -0.155 && -0.247 & -0.247 && -0.061 & -0.065 && * & * \\
                 & Cardinal          &  0.001 & No        &&  1.189 &  1.222 &&  0.862 &  0.927 &&  1.466 &  1.496 && * & * \\
                 & Genitives         &  0.001 &           && -0.693 & -0.711 && -0.768 & -0.801 && -0.592 & -0.616 && * & * \\
                 & Measurecase       &  0.001 & Acc       &&  0.030 &  0.031 && -0.150 & -0.159 &&  0.212 &  0.222 &&   &   \\
                 &                   &        & Dat       &&  0.705 &  0.729 &&  0.455 &  0.465 &&  0.944 &  0.995 && * & * \\[0.5\baselineskip]

       Second    & Kindattraction    &  0.020 &           &&  0.225 &  0.244 &&  0.049 &  0.056 &&  0.393 &  0.422 && * & * \\
       (Kind)    & Kindfreq          &  0.095 &           &&  0.146 &  0.164 && -0.023 & -0.016 &&  0.301 &  0.341 &&   &   \\
                 & Kindgender        &  0.001 & Neut      &&  0.021 &  0.013 && -0.367 & -0.409 &&  0.392 &  0.435 &&   &   \\
                 &                   &        & Fem       &&  1.269 &  1.289 &&  0.800 &  0.788 &&  1.709 &  1.783 && * & * \\[0.5\baselineskip]

       Second    & Measureattraction &  0.001 &           &&  0.282 &  0.299 &&  0.106 &  0.102 &&  0.447 &  0.515 && * & * \\
       (Measure) & Measureclass      &  0.001 & Container &&  0.252 &  0.257 && -0.265 & -0.303 &&  0.788 &  0.813 &&   &   \\
                 &                   &        & Rest      &&  0.421 &  0.379 && -0.209 & -0.378 &&  1.063 &  1.091 &&   &   \\
                 &                   &        & Amount    &&  0.831 &  0.889 &&  0.215 &  0.220 &&  1.432 &  1.569 && * & * \\
                 &                   &        & Portion   &&  1.217 &  1.253 &&  0.675 &  0.689 &&  1.684 &  1.840 && * & * \\
                 & Measurefreq       &  0.005 &           && -0.231 & -0.232 && -0.363 & -0.395 && -0.079 & -0.073 && * & * \\

  \end{tabular}
  }
  \caption{Coefficient table comparing Maximum Likelihood Estimation (MLE, with 95\% bootstrap confidence interval) and `Bayesian' Markov-Chain Monte Carlo estimation (MCMC); the intercept is -4.328 (MLE) and -4.441 (MCMC)}
  \label{tab:bigtable}
\end{sidewaystable}

Standard diagnostics for MLE show that the model quality is quite good.
Generalised variance inflation factors for the regressors were calculated to check for multicollinearity \citep{FoxMonette1992,ZuurEa2010}, and none of the corrected $\text{GVIF}^{1/2\text{df}}$ was higher than 1.6.
Nakagawa \& Schielzeth's pseudo-coefficient of determination is $R_m^2=0.409$ and $R^2_c=0.495$ (see \citealp{Gries2015} for a basic introduction to these $R^2$ measures, or else \citealp{NakagawaSchielzeth2013}).
The rate of correct predictions is 0.843, which means a proportional reduction of error of $\lambda=0.297$.
The lemma intercepts have standard deviations of $\sigma_{\text{Measurelemma}}=0.448$ and $\sigma_{\text{Kindlemma}}=0.604$.

The coefficient estimates are specified in Table~\ref{tab:bigtable} for each regressor (or regressor level) in the columns labelled \textit{Coefficient}.
For a robust quantification of the precision of the estimation, I ran a parametric bootstrap (using the \mbox{\textit{confint.merMod}} function from \textit{lme4}) with 1,000 replications and using the percentile method for the calculation of the intervals.
The resulting 95\% bootstrap confidence intervals are reported in Table~\ref{tab:bigtable} in the columns labelled \textit{CI low} and \textit{CI high} (= upper and lower 2.5th percentiles).
The column \textit{CI excludes 0} shows an asterisk for those intervals that do not include 0.
Furthermore, for each regressor, a p-value was obtained by dropping the regressor from the full model, re-estimating the nested model, and comparing it to the full model.
Instead of inexact Wald approximations and Likelihood Ratio Tests, I used a drop-in bootstrap replacement for the Likelihood Ratio Test from the function \textit{PBmodcomp} from the \textit{pbkrtest} package \citep{HalekohHojsgaard2014}.
I call the corresponding value $p_{\text{PB}}$, and it is given in the respective columns in Table~\ref{tab:bigtable}.
Only \textit{Kindfreq} ($\mpPB=0.095$) can be seen as slightly too high to be convincing (non-significant).

Finally, I show that MCMC does not necessarily lead to different results.
Actually, given the low complexity of the model and the large sample size, it would be surprising if they did (see Section~\ref{sec:rightstatistics}).
The model was re-estimated using the \textit{stan\_glmer} function from the \textit{rstanarm} package, which provides an \textit{lme4}-compatible syntax for estimating common model types with the \textit{stan} software \citep{CarpenterEa2017}.
The algorithm was run with 4 chains and 1,000 iterations, and I used plausible default priors.
Most notably, priors for coefficients were specified as $\mathcal{N}(0,10)$ because coefficients higher than 10 or lower than $-10$ are extremely rare in well-specified models on appropriate data.%
\footnote{Consider that with a coefficient of 10, each increase by 1 in the regressor variable increases the odds by $exp(10)=22,026.47$.
To reliably estimate such coefficients, extremely large samples would be required.}
The algorithm converged, and for all coefficients, the $\hat{R}$ diagnostic was exactly 1.
The resulting coefficients and intervals as well as an * for ``confidence interval does not contain 0'' are also given in Table~\ref{tab:bigtable} in the columns labelled \textit{MCMC}.
Both methods lead to exactly the same results (minus negligible numerical differences) as expected given the modest complexity of the model structure and the large sample size.
The signs and magnitudes of the coefficients are identical, and the confidence intervals have the same width and symmetry properties.
Figure~\ref{fig:fixeffs} illustrates this by also showing both estimates.
Since both estimators converge, I only interpret the MLE model in the next section.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Interpretation}
\label{sec:interpretation}

% ATTRACTION

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.5\textwidth]{../R/output/corpus_Measureattraction}~\includegraphics[width=0.5\textwidth]{../R/output/corpus_Kindattraction}
  \caption{Effect plots for the regressors \textit{Measureattraction} and \textit{Kindattraction}; y-axes are not aligned}
  \label{fig:eff:attraction}
\end{figure}

The results reported in Section~\ref{sec:corpushierarchicalmodel} generally confirm the hypotheses from Section~\ref{sec:analyses}.
First, the prototypicality effect related to the non-alternating \PGCd\ and \NACb\ can be shown (see the effect plots in Figure~\ref{fig:eff:attraction}).%
\footnote{Effect plots were created using the \textit{effects} package \citep{Fox2003}.
They show the changes in probability for the outcome (y-axis) dependent on values of a regressor (x-axis) at typical values of all other regressors.
The vertical bars (categorical variables), and the grey areas (continuous variables) are asymptotic 95\% confidence intervals calculated from \textit{glmer}.
They are not bootstrapped.
Readers should be aware that the axes are specifically scaled so as to result in a linear plot, and that the range of the axes varies between plots.}
The effect is as expected:
if a lemma appears relatively more often in the \PGCd\ (compared to its frequency in the \NACb), the \PGCa\ tends to be chosen over the \NACa\ with this specific lemma.
The effect for measure nouns is stronger, and it was estimated with higher precision.

An interesting picture emerges for the lemma frequencies.
A higher-than-average lemma frequency of measure nouns favours the \NACa\ ($\beta_{\text{Measurefreq}}=-0.231$, $\mpPB=0.005$), which is as expected if we assume at least a tendency for highly grammaticalised items to be more frequent.
With kind nouns, higher frequency seems to favour the \PGCa ($\beta_{\text{Kindfreq}}=0.146$, $\mpPB=0.095$).
However, there is no clear theoretical interpretation (see Section~\ref{sec:analyses}), and the estimate is imprecise (not significant at $\alpha=0.05$, see above).
The effect can therefore be ignored or treated as a nuisance variable.

% GRAMMATICALISATION
% => lemmas and lemma classes

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.5\textwidth]{../R/output/corpus_Measureclass}
  \caption{Effect plot for the regressor \textit{Measureclass}}
  \label{fig:eff:measureattraction}
\end{figure}

In Section~\ref{sec:analyses}, it was also hypothesised that classes of measure nouns with a higher degree of grammaticalisation should favour the \NACa.
The \textit{Measureclass} second-level predictor was successfully estimated ($\mpPB=0.001$).
Looking at the effect plot in Figure~\ref{fig:eff:measureattraction}, it is evident that abstract non-referential physical measure nouns (such as \textit{Gramm} `gram' or \textit{Liter} `litre') with a high degree of grammaticalisation favour the \NACa.
At the other end of the scale, nouns denoting natural portions like \textit{Haufen} `heap', \textit{Bündel} `bundle', \textit{Schluck} `gulp' favour the \PGCa.
These are referential nouns, confirming the hypothesis that it is prototypical of the PGC to contain two referential nouns, while the NAC prototypically only contains one (the kind noun).

% CARDINALS

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.5\textwidth]{../R/output/corpus_Cardinal}
  \caption{Effect plot for the regressor \textit{Cardinal}}
  \label{fig:eff:leftcontext}
\end{figure}

I now turn to the predicted effect of cardinals as modifiers of the measure noun.
Figure~\ref{fig:eff:leftcontext} shows that cardinals indeed influence the choice of the alternant ($\mpPB=0.001$), and that cardinals have a strong tendency to co-occur with the \NACa.
This effect was predicted in Section~\ref{sec:analyses}.

% REGISTER
% => Badness
% => Genitives

The style-related proxy variables point to the expected direction.
Increased \textit{Badness} of the document favours the \NACa\ ($\beta_{\text{Badness}}=-0.152$, $\mpPB=0.002$), and so does a lower density of genitives ($\beta=-0.693$, $\mpPB=0.001$).
While these are merely proxies to style (and partially circular in the case of \textit{Genitives}), this result can at least encourage future work into stylistic effects.

% NUISANCE
% => dative effect
% => frequency

The influence of \textit{Measurecase} ($\mpPB=0.001$) is as predicted in previous analyses (see Section~\ref{sec:analyses}).
A measure noun in the dative favours the \PGCa\ with $\beta_{\text{MeasurecaseDat}}=0.705$ (compared to the nominative, which is on the intercept).
Although \textit{Measurecase} is a nuisance variable in the context of this study, convergence with previous work strengthens its validity.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2017-12-22 Latex/sections/4experiments.tex
Sections in separate files.

\section{Experiments}
\label{sec:experimental}

\subsection{Experiment 1: forced-choice}
\label{sec:exp:fc}

In the two experiments reported in this section, I use probabilities for the alternating constructions calculated for attested material, and I correlate these probabilities with the participants' reactions.
Thus, a direct link can be established between output material found in corpora and the behaviour of linguistic agents (see also Section~\ref{sec:cogocl}).
Both experiments use sentences containing attested MNPs from the corpus sample (embedded into simplified sentences) as stimuli.
Also, the probabilities that the corpus-based model assigns to the two alternants in these sentences are used as the main regressor in both studies.

The first experiment tests preferences for constructions explicitly.
\cite{FordBresnan2013} use the \textit{split-100} task in which participants have to distribute 100 points between the alternatives, assigning more points to more natural sounding alternative.
In essence, participants distribute a probability mass between two alternants, which is intended to produce more subtle results compared to a two-alternative forced-choice task such as in \cite{Rosenbach2013}, where participants have to choose one of two alternants.
The split-100 paradigm has been criticised in \cite{ArppeJaervikivi2007}.
The criticism was reiterated in \cite{DivjakEa2016}, where they use a forced-choice task.
In \cite{VerhoevenTemme2017}, it was shown that results from forced-choice and split-100 experiments mostly converge (with some numeric intricacies related to the non-linear distribution of preferences for alternants).
I present a forced-choice experiment in this Section and not a split-100 experiment, mainly because in a dry run of the experiment, participants complained about the unnaturalness of distributing a probability mass across two alternants and tended to produce ratings of 0 and 100 (and to a lesser degree 50).
Participants had to choose between two sentences that differed only in that one contained the \NACa\, and the other one contained the \PGCa.
The analysis compares the probabilities assigned to the stimuli by the corpus-derived model with the frequency with which participants chose the alternants for the same stimuli.

There were 24 participants (native speakers of German without reading or writing disabilities) aged 19 to 30 living permanently in Berlin, who were recruited from introductory linguistics courses at Freie Universität Berlin.
Although the experiment was conducted in the last four weeks of their first semester, participants had no deeper explicit knowledge of linguistics, grammar, or experimental methods.
None of them had ever participated in a forced-choice experiment before.
Participation was voluntary but participants received credit in partial fulfillment of course requirements.

As stimuli, attested MNPs from the corpus study were used, but the sentences were radically simplified to avoid influences from contextual nuisance factors as much as possible.
The approach is also justified because according to the theoretical assessment in Section~\ref{sec:analyses}, the choice of alternants depends mostly on a very local constructional context.
I sampled 16 MNPs from the concordance and made sure that the simplifications and normalisations did not affect any of the regressors used in the corpus study.
In the simplified sentences, the case, number, etc. of the MNP remained the same as in the attested sentence, as did the choice of lexical material within the MNP.
Eight sentences contained masculine or neuter kind nouns, and the other eight contained feminine kind nouns.
Furthermore, in each of the masculine\slash neuter and feminine groups, four sentences originally containing the \NACa\ and four sentences originally containing the \PGCa\ were chosen.
More precisely, the sentences were sampled as \textit{highly prototypical examples} of \PGCa\ (high probability assigned by GLMM) and \NACa\ (low probability assigned by GLMM), respectively.%
\footnote{Remember from Section~\ref{sec:corpusstudies} that the model predicts the probability that the \PGCa\ is chosen over the \NACa.}
High and low probabilities were defined as the top and bottom 20\% of all probabilities assigned by the GLMM.
Lemmas and feature combinations were made unique within each group whenever possible.
The design is summarised in Table \ref{tab:experiment1:design}.

\begin{table}
  \centering
  \begin{tabular}[h]{lll}
     & Masculine\slash Neuter & Feminine \\
     \midrule
     \textbf{high prob.\ for PGC\Subsf{adj}} & 4 sentences & 4 sentences \\
     \textbf{low prob.\ for PGC\Subsf{adj}} & 4 sentences & 4 sentences \\
  \end{tabular}
  \caption{The four groups of sentences chosen as stimuli; in each group of four sentences, combinations of important factor values were made unique whenever possible}
  \label{tab:experiment1:design}
\end{table}

The final pairs of stimuli were the sentence containing the attested and preferred alternant (according to the corpus GLMM) on the one hand and a modified version containing the dispreferred alternant on the other hand.
They were presented next to each other, and a 20 second time limit for each choice was set.%
\footnote{No participant ever exceeded the time limit.}
The position on the screen (left\slash right) and the order of sentences were randomised for each participant.
As fillers, 23 pairs of sentences exemplifying similar but unrelated alternation phenomena from German morpho-syntax were used.
Thus, participants saw 39 pairs of sentences and 78 sentences in total.
They were instructed to select from each pair of sentences the one that seemed more natural to them in the sense that they would use it rather than the other one.
The experiment was conducted using \textit{PsychoPy} \citep{Peirce2007}.

\begin{figure}[hb!]
\centering
\includegraphics[width=0.5\textwidth]{../R/output/fc_effects}
\caption{Effect plot for the multilevel logistic regression in the forced-choice experiment: predictability of participants' choices using the probabilities derived from the corpus-based GLMM}
\label{fig:afc:effects}
\end{figure}

Then, a multilevel logistic regression was specified with the probability of the \PGCa\ predicted for each sentence by the corpus-based GLMM as the only fixed effect \textit{Modelprediction}.%
\footnote{The document-level variables \textit{Badness} and \textit{Genitives} were set to 0, which is the mean for z-transformed variables.}
A random intercept and slope were added for the individual sentence (item) in order to catch idiosyncrasies of single sentences.
Also, a random intercept and slope for participants was added.%
\footnote{The random slopes were added to comply with \citet[257]{BarrEa2013} who predict \textit{catastrophically high Type I error rates} for experimental designs with within-subject manipulations if random effects structures are not kept maximal.
Notice that the model reported here was estimated with conceptually identical results with regard to the predictor of interest (\textit{Modelprediction}) if only random intercepts were used.}
Coefficients were estimated with Maximum Likelihood Estimation (\textit{lmer} function from \textit{lme4}).
The number of observations was \textit{n=}384.

A certain amount of the variance can be accounted for by idiosyncrasies of single sentences ($\sigma_{\text{Sentence}}=1.785$, $\sigma^{\text{Modelprediction}}_{\text{Sentence}}=5.996$, 16 levels).%
\footnote{I use $\sigma^f_r$ to denote the standard deviation of the random intercepts for the fixed effect $f$ varying by random effect $r$.}
Also, among participants, there are clearly different preferences ($\sigma_{\text{Participant}}=0.781$, $\sigma^{\text{Modelprediction}}_{\text{Participant}}=0.484$, 24 levels).
On the extreme ends, one participant chose the \PGCa\ in 13 of 16 cases, and two participants only chose it in 5 of 16 cases.
The regressor \textit{Modelprediction} achieves $\mpPB=0.007$ (1,000 replications) and is estimated at 5.408 relative to an intercept of -1.304.
The confidence interval from a parametric bootstrap (1,000 replications, percentile method) for the regressor is acceptable but slightly large with a lower bound of 1.626 and an upper bound of 8.397.
The pseudo-coefficients of determination are $R^2_{m}=0.227$ and $R^2_{c}=0.561$, which means that over 22\% of the variance in the data can be explained by considering only the predictions from the corpus-based GLMM.
The effect display for the single fixed regressor \textit{Modelprediction} is given in Figure \ref{fig:afc:effects}.
The result is very clear.
The higher the probability of the \PGCa\ predicted from usage data, the more often participants chose the \PGCa\ alternant in the forced-choice task.
In summary, the forced-choice experiment clearly succeeded in corroborating the results from the corpus study in as much as the preferences extracted from usage data correspond to native speakers' choices.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Experiment 2: self-paced reading}
\label{sec:exp:spr}

The second experiment tests preferences more implicitly.
It is expected that reading less prototypical alternants (the one assigned a low probability by the corpus-derived model) in a given context and with given lexical material incurs a processing overhead for the reader (\citealp{Kaiser2013}).
In this section, a self-paced reading experiment is therefore presented.
In a very similar fashion, \cite{DivjakEa2016} apply the self-paced reading paradigm in the validation of corpus-based models.
The analysis compares the corpus-derived probabilities with potential lags in reading time for sentences with the preferred and the non-preferred constructions.

Concretely, the exact same stimuli as in the forced-choice experiments were used.
Each participant read both the 16 sentences with the alternant predicted by the corpus model and the 16 modified sentences with the alternant that the corpus model did not predict.%
\footnote{Notice that lemmas and their frequencies as well as lemma classes are included as regressors in the corpus-based GLMM, and there was consequently no additional controlling of lemma frequencies, etc.}
To minimise repetition effects, the stimuli for each participant were separated into two blocks of 16 targets and 33 fillers per block.
In the experiment, participants first read all sentences from the first block, then all sentences from the second block.
From each target sentence pair, one sentence was assigned to the first block and the other sentence to the other block.
The assignment of members of the individual sentence pairs to the blocks was randomised for each participant individually, as was the order within each block.
The sentences from each pair of alternants were kept as far apart as possible.
The fillers also came in pairs such that the second block exclusively contained sentences to which participants had been exposed in the first block in slightly modified form.
In total, each participant read 98 sentences.
After each sentence, participants had to answer simple (non-metalinguistic) yes-no questions about the previous sentence as distractors.
The distractor questions were different between the first and the second blocks.
There were 38 participants recruited in exactly the same manner as for the experiment reported in Section~\ref{sec:exp:fc}.
None of them had ever participated in any kind of reading experiment, and none of them took part in the first experiment.
The experiment was conducted using \textit{PsychoPy}.

The reading times were residualised per speaker based on the reading times of all words (not just the targets) by that speaker.
The adjective and the kind noun (\ie\ the constituents bearing the critical case markers) were used as the target region, for instance the bracketed words in the example \textit{zwei Gläser} [\textit{spru\-deln\-des Wasser}] `two glasses of sparkling water'.
Outliers farther than 2 inter-quartile ranges from the mean logarithmised residualised reading time were removed (64 data points), resulting in a total number of \textit{n=}1,152 observations.
An LMM was specified with the logarithmised residual reading times as the response variable.

The probabilities derived from the corpus GLMM (\textit{Modelprediction}) were added as the main regressor of interest.
It should be remembered that the corpus GLMM predicts the probability of the \PGCa.
As a consequence, the higher the GLMM prediction is, the more prototypical the sentence is for containing the \PGCa.
It is therefore expected that reading times are higher when the value of \textit{Modelprediction} is higher but the sentence contains the \NACa.
However, when the sentence contains the \PGCa, reading times should be lower when \textit{Modelprediction} is higher.
To account for this, an interaction between \textit{Modelprediction} and \textit{Construction} (levels \textit{PGCa} and \textit{NACa}) was added to the model.

Furthermore, the position (1--98) of the sentence in the individual experiment (\textit{Position}) was included as a fixed effect to control for the usual increase in reading speed during an experiment run.
Random intercepts were specified for \textit{Participant} and \textit{Item} (the 16 sentence pairs are one \textit{Item} each).%
\footnote{I tried random slopes in order to keep the random effect structure maximal \citep{BarrEa2013} but it was impossible to get the algorithm to converge due to the added complexity of the interaction.}

\begin{table}
  \centering
  \begin{tabular}{lrrrc}
    Regressor & \multicolumn{1}{r}{Coefficient} & \multicolumn{1}{r}{CI low} & \multicolumn{1}{r}{CI high} & \multicolumn{1}{r}{CI excludes 0} \\ \midrule
    ConstructionPGCa                 &  0.054 &  0.012 &  0.095 &  *  \\ 
    Modelprediction                  & -0.006 & -0.113 &  0.110 &     \\ 
    Position                         & -0.005 & -0.005 &  0.004 &     \\ 
    ConstructionPGCa:Modelprediction & -0.125 & -0.234 & -0.023 &  *  \\ 
  \end{tabular}
  \caption{Fixed effect coefficient table for the LMM used to analyse the self-paced reading experiment; the intercept is 0.829}
  \label{tab:exp:spr}
\end{table}

Table~\ref{tab:exp:spr} shows the coefficient estimates with a 95\% parametric bootstrap confidence interval (1,000 replications, percentile method).
The standard deviation of the participant intercepts is $\sigma_{\text{Participant}}=0.079$ and of the item intercepts $\sigma_{\text{Item}}=0.037$.
Comparing the full model to a model without the main regressor \textit{Modelprediction} (and consequently also without the interaction with \textit{Construction}) in a PB test gives $\mpPB=0.036$.
The pseudo-coefficients of determination are $R^2_m=0.237$ and $R^2_c=0.346$.
There is thus some inter-subject variation and a noteworthy correlation between the reactions of the subjects and the corpus-derived probabilities.
\hl{The fact that the correlation is not stronger can be explained in parts \ldots}

\begin{figure}[h]
\centering
\includegraphics[width=0.5\textwidth]{../R/output/spr_effects}
\caption{Effect plot for the LMM in the self-paced reading experiment: modeling participants' residualised log reading times on the probabilities given by the corpus-based GLMM}
\label{fig:spr:effects}
\end{figure}

The effect plot for the effect of interest is shown in Figure~\ref{fig:spr:effects}.
The estimate for the sentences with \NACa\ is obviously imprecise and no differences in reading times are observed.
There is a clearer effect in the sentences with \PGCa, which is also confirmed by the significant results from the bootstrapped confidence intervals (see Table~\ref{tab:exp:spr}) and from the PB test reported above.
The \PGCa\ brings about an increased reading time ($\beta_{\text{ConstructionPGCa}}=0.054$), which is plausible because it is the much rarer construction (see Section~\ref{sec:corpusstudies}).
However, if it occurs in a prototypical context and with prototypical lexical material, reading times drop ($\beta_{\text{ConstructionPGCa:Modelprediction}}=-0.125$).
This can be seen in the downward slope in the right panel of Figure~\ref{fig:spr:effects}.
This fits into the general picture inasmuch as the construction with the lower frequency might be developing towards a more sharply defined prototype.%
\footnote{In this context, it should be remembered from Section~\ref{sec:gettingdata} that even the \PGCd\ is much rarer that the \NACb\ (17,252 vs.\ 315,635 occurrences in the auxiliary corpus samples).}
Conversely, the \NACa\ (like the NAC in general) might be the highly frequent default which does not incur a reading time penalty, even if it is not the optimal choice in the given context and with the given lexical material.

This concludes the report of the two experiments.
In the final section, I take stock and summarise the contribution of the present study to the research on alternations in cognitive linguistics.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2017-12-22 Latex/sections/5conclusions.tex
Sections in separate files.

\section{Conclusions}
\label{sec:conclusion}

This paper stands in a now ten year-old tradition of research on grammatical alternations using corpus and experimental data.
In my view, the main tenets of this line of research are:
(i)~Language, viewed from a cognitively realistic angle, is a probabilistic phenomenon and cannot be modelled appropriately within Aristotelian frameworks that assume discrete categories.
(ii)~Corpora are collections of usage events (language production) and can therefore be used to evaluate both the claim made in (i) and specific theoretical claims (in case studies) about factors influencing speakers' decisions to use specific forms or constructions.
(iii)~Given (ii), we expect results from corpus analyses and from appropriate experiments to yield similar results, not necessarily as a form of validation of the corpus-based findings, but as converging evidence.
I consider these three points to be of utmost importance because they clearly set this approach to linguistic research apart from \textit{both} Aristotelian frameworks \textit{and} introspective, non-empirical, and anti-quantitative versions of cognitive linguistics (see \citealp{Dabrowska2016} for a pithy philippic against such approaches).

The present paper adds to the evidence that all of the aforementioned three points are correct.
A grammatical alternation in German measure NPs was examined using corpus data based on factors partly derived from existing accounts, formulated in terms of construction prototypes.
The preferences extracted from the DECOW web corpus were confirmed in a forced-choice experiment, in which participants explicitly chose alternants in line with the probabilities derived from the corpus-based model.
In a more implicit self-paced reading experiment, it was shown that the much rarer alternant brings about a reading time penalty except in cases for which the corpus model predicts very high probability for this alternant.

Future work could extend these results and provide a general picture of the constructions expressing measurements (see Section~\ref{sec:descriptive}).
This would be a much more complicated task given that the choices then would no longer be binary and that the meaning of the alternative ways of expressing measurements are semantically more varied.
Finally, I want to point out that German is mildly under-researched in the specific framework used here.
This is quite surprising given the fact that German morpho-syntax is famous for its alternations, which are usually called \textit{Zweifelsfälle} (`cases of doubt') in the traditional literature \citep{Duden09,Klein2009}.
Instead of being drowned in normative, descriptive, or didactic discussions, they could serve as ideal test cases in cognitive linguistics.

### 2017-12-22 Latex/article.tex
More reviewer-induced changes to Sec. 1.

@@
   \affil[1]{Freie Universität Berlin}
-  \title{Competing Constructions for German Measure Noun Phrases}
-  \runningtitle{Competing Constructions for German Measure Noun Phrases}
+  \title{Abstractions and exemplar effects in the German measure noun phrase alternation}
+  \runningtitle{Abstractions and exemplar effects in the German measure noun phrase alternation}
   %\subtitle{...}

### 2017-12-22 Latex/sections/1intro.tex
More reviewer-induced changes to Sec. 1.

@@
-\section{Cognitively oriented corpus linguistics}
+\section{Prototypes, and exemplars, and corpora}
 \label{sec:cogocl}
@@
 By \textit{alternation} I refer to a situation where two or more forms or constructions are available with no clear difference in acceptability, function, or meaning.
-The study of alternations has a long history in cognitively oriented corpus linguistics (\citealp{BresnanEa2007,BresnanHay2010,BresnanFord2010,DivjakArppe2013,Gries2015,NessetJanda2010}, to name just a few publications).
+The study of alternations has a long history in cognitively oriented corpus linguistics (for example, \citealp{BresnanEa2007,BresnanHay2010,BresnanFord2010,DivjakArppe2013,Gries2015,NessetJanda2010}).
 This area of research is based on the assumption that language is a probabilistic phenomenon \citep{Bresnan2007} where alternants are chosen neither deterministically nor fully at random.
@@
 The estimation of the model coefficients quantifies the influence that the factors have on the probability that either alternant is chosen.
+There are two fundamental issues to consider.
+First, there is a question of how corpus data relates to experimental findings, which provide more direct evidence of mental representations and processes.
+Second, the appropriate modelling of such results in cognitive linguistics is a key issue.
 
-For such probabilistic generalisations estimated on corpus data, there has always been an interest in correlating the findings with results from experimental work (for example, \citealp{ArppeJaervikivi2007,BresnanEa2007,BresnanFord2010,DivjakGries2008,DivjakEa2016,FordBresnan2013}).
+As for the first issue, there has always been an interest in correlating of probabilistic generalisations extracted from corpus data with results from experimental work (for example, \citealp{ArppeJaervikivi2007,BresnanEa2007,BresnanFord2010,DivjakGries2008,DivjakEa2016,FordBresnan2013}).
 This is often called a \textit{validation} of the corpus-derived findings, but \citet[303]{Divjak2016a} rightly criticises this choice of words ``because it creates the impression that behavioral experimental data is inherently more valuable than textual data,'' citing \cite{TummersEa2005}, who state that a corpus is ``a sample of spontaneous language use that is (generally) realized by native speakers.''
-However, as \citet[486--487]{Dabrowska2016} convincingly argues, this does not imply that we can in some way ``deduce mental representations from patterns of use,'' \ie\ from corpus data.%
-\footnote{In its colloquial meaning, the verb \textit{deduce} seems appropriate here.
-However, in terms of the scientific method, we are most likely talking about \textit{induction} rather than \textit{deduction}.}
+However, as \citet[486--487]{Dabrowska2016} convincingly argues, this does not imply that we can in some way ``deduce mental representations from patterns of use,'' \ie\ from corpus data.
 It would be highly surprising if this were possible, and the same holds for experimental methods.
@@
 Then, predictions are derived from this theory \textit{before} the experiment or the corpus study is conducted to \textit{test} the theory.
-Technical problems of statistical inference aside, a successful experiment corroborates the theory, and a failed experiment weakens the theory.
-The simplest assumption would be that the same approach is adequate for corpus studies.
+While the same approach is by and large applicable to corpus data.
 
-The present study is conducted within the general paradigm of cognitive corpus linguistics and includes a comparison of the corpus findings with results from two experiments.
-I assume a model of similarity-based classification in the form of Prototype Theory for modeling alternation.
-Protoype Theory has been used in alternation research in cognitively oriented corpus linguistics (see \citealp{DivjakArppe2013}; \citealp{Gries2003}).
-It is assumed that the variables annotated in a corpus study on an alternation phenomenon define prototypes for the alternants.
-The similarity of a given feature vector (in a concrete sentence where one of the alternants is used) influences to a large extent which alternant is chosen by speakers.
-This entails that a variant of Prototype Theory with features is assumed \citep{Rosch1978}, instead of the monolithic prototypes of earlier versions.
-Prototype Theory is well suited for modeling constructional choices but it is just one of several similarity-based theories of classification, the most prominent other framework being Exemplar Theory (\citealp{MedinSchaffer1978,Hintzman1986}; see \citealp{StormsEa2000} for a comparison of the theories in experimental settings).
-Prototype Theory and Exemplar Theory model essentially the same types of effects and differ mainly in whether they assume higher-level abstractions as part of the cognitive representation (Prototype Theory) or try to do without them (Exemplar Theory).
-I am sceptical that corpus analysis alone could ever decide which theory is more suitable, given that substantial doubt has been voiced whether even experimental methods are ultimately able to do so \citep{Barsalou1990}.%
-\footnote{In \cite{DivjakArppe2013}, it was shown using corpus data how both models form a converging picture.}
-Prototype Theory (with feature abstractions) is preferred here simply because it fits the established alternation modeling paradigm, which relies on features being weighted in statistical models.
-In Section~\ref{sec:analyses}, a prototype-theoretical parlance is therefore adopted.
+\hl{Discussion of non-convergence here.}
 
-In the remainder of the paper, I introduce the alternation phenomenon (Section~\ref{sec:descriptive}) and suggest a theory-driven set of factors influencing the alternation (Section~\ref{sec:analyses}).
-Then the corpus study is presented, including an appropriate statistical analysis (Section~\ref{sec:corpusstudies}).
-Two experiments which corroborate the corpus-based findings are then reported in Section~\ref{sec:experimental} before I sum up the paper in Section~\ref{sec:conclusion}.
+\hl{Turning to the second issue, }
+This entails that a variant of prototype theory with features is assumed \citep{Rosch1978}.
+Prototype theory is well suited for modeling constructional choices, but it is just one of several similarity-based theories of classification, the most prominent other framework being Exemplar Theory (\citealp{MedinSchaffer1978,Hintzman1986}; see \citealp{StormsEa2000} for a comparison of the theories in experimental settings).
+Prototype Theory and Exemplar Theory model essentially the same types of effects and differ mainly in whether they assume higher-level abstractions as part of the cognitive representation of categories (prototype theory) or assume that categories emerge through the storage of many exemplars and similarity classification on those exemplars (exemplar theory).
+As \cite{Barsalou1990} shows, the two theories model the same types of effects and are informationally equivalent.
+Consequently, experiments which favour one theory over the other use procedural behaviour of subjects, for example the speed of category retrieval.
+In very early experiments, \cite{PosnerKeele1968} showed, for example, that highly prototypical unseen exemplars were categorised more easily by subjects compared to less prototypical ones, even if these were included in the learning data.
+Since corpus data only show artefacts of production events and we have no experimental access to the speakers or writers performance, one should be sceptical whether corpus analysis alone could ever decide which theory of mental representation is more suitable (see also \citealp[22]{Gries2003} and the \citealp[486--487]{Dabrowska2016} quote above).
+%The corpus-based approach in \cite{DivjakArppe2013}, for example, primarily shows that corpus data are compatible with aspects of prototype and exemplar theory.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\cite{VanpaemelStorms2008,LeeVanpaemel2008} proposed the \textit{varying abstraction model} (VAM) which ``attempt[s] to balance economy and informativeness'' \citep[745]{LeeVanpaemel2008}, treating models with full abstraction (prototypes) and no abstraction at all (radical exemplar theory) as special cases of a model which allows for both abstraction and exemplar effects.
+However, \cite[183--184]{Vanpaemel2016} describes the state of affairs between adherents of neo-prototype theory (such as \citealp{MindaSmith2001,MindaSmith2002}) and exemplar theory as a stalemate, and suggests yet another approach to solve the divide.
+
+\cite{ConawayKurtz2016} show that reference point approaches to category learning such as prototype and exemplar theory fail to explain certain experimental results where subjects learn to generalise beyond the input in a way that cannot be explained by similarity.
+
+\cite{GriffithsEa2009} present a computational model based on the hierarchichal Dirirchlet process which is able to chose the appropriate complexity of representation for a given category.
+
+In \cite{HahnEa2010}, the authors show that subjects even use similarity (exemplar) effects when they are given very simple explicit rules, which is highly relevant because most studies focus on implicit rule-based knowledge, which involves many auxiliary assumptions in actual experiments \citep[2]{HahnEa2010}.
+
+An aspect generally neglected in models of categorisation is how categorisation feeds exemplar generation and creativity (\ie category generation), two subjects of utmost importance to many linguists.
+In \cite{JernKemp2013}, a model is assumed where cognitive agents sample from learned distributions over both categories and exemplars to generate new ones.
+
+One thing that should be kept in mind is that most of the research on categorisation in cognitive science uses experiments with highly simplified stimuli and very simple tasks.
+It is remarkable in this context that \cite{VoorspoelsEa2011} consider the experimental task reported by them -- assigning typicality scores to nouns from the domains of \textit{animals} and \textit{artefacts} to categories like \textit{bird}, \textit{fish}, \textit{clothing}, or \textit{tools} -- a study of ``superordinate natural language categories, whereas most evidence supporting exemplar representations has been found in artificial categories of a more subordinate level'' \citep[1013]{VoorspoelsEa2011}.
+Linguists (not just of the cognitive variety) are usually interested in high-level generalisations and complex categories, especially in (morpho-)syntax.%
+\footnote{Notice that recently, approaches have emerged which solve at least some problems by abandoning typically linguistic high-level features altogether \citep{BaayenEa2016,RamscarPort2016}.}
+Since, as a consequence, formal models of categorisation in cognitive linguistics have not been spelled out at the same level of mathematical detail as in cognitive science, I propose that the primary focus should be on determining which factors influence choices made by speakers.
+As \cite[22]{Gries2003} put it with reference to classic prototype theory, ``even if the form of analysis does not translate into statements on mental representations, the high predictive power [...] shows that the cognitive factors underlying the choice of construction have been identified properly and weighted in accordance with their importance for actual usage.''
+A major advancement since this early time has been the focus on which effects are at least more plausibly modelled as prototype\slash abstraction and exemplar effects, and how observed data fits the two approaches (see, for example, \citealp{DivjakArppe2013}).
+This paper contributes this discussion.
+
+% The present study is conducted within the general paradigm of cognitive corpus linguistics and includes a comparison of the corpus findings with results from two experiments.
+% I assume a model of similarity-based classification in the form of Prototype Theory for modeling alternation.
+% Protoype Theory has been used in alternation research in cognitively oriented corpus linguistics (see \citealp{DivjakArppe2013}; \citealp{Gries2003}).
+% It is assumed that the variables annotated in a corpus study on an alternation phenomenon define prototypes for the alternants.
+% The similarity of a given feature vector (in a concrete sentence where one of the alternants is used) influences to a large extent which alternant is chosen by speakers.
+% This entails that a variant of Prototype Theory with features is assumed \citep{Rosch1978}, instead of the monolithic prototypes of earlier versions.
+% Prototype Theory is well suited for modeling constructional choices but it is just one of several similarity-based theories of classification, the most prominent other framework being Exemplar Theory (\citealp{MedinSchaffer1978,Hintzman1986}; see \citealp{StormsEa2000} for a comparison of the theories in experimental settings).
+% Prototype Theory and Exemplar Theory model essentially the same types of effects and differ mainly in whether they assume higher-level abstractions as part of the cognitive representation (Prototype Theory) or try to do without them (Exemplar Theory).
+% I am sceptical that corpus analysis alone could ever decide which theory is more suitable, given that substantial doubt has been voiced whether even experimental methods are ultimately able to do so \citep{Barsalou1990}.%
+% \footnote{In \cite{DivjakArppe2013}, it was shown using corpus data how both models form a converging picture.}
+% Prototype Theory (with feature abstractions) is preferred here simply because it fits the established alternation modeling paradigm, which relies on features being weighted in statistical models.
+% In Section~\ref{sec:analyses}, a prototype-theoretical parlance is therefore adopted.
+% 
+% In the remainder of the paper, I introduce the alternation phenomenon (Section~\ref{sec:descriptive}) and suggest a theory-driven set of factors influencing the alternation (Section~\ref{sec:analyses}).
+% Then the corpus study is presented, including an appropriate statistical analysis (Section~\ref{sec:corpusstudies}).
+% Two experiments which corroborate the corpus-based findings are then reported in Section~\ref{sec:experimental} before I sum up the paper in Section~\ref{sec:conclusion}.

### 2017-12-23 Latex/sections/1intro.tex
More editing in Sec. 1

@@
 
-As for the first issue, there has always been an interest in correlating of probabilistic generalisations extracted from corpus data with results from experimental work (for example, \citealp{ArppeJaervikivi2007,BresnanEa2007,BresnanFord2010,DivjakGries2008,DivjakEa2016,FordBresnan2013}).
+As for the first issue, there has always been an interest in correlating probabilistic generalisations extracted from corpus data with results from experimental work (for example, \citealp{ArppeJaervikivi2007,BresnanEa2007,BresnanFord2010,DivjakGries2008,DivjakEa2016,FordBresnan2013}).
 This is often called a \textit{validation} of the corpus-derived findings, but \citet[303]{Divjak2016a} rightly criticises this choice of words ``because it creates the impression that behavioral experimental data is inherently more valuable than textual data,'' citing \cite{TummersEa2005}, who state that a corpus is ``a sample of spontaneous language use that is (generally) realized by native speakers.''
@@
 This entails that a variant of prototype theory with features is assumed \citep{Rosch1978}.
-Prototype theory is well suited for modeling constructional choices, but it is just one of several similarity-based theories of classification, the most prominent other framework being Exemplar Theory (\citealp{MedinSchaffer1978,Hintzman1986}; see \citealp{StormsEa2000} for a comparison of the theories in experimental settings).
-Prototype Theory and Exemplar Theory model essentially the same types of effects and differ mainly in whether they assume higher-level abstractions as part of the cognitive representation of categories (prototype theory) or assume that categories emerge through the storage of many exemplars and similarity classification on those exemplars (exemplar theory).
-As \cite{Barsalou1990} shows, the two theories model the same types of effects and are informationally equivalent.
-Consequently, experiments which favour one theory over the other use procedural behaviour of subjects, for example the speed of category retrieval.
+Prototype theory is well suited for modeling constructional choices, but it is just one of several similarity-based theories of classification, the most prominent other framework being exemplar theory (\citealp{MedinSchaffer1978,Hintzman1986}; see \citealp{StormsEa2000} for a comparison of the theories in different experimental settings).
+Prototype theory and exemplar theory model essentially the same types of effects but differ significantly in whether they assume higher-level abstractions in the form of single maximally prototypical exemplars or their features (prototype theory) or assume that categories emerge through the storage of many exemplars and similarity classification on those exemplars (exemplar theory).
+\hl{Parallel to Langacker here}
+
+As \cite{Barsalou1990} already showed, however, prototype and exemplar theory model the same types of effects and are informationally equivalent.
+Consequently, experiments which favour one theory over the other use procedural behaviour of subjects, for example the speed of category retrieval, and not mere output data.
 In very early experiments, \cite{PosnerKeele1968} showed, for example, that highly prototypical unseen exemplars were categorised more easily by subjects compared to less prototypical ones, even if these were included in the learning data.
-Since corpus data only show artefacts of production events and we have no experimental access to the speakers or writers performance, one should be sceptical whether corpus analysis alone could ever decide which theory of mental representation is more suitable (see also \citealp[22]{Gries2003} and the \citealp[486--487]{Dabrowska2016} quote above).
-%The corpus-based approach in \cite{DivjakArppe2013}, for example, primarily shows that corpus data are compatible with aspects of prototype and exemplar theory.
+Since corpus data only show artefacts of production events and we have no experimental access to the speaker's or writer's performance, one should be sceptical whether corpus analysis alone could ever decide which theory of mental representation is more suitable (see also \citealp[22]{Gries2003} and the \citealp[486--487]{Dabrowska2016} quote above).
 
+In cognitive science, it is mostly accepted that exemplar theories have greater explanatory power \citep[184]{Vanpaemel2016}, and that abstraction is only needed marginally, if at all.%
+\footnote{The hard empirical evidence in favour of exemplar models is substantial.
+  For example, in \cite{HahnEa2010}, the authors show that subjects even use exemplar similarity over abstract knowledge even when they are given very simple explicit rules.
+  This is highly relevant because most other studies focus on the learning of implicit rule-based knowledge, which involves many auxiliary assumptions in actual experiments \citep[2]{HahnEa2010}.
+On the other hand, there is evidence that neither theory is fully adequate to model humans' capabilities to form categories.
+For example, \cite{ConawayKurtz2016} show that reference point approaches to category learning such as prototype and exemplar theory fail to explain certain experimental results where subjects learn to generalise beyond the input in a way that cannot be explained by similarity.
+}
+Still, various attempts have been made over the past decades to settle the dispute between abstraction-based models (models with rules or prototypes) and exemplar models or to find models which unite the two extremes.
 \cite{VanpaemelStorms2008,LeeVanpaemel2008} proposed the \textit{varying abstraction model} (VAM) which ``attempt[s] to balance economy and informativeness'' \citep[745]{LeeVanpaemel2008}, treating models with full abstraction (prototypes) and no abstraction at all (radical exemplar theory) as special cases of a model which allows for both abstraction and exemplar effects.
-However, \cite[183--184]{Vanpaemel2016} describes the state of affairs between adherents of neo-prototype theory (such as \citealp{MindaSmith2001,MindaSmith2002}) and exemplar theory as a stalemate, and suggests yet another approach to solve the divide.
+The mixture model of categorisation (MMC) by \cite{Rosseel2002} is a model with abstraction in the form of hierarchical clusters of exemplars, and these clusters of objects are characterised by a probability distribution over their features, and categorising new objects is a process of estimating the probability of this object being from one of the clusters.
+\cite{GriffithsEa2009} go further and present a computational model based on the hierarchical Dirirchlet process which is able to chose the appropriate complexity of representation for a given category.
+However, despite these (and more) attempts to , \cite[183--184]{Vanpaemel2016} describes the state of affairs between adherents of neo-prototype theory (such as \citealp{MindaSmith2001,MindaSmith2002}) and exemplar theory as a stalemate, and suggests yet another approach to solve the divide.
 
-\cite{ConawayKurtz2016} show that reference point approaches to category learning such as prototype and exemplar theory fail to explain certain experimental results where subjects learn to generalise beyond the input in a way that cannot be explained by similarity.
+%An aspect generally neglected in models of categorisation is how categorisation feeds exemplar generation and creativity (\ie category generation), two subjects of utmost importance to many linguists.
+%In \cite{JernKemp2013}, a model is assumed where cognitive agents sample from learned distributions over both categories and exemplars to generate new ones.
 
-\cite{GriffithsEa2009} present a computational model based on the hierarchichal Dirirchlet process which is able to chose the appropriate complexity of representation for a given category.
-
-In \cite{HahnEa2010}, the authors show that subjects even use similarity (exemplar) effects when they are given very simple explicit rules, which is highly relevant because most studies focus on implicit rule-based knowledge, which involves many auxiliary assumptions in actual experiments \citep[2]{HahnEa2010}.
-
-An aspect generally neglected in models of categorisation is how categorisation feeds exemplar generation and creativity (\ie category generation), two subjects of utmost importance to many linguists.
-In \cite{JernKemp2013}, a model is assumed where cognitive agents sample from learned distributions over both categories and exemplars to generate new ones.
+In cognitive linguistics, \cite{DivjakArppe2013} is a very rare example of a paper where such issues are taken up.
+Their corpus-based approach shows ``one way of systematically analyzing usage data as contained in corpora to yield a scheme, compatible with usage-based theories of language, by which the assumptions of both the prototype and exemplar theories can be operationalized'' \citep[267]{DivjakArppe2013}.
+Their approach to implementing a varying abstraction model \citep[254--260]{DivjakArppe2013} is based on hierarchical clustering of annotated properties of sentences.
+They hierarchically cluster sentences containing Russian verbs of trying.
+Then, they single out one sentence from each cluster which scores the highest probability for any of the six \textit{try} verbs according to a polytomous regression model estimated on the same data.
+The clusters are interpreted as intermediate-level exemplar-derived abstractions of typical contexts for these high-probability verbs (typically more than one cluster for each verb; \citealp[255--256]{DivjakArppe2013}).
+This is pioneering work, but the crucial difference between data-driven corpus-based analyses and experiments in cognitive science (they use \citealp{VerbeemenEa2007} as their reference) is that cognitive experimental research is based on experiments where subjects produce category assignments, and in corpus studies, the categories and category membership is determined purely from the data.
 
@@
 It is remarkable in this context that \cite{VoorspoelsEa2011} consider the experimental task reported by them -- assigning typicality scores to nouns from the domains of \textit{animals} and \textit{artefacts} to categories like \textit{bird}, \textit{fish}, \textit{clothing}, or \textit{tools} -- a study of ``superordinate natural language categories, whereas most evidence supporting exemplar representations has been found in artificial categories of a more subordinate level'' \citep[1013]{VoorspoelsEa2011}.
-Linguists (not just of the cognitive variety) are usually interested in high-level generalisations and complex categories, especially in (morpho-)syntax.%
-\footnote{Notice that recently, approaches have emerged which solve at least some problems by abandoning typically linguistic high-level features altogether \citep{BaayenEa2016,RamscarPort2016}.}
-Since, as a consequence, formal models of categorisation in cognitive linguistics have not been spelled out at the same level of mathematical detail as in cognitive science, I propose that the primary focus should be on determining which factors influence choices made by speakers.
+Linguists (not just of the cognitive variety) are usually interested in much more complex high-level categories and use large and complex feature sets, especially in (morpho-)syntax.%
+\footnote{Notice that recently, approaches have emerged which solve at least some problems by abandoning linguistic high-level features altogether \citep{BaayenEa2016,RamscarPort2016}.}
+
+Most likely a consequence of linguists' interest in  is that formal models of categorisation in cognitive linguistics have not been spelled out at the same level of mathematical detail as in cognitive science.%
+\footnote{Again using \cite{DivjakArppe2013} as an example, it is obvious that they use advanced statistics in the form of polytomous regression and hierarchical agglomerative clustering.
+However, models like the one proposed in \cite{GriffithsEa2009} are fully spelled out probability models of the behaviour of cognitive agents.}
+However, the fact that linguists deal with real data should be seen as an advantage in that it greatly improves the \textit{external validity} of studies, \ie the generalisability of the findings.
+Typical artificial tasks in cognitive science have been criticised exactly for the lack of external validity (\eg \citealp{Murphy2003}).
+
+I propose that the primary focus should be on determining which factors influence choices made by speakers.
 As \cite[22]{Gries2003} put it with reference to classic prototype theory, ``even if the form of analysis does not translate into statements on mental representations, the high predictive power [...] shows that the cognitive factors underlying the choice of construction have been identified properly and weighted in accordance with their importance for actual usage.''

### 2017-12-25 Latex/article.tex
Nearing completion of Sec. 1 rewrite.

--- 2017-12-22 Latex/article.tex
@@
   \affil[1]{Freie Universität Berlin}
-  \title{Competing Constructions for German Measure Noun Phrases}
-  \runningtitle{Competing Constructions for German Measure Noun Phrases}
+  \title{Abstractions and items: the measure noun phrase alternation in German}
+  \runningtitle{Abstraction and items: the measure noun phrase alternation German}
   %\subtitle{...}
@@
   I thank (in alphabetical order) Felix Bildhauer, Susanne Flach, Elizabeth Pankratz, Samuel Reichert, Ulrike Sayatz, and Christian Zimmer for valuable discussions and comments.
-  Also, I would like to thank two reviewers for Cognitive Linguistics with deep linguistics knowledge and knowledge of German grammar for comments which helped to improve the paper a lot.
-  I thank Dagmar Divjak who, in her role as editor for Cognitive Linguistics, also read the paper thoroughly and provided very useful comments.
-
+  Also, I would like to thank the reviewers for Cognitive Linguistics as well as associate editor Dagmar Divjak for insightful comments which helped to improve the quality of this paper a lot.
   Furthermore, I thank Ulrike Sayatz for helping me to recruit the participants for the experiments.

### 2017-12-25 Latex/sections/1intro.tex
Nearing completion of Sec. 1 rewrite.

@@
 By \textit{alternation} I refer to a situation where two or more forms or constructions are available with no clear difference in acceptability, function, or meaning.
-The study of alternations has a long history in cognitively oriented corpus linguistics (for example, \citealp{BresnanEa2007,BresnanHay2010,BresnanFord2010,DivjakArppe2013,Gries2015,NessetJanda2010}).
+The study of constructional alternations has a long history in cognitively oriented corpus linguistics (for example, \citealp{BresnanEa2007,BresnanHay2010,BresnanFord2010,DivjakArppe2013,Gries2015,NessetJanda2010,Wulff2003}).
 This area of research is based on the assumption that language is a probabilistic phenomenon \citep{Bresnan2007} where alternants are chosen neither deterministically nor fully at random.
@@
 The estimation of the model coefficients quantifies the influence that the factors have on the probability that either alternant is chosen.
-There are two fundamental issues to consider.
-First, there is a question of how corpus data relates to experimental findings, which provide more direct evidence of mental representations and processes.
+There are two fundamental issues to consider with respect to this tradition as a part of cognitive linguistics in the broad sense.
+First, there is the question of how corpus data relates to experimental findings (which provide more direct evidence of mental representations and processes but are often intrinsically narrower in scope), and how well the two converge or should converge.
 Second, the appropriate modelling of such results in cognitive linguistics is a key issue.
@@
 As for the first issue, there has always been an interest in correlating probabilistic generalisations extracted from corpus data with results from experimental work (for example, \citealp{ArppeJaervikivi2007,BresnanEa2007,BresnanFord2010,DivjakGries2008,DivjakEa2016,FordBresnan2013}).
-This is often called a \textit{validation} of the corpus-derived findings, but \citet[303]{Divjak2016a} rightly criticises this choice of words ``because it creates the impression that behavioral experimental data is inherently more valuable than textual data,'' citing \cite{TummersEa2005}, who state that a corpus is ``a sample of spontaneous language use that is (generally) realized by native speakers.''
-However, as \citet[486--487]{Dabrowska2016} convincingly argues, this does not imply that we can in some way ``deduce mental representations from patterns of use,'' \ie\ from corpus data.
+This is often called a \textit{validation} of the corpus-derived findings, but \citet[303]{Divjak2016a} criticises this choice of words ``because it creates the impression that behavioral experimental data is inherently more valuable than textual data,'' citing \cite{TummersEa2005}, who state that a corpus is ``a sample of spontaneous language use that is (generally) realized by native speakers''.
+However, as \citet[486--487]{Dabrowska2016} convincingly argues, this does not imply that we can in some way ``deduce mental representations from patterns of use,'' \ie\ from corpus data.%
+\footnote{Epistemologically, I think she refers to acts of \textit{induction} rather than \textit{deduction}.}
 It would be highly surprising if this were possible, and the same holds for experimental methods.
@@
 Rather, under the standard approach, a theory of cognitive representation is pre-specified.
-Then, predictions are derived from this theory \textit{before} the experiment or the corpus study is conducted to \textit{test} the theory.
+Then, predictions are derived from this theory \textit{before} the experiment or the corpus study is conducted in order to \textit{test} the theory.
 While the same approach is by and large applicable to corpus data.
 
-\hl{Discussion of non-convergence here.}
+An important question, however, is whether the usage data found in corpora are truly predictive of speakers' and writers' cognitive representation of language and\slash or of their overall linguistic behaviour.
+An overview of this issue was given by \cite{NewmanSorensenduncan2015}, who enumerate a number of studies which show how corpus data and experimental data converge (such as \citealp{DurrantDoherty2010,BresnanEa2007,GriesWulff2005,GriesEa2005}) and a number of studies where the two types of data did not lead to diverging or only partially converging results (such as \citealp{ArppeJaervikivi2007,Dabrowska2014,Mollin2009}).
+When researchers do not reach convergence, they mostly try to explain this by differentiating between the actual cognitive construct and what the pooled usage data as found in corpora represent, for example \citet[411]{Dabrowska2014}, who lists a long number of possible reasons to explain why subjects in her experiment diverged in their word association preferences from collocation measures extracted from corpora.
+In some cases, researchers argue for a more adequate statistical analysis to increase the fit between corpus data and subjects' reactions in experiments, for example \cite{DivjakEa2016}, who show that generalised additive models (GAMs) are better suited than generalised linear models (GLMs) in the analysis of their reading-time experiment based on data extracted from previous corpus studies.
+Much more optimistically compared to these publications dealing with possible frictions between usage and experimental data, \cite{StefanowitschFlach2016} recently proposed a straightforward positivist perspective of corpora as representing the input of an average adult speaker, thus licensing inferences from corpus data to cognitive representations under a ``corpus-as-input'' view.
+They state that ``in this wider context, large, register-mixed corpora such as the British National Corpus [\ldots] may not be perfect models of the linguistic experience of adult speakers, but they are reasonably close to the input of an idealized average member of the relevant speech community'' \citep[104]{StefanowitschFlach2016}.%
+\footnote{This, of course, raises the old question of how corpora should be composed to represent what they are intended to represent.
+As points of reference for determining ``representativeness'', at least three major criteria have been used, namely the distribution of texts or text types in the output of all speakers of a language (production-based), the distribution of the relevance of texts or text types in the whole speech community (relevance-based), and the distribution of speakers' exposure to different texts or text types (perception-based).
+See \citet{Biber1993}, \citet{MceneryEa2006}, \citet{Leech2007}, \citet{Hunston2008} for overviews from different perspectives.}
+Clearly, no clear picture has yet emerged, which is not surprising given the vast number of cognitive constructs assumed at diverse levels (such as the level of words in collocation research and much more complex constructional levels in alternation research involving lexical, syntagmatic, and contextual factors), the problems of corpus composition, the operationalisations involved in making experiments, and in the choice of statistical tools.
+While far from providing definitive solutions, my discussion in Section~\ref{sec:conclusion} will provide possible explanations for the goodness of fit between the corpus data and the experimental data reported in Sections~\ref{sec:corpusstudies} and~\ref{sec:experimental}, much in the spirit of \citet{Dabrowska2014}.
 
-\hl{Turning to the second issue, }
-This entails that a variant of prototype theory with features is assumed \citep{Rosch1978}.
-Prototype theory is well suited for modeling constructional choices, but it is just one of several similarity-based theories of classification, the most prominent other framework being exemplar theory (\citealp{MedinSchaffer1978,Hintzman1986}; see \citealp{StormsEa2000} for a comparison of the theories in different experimental settings).
+I now turn to the second issue, namely which type of cognitive representation research on alternations provides evidence for.
+The typical approach in alternation research is to annotate a large number of corpus sentences with linguistic features and to model the probability of either variant being chosen given these features.
+The idea is that a variant is chosen when the influencing features cumulatively assume \textit{prototypical} values for that variant.
+I other words, a variant of prototype theory with features \citep{Rosch1978} suggests itself, and some researchers such as \cite{Gries2003}, \cite{NessetJanda2010}, or \cite{Schaefer2016c} indeed commit to prototype theory in alternation modelling.
+Under prototype theory, category membership is defined by similarity to an ideal exemplar, which is usually identified with characterising features (see \citealp{Taylor2008} for an overview).
+While prototype theory is well suited for modeling constructional choices, it is just one of at least two major similarity-based theories of classification, the most prominent other framework being exemplar theory (\citealp{MedinSchaffer1978,Hintzman1986}; see \citealp{StormsEa2000} for a comparison of the theories in different experimental settings).
 Prototype theory and exemplar theory model essentially the same types of effects but differ significantly in whether they assume higher-level abstractions in the form of single maximally prototypical exemplars or their features (prototype theory) or assume that categories emerge through the storage of many exemplars and similarity classification on those exemplars (exemplar theory).
@@
 As \cite{Barsalou1990} already showed, however, prototype and exemplar theory model the same types of effects and are informationally equivalent.
-Consequently, experiments which favour one theory over the other use procedural behaviour of subjects, for example the speed of category retrieval, and not mere output data.
+Consequently, experiments which favour one theory over the other use procedural behaviour of subjects in experiments, for example the speed of category retrieval, as opposed to mere output data.
 In very early experiments, \cite{PosnerKeele1968} showed, for example, that highly prototypical unseen exemplars were categorised more easily by subjects compared to less prototypical ones, even if these were included in the learning data.

### 2017-12-26 Latex/sections/1intro.tex
Nearing completion of Sec. 1 rewrite.

@@
 There are two fundamental issues to consider with respect to this tradition as a part of cognitive linguistics in the broad sense.
-First, there is the question of how corpus data relates to experimental findings (which provide more direct evidence of mental representations and processes but are often intrinsically narrower in scope), and how well the two converge or should converge.
-Second, the appropriate modelling of such results in cognitive linguistics is a key issue.
+First, there is the question of whether corpus data do provide any insight into cognitive representations at all.
+This question can and should be answered by testing how well corpus-derived models converge with or diverge from experimental findings, which provide more direct evidence of mental representations and processes but are often intrinsically narrower in scope.
+Second and much related to the first question, the appropriate modelling of such results in cognitive linguistics is a key issue.
 
-As for the first issue, there has always been an interest in correlating probabilistic generalisations extracted from corpus data with results from experimental work (for example, \citealp{ArppeJaervikivi2007,BresnanEa2007,BresnanFord2010,DivjakGries2008,DivjakEa2016,FordBresnan2013}).
+As for the first point, there has for a long time been an interest in correlating probabilistic generalisations extracted from corpus data with results from experimental work (for example, \citealp{ArppeJaervikivi2007,BresnanEa2007,BresnanFord2010,DivjakGries2008,DivjakEa2016,FordBresnan2013}).
 This is often called a \textit{validation} of the corpus-derived findings, but \citet[303]{Divjak2016a} criticises this choice of words ``because it creates the impression that behavioral experimental data is inherently more valuable than textual data,'' citing \cite{TummersEa2005}, who state that a corpus is ``a sample of spontaneous language use that is (generally) realized by native speakers''.
@@
 \footnote{Epistemologically, I think she refers to acts of \textit{induction} rather than \textit{deduction}.}
-It would be highly surprising if this were possible, and the same holds for experimental methods.
+It would be highly surprising if this were possible, and the same holds for experimental methods, albeit to a different degree.
 Nobody assumes that we can inductively infer mental representations from experiments, which -- as opposed to corpus studies -- even allow for direct access to the cognitive agent and offer much better possibilities to control experimental conditions and nuisance variables.
@@
 Then, predictions are derived from this theory \textit{before} the experiment or the corpus study is conducted in order to \textit{test} the theory.
-While the same approach is by and large applicable to corpus data.
+While the same approach is by and large applicable to corpus data, I know discuss some relevant differences.
 
-An important question, however, is whether the usage data found in corpora are truly predictive of speakers' and writers' cognitive representation of language and\slash or of their overall linguistic behaviour.
+As mentioned above, the central question is whether usage data as found in corpora are truly predictive of speakers' and writers' cognitive representation of language and\slash or of their overall linguistic behaviour; and this is where the experimental validation (or, more neutrally \textit{corroboration}) comes into play.
 An overview of this issue was given by \cite{NewmanSorensenduncan2015}, who enumerate a number of studies which show how corpus data and experimental data converge (such as \citealp{DurrantDoherty2010,BresnanEa2007,GriesWulff2005,GriesEa2005}) and a number of studies where the two types of data did not lead to diverging or only partially converging results (such as \citealp{ArppeJaervikivi2007,Dabrowska2014,Mollin2009}).
 When researchers do not reach convergence, they mostly try to explain this by differentiating between the actual cognitive construct and what the pooled usage data as found in corpora represent, for example \citet[411]{Dabrowska2014}, who lists a long number of possible reasons to explain why subjects in her experiment diverged in their word association preferences from collocation measures extracted from corpora.
-In some cases, researchers argue for a more adequate statistical analysis to increase the fit between corpus data and subjects' reactions in experiments, for example \cite{DivjakEa2016}, who show that generalised additive models (GAMs) are better suited than generalised linear models (GLMs) in the analysis of their reading-time experiment based on data extracted from previous corpus studies.
+In some cases, researchers simply argue for a more adequate statistical analysis to increase the fit between corpus data and subjects' reactions in experiments, for example \cite{DivjakEa2016}, who show that generalised additive models (GAMs) are better suited than generalised linear models (GLMs) in the analysis of their reading-time experiment based on data extracted from previous corpus studies.
 Much more optimistically compared to these publications dealing with possible frictions between usage and experimental data, \cite{StefanowitschFlach2016} recently proposed a straightforward positivist perspective of corpora as representing the input of an average adult speaker, thus licensing inferences from corpus data to cognitive representations under a ``corpus-as-input'' view.
@@
 In very early experiments, \cite{PosnerKeele1968} showed, for example, that highly prototypical unseen exemplars were categorised more easily by subjects compared to less prototypical ones, even if these were included in the learning data.
+This was (at the time) taken as evidence that subjects categorise by prototypes.
 Since corpus data only show artefacts of production events and we have no experimental access to the speaker's or writer's performance, one should be sceptical whether corpus analysis alone could ever decide which theory of mental representation is more suitable (see also \citealp[22]{Gries2003} and the \citealp[486--487]{Dabrowska2016} quote above).
@@
 On the other hand, there is evidence that neither theory is fully adequate to model humans' capabilities to form categories.
-For example, \cite{ConawayKurtz2016} show that reference point approaches to category learning such as prototype and exemplar theory fail to explain certain experimental results where subjects learn to generalise beyond the input in a way that cannot be explained by similarity.
+For example, \cite{ConawayKurtz2016} show that both prototype theory and exemplar theory fail to explain certain experimental results where subjects learn to generalise beyond the input in a way that cannot be explained by similarity.
 }
@@
 \cite{VanpaemelStorms2008,LeeVanpaemel2008} proposed the \textit{varying abstraction model} (VAM) which ``attempt[s] to balance economy and informativeness'' \citep[745]{LeeVanpaemel2008}, treating models with full abstraction (prototypes) and no abstraction at all (radical exemplar theory) as special cases of a model which allows for both abstraction and exemplar effects.
-The mixture model of categorisation (MMC) by \cite{Rosseel2002} is a model with abstraction in the form of hierarchical clusters of exemplars, and these clusters of objects are characterised by a probability distribution over their features, and categorising new objects is a process of estimating the probability of this object being from one of the clusters.
+The mixture model of categorisation (MMC) by \cite{Rosseel2002} is a model with abstraction in the form of hierarchical clusters of exemplars, and these clusters of objects are characterised by a probability distribution over their features, and categorising new objects is a process of estimating the probability of this object belonging to one of the clusters.
 \cite{GriffithsEa2009} go further and present a computational model based on the hierarchical Dirirchlet process which is able to chose the appropriate complexity of representation for a given category.
-However, despite these (and more) attempts to , \cite[183--184]{Vanpaemel2016} describes the state of affairs between adherents of neo-prototype theory (such as \citealp{MindaSmith2001,MindaSmith2002}) and exemplar theory as a stalemate, and suggests yet another approach to solve the divide.
+However, despite these (and more) attempts to reconcile or unite the two approaches while developing spelled-out mathematical models, \cite[183--184]{Vanpaemel2016} describes the state of affairs between adherents of neo-prototype theory (such as \citealp{MindaSmith2001,MindaSmith2002}) and exemplar theory as a stalemate.
 
@@
 The clusters are interpreted as intermediate-level exemplar-derived abstractions of typical contexts for these high-probability verbs (typically more than one cluster for each verb; \citealp[255--256]{DivjakArppe2013}).
-This is pioneering work, but the crucial difference between data-driven corpus-based analyses and experiments in cognitive science (they use \citealp{VerbeemenEa2007} as their reference) is that cognitive experimental research is based on experiments where subjects produce category assignments, and in corpus studies, the categories and category membership is determined purely from the data.
+The crucial difference between such data-driven corpus-based analyses and experiments in cognitive science (\citealp{DivjakArppe2013
+} use \citealp{VerbeemenEa2007} as their reference) is that cognitive experimental research is based on experiments where subjects produce actual category assignments or similarity judgements, and in corpus studies, the categories and category membership is determined purely from existing data.
+The experimental approach with reduced and\slash or artificial stimuli makes it much easier to examine very specific effects in the behaviour of the subjects.
+While I am convinced that the results presented in \citet{DivjakArppe2013} are valid and important (especially given the previous and subsequent research the authors have conducted on the data, including experimental work as in \citealp{DivjakEa2016}), any data set can be clustered to yield a certain number of clusters.
+Thus, the study does not in any way ensure that the clusters emerging from the data in any way correspond to any speaker's cognitive representation.%
+\footnote{See, again, \citet[486--487]{Dabrowska2016} and \citet[22]{Gries2003}.}
 
-One thing that should be kept in mind is that most of the research on categorisation in cognitive science uses experiments with highly simplified stimuli and very simple tasks.
-It is remarkable in this context that \cite{VoorspoelsEa2011} consider the experimental task reported by them -- assigning typicality scores to nouns from the domains of \textit{animals} and \textit{artefacts} to categories like \textit{bird}, \textit{fish}, \textit{clothing}, or \textit{tools} -- a study of ``superordinate natural language categories, whereas most evidence supporting exemplar representations has been found in artificial categories of a more subordinate level'' \citep[1013]{VoorspoelsEa2011}.
-Linguists (not just of the cognitive variety) are usually interested in much more complex high-level categories and use large and complex feature sets, especially in (morpho-)syntax.%
-\footnote{Notice that recently, approaches have emerged which solve at least some problems by abandoning linguistic high-level features altogether \citep{BaayenEa2016,RamscarPort2016}.}
-
-Most likely a consequence of linguists' interest in  is that formal models of categorisation in cognitive linguistics have not been spelled out at the same level of mathematical detail as in cognitive science.%
-\footnote{Again using \cite{DivjakArppe2013} as an example, it is obvious that they use advanced statistics in the form of polytomous regression and hierarchical agglomerative clustering.
-However, models like the one proposed in \cite{GriffithsEa2009} are fully spelled out probability models of the behaviour of cognitive agents.}
-However, the fact that linguists deal with real data should be seen as an advantage in that it greatly improves the \textit{external validity} of studies, \ie the generalisability of the findings.
-Typical artificial tasks in cognitive science have been criticised exactly for the lack of external validity (\eg \citealp{Murphy2003}).
+However, the trade-off with experiments with highly simplified stimuli and very simple tasks is their lack of external validity (\ie their generalisability) and their high dependence on problematic operationalisations of constructs, control of confounding factors, etc.
+(in other words, dependence on construct and internal validity).
+Tasks in cognitive science have been criticised exactly for their lack of external validity, for example by \citet{Murphy2003}.
+From a linguistic perspective, it is remarkable in this context that \cite{VoorspoelsEa2011} consider their experimental task -- which is the assignment of typicality scores to nouns from the domains of \textit{animals} and \textit{artefacts} to categories like \textit{bird}, \textit{fish}, \textit{clothing}, or \textit{tools} -- a study of ``superordinate natural language categories, whereas most evidence supporting exemplar representations has been found in artificial categories of a more subordinate level'' \citep[1013]{VoorspoelsEa2011}.
+Corpus linguists interested in probabilistic alternation modelling deal with much more complex high-level categories and use large and complex feature sets, especially in (morpho-)syntax.%
+\footnote{Notice, however, that recently, approaches have emerged which solve at least some problems by abandoning linguistic high-level features altogether \citep{BaayenEa2016,RamscarPort2016}.
+Clearly, they have not (or at least not yet) reached mainstream popularity.}
+It is thus an advantage of much linguistic work on categorisation that it deals with complex and realistically produced data, because this greatly improves the \textit{external validity} of studies, \ie the generalisability of the findings.
 

### 2017-12-27 Latex/article.tex
Minor progess w/ revision.

@@
   \affil[1]{Freie Universität Berlin}
-  \title{Abstractions and items: the measure noun phrase alternation in German}
-  \runningtitle{Abstraction and items: the measure noun phrase alternation German}
+  \title{Abstractions and exemplars: the measure noun phrase alternation in German}
+  \runningtitle{Abstraction and exemplars: the measure noun phrase alternation German}
   %\subtitle{...}

### 2017-12-27 Latex/sections/1intro.tex
Minor progess w/ revision.

@@
-\section{Prototypes, and exemplars, and corpora}
+\section{Prototypes, exemplars, and corpora in cognitive linguistics}
 \label{sec:cogocl}
@@
 By \textit{alternation} I refer to a situation where two or more forms or constructions are available with no clear difference in acceptability, function, or meaning.
-The study of constructional alternations has a long history in cognitively oriented corpus linguistics (for example, \citealp{BresnanEa2007,BresnanHay2010,BresnanFord2010,DivjakArppe2013,Gries2015,NessetJanda2010,Wulff2003}).
+The study of lexical and constructional alternations has a long history in cognitively oriented corpus linguistics (for example, \citealp{BresnanEa2007,BresnanHay2010,BresnanFord2010,DivjakArppe2013,Gries2015,NessetJanda2010,Wulff2003}).
 This area of research is based on the assumption that language is a probabilistic phenomenon \citep{Bresnan2007} where alternants are chosen neither deterministically nor fully at random.
@@
 This is often called a \textit{validation} of the corpus-derived findings, but \citet[303]{Divjak2016a} criticises this choice of words ``because it creates the impression that behavioral experimental data is inherently more valuable than textual data,'' citing \cite{TummersEa2005}, who state that a corpus is ``a sample of spontaneous language use that is (generally) realized by native speakers''.
+See also \citet{Newman2011} for a very positive view of corpora as a source of data in cognitive linguistics in their own right.
 However, as \citet[486--487]{Dabrowska2016} convincingly argues, this does not imply that we can in some way ``deduce mental representations from patterns of use,'' \ie\ from corpus data.%
@@
 Then, predictions are derived from this theory \textit{before} the experiment or the corpus study is conducted in order to \textit{test} the theory.
-While the same approach is by and large applicable to corpus data, I know discuss some relevant differences.
+While the same approach is by and large applicable to corpus data, I now discuss some relevant differences.
 
 As mentioned above, the central question is whether usage data as found in corpora are truly predictive of speakers' and writers' cognitive representation of language and\slash or of their overall linguistic behaviour; and this is where the experimental validation (or, more neutrally \textit{corroboration}) comes into play.
-An overview of this issue was given by \cite{NewmanSorensenduncan2015}, who enumerate a number of studies which show how corpus data and experimental data converge (such as \citealp{DurrantDoherty2010,BresnanEa2007,GriesWulff2005,GriesEa2005}) and a number of studies where the two types of data did not lead to diverging or only partially converging results (such as \citealp{ArppeJaervikivi2007,Dabrowska2014,Mollin2009}).
-When researchers do not reach convergence, they mostly try to explain this by differentiating between the actual cognitive construct and what the pooled usage data as found in corpora represent, for example \citet[411]{Dabrowska2014}, who lists a long number of possible reasons to explain why subjects in her experiment diverged in their word association preferences from collocation measures extracted from corpora.
+An overview of this issue was given by \cite{NewmanSorensenduncan2015}, who enumerate a number of studies which show how corpus data and experimental data converge (such as \citealp{BresnanEa2007,DurrantDoherty2010,GriesWulff2005,GriesEa2005}) and a number of studies where the two types of data did not lead to diverging or only partially converging results (such as \citealp{ArppeJaervikivi2007,Dabrowska2014,Mollin2009}).
+When researchers do not achieve convergence, they mostly try to explain this by differentiating between the actual cognitive construct and what the pooled usage data as found in corpora represent, for example \citet[411]{Dabrowska2014}, who lists a long number of possible reasons to explain why subjects in her experiment diverged in their word association preferences from collocation measures extracted from corpora.
 In some cases, researchers simply argue for a more adequate statistical analysis to increase the fit between corpus data and subjects' reactions in experiments, for example \cite{DivjakEa2016}, who show that generalised additive models (GAMs) are better suited than generalised linear models (GLMs) in the analysis of their reading-time experiment based on data extracted from previous corpus studies.
-Much more optimistically compared to these publications dealing with possible frictions between usage and experimental data, \cite{StefanowitschFlach2016} recently proposed a straightforward positivist perspective of corpora as representing the input of an average adult speaker, thus licensing inferences from corpus data to cognitive representations under a ``corpus-as-input'' view.
+Much more optimistically compared to these publications dealing with possible frictions between usage and experimental data, \cite{StefanowitschFlach2016} recently proposed a straightforward positivist perspective of corpora as representing the input of an average adult speaker, thus licensing inferences from corpus data to cognitive representations learned from such data under a ``corpus-as-input'' view.
 They state that ``in this wider context, large, register-mixed corpora such as the British National Corpus [\ldots] may not be perfect models of the linguistic experience of adult speakers, but they are reasonably close to the input of an idealized average member of the relevant speech community'' \citep[104]{StefanowitschFlach2016}.%
@@
 See \citet{Biber1993}, \citet{MceneryEa2006}, \citet{Leech2007}, \citet{Hunston2008} for overviews from different perspectives.}
-Clearly, no clear picture has yet emerged, which is not surprising given the vast number of cognitive constructs assumed at diverse levels (such as the level of words in collocation research and much more complex constructional levels in alternation research involving lexical, syntagmatic, and contextual factors), the problems of corpus composition, the operationalisations involved in making experiments, and in the choice of statistical tools.
+In essence, this would entail that any generalisation extracted from the BNC can be attributed with some kind of mental reality.
+
+Obviously, no clear picture has emerged yet, which is not surprising given the vast number of cognitive constructs assumed at diverse levels (such as the level of words in collocation research and much more complex constructional levels in alternation research involving lexical, syntagmatic, and contextual factors), the problems of corpus composition, the operationalisations involved in making experiments, and in the choice of statistical tools.
 While far from providing definitive solutions, my discussion in Section~\ref{sec:conclusion} will provide possible explanations for the goodness of fit between the corpus data and the experimental data reported in Sections~\ref{sec:corpusstudies} and~\ref{sec:experimental}, much in the spirit of \citet{Dabrowska2014}.
@@
 I now turn to the second issue, namely which type of cognitive representation research on alternations provides evidence for.
-The typical approach in alternation research is to annotate a large number of corpus sentences with linguistic features and to model the probability of either variant being chosen given these features.
+The typical approach in alternation research is to annotate a large number of corpus sentences with linguistic features and to model the probability of the variants being chosen given these features.
 The idea is that a variant is chosen when the influencing features cumulatively assume \textit{prototypical} values for that variant.
@@
 Under prototype theory, category membership is defined by similarity to an ideal exemplar, which is usually identified with characterising features (see \citealp{Taylor2008} for an overview).
-While prototype theory is well suited for modeling constructional choices, it is just one of at least two major similarity-based theories of classification, the most prominent other framework being exemplar theory (\citealp{MedinSchaffer1978,Hintzman1986}; see \citealp{StormsEa2000} for a comparison of the theories in different experimental settings).
+While prototype theory is well suited for modeling constructional choices, it is just one of at least two major similarity-based theories of classification, the most prominent other framework being exemplar theory (\citealp{MedinSchaffer1978,Hintzman1986}).
 Prototype theory and exemplar theory model essentially the same types of effects but differ significantly in whether they assume higher-level abstractions in the form of single maximally prototypical exemplars or their features (prototype theory) or assume that categories emerge through the storage of many exemplars and similarity classification on those exemplars (exemplar theory).
-\hl{Parallel to Langacker here}
+Similarity-based categorisation is seen as the central and conceptually sufficient approach to linguistic categorisation by many cognitive linguists; see for example \cite{Taylor2003} for a comprehensive treatment in terms of prototype theory.
+In the influential framework of cognitive grammar \citep{Langacker1987}, prototypes (which, as should be remembered, represent abstractions already in cognitive science) are literally taken as prototypical exemplars, and there is an additional level of fully discrete abstractions in the form of schemas.
+Schemas are characterised by the properties common to all members of the category, whereas a prototypical category member might have very specific additional properties not at all shared by all or even most members \citep[371-375]{Langacker1987}.
+The prototype can serve as a reference point when classifying new objects which do not share all properties of the schema, but this would (if repeated) lead to the creation of an even more abstract (hierarchically higher and less specific) schema which describes the new member and the ones belonging to the previous schema.
+As pointed out by \citet[136--137]{Langacker1987}, schemas and prototypes thus fulfil different roles and can be assumed to co-exist.
+A strict exemplar view of language is incompatible, as far as I can see, with Langacker's view of schemas, but any theory of categorisation that allows for at least some kind of abstraction, is not in principled contradiction with it.%
+\footnote{See especially theories with varying intermediate levels of abstraction described below, which seem compatible with Langacker's formation of schema hierarchies through classification by prototype similarity.}
+In the remainder of this paper, I do not use schemas in my descriptions of the relevant categories, mostly because the aspect of similarity and fuzzy classification is central to my point, and a formulation in terms of schemas would bring about an unnecessarily high degree of abstraction (see \citealp[70--71]{Taylor2003} for a parallel argument).
+However, I fully describe the abstract features of the prototypes, of course, while accounting for item-specific and exemplar effects.
 
-As \cite{Barsalou1990} already showed, however, prototype and exemplar theory model the same types of effects and are informationally equivalent.
+Turning back to the prototype vs.\ schema debate, \cite{Barsalou1990} already showed that prototype and exemplar theory model the same types of effects and are informationally equivalent.
 Consequently, experiments which favour one theory over the other use procedural behaviour of subjects in experiments, for example the speed of category retrieval, as opposed to mere output data.
-In very early experiments, \cite{PosnerKeele1968} showed, for example, that highly prototypical unseen exemplars were categorised more easily by subjects compared to less prototypical ones, even if these were included in the learning data.
-This was (at the time) taken as evidence that subjects categorise by prototypes.
-Since corpus data only show artefacts of production events and we have no experimental access to the speaker's or writer's performance, one should be sceptical whether corpus analysis alone could ever decide which theory of mental representation is more suitable (see also \citealp[22]{Gries2003} and the \citealp[486--487]{Dabrowska2016} quote above).
+In very early experiments, \cite{PosnerKeele1968} showed, for example, that highly prototypical unseen exemplars were categorised more easily by subjects compared to less prototypical ones which had been included in the learning data.
+This was (at the time) taken as evidence that subjects categorise by prototypes.%
+\footnote{See \citet{StormsEa2000} for a comparison of the theories in different experimental settings.}
+Since corpus data only show artefacts of production events and we have no experimental access to the speaker's or writer's performance and their actual similarity judgements, one should be sceptical whether corpus analysis alone could ever decide which theory of mental representation is more suitable (see also \citealp[22]{Gries2003} and the \citealp[486--487]{Dabrowska2016} quote above).
 
@@
 \footnote{The hard empirical evidence in favour of exemplar models is substantial.
-  For example, in \cite{HahnEa2010}, the authors show that subjects even use exemplar similarity over abstract knowledge even when they are given very simple explicit rules.
+  For example, in \cite{HahnEa2010}, the authors show that subjects even use exemplar similarity over abstract knowledge even when they are given very simple explicit rules to be learned.
   This is highly relevant because most other studies focus on the learning of implicit rule-based knowledge, which involves many auxiliary assumptions in actual experiments \citep[2]{HahnEa2010}.
@@
 Still, various attempts have been made over the past decades to settle the dispute between abstraction-based models (models with rules or prototypes) and exemplar models or to find models which unite the two extremes.
-\cite{VanpaemelStorms2008,LeeVanpaemel2008} proposed the \textit{varying abstraction model} (VAM) which ``attempt[s] to balance economy and informativeness'' \citep[745]{LeeVanpaemel2008}, treating models with full abstraction (prototypes) and no abstraction at all (radical exemplar theory) as special cases of a model which allows for both abstraction and exemplar effects.
+\cite{VanpaemelStorms2008} and \citet{LeeVanpaemel2008} proposed the \textit{varying abstraction model} (VAM) which ``attempt[s] to balance economy and informativeness'' \citep[745]{LeeVanpaemel2008}, treating models with full abstraction (prototypes) and no abstraction at all (radical exemplar theory) as special cases of a model which allows for both abstraction and exemplar effects.
 The mixture model of categorisation (MMC) by \cite{Rosseel2002} is a model with abstraction in the form of hierarchical clusters of exemplars, and these clusters of objects are characterised by a probability distribution over their features, and categorising new objects is a process of estimating the probability of this object belonging to one of the clusters.
-\cite{GriffithsEa2009} go further and present a computational model based on the hierarchical Dirirchlet process which is able to chose the appropriate complexity of representation for a given category.
+\cite{GriffithsEa2009} go further and present a computational model which is able to chose the appropriate complexity of representation for a given category.
 However, despite these (and more) attempts to reconcile or unite the two approaches while developing spelled-out mathematical models, \cite[183--184]{Vanpaemel2016} describes the state of affairs between adherents of neo-prototype theory (such as \citealp{MindaSmith2001,MindaSmith2002}) and exemplar theory as a stalemate.
@@
 
-In cognitive linguistics, \cite{DivjakArppe2013} is a very rare example of a paper where such issues are taken up.
+In cognitive linguistics, \cite{DivjakArppe2013} is a very rare example of a paper where such issues are taken up with reference to the current research in cognitive science.
 Their corpus-based approach shows ``one way of systematically analyzing usage data as contained in corpora to yield a scheme, compatible with usage-based theories of language, by which the assumptions of both the prototype and exemplar theories can be operationalized'' \citep[267]{DivjakArppe2013}.
 Their approach to implementing a varying abstraction model \citep[254--260]{DivjakArppe2013} is based on hierarchical clustering of annotated properties of sentences.
-They hierarchically cluster sentences containing Russian verbs of trying.
-Then, they single out one sentence from each cluster which scores the highest probability for any of the six \textit{try} verbs according to a polytomous regression model estimated on the same data.
+They cluster sentences containing Russian verbs of trying.
+Then, they single out the one sentence from each cluster which scores the highest probability for any of the six \textit{try} verbs according to a polytomous regression model estimated on the same data.
 The clusters are interpreted as intermediate-level exemplar-derived abstractions of typical contexts for these high-probability verbs (typically more than one cluster for each verb; \citealp[255--256]{DivjakArppe2013}).
-The crucial difference between such data-driven corpus-based analyses and experiments in cognitive science (\citealp{DivjakArppe2013
-} use \citealp{VerbeemenEa2007} as their reference) is that cognitive experimental research is based on experiments where subjects produce actual category assignments or similarity judgements, and in corpus studies, the categories and category membership is determined purely from existing data.
+The crucial difference between such data-driven corpus-based analyses and experiments in cognitive science (\citealp{DivjakArppe2013} use \citealp{VerbeemenEa2007} as their reference) is that cognitive research is based on experiments where subjects produce actual category assignments or similarity judgements, and in corpus studies, the categories and category membership are determined purely from existing data.
 The experimental approach with reduced and\slash or artificial stimuli makes it much easier to examine very specific effects in the behaviour of the subjects.
-While I am convinced that the results presented in \citet{DivjakArppe2013} are valid and important (especially given the previous and subsequent research the authors have conducted on the data, including experimental work as in \citealp{DivjakEa2016}), any data set can be clustered to yield a certain number of clusters.
-Thus, the study does not in any way ensure that the clusters emerging from the data in any way correspond to any speaker's cognitive representation.%
+While I am convinced that the results presented in \citet{DivjakArppe2013} are valid and important (especially given the previous and subsequent research the authors have conducted on the data, including experimental work presented in \citealp{DivjakEa2016}), any data set can be clustered to yield a certain number of clusters.
+Thus, the study does not in any way ensure that the clusters emerging from the data correspond to any speaker's cognitive representation.%
 \footnote{See, again, \citet[486--487]{Dabrowska2016} and \citet[22]{Gries2003}.}
 
-However, the trade-off with experiments with highly simplified stimuli and very simple tasks is their lack of external validity (\ie their generalisability) and their high dependence on problematic operationalisations of constructs, control of confounding factors, etc.
-(in other words, dependence on construct and internal validity).
+On the other hand, the trade-off one has to accept when doing experiments with highly simplified stimuli and very simple tasks is their lack of \textit{external validity} (\ie their generalisability) and their high dependence on potentially problematic operationalisations of constructs, control of confounding factors in the face of a limited number of available subjects, etc.
+(in other words, critical dependence on \textit{construct validity} and \textit{internal validity}).%
+\footnote{An accessible overview of the different types of validity can be found in Chapter~1 of \citet{MaxwellDelaney2004}.}
 Tasks in cognitive science have been criticised exactly for their lack of external validity, for example by \citet{Murphy2003}.
@@
 \footnote{Notice, however, that recently, approaches have emerged which solve at least some problems by abandoning linguistic high-level features altogether \citep{BaayenEa2016,RamscarPort2016}.
-Clearly, they have not (or at least not yet) reached mainstream popularity.}
-It is thus an advantage of much linguistic work on categorisation that it deals with complex and realistically produced data, because this greatly improves the \textit{external validity} of studies, \ie the generalisability of the findings.
+Clearly, they have not (or at least not yet) reached mainstream popularity, and it remains to be seen how well they perform on a broader range of questions.}
+It is thus an advantage of much linguistic work on categorisation that it deals with complex and realistically produced data, because this greatly improves the external validity of studies.
+An ideal contribution to the research on (levels of) category abstraction in the human mind by cognitive corpus linguists is thus if we provide analyses which have great external validity and complexity while carefully making sure whether and to which degree these finding correlate with reactions from cognitive agents under more controlled experimental conditions.
 
-I propose that the primary focus should be on determining which factors influence choices made by speakers.
-As \cite[22]{Gries2003} put it with reference to classic prototype theory, ``even if the form of analysis does not translate into statements on mental representations, the high predictive power [...] shows that the cognitive factors underlying the choice of construction have been identified properly and weighted in accordance with their importance for actual usage.''
-A major advancement since this early time has been the focus on which effects are at least more plausibly modelled as prototype\slash abstraction and exemplar effects, and how observed data fits the two approaches (see, for example, \citealp{DivjakArppe2013}).
-This paper contributes this discussion.
-
-% The present study is conducted within the general paradigm of cognitive corpus linguistics and includes a comparison of the corpus findings with results from two experiments.
-% I assume a model of similarity-based classification in the form of Prototype Theory for modeling alternation.
-% Protoype Theory has been used in alternation research in cognitively oriented corpus linguistics (see \citealp{DivjakArppe2013}; \citealp{Gries2003}).
-% It is assumed that the variables annotated in a corpus study on an alternation phenomenon define prototypes for the alternants.
-% The similarity of a given feature vector (in a concrete sentence where one of the alternants is used) influences to a large extent which alternant is chosen by speakers.
-% This entails that a variant of Prototype Theory with features is assumed \citep{Rosch1978}, instead of the monolithic prototypes of earlier versions.
-% Prototype Theory is well suited for modeling constructional choices but it is just one of several similarity-based theories of classification, the most prominent other framework being Exemplar Theory (\citealp{MedinSchaffer1978,Hintzman1986}; see \citealp{StormsEa2000} for a comparison of the theories in experimental settings).
-% Prototype Theory and Exemplar Theory model essentially the same types of effects and differ mainly in whether they assume higher-level abstractions as part of the cognitive representation (Prototype Theory) or try to do without them (Exemplar Theory).
-% I am sceptical that corpus analysis alone could ever decide which theory is more suitable, given that substantial doubt has been voiced whether even experimental methods are ultimately able to do so \citep{Barsalou1990}.%
-% \footnote{In \cite{DivjakArppe2013}, it was shown using corpus data how both models form a converging picture.}
-% Prototype Theory (with feature abstractions) is preferred here simply because it fits the established alternation modeling paradigm, which relies on features being weighted in statistical models.
-% In Section~\ref{sec:analyses}, a prototype-theoretical parlance is therefore adopted.
-% 
-% In the remainder of the paper, I introduce the alternation phenomenon (Section~\ref{sec:descriptive}) and suggest a theory-driven set of factors influencing the alternation (Section~\ref{sec:analyses}).
-% Then the corpus study is presented, including an appropriate statistical analysis (Section~\ref{sec:corpusstudies}).
-% Two experiments which corroborate the corpus-based findings are then reported in Section~\ref{sec:experimental} before I sum up the paper in Section~\ref{sec:conclusion}.
+This paper contributes to this endeavour in many ways.
+After a thorough description of the alternation under examination, I discuss potential influencing factors comprising high-level schematic semantic generalisations as well as exemplar-similarity and item-specific effects in Section~\ref{sec:germanmeasurenps}.
+I will argue that
+there are lemma-specific and exemplar effects but also generalisations at the level of the construction as well as generalisations overlaying the lemma-specific effects, leading to a complex hierarchical structure.
+In Section~\ref{sec:corpusstudies}, I present a corpus study and report a true multilevel generalised linear model with the appropriate hierarchical structure for the hypothesised effects.
+In Section~\ref{sec:experimental}, I test the predictions of the corpus-based model in two experimental paradigms (forced-choice and self-paced reading), showing that they indeed converge, albeit with intermediate effect strength.
+In Section~\ref{sec:conclusion}, I interpret the findings in the light of the issues of cognitive representations and corpus data and the convergence of usage data and experimental data.

### 2017-12-27 Latex/sections/2descriptive.tex
More reviewer-induced changes to Sec. 1.
Minor progess w/ revision.

@@
-\section{Case assignment in German measure NPs}
+\section{Case in German measure NPs}
 \label{sec:germanmeasurenps}
 
-\subsection{Two stable cases and a case alternation}
+\subsection{Alternating and non-alternating measure NP constructions}
 \label{sec:descriptive}
@@
 
-It is difficult to determine whether the bare kind noun in the narrow apposition construction as in (\ref{ex:intro:narrowapposition1}) bears no case at all, a generic case, or agrees in case with the measure noun.%
-\footnote{I would like to thank one of the reviewers for pointing this out.}
+It is difficult to determine whether the bare kind noun in the narrow apposition construction as in (\ref{ex:intro:narrowapposition1}) bears no case at all, a generic case, or agrees in case with the measure noun.
 When there is an adjective as in (\ref{ex:intro:alternation2}), the embedded kind NP clearly agrees in case, but due to the overall absence of markers of case in the singular, bare nouns mostly show no indication of their case.
@@
 However, a few of them can be coerced into a mass noun, such as \textit{Hase} `rabbit' (meaning `rabbit meat').
-It then appears as if the uninflected form is preferred, but the inflected form is not excluded.
+It then appears as if the uninflected form is preferred, but the inflected form is not excluded, at least for some speakers.
 In (\ref{ex:macabre1}), the clearly acceptable form \textit{Hase} can only be a nominative singular or caseless.
@@
 Even a full unacceptability of (\ref{ex:macabre2}) would not be conclusive, however, as a possible aversion of speakers towards the case-marked form might be due to the fact that it is at least potentially also a genitive, in which case the constraint against bare genitive nouns would apply.
-With plural kind nouns, the obligatory marking of the dative plural with \textit{-en} might provide some clues.
+With plural kind nouns, the obligatory marking of all dative plurals (except those where suffixing \textit{-en} is phonotactically impossible) with \textit{-en} might provide some clues.
 However, plural kind nouns do not behave like singular ones in measure phrases anyway, as will be argued in Section~\ref{sec:neighbouringcases}.
@@
 Descriptive grammars seem to favour an analysis in terms of caselessness (for example, \citealp[1981]{ZifonunEa1997c}).
-The hazy picture of the case of bare kind NPs is most likely due to the fact that case is so rarely marked on German nouns, where case is marked mostly on determiners and to some degree adjectives.
-The uncertainty in the few cases where case can be marked (weak nouns in the singular and dative plurals) would thus be a direct consequence of the fact that the construction is not very specific with respect to the case of the kind noun.
+The hazy picture of the case of bare kind NPs is most likely due to the fact that case is so sparsely marked on nouns in contemporary German, where case is marked mostly on determiners and to some degree adjectives.
+The uncertainty in the few cases where case can be marked (weak nouns in the singular and dative plurals as described above) would thus be a direct consequence of the fact that the construction is not very specific with respect to the case of the kind noun.
 
@@
 
-It was shown empirically in \cite{Hentschel1993} that such sentences are de facto not acceptable.
+It was shown empirically in \cite{Hentschel1993} that such sentences are de facto not acceptable, and the results in \citet{Zimmer2015} are in line with her findings.
 Also, in my corpus sample, they simply did not occur.
@@
 
-\subsection{Factors controlling the alternation}
+\subsection{Prototype and exemplar effects}
 \label{sec:analyses}

### 2017-12-27 Latex/sections/3corpus.tex
More reviewer-induced changes to Sec. 1.
Minor progess w/ revision.

@@
 
-\subsection{On statistical analysis}
-\label{sec:rightstatistics}
-
-In this section, I justify my choice of statistical models used in Section~\ref{sec:corpushierarchicalmodel}.
-Readers might think that the method used here for modeling grammatical alternations -- namely (Hierarchical) Logistic Regression\slash Generalised Linear (Mixed) Models or GL(M)Ms -- does not require much justification.
-After all, GLMMs have been established as the major tool in the analysis of alternation phenomena.
-All studies mentioned at the outset of Section~\ref{sec:cogocl} use some form of regression\slash GLMM.
-Over the past few years, however, modified or alternative methods have been proposed.
-While it is impossible to discuss all of these methods here, I just make a few remarks on Bayesian estimation (see \citealp{GelmanEa2014}) as it was proposed in \cite{Levshina2016} and \cite{Divjak2016a}, for example.
-Conceptually, I see three points of discussion that should be kept apart.
-First, Bayesian methods are sometimes touted as superior tools for scientific inference compared to frequentist methods.
-Second, it has been proposed that the Bayesian interpretation of probability is more cognitively adequate for the modeling of linguistic data \cite[301--302]{Divjak2016a}.
-Third and very specific to this paper, given established methods in the modeling of alternation and variation, it has to be decided whether so-called Bayesian methods lead to substantially different results.
-
-% As for the first point, this is not quite the place to discuss it fully.
-% The basic distinction is a philosophical one and related to the concepts of \textit{direct} and \textit{inverse probability} (\eg\ \citealp{Senn2011}).
-% Frequentists assume that models and parameters are fixed, for example a model specifying that a coin is fair.
-% We can then calculate for observed data (for example a measurement of 3 heads in an experiment with 10 tosses) how often such a result or a more extreme result would occur if the model were true and we repeated the experiment arbitrarily often.
-% This is essentially the frequentist notion of direct probability, \ie\ long-run frequencies under replication.
-% Standard tests in the Fisher and Neyman-Pearson traditions as well as Neyman confidence intervals are based on this concept of probability.
-% Bayesian approaches (in the now common interpretation), on the other hand, condition on the particular data and quantify inductively the probability of model parameters given the data.
-% The parameters are thus not fixed, and the probability is usually equated with researchers' posterior beliefs about model parameters.
-% There is actually a debate among Bayesians about the proper interpretation of Bayesian methods, and whether a notion of hypothesis testing is compatible (or even already contained) in the Bayesian approach.
-% In \citet[10]{GelmanShalizi2013}, the authors -- prominent Bayesians themselves -- acknowledge that a theory of statistical testing is a desideratum, and they state about the standard inductive interpretation of Bayesianism that ``most of this received view of Bayesian inference is wrong,'' and they develop a Bayesian notion of p-values (see also \citealp{Mayo2013}, for a frequentist reply; also \citealp{Senn2011} on different strands of Bayesianism and their stance on inductive vs.\ deductive reasoning, and \citealp{Mayo2011}, for a critical reply to \citealp{Senn2011}).
-% Clearly, in such quarrels between and among camps of philosophers of science and statisticians, it is difficult for mere practitioners to take sides.
+% \subsection{On statistical analysis}
+% \label{sec:rightstatistics}
 % 
-% These quarrels relate to the second point, however.
-% \citet[301--302]{Divjak2016a} speaks favourably of Bayesian methods because the Bayesian concept of probability is more adequate for cognitive modeling compared to the frequentist one.
-% Her argument is part of a larger body of literature asking for cognitively plausible modeling techniques, for example Naive Discriminative Learning (NDL; \citealp{Baayen2011,BaayenEa2013,MilinEa2016,TheijssenEa2013}).
-% On p.\ 303 of \cite{Divjak2016a}, the author goes on to explicitly mention NDL as well.
-% Yet, neither frequentist nor Bayesian methods were conceived as cognitive models, but as systems of inference for scientists (see above, and see also \citealp[302]{Divjak2016a}).
-% The fundamental question that lurks behind such arguments is how we interpret our statistical models (estimated on corpus data).
-% Are they inductive models of cognitive representations that also human learners would infer from being exposed to the corpus data?%
-% \footnote{In which case we are doing ``data science in language research'' in the words of \citealp{MilinEa2016}.
-% I see this as standing in contradiction to the view advocated in \cite{Dabrowska2016} as cited above.}
-% Or are they tests of theories that are pre-specified and merely tested for predictive accuracy on linguistic output data contained in corpora?
-% In the former case, we adopt a strong \textit{corpus as input} hypothesis and should definitely resort to methods like NDL.
-% However, this would most likely require us to toss most previous work done in (cognitive) linguistics into the bin and to abandon all high-level generalisations that most existing studies have been based upon.
-% In fact, arguments to this extreme effect and against using high-level generalisation have been made, for example in \cite{BaayenEa2016}, \citet[299--300]{Divjak2016a}, \cite{RamscarPort2016}, \cite{TheijssenEa2013}.
-% In the latter and less extreme case, the cognitive commitment does, however, not necessarily extend to the statistical methods used.
-% These methods then do not need to be any more cognitively plausible than an ANOVA used to analyse the results from an experiment.
-% I view my own work (see Sections~\ref{sec:germanmeasurenps}--\ref{sec:experimental}) in the tradition of testing theories (which embrace high-level generalisations), and I agree to provisionally use Likelihood methods (see below).
-% I remain fully agnostic with respect to the question, which approach is \textit{right}.
-% The best strategy for cognitive linguistics as a field might be to cultivate many methods while making sure that each method is applied carefully and competently.
-
-The first and second point cannot be discussed in detail here.%
-\footnote{However, there is a number of critical papers by statisticians (even prominent Bayesians) and philosophers of science in which the extreme credibility that has recently been assigned to (subjective and automatic) Bayesian approaches (and especially the standard inductivist interpretation of Bayesianism) is questioned (for example, \citealp{GelmanShalizi2013,Mayo2011,Senn2011}).}
-With regard to the third point, \citet[251--252]{Levshina2016} argues for Bayesian estimation in mixed regression settings.
-First, she claims that ``while frequentist statistics only allows one to test whether the null hypothesis can be rejected, Bayesian statistics enables one both to test the null hypothesis and to estimate the probability of specific parameter values given the data.''
-This does not do justice to frequentist methods in that a strong focus on the rejection of the null hypothesis is characteristic only of Fisher's approach.
-In the Neyman-Pearson approach, results ideally \textit{favour} the main hypothesis vis-à-vis the alternative hypothesis (cf.\ \citealp{Lehmann1993,Lehmann2011,Perezgonzalez2015}).
-Also, especially Neyman-style frequentism has well-known extensions to estimation, for example in the form of confidence intervals (see \citealp{GreenlandEa2016}, esp.\ p.\ 340).
-She then explains that a ``distinctive feature of Bayesian statistics is the use of so-called priors'' and that ``posterior probabilities depend on both the prior beliefs and the data, whereas the results of a frequentist model depend only on the data'' \cite[252]{Levshina2016}.
-Remarkably, given this statement, she does \textit{not} use informative priors and in her footnote 8 \cite[252]{Levshina2016} admits that priors were probed using trial and error.
-So, the proclaimed major advantage of Bayesian modeling was apparently not taken advantage of.%
-\footnote{In the words of \cite{Senn2011}: ``You may believe you are a Bayesian but you are probably wrong.''
-\citet[347--348]{GelmanHill2006} ``view any noninformative prior distribution as inherently provisional'' and give recommendations how to proceed once posteriors have been obtained from noninformative priors.}
-Now, Maximum Likelihood Estimation (MLE) -- the traditional method which could have been used instead -- is not exactly \textit{frequentist} in the sense of Neyman-Pearson testing theory.
-MLE, like inductive Bayesianism, conditions on the particular data inasmuch as it searches for the most likely set of parameters given the data.
-What is more, Bayesian estimators are in fact based on the Likelihood and merely multiply it by the prior \citep[6--8]{GelmanEa2014}.
-If the prior is flat, results converge (see also \citealp[347]{GelmanHill2006}).
-The same is true if the sample size is large compared to the number of parameters, at least for finite-dimensional parameter models \citep[1119--1120]{Freedman1999}, a well-established result known as the \textit{Bernstein-von Mises theorem}.
-With a modest model structure including 17 fixed effects and 2,646 data points in \cite{Levshina2016}, it is highly likely that the same results would have been obtained with Maximum Likelihood methods.
-In fact, she admits that changing the priors did not lead to substantially different results in her footnote 8.
-This is a clear sign that the prior is ``swamped by the data'' \citep[1119]{Freedman1999}.
-% Going back to the argument that Bayesian methods are allegedly more cognitively plausible than `frequentist' methods, the claim might still be true.
-% However, most similar studies (dating back to \citealp{BresnanEa2007}) use an estimator (MLE) that is not frequentist in a narrow sense, and which is as cognitively plausible as Bayesian estimators because it most likely produces the same results in these studies.
-
-If there had been evidence in Levshina's study that Bayesian and MLE methods did \textit{not} converge, it would have been an occasion to demonstrate the selective superiority of the algorithms used in Bayesian estimation.
-After all, there are situations where Bayesian estimators can be more robust, namely with heavily censored data, complex hierarchical models, perfect separation, etc.
+% In this section, I justify my choice of statistical models used in Section~\ref{sec:corpushierarchicalmodel}.
+% Readers might think that the method used here for modeling grammatical alternations -- namely (Hierarchical) Logistic Regression\slash Generalised Linear (Mixed) Models or GL(M)Ms -- does not require much justification.
+% After all, GLMMs have been established as the major tool in the analysis of alternation phenomena.
+% All studies mentioned at the outset of Section~\ref{sec:cogocl} use some form of regression\slash GLMM.
+% Over the past few years, however, modified or alternative methods have been proposed.
+% While it is impossible to discuss all of these methods here, I just make a few remarks on Bayesian estimation (see \citealp{GelmanEa2014}) as it was proposed in \cite{Levshina2016} and \cite{Divjak2016a}, for example.
+% Conceptually, I see three points of discussion that should be kept apart.
+% First, Bayesian methods are sometimes touted as superior tools for scientific inference compared to frequentist methods.
+% Second, it has been proposed that the Bayesian interpretation of probability is more cognitively adequate for the modeling of linguistic data \cite[301--302]{Divjak2016a}.
+% Third and very specific to this paper, given established methods in the modeling of alternation and variation, it has to be decided whether so-called Bayesian methods lead to substantially different results.
+% 
+% % As for the first point, this is not quite the place to discuss it fully.
+% % The basic distinction is a philosophical one and related to the concepts of \textit{direct} and \textit{inverse probability} (\eg\ \citealp{Senn2011}).
+% % Frequentists assume that models and parameters are fixed, for example a model specifying that a coin is fair.
+% % We can then calculate for observed data (for example a measurement of 3 heads in an experiment with 10 tosses) how often such a result or a more extreme result would occur if the model were true and we repeated the experiment arbitrarily often.
+% % This is essentially the frequentist notion of direct probability, \ie\ long-run frequencies under replication.
+% % Standard tests in the Fisher and Neyman-Pearson traditions as well as Neyman confidence intervals are based on this concept of probability.
+% % Bayesian approaches (in the now common interpretation), on the other hand, condition on the particular data and quantify inductively the probability of model parameters given the data.
+% % The parameters are thus not fixed, and the probability is usually equated with researchers' posterior beliefs about model parameters.
+% % There is actually a debate among Bayesians about the proper interpretation of Bayesian methods, and whether a notion of hypothesis testing is compatible (or even already contained) in the Bayesian approach.
+% % In \citet[10]{GelmanShalizi2013}, the authors -- prominent Bayesians themselves -- acknowledge that a theory of statistical testing is a desideratum, and they state about the standard inductive interpretation of Bayesianism that ``most of this received view of Bayesian inference is wrong,'' and they develop a Bayesian notion of p-values (see also \citealp{Mayo2013}, for a frequentist reply; also \citealp{Senn2011} on different strands of Bayesianism and their stance on inductive vs.\ deductive reasoning, and \citealp{Mayo2011}, for a critical reply to \citealp{Senn2011}).
+% % Clearly, in such quarrels between and among camps of philosophers of science and statisticians, it is difficult for mere practitioners to take sides.
+% % 
+% % These quarrels relate to the second point, however.
+% % \citet[301--302]{Divjak2016a} speaks favourably of Bayesian methods because the Bayesian concept of probability is more adequate for cognitive modeling compared to the frequentist one.
+% % Her argument is part of a larger body of literature asking for cognitively plausible modeling techniques, for example Naive Discriminative Learning (NDL; \citealp{Baayen2011,BaayenEa2013,MilinEa2016,TheijssenEa2013}).
+% % On p.\ 303 of \cite{Divjak2016a}, the author goes on to explicitly mention NDL as well.
+% % Yet, neither frequentist nor Bayesian methods were conceived as cognitive models, but as systems of inference for scientists (see above, and see also \citealp[302]{Divjak2016a}).
+% % The fundamental question that lurks behind such arguments is how we interpret our statistical models (estimated on corpus data).
+% % Are they inductive models of cognitive representations that also human learners would infer from being exposed to the corpus data?%
+% % \footnote{In which case we are doing ``data science in language research'' in the words of \citealp{MilinEa2016}.
+% % I see this as standing in contradiction to the view advocated in \cite{Dabrowska2016} as cited above.}
+% % Or are they tests of theories that are pre-specified and merely tested for predictive accuracy on linguistic output data contained in corpora?
+% % In the former case, we adopt a strong \textit{corpus as input} hypothesis and should definitely resort to methods like NDL.
+% % However, this would most likely require us to toss most previous work done in (cognitive) linguistics into the bin and to abandon all high-level generalisations that most existing studies have been based upon.
+% % In fact, arguments to this extreme effect and against using high-level generalisation have been made, for example in \cite{BaayenEa2016}, \citet[299--300]{Divjak2016a}, \cite{RamscarPort2016}, \cite{TheijssenEa2013}.
+% % In the latter and less extreme case, the cognitive commitment does, however, not necessarily extend to the statistical methods used.
+% % These methods then do not need to be any more cognitively plausible than an ANOVA used to analyse the results from an experiment.
+% % I view my own work (see Sections~\ref{sec:germanmeasurenps}--\ref{sec:experimental}) in the tradition of testing theories (which embrace high-level generalisations), and I agree to provisionally use Likelihood methods (see below).
+% % I remain fully agnostic with respect to the question, which approach is \textit{right}.
+% % The best strategy for cognitive linguistics as a field might be to cultivate many methods while making sure that each method is applied carefully and competently.
+% 
+% The first and second point cannot be discussed in detail here.%
+% \footnote{However, there is a number of critical papers by statisticians (even prominent Bayesians) and philosophers of science in which the extreme credibility that has recently been assigned to (subjective and automatic) Bayesian approaches (and especially the standard inductivist interpretation of Bayesianism) is questioned (for example, \citealp{GelmanShalizi2013,Mayo2011,Senn2011}).}
+% With regard to the third point, \citet[251--252]{Levshina2016} argues for Bayesian estimation in mixed regression settings.
+% First, she claims that ``while frequentist statistics only allows one to test whether the null hypothesis can be rejected, Bayesian statistics enables one both to test the null hypothesis and to estimate the probability of specific parameter values given the data.''
+% This does not do justice to frequentist methods in that a strong focus on the rejection of the null hypothesis is characteristic only of Fisher's approach.
+% In the Neyman-Pearson approach, results ideally \textit{favour} the main hypothesis vis-à-vis the alternative hypothesis (cf.\ \citealp{Lehmann1993,Lehmann2011,Perezgonzalez2015}).
+% Also, especially Neyman-style frequentism has well-known extensions to estimation, for example in the form of confidence intervals (see \citealp{GreenlandEa2016}, esp.\ p.\ 340).
+% She then explains that a ``distinctive feature of Bayesian statistics is the use of so-called priors'' and that ``posterior probabilities depend on both the prior beliefs and the data, whereas the results of a frequentist model depend only on the data'' \cite[252]{Levshina2016}.
+% Remarkably, given this statement, she does \textit{not} use informative priors and in her footnote 8 \cite[252]{Levshina2016} admits that priors were probed using trial and error.
+% So, the proclaimed major advantage of Bayesian modeling was apparently not taken advantage of.%
+% \footnote{In the words of \cite{Senn2011}: ``You may believe you are a Bayesian but you are probably wrong.''
+% \citet[347--348]{GelmanHill2006} ``view any noninformative prior distribution as inherently provisional'' and give recommendations how to proceed once posteriors have been obtained from noninformative priors.}
+% Now, Maximum Likelihood Estimation (MLE) -- the traditional method which could have been used instead -- is not exactly \textit{frequentist} in the sense of Neyman-Pearson testing theory.
+% MLE, like inductive Bayesianism, conditions on the particular data inasmuch as it searches for the most likely set of parameters given the data.
+% What is more, Bayesian estimators are in fact based on the Likelihood and merely multiply it by the prior \citep[6--8]{GelmanEa2014}.
+% If the prior is flat, results converge (see also \citealp[347]{GelmanHill2006}).
+% The same is true if the sample size is large compared to the number of parameters, at least for finite-dimensional parameter models \citep[1119--1120]{Freedman1999}, a well-established result known as the \textit{Bernstein-von Mises theorem}.
+% With a modest model structure including 17 fixed effects and 2,646 data points in \cite{Levshina2016}, it is highly likely that the same results would have been obtained with Maximum Likelihood methods.
+% In fact, she admits that changing the priors did not lead to substantially different results in her footnote 8.
+% This is a clear sign that the prior is ``swamped by the data'' \citep[1119]{Freedman1999}.
+% % Going back to the argument that Bayesian methods are allegedly more cognitively plausible than `frequentist' methods, the claim might still be true.
+% % However, most similar studies (dating back to \citealp{BresnanEa2007}) use an estimator (MLE) that is not frequentist in a narrow sense, and which is as cognitively plausible as Bayesian estimators because it most likely produces the same results in these studies.
+% 
+% If there had been evidence in Levshina's study that Bayesian and MLE methods did \textit{not} converge, it would have been an occasion to demonstrate the selective superiority of the algorithms used in Bayesian estimation.
+% After all, there are situations where Bayesian estimators can be more robust, namely with heavily censored data, complex hierarchical models, perfect separation, etc.
 (see \citealp{Freedman1999}, \citealp[345--348]{GelmanHill2006}).
-I want to state clearly that these points do not in any way invalidate the results presented in \cite{Levshina2016}.
-However, being ``Bayesian'' is most likely not among its selling points.
-Additionally, I want to voice the concern that many practitioners are probably already struggling with getting an adequate grasp of advanced statistical methods and that it might therefore be wise to use the more conservative and better understood method if the alternative method is not absolutely required for substantive reasons.
-In the next section, I estimate the parameters of my hierarchical model with MLE and Markov-Chain Monte Carlo (MCMC) methods (the currently most prominent estimator used in Bayesian settings) to demonstrate their expectable convergence.
-
+% I want to state clearly that these points do not in any way invalidate the results presented in \cite{Levshina2016}.
+% However, being ``Bayesian'' is most likely not among its selling points.
+% Additionally, I want to voice the concern that many practitioners are probably already struggling with getting an adequate grasp of advanced statistical methods and that it might therefore be wise to use the more conservative and better understood method if the alternative method is not absolutely required for substantive reasons.
+% In the next section, I estimate the parameters of my hierarchical model with MLE and Markov-Chain Monte Carlo (MCMC) methods (the currently most prominent estimator used in Bayesian settings) to demonstrate their expectable convergence.
+% 
+% 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2017-12-27 Latex/sections/5conclusions.tex
More reviewer-induced changes to Sec. 1.
Nearing completion of Sec. 1 rewrite.
Minor progess w/ revision.

@@
 \label{sec:conclusion}
+
+\hl{Mention pre-registration.}
+
+\hl{The fact that the correlation is not stronger can be explained in parts \ldots}
+
+\hl{Wen need: more case studies +  diverse and different corpora + focus on register/style-specific aspects + inter-speaker similarity + focus on different experimental paradigms and what they really tell us + more clarity wrt statistical methods + ultimately fully specified mathematical models as in cognitive science}
 

### 2017-12-29 Latex/sections/2descriptive.tex
Revisions.

@@
-\section{Case in German measure NPs}
+\section{Modelling the measure noun alternation}
 \label{sec:germanmeasurenps}
@@
 
+\subsubsection{A closer look at the syntax of the alternating constructions}
+\label{sec:syntax}
+
+In order to understand the MNP alternation, it is vital to consider how in the relevant syntactic structure [N\Sub{1}~[A~N\Sub{2}]\Sub{NP\Sub{2}}]\Sub{NP\Sub{1}} (as instantiated in the \NACa\ and the \PGCa), the adjective is morpho-syntactically ambiguous between a determiner and a modifying adjective.
+To show this, the strong\slash weak inflection patterns of adjectives needs to be taken into account.
+In NPs with a strongly inflected determiner, attributive adjectives inflect according to the massively syncretic \textit{weak} pattern.\footnote{The weak pattern has only two suffixes.
+The structural cases (nominative and accusative) in the singular are marked with \textit{-e}, the oblique cases (dative and genitive) as well as the entire plural are marked with \textit{-en}.
+The only exception to this rule is the masculine accusative singular, which is also marked with \textit{-en}.}
+If there is no determiner (as is the case in the alternating constructions), attributive adjectives inflect like determiners themselves.
+This is called the \textit{strong} inflectional pattern.
+For example, in the dative (governed here by the preposition \textit{mit}), the strong suffix \textit{-em} is on the determiner in (\ref{ex:starkschwach:a}) and the adjective bears the weak suffix \textit{-en}.
+In (\ref{ex:starkschwach:b}), the strong suffix \textit{-em} appears on the adjective because there is no determiner.%
+\footnote{In the masculine and neuter singular genitive, the strong and weak forms are indistinguishable.
+Since the alternation itself does not occur in the genitive (see Section~\ref{sec:descriptive}), this does not create any complications for the present study.}
+
+\begin{exe}
+  \ex\label{ex:starkschwach} 
+  \begin{xlist}
+    \ex{\label{ex:starkschwach:a}\gll mit ein\textbf{-em} stark\textbf{-en} Kaffee\\
+                                      with a strong coffee\\}
+    \ex{\label{ex:starkschwach:b}\gll mit stark\textbf{-em} Kaffee\\
+                                      with strong coffee\\}
+  \end{xlist}
+\end{exe}
+
+Thus, the adjectives in the \NACa\ and the \PGCa\ have properties of adjectives as well as determiners.
+On the one hand, they are lexical adjectives and function as attributive modifiers.
+On the other hand, they are inflected like determiners, and they are the leftmost element in the NP, which is typical of determiners.
+This unusual double nature of adjectives in NPs without determiners leads to a plausible interpretation of the pattern shown in Table~\ref{tab:constructions}.
+Whenever speakers classify the adjective in the kind noun phrase more as a determiner, they have to use the \PGCa\ because, if there is a determiner, the PGC is the only option.
+When they classify the adjective more as an adjective, the kind NP has no determiner, and they have to use the \NACa.%
+\footnote{While the generative analysis presented in \cite{Bhatt1990} cannot properly deal with probabilistic effects, Bhatt comes close to this interpretation by analysing the kind NP in the PGC as a DP and in the NAC as an NP.}
+This morpho-syntactic ambiguity means that the \NACa\ is in fact a \NACb\ in disguise, and the \PGCa\ is a \PGCd\ in disguise.
+This explains why the alternation arises in the first place.
+In Section~\ref{sec:analyses}, I will argue that the NAC and the PGC constructions also have semantically distinguishable prototypes, and that the alternation between \PGCa\ and \NACa\ is an alternation between these prototypes.
+
 \subsubsection{Neighbouring cases}
@@
 
-I now turn to some more subtle issues related to the measure noun case alternation in order to delimit the scope of the study.
+I now finally turn to some more subtle issues related to the measure noun case alternation in order to delimit the scope of the study.
 First, there is a claim found in some grammars that a generic nominative, accusative, and even dative on the kind noun are used instead of the genitive (\PGCa) or case agreement (\NACa).
@@
 
-\subsection{Prototype and exemplar effects}
+\subsection{What controls the MNP alternation?}
 \label{sec:analyses}
 
-This section briefly reviews existing analyses of the \PGCa\ vs.\ \NACa\ alternation and related issues.
-I also develop my own analysis and the appropriate hypotheses for the empirical studies presented in Sections~\ref{sec:corpusstudies} and \ref{sec:experimental}.
+In this section, I develop my analysis of the MNP alternation and the appropriate hypotheses for the empirical studies presented in Sections~\ref{sec:corpusstudies} and \ref{sec:experimental}.
+I also review some existing analyses of the alternation and related issues.
+It should be remembered from Section~\ref{sec:syntax} that thanks to the syntactic ambiguity of strongly inflected adjectives, the \PGCa\ is plausibly a \PGCd\ in disguise and the \NACa\ a \NACb\ in disguise.
 
-.%
+\subsubsection{Prototype effects}
+\label{sec:prototypeeffects}
+
+My analysis is based on the idea that the PGC prototype express a pseudo-partitive where two discernible entities -- the measure and the substance -- are referenced.
+The NAC prototype, on the other hand, expresses a mere quantity where the measure is not referenced as an entitiy in its own right.
+Prototypical exemplars are given in (\ref{ex:prototypes1}) for the PGC and (\ref{ex:prototypes2}) for the NAC.
+
+\begin{exe}
+  \ex\label{ex:prototypes} 
+  \begin{xlist}
+    \ex{\label{ex:prototypes1}\gll Sie nahmen [einen Löffel [irgendeiner Medizin]\Sub{Gen}]\Sub{Acc}.\\
+    they took a spoon some medication \\
+    \trans They took a spoon(ful) of some medication.
+  }
+    \ex{\label{ex:prototypes2}\gll Sie kauften [drei Liter [Öl]\Sub{Acc\slash caseless}]\Sub{Acc}.\\
+    they bought three liters oil \\
+    \trans They bought three liters of oil.
+  }
+  \end{xlist}
+\end{exe}
+
+While the \PGCd\ in (\ref{ex:prototypes1}) allows an interpretation where speakers conceptualise the substance itself, \ie the medication, and a spoon used to take a quantity from the medication, the \NACb\ in (\ref{ex:prototypes2}) does not allow such an interpretation.
+While in the given examples, the effect is strongly supported by the choice of the measure noun lemmas, I argue that the meaning of the prototypes is independent of this.%
+\footnote{A reviewer asked for a specification of the \textit{schematic meaning} of the superordinate constructions in terms of \citet{Langacker1987}.
+However, none of the constructions (\PGCd, \PGCa, \NACb, and \NACa) are exclusively associated with one of the discussed meanings, and thus I do not see how one could specify schemas, for which ``membership is not a matter of degree'' and whose properties are ``fully compatible with all the members of the category'' \citep[371]{Langacker1987}.
+A modelling in terms of a probabilistic similarity effects is more appropriate.
+See also the argument in \citealp[70--71]{Taylor2003} and my discussion in Section~\ref{sec:cogocl}.}
+The independent meanings of the prototypes are, as I will show, a result of diachronic developments and grammaticalisation processes.
+Furthermore, I argue that the prototypical meanings are reflected in their usage patterns, which will be tested in the corpus study in Section~\ref{sec:corpusstudies}.
+
+I begin by showing how the two meanings of the constructions emerge as a consequence of grammaticalisation processes of partitives and similar constructions.
+It is often assumed that pseudo-partitives and quantity constructions arise as a form of grammaticalised partitives (\eg\ \citealp[536--539]{Koptjevskaja2001} for Finnish and Estonian, \citealp[559]{Koptjevskaja2001} for European languages in general).
+The grammaticalisation paths uncovered by \citet[esp.\ 526--530]{Koptjevskaja2001} are relevant for the case at hand.
+The grammaticalisation path can start out (in some languages) with constructions involving two referential nouns (not even necessarily forming a single and contiguous NP) and a \textit{separative} meaning as in \textit{(cut) two slices from the cake} \citep[535]{Koptjevskaja2001}.
+In this type of construction, it is most obvious that two separate referents (in the given example the cake and the slices) are conceptualised.
+The \textit{part-of} meaning of true partitives as in \textit{a slice of the cake} represents the first stage of a development wherein the measure noun can already lose some semantic content, when for example words like \textit{bite} are no longer necessarily interpreted as a piece literally bitten out of something.
+The pseudo-partitive stage finally instantiates a \textit{quantity-of} relation, potentially even leading to fully grammaticalised quantifiers such as \textit{a lot}.
+In German, the PGC is clearly the older construction \citep{Zimmer2015}.
+As predicted by the grammaticalisation pattern just described, it still has the potential to form a true partitive (if the kind noun is definite).
+Conversely, the NAC lacks this ability to form true partitives, and has gone further down the grammaticalisation path.
+It is thus not surprising that it has lost (at least prototypically) the semantics which allows both the measure and the substance to be conceptually accessible as independent referents.
+As a consequence, we should expect the NAC constructions to be typical hosting constructions for more strongly grammaticalised measure nouns.
+For example, highly grammaticalised non-referential nouns like \textit{Gramm} `gram' and \textit{Meter} `metre' should occur proportionally more often in NAC constructions than in PGC constructions.
+If such preferences can actually shown in usage patterns, it would lend strong support for the hypothesised difference in the meanings of the prototypes.
+In the corpus study, measure lemmas will therefore be annotated with appropriate semantic class labels to check whether semantic classes of measure nouns have different affinities to the two variants.
+
+A second preference should also be observable as a consequence of the different meanings.
+As described above, the grammaticalisation path leads from NPs denoting individuated objects standing in a \textit{part-of} relation to a construction with a more diffuse \textit{quantity-of} relation.
+Both types of relations can be numerically quantified -- inasmuch as a precise number of \textit{parts} or a numerically exact \textit{quantity} can be specified.
+However, it is much more typical of quantities to be specified with numerical precision.
+This is most obviously so with the highly grammaticalised physical measure nouns like \textit{centilitre} which are very typically used with exact numerals instead of unspecific quantifiers, although both options are available in principle (\textit{three centilitres} vs.\ \textit{several centilitres}).
+Since the \NACa\ is more closely associated with the \textit{quantity-of} relation, cardinals as attributes of the measure noun are expected to have a higher proportional frequency in the \NACa.
+For illustration, (\ref{ex:determiners}) shows the expected alternants under this hypothesis.
+
+\begin{exe}
+  \ex\label{ex:determiners} 
+  \begin{xlist}
+    \ex{\label{ex:determiners:nac}\gll [[Drei Centiliter]\Sub{Nom} [heißer Rum]\Sub{Nom}]\Sub{Nom} sind genug.\\
+                                       three centilitres hot rum are enough\\
+				     \glt Three centilitres of hot rum is enough.}
+    \ex{\label{ex:determiners:pgc}\gll [[Einige Centiliter]\Sub{Nom} [heißen Rums]\Sub{Gen}]\Sub{Nom} sind genug.\\
+                                       some   centilitres hot    rum  are   enough\\
+				     \glt A few centilitres of hot rum is enough.}
+  \end{xlist}
+\end{exe}
+
+In (\ref{ex:determiners:nac}), the measure noun is modified by a cardinal \textit{drei} `three', and hence the \NACa\ is preferred.
+In (\ref{ex:determiners:pgc}), the measure noun is modified by a non-cardinal determiner \textit{einige} `some', and the \PGCa\ is preferred.
+Especially with exact physical measure nouns (like \textit{centilitre}), exact numerical quantification is invited.
+By hypothesis, however, this goes beyond a selection effect tied to measure lemmas, and cardinal quantifiers are expected to co-occur relatively more often with the \NACa.
+
+Finally, it was found that the \PGCa\ is more typical of higher stylistic levels (distinctly edited, closer to the non-regional standard, more formal) and\slash or even exclusive to written language (see \citealp[320--323]{Hentschel1993}).
+This is not surprising, given that the genitive -- an intrinsic part of the PGC -- is generally underrepresented in colloquial vernacular variants of German as a result of a diachronic process wherein many (but by no means all) uses of the genitive are being replaced by other cases or periphrastic constructions \citep{FleischerSchallert2011}.
+Under an integral view of prototypes, which incorporate effects related to larger contexts, registers, and styles, such preferences can be part of what defines the construction prototypes, and the \PGCa\ should occur proportionally more often in more elaborate styles closer to the standard.%
 \footnote{A reviewer mentioned that she or he had the impression that dialectal variation is also a factor influencing the alternation.
 This is definitely true as some dialects (such as Alemannic) tend to have no genitive at all.
-While this is an interesting aspect for further research, the main obstacle is that there are no corpora of German which are both large enough and annotated with reliable meta data about regional variants.
-In the annotation of the corpus study, documents obviously written in a regional variant were excluded.
-}
+While this is an interesting aspect for further research, the main obstacle for the present study is that there are no corpora of German which are both large enough and annotated with reliable meta data about regional variants.
+In the annotation of the corpus study, documents obviously written in a regional variant were excluded.}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Items, exemplar effects, and hierarchical models}
+\label{sec:itemandexemplareffects}
+
+As discussed in Section~\ref{sec:cogocl}, prototype-based and exemplar-based approaches are merely the endpoints of a spectrum of theories allowing abstraction in cognitive representations to varying degrees.
+While the prototypes for the PGC and the NAC were clearly specified with reference to abstract meaning in Section~\ref{sec:prototypeeffects}, at least one exemplar-driven similarity effect can also be expected to influence the MNP alternation.
+The measure and kind noun lemmas which occur in the alternating constructions obviously also occur in the non-alternating constructions.
+The relative frequency with which they occur in these stable cases -- where choosing an alternative is impossible -- could thus be a factor influencing the alternating, less stable case.
+Should this be confirmed, it would be highly implausible to conceive of such effects as prototype effects.
+In the corpus study reported in Section~\ref{sec:corpusstudies}, a measure quantifying this influence will therefore be included as the \textit{attraction strength}.
+
+It should be noted that such and exemplar-type effect would have to exist in addition to the predicted semantic prototype effect for measure lemmas described in Section~\ref{sec:prototypeeffects}, namely the effect of semantic classes of measure nouns.
+
+To control for the possibility that a potential lemma attraction strength effect is merely an artefact of lemma-specific tendencies, such simple lemma-specific tendencies will have to be accounted for in the modelling as well.
+Additionally,
+
+This leads to a more complex model structure if a generalised linear mixed model (GLMM) is specified, as is customary in alternation modelling (see \citealp{Gries2015} for an argument in favour of such models in corpus linguistic studies; \citealp{GelmanHill2006} (especially part 2) for a comprehensive text book on the subject).
+A GLMM models the influence which several variables (\textit{predictors}) have on the probability that an alternant is chosen.
+The so-called \textit{fixed effects} are assumed to be fixed population parameters which quantify the strength and the direction of the effect which, for example, a specific feature of the syntactic context or the stylistic level of the text have on the choice of the alternants.
+The so-called \textit{random effects} are not fixed population parameters, but they vary by group.
+The simplest random effect is a varying intercept, which predicts for groups of observations defined by lemmas, genres, speakers, etc. a constant term to be added to the model.
+In a truly hierarchical model, however, the varying intercept itself comes with an additional linear model where second-level fixed effects predict the group-level effect.
+To illustrate, the simple model specification in (\ref{eq:glmm}) is a hierarchical logistic regression model.
+
+\begin{equation}
+  Pr(y_i=1)=logit^{-1}(\beta^0_{j[i]}+\beta^1\cdot x^1_i+\beta^2\cdot x^2_i)
+  \label{eq:glmm}
+\end{equation}
+
+The varying intercept $\beta^0_{j[i]}$ adds a constant term for group $j$ to which $i$ belongs (encoded here as $j[i]$, a notation borrowed from \citealp{GelmanHill2006}).
+Each $\beta$ is a fixed-effect coefficient, and each $x$ is an observation-level variable.
+(Here, two fixed effects are included for illustration purposes.)
+The observations (for example, the lines of a concordance) are indexed by $i$ and the groups\slash random intercepts by $j$.
+Let us say the groups are defined by lemmas and we also include lemma frequency as a control variable.
+In this case, a second-level model would be given similar to (\ref{eq:level2}).
+
+\begin{equation}
+  \beta^0_j\sim\mathcal{N}(\gamma_0+\gamma^1\cdot u_j, \sigma)
+  \label{eq:level2}
+\end{equation}
+
+This says that the intercepts follow a normal distribution with a standard deviation of $\sigma$ and a mean predicted from the model $\gamma_0+\gamma^1\cdot u_j$, where $\gamma^0$ is the second-level intercept, other $\gamma$s the second-level coefficients, and each $u$ a second-level predictor variable.
+If $\beta^0_{j[i]}$ were a lemma random effect and we also included, for example, lemma frequency as a control variable, then $u_j$ could be the lemma frequency predictor which helps to predict the lemma frequency instead of just having simple per-lemma constants.

### 2017-12-29 Latex/sections/3corpus.tex
Revisions.

--- 2017-12-22 Latex/sections/3corpus.tex
@@
 
-\subsection{Corpus choice and sampling}
+\subsection{Corpus choice}
 \label{sec:gettingdata}
-
-% TODO Quote:
-% Brems (2016:171--175) for using GOOGLE data, including COUNTS!
 
@@
 Therefore, DECOW is a valid choice for this study.
+
+\subsection{Pre-study: main prototype effects in the non-alternating constructions}
+
+\subsection{Main study: alternating constructions}
+
+\subsubsection{Sampling}
 
@@
 
-\subsection{Variables and annotation}
+\subsubsection{Variables and annotation}
 \label{sec:annotation}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Model}
+\label{sec:corpushierarchicalmodel}
 
-\subsection{On statistical analysis}
-\label{sec:rightstatistics}
-
-In this section, I justify my choice of statistical models used in Section~\ref{sec:corpushierarchicalmodel}.
-Readers might think that the method used here for modeling grammatical alternations -- namely (Hierarchical) Logistic Regression\slash Generalised Linear (Mixed) Models or GL(M)Ms -- does not require much justification.
-After all, GLMMs have been established as the major tool in the analysis of alternation phenomena.
-All studies mentioned at the outset of Section~\ref{sec:cogocl} use some form of regression\slash GLMM.
-Over the past few years, however, modified or alternative methods have been proposed.
-While it is impossible to discuss all of these methods here, I just make a few remarks on Bayesian estimation (see \citealp{GelmanEa2014}) as it was proposed in \cite{Levshina2016} and \cite{Divjak2016a}, for example.
-Conceptually, I see three points of discussion that should be kept apart.
-First, Bayesian methods are sometimes touted as superior tools for scientific inference compared to frequentist methods.
-Second, it has been proposed that the Bayesian interpretation of probability is more cognitively adequate for the modeling of linguistic data \cite[301--302]{Divjak2016a}.
-Third and very specific to this paper, given established methods in the modeling of alternation and variation, it has to be decided whether so-called Bayesian methods lead to substantially different results.
-
-% As for the first point, this is not quite the place to discuss it fully.
-% The basic distinction is a philosophical one and related to the concepts of \textit{direct} and \textit{inverse probability} (\eg\ \citealp{Senn2011}).
-% Frequentists assume that models and parameters are fixed, for example a model specifying that a coin is fair.
-% We can then calculate for observed data (for example a measurement of 3 heads in an experiment with 10 tosses) how often such a result or a more extreme result would occur if the model were true and we repeated the experiment arbitrarily often.
-% This is essentially the frequentist notion of direct probability, \ie\ long-run frequencies under replication.
-% Standard tests in the Fisher and Neyman-Pearson traditions as well as Neyman confidence intervals are based on this concept of probability.
-% Bayesian approaches (in the now common interpretation), on the other hand, condition on the particular data and quantify inductively the probability of model parameters given the data.
-% The parameters are thus not fixed, and the probability is usually equated with researchers' posterior beliefs about model parameters.
-% There is actually a debate among Bayesians about the proper interpretation of Bayesian methods, and whether a notion of hypothesis testing is compatible (or even already contained) in the Bayesian approach.
-% In \citet[10]{GelmanShalizi2013}, the authors -- prominent Bayesians themselves -- acknowledge that a theory of statistical testing is a desideratum, and they state about the standard inductive interpretation of Bayesianism that ``most of this received view of Bayesian inference is wrong,'' and they develop a Bayesian notion of p-values (see also \citealp{Mayo2013}, for a frequentist reply; also \citealp{Senn2011} on different strands of Bayesianism and their stance on inductive vs.\ deductive reasoning, and \citealp{Mayo2011}, for a critical reply to \citealp{Senn2011}).
-% Clearly, in such quarrels between and among camps of philosophers of science and statisticians, it is difficult for mere practitioners to take sides.
-% 
-% These quarrels relate to the second point, however.
-% \citet[301--302]{Divjak2016a} speaks favourably of Bayesian methods because the Bayesian concept of probability is more adequate for cognitive modeling compared to the frequentist one.
-% Her argument is part of a larger body of literature asking for cognitively plausible modeling techniques, for example Naive Discriminative Learning (NDL; \citealp{Baayen2011,BaayenEa2013,MilinEa2016,TheijssenEa2013}).
-% On p.\ 303 of \cite{Divjak2016a}, the author goes on to explicitly mention NDL as well.
-% Yet, neither frequentist nor Bayesian methods were conceived as cognitive models, but as systems of inference for scientists (see above, and see also \citealp[302]{Divjak2016a}).
-% The fundamental question that lurks behind such arguments is how we interpret our statistical models (estimated on corpus data).
-% Are they inductive models of cognitive representations that also human learners would infer from being exposed to the corpus data?%
-% \footnote{In which case we are doing ``data science in language research'' in the words of \citealp{MilinEa2016}.
-% I see this as standing in contradiction to the view advocated in \cite{Dabrowska2016} as cited above.}
-% Or are they tests of theories that are pre-specified and merely tested for predictive accuracy on linguistic output data contained in corpora?
-% In the former case, we adopt a strong \textit{corpus as input} hypothesis and should definitely resort to methods like NDL.
-% However, this would most likely require us to toss most previous work done in (cognitive) linguistics into the bin and to abandon all high-level generalisations that most existing studies have been based upon.
-% In fact, arguments to this extreme effect and against using high-level generalisation have been made, for example in \cite{BaayenEa2016}, \citet[299--300]{Divjak2016a}, \cite{RamscarPort2016}, \cite{TheijssenEa2013}.
-% In the latter and less extreme case, the cognitive commitment does, however, not necessarily extend to the statistical methods used.
-% These methods then do not need to be any more cognitively plausible than an ANOVA used to analyse the results from an experiment.
-% I view my own work (see Sections~\ref{sec:germanmeasurenps}--\ref{sec:experimental}) in the tradition of testing theories (which embrace high-level generalisations), and I agree to provisionally use Likelihood methods (see below).
-% I remain fully agnostic with respect to the question, which approach is \textit{right}.
-% The best strategy for cognitive linguistics as a field might be to cultivate many methods while making sure that each method is applied carefully and competently.
-
-The first and second point cannot be discussed in detail here.%
-\footnote{However, there is a number of critical papers by statisticians (even prominent Bayesians) and philosophers of science in which the extreme credibility that has recently been assigned to (subjective and automatic) Bayesian approaches (and especially the standard inductivist interpretation of Bayesianism) is questioned (for example, \citealp{GelmanShalizi2013,Mayo2011,Senn2011}).}
-With regard to the third point, \citet[251--252]{Levshina2016} argues for Bayesian estimation in mixed regression settings.
-First, she claims that ``while frequentist statistics only allows one to test whether the null hypothesis can be rejected, Bayesian statistics enables one both to test the null hypothesis and to estimate the probability of specific parameter values given the data.''
-This does not do justice to frequentist methods in that a strong focus on the rejection of the null hypothesis is characteristic only of Fisher's approach.
-In the Neyman-Pearson approach, results ideally \textit{favour} the main hypothesis vis-à-vis the alternative hypothesis (cf.\ \citealp{Lehmann1993,Lehmann2011,Perezgonzalez2015}).
-Also, especially Neyman-style frequentism has well-known extensions to estimation, for example in the form of confidence intervals (see \citealp{GreenlandEa2016}, esp.\ p.\ 340).
-She then explains that a ``distinctive feature of Bayesian statistics is the use of so-called priors'' and that ``posterior probabilities depend on both the prior beliefs and the data, whereas the results of a frequentist model depend only on the data'' \cite[252]{Levshina2016}.
-Remarkably, given this statement, she does \textit{not} use informative priors and in her footnote 8 \cite[252]{Levshina2016} admits that priors were probed using trial and error.
-So, the proclaimed major advantage of Bayesian modeling was apparently not taken advantage of.%
-\footnote{In the words of \cite{Senn2011}: ``You may believe you are a Bayesian but you are probably wrong.''
-\citet[347--348]{GelmanHill2006} ``view any noninformative prior distribution as inherently provisional'' and give recommendations how to proceed once posteriors have been obtained from noninformative priors.}
-Now, Maximum Likelihood Estimation (MLE) -- the traditional method which could have been used instead -- is not exactly \textit{frequentist} in the sense of Neyman-Pearson testing theory.
-MLE, like inductive Bayesianism, conditions on the particular data inasmuch as it searches for the most likely set of parameters given the data.
-What is more, Bayesian estimators are in fact based on the Likelihood and merely multiply it by the prior \citep[6--8]{GelmanEa2014}.
-If the prior is flat, results converge (see also \citealp[347]{GelmanHill2006}).
-The same is true if the sample size is large compared to the number of parameters, at least for finite-dimensional parameter models \citep[1119--1120]{Freedman1999}, a well-established result known as the \textit{Bernstein-von Mises theorem}.
-With a modest model structure including 17 fixed effects and 2,646 data points in \cite{Levshina2016}, it is highly likely that the same results would have been obtained with Maximum Likelihood methods.
-In fact, she admits that changing the priors did not lead to substantially different results in her footnote 8.
-This is a clear sign that the prior is ``swamped by the data'' \citep[1119]{Freedman1999}.
-% Going back to the argument that Bayesian methods are allegedly more cognitively plausible than `frequentist' methods, the claim might still be true.
-% However, most similar studies (dating back to \citealp{BresnanEa2007}) use an estimator (MLE) that is not frequentist in a narrow sense, and which is as cognitively plausible as Bayesian estimators because it most likely produces the same results in these studies.
-
-If there had been evidence in Levshina's study that Bayesian and MLE methods did \textit{not} converge, it would have been an occasion to demonstrate the selective superiority of the algorithms used in Bayesian estimation.
-After all, there are situations where Bayesian estimators can be more robust, namely with heavily censored data, complex hierarchical models, perfect separation, etc.
-(see \citealp{Freedman1999}, \citealp[345--348]{GelmanHill2006}).
-I want to state clearly that these points do not in any way invalidate the results presented in \cite{Levshina2016}.
-However, being ``Bayesian'' is most likely not among its selling points.
-Additionally, I want to voice the concern that many practitioners are probably already struggling with getting an adequate grasp of advanced statistical methods and that it might therefore be wise to use the more conservative and better understood method if the alternative method is not absolutely required for substantive reasons.
-In the next section, I estimate the parameters of my hierarchical model with MLE and Markov-Chain Monte Carlo (MCMC) methods (the currently most prominent estimator used in Bayesian settings) to demonstrate their expectable convergence.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{A hierarchical model of the measure noun alternation}
-\label{sec:corpushierarchicalmodel}
+ using R \citep{R}, \textit{lme4} \citep{lme4}
 
@@
   \centering
-  \includegraphics[width=0.85\textwidth]{../R/output/corpus_fixeffs_mle+mcmc}
-  \caption{Coefficients (MLE and MCMC) with 95\% confidence intervals (for details see text); the intercept is -4.328 (MLE) and -4.441 (MCMC)}
+  \includegraphics[width=0.85\textwidth]{../R/output/corpus_fixeffs}
+  \caption{Coefficients with 95\% confidence intervals (for details see text); the intercept is -4.328}
   \label{fig:fixeffs}
@@
 
-In this section, I report the results of fitting a multilevel model to the data using R \citep{R}, \textit{lme4} \citep{lme4} for Maximum Likelihood Estimation (MLE) and \textit{rstanarm} \citep{rstanarm} for Bayesian Markov-Chain Monte Carlo (MCMC) estimation (see Section~\ref{sec:rightstatistics}).
+In this section, I report the results of fitting a multilevel model to the alternation data.
 The purpose is to model the influence of the regressors specified in Table~\ref{tab:variables} on the probability that the \PGCa\ is chosen over the \NACa.
@@
 The intercept comprises \textit{Cardinal=Yes}, \textit{Measurecase=Nom}, \textit{Kindgender=Masc}, \textit{Measureclass=Physical}, and 0 for all numeric z-transformed regressors.
-It is -4.328 for the ML estimate and -4.441 for the MCMC estimate.
+It was estimated at -4.328.
 
@@
 
-\begin{sidewaystable}
+\begin{table}
   \centering
-  \resizebox{\textheight}{!}{
-  \begin{tabular}{llrlp{0.5em}rrp{0.5em}rrp{0.5em}rrp{0.5em}cc}
-      Model & Regressor & \multicolumn{1}{l}{\pPB} & Factor && \multicolumn{2}{l}{Coefficient} && \multicolumn{2}{l}{CI low} && \multicolumn{2}{l}{CI high} && \multicolumn{2}{l}{CI excludes 0} \\
-      level &                   &       & level     && \multicolumn{1}{l}{MLE} & \multicolumn{1}{l}{MCMC} && \multicolumn{1}{l}{MLE} & \multicolumn{1}{l}{MCMC} && \multicolumn{1}{l}{MLE} & \multicolumn{1}{l}{MCMC} && \multicolumn{1}{l}{MLE} & \multicolumn{1}{l}{MCMC} \\\midrule
-       First     & Badness           &  0.002 &           && -0.152 & -0.155 && -0.247 & -0.247 && -0.061 & -0.065 && * & * \\
-                 & Cardinal          &  0.001 & No        &&  1.189 &  1.222 &&  0.862 &  0.927 &&  1.466 &  1.496 && * & * \\
-                 & Genitives         &  0.001 &           && -0.693 & -0.711 && -0.768 & -0.801 && -0.592 & -0.616 && * & * \\
-                 & Measurecase       &  0.001 & Acc       &&  0.030 &  0.031 && -0.150 & -0.159 &&  0.212 &  0.222 &&   &   \\
-                 &                   &        & Dat       &&  0.705 &  0.729 &&  0.455 &  0.465 &&  0.944 &  0.995 && * & * \\[0.5\baselineskip]
+  \resizebox{\textwidth}{!}{
+    \begin{tabular}{llrlrrrc}
+    Model level  & Regressor         & $\text{p}_{\text{PB}}$ & Factor level & Coefficient & CI low & CI high & CI excludes 0 \\
+    \midrule
+    First        & Badness           &  0.002                 &              & -0.152      & -0.247 & -0.061  & *             \\
+                 & Cardinal          &  0.001                 & No           &  1.189      &  0.862 &  1.466  & *             \\
+                 & Genitives         &  0.001                 &              & -0.693      & -0.768 & -0.592  & *             \\
+                 & Measurecase       &  0.001                 & Acc          &  0.030      & -0.150 &  0.212  &               \\
+                 &                   &                        & Dat          &  0.705      &  0.455 &  0.944  & *             \\[0.5\baselineskip]
 
-       Second    & Kindattraction    &  0.020 &           &&  0.225 &  0.244 &&  0.049 &  0.056 &&  0.393 &  0.422 && * & * \\
-       (Kind)    & Kindfreq          &  0.095 &           &&  0.146 &  0.164 && -0.023 & -0.016 &&  0.301 &  0.341 &&   &   \\
-                 & Kindgender        &  0.001 & Neut      &&  0.021 &  0.013 && -0.367 & -0.409 &&  0.392 &  0.435 &&   &   \\
-                 &                   &        & Fem       &&  1.269 &  1.289 &&  0.800 &  0.788 &&  1.709 &  1.783 && * & * \\[0.5\baselineskip]
+    Second       & Kindattraction    &  0.020                 &              &  0.225      &  0.049 &  0.393  & *             \\
+    (Kind)       & Kindfreq          &  0.095                 &              &  0.146      & -0.023 &  0.301  &               \\
+                 & Kindgender        &  0.001                 & Neut         &  0.021      & -0.367 &  0.392  &               \\
+                 &                   &                        & Fem          &  1.269      &  0.800 &  1.709  & *             \\[0.5\baselineskip]
 
-       Second    & Measureattraction &  0.001 &           &&  0.282 &  0.299 &&  0.106 &  0.102 &&  0.447 &  0.515 && * & * \\
-       (Measure) & Measureclass      &  0.001 & Container &&  0.252 &  0.257 && -0.265 & -0.303 &&  0.788 &  0.813 &&   &   \\
-                 &                   &        & Rest      &&  0.421 &  0.379 && -0.209 & -0.378 &&  1.063 &  1.091 &&   &   \\
-                 &                   &        & Amount    &&  0.831 &  0.889 &&  0.215 &  0.220 &&  1.432 &  1.569 && * & * \\
-                 &                   &        & Portion   &&  1.217 &  1.253 &&  0.675 &  0.689 &&  1.684 &  1.840 && * & * \\
-                 & Measurefreq       &  0.005 &           && -0.231 & -0.232 && -0.363 & -0.395 && -0.079 & -0.073 && * & * \\
+    Second       & Measureattraction &  0.001                 &              &  0.282      &  0.106 &  0.447  & *             \\
+    (Measure)    & Measureclass      &  0.001                 & Container    &  0.252      & -0.265 &  0.788  &               \\
+                 &                   &                        & Rest         &  0.421      & -0.209 &  1.063  &               \\
+                 &                   &                        & Amount       &  0.831      &  0.215 &  1.432  & *             \\
+                 &                   &                        & Portion      &  1.217      &  0.675 &  1.684  & *             \\
+                 & Measurefreq       &  0.005                 &              & -0.231      & -0.363 & -0.079  & *             \\
 
@@
   }
-  \caption{Coefficient table comparing Maximum Likelihood Estimation (MLE, with 95\% bootstrap confidence interval) and `Bayesian' Markov-Chain Monte Carlo estimation (MCMC); the intercept is -4.328 (MLE) and -4.441 (MCMC)}
+  \caption{Coefficient table with 95\% bootstrap confidence intervals; the intercept is -4.328}
   \label{tab:bigtable}
-\end{sidewaystable}
+\end{table}
 
-Standard diagnostics for MLE show that the model quality is quite good.
-Generalised variance inflation factors for the regressors were calculated to check for multicollinearity \citep{FoxMonette1992,ZuurEa2010}, and none of the corrected $\text{GVIF}^{1/2\text{df}}$ was higher than 1.6.
+Standard diagnostics show that the model quality is quite good.
 Nakagawa \& Schielzeth's pseudo-coefficient of determination is $R_m^2=0.409$ and $R^2_c=0.495$ (see \citealp{Gries2015} for a basic introduction to these $R^2$ measures, or else \citealp{NakagawaSchielzeth2013}).
 The rate of correct predictions is 0.843, which means a proportional reduction of error of $\lambda=0.297$.
+Generalised variance inflation factors for the regressors were calculated to check for multicollinearity \citep{FoxMonette1992,ZuurEa2010}, and none of the corrected $\text{GVIF}^{1/2\text{df}}$ was higher than 1.6.
 The lemma intercepts have standard deviations of $\sigma_{\text{Measurelemma}}=0.448$ and $\sigma_{\text{Kindlemma}}=0.604$.
@@
 
-Finally, I show that MCMC does not necessarily lead to different results.
-Actually, given the low complexity of the model and the large sample size, it would be surprising if they did (see Section~\ref{sec:rightstatistics}).
-The model was re-estimated using the \textit{stan\_glmer} function from the \textit{rstanarm} package, which provides an \textit{lme4}-compatible syntax for estimating common model types with the \textit{stan} software \citep{CarpenterEa2017}.
-The algorithm was run with 4 chains and 1,000 iterations, and I used plausible default priors.
-Most notably, priors for coefficients were specified as $\mathcal{N}(0,10)$ because coefficients higher than 10 or lower than $-10$ are extremely rare in well-specified models on appropriate data.%
-\footnote{Consider that with a coefficient of 10, each increase by 1 in the regressor variable increases the odds by $exp(10)=22,026.47$.
-To reliably estimate such coefficients, extremely large samples would be required.}
-The algorithm converged, and for all coefficients, the $\hat{R}$ diagnostic was exactly 1.
-The resulting coefficients and intervals as well as an * for ``confidence interval does not contain 0'' are also given in Table~\ref{tab:bigtable} in the columns labelled \textit{MCMC}.
-Both methods lead to exactly the same results (minus negligible numerical differences) as expected given the modest complexity of the model structure and the large sample size.
-The signs and magnitudes of the coefficients are identical, and the confidence intervals have the same width and symmetry properties.
-Figure~\ref{fig:fixeffs} illustrates this by also showing both estimates.
-Since both estimators converge, I only interpret the MLE model in the next section.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Interpretation}
+\subsubsection{Interpretation}
 \label{sec:interpretation}

### 2017-12-31 Latex/sections/1intro.tex
Statistics final. More revisions in Sec. 3.

@@
 The mixture model of categorisation (MMC) by \cite{Rosseel2002} is a model with abstraction in the form of hierarchical clusters of exemplars, and these clusters of objects are characterised by a probability distribution over their features, and categorising new objects is a process of estimating the probability of this object belonging to one of the clusters.
-\cite{GriffithsEa2009} go further and present a computational model which is able to chose the appropriate complexity of representation for a given category.
+\cite{GriffithsEa2009} go further and present a computational model which is able to choose the appropriate complexity of representation for a given category.
 However, despite these (and more) attempts to reconcile or unite the two approaches while developing spelled-out mathematical models, \cite[183--184]{Vanpaemel2016} describes the state of affairs between adherents of neo-prototype theory (such as \citealp{MindaSmith2001,MindaSmith2002}) and exemplar theory as a stalemate.
@@
 While I am convinced that the results presented in \citet{DivjakArppe2013} are valid and important (especially given the previous and subsequent research the authors have conducted on the data, including experimental work presented in \citealp{DivjakEa2016}), any data set can be clustered to yield a certain number of clusters.
-Thus, the study does not in any way ensure that the clusters emerging from the data correspond to any speaker's cognitive representation.%
+Thus, the study does not ensure that the clusters emerging from the data correspond to any speaker's cognitive representation.%
 \footnote{See, again, \citet[486--487]{Dabrowska2016} and \citet[22]{Gries2003}.}
@@
 In Section~\ref{sec:corpusstudies}, I present a corpus study and report a true multilevel generalised linear model with the appropriate hierarchical structure for the hypothesised effects.
-In Section~\ref{sec:experimental}, I test the predictions of the corpus-based model in two experimental paradigms (forced-choice and self-paced reading), showing that they indeed converge, albeit with intermediate effect strength.
+In Section~\ref{sec:experimental}, I test the predictions of the corpus-based model in two experimental paradigms (forced-choice and self-paced reading), showing that they indeed converge, albeit with no more than intermediate effect strength.
 In Section~\ref{sec:conclusion}, I interpret the findings in the light of the issues of cognitive representations and corpus data and the convergence of usage data and experimental data.

### 2017-12-31 Latex/sections/2descriptive.tex
Statistics final. More revisions in Sec. 3.

@@
 \footnote{A reviewer asked for a specification of the \textit{schematic meaning} of the superordinate constructions in terms of \citet{Langacker1987}.
-However, none of the constructions (\PGCd, \PGCa, \NACb, and \NACa) are exclusively associated with one of the discussed meanings, and thus I do not see how one could specify schemas, for which ``membership is not a matter of degree'' and whose properties are ``fully compatible with all the members of the category'' \citep[371]{Langacker1987}.
+However, none of the constructions (\PGCd, \PGCa, \NACb, and \NACa) are exclusively associated with one of the discussed meanings, and thus I do not see how one could specify schemas (even at a superordinate level), for which ``membership is not a matter of degree'' and whose properties are ``fully compatible with all the members of the category'' \citep[371]{Langacker1987}.
 A modelling in terms of a probabilistic similarity effects is more appropriate.
-See also the argument in \citealp[70--71]{Taylor2003} and my discussion in Section~\ref{sec:cogocl}.}
+See also the argument in \citet[70--71]{Taylor2003} and my discussion in Section~\ref{sec:cogocl}.}
 The independent meanings of the prototypes are, as I will show, a result of diachronic developments and grammaticalisation processes.
@@
 In the corpus study, measure lemmas will therefore be annotated with appropriate semantic class labels to check whether semantic classes of measure nouns have different affinities to the two variants.
+The actual classification was inspired by the list in \citet[530]{Koptjevskaja2001}, but due to the low frequencies of many of the potential classes, a very coarse classification was finally used.
+With typical examples, the classes are:
+\textit{Physical} (abstract precisely measurable units such as \textit{Liter} `litre', \textit{Meter} `metre', \textit{Gramm} `gram'),
+\textit{Container} (\textit{Eimer} `bucket'),
+\textit{Amount} (\textit{Menge} `amount'), 
+\textit{Portion} (natural portions like \textit{Happen} `bite' or \textit{Krümel} `crumb').
+The lemmas that did not fit into either of these classes were labelled \textit{Rest}.
 
@@
 
-\subsubsection{Items, exemplar effects, and hierarchical models}
+\subsubsection{Items, exemplar effects, and multilevel models}
 \label{sec:itemandexemplareffects}
@@
 
-It should be noted that such and exemplar-type effect would have to exist in addition to the predicted semantic prototype effect for measure lemmas described in Section~\ref{sec:prototypeeffects}, namely the effect of semantic classes of measure nouns.
+It should be noted that such and exemplar-type effect would have to be confirmed \textit{in addition} to the predicted semantic prototype effect for measure lemmas described in Section~\ref{sec:prototypeeffects}, namely the effect of semantic classes of measure nouns.
+This means that the statistical model needs to track an abstraction effect and an exemplar-like effect for measure nouns.
+Additionally, controlling for simple lemma frequency effects is always a good idea, and it should be done for kind and measure nouns.
+While, for example, very frequent kind nouns might preserve the older \PGCa, low-frequency kind nouns might tend to occur more often in the \NACa.
+For measure nouns, high frequency might lead to a higher affinity to the \NACa\ if we assume at least a tendency for highly grammaticalised items to be more frequent.
+Of course, the model must be specified in a way such that we can be sure that any detected frequency effect goes beyond the semantic effect, which is captured in the semantic classes described in Section~\ref{sec:prototypeeffects}.
+Such frequency effects would also be very difficult to describe as abstractions.
 
-To control for the possibility that a potential lemma attraction strength effect is merely an artefact of lemma-specific tendencies, such simple lemma-specific tendencies will have to be accounted for in the modelling as well.
-Additionally,
-
-This leads to a more complex model structure if a generalised linear mixed model (GLMM) is specified, as is customary in alternation modelling (see \citealp{Gries2015} for an argument in favour of such models in corpus linguistic studies; \citealp{GelmanHill2006} (especially part 2) for a comprehensive text book on the subject).
+With all this, it must also be made sure that these influences at the lemma level (lemma type frequency, attraction strength, and semantic classes) are not spurious and just artefacts of mere lemma idiosyncrasies.
+This leads to a rather complex and truly multilevel model structure for the generalised linear mixed models (GLMMs) customarily used in alternation modelling.%
+\footnote{See \citet{Gries2015} for an argument in favour of varying-slope and varying-intercept models in corpus linguistic studies.
+See \citet{GelmanHill2006} (especially part 2) for a comprehensive text book on the subject including multilevel models.}
 A GLMM models the influence which several variables (\textit{predictors}) have on the probability that an alternant is chosen.
@@
 The simplest random effect is a varying intercept, which predicts for groups of observations defined by lemmas, genres, speakers, etc. a constant term to be added to the model.
-In a truly hierarchical model, however, the varying intercept itself comes with an additional linear model where second-level fixed effects predict the group-level effect.
-To illustrate, the simple model specification in (\ref{eq:glmm}) is a hierarchical logistic regression model.
+In a true multilevel model, however, the varying intercept itself comes with an additional linear model where second-level fixed effects predict the group-level effect, and this is exactly what is needed here.
+To illustrate, the simple model specification in (\ref{eq:glmm}) is a multilevel logistic regression model.
 
 \begin{equation}
-  Pr(y_i=1)=logit^{-1}(\beta^0_{j[i]}+\beta^1\cdot x^1_i+\beta^2\cdot x^2_i)
+  Pr(y_i=1)=logit^{-1}(\beta^0_{j[i]}+\beta^1\cdot x^1_i)
   \label{eq:glmm}
@@
 
-The varying intercept $\beta^0_{j[i]}$ adds a constant term for group $j$ to which $i$ belongs (encoded here as $j[i]$, a notation borrowed from \citealp{GelmanHill2006}).
-Each $\beta$ is a fixed-effect coefficient, and each $x$ is an observation-level variable.
-(Here, two fixed effects are included for illustration purposes.)
+$Pr(y_1=1)$ is the predicted probability that the variable for the construction type in an observation (indexed by $i$) takes on the value 1 ($y_i=1$).
+The varying intercept $\beta^0_{j[i]}$ adds a constant for group $j$ to which $i$ belongs (encoded here as $j[i]$, a notation borrowed from \citealp{GelmanHill2006}) to the linear term.%
+\footnote{The linear term is the argument to the $logit^{-1}$ link function which transforms it into something which is interpretable as a probability.}
+Any other $\beta$ is a fixed-effect coefficient, and each $x$ is an observation-level variable.
+Here, just one fixed effect is included for illustration purposes, and $x^1$ could be, for example, the variable encoding whether a cardinal modifies the measure noun.
 The observations (for example, the lines of a concordance) are indexed by $i$ and the groups\slash random intercepts by $j$.
@@
 
-This says that the intercepts follow a normal distribution with a standard deviation of $\sigma$ and a mean predicted from the model $\gamma_0+\gamma^1\cdot u_j$, where $\gamma^0$ is the second-level intercept, other $\gamma$s the second-level coefficients, and each $u$ a second-level predictor variable.
-If $\beta^0_{j[i]}$ were a lemma random effect and we also included, for example, lemma frequency as a control variable, then $u_j$ could be the lemma frequency predictor which helps to predict the lemma frequency instead of just having simple per-lemma constants.
+This says that the intercepts follow a normal distribution with a standard deviation of $\sigma$ and a mean predicted from the model $\gamma_0+\gamma^1\cdot u_j$, where $\gamma^0$ is the second-level intercept, the other $\gamma$ are the second-level coefficients, and each $u$ is a second-level predictor variable.
+If $\beta^0_{j}$ were a lemma random effect, then $u_j$ would be the lemma frequency predictor which helps to predict the lemma intercept instead of just having a simple per-lemma constant.%
+\footnote{Since the standard \textit{lme4} package takes care of multilevel modelling automatically, an R formula for (\ref{eq:glmm}) and (\ref{eq:level2}) could be \texttt{Construction\textasciitilde Cardinal+Lemmafrequency+(1|Lemma)} given that in the data set, the values of \texttt{Lemmafrequency} are unique for each \texttt{Lemma}.}
+
+While multilevel models are rarely used in corpus linguistics, and \citet{Gries2015} even calls the simpler varying-intercepts and varying-slopes models ``underused'', they provide an excellent tool to describe situations where preferences at the sentence level and at lexical levels need to be integrated.
+The models do not differentiate in and of themselves between abstraction and exemplar effects, but they allow researchers to tune the degree of complexity of models incorporating both types of effects according to their theory and the phenomenon at hand.
+In Section~\ref{sec:corpushierarchicalmodel}, a full multilevel model will therefore be used.

### 2017-12-31 Latex/sections/3corpus.tex
Statistics final. More revisions in Sec. 3.

@@
 
-\subsection{Corpus choice}
+\subsection{Preliminiaries}
+
+\subsubsection{Corpus choice}
 \label{sec:gettingdata}
@@
 \footnote{The COW corpora (Dutch, English, French, German, Spanish, Swedish) are made available for free at \url{https://www.webcorpora.org}.
-At the time of this writing, a newer 2016 version DECOW16 has already been released.}
+At the time of this writing, a newer 2016 version DECOW16A has already been released.}
 I chose this corpus for two main reasons.%
@@
 Careless use of search engine results is still found, however, see for example \citet[171--175]{DeclerckBrems2016}.}
-First, the external validity of any study is increased through a higher heterogeneity of the sample \citep[30]{MaxwellDelaney2004}, and the DECOW corpus has clearly a much more heterogeneous composition compared to the only other very large corpus of German, the DeReKo \citep{KupietzEa2010} of the Institute for the German Language (IDS), which contains almost exclusively newspaper texts.%
-\footnote{It was shown in \cite{W16-2601} that, for example, the range of topics covered is much smaller in DeReKo compared to DECOW.}
+First, the external validity of any study is increased through a higher heterogeneity of the sample \citep[30]{MaxwellDelaney2004}, and the DECOW14A corpus has clearly a much more heterogeneous composition compared to the only other very large corpus of German, the DeReKo \citep{KupietzEa2010} of the Institute for the German Language (IDS), which contains almost exclusively newspaper texts.%
+\footnote{It was shown in \cite{W16-2601} that, for example, the range of topics covered is much smaller in DeReKo compared to DECOW14A.}
 Second, it was already mentioned that normative grammars often adopt clear positions regarding the grammaticality of either the \NACa\ or the \PGCa.
@@
 For these or similar reasons, COW corpora have been used in a number of peer-reviewed publications, for example \cite{VanGoethemHiligsmann2014}, \cite{VanGoethemHuening2015}, \cite{MuellerS2014}, \cite{Schaefer2016c}, \cite{SchaeferSayatz2014}, \cite{SchaeferSayatz2016}, and \cite{Zimmer2015}.
-Therefore, DECOW is a valid choice for this study.
+Therefore, DECOW14A is a valid choice for this study.
 
-\subsection{Pre-study: main prototype effects in the non-alternating constructions}
-
-\subsection{Main study: alternating constructions}
-
-\subsubsection{Sampling}
+\subsubsection{Bootstrapping pairs of lemmas}
+\label{sec:bootstrappinlemmapairs}
 
@@
 Therefore, it was highly desirable to obtain a sample in which most of the highly frequent actually-occurring combinations of kind nouns and measure nouns were represented.
-I applied a three-stage process in order to obtain such a sample, which consisted of the following steps:
+I applied a two-stage process in order to obtain a list of actually co-occurring measure nouns and kind nouns.
+First, I generate a list of the one hundred most frequent mass nouns,
+Second, I derived a list of all measure nouns with which the mass nouns co-occurred in the \NACb.
 
-\begin{enumerate}[i.]
-  \item\label{enum:data:step1} generating a list of the one hundred most frequent mass nouns,
-  \item\label{enum:data:step2} deriving a list of all measure nouns with which the mass nouns co-occur in the \NACb, and 
-  \item\label{enum:data:step3} sampling the target constructions by querying each combination of mass noun and measure noun found in step (\ref{enum:data:step2}).
-\end{enumerate}
-
-\vspace{-1\baselineskip}
-
-In step (\ref{enum:data:step1}), I exported a list of all nouns in the DECOW14A01 sub-corpus sorted by their token frequency and manually went through it from the most frequent noun downwards, selecting the first one hundred mass nouns that occurred in the list.%
+In the first step, I exported a list of all nouns in the DECOW14A01 sub-corpus sorted by their token frequency and manually went through it from the most frequent noun downwards, selecting the first one hundred mass nouns that occurred in the list.%
 \footnote{DECOW14A01 is the first slice (roughly a twentieth) of the complete DECOW14A corpus.
@@
 
-This list of mass nouns was used in step (\ref{enum:data:step2}) to derive a list of measure nouns co-occurring with the mass nouns. 
+This list of mass nouns was used in the second step to derive a list of measure nouns co-occurring with the mass nouns. 
 In order to generate this list, I utilised the fact that a direct sequence of two nouns almost always instantiates the bare-noun NAC if the second noun is a mass noun.
-Hence, I searched for all sequences N\Sub{1}N\Sub{2} where N\Sub{2} was one of the mass noun lemmas extracted in step (\ref{enum:data:step1}).
+Hence, I searched for all sequences N\Sub{1}N\Sub{2} where N\Sub{2} was one of the mass noun lemmas extracted in in the first step.
 Then, the resulting 100 lists of noun-noun combinations were each sorted by frequency in descending order and sieved manually to remove erroneous hits.
@@
 
-In step (\ref{enum:data:step3}), each of these 2,365 noun–noun combinations was queried in the target constructions (\PGCa\ and \NACa) individually in each of the first ten slices of DECOW (roughly 10 billion tokens).
-In order to reduce the sample size for the manual annotation process, the final concordance was sampled from the results of these 2,365 queries.
-Since the mass nouns in the sample were distributed according to the usual power law (often referred to as a \textit{Zipfian} distribution), I used all hits for nouns with a frequency up to 100 and a sample of 100 of all those with higher frequency.
-The final sample contained 6,843 sentences, which was reduced to 5,063 in the manual annotation process due to removal of noisy material, erroneous hits and uninformative cases where the measure noun was in the genitive, in which case the \NACa\ cannot be distinguished from the \PGCa.
-Given the careful sampling procedure described in this section, we can be highly sure that it contains all relevant and reasonably frequent noun–noun combinations in the target constructions.%
-\footnote{In a similar fashion, the 100 most frequent measure nouns occurring with plural kind nouns were listed and queried, resulting in a sample of 871 sentences.
-As stated in Section~\ref{sec:germanmeasurenps}, the \NACa\ is virtually never used with plural kind nouns, and this sample was not used except for quantifying the frequency of occurrence of the constructions (67 times \NACa\ and 794 times \PGCa).
-However, the sample is distributed with the data package accompanying this paper.
-}
+\subsubsection{Variables and annotation}
+\label{sec:variablesandannotation}
 
-Finally, two auxiliary samples were also drawn.
-As mentioned in Section~\ref{sec:analyses}, the distribution of the measure noun and kind noun lemmas in the \NACb\ and the \PGCd\ with a determiner will be modelled as factors influencing the alternation.
-Therefore, all noun-noun pairs from the process described above were also queried in the two non-alternating constructions, resulting in 17,252 hits for the \PGCd\ and 315,635 hits for the \NACb.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsubsection{Variables and annotation}
-\label{sec:annotation}
-
-The full set of manually annotated variables for the main sample is given in Table~\ref{tab:variables}, and I briefly discuss it now.%
-\footnote{All numeric variables were also z-transformed (\ie\ centered to the mean and rescaled such that they have a standard deviation of 1) to facilitate their interpretation in the regression models reported in the next section.}
-Notice first that \textit{Construction} is the response variable (or `dependent variable') with the values \textit{PGCa} and \textit{NACa}.
+The full set of manually annotated variables for the main study given in Table~\ref{tab:variables}.%
+\footnote{All numeric variables were z-transformed (\ie\ centered to the mean and rescaled such that they have a standard deviation of 1) to facilitate their interpretation in the regression models.}
+Notice first that \textit{Construction} is the response variable with the values \textit{PGCadj} and \textit{NACadj}.
 
@@
     Sentence       & Cardinal                         & factor  & Yes, No                   \\
-                   & \textbf{Construction (response)} & factor  & NACa, PGCa                \\
+                   & \textbf{Construction (response)} & factor  & NACadj, PGCadj            \\
                    & Measurecase                      & factor  & Nom, Acc, Dat             \\
     Kind lemma     & Kindattraction                   & numeric &                           \\
+                   & (Kindcollo)                      & numeric &                           \\
                    & Kindfreq                         & numeric &                           \\
+		   & Kindgender                       & factor  & MascNeut, Fem             \\
     Measure lemma  & Measureattraction                & numeric &                           \\
+                   & (Measurecollo)                   & numeric &                           \\
                    & Measureclass                     & factor  & Physical, Container,      \\
@@
   \end{tabular}
-  \caption{Annotated variables for the main sample}
+  \caption{Annotated variables for the main study}
   \label{tab:variables}
@@
 They were calculated from the auxiliary samples described at the end of Section~\ref{sec:gettingdata} as a log-transformed quotient.
-The higher the value, the more often the noun occurs in the \PGCd (proportionally).%
-\footnote{
-  It could be argued that some more advanced measure of attraction strength should be used, as is done in Collostructional Analysis \citep{StefanowitschGries2003,GriesStefanowitsch2004}, see also \cite{Gries2015a}.
-  Three main points speak against such an approach in the present case.
-  First, the goal here is to quantify how often lemmas occur in the \PGCd\ and the \NACb, and these constructions do not compete at all but are rather mutually exclusive.
-  Collostructional approaches are not made for such scenarios.
-  Second, the attraction values will be used as regressors in a hierarchical logistic regression and the values resulting from collostructional analysis, \ie\ logarithmised Fisher p-values, have a very unfavourable distribution in the case at hand.
-  They cluster around 0 and they include values of $-\infty$.
-  Third, I tried using collexeme strength as a regressor (with smoothing to remedy the mathematical problems), and the results were unsatisfactory compared to the simple quotient used here.
-}
-Additionally, \textit{Kindfreq} and \textit{Measurefreq} are the logarithm-transformed frequencies per 1,000,000 words of each lemma, extracted from the frequency lists distributed by the DECOW corpus creators on their web page.
-They were added to control for basic frequency effects.
-
+The higher the value, the more often the noun occurs in the \PGCd (proportionally).
+It could be argued that other measures of attraction strength could be used, for example those popularised in collostructional analysis (\citealp{StefanowitschGries2003,GriesStefanowitsch2004}; see also \citealp{Gries2015a}).
+However, the goal here is to quantify how often lemmas occur in the \PGCd\ and the \NACb, and these constructions do not compete at all but are rather mutually exclusive.
+While this does not preclude the use of collostructional analysis, it is an open question whether the marginals (usually called the ``expected frequencies'' in the collostructional literature) given the overall frequency of the constructions and the lemmas are cognitively relevant in this case.
+After all, the main difference between the quotient used here and the logarithmised p-values from a Fisher test is that they take these marginals into account. 
+However, using collexeme strength instead of the raw frequency quotient was tried as an alternative (variables \textit{Kindcollo} and \textit{Measurecollo}); see Sections~\ref{sec:corpushierarchicalmodel} and~\ref{sec:interpretation}.
+Additionally, \textit{Kindfreq} and \textit{Measurefreq} are the logarithm-transformed frequencies per 1,000,000 words of each lemma, extracted from the frequency lists distributed by the DECOW14A corpus creators on their web page.
 In Section~\ref{sec:analyses}, it was hypothesised that classes of measure lemmas might have different preferences for the two alternants.
-To capture this, class information was annotated for measure lemmas.
-The classification was inspired by the list in \citet[530]{Koptjevskaja2001} but due to the low frequencies of many of the potential classes, a very coarse classification was finally used.
-With typical examples and their frequencies in the final sample, the classes are:
-\textit{Physical} (abstract precisely measurable units such as \textit{Liter} `litre', \textit{Meter} `metre', \textit{Gramm} `gram'; \textit{f=}1,968),
-\textit{Container} (\textit{Eimer} `bucket'; \textit{f=}740),
-\textit{Amount} (\textit{Menge} `amount'; \textit{f=}1,364), 
-\textit{Portion} (natural portions like \textit{Happen} `bite' or \textit{Krümel} `crumb'; \textit{f=}713).
-The lemmas that did not fit into either of these classes were labelled \textit{Rest} (\textit{f=}278).
-
-The variable \textit{Cardinal} encodes whether the measure noun is modified by a cardinal (\textit{f=}1,939) or not (\textit{f=}3,124).
-The purpose of this variable is to test whether cardinals really favour the \NACa\ as hypothesised in Section~\ref{sec:analyses}.
+To capture this, class information was annotated for measure lemmas as \textit{Measureclass}.
+The variable \textit{Cardinal} encodes whether the MNP is modified by a cardinal.
 
 To capture the influence of style mentioned in Section~\ref{sec:analyses}, two proxy variables were used.
-At the document level, the DECOW corpus has an annotation for \textit{Badness}.
+At the document level, the DECOW14A corpus has an annotation for \textit{Badness}.
 As described in \cite{SchaeferEa2013}, \textit{Badness} measures how well the distribution of highly frequent short words in the document matches a pre-generated language model for German.
+They also show that the Badness score corresponds robustly with human raters' intuitions about text coherence and text quality (see the paper for an operationalisation of this notion).
 Documents with higher Badness usually contain more incoherent language, shorter sentences, etc.
-If the \PGCa\ actually favours higher stylistic levels, a high \textit{Badness} should be correlated with fewer occurrences.
+If the \PGCa\ actually favours more elaborate stylistic levels, a high \textit{Badness} should be correlated with fewer occurrences.
 Documents in DECOW14 have also been annotated with a variable called \textit{Genitives}.
 The higher the values of this variable, the lower the proportion of genitives among all case-bearing forms is.
-A high number of genitives is indicative of higher levels of style.
+A high number of genitives is also indicative of a more formal, elaborate style close to the written standard.
 However, the use of this variable as a regressor in the present study might be considered problematic.
 Since the \PGCa\ contains a genitive itself, the regressor variable \textit{Genitive} and the document-level variable \textit{Genitives} are not fully independent.
-Since instances of the \PGCa\ make up for only a minute fraction of all genitives, I still use \textit{Genitives} as a regressor with the appropriate caveats.
+Since instances of the \PGCa\ make up for only a minute fraction of all genitives, I still use \textit{Genitives} as a regressor.%
+\footnote{A reviewer suggested that a fully fledged multidimensional analysis \citep{Biber1988} would help to improve the operationalisation of the influence of style.
+This sounds plausible, but there simply is no large corpus of German for which similar data have been published.
+In the meantime, the creators of COW and the Institut für Deutsche Sprache (IDS) Mannheim have developed a similar annotation framework (COReX), and the COW creators have pre-released a non-public beta version of the corresponding data base for DECOW16A to their users.
+This data base contains 118 automatically extracted lexico-grammatical features for each document.
+As of today, it is not recommended for use in published research, however.
+Also, the genitive is generally considered a good indicator of style in German, and its frequency is usually high enough such that the \textit{Genitives} score is a stable measure even for shorter documents (compared to first or second person pronouns, for example, which are totally absent in a majority of documents).
+The genitive is still the dominant attributive case in German.}
 
-Finally, one variable was added as nuisance variable in the context of the present study.
-It was reported in the literature that MNPs in the dative and with a masculine or neuter kind noun favour the \PGCa\ more than the corresponding nominative and accusative MNPs \citep{Hentschel1993,Zimmer2015}.
+Finally, two variables were added as nuisance variables in the context of the present study.
+First, it was reported in the literature that MNPs in the dative and with a masculine or neuter kind noun favour the \PGCa\ more than the corresponding nominative and accusative MNPs \citep{Hentschel1993,Zimmer2015}.
 As an example, \textit{mit einem Stück frischen Brots} `with a piece of fresh bread' (\PGCa) would be preferred more strongly against \textit{mit einem Stück frischem Brot} (\NACa).
@@
 To control for this effect, the case of the measure noun was manually annotated (variable \textit{Measurecase}).
+Second, due to differences in case syncretisms, it is possible that feminine kind nouns have slightly different preferences than masculine and neuter ones, and the appropriate variable \textit{Kindgender} was annotated.
 
-\subsubsection{Model}
+\subsection{Pre-study: main prototype effects in the non-alternating constructions}
+
+The main study to be reported below deals exclusively with the alternating constructions, as is customary in alternation modelling.
+However, the prototypical features described in Section~\ref{sec:prototypeeffects} should be observable in the non-alternating cases as well if the theory laid out in Section~\ref{sec:analyses} is correct.
+Therefore, in this section I examine the distribution of the prototypical features in the non-alternating cases to see whether they are in accord with the theoretical predictions.
+
+For this pre-study, each of the 2,365 combinations of measure noun lemma and kind noun lemma were queried in the \NACb\ and the \PGCd.
+Each of the resulting 2,365 concordances was scaled down randomly to a size of maximally 100 sentences, and from the resulting 35,766 sentences, 5,000 were randomly sampled for the statistical analysis.
+All features described in Section~\ref{sec:variablesandannotation} were annotated except for the ones which do not apply in the non-alternating case (\textit{Kindattraction} and \textit{Measureattraction} and \textit{Measurecase}).
+The aggregated counts for each concordance were analysed in the form of a multilevel logistic regression model using R \citep{R} and the \textit{lme4} package \citep{Bates2010,BatesEa2015}.%
+\footnote{An alternative \textit{BOBYQA} optimiser from the \textit{nloptr} package \citep{Johnson2017} was used for all fits with \textit{lme4} reported in this paper.}
+
+\begin{table}
+  \centering
+  \resizebox{\textwidth}{!}{
+    \begin{tabular}{llrlrrrc}
+    Model level  & Regressor         & $\text{p}_{\text{PB}}$ & Factor level & Coefficient & CI low & CI high & CI excludes 0 \\
+    \midrule
+    First        & Badness           &                 &              &       &  &   &              \\
+                 & Cardinal          &                 & No           &       &  &   &              \\
+                 & Genitives         &                 &              &       &  &   &              \\
+
+    Second       & (Kindfreq)        &                 &              &       &  &   &              \\
+    (Kind)       & (Kindgender)      &                 &              &       &  &   &              \\[0.5\baselineskip]
+
+    Second       & Measureclass      &                 & Container    &       &  &   &              \\
+    (Measure)    &                   &                 & Rest         &       &  &   &              \\
+                 &                   &                 & Amount       &       &  &   &              \\
+                 &                   &                 & Portion      &       &  &   &              \\
+                 & (Measurefreq)     &                 &              &       &  &   &              \\
+
+  \end{tabular}
+  }
+  \caption{Coefficient table with 95\% bootstrap confidence intervals for the pre-study; the intercept is XXX}
+  \label{tab:smalltable}
+\end{table}
+
+The coefficient estimates are specified in Table~\ref{tab:smalltable} for each regressor (or regressor level) in the columns labelled \textit{Coefficient}.
+Given the coding of the response variable, coefficients leaning to the positive side can be interpreted as describing a context typical of the \PGCd.
+For a robust quantification of the precision of the estimation, I ran a parametric bootstrap (using the \mbox{\textit{confint.merMod}} function from \textit{lme4}) with 1,000 replications and using the percentile method for the calculation of the intervals.
+The resulting 95\% bootstrap confidence intervals are reported in Table~\ref{tab:smalltable} in the columns labelled \textit{CI low} and \textit{CI high} (= upper and lower 2.5th percentiles).
+The column \textit{CI excludes 0} shows an asterisk for those intervals that do not include 0.
+Furthermore, for each regressor, a p-value was obtained by dropping the regressor from the full model, re-estimating the nested model, and comparing it to the full model.
+Instead of inexact Wald approximations and Likelihood Ratio Tests, I used a drop-in bootstrap replacement for the Likelihood Ratio Test from the function \textit{PBmodcomp} from the \textit{pbkrtest} package \citep{HalekohHojsgaard2014}.
+I call the corresponding value $p_{\text{PB}}$, and it is given in the respective columns in Table~\ref{tab:smalltable}.
+Regressors which did not reach sig=0.05 in pbkrtest (\textit{Kindfreq}, \textit{Kindgender}, and \textit{Measurefreq}) were removed from the model, appear in brackets in Table~\ref{tab:smalltable}, and have no coefficient estimates.
+
+\subsection{Main study: the alternation}
+\label{sec:annotation}
 \label{sec:corpushierarchicalmodel}
 
- using R \citep{R}, \textit{lme4} \citep{lme4}
+For the main study, each of the 2,365 noun–noun combinations was queried in the alternating constructions \PGCa\ and \NACa\ in DECOW14A.%
+\footnote{In fact, due to processing considerations with the COW interface at the time, only ten slices of DECOW14A were used, which add up to approximately 10 billion tokens.}
+For the manual annotation process, a random sub-sample was generated.
+For each mass noun, the concordance was downsampled randomly to maximally 100 sentences, which resulted in 6,843 sentences without further downsampling.
+In the manual annotation process the size was further reduced to 5,063 sentences due to removal of noisy material, erroneous hits and uninformative cases where the measure noun was in the genitive, in which case the \NACa\ cannot be distinguished from the \PGCa.
+Given the careful sampling procedure described in this section, we can be highly sure that it contains all relevant and reasonably frequent noun–noun combinations in the target constructions.%
+\footnote{In a similar fashion, the 100 most frequent measure nouns occurring with plural kind nouns were listed and queried, resulting in a sample of 871 sentences.
+As stated in Section~\ref{sec:germanmeasurenps}, the \NACa\ is virtually never used with plural kind nouns, and this sample was not used except for quantifying the frequency of occurrence of the constructions (67 times \NACa\ and 794 times \PGCa).
+The sample is distributed with the data package accompanying this paper.}
+
+Finally, two auxiliary samples were also drawn.
+As mentioned in Section~\ref{sec:analyses}, the distribution of the measure noun and kind noun lemmas in the \NACb\ and the \PGCd\ with a determiner will be modelled as factors influencing the alternation.
+Therefore, all noun-noun pairs from the process described above were also queried in the two non-alternating constructions, resulting in 17,252 hits for the \PGCd\ and 315,635 hits for the \NACb.
 
@@
 
-In this section, I report the results of fitting a multilevel model to the alternation data.
-The purpose is to model the influence of the regressors specified in Table~\ref{tab:variables} on the probability that the \PGCa\ is chosen over the \NACa.
+Then, a multilevel logistic regression model was fit which models the influence of the regressors specified in Table~\ref{tab:variables} on the probability that the \PGCa\ is chosen over the \NACa.
 All regressors from Table~\ref{tab:variables} were included, and the measure lemma and the kind noun lemma were specified as varying-intercept random effects.
@@
 
-The regressors with the measure lemma as their unit of reference have no within-measure lemma variance, and the \textit{glmer} function automatically estimates them as \textit{group level predictors} (or \textit{second-level effects}), cf.\ \citet[265--269,302--304]{GelmanHill2006}.
+The regressors with the measure lemma as their unit of reference have no within-measure lemma variance, and the \textit{glmer} function automatically estimates them as \textit{group level predictors} (or \textit{second-level effects}), cf.\ \citet[265--269,302--304]{GelmanHill2006} and Section~\ref{sec:itemandexemplareffects}.
 The same goes for those listed with the kind lemma as their unit of reference.
@@
   }
-  \caption{Coefficient table with 95\% bootstrap confidence intervals; the intercept is -4.328}
+  \caption{Coefficient table with 95\% bootstrap confidence intervals for the main study; the intercept is -4.328}
   \label{tab:bigtable}
@@
 
-Standard diagnostics show that the model quality is quite good.
+Standard diagnostics show that the model quality is quite good.%
+\footnote{Notice that no aggressive model selection was applied, especially not upward model selection including probing for interactions.
+I am convinced that the model should be specified based on theoretical considerations (and known nuisance effects) to avoid the dangers of data dredging and to avoid models which are difficult to interpret.
+Removal of useless regressors (judging by, for example, the pbkrtest and coefficients of determination) is the only type of model selection I allow in my research.}
 Nakagawa \& Schielzeth's pseudo-coefficient of determination is $R_m^2=0.409$ and $R^2_c=0.495$ (see \citealp{Gries2015} for a basic introduction to these $R^2$ measures, or else \citealp{NakagawaSchielzeth2013}).
@@
 The lemma intercepts have standard deviations of $\sigma_{\text{Measurelemma}}=0.448$ and $\sigma_{\text{Kindlemma}}=0.604$.
+Only \textit{Kindfreq} ($\mpPB=0.095$) can be seen as slightly too high to be convincing, failing at sig=0.05.
 
-The coefficient estimates are specified in Table~\ref{tab:bigtable} for each regressor (or regressor level) in the columns labelled \textit{Coefficient}.
-For a robust quantification of the precision of the estimation, I ran a parametric bootstrap (using the \mbox{\textit{confint.merMod}} function from \textit{lme4}) with 1,000 replications and using the percentile method for the calculation of the intervals.
-The resulting 95\% bootstrap confidence intervals are reported in Table~\ref{tab:bigtable} in the columns labelled \textit{CI low} and \textit{CI high} (= upper and lower 2.5th percentiles).
-The column \textit{CI excludes 0} shows an asterisk for those intervals that do not include 0.
-Furthermore, for each regressor, a p-value was obtained by dropping the regressor from the full model, re-estimating the nested model, and comparing it to the full model.
-Instead of inexact Wald approximations and Likelihood Ratio Tests, I used a drop-in bootstrap replacement for the Likelihood Ratio Test from the function \textit{PBmodcomp} from the \textit{pbkrtest} package \citep{HalekohHojsgaard2014}.
-I call the corresponding value $p_{\text{PB}}$, and it is given in the respective columns in Table~\ref{tab:bigtable}.
-Only \textit{Kindfreq} ($\mpPB=0.095$) can be seen as slightly too high to be convincing (non-significant).
+Using collexeme strength (\textit{Measurecollo} and \textit{Kindcollo}) instead of the quotient for the attraction strength (\textit{Measureattraction} and \textit{Kindattraction}) was not successful.
+While the attraction measures reach satisfying sig levels, the $\text{p}_{\text{PB}}$ value for \textit{Kindcollo} was 0.191 and the one for \textit{Measurecollo} was 0.443.
+The coefficients of determination drop to $R_m^2=0.383$ and $R^2_c=0.490$.
 
-\subsubsection{Interpretation}
+\subsection{Interpretation}
 \label{sec:interpretation}
+
+%\hl{As collostructional analysis is relatively popular in some cognitively oriented areas of corpus linguistics, I tried using collexeme strength as a regressor (with smoothing to remedy the mathematical problems), and the results were unsatisfactory compared to the simple quotient used here.
+%The corresponding regressor did not even reach sig=0.1 in the pbkrtest.
+%Both the cognitive underpinnings and the math behind collostructional analysis are still much debated (see prominently \citealp{SchmidKuechenhoff2013,KuechenhoffSchmid2015}), I see this result }
 
@@
 Although \textit{Measurecase} is a nuisance variable in the context of this study, convergence with previous work strengthens its validity.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2017-12-31 R/0_wrapper.R

@@ -1,8 +1,5 @@
-#
-# This calls all scripts in the right order to replicate results from my paper
-# 'Competing Constructions for German Measure Phrases'
 
-# Scripts for the analysis of the experiments need to be run manually.
+# Scripts for the analysis of the experiments and the pre-study need to be run manually.
 
 rm(list = ls())
 
@@ -17,11 +14,8 @@ save.persistent        <- T                # Write to files instead of screen/co
 out.dir                <- "output/"        # Put generated files here.
 precision              <- 3                # Precision in result table.
 the.nAGQ               <- 0                # 0 for speed, 1 for precision and eytremely slow computation.
-ci.boot.nsim           <- 1000             # Even 100 can be extremely slow. Set to 1000 for production.
-ci.boot.modelcomp.nsim <- 1000             # This takes ages, even with 10. Set to 1000 for for production.
-chains                 <- 4                # MCMC.
-seed                   <- 6976             # MCMC.
-iter                   <- 1000             # MCMC. Set to 1000 for production.
+ci.boot.nsim           <- 1000            # Even 100 can be extremely slow. Set to 1000 for production.
+ci.boot.modelcomp.nsim <- 1000            # This takes ages, even with 10. Set to 1000 for for production.
 
 
 # Delete old output.
@@ -47,8 +41,8 @@ source('2_attraction.R') # calculate attraction factors
 cat("\n\n GLMM ...\n")
 source('3_glmm.R')       # full GLMMs
 
-cat("\n\n GLMM with MCMC ...\n")
-source('4_glmm-mcmc.R')  # same GLMMs estimated with Stan (Bayesian/MCMC)
+cat("\n\n GLMM with collexeme strength instead of attraction...\n")
+source('3a_glmm_collo.R')       # full GLMMs will collonoscopy instead of attraction quotient.
 
 cat("\n\n Prediction for experimental stimuli ...\n")
 source('5_stimuli_predict.R')

### 2017-12-31 R/1_load.R

@@ -18,7 +18,6 @@ measure <- measure[,c("Measurecase",
           "Measurelemma",
           "Measureclass",
           "Measurefreq",
-          "Measurenumber",
           "Minus1pos",
           "Genitives",
           "Badness",
@@ -80,7 +79,6 @@ fax  <- c("Measurecase",
           "Kindcase",
           "Measurelemma",
           "Measureclass",
-          "Measurenumber",
           "Leftcontext")
 
 for (fak in fax) {

### 2017-12-31 R/3_glmm.R

@@ -17,15 +17,12 @@ measure.glmm <- glmer(Construction~1
                  +Cardinal
                  +Genitives
                  +Measurecase
-                 +Measurenumber
 
                  +Kindattraction
-                 #+Kindcollo
                  +Kindfreq
                  +Kindgender
 
                  +Measureattraction
-                 #+Measurecollo
                  +Measureclass
                  +Measurefreq
 
@@ -56,7 +53,7 @@ measure.glmm.lemrand.names <- rownames(coef(measure.glmm)$Kindlemma)[measure.glm
 
 # PB test.
 bootcomp.regs <- c("(1|Measurelemma)", "(1|Kindlemma)", 
-                   "Badness", "Genitives", "Cardinal", "Measurecase", "Measurenumber",
+                   "Badness", "Genitives", "Cardinal", "Measurecase",
                    "Kindattraction", "Kindfreq", "Kindgender", 
                    "Measureattraction", "Measureclass", "Measurefreq")
 modelcomp <- lmer.modelcomparison(model = measure.glmm, regressors = bootcomp.regs,
@@ -87,7 +84,7 @@ if (save.persistent) sink()
 
 
 # Get the bootstrapped CIs.
-opts.ci.95 <- list(level = 0.95, method = "boot", boot.type = "perc", nsim = ci.boot.nsim, parallel="multicore", ncpus=8)
+opts.ci.95 <- list(level = 0.95, method = "boot", boot.type = "perc", nsim = ci.boot.nsim, parallel="multicore", ncpus=4)
 measure.ci.95 <- do.call(confint.merMod, c(opts.ci.95, list(object = measure.glmm, parm = names(fixef(measure.glmm)))))
 measure.ci.95 <- measure.ci.95[nrow(measure.ci.95):2,]
 
@@ -114,7 +111,7 @@ if (save.persistent) dev.off()
 
 
 # Effect plots.
-effs <- c("Badness", "Genitives", "Cardinal", "Measurecase", "Measurenumber",
+effs <- c("Badness", "Genitives", "Cardinal", "Measurecase", 
           "Kindattraction", "Kindfreq", "Kindgender", 
           "Measureattraction", "Measureclass", "Measurefreq")
 

### 2017-12-31 R/5_stimuli_predict.R

@@ -9,7 +9,6 @@ stims.fax  <- c(
   "Kindlemma",
   "Measurelemma",
   "Measureclass",
-  "Measurenumber",
   "Cardinal",
   "Kindgender"
   )

### 2017-12-31 R/7_exp-spr.R

@@ -27,6 +27,8 @@ data.files.names <- list.files(data.dir)
 data.files.names <- data.files.names[grep('csv$', data.files.names)]
 data.files       <- paste(data.dir, data.files.names, sep="")
 
+# Helpers.
+source('glmtools.R')
 
 load("output/corpus_stims_pred.RData")
 
@@ -160,13 +162,22 @@ if (save.persistent) sink()
 # Make formula and estimate model.
 # ################################
 
+# Overparametrised model with abundant random effects structure.
+formule.p1 <- paste(target, "~Construction*Modelprediction+Position+(1+Modelprediction|Participant)+(1|Item)", sep = "")
+formule.p2 <- paste(target, "~Construction*Modelprediction+Position+(1+Construction*Modelprediction|Participant)", sep = "")
+rt.model.p1 <- lmer(formule.p1, data = rt.all, REML = T,
+                 control = lmerControl(optimizer = "nloptwrap2", calc.derivs = FALSE))
+rt.model.p2 <- lmer(formule.p2, data = rt.all, REML = T,
+                    control = lmerControl(optimizer = "nloptwrap2", calc.derivs = FALSE))
+
 # Model with coprpus-based probs as input.
-#formule.rsl <- paste(target, "~Construction*Modelprediction+Position+(1+Construction*Modelprediction|Participant)", sep = "")
 formule <- paste(target, "~Construction*Modelprediction+Position+(1|Participant)+(1|Item)", sep = "")
 formule.0 <- paste(target, "~Construction+Position+(1|Participant)+(1|Item)", sep = "")
 #rt.model.rsl <- lmer(formule.rsl, data = rt.all, REML = T)
-rt.model <- lmer(formule, data = rt.all, REML = T)
-rt.model.0 <- lmer(formule.0, data = rt.all, REML = T)
+rt.model <- lmer(formule, data = rt.all, REML = T,
+                 control = lmerControl(optimizer = "nloptwrap2", calc.derivs = FALSE))
+rt.model.0 <- lmer(formule.0, data = rt.all, REML = T,
+                   control = lmerControl(optimizer = "nloptwrap2", calc.derivs = FALSE))
 
 
 # ######
@@ -175,13 +186,12 @@ rt.model.0 <- lmer(formule.0, data = rt.all, REML = T)
 
 if (save.persistent) sink(paste(out.dir, "results.txt", sep=""), append = T)
 
[…]

### 2017-12-31 R/3a_glmm_collo.R
Adding untracked files.

@@ -1,0 +1,86 @@
+require(lme4)
+require(gridExtra)
+require(MuMIn)
+require(effects)
+require(gridExtra)
+require(car)
+require(pbkrtest)
+require(lattice)
+
+measure$Kindcollo <- z.transform(measure$Kindcollo)
+measure$Measurecollo <- z.transform(measure$Measurecollo)
+
+measure.glmm.collo <- glmer(Construction~1
+                      
+                      +(1|Measurelemma)
+                      +(1|Kindlemma)
+                      
+                      +Badness                 
+                      +Cardinal
+                      +Genitives
+                      +Measurecase
+
+                      +Kindcollo
+                      +Kindfreq
+                      +Kindgender
+                      
+                      +Measurecollo
+                      +Measureclass
+                      +Measurefreq
+                      
+                      , data=measure, family=binomial(link=logit), na.action = na.fail, nAGQ=the.nAGQ,
+                      control=glmerControl(optimizer="nloptwrap2", optCtrl=list(maxfun=2e5)))
+
+
+# Standard diagnostics.
+measure.glmm.collo.wald <- Anova(measure.glmm.collo, type="II")
+measure.glmm.collo.r2 <- r.squaredGLMM(measure.glmm.collo)
+
+
+# Predict and analyze with optimal cutoff.
+all.cut           <- seq(-2,2,0.01)[which.max(unlist(lapply(seq(-2,2,0.01), function(x) {corr.prop(measure.glmm.collo, measure$Construction, x)})))]
+measure.glmm.collo.pred <- ifelse(predict(measure.glmm.collo) < all.cut, 0, 1)
+measure.glmm.collo.cmat <- table(measure.glmm.collo.pred, measure$Construction, dnn=c("Pred","Obs"))
+measure.glmm.collo.corr <- sum(diag(measure.glmm.collo.cmat))/sum(measure.glmm.collo.cmat)
+measure.glmm.collo.base <- length(which(measure$Construction == "NACa"))/length(measure$Construction)
+measure.glmm.collo.pre  <- (measure.glmm.collo.corr-measure.glmm.collo.base)/(1-measure.glmm.collo.base)
+
+
+# Extract random effects.
[…]

### 2017-12-31 Latex/article.tex
Proofread acknowledgements, Sec 1.

@@
   I thank (in alphabetical order) Felix Bildhauer, Susanne Flach, Elizabeth Pankratz, Samuel Reichert, Ulrike Sayatz, and Christian Zimmer for valuable discussions and comments.
-  Also, I would like to thank the reviewers for Cognitive Linguistics as well as associate editor Dagmar Divjak for insightful comments which helped to improve the quality of this paper a lot.
+  Also, I would like to thank the reviewers for Cognitive Linguistics as well as associate editor Dagmar Divjak for insightful comments which helped to improve the quality of this paper significantly.
   Furthermore, I thank Ulrike Sayatz for helping me to recruit the participants for the experiments.
   Elizabeth Pankratz thankfully also fixed my English.
-  Finally, I am grateful to my student assistants Kim Maser for her work on the annotation of the concordances and Luise Rißmann for supervising most of the experiments.
+  Finally, I am grateful to my student assistants Kim Maser, for her work on the annotation of the concordances, and Luise Rißmann, for supervising most of the experiments.
   The research presented here was made possible in part through funding from the \textit{Deutsche Forschungsgemeinschaft} (DFG, personal grant SCHA1916/1-1).

### 2017-12-31 Latex/sections/1intro.tex
Proofread acknowledgements, Sec 1.

@@
 
-This paper deals with a morpho-syntactic alternation between two constructions that occurs only in a very specific type of measure noun phrase in German.
+\hl{This paper deals with a morpho-syntactic alternation between two constructions that occurs only in a very specific type of measure noun phrase in German.}
 By \textit{alternation} I refer to a situation where two or more forms or constructions are available with no clear difference in acceptability, function, or meaning.
@@
 This question can and should be answered by testing how well corpus-derived models converge with or diverge from experimental findings, which provide more direct evidence of mental representations and processes but are often intrinsically narrower in scope.
-Second and much related to the first question, the appropriate modelling of such results in cognitive linguistics is a key issue.
+Second and closely related to the first question, the appropriate modelling of such results in cognitive linguistics is a key issue.
 
-As for the first point, there has for a long time been an interest in correlating probabilistic generalisations extracted from corpus data with results from experimental work (for example, \citealp{ArppeJaervikivi2007,BresnanEa2007,BresnanFord2010,DivjakGries2008,DivjakEa2016,FordBresnan2013}).
+Concerning the first point, there has for a long time been an interest in correlating probabilistic generalisations extracted from corpus data with results from experimental work (for example, \citealp{ArppeJaervikivi2007,BresnanEa2007,BresnanFord2010,DivjakGries2008,DivjakEa2016,FordBresnan2013}).
 This is often called a \textit{validation} of the corpus-derived findings, but \citet[303]{Divjak2016a} criticises this choice of words ``because it creates the impression that behavioral experimental data is inherently more valuable than textual data,'' citing \cite{TummersEa2005}, who state that a corpus is ``a sample of spontaneous language use that is (generally) realized by native speakers''.
 See also \citet{Newman2011} for a very positive view of corpora as a source of data in cognitive linguistics in their own right.
-However, as \citet[486--487]{Dabrowska2016} convincingly argues, this does not imply that we can in some way ``deduce mental representations from patterns of use,'' \ie\ from corpus data.%
-\footnote{Epistemologically, I think she refers to acts of \textit{induction} rather than \textit{deduction}.}
+However, as \citet[486--487]{Dabrowska2016} convincingly argues, this does not mean that we can in some way ``deduce mental representations from patterns of use,'' \ie\ from corpus data.%
+\footnote{Epistemologically, I believe she refers to acts of \textit{induction} rather than \textit{deduction}.}
 It would be highly surprising if this were possible, and the same holds for experimental methods, albeit to a different degree.
@@
 
-As mentioned above, the central question is whether usage data as found in corpora are truly predictive of speakers' and writers' cognitive representation of language and\slash or of their overall linguistic behaviour; and this is where the experimental validation (or, more neutrally \textit{corroboration}) comes into play.
-An overview of this issue was given by \cite{NewmanSorensenduncan2015}, who enumerate a number of studies which show how corpus data and experimental data converge (such as \citealp{BresnanEa2007,DurrantDoherty2010,GriesWulff2005,GriesEa2005}) and a number of studies where the two types of data did not lead to diverging or only partially converging results (such as \citealp{ArppeJaervikivi2007,Dabrowska2014,Mollin2009}).
+As mentioned above, the central question is whether usage data as found in corpora are truly predictive of speakers' and writers' cognitive representation of language and\slash or of their overall linguistic behaviour; and this is where the experimental validation (or, more neutrally, \textit{corroboration}) comes into play.
+An overview of this issue was given by \cite{NewmanSorensenduncan2015}, who enumerate a number of studies which show how corpus data and experimental data converge (such as \citealp{BresnanEa2007,DurrantDoherty2010,GriesWulff2005,GriesEa2005}) and a number of studies where the two types of data \hl{did not lead to diverging} or only partially converging results (such as \citealp{ArppeJaervikivi2007,Dabrowska2014,Mollin2009}).% converge = did not diverge.
+What do you mean by the second point?
 When researchers do not achieve convergence, they mostly try to explain this by differentiating between the actual cognitive construct and what the pooled usage data as found in corpora represent, for example \citet[411]{Dabrowska2014}, who lists a long number of possible reasons to explain why subjects in her experiment diverged in their word association preferences from collocation measures extracted from corpora.
@@
 See \citet{Biber1993}, \citet{MceneryEa2006}, \citet{Leech2007}, \citet{Hunston2008} for overviews from different perspectives.}
-In essence, this would entail that any generalisation extracted from the BNC can be attributed with some kind of mental reality.
+In essence, this would entail that any generalisation extracted from the BNC can be attributed \hl{with (to?)} some kind of mental reality.
+% "attributed with" doesn't mean anything.
+What do you mean to say here?
 
-Obviously, no clear picture has emerged yet, which is not surprising given the vast number of cognitive constructs assumed at diverse levels (such as the level of words in collocation research and much more complex constructional levels in alternation research involving lexical, syntagmatic, and contextual factors), the problems of corpus composition, the operationalisations involved in making experiments, and in the choice of statistical tools.
+Obviously, no clear picture has emerged yet, which is not surprising given the vast number of cognitive constructs assumed at diverse levels (such as the level of words in collocation research and much more complex constructional levels in alternation research involving lexical, syntagmatic, and contextual factors), the problems of corpus composition, \hl{the operationalisations involved in making experiments, and in the choice of statistical tools.} % Is the highlighted bit one coordinated unit?
+i.e.
+operationalisations involved in making experiments and choice of statistical tools?
+Then: "...the problems of corpus composition, and the operationalisations involved in making experiments and in the choice...".
+If not, take out the "in" fom "in the choice of statistical tools" and leave everything else the same.
 While far from providing definitive solutions, my discussion in Section~\ref{sec:conclusion} will provide possible explanations for the goodness of fit between the corpus data and the experimental data reported in Sections~\ref{sec:corpusstudies} and~\ref{sec:experimental}, much in the spirit of \citet{Dabrowska2014}.
@@
 The idea is that a variant is chosen when the influencing features cumulatively assume \textit{prototypical} values for that variant.
-I other words, a variant of prototype theory with features \citep{Rosch1978} suggests itself, and some researchers such as \cite{Gries2003}, \cite{NessetJanda2010}, or \cite{Schaefer2016c} indeed commit to prototype theory in alternation modelling.
+In other words, a variant of prototype theory with features \citep{Rosch1978} could be applied, and some researchers such as \cite{Gries2003}, \cite{NessetJanda2010}, or \cite{Schaefer2016c} indeed commit to prototype theory in alternation modelling.
 Under prototype theory, category membership is defined by similarity to an ideal exemplar, which is usually identified with characterising features (see \citealp{Taylor2008} for an overview).
@@
 
-Turning back to the prototype vs.\ schema debate, \cite{Barsalou1990} already showed that prototype and exemplar theory model the same types of effects and are informationally equivalent.
+Turning back to the prototype vs.\ \hl{schema} debate, \cite{Barsalou1990} already showed that prototype and exemplar theory model the same types of effects and are informationally equivalent.
+% Shouldn't this be prototype vs.
+exemplar?
+Since schemas are Langacker's thing and you don't mention them in this paragraph, but instead talk about exemplars?
 Consequently, experiments which favour one theory over the other use procedural behaviour of subjects in experiments, for example the speed of category retrieval, as opposed to mere output data.
@@
 It is thus an advantage of much linguistic work on categorisation that it deals with complex and realistically produced data, because this greatly improves the external validity of studies.
-An ideal contribution to the research on (levels of) category abstraction in the human mind by cognitive corpus linguists is thus if we provide analyses which have great external validity and complexity while carefully making sure whether and to which degree these finding correlate with reactions from cognitive agents under more controlled experimental conditions.
+An ideal contribution by cognitive corpus linguists to the research on (levels of) category abstraction in the human mind would thus be to provide analyses which have great external validity and complexity while carefully making sure that (and determining to what extent) these finding correlate with reactions from cognitive agents under more controlled experimental conditions.
 

### 2017-12-31 Latex/sections/1intro.tex
Removed one highlight.

@@
 
-\hl{This paper deals with a morpho-syntactic alternation between two constructions that occurs only in a very specific type of measure noun phrase in German.}
+This paper deals with a morpho-syntactic alternation between two constructions that occurs only in a very specific type of measure noun phrase in German.
 By \textit{alternation} I refer to a situation where two or more forms or constructions are available with no clear difference in acceptability, function, or meaning.

### 2017-12-31 Latex/sections/2descriptive.tex
Proofread.

@@
 I call the kind-denoting noun the \textit{kind noun} and the quantity-denoting noun the \textit{measure noun}.
-For illustration purposes, in the English \textit{a glass of good wine}, \textit{glass} is the measure noun and \textit{wine} is the kind noun.
+For illustration purposes, in the English phrase \textit{a glass of good wine}, \textit{glass} is the measure noun and \textit{wine} is the kind noun.
 Measure nouns can be all sorts of nouns which denote a quantity (such as \textit{litre} or \textit{amount}) but also those denoting containers, collections, etc.
 (such as \textit{glass} or \textit{bucket}).
-Like \citet[284]{Brems2003}, I consider nouns as measure nouns ``which, strictly speaking, do not designate a `measure', but display a more nebulous potential for quantification'' (also \citealp[530]{Koptjevskaja2001}, and \citealp[338]{Rutkowski2007}).
+Like \citet[284]{Brems2003}, I also consider nouns ``which, strictly speaking, do not designate a `measure', but display a more nebulous potential for quantification'' to be measure nouns (also \citealp[530]{Koptjevskaja2001}, and \citealp[338]{Rutkowski2007}).
 
@@
 
-Three different syntactic configurations within MNPs need to be distinguished, and the case alternation occurs only in one of them.
+Three different syntactic configurations within MNPs need to be distinguished, and case alternation occurs only in one of them.
 It occurs only when the kind noun is modified by an attributive adjective and there is no determiner, as in (\ref{ex:intro:alternation}).
@@
 With plural kind nouns, the obligatory marking of all dative plurals (except those where suffixing \textit{-en} is phonotactically impossible) with \textit{-en} might provide some clues.
-However, plural kind nouns do not behave like singular ones in measure phrases anyway, as will be argued in Section~\ref{sec:neighbouringcases}.
+However, plural kind nouns do not behave like singular ones in measure phrases, as will be argued in Section~\ref{sec:neighbouringcases}.
 Also, judgements vary between (\ref{ex:datpl1}) and (\ref{ex:datpl2}), and both are found in corpora.%
@@
 
-To summarise, the case patterns in the NAC and in the PGC (depending on the structure of the kind NP) is given in Table~\ref{tab:constructions}.
+To summarise, the case patterns in the NAC and in the PGC (depending on the structure of the kind NP) are given in Table~\ref{tab:constructions}.
 I call the narrow apposition construction with a bare kind noun the \NACb\ and the partitive genitive with a determiner in the kind NP the \PGCd.
@@
 I have demonstrated that there is an alternation between two measure noun constructions in a narrow syntactic configuration (kind NP with an adjective but without a determiner), and that the two constructions differ in the case of the kind noun (case agreement with the measure noun or genitive).
-I turn to more theory-oriented discussion of the alternation in the next section.
+I turn to a more theory-oriented discussion of the alternation in the next section.
 
@@
 I also review some existing analyses of the alternation and related issues.
-It should be remembered from Section~\ref{sec:syntax} that thanks to the syntactic ambiguity of strongly inflected adjectives, the \PGCa\ is plausibly a \PGCd\ in disguise and the \NACa\ a \NACb\ in disguise.
+It should be remembered from Section~\ref{sec:syntax} that, thanks to the syntactic ambiguity of strongly inflected adjectives, the \PGCa\ is plausibly a \PGCd\ in disguise and the \NACa\ a \NACb\ in disguise.
 
@@
 
-My analysis is based on the idea that the PGC prototype express a pseudo-partitive where two discernible entities -- the measure and the substance -- are referenced.
-The NAC prototype, on the other hand, expresses a mere quantity where the measure is not referenced as an entitiy in its own right.
+My analysis is based on the idea that the PGC prototype expresses a pseudo-partitive where two discernible entities -- the measure and the substance -- are referenced.
+The NAC prototype, on the other hand, merely expresses a quantity, and the measure is not referenced as an entity in its own right.
 Prototypical exemplars are given in (\ref{ex:prototypes1}) for the PGC and (\ref{ex:prototypes2}) for the NAC.
@@
 
-While the \PGCd\ in (\ref{ex:prototypes1}) allows an interpretation where speakers conceptualise the substance itself, \ie the medication, and a spoon used to take a quantity from the medication, the \NACb\ in (\ref{ex:prototypes2}) does not allow such an interpretation.
+While the \PGCd\ in (\ref{ex:prototypes1}) allows an interpretation where speakers conceptualise the substance itself, \ie\ the medication, and a spoon used to take a quantity from the medication, the \NACb\ in (\ref{ex:prototypes2}) does not allow such an interpretation.
 While in the given examples, the effect is strongly supported by the choice of the measure noun lemmas, I argue that the meaning of the prototypes is independent of this.%
@@
 In this type of construction, it is most obvious that two separate referents (in the given example the cake and the slices) are conceptualised.
-The \textit{part-of} meaning of true partitives as in \textit{a slice of the cake} represents the first stage of a development wherein the measure noun can already lose some semantic content, when for example words like \textit{bite} are no longer necessarily interpreted as a piece literally bitten out of something.
+The \textit{part-of} meaning of true partitives as in \textit{a slice of the cake} represents the first stage of a development wherein the measure noun can already lose some semantic content, when, for example, words like \textit{bite} are no longer necessarily interpreted as a piece literally bitten out of something.
 The pseudo-partitive stage finally instantiates a \textit{quantity-of} relation, potentially even leading to fully grammaticalised quantifiers such as \textit{a lot}.
@@
 As predicted by the grammaticalisation pattern just described, it still has the potential to form a true partitive (if the kind noun is definite).
-Conversely, the NAC lacks this ability to form true partitives, and has gone further down the grammaticalisation path.
+Conversely, the NAC lacks this ability to form true partitives and has gone further down the grammaticalisation path.
 It is thus not surprising that it has lost (at least prototypically) the semantics which allows both the measure and the substance to be conceptually accessible as independent referents.
@@
 In the corpus study, measure lemmas will therefore be annotated with appropriate semantic class labels to check whether semantic classes of measure nouns have different affinities to the two variants.
-The actual classification was inspired by the list in \citet[530]{Koptjevskaja2001}, but due to the low frequencies of many of the potential classes, a very coarse classification was finally used.
+The actual classification was inspired by the list in \citet[530]{Koptjevskaja2001}, but due to the low frequencies of many of the potential classes, a very coarse classification was used in the end.
 With typical examples, the classes are:
@@
 However, it is much more typical of quantities to be specified with numerical precision.
-This is most obviously so with the highly grammaticalised physical measure nouns like \textit{centilitre} which are very typically used with exact numerals instead of unspecific quantifiers, although both options are available in principle (\textit{three centilitres} vs.\ \textit{several centilitres}).
+This is most obviously so with the highly grammaticalised physical measure nouns like \textit{centilitre}, which are very typically used with exact numerals instead of unspecific quantifiers, although both options are available in principle (\textit{three centilitres} vs.\ \textit{several centilitres}).
 Since the \NACa\ is more closely associated with the \textit{quantity-of} relation, cardinals as attributes of the measure noun are expected to have a higher proportional frequency in the \NACa.
@@
 This is not surprising, given that the genitive -- an intrinsic part of the PGC -- is generally underrepresented in colloquial vernacular variants of German as a result of a diachronic process wherein many (but by no means all) uses of the genitive are being replaced by other cases or periphrastic constructions \citep{FleischerSchallert2011}.
-Under an integral view of prototypes, which incorporate effects related to larger contexts, registers, and styles, such preferences can be part of what defines the construction prototypes, and the \PGCa\ should occur proportionally more often in more elaborate styles closer to the standard.%
+Under an integral view of prototypes, which incorporates effects related to larger contexts, registers, and styles, such preferences can be part of what defines the construction prototypes, and the \PGCa\ should occur proportionally more often in more elaborate styles closer to the standard.%
 \footnote{A reviewer mentioned that she or he had the impression that dialectal variation is also a factor influencing the alternation.
-This is definitely true as some dialects (such as Alemannic) tend to have no genitive at all.
-While this is an interesting aspect for further research, the main obstacle for the present study is that there are no corpora of German which are both large enough and annotated with reliable meta data about regional variants.
+This is definitely true, as some dialects (such as Alemannic) tend to have no genitive at all.
+While this is an interesting aspect for further research, the main obstacle for the present study is that there are no corpora of German which are both large enough and annotated with reliable metadata about regional variants.
 In the annotation of the corpus study, documents obviously written in a regional variant were excluded.}
@@
 
-It should be noted that such and exemplar-type effect would have to be confirmed \textit{in addition} to the predicted semantic prototype effect for measure lemmas described in Section~\ref{sec:prototypeeffects}, namely the effect of semantic classes of measure nouns.
+It should be noted that such an exemplar-type effect would have to be confirmed \textit{in addition} to the predicted semantic prototype effect for measure lemmas described in Section~\ref{sec:prototypeeffects}, namely the effect of semantic classes of measure nouns.
 This means that the statistical model needs to track an abstraction effect and an exemplar-like effect for measure nouns.
@@
 
-With all this, it must also be made sure that these influences at the lemma level (lemma type frequency, attraction strength, and semantic classes) are not spurious and just artefacts of mere lemma idiosyncrasies.
+With all this, it must also be ensured that these influences at the lemma level (lemma type frequency, attraction strength, and semantic classes) are not spurious and just artefacts of mere lemma idiosyncrasies.
 This leads to a rather complex and truly multilevel model structure for the generalised linear mixed models (GLMMs) customarily used in alternation modelling.%
@@
 
-While multilevel models are rarely used in corpus linguistics, and \citet{Gries2015} even calls the simpler varying-intercepts and varying-slopes models ``underused'', they provide an excellent tool to describe situations where preferences at the sentence level and at lexical levels need to be integrated.
-The models do not differentiate in and of themselves between abstraction and exemplar effects, but they allow researchers to tune the degree of complexity of models incorporating both types of effects according to their theory and the phenomenon at hand.
+While multilevel models are rarely used in corpus linguistics -- \citet{Gries2015} even calls the simpler varying-intercepts and varying-slopes models ``underused'' -- they provide an excellent tool to describe situations where preferences at the sentence level and at lexical levels need to be integrated.
+The models do not differentiate in and of themselves between abstraction and exemplar effects, but they allow researchers to tune the degree of complexity of models, incorporating both types of effects according to their theory and the phenomenon at hand.
 In Section~\ref{sec:corpushierarchicalmodel}, a full multilevel model will therefore be used.

### 2017-12-31 Latex/sections/3corpus.tex
Proofread.

@@
 I applied a two-stage process in order to obtain a list of actually co-occurring measure nouns and kind nouns.
-First, I generate a list of the one hundred most frequent mass nouns,
+First, I generated a list of the one hundred most frequent mass nouns.
 Second, I derived a list of all measure nouns with which the mass nouns co-occurred in the \NACb.
@@
 In order to generate this list, I utilised the fact that a direct sequence of two nouns almost always instantiates the bare-noun NAC if the second noun is a mass noun.
-Hence, I searched for all sequences N\Sub{1}N\Sub{2} where N\Sub{2} was one of the mass noun lemmas extracted in in the first step.
+I therefore searched for all sequences N\Sub{1}N\Sub{2} where N\Sub{2} was one of the mass noun lemmas extracted in in the first step.
 Then, the resulting 100 lists of noun-noun combinations were each sorted by frequency in descending order and sieved manually to remove erroneous hits.
@@
 
-The full set of manually annotated variables for the main study given in Table~\ref{tab:variables}.%
+The full set of manually annotated variables for the main study is given in Table~\ref{tab:variables}.%
 \footnote{All numeric variables were z-transformed (\ie\ centered to the mean and rescaled such that they have a standard deviation of 1) to facilitate their interpretation in the regression models.}
@@
 They were calculated from the auxiliary samples described at the end of Section~\ref{sec:gettingdata} as a log-transformed quotient.
-The higher the value, the more often the noun occurs in the \PGCd (proportionally).
+The higher the value, the more often the noun occurs in the \PGCd\ (proportionally).
 It could be argued that other measures of attraction strength could be used, for example those popularised in collostructional analysis (\citealp{StefanowitschGries2003,GriesStefanowitsch2004}; see also \citealp{Gries2015a}).
 However, the goal here is to quantify how often lemmas occur in the \PGCd\ and the \NACb, and these constructions do not compete at all but are rather mutually exclusive.
-While this does not preclude the use of collostructional analysis, it is an open question whether the marginals (usually called the ``expected frequencies'' in the collostructional literature) given the overall frequency of the constructions and the lemmas are cognitively relevant in this case.
+While this does not preclude the use of collostructional analysis, it is an open question whether the marginals (usually called the ``expected frequencies'' in the collostructional literature), given the overall frequency of the constructions and the lemmas, are cognitively relevant in this case.
 After all, the main difference between the quotient used here and the logarithmised p-values from a Fisher test is that they take these marginals into account. 
@@
 Since the \PGCa\ contains a genitive itself, the regressor variable \textit{Genitive} and the document-level variable \textit{Genitives} are not fully independent.
-Since instances of the \PGCa\ make up for only a minute fraction of all genitives, I still use \textit{Genitives} as a regressor.%
+Since instances of the \PGCa\ make up for only a minute fraction of all genitives, however, I still use \textit{Genitives} as a regressor.%
 \footnote{A reviewer suggested that a fully fledged multidimensional analysis \citep{Biber1988} would help to improve the operationalisation of the influence of style.
@@
 This data base contains 118 automatically extracted lexico-grammatical features for each document.
-As of today, it is not recommended for use in published research, however.
+However, at this time, it is not recommended for use in published research.
 Also, the genitive is generally considered a good indicator of style in German, and its frequency is usually high enough such that the \textit{Genitives} score is a stable measure even for shorter documents (compared to first or second person pronouns, for example, which are totally absent in a majority of documents).
@@
 Each of the resulting 2,365 concordances was scaled down randomly to a size of maximally 100 sentences, and from the resulting 35,766 sentences, 5,000 were randomly sampled for the statistical analysis.
-All features described in Section~\ref{sec:variablesandannotation} were annotated except for the ones which do not apply in the non-alternating case (\textit{Kindattraction} and \textit{Measureattraction} and \textit{Measurecase}).
+All features described in Section~\ref{sec:variablesandannotation} were annotated except for the ones which do not apply in the non-alternating case (\textit{Kindattraction}, \textit{Measureattraction}, and \textit{Measurecase}).
 The aggregated counts for each concordance were analysed in the form of a multilevel logistic regression model using R \citep{R} and the \textit{lme4} package \citep{Bates2010,BatesEa2015}.%
@@
 For the main study, each of the 2,365 noun–noun combinations was queried in the alternating constructions \PGCa\ and \NACa\ in DECOW14A.%
-\footnote{In fact, due to processing considerations with the COW interface at the time, only ten slices of DECOW14A were used, which add up to approximately 10 billion tokens.}
+\footnote{Due to processing considerations with the COW interface at the time, only ten slices of DECOW14A were used, which add up to approximately 10 billion tokens.}
 For the manual annotation process, a random sub-sample was generated.
 For each mass noun, the concordance was downsampled randomly to maximally 100 sentences, which resulted in 6,843 sentences without further downsampling.
-In the manual annotation process the size was further reduced to 5,063 sentences due to removal of noisy material, erroneous hits and uninformative cases where the measure noun was in the genitive, in which case the \NACa\ cannot be distinguished from the \PGCa.
-Given the careful sampling procedure described in this section, we can be highly sure that it contains all relevant and reasonably frequent noun–noun combinations in the target constructions.%
+In the manual annotation process the size was further reduced to 5,063 sentences due to removal of noisy material, erroneous hits, and uninformative cases where the measure noun was in the genitive, in which case the \NACa\ cannot be distinguished from the \PGCa.
+Given the careful sampling procedure described in this section, we can be highly certain that it contains all relevant and reasonably frequent noun–noun combinations in the target constructions.%
 \footnote{In a similar fashion, the 100 most frequent measure nouns occurring with plural kind nouns were listed and queried, resulting in a sample of 871 sentences.
@@
 
-The style-related proxy variables point to the expected direction.
+The style-related proxy variables point in the expected direction.
 Increased \textit{Badness} of the document favours the \NACa\ ($\beta_{\text{Badness}}=-0.152$, $\mpPB=0.002$), and so does a lower density of genitives ($\beta=-0.693$, $\mpPB=0.001$).

### 2017-12-31 Latex/sections/1intro.tex
Finalising corrections in Sec. 1-3.

@@
 As mentioned above, the central question is whether usage data as found in corpora are truly predictive of speakers' and writers' cognitive representation of language and\slash or of their overall linguistic behaviour; and this is where the experimental validation (or, more neutrally, \textit{corroboration}) comes into play.
-An overview of this issue was given by \cite{NewmanSorensenduncan2015}, who enumerate a number of studies which show how corpus data and experimental data converge (such as \citealp{BresnanEa2007,DurrantDoherty2010,GriesWulff2005,GriesEa2005}) and a number of studies where the two types of data \hl{did not lead to diverging} or only partially converging results (such as \citealp{ArppeJaervikivi2007,Dabrowska2014,Mollin2009}).% converge = did not diverge.
-What do you mean by the second point?
+An overview of this issue was given by \cite{NewmanSorensenduncan2015}, who enumerate a number of studies which show how corpus data and experimental data converge (such as \citealp{BresnanEa2007,DurrantDoherty2010,GriesWulff2005,GriesEa2005}) and a number of studies where the two types of data led to diverging or only partially converging results (such as \citealp{ArppeJaervikivi2007,Dabrowska2014,Mollin2009}).
 When researchers do not achieve convergence, they mostly try to explain this by differentiating between the actual cognitive construct and what the pooled usage data as found in corpora represent, for example \citet[411]{Dabrowska2014}, who lists a long number of possible reasons to explain why subjects in her experiment diverged in their word association preferences from collocation measures extracted from corpora.
@@
 See \citet{Biber1993}, \citet{MceneryEa2006}, \citet{Leech2007}, \citet{Hunston2008} for overviews from different perspectives.}
-In essence, this would entail that any generalisation extracted from the BNC can be attributed \hl{with (to?)} some kind of mental reality.
-% "attributed with" doesn't mean anything.
-What do you mean to say here?
+In essence, this would entail that any generalisation extracted from the BNC could be assumed to have some kind of mental reality, which is at least doubtful.
 
-Obviously, no clear picture has emerged yet, which is not surprising given the vast number of cognitive constructs assumed at diverse levels (such as the level of words in collocation research and much more complex constructional levels in alternation research involving lexical, syntagmatic, and contextual factors), the problems of corpus composition, \hl{the operationalisations involved in making experiments, and in the choice of statistical tools.} % Is the highlighted bit one coordinated unit?
-i.e.
-operationalisations involved in making experiments and choice of statistical tools?
-Then: "...the problems of corpus composition, and the operationalisations involved in making experiments and in the choice...".
-If not, take out the "in" fom "in the choice of statistical tools" and leave everything else the same.
+Obviously, no clear picture has emerged yet, which is not surprising given the vast number of cognitive constructs assumed at diverse levels (such as the level of words in collocation research and much more complex constructional levels in alternation research involving lexical, syntagmatic, and contextual factors), the problems of corpus composition, the operationalisations involved in making experiments, and the choice of statistical tools.
 While far from providing definitive solutions, my discussion in Section~\ref{sec:conclusion} will provide possible explanations for the goodness of fit between the corpus data and the experimental data reported in Sections~\ref{sec:corpusstudies} and~\ref{sec:experimental}, much in the spirit of \citet{Dabrowska2014}.
@@
 
-Turning back to the prototype vs.\ \hl{schema} debate, \cite{Barsalou1990} already showed that prototype and exemplar theory model the same types of effects and are informationally equivalent.
-% Shouldn't this be prototype vs.
-exemplar?
-Since schemas are Langacker's thing and you don't mention them in this paragraph, but instead talk about exemplars?
+Turning back to the prototype vs.\ exemplar debate, \cite{Barsalou1990} already showed that prototype and exemplar theory model the same types of effects and are informationally equivalent.
 Consequently, experiments which favour one theory over the other use procedural behaviour of subjects in experiments, for example the speed of category retrieval, as opposed to mere output data.
@@
 
-On the other hand, the trade-off one has to accept when doing experiments with highly simplified stimuli and very simple tasks is their lack of \textit{external validity} (\ie their generalisability) and their high dependence on potentially problematic operationalisations of constructs, control of confounding factors in the face of a limited number of available subjects, etc.
+On the other hand, the trade-off one has to accept when doing experiments with highly simplified stimuli and very simple tasks is their lower \textit{external validity} (\ie\ their lower degree of generalisability) and their high dependence on potentially problematic operationalisations of constructs, control of confounding factors in the face of a limited number of available subjects, etc.
 (in other words, critical dependence on \textit{construct validity} and \textit{internal validity}).%
@@
 This paper contributes to this endeavour in many ways.
-After a thorough description of the alternation under examination, I discuss potential influencing factors comprising high-level schematic semantic generalisations as well as exemplar-similarity and item-specific effects in Section~\ref{sec:germanmeasurenps}.
+After a thorough description of the alternation under examination, I discuss potential influencing factors comprising high-level abstract semantic generalisations as well as exemplar-similarity and item-specific effects in Section~\ref{sec:germanmeasurenps}.
 I will argue that

### 2017-12-31 Latex/sections/2descriptive.tex
Finalising corrections in Sec. 1-3.

@@
 Even a full unacceptability of (\ref{ex:macabre2}) would not be conclusive, however, as a possible aversion of speakers towards the case-marked form might be due to the fact that it is at least potentially also a genitive, in which case the constraint against bare genitive nouns would apply.
-With plural kind nouns, the obligatory marking of all dative plurals (except those where suffixing \textit{-en} is phonotactically impossible) with \textit{-en} might provide some clues.
+With plural kind nouns, the obligatory marking of all dative plurals with \textit{-en} (except those where this is phonotactically impossible) might provide some clues.
 However, plural kind nouns do not behave like singular ones in measure phrases, as will be argued in Section~\ref{sec:neighbouringcases}.
@@
 This unusual double nature of adjectives in NPs without determiners leads to a plausible interpretation of the pattern shown in Table~\ref{tab:constructions}.
-Whenever speakers classify the adjective in the kind noun phrase more as a determiner, they have to use the \PGCa\ because, if there is a determiner, the PGC is the only option.
-When they classify the adjective more as an adjective, the kind NP has no determiner, and they have to use the \NACa.%
+Whenever speakers classify the adjective in the kind noun phrase more as a determiner, they use the \PGCa\ because, if there is a determiner, the PGC is the only option.
+When they classify the adjective more as an adjective, the kind NP has no determiner, and they use the \NACa.%
 \footnote{While the generative analysis presented in \cite{Bhatt1990} cannot properly deal with probabilistic effects, Bhatt comes close to this interpretation by analysing the kind NP in the PGC as a DP and in the NAC as an NP.}
@@
 For example, highly grammaticalised non-referential nouns like \textit{Gramm} `gram' and \textit{Meter} `metre' should occur proportionally more often in NAC constructions than in PGC constructions.
-If such preferences can actually shown in usage patterns, it would lend strong support for the hypothesised difference in the meanings of the prototypes.
+If such preferences can actually be shown in usage patterns, it would lend strong support for the hypothesised difference in the meanings of the prototypes.
 In the corpus study, measure lemmas will therefore be annotated with appropriate semantic class labels to check whether semantic classes of measure nouns have different affinities to the two variants.
-The actual classification was inspired by the list in \citet[530]{Koptjevskaja2001}, but due to the low frequencies of many of the potential classes, a very coarse classification was used in the end.
+The actual classification is based on the list in \citet[530]{Koptjevskaja2001}, but due to the low frequencies of many of the potential classes, a very coarse classification was used in the end.
 With typical examples, the classes are:
@@
 This is not surprising, given that the genitive -- an intrinsic part of the PGC -- is generally underrepresented in colloquial vernacular variants of German as a result of a diachronic process wherein many (but by no means all) uses of the genitive are being replaced by other cases or periphrastic constructions \citep{FleischerSchallert2011}.
-Under an integral view of prototypes, which incorporates effects related to larger contexts, registers, and styles, such preferences can be part of what defines the construction prototypes, and the \PGCa\ should occur proportionally more often in more elaborate styles closer to the standard.%
+Under an integral view of prototypes, which incorporates effects related to larger contexts and styles, such preferences can be part of what defines the construction prototypes, and the \PGCa\ should occur proportionally more often in more elaborate styles closer to the standard.%
 \footnote{A reviewer mentioned that she or he had the impression that dialectal variation is also a factor influencing the alternation.
@@
 As discussed in Section~\ref{sec:cogocl}, prototype-based and exemplar-based approaches are merely the endpoints of a spectrum of theories allowing abstraction in cognitive representations to varying degrees.
-While the prototypes for the PGC and the NAC were clearly specified with reference to abstract meaning in Section~\ref{sec:prototypeeffects}, at least one exemplar-driven similarity effect can also be expected to influence the MNP alternation.
+While the prototypes for the PGC and the NAC were clearly specified with reference to abstract meaning in Section~\ref{sec:prototypeeffects}, at least one exemplar-driven similarity effect can also be expected to influence the MNP alternation, leading to a mixed model with abstractions as well as exemplar effects.
 The measure and kind noun lemmas which occur in the alternating constructions obviously also occur in the non-alternating constructions.
@@
 For measure nouns, high frequency might lead to a higher affinity to the \NACa\ if we assume at least a tendency for highly grammaticalised items to be more frequent.
-Of course, the model must be specified in a way such that we can be sure that any detected frequency effect goes beyond the semantic effect, which is captured in the semantic classes described in Section~\ref{sec:prototypeeffects}.
+Of course, the model must be specified in a way such that we can be sure that any detected frequency effect also goes beyond the semantic effect, which is captured in the semantic classes described in Section~\ref{sec:prototypeeffects}.
 Such frequency effects would also be very difficult to describe as abstractions.
@@
 See \citet{GelmanHill2006} (especially part 2) for a comprehensive text book on the subject including multilevel models.}
-A GLMM models the influence which several variables (\textit{predictors}) have on the probability that an alternant is chosen.
+A GLMM models the influence which several variables (\textit{predictors}) have on the probability that an alternant is chosen (the \textit{response}).
 The so-called \textit{fixed effects} are assumed to be fixed population parameters which quantify the strength and the direction of the effect which, for example, a specific feature of the syntactic context or the stylistic level of the text have on the choice of the alternants.
@@
 In a true multilevel model, however, the varying intercept itself comes with an additional linear model where second-level fixed effects predict the group-level effect, and this is exactly what is needed here.
-To illustrate, the simple model specification in (\ref{eq:glmm}) is a multilevel logistic regression model.
+To illustrate, the simple model specification in (\ref{eq:glmm}) represents the first level of a multilevel logistic regression model.
 

### 2017-12-31 Latex/sections/3corpus.tex
Finalising corrections in Sec. 1-3.

@@
 
-I now turn to the sampling procedure applied to obtain concordances for manual annotation and statistical analysis.
 Among the factors potentially influencing the alternation (see Section~\ref{sec:germanmeasurenps}) were lemma-specific preference effects.
@@
 I applied a two-stage process in order to obtain a list of actually co-occurring measure nouns and kind nouns.
-First, I generated a list of the one hundred most frequent mass nouns.
+First, I generated a list of the one hundred most frequent genuine mass nouns.
 Second, I derived a list of all measure nouns with which the mass nouns co-occurred in the \NACb.
@@
 It contains just over one billion tokens.}
-Mass nouns were defined as concrete nouns which denote a substance in the broad sense, combine with uninflected mass quantifiers such as \textit{viel} `much' and \textit{wenig} `little' (\textit{viel Bier} `much beer'), and form only sortal and unit plurals (such as the plural \textit{Biere} `types of beer' or `glasses of beer').
+Mass nouns were defined as concrete nouns which denote a substance genuinely (without coercion), combine with uninflected mass quantifiers such as \textit{viel} `much' and \textit{wenig} `little' (\textit{viel Bier} `much beer'), and form only sortal and unit plurals (such as the plural \textit{Biere} `types of beer' or `glasses of beer').
 Abstract nouns which partially behave like mass nouns (like \textit{Spaß} `fun’ or \textit{Gefahr} `danger’) were excluded because they are usually not quantified in the same way as concrete mass nouns.
@@
 \subsection{Pre-study: main prototype effects in the non-alternating constructions}
+\label{sec:prestudy}
 
@@
 I call the corresponding value $p_{\text{PB}}$, and it is given in the respective columns in Table~\ref{tab:smalltable}.
-Regressors which did not reach sig=0.05 in pbkrtest (\textit{Kindfreq}, \textit{Kindgender}, and \textit{Measurefreq}) were removed from the model, appear in brackets in Table~\ref{tab:smalltable}, and have no coefficient estimates.
+Regressors which did not reach sig=0.05 in in the pbkrtest (\textit{Kindfreq}, \textit{Kindgender}, and \textit{Measurefreq}) were removed from the model, appear in brackets in Table~\ref{tab:smalltable}, and have no coefficient estimates.
 
@@
 In the manual annotation process the size was further reduced to 5,063 sentences due to removal of noisy material, erroneous hits, and uninformative cases where the measure noun was in the genitive, in which case the \NACa\ cannot be distinguished from the \PGCa.
-Given the careful sampling procedure described in this section, we can be highly certain that it contains all relevant and reasonably frequent noun–noun combinations in the target constructions.%
+Given the careful sampling procedure described in this section and Section~\ref{sec:bootstrappinlemmapairs}, we can be highly certain that it contains all relevant and reasonably frequent noun–noun combinations in the target constructions.%
 \footnote{In a similar fashion, the 100 most frequent measure nouns occurring with plural kind nouns were listed and queried, resulting in a sample of 871 sentences.

### 2017-12-31 Latex/sections/2descriptive.tex
Revisions in Sec.3 almost and in Sec.4 cmpltly done.

@@
 \begin{exe}
-%  \ex\label{ex:intro:pseudopartitive}
-%  \begin{xlist}
   \ex[ ]{\label{ex:intro:pseudopartitive1}\gll Wir trinken [[ein Glas]\Sub{Acc} [dieses Weins]\Sub{Gen}]\Sub{Acc}.\\
@@
   \glt We drink a glass of this wine.}
-%    \ex[*]{\label{ex:intro:pseudopartitive2} Wir trinken [[eine Tasse]\Sub{Acc} [einen leckeren Kaffee]\Sub{Acc}]\Sub{Acc}.}
-%  \end{xlist}
 \end{exe}
@@
 
-% namely:
-% 
-% \begin{enumerate}[i.]
-%   \item{\label{it:intro:idiot} alleged alternatives to the \NACa\ and the \PGCa,}
-% %  \item{\label{it:intro:datsg} alternative \textit{weak} forms of dative singular neuter adjectives,}
-% %  \item{\label{it:intro:femsg} case syncretism in feminine NPs,}
-%   \item{\label{it:intro:plurl} similar constructions with plural/collective kind nouns,}
-%   \item{\label{it:intro:noifl} grammaticalised non-inflected measure nouns, and}
-%   \item{\label{it:intro:other} alternative constructions for expressing quantities.}
-% \end{enumerate}
-%
-% \vspace{-1\baselineskip}
-
 I now finally turn to some more subtle issues related to the measure noun case alternation in order to delimit the scope of the study.
@@
 
-% As for (\ref{it:intro:datsg}), a complication with neuter kind nouns in the dative is mentioned by \citet[20--22]{Zimmer2015}.
-% In the \NACa, the adjective normally inflects with the so-called \textit{strong} case and number suffixes, which are used when no determiner with strong inflection is present in the NP.
-% \cite{Zimmer2015} reports a high number of occurrences of adjectives being inflected with the \textit{weak} inflectional marker, which are normally only used if a strongly inflected determiner precedes the adjective: \textit{kalt-en Wasser} `cold water' instead of \textit{kalt-em Wasser}.%
-% \footnote{See Section~\ref{sec:analyses} for more discussion of adjectival inflection.}
-% In my corpus sample, this tendency was not nearly as clear, and there was a high number of very noisy sentences among those potentially showing this pattern.
-% Hence, I do not discuss these forms.%
-% \footnote{All corpus data and scripts used for this paper will be released freely, so further examination of the few cases maybe showing this kind of inflection is possible.}
-% 
-% \label{page:femininesyncretism}
-% With feminine kind nouns, (\ref{it:intro:femsg}) is relevant.
-% Whereas the singular masculine and neuter nominal subsystem is a four case system (at least for NPs with adjectives and determiners), feminine singular NPs show nominative-accusative and dative-genitive syncretisms.
-% Table~\ref{tab:syncretisms} shows this as well as the partial syncretism in the plural, where no gender distinctions are made.
-% This means that with dative measure nouns, the alternation is unobservable if the kind noun is feminine, and it is problematic to speak of a \textit{genitive} kind noun construction in this case.
-% This is taken into account in the statistical modeling and the discussion presented in Section~\ref{sec:corpusstudies}.
-% 
-% \begin{table}
-%   \centering
-%   \begin{tabular}{llll}
-%      & Masculine\slash Neuter & Feminine & Plural \\
-%      \midrule
-%      nominative & rot-er Wein    & \multirow{2}{*}{frisch-e Sahne}   & \multirow{2}{*}{klein-e Äpfel} \\
-%      accusative & rot-en Wein    &                                   &                                \\
-%      dative     & rot-em Wein    & \multirow{2}{*}{frisch-er Sahne}  & klein-en Äpfel-n               \\
-%      genitive   & rot-en Wein-es &                                   & klein-er Äpfel                 \\
-%   \end{tabular}
-%   \caption{Case syncretisms in NPs with an adjective and without a determiner (\textit{roter Wein} `red wine', \textit{frische Sahne} `fresh cream', \textit{kleine Äpfel} `small apples')}
-%   \label{tab:syncretisms}
-% \end{table}
-
 Second, we find that, if the kind noun is a plural count noun as in \textit{ein Sack kleine Äpfel} (\NACa) or \textit{ein Sack kleiner Äpfel} (\PGCa) `a bag of small apples', a similar alternation between PGC and NAC can be observed.
 In line with experimental results reported in \citet[15--16]{Zimmer2015}, I found that the PGC is so dominant with plural kind nouns (794 of 861 cases, or over 92\%, cf.\ Section~\ref{sec:corpusstudies}) that the alternation cannot be analysed in the same way as in the singular.
-%\footnote{More concretely, if these cases were included in the regression analysis of the corpus data along with a factor encoding the number of the kind noun, this factor would most assuredly override any other regressor for the data points with a plural kind noun.}
 While this will play a role in the interpretation of the corpus findings, MNPs with plural kind nouns will not be included in the corpus study and the experiments reported in Sections~\ref{sec:corpusstudies} and \ref{sec:experimental}.
@@
 They cannot be used in normal MNPs.}
-%\footnote{With neglectable frequency, this effect is extended to container nouns like \textit{Glas} `glass', but with a twist.
-%A non-inflected version \textit{zwei Glas roter Wein} `two glasses of wine' and an inflected version \textit{zwei Gläser roter Wein} co-exist.
-%The non-inflected version denotes the quantity of wine fitting into two glasses, the inflected version denotes two glasses filled with wine.}
 I treat these cases like other measure nouns because they enter into both the \NACa\ as in (\ref{ex:uninflectedmeasures:a}) and the \PGCa\ as in (\ref{ex:uninflectedmeasures:b}).
@@
 I turn to a more theory-oriented discussion of the alternation in the next section.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2017-12-31 Latex/sections/3corpus.tex
Revisions in Sec.3 almost and in Sec.4 cmpltly done.

@@
     \midrule
-    First        & Badness           &                 &              &       &  &   &              \\
-                 & Cardinal          &                 & No           &       &  &   &              \\
-                 & Genitives         &                 &              &       &  &   &              \\
+    1                & Badness           & 0.042                  &              &  0.120      & -0.002 &  0.228  &              \\
+                     & Cardinal          & 0.001                  & No           &  1.419      &  0.869 &  1.993  &  *           \\
+                     & Genitives         & 0.001                  &              & -0.710      & -0.815 & -0.556  &  *           \\[0.5\baselineskip]
 
-    Second       & (Kindfreq)        &                 &              &       &  &   &              \\
-    (Kind)       & (Kindgender)      &                 &              &       &  &   &              \\[0.5\baselineskip]
+    2 (Kindlemma)    & (Kindfreq)        & (0.781)                &              &             &        &         &              \\
+                     & (Kindgender)      & (0.199)                &              &             &        &         &              \\[0.5\baselineskip]
 
-    Second       & Measureclass      &                 & Container    &       &  &   &              \\
-    (Measure)    &                   &                 & Rest         &       &  &   &              \\
-                 &                   &                 & Amount       &       &  &   &              \\
-                 &                   &                 & Portion      &       &  &   &              \\
-                 & (Measurefreq)     &                 &              &       &  &   &              \\
+    2 (Measurelemma) & Measureclass      & 0.005                  & Container    & 0.445       & -0.579 &  1.532  &              \\
+                     &                   &                        & Rest         & 1.745       &  0.649 &  2.831  &  *           \\
+                     &                   &                        & Amount       & 1.597       &  0.125 &  2.751  &  *           \\
+                     &                   &                        & Portion      & 1.782       &  0.771 &  2.690  &  *           \\
+                     & (Measurefreq)     & (0.265)                &              &             &        &         &              \\
 
@@
   }
-  \caption{Coefficient table with 95\% bootstrap confidence intervals for the pre-study; the intercept is XXX}
+  \caption{Coefficient table with 95\% bootstrap confidence intervals for the pre-study; the intercept is -5.370}
   \label{tab:smalltable}
@@
 The coefficient estimates are specified in Table~\ref{tab:smalltable} for each regressor (or regressor level) in the columns labelled \textit{Coefficient}.
-Given the coding of the response variable, coefficients leaning to the positive side can be interpreted as describing a context typical of the \PGCd.
+Given the coding of the response variable, coefficients leaning to the positive side can be interpreted as characterising a configuration typical of the \PGCd.
 For a robust quantification of the precision of the estimation, I ran a parametric bootstrap (using the \mbox{\textit{confint.merMod}} function from \textit{lme4}) with 1,000 replications and using the percentile method for the calculation of the intervals.
@@
 I call the corresponding value $p_{\text{PB}}$, and it is given in the respective columns in Table~\ref{tab:smalltable}.
-Regressors which did not reach sig=0.05 in in the pbkrtest (\textit{Kindfreq}, \textit{Kindgender}, and \textit{Measurefreq}) were removed from the model, appear in brackets in Table~\ref{tab:smalltable}, and have no coefficient estimates.
+Regressors which did not reach sig=0.05 in in the PBmodcomp test (\textit{Kindfreq}, \textit{Kindgender}, and \textit{Measurefreq}) were removed from the model, appear in brackets in Table~\ref{tab:smalltable}, and have no coefficient estimates.
+The overall intercept is -5.370 and reoresents \textit{Cardinal=Yes}, \textit{Measureclass=Physical} and 0 for all (z-transformed) numeric regressors.
+The standard deviation in the random intercepts is 1.157 for \textit{Measurelemma} ($p_{\text{PB}}=0.001$) and 0.923 for \textit{Kindlemma} ($p_{\text{PB}}=0.0.002$).
+Nakagawa \& Schielzeth's pseudo-coefficient of determination is $R_m^2=0.278$ and $R^2_c=0.566$ (see \citealp{Gries2015} for a basic introduction to these $R^2$ measures, or else \citealp{NakagawaSchielzeth2013}).
+This means that the effect of the fixed effects (marginal $R^2$) is weak to intermediate, and that lemma-specific effects (conditional minus marginal $R^2$) are equally strong with a difference between the two coefficients of 0.289.
+
+This show that the main prototypical factors \textit{Cardinal} and \textit{Measureclass} indeed play the predicted role
+The abstract non-referential physical measure nouns (such as \textit{Gramm} `gram' or \textit{Liter} `litre') with a high degree of grammaticalisation favour the \NACb.
+At the other end of the scale, nouns denoting natural portions like \textit{Haufen} `heap', \textit{Bündel} `bundle', \textit{Schluck} `gulp' favour the \PGCd.
+Also, The presence of a cardinal modifier clearly favours the \NACb.
+The stylistic factor \textit{Genitives} is confirmed to have an influence as predicted inasmuch as a general lack of genitives (which is encoded as a \textit{higher} value of the variable) favours the \NACb.
+\textit{Badness} points into the wrong direction, but it barely reaches sig=0.05, and its confidence interval includes 0.
+The lemma frequency regressors both clearly fail the PBmodcomp test.
+I will return to these results in Section~\ref{sec:interpretation}.
 
@@
 The intercept comprises \textit{Cardinal=Yes}, \textit{Measurecase=Nom}, \textit{Kindgender=Masc}, \textit{Measureclass=Physical}, and 0 for all numeric z-transformed regressors.
-It was estimated at -4.328.
+It was estimated at -3.548.
 
-The regressors with the measure lemma as their unit of reference have no within-measure lemma variance, and the \textit{glmer} function automatically estimates them as \textit{group level predictors} (or \textit{second-level effects}), cf.\ \citet[265--269,302--304]{GelmanHill2006} and Section~\ref{sec:itemandexemplareffects}.
+The regressors with the measure lemma as their unit of reference have no within-measure lemma variance, and the \textit{glmer} function automatically estimates them as group-level predictors (or second-level effects), cf.\ \citet[265--269,302--304]{GelmanHill2006} and Section~\ref{sec:itemandexemplareffects}.
 The same goes for those listed with the kind lemma as their unit of reference.
@@
     \midrule
-    First        & Badness           &  0.002                 &              & -0.152      & -0.247 & -0.061  & *             \\
+    1               & Badness           &  0.002                 &              & -0.152      & -0.247 & -0.061  & *             \\
                     & Cardinal          &  0.001                 & No           &  1.189      &  0.862 &  1.466  & *             \\
@@
 
-    Second       & Kindattraction    &  0.020                 &              &  0.225      &  0.049 &  0.393  & *             \\
-    (Kind)       & Kindfreq          &  0.095                 &              &  0.146      & -0.023 &  0.301  &               \\
+    2 (Kindlemma)   & Kindattraction    &  0.020                 &              &  0.225      &  0.049 &  0.393  & *             \\
+                    & Kindfreq          &  0.095                 &              &  0.146      & -0.023 &  0.301  &               \\
                     & Kindgender        &  0.001                 & Neut         &  0.021      & -0.367 &  0.392  &               \\
@@
 
-    Second       & Measureattraction &  0.001                 &              &  0.282      &  0.106 &  0.447  & *             \\
-    (Measure)    & Measureclass      &  0.001                 & Container    &  0.252      & -0.265 &  0.788  &               \\
+    2 (Measurelemma & Measureattraction &  0.001                 &              &  0.282      &  0.106 &  0.447  & *             \\
+                    & Measureclass      &  0.001                 & Container    &  0.252      & -0.265 &  0.788  &               \\
                     &                   &                        & Rest         &  0.421      & -0.209 &  1.063  &               \\
@@
   }
-  \caption{Coefficient table with 95\% bootstrap confidence intervals for the main study; the intercept is -4.328}
+  \caption{Coefficient table with 95\% bootstrap confidence intervals for the main study; the intercept is -3.548}
   \label{tab:bigtable}
@@
 I am convinced that the model should be specified based on theoretical considerations (and known nuisance effects) to avoid the dangers of data dredging and to avoid models which are difficult to interpret.
-Removal of useless regressors (judging by, for example, the pbkrtest and coefficients of determination) is the only type of model selection I allow in my research.}
-Nakagawa \& Schielzeth's pseudo-coefficient of determination is $R_m^2=0.409$ and $R^2_c=0.495$ (see \citealp{Gries2015} for a basic introduction to these $R^2$ measures, or else \citealp{NakagawaSchielzeth2013}).
+Removal of useless regressors (judging by, for example, the PBmodcomp test and coefficients of determination) is the only type of model selection I allow in my research.}
+Nakagawa \& Schielzeth's pseudo-coefficient of determination is $R_m^2=0.409$ and $R^2_c=0.495$.
 The rate of correct predictions is 0.843, which means a proportional reduction of error of $\lambda=0.297$.
-Generalised variance inflation factors for the regressors were calculated to check for multicollinearity \citep{FoxMonette1992,ZuurEa2010}, and none of the corrected $\text{GVIF}^{1/2\text{df}}$ was higher than 1.6.
+Generalised variance inflation factors for the regressors were calculated to check for multicollinearity \citep{FoxMonette1992,ZuurEa2010}, and the highest corrected $\text{GVIF}^{1/2\text{df}}$ was 1.520 for \textit{Cardinal}.
 The lemma intercepts have standard deviations of $\sigma_{\text{Measurelemma}}=0.448$ and $\sigma_{\text{Kindlemma}}=0.604$.
-Only \textit{Kindfreq} ($\mpPB=0.095$) can be seen as slightly too high to be convincing, failing at sig=0.05.
+Only \textit{Kindfreq} ($\mpPB=0.095$) could be seen as slightly too high to be convincing, failing at sig=0.05.
 
-Using collexeme strength (\textit{Measurecollo} and \textit{Kindcollo}) instead of the quotient for the attraction strength (\textit{Measureattraction} and \textit{Kindattraction}) was not successful.
-While the attraction measures reach satisfying sig levels, the $\text{p}_{\text{PB}}$ value for \textit{Kindcollo} was 0.191 and the one for \textit{Measurecollo} was 0.443.
-The coefficients of determination drop to $R_m^2=0.383$ and $R^2_c=0.490$.
+Using signed logarithmised collexeme strength (\textit{Measurecollo} and \textit{Kindcollo}) (with smoothing to avoid log(0) operations) instead of the quotient for the attraction strength (\textit{Measureattraction} and \textit{Kindattraction}) (see Section~\ref{sec:variablesandannotation}) was not successful.
+While the attraction measures reach satisfying sig levels in the PBmodcomp test (0.020 for \textit{Kindattraction} and 0.001 for \textit{Measureattraction}), the $p_{\text{PB}}$ value for \textit{Kindcollo} was 0.191 and the one for \textit{Measurecollo} was 0.443.
+The coefficients of determination drop to $R_m^2=0.135$ and $R^2_c=0.636$.
 
@@
 %\hl{As collostructional analysis is relatively popular in some cognitively oriented areas of corpus linguistics, I tried using collexeme strength as a regressor (with smoothing to remedy the mathematical problems), and the results were unsatisfactory compared to the simple quotient used here.
-%The corresponding regressor did not even reach sig=0.1 in the pbkrtest.
+%The corresponding regressor did not even reach sig=0.1 in the PBmodcomp test.
 %Both the cognitive underpinnings and the math behind collostructional analysis are still much debated (see prominently \citealp{SchmidKuechenhoff2013,KuechenhoffSchmid2015}), I see this result }

### 2017-12-31 Latex/sections/4experiments.tex
More reviewer-induced changes to Sec. 1.
Minor progess w/ revision.
Revisions in Sec.3 almost and in Sec.4 cmpltly done.

@@
 In \cite{VerhoevenTemme2017}, it was shown that results from forced-choice and split-100 experiments mostly converge (with some numeric intricacies related to the non-linear distribution of preferences for alternants).
-I present a forced-choice experiment in this Section and not a split-100 experiment, mainly because in a dry run of the experiment, participants complained about the unnaturalness of distributing a probability mass across two alternants and tended to produce ratings of 0 and 100 (and to a lesser degree 50).
+I present a forced-choice experiment in this Section instead of a split-100 experiment not only because of some researchers' critical attitude towards the split-100 task, but also because in a dry run of the experiment, participants complained about the unnaturalness of distributing a probability mass across two alternants and tended to produce ratings of 0 and 100 (and to a lesser degree 50).
 Participants had to choose between two sentences that differed only in that one contained the \NACa\, and the other one contained the \PGCa.
@@
 Furthermore, in each of the masculine\slash neuter and feminine groups, four sentences originally containing the \NACa\ and four sentences originally containing the \PGCa\ were chosen.
-More precisely, the sentences were sampled as \textit{highly prototypical examples} of \PGCa\ (high probability assigned by GLMM) and \NACa\ (low probability assigned by GLMM), respectively.%
+More precisely, the sentences were sampled as \textit{highly typical examples} of \PGCa\ (high probability assigned by GLMM) and \NACa\ (low probability assigned by GLMM), respectively.%
 \footnote{Remember from Section~\ref{sec:corpusstudies} that the model predicts the probability that the \PGCa\ is chosen over the \NACa.}
@@
 
-\begin{figure}[hb!]
+\begin{figure}[htbp!]
 \centering
@@
 A random intercept and slope were added for the individual sentence (item) in order to catch idiosyncrasies of single sentences.
-Also, a random intercept and slope for participants was added.%
-\footnote{The random slopes were added to comply with \citet[257]{BarrEa2013} who predict \textit{catastrophically high Type I error rates} for experimental designs with within-subject manipulations if random effects structures are not kept maximal.
-Notice that the model reported here was estimated with conceptually identical results with regard to the predictor of interest (\textit{Modelprediction}) if only random intercepts were used.}
 Coefficients were estimated with Maximum Likelihood Estimation (\textit{lmer} function from \textit{lme4}).
@@
 
-A certain amount of the variance can be accounted for by idiosyncrasies of single sentences ($\sigma_{\text{Sentence}}=1.785$, $\sigma^{\text{Modelprediction}}_{\text{Sentence}}=5.996$, 16 levels).%
-\footnote{I use $\sigma^f_r$ to denote the standard deviation of the random intercepts for the fixed effect $f$ varying by random effect $r$.}
-Also, among participants, there are clearly different preferences ($\sigma_{\text{Participant}}=0.781$, $\sigma^{\text{Modelprediction}}_{\text{Participant}}=0.484$, 24 levels).
+\begin{figure}[htbp!]
+  \centering
+  \includegraphics[width=0.5\textwidth]{../R/output/fc_proportions}
+  \caption{Spineplot of the proportion of responses plotted against the predictions from the corpus-based model in the forced-choice experiment}
+  \label{fig:spines}
+\end{figure}
+
+A good amount of the variance can be accounted for by idiosyncrasies of single sentences ($\sigma_{\text{Item}}=1.217$).%
+Also, among participants, there are clearly different preferences ($\sigma_{\text{Participant}}=0.412$).
 On the extreme ends, one participant chose the \PGCa\ in 13 of 16 cases, and two participants only chose it in 5 of 16 cases.
-The regressor \textit{Modelprediction} achieves $\mpPB=0.007$ (1,000 replications) and is estimated at 5.408 relative to an intercept of -1.304.
-The confidence interval from a parametric bootstrap (1,000 replications, percentile method) for the regressor is acceptable but slightly large with a lower bound of 1.626 and an upper bound of 8.397.
-The pseudo-coefficients of determination are $R^2_{m}=0.227$ and $R^2_{c}=0.561$, which means that over 22\% of the variance in the data can be explained by considering only the predictions from the corpus-based GLMM.
+The regressor \textit{Modelprediction} achieves $\mpPB=0.003$ (1,000 replications) and is estimated at 4.389 relative to an intercept of -1.270.
+The confidence interval from a parametric bootstrap (1,000 replications, percentile method) for the regressor is acceptable but slightly large with a lower bound of 1.788 and an upper bound of 6.599.
+The pseudo-coefficients of determination are $R^2_{m}=0.185$ and $R^2_{c}=0.455$, which means that roughly 19\% of the variance in the data can be explained by considering only the predictions from the corpus-based GLMM.
+This is a weak result for the fixed effects part, but clearly worthy of mention.
 The effect display for the single fixed regressor \textit{Modelprediction} is given in Figure \ref{fig:afc:effects}.
-The result is very clear.
 The higher the probability of the \PGCa\ predicted from usage data, the more often participants chose the \PGCa\ alternant in the forced-choice task.
-In summary, the forced-choice experiment clearly succeeded in corroborating the results from the corpus study in as much as the preferences extracted from usage data correspond to native speakers' choices.
+A closer look at the results in the form of the spineplot in Figure~\ref{fig:spines} shows, however, that it was likely idiosyncracies in the sentences with a model prediction between 0.5 and 0.6 which spoiled an otherwise much better correlation.
+Clearly, a revised and improved experiment might lead to a much better fit in future research.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+In principle, a per-stimulus random effect for \textit{Modelprediction} could also remedy the problem with individual stimuli at least partially.
+Therefore, a model with random slopes for \textit{Modelprediction} and both random effects (\textit{Participant} and \textit{Item}) was specified and estimated.
+The random slope for participants was added to comply with \citet[257]{BarrEa2013} who predict \textit{catastrophically high Type I error rates} for experimental designs with within-subject manipulations if random effects structures are not kept maximal.
+The coefficient of the fixed effect changed noticeably but not enough to change the interpretation (5.408 relative to an intercept of -1.304), and the marginal $R^2_m$ rises to 0.213 ($R^2_c=0.488$).
+In line with expectations, the standard deviation in the random slopes for \textit{Item} is high at 5.996.
+However, the covariance parameters were estimated at -1.0, which is a clear sign that the variance-covariance matrix could not be estimated successfully.
+The same was true for models with only \textit{Item} and \textit{Participant} random slopes.
+This is exactly the kind of model overparametrisation criticised in \citet{BatesEa2015a} and \citet{MatuschekEa2017}.
+The available data is simply insufficient to estimate the parameters of the more complex model with varying slopes.%
+\footnote{\citet[1]{BatesEa2015a} state: ``We show that failure to converge typically is not due to a suboptimal estimation algorithm, but is a consequence of attempting to fit a model that is too complex to be properly supported by the data, irrespective of whether estimation is based on maximum likelihood or on Bayesian hierarchical modeling with uninformative or weakly informative priors.
+Importantly, even under convergence, overparameterization may lead to uninterpretable models.''
+}
+
+In summary, the forced-choice experiment succeeded in corroborating the results from the corpus study in as much as the preferences extracted from usage data correspond to native speakers' choices, but the correlation is weak, likely due to problems with individual test items and\slash or too few data.
 
@@
 The second experiment tests preferences more implicitly.
-It is expected that reading less prototypical alternants (the one assigned a low probability by the corpus-derived model) in a given context and with given lexical material incurs a processing overhead for the reader (\citealp{Kaiser2013}).
+It is expected that reading less typical alternants (the one assigned a low probability by the corpus-derived model) in a given context and with given lexical material incurs a processing overhead for the reader (\citealp{Kaiser2013}).
 In this section, a self-paced reading experiment is therefore presented.
@@
 It should be remembered that the corpus GLMM predicts the probability of the \PGCa.
-As a consequence, the higher the GLMM prediction is, the more prototypical the sentence is for containing the \PGCa.
+As a consequence, the higher the GLMM prediction is, the more typical the sentence is for containing the \PGCa.
 It is therefore expected that reading times are higher when the value of \textit{Modelprediction} is higher but the sentence contains the \NACa.
 However, when the sentence contains the \PGCa, reading times should be lower when \textit{Modelprediction} is higher.
-To account for this, an interaction between \textit{Modelprediction} and \textit{Construction} (levels \textit{PGCa} and \textit{NACa}) was added to the model.
+To account for this, an interaction between \textit{Modelprediction} and \textit{Construction} (levels \textit{PGCadj} and \textit{NACadj}) was added to the model.
 
@@
 Random intercepts were specified for \textit{Participant} and \textit{Item} (the 16 sentence pairs are one \textit{Item} each).%
-\footnote{I tried random slopes in order to keep the random effect structure maximal \citep{BarrEa2013} but it was impossible to get the algorithm to converge due to the added complexity of the interaction.}
+\footnote{Again, all attempts to include random slopes resulted in the variance-covariance matrix not being properly estimated (1 or -1 covariance parameters).}
 
@@
     Regressor & \multicolumn{1}{r}{Coefficient} & \multicolumn{1}{r}{CI low} & \multicolumn{1}{r}{CI high} & \multicolumn{1}{r}{CI excludes 0} \\ \midrule
-    ConstructionPGCa                 &  0.054 &  0.012 &  0.095 &  *  \\ 
-    Modelprediction                  & -0.006 & -0.113 &  0.110 &     \\ 
+    Construction=PGCadj                 &  0.054 &  0.012 &  0.095 &  *  \\ 
+    Modelprediction                     & -0.003 & -0.113 &  0.110 &     \\ 
     Position                            & -0.005 & -0.005 &  0.004 &     \\ 
-    ConstructionPGCa:Modelprediction & -0.125 & -0.234 & -0.023 &  *  \\ 
+    Construction=PGCadj:Modelprediction & -0.125 & -0.234 & -0.023 &  *  \\ 
   \end{tabular}
@@
 Comparing the full model to a model without the main regressor \textit{Modelprediction} (and consequently also without the interaction with \textit{Construction}) in a PB test gives $\mpPB=0.036$.
-The pseudo-coefficients of determination are $R^2_m=0.237$ and $R^2_c=0.346$.
+The pseudo-coefficients of determination are $R^2_m=0.239$ and $R^2_c=0.346$.
 There is thus some inter-subject variation and a noteworthy correlation between the reactions of the subjects and the corpus-derived probabilities.
-\hl{The fact that the correlation is not stronger can be explained in parts \ldots}
 
-\begin{figure}[h]
+\begin{figure}[htbp!]
 \centering
@@
 
-The effect plot for the effect of interest is shown in Figure~\ref{fig:spr:effects}.
-The estimate for the sentences with \NACa\ is obviously imprecise and no differences in reading times are observed.
+The effect plot for the interaction of interest is shown in Figure~\ref{fig:spr:effects}.
+The estimate for the sentences with \NACa\ is obviously imprecise and no significant differences in reading times are observed.
 There is a clearer effect in the sentences with \PGCa, which is also confirmed by the significant results from the bootstrapped confidence intervals (see Table~\ref{tab:exp:spr}) and from the PB test reported above.
-The \PGCa\ brings about an increased reading time ($\beta_{\text{ConstructionPGCa}}=0.054$), which is plausible because it is the much rarer construction (see Section~\ref{sec:corpusstudies}).
-However, if it occurs in a prototypical context and with prototypical lexical material, reading times drop ($\beta_{\text{ConstructionPGCa:Modelprediction}}=-0.125$).
+The \PGCa\ brings about an increased reading time, which is plausible because it is the much rarer construction (see Section~\ref{sec:corpusstudies}).
+However, if it occurs in a prototypical context and with prototypical lexical material, reading times drop.
 This can be seen in the downward slope in the right panel of Figure~\ref{fig:spr:effects}.
 This fits into the general picture inasmuch as the construction with the lower frequency might be developing towards a more sharply defined prototype.%
-\footnote{In this context, it should be remembered from Section~\ref{sec:gettingdata} that even the \PGCd\ is much rarer that the \NACb\ (17,252 vs.\ 315,635 occurrences in the auxiliary corpus samples).}
+\footnote{In this context, it should be remembered from Section~\ref{sec:corpusstudies} that even the \PGCd\ is much rarer that the \NACb\ (17,252 vs.\ 315,635 occurrences in the auxiliary corpus samples).}
 Conversely, the \NACa\ (like the NAC in general) might be the highly frequent default which does not incur a reading time penalty, even if it is not the optimal choice in the given context and with the given lexical material.
 
-This concludes the report of the two experiments.
-In the final section, I take stock and summarise the contribution of the present study to the research on alternations in cognitive linguistics.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+In Section~\ref{sec:conclusion}, I take stock and summarise the contribution of the present study to the research on alternations in cognitive linguistics.

### 2017-12-31 R/3a_glmm_collo.R

@@ -72,8 +72,8 @@ print(summary(measure.glmm.collo), correlation=F)
 cat("\n\n")
 print(measure.glmm.collo.wald)
 cat("\n\nModel comparison with LR and PB test\n")
-print(modelcomp)
-print(round(modelcomp$PB.p, precision))
+print(pbmc.measurecollo)
+print(pbmc.kindcollo)
 cat("\n\n")
 print(round(measure.glmm.collo.r2, precision))
 cat("\n\n")

### 2017-12-31 R/6_exp-fc.R

@@ -63,9 +63,18 @@ responses.df$Chosenconstruction <- as.factor(ifelse(responses.df$Chosenconstruct
 # REPORT
 if (save.persistent) sink(paste(out.dir, "results.txt", sep=""))
 
+
+# Overparametrised model.
+cat("\n\nGLMM, predicting reactions from corpus model: OVERPARAMETRISED \n\n")
+model.fc.op <- glmer(Chosenconstruction~Modelprediction+(1+Modelprediction|Item)+(1+Modelprediction|Participant),
+                  family = binomial(link = "logit"), data = responses.df, nAGQ=the.nAGQ,
+                  control=glmerControl(optimizer="nloptwrap2", optCtrl=list(maxfun=2e5)))
+print(summary(model.fc.op))
+
+
 # Actual model.
-cat("\n\nGLMM, predicting reactions from corpus model\n\n")
-model.fc <- glmer(Chosenconstruction~Modelprediction+(1+Modelprediction|Item)+(1+Modelprediction|Participant),
+cat("\n\nGLMM, predicting reactions from corpus model: PARSIMOUNIOUS\n\n")
+model.fc <- glmer(Chosenconstruction~Modelprediction+(1|Item)+(1|Participant),
                   family = binomial(link = "logit"), data = responses.df, nAGQ=the.nAGQ,
                   control=glmerControl(optimizer="nloptwrap2", optCtrl=list(maxfun=2e5)))
 print(summary(model.fc))
@@ -128,9 +137,11 @@ if (save.persistent) dev.off()
 
 # Descriptive plot of responses model prediction.
 if (save.persistent) pdf(paste(out.dir, "proportions.pdf", sep=""))
-plot(responses.df$Chosenconstruction~responses.df$Modelprediction,
-     main = "Forced choice: distribution of responses\nby (binned) predictions from corpus-based model",
-     xlab = "Probability for PGCadj from corpus-based model", ylab="Proportion of responses",
+par(cex.lab = 1.5, cex.axis = 1.5, mar=c(5,5,5,5))
+spineplot(responses.df$Chosenconstruction~responses.df$Modelprediction,
+     #main = "Forced choice: distribution of responses\nby (binned) predictions from corpus-based model",
+     xlab = "Probability for PGCadj from corpus-based model",
+     ylab = "Proportion of responses",
      col = c("gray30", "white"))
 if (save.persistent) dev.off()
 

### 2018-01-01 Latex/article.tex
Done with everything except conclusion.

@@
   Elizabeth Pankratz thankfully also fixed my English.
-  Finally, I am grateful to my student assistants Kim Maser, for her work on the annotation of the concordances, and Luise Rißmann, for supervising most of the experiments.
+  Finally, I am grateful to my student assistants Kim Maser for her work on the annotation of the concordances and Luise Rißmann for supervising most of the experiments.
   The research presented here was made possible in part through funding from the \textit{Deutsche Forschungsgemeinschaft} (DFG, personal grant SCHA1916/1-1).

### 2018-01-01 Latex/sections/3corpus.tex
Done with everything except conclusion.

@@
 
+\subsubsection{Sampling and annotation}
+
 The main study to be reported below deals exclusively with the alternating constructions, as is customary in alternation modelling.
@@
 All features described in Section~\ref{sec:variablesandannotation} were annotated except for the ones which do not apply in the non-alternating case (\textit{Kindattraction}, \textit{Measureattraction}, and \textit{Measurecase}).
-The aggregated counts for each concordance were analysed in the form of a multilevel logistic regression model using R \citep{R} and the \textit{lme4} package \citep{Bates2010,BatesEa2015}.%
+
+\subsubsection{Model}
+
+The annotated concordance was analysed in the form of a multilevel logistic regression model using R \citep{R} and the \textit{lme4} package \citep{Bates2010,BatesEa2015}.%
 \footnote{An alternative \textit{BOBYQA} optimiser from the \textit{nloptr} package \citep{Johnson2017} was used for all fits with \textit{lme4} reported in this paper.}
@@
 Nakagawa \& Schielzeth's pseudo-coefficient of determination is $R_m^2=0.278$ and $R^2_c=0.566$ (see \citealp{Gries2015} for a basic introduction to these $R^2$ measures, or else \citealp{NakagawaSchielzeth2013}).
-This means that the effect of the fixed effects (marginal $R^2$) is weak to intermediate, and that lemma-specific effects (conditional minus marginal $R^2$) are equally strong with a difference between the two coefficients of 0.289.
 
-This show that the main prototypical factors \textit{Cardinal} and \textit{Measureclass} indeed play the predicted role
+\subsubsection{Interpretation}
+
+The coefficients of determination indicate that the effect of the fixed effects (marginal $R^2$) is weak to intermediate, and that lemma-specific effects (conditional minus marginal $R^2$) are equally strong with a difference between the two coefficients of 0.289.
+
+The main prototypical factors \textit{Cardinal} and \textit{Measureclass} bot pass the PBmodcomp test at sig=0.05 and play the expected role.
 The abstract non-referential physical measure nouns (such as \textit{Gramm} `gram' or \textit{Liter} `litre') with a high degree of grammaticalisation favour the \NACb.
@@
 The lemma frequency regressors both clearly fail the PBmodcomp test.
-I will return to these results in Section~\ref{sec:interpretation}.
+I will return to these results in Section~\ref{sec:conclusion}.
 
@@
 \label{sec:corpushierarchicalmodel}
+
+\subsubsection{Sampling and annotation}
 
@@
 Therefore, all noun-noun pairs from the process described above were also queried in the two non-alternating constructions, resulting in 17,252 hits for the \PGCd\ and 315,635 hits for the \NACb.
+
+\subsubsection{Model}
 
@@
 Using signed logarithmised collexeme strength (\textit{Measurecollo} and \textit{Kindcollo}) (with smoothing to avoid log(0) operations) instead of the quotient for the attraction strength (\textit{Measureattraction} and \textit{Kindattraction}) (see Section~\ref{sec:variablesandannotation}) was not successful.
-While the attraction measures reach satisfying sig levels in the PBmodcomp test (0.020 for \textit{Kindattraction} and 0.001 for \textit{Measureattraction}), the $p_{\text{PB}}$ value for \textit{Kindcollo} was 0.191 and the one for \textit{Measurecollo} was 0.443.
+While the attraction measures reach satisfying sig levels in the PBmodcomp test (0.020 for \textit{Kindattraction} and 0.001 for \textit{Measureattraction}), the $p_{\text{PB}}$ value for \textit{Kindcollo} was 0.191 and the one for \textit{Measurecollo} was 0.443.%
+\footnote{My approach to statistical inference is essentially Fisherian \citep{Fisher1935,Fisher1959}, and hence I use the term \textit{sig level} instead of the term \textit{$\alpha$ level} known from null hypothesis significance testing (NHST).
+For overviews of systems of statistical inference with special focus on this and related questions, see \citet{Lehmann1993,Lehmann2011} or the shorter \citet{Perezgonzalez2015}.}
 The coefficients of determination drop to $R_m^2=0.135$ and $R^2_c=0.636$.
 
-\subsection{Interpretation}
-\label{sec:interpretation}
-
-%\hl{As collostructional analysis is relatively popular in some cognitively oriented areas of corpus linguistics, I tried using collexeme strength as a regressor (with smoothing to remedy the mathematical problems), and the results were unsatisfactory compared to the simple quotient used here.
-%The corresponding regressor did not even reach sig=0.1 in the PBmodcomp test.
-%Both the cognitive underpinnings and the math behind collostructional analysis are still much debated (see prominently \citealp{SchmidKuechenhoff2013,KuechenhoffSchmid2015}), I see this result }
-
-% ATTRACTION
+\subsubsection{Interpretation}
 
@@
 An interesting picture emerges for the lemma frequencies.
-A higher-than-average lemma frequency of measure nouns favours the \NACa\ ($\beta_{\text{Measurefreq}}=-0.231$, $\mpPB=0.005$), which is as expected if we assume at least a tendency for highly grammaticalised items to be more frequent.
-With kind nouns, higher frequency seems to favour the \PGCa ($\beta_{\text{Kindfreq}}=0.146$, $\mpPB=0.095$).
-However, there is no clear theoretical interpretation (see Section~\ref{sec:analyses}), and the estimate is imprecise (not significant at $\alpha=0.05$, see above).
+A higher-than-average lemma frequency of measure nouns favours the \NACa\, which is as expected if we assume at least a tendency for highly grammaticalised items to be more frequent.
+With kind nouns, higher frequency seems to favour the \PGCa.
+However, there is no clear theoretical interpretation (see Section~\ref{sec:analyses}), and the estimate is imprecise (not significant at sig=0.05).
 The effect can therefore be ignored or treated as a nuisance variable.
-
-% GRAMMATICALISATION
-% => lemmas and lemma classes
 
@@
 In Section~\ref{sec:analyses}, it was also hypothesised that classes of measure nouns with a higher degree of grammaticalisation should favour the \NACa.
-The \textit{Measureclass} second-level predictor was successfully estimated ($\mpPB=0.001$).
+The \textit{Measureclass} second-level predictor reaches sig=0.05 in the PBmodcomp test.
 Looking at the effect plot in Figure~\ref{fig:eff:measureattraction}, it is evident that abstract non-referential physical measure nouns (such as \textit{Gramm} `gram' or \textit{Liter} `litre') with a high degree of grammaticalisation favour the \NACa.
@@
 These are referential nouns, confirming the hypothesis that it is prototypical of the PGC to contain two referential nouns, while the NAC prototypically only contains one (the kind noun).
-
-% CARDINALS
 
@@
 
-I now turn to the predicted effect of cardinals as modifiers of the measure noun.
-Figure~\ref{fig:eff:leftcontext} shows that cardinals indeed influence the choice of the alternant ($\mpPB=0.001$), and that cardinals have a strong tendency to co-occur with the \NACa.
+Figure~\ref{fig:eff:leftcontext} shows that cardinals indeed influence the choice of the alternant, and that cardinals have a strong tendency to co-occur with the \NACa.
 This effect was predicted in Section~\ref{sec:analyses}.
 
-% REGISTER
-% => Badness
-% => Genitives
-
 The style-related proxy variables point in the expected direction.
-Increased \textit{Badness} of the document favours the \NACa\ ($\beta_{\text{Badness}}=-0.152$, $\mpPB=0.002$), and so does a lower density of genitives ($\beta=-0.693$, $\mpPB=0.001$).
+Increased \textit{Badness} of the document favours the \NACa, and so does a lower density of genitives.
 While these are merely proxies to style (and partially circular in the case of \textit{Genitives}), this result can at least encourage future work into stylistic effects.
 
-% NUISANCE
-% => dative effect
-% => frequency
-
-The influence of \textit{Measurecase} ($\mpPB=0.001$) is as predicted in previous analyses (see Section~\ref{sec:analyses}).
-A measure noun in the dative favours the \PGCa\ with $\beta_{\text{MeasurecaseDat}}=0.705$ (compared to the nominative, which is on the intercept).
+The influence of \textit{Measurecase} is as predicted in previous analyses (see Section~\ref{sec:analyses}).
+A measure noun in the dative favours the \PGCa\ (compared to the nominative, which is on the intercept).
 Although \textit{Measurecase} is a nuisance variable in the context of this study, convergence with previous work strengthens its validity.

### 2018-01-01 Latex/sections/4experiments.tex
Done with everything except conclusion.

@@
 \label{sec:exp:fc}
+
+\subsubsection{Setup, stimuli, and participants}
 
@@
 
+\subsubsection{Model}
+
 \begin{figure}[htbp!]
@@
 
-Then, a multilevel logistic regression was specified with the probability of the \PGCa\ predicted for each sentence by the corpus-based GLMM as the only fixed effect \textit{Modelprediction}.%
+A multilevel logistic regression was specified with the probability of the \PGCa\ predicted for each sentence by the corpus-based GLMM as the only fixed effect \textit{Modelprediction}.%
 \footnote{The document-level variables \textit{Badness} and \textit{Genitives} were set to 0, which is the mean for z-transformed variables.}
@@
 The pseudo-coefficients of determination are $R^2_{m}=0.185$ and $R^2_{c}=0.455$, which means that roughly 19\% of the variance in the data can be explained by considering only the predictions from the corpus-based GLMM.
-This is a weak result for the fixed effects part, but clearly worthy of mention.
+
+\subsubsection{Interpretation}
+
+The coefficients of determination indicate a weak result for the fixed effects part, which is nonetheless worthy of mention (very close to 0.2).
 The effect display for the single fixed regressor \textit{Modelprediction} is given in Figure \ref{fig:afc:effects}.
@@
 \label{sec:exp:spr}
+
+\subsubsection{Setup, stimuli, and participants}
 
@@
 Outliers farther than 2 inter-quartile ranges from the mean logarithmised residualised reading time were removed (64 data points), resulting in a total number of \textit{n=}1,152 observations.
+
+\subsubsection{Model}
+
 An LMM was specified with the logarithmised residual reading times as the response variable.
-
 The probabilities derived from the corpus GLMM (\textit{Modelprediction}) were added as the main regressor of interest.
@@
 The pseudo-coefficients of determination are $R^2_m=0.239$ and $R^2_c=0.346$.
-There is thus some inter-subject variation and a noteworthy correlation between the reactions of the subjects and the corpus-derived probabilities.
+
+An alternative Gaussian generalised additive model with an identity link was also fit (see \citealp{DivjakEa2016}) using the \textit{mgcv} package \citep{Wood2011}.
+The full results are included in the data package for this paper, but the fit was not better than with the LMM reported above.
+The estimated smoother for the \textit{Modelprediction} variable is essentially linear, and the $R^2$ (corresponding to the marginal $R^2$ of the LMM) was 0.237.
+
+\subsubsection{Interpretation}
+
+The coefficients of determination indicate that there is a noteworthy correlation between the reactions of the subjects and the corpus-derived probabilities (marginal $R^2$ and that there some inter-subject variation (conditional minus marginal $R^2$).
 

### 2018-01-01 R/6_exp-fc.R

@@ -137,7 +137,7 @@ if (save.persistent) dev.off()
 
 # Descriptive plot of responses model prediction.
 if (save.persistent) pdf(paste(out.dir, "proportions.pdf", sep=""))
-par(cex.lab = 1.5, cex.axis = 1.5, mar=c(5,5,5,5))
+par(cex.lab = 1.75, cex.axis = 1.75, mar=c(5,5,5,5))
 spineplot(responses.df$Chosenconstruction~responses.df$Modelprediction,
      #main = "Forced choice: distribution of responses\nby (binned) predictions from corpus-based model",
      xlab = "Probability for PGCadj from corpus-based model",

### 2018-01-01 R/7_exp-spr.R

@@ -313,7 +313,7 @@ rt.gamm <- gamm(tf.rt.resid ~ s(Position) + s(Modelprediction, by = Construction
   niterPQL = 1000
 )
 
-# ... with NO result. At least not different from LM.
+# ... with NO different resultcompared to LMM.
 # Notice that the relevant smoother (only significant term)
 # is virtually linear.
 

### 2018-01-01 Latex/sections/1intro.tex
Revisions in Sec.3 almost and in Sec.4 cmpltly done.
Working on conclusion.

@@
 This paper deals with a morpho-syntactic alternation between two constructions that occurs only in a very specific type of measure noun phrase in German.
-By \textit{alternation} I refer to a situation where two or more forms or constructions are available with no clear difference in acceptability, function, or meaning.
+By \textit{alternation} I refer to a situation where two or more forms or constructions are available with no clear (but potentially a subtle) difference in acceptability, function, or meaning.
 The study of lexical and constructional alternations has a long history in cognitively oriented corpus linguistics (for example, \citealp{BresnanEa2007,BresnanHay2010,BresnanFord2010,DivjakArppe2013,Gries2015,NessetJanda2010,Wulff2003}).
@@
 This question can and should be answered by testing how well corpus-derived models converge with or diverge from experimental findings, which provide more direct evidence of mental representations and processes but are often intrinsically narrower in scope.
-Second and closely related to the first question, the appropriate modelling of such results in cognitive linguistics is a key issue.
+Second and closely related to the first question, the appropriate modelling of such results in cognitive linguistics (\ie\ the assumed underlying constructs) is a key issue.
 
@@
 
-%An aspect generally neglected in models of categorisation is how categorisation feeds exemplar generation and creativity (\ie category generation), two subjects of utmost importance to many linguists.
-%In \cite{JernKemp2013}, a model is assumed where cognitive agents sample from learned distributions over both categories and exemplars to generate new ones.
-
 In cognitive linguistics, \cite{DivjakArppe2013} is a very rare example of a paper where such issues are taken up with reference to the current research in cognitive science.

### 2018-01-01 Latex/sections/2descriptive.tex
Working on conclusion.

@@
 I describe the narrowly defined syntactic configuration in which the alternation occurs, and I motivate the focus on \textit{only} this narrow range rather than, for example, the whole range of nominal constructions expressing quantities.
+Finally, I develop a model with prototype effects as well as exemplar effects based in existing research and my own theorising.
 
@@
 The relative frequency with which they occur in these stable cases -- where choosing an alternative is impossible -- could thus be a factor influencing the alternating, less stable case.
-Should this be confirmed, it would be highly implausible to conceive of such effects as prototype effects.
+Should this be confirmed, it would be highly implausible to conceive of such an effect as a prototype effect.
 In the corpus study reported in Section~\ref{sec:corpusstudies}, a measure quantifying this influence will therefore be included as the \textit{attraction strength}.

### 2018-01-01 Latex/sections/3corpus.tex
Working on conclusion.

@@
                    & Kindfreq                         & numeric &                           \\
-		   & Kindgender                       & factor  & MascNeut, Fem             \\
+		   & Kindgender                       & factor  & Masc, Neut, Fem             \\
     Measure lemma  & Measureattraction                & numeric &                           \\
@@
 After all, the main difference between the quotient used here and the logarithmised p-values from a Fisher test is that they take these marginals into account. 
-However, using collexeme strength instead of the raw frequency quotient was tried as an alternative (variables \textit{Kindcollo} and \textit{Measurecollo}); see Sections~\ref{sec:corpushierarchicalmodel} and~\ref{sec:interpretation}.
+However, using collexeme strength instead of the raw frequency quotient was tried as an alternative (variables \textit{Kindcollo} and \textit{Measurecollo}); see Sections~\ref{sec:corpushierarchicalmodel} and~\ref{sec:interpretation:corpus}.
 Additionally, \textit{Kindfreq} and \textit{Measurefreq} are the logarithm-transformed frequencies per 1,000,000 words of each lemma, extracted from the frequency lists distributed by the DECOW14A corpus creators on their web page.
@@
 
-\subsubsection{Model}
+\subsubsection{Statistical model}
 
@@
 
-The coefficients of determination indicate that the effect of the fixed effects (marginal $R^2$) is weak to intermediate, and that lemma-specific effects (conditional minus marginal $R^2$) are equally strong with a difference between the two coefficients of 0.289.
+The coefficients of determination indicate that the effect of the fixed effects (marginal $R^2$) is adequately strong, and that lemma-specific effects (conditional minus marginal $R^2$) are equally strong with a difference between the two coefficients of 0.289.
 

### 2018-01-01 Latex/sections/4experiments.tex
Working on conclusion.

@@
 
-\subsubsection{Setup, stimuli, and participants}
-
 In the two experiments reported in this section, I use probabilities for the alternating constructions calculated for attested material, and I correlate these probabilities with the participants' reactions.
@@
 Also, the probabilities that the corpus-based model assigns to the two alternants in these sentences are used as the main regressor in both studies.
+
+Readers might wonder why the stimuli were chosen in such a way.
+Alternatively, one could have created stimuli where the features favouring the alternatives were permuted appropriately and thus tested directly.
+The answer is related to what was said in Section~\ref{sec:cogocl} about the rich multi-factorial data sets used in corpus studies and the comparatively restricted ones in experiments.
+Looking at Table~\ref{tab:variables}, we see one binary factor (\textit{Cardinal}), two three-level factors (\textit{Measurecase} and \textit{Kindgender}) and one five-level factor (\textit{Measureclass}).
+Since the dependent variable is also binary (\textit{Construction}), permuting these factors alone leads to 180 different possible permutations.
+On top of this, the numeric variables \textit{Kindattraction} and \textit{Measureattraction} would also have to be tested at at least two or three different values each.
+It is obvious that this is impossible in a controlled experiment, given that participants can usually be exposed to something in the region of one hundred sentences (including many more fillers than target stimuli) within thirty minutes to an hour.
+Thus, while the present approach only allows for a global test of the corpus-derived model, this appears to me to be the only feasible way.%
+\footnote{Obviously, the experimental data are much too sparse to perform even post-hoc analyses with respect to the single regressors in a statistically and scientifically sound way.}
+It is, in the sense of Section~\ref{sec:cogocl}, an optimal synthesis of data-rich multifactorial corpus studies and experimental validation, and it is one that has been used (although rarely; see \citealp[3--4]{DivjakEa2016a}) since the seminal \citet{BresnanEa2007} paper.
+
+\subsubsection{Setup, stimuli, and participants}
 
@@
 The number of observations was \textit{n=}384.
-
-\begin{figure}[htbp!]
-  \centering
-  \includegraphics[width=0.5\textwidth]{../R/output/fc_proportions}
-  \caption{Spineplot of the proportion of responses plotted against the predictions from the corpus-based model in the forced-choice experiment}
-  \label{fig:spines}
-\end{figure}
 
@@
 Clearly, a revised and improved experiment might lead to a much better fit in future research.
+
+\begin{figure}[htbp!]
+  \centering
+  \includegraphics[width=0.5\textwidth]{../R/output/fc_proportions}
+  \caption{Spineplot of the proportion of responses plotted against the predictions from the corpus-based model in the forced-choice experiment}
+  \label{fig:spines}
+\end{figure}
 

### 2018-01-01 Latex/sections/5conclusions.tex
Revisions.
Working on conclusion.

--- 2017-12-22 Latex/sections/5conclusions.tex
@@
 
-This paper stands in a now ten year-old tradition of research on grammatical alternations using corpus and experimental data.
-In my view, the main tenets of this line of research are:
-(i)~Language, viewed from a cognitively realistic angle, is a probabilistic phenomenon and cannot be modelled appropriately within Aristotelian frameworks that assume discrete categories.
-(ii)~Corpora are collections of usage events (language production) and can therefore be used to evaluate both the claim made in (i) and specific theoretical claims (in case studies) about factors influencing speakers' decisions to use specific forms or constructions.
-(iii)~Given (ii), we expect results from corpus analyses and from appropriate experiments to yield similar results, not necessarily as a form of validation of the corpus-based findings, but as converging evidence.
-I consider these three points to be of utmost importance because they clearly set this approach to linguistic research apart from \textit{both} Aristotelian frameworks \textit{and} introspective, non-empirical, and anti-quantitative versions of cognitive linguistics (see \citealp{Dabrowska2016} for a pithy philippic against such approaches).
+%%% PROTOYPE VS EXEMPLARS
 
-The present paper adds to the evidence that all of the aforementioned three points are correct.
-A grammatical alternation in German measure NPs was examined using corpus data based on factors partly derived from existing accounts, formulated in terms of construction prototypes.
-The preferences extracted from the DECOW web corpus were confirmed in a forced-choice experiment, in which participants explicitly chose alternants in line with the probabilities derived from the corpus-based model.
-In a more implicit self-paced reading experiment, it was shown that the much rarer alternant brings about a reading time penalty except in cases for which the corpus model predicts very high probability for this alternant.
+% Shown how prototype and exemplar effects control the alternation.
 
-Future work could extend these results and provide a general picture of the constructions expressing measurements (see Section~\ref{sec:descriptive}).
+% Unnecessary to follow extreme routes for no-abstraction or full-abstraction model.
+
+% Show how multilevel modelling helps to specify complex interactions between abstract features, item-specific tendencies and exemplar effects.
+
+% All in all, clear evidence for probabilistic nature of the phenomenon.
+
+%%% CONVERGENCE AND NON-CONVERGENCE
+
+% There are now a number of published multivariate models that use data, extracted from corpora and annotated for a multitude of morphological, syntac- tic, semantic and pragmatic parameters, to predict the choice for one morpheme, lexeme or construction over another.
+However, most of these studies are con- cerned with phenomena that involve binary choices (Gries 2003; De Sutter et al.
+2008) and only a small number of these1 corpus-based studies have been cross-validated (Keller 2000; Wasow and Arnold 2003; Sorace and Keller 2005; Roland et al.
+2006; Arppe and Järvikivi 2007; Divjak and Gries 2008).2 Of these cross- validated studies, few have directly evaluated the prediction accuracy of a complex, multivariate corpus-based model on humans using authentic corpus sentences (with the exception of Bresnan 2007; Bresnan and Ford 2010; Ford et al.
+2013a; Ford et al.
+2013b), and even fewer have attempted to evaluate the prediction accuracy of a polytomous corpus-based model in this way (but see Arppe and Abdulrahim 2013 for a first attempt).
+Below we will review the latter two types of cross-validated studies.
+
+% Pre-study and main study mostly converge, although Badness is weak in pre-study. Makes it highly likely that the postulated factors are substantial.
+% Also: Genitives is a good measure, corroborated through pre-study.
+
+% Notice that overall fit of main study is good R2m>0.4 but not anywhere near perfect predictability.
+% Due to more subtle preferences and more free variation. Rarely do speakers feel one alternative is odd or bad. => More experiments.
+
+% Important: Nuisance factor (case) in line with Zimmer2015, which improves the validity of the presented results.
+
+% Weak convergence with experiments: Consider lower R2 in main study as level to begin with. No wonder that experimental corroboration is even weaker.
+% See also extreme inter-speaker variation, and possibly problematic stimuli which marred results. => More experiments.
+% Why self-paced reading better? Less room for normative evaluations, and genitives are a hot matter of public debate.
+
+% (External) things that were NOT corroborated:
+% Collexeme strength failed against raw frequency-based measure.
+% GAM did not improve fit in reading time experiment. Effect was quite linear.
+% Overparametrisation was ill-advised as predicted by BatesEa2015a, MatuschekEa2017.
+
+% Mention pre-registration.
+
+% We need: more case studies +  diverse and different corpora + focus on register/style-specific aspects + inter-speaker similarity + focus on different experimental paradigms and what they really tell us + more clarity wrt statistical methods + ultimately fully specified mathematical models as in cognitive science
+
+Besides the methodological aspects mentioned above, future work could extend the results and provide a general picture of the constructions expressing measurements (see Section~\ref{sec:descriptive}).
 This would be a much more complicated task given that the choices then would no longer be binary and that the meaning of the alternative ways of expressing measurements are semantically more varied.
+
+% Potential refinement: more detailed analysis instead of CARD vs. not CARD
+% Potential refinement: postnominal modification (= true partitives)
+% Regional variation (experiments)
+% Mention R1 and his ``lexicalised A+N combinations''
+
 Finally, I want to point out that German is mildly under-researched in the specific framework used here.

### 2018-01-02 Latex/sections/1intro.tex
Done.

@@
 Concerning the first point, there has for a long time been an interest in correlating probabilistic generalisations extracted from corpus data with results from experimental work (for example, \citealp{ArppeJaervikivi2007,BresnanEa2007,BresnanFord2010,DivjakGries2008,DivjakEa2016,FordBresnan2013}).
-This is often called a \textit{validation} of the corpus-derived findings, but \citet[303]{Divjak2016a} criticises this choice of words ``because it creates the impression that behavioral experimental data is inherently more valuable than textual data,'' citing \cite{TummersEa2005}, who state that a corpus is ``a sample of spontaneous language use that is (generally) realized by native speakers''.
+This is often called a \textit{validation} of the corpus-derived findings, but \citet[303]{Divjak2016a} criticises this choice of words ``because it creates the impression that behavioral experimental data is inherently more valuable than textual data'', citing \cite{TummersEa2005}, who state that a corpus is ``a sample of spontaneous language use that is (generally) realized by native speakers''.
 See also \citet{Newman2011} for a very positive view of corpora as a source of data in cognitive linguistics in their own right.
-However, as \citet[486--487]{Dabrowska2016} convincingly argues, this does not mean that we can in some way ``deduce mental representations from patterns of use,'' \ie\ from corpus data.%
+However, as \citet[486--487]{Dabrowska2016} convincingly argues, this does not mean that we can in some way ``deduce mental representations from patterns of use'', \ie\ from corpus data.%
 \footnote{Epistemologically, I believe she refers to acts of \textit{induction} rather than \textit{deduction}.}
@@
 In the remainder of this paper, I do not use schemas in my descriptions of the relevant categories, mostly because the aspect of similarity and fuzzy classification is central to my point, and a formulation in terms of schemas would bring about an unnecessarily high degree of abstraction (see \citealp[70--71]{Taylor2003} for a parallel argument).
-However, I fully describe the abstract features of the prototypes, of course, while accounting for item-specific and exemplar effects.
+However, I fully describe the abstract features of the prototypes.
 
@@
 Since corpus data only show artefacts of production events and we have no experimental access to the speaker's or writer's performance and their actual similarity judgements, one should be sceptical whether corpus analysis alone could ever decide which theory of mental representation is more suitable (see also \citealp[22]{Gries2003} and the \citealp[486--487]{Dabrowska2016} quote above).
+However, as I will show, some effects are more naturally analysed as prototype effects while others are almost necessarily exemplar effects.
 
@@
 Still, various attempts have been made over the past decades to settle the dispute between abstraction-based models (models with rules or prototypes) and exemplar models or to find models which unite the two extremes.
-\cite{VanpaemelStorms2008} and \citet{LeeVanpaemel2008} proposed the \textit{varying abstraction model} (VAM) which ``attempt[s] to balance economy and informativeness'' \citep[745]{LeeVanpaemel2008}, treating models with full abstraction (prototypes) and no abstraction at all (radical exemplar theory) as special cases of a model which allows for both abstraction and exemplar effects.
+\cite{VanpaemelStorms2008} and \citet{LeeVanpaemel2008} proposed the \textit{varying abstraction model} (VAM) which ``attempt[s] to balance economy and informativeness'' \citep[745]{LeeVanpaemel2008}, treating models with full abstraction (radical prototype theory) and no abstraction at all (radical exemplar theory) as special cases of a model which allows for both abstraction and exemplar effects.
 The mixture model of categorisation (MMC) by \cite{Rosseel2002} is a model with abstraction in the form of hierarchical clusters of exemplars, and these clusters of objects are characterised by a probability distribution over their features, and categorising new objects is a process of estimating the probability of this object belonging to one of the clusters.

### 2018-01-02 Latex/sections/3corpus.tex
Done.

@@
 The variables \textit{Kindattraction} and \textit{Measureattraction} encode the ratio with which a given kind noun lemma or measure noun lemma occurs in the \PGCd\ and the \NACb.
-They were calculated from the auxiliary samples described at the end of Section~\ref{sec:gettingdata} as a log-transformed quotient.
+They were calculated from auxiliary samples to be described in Section~\ref{sec:mainstudysampling} as a log-transformed quotient.
 The higher the value, the more often the noun occurs in the \PGCd\ (proportionally).
@@
 After all, the main difference between the quotient used here and the logarithmised p-values from a Fisher test is that they take these marginals into account. 
-However, using collexeme strength instead of the raw frequency quotient was tried as an alternative (variables \textit{Kindcollo} and \textit{Measurecollo}); see Sections~\ref{sec:corpushierarchicalmodel} and~\ref{sec:interpretation:corpus}.
+However, using collexeme strength instead of the raw frequency quotient was tried as an alternative (variables \textit{Kindcollo} and \textit{Measurecollo}); see Section~\ref{sec:corpushierarchicalmodel} for a discussion of the negative results.
+
 Additionally, \textit{Kindfreq} and \textit{Measurefreq} are the logarithm-transformed frequencies per 1,000,000 words of each lemma, extracted from the frequency lists distributed by the DECOW14A corpus creators on their web page.
@@
 I call the corresponding value $p_{\text{PB}}$, and it is given in the respective columns in Table~\ref{tab:smalltable}.
-Regressors which did not reach sig=0.05 in in the PBmodcomp test (\textit{Kindfreq}, \textit{Kindgender}, and \textit{Measurefreq}) were removed from the model, appear in brackets in Table~\ref{tab:smalltable}, and have no coefficient estimates.
+Regressors which did not reach sig=0.05 in in the PBmodcomp test (\textit{Kindfreq}, \textit{Kindgender}, and \textit{Measurefreq}) were removed from the model, appear in brackets in Table~\ref{tab:smalltable}, and consequently have no coefficient estimates.%
+\footnote{My approach to statistical inference is essentially Fisherian \citep{Fisher1935,Fisher1959}, and hence I use the term \textit{sig level} instead of the term \textit{$\alpha$ level} known from null hypothesis significance testing (NHST).
+For overviews of systems of statistical inference with special focus on this and related questions, see \citet{Lehmann1993,Lehmann2011} or the shorter \citet{Perezgonzalez2015}.}
 The overall intercept is -5.370 and reoresents \textit{Cardinal=Yes}, \textit{Measureclass=Physical} and 0 for all (z-transformed) numeric regressors.
@@
 The lemma frequency regressors both clearly fail the PBmodcomp test.
-I will return to these results in Section~\ref{sec:conclusion}.
+I will return to these results in Section~\ref{sec:mainstudyinterpretation}.
 
@@
 \subsubsection{Sampling and annotation}
+\label{sec:mainstudysampling}
 
@@
 \subsubsection{Statistical model}
+\label{sec:mainstudystatisticalmodel}
 
@@
 \footnote{Notice that no aggressive model selection was applied, especially not upward model selection including probing for interactions.
-I am convinced that the model should be specified based on theoretical considerations (and known nuisance effects) to avoid the dangers of data dredging and to avoid models which are difficult to interpret.
+I am convinced that models should be specified based on theoretical considerations (including known nuisance effects) to avoid the dangers of data dredging and to avoid models which are difficult to interpret.
 Removal of useless regressors (judging by, for example, the PBmodcomp test and coefficients of determination) is the only type of model selection I allow in my research.}
@@
 
-Using signed logarithmised collexeme strength (\textit{Measurecollo} and \textit{Kindcollo}) (with smoothing to avoid log(0) operations) instead of the quotient for the attraction strength (\textit{Measureattraction} and \textit{Kindattraction}) (see Section~\ref{sec:variablesandannotation}) was not successful.
-While the attraction measures reach satisfying sig levels in the PBmodcomp test (0.020 for \textit{Kindattraction} and 0.001 for \textit{Measureattraction}), the $p_{\text{PB}}$ value for \textit{Kindcollo} was 0.191 and the one for \textit{Measurecollo} was 0.443.%
-\footnote{My approach to statistical inference is essentially Fisherian \citep{Fisher1935,Fisher1959}, and hence I use the term \textit{sig level} instead of the term \textit{$\alpha$ level} known from null hypothesis significance testing (NHST).
-For overviews of systems of statistical inference with special focus on this and related questions, see \citet{Lehmann1993,Lehmann2011} or the shorter \citet{Perezgonzalez2015}.}
-The coefficients of determination drop to $R_m^2=0.135$ and $R^2_c=0.636$.
+Using signed logarithmised collexeme strength (\textit{Measurecollo} and \textit{Kindcollo}) (with smoothing to avoid \textit{log(0)} errors) instead of the quotient for the attraction strength (\textit{Measureattraction} and \textit{Kindattraction}) (see Section~\ref{sec:variablesandannotation}) was not successful.
+While the attraction measures reach satisfying sig levels in the PBmodcomp test (0.020 for \textit{Kindattraction} and 0.001 for \textit{Measureattraction}), the $p_{\text{PB}}$ value for \textit{Kindcollo} was 0.191 and the one for \textit{Measurecollo} was 0.443.
+The coefficients of determination drop to $R_m^2=0.376$ and $R^2_c=0.480$.
 
 \subsubsection{Interpretation}
+\label{sec:mainstudyinterpretation}
 
@@
 An interesting picture emerges for the lemma frequencies.
-A higher-than-average lemma frequency of measure nouns favours the \NACa\, which is as expected if we assume at least a tendency for highly grammaticalised items to be more frequent.
+A higher-than-average lemma frequency of measure nouns favours the \NACa, which is as expected if we assume at least a tendency for highly grammaticalised items to be more frequent.
 With kind nouns, higher frequency seems to favour the \PGCa.
@@
 Although \textit{Measurecase} is a nuisance variable in the context of this study, convergence with previous work strengthens its validity.
+
+To close this section, I now compare the results of the pre-study and the main study.
+Even though the non-alternating constructions are surely subject to additional constraints (especially because they are much less interchangeable than the alternating constructions), the coefficients align neatly in many cases.
+First of all, the intercepts encode a similar overall dominance of the NAC constructions (pre-study: -5.370; main study: -3.548).
+For the \textit{Cardinal} effect, the coefficients 1.419 (pre-study) and 1.189 (main study) have the same sign and magnitude and mostly overlapping confidence intervals.
+The levels of \textit{Measureclass} have comparable coefficients, although the divergence is larger.
+In both studies, the abstract measure nouns in the \textit{Physical} class (on the intercept) are most clearly associated with the NAC constructions.
+Also, in both cases, the \textit{Container} class is closest to \textit{Physical} with the same sign and magnitude (pre-study: 0.445; main study: 0.252).
+The main difference -- if we ignore the \textit{Rest} class which can be expected to show no clear tendency -- is that \textit{Amount} and \textit{Portion} are slightly closer together in their tendency to favour the PGC constructions in the pre-study (\textit{Amount} 1.597 and \textit{Portion} 1.782 in the pre-study vs.\ \textit{Amount} 0.831 and \textit{Portion} 1.217 in the main study).
+The difference is not huge, and the overall order of the coefficients is the same.
+The \textit{Genitives} effect also converges with -0.710 in the pre-study and -0.693 in the main study.
+The two studies do not converge with respect to the \textit{Badness} variable, which is not significant in the pre-study.
+It would, however, surprise to achieve perfectly converging results given that even real effects are missed at certain rates.
+In the case of \textit{Badness}, we see that even in the main study, the effect size is small (-0.152), and thus even at the impressive sample size of roughly 5,000 in both studies, the effect might simply be too weak to be detected reliably.
+Finally, the frequency effect \textit{Measurefreq} is detected in the main study but not in the pre-study, while the \textit{Kindfreq} effect is essentially absent in both studies.
+In connection with this, it is revealing to look at the coefficients of determination.
+In the pre-study, a much greater proportion of the variance is explained by taking the lemma random effects into account ($R^2_m=0.278$ and $R^2_c=0.566$) than it is in the main study ($R^2_m=0.376$ and $R^2_c=0.480$).
+Thus, the frequency effect for measure lemmas might be swamped by the lemma random intercepts.
+All things considered, the studies have shown that the predicted prototype effects are mostly found in usage data for both the alternating and the non-alternating constructions.

### 2018-01-02 Latex/sections/4experiments.tex
Done.

@@
 In \cite{VerhoevenTemme2017}, it was shown that results from forced-choice and split-100 experiments mostly converge (with some numeric intricacies related to the non-linear distribution of preferences for alternants).
-I present a forced-choice experiment in this Section instead of a split-100 experiment not only because of some researchers' critical attitude towards the split-100 task, but also because in a dry run of the experiment, participants complained about the unnaturalness of distributing a probability mass across two alternants and tended to produce ratings of 0 and 100 (and to a lesser degree 50).
+I present a forced-choice experiment in this Section instead of a split-100 experiment not only because of some researchers' critical attitude towards the split-100 task, but also because in a dry run of the experiment, participants complained about the unnaturalness of distributing a probability mass across two alternants and tended to produce ratings of 0 or 100 (or 50, although to a lesser degree).
 Participants had to choose between two sentences that differed only in that one contained the \NACa\, and the other one contained the \PGCa.
@@
 The regressor \textit{Modelprediction} achieves $\mpPB=0.003$ (1,000 replications) and is estimated at 4.389 relative to an intercept of -1.270.
-The confidence interval from a parametric bootstrap (1,000 replications, percentile method) for the regressor is acceptable but slightly large with a lower bound of 1.788 and an upper bound of 6.599.
+The confidence interval from a parametric bootstrap (1,000 replications, percentile method) for the regressor is acceptable but a tad large with a lower bound of 1.788 and an upper bound of 6.599.
 The pseudo-coefficients of determination are $R^2_{m}=0.185$ and $R^2_{c}=0.455$, which means that roughly 19\% of the variance in the data can be explained by considering only the predictions from the corpus-based GLMM.
@@
 
-The coefficients of determination indicate a weak result for the fixed effects part, which is nonetheless worthy of mention (very close to 0.2).
+The coefficients of determination indicate a weak result for the fixed effects part of the model, which is nonetheless worthy of mention (very close to 0.2).
 The effect display for the single fixed regressor \textit{Modelprediction} is given in Figure \ref{fig:afc:effects}.
 The higher the probability of the \PGCa\ predicted from usage data, the more often participants chose the \PGCa\ alternant in the forced-choice task.
-A closer look at the results in the form of the spineplot in Figure~\ref{fig:spines} shows, however, that it was likely idiosyncracies in the sentences with a model prediction between 0.5 and 0.6 which spoiled an otherwise much better correlation.
+A closer look at the results in the form of the spineplot in Figure~\ref{fig:spines} shows, however, that it was likely an idiosyncracy of the sentence with a model prediction between 0.5 and 0.6 which spoiled an otherwise much better correlation.
+The problematic sentence with a model prediction of 0.548 is given in (\ref{ex:problem}) in the \PGCa\ variant.
+
+\begin{exe}
+  \ex{\label{ex:problem}\gll Man machte mal wieder viel Lärm um [jede Menge [heißer Luft]\Sub{Gen}]\Sub{Acc}.\\
+  one made once again much noise about any amount hot air\\
+  \trans People made much ado about nothing once again.}
+\end{exe}
+
+In retrospect, this stimulus was badly chosen because \textit{heiße Luft} `ado' (literally `hot air') is a fixed metaphorical expression.
+This obviously influences the reactions of participants, and I will return to this in the discussion in Section~\ref{sec:conclusion}.
 Clearly, a revised and improved experiment might lead to a much better fit in future research.
@@
 
-In principle, a per-stimulus random effect for \textit{Modelprediction} could also remedy the problem with individual stimuli at least partially.
-Therefore, a model with random slopes for \textit{Modelprediction} and both random effects (\textit{Participant} and \textit{Item}) was specified and estimated.
+In principle, random slope for \textit{Modelprediction} varying by stimulus could also remedy problems with individual stimuli at least partially.
+Therefore, a model with random slopes for \textit{Modelprediction} varying by both random effects (\textit{Participant} and \textit{Item}) was specified and estimated.
 The random slope for participants was added to comply with \citet[257]{BarrEa2013} who predict \textit{catastrophically high Type I error rates} for experimental designs with within-subject manipulations if random effects structures are not kept maximal.

### 2018-01-02 Latex/sections/5conclusions.tex
Done.

--- 2017-12-22 Latex/sections/5conclusions.tex
@@
 
-This paper stands in a now ten year-old tradition of research on grammatical alternations using corpus and experimental data.
-In my view, the main tenets of this line of research are:
-(i)~Language, viewed from a cognitively realistic angle, is a probabilistic phenomenon and cannot be modelled appropriately within Aristotelian frameworks that assume discrete categories.
-(ii)~Corpora are collections of usage events (language production) and can therefore be used to evaluate both the claim made in (i) and specific theoretical claims (in case studies) about factors influencing speakers' decisions to use specific forms or constructions.
-(iii)~Given (ii), we expect results from corpus analyses and from appropriate experiments to yield similar results, not necessarily as a form of validation of the corpus-based findings, but as converging evidence.
-I consider these three points to be of utmost importance because they clearly set this approach to linguistic research apart from \textit{both} Aristotelian frameworks \textit{and} introspective, non-empirical, and anti-quantitative versions of cognitive linguistics (see \citealp{Dabrowska2016} for a pithy philippic against such approaches).
+The empirical studies in Sections~\ref{sec:corpusstudies} and~\ref{sec:experimental} confirm the theoretical model proposed in Section~\ref{sec:germanmeasurenps}, which predicted prototype effects and exemplar effects to jointly determine which of the two alternating constructions is chosen.
+More specifically, the pre-study and the main study provided evidence that the NAC constructions favour more strongly grammaticalised measure nouns, modification by cardinals, and styles where the genitive is underrepresented and the text is less coherently and elaborately written, as captured in the \textit{Badness} variable.%
+\footnote{Although, as pointed out in Section~\ref{sec:mainstudyinterpretation}, the \textit{Badness} effect is weak and not detected in the pre-study, probably as a consequence of its weakness.}
+The overall convergence between the pre-study and the main study and also the expected result for \textit{Measurecase} given previous research \citep{Zimmer2015} -- although it was only a control variable here, strengthen the validity of this study even without the experimental validation.
 
-The present paper adds to the evidence that all of the aforementioned three points are correct.
-A grammatical alternation in German measure NPs was examined using corpus data based on factors partly derived from existing accounts, formulated in terms of construction prototypes.
-The preferences extracted from the DECOW web corpus were confirmed in a forced-choice experiment, in which participants explicitly chose alternants in line with the probabilities derived from the corpus-based model.
-In a more implicit self-paced reading experiment, it was shown that the much rarer alternant brings about a reading time penalty except in cases for which the corpus model predicts very high probability for this alternant.
+While the results do not \textit{prove} -- due to the intrinsic equivalence of prototype and exemplar models with respect to the output they produce -- that the model proposed in Section~\ref{sec:germanmeasurenps} is congruent with any speaker's mental representation, the two attraction features (\textit{Kindattraction} and \textit{Measureattraction}) are virtually impossible to conceive of as prototype effects.
+As discussed in Section~\ref{sec:cogocl}, it is unnecessary to follow an extreme route, be it radical prototype theory or radical exemplar theory given recent developments in cognitive science and (less prominently) cognitive linguistics.
+While it is surely an intentionally overstated comment, \citet[15]{Kapatsinski2014} suggests that ``[i]n the extreme, some speakers’ heads could host exemplar models, and some could contain fairly abstract grammars, and the produced output would be essentially identical''.
+If this is (to a probably less extreme degree) the case, pooled data from corpora might only provide (still highly informative) data averaged over levels of abstraction.
+Clearly, more research is required in this direction.
 
-Future work could extend these results and provide a general picture of the constructions expressing measurements (see Section~\ref{sec:descriptive}).
+Another important aspect of the research presented here is the validation of the results derived from corpus data in two experimental paradigms.
+While such cross-validations have been done (with varying success, see Section~\ref{sec:cogocl}) for over a decade, they have not yet become standard procedure.
+As \citet[3--4]{DivjakEa2016a} put it:
+
+\begin{quote}
+  There are now a number of published multivariate models that use data, extracted from corpora [\ldots] to predict the choice for one morpheme, lexeme or construction over another.
+  However, [\ldots] only a small number of these corpus-based studies have been cross-validated [\ldots].
+  Of these cross- validated studies, few have directly evaluated the prediction accuracy of a complex, multivariate corpus-based model on humans using authentic corpus sentences [\ldots].
+\end{quote}
+
+The question now arises whether in the present case convergence was reached, and the answer is a clear yes.
+The predictions made by the corpus-based model were a significant factor, both in the more explicit paradigm (forced choice) and the implicit paradigm (self-paced reading), showing that the model indeed predicts the variant which language users (readers in this case) expect.
+The overall effect as measured in the $R^2$ was weak (roughly 0.2) in both cases, however.
+While part of this could be traced to a suboptimally chosen stimulus, we really need to consider what kind of convergence we expect to see between corpus data and experiments.
+The main corpus study had $R^2_m=0.409$ and $R^2_c=0.495$, which is good but not anywhere near a perfect fit.%
+\footnote{See below for potentially missing variables.}
+This is also obvious from the prediction accuracy of 0.843, which corresponds to a reduction of the prediction error of 0.297.
+With the added inter-speaker variability brought into play by the experimental setup, a perfect fit cannot be expected.
+Like many phenomena in German which are often called \textit{Zweifelsfälle} (`cases of doubt') in the traditional literature \citep{Duden09,Klein2009} the MNP alternation is a case where speakers very often have no clear intuition and a lot of free variation seems to be involved.
+Rarely do speakers feel that one alternative is clearly odd or bad when presented in isolation.
+Additionally, cases of doubt involving the genitive are often a matter of fierce public debate, and especially the forced choice experiment does not prevent participants from making normative judgements.
+This might account for the slightly better fit in the self-paced reading experiment, although the difference might be too small to warrant an interpretation.
+Thus, the present study shows that what counts as convergence between corpus and experimental results should be gauged considering the nature of the phenomenon at hand, the source of the corpus data, and the experimental paradigm.
+
+Clearly, more case studies using diverse and different corpora are needed, and it should become standard practice to cross-validate them using experimental methods.
+Given that a single failure to achieve convergence does not provide any evidence for divergence, many more studies need to be published to the point that meta-analysis becomes possible.%
+\footnote{This is even more vital considering the variation in results from experimental work.
+See, for example, the impressive list of different reading time results from ten papers on Chinese relative clause processing in \citet[8]{Vasishth2015}.}
+
+On a larger methodological scale, this paper also makes a number of contributions.
+The statistical model was a true multilevel model (see Section~\ref{sec:itemandexemplareffects} and Section~\ref{sec:mainstudystatisticalmodel}), demonstrating how multilevel modelling helps to specify complex hierarchies of abstract features, item-specific tendencies, and exemplar effects at the level of observations (sentences; first level) and lemmas (second level).
+While \cite{Gries2015} still calls mixed models ``underused'' in corpus linguistics, multilevel models are at least equally underused tools.
+At the same time, the effects of overparametrisation of mixed models with varying slopes as criticised by \cite{BatesEa2015a} was demonstrated.
+Furthermore, fitting an additive model to the reading time data did not improve the fit as there were no non-linearities in the data.
+While \cite{DivjakEa2016} also use attested sentences, they find that an additive model helped to deal with non-linearities.
+This is clearly another area where only more studies can lead to clarification.
+Finally, much like \cite{Dabrowska2014} found that speakers' knowledge of collocations was not matched by a set of standard measures of collocation strength extracted from corpora, a simple quotient based on raw frequencies for the attraction strength performed much better in the present study compared to collexeme strength (see Section~\ref{sec:corpusstudies}).
+While this does not allow the conclusion that collexeme strength has no cognitive reality, it might indicate that for measuring attraction effects exerted by non-alternating constructions on alternating constructions, other measures might be more appropriate.
+
+Besides the methodological aspects mentioned above, future work could extend the results presented here to provide a more general picture of the constructions expressing measurements (see Section~\ref{sec:descriptive}).
 This would be a much more complicated task given that the choices then would no longer be binary and that the meaning of the alternative ways of expressing measurements are semantically more varied.
-Finally, I want to point out that German is mildly under-researched in the specific framework used here.
-This is quite surprising given the fact that German morpho-syntax is famous for its alternations, which are usually called \textit{Zweifelsfälle} (`cases of doubt') in the traditional literature \citep{Duden09,Klein2009}.
-Instead of being drowned in normative, descriptive, or didactic discussions, they could serve as ideal test cases in cognitive linguistics.
+Also, a more in-depth look at stylistic or register effects might be possible using new Biber-style annotations which will be released with a new version of the DECOW corpus soon (see Section~\ref{sec:variablesandannotation}.
+Also, regional variation as a potentially influencing factor was already mentioned in Section~\ref{sec:prototypeeffects}.
+As it would be infeasible to examine this with existing corpora, experimental work with participants from various regions might be the ideal approach.
+A reviewer also pointed out that strongly lexicalised adjective-noun combinations like \textit{schwarzer Tee} `black tea' might have a tendency to occur in the \NACa\ because the have more compound-like qualities, blocking strong case inflection in between.
+While this potential effect was impossible to incorporate into the present study, it could be integrated into future research on the phenomenon.
+
+In closing, I want to point out that the so-called \textit{case of doubt} in German morpho-syntax -- like the measure noun phrase alternation -- are in fact ideal test cases for probabilistic modelling of alternation phenomena in cognitive linguistics.
+Doing more research on them could help to provide answers to many of the fundamental and methodological issues mentioned above.

### 2018-01-03 Latex/article.tex
Corrections in 3+4. New abstract.

@@
 
+\usepackage{lineno}
+
 \begin{document}
+
+%\runninglinenumbers
+%\modulolinenumbers[1]
+%\setlength\linenumbersep{20pt}
 
@@
   %\subtitle{...}
-  \abstract{In this paper, a case alternation in German measure noun phrases is examined from the perspective of cognitively oriented corpus linguistics.
-    In certain pseudo-partitive constructions, the embedded kind-denoting noun either agrees in its case with the measure head noun (\textit{eine Tasse guter Kaffee} `a cup of good coffee') or it stands in the genitive (\textit{eine Tasse guten Kaffees}).
-    My analysis is formulated in terms of Prototype Theory.
-    I assume that the choice of alternants is influenced by non-alternating neighbouring constructions representing the prototypes more directly.
-    I argue that the frequencies with which individual lemmas occur in these prototypical non-alternating constructions partially predict which alternant is chosen and that the genitive alternant is prototypical of strongly grammaticalised measure nouns.
-    I present a large-scale corpus study using the DECOW corpus to support this theory.
-    In the statistical analysis, I compare Maximum Likelihood estimators to Bayesian estimators (recently proposed for similar studies), showing that results \textit{predictably} do not differ.
-    Finally, two experiments (forced choice and self-paced reading) are reported.
-    The usage-based findings show a clear correlation with the behaviour of cognitive agents in both experiments.
-    The present study therefore contributes to the well-established field of research into alternations using corpus and experimental methods.}
-  \keywords{corpus methods and experimental methods, alternations, hierarchical models, pseudo-partitives, German}
+  \abstract{In this paper, an alternation in German measure noun phrases is examined under a varying-abstraction perspective.
+    In a specific measure NP construction, the embedded kind-denoting noun either agrees in case with the measure noun (\textit{eine Tasse guter Kaffee} `a cup of good coffee') or it stands in the genitive (\textit{eine Tasse guten Kaffees}).
+    Each of the two alternants is syntactically similar to a non-alternating variant.
+    I propose a prototype model which assigns a common prototypical meaning to each of the alternants and its corresponding non-alternating construction.
+    Based on that, I predict that lexical, morpho-syntactic, and stylistic features help to predict the choice of the alternant.
+    A large corpus study is presented which supports this theory.
+    However, in addition to the prototype effects, an exemplar effect is also shown to influence the choice, namely the relative frequencies with which lemmas occur in the non-alternating constructions.
+    I argue that allowing both prototype and exemplar effects is more adequate than following radical prototype or exemplar approaches.
+    It is also verified in two experiments that the corpus-derived model corresponds to the behaviour of native speakers.
+    The weak effect size of the experimental validation is discussed in the context of corpus-based cognitive linguistics and the validation of corpus-derived models.
+  }
+  \keywords{prototypes vs.\ exemplars, corpus methods and experimental validation, alternations, hierarchical models, pseudo-partitives, German}
   %\classification[PACS]{...}
@@
   %\accepted{...}
-  \journalname{Submitted April 2017. Do not quote.}
+  \journalname{Submitted April 2017, revisions submitted January 2018. Do not quote.}
   \journalyear{}

### 2018-01-03 Latex/sections/1intro.tex
Corrections in 3+4. New abstract.

@@
 Much more optimistically compared to these publications dealing with possible frictions between usage and experimental data, \cite{StefanowitschFlach2016} recently proposed a straightforward positivist perspective of corpora as representing the input of an average adult speaker, thus licensing inferences from corpus data to cognitive representations learned from such data under a ``corpus-as-input'' view.
-They state that ``in this wider context, large, register-mixed corpora such as the British National Corpus [\ldots] may not be perfect models of the linguistic experience of adult speakers, but they are reasonably close to the input of an idealized average member of the relevant speech community'' \citep[104]{StefanowitschFlach2016}.%
-\footnote{This, of course, raises the old question of how corpora should be composed to represent what they are intended to represent.
-As points of reference for determining ``representativeness'', at least three major criteria have been used, namely the distribution of texts or text types in the output of all speakers of a language (production-based), the distribution of the relevance of texts or text types in the whole speech community (relevance-based), and the distribution of speakers' exposure to different texts or text types (perception-based).
-See \citet{Biber1993}, \citet{MceneryEa2006}, \citet{Leech2007}, \citet{Hunston2008} for overviews from different perspectives.}
+They state that ``in this wider context, large, register-mixed corpora such as the British National Corpus [\ldots] may not be perfect models of the linguistic experience of adult speakers, but they are reasonably close to the input of an idealized average member of the relevant speech community'' \citep[104]{StefanowitschFlach2016}.
+% \footnote{This, of course, raises the old question of how corpora should be composed to represent what they are intended to represent.
+% As points of reference for determining ``representativeness'', at least three major criteria have been used, namely the distribution of texts or text types in the output of all speakers of a language (production-based), the distribution of the relevance of texts or text types in the whole speech community (relevance-based), and the distribution of speakers' exposure to different texts or text types (perception-based).
+% See \citet{Biber1993}, \citet{MceneryEa2006}, \citet{Leech2007}, \citet{Hunston2008} for overviews from different perspectives.}
 In essence, this would entail that any generalisation extracted from the BNC could be assumed to have some kind of mental reality, which is at least doubtful.
@@
 
-In cognitive science, it is mostly accepted that exemplar theories have greater explanatory power \citep[184]{Vanpaemel2016}, and that abstraction is only needed marginally, if at all.%
-\footnote{The hard empirical evidence in favour of exemplar models is substantial.
-  For example, in \cite{HahnEa2010}, the authors show that subjects even use exemplar similarity over abstract knowledge even when they are given very simple explicit rules to be learned.
-  This is highly relevant because most other studies focus on the learning of implicit rule-based knowledge, which involves many auxiliary assumptions in actual experiments \citep[2]{HahnEa2010}.
-On the other hand, there is evidence that neither theory is fully adequate to model humans' capabilities to form categories.
-For example, \cite{ConawayKurtz2016} show that both prototype theory and exemplar theory fail to explain certain experimental results where subjects learn to generalise beyond the input in a way that cannot be explained by similarity.
-}
+In cognitive science, it is mostly accepted that exemplar theories have greater explanatory power \citep[184]{Vanpaemel2016}, and that abstraction is only needed marginally, if at all.
+% \footnote{The hard empirical evidence in favour of exemplar models is substantial.
+%   For example, in \cite{HahnEa2010}, the authors show that subjects even use exemplar similarity over abstract knowledge even when they are given very simple explicit rules to be learned.
+%   This is highly relevant because most other studies focus on the learning of implicit rule-based knowledge, which involves many auxiliary assumptions in actual experiments \citep[2]{HahnEa2010}.
+% On the other hand, there is evidence that neither theory is fully adequate to model humans' capabilities to form categories.
+% For example, \cite{ConawayKurtz2016} show that both prototype theory and exemplar theory fail to explain certain experimental results where subjects learn to generalise beyond the input in a way that cannot be explained by similarity.
+% }
 Still, various attempts have been made over the past decades to settle the dispute between abstraction-based models (models with rules or prototypes) and exemplar models or to find models which unite the two extremes.

### 2018-01-03 Latex/sections/2descriptive.tex
Done.
Corrections in 3+4. New abstract.

@@
 
-This construction is usually classified as a \textit{Narrow Apposition Construction} \citep{Loebel1986}, henceforth NAC.%
-\footnote{The construction as in (\ref{ex:intro:narrowapposition1}) is also referred to as the \textit{Direct Partitive Construction} for other Germanic languages in which the PGC with the synthetic genitive is not available.
-This nomenclature makes sense in contrast to the \textit{Indirect Partitive Construction} with prepositional linkers translating to \textit{of} -- \ie\ analytic genitives -- in such languages, see \cite{HankamerMikkelsen2008} for Danish.
-For German, this terminology is not distinctive enough, which is why I use the terms NAC and PGC.}
+This construction is usually classified as a \textit{Narrow Apposition Construction} \citep{Loebel1986}, henceforth NAC.
+% \footnote{The construction as in (\ref{ex:intro:narrowapposition1}) is also referred to as the \textit{Direct Partitive Construction} for other Germanic languages in which the PGC with the synthetic genitive is not available.
+% This nomenclature makes sense in contrast to the \textit{Indirect Partitive Construction} with prepositional linkers translating to \textit{of} -- \ie\ analytic genitives -- in such languages, see \cite{HankamerMikkelsen2008} for Danish.
+% For German, this terminology is not distinctive enough, which is why I use the terms NAC and PGC.}
 Notice that the unavailability of the genitive on the kind noun follows independently from a constraint that genitive NPs in German require the presence of some strongly case-marked element (determiner or adjective) in addition to the head noun in order to be acceptable (\citealp{GallmannLindauer1994,Schachtl1989}; see also \citealp[160]{Eisenberg2013b}).
@@
 
-\subsubsection{A closer look at the syntax of the alternating constructions}
+\subsubsection{Syntax of the alternating constructions}
 \label{sec:syntax}
@@
 To show this, the strong\slash weak inflection patterns of adjectives needs to be taken into account.
-In NPs with a strongly inflected determiner, attributive adjectives inflect according to the massively syncretic \textit{weak} pattern.\footnote{The weak pattern has only two suffixes.
-The structural cases (nominative and accusative) in the singular are marked with \textit{-e}, the oblique cases (dative and genitive) as well as the entire plural are marked with \textit{-en}.
-The only exception to this rule is the masculine accusative singular, which is also marked with \textit{-en}.}
+In NPs with a strongly inflected determiner, attributive adjectives inflect according to the massively syncretic \textit{weak} pattern.
+% \footnote{The weak pattern has only two suffixes.
+% The structural cases (nominative and accusative) in the singular are marked with \textit{-e}, the oblique cases (dative and genitive) as well as the entire plural are marked with \textit{-en}.
+% The only exception to this rule is the masculine accusative singular, which is also marked with \textit{-en}.}
 If there is no determiner (as is the case in the alternating constructions), attributive adjectives inflect like determiners themselves.
@@
 \textit{Physical} (abstract precisely measurable units such as \textit{Liter} `litre', \textit{Meter} `metre', \textit{Gramm} `gram'),
-\textit{Container} (\textit{Eimer} `bucket'),
+\textit{Container} (\textit{Tasse} `cup'),
 \textit{Amount} (\textit{Menge} `amount'), 
 \textit{Portion} (natural portions like \textit{Happen} `bite' or \textit{Krümel} `crumb').
+Notice that \textit{Container} is the class containing words like \textit{spoon} or \textit{cup}, which often develop into partially grammaticalised abstract measure nouns.
 The lemmas that did not fit into either of these classes were labelled \textit{Rest}.

### 2018-01-03 Latex/sections/3corpus.tex
Corrections in 3+4. New abstract.

@@
 The overall intercept is -5.370 and reoresents \textit{Cardinal=Yes}, \textit{Measureclass=Physical} and 0 for all (z-transformed) numeric regressors.
-The standard deviation in the random intercepts is 1.157 for \textit{Measurelemma} ($p_{\text{PB}}=0.001$) and 0.923 for \textit{Kindlemma} ($p_{\text{PB}}=0.0.002$).
+The standard deviation in the random intercepts is 1.157 for \textit{Measurelemma} ($p_{\text{PB}}=0.001$) and 0.923 for \textit{Kindlemma} ($p_{\text{PB}}=0.002$).
 Nakagawa \& Schielzeth's pseudo-coefficient of determination is $R_m^2=0.278$ and $R^2_c=0.566$ (see \citealp{Gries2015} for a basic introduction to these $R^2$ measures, or else \citealp{NakagawaSchielzeth2013}).
@@
 To close this section, I now compare the results of the pre-study and the main study.
-Even though the non-alternating constructions are surely subject to additional constraints (especially because they are much less interchangeable than the alternating constructions), the coefficients align neatly in many cases.
+Even though the non-alternating constructions are surely subject to additional constraints (especially because they are not interchangeable like the alternating constructions), the coefficients align neatly in many cases.
 First of all, the intercepts encode a similar overall dominance of the NAC constructions (pre-study: -5.370; main study: -3.548).
@@
 The two studies do not converge with respect to the \textit{Badness} variable, which is not significant in the pre-study.
-It would, however, surprise to achieve perfectly converging results given that even real effects are missed at certain rates.
+It would, however, surprise to achieve perfectly converging results given that even real effects are missed at certain rates in empirical studies.
 In the case of \textit{Badness}, we see that even in the main study, the effect size is small (-0.152), and thus even at the impressive sample size of roughly 5,000 in both studies, the effect might simply be too weak to be detected reliably.
@@
 Thus, the frequency effect for measure lemmas might be swamped by the lemma random intercepts.
-All things considered, the studies have shown that the predicted prototype effects are mostly found in usage data for both the alternating and the non-alternating constructions.
+All things considered, the studies have shown that the predicted prototype effects are found in usage data for both the alternating and the non-alternating constructions.

### 2018-01-03 Latex/sections/4experiments.tex
Corrections in 3+4. New abstract.

@@
 \footnote{Obviously, the experimental data are much too sparse to perform even post-hoc analyses with respect to the single regressors in a statistically and scientifically sound way.}
-It is, in the sense of Section~\ref{sec:cogocl}, an optimal synthesis of data-rich multifactorial corpus studies and experimental validation, and it is one that has been used (although rarely; see \citealp[3--4]{DivjakEa2016a}) since the seminal \citet{BresnanEa2007} paper.
+It is, in the sense of Section~\ref{sec:cogocl}, an optimal synthesis of data-rich multifactorial corpus studies and experimental validation, and it is one that has been used (although rarely; see \citealp[3--4]{DivjakEa2016a}) at least since the seminal \citet{BresnanEa2007} paper.
 
@@
 The higher the probability of the \PGCa\ predicted from usage data, the more often participants chose the \PGCa\ alternant in the forced-choice task.
-A closer look at the results in the form of the spineplot in Figure~\ref{fig:spines} shows, however, that it was likely an idiosyncracy of the sentence with a model prediction between 0.5 and 0.6 which spoiled an otherwise much better correlation.
+A closer look at the results in the form of the spineplot in Figure~\ref{fig:spines} shows, however, that it was likely an idiosyncracy of a single sentence with a probability predicted by the model between 0.5 and 0.6 which spoiled an otherwise much better correlation.
 The problematic sentence with a model prediction of 0.548 is given in (\ref{ex:problem}) in the \PGCa\ variant.
@@
 In retrospect, this stimulus was badly chosen because \textit{heiße Luft} `ado' (literally `hot air') is a fixed metaphorical expression.
-This obviously influences the reactions of participants, and I will return to this in the discussion in Section~\ref{sec:conclusion}.
-Clearly, a revised and improved experiment might lead to a much better fit in future research.
+This obviously influences the reactions of participants, and a revised and improved experiment might lead to a much better fit in future research.
 
@@
 
-The coefficients of determination indicate that there is a noteworthy correlation between the reactions of the subjects and the corpus-derived probabilities (marginal $R^2$ and that there some inter-subject variation (conditional minus marginal $R^2$).
+The coefficients of determination indicate that there is a noteworthy correlation between the reactions of the subjects and the corpus-derived probabilities (marginal $R^2$) and that there some between-subject variation (conditional minus marginal $R^2$).
 

### 2018-01-03 Latex/sections/5conclusions.tex
Corrections in 3+4. New abstract.

@@
 \footnote{Although, as pointed out in Section~\ref{sec:mainstudyinterpretation}, the \textit{Badness} effect is weak and not detected in the pre-study, probably as a consequence of its weakness.}
-The overall convergence between the pre-study and the main study and also the expected result for \textit{Measurecase} given previous research \citep{Zimmer2015} -- although it was only a control variable here, strengthen the validity of this study even without the experimental validation.
+The overall convergence between the pre-study and the main study and also the expected result for \textit{Measurecase} given previous research \citep{Zimmer2015} -- although it was only a control variable here -- strengthen the validity of this study even without the experimental validation.
 
@@
 
-The question now arises whether in the present case convergence was reached, and the answer is a clear yes.
-The predictions made by the corpus-based model were a significant factor, both in the more explicit paradigm (forced choice) and the implicit paradigm (self-paced reading), showing that the model indeed predicts the variant which language users (readers in this case) expect.
+The question now arises whether convergence was reached in the present case.
+The answer is a clear yes.
+The predictions made by the corpus-based model were a significant factor, both in the more explicit paradigm (forced choice) and the implicit paradigm (self-paced reading).
+This shows that the model indeed predicts the variant which language users expect.
 The overall effect as measured in the $R^2$ was weak (roughly 0.2) in both cases, however.
-While part of this could be traced to a suboptimally chosen stimulus, we really need to consider what kind of convergence we expect to see between corpus data and experiments.
-The main corpus study had $R^2_m=0.409$ and $R^2_c=0.495$, which is good but not anywhere near a perfect fit.%
-\footnote{See below for potentially missing variables.}
-This is also obvious from the prediction accuracy of 0.843, which corresponds to a reduction of the prediction error of 0.297.
-With the added inter-speaker variability brought into play by the experimental setup, a perfect fit cannot be expected.
+While part of this could be traced back to a suboptimally chosen stimulus, we really need to consider what kind of convergence we expect to see between corpus data and experiments.
+The main corpus study had $R^2_m=0.409$ and $R^2_c=0.495$, which is good but not anywhere near a perfect fit.
+With the added inter-speaker variability brought into play by the experimental setup (compared to the averaging across thousands of speakers in the corpus study), a perfect fit cannot be expected.
 Like many phenomena in German which are often called \textit{Zweifelsfälle} (`cases of doubt') in the traditional literature \citep{Duden09,Klein2009} the MNP alternation is a case where speakers very often have no clear intuition and a lot of free variation seems to be involved.
-Rarely do speakers feel that one alternative is clearly odd or bad when presented in isolation.
-Additionally, cases of doubt involving the genitive are often a matter of fierce public debate, and especially the forced choice experiment does not prevent participants from making normative judgements.
-This might account for the slightly better fit in the self-paced reading experiment, although the difference might be too small to warrant an interpretation.
+Rarely do speakers feel that one alternative is clearly odd or bad.
+Additionally, cases of doubt involving the genitive are often a matter of fierce public debate, and especially the forced choice paradigm does not effectively prevent participants from making normative judgements.
+This might even account for the slightly better fit in the self-paced reading experiment, where normative considerations are suppressed.
 Thus, the present study shows that what counts as convergence between corpus and experimental results should be gauged considering the nature of the phenomenon at hand, the source of the corpus data, and the experimental paradigm.
-
 Clearly, more case studies using diverse and different corpora are needed, and it should become standard practice to cross-validate them using experimental methods.
@@
 While \cite{Gries2015} still calls mixed models ``underused'' in corpus linguistics, multilevel models are at least equally underused tools.
-At the same time, the effects of overparametrisation of mixed models with varying slopes as criticised by \cite{BatesEa2015a} was demonstrated.
+At the same time, the effects of overparametrisation of mixed models with varying slopes as criticised by \cite{BatesEa2015a} were demonstrated.
 Furthermore, fitting an additive model to the reading time data did not improve the fit as there were no non-linearities in the data.
@@
 
-Besides the methodological aspects mentioned above, future work could extend the results presented here to provide a more general picture of the constructions expressing measurements (see Section~\ref{sec:descriptive}).
-This would be a much more complicated task given that the choices then would no longer be binary and that the meaning of the alternative ways of expressing measurements are semantically more varied.
-Also, a more in-depth look at stylistic or register effects might be possible using new Biber-style annotations which will be released with a new version of the DECOW corpus soon (see Section~\ref{sec:variablesandannotation}.
+Besides the methodological aspects mentioned above, future work could extend the validity and the scope of the results presented here to provide a more general picture of the constructions expressing measurements (see Section~\ref{sec:descriptive}).
+This would be a much more complicated task given that the choices would no longer be binary and that the meaning of the alternative ways of expressing measurements are semantically more varied.
+Also, a more in-depth look at stylistic or register effects might be possible using the new Biber-style annotations which will be released with a new version of the DECOW corpus soon (see Section~\ref{sec:variablesandannotation}).
 Also, regional variation as a potentially influencing factor was already mentioned in Section~\ref{sec:prototypeeffects}.
 As it would be infeasible to examine this with existing corpora, experimental work with participants from various regions might be the ideal approach.
-A reviewer also pointed out that strongly lexicalised adjective-noun combinations like \textit{schwarzer Tee} `black tea' might have a tendency to occur in the \NACa\ because the have more compound-like qualities, blocking strong case inflection in between.
+A reviewer also pointed out that strongly lexicalised adjective-noun combinations like \textit{schwarzer Tee} `black tea' might have a tendency to occur in the \NACa\ because they have more compound-like qualities, blocking strong case inflection in between them.
 While this potential effect was impossible to incorporate into the present study, it could be integrated into future research on the phenomenon.
@@
 In closing, I want to point out that the so-called \textit{case of doubt} in German morpho-syntax -- like the measure noun phrase alternation -- are in fact ideal test cases for probabilistic modelling of alternation phenomena in cognitive linguistics.
-Doing more research on them could help to provide answers to many of the fundamental and methodological issues mentioned above.
+Doing more research on them could help to provide answers to many of the fundamental and methodological issues raised here.

### 2018-01-03 Latex/article.tex
Final fixes and anonymisation.

@@
 
-%\runninglinenumbers
-%\modulolinenumbers[1]
-%\setlength\linenumbersep{20pt}
+\runninglinenumbers
+\modulolinenumbers[1]
+\setlength\linenumbersep{20pt}
 
@@
 
-  \author*[1]{Roland Schäfer}
-  \runningauthor{Roland Schäfer. Submitted April 2017. Do not quote.}
-  %\affil[1]{}
-  \affil[1]{Freie Universität Berlin}
+  %\author*[1]{Roland Schäfer}
+  %\runningauthor{Roland Schäfer. Submitted April 2017. Do not quote.}
+  %\affil[1]{Freie Universität Berlin}
+
   \title{Abstractions and exemplars: the measure noun phrase alternation in German}
@@
   %\subtitle{...}
+
   \abstract{In this paper, an alternation in German measure noun phrases is examined under a varying-abstraction perspective.
@@
   %\accepted{...}
-  \journalname{Submitted April 2017, revisions submitted January 2018. Do not quote.}
-  \journalyear{}
+  %\journalname{Submitted April 2017, revised version submitted January 2018. Do not quote.}
+  %\journalyear{}
   %\journalvolume{...}
@@
 
-\begin{acknowledgement}
-  I thank (in alphabetical order) Felix Bildhauer, Susanne Flach, Elizabeth Pankratz, Samuel Reichert, Ulrike Sayatz, and Christian Zimmer for valuable discussions and comments.
-  Also, I would like to thank the reviewers for Cognitive Linguistics as well as associate editor Dagmar Divjak for insightful comments which helped to improve the quality of this paper significantly.
-  Furthermore, I thank Ulrike Sayatz for helping me to recruit the participants for the experiments.
-  Elizabeth Pankratz thankfully also fixed my English.
-  Finally, I am grateful to my student assistants Kim Maser for her work on the annotation of the concordances and Luise Rißmann for supervising most of the experiments.
-  The research presented here was made possible in part through funding from the \textit{Deutsche Forschungsgemeinschaft} (DFG, personal grant SCHA1916/1-1).
-\end{acknowledgement}
+% \begin{acknowledgement}
+%   I thank (in alphabetical order) Felix Bildhauer, Susanne Flach, Elizabeth Pankratz, Samuel Reichert, Ulrike Sayatz, and Christian Zimmer for valuable discussions and comments.
+%   Also, I would like to thank the reviewers for Cognitive Linguistics as well as associate editor Dagmar Divjak for insightful comments which helped to improve the quality of this paper significantly.
+%   Furthermore, I thank Ulrike Sayatz for helping me to recruit the participants for the experiments.
+%   Elizabeth Pankratz thankfully also fixed my English.
+%   Finally, I am grateful to my student assistants Kim Maser for her work on the annotation of the concordances and Luise Rißmann for supervising most of the experiments.
+%   The research presented here was made possible in part through funding from the \textit{Deutsche Forschungsgemeinschaft} (DFG, personal grant SCHA1916/1-1).
+% \end{acknowledgement}
 

### 2018-01-03 Latex/sections/2descriptive.tex
Final fixes and anonymisation.

@@
 In principle, this paper is about the middle column of Table~\ref{tab:constructions}, \ie\ the syntactic configuration in which two different case patterns are acceptable.
-However, in Section~\ref{sec:analyses}, the outer columns (\NACb\ and \PGCd) will still play a major role when the factors controlling the alternation are discussed.
+However, the outer columns (\NACb\ and \PGCd) will still play a major role when the factors controlling the alternation are discussed.
 
@@
 The models do not differentiate in and of themselves between abstraction and exemplar effects, but they allow researchers to tune the degree of complexity of models, incorporating both types of effects according to their theory and the phenomenon at hand.
-In Section~\ref{sec:corpushierarchicalmodel}, a full multilevel model will therefore be used.
+In Sections~\ref{sec:prestudy} and~\ref{sec:corpushierarchicalmodel}, multilevel models will therefore be used.

### 2018-01-03 Latex/sections/3corpus.tex
Final fixes and anonymisation.

@@
 After all, the main difference between the quotient used here and the logarithmised p-values from a Fisher test is that they take these marginals into account. 
-However, using collexeme strength instead of the raw frequency quotient was tried as an alternative (variables \textit{Kindcollo} and \textit{Measurecollo}); see Section~\ref{sec:corpushierarchicalmodel} for a discussion of the negative results.
+However, using collexeme strength instead of the raw frequency quotient was tried as an alternative (variables \textit{Kindcollo} and \textit{Measurecollo}); see Section~\ref{sec:mainstudystatisticalmodel} for a discussion of the negative results.
 

### 2018-01-03 Latex/sections/4experiments.tex
Final fixes and anonymisation.

@@
 In the two experiments reported in this section, I use probabilities for the alternating constructions calculated for attested material, and I correlate these probabilities with the participants' reactions.
-Thus, a direct link can be established between output material found in corpora and the behaviour of linguistic agents (see also Section~\ref{sec:cogocl}).
+Thus, a direct link can be established between output material found in corpora and the behaviour of linguistic agents.
 Both experiments use sentences containing attested MNPs from the corpus sample (embedded into simplified sentences) as stimuli.
-Also, the probabilities that the corpus-based model assigns to the two alternants in these sentences are used as the main regressor in both studies.
 
@@
 Looking at Table~\ref{tab:variables}, we see one binary factor (\textit{Cardinal}), two three-level factors (\textit{Measurecase} and \textit{Kindgender}) and one five-level factor (\textit{Measureclass}).
-Since the dependent variable is also binary (\textit{Construction}), permuting these factors alone leads to 180 different possible permutations.
+Since the dependent variable is binary (\textit{Construction}), permuting these factors alone leads to 180 different possible permutations.
 On top of this, the numeric variables \textit{Kindattraction} and \textit{Measureattraction} would also have to be tested at at least two or three different values each.
@@
 
-The first experiment tests preferences for constructions explicitly.
-\cite{FordBresnan2013} use the \textit{split-100} task in which participants have to distribute 100 points between the alternatives, assigning more points to more natural sounding alternative.
-In essence, participants distribute a probability mass between two alternants, which is intended to produce more subtle results compared to a two-alternative forced-choice task such as in \cite{Rosenbach2013}, where participants have to choose one of two alternants.
-The split-100 paradigm has been criticised in \cite{ArppeJaervikivi2007}.
-The criticism was reiterated in \cite{DivjakEa2016}, where they use a forced-choice task.
-In \cite{VerhoevenTemme2017}, it was shown that results from forced-choice and split-100 experiments mostly converge (with some numeric intricacies related to the non-linear distribution of preferences for alternants).
-I present a forced-choice experiment in this Section instead of a split-100 experiment not only because of some researchers' critical attitude towards the split-100 task, but also because in a dry run of the experiment, participants complained about the unnaturalness of distributing a probability mass across two alternants and tended to produce ratings of 0 or 100 (or 50, although to a lesser degree).
+The first experiment tests preferences for constructions explicitly in a forced choice task.
+% \cite{FordBresnan2013} use the \textit{split-100} task in which participants have to distribute 100 points between the alternatives, assigning more points to more natural sounding alternative.
+% In essence, participants distribute a probability mass between two alternants, which is intended to produce more subtle results compared to a two-alternative forced-choice task such as in \cite{Rosenbach2013}, where participants have to choose one of two alternants.
+% The split-100 paradigm has been criticised in \cite{ArppeJaervikivi2007}.
+% The criticism was reiterated in \cite{DivjakEa2016}, where they use a forced-choice task.
+% In \cite{VerhoevenTemme2017}, it was shown that results from forced-choice and split-100 experiments mostly converge (with some numeric intricacies related to the non-linear distribution of preferences for alternants).
+% I present a forced-choice experiment in this section instead of a split-100 experiment not only because of some researchers' critical attitude towards the split-100 task, but also because in a dry run of the experiment, participants complained about the unnaturalness of distributing a probability mass across two alternants and tended to produce ratings of 0 or 100 (or 50, although to a lesser degree).
 Participants had to choose between two sentences that differed only in that one contained the \NACa\, and the other one contained the \PGCa.
-The analysis compares the probabilities assigned to the stimuli by the corpus-derived model with the frequency with which participants chose the alternants for the same stimuli.
-
-There were 24 participants (native speakers of German without reading or writing disabilities) aged 19 to 30 living permanently in Berlin, who were recruited from introductory linguistics courses at Freie Universität Berlin.
+The analysis compares the probabilities assigned to the stimuli by the corpus-derived model with the frequency with which participants chose the alternants.
+There were 24 participants (native speakers of German without reading or writing disabilities) aged 19 to 30 living permanently in [NAME OF CITY REDACTED], who were recruited from introductory linguistics courses at [NAME OF UNIVERSITY REDACTED].
 Although the experiment was conducted in the last four weeks of their first semester, participants had no deeper explicit knowledge of linguistics, grammar, or experimental methods.
@@
 \footnote{The document-level variables \textit{Badness} and \textit{Genitives} were set to 0, which is the mean for z-transformed variables.}
-A random intercept and slope were added for the individual sentence (item) in order to catch idiosyncrasies of single sentences.
+A random intercept and slope were added for the individual sentence pair (item) in order to catch idiosyncrasies of single sentences.
 Coefficients were estimated with Maximum Likelihood Estimation (\textit{lmer} function from \textit{lme4}).
@@
 
-A good amount of the variance can be accounted for by idiosyncrasies of single sentences ($\sigma_{\text{Item}}=1.217$).%
+A good amount of the variance can be accounted for by idiosyncrasies of single sentences ($\sigma_{\text{Item}}=1.217$).
 Also, among participants, there are clearly different preferences ($\sigma_{\text{Participant}}=0.412$).
@@
 
-In principle, random slope for \textit{Modelprediction} varying by stimulus could also remedy problems with individual stimuli at least partially.
+In principle, random slope for \textit{Modelprediction} varying by item could also remedy problems with individual stimuli at least partially.
 Therefore, a model with random slopes for \textit{Modelprediction} varying by both random effects (\textit{Participant} and \textit{Item}) was specified and estimated.
-The random slope for participants was added to comply with \citet[257]{BarrEa2013} who predict \textit{catastrophically high Type I error rates} for experimental designs with within-subject manipulations if random effects structures are not kept maximal.
+The random slope for participants was added to comply with \citet[257]{BarrEa2013} who predict ``catastrophically high Type I error rates'' for experimental designs with within-subject manipulations if random effects structures are not kept maximal.
 The coefficient of the fixed effect changed noticeably but not enough to change the interpretation (5.408 relative to an intercept of -1.304), and the marginal $R^2_m$ rises to 0.213 ($R^2_c=0.488$).
@@
 However, the covariance parameters were estimated at -1.0, which is a clear sign that the variance-covariance matrix could not be estimated successfully.
-The same was true for models with only \textit{Item} and \textit{Participant} random slopes.
+The same was true for models with only an \textit{Item} and a \textit{Participant} random slope.
 This is exactly the kind of model overparametrisation criticised in \citet{BatesEa2015a} and \citet{MatuschekEa2017}.
-The available data is simply insufficient to estimate the parameters of the more complex model with varying slopes.%
+The available data are insufficient to estimate the parameters of the more complex model with varying slopes.%
 \footnote{\citet[1]{BatesEa2015a} state: ``We show that failure to converge typically is not due to a suboptimal estimation algorithm, but is a consequence of attempting to fit a model that is too complex to be properly supported by the data, irrespective of whether estimation is based on maximum likelihood or on Bayesian hierarchical modeling with uninformative or weakly informative priors.
@@
 
-In summary, the forced-choice experiment succeeded in corroborating the results from the corpus study in as much as the preferences extracted from usage data correspond to native speakers' choices, but the correlation is weak, likely due to problems with individual test items and\slash or too few data.
+In summary, the forced-choice experiment succeeded in corroborating the results from the corpus study inasmuch as the preferences extracted from usage data correspond to native speakers' choices, but the correlation is weak, likely at least in part due to problems with individual test items and\slash or too sparse data.
 
@@
 The second experiment tests preferences more implicitly.
-It is expected that reading less typical alternants (the one assigned a low probability by the corpus-derived model) in a given context and with given lexical material incurs a processing overhead for the reader (\citealp{Kaiser2013}).
+It is expected that reading less typical alternants in a given context and with given lexical material incurs a processing overhead for the reader (\citealp{Kaiser2013}).
 In this section, a self-paced reading experiment is therefore presented.
@@
 The assignment of members of the individual sentence pairs to the blocks was randomised for each participant individually, as was the order within each block.
-The sentences from each pair of alternants were kept as far apart as possible.
 The fillers also came in pairs such that the second block exclusively contained sentences to which participants had been exposed in the first block in slightly modified form.
@@
 The standard deviation of the participant intercepts is $\sigma_{\text{Participant}}=0.079$ and of the item intercepts $\sigma_{\text{Item}}=0.037$.
-Comparing the full model to a model without the main regressor \textit{Modelprediction} (and consequently also without the interaction with \textit{Construction}) in a PB test gives $\mpPB=0.036$.
+Comparing the full model to a model without the main regressor \textit{Modelprediction} (and consequently also without the interaction with \textit{Construction}) in a PBmodcomp test gives $\mpPB=0.036$.
 The pseudo-coefficients of determination are $R^2_m=0.239$ and $R^2_c=0.346$.
@@
 
-The coefficients of determination indicate that there is a noteworthy correlation between the reactions of the subjects and the corpus-derived probabilities (marginal $R^2$) and that there some between-subject variation (conditional minus marginal $R^2$).
+The coefficients of determination indicate that there is a noteworthy correlation between the reactions of the subjects and the corpus-derived probabilities (marginal $R^2$) and that there is some between-subject variation (conditional minus marginal $R^2$).
 
@@
 The effect plot for the interaction of interest is shown in Figure~\ref{fig:spr:effects}.
-The estimate for the sentences with \NACa\ is obviously imprecise and no significant differences in reading times are observed.
-There is a clearer effect in the sentences with \PGCa, which is also confirmed by the significant results from the bootstrapped confidence intervals (see Table~\ref{tab:exp:spr}) and from the PB test reported above.
+The estimate for the sentences with \NACa\ is obviously imprecise, and no significant differences in reading times are observed.
+There is a clearer effect in the sentences with \PGCa, which is also confirmed by the significant results from the bootstrapped confidence intervals (see Table~\ref{tab:exp:spr}) and from the PBmodcomp test reported above.
 The \PGCa\ brings about an increased reading time, which is plausible because it is the much rarer construction (see Section~\ref{sec:corpusstudies}).
-However, if it occurs in a prototypical context and with prototypical lexical material, reading times drop.
+However, if it occurs in a prototypical context and with typical lexical material, reading times drop.
 This can be seen in the downward slope in the right panel of Figure~\ref{fig:spr:effects}.
@@
 
-In Section~\ref{sec:conclusion}, I take stock and summarise the contribution of the present study to the research on alternations in cognitive linguistics.
+In Section~\ref{sec:conclusion}, I take stock and summarise the contributions of the present study to the research on alternations in cognitive linguistics.

### 2018-01-03 Latex/sections/5conclusions.tex
Final fixes and anonymisation.

@@
 \footnote{Although, as pointed out in Section~\ref{sec:mainstudyinterpretation}, the \textit{Badness} effect is weak and not detected in the pre-study, probably as a consequence of its weakness.}
-The overall convergence between the pre-study and the main study and also the expected result for \textit{Measurecase} given previous research \citep{Zimmer2015} -- although it was only a control variable here -- strengthen the validity of this study even without the experimental validation.
+The overall convergence between the pre-study and the main study and also the expected result for \textit{Measurecase} in line with previous research \citep{Zimmer2015} -- although it was only a control variable here -- strengthen the validity of this study even without the experimental validation.
 
@@
 As discussed in Section~\ref{sec:cogocl}, it is unnecessary to follow an extreme route, be it radical prototype theory or radical exemplar theory given recent developments in cognitive science and (less prominently) cognitive linguistics.
-While it is surely an intentionally overstated comment, \citet[15]{Kapatsinski2014} suggests that ``[i]n the extreme, some speakers’ heads could host exemplar models, and some could contain fairly abstract grammars, and the produced output would be essentially identical''.
-If this is (to a probably less extreme degree) the case, pooled data from corpora might only provide (still highly informative) data averaged over levels of abstraction.
+While it is surely an intentionally overstated comment, \citet[15]{Kapatsinski2014} even suggests that ``[i]n the extreme, some speakers’ heads could host exemplar models, and some could contain fairly abstract grammars, and the produced output would be essentially identical''.
+If this is (to a probably less extreme degree) the case, corpora would only provide (still informative) pooled data averaged over varying levels of abstraction.
 Clearly, more research is required in this direction.
@@
 \begin{quote}
-  There are now a number of published multivariate models that use data, extracted from corpora [\ldots] to predict the choice for one morpheme, lexeme or construction over another.
+  There are now a number of published multivariate models that use data[,] extracted from corpora [\ldots] to predict the choice for one morpheme, lexeme or construction over another.
   However, [\ldots] only a small number of these corpus-based studies have been cross-validated [\ldots].
@@
 
-The question now arises whether convergence was reached in the present case.
+The question now arises whether convergence was reached in the present case or not.
 The answer is a clear yes.
@@
 The overall effect as measured in the $R^2$ was weak (roughly 0.2) in both cases, however.
-While part of this could be traced back to a suboptimally chosen stimulus, we really need to consider what kind of convergence we expect to see between corpus data and experiments.
+While part of this could be traced back to one suboptimally chosen stimulus, we really need to consider what kind of convergence we expect to see between corpus data and experiments.
 The main corpus study had $R^2_m=0.409$ and $R^2_c=0.495$, which is good but not anywhere near a perfect fit.
@@
 Rarely do speakers feel that one alternative is clearly odd or bad.
-Additionally, cases of doubt involving the genitive are often a matter of fierce public debate, and especially the forced choice paradigm does not effectively prevent participants from making normative judgements.
+Additionally, cases of doubt involving the genitive are often a matter of fierce normative public debate, and especially the forced choice paradigm does not effectively prevent participants from making normative judgements.
 This might even account for the slightly better fit in the self-paced reading experiment, where normative considerations are suppressed.
@@
 The statistical model was a true multilevel model (see Section~\ref{sec:itemandexemplareffects} and Section~\ref{sec:mainstudystatisticalmodel}), demonstrating how multilevel modelling helps to specify complex hierarchies of abstract features, item-specific tendencies, and exemplar effects at the level of observations (sentences; first level) and lemmas (second level).
-While \cite{Gries2015} still calls mixed models ``underused'' in corpus linguistics, multilevel models are at least equally underused tools.
+While \cite{Gries2015} still calls mixed models ``underused'' in corpus linguistics, multilevel models are consequently at least equally underused tools.
 At the same time, the effects of overparametrisation of mixed models with varying slopes as criticised by \cite{BatesEa2015a} were demonstrated.

### 2018-01-04 Latex/sections/2descriptive.tex
About to submit revision.

@@
 
-Finally, there are alternative ways of expressing similar quantificational meanings.
-In the variationist tradition, which is strongly influenced by Labovian sociolinguistics, the \textit{principle of accountability} would dictate that proper studies should examine a variationist \textit{variable}, \ie\ all different \textit{ways of saying the same thing} (\citealp{Labov1966}, \citealp{Labov1969}, for an overview see \citealp{Tagliamonte2012}).
-In the case at hand, the variable might be something like \textit{measurement of quantities of substances and collections}.
-I argue that it is fully justified to focus narrowly on \NACa\ and \PGCa\ with their well-defined morpho-syntactic properties, mostly because the alternative constructions are not used in the same range of contexts.
-Two major alternatives might be considered.
-There is an analytic (pseudo-)partitive with \textit{von} `of'.
-It is only available as an alternative to the PGC if the kind noun phrase contains a (definite or indefinite) determiner as in (\ref{ex:analyticalpartitive}).
-
-\begin{exe}
-  \ex\label{ex:analyticalpartitive} 
-  \begin{xlist}
-    \ex[ ]{\label{ex:analyticalpartitive:a}\gll ein Glas von dem roten Wein\\
-                                                a glass of the red wine\\
-                                                \glt a glass of the red wine
-       }
-    \ex[*]{\label{ex:analyticalpartitive:b}\gll ein Glas von rotem Wein\\
-                                                a glass of red wine\\
-                                                \glt a glass of red wine
-       }
-  \end{xlist}
-\end{exe}
-
-This means that the \textit{von} (pseudo-)partitive does not compete with the \NACa\ and the \PGCa.
-In fact, there is not a single context in which they can be interchanged.
-
-Additionally, we find constructions with \textit{voll} (\textit{von})\slash\textit{voller} `full (of)' and \textit{mit} `with' are available, see (\ref{ex:alternatives}).
-
-\begin{exe}
-  \ex\label{ex:alternatives}
-  \begin{xlist}
-  \ex{\label{ex:voll}\gll   ein Glas voll von rotem Wein\\
-                            a glass full of red wine\\
-			    \glt a glass full of red wine}
-  \ex{\label{ex:voller}\gll ein Glas {voll\slash voller} {rotem\slash roter} Wein\\
-                            a glass full-of red wine\\
-			    \glt a glass full of red wine}
-  \ex{\label{ex:mit}\gll    ein Glas mit rotem Wein\\
-                            a glass with red wine\\
-			    \glt a glass with red wine in it\slash a glass filled with red wine}
-  \end{xlist}
-\end{exe}
-
-These construction have very idiosyncratic properties.
-The construction with \textit{voller} is discussed in \cite{Zeldes2018} and the construction with \textit{mit} in \cite{Bhatt1990}.
-They are not semantically equivalent to the NAC and the PGC, and they are only available for a small subset of measure nouns.
-All of these constructions are used only with measure nouns denoting containers, where the whole NP typically refers to the container and never to the quantity contained in it.
-They are incompatible with measure nouns denoting natural portions such as \textit{Schluck} `gulp' or \textit{Haufen} `heap' and strongly grammaticalised nouns such as \textit{Gramm} `gram'.
-This disqualifies them as alternatives to the \NACa\ or \PGCa\ in most contexts, and they are thus clearly not just more ways of saying the same thing. 
-It is therefore reasonable to focus on the two well-defined alternants.
+% Finally, there are alternative ways of expressing similar quantificational meanings.
+% In the variationist tradition, which is strongly influenced by Labovian sociolinguistics, the \textit{principle of accountability} would dictate that proper studies should examine a variationist \textit{variable}, \ie\ all different \textit{ways of saying the same thing} (\citealp{Labov1966}, \citealp{Labov1969}, for an overview see \citealp{Tagliamonte2012}).
+% In the case at hand, the variable might be something like \textit{measurement of quantities of substances and collections}.
+% I argue that it is fully justified to focus narrowly on \NACa\ and \PGCa\ with their well-defined morpho-syntactic properties, mostly because the alternative constructions are not used in the same range of contexts.
+% Two major alternatives might be considered.
+% There is an analytic (pseudo-)partitive with \textit{von} `of'.
+% It is only available as an alternative to the PGC if the kind noun phrase contains a (definite or indefinite) determiner as in (\ref{ex:analyticalpartitive}).
+% 
+% \begin{exe}
+%   \ex\label{ex:analyticalpartitive} 
+%   \begin{xlist}
+%     \ex[ ]{\label{ex:analyticalpartitive:a}\gll ein Glas von dem roten Wein\\
+%                                                 a glass of the red wine\\
+%                                                 \glt a glass of the red wine
+%        }
+%     \ex[*]{\label{ex:analyticalpartitive:b}\gll ein Glas von rotem Wein\\
+%                                                 a glass of red wine\\
+%                                                 \glt a glass of red wine
+%        }
+%   \end{xlist}
+% \end{exe}
+% 
+% This means that the \textit{von} (pseudo-)partitive does not compete with the \NACa\ and the \PGCa.
+% In fact, there is not a single context in which they can be interchanged. 
+% 
+% Additionally, we find constructions with \textit{voll} (\textit{von})\slash\textit{voller} `full (of)' and \textit{mit} `with' are available, see (\ref{ex:alternatives}).
+% 
+% \begin{exe}
+%   \ex\label{ex:alternatives}
+%   \begin{xlist}
+%   \ex{\label{ex:voll}\gll   ein Glas voll von rotem Wein\\
+%                             a glass full of red wine\\
+% 			    \glt a glass full of red wine}
+%   \ex{\label{ex:voller}\gll ein Glas {voll\slash voller} {rotem\slash roter} Wein\\
+%                             a glass full-of red wine\\
+% 			    \glt a glass full of red wine}
+%   \ex{\label{ex:mit}\gll    ein Glas mit rotem Wein\\
+%                             a glass with red wine\\
+% 			    \glt a glass with red wine in it\slash a glass filled with red wine}
+%   \end{xlist}
+% \end{exe}
+% 
+% These construction have very idiosyncratic properties.
+% The construction with \textit{voller} is discussed in \cite{Zeldes2018} and the construction with \textit{mit} in \cite{Bhatt1990}.
+% They are not semantically equivalent to the NAC and the PGC, and they are only available for a small subset of measure nouns.
+% All of these constructions are used only with measure nouns denoting containers, where the whole NP typically refers to the container and never to the quantity contained in it.
+% They are incompatible with measure nouns denoting natural portions such as \textit{Schluck} `gulp' or \textit{Haufen} `heap' and strongly grammaticalised nouns such as \textit{Gramm} `gram'.
+% This disqualifies them as alternatives to the \NACa\ or \PGCa\ in most contexts, and they are thus clearly not just more ways of saying the same thing. 
+% It is therefore reasonable to focus on the two well-defined alternants.
 

### 2018-01-04 R/3_glmm.R

@@ -125,7 +125,7 @@ for (eff in effs) {
   p <- plot(effect(eff, measure.glmm),
             rug=F,
             main = NULL,
-            ylab = "Probability of PGCa",
+            ylab = "Probability of PGCadj",
             colors = c("black", "darkblue")
             )
   print(p)

### 2018-01-04 R/7_exp-spr.R

@@ -142,7 +142,7 @@ rt.all$rt.tt.resid     <- as.numeric(rt.all$t1.rt.resid + rt.all$t1.rt.resid)
 
 
 # Fix factor levels.
-revalue(rt.all$Construction, c("NACa" = "NACadj", "PGCa" = "PGCadj"))
+rt.all$Construction <- revalue(rt.all$Construction, c("NACa" = "NACadj", "PGCa" = "PGCadj"))
 
 
 # ################
@@ -189,7 +189,7 @@ if (save.persistent) sink(paste(out.dir, "results.txt", sep=""), append = T)
 cat("\n\n##### Models w/ random slopes #####\n\n")
 print(summary(rt.model.p1))
 print(summary(rt.model.p2))
-cat("\n\nAs usual, dear R2, if it doesn't converge, it's for a reason, and tweaking to optimiser does not help!")
+cat("\n\nAs usual, dear R2, if it doesn't converge, it's for a reason, and tweaking to optimiser does not help.")
 
 
 

### 2018-01-05 Latex/sections/1intro.tex
Ready to submit.

@@
 
-This paper deals with a morpho-syntactic alternation between two constructions that occurs only in a very specific type of measure noun phrase in German.
+This paper deals with a morpho-syntactic alternation between two constructions which occurs only in a very specific type of measure noun phrase in German.
 By \textit{alternation} I refer to a situation where two or more forms or constructions are available with no clear (but potentially a subtle) difference in acceptability, function, or meaning.
-The study of lexical and constructional alternations has a long history in cognitively oriented corpus linguistics (for example, \citealp{BresnanEa2007,BresnanHay2010,BresnanFord2010,DivjakArppe2013,Gries2015,NessetJanda2010,Wulff2003}).
+The study of lexical and constructional alternations has a long history in cognitively oriented corpus linguistics (for example, \citealp{BresnanEa2007,BresnanHay2010,BresnanFord2010,DivjakArppe2013,Gries2015,NessetJanda2010}).
 This area of research is based on the assumption that language is a probabilistic phenomenon \citep{Bresnan2007} where alternants are chosen neither deterministically nor fully at random.
-Instead, multifactorial models are constructed which incorporate influencing factors from diverse levels, including contextual factors.
+Instead, multifactorial models are constructed which incorporate influencing factors from diverse levels, including lexical and contextual factors.
 The estimation of the model coefficients quantifies the influence that the factors have on the probability that either alternant is chosen.
@@
 Nobody assumes that we can inductively infer mental representations from experiments, which -- as opposed to corpus studies -- even allow for direct access to the cognitive agent and offer much better possibilities to control experimental conditions and nuisance variables.
-Rather, under the standard approach, a theory of cognitive representation is pre-specified.
+Rather, a theory of cognitive representation is pre-specified.
 Then, predictions are derived from this theory \textit{before} the experiment or the corpus study is conducted in order to \textit{test} the theory.
@@
 An overview of this issue was given by \cite{NewmanSorensenduncan2015}, who enumerate a number of studies which show how corpus data and experimental data converge (such as \citealp{BresnanEa2007,DurrantDoherty2010,GriesWulff2005,GriesEa2005}) and a number of studies where the two types of data led to diverging or only partially converging results (such as \citealp{ArppeJaervikivi2007,Dabrowska2014,Mollin2009}).
-When researchers do not achieve convergence, they mostly try to explain this by differentiating between the actual cognitive construct and what the pooled usage data as found in corpora represent, for example \citet[411]{Dabrowska2014}, who lists a long number of possible reasons to explain why subjects in her experiment diverged in their word association preferences from collocation measures extracted from corpora.
+When researchers do not achieve convergence, they mostly try to explain this by differentiating between the actual cognitive construct and what the pooled usage data as found in corpora represent.
+For example, \citet[411]{Dabrowska2014} lists a long number of possible reasons to explain why subjects in her experiment diverged in their word association preferences from collocation measures extracted from corpora.
 In some cases, researchers simply argue for a more adequate statistical analysis to increase the fit between corpus data and subjects' reactions in experiments, for example \cite{DivjakEa2016}, who show that generalised additive models (GAMs) are better suited than generalised linear models (GLMs) in the analysis of their reading-time experiment based on data extracted from previous corpus studies.
@@
 The typical approach in alternation research is to annotate a large number of corpus sentences with linguistic features and to model the probability of the variants being chosen given these features.
-The idea is that a variant is chosen when the influencing features cumulatively assume \textit{prototypical} values for that variant.
-In other words, a variant of prototype theory with features \citep{Rosch1978} could be applied, and some researchers such as \cite{Gries2003}, \cite{NessetJanda2010}, or \cite{Schaefer2016c} indeed commit to prototype theory in alternation modelling.
+The idea is that a variant is chosen when the influencing features cumulatively assume typical values for that variant.
+In other words, a variant of prototype theory with features \citep{Rosch1978} is the appropriate model, also because the features used to feed the model are more often than not abstract high-level linguistic features.
+Some researchers such as \cite{Gries2003}, \cite{NessetJanda2010}, or \cite{Schaefer2016c} indeed commit to prototype theory in alternation modelling.
 Under prototype theory, category membership is defined by similarity to an ideal exemplar, which is usually identified with characterising features (see \citealp{Taylor2008} for an overview).
@@
 As pointed out by \citet[136--137]{Langacker1987}, schemas and prototypes thus fulfil different roles and can be assumed to co-exist.
-A strict exemplar view of language is incompatible, as far as I can see, with Langacker's view of schemas, but any theory of categorisation that allows for at least some kind of abstraction, is not in principled contradiction with it.%
-\footnote{See especially theories with varying intermediate levels of abstraction described below, which seem compatible with Langacker's formation of schema hierarchies through classification by prototype similarity.}
+A strict exemplar view of language is incompatible, as far as I can see, with Langacker's view of schemas, but any theory of categorisation that allows for at least some kind of abstraction, is not in principled contradiction with it.
 In the remainder of this paper, I do not use schemas in my descriptions of the relevant categories, mostly because the aspect of similarity and fuzzy classification is central to my point, and a formulation in terms of schemas would bring about an unnecessarily high degree of abstraction (see \citealp[70--71]{Taylor2003} for a parallel argument).
@@
 Clearly, they have not (or at least not yet) reached mainstream popularity, and it remains to be seen how well they perform on a broader range of questions.}
-It is thus an advantage of much linguistic work on categorisation that it deals with complex and realistically produced data, because this greatly improves the external validity of studies.
-An ideal contribution by cognitive corpus linguists to the research on (levels of) category abstraction in the human mind would thus be to provide analyses which have great external validity and complexity while carefully making sure that (and determining to what extent) these finding correlate with reactions from cognitive agents under more controlled experimental conditions.
+It is thus an advantage of much linguistic work on categorisation that it deals with complex and realistically produced data, because this greatly improves the external validity of studies, although by sacrificing some construct validity.
+An ideal contribution by cognitive corpus linguists to the research on (levels of) category abstraction in the human mind would thus be to provide analyses which have great external validity and complexity while carefully making sure that (and determining to what extent) these finding correlate with reactions from cognitive agents under more controlled experimental conditions, which increases the construct validity.
 
@@
 In Section~\ref{sec:corpusstudies}, I present a corpus study and report a true multilevel generalised linear model with the appropriate hierarchical structure for the hypothesised effects.
-In Section~\ref{sec:experimental}, I test the predictions of the corpus-based model in two experimental paradigms (forced-choice and self-paced reading), showing that they indeed converge, albeit with no more than intermediate effect strength.
+In Section~\ref{sec:experimental}, I test the predictions of the corpus-based model in two experimental paradigms (forced-choice and self-paced reading), showing that they indeed converge, albeit with low effect strength.
 In Section~\ref{sec:conclusion}, I interpret the findings in the light of the issues of cognitive representations and corpus data and the convergence of usage data and experimental data.

### 2018-01-05 Latex/sections/2descriptive.tex
Ready to submit.

@@
 
+In this section, I introduce and illustrate the relevant alternating constructions in Section~\ref{sec:descriptive}.
+I describe the narrowly defined syntactic configuration in which the alternation occurs.
+Then, I develop a model with prototype effects as well as exemplar effects based in existing research and my own theorising in Section~\ref{sec:analyses}.
+
 \subsection{Alternating and non-alternating measure NP constructions}
 \label{sec:descriptive}
-
-In this section, I introduce and illustrate the relevant alternating constructions.
-I describe the narrowly defined syntactic configuration in which the alternation occurs, and I motivate the focus on \textit{only} this narrow range rather than, for example, the whole range of nominal constructions expressing quantities.
-Finally, I develop a model with prototype effects as well as exemplar effects based in existing research and my own theorising.
 
@@
 
-\subsubsection{Neighbouring cases}
+\subsubsection{Minor issues}
 \label{sec:neighbouringcases}
@@
 I now finally turn to some more subtle issues related to the measure noun case alternation in order to delimit the scope of the study.
-First, there is a claim found in some grammars that a generic nominative, accusative, and even dative on the kind noun are used instead of the genitive (\PGCa) or case agreement (\NACa).
+First, there is a claim found in some grammars that a generic nominative, accusative, and even dative on the kind noun can be used instead of the genitive (\PGCa) or case agreement (\NACa).
 Overviews can be found in \cite{Hentschel1993} and \cite{Zimmer2015}.
@@
   \begin{xlist}
-    \ex{\label{ex:uninflectedmeasures:a} zwei Gramm brauner Zucker}
-    \ex{\label{ex:uninflectedmeasures:b}\gll zwei Gramm braunen Zuckers\\
+    \ex{\label{ex:uninflectedmeasures:a} [zwei Gramm [brauner Zucker]\Sub{Nom}]\Sub{Nom}}
+    \ex{\label{ex:uninflectedmeasures:b}\gll [zwei Gramm [braunen Zuckers]\Sub{Gen}]\Sub{Nom}\\
                                          two gram brown sugar\\
@@
 I also review some existing analyses of the alternation and related issues.
-It should be remembered from Section~\ref{sec:syntax} that, thanks to the syntactic ambiguity of strongly inflected adjectives, the \PGCa\ is plausibly a \PGCd\ in disguise and the \NACa\ a \NACb\ in disguise.
 
@@
 With typical examples, the classes are:
-\textit{Physical} (abstract precisely measurable units such as \textit{Liter} `litre', \textit{Meter} `metre', \textit{Gramm} `gram'),
+\textit{Physical} (typically non-referential precisely measurable units such as \textit{Liter} `litre', \textit{Meter} `metre', \textit{Gramm} `gram'),
 \textit{Container} (\textit{Tasse} `cup'),
@@
 \textit{Portion} (natural portions like \textit{Happen} `bite' or \textit{Krümel} `crumb').
-Notice that \textit{Container} is the class containing words like \textit{spoon} or \textit{cup}, which often develop into partially grammaticalised abstract measure nouns.
+Notice that \textit{Container} is the class containing words like \textit{spoon} or \textit{cup}, which often develop into partially grammaticalised physical measure nouns.
 The lemmas that did not fit into either of these classes were labelled \textit{Rest}.
@@
 This means that the statistical model needs to track an abstraction effect and an exemplar-like effect for measure nouns.
-Additionally, controlling for simple lemma frequency effects is always a good idea, and it should be done for kind and measure nouns.
+Additionally, controlling for raw lemma frequency effects is always a good idea, and it should be done for kind and measure nouns.
 While, for example, very frequent kind nouns might preserve the older \PGCa, low-frequency kind nouns might tend to occur more often in the \NACa.
 For measure nouns, high frequency might lead to a higher affinity to the \NACa\ if we assume at least a tendency for highly grammaticalised items to be more frequent.
-Of course, the model must be specified in a way such that we can be sure that any detected frequency effect also goes beyond the semantic effect, which is captured in the semantic classes described in Section~\ref{sec:prototypeeffects}.
+Of course, the model must be specified in a way such that we can be sure that any detected frequency effect also goes beyond the semantic effect which is captured in the semantic classes described in Section~\ref{sec:prototypeeffects}.
 Such frequency effects would also be very difficult to describe as abstractions.
@@
 
-$Pr(y_1=1)$ is the predicted probability that the variable for the construction type in an observation (indexed by $i$) takes on the value 1 ($y_i=1$).
+$Pr(y_i=1)$ is the probability that the variable for the construction type in observation $i$ takes on the value 1 ($y_i=1$).
 The varying intercept $\beta^0_{j[i]}$ adds a constant for group $j$ to which $i$ belongs (encoded here as $j[i]$, a notation borrowed from \citealp{GelmanHill2006}) to the linear term.%
@@
 Any other $\beta$ is a fixed-effect coefficient, and each $x$ is an observation-level variable.
-Here, just one fixed effect is included for illustration purposes, and $x^1$ could be, for example, the variable encoding whether a cardinal modifies the measure noun.
+Here, just one fixed effect is included for illustration purposes.
+$x^1$ could be, for example, the variable encoding whether a cardinal modifies the measure noun.
 The observations (for example, the lines of a concordance) are indexed by $i$ and the groups\slash random intercepts by $j$.

### 2018-01-05 Latex/sections/3corpus.tex
Ready to submit.

@@
 Therefore, it was highly desirable to obtain a sample in which most of the highly frequent actually-occurring combinations of kind nouns and measure nouns were represented.
-I applied a two-stage process in order to obtain a list of actually co-occurring measure nouns and kind nouns.
+I applied a two-stage process in order to obtain a list of co-occurring measure nouns and kind nouns.
 First, I generated a list of the one hundred most frequent genuine mass nouns.
@@
 In order to generate this list, I utilised the fact that a direct sequence of two nouns almost always instantiates the bare-noun NAC if the second noun is a mass noun.
-I therefore searched for all sequences N\Sub{1}N\Sub{2} where N\Sub{2} was one of the mass noun lemmas extracted in in the first step.
+I therefore searched for all sequences N\Sub{1}N\Sub{2} where N\Sub{2} was one of the mass noun lemmas extracted in the first step.
 Then, the resulting 100 lists of noun-noun combinations were each sorted by frequency in descending order and sieved manually to remove erroneous hits.
@@
 While this does not preclude the use of collostructional analysis, it is an open question whether the marginals (usually called the ``expected frequencies'' in the collostructional literature), given the overall frequency of the constructions and the lemmas, are cognitively relevant in this case.
-After all, the main difference between the quotient used here and the logarithmised p-values from a Fisher test is that they take these marginals into account. 
+After all, the main difference between the quotient used here and signed logarithmised p-values from a Fisher test is that they take these marginals into account. 
 However, using collexeme strength instead of the raw frequency quotient was tried as an alternative (variables \textit{Kindcollo} and \textit{Measurecollo}); see Section~\ref{sec:mainstudystatisticalmodel} for a discussion of the negative results.
@@
 
-\subsubsection{Sampling and annotation}
-
 The main study to be reported below deals exclusively with the alternating constructions, as is customary in alternation modelling.
@@
 Therefore, in this section I examine the distribution of the prototypical features in the non-alternating cases to see whether they are in accord with the theoretical predictions.
+
+\subsubsection{Sampling and annotation}
 
@@
 Furthermore, for each regressor, a p-value was obtained by dropping the regressor from the full model, re-estimating the nested model, and comparing it to the full model.
-Instead of inexact Wald approximations and Likelihood Ratio Tests, I used a drop-in bootstrap replacement for the Likelihood Ratio Test from the function \textit{PBmodcomp} from the \textit{pbkrtest} package \citep{HalekohHojsgaard2014}.
+Instead of inexact Wald approximations and Likelihood Ratio Tests, I used a drop-in bootstrap replacement for the Likelihood Ratio Test in the form of the function \textit{PBmodcomp} from the \textit{pbkrtest} package \citep{HalekohHojsgaard2014}.
 I call the corresponding value $p_{\text{PB}}$, and it is given in the respective columns in Table~\ref{tab:smalltable}.
@@
 
-The main prototypical factors \textit{Cardinal} and \textit{Measureclass} bot pass the PBmodcomp test at sig=0.05 and play the expected role.
-The abstract non-referential physical measure nouns (such as \textit{Gramm} `gram' or \textit{Liter} `litre') with a high degree of grammaticalisation favour the \NACb.
+The main prototypical factors \textit{Cardinal} and \textit{Measureclass} both pass the PBmodcomp test at sig=0.05 and play the expected role.
+The non-referential physical measure nouns (such as \textit{Gramm} `gram' or \textit{Liter} `litre') with a high degree of grammaticalisation favour the \NACb.
 At the other end of the scale, nouns denoting natural portions like \textit{Haufen} `heap', \textit{Bündel} `bundle', \textit{Schluck} `gulp' favour the \PGCd.
@@
   \includegraphics[width=0.85\textwidth]{../R/output/corpus_fixeffs}
-  \caption{Coefficients with 95\% confidence intervals (for details see text); the intercept is -4.328}
+  \caption{Coefficients with 95\% confidence intervals (for details see text); the intercept is -3.548}
   \label{fig:fixeffs}
@@
 Increased \textit{Badness} of the document favours the \NACa, and so does a lower density of genitives.
-While these are merely proxies to style (and partially circular in the case of \textit{Genitives}), this result can at least encourage future work into stylistic effects.
+While these are merely proxies to style, this result can at least encourage future work into stylistic effects.
 
@@
 To close this section, I now compare the results of the pre-study and the main study.
-Even though the non-alternating constructions are surely subject to additional constraints (especially because they are not interchangeable like the alternating constructions), the coefficients align neatly in many cases.
+Even though the non-alternating constructions are surely subject to additional constraints, the coefficients align neatly in many cases.
 First of all, the intercepts encode a similar overall dominance of the NAC constructions (pre-study: -5.370; main study: -3.548).
@@
 The levels of \textit{Measureclass} have comparable coefficients, although the divergence is larger.
-In both studies, the abstract measure nouns in the \textit{Physical} class (on the intercept) are most clearly associated with the NAC constructions.
+In both studies, the non-referential measure nouns in the \textit{Physical} class (on the intercept) are most clearly associated with the NAC constructions.
 Also, in both cases, the \textit{Container} class is closest to \textit{Physical} with the same sign and magnitude (pre-study: 0.445; main study: 0.252).

### 2018-01-05 Latex/sections/4experiments.tex
Ready to submit.

@@
 
-As stimuli, attested MNPs from the corpus study were used, but the sentences were radically simplified to avoid influences from contextual nuisance factors as much as possible.
+As stimuli, attested MNPs from the corpus study were used, but the sentences were simplified to avoid influences from contextual nuisance factors as much as possible.
 The approach is also justified because according to the theoretical assessment in Section~\ref{sec:analyses}, the choice of alternants depends mostly on a very local constructional context.
@@
 
-The coefficients of determination indicate a weak result for the fixed effects part of the model, which is nonetheless worthy of mention (very close to 0.2).
+The marginal $R^2$ indicates a weak result for the fixed effects part of the model, which is nonetheless worthy of mention (close to 0.2).
 The effect display for the single fixed regressor \textit{Modelprediction} is given in Figure \ref{fig:afc:effects}.

### 2018-01-05 Latex/sections/5conclusions.tex
Ready to submit.

@@
 While it is surely an intentionally overstated comment, \citet[15]{Kapatsinski2014} even suggests that ``[i]n the extreme, some speakers’ heads could host exemplar models, and some could contain fairly abstract grammars, and the produced output would be essentially identical''.
-If this is (to a probably less extreme degree) the case, corpora would only provide (still informative) pooled data averaged over varying levels of abstraction.
+If this is the case (even to a less extreme degree), corpora only provide (still informative) pooled data averaged over speaker grammars with varying levels of abstraction.
 Clearly, more research is required in this direction.
@@
 With the added inter-speaker variability brought into play by the experimental setup (compared to the averaging across thousands of speakers in the corpus study), a perfect fit cannot be expected.
-Like many phenomena in German which are often called \textit{Zweifelsfälle} (`cases of doubt') in the traditional literature \citep{Duden09,Klein2009} the MNP alternation is a case where speakers very often have no clear intuition and a lot of free variation seems to be involved.
+Like many phenomena in German which are often called \textit{Zweifelsfälle} (`cases of doubt') in the traditional literature \citep{Duden09,Klein2009} the MNP alternation is one of the cases where speakers very often have no clear intuition and a lot of free variation seems to be involved.
 Rarely do speakers feel that one alternative is clearly odd or bad.
@@
 Clearly, more case studies using diverse and different corpora are needed, and it should become standard practice to cross-validate them using experimental methods.
-Given that a single failure to achieve convergence does not provide any evidence for divergence, many more studies need to be published to the point that meta-analysis becomes possible.%
+Given that a single failure to achieve convergence does not provide conclusive evidence for divergence, many more studies need to be published to the point that meta-analysis becomes possible.%
 \footnote{This is even more vital considering the variation in results from experimental work.
@@
 
-Besides the methodological aspects mentioned above, future work could extend the validity and the scope of the results presented here to provide a more general picture of the constructions expressing measurements (see Section~\ref{sec:descriptive}).
-This would be a much more complicated task given that the choices would no longer be binary and that the meaning of the alternative ways of expressing measurements are semantically more varied.
-Also, a more in-depth look at stylistic or register effects might be possible using the new Biber-style annotations which will be released with a new version of the DECOW corpus soon (see Section~\ref{sec:variablesandannotation}).
+Besides the methodological aspects mentioned above, future work could extend the validity of the results presented here through a more in-depth look at stylistic or register effects, for example using the new Biber-style annotations \citep{Biber1988} which will be released with a new version of the DECOW corpus soon.
 Also, regional variation as a potentially influencing factor was already mentioned in Section~\ref{sec:prototypeeffects}.

### 2018-01-06 Latex/sections/2descriptive.tex
Removed DOIs from bibliography.

@@
 Finally, it was found that the \PGCa\ is more typical of higher stylistic levels (distinctly edited, closer to the non-regional standard, more formal) and\slash or even exclusive to written language (see \citealp[320--323]{Hentschel1993}).
-This is not surprising, given that the genitive -- an intrinsic part of the PGC -- is generally underrepresented in colloquial vernacular variants of German as a result of a diachronic process wherein many (but by no means all) uses of the genitive are being replaced by other cases or periphrastic constructions \citep{FleischerSchallert2011}.
+The genitive -- an intrinsic part of the PGC -- is rarer in colloquial vernacular variants of German compared to the written standard.
+This is the result of a diachronic process wherein some (but by no means all) uses of the genitive are being replaced by other cases or periphrastic constructions \citep{FleischerSchallert2011}.
 Under an integral view of prototypes, which incorporates effects related to larger contexts and styles, such preferences can be part of what defines the construction prototypes, and the \PGCa\ should occur proportionally more often in more elaborate styles closer to the standard.%

### 2018-01-06 Latex/sections/3corpus.tex
Mini fix.

@@
 
-Using signed logarithmised collexeme strength (\textit{Measurecollo} and \textit{Kindcollo}) (with smoothing to avoid \textit{log(0)} errors) instead of the quotient for the attraction strength (\textit{Measureattraction} and \textit{Kindattraction}) (see Section~\ref{sec:variablesandannotation}) was not successful.
+Using signed logarithmised collexeme strength (\textit{Measurecollo} and \textit{Kindcollo}) (with smoothing to avoid arithmetic problems) instead of the quotient for the attraction strength (\textit{Measureattraction} and \textit{Kindattraction}) (see Section~\ref{sec:variablesandannotation}) was not successful.
 While the attraction measures reach satisfying sig levels in the PBmodcomp test (0.020 for \textit{Kindattraction} and 0.001 for \textit{Measureattraction}), the $p_{\text{PB}}$ value for \textit{Kindcollo} was 0.191 and the one for \textit{Measurecollo} was 0.443.

### 2018-01-06 Latex/sections/4experiments.tex
Mini fix.

@@
 The analysis compares the probabilities assigned to the stimuli by the corpus-derived model with the frequency with which participants chose the alternants.
-There were 24 participants (native speakers of German without reading or writing disabilities) aged 19 to 30 living permanently in [NAME OF CITY REDACTED], who were recruited from introductory linguistics courses at [NAME OF UNIVERSITY REDACTED].
+There were 24 participants (native speakers of German without reading or writing disabilities) aged 19 to 30 living permanently in [name of city redacted], who were recruited from introductory linguistics courses at [name of university redacted].
 Although the experiment was conducted in the last four weeks of their first semester, participants had no deeper explicit knowledge of linguistics, grammar, or experimental methods.

### 2018-01-11 Latex/article.tex
About to submit revision.
Fixing replies to comments.
Minor corrections.

@@
 
-\runninglinenumbers
-\modulolinenumbers[1]
-\setlength\linenumbersep{20pt}
+% \runninglinenumbers
+% \modulolinenumbers[1]
+% \setlength\linenumbersep{20pt}
 
@@
     In a specific measure NP construction, the embedded kind-denoting noun either agrees in case with the measure noun (\textit{eine Tasse guter Kaffee} `a cup of good coffee') or it stands in the genitive (\textit{eine Tasse guten Kaffees}).
-    Each of the two alternants is syntactically similar to a non-alternating variant.
+    Each of the two alternants is syntactically similar to a non-alternating construction.
     I propose a prototype model which assigns a common prototypical meaning to each of the alternants and its corresponding non-alternating construction.
-    Based on that, I predict that lexical, morpho-syntactic, and stylistic features help to predict the choice of the alternant.
+    Based on this, I argue that lexical, morpho-syntactic, and stylistic features help to predict the choice of the alternant.
     A large corpus study is presented which supports this theory.

### 2018-01-11 Latex/sections/1intro.tex
Minor corrections.

@@
 Clearly, they have not (or at least not yet) reached mainstream popularity, and it remains to be seen how well they perform on a broader range of questions.}
-It is thus an advantage of much linguistic work on categorisation that it deals with complex and realistically produced data, because this greatly improves the external validity of studies, although by sacrificing some construct validity.
+It is thus an advantage of much linguistic work on categorisation that it deals with complex and realistically produced data because this greatly improves the external validity of studies, although by sacrificing some construct validity.
 An ideal contribution by cognitive corpus linguists to the research on (levels of) category abstraction in the human mind would thus be to provide analyses which have great external validity and complexity while carefully making sure that (and determining to what extent) these finding correlate with reactions from cognitive agents under more controlled experimental conditions, which increases the construct validity.

### 2018-04-08 Latex/article.tex
Before radical shortening.

--- 2018-01-03 Latex/article.tex
@@
 % \usepackage[firstpage]{draftwatermark}
-% \SetWatermarkFontSize{30pt}
+% \SetWatermarkFontSize{24pt}
 % \SetWatermarkHorCenter{300pt}
-% \SetWatermarkColor[rgb]{0.9,0.5,0.5}
-% \SetWatermarkText{\textsf{Submitted April 2017. Do not quote.}}
+% \SetWatermarkColor[rgb]{0.9,0.2,0.2}
+% \SetWatermarkText{\textsf{Revision 2}}
 
@@
 
-%\runninglinenumbers
-%\modulolinenumbers[1]
-%\setlength\linenumbersep{20pt}
+% \runninglinenumbers
+% \modulolinenumbers[1]
+% \setlength\linenumbersep{20pt}
 
@@
   \author*[1]{Roland Schäfer}
-  \runningauthor{Roland Schäfer. Submitted April 2017. Do not quote.}
-  %\affil[1]{}
+  \runningauthor{Roland Schäfer}
   \affil[1]{Freie Universität Berlin}
+
   \title{Abstractions and exemplars: the measure noun phrase alternation in German}
@@
   %\subtitle{...}
+
   \abstract{In this paper, an alternation in German measure noun phrases is examined under a varying-abstraction perspective.
     In a specific measure NP construction, the embedded kind-denoting noun either agrees in case with the measure noun (\textit{eine Tasse guter Kaffee} `a cup of good coffee') or it stands in the genitive (\textit{eine Tasse guten Kaffees}).
-    Each of the two alternants is syntactically similar to a non-alternating variant.
+    Each of the two alternants is syntactically similar to a non-alternating construction.
     I propose a prototype model which assigns a common prototypical meaning to each of the alternants and its corresponding non-alternating construction.
-    Based on that, I predict that lexical, morpho-syntactic, and stylistic features help to predict the choice of the alternant.
+    Based on this, I argue that lexical, morpho-syntactic, and stylistic features help to predict the choice of the alternant.
     A large corpus study is presented which supports this theory.
@@
   %\accepted{...}
-  \journalname{Submitted April 2017, revisions submitted January 2018. Do not quote.}
-  \journalyear{}
+  %\journalname{...}
+  %\journalyear{}
   %\journalvolume{...}

### 2018-04-08 Latex/sections/1intro.tex
Before radical shortening.

@@
 While prototype theory is well suited for modeling constructional choices, it is just one of at least two major similarity-based theories of classification, the most prominent other framework being exemplar theory (\citealp{MedinSchaffer1978,Hintzman1986}).
-Prototype theory and exemplar theory model essentially the same types of effects but differ significantly in whether they assume higher-level abstractions in the form of single maximally prototypical exemplars or their features (prototype theory) or assume that categories emerge through the storage of many exemplars and similarity classification on those exemplars (exemplar theory).
-Similarity-based categorisation is seen as the central and conceptually sufficient approach to linguistic categorisation by many cognitive linguists; see for example \cite{Taylor2003} for a comprehensive treatment in terms of prototype theory.
+Prototype theory and exemplar theory model essentially the same types of effects when only output data are considered.
+Corpus data (representing purely output data, usually from many speakers) might serve as evidence that some form of similarity-based classification is appropriate.
+However, the theories differ significantly in whether they assume higher-level abstractions in the form of single maximally prototypical exemplars or their features (prototype theory) or assume that categories emerge through the storage of many exemplars and similarity classification on those exemplars (exemplar theory).
+Similarity-based categorisation is seen as the central and conceptually sufficient approach to linguistic categorisation by many cognitive linguists.
+For example, see \cite{Taylor2003} for a comprehensive treatment in terms of prototype theory and \cite{Taylor2012} for a detailed discussion of an exemplar-based approach.
 In the influential framework of cognitive grammar \citep{Langacker1987}, prototypes (which, as should be remembered, represent abstractions already in cognitive science) are literally taken as prototypical exemplars, and there is an additional level of fully discrete abstractions in the form of schemas.

### 2018-04-08 Latex/sections/2descriptive.tex
Before radical shortening.

@@
 This unusual double nature of adjectives in NPs without determiners leads to a plausible interpretation of the pattern shown in Table~\ref{tab:constructions}.
-Whenever speakers classify the adjective in the kind noun phrase more as a determiner, they use the \PGCa\ because, if there is a determiner, the PGC is the only option.
-When they classify the adjective more as an adjective, the kind NP has no determiner, and they use the \NACa.%
+If the adjective is classified as a determiner by virtue of carrying the inflectional markers which are otherwise characteristic of determiners, the \PGCa\ is the appropriate choice.
+However, if it is classified as an adjective, the \NACa\ suggests itself because of the lack of a determiner.%
 \footnote{While the generative analysis presented in \cite{Bhatt1990} cannot properly deal with probabilistic effects, Bhatt comes close to this interpretation by analysing the kind NP in the PGC as a DP and in the NAC as an NP.}
@@
 The pseudo-partitive stage finally instantiates a \textit{quantity-of} relation, potentially even leading to fully grammaticalised quantifiers such as \textit{a lot}.
-In German, the PGC is clearly the older construction \citep{Zimmer2015}.
+In German, the two (now clearly distinct) available constructions have emerged diachronically from a single source through a complex reanalysis process, and the PGC is clearly the older construction \citep[2--4]{Zimmer2015}.
 As predicted by the grammaticalisation pattern just described, it still has the potential to form a true partitive (if the kind noun is definite).

### 2018-04-08 R/3_glmm.R

@@ -136,7 +136,7 @@ for (eff in effs) {
 
 
 # Selective ranef plots.
-opts.dotchart <- list(pch=19, col="black", cex=1, xlab="Estimate of intercept")
+opts.dotchart <- list(pch=19, col="black", cex=1, xlab="Prediction of conditional mode")
 n.select <- 30
 main.dotchart.meas <- "Meas. rand. eff."
 main.dotchart.kind <- "Kind rand. eff."

### 2018-04-08 R/7_exp-spr.R

@@ -313,7 +313,7 @@ rt.gamm <- gamm(tf.rt.resid ~ s(Position) + s(Modelprediction, by = Construction
   niterPQL = 1000
 )
 
-# ... with NO different resultcompared to LMM.
+# ... with NO different result compared to LMM.
 # Notice that the relevant smoother (only significant term)
 # is virtually linear.
 

### 2018-04-08 Latex/sections/1intro.tex
First round of shortening.

@@
 See also \citet{Newman2011} for a very positive view of corpora as a source of data in cognitive linguistics in their own right.
-However, as \citet[486--487]{Dabrowska2016} convincingly argues, this does not mean that we can in some way ``deduce mental representations from patterns of use'', \ie\ from corpus data.%
-\footnote{Epistemologically, I believe she refers to acts of \textit{induction} rather than \textit{deduction}.}
+However, as \citet[486--487]{Dabrowska2016} convincingly argues, this does not mean that we can in some way ``deduce mental representations from patterns of use'', \ie\ from corpus data.
 It would be highly surprising if this were possible, and the same holds for experimental methods, albeit to a different degree.
@@
 They state that ``in this wider context, large, register-mixed corpora such as the British National Corpus [\ldots] may not be perfect models of the linguistic experience of adult speakers, but they are reasonably close to the input of an idealized average member of the relevant speech community'' \citep[104]{StefanowitschFlach2016}.
-% \footnote{This, of course, raises the old question of how corpora should be composed to represent what they are intended to represent.
-% As points of reference for determining ``representativeness'', at least three major criteria have been used, namely the distribution of texts or text types in the output of all speakers of a language (production-based), the distribution of the relevance of texts or text types in the whole speech community (relevance-based), and the distribution of speakers' exposure to different texts or text types (perception-based).
-% See \citet{Biber1993}, \citet{MceneryEa2006}, \citet{Leech2007}, \citet{Hunston2008} for overviews from different perspectives.}
 In essence, this would entail that any generalisation extracted from the BNC could be assumed to have some kind of mental reality, which is at least doubtful.
@@
 In cognitive science, it is mostly accepted that exemplar theories have greater explanatory power \citep[184]{Vanpaemel2016}, and that abstraction is only needed marginally, if at all.
-% \footnote{The hard empirical evidence in favour of exemplar models is substantial.
-%   For example, in \cite{HahnEa2010}, the authors show that subjects even use exemplar similarity over abstract knowledge even when they are given very simple explicit rules to be learned.
-%   This is highly relevant because most other studies focus on the learning of implicit rule-based knowledge, which involves many auxiliary assumptions in actual experiments \citep[2]{HahnEa2010}.
-% On the other hand, there is evidence that neither theory is fully adequate to model humans' capabilities to form categories.
-% For example, \cite{ConawayKurtz2016} show that both prototype theory and exemplar theory fail to explain certain experimental results where subjects learn to generalise beyond the input in a way that cannot be explained by similarity.
-% }
 Still, various attempts have been made over the past decades to settle the dispute between abstraction-based models (models with rules or prototypes) and exemplar models or to find models which unite the two extremes.

### 2018-04-08 Latex/sections/2descriptive.tex
First round of shortening.

@@
 This construction is usually classified as a \textit{Narrow Apposition Construction} \citep{Loebel1986}, henceforth NAC.
-% \footnote{The construction as in (\ref{ex:intro:narrowapposition1}) is also referred to as the \textit{Direct Partitive Construction} for other Germanic languages in which the PGC with the synthetic genitive is not available.
-% This nomenclature makes sense in contrast to the \textit{Indirect Partitive Construction} with prepositional linkers translating to \textit{of} -- \ie\ analytic genitives -- in such languages, see \cite{HankamerMikkelsen2008} for Danish.
-% For German, this terminology is not distinctive enough, which is why I use the terms NAC and PGC.}
 Notice that the unavailability of the genitive on the kind noun follows independently from a constraint that genitive NPs in German require the presence of some strongly case-marked element (determiner or adjective) in addition to the head noun in order to be acceptable (\citealp{GallmannLindauer1994,Schachtl1989}; see also \citealp[160]{Eisenberg2013b}).
 
-It is difficult to determine whether the bare kind noun in the narrow apposition construction as in (\ref{ex:intro:narrowapposition1}) bears no case at all, a generic case, or agrees in case with the measure noun.
+It is difficult to determine whether the bare kind noun in the NAC as in (\ref{ex:intro:narrowapposition1}) bears no case at all, a generic case, or agrees in case with the measure noun.
 When there is an adjective as in (\ref{ex:intro:alternation2}), the embedded kind NP clearly agrees in case, but due to the overall absence of markers of case in the singular, bare nouns mostly show no indication of their case.
-The only nouns which do have case markers in the singular are the so-called weak nouns \citep{Koepcke1995,Schaefer2016c}, which have something like a non-nominative \textit{-en} marker in the singular.
-Unfortunately, there are no genuine mass nouns among the weak nouns.
-However, a few of them can be coerced into a mass noun, such as \textit{Hase} `rabbit' (meaning `rabbit meat').
-It then appears as if the uninflected form is preferred, but the inflected form is not excluded, at least for some speakers.
-In (\ref{ex:macabre1}), the clearly acceptable form \textit{Hase} can only be a nominative singular or caseless.
-In (\ref{ex:macabre2}), the form \textit{Hasen} could be an accusative, dative, or genitive.
-
-\begin{exe}
-  \ex\label{ex:macabre}
-  \begin{xlist}
-    \ex[ ]{\label{ex:macabre1}\gll Niemand will [ein Stück [Hase]\Sub{Nom\slash caseless}]\Sub{Acc} essen.\\
-    nobody wants a piece rabbit eat\\
-  \trans Nobody wants to eat a piece of rabbit.}
-    \ex[?]{\label{ex:macabre2}\gll Niemand will [ein Stück [Hasen]\Sub{Acc\slash Dat\slash Gen}]\Sub{Acc} essen.\\
-    nobody wants a piece rabbit eat\\}
-  \end{xlist}
-\end{exe}
-
-Even a full unacceptability of (\ref{ex:macabre2}) would not be conclusive, however, as a possible aversion of speakers towards the case-marked form might be due to the fact that it is at least potentially also a genitive, in which case the constraint against bare genitive nouns would apply.
-With plural kind nouns, the obligatory marking of all dative plurals with \textit{-en} (except those where this is phonotactically impossible) might provide some clues.
-However, plural kind nouns do not behave like singular ones in measure phrases, as will be argued in Section~\ref{sec:neighbouringcases}.
-Also, judgements vary between (\ref{ex:datpl1}) and (\ref{ex:datpl2}), and both are found in corpora.%
-\footnote{For example, the variant in (\ref{ex:datpl1}) with the lemmas \textit{Sack} and \textit{Apfel} occurs four times, and the one in (\ref{ex:datpl2}) twice in the very large DECOW corpus (see Section~\ref{sec:gettingdata}).
-Similarly, for [\textit{Kiste} [\textit{Äpfel}]\Sub{Nom\slash Acc\slash Gen}]\Sub{Dat} `box of apples' (no case identity), I find three examples, and [\textit{Kiste} [\textit{Äpfeln}]\Sub{Dat}]\Sub{Dat} (clearly marked case identity), I find six examples.
-}
-
-\begin{exe}
-  \ex\label{ex:datpl} 
-  \begin{xlist}
-  \ex{\label{ex:datpl1}\gll mit [zwei Säcken [Äpfel]\Sub{Nom\slash Acc\slash Gen}]\Sub{Dat}\\
-     with a sack apples\\
-    \trans with a sack of apples}
-    \ex{\label{ex:datpl2}\gll mit [zwei Säcken [Äpfeln]\Sub{Dat}]\Sub{Dat}\\
-    with a sack apples\\}
-  \end{xlist}
-\end{exe}
-
 Descriptive grammars seem to favour an analysis in terms of caselessness (for example, \citealp[1981]{ZifonunEa1997c}).
-The hazy picture of the case of bare kind NPs is most likely due to the fact that case is so sparsely marked on nouns in contemporary German, where case is marked mostly on determiners and to some degree adjectives.
-The uncertainty in the few cases where case can be marked (weak nouns in the singular and dative plurals as described above) would thus be a direct consequence of the fact that the construction is not very specific with respect to the case of the kind noun.
+Since the concrete case status of the kind NP in the NAC is not relevant to the argument presented here, the question is left open.
 
@@
 In NPs with a strongly inflected determiner, attributive adjectives inflect according to the massively syncretic \textit{weak} pattern.
-% \footnote{The weak pattern has only two suffixes.
-% The structural cases (nominative and accusative) in the singular are marked with \textit{-e}, the oblique cases (dative and genitive) as well as the entire plural are marked with \textit{-en}.
-% The only exception to this rule is the masculine accusative singular, which is also marked with \textit{-en}.}
 If there is no determiner (as is the case in the alternating constructions), attributive adjectives inflect like determiners themselves.
@@
 In Section~\ref{sec:analyses}, I will argue that the NAC and the PGC constructions also have semantically distinguishable prototypes, and that the alternation between \PGCa\ and \NACa\ is an alternation between these prototypes.
-
-\subsubsection{Minor issues}
-\label{sec:neighbouringcases}
-
-I now finally turn to some more subtle issues related to the measure noun case alternation in order to delimit the scope of the study.
-First, there is a claim found in some grammars that a generic nominative, accusative, and even dative on the kind noun can be used instead of the genitive (\PGCa) or case agreement (\NACa).
-Overviews can be found in \cite{Hentschel1993} and \cite{Zimmer2015}.
-Sentence (\ref{ex:intro:gennom}) shows a putative generic nominative on the kind noun inside an accusative MNP.
-
-\begin{exe}
-    \ex[*]{\label{ex:intro:gennom} Wir trinken [[eine Tasse]\Sub{Acc} [heißer Kaffee]\Sub{Nom}]\Sub{Acc}.}
-\end{exe}
-
-It was shown empirically in \cite{Hentschel1993} that such sentences are de facto not acceptable, and the results in \citet{Zimmer2015} are in line with her findings.
-Also, in my corpus sample, they simply did not occur.
-Even if they are accepted by some speakers, their extremely low frequency makes it virtually impossible to study them using corpus linguistic methods (Section~\ref{sec:corpusstudies}), and I consequently do not discuss them further.
-
-Second, we find that, if the kind noun is a plural count noun as in \textit{ein Sack kleine Äpfel} (\NACa) or \textit{ein Sack kleiner Äpfel} (\PGCa) `a bag of small apples', a similar alternation between PGC and NAC can be observed.
-In line with experimental results reported in \citet[15--16]{Zimmer2015}, I found that the PGC is so dominant with plural kind nouns (794 of 861 cases, or over 92\%, cf.\ Section~\ref{sec:corpusstudies}) that the alternation cannot be analysed in the same way as in the singular.
-While this will play a role in the interpretation of the corpus findings, MNPs with plural kind nouns will not be included in the corpus study and the experiments reported in Sections~\ref{sec:corpusstudies} and \ref{sec:experimental}.
-
-Third, some measure nouns have been grammaticalised in a way that they always appear in their non-inflected form.
-They are typical measure nouns like \textit{Gramm} `gram', \textit{Pfund} `pound' or \textit{Prozent} `percent', which have no normal plural forms at all.%
-\footnote{Plurals like \textit{Pfunde} and \textit{Prozente} have special meanings and have very restricted uses, mostly in idiomatic expressions such as \textit{Pfunde verlieren} `lose pounds' and \textit{Prozente machen} `make a profit'.
-They cannot be used in normal MNPs.}
-I treat these cases like other measure nouns because they enter into both the \NACa\ as in (\ref{ex:uninflectedmeasures:a}) and the \PGCa\ as in (\ref{ex:uninflectedmeasures:b}).
-In Section~\ref{sec:analyses}, however, degrees of grammaticalisation as a factor influencing the alternation will be discussed prominently.
-
-\begin{exe}
-  \ex\label{ex:uninflectedmeasures}
-  \begin{xlist}
-    \ex{\label{ex:uninflectedmeasures:a} [zwei Gramm [brauner Zucker]\Sub{Nom}]\Sub{Nom}}
-    \ex{\label{ex:uninflectedmeasures:b}\gll [zwei Gramm [braunen Zuckers]\Sub{Gen}]\Sub{Nom}\\
-                                         two gram brown sugar\\
-				         two grams of brown sugar}
-  \end{xlist}
-\end{exe}
-
-% Finally, there are alternative ways of expressing similar quantificational meanings.
-% In the variationist tradition, which is strongly influenced by Labovian sociolinguistics, the \textit{principle of accountability} would dictate that proper studies should examine a variationist \textit{variable}, \ie\ all different \textit{ways of saying the same thing} (\citealp{Labov1966}, \citealp{Labov1969}, for an overview see \citealp{Tagliamonte2012}).
-% In the case at hand, the variable might be something like \textit{measurement of quantities of substances and collections}.
-% I argue that it is fully justified to focus narrowly on \NACa\ and \PGCa\ with their well-defined morpho-syntactic properties, mostly because the alternative constructions are not used in the same range of contexts.
-% Two major alternatives might be considered.
-% There is an analytic (pseudo-)partitive with \textit{von} `of'.
-% It is only available as an alternative to the PGC if the kind noun phrase contains a (definite or indefinite) determiner as in (\ref{ex:analyticalpartitive}).
-% 
-% \begin{exe}
-%   \ex\label{ex:analyticalpartitive} 
-%   \begin{xlist}
-%     \ex[ ]{\label{ex:analyticalpartitive:a}\gll ein Glas von dem roten Wein\\
-%                                                 a glass of the red wine\\
-%                                                 \glt a glass of the red wine
-%        }
-%     \ex[*]{\label{ex:analyticalpartitive:b}\gll ein Glas von rotem Wein\\
-%                                                 a glass of red wine\\
-%                                                 \glt a glass of red wine
-%        }
-%   \end{xlist}
-% \end{exe}
-% 
-% This means that the \textit{von} (pseudo-)partitive does not compete with the \NACa\ and the \PGCa.
-% In fact, there is not a single context in which they can be interchanged. 
-% 
-% Additionally, we find constructions with \textit{voll} (\textit{von})\slash\textit{voller} `full (of)' and \textit{mit} `with' are available, see (\ref{ex:alternatives}).
-% 
-% \begin{exe}
-%   \ex\label{ex:alternatives}
-%   \begin{xlist}
-%   \ex{\label{ex:voll}\gll   ein Glas voll von rotem Wein\\
-%                             a glass full of red wine\\
-% 			    \glt a glass full of red wine}
-%   \ex{\label{ex:voller}\gll ein Glas {voll\slash voller} {rotem\slash roter} Wein\\
-%                             a glass full-of red wine\\
-% 			    \glt a glass full of red wine}
-%   \ex{\label{ex:mit}\gll    ein Glas mit rotem Wein\\
-%                             a glass with red wine\\
-% 			    \glt a glass with red wine in it\slash a glass filled with red wine}
-%   \end{xlist}
-% \end{exe}
-% 
-% These construction have very idiosyncratic properties.
-% The construction with \textit{voller} is discussed in \cite{Zeldes2018} and the construction with \textit{mit} in \cite{Bhatt1990}.
-% They are not semantically equivalent to the NAC and the PGC, and they are only available for a small subset of measure nouns.
-% All of these constructions are used only with measure nouns denoting containers, where the whole NP typically refers to the container and never to the quantity contained in it.
-% They are incompatible with measure nouns denoting natural portions such as \textit{Schluck} `gulp' or \textit{Haufen} `heap' and strongly grammaticalised nouns such as \textit{Gramm} `gram'.
-% This disqualifies them as alternatives to the \NACa\ or \PGCa\ in most contexts, and they are thus clearly not just more ways of saying the same thing. 
-% It is therefore reasonable to focus on the two well-defined alternants.
-
-This concludes the descriptive overview of the phenomenon.
-I have demonstrated that there is an alternation between two measure noun constructions in a narrow syntactic configuration (kind NP with an adjective but without a determiner), and that the two constructions differ in the case of the kind noun (case agreement with the measure noun or genitive).
-I turn to a more theory-oriented discussion of the alternation in the next section.
 
@@
 While the \PGCd\ in (\ref{ex:prototypes1}) allows an interpretation where speakers conceptualise the substance itself, \ie\ the medication, and a spoon used to take a quantity from the medication, the \NACb\ in (\ref{ex:prototypes2}) does not allow such an interpretation.
-While in the given examples, the effect is strongly supported by the choice of the measure noun lemmas, I argue that the meaning of the prototypes is independent of this.%
-\footnote{A reviewer asked for a specification of the \textit{schematic meaning} of the superordinate constructions in terms of \citet{Langacker1987}.
-However, none of the constructions (\PGCd, \PGCa, \NACb, and \NACa) are exclusively associated with one of the discussed meanings, and thus I do not see how one could specify schemas (even at a superordinate level), for which ``membership is not a matter of degree'' and whose properties are ``fully compatible with all the members of the category'' \citep[371]{Langacker1987}.
-A modelling in terms of a probabilistic similarity effects is more appropriate.
-See also the argument in \citet[70--71]{Taylor2003} and my discussion in Section~\ref{sec:cogocl}.}
+While in the given examples, the effect is strongly supported by the choice of the measure noun lemmas, I argue that the meaning of the prototypes is independent of this.
 The independent meanings of the prototypes are, as I will show, a result of diachronic developments and grammaticalisation processes.
@@
 This is the result of a diachronic process wherein some (but by no means all) uses of the genitive are being replaced by other cases or periphrastic constructions \citep{FleischerSchallert2011}.
-Under an integral view of prototypes, which incorporates effects related to larger contexts and styles, such preferences can be part of what defines the construction prototypes, and the \PGCa\ should occur proportionally more often in more elaborate styles closer to the standard.%
-\footnote{A reviewer mentioned that she or he had the impression that dialectal variation is also a factor influencing the alternation.
-This is definitely true, as some dialects (such as Alemannic) tend to have no genitive at all.
-While this is an interesting aspect for further research, the main obstacle for the present study is that there are no corpora of German which are both large enough and annotated with reliable metadata about regional variants.
-In the annotation of the corpus study, documents obviously written in a regional variant were excluded.}
+Under an integral view of prototypes, which incorporates effects related to larger contexts and styles, such preferences can be part of what defines the construction prototypes, and the \PGCa\ should occur proportionally more often in more elaborate styles closer to the standard.
 
@@
 With all this, it must also be ensured that these influences at the lemma level (lemma type frequency, attraction strength, and semantic classes) are not spurious and just artefacts of mere lemma idiosyncrasies.
-This leads to a rather complex and truly multilevel model structure for the generalised linear mixed models (GLMMs) customarily used in alternation modelling.%
+This leads to a rather complex and truly multilevel model structure for the generalised linear mixed models (GLMMs).%
 \footnote{See \citet{Gries2015} for an argument in favour of varying-slope and varying-intercept models in corpus linguistic studies.
@@
 $Pr(y_i=1)$ is the probability that the variable for the construction type in observation $i$ takes on the value 1 ($y_i=1$).
-The varying intercept $\beta^0_{j[i]}$ adds a constant for group $j$ to which $i$ belongs (encoded here as $j[i]$, a notation borrowed from \citealp{GelmanHill2006}) to the linear term.%
-\footnote{The linear term is the argument to the $logit^{-1}$ link function which transforms it into something which is interpretable as a probability.}
+The varying intercept $\beta^0_{j[i]}$ adds a constant for group $j$ to which $i$ belongs (encoded here as $j[i]$, a notation borrowed from \citealp{GelmanHill2006}) to the linear term.
 Any other $\beta$ is a fixed-effect coefficient, and each $x$ is an observation-level variable.
@@
 This says that the intercepts follow a normal distribution with a standard deviation of $\sigma$ and a mean predicted from the model $\gamma_0+\gamma^1\cdot u_j$, where $\gamma^0$ is the second-level intercept, the other $\gamma$ are the second-level coefficients, and each $u$ is a second-level predictor variable.
-If $\beta^0_{j}$ were a lemma random effect, then $u_j$ would be the lemma frequency predictor which helps to predict the lemma intercept instead of just having a simple per-lemma constant.%
-\footnote{Since the standard \textit{lme4} package takes care of multilevel modelling automatically, an R formula for (\ref{eq:glmm}) and (\ref{eq:level2}) could be \texttt{Construction\textasciitilde Cardinal+Lemmafrequency+(1|Lemma)} given that in the data set, the values of \texttt{Lemmafrequency} are unique for each \texttt{Lemma}.}
+If $\beta^0_{j}$ were a lemma random effect, then $u_j$ would be the lemma frequency predictor which helps to predict the lemma intercept instead of just having a simple per-lemma constant.
 

### 2018-04-08 Latex/sections/3corpus.tex
Minor corrections.
First round of shortening.

@@
 At the time of this writing, a newer 2016 version DECOW16A has already been released.}
-I chose this corpus for two main reasons.%
-\footnote{The use of web data for linguistic research does require explicit and careful justification.
-Due to the noisy nature and unknown composition of the web, only carefully designed and established web corpora like the COW corpora or the SketchEngine corpora \citep{KilgarriffEa2014} should be used.
-Clearly, using search engine results is ``bad science'' for many reasons, most prominently total non-replicability of results, as \cite{Kilgarriff2006} pointed out more than ten years ago.
-Careless use of search engine results is still found, however, see for example \citet[171--175]{DeclerckBrems2016}.}
+I chose this corpus for two main reasons.
 First, the external validity of any study is increased through a higher heterogeneity of the sample \citep[30]{MaxwellDelaney2004}, and the DECOW14A corpus has clearly a much more heterogeneous composition compared to the only other very large corpus of German, the DeReKo \citep{KupietzEa2010} of the Institute for the German Language (IDS), which contains almost exclusively newspaper texts.%
@@
   \end{tabular}
-  \caption{Annotated variables for the main study}
+  \caption{Annotated variables for the corpus studies}
   \label{tab:variables}
@@
 Since the \PGCa\ contains a genitive itself, the regressor variable \textit{Genitive} and the document-level variable \textit{Genitives} are not fully independent.
-Since instances of the \PGCa\ make up for only a minute fraction of all genitives, however, I still use \textit{Genitives} as a regressor.%
-\footnote{A reviewer suggested that a fully fledged multidimensional analysis \citep{Biber1988} would help to improve the operationalisation of the influence of style.
-This sounds plausible, but there simply is no large corpus of German for which similar data have been published.
-In the meantime, the creators of COW and the Institut für Deutsche Sprache (IDS) Mannheim have developed a similar annotation framework (COReX), and the COW creators have pre-released a non-public beta version of the corresponding data base for DECOW16A to their users.
-This data base contains 118 automatically extracted lexico-grammatical features for each document.
-However, at this time, it is not recommended for use in published research.
-Also, the genitive is generally considered a good indicator of style in German, and its frequency is usually high enough such that the \textit{Genitives} score is a stable measure even for shorter documents (compared to first or second person pronouns, for example, which are totally absent in a majority of documents).
-The genitive is still the dominant attributive case in German.}
+Since instances of the \PGCa\ make up for only a minute fraction of all genitives, however, I still use \textit{Genitives} as a regressor.
 
@@
 I call the corresponding value $p_{\text{PB}}$, and it is given in the respective columns in Table~\ref{tab:smalltable}.
-Regressors which did not reach sig=0.05 in in the PBmodcomp test (\textit{Kindfreq}, \textit{Kindgender}, and \textit{Measurefreq}) were removed from the model, appear in brackets in Table~\ref{tab:smalltable}, and consequently have no coefficient estimates.%
-\footnote{My approach to statistical inference is essentially Fisherian \citep{Fisher1935,Fisher1959}, and hence I use the term \textit{sig level} instead of the term \textit{$\alpha$ level} known from null hypothesis significance testing (NHST).
-For overviews of systems of statistical inference with special focus on this and related questions, see \citet{Lehmann1993,Lehmann2011} or the shorter \citet{Perezgonzalez2015}.}
-The overall intercept is -5.370 and reoresents \textit{Cardinal=Yes}, \textit{Measureclass=Physical} and 0 for all (z-transformed) numeric regressors.
+Regressors which did not reach sig=0.05 in in the PBmodcomp test (\textit{Kindfreq}, \textit{Kindgender}, and \textit{Measurefreq}) were removed from the model, appear in brackets in Table~\ref{tab:smalltable}, and consequently have no coefficient estimates.
+The overall intercept is -5.370 and represents \textit{Cardinal=Yes}, \textit{Measureclass=Physical} and 0 for all (z-transformed) numeric regressors.
 The standard deviation in the random intercepts is 1.157 for \textit{Measurelemma} ($p_{\text{PB}}=0.001$) and 0.923 for \textit{Kindlemma} ($p_{\text{PB}}=0.002$).
@@
 For the main study, each of the 2,365 noun–noun combinations was queried in the alternating constructions \PGCa\ and \NACa\ in DECOW14A.%
-\footnote{Due to processing considerations with the COW interface at the time, only ten slices of DECOW14A were used, which add up to approximately 10 billion tokens.}
+\footnote{Due to efficiency problems with the old COW interface at the time, only ten slices of DECOW14A were used, which add up to approximately 10 billion tokens.}
 For the manual annotation process, a random sub-sample was generated.
 For each mass noun, the concordance was downsampled randomly to maximally 100 sentences, which resulted in 6,843 sentences without further downsampling.
-In the manual annotation process the size was further reduced to 5,063 sentences due to removal of noisy material, erroneous hits, and uninformative cases where the measure noun was in the genitive, in which case the \NACa\ cannot be distinguished from the \PGCa.
+In the manual annotation process the sample was further reduced to 5,063 sentences due to removal of noisy material, erroneous hits, and uninformative cases where the measure noun was in the genitive, in which case the \NACa\ cannot be distinguished from the \PGCa.
 Given the careful sampling procedure described in this section and Section~\ref{sec:bootstrappinlemmapairs}, we can be highly certain that it contains all relevant and reasonably frequent noun–noun combinations in the target constructions.%
@@
 
-Standard diagnostics show that the model quality is quite good.%
-\footnote{Notice that no aggressive model selection was applied, especially not upward model selection including probing for interactions.
-I am convinced that models should be specified based on theoretical considerations (including known nuisance effects) to avoid the dangers of data dredging and to avoid models which are difficult to interpret.
-Removal of useless regressors (judging by, for example, the PBmodcomp test and coefficients of determination) is the only type of model selection I allow in my research.}
+Standard diagnostics show that the model quality is quite good.
 Nakagawa \& Schielzeth's pseudo-coefficient of determination is $R_m^2=0.409$ and $R^2_c=0.495$.
@@
 First, the prototypicality effect related to the non-alternating \PGCd\ and \NACb\ can be shown (see the effect plots in Figure~\ref{fig:eff:attraction}).%
-\footnote{Effect plots were created using the \textit{effects} package \citep{Fox2003}.
-They show the changes in probability for the outcome (y-axis) dependent on values of a regressor (x-axis) at typical values of all other regressors.
-The vertical bars (categorical variables), and the grey areas (continuous variables) are asymptotic 95\% confidence intervals calculated from \textit{glmer}.
-They are not bootstrapped.
-Readers should be aware that the axes are specifically scaled so as to result in a linear plot, and that the range of the axes varies between plots.}
+\footnote{Effect plots were created using the \textit{effects} package \citep{Fox2003}.}
 The effect is as expected:

### 2018-04-08 Latex/sections/4experiments.tex
First round of shortening.

@@
 The first experiment tests preferences for constructions explicitly in a forced choice task.
-% \cite{FordBresnan2013} use the \textit{split-100} task in which participants have to distribute 100 points between the alternatives, assigning more points to more natural sounding alternative.
-% In essence, participants distribute a probability mass between two alternants, which is intended to produce more subtle results compared to a two-alternative forced-choice task such as in \cite{Rosenbach2013}, where participants have to choose one of two alternants.
-% The split-100 paradigm has been criticised in \cite{ArppeJaervikivi2007}.
-% The criticism was reiterated in \cite{DivjakEa2016}, where they use a forced-choice task.
-% In \cite{VerhoevenTemme2017}, it was shown that results from forced-choice and split-100 experiments mostly converge (with some numeric intricacies related to the non-linear distribution of preferences for alternants).
-% I present a forced-choice experiment in this section instead of a split-100 experiment not only because of some researchers' critical attitude towards the split-100 task, but also because in a dry run of the experiment, participants complained about the unnaturalness of distributing a probability mass across two alternants and tended to produce ratings of 0 or 100 (or 50, although to a lesser degree).
 Participants had to choose between two sentences that differed only in that one contained the \NACa\, and the other one contained the \PGCa.
 The analysis compares the probabilities assigned to the stimuli by the corpus-derived model with the frequency with which participants chose the alternants.
-There were 24 participants (native speakers of German without reading or writing disabilities) aged 19 to 30 living permanently in [name of city redacted], who were recruited from introductory linguistics courses at [name of university redacted].
+There were 24 participants (native speakers of German without reading or writing disabilities) aged 19 to 30 living permanently in Berlin, who were recruited from introductory linguistics courses at Freie Universität Berlin.
 Although the experiment was conducted in the last four weeks of their first semester, participants had no deeper explicit knowledge of linguistics, grammar, or experimental methods.

### 2018-04-08 Latex/sections/5conclusions.tex
First round of shortening.

@@
 
-Besides the methodological aspects mentioned above, future work could extend the validity of the results presented here through a more in-depth look at stylistic or register effects, for example using the new Biber-style annotations \citep{Biber1988} which will be released with a new version of the DECOW corpus soon.
+Besides the methodological aspects mentioned above, future work could extend the validity of the results presented here through a more in-depth look at stylistic or register effects, for example using the new Biber-style annotations \citep{Biber1988} which will be released with a new version of the DECOW corpus soon according to its creators.
 Also, regional variation as a potentially influencing factor was already mentioned in Section~\ref{sec:prototypeeffects}.

### 2018-04-09 Latex/sections/1intro.tex
First version under 13,000 words.

@@
 
-This paper deals with a morpho-syntactic alternation between two constructions which occurs only in a very specific type of measure noun phrase in German.
-By \textit{alternation} I refer to a situation where two or more forms or constructions are available with no clear (but potentially a subtle) difference in acceptability, function, or meaning.
-The study of lexical and constructional alternations has a long history in cognitively oriented corpus linguistics (for example, \citealp{BresnanEa2007,BresnanHay2010,BresnanFord2010,DivjakArppe2013,Gries2015,NessetJanda2010}).
-This area of research is based on the assumption that language is a probabilistic phenomenon \citep{Bresnan2007} where alternants are chosen neither deterministically nor fully at random.
-Instead, multifactorial models are constructed which incorporate influencing factors from diverse levels, including lexical and contextual factors.
+\subsection{Alternations}
+\label{sec:alternations}
+
+This paper deals with a morpho-syntactic alternation between two measure noun constructions in German.
+By \textit{alternation} I refer to a situation where two or more forms or constructions are available with no clear difference in acceptability, function, or meaning.
+The study of lexical and constructional alternations has a long history in cognitive corpus linguistics (for example, \citealp{BresnanEa2007,BresnanHay2010,BresnanFord2010,DivjakArppe2013,Gries2015,NessetJanda2010}).
+This research is based on the assumption that language is a probabilistic phenomenon where alternants are chosen neither deterministically nor fully at random \citep{Bresnan2007}.
+Multifactorial models are constructed which incorporate influencing factors from diverse levels, including lexical and contextual factors.
 The estimation of the model coefficients quantifies the influence that the factors have on the probability that either alternant is chosen.
-There are two fundamental issues to consider with respect to this tradition as a part of cognitive linguistics in the broad sense.
+There are two fundamental issues to consider with respect to this tradition.
 First, there is the question of whether corpus data do provide any insight into cognitive representations at all.
-This question can and should be answered by testing how well corpus-derived models converge with or diverge from experimental findings, which provide more direct evidence of mental representations and processes but are often intrinsically narrower in scope.
-Second and closely related to the first question, the appropriate modelling of such results in cognitive linguistics (\ie\ the assumed underlying constructs) is a key issue.
+This question can and should be answered by testing how well corpus-derived models converge with or diverge from experimental findings, which provide more direct evidence of mental representations and processes (see Secion~\ref{sec:usagedataincognitivelinguistics}).
+Second, the appropriate modelling of such results in cognitive linguistics (\ie\ the assumed constructs) is a key issue (see Section~\ref{sec:prototypesexemplarscorporaandcontrolledexperiments}).
 
-Concerning the first point, there has for a long time been an interest in correlating probabilistic generalisations extracted from corpus data with results from experimental work (for example, \citealp{ArppeJaervikivi2007,BresnanEa2007,BresnanFord2010,DivjakGries2008,DivjakEa2016,FordBresnan2013}).
+\subsection{Usage data in cognitive linguistics}
+\label{sec:usagedataincognitivelinguistics}
+
+For some time, there has been an interest in correlating probabilistic generalisations extracted from corpus data and results from experimental work (for example, \citealp{ArppeJaervikivi2007,BresnanEa2007,BresnanFord2010,DivjakGries2008,DivjakEa2016,FordBresnan2013}).
 This is often called a \textit{validation} of the corpus-derived findings, but \citet[303]{Divjak2016a} criticises this choice of words ``because it creates the impression that behavioral experimental data is inherently more valuable than textual data'', citing \cite{TummersEa2005}, who state that a corpus is ``a sample of spontaneous language use that is (generally) realized by native speakers''.
-See also \citet{Newman2011} for a very positive view of corpora as a source of data in cognitive linguistics in their own right.
-However, as \citet[486--487]{Dabrowska2016} convincingly argues, this does not mean that we can in some way ``deduce mental representations from patterns of use'', \ie\ from corpus data.
-It would be highly surprising if this were possible, and the same holds for experimental methods, albeit to a different degree.
-Nobody assumes that we can inductively infer mental representations from experiments, which -- as opposed to corpus studies -- even allow for direct access to the cognitive agent and offer much better possibilities to control experimental conditions and nuisance variables.
-Rather, a theory of cognitive representation is pre-specified.
-Then, predictions are derived from this theory \textit{before} the experiment or the corpus study is conducted in order to \textit{test} the theory.
-While the same approach is by and large applicable to corpus data, I now discuss some relevant differences.
+See also \citet{Newman2011} for a positive view of corpora as a source of data in cognitive linguistics in their own right.
+However, as \citet[486--487]{Dabrowska2016} argues, this does not mean that we can in some way ``deduce mental representations from patterns of use'', \ie\ from corpus data.
+It would indeed be highly surprising if this were possible, and the same holds for experimental methods.
+Nobody assumes that we can inductively infer mental representations from experiments, which even allow for direct access to the cognitive agent and offer much better ways of controlling experimental conditions and nuisance variables.
+Rather, predictions are derived from existing theories in order to \textit{test} the theory.
+While the same approach is applicable to corpus data, I discuss some relevant differences.
 
-As mentioned above, the central question is whether usage data as found in corpora are truly predictive of speakers' and writers' cognitive representation of language and\slash or of their overall linguistic behaviour; and this is where the experimental validation (or, more neutrally, \textit{corroboration}) comes into play.
-An overview of this issue was given by \cite{NewmanSorensenduncan2015}, who enumerate a number of studies which show how corpus data and experimental data converge (such as \citealp{BresnanEa2007,DurrantDoherty2010,GriesWulff2005,GriesEa2005}) and a number of studies where the two types of data led to diverging or only partially converging results (such as \citealp{ArppeJaervikivi2007,Dabrowska2014,Mollin2009}).
-When researchers do not achieve convergence, they mostly try to explain this by differentiating between the actual cognitive construct and what the pooled usage data as found in corpora represent.
-For example, \citet[411]{Dabrowska2014} lists a long number of possible reasons to explain why subjects in her experiment diverged in their word association preferences from collocation measures extracted from corpora.
-In some cases, researchers simply argue for a more adequate statistical analysis to increase the fit between corpus data and subjects' reactions in experiments, for example \cite{DivjakEa2016}, who show that generalised additive models (GAMs) are better suited than generalised linear models (GLMs) in the analysis of their reading-time experiment based on data extracted from previous corpus studies.
-Much more optimistically compared to these publications dealing with possible frictions between usage and experimental data, \cite{StefanowitschFlach2016} recently proposed a straightforward positivist perspective of corpora as representing the input of an average adult speaker, thus licensing inferences from corpus data to cognitive representations learned from such data under a ``corpus-as-input'' view.
+As mentioned above, the central question is whether usage data as found in corpora are truly predictive of speakers' and writers' cognitive representation of language and\slash or of their linguistic behaviour.
+This is where the experimental validation (or \textit{corroboration}) comes into play.
+An overview was given in \cite{NewmanSorensenduncan2015}, who enumerate a number of studies showing how corpus data and experimental data converge (such as \citealp{BresnanEa2007,DurrantDoherty2010,GriesWulff2005,GriesEa2005}) and a number of studies where the two types of data led to diverging or only partially converging results (such as \citealp{ArppeJaervikivi2007,Dabrowska2014,Mollin2009}).
+When researchers do not achieve convergence, they often try to explain this by differentiating between the actual cognitive construct and whatever the pooled usage data as found in corpora represent.
+For example, \citet[411]{Dabrowska2014} lists a number of possible reasons to explain why subjects in her experiment diverged in their word association preferences from collocation measures extracted from corpora.
+Alternatively, researchers argue for a more adequate statistical analysis to increase the fit between corpus data and experimental data.
+See, for example \cite{DivjakEa2016}, who show that generalised additive models (GAMs) are better suited than generalised linear models (GLMs) for correlating reading times and corpus data. 
+Much more optimistically, \cite{StefanowitschFlach2016} recently proposed a straightforward positivist perspective of corpora as representing the input of an average adult speaker, thus licensing inferences from corpus data to cognitive representations under a ``corpus-as-input'' view.
 They state that ``in this wider context, large, register-mixed corpora such as the British National Corpus [\ldots] may not be perfect models of the linguistic experience of adult speakers, but they are reasonably close to the input of an idealized average member of the relevant speech community'' \citep[104]{StefanowitschFlach2016}.
-In essence, this would entail that any generalisation extracted from the BNC could be assumed to have some kind of mental reality, which is at least doubtful.
+In essence, this entails that any generalisation extracted from the BNC can be assumed to have some kind of mental reality, which is at least doubtful.
 
-Obviously, no clear picture has emerged yet, which is not surprising given the vast number of cognitive constructs assumed at diverse levels (such as the level of words in collocation research and much more complex constructional levels in alternation research involving lexical, syntagmatic, and contextual factors), the problems of corpus composition, the operationalisations involved in making experiments, and the choice of statistical tools.
-While far from providing definitive solutions, my discussion in Section~\ref{sec:conclusion} will provide possible explanations for the goodness of fit between the corpus data and the experimental data reported in Sections~\ref{sec:corpusstudies} and~\ref{sec:experimental}, much in the spirit of \citet{Dabrowska2014}.
+No general consensus has emerged yet, which is not surprising given the number of cognitive constructs assumed at diverse levels, the problems of corpus composition, the operationalisations involved in experiments, and the choice of statistical tools.
+While far from providing definitive solutions, my discussion in Section~\ref{sec:conclusion} will provide possible explanations for the quality of the fit between the corpus data and the experimental data reported in Sections~\ref{sec:corpusstudies} and~\ref{sec:experimental}, much in the spirit of \citet{Dabrowska2014}.
 
-I now turn to the second issue, namely which type of cognitive representation research on alternations provides evidence for.
+\subsection{Prototypes, exemplars, corpora, and controlled experiments}
+\label{sec:prototypesexemplarscorporaandcontrolledexperiments}
+
+I now turn to cognitive representations as constructs in alternation research.
 The typical approach in alternation research is to annotate a large number of corpus sentences with linguistic features and to model the probability of the variants being chosen given these features.
-The idea is that a variant is chosen when the influencing features cumulatively assume typical values for that variant.
-In other words, a variant of prototype theory with features \citep{Rosch1978} is the appropriate model, also because the features used to feed the model are more often than not abstract high-level linguistic features.
+The idea is that a variant is chosen when the influencing features have certain typical values.
+In other words, a variant of prototype theory with features \citep{Rosch1978} is the appropriate model, also because the features used to feed the model are usually abstract high-level linguistic features.
 Some researchers such as \cite{Gries2003}, \cite{NessetJanda2010}, or \cite{Schaefer2016c} indeed commit to prototype theory in alternation modelling.
-Under prototype theory, category membership is defined by similarity to an ideal exemplar, which is usually identified with characterising features (see \citealp{Taylor2008} for an overview).
-While prototype theory is well suited for modeling constructional choices, it is just one of at least two major similarity-based theories of classification, the most prominent other framework being exemplar theory (\citealp{MedinSchaffer1978,Hintzman1986}).
-Prototype theory and exemplar theory model essentially the same types of effects when only output data are considered.
-Corpus data (representing purely output data, usually from many speakers) might serve as evidence that some form of similarity-based classification is appropriate.
-However, the theories differ significantly in whether they assume higher-level abstractions in the form of single maximally prototypical exemplars or their features (prototype theory) or assume that categories emerge through the storage of many exemplars and similarity classification on those exemplars (exemplar theory).
-Similarity-based categorisation is seen as the central and conceptually sufficient approach to linguistic categorisation by many cognitive linguists.
-For example, see \cite{Taylor2003} for a comprehensive treatment in terms of prototype theory and \cite{Taylor2012} for a detailed discussion of an exemplar-based approach.
-In the influential framework of cognitive grammar \citep{Langacker1987}, prototypes (which, as should be remembered, represent abstractions already in cognitive science) are literally taken as prototypical exemplars, and there is an additional level of fully discrete abstractions in the form of schemas.
-Schemas are characterised by the properties common to all members of the category, whereas a prototypical category member might have very specific additional properties not at all shared by all or even most members \citep[371-375]{Langacker1987}.
-The prototype can serve as a reference point when classifying new objects which do not share all properties of the schema, but this would (if repeated) lead to the creation of an even more abstract (hierarchically higher and less specific) schema which describes the new member and the ones belonging to the previous schema.
-As pointed out by \citet[136--137]{Langacker1987}, schemas and prototypes thus fulfil different roles and can be assumed to co-exist.
-A strict exemplar view of language is incompatible, as far as I can see, with Langacker's view of schemas, but any theory of categorisation that allows for at least some kind of abstraction, is not in principled contradiction with it.
-In the remainder of this paper, I do not use schemas in my descriptions of the relevant categories, mostly because the aspect of similarity and fuzzy classification is central to my point, and a formulation in terms of schemas would bring about an unnecessarily high degree of abstraction (see \citealp[70--71]{Taylor2003} for a parallel argument).
-However, I fully describe the abstract features of the prototypes.
+Under prototype theory, category membership is defined by similarity to an ideal exemplar and its characterising features (see \citealp{Taylor2008} for an overview).
 
-Turning back to the prototype vs.\ exemplar debate, \cite{Barsalou1990} already showed that prototype and exemplar theory model the same types of effects and are informationally equivalent.
-Consequently, experiments which favour one theory over the other use procedural behaviour of subjects in experiments, for example the speed of category retrieval, as opposed to mere output data.
-In very early experiments, \cite{PosnerKeele1968} showed, for example, that highly prototypical unseen exemplars were categorised more easily by subjects compared to less prototypical ones which had been included in the learning data.
-This was (at the time) taken as evidence that subjects categorise by prototypes.%
+While prototype theory is well suited for modeling constructional choices, it is just one of at least two major similarity-based theories of classification, the prominent alternative being exemplar theory (\citealp{MedinSchaffer1978,Hintzman1986}).
+Prototype theory and exemplar theory model essentially the same effects when only output data are considered.
+The theories differ significantly in whether they assume higher-level abstractions in the form of single maximally prototypical exemplars (prototype theory) or categories emerging through the storage of many exemplars and similarity-based classification over those sets of exemplars (exemplar theory).
+\cite{Barsalou1990} already showed that prototype and exemplar theory model the same types of surface effects and are informationally equivalent.
+Consequently, experiments which favour one theory over the other use procedural behaviour of subjects in experiments, for example the speed of category retrieval, as opposed to mere output data.%
 \footnote{See \citet{StormsEa2000} for a comparison of the theories in different experimental settings.}
@@
 In cognitive science, it is mostly accepted that exemplar theories have greater explanatory power \citep[184]{Vanpaemel2016}, and that abstraction is only needed marginally, if at all.
-Still, various attempts have been made over the past decades to settle the dispute between abstraction-based models (models with rules or prototypes) and exemplar models or to find models which unite the two extremes.
-\cite{VanpaemelStorms2008} and \citet{LeeVanpaemel2008} proposed the \textit{varying abstraction model} (VAM) which ``attempt[s] to balance economy and informativeness'' \citep[745]{LeeVanpaemel2008}, treating models with full abstraction (radical prototype theory) and no abstraction at all (radical exemplar theory) as special cases of a model which allows for both abstraction and exemplar effects.
-The mixture model of categorisation (MMC) by \cite{Rosseel2002} is a model with abstraction in the form of hierarchical clusters of exemplars, and these clusters of objects are characterised by a probability distribution over their features, and categorising new objects is a process of estimating the probability of this object belonging to one of the clusters.
+Still, various attempts have been made to settle the dispute between abstraction-based models (models with rules or prototypes) and exemplar models.
+\cite{VanpaemelStorms2008} and \citet{LeeVanpaemel2008} proposed the \textit{varying abstraction model} (VAM) which ``attempt[s] to balance economy and informativeness'' \citep[745]{LeeVanpaemel2008}, treating models with full abstraction (radical prototype theory) and no abstraction at all (radical exemplar theory) as special cases of a flexible approach which allows both types of mechanisms.
+The mixture model of categorisation (MMC) by \cite{Rosseel2002} is a model with abstraction in the form of hierarchical clusters of exemplars.
+These clusters of objects are characterised by a probability distribution over their features, and categorising new objects is a process of estimating the probability of this object belonging to one of the clusters.
 \cite{GriffithsEa2009} go further and present a computational model which is able to choose the appropriate complexity of representation for a given category.
-However, despite these (and more) attempts to reconcile or unite the two approaches while developing spelled-out mathematical models, \cite[183--184]{Vanpaemel2016} describes the state of affairs between adherents of neo-prototype theory (such as \citealp{MindaSmith2001,MindaSmith2002}) and exemplar theory as a stalemate.
+However, despite these and more attempts to reconcile or unite the two approaches \citet[183--184]{Vanpaemel2016} describes the state of affairs between adherents of neo-prototype theory (such as \citealp{MindaSmith2001,MindaSmith2002}) and exemplar theory as a stalemate.
 
-In cognitive linguistics, \cite{DivjakArppe2013} is a very rare example of a paper where such issues are taken up with reference to the current research in cognitive science.
+In cognitive linguistics, similarity-based categorisation is often seen as the central and conceptually sufficient approach to linguistic categorisation.
+See \cite{Taylor2003} for a comprehensive treatment in terms of prototype theory and \cite{Taylor2012} for a detailed discussion of an exemplar-based approach.
+\cite{DivjakArppe2013} is a rare example of a paper in cognitive linguistics where a synthesis between prototype and exemplar models is proposed.
 Their corpus-based approach shows ``one way of systematically analyzing usage data as contained in corpora to yield a scheme, compatible with usage-based theories of language, by which the assumptions of both the prototype and exemplar theories can be operationalized'' \citep[267]{DivjakArppe2013}.
@@
 While I am convinced that the results presented in \citet{DivjakArppe2013} are valid and important (especially given the previous and subsequent research the authors have conducted on the data, including experimental work presented in \citealp{DivjakEa2016}), any data set can be clustered to yield a certain number of clusters.
-Thus, the study does not ensure that the clusters emerging from the data correspond to any speaker's cognitive representation.%
-\footnote{See, again, \citet[486--487]{Dabrowska2016} and \citet[22]{Gries2003}.}
+Thus, the study does not ensure that the clusters emerging from the data correspond to any speaker's cognitive representation.
 
-On the other hand, the trade-off one has to accept when doing experiments with highly simplified stimuli and very simple tasks is their lower \textit{external validity} (\ie\ their lower degree of generalisability) and their high dependence on potentially problematic operationalisations of constructs, control of confounding factors in the face of a limited number of available subjects, etc.
-(in other words, critical dependence on \textit{construct validity} and \textit{internal validity}).%
+On the other hand, the trade-off one has to accept when doing experiments with highly simplified stimuli is lower \textit{external validity} (\ie\ decreased generalisability) and higher dependence on proper operationalisations of constructs (\textit{construct validity}).%
 \footnote{An accessible overview of the different types of validity can be found in Chapter~1 of \citet{MaxwellDelaney2004}.}
 Tasks in cognitive science have been criticised exactly for their lack of external validity, for example by \citet{Murphy2003}.
-From a linguistic perspective, it is remarkable in this context that \cite{VoorspoelsEa2011} consider their experimental task -- which is the assignment of typicality scores to nouns from the domains of \textit{animals} and \textit{artefacts} to categories like \textit{bird}, \textit{fish}, \textit{clothing}, or \textit{tools} -- a study of ``superordinate natural language categories, whereas most evidence supporting exemplar representations has been found in artificial categories of a more subordinate level'' \citep[1013]{VoorspoelsEa2011}.
-Corpus linguists interested in probabilistic alternation modelling deal with much more complex high-level categories and use large and complex feature sets, especially in (morpho-)syntax.%
-\footnote{Notice, however, that recently, approaches have emerged which solve at least some problems by abandoning linguistic high-level features altogether \citep{BaayenEa2016,RamscarPort2016}.
-Clearly, they have not (or at least not yet) reached mainstream popularity, and it remains to be seen how well they perform on a broader range of questions.}
+From a linguistic perspective, it seems remarkable that \citet[1013]{VoorspoelsEa2011} consider their experimental task -- the assignment of typicality scores to nouns from the domains of \textit{animals} and \textit{artefacts} to categories like \textit{bird}, \textit{fish}, \textit{clothing}, or \textit{tools} -- a study of ``superordinate natural language categories, whereas most evidence supporting exemplar representations has been found in artificial categories of a more subordinate level''.
+Corpus linguists interested in probabilistic alternation modelling investigate significantly more complex high-level categories and use large and complex feature sets.%
+\footnote{However, approaches have emerged which model linguistic phenomena without reference to high-level linguistic features altogether \citep{BaayenEa2016,RamscarPort2016}.}
 It is thus an advantage of much linguistic work on categorisation that it deals with complex and realistically produced data because this greatly improves the external validity of studies, although by sacrificing some construct validity.
-An ideal contribution by cognitive corpus linguists to the research on (levels of) category abstraction in the human mind would thus be to provide analyses which have great external validity and complexity while carefully making sure that (and determining to what extent) these finding correlate with reactions from cognitive agents under more controlled experimental conditions, which increases the construct validity.
+Thus, an ideal contribution to the research on category abstraction by cognitive corpus linguists is to provide analyses which have great external validity and complexity while carefully making sure that these finding correlate with actions and reactions under more controlled experimental conditions, thus increasing the construct validity.
 
-This paper contributes to this endeavour in many ways.
+This paper contributes to solving the questions raised in this section in many ways.
 After a thorough description of the alternation under examination, I discuss potential influencing factors comprising high-level abstract semantic generalisations as well as exemplar-similarity and item-specific effects in Section~\ref{sec:germanmeasurenps}.
-I will argue that
-there are lemma-specific and exemplar effects but also generalisations at the level of the construction as well as generalisations overlaying the lemma-specific effects, leading to a complex hierarchical structure.
+I will argue that there are lemma-specific and exemplar effects but also generalisations at the level of the construction as well as generalisations overlaying the lemma-specific effects, leading to a complex hierarchical structure.
 In Section~\ref{sec:corpusstudies}, I present a corpus study and report a true multilevel generalised linear model with the appropriate hierarchical structure for the hypothesised effects.

### 2018-04-09 Latex/sections/2descriptive.tex
First version under 13,000 words.

@@
 Under an integral view of prototypes, which incorporates effects related to larger contexts and styles, such preferences can be part of what defines the construction prototypes, and the \PGCa\ should occur proportionally more often in more elaborate styles closer to the standard.
-
-\subsubsection{Items, exemplar effects, and multilevel models}
-\label{sec:itemandexemplareffects}
-
-As discussed in Section~\ref{sec:cogocl}, prototype-based and exemplar-based approaches are merely the endpoints of a spectrum of theories allowing abstraction in cognitive representations to varying degrees.
-While the prototypes for the PGC and the NAC were clearly specified with reference to abstract meaning in Section~\ref{sec:prototypeeffects}, at least one exemplar-driven similarity effect can also be expected to influence the MNP alternation, leading to a mixed model with abstractions as well as exemplar effects.
-The measure and kind noun lemmas which occur in the alternating constructions obviously also occur in the non-alternating constructions.
-The relative frequency with which they occur in these stable cases -- where choosing an alternative is impossible -- could thus be a factor influencing the alternating, less stable case.
-Should this be confirmed, it would be highly implausible to conceive of such an effect as a prototype effect.
-In the corpus study reported in Section~\ref{sec:corpusstudies}, a measure quantifying this influence will therefore be included as the \textit{attraction strength}.
-
-It should be noted that such an exemplar-type effect would have to be confirmed \textit{in addition} to the predicted semantic prototype effect for measure lemmas described in Section~\ref{sec:prototypeeffects}, namely the effect of semantic classes of measure nouns.
-This means that the statistical model needs to track an abstraction effect and an exemplar-like effect for measure nouns.
-Additionally, controlling for raw lemma frequency effects is always a good idea, and it should be done for kind and measure nouns.
-While, for example, very frequent kind nouns might preserve the older \PGCa, low-frequency kind nouns might tend to occur more often in the \NACa.
-For measure nouns, high frequency might lead to a higher affinity to the \NACa\ if we assume at least a tendency for highly grammaticalised items to be more frequent.
-Of course, the model must be specified in a way such that we can be sure that any detected frequency effect also goes beyond the semantic effect which is captured in the semantic classes described in Section~\ref{sec:prototypeeffects}.
-Such frequency effects would also be very difficult to describe as abstractions.
-
-With all this, it must also be ensured that these influences at the lemma level (lemma type frequency, attraction strength, and semantic classes) are not spurious and just artefacts of mere lemma idiosyncrasies.
-This leads to a rather complex and truly multilevel model structure for the generalised linear mixed models (GLMMs).%
-\footnote{See \citet{Gries2015} for an argument in favour of varying-slope and varying-intercept models in corpus linguistic studies.
-See \citet{GelmanHill2006} (especially part 2) for a comprehensive text book on the subject including multilevel models.}
-A GLMM models the influence which several variables (\textit{predictors}) have on the probability that an alternant is chosen (the \textit{response}).
-The so-called \textit{fixed effects} are assumed to be fixed population parameters which quantify the strength and the direction of the effect which, for example, a specific feature of the syntactic context or the stylistic level of the text have on the choice of the alternants.
-The so-called \textit{random effects} are not fixed population parameters, but they vary by group.
-The simplest random effect is a varying intercept, which predicts for groups of observations defined by lemmas, genres, speakers, etc. a constant term to be added to the model.
-In a true multilevel model, however, the varying intercept itself comes with an additional linear model where second-level fixed effects predict the group-level effect, and this is exactly what is needed here.
-To illustrate, the simple model specification in (\ref{eq:glmm}) represents the first level of a multilevel logistic regression model.
-
-\begin{equation}
-  Pr(y_i=1)=logit^{-1}(\beta^0_{j[i]}+\beta^1\cdot x^1_i)
-  \label{eq:glmm}
-\end{equation}
-
-$Pr(y_i=1)$ is the probability that the variable for the construction type in observation $i$ takes on the value 1 ($y_i=1$).
-The varying intercept $\beta^0_{j[i]}$ adds a constant for group $j$ to which $i$ belongs (encoded here as $j[i]$, a notation borrowed from \citealp{GelmanHill2006}) to the linear term.
-Any other $\beta$ is a fixed-effect coefficient, and each $x$ is an observation-level variable.
-Here, just one fixed effect is included for illustration purposes.
-$x^1$ could be, for example, the variable encoding whether a cardinal modifies the measure noun.
-The observations (for example, the lines of a concordance) are indexed by $i$ and the groups\slash random intercepts by $j$.
-Let us say the groups are defined by lemmas and we also include lemma frequency as a control variable.
-In this case, a second-level model would be given similar to (\ref{eq:level2}).
-
-\begin{equation}
-  \beta^0_j\sim\mathcal{N}(\gamma_0+\gamma^1\cdot u_j, \sigma)
-  \label{eq:level2}
-\end{equation}
-
-This says that the intercepts follow a normal distribution with a standard deviation of $\sigma$ and a mean predicted from the model $\gamma_0+\gamma^1\cdot u_j$, where $\gamma^0$ is the second-level intercept, the other $\gamma$ are the second-level coefficients, and each $u$ is a second-level predictor variable.
-If $\beta^0_{j}$ were a lemma random effect, then $u_j$ would be the lemma frequency predictor which helps to predict the lemma intercept instead of just having a simple per-lemma constant.
-
-While multilevel models are rarely used in corpus linguistics -- \citet{Gries2015} even calls the simpler varying-intercepts and varying-slopes models ``underused'' -- they provide an excellent tool to describe situations where preferences at the sentence level and at lexical levels need to be integrated.
-The models do not differentiate in and of themselves between abstraction and exemplar effects, but they allow researchers to tune the degree of complexity of models, incorporating both types of effects according to their theory and the phenomenon at hand.
-In Sections~\ref{sec:prestudy} and~\ref{sec:corpushierarchicalmodel}, multilevel models will therefore be used.

### 2018-04-09 Latex/sections/3corpus.tex
First version under 13,000 words.

@@
 
-Among the factors potentially influencing the alternation (see Section~\ref{sec:germanmeasurenps}) were lemma-specific preference effects.
-Therefore, it was highly desirable to obtain a sample in which most of the highly frequent actually-occurring combinations of kind nouns and measure nouns were represented.
+Among the factors potentially influencing the alternation (see Section~\ref{sec:germanmeasurenps}) are lemma-specific preference effects.
+Therefore, it was desirable to obtain a sample in which most of the highly frequent actually-occurring combinations of kind nouns and measure nouns were represented.
 I applied a two-stage process in order to obtain a list of co-occurring measure nouns and kind nouns.
@@
 
-In the first step, I exported a list of all nouns in the DECOW14A01 sub-corpus sorted by their token frequency and manually went through it from the most frequent noun downwards, selecting the first one hundred mass nouns that occurred in the list.%
+In the first step, I exported a list of all nouns in the DECOW14A01 sub-corpus sorted by their token frequency.
+The first one hundred mass nouns were extracted manually from this list.%
 \footnote{DECOW14A01 is the first slice (roughly a twentieth) of the complete DECOW14A corpus.
-It contains just over one billion tokens.}
+It contains roughly one billion tokens.}
 Mass nouns were defined as concrete nouns which denote a substance genuinely (without coercion), combine with uninflected mass quantifiers such as \textit{viel} `much' and \textit{wenig} `little' (\textit{viel Bier} `much beer'), and form only sortal and unit plurals (such as the plural \textit{Biere} `types of beer' or `glasses of beer').
@@
 In order to generate this list, I utilised the fact that a direct sequence of two nouns almost always instantiates the bare-noun NAC if the second noun is a mass noun.
-I therefore searched for all sequences N\Sub{1}N\Sub{2} where N\Sub{2} was one of the mass noun lemmas extracted in the first step.
+I searched for all sequences N\Sub{1}N\Sub{2} where N\Sub{2} was one of the mass noun lemmas extracted in the first step.
 Then, the resulting 100 lists of noun-noun combinations were each sorted by frequency in descending order and sieved manually to remove erroneous hits.
@@
 The full set of manually annotated variables for the main study is given in Table~\ref{tab:variables}.%
-\footnote{All numeric variables were z-transformed (\ie\ centered to the mean and rescaled such that they have a standard deviation of 1) to facilitate their interpretation in the regression models.}
+\footnote{All numeric variables were z-transformed to facilitate their interpretation in the regression models.}
 Notice first that \textit{Construction} is the response variable with the values \textit{PGCadj} and \textit{NACadj}.
@@
 The variables \textit{Kindattraction} and \textit{Measureattraction} encode the ratio with which a given kind noun lemma or measure noun lemma occurs in the \PGCd\ and the \NACb.
-They were calculated from auxiliary samples to be described in Section~\ref{sec:mainstudysampling} as a log-transformed quotient.
-The higher the value, the more often the noun occurs in the \PGCd\ (proportionally).
-It could be argued that other measures of attraction strength could be used, for example those popularised in collostructional analysis (\citealp{StefanowitschGries2003,GriesStefanowitsch2004}; see also \citealp{Gries2015a}).
+They were calculated from auxiliary samples described in Section~\ref{sec:mainstudysampling} as a log-transformed quotient.
+The higher the value, the more often the noun occurs in the \PGCd.
+It could be argued that other measures of attraction strength should be used, for example those popularised in collostructional analysis (CA; \citealp{StefanowitschGries2003,GriesStefanowitsch2004}; see also \citealp{Gries2015a}).
 However, the goal here is to quantify how often lemmas occur in the \PGCd\ and the \NACb, and these constructions do not compete at all but are rather mutually exclusive.
-While this does not preclude the use of collostructional analysis, it is an open question whether the marginals (usually called the ``expected frequencies'' in the collostructional literature), given the overall frequency of the constructions and the lemmas, are cognitively relevant in this case.
-After all, the main difference between the quotient used here and signed logarithmised p-values from a Fisher test is that they take these marginals into account. 
-However, using collexeme strength instead of the raw frequency quotient was tried as an alternative (variables \textit{Kindcollo} and \textit{Measurecollo}); see Section~\ref{sec:mainstudystatisticalmodel} for a discussion of the negative results.
+While this does not preclude the use of collostructional analysis, it is an open question whether the marginals are cognitively relevant in this case.
+After all, the main difference between the quotient used here and measures used in CA (originally signed logarithmised p-values from a Fisher test) is that they take the marginals into account. 
+However, using Fisher-$p$-based collexeme strength instead of the raw frequency quotient was tried as an alternative (variables \textit{Kindcollo} and \textit{Measurecollo}); see Section~\ref{sec:mainstudystatisticalmodel} for a discussion of the negative results.
 
-Additionally, \textit{Kindfreq} and \textit{Measurefreq} are the logarithm-transformed frequencies per 1,000,000 words of each lemma, extracted from the frequency lists distributed by the DECOW14A corpus creators on their web page.
 In Section~\ref{sec:analyses}, it was hypothesised that classes of measure lemmas might have different preferences for the two alternants.
@@
 At the document level, the DECOW14A corpus has an annotation for \textit{Badness}.
-As described in \cite{SchaeferEa2013}, \textit{Badness} measures how well the distribution of highly frequent short words in the document matches a pre-generated language model for German.
-They also show that the Badness score corresponds robustly with human raters' intuitions about text coherence and text quality (see the paper for an operationalisation of this notion).
-Documents with higher Badness usually contain more incoherent language, shorter sentences, etc.
+As described in \cite{SchaeferEa2013}, \textit{Badness} measures how well the distribution of highly frequent short words in the document matches a language model of standard German.
+The paper also shows that the Badness score corresponds robustly with human raters' intuitions about text coherence and text quality.
+Documents with higher Badness usually contain more incoherent language and shorter sentences.
 If the \PGCa\ actually favours more elaborate stylistic levels, a high \textit{Badness} should be correlated with fewer occurrences.
@@
 A high number of genitives is also indicative of a more formal, elaborate style close to the written standard.
-However, the use of this variable as a regressor in the present study might be considered problematic.
-Since the \PGCa\ contains a genitive itself, the regressor variable \textit{Genitive} and the document-level variable \textit{Genitives} are not fully independent.
-Since instances of the \PGCa\ make up for only a minute fraction of all genitives, however, I still use \textit{Genitives} as a regressor.
+However, since the \PGCa\ contains a genitive itself, the regressor variable \textit{Genitive} and the document-level variable \textit{Genitives} are not fully independent.
+Since instances of the \PGCa\ make up for only a minute fraction of all genitives, I still use \textit{Genitives} as a regressor with the appropriate caveats.
 
-Finally, two variables were added as nuisance variables in the context of the present study.
+Finally, to control for simple frequency effects, \textit{Kindfreq} and \textit{Measurefreq} are the logarithm-transformed frequencies per 1,000,000 words of each lemma, extracted from the frequency lists distributed by the DECOW14A corpus creators on their web page.
+Also, two variables were added as nuisance variables in the context of the present study.
 First, it was reported in the literature that MNPs in the dative and with a masculine or neuter kind noun favour the \PGCa\ more than the corresponding nominative and accusative MNPs \citep{Hentschel1993,Zimmer2015}.
@@
 As with all the examples, native speakers of German will most likely notice that differences are subtle.
-To control for this effect, the case of the measure noun was manually annotated (variable \textit{Measurecase}).
+The case of the measure noun was manually annotated (variable \textit{Measurecase}).
 Second, due to differences in case syncretisms, it is possible that feminine kind nouns have slightly different preferences than masculine and neuter ones, and the appropriate variable \textit{Kindgender} was annotated.
@@
 The main study to be reported below deals exclusively with the alternating constructions, as is customary in alternation modelling.
-However, the prototypical features described in Section~\ref{sec:prototypeeffects} should be observable in the non-alternating cases as well if the theory laid out in Section~\ref{sec:analyses} is correct.
+However, the prototypical features described in Section~\ref{sec:prototypeeffects} should be observable in the non-alternating cases as well.
 Therefore, in this section I examine the distribution of the prototypical features in the non-alternating cases to see whether they are in accord with the theoretical predictions.
@@
 For this pre-study, each of the 2,365 combinations of measure noun lemma and kind noun lemma were queried in the \NACb\ and the \PGCd.
-Each of the resulting 2,365 concordances was scaled down randomly to a size of maximally 100 sentences, and from the resulting 35,766 sentences, 5,000 were randomly sampled for the statistical analysis.
+Each of the resulting 2,365 concordances was scaled down randomly to a size of maximally 100 sentences.
+From the resulting 35,766 sentences, 5,000 were randomly sampled for the statistical analysis.
 All features described in Section~\ref{sec:variablesandannotation} were annotated except for the ones which do not apply in the non-alternating case (\textit{Kindattraction}, \textit{Measureattraction}, and \textit{Measurecase}).
@@
 
-The annotated concordance was analysed in the form of a multilevel logistic regression model using R \citep{R} and the \textit{lme4} package \citep{Bates2010,BatesEa2015}.%
-\footnote{An alternative \textit{BOBYQA} optimiser from the \textit{nloptr} package \citep{Johnson2017} was used for all fits with \textit{lme4} reported in this paper.}
+The annotated concordance was analysed in the form of a multilevel logistic regression model using R \citep{R} and the \textit{lme4} package \citep{Bates2010,BatesEa2015}.
+An alternative \textit{BOBYQA} optimiser from the \textit{nloptr} package \citep{Johnson2017} was used for all fits with \textit{lme4} reported in this paper.
 
@@
 
-The coefficient estimates are specified in Table~\ref{tab:smalltable} for each regressor (or regressor level) in the columns labelled \textit{Coefficient}.
+The coefficient estimates are specified in Table~\ref{tab:smalltable} for each regressor (column \textit{Coefficient}).
 Given the coding of the response variable, coefficients leaning to the positive side can be interpreted as characterising a configuration typical of the \PGCd.
-For a robust quantification of the precision of the estimation, I ran a parametric bootstrap (using the \mbox{\textit{confint.merMod}} function from \textit{lme4}) with 1,000 replications and using the percentile method for the calculation of the intervals.
-The resulting 95\% bootstrap confidence intervals are reported in Table~\ref{tab:smalltable} in the columns labelled \textit{CI low} and \textit{CI high} (= upper and lower 2.5th percentiles).
+For a robust quantification of the precision of the estimation, I ran a parametric bootstrap (using the \mbox{\textit{confint.merMod}} function from \textit{lme4}) with 1,000 replications and using the percentile method.
+The resulting 95\% bootstrap confidence intervals are reported in Table~\ref{tab:smalltable} (columns \textit{CI low} and \textit{CI high}).
 The column \textit{CI excludes 0} shows an asterisk for those intervals that do not include 0.
@@
 Instead of inexact Wald approximations and Likelihood Ratio Tests, I used a drop-in bootstrap replacement for the Likelihood Ratio Test in the form of the function \textit{PBmodcomp} from the \textit{pbkrtest} package \citep{HalekohHojsgaard2014}.
-I call the corresponding value $p_{\text{PB}}$, and it is given in the respective columns in Table~\ref{tab:smalltable}.
-Regressors which did not reach sig=0.05 in in the PBmodcomp test (\textit{Kindfreq}, \textit{Kindgender}, and \textit{Measurefreq}) were removed from the model, appear in brackets in Table~\ref{tab:smalltable}, and consequently have no coefficient estimates.
-The overall intercept is -5.370 and represents \textit{Cardinal=Yes}, \textit{Measureclass=Physical} and 0 for all (z-transformed) numeric regressors.
+I call the corresponding value $p_{\text{PB}}$ (column $p_{\text{PB}}$).
+Regressors which did not reach sig=0.05 in in the PBmodcomp test (\textit{Kindfreq}, \textit{Kindgender}, and \textit{Measurefreq}) were removed from the model, appear in brackets in Table~\ref{tab:smalltable}, and have no coefficient estimates.
+
+The overall intercept is -5.370, and it represents \textit{Cardinal=Yes}, \textit{Measureclass=Physical} as 0 for all (z-transformed) numeric regressors.
 The standard deviation in the random intercepts is 1.157 for \textit{Measurelemma} ($p_{\text{PB}}=0.001$) and 0.923 for \textit{Kindlemma} ($p_{\text{PB}}=0.002$).
-Nakagawa \& Schielzeth's pseudo-coefficient of determination is $R_m^2=0.278$ and $R^2_c=0.566$ (see \citealp{Gries2015} for a basic introduction to these $R^2$ measures, or else \citealp{NakagawaSchielzeth2013}).
+Nakagawa \& Schielzeth's pseudo-coefficient of determination is $R_m^2=0.278$ and $R^2_c=0.566$ (see \citealp{Gries2015} or \citealp{NakagawaSchielzeth2013}).
 
@@
 
-The coefficients of determination indicate that the effect of the fixed effects (marginal $R^2$) is adequately strong, and that lemma-specific effects (conditional minus marginal $R^2$) are equally strong with a difference between the two coefficients of 0.289.
+The coefficients of determination indicate that the effect of the fixed effects (marginal $R^2$) is adequately strong, and that lemma-specific effects are equally strong with a difference between the two coefficients of 0.289.
 
 The main prototypical factors \textit{Cardinal} and \textit{Measureclass} both pass the PBmodcomp test at sig=0.05 and play the expected role.
-The non-referential physical measure nouns (such as \textit{Gramm} `gram' or \textit{Liter} `litre') with a high degree of grammaticalisation favour the \NACb.
+Non-referential physical measure nouns (such as \textit{Gramm} `gram' or \textit{Liter} `litre') with a high degree of grammaticalisation favour the \NACb.
 At the other end of the scale, nouns denoting natural portions like \textit{Haufen} `heap', \textit{Bündel} `bundle', \textit{Schluck} `gulp' favour the \PGCd.

### 2018-04-09 Latex/sections/2descriptive.tex
Virtually done, some micro-shortening ahead.

@@
 I call the kind-denoting noun the \textit{kind noun} and the quantity-denoting noun the \textit{measure noun}.
-For illustration purposes, in the English phrase \textit{a glass of good wine}, \textit{glass} is the measure noun and \textit{wine} is the kind noun.
+For illustration purposes, in the English phrase \textit{a glass of good wine}, the measure noun is \textit{glass} and the kind noun is \textit{wine}.
 Measure nouns can be all sorts of nouns which denote a quantity (such as \textit{litre} or \textit{amount}) but also those denoting containers, collections, etc.
@@
 
-Three different syntactic configurations within MNPs need to be distinguished, and case alternation occurs only in one of them.
-It occurs only when the kind noun is modified by an attributive adjective and there is no determiner, as in (\ref{ex:intro:alternation}).
+Three different syntactic configurations within MNPs need to be distinguished.
+The case alternation only occurs when the kind noun is modified by an attributive adjective and there is no determiner, as in (\ref{ex:intro:alternation}).
 Superficially, the sentences are functionally and semantically equivalent either with the kind noun in the genitive (\ref{ex:intro:alternation1}) or in the same case as the measure noun, an accusative in the case of (\ref{ex:intro:alternation2}).%
@@
 See \cite{Hentschel1993} and \cite{Zimmer2015} for analyses of the sometimes absurd stances taken in grammars of German.
-As will be shown (especially in Section~\ref{sec:exp:fc}), there might be preferences, but we cannot assume either construction to be unacceptable.
 }
@@
 
-Second, if the kind noun is bare -- \ie\ if it comes neither with a determiner nor a modifying adjective -- it is uninflected as in (\ref{ex:intro:narrowapposition1}), and the genitive as seen in the PGC is not acceptable, see (\ref{ex:intro:narrowapposition2}).
+Second, if the kind noun is bare -- \ie\ if it comes neither with a determiner nor a modifying adjective -- it is uninflected as in (\ref{ex:intro:narrowapposition1}), and the genitive as seen in the PGC is not acceptable, see (\ref{ex:intro:narrowapposition2}).%
+\footnote{
+It is difficult to determine whether the bare kind noun in (\ref{ex:intro:narrowapposition1}) bears no case at all, a generic case, or agrees in case with the measure noun.
+When there is an adjective as in (\ref{ex:intro:alternation2}), the embedded kind NP clearly agrees in case, but due to the overall absence of markers of case in the singular, bare nouns mostly show no indication of their case.
+Descriptive grammars seem to favour an analysis in terms of caselessness (for example, \citealp[1981]{ZifonunEa1997c}).
+Since it is not relevant to the argument presented here, the question is left open. 
+}
 
@@
 Notice that the unavailability of the genitive on the kind noun follows independently from a constraint that genitive NPs in German require the presence of some strongly case-marked element (determiner or adjective) in addition to the head noun in order to be acceptable (\citealp{GallmannLindauer1994,Schachtl1989}; see also \citealp[160]{Eisenberg2013b}).
-
-It is difficult to determine whether the bare kind noun in the NAC as in (\ref{ex:intro:narrowapposition1}) bears no case at all, a generic case, or agrees in case with the measure noun.
-When there is an adjective as in (\ref{ex:intro:alternation2}), the embedded kind NP clearly agrees in case, but due to the overall absence of markers of case in the singular, bare nouns mostly show no indication of their case.
-Descriptive grammars seem to favour an analysis in terms of caselessness (for example, \citealp[1981]{ZifonunEa1997c}).
-Since the concrete case status of the kind NP in the NAC is not relevant to the argument presented here, the question is left open.
 
@@
 In order to understand the MNP alternation, it is vital to consider how in the relevant syntactic structure [N\Sub{1}~[A~N\Sub{2}]\Sub{NP\Sub{2}}]\Sub{NP\Sub{1}} (as instantiated in the \NACa\ and the \PGCa), the adjective is morpho-syntactically ambiguous between a determiner and a modifying adjective.
-To show this, the strong\slash weak inflection patterns of adjectives needs to be taken into account.
+To show this, the so-called strong and weak inflection patterns of German adjectives needs to be taken into account.
 In NPs with a strongly inflected determiner, attributive adjectives inflect according to the massively syncretic \textit{weak} pattern.
@@
 However, if it is classified as an adjective, the \NACa\ suggests itself because of the lack of a determiner.%
-\footnote{While the generative analysis presented in \cite{Bhatt1990} cannot properly deal with probabilistic effects, Bhatt comes close to this interpretation by analysing the kind NP in the PGC as a DP and in the NAC as an NP.}
+\footnote{The generative analysis presented in \cite{Bhatt1990} comes close to this interpretation by analysing the kind NP in the PGC as a DP and in the NAC as an NP.}
 This morpho-syntactic ambiguity means that the \NACa\ is in fact a \NACb\ in disguise, and the \PGCa\ is a \PGCd\ in disguise.
@@
 
-In this section, I develop my analysis of the MNP alternation and the appropriate hypotheses for the empirical studies presented in Sections~\ref{sec:corpusstudies} and \ref{sec:experimental}.
-I also review some existing analyses of the alternation and related issues.
-
 \subsubsection{Prototype effects}
@@
 
-My analysis is based on the idea that the PGC prototype expresses a pseudo-partitive where two discernible entities -- the measure and the substance -- are referenced.
+My analysis of what controls the MNP alternation is based first and foremost on the idea that the PGC prototype expresses a pseudo-partitive where two discernible entities -- the measure and the substance -- are referenced.
 The NAC prototype, on the other hand, merely expresses a quantity, and the measure is not referenced as an entity in its own right.
@@
 While in the given examples, the effect is strongly supported by the choice of the measure noun lemmas, I argue that the meaning of the prototypes is independent of this.
-The independent meanings of the prototypes are, as I will show, a result of diachronic developments and grammaticalisation processes.
+The independent meanings of the prototypes are the result of diachronic developments and grammaticalisation processes.
 Furthermore, I argue that the prototypical meanings are reflected in their usage patterns, which will be tested in the corpus study in Section~\ref{sec:corpusstudies}.
@@
 It is often assumed that pseudo-partitives and quantity constructions arise as a form of grammaticalised partitives (\eg\ \citealp[536--539]{Koptjevskaja2001} for Finnish and Estonian, \citealp[559]{Koptjevskaja2001} for European languages in general).
-The grammaticalisation paths uncovered by \citet[esp.\ 526--530]{Koptjevskaja2001} are relevant for the case at hand.
+The grammaticalisation paths described by \citet[esp.\ 526--530]{Koptjevskaja2001} are relevant for the case at hand.
 The grammaticalisation path can start out (in some languages) with constructions involving two referential nouns (not even necessarily forming a single and contiguous NP) and a \textit{separative} meaning as in \textit{(cut) two slices from the cake} \citep[535]{Koptjevskaja2001}.
 In this type of construction, it is most obvious that two separate referents (in the given example the cake and the slices) are conceptualised.
-The \textit{part-of} meaning of true partitives as in \textit{a slice of the cake} represents the first stage of a development wherein the measure noun can already lose some semantic content, when, for example, words like \textit{bite} are no longer necessarily interpreted as a piece literally bitten out of something.
+The \textit{part-of} meaning of true partitives as in \textit{a slice of the cake} represents the first stage of a development wherein the measure noun can already lose some semantic content, especially when words like \textit{bite} are no longer necessarily interpreted as a piece literally bitten out of something.
 The pseudo-partitive stage finally instantiates a \textit{quantity-of} relation, potentially even leading to fully grammaticalised quantifiers such as \textit{a lot}.
@@
 It is thus not surprising that it has lost (at least prototypically) the semantics which allows both the measure and the substance to be conceptually accessible as independent referents.
-As a consequence, we should expect the NAC constructions to be typical hosting constructions for more strongly grammaticalised measure nouns.
+As a consequence, we should expect the NAC constructions to host more strongly grammaticalised measure nouns.
 For example, highly grammaticalised non-referential nouns like \textit{Gramm} `gram' and \textit{Meter} `metre' should occur proportionally more often in NAC constructions than in PGC constructions.
@@
 In the corpus study, measure lemmas will therefore be annotated with appropriate semantic class labels to check whether semantic classes of measure nouns have different affinities to the two variants.
-The actual classification is based on the list in \citet[530]{Koptjevskaja2001}, but due to the low frequencies of many of the potential classes, a very coarse classification was used in the end.
+The actual classification is based on the list in \citet[530]{Koptjevskaja2001}, but due to the low token frequencies in many of the potential classes, a very coarse classification was used in the end.
 With typical examples, the classes are:
@@
 However, it is much more typical of quantities to be specified with numerical precision.
-This is most obviously so with the highly grammaticalised physical measure nouns like \textit{centilitre}, which are very typically used with exact numerals instead of unspecific quantifiers, although both options are available in principle (\textit{three centilitres} vs.\ \textit{several centilitres}).
+This is most obviously so with the highly grammaticalised physical measure nouns like \textit{centilitre}, which are very typically used with numerals instead of unspecific quantifiers, although both options are available in principle (\textit{three centilitres} vs.\ \textit{several centilitres}).
 Since the \NACa\ is more closely associated with the \textit{quantity-of} relation, cardinals as attributes of the measure noun are expected to have a higher proportional frequency in the \NACa.
@@
 Under an integral view of prototypes, which incorporates effects related to larger contexts and styles, such preferences can be part of what defines the construction prototypes, and the \PGCa\ should occur proportionally more often in more elaborate styles closer to the standard.
+
+\subsubsection{Exemplar effects}
+\label{sec:exemplareffects}
+
+While the prototypes for the PGC and the NAC were specified with reference to high-level features in Section~\ref{sec:prototypeeffects}, at least one exemplar-driven similarity effect can also be expected to influence the MNP alternation.
+The measure and kind noun lemmas which occur in the alternating constructions obviously also occur in the non-alternating constructions.
+The relative frequency with which they occur in these stable and highly frequent cases -- where choosing an alternative is impossible -- could thus be a factor influencing the alternating, less stable case.
+Should this be confirmed, it would be highly implausible to conceive of such an effect as a prototype effect.
+In the corpus study reported in Section~\ref{sec:corpusstudies}, a measure quantifying this influence will therefore be included as the \textit{attraction strength}.
+
+It should be noted that such an exemplar-type effect would have to be confirmed \textit{in addition} to the predicted semantic prototype effect for measure lemmas described in Section~\ref{sec:prototypeeffects}, namely the effect of semantic classes of measure nouns.
+It must also be ensured that these influences at the lemma level are not spurious and just artefacts of mere lemma idiosyncrasies.
+This leads to a rather complex and truly multilevel model structure for the generalised linear mixed models (GLMMs) to be used in Section~\ref{sec:corpusstudies}.%
+\footnote{See \citet{GelmanHill2006} (especially part 2) for a comprehensive text book with a focus on multilevel models.}
+While true multilevel models are not used very often in corpus linguistics -- \citet{Gries2015} even calls the simpler varying-intercepts and varying-slopes models ``underused'' -- they provide an excellent tool to describe situations where preferences at the sentence level and at lexical levels need to be integrated.
+The models do not differentiate in and of themselves between abstraction and exemplar effects, but they allow researchers to tune the degree of complexity of models, incorporating both types of effects according to their theory and the phenomenon at hand.
+In Sections~\ref{sec:prestudy} and~\ref{sec:corpushierarchicalmodel}, multilevel models will therefore be used.

### 2018-04-09 Latex/sections/3corpus.tex
Virtually done, some micro-shortening ahead.

@@
 
-The regressors with the measure lemma as their unit of reference have no within-measure lemma variance, and the \textit{glmer} function automatically estimates them as group-level predictors (or second-level effects), cf.\ \citet[265--269,302--304]{GelmanHill2006} and Section~\ref{sec:itemandexemplareffects}.
+The regressors with the measure lemma as their unit of reference have no within-measure lemma variance, and the \textit{glmer} function automatically estimates them as group-level predictors (or second-level effects), cf.\ \citet[265--269,302--304]{GelmanHill2006}.
 The same goes for those listed with the kind lemma as their unit of reference.

### 2018-04-09 Latex/sections/5conclusions.tex
Virtually done, some micro-shortening ahead.

@@
 On a larger methodological scale, this paper also makes a number of contributions.
-The statistical model was a true multilevel model (see Section~\ref{sec:itemandexemplareffects} and Section~\ref{sec:mainstudystatisticalmodel}), demonstrating how multilevel modelling helps to specify complex hierarchies of abstract features, item-specific tendencies, and exemplar effects at the level of observations (sentences; first level) and lemmas (second level).
+The statistical model was a true multilevel model (see Sections~\ref{sec:exemplareffects} and~\ref{sec:mainstudystatisticalmodel}), demonstrating how multilevel modelling helps to specify complex hierarchies of abstract features, item-specific tendencies, and exemplar effects at the level of observations (sentences; first level) and lemmas (second level).
 While \cite{Gries2015} still calls mixed models ``underused'' in corpus linguistics, multilevel models are consequently at least equally underused tools.

### 2018-04-09 Latex/sections/1intro.tex
Basically done and below 13,000 words.

@@
 
-While prototype theory is well suited for modeling constructional choices, it is just one of at least two major similarity-based theories of classification, the prominent alternative being exemplar theory (\citealp{MedinSchaffer1978,Hintzman1986}).
+While prototype theory is well suited for modelling constructional choices, it is just one of at least two major similarity-based theories of classification, the prominent alternative being exemplar theory (\citealp{MedinSchaffer1978,Hintzman1986}).
 Prototype theory and exemplar theory model essentially the same effects when only output data are considered.

### 2018-04-09 Latex/sections/3corpus.tex
Basically done and below 13,000 words.

@@
 I applied a two-stage process in order to obtain a list of co-occurring measure nouns and kind nouns.
-First, I generated a list of the one hundred most frequent genuine mass nouns.
-Second, I derived a list of all measure nouns with which the mass nouns co-occurred in the \NACb.
 
-In the first step, I exported a list of all nouns in the DECOW14A01 sub-corpus sorted by their token frequency.
-The first one hundred mass nouns were extracted manually from this list.%
-\footnote{DECOW14A01 is the first slice (roughly a twentieth) of the complete DECOW14A corpus.
-It contains roughly one billion tokens.}
-Mass nouns were defined as concrete nouns which denote a substance genuinely (without coercion), combine with uninflected mass quantifiers such as \textit{viel} `much' and \textit{wenig} `little' (\textit{viel Bier} `much beer'), and form only sortal and unit plurals (such as the plural \textit{Biere} `types of beer' or `glasses of beer').
-Abstract nouns which partially behave like mass nouns (like \textit{Spaß} `fun’ or \textit{Gefahr} `danger’) were excluded because they are usually not quantified in the same way as concrete mass nouns.
-The hundredth selected mass noun was \textit{Schmuck} `jewellery’, which is the 3,054th most frequent noun in the original frequency list.
+In the first step, I exported a list of all nouns from the DECOW14A01 sub-corpus (one billion tokens large).
+The one hundred most frequent mass nouns were extracted manually from this list.
+Mass nouns were defined as concrete nouns which genuinely denote a substance, combine with uninflected mass quantifiers such as \textit{viel} `much' and \textit{wenig} `little' (\textit{viel Bier} `much beer'), and form only sortal and unit plurals (such as the plural \textit{Biere} `types of beer' or `glasses of beer').
+Abstract nouns which partially behave like mass nouns (like \textit{Spaß} `fun’ or \textit{Gefahr} `danger’) were excluded because they are not quantified in the same way as concrete mass nouns.
 
@@
 The variables \textit{Kindattraction} and \textit{Measureattraction} encode the ratio with which a given kind noun lemma or measure noun lemma occurs in the \PGCd\ and the \NACb.
-They were calculated from auxiliary samples described in Section~\ref{sec:mainstudysampling} as a log-transformed quotient.
-The higher the value, the more often the noun occurs in the \PGCd.
+They were calculated from auxiliary samples to be described in Section~\ref{sec:mainstudysampling} as a log-transformed quotient.
+The higher the value for some noun, the (relatively) more often this noun occurs in the \PGCd.
 It could be argued that other measures of attraction strength should be used, for example those popularised in collostructional analysis (CA; \citealp{StefanowitschGries2003,GriesStefanowitsch2004}; see also \citealp{Gries2015a}).
@@
 
-Finally, to control for simple frequency effects, \textit{Kindfreq} and \textit{Measurefreq} are the logarithm-transformed frequencies per 1,000,000 words of each lemma, extracted from the frequency lists distributed by the DECOW14A corpus creators on their web page.
-Also, two variables were added as nuisance variables in the context of the present study.
+Furthermore, to control for simple frequency effects, \textit{Kindfreq} and \textit{Measurefreq} are the logarithm-transformed frequencies per 1,000,000 words of each lemma, extracted from the frequency lists distributed by the DECOW14A corpus creators on their web page.
+
+Finally, two variables were added as nuisance variables in the context of the present study.
 First, it was reported in the literature that MNPs in the dative and with a masculine or neuter kind noun favour the \PGCa\ more than the corresponding nominative and accusative MNPs \citep{Hentschel1993,Zimmer2015}.
@@
 The case of the measure noun was manually annotated (variable \textit{Measurecase}).
-Second, due to differences in case syncretisms, it is possible that feminine kind nouns have slightly different preferences than masculine and neuter ones, and the appropriate variable \textit{Kindgender} was annotated.
+Second, due to differences in case syncretisms, it is likely that feminine kind nouns have slightly different preferences than masculine and neuter ones, and hence the appropriate variable \textit{Kindgender} was annotated.
 
@@
 
-For this pre-study, each of the 2,365 combinations of measure noun lemma and kind noun lemma were queried in the \NACb\ and the \PGCd.
-Each of the resulting 2,365 concordances was scaled down randomly to a size of maximally 100 sentences.
+For this pre-study, each of the 2,365 combinations of measure noun lemma and kind noun lemma were queried in the \NACb\ and the \PGCd\ with a limit of 100 randomly chosen results for each single query.
 From the resulting 35,766 sentences, 5,000 were randomly sampled for the statistical analysis.
@@
 
-For the main study, each of the 2,365 noun–noun combinations was queried in the alternating constructions \PGCa\ and \NACa\ in DECOW14A.%
-\footnote{Due to efficiency problems with the old COW interface at the time, only ten slices of DECOW14A were used, which add up to approximately 10 billion tokens.}
-For the manual annotation process, a random sub-sample was generated.
-For each mass noun, the concordance was downsampled randomly to maximally 100 sentences, which resulted in 6,843 sentences without further downsampling.
-In the manual annotation process the sample was further reduced to 5,063 sentences due to removal of noisy material, erroneous hits, and uninformative cases where the measure noun was in the genitive, in which case the \NACa\ cannot be distinguished from the \PGCa.
-Given the careful sampling procedure described in this section and Section~\ref{sec:bootstrappinlemmapairs}, we can be highly certain that it contains all relevant and reasonably frequent noun–noun combinations in the target constructions.%
+For the main study, each of the 2,365 noun–noun combinations was queried in the alternating constructions \PGCa\ and \NACa\ using a sub-corpus of DECOW14A (slices DECOW14A01 to DECOW14A10 with a total size of 10 billion tokens).
+For further manual annotation, a random sub-sample was generated.
+For each mass noun, the concordance was downsampled randomly to maximally 100 sentences, which resulted in 6,843 sentences.
+In the manual annotation, the sample was further reduced to 5,063 sentences due to removal of noisy material, erroneous hits, and uninformative cases where the measure noun was in the genitive, in which case the \NACa\ cannot be distinguished from the \PGCa.%
 \footnote{In a similar fashion, the 100 most frequent measure nouns occurring with plural kind nouns were listed and queried, resulting in a sample of 871 sentences.
@@
 
-Then, a multilevel logistic regression model was fit which models the influence of the regressors specified in Table~\ref{tab:variables} on the probability that the \PGCa\ is chosen over the \NACa.
+A multilevel logistic regression model was fit which models the influence of the regressors specified in Table~\ref{tab:variables} on the probability that the \PGCa\ is chosen over the \NACa.
 All regressors from Table~\ref{tab:variables} were included, and the measure lemma and the kind noun lemma were specified as varying-intercept random effects.
@@
 
-    2 (Measurelemma & Measureattraction &  0.001                 &              &  0.282      &  0.106 &  0.447  & *             \\
+    2 (Measurelemma) & Measureattraction &  0.001                 &              &  0.282      &  0.106 &  0.447  & *             \\
                     & Measureclass      &  0.001                 & Container    &  0.252      & -0.265 &  0.788  &               \\
@@
 
-Using signed logarithmised collexeme strength (\textit{Measurecollo} and \textit{Kindcollo}) (with smoothing to avoid arithmetic problems) instead of the quotient for the attraction strength (\textit{Measureattraction} and \textit{Kindattraction}) (see Section~\ref{sec:variablesandannotation}) was not successful.
+Using signed logarithmised Fisher p-values as a measure of collexeme strength (\textit{Measurecollo} and \textit{Kindcollo}) instead of the quotient for the attraction strength (\textit{Measureattraction} and \textit{Kindattraction}) (see Section~\ref{sec:variablesandannotation}) was not successful.
 While the attraction measures reach satisfying sig levels in the PBmodcomp test (0.020 for \textit{Kindattraction} and 0.001 for \textit{Measureattraction}), the $p_{\text{PB}}$ value for \textit{Kindcollo} was 0.191 and the one for \textit{Measurecollo} was 0.443.
@@
 With kind nouns, higher frequency seems to favour the \PGCa.
-However, there is no clear theoretical interpretation (see Section~\ref{sec:analyses}), and the estimate is imprecise (not significant at sig=0.05).
+However, this effect has no clear theoretical interpretation, and its coefficient estimate is imprecise (not significant at sig=0.05).
 The effect can therefore be ignored or treated as a nuisance variable.
@@
 The two studies do not converge with respect to the \textit{Badness} variable, which is not significant in the pre-study.
-It would, however, surprise to achieve perfectly converging results given that even real effects are missed at certain rates in empirical studies.
+However, it would be surprising if we achieved perfectly converging results given that even real effects are missed at certain rates in empirical studies.
 In the case of \textit{Badness}, we see that even in the main study, the effect size is small (-0.152), and thus even at the impressive sample size of roughly 5,000 in both studies, the effect might simply be too weak to be detected reliably.
@@
 Thus, the frequency effect for measure lemmas might be swamped by the lemma random intercepts.
-All things considered, the studies have shown that the predicted prototype effects are found in usage data for both the alternating and the non-alternating constructions.
+All things considered, the studies have shown that the predicted prototype and exemplar effects are reflected in usage data for both the alternating and the non-alternating constructions.

### 2018-04-09 Latex/sections/4experiments.tex
Basically done and below 13,000 words.

@@
 
-In the two experiments reported in this section, I use probabilities for the alternating constructions calculated for attested material, and I correlate these probabilities with the participants' reactions.
-Thus, a direct link can be established between output material found in corpora and the behaviour of linguistic agents.
-Both experiments use sentences containing attested MNPs from the corpus sample (embedded into simplified sentences) as stimuli.
+In this section, the probabilities of the alternants as calculated for attested material are correlated with the reactions of participants in controlled experiments.
+Thus, a direct link is established between output material found in corpora and the behaviour of linguistic agents.
 
-Readers might wonder why the stimuli were chosen in such a way.
+The decision to use attested material and perform a global validation of the corpus-based model was not arbitrary.
 Alternatively, one could have created stimuli where the features favouring the alternatives were permuted appropriately and thus tested directly.
@@
 Since the dependent variable is binary (\textit{Construction}), permuting these factors alone leads to 180 different possible permutations.
-On top of this, the numeric variables \textit{Kindattraction} and \textit{Measureattraction} would also have to be tested at at least two or three different values each.
-It is obvious that this is impossible in a controlled experiment, given that participants can usually be exposed to something in the region of one hundred sentences (including many more fillers than target stimuli) within thirty minutes to an hour.
+On top of this, the numeric variables \textit{Kindattraction} and \textit{Measureattraction} would also have to be tested at several values.
+It is obvious that this is impossible in a controlled experiment, given that a participant can be exposed to no more than roughly one hundred sentences, most of which would be fillers rather than target sentences.
 Thus, while the present approach only allows for a global test of the corpus-derived model, this appears to me to be the only feasible way.%
-\footnote{Obviously, the experimental data are much too sparse to perform even post-hoc analyses with respect to the single regressors in a statistically and scientifically sound way.}
+\footnote{For the same reasons, the experimental data are also much too sparse to perform even post-hoc analyses with respect to the single regressors in a sound way.}
 It is, in the sense of Section~\ref{sec:cogocl}, an optimal synthesis of data-rich multifactorial corpus studies and experimental validation, and it is one that has been used (although rarely; see \citealp[3--4]{DivjakEa2016a}) at least since the seminal \citet{BresnanEa2007} paper.
@@
 
-The first experiment tests preferences for constructions explicitly in a forced choice task.
+The first experiment tests preferences for constructions explicitly in a forced-choice task.
 Participants had to choose between two sentences that differed only in that one contained the \NACa\, and the other one contained the \PGCa.
-The analysis compares the probabilities assigned to the stimuli by the corpus-derived model with the frequency with which participants chose the alternants.
-There were 24 participants (native speakers of German without reading or writing disabilities) aged 19 to 30 living permanently in Berlin, who were recruited from introductory linguistics courses at Freie Universität Berlin.
+In the analysis, the probabilities assigned to the stimuli by the corpus-derived model were compared to the frequency with which participants chose the alternants.
+There participants were 24 native speakers of German without reading or writing disabilities, aged 19 to 30, and living permanently in Berlin.
+They were recruited from introductory linguistics courses at Freie Universität Berlin.
 Although the experiment was conducted in the last four weeks of their first semester, participants had no deeper explicit knowledge of linguistics, grammar, or experimental methods.
 None of them had ever participated in a forced-choice experiment before.
-Participation was voluntary but participants received credit in partial fulfillment of course requirements.
+Participation was voluntary, but participants received credit in partial fulfillment of course requirements.
 
 As stimuli, attested MNPs from the corpus study were used, but the sentences were simplified to avoid influences from contextual nuisance factors as much as possible.
-The approach is also justified because according to the theoretical assessment in Section~\ref{sec:analyses}, the choice of alternants depends mostly on a very local constructional context.
-I sampled 16 MNPs from the concordance and made sure that the simplifications and normalisations did not affect any of the regressors used in the corpus study.
+16 MNPs were sampled from the concordance, making sure that any simplifications and normalisations did not affect any of the regressors used in the corpus study.
 In the simplified sentences, the case, number, etc. of the MNP remained the same as in the attested sentence, as did the choice of lexical material within the MNP.
@@
 Furthermore, in each of the masculine\slash neuter and feminine groups, four sentences originally containing the \NACa\ and four sentences originally containing the \PGCa\ were chosen.
-More precisely, the sentences were sampled as \textit{highly typical examples} of \PGCa\ (high probability assigned by GLMM) and \NACa\ (low probability assigned by GLMM), respectively.%
+Moreover, the sentences were sampled as typical examples of \PGCa\ (high probability assigned by GLMM) and \NACa\ (low probability assigned by GLMM), respectively.%
 \footnote{Remember from Section~\ref{sec:corpusstudies} that the model predicts the probability that the \PGCa\ is chosen over the \NACa.}
-High and low probabilities were defined as the top and bottom 20\% of all probabilities assigned by the GLMM.
+High and low probabilities were defined (roughly) as the top and bottom 20\% of all probabilities assigned by the GLMM.
 Lemmas and feature combinations were made unique within each group whenever possible.
+For calculating the model predictions, the document-level variables \textit{Badness} and \textit{Genitives} were set to 0, which is the mean for z-transformed variables.
 The design is summarised in Table \ref{tab:experiment1:design}.
@@
 \footnote{No participant ever exceeded the time limit.}
-The position on the screen (left\slash right) and the order of sentences were randomised for each participant.
+The order of sentences was randomised for each participant, and the position of the alternants on the screen (left\slash right) was randomised per participant and sentence.
 As fillers, 23 pairs of sentences exemplifying similar but unrelated alternation phenomena from German morpho-syntax were used.
@@
 
-\begin{figure}[htbp!]
-\centering
-\includegraphics[width=0.5\textwidth]{../R/output/fc_effects}
-\caption{Effect plot for the multilevel logistic regression in the forced-choice experiment: predictability of participants' choices using the probabilities derived from the corpus-based GLMM}
-\label{fig:afc:effects}
-\end{figure}
-
-A multilevel logistic regression was specified with the probability of the \PGCa\ predicted for each sentence by the corpus-based GLMM as the only fixed effect \textit{Modelprediction}.%
-\footnote{The document-level variables \textit{Badness} and \textit{Genitives} were set to 0, which is the mean for z-transformed variables.}
-A random intercept and slope were added for the individual sentence pair (item) in order to catch idiosyncrasies of single sentences.
+A multilevel logistic regression was specified with the probability of the \PGCa\ predicted for each sentence by the corpus-based GLMM as the only fixed effect \textit{Modelprediction}.
+A random intercept was added for the individual sentence pair (\textit{Item}) in order to catch idiosyncrasies of single sentences.
 Coefficients were estimated with Maximum Likelihood Estimation (\textit{lmer} function from \textit{lme4}).
@@
 \subsubsection{Interpretation}
+
+\begin{figure}[htbp!]
+\centering
+\includegraphics[width=0.5\textwidth]{../R/output/fc_effects}
+\caption{Effect plot for the multilevel logistic regression in the forced-choice experiment: predictability of participants' choices using the probabilities derived from the corpus-based GLMM}
+\label{fig:afc:effects}
+\end{figure}
 
@@
 In line with expectations, the standard deviation in the random slopes for \textit{Item} is high at 5.996.
-However, the covariance parameters were estimated at -1.0, which is a clear sign that the variance-covariance matrix could not be estimated successfully.
+However, the covariance parameters were estimated at exactly -1, which is a clear sign that the variance-covariance matrix could not be estimated successfully.
 The same was true for models with only an \textit{Item} and a \textit{Participant} random slope.
@@
 The available data are insufficient to estimate the parameters of the more complex model with varying slopes.%
-\footnote{\citet[1]{BatesEa2015a} state: ``We show that failure to converge typically is not due to a suboptimal estimation algorithm, but is a consequence of attempting to fit a model that is too complex to be properly supported by the data, irrespective of whether estimation is based on maximum likelihood or on Bayesian hierarchical modeling with uninformative or weakly informative priors.
+\footnote{\citet[1]{BatesEa2015a} state: ``We show that failure to converge typically is not due to a suboptimal estimation algorithm, but is a consequence of attempting to fit a model that is too complex to be properly supported by the data, irrespective of whether estimation is based on maximum likelihood or on Bayesian hierarchical modelling with uninformative or weakly informative priors.
 Importantly, even under convergence, overparameterization may lead to uninterpretable models.''
@@
 
-The second experiment tests preferences more implicitly.
+The second experiment was conducted in a more implicit paradigm.
 It is expected that reading less typical alternants in a given context and with given lexical material incurs a processing overhead for the reader (\citealp{Kaiser2013}).
@@
 
-Concretely, the exact same stimuli as in the forced-choice experiments were used.
+The exact same stimuli as in the forced-choice experiments were used.
 Each participant read both the 16 sentences with the alternant predicted by the corpus model and the 16 modified sentences with the alternant that the corpus model did not predict.%
@@
 The probabilities derived from the corpus GLMM (\textit{Modelprediction}) were added as the main regressor of interest.
-It should be remembered that the corpus GLMM predicts the probability of the \PGCa.
-As a consequence, the higher the GLMM prediction is, the more typical the sentence is for containing the \PGCa.
+It should be remembered that the higher the GLMM prediction is, the more typical the sentence is for containing the \PGCa.
 It is therefore expected that reading times are higher when the value of \textit{Modelprediction} is higher but the sentence contains the \NACa.
@@
 
-The coefficients of determination indicate that there is a noteworthy correlation between the reactions of the subjects and the corpus-derived probabilities (marginal $R^2$) and that there is some between-subject variation (conditional minus marginal $R^2$).
+The coefficients of determination indicate that there is a noteworthy correlation between the reactions of the subjects and the corpus-derived probabilities (marginal $R^2$) and that there is some between-subject variation (conditional minus marginal $R^2$, which is 0.107).
 
@@
 \includegraphics[width=0.5\textwidth]{../R/output/spr_effects}
-\caption{Effect plot for the LMM in the self-paced reading experiment: modeling participants' residualised log reading times on the probabilities given by the corpus-based GLMM}
+\caption{Effect plot for the LMM in the self-paced reading experiment: modelling participants' residualised log reading times on the probabilities given by the corpus-based GLMM}
 \label{fig:spr:effects}

### 2018-04-09 Latex/sections/5conclusions.tex
Basically done and below 13,000 words.

@@
 
-The empirical studies in Sections~\ref{sec:corpusstudies} and~\ref{sec:experimental} confirm the theoretical model proposed in Section~\ref{sec:germanmeasurenps}, which predicted prototype effects and exemplar effects to jointly determine which of the two alternating constructions is chosen.
-More specifically, the pre-study and the main study provided evidence that the NAC constructions favour more strongly grammaticalised measure nouns, modification by cardinals, and styles where the genitive is underrepresented and the text is less coherently and elaborately written, as captured in the \textit{Badness} variable.%
-\footnote{Although, as pointed out in Section~\ref{sec:mainstudyinterpretation}, the \textit{Badness} effect is weak and not detected in the pre-study, probably as a consequence of its weakness.}
-The overall convergence between the pre-study and the main study and also the expected result for \textit{Measurecase} in line with previous research \citep{Zimmer2015} -- although it was only a control variable here -- strengthen the validity of this study even without the experimental validation.
+The studies reported in Sections~\ref{sec:corpusstudies} and~\ref{sec:experimental} have confirmed the theoretical model proposed in Section~\ref{sec:germanmeasurenps}, which predicted prototype effects and exemplar effects to jointly determine which of the two alternating constructions is chosen.
+More specifically, the pre-study and the main study provided evidence that the NAC constructions favour more strongly grammaticalised measure nouns, modification by cardinals, and styles where the genitive is underrepresented and the text is less coherently and elaborately written.
+The overall convergence between the pre-study and the main study and the convergence between corpus data and the experimental cross-validation strengthen the validity of this study.
+Also, the fact that the result for \textit{Measurecase} (although it was only a control variable in the present context) converged with previous research \citep{Zimmer2015} lends credibility to the results.
 
-While the results do not \textit{prove} -- due to the intrinsic equivalence of prototype and exemplar models with respect to the output they produce -- that the model proposed in Section~\ref{sec:germanmeasurenps} is congruent with any speaker's mental representation, the two attraction features (\textit{Kindattraction} and \textit{Measureattraction}) are virtually impossible to conceive of as prototype effects.
-As discussed in Section~\ref{sec:cogocl}, it is unnecessary to follow an extreme route, be it radical prototype theory or radical exemplar theory given recent developments in cognitive science and (less prominently) cognitive linguistics.
+Many relevant conclusions can be drawn based on the research presented here.
+Although direct inferences from corpus data to mental representations are problematic (see Section~\ref{sec:cogocl}), the two attraction features (\textit{Kindattraction} and \textit{Measureattraction}) are virtually impossible to conceive of as prototype effects and thus favour an exemplar view or a mixed exemplar-prototype view.
+As discussed in Section~\ref{sec:cogocl}, it is unnecessary to follow an extreme route, be it radical prototype theory or radical exemplar theory, given recent developments in cognitive science and cognitive linguistics.
 While it is surely an intentionally overstated comment, \citet[15]{Kapatsinski2014} even suggests that ``[i]n the extreme, some speakers’ heads could host exemplar models, and some could contain fairly abstract grammars, and the produced output would be essentially identical''.
-If this is the case (even to a less extreme degree), corpora only provide (still informative) pooled data averaged over speaker grammars with varying levels of abstraction.
+If this is the case (even to a less extreme degree), corpora only provide pooled data averaged over speaker grammars with varying levels of abstraction.
 Clearly, more research is required in this direction.
@@
   However, [\ldots] only a small number of these corpus-based studies have been cross-validated [\ldots].
-  Of these cross- validated studies, few have directly evaluated the prediction accuracy of a complex, multivariate corpus-based model on humans using authentic corpus sentences [\ldots].
+  Of these cross-validated studies, few have directly evaluated the prediction accuracy of a complex, multivariate corpus-based model on humans using authentic corpus sentences [\ldots].
 \end{quote}
@@
 The predictions made by the corpus-based model were a significant factor, both in the more explicit paradigm (forced choice) and the implicit paradigm (self-paced reading).
-This shows that the model indeed predicts the variant which language users expect.
+This shows that the model indeed partially predicts the variant which language users expect.
 The overall effect as measured in the $R^2$ was weak (roughly 0.2) in both cases, however.
@@
 With the added inter-speaker variability brought into play by the experimental setup (compared to the averaging across thousands of speakers in the corpus study), a perfect fit cannot be expected.
-Like many phenomena in German which are often called \textit{Zweifelsfälle} (`cases of doubt') in the traditional literature \citep{Duden09,Klein2009} the MNP alternation is one of the cases where speakers very often have no clear intuition and a lot of free variation seems to be involved.
+Like many phenomena in German which are often called \textit{Zweifelsfälle} (`cases of doubt') in the traditional literature \citep{Duden09,Klein2009} the MNP alternation is a case where speakers often have no clear intuition, and a lot of free variation seems to be involved.
 Rarely do speakers feel that one alternative is clearly odd or bad.
@@
 The statistical model was a true multilevel model (see Sections~\ref{sec:exemplareffects} and~\ref{sec:mainstudystatisticalmodel}), demonstrating how multilevel modelling helps to specify complex hierarchies of abstract features, item-specific tendencies, and exemplar effects at the level of observations (sentences; first level) and lemmas (second level).
-While \cite{Gries2015} still calls mixed models ``underused'' in corpus linguistics, multilevel models are consequently at least equally underused tools.
+\cite{Gries2015} still calls mixed models ``underused'' in corpus linguistics, and multilevel models are consequently also an underused tools.
 At the same time, the effects of overparametrisation of mixed models with varying slopes as criticised by \cite{BatesEa2015a} were demonstrated.
@@
 This is clearly another area where only more studies can lead to clarification.
-Finally, much like \cite{Dabrowska2014} found that speakers' knowledge of collocations was not matched by a set of standard measures of collocation strength extracted from corpora, a simple quotient based on raw frequencies for the attraction strength performed much better in the present study compared to collexeme strength (see Section~\ref{sec:corpusstudies}).
-While this does not allow the conclusion that collexeme strength has no cognitive reality, it might indicate that for measuring attraction effects exerted by non-alternating constructions on alternating constructions, other measures might be more appropriate.
+Finally, much like \cite{Dabrowska2014} found that speakers' knowledge of collocations was not matched by a set of standard measures of collocation strength extracted from corpora, a simple quotient based on raw frequencies for the attraction strength performed much better in the present study compared to collexeme strength in the form of logarithmised Fisher p-values (see Section~\ref{sec:corpusstudies}).
+While this does not allow the conclusion that collexeme strength has no cognitive reality, it might indicate that for measuring attraction effects exerted by non-alternating constructions on alternating constructions, taking the marginals into account might not be crucial.%
+\footnote{However, many different measures have been proposed within the collostructional framework (see \citealp{Gries2015b} for an overview of them and a defense of the framework).
+Future research might show significant differences between them for the problem at hand.}
 
 Besides the methodological aspects mentioned above, future work could extend the validity of the results presented here through a more in-depth look at stylistic or register effects, for example using the new Biber-style annotations \citep{Biber1988} which will be released with a new version of the DECOW corpus soon according to its creators.
-Also, regional variation as a potentially influencing factor was already mentioned in Section~\ref{sec:prototypeeffects}.
-As it would be infeasible to examine this with existing corpora, experimental work with participants from various regions might be the ideal approach.
 A reviewer also pointed out that strongly lexicalised adjective-noun combinations like \textit{schwarzer Tee} `black tea' might have a tendency to occur in the \NACa\ because they have more compound-like qualities, blocking strong case inflection in between them.

### 2018-04-11 Latex/article.tex
First round of shortening.
First version under 13,000 words.
Basically done and below 13,000 words.
Final author fixes. Off to proofreading.

@@
     Based on this, I argue that lexical, morpho-syntactic, and stylistic features help to predict the choice of the alternant.
-    A large corpus study is presented which supports this theory.
+    A large corpus study is presented which supports this analysis.
     However, in addition to the prototype effects, an exemplar effect is also shown to influence the choice, namely the relative frequencies with which lemmas occur in the non-alternating constructions.
@@
   }
-  \keywords{prototypes vs.\ exemplars, corpus methods and experimental validation, alternations, hierarchical models, pseudo-partitives, German}
+  \keywords{prototypes vs.\ exemplars, corpus methods and experimental validation, alternations, multilevel modeling, pseudo-partitives, German}
   %\classification[PACS]{...}
@@
 \maketitle
+
+% For word counting.
+%\raggedright
+%\pagestyle{empty}
 

### 2018-04-11 Latex/sections/1intro.tex
Final author fixes. Off to proofreading.

@@
 There are two fundamental issues to consider with respect to this tradition.
-First, there is the question of whether corpus data do provide any insight into cognitive representations at all.
-This question can and should be answered by testing how well corpus-derived models converge with or diverge from experimental findings, which provide more direct evidence of mental representations and processes (see Secion~\ref{sec:usagedataincognitivelinguistics}).
-Second, the appropriate modelling of such results in cognitive linguistics (\ie\ the assumed constructs) is a key issue (see Section~\ref{sec:prototypesexemplarscorporaandcontrolledexperiments}).
+First, there is the question of whether corpus data provide any insight into cognitive representations at all.
+This question can and should be answered by testing how well corpus-derived models converge with or diverge from experimental findings (Secion~\ref{sec:usagedataincognitivelinguistics}).
+Second, the appropriate modelling of such results in cognitive linguistics -- \ie\ the assumed constructs -- is a key issue (Section~\ref{sec:prototypesexemplarscorporaandcontrolledexperiments}).
 
@@
 Nobody assumes that we can inductively infer mental representations from experiments, which even allow for direct access to the cognitive agent and offer much better ways of controlling experimental conditions and nuisance variables.
-Rather, predictions are derived from existing theories in order to \textit{test} the theory.
-While the same approach is applicable to corpus data, I discuss some relevant differences.
+Rather, predictions are derived from existing theories in order to \textit{test} the theories.
+While this approach is applicable to corpus data, I discuss some relevant differences.
 
@@
 This is where the experimental validation (or \textit{corroboration}) comes into play.
-An overview was given in \cite{NewmanSorensenduncan2015}, who enumerate a number of studies showing how corpus data and experimental data converge (such as \citealp{BresnanEa2007,DurrantDoherty2010,GriesWulff2005,GriesEa2005}) and a number of studies where the two types of data led to diverging or only partially converging results (such as \citealp{ArppeJaervikivi2007,Dabrowska2014,Mollin2009}).
+A review of the state of the art was provided by \cite{NewmanSorensenduncan2015}, who enumerate a number of studies showing how corpus data and experimental data converge (such as \citealp{BresnanEa2007,DurrantDoherty2010,GriesWulff2005,GriesEa2005}) and a number of studies where the two types of data led to diverging or only partially converging results (such as \citealp{ArppeJaervikivi2007,Dabrowska2014,Mollin2009}).
 When researchers do not achieve convergence, they often try to explain this by differentiating between the actual cognitive construct and whatever the pooled usage data as found in corpora represent.
@@
 See, for example \cite{DivjakEa2016}, who show that generalised additive models (GAMs) are better suited than generalised linear models (GLMs) for correlating reading times and corpus data. 
-Much more optimistically, \cite{StefanowitschFlach2016} recently proposed a straightforward positivist perspective of corpora as representing the input of an average adult speaker, thus licensing inferences from corpus data to cognitive representations under a ``corpus-as-input'' view.
+Much more optimistically, \cite{StefanowitschFlach2016} proposed a straightforward positivist perspective of corpora as representing the input of an average adult speaker, thus licensing inferences from corpus data to cognitive representations under a ``corpus-as-input'' view.
 They state that ``in this wider context, large, register-mixed corpora such as the British National Corpus [\ldots] may not be perfect models of the linguistic experience of adult speakers, but they are reasonably close to the input of an idealized average member of the relevant speech community'' \citep[104]{StefanowitschFlach2016}.
@@
 \footnote{See \citet{StormsEa2000} for a comparison of the theories in different experimental settings.}
-Since corpus data only show artefacts of production events and we have no experimental access to the speaker's or writer's performance and their actual similarity judgements, one should be sceptical whether corpus analysis alone could ever decide which theory of mental representation is more suitable (see also \citealp[22]{Gries2003} and the \citealp[486--487]{Dabrowska2016} quote above).
-However, as I will show, some effects are more naturally analysed as prototype effects while others are almost necessarily exemplar effects.
+Since corpus data only show artefacts of production events and we have no experimental access to the speaker's or writer's performance and their actual similarity judgements, one should be sceptical whether corpus analysis alone could ever help to decide which theory of mental representation is more suitable (see also \citealp[22]{Gries2003} and the \citealp[486--487]{Dabrowska2016} quote above).
+However, as I will show, some effects are naturally analysed as prototype effects while others are almost necessarily exemplar effects.
 
@@
 \footnote{However, approaches have emerged which model linguistic phenomena without reference to high-level linguistic features altogether \citep{BaayenEa2016,RamscarPort2016}.}
-It is thus an advantage of much linguistic work on categorisation that it deals with complex and realistically produced data because this greatly improves the external validity of studies, although by sacrificing some construct validity.
+It is thus an advantage of much linguistic work on categorisation that it deals with complex and realistically produced data because this greatly improves the external validity of studies, although by potentially sacrificing some construct validity.
 Thus, an ideal contribution to the research on category abstraction by cognitive corpus linguists is to provide analyses which have great external validity and complexity while carefully making sure that these finding correlate with actions and reactions under more controlled experimental conditions, thus increasing the construct validity.

### 2018-04-11 Latex/sections/2descriptive.tex
Final author fixes. Off to proofreading.

@@
 In this section, I introduce and illustrate the relevant alternating constructions in Section~\ref{sec:descriptive}.
-I describe the narrowly defined syntactic configuration in which the alternation occurs.
 Then, I develop a model with prototype effects as well as exemplar effects based in existing research and my own theorising in Section~\ref{sec:analyses}.
@@
 
-\subsubsection{Core structures related to the alternation}
+\subsubsection{The measure noun phrase alternation}
+\label{sec:themeasurenounphrasealternation}
 
@@
 The case alternation only occurs when the kind noun is modified by an attributive adjective and there is no determiner, as in (\ref{ex:intro:alternation}).
-Superficially, the sentences are functionally and semantically equivalent either with the kind noun in the genitive (\ref{ex:intro:alternation1}) or in the same case as the measure noun, an accusative in the case of (\ref{ex:intro:alternation2}).%
+Superficially, the sentences are functionally and semantically equivalent either with the kind noun in the genitive (\ref{ex:intro:alternation1}) or in the same case as the measure noun -- an accusative in the case of (\ref{ex:intro:alternation2}).%
 \footnote{Some descriptive and normative grammars take stronger positions with regard to the acceptability of the two options.
@@
 }
+Notice that the genitive is virtually always used when the kind noun is a plural count noun, and the corresponding cases are consequently not included in the present study.
 
@@
   \begin{xlist}
-    \ex{\label{ex:starkschwach:a}\gll mit ein\textbf{-em} stark\textbf{-en} Kaffee\\
+    \ex{\label{ex:starkschwach:a}\gll mit ein-em stark-en Kaffee\\
                                       with a strong coffee\\}
-    \ex{\label{ex:starkschwach:b}\gll mit stark\textbf{-em} Kaffee\\
+    \ex{\label{ex:starkschwach:b}\gll mit stark-em Kaffee\\
                                       with strong coffee\\}
@@
 
-While the \PGCd\ in (\ref{ex:prototypes1}) allows an interpretation where speakers conceptualise the substance itself, \ie\ the medication, and a spoon used to take a quantity from the medication, the \NACb\ in (\ref{ex:prototypes2}) does not allow such an interpretation.
+While the \PGCd\ in (\ref{ex:prototypes1}) allows an interpretation where speakers conceptualise the substance itself (\ie\ the medication) and an actual spoon used to measure the quantity of the medication, the \NACb\ in (\ref{ex:prototypes2}) does not allow such an interpretation.
 While in the given examples, the effect is strongly supported by the choice of the measure noun lemmas, I argue that the meaning of the prototypes is independent of this.
 The independent meanings of the prototypes are the result of diachronic developments and grammaticalisation processes.
-Furthermore, I argue that the prototypical meanings are reflected in their usage patterns, which will be tested in the corpus study in Section~\ref{sec:corpusstudies}.
+Furthermore, I argue that the prototypical meanings are reflected in their usage patterns.
 
@@
 It is often assumed that pseudo-partitives and quantity constructions arise as a form of grammaticalised partitives (\eg\ \citealp[536--539]{Koptjevskaja2001} for Finnish and Estonian, \citealp[559]{Koptjevskaja2001} for European languages in general).
-The grammaticalisation paths described by \citet[esp.\ 526--530]{Koptjevskaja2001} are relevant for the case at hand.
-The grammaticalisation path can start out (in some languages) with constructions involving two referential nouns (not even necessarily forming a single and contiguous NP) and a \textit{separative} meaning as in \textit{(cut) two slices from the cake} \citep[535]{Koptjevskaja2001}.
+The grammaticalisation path described by \citet[esp.\ 526--530]{Koptjevskaja2001} can start out (in some languages) with constructions involving two referential nouns (not even necessarily forming a single and contiguous NP) and a \textit{separative} meaning as in \textit{(cut) two slices from the cake} \citep[535]{Koptjevskaja2001}.
 In this type of construction, it is most obvious that two separate referents (in the given example the cake and the slices) are conceptualised.
@@
 It is thus not surprising that it has lost (at least prototypically) the semantics which allows both the measure and the substance to be conceptually accessible as independent referents.
-As a consequence, we should expect the NAC constructions to host more strongly grammaticalised measure nouns.
+As a consequence, we should expect a tendency for the NAC constructions to host more strongly grammaticalised measure nouns.
 For example, highly grammaticalised non-referential nouns like \textit{Gramm} `gram' and \textit{Meter} `metre' should occur proportionally more often in NAC constructions than in PGC constructions.
@@
 In the corpus study, measure lemmas will therefore be annotated with appropriate semantic class labels to check whether semantic classes of measure nouns have different affinities to the two variants.
-The actual classification is based on the list in \citet[530]{Koptjevskaja2001}, but due to the low token frequencies in many of the potential classes, a very coarse classification was used in the end.
+The actual classification used is based on the list in \citet[530]{Koptjevskaja2001}, but due to the low token frequencies in many of the potential classes, a very coarse classification was used in the end.
 With typical examples, the classes are:
@@
 
-Finally, it was found that the \PGCa\ is more typical of higher stylistic levels (distinctly edited, closer to the non-regional standard, more formal) and\slash or even exclusive to written language (see \citealp[320--323]{Hentschel1993}).
+Finally, it was found that the \PGCa\ is more typical of higher stylistic levels (edited, closer to the non-regional standard, more formal) and\slash or even exclusive to written language (see \citealp[320--323]{Hentschel1993}).
 The genitive -- an intrinsic part of the PGC -- is rarer in colloquial vernacular variants of German compared to the written standard.
@@
 While true multilevel models are not used very often in corpus linguistics -- \citet{Gries2015} even calls the simpler varying-intercepts and varying-slopes models ``underused'' -- they provide an excellent tool to describe situations where preferences at the sentence level and at lexical levels need to be integrated.
-The models do not differentiate in and of themselves between abstraction and exemplar effects, but they allow researchers to tune the degree of complexity of models, incorporating both types of effects according to their theory and the phenomenon at hand.
+The models do not differentiate in and of themselves between abstraction and exemplar effects, but they allow researchers to tune the degree of complexity of models, incorporating both types of effects according to their analysis and the phenomenon at hand.
 In Sections~\ref{sec:prestudy} and~\ref{sec:corpushierarchicalmodel}, multilevel models will therefore be used.

### 2018-04-11 Latex/sections/3corpus.tex
Final author fixes. Off to proofreading.

@@
 I chose this corpus for two main reasons.
-First, the external validity of any study is increased through a higher heterogeneity of the sample \citep[30]{MaxwellDelaney2004}, and the DECOW14A corpus has clearly a much more heterogeneous composition compared to the only other very large corpus of German, the DeReKo \citep{KupietzEa2010} of the Institute for the German Language (IDS), which contains almost exclusively newspaper texts.%
-\footnote{It was shown in \cite{W16-2601} that, for example, the range of topics covered is much smaller in DeReKo compared to DECOW14A.}
+First, the external validity of any study is increased through a higher heterogeneity of the sample \citep[30]{MaxwellDelaney2004}, and the DECOW14A corpus has clearly a much more heterogeneous composition compared to the only other very large corpus of German, the DeReKo of the Institute for the German Language \citep{KupietzEa2010}, which contains almost exclusively newspaper texts.%
+\footnote{It was shown in \cite{W16-2601} that, for example, the range of topics is much smaller in DeReKo compared to DECOW14A.}
 Second, it was already mentioned that normative grammars often adopt clear positions regarding the grammaticality of either the \NACa\ or the \PGCa.
-Thus, newspaper text or any other text that conforms strongly to normative grammars might not represent the alternation phenomenon fully (and without bias) because authors and proofreaders who must adhere to normative guidelines might favour one alternative or the other explicitly.
-Web corpora, on the other hand, contain at least some amount of non-standard language from forums and similar sources.
-For these or similar reasons, COW corpora have been used in a number of peer-reviewed publications, for example \cite{VanGoethemHiligsmann2014}, \cite{VanGoethemHuening2015}, \cite{MuellerS2014}, \cite{Schaefer2016c}, \cite{SchaeferSayatz2014}, \cite{SchaeferSayatz2016}, and \cite{Zimmer2015}.
+Thus, newspaper text or any other text that conforms strongly to normative grammars might not represent the alternation phenomenon fully and without bias because authors and proofreaders who must adhere to normative guidelines might favour one alternative or the other explicitly.
+Web corpora, on the other hand, contain a significant amount of non-standard language from forums and similar sources.
+For such reasons, COW corpora have been used in a number of peer-reviewed publications, for example \cite{VanGoethemHiligsmann2014}, \cite{VanGoethemHuening2015}, \cite{MuellerS2014}, \cite{Schaefer2016c}, \cite{SchaeferSayatz2014}, \cite{SchaeferSayatz2016}, and \cite{Zimmer2015}.
 Therefore, DECOW14A is a valid choice for this study.
@@
 
-Among the factors potentially influencing the alternation (see Section~\ref{sec:germanmeasurenps}) are lemma-specific preference effects.
+Among the factors potentially influencing the alternation (see Section~\ref{sec:germanmeasurenps}) are lemma-specific effects.
 Therefore, it was desirable to obtain a sample in which most of the highly frequent actually-occurring combinations of kind nouns and measure nouns were represented.
@@
 In the first step, I exported a list of all nouns from the DECOW14A01 sub-corpus (one billion tokens large).
-The one hundred most frequent mass nouns were extracted manually from this list.
+Then, the one hundred most frequent mass nouns were extracted manually from this list.
 Mass nouns were defined as concrete nouns which genuinely denote a substance, combine with uninflected mass quantifiers such as \textit{viel} `much' and \textit{wenig} `little' (\textit{viel Bier} `much beer'), and form only sortal and unit plurals (such as the plural \textit{Biere} `types of beer' or `glasses of beer').
-Abstract nouns which partially behave like mass nouns (like \textit{Spaß} `fun’ or \textit{Gefahr} `danger’) were excluded because they are not quantified in the same way as concrete mass nouns.
+Abstract nouns which partially behave like mass nouns (like \textit{Spaß} `fun’ or \textit{Gefahr} `danger’) were excluded because they are not measured in the same way as concrete mass nouns.
 
@@
 From each of the 100 lists, I also removed noun-noun combinations that had a frequency below 2, except if the individual list would have otherwise been shorter than 20 noun-noun combinations.
-The result was a list of the most frequent 2,365 individual combinations of a measure noun and a mass noun.
+The result was a list of the most frequent 2,365 individual combinations of measure and mass nouns.
 
@@
 However, the goal here is to quantify how often lemmas occur in the \PGCd\ and the \NACb, and these constructions do not compete at all but are rather mutually exclusive.
-While this does not preclude the use of collostructional analysis, it is an open question whether the marginals are cognitively relevant in this case.
+While this does not preclude the use of CA, it is an open question whether the marginals are cognitively relevant in this case.
 After all, the main difference between the quotient used here and measures used in CA (originally signed logarithmised p-values from a Fisher test) is that they take the marginals into account. 
-However, using Fisher-$p$-based collexeme strength instead of the raw frequency quotient was tried as an alternative (variables \textit{Kindcollo} and \textit{Measurecollo}); see Section~\ref{sec:mainstudystatisticalmodel} for a discussion of the negative results.
+However, using Fisher-p-based collexeme strength instead of the raw frequency quotient was tried as an alternative (variables \textit{Kindcollo} and \textit{Measurecollo}).
+See Section~\ref{sec:mainstudystatisticalmodel} for a discussion of the negative results.
 
@@
 
-Furthermore, to control for simple frequency effects, \textit{Kindfreq} and \textit{Measurefreq} are the logarithm-transformed frequencies per 1,000,000 words of each lemma, extracted from the frequency lists distributed by the DECOW14A corpus creators on their web page.
+Furthermore, to control for simple frequency effects, \textit{Kindfreq} and \textit{Measurefreq} are included as the logarithm-transformed frequencies per 1,000,000 words of each lemma, extracted from the frequency lists distributed by the DECOW14A corpus creators on their web page.
 
@@
 
-The overall intercept is -5.370, and it represents \textit{Cardinal=Yes}, \textit{Measureclass=Physical} as 0 for all (z-transformed) numeric regressors.
-The standard deviation in the random intercepts is 1.157 for \textit{Measurelemma} ($p_{\text{PB}}=0.001$) and 0.923 for \textit{Kindlemma} ($p_{\text{PB}}=0.002$).
-Nakagawa \& Schielzeth's pseudo-coefficient of determination is $R_m^2=0.278$ and $R^2_c=0.566$ (see \citealp{Gries2015} or \citealp{NakagawaSchielzeth2013}).
+The overall intercept is -5.370, and it represents \textit{Cardinal=Yes}, \textit{Measureclass=Physical} and 0 for all (z-transformed) numeric regressors.
+The standard deviation of the random intercepts is 1.157 for \textit{Measurelemma} ($p_{\text{PB}}=0.001$) and 0.923 for \textit{Kindlemma} ($p_{\text{PB}}=0.002$).
+Nakagawa \& Schielzeth's pseudo-coefficients of determination are $R_m^2=0.278$ and $R^2_c=0.566$ (\citealp{Gries2015}; \citealp{NakagawaSchielzeth2013}).
 
@@
 
-The coefficients of determination indicate that the effect of the fixed effects (marginal $R^2$) is adequately strong, and that lemma-specific effects are equally strong with a difference between the two coefficients of 0.289.
+The coefficients of determination indicate that the influence of the fixed effects (marginal $R^2$) is adequately strong, and that lemma-specific effects are equally strong with a difference between the two coefficients of 0.289.
 
@@
 Also, The presence of a cardinal modifier clearly favours the \NACb.
-The stylistic factor \textit{Genitives} is confirmed to have an influence as predicted inasmuch as a general lack of genitives (which is encoded as a \textit{higher} value of the variable) favours the \NACb.
+The stylistic factor \textit{Genitives} shows the predicted influence inasmuch as a general lack of genitives (which is encoded as a \textit{higher} value of the variable) favours the \NACb.
 \textit{Badness} points into the wrong direction, but it barely reaches sig=0.05, and its confidence interval includes 0.
@@
 \footnote{In a similar fashion, the 100 most frequent measure nouns occurring with plural kind nouns were listed and queried, resulting in a sample of 871 sentences.
-As stated in Section~\ref{sec:germanmeasurenps}, the \NACa\ is virtually never used with plural kind nouns, and this sample was not used except for quantifying the frequency of occurrence of the constructions (67 times \NACa\ and 794 times \PGCa).
+As stated in Section~\ref{sec:themeasurenounphrasealternation}, the \NACa\ is virtually never used with plural kind nouns, and this sample was not used except for quantifying the frequency of occurrence of the constructions (67 times \NACa\ and 794 times \PGCa).
 The sample is distributed with the data package accompanying this paper.}
@@
 Finally, two auxiliary samples were also drawn.
-As mentioned in Section~\ref{sec:analyses}, the distribution of the measure noun and kind noun lemmas in the \NACb\ and the \PGCd\ with a determiner will be modelled as factors influencing the alternation.
+As mentioned in Section~\ref{sec:analyses}, the distribution of the measure and kind noun lemmas in the \NACb\ and the \PGCd\ with a determiner will be modelled as factors influencing the alternation.
 Therefore, all noun-noun pairs from the process described above were also queried in the two non-alternating constructions, resulting in 17,252 hits for the \PGCd\ and 315,635 hits for the \NACb.
@@
 The sample size was \textit{n=}5,063 with 1,134 cases of \PGCa\ and 3,929 cases of \NACa.
-The results of the estimation are shown in Table~\ref{tab:bigtable} and in Figure~\ref{fig:fixeffs}.
+The results of the estimation are shown in Figure~\ref{fig:fixeffs} and in Table~\ref{tab:bigtable}.
 The intercept comprises \textit{Cardinal=Yes}, \textit{Measurecase=Nom}, \textit{Kindgender=Masc}, \textit{Measureclass=Physical}, and 0 for all numeric z-transformed regressors.
@@
 
-The regressors with the measure lemma as their unit of reference have no within-measure lemma variance, and the \textit{glmer} function automatically estimates them as group-level predictors (or second-level effects), cf.\ \citet[265--269,302--304]{GelmanHill2006}.
+The regressors with the measure lemma as their unit of reference have no within-measure lemma variance, and the \textit{glmer} function automatically estimates them as group-level predictors (second-level effects), cf.\ \citet[265--269,302--304]{GelmanHill2006}.
 The same goes for those listed with the kind lemma as their unit of reference.
@@
 Standard diagnostics show that the model quality is quite good.
-Nakagawa \& Schielzeth's pseudo-coefficient of determination is $R_m^2=0.409$ and $R^2_c=0.495$.
+Nakagawa \& Schielzeth's pseudo-coefficients of determination are $R_m^2=0.409$ and $R^2_c=0.495$.
 The rate of correct predictions is 0.843, which means a proportional reduction of error of $\lambda=0.297$.
@@
 Using signed logarithmised Fisher p-values as a measure of collexeme strength (\textit{Measurecollo} and \textit{Kindcollo}) instead of the quotient for the attraction strength (\textit{Measureattraction} and \textit{Kindattraction}) (see Section~\ref{sec:variablesandannotation}) was not successful.
-While the attraction measures reach satisfying sig levels in the PBmodcomp test (0.020 for \textit{Kindattraction} and 0.001 for \textit{Measureattraction}), the $p_{\text{PB}}$ value for \textit{Kindcollo} was 0.191 and the one for \textit{Measurecollo} was 0.443.
-The coefficients of determination drop to $R_m^2=0.376$ and $R^2_c=0.480$.
+The $p_{\text{PB}}$ value for \textit{Kindcollo} was 0.191 and the one for \textit{Measurecollo} was 0.443.
+The coefficients of determination dropped to $R_m^2=0.376$ and $R^2_c=0.480$.
 
@@
 
-The results reported in Section~\ref{sec:corpushierarchicalmodel} generally confirm the hypotheses from Section~\ref{sec:analyses}.
-First, the prototypicality effect related to the non-alternating \PGCd\ and \NACb\ can be shown (see the effect plots in Figure~\ref{fig:eff:attraction}).%
+The results generally confirm the hypotheses from Section~\ref{sec:analyses}.
+First, the exemplar effect (the influence of the non-alternating \PGCd\ and \NACb) was shown (see the effect plots in Figure~\ref{fig:eff:attraction}).%
 \footnote{Effect plots were created using the \textit{effects} package \citep{Fox2003}.}
@@
 An interesting picture emerges for the lemma frequencies.
-A higher-than-average lemma frequency of measure nouns favours the \NACa, which is as expected if we assume at least a tendency for highly grammaticalised items to be more frequent.
+A higher-than-average lemma frequency of measure nouns favours the \NACa, which is as expected if we assume at least a tendency for highly grammaticalised words to be more frequent.
 With kind nouns, higher frequency seems to favour the \PGCa.
@@
 Figure~\ref{fig:eff:leftcontext} shows that cardinals indeed influence the choice of the alternant, and that cardinals have a strong tendency to co-occur with the \NACa.
-This effect was predicted in Section~\ref{sec:analyses}.
 
@@
 
-The influence of \textit{Measurecase} is as predicted in previous analyses (see Section~\ref{sec:analyses}).
+The influence of \textit{Measurecase} is as predicted in previous analyses.
 A measure noun in the dative favours the \PGCa\ (compared to the nominative, which is on the intercept).
@@
 However, it would be surprising if we achieved perfectly converging results given that even real effects are missed at certain rates in empirical studies.
-In the case of \textit{Badness}, we see that even in the main study, the effect size is small (-0.152), and thus even at the impressive sample size of roughly 5,000 in both studies, the effect might simply be too weak to be detected reliably.
+In the case of \textit{Badness}, we see that even in the main study, the post-hoc effect size is small (-0.152), and thus even at the impressive sample size of roughly 5,000 in both studies, the effect might simply be too weak to be detected reliably.
 Finally, the frequency effect \textit{Measurefreq} is detected in the main study but not in the pre-study, while the \textit{Kindfreq} effect is essentially absent in both studies.

### 2018-04-11 Latex/sections/4experiments.tex
Final author fixes. Off to proofreading.

@@
 
-In this section, the probabilities of the alternants as calculated for attested material are correlated with the reactions of participants in controlled experiments.
+In this section, the probabilities of the alternants (as calculated from attested material in the corpus study) are correlated with the reactions of participants in controlled experiments.
 Thus, a direct link is established between output material found in corpora and the behaviour of linguistic agents.
@@
 The decision to use attested material and perform a global validation of the corpus-based model was not arbitrary.
-Alternatively, one could have created stimuli where the features favouring the alternatives were permuted appropriately and thus tested directly.
+Alternatively, one could have created stimuli where the features favouring the alternatives were permuted and thus tested directly.
 The answer is related to what was said in Section~\ref{sec:cogocl} about the rich multi-factorial data sets used in corpus studies and the comparatively restricted ones in experiments.
@@
 Since the dependent variable is binary (\textit{Construction}), permuting these factors alone leads to 180 different possible permutations.
-On top of this, the numeric variables \textit{Kindattraction} and \textit{Measureattraction} would also have to be tested at several values.
+Additionally, the numeric variables \textit{Kindattraction} and \textit{Measureattraction} would also have to be tested at several values.
 It is obvious that this is impossible in a controlled experiment, given that a participant can be exposed to no more than roughly one hundred sentences, most of which would be fillers rather than target sentences.
 Thus, while the present approach only allows for a global test of the corpus-derived model, this appears to me to be the only feasible way.%
-\footnote{For the same reasons, the experimental data are also much too sparse to perform even post-hoc analyses with respect to the single regressors in a sound way.}
-It is, in the sense of Section~\ref{sec:cogocl}, an optimal synthesis of data-rich multifactorial corpus studies and experimental validation, and it is one that has been used (although rarely; see \citealp[3--4]{DivjakEa2016a}) at least since the seminal \citet{BresnanEa2007} paper.
+\footnote{For the same reasons, the experimental data are also much too sparse to perform post-hoc analyses with respect to the single regressors in a sound way.}
+It is, in the sense of Section~\ref{sec:cogocl}, an optimal synthesis of data-rich multifactorial corpus studies and experimental validation, and it is one that has been used at least since the seminal \citet{BresnanEa2007} paper -- although rarely, as \citet[3--4]{DivjakEa2016a} point out.
 
@@
 They were recruited from introductory linguistics courses at Freie Universität Berlin.
-Although the experiment was conducted in the last four weeks of their first semester, participants had no deeper explicit knowledge of linguistics, grammar, or experimental methods.
+Although the experiment was conducted in the last four weeks of their first semester, participants had no deeper knowledge of linguistics, grammar, or experimental methods.
 None of them had ever participated in a forced-choice experiment before.
@@
 The higher the probability of the \PGCa\ predicted from usage data, the more often participants chose the \PGCa\ alternant in the forced-choice task.
-A closer look at the results in the form of the spineplot in Figure~\ref{fig:spines} shows, however, that it was likely an idiosyncracy of a single sentence with a probability predicted by the model between 0.5 and 0.6 which spoiled an otherwise much better correlation.
+A closer look at the results in the form of the spineplot in Figure~\ref{fig:spines} shows, however, that it was likely an idiosyncracy of a single sentence which spoiled an otherwise much better correlation.
 The problematic sentence with a model prediction of 0.548 is given in (\ref{ex:problem}) in the \PGCa\ variant.
@@
 It is expected that reading less typical alternants in a given context and with given lexical material incurs a processing overhead for the reader (\citealp{Kaiser2013}).
-In this section, a self-paced reading experiment is therefore presented.
+In this section, a self-paced reading experiment is presented.
 In a very similar fashion, \cite{DivjakEa2016} apply the self-paced reading paradigm in the validation of corpus-based models.
@@
 
-The reading times were residualised per speaker based on the reading times of all words (not just the targets) by that speaker.
+The reading times were residualised per speaker based on the reading times of all words (not just the targets) for that speaker.
 The adjective and the kind noun (\ie\ the constituents bearing the critical case markers) were used as the target region, for instance the bracketed words in the example \textit{zwei Gläser} [\textit{spru\-deln\-des Wasser}] `two glasses of sparkling water'.
@@
 
-Furthermore, the position (1--98) of the sentence in the individual experiment (\textit{Position}) was included as a fixed effect to control for the usual increase in reading speed during an experiment run.
+Furthermore, the position (1--98) of the sentence in the individual experiment (\textit{Position}) was included as a fixed effect to control for the usual increase in reading speed during an experiment.
 Random intercepts were specified for \textit{Participant} and \textit{Item} (the 16 sentence pairs are one \textit{Item} each).%
@@
 
-The coefficients of determination indicate that there is a noteworthy correlation between the reactions of the subjects and the corpus-derived probabilities (marginal $R^2$) and that there is some between-subject variation (conditional minus marginal $R^2$, which is 0.107).
+The coefficients of determination indicate that there is a noteworthy correlation between the reactions of the subjects and the corpus-derived probabilities (marginal $R^2$) and that there is some between-subject variation: the difference between $R^2_m$ and $R^2_c$ is 0.107.
 

### 2018-04-11 Latex/sections/5conclusions.tex
Final author fixes. Off to proofreading.

@@
 The overall convergence between the pre-study and the main study and the convergence between corpus data and the experimental cross-validation strengthen the validity of this study.
-Also, the fact that the result for \textit{Measurecase} (although it was only a control variable in the present context) converged with previous research \citep{Zimmer2015} lends credibility to the results.
+Also, the fact that the result for \textit{Measurecase} converged with previous research \citep{Zimmer2015} lends credibility to the results.
 
@@
 This shows that the model indeed partially predicts the variant which language users expect.
-The overall effect as measured in the $R^2$ was weak (roughly 0.2) in both cases, however.
-While part of this could be traced back to one suboptimally chosen stimulus, we really need to consider what kind of convergence we expect to see between corpus data and experiments.
+The overall effect as measured in the $R^2$ was acceptable but not strong (roughly 0.2) in both cases, however.
+While this could be traced back at least partially to one suboptimally chosen stimulus, we really need to consider what kind of convergence we expect to see between corpus data and experiments.
 The main corpus study had $R^2_m=0.409$ and $R^2_c=0.495$, which is good but not anywhere near a perfect fit.
 With the added inter-speaker variability brought into play by the experimental setup (compared to the averaging across thousands of speakers in the corpus study), a perfect fit cannot be expected.
-Like many phenomena in German which are often called \textit{Zweifelsfälle} (`cases of doubt') in the traditional literature \citep{Duden09,Klein2009} the MNP alternation is a case where speakers often have no clear intuition, and a lot of free variation seems to be involved.
-Rarely do speakers feel that one alternative is clearly odd or bad.
+Furthermore, like many alternations in German which are often labelled as \textit{Zweifelsfälle} (`cases of doubt') in the traditional literature \citep{Duden09,Klein2009}, the MNP alternation is a case where speakers often have no clear intuition, and a lot of free variation seems to be involved.
+Rarely do speakers feel that one alternative is clearly odd or ungrammatical.
 Additionally, cases of doubt involving the genitive are often a matter of fierce normative public debate, and especially the forced choice paradigm does not effectively prevent participants from making normative judgements.
@@
 On a larger methodological scale, this paper also makes a number of contributions.
-The statistical model was a true multilevel model (see Sections~\ref{sec:exemplareffects} and~\ref{sec:mainstudystatisticalmodel}), demonstrating how multilevel modelling helps to specify complex hierarchies of abstract features, item-specific tendencies, and exemplar effects at the level of observations (sentences; first level) and lemmas (second level).
+The statistical model was a true multilevel model (see Section~\ref{sec:mainstudystatisticalmodel}), demonstrating how multilevel modelling helps to specify complex hierarchies of abstract features, item-specific tendencies, and exemplar effects at the level of observations (sentences; first level) and lemmas (second level).
 \cite{Gries2015} still calls mixed models ``underused'' in corpus linguistics, and multilevel models are consequently also an underused tools.
-At the same time, the effects of overparametrisation of mixed models with varying slopes as criticised by \cite{BatesEa2015a} were demonstrated.
+At the same time, the ill effects of overparametrisation of mixed models with varying slopes as criticised by \cite{BatesEa2015a} were demonstrated.
 Furthermore, fitting an additive model to the reading time data did not improve the fit as there were no non-linearities in the data.
@@
 This is clearly another area where only more studies can lead to clarification.
-Finally, much like \cite{Dabrowska2014} found that speakers' knowledge of collocations was not matched by a set of standard measures of collocation strength extracted from corpora, a simple quotient based on raw frequencies for the attraction strength performed much better in the present study compared to collexeme strength in the form of logarithmised Fisher p-values (see Section~\ref{sec:corpusstudies}).
+Finally, much like \cite{Dabrowska2014} found that speakers' knowledge of collocations was not matched by a set of standard measures of collocation strength extracted from corpora, a simple quotient based on raw frequencies for the attraction strength performed much better in the present study compared to collexeme strength in the form of logarithmised Fisher p-values.
 While this does not allow the conclusion that collexeme strength has no cognitive reality, it might indicate that for measuring attraction effects exerted by non-alternating constructions on alternating constructions, taking the marginals into account might not be crucial.%

### 2018-04-18 Latex/sections/3corpus.tex
Important fix (R2 value mix-up).

@@
 In connection with this, it is revealing to look at the coefficients of determination.
-In the pre-study, a much greater proportion of the variance is explained by taking the lemma random effects into account ($R^2_m=0.278$ and $R^2_c=0.566$) than it is in the main study ($R^2_m=0.376$ and $R^2_c=0.480$).
+In the pre-study, a much greater proportion of the variance is explained by taking the lemma random effects into account ($R^2_m=0.278$ and $R^2_c=0.566$) than it is in the main study ($R^2_m=0.409$ and $R^2_c=0.495$).
 Thus, the frequency effect for measure lemmas might be swamped by the lemma random intercepts.

### 2018-04-21 Latex/sections/1intro.tex
First half of proofreading.

@@
 This question can and should be answered by testing how well corpus-derived models converge with or diverge from experimental findings (Secion~\ref{sec:usagedataincognitivelinguistics}).
-Second, the appropriate modelling of such results in cognitive linguistics -- \ie\ the assumed constructs -- is a key issue (Section~\ref{sec:prototypesexemplarscorporaandcontrolledexperiments}).
+Second, the appropriate modelling of such results in cognitive linguistics is a key issue (Section~\ref{sec:prototypesexemplarscorporaandcontrolledexperiments}).
 
@@
 
-For some time, there has been an interest in correlating probabilistic generalisations extracted from corpus data and results from experimental work (for example, \citealp{ArppeJaervikivi2007,BresnanEa2007,BresnanFord2010,DivjakGries2008,DivjakEa2016,FordBresnan2013}).
+For some time, there has been an interest in correlating probabilistic generalisations extracted from corpus data and results from experimental work \citep{ArppeJaervikivi2007,BresnanEa2007,BresnanFord2010,DivjakGries2008,DivjakEa2016,FordBresnan2013}.
 This is often called a \textit{validation} of the corpus-derived findings, but \citet[303]{Divjak2016a} criticises this choice of words ``because it creates the impression that behavioral experimental data is inherently more valuable than textual data'', citing \cite{TummersEa2005}, who state that a corpus is ``a sample of spontaneous language use that is (generally) realized by native speakers''.
@@
 Rather, predictions are derived from existing theories in order to \textit{test} the theories.
-While this approach is applicable to corpus data, I discuss some relevant differences.
+While this approach is applicable to corpus data, I discuss some relevant differences below.
 
@@
 They state that ``in this wider context, large, register-mixed corpora such as the British National Corpus [\ldots] may not be perfect models of the linguistic experience of adult speakers, but they are reasonably close to the input of an idealized average member of the relevant speech community'' \citep[104]{StefanowitschFlach2016}.
-In essence, this entails that any generalisation extracted from the BNC can be assumed to have some kind of mental reality, which is at least doubtful.
+If this were a valid assumption then any generalisation extracted from the BNC could be assumed to have some kind of mental reality, which is doubtful.
 
@@
 While I am convinced that the results presented in \citet{DivjakArppe2013} are valid and important (especially given the previous and subsequent research the authors have conducted on the data, including experimental work presented in \citealp{DivjakEa2016}), any data set can be clustered to yield a certain number of clusters.
-Thus, the study does not ensure that the clusters emerging from the data correspond to any speaker's cognitive representation.
+Thus, the study does not ensure that the clusters emerging from the data correspond to any particular speaker's cognitive representation.
 
@@
 Tasks in cognitive science have been criticised exactly for their lack of external validity, for example by \citet{Murphy2003}.
-From a linguistic perspective, it seems remarkable that \citet[1013]{VoorspoelsEa2011} consider their experimental task -- the assignment of typicality scores to nouns from the domains of \textit{animals} and \textit{artefacts} to categories like \textit{bird}, \textit{fish}, \textit{clothing}, or \textit{tools} -- a study of ``superordinate natural language categories, whereas most evidence supporting exemplar representations has been found in artificial categories of a more subordinate level''.
+From a linguistic perspective, it seems remarkable that \citet[1013]{VoorspoelsEa2011} consider their experimental task -- the assignment of typicality scores to nouns from the domains of \textit{animals} and \textit{artefacts} to categories such as \textit{bird}, \textit{fish}, \textit{clothing}, or \textit{tools} -- a study of ``superordinate natural language categories, whereas most evidence supporting exemplar representations has been found in artificial categories of a more subordinate level''.
 Corpus linguists interested in probabilistic alternation modelling investigate significantly more complex high-level categories and use large and complex feature sets.%
@@
 It is thus an advantage of much linguistic work on categorisation that it deals with complex and realistically produced data because this greatly improves the external validity of studies, although by potentially sacrificing some construct validity.
-Thus, an ideal contribution to the research on category abstraction by cognitive corpus linguists is to provide analyses which have great external validity and complexity while carefully making sure that these finding correlate with actions and reactions under more controlled experimental conditions, thus increasing the construct validity.
+Thus, the ideal contribution to the research on category abstraction by cognitive corpus linguists is to provide analyses which have high external validity and complexity while carefully making sure that these finding correlate with actions and reactions under more controlled experimental conditions, thus increasing the construct validity.
 
@@
 In Section~\ref{sec:corpusstudies}, I present a corpus study and report a true multilevel generalised linear model with the appropriate hierarchical structure for the hypothesised effects.
-In Section~\ref{sec:experimental}, I test the predictions of the corpus-based model in two experimental paradigms (forced-choice and self-paced reading), showing that they indeed converge, albeit with low effect strength.
+In Section~\ref{sec:experimental}, I test the predictions of the corpus-based model in two experimental paradigms (forced-choice and self-paced reading), demonstrating that they indeed converge, albeit with low effect strength.
 In Section~\ref{sec:conclusion}, I interpret the findings in the light of the issues of cognitive representations and corpus data and the convergence of usage data and experimental data.

### 2018-04-21 Latex/sections/2descriptive.tex
First half of proofreading.

@@
 In this section, I introduce and illustrate the relevant alternating constructions in Section~\ref{sec:descriptive}.
-Then, I develop a model with prototype effects as well as exemplar effects based in existing research and my own theorising in Section~\ref{sec:analyses}.
+Then, I develop a model with prototype effects as well as exemplar effects based on existing research and my own theory in Section~\ref{sec:analyses}.
 
@@
 In NPs with a strongly inflected determiner, attributive adjectives inflect according to the massively syncretic \textit{weak} pattern.
-If there is no determiner (as is the case in the alternating constructions), attributive adjectives inflect like determiners themselves.
+If there is no determiner (as is the case in the alternating constructions), attributive adjectives inflect as determiners themselves.
 This is called the \textit{strong} inflectional pattern.
@@
 On the one hand, they are lexical adjectives and function as attributive modifiers.
-On the other hand, they are inflected like determiners, and they are the leftmost element in the NP, which is typical of determiners.
+On the other hand, they are inflected as are determiners, and they are the leftmost element in the NP, which is typical of determiners.
 This unusual double nature of adjectives in NPs without determiners leads to a plausible interpretation of the pattern shown in Table~\ref{tab:constructions}.
@@
     they took a spoon some medication \\
-    \trans They took a spoon(ful) of some medication.
+    \trans They took a spoon(ful) of any old medication.
   }
@@
 As predicted by the grammaticalisation pattern just described, it still has the potential to form a true partitive (if the kind noun is definite).
-Conversely, the NAC lacks this ability to form true partitives and has gone further down the grammaticalisation path.
+Conversely, the NAC lacks this ability to form true partitives and has advanced further down the grammaticalisation path.
 It is thus not surprising that it has lost (at least prototypically) the semantics which allows both the measure and the substance to be conceptually accessible as independent referents.

### 2018-04-21 Latex/sections/3corpus.tex
Second half of proofreading.

@@
 
-For the present study, I used the German \textit{Corpus from the Web} (COW) in its 2014 version DECOW14A (\citealp{SchaeferBildhauer2012full}, and \citealp{Schaefer2015b}, as well as \citealp{BiemannEa2013}, and \citealp{SchaeferBildhauer2013}, for overviews of web corpora in general and the methodology of their construction), which contains almost 21 billion tokens.%
+For the present study, I used the 21 billion token German \textit{Corpus from the Web} (COW) in its 2014 version DECOW14A (\citealp{SchaeferBildhauer2012full}, and \citealp{Schaefer2015b}, as well as \citealp{BiemannEa2013}, and \citealp{SchaeferBildhauer2013}, for overviews of web corpora in general and the methodology of their construction).%
 \footnote{The COW corpora (Dutch, English, French, German, Spanish, Swedish) are made available for free at \url{https://www.webcorpora.org}.
 At the time of this writing, a newer 2016 version DECOW16A has already been released.}
-I chose this corpus for two main reasons.
-First, the external validity of any study is increased through a higher heterogeneity of the sample \citep[30]{MaxwellDelaney2004}, and the DECOW14A corpus has clearly a much more heterogeneous composition compared to the only other very large corpus of German, the DeReKo of the Institute for the German Language \citep{KupietzEa2010}, which contains almost exclusively newspaper texts.%
-\footnote{It was shown in \cite{W16-2601} that, for example, the range of topics is much smaller in DeReKo compared to DECOW14A.}
-Second, it was already mentioned that normative grammars often adopt clear positions regarding the grammaticality of either the \NACa\ or the \PGCa.
-Thus, newspaper text or any other text that conforms strongly to normative grammars might not represent the alternation phenomenon fully and without bias because authors and proofreaders who must adhere to normative guidelines might favour one alternative or the other explicitly.
+This corpus was chosen first because the external validity of any study is increased through a more heterogeneous sample \citep[30]{MaxwellDelaney2004}.
+DECOW14A clearly has a much more heterogeneous composition compared with the only other very large corpus of German, the DeReKo of the Institute for the German Language \citep{KupietzEa2010}.
+DeReKo contains almost exclusively newspaper texts.%
+\footnote{It was shown in \cite{W16-2601} that, for example, the range of topics is much smaller in DeReKo than in DECOW14A.}
+The second reason for choosing this particular corpus is that normative grammars often adopt clear positions regarding the grammaticality of either the \NACa\ or the \PGCa.
+Thus, newspaper text or any other texts which conform strongly to normative grammars may not represent the alternation phenomenon fully and without bias because authors and proofreaders who must adhere to normative guidelines might explicitly favour one alternative over another.
 Web corpora, on the other hand, contain a significant amount of non-standard language from forums and similar sources.
-For such reasons, COW corpora have been used in a number of peer-reviewed publications, for example \cite{VanGoethemHiligsmann2014}, \cite{VanGoethemHuening2015}, \cite{MuellerS2014}, \cite{Schaefer2016c}, \cite{SchaeferSayatz2014}, \cite{SchaeferSayatz2016}, and \cite{Zimmer2015}.
-Therefore, DECOW14A is a valid choice for this study.
+For such reasons, COW corpora have been used in a number of peer-reviewed publications \citep{VanGoethemHiligsmann2014,VanGoethemHuening2015,MuellerS2014,Schaefer2016c,SchaeferSayatz2014,SchaeferSayatz2016,Zimmer2015}.
 
@@
 Among the factors potentially influencing the alternation (see Section~\ref{sec:germanmeasurenps}) are lemma-specific effects.
-Therefore, it was desirable to obtain a sample in which most of the highly frequent actually-occurring combinations of kind nouns and measure nouns were represented.
-I applied a two-stage process in order to obtain a list of co-occurring measure nouns and kind nouns.
+Therefore, it was necessary to obtain a sample in which most of the highly frequent actually-occurring combinations of kind nouns and measure nouns were represented.
+Hence, I applied a two-stage process in order to obtain a list of co-occurring measure nouns and kind nouns.
 
@@
 Mass nouns were defined as concrete nouns which genuinely denote a substance, combine with uninflected mass quantifiers such as \textit{viel} `much' and \textit{wenig} `little' (\textit{viel Bier} `much beer'), and form only sortal and unit plurals (such as the plural \textit{Biere} `types of beer' or `glasses of beer').
-Abstract nouns which partially behave like mass nouns (like \textit{Spaß} `fun’ or \textit{Gefahr} `danger’) were excluded because they are not measured in the same way as concrete mass nouns.
+Abstract nouns which partially behave as mass nouns (for example, \textit{Spaß} `fun’ or \textit{Gefahr} `danger’) were excluded because they are not measured in the same way as concrete mass nouns.
 
@@
 Documents in DECOW14 have also been annotated with a variable called \textit{Genitives}.
-The higher the values of this variable, the lower the proportion of genitives among all case-bearing forms is.
+For each document it is the case that the higher the values of this variable, the lower the proportion of genitives among all case-bearing forms.
 A high number of genitives is also indicative of a more formal, elaborate style close to the written standard.
@@
 
-Finally, two variables were added as nuisance variables in the context of the present study.
-First, it was reported in the literature that MNPs in the dative and with a masculine or neuter kind noun favour the \PGCa\ more than the corresponding nominative and accusative MNPs \citep{Hentschel1993,Zimmer2015}.
+Finally, two variables have been added as nuisance variables in the context of the present study.
+First, the case of the measure noun was manually annotated (variable \textit{Measurecase}).
+This was carried out because it was reported in the literature that MNPs in the dative containing a masculine or neuter kind noun favour the \PGCa\ more than the corresponding nominative and accusative MNPs \citep{Hentschel1993,Zimmer2015}.
 As an example, \textit{mit einem Stück frischen Brots} `with a piece of fresh bread' (\PGCa) would be preferred more strongly against \textit{mit einem Stück frischem Brot} (\NACa).
 As with all the examples, native speakers of German will most likely notice that differences are subtle.
-The case of the measure noun was manually annotated (variable \textit{Measurecase}).
 Second, due to differences in case syncretisms, it is likely that feminine kind nouns have slightly different preferences than masculine and neuter ones, and hence the appropriate variable \textit{Kindgender} was annotated.
@@
 The results of the estimation are shown in Figure~\ref{fig:fixeffs} and in Table~\ref{tab:bigtable}.
-The intercept comprises \textit{Cardinal=Yes}, \textit{Measurecase=Nom}, \textit{Kindgender=Masc}, \textit{Measureclass=Physical}, and 0 for all numeric z-transformed regressors.
-It was estimated at -3.548.
+The intercept estimated at -3.548 comprises \textit{Cardinal=Yes}, \textit{Measurecase=Nom}, \textit{Kindgender=Masc}, \textit{Measureclass=Physical}, and 0 for all numeric z-transformed regressors.
 
@@
 
-Using signed logarithmised Fisher p-values as a measure of collexeme strength (\textit{Measurecollo} and \textit{Kindcollo}) instead of the quotient for the attraction strength (\textit{Measureattraction} and \textit{Kindattraction}) (see Section~\ref{sec:variablesandannotation}) was not successful.
+The use of signed logarithmised Fisher p-values as a measure of collexeme strength (\textit{Measurecollo} and \textit{Kindcollo}) instead of the quotient for the attraction strength (\textit{Measureattraction} and \textit{Kindattraction}) (see Section~\ref{sec:variablesandannotation}) was not successful.
 The $p_{\text{PB}}$ value for \textit{Kindcollo} was 0.191 and the one for \textit{Measurecollo} was 0.443.
@@
 The results generally confirm the hypotheses from Section~\ref{sec:analyses}.
-First, the exemplar effect (the influence of the non-alternating \PGCd\ and \NACb) was shown (see the effect plots in Figure~\ref{fig:eff:attraction}).%
+First, the exemplar effect (the influence of the non-alternating \PGCd\ and \NACb) was validated (see the effect plots in Figure~\ref{fig:eff:attraction}).%
 \footnote{Effect plots were created using the \textit{effects} package \citep{Fox2003}.}
@@
 if a lemma appears relatively more often in the \PGCd\ (compared to its frequency in the \NACb), the \PGCa\ tends to be chosen over the \NACa\ with this specific lemma.
-The effect for measure nouns is stronger, and it was estimated with higher precision.
+The effect for measure nouns is even stronger and could be estimated with higher precision.
 

### 2018-04-21 Latex/sections/4experiments.tex
Second half of proofreading.

@@
 Alternatively, one could have created stimuli where the features favouring the alternatives were permuted and thus tested directly.
-The answer is related to what was said in Section~\ref{sec:cogocl} about the rich multi-factorial data sets used in corpus studies and the comparatively restricted ones in experiments.
+The answer is related to the statement in Section~\ref{sec:cogocl} about the rich multi-factorial data sets used in corpus studies and the comparatively restricted ones in experiments.
 Looking at Table~\ref{tab:variables}, we see one binary factor (\textit{Cardinal}), two three-level factors (\textit{Measurecase} and \textit{Kindgender}) and one five-level factor (\textit{Measureclass}).
@@
 \footnote{For the same reasons, the experimental data are also much too sparse to perform post-hoc analyses with respect to the single regressors in a sound way.}
-It is, in the sense of Section~\ref{sec:cogocl}, an optimal synthesis of data-rich multifactorial corpus studies and experimental validation, and it is one that has been used at least since the seminal \citet{BresnanEa2007} paper -- although rarely, as \citet[3--4]{DivjakEa2016a} point out.
+It is, in the sense of Section~\ref{sec:cogocl}, an optimal synthesis of data-rich multifactorial corpus studies and experimental validation, and it is one that has been used at least since the seminal \citet{BresnanEa2007} paper.
+As \citet[3--4]{DivjakEa2016a} pointed out, however, the numbers of published studies including validation is still inadequate.
 
@@
 
-The final pairs of stimuli were the sentence containing the attested and preferred alternant (according to the corpus GLMM) on the one hand and a modified version containing the dispreferred alternant on the other hand.
+The final pairs of stimuli were the sentences containing the attested and preferred alternant (according to the corpus GLMM) and a modified version containing the dispreferred alternant.
 They were presented next to each other, and a 20 second time limit for each choice was set.%

### 2018-04-23 Latex/article.tex
Final spellchecking.

@@
   I thank (in alphabetical order) Felix Bildhauer, Susanne Flach, Elizabeth Pankratz, Samuel Reichert, Ulrike Sayatz, and Christian Zimmer for valuable discussions and comments.
-  Also, I would like to thank the reviewers for Cognitive Linguistics as well as associate editor Dagmar Divjak for insightful comments which helped to improve the quality of this paper significantly.
+  Also, I would like to thank the three reviewers for Cognitive Linguistics as well as associate editor Dagmar Divjak for insightful comments which helped to improve the quality of this paper significantly.
   Furthermore, I thank Ulrike Sayatz for helping me to recruit the participants for the experiments.
-  Elizabeth Pankratz thankfully also fixed my English.
+  I am grateful to Sonya Faber and Elizabeth Pankratz (in alphabetical order) for fixing my English.
   Finally, I am grateful to my student assistants Kim Maser for her work on the annotation of the concordances and Luise Rißmann for supervising most of the experiments.

### 2018-04-23 Latex/sections/1intro.tex
Final spellchecking.

@@
 First, there is the question of whether corpus data provide any insight into cognitive representations at all.
-This question can and should be answered by testing how well corpus-derived models converge with or diverge from experimental findings (Secion~\ref{sec:usagedataincognitivelinguistics}).
+This question can and should be answered by testing how well corpus-derived models converge with or diverge from experimental findings (Section~\ref{sec:usagedataincognitivelinguistics}).
 Second, the appropriate modelling of such results in cognitive linguistics is a key issue (Section~\ref{sec:prototypesexemplarscorporaandcontrolledexperiments}).

### 2018-04-23 Latex/sections/4experiments.tex
Final spellchecking.

@@
 The higher the probability of the \PGCa\ predicted from usage data, the more often participants chose the \PGCa\ alternant in the forced-choice task.
-A closer look at the results in the form of the spineplot in Figure~\ref{fig:spines} shows, however, that it was likely an idiosyncracy of a single sentence which spoiled an otherwise much better correlation.
+A closer look at the results in the form of the spine plot in Figure~\ref{fig:spines} shows, however, that it was likely an idiosyncracy of a single sentence which spoiled an otherwise much better correlation.
 The problematic sentence with a model prediction of 0.548 is given in (\ref{ex:problem}) in the \PGCa\ variant.
@@
   \includegraphics[width=0.5\textwidth]{../R/output/fc_proportions}
-  \caption{Spineplot of the proportion of responses plotted against the predictions from the corpus-based model in the forced-choice experiment}
+  \caption{Spine plot of the proportion of responses plotted against the predictions from the corpus-based model in the forced-choice experiment}
   \label{fig:spines}
@@
 \footnote{\citet[1]{BatesEa2015a} state: ``We show that failure to converge typically is not due to a suboptimal estimation algorithm, but is a consequence of attempting to fit a model that is too complex to be properly supported by the data, irrespective of whether estimation is based on maximum likelihood or on Bayesian hierarchical modelling with uninformative or weakly informative priors.
-Importantly, even under convergence, overparameterization may lead to uninterpretable models.''
+Importantly, even under convergence, over-parameterisation may lead to uninterpretable models.''
 }

### 2018-04-23 Latex/article.tex
Final version.

@@
   I thank (in alphabetical order) Felix Bildhauer, Susanne Flach, Elizabeth Pankratz, Samuel Reichert, Ulrike Sayatz, and Christian Zimmer for valuable discussions and comments.
-  Also, I would like to thank the three reviewers for Cognitive Linguistics as well as associate editor Dagmar Divjak for insightful comments which helped to improve the quality of this paper significantly.
+  Also, I would like to thank the three reviewers for Cognitive Linguistics as well as associate editor Dagmar \mbox{Divjak} for insightful comments which helped to improve the quality of this paper significantly.
   Furthermore, I thank Ulrike Sayatz for helping me to recruit the participants for the experiments.

### 2018-04-23 Latex/sections/2descriptive.tex
Final version.

@@
 
-In this section, I introduce and illustrate the relevant alternating constructions in Section~\ref{sec:descriptive}.
+I introduce and illustrate the relevant alternating constructions in Section~\ref{sec:descriptive}.
 Then, I develop a model with prototype effects as well as exemplar effects based on existing research and my own theory in Section~\ref{sec:analyses}.
@@
 \footnote{If the kind noun is definite, the construction instantiates a true partitive.
-Whereas partitives are constructions denoting a proper part-of relation as in \textit{a sip of the wine}, pseudo partitives -- albeit syntactically similar and diachronically related to partitives in many languages -- merely denote quantities and contain indefinite kind nouns as in \textit{a sip of wine}.
+Whereas partitives are constructions denoting a proper part-of relation as in \textit{a sip of the wine}, pseudo partitives merely denote quantities and contain indefinite kind nouns as in \textit{a sip of wine}.
 In the literature on German, some authors incorrectly call the pseudo-partitive a \textit{partitive} \citep{Hentschel1993} while some realise the difference and at least mention it \citep{Eschenbach1994,GallmannLindauer1994,Loebel1989,Zimmer2015}.
@@
 \footnote{In the masculine and neuter singular genitive, the strong and weak forms are indistinguishable.
-Since the alternation itself does not occur in the genitive (see Section~\ref{sec:descriptive}), this does not create any complications for the present study.}
+Since the alternation itself does not occur in the genitive, this does not create any complications for the present study.}
 
@@
     \ex{\label{ex:prototypes1}\gll Sie nahmen [einen Löffel [irgendeiner Medizin]\Sub{Gen}]\Sub{Acc}.\\
-    they took a spoon some medication \\
+    they took a spoon {some\slash any} medication \\
     \trans They took a spoon(ful) of any old medication.
@@
 
-While the prototypes for the PGC and the NAC were specified with reference to high-level features in Section~\ref{sec:prototypeeffects}, at least one exemplar-driven similarity effect can also be expected to influence the MNP alternation.
+While the prototypes for the PGC and the NAC were specified with reference to high-level generalisations in Section~\ref{sec:prototypeeffects}, at least one exemplar-driven similarity effect can also be expected to influence the MNP alternation.
 The measure and kind noun lemmas which occur in the alternating constructions obviously also occur in the non-alternating constructions.
@@
 
-It should be noted that such an exemplar-type effect would have to be confirmed \textit{in addition} to the predicted semantic prototype effect for measure lemmas described in Section~\ref{sec:prototypeeffects}, namely the effect of semantic classes of measure nouns.
+It should be noted that such an exemplar-type effect would have to be confirmed \textit{in addition to} the predicted semantic prototype effect for measure lemmas described in Section~\ref{sec:prototypeeffects}, namely the effect of semantic classes of measure nouns.
 It must also be ensured that these influences at the lemma level are not spurious and just artefacts of mere lemma idiosyncrasies.

### 2018-04-23 Latex/sections/3corpus.tex
Final spellchecking.
Final version.

@@
 
-\subsection{Preliminiaries}
+\subsection{Preliminaries}
 
@@
 This list of mass nouns was used in the second step to derive a list of measure nouns co-occurring with the mass nouns. 
-In order to generate this list, I utilised the fact that a direct sequence of two nouns almost always instantiates the bare-noun NAC if the second noun is a mass noun.
+In order to generate this list, I utilised the fact that a direct sequence of two nouns almost always instantiates the \NACb\ if the second noun is a mass noun.
 I searched for all sequences N\Sub{1}N\Sub{2} where N\Sub{2} was one of the mass noun lemmas extracted in the first step.
@@
 With kind nouns, higher frequency seems to favour the \PGCa.
-However, this effect has no clear theoretical interpretation, and its coefficient estimate is imprecise (not significant at sig=0.05).
-The effect can therefore be ignored or treated as a nuisance variable.
+However, this effect has no clear theoretical interpretation, and its coefficient estimate is imprecise (not reaching sig=0.05).
+The effect can therefore be dismissed or treated as a nuisance variable.
 

### 2018-04-23 Latex/sections/5conclusions.tex
Final version.

@@
 
-In closing, I want to point out that the so-called \textit{case of doubt} in German morpho-syntax -- like the measure noun phrase alternation -- are in fact ideal test cases for probabilistic modelling of alternation phenomena in cognitive linguistics.
+In closing, I want to point out that the so-called \textit{cases of doubt} in German morpho-syntax -- like the measure noun phrase alternation -- are in fact ideal test cases for probabilistic modelling of alternation phenomena in cognitive linguistics.
 Doing more research on them could help to provide answers to many of the fundamental and methodological issues raised here.

### 2018-04-24 Latex/article.tex
Public pre-print version.

@@
 
-% \usepackage[firstpage]{draftwatermark}
-% \SetWatermarkFontSize{24pt}
-% \SetWatermarkHorCenter{300pt}
-% \SetWatermarkColor[rgb]{0.9,0.2,0.2}
-% \SetWatermarkText{\textsf{Revision 2}}
+\usepackage[firstpage]{draftwatermark}
+\SetWatermarkFontSize{24pt}
+\SetWatermarkHorCenter{300pt}
+\SetWatermarkColor[rgb]{0.9,0.2,0.2}
+\SetWatermarkText{\textsf{Final revision. Quote as draft and to appear in COGL.}}
 
@@
     It is also verified in two experiments that the corpus-derived model corresponds to the behaviour of native speakers.
-    The weak effect size of the experimental validation is discussed in the context of corpus-based cognitive linguistics and the validation of corpus-derived models.
-  }
+    The weak effect size of the experimental validation is discussed in the context of corpus-based cognitive linguistics and the validation of corpus-derived models.}
   \keywords{prototypes vs.\ exemplars, corpus methods and experimental validation, alternations, multilevel modeling, pseudo-partitives, German}
@@
   %\dedication{...}
-  %\received{...}
-  %\accepted{...}
-  %\journalname{...}
+  \received{18 April 2017}
+  \accepted{3 April 2018}
+  \journalname{To appear in Cognitive Linguistics}
   %\journalyear{}
@@
 % For word counting.
-%\raggedright
-%\pagestyle{empty}
+% \raggedright
+% \pagestyle{empty}
 

### 2018-04-24 Latex/sections/1intro.tex
Public pre-print version.

@@
+\begin{center}
+  Data and scripts: \url{https://github.com/rsling/measurenps}
+\end{center}
+
 \section{Prototypes, exemplars, and corpora in cognitive linguistics}
@@
 Multifactorial models are constructed which incorporate influencing factors from diverse levels, including lexical and contextual factors.
-The estimation of the model coefficients quantifies the influence that the factors have on the probability that either alternant is chosen.
+The estimation of the model coefficients quantifies the influence which the factors have on the probability that either alternant is chosen.
 There are two fundamental issues to consider with respect to this tradition.
@@
 Nobody assumes that we can inductively infer mental representations from experiments, which even allow for direct access to the cognitive agent and offer much better ways of controlling experimental conditions and nuisance variables.
-Rather, predictions are derived from existing theories in order to \textit{test} the theories.
+Rather, predictions are derived from existing theories in order to test these theories.
 While this approach is applicable to corpus data, I discuss some relevant differences below.
@@
 As mentioned above, the central question is whether usage data as found in corpora are truly predictive of speakers' and writers' cognitive representation of language and\slash or of their linguistic behaviour.
-This is where the experimental validation (or \textit{corroboration}) comes into play.
+This is where the experimental validation (or corroboration) comes into play.
 A review of the state of the art was provided by \cite{NewmanSorensenduncan2015}, who enumerate a number of studies showing how corpus data and experimental data converge (such as \citealp{BresnanEa2007,DurrantDoherty2010,GriesWulff2005,GriesEa2005}) and a number of studies where the two types of data led to diverging or only partially converging results (such as \citealp{ArppeJaervikivi2007,Dabrowska2014,Mollin2009}).
@@
 \footnote{See \citet{StormsEa2000} for a comparison of the theories in different experimental settings.}
-Since corpus data only show artefacts of production events and we have no experimental access to the speaker's or writer's performance and their actual similarity judgements, one should be sceptical whether corpus analysis alone could ever help to decide which theory of mental representation is more suitable (see also \citealp[22]{Gries2003} and the \citealp[486--487]{Dabrowska2016} quote above).
+Since corpus data only show artefacts of production events and we have no experimental access to the speaker's or writer's performance and their actual similarity judgements, one should be sceptical whether corpus analysis alone could ever help to decide which theory of mental representation is more suitable (see also \citealp[22]{Gries2003} and the quote from \citealp[486--487]{Dabrowska2016} as cited above).
 However, as I will show, some effects are naturally analysed as prototype effects while others are almost necessarily exemplar effects.
@@
 While I am convinced that the results presented in \citet{DivjakArppe2013} are valid and important (especially given the previous and subsequent research the authors have conducted on the data, including experimental work presented in \citealp{DivjakEa2016}), any data set can be clustered to yield a certain number of clusters.
-Thus, the study does not ensure that the clusters emerging from the data correspond to any particular speaker's cognitive representation.
+Thus, the corpus study alone does not ensure that the clusters emerging from the data correspond to any particular speaker's cognitive representation.
 
@@
 From a linguistic perspective, it seems remarkable that \citet[1013]{VoorspoelsEa2011} consider their experimental task -- the assignment of typicality scores to nouns from the domains of \textit{animals} and \textit{artefacts} to categories such as \textit{bird}, \textit{fish}, \textit{clothing}, or \textit{tools} -- a study of ``superordinate natural language categories, whereas most evidence supporting exemplar representations has been found in artificial categories of a more subordinate level''.
-Corpus linguists interested in probabilistic alternation modelling investigate significantly more complex high-level categories and use large and complex feature sets.%
+Corpus linguists interested in probabilistic alternation modelling often investigate significantly more complex high-level categories and use large and complex feature sets.%
 \footnote{However, approaches have emerged which model linguistic phenomena without reference to high-level linguistic features altogether \citep{BaayenEa2016,RamscarPort2016}.}

### 2018-05-16 R/4_glmm-mcmc.R
Adapted MCMC script for introduction to habilitation.

@@ -1,10 +1,36 @@
 require(rstanarm)
 require(rstan)
 
+rm(list = ls())
 
 rstan_options(auto_write = TRUE)
 options(mc.cores = parallel::detectCores())
 
+# Set this to the location of R script sources.
+setwd('/Users/user/Workingcopies/measurenps/GitHub/R')
+
+# Load data as published in COGL (no MCMC).
+load(file=paste("output/corpus_workspace.RData", sep=""))
+
+# Set font for Habil Rahmentext (Glossa style).
+library(showtext)
+font_paths("/Library/Fonts")
+font_add("Fira Sans",
+         regular    = "FiraSans-Book.otf",
+         bold       = "FiraSans-Bold.otf",
+         italic     = "FiraSans-BookItalic.otf",
+         bolditalic = "FiraSans-BoldItalic.otf")
+showtext_auto()
+
+# Options.
+set.seed(239642)                             # Do NOT change, or results will differ.
+save.persistent        <- T                  # Write to files instead of screen/console.
+out.dir <- paste(out.dir, "mcmc_", sep="")   # Put generated files here.
+precision              <- 3                  # Precision in result table.
+
+chains <- 4
+iter <- 1000
+seed <- 2346
 
 measure.glmm.mcmc <- stan_glmer(Construction~1
                                 
@@ -15,15 +41,12 @@ measure.glmm.mcmc <- stan_glmer(Construction~1
                                 +Cardinal
                                 +Genitives
                                 +Measurecase
-                                +Measurenumber
                                 
                                 +Kindattraction
-                                #+Kindcollo
                                 +Kindfreq
                                 +Kindgender
                                 
                                 +Measureattraction
-                                #+Measurecollo
                                 +Measureclass
                                 +Measurefreq
                            ,
@@ -78,7 +101,7 @@ dotchart(rep(-100, length(measure.fixeffs)), pch=20,
          lcolor = "gray",
          cex = 1.2,
 #         main=paste("MLE and MCMC oefficient estimates\n with 95% confidence intervals", sep="")
-         main=NULL
+         main="MLE and MCMC oefficient estimates\n with 95% confidence intervals"
          )
 lines(c(0,0), c(0,length(measure.ci.95)), col="gray")
 
@@ -94,3 +117,6 @@ for (i in 1:nrow(measure.ci.95)) {
[…]

### 2018-05-29 Latex/sections/3corpus.tex
Final touches after final acceptance.

@@
 DeReKo contains almost exclusively newspaper texts.%
-\footnote{It was shown in \cite{W16-2601} that, for example, the range of topics is much smaller in DeReKo than in DECOW14A.}
+\footnote{It was shown in \cite{BildhauerSchaefer2016} that, for example, the range of topics is much smaller in DeReKo than in DECOW14A.}
 The second reason for choosing this particular corpus is that normative grammars often adopt clear positions regarding the grammaticality of either the \NACa\ or the \PGCa.
@@
 
-For this pre-study, each of the 2,365 combinations of measure noun lemma and kind noun lemma were queried in the \NACb\ and the \PGCd\ with a limit of 100 randomly chosen results for each single query.
+For this pre-study, each of the 2,365 combinations of measure noun lemma and kind noun lemma were queried in the \NACb\ and the \PGCd\ with a limit of 100 randomly chosen results for each single query.%
+\footnote{The data package for this paper (including all concordances and R scripts) is made available under the DOI \doi{10.5281/zenodo.1254871}.}
 From the resulting 35,766 sentences, 5,000 were randomly sampled for the statistical analysis.

### 2018-05-30 Latex/article.tex
Adapted MCMC script for introduction to habilitation.
Final touches after final acceptance.
Final final touches.

@@
 
-%\usepackage[utf8]{inputenc}
 \RequirePackage[no-math]{fontspec}
@@
 \usepackage{soul}
-
-\usepackage[firstpage]{draftwatermark}
-\SetWatermarkFontSize{24pt}
-\SetWatermarkHorCenter{300pt}
-\SetWatermarkColor[rgb]{0.9,0.2,0.2}
-\SetWatermarkText{\textsf{Final revision. Quote as draft and to appear in COGL.}}
 
@@
 
-% \runninglinenumbers
-% \modulolinenumbers[1]
-% \setlength\linenumbersep{20pt}
-
   %\articletype{...}
@@
   \runningauthor{Roland Schäfer}
-  \affil[1]{Freie Universität Berlin}
+  \affil[1]{Deutsche und niederländische Philologie, Freie Universität Berlin, email: roland.schaefer@fu-berlin.de}
 
@@
     The weak effect size of the experimental validation is discussed in the context of corpus-based cognitive linguistics and the validation of corpus-derived models.}
-  \keywords{prototypes vs.\ exemplars, corpus methods and experimental validation, alternations, multilevel modeling, pseudo-partitives, German}
+  \keywords{prototypes vs.\ exemplars, corpus methods and experimental validation, morpho-syntactic alternations, multilevel modeling, pseudo-partitives, German}
   %\classification[PACS]{...}
@@
   %\dedication{...}
-  \received{18 April 2017}
-  \accepted{3 April 2018}
-  \journalname{To appear in Cognitive Linguistics}
+  %\received{}
+  %\accepted{}
+  %\journalname{}
   %\journalyear{}
@@
 \maketitle
-
-% For word counting.
-% \raggedright
-% \pagestyle{empty}
 

### 2018-05-30 Latex/sections/1intro.tex
Final touches after final acceptance.
Final final touches.

@@
-\begin{center}
-  Data and scripts: \url{https://github.com/rsling/measurenps}
-\end{center}
-
 \section{Prototypes, exemplars, and corpora in cognitive linguistics}
@@
 The experimental approach with reduced and\slash or artificial stimuli makes it much easier to examine very specific effects in the behaviour of the subjects.
-While I am convinced that the results presented in \citet{DivjakArppe2013} are valid and important (especially given the previous and subsequent research the authors have conducted on the data, including experimental work presented in \citealp{DivjakEa2016}), any data set can be clustered to yield a certain number of clusters.
+While I am convinced that the results presented in \citet{DivjakArppe2013} are valid and important (especially given the previous and subsequent research the authors have conducted on the data, including experimental work presented in \citealp{DivjakGries2008,DivjakEa2016}), any data set can be clustered to yield a certain number of clusters.
 Thus, the corpus study alone does not ensure that the clusters emerging from the data correspond to any particular speaker's cognitive representation.

### 2018-05-30 Latex/sections/3corpus.tex
Final final touches.

@@
 For this pre-study, each of the 2,365 combinations of measure noun lemma and kind noun lemma were queried in the \NACb\ and the \PGCd\ with a limit of 100 randomly chosen results for each single query.%
-\footnote{The data package for this paper (including all concordances and R scripts) is made available under the DOI \doi{10.5281/zenodo.1254871}.}
+\footnote{The data package for this paper (including all concordances and R scripts) is made available under the DOI \url{https://dx.doi.org/10.5281/zenodo.1254871}.}
 From the resulting 35,766 sentences, 5,000 were randomly sampled for the statistical analysis.

### 2018-05-30 Latex/sections/4experiments.tex
Final version.
Final final touches.

@@
 In retrospect, this stimulus was badly chosen because \textit{heiße Luft} `ado' (literally `hot air') is a fixed metaphorical expression.
-This obviously influences the reactions of participants, and a revised and improved experiment might lead to a much better fit in future research.
+This obviously influences the reactions of participants, and a revised and improved experiment might lead to a much better fit in future research.%
+\footnote{Since I am convinced that the statistical analysis of an experiment has to be determined as part of the experiment's design before the experiment is conducted, removing the data points corresponding to the problematic stimulus from the analysis is not an option.
+Future experiments with improved stimuli is the proper route to take.}
 
@@
 However, the covariance parameters were estimated at exactly -1, which is a clear sign that the variance-covariance matrix could not be estimated successfully.
-The same was true for models with only an \textit{Item} and a \textit{Participant} random slope.
+The same was true for models with only an \textit{Item} or only a \textit{Participant} random slope.
 This is exactly the kind of model overparametrisation criticised in \citet{BatesEa2015a} and \citet{MatuschekEa2017}.

### 2018-05-30 Latex/sections/3corpus.tex
Final final final touches.

--- 2018-05-29 Latex/sections/3corpus.tex
@@
 For this pre-study, each of the 2,365 combinations of measure noun lemma and kind noun lemma were queried in the \NACb\ and the \PGCd\ with a limit of 100 randomly chosen results for each single query.%
-\footnote{The data package for this paper (including all concordances and R scripts) is made available under the DOI \doi{10.5281/zenodo.1254871}.}
+\footnote{The data package for this paper (including all concordances and R scripts) is made available under the following DOI\slash URL: \url{https://dx.doi.org/10.5281/zenodo.1254871}.}
 From the resulting 35,766 sentences, 5,000 were randomly sampled for the statistical analysis.

### 2018-10-08 Latex/article.tex
Preprint w/ watermark after submission.
Adding planned vol. & issue.
Correcting planned vol. & issue.
Copy-editing hell.

@@
 
-\usepackage{gb4e-}
+\usepackage{langsci-gb4e}
 
@@
 
-\usepackage{natbib}
-\bibliographystyle{./unified}
+% \usepackage{natbib}
+% \bibliographystyle{./unified}
 
-\usepackage{lineno}
+% This will most likely be changed by the typesetters.
+% I put it in to get through LaTeX-unaware copy-editing.
+\usepackage[maxbibnames=99,
+  maxcitenames=2,
+  uniquelist=false,
+  backend=biber,
+  bibstyle=biblatex-sp-unified,
+  citestyle=sp-authoryear-comp]{biblatex}
+\newcommand{\citealp}[2][]{\citealt[#1]{#2}}
+\addbibresource{article.bib}
+\DefineBibliographyStrings{english}{%
+  volume  = {Vol.},
+}
+\renewcommand*\finalnamedelim{\addspace and\space}
+\renewcommand*{\multicitedelim}{\addsemicolon\space}
 
@@
 
-  %\articletype{...}
-
   \author*[1]{Roland Schäfer}
   \runningauthor{Roland Schäfer}
-  \affil[1]{Deutsche und niederländische Philologie, Freie Universität Berlin, email: roland.schaefer@fu-berlin.de}
+  \affil[1]{Deutsche und niederländische Philologie, Freie Universität Berlin, Germany, E-mail: roland.schaefer@fu-berlin.de}
 
-  \title{Abstractions and exemplars: the measure noun phrase alternation in German}
-  \runningtitle{Abstraction and exemplars: the measure noun phrase alternation German}
-  %\subtitle{...}
+  \title{Abstractions and exemplars: The measure noun phrase alternation in German}
+  \runningtitle{Abstraction and exemplars: The measure noun phrase alternation German}
 
@@
     I propose a prototype model which assigns a common prototypical meaning to each of the alternants and its corresponding non-alternating construction.
-    Based on this, I argue that lexical, morpho-syntactic, and stylistic features help to predict the choice of the alternant.
+    Based on this, I argue that lexical, morphosyntactic, and stylistic features help to predict the choice of the alternant.
     A large corpus study is presented which supports this analysis.
@@
     The weak effect size of the experimental validation is discussed in the context of corpus-based cognitive linguistics and the validation of corpus-derived models.}
-  \keywords{prototypes vs.\ exemplars, corpus methods and experimental validation, morpho-syntactic alternations, multilevel modeling, pseudo-partitives, German}
-  %\classification[PACS]{...}
-  %\communicated{...}
-  %\dedication{...}
-  %\received{}
-  %\accepted{}
-  %\journalname{}
-  %\journalyear{}
-  %\journalvolume{...}
-  %\journalissue{...}
-  %\startpage{...}
-  %\aop
-  %\DOI{...}
+  \keywords{prototypes vs.\ exemplars, corpus methods and experimental validation, morphosyntactic alternations, multilevel modeling, pseudo-partitives, German}
+
+  \received{April 18, 2017}
+  \revised{January 8, 2018}
+  \accepted{May 30, 2018}
 
@@
 
-\bibliography{rs}
+\renewcommand*\finalnamedelim{\addspace\&\space}
+%\bibliography{article}
+\printbibliography
 

### 2018-10-08 Latex/sections/1intro.tex
Copy-editing hell.

@@
 
-This paper deals with a morpho-syntactic alternation between two measure noun constructions in German.
+This paper deals with a morphosyntactic alternation between two measure noun constructions in German.
 By \textit{alternation} I refer to a situation where two or more forms or constructions are available with no clear difference in acceptability, function, or meaning.
-The study of lexical and constructional alternations has a long history in cognitive corpus linguistics (for example, \citealp{BresnanEa2007,BresnanHay2010,BresnanFord2010,DivjakArppe2013,Gries2015,NessetJanda2010}).
+The study of lexical and constructional alternations has a long history in cognitive corpus linguistics (for example, \citealp{BresnanEa2007}; \citealt{BresnanHay2010}; \citealt{BresnanFord2010}; \citealt{DivjakArppe2013}; \citealt{Gries2015}; \citealt{NessetJanda2010}).
 This research is based on the assumption that language is a probabilistic phenomenon where alternants are chosen neither deterministically nor fully at random \citep{Bresnan2007}.
@@
 For some time, there has been an interest in correlating probabilistic generalisations extracted from corpus data and results from experimental work \citep{ArppeJaervikivi2007,BresnanEa2007,BresnanFord2010,DivjakGries2008,DivjakEa2016,FordBresnan2013}.
-This is often called a \textit{validation} of the corpus-derived findings, but \citet[303]{Divjak2016a} criticises this choice of words ``because it creates the impression that behavioral experimental data is inherently more valuable than textual data'', citing \cite{TummersEa2005}, who state that a corpus is ``a sample of spontaneous language use that is (generally) realized by native speakers''.
+This is often called a \textit{validation} of the corpus-derived findings, but \citet[303]{Divjak2016a} criticises this choice of words ``because it creates the impression that behavioral experimental data is inherently more valuable than textual data'', citing \citet{TummersEa2005}, who state that a corpus is ``a sample of spontaneous language use that is (generally) realized by native speakers''.
 See also \citet{Newman2011} for a positive view of corpora as a source of data in cognitive linguistics in their own right.
@@
 This is where the experimental validation (or corroboration) comes into play.
-A review of the state of the art was provided by \cite{NewmanSorensenduncan2015}, who enumerate a number of studies showing how corpus data and experimental data converge (such as \citealp{BresnanEa2007,DurrantDoherty2010,GriesWulff2005,GriesEa2005}) and a number of studies where the two types of data led to diverging or only partially converging results (such as \citealp{ArppeJaervikivi2007,Dabrowska2014,Mollin2009}).
+A review of the state of the art was provided by \citet{NewmanSorensenduncan2015}, who enumerate a number of studies showing how corpus data and experimental data converge (such as \citealp{BresnanEa2007,DurrantDoherty2010,GriesWulff2005,GriesEa2005}) and a number of studies where the two types of data led to diverging or only partially converging results (such as \citealp{ArppeJaervikivi2007,Dabrowska2014,Mollin2009}).
 When researchers do not achieve convergence, they often try to explain this by differentiating between the actual cognitive construct and whatever the pooled usage data as found in corpora represent.
@@
 Alternatively, researchers argue for a more adequate statistical analysis to increase the fit between corpus data and experimental data.
-See, for example \cite{DivjakEa2016}, who show that generalised additive models (GAMs) are better suited than generalised linear models (GLMs) for correlating reading times and corpus data. 
-Much more optimistically, \cite{StefanowitschFlach2016} proposed a straightforward positivist perspective of corpora as representing the input of an average adult speaker, thus licensing inferences from corpus data to cognitive representations under a ``corpus-as-input'' view.
+See, for example \citet{DivjakEa2016}, who show that generalised additive models (GAMs) are better suited than generalised linear models (GLMs) for correlating reading times and corpus data. 
+Much more optimistically, \citet{StefanowitschFlach2016} proposed a straightforward positivist perspective of corpora as representing the input of an average adult speaker, thus licensing inferences from corpus data to cognitive representations under a ``corpus-as-input'' view.
 They state that ``in this wider context, large, register-mixed corpora such as the British National Corpus [\ldots] may not be perfect models of the linguistic experience of adult speakers, but they are reasonably close to the input of an idealized average member of the relevant speech community'' \citep[104]{StefanowitschFlach2016}.
@@
 In other words, a variant of prototype theory with features \citep{Rosch1978} is the appropriate model, also because the features used to feed the model are usually abstract high-level linguistic features.
-Some researchers such as \cite{Gries2003}, \cite{NessetJanda2010}, or \cite{Schaefer2016c} indeed commit to prototype theory in alternation modelling.
+Some researchers such as \citet{Gries2003}, \citet{NessetJanda2010}, or \citet{Schaefer2016c} indeed commit to prototype theory in alternation modelling.
 Under prototype theory, category membership is defined by similarity to an ideal exemplar and its characterising features (see \citealp{Taylor2008} for an overview).
@@
 The theories differ significantly in whether they assume higher-level abstractions in the form of single maximally prototypical exemplars (prototype theory) or categories emerging through the storage of many exemplars and similarity-based classification over those sets of exemplars (exemplar theory).
-\cite{Barsalou1990} already showed that prototype and exemplar theory model the same types of surface effects and are informationally equivalent.
+\citet{Barsalou1990} already showed that prototype and exemplar theory model the same types of surface effects and are informationally equivalent.
 Consequently, experiments which favour one theory over the other use procedural behaviour of subjects in experiments, for example the speed of category retrieval, as opposed to mere output data.%
@@
 Still, various attempts have been made to settle the dispute between abstraction-based models (models with rules or prototypes) and exemplar models.
-\cite{VanpaemelStorms2008} and \citet{LeeVanpaemel2008} proposed the \textit{varying abstraction model} (VAM) which ``attempt[s] to balance economy and informativeness'' \citep[745]{LeeVanpaemel2008}, treating models with full abstraction (radical prototype theory) and no abstraction at all (radical exemplar theory) as special cases of a flexible approach which allows both types of mechanisms.
-The mixture model of categorisation (MMC) by \cite{Rosseel2002} is a model with abstraction in the form of hierarchical clusters of exemplars.
+\citet{VanpaemelStorms2008} and \citet{LeeVanpaemel2008} proposed the \textit{varying abstraction model} (VAM) which ``attempt[s] to balance economy and informativeness'' \citep[745]{LeeVanpaemel2008}, treating models with full abstraction (radical prototype theory) and no abstraction at all (radical exemplar theory) as special cases of a flexible approach which allows both types of mechanisms.
+The mixture model of categorisation (MMC) by \citet{Rosseel2002} is a model with abstraction in the form of hierarchical clusters of exemplars.
 These clusters of objects are characterised by a probability distribution over their features, and categorising new objects is a process of estimating the probability of this object belonging to one of the clusters.
-\cite{GriffithsEa2009} go further and present a computational model which is able to choose the appropriate complexity of representation for a given category.
+\citet{GriffithsEa2009} go further and present a computational model which is able to choose the appropriate complexity of representation for a given category.
 However, despite these and more attempts to reconcile or unite the two approaches \citet[183--184]{Vanpaemel2016} describes the state of affairs between adherents of neo-prototype theory (such as \citealp{MindaSmith2001,MindaSmith2002}) and exemplar theory as a stalemate.
@@
 In cognitive linguistics, similarity-based categorisation is often seen as the central and conceptually sufficient approach to linguistic categorisation.
-See \cite{Taylor2003} for a comprehensive treatment in terms of prototype theory and \cite{Taylor2012} for a detailed discussion of an exemplar-based approach.
-\cite{DivjakArppe2013} is a rare example of a paper in cognitive linguistics where a synthesis between prototype and exemplar models is proposed.
+See \citet{Taylor2003} for a comprehensive treatment in terms of prototype theory and \citet{Taylor2012} for a detailed discussion of an exemplar-based approach.
+\citet{DivjakArppe2013} is a rare example of a paper in cognitive linguistics where a synthesis between prototype and exemplar models is proposed.
 Their corpus-based approach shows ``one way of systematically analyzing usage data as contained in corpora to yield a scheme, compatible with usage-based theories of language, by which the assumptions of both the prototype and exemplar theories can be operationalized'' \citep[267]{DivjakArppe2013}.

### 2018-10-08 Latex/sections/2descriptive.tex
Copy-editing hell.

@@
 \footnote{Some descriptive and normative grammars take stronger positions with regard to the acceptability of the two options.
-See \cite{Hentschel1993} and \cite{Zimmer2015} for analyses of the sometimes absurd stances taken in grammars of German.
+See \citet{Hentschel1993} and \citet{Zimmer2015} for analyses of the sometimes absurd stances taken in grammars of German.
 }
@@
     we drink a glass good wine \\
-    \glt We drink a glass of good wine.}
+    \glt `We drink a glass of good wine.'}
     \ex[ ]{\label{ex:intro:alternation2} Wir trinken [[ein Glas]\Sub{Acc} [guten Wein]\Sub{Acc}]\Sub{Acc}.}
@@
 
-This specific configuration has to be seen in the context of two other configurations, to which I turn now.
+\noindent This specific configuration has to be seen in the context of two other configurations, to which I turn now.
 First, if the kind noun forms an NP with a determiner, the construction resembles (and is usually called) a \textit{pseudo-partitive} (on partitives and pseudo-partitives see, e.\,g., \citealp{Barker1998,Selkirk1977,Stickney2007,Vos1999}; for a recent application of the terminology to German, see \citealp{Gerstenberger2015}).%
@@
 Whereas partitives are constructions denoting a proper part-of relation as in \textit{a sip of the wine}, pseudo partitives merely denote quantities and contain indefinite kind nouns as in \textit{a sip of wine}.
-In the literature on German, some authors incorrectly call the pseudo-partitive a \textit{partitive} \citep{Hentschel1993} while some realise the difference and at least mention it \citep{Eschenbach1994,GallmannLindauer1994,Loebel1989,Zimmer2015}.
+In the literature on German, some authors incorrectly call the pseudo-partitive a partitive \citep{Hentschel1993} while some realise the difference and at least mention it \citep{Eschenbach1994,GallmannLindauer1994,Loebel1989,Zimmer2015}.
 }
@@
   we drink a glass this wine\\
-  \glt We drink a glass of this wine.}
+  \glt `We drink a glass of this wine.'}
 \end{exe}
 
-Second, if the kind noun is bare -- \ie\ if it comes neither with a determiner nor a modifying adjective -- it is uninflected as in (\ref{ex:intro:narrowapposition1}), and the genitive as seen in the PGC is not acceptable, see (\ref{ex:intro:narrowapposition2}).%
+\noindent Second, if the kind noun is bare -- \ie\ if it comes neither with a determiner nor a modifying adjective -- it is uninflected as in (\ref{ex:intro:narrowapposition1}), and the genitive as seen in the PGC is not acceptable, see (\ref{ex:intro:narrowapposition2}).%
 \footnote{
@@
     we drink a glass wine\\
-    \glt We drink a glass of wine.}
+    \glt `We drink a glass of wine.'}
     \ex[*]{\label{ex:intro:narrowapposition2} Wir trinken [[ein Glas]\Sub{Acc} [Weins]\Sub{Gen}]\Sub{Acc}.}
@@
 
-This construction is usually classified as a \textit{Narrow Apposition Construction} \citep{Loebel1986}, henceforth NAC.
+\noindent This construction is usually classified as a Narrow Apposition Construction \citep{Loebel1986}, henceforth NAC.
 Notice that the unavailability of the genitive on the kind noun follows independently from a constraint that genitive NPs in German require the presence of some strongly case-marked element (determiner or adjective) in addition to the head noun in order to be acceptable (\citealp{GallmannLindauer1994,Schachtl1989}; see also \citealp[160]{Eisenberg2013b}).
@@
   \centering
+  \caption{NAC and PGC constructions in different NP structures with examples and references to full example sentences}
   \begin{tabular}{llll}
@@
   \end{tabular}
-  \caption{NAC and PGC constructions in different NP structures with examples and references to full example sentences}
   \label{tab:constructions}
@@
 
-In order to understand the MNP alternation, it is vital to consider how in the relevant syntactic structure [N\Sub{1}~[A~N\Sub{2}]\Sub{NP\Sub{2}}]\Sub{NP\Sub{1}} (as instantiated in the \NACa\ and the \PGCa), the adjective is morpho-syntactically ambiguous between a determiner and a modifying adjective.
+In order to understand the MNP alternation, it is vital to consider how in the relevant syntactic structure [N\Sub{1}~[A~N\Sub{2}]\Sub{NP\Sub{2}}]\Sub{NP\Sub{1}} (as instantiated in the \NACa\ and the \PGCa), the adjective is morphosyntactically ambiguous between a determiner and a modifying adjective.
 To show this, the so-called strong and weak inflection patterns of German adjectives needs to be taken into account.
@@
 
-Thus, the adjectives in the \NACa\ and the \PGCa\ have properties of adjectives as well as determiners.
+\noindent Thus, the adjectives in the \NACa\ and the \PGCa\ have properties of adjectives as well as determiners.
 On the one hand, they are lexical adjectives and function as attributive modifiers.
@@
 However, if it is classified as an adjective, the \NACa\ suggests itself because of the lack of a determiner.%
-\footnote{The generative analysis presented in \cite{Bhatt1990} comes close to this interpretation by analysing the kind NP in the PGC as a DP and in the NAC as an NP.}
-This morpho-syntactic ambiguity means that the \NACa\ is in fact a \NACb\ in disguise, and the \PGCa\ is a \PGCd\ in disguise.
+\footnote{The generative analysis presented in \citet{Bhatt1990} comes close to this interpretation by analysing the kind NP in the PGC as a DP and in the NAC as an NP.}
+This morphosyntactic ambiguity means that the \NACa\ is in fact a \NACb\ in disguise, and the \PGCa\ is a \PGCd\ in disguise.
 This explains why the alternation arises in the first place.
@@
     they took a spoon {some\slash any} medication \\
-    \trans They took a spoon(ful) of any old medication.
+    \trans `They took a spoon(ful) of any old medication.'
   }
@@
     they bought three liters oil \\
-    \trans They bought three liters of oil.
+    \trans `They bought three liters of oil.'
   }
@@
 
-While the \PGCd\ in (\ref{ex:prototypes1}) allows an interpretation where speakers conceptualise the substance itself (\ie\ the medication) and an actual spoon used to measure the quantity of the medication, the \NACb\ in (\ref{ex:prototypes2}) does not allow such an interpretation.
+\noindent While the \PGCd\ in (\ref{ex:prototypes1}) allows an interpretation where speakers conceptualise the substance itself (\ie\ the medication) and an actual spoon used to measure the quantity of the medication, the \NACb\ in (\ref{ex:prototypes2}) does not allow such an interpretation.
 While in the given examples, the effect is strongly supported by the choice of the measure noun lemmas, I argue that the meaning of the prototypes is independent of this.
@@
                                        three centilitres hot rum are enough\\
-				     \glt Three centilitres of hot rum is enough.}
+				     \glt `Three centilitres of hot rum is enough.'}
     \ex{\label{ex:determiners:pgc}\gll [[Einige Centiliter]\Sub{Nom} [heißen Rums]\Sub{Gen}]\Sub{Nom} sind genug.\\
                                        some   centilitres hot    rum  are   enough\\
-				     \glt A few centilitres of hot rum is enough.}
+				     \glt `A few centilitres of hot rum is enough.'}
   \end{xlist}
@@
 
-In (\ref{ex:determiners:nac}), the measure noun is modified by a cardinal \textit{drei} `three', and hence the \NACa\ is preferred.
+\noindent In (\ref{ex:determiners:nac}), the measure noun is modified by a cardinal \textit{drei} `three', and hence the \NACa\ is preferred.
 In (\ref{ex:determiners:pgc}), the measure noun is modified by a non-cardinal determiner \textit{einige} `some', and the \PGCa\ is preferred.
@@
 Should this be confirmed, it would be highly implausible to conceive of such an effect as a prototype effect.
-In the corpus study reported in Section~\ref{sec:corpusstudies}, a measure quantifying this influence will therefore be included as the \textit{attraction strength}.
+In the corpus study reported in Section~\ref{sec:corpusstudies}, a measure quantifying this influence will therefore be included as the attraction strength.
 
-It should be noted that such an exemplar-type effect would have to be confirmed \textit{in addition to} the predicted semantic prototype effect for measure lemmas described in Section~\ref{sec:prototypeeffects}, namely the effect of semantic classes of measure nouns.
+It should be noted that such an exemplar-type effect would have to be confirmed in addition to the predicted semantic prototype effect for measure lemmas described in Section~\ref{sec:prototypeeffects}, namely the effect of semantic classes of measure nouns.
 It must also be ensured that these influences at the lemma level are not spurious and just artefacts of mere lemma idiosyncrasies.

### 2018-10-08 Latex/sections/3corpus.tex
Copy-editing hell.

@@
 
-For the present study, I used the 21 billion token German \textit{Corpus from the Web} (COW) in its 2014 version DECOW14A (\citealp{SchaeferBildhauer2012full}, and \citealp{Schaefer2015b}, as well as \citealp{BiemannEa2013}, and \citealp{SchaeferBildhauer2013}, for overviews of web corpora in general and the methodology of their construction).%
+For the present study, I used the 21 billion token German \textit{Corpus from the Web} (COW) in its 2014 version DECOW14A (\citealp{SchaeferBildhauer2012full}, and \citealp{Schaefer2015b}, as well as \citealp{BiemannEa2013}, and \citealp{SchaeferBildhauer2013alt}, for overviews of web corpora in general and the methodology of their construction).%
 \footnote{The COW corpora (Dutch, English, French, German, Spanish, Swedish) are made available for free at \url{https://www.webcorpora.org}.
@@
 DeReKo contains almost exclusively newspaper texts.%
-\footnote{It was shown in \cite{BildhauerSchaefer2016} that, for example, the range of topics is much smaller in DeReKo than in DECOW14A.}
+\footnote{It was shown in \citet{BildhauerSchaefer2016} that, for example, the range of topics is much smaller in DeReKo than in DECOW14A.}
 The second reason for choosing this particular corpus is that normative grammars often adopt clear positions regarding the grammaticality of either the \NACa\ or the \PGCa.
@@
   \centering
+  \caption{Annotated variables for the corpus studies}
   \begin{tabular}{llll}
@@
   \end{tabular}
-  \caption{Annotated variables for the corpus studies}
   \label{tab:variables}
@@
 At the document level, the DECOW14A corpus has an annotation for \textit{Badness}.
-As described in \cite{SchaeferEa2013}, \textit{Badness} measures how well the distribution of highly frequent short words in the document matches a language model of standard German.
+As described in \citet{SchaeferEa2013}, \textit{Badness} measures how well the distribution of highly frequent short words in the document matches a language model of standard German.
 The paper also shows that the Badness score corresponds robustly with human raters' intuitions about text coherence and text quality.
@@
 
-\subsection{Pre-study: main prototype effects in the non-alternating constructions}
+\subsection{Pre-study: Main prototype effects in the non-alternating constructions}
 \label{sec:prestudy}
@@
   \centering
+  \caption{Coefficient table with 95\% bootstrap confidence intervals for the pre-study; the intercept is -5.370}
   \resizebox{\textwidth}{!}{
@@
   }
-  \caption{Coefficient table with 95\% bootstrap confidence intervals for the pre-study; the intercept is -5.370}
   \label{tab:smalltable}
@@
 
-\subsection{Main study: the alternation}
+\subsection{Main study: The alternation}
 \label{sec:annotation}
@@
   \centering
+  \caption{Coefficient table with 95\% bootstrap confidence intervals for the main study; the intercept is -3.548}
   \resizebox{\textwidth}{!}{
@@
   }
-  \caption{Coefficient table with 95\% bootstrap confidence intervals for the main study; the intercept is -3.548}
   \label{tab:bigtable}

### 2018-10-08 Latex/sections/4experiments.tex
Final final final touches.
Copy-editing hell.

@@
 
-\subsection{Experiment 1: forced-choice}
+\subsection{Experiment 1: Forced-choice}
 \label{sec:exp:fc}
@@
 In the analysis, the probabilities assigned to the stimuli by the corpus-derived model were compared to the frequency with which participants chose the alternants.
-There participants were 24 native speakers of German without reading or writing disabilities, aged 19 to 30, and living permanently in Berlin.
+The participants were 24 native speakers of German without reading or writing disabilities, aged 19 to 30, and living permanently in Berlin.
 They were recruited from introductory linguistics courses at Freie Universität Berlin.
@@
   \centering
+  \caption{The four groups of sentences chosen as stimuli; in each group of four sentences, combinations of important factor values were made unique whenever possible}
   \begin{tabular}[h]{lll}
@@
   \end{tabular}
-  \caption{The four groups of sentences chosen as stimuli; in each group of four sentences, combinations of important factor values were made unique whenever possible}
   \label{tab:experiment1:design}
@@
 The order of sentences was randomised for each participant, and the position of the alternants on the screen (left\slash right) was randomised per participant and sentence.
-As fillers, 23 pairs of sentences exemplifying similar but unrelated alternation phenomena from German morpho-syntax were used.
+As fillers, 23 pairs of sentences exemplifying similar but unrelated alternation phenomena from German morphosyntax were used.
 Thus, participants saw 39 pairs of sentences and 78 sentences in total.
@@
 
-In retrospect, this stimulus was badly chosen because \textit{heiße Luft} `ado' (literally `hot air') is a fixed metaphorical expression.
+\noindent In retrospect, this stimulus was badly chosen because \textit{heiße Luft} `ado' (literally `hot air') is a fixed metaphorical expression.
 This obviously influences the reactions of participants, and a revised and improved experiment might lead to a much better fit in future research.%
 \footnote{Since I am convinced that the statistical analysis of an experiment has to be determined as part of the experiment's design before the experiment is conducted, removing the data points corresponding to the problematic stimulus from the analysis is not an option.
-Future experiments with improved stimuli is the proper route to take.}
+Future research with improved stimuli is the proper route to take.}
 
@@
 
-\subsection{Experiment 2: self-paced reading}
+\subsection{Experiment 2: Self-paced reading}
 \label{sec:exp:spr}
@@
 In this section, a self-paced reading experiment is presented.
-In a very similar fashion, \cite{DivjakEa2016} apply the self-paced reading paradigm in the validation of corpus-based models.
+In a very similar fashion, \citet{DivjakEa2016} apply the self-paced reading paradigm in the validation of corpus-based models.
 The analysis compares the corpus-derived probabilities with potential lags in reading time for sentences with the preferred and the non-preferred constructions.
@@
   \centering
+  \caption{Fixed effect coefficient table for the LMM used to analyse the self-paced reading experiment; the intercept is 0.829}
   \begin{tabular}{lrrrc}
@@
   \end{tabular}
-  \caption{Fixed effect coefficient table for the LMM used to analyse the self-paced reading experiment; the intercept is 0.829}
   \label{tab:exp:spr}

### 2018-10-08 Latex/sections/5conclusions.tex
Copy-editing hell.

--- 2018-04-11 Latex/sections/5conclusions.tex
@@
 As discussed in Section~\ref{sec:cogocl}, it is unnecessary to follow an extreme route, be it radical prototype theory or radical exemplar theory, given recent developments in cognitive science and cognitive linguistics.
-While it is surely an intentionally overstated comment, \citet[15]{Kapatsinski2014} even suggests that ``[i]n the extreme, some speakers’ heads could host exemplar models, and some could contain fairly abstract grammars, and the produced output would be essentially identical''.
+While it is surely an intentionally overstated comment, \citet[15]{Kapatsinski2014} even suggests that ``in the extreme, some speakers’ heads could host exemplar models, and some could contain fairly abstract grammars, and the produced output would be essentially identical''.
 If this is the case (even to a less extreme degree), corpora only provide pooled data averaged over speaker grammars with varying levels of abstraction.
@@
 \begin{quote}
-  There are now a number of published multivariate models that use data[,] extracted from corpora [\ldots] to predict the choice for one morpheme, lexeme or construction over another.
+  There are now a number of published multivariate models that use data, extracted from corpora [\ldots] to predict the choice for one morpheme, lexeme or construction over another.
   However, [\ldots] only a small number of these corpus-based studies have been cross-validated [\ldots].
@@
 The statistical model was a true multilevel model (see Section~\ref{sec:mainstudystatisticalmodel}), demonstrating how multilevel modelling helps to specify complex hierarchies of abstract features, item-specific tendencies, and exemplar effects at the level of observations (sentences; first level) and lemmas (second level).
-\cite{Gries2015} still calls mixed models ``underused'' in corpus linguistics, and multilevel models are consequently also an underused tools.
-At the same time, the ill effects of overparametrisation of mixed models with varying slopes as criticised by \cite{BatesEa2015a} were demonstrated.
+\citet{Gries2015} still calls mixed models ``underused'' in corpus linguistics, and multilevel models are consequently also an underused tools.
+At the same time, the ill effects of overparametrisation of mixed models with varying slopes as criticised by \citet{BatesEa2015a} were demonstrated.
 Furthermore, fitting an additive model to the reading time data did not improve the fit as there were no non-linearities in the data.
-While \cite{DivjakEa2016} also use attested sentences, they find that an additive model helped to deal with non-linearities.
+While \citet{DivjakEa2016} also use attested sentences, they find that an additive model helped to deal with non-linearities.
 This is clearly another area where only more studies can lead to clarification.
-Finally, much like \cite{Dabrowska2014} found that speakers' knowledge of collocations was not matched by a set of standard measures of collocation strength extracted from corpora, a simple quotient based on raw frequencies for the attraction strength performed much better in the present study compared to collexeme strength in the form of logarithmised Fisher p-values.
+Finally, much like \citet{Dabrowska2014} found that speakers' knowledge of collocations was not matched by a set of standard measures of collocation strength extracted from corpora, a simple quotient based on raw frequencies for the attraction strength performed much better in the present study compared to collexeme strength in the form of logarithmised Fisher p-values.
 While this does not allow the conclusion that collexeme strength has no cognitive reality, it might indicate that for measuring attraction effects exerted by non-alternating constructions on alternating constructions, taking the marginals into account might not be crucial.%
@@
 
-In closing, I want to point out that the so-called \textit{case of doubt} in German morpho-syntax -- like the measure noun phrase alternation -- are in fact ideal test cases for probabilistic modelling of alternation phenomena in cognitive linguistics.
+In closing, I want to point out that the so-called cases of doubt in German morphosyntax -- like the measure noun phrase alternation -- are in fact ideal test cases for probabilistic modelling of alternation phenomena in cognitive linguistics.
 Doing more research on them could help to provide answers to many of the fundamental and methodological issues raised here.
